### 2013-01-23 stackExchange/writeups/iccm-2013/template/template.tex
foo
Former-commit-id: 2844846a28efaa6e5a743f82168a0af0f799331f [formerly f1e29f6386ecccea3e610f2fb245e0bbe1c62d14]
Former-commit-id: 7a329ae42548c04044a7b3363045fa3bbd56e1b5

% 
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified: Niels Taatgen (taatgen@cmu.edu) 10/24/2006

%% Change ``a4paper'' in the following line to ``letterpaper'' if you are
%% producing a letter-format document.

\documentclass[10pt,letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}

\title{How to Make a Camera-Ready Proceedings Contribution}

\author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\
  Department of Psychology, 1202 W. Johnson Street \\
  Madison, WI 53706 USA
  \AND {\large \bf Sharon J. Derry (SDJ@Macc.Wisc.Edu)} \\
  Department of Educational Psychology, 1025 W. Johnson Street \\
  Madison, WI 53706 USA}

\begin{document}

\maketitle

\begin{abstract}
The abstract should be one paragraph, indented 1/8~inch on both sides,
in 9~point font with single spacing. The heading {\bf Abstract} should
be 10~point, bold, centered, with one line space below it. This
one-paragraph abstract section is required only for standard spoken
papers and standard posters (i.e., those presentations that will be
represented by six page papers in the Proceedings).

\textbf{Keywords:} 
Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term.
\end{abstract}

\section{General Formatting Instructions}

For standard spoken papers and standard posters, the entire contribution (including figures, references, everything) can be no longer than six pages. For abstract posters, the entire contribution can be no longer than one page. For symposia, the entire contribution can be no longer than two pages.

The text of the paper should be formatted in two columns with an
overall width of 7 inches (17.8 cm) and length of 9.25 inches (23.5
cm), with 0.25 inches between the columns. Leave two line spaces
between the last author listed and the text of the paper. The left
margin should be 0.75 inches and the top margin should be 1 inch.
\textbf{The right and bottom margins will depend on whether you use
  U.S. letter or A4 paper, so you must be sure to measure the width of
  the printed text.} Use 10~point Times Roman with 12~point vertical
spacing, unless otherwise specified.

The title should be in 14~point, bold, and centered. The title should
be formatted with initial caps (the first letter of content words
capitalized and the rest lower case). Each author's name should appear
on a separate line, 11~point bold, and centered, with the author's
email address in parentheses. Under each author's name list the
author's affiliation and postal address in ordinary 10~point type.

Indent the first line of each paragraph by 1/8~inch (except for the
first paragraph of a new section). Do not add extra vertical space
between paragraphs.

\section{First-Level Headings}

First level headings should be in 12~point, initial caps, bold and
centered. Leave one line space above the heading and 1/4~ line space
below the heading.

\subsection{Second-Level Headings}

Second level headings should be 11~point, initial caps, bold, and
flush left. Leave one line space above the heading and 1/4~ line
space below the heading.

\subsubsection{Third-Level Headings}

Third-level headings should be 10~point, initial caps, bold, and flush
left. Leave one line space above the heading, but no space after the
heading.

\section{Formalities, Footnotes, and Floats}

Use standard APA citation format. Citations within the text should
include the author's last name and year. If the authors' names are
included in the sentence, place only the year in parentheses, as in
\citeA{NewellSimon1972a}, but otherwise place the entire reference in
parentheses with the authors and year separated by a comma
\cite{NewellSimon1972a}. List multiple references alphabetically and
separate them by semicolons
\cite{ChalnickBillman1988a,NewellSimon1972a}. Use the
et~al. construction only after listing all the authors to a
publication in an earlier reference and for citations with four or
more authors.

\subsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first
  footnote.} in the text. Place the footnotes in 9~point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.}

\subsection{Tables}

Number tables consecutively; place the table number and title (in
10~point) above the table with one line space above the caption and
one line space below it, as in Table~\ref{sample-table}. You may float
tables to the top or bottom of a column, set wide tables across both
columns.

\begin{table}[!ht]
\begin{center} 
\caption{Sample table title.} 
\label{sample-table} 
\vskip 0.12in
\begin{tabular}{ll} 
\hline
Error type    &  Example \\
\hline
Take smaller        &   63 - 44 = 21 \\
Always borrow~~~~   &   96 - 42 = 34 \\
0 - N = N           &   70 - 47 = 37 \\
0 - N = 0           &   70 - 47 = 30 \\
\hline
\end{tabular} 
\end{center} 
\end{table}

\subsection{Figures}

All artwork must be very dark for purposes of reproduction and should
not be hand drawn. Number figures sequentially, placing the figure
number and caption, in 10~point, after the figure with one line space
above the caption and one line space below it, as in
Figure~\ref{sample-figure}. If necessary, leave extra white space at
the bottom of the page to avoid splitting the figure and figure
caption. You may float figures to the top or bottom of a column, or
set wide figures across both columns.

\begin{figure}[ht]
\begin{center}
\fbox{CoGNiTiVe ScIeNcE}
\end{center}
\caption{This is a figure.} 
\label{sample-figure}
\end{figure}

\section{Acknowledgments}

Place acknowledgments (including funding information) in a section at
the end of the paper.

\section{References Instructions}

Follow the APA Publication Manual for citation format, both within the
text and in the reference list, with the following exceptions: (a) do
not cite the page numbers of any book, including chapters in edited
volumes; (b) use the same format for unpublished references as for
published ones. Alphabetize references by the surnames of the authors,
with single author entries preceding multiple author entries. Order
references by the same authors by the year of publication, with the
earliest first.

Use a first level section heading for the reference list. Use a
hanging indent style, with the first line of the reference flush
against the left margin and subsequent lines indented by 1/8~inch.
Below are example references for a conference paper, book chapter,
journal article, technical report, dissertation, book, and edited
volume, respectively.

\nocite{ChalnickBillman1988a}
\nocite{Feigenbaum1963a}
\nocite{Hill1983a}
\nocite{OhlssonLangley1985a}
\nocite{Lewis1978a}
\nocite{NewellSimon1972a}
\nocite{ShragerLangley1990a}

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{bibliography}

\end{document}

### 2013-01-23 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: be9556d67cf02bd7a67262aa7b533049b1b7b272 [formerly 58ad74fc1d7e4305c5437a916e066921b5fe45ee]
Former-commit-id: 3ea097c8ac74346ee72588c3c6685f17a852ae77

% 
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified: Niels Taatgen (taatgen@cmu.edu) 10/24/2006

%% Change ``a4paper'' in the following line to ``letterpaper'' if you are
%% producing a letter-format document.

\documentclass[10pt,letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage[group-separator={,}]{siunitx}

\title{FIXME}

\author{{\large \bf Clayton Stanley (clayton.stanley@rice.edu)} \\
  Department of Psychology, 6100 Main Street \\
  Houston, TX 77005 USA 
  \AND {\large \bf Mike Byrne (byrne@rice.edu)} \\
  Department of Psychology and Computer Science, 6100 Main Street \\
  Houston, TX 77005 USA }

\begin{document}

\maketitle

\begin{abstract}
The abstract should be one paragraph, indented 1/8~inch on both sides,
in 9~point font with single spacing. The heading {\bf Abstract} should
be 10~point, bold, centered, with one line space below it. This
one-paragraph abstract section is required only for standard spoken
papers and standard posters (i.e., those presentations that will be
represented by six page papers in the Proceedings).

\textbf{Keywords:} 
Machine Learning; ACT-R; Large-Scale Declarative Memory
\end{abstract}

\section{Introduction}

Human-based tagging of online content has increased in recent years (\cite{Chang2010}).
A large portion of the growing number of social media sites support some sort of human-directed tagging of posts
Additionally, an increasing number of individuals are using tags as a form of query-based search to find information
(e.g., \cite{Diakopoulos2010}).

Although much good work has been done to characterize the lifetime and growth characteristics of tags for real-time social networking sites such as Twitter
(e.g., \cite{Bauer2012}, \cite{Tsur2012}, \cite{Chang2010})
less is known about a person's motivation for choosing particular tags during the creation of the tweet, post, etc.
Modeling tag creation accurately (as opposed to modeing tag growth and behavior after creation) could allow a system to predict the hashtag that a user should use, given the contextual clues available.
The benefits of this type of prediction are numerous.
For example, users could be introduced to tags (and potentially comunities surrounding those tags) that are of interest to them.
Also tag cleanup systems could be implemented where mistags are identified and corrected.

A tag-prediction system was built here for the StackOverflow dataset that can solve precisely these sorts of problems.
The StackOverflow dataset was chosen because it is relatively constrained, contains a large number of posts and associated tags, and the data are easy to obtain.
It is constrained because there is a fully enumerated and limited set of tags that a poster can use, as opposed to a less-constrained domain such as Twitter.
The site also makes public all posting data on a quarterly basis, which can be easily downloaded and loaded into a relational database for further analysis.

\section{Method}

ACT-R's declarative memory retrieval theory was used to model tag retrievals for StackOverflow posts.
The hypothesis is that the tags used for posts are based on a tag's prior use and the content of words in the title and body of the post.
The model was trained on 2/3 of the dataset (1M posts), and cross validated on 10 samples of 1,000 posts not contained within the training dataset.
A logistic regression statistical technique was used to calibrate model parameters and optimize performance.
Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).

\subsection{Stack Overflow Dataset}

The StackOverflow dataset consists of all questions and answers posted on stackoverflow.com.
The questions are all programming related and can be posted by anyone with a StackOverflow account.
Each question is tagged by the poster with the tags that are most representative of the post.
Example tags are programming languages (PHP, Common-Lisp, MySQL, C\#) and also general topics (databases, optimization, arrays).
An Example question from the stackoverflow.com site is included in Figure FIXME.
Note that Figure FIXME is tagged with 'PHP' and 'magic-quotes' for example.

The Sep' 11 dataset (\cite{DataDump2011}) was downloaded and imported into the MySQL relational database system.
The dataset contains FIXME posts from FIXME unique users.

\subsection{Measuring Co-occurrences}

Model tag predictions are based on the amount of word co-occurrences between words in the title and tags used for that post.
To build the co-occurrence matrix between title words and tags, all title text and associated tags were extracted from the MySQL database.
The Python Natural Language Processing Toolkit (\cite{Bird2009}) was used to chunk (i.e., tokenize and lemmetize) the title text.
The tags were already chunked so those words were not processed any further.
Afterwords the processed words were fed back in to the MySQL database.

A MySQL query was then used to build the title-word, tag word co-occurrence matrix.
For a given word found in any title, the query counts the number of times that word appears with each tag.
Each of those counts is recorded as the word co-occurrence for that title word and each associated tag.
This process is done across all words found in any title.
The result is a comma-separated value (csv) file of title-word, tag, count for each title-word, tag combination.

\subsection{Measuring Tag Occurrences}

Model tag predictions are also based on the frequency that each tag has been used.
More often used tags more likely to have higher activation.
A MySQL query was used to build a similar csv for tag occurrences.
This csv for example would show that C\# is the most often used tag, where FIXME posts were tagged with C\#.

\subsection{Generating Model Predictions}

For a given title in a post, the model returns the activations for each possible tag.
The tag with the highest activation is the most likely tag to be associated with that post.
Tag activation is a function of three components:
The tag's base level activation (prior odds of being used),
contextual activation for words in the title (word co-occurrence weighting),
and contextual activation for words in the body.

Contextual activation is computed by comparing the words in the title with each tag's word co-occurrence vector.
If a tag's co-occurrences correlate strongly with the words in the title, then that tag's activation will be high.
If there is a low correlation between the observed co-occurrences for a tag and the observed words in the title, than that tag's activation will be low.
The equations that operationalize this process will be included in the Model section.

The R statistical programming environment was used to build the model and generate model predictions.
This was done by feeding the word co-occurrence and tag occurrence csv's into a separate sparse matrix and sparse vector in R.
A hash was computed for each unique word, so that indices for the sparse matrix and vector could be numeric.
The dimensions for the co-occurrence sparse matrix were context-word, tag, and the tag occurrence vector was 1 dimensional (tag).

1 million of the ~1.5 million posts were fed into the sparse matrix and vector to build the population co-occurrences and tag occurrences.
An interface to the model was developed so that one could feed in a vector of words, and the model would return a sparse vector of tag activations given that context.
In order to generate a model prediction, tag activations for title and body context were computed separately, and then added to each tag's prior base-level activation.
The resulting tags were rank sorted, and the one with the highest activation would then be the model's prediction of the most likely tag associated with that post.

\subsection{Measuring Model Fit}

For a given post, the model produces a sparse vector of tag activations.
For that post, the poster's chosen tags are known (see Figures \ref{fig:screenTreeview}, \ref{fig:screenMagic}, and \ref{fig:screenConst}).
So model performance can be measured by comparing the model's tag activations with chosen tags for a post, and then aggregated across posts to get an overall average performance.
If the model consistently produces higher tag activations for chosen tags than non-chosen tags, the model should be able to accurately predict the chosen tags.

Logistic regression was used in order to operationalize this fit measure.
For each post, the model's tag activations for the highest 100 tags were recorded, as well as the tag activations for all tags that were chosen by the poster.
Each of these observations was assigned a category of 0 or 1 depending on if the tag was a chosen tag or not for that post (0 is not chosen).
Conceptually, if the model has good categorization power, the observations assigned 1 should have higher tag activations than the observations assigned 0.

These vectors of tag activation, categorization were collected across 10 sets of 1,000 posts.
Each of those sets did not have any posts that were present in the 1 million posts used to train the model.
So a form of cross validation was used to measure model fit, although the more robust full k-fold cross validation technique was not used (due mostly to resource and time limitations).

R was used to perform the logistic regression.
The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from two predictor variables: Tag occurrence activation and word co-occurrence activation.
Further discussion and operationalization of the predictor variables is included in the Model section.

\section{Model}

The tag-prediction model is a cognitively-inspired Bayesian probabilistic model, based on ACT-R's declarative memory retrieval mechanisms (\cite{Anderson2004}).
These equations describe the likelihood that a chunk from memory will be retrieved, given the current context.
In this way the task of retrieving the tag with the highest activation from the words in the title is analogous to performing a declarative memory retrieval request for a tag, given the context of title words.

This model is similar in design to the approach used by SNIF-ACT (\cite{Fu2007}, \cite{Pirolli2003}). 
SNIF-ACT can predict link traversal for search queries.
That is, given a search query (goal state), fetched results, and words included in those fetched results, the model predicts the most likely link that a person will click on.
This StackOverflow model is similar because it leverages ACT-R's declarative memory retrieval mechanisms to generate predictions (tag activations instead of link activations).
However, it is slightly different because it takes into account prior odds of tag activations (set to zero in SNIF-ACT).
This was set to zero in SNIF-ACT because each link has multiple words, so looking at word co-occurrences between the search query and link words produces a stable predictor.
It's also the case that search engines already rank order search results.
So even if prior odds for link activations were used in the SNIF-ACT model, the prior odds for the top few links would be virtually identical, so model predictions would not change much.
For the StackOverflow dataset, each tag is a single word, and those tags are not already rank ordered.
So ACT-R's prior odds component of the declarative memory retrieval equations was included for the StackOverflow model.

\subsection{ACT-R}

ACT-R's declarative memory retrieval equations are included below:

$A_{i} = B_{i} + \sum_{j\in C}^{ } W_{j} S_{ji}$

$B_{i} = log \frac{p_{i}}{1-p_{i}}$

$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{p(j,i)}{p(j)p(i)} = log \frac{NN(j,i)}{N_{Row}(j)N_{Col}(i)}$

$W_{j} = WE_{j}$

$E_{j} = 1-\frac{H_{j}}{H_{max}}$

$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$

$H_{max} = maxH$

Here, the $_{i}$ subscript denotes activation for a particular tag i, and the $_{j}$ subscript is for context.
$A_{i}$ is total activation for tag i.
That activation is a function of two components: Base-level activation $B_{i}$ and contextual activation $S_{ji}$.

\subsubsection{Base-Level Activation}

Base-level activation ($B_{i}$) is the log prior odds of observing tag i in the world.
So the tag with the highest frequency (C\#) would have the highest base-level activation.
The distribution of base level tag activations is included in Figure \ref{fig:tagActDis}.
Note that the majority of tags have very low base level activations, meaning that many of the tags are not frequently used.

In order to better visualize how often each tag is used, the tags' base-level activations were sorted and plotted in Figure \ref{fig:tagActSorted}.
The tag with the highest frequency of use (C\#) has the highest activation, and is the data point at the far left of the plot.
Note in particular how sharply tag activation decreases, meaning that only a small set of tags are the ones used almost entirely to tag posts.
This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.
Otherwise without context, the likelihood of choosing C\# as a tag would be the same as choosing ora-00997 (one of the many tags with only a single occurrence of use).
One might argue that with context, it's unlikely that the words associated with ora-00997 will occur again, so pure context would be able to still often predict C\#.
However, the ACT-R theory asserts that pure context is often not enough to overcome high base-level activations (e.g., it can be difficult to break a bad habit).

\subsubsection{Contextual Activation}

Contextual activation ($S_{ji}$) can be thought of as the amount of odds adjustment to base levels, given the current context $_{j}$.
For example, if the words in the title contained only 'PHP', 'PHP5', and 'PHPUnit', then the number of $_{j}$'s is three.
And for the 'PHP' tag, those three words are often in the title of PHP-tagged posts.
So the sum of the three $S_{ji}$'s for the tag PHP will be high, meaning that the odds adjustment for PHP given the current context will be high.
Since PHP already has a high base-level activation (frequently used tag), the model should produce a high total activation for the tag PHP, given the context.

The distribution of contextual activations is included in Figure \ref{fig:sjiActDis}.
Note the negative skew of activations, indicating that there are fewer word co-occurrences that occur very often than those that occur only once.
However the frequency of word co-occurrences isn't precisely what is used for the $S_{ji}$ calculation.
Note that the co-occurrences ($N(j,i)$) are scaled by total occurrences (N) as well as total occurrences for the context chunk j ($N_{Row}(j)$) and tag i ($N_{Col}(i)$).

It is also the case that there are many context, tag pairings that are never observed.
That is, the sparse matrix of $S_{ji}$'s is very sparse.
To visualize this, a scatter of the $S_{ji}$'s is included in Figure \ref{fig:sjiActScatter}, where a dot is included if the $S_{ji}$ at that location is non-zero.
Note that the visualization isn't perfect, since the thickness of the dot can influence how sparse the visualization looks.
But there are a few interesting things to note in this plot.
First, the context words that are tags often co-occur with that tag (line across main diagonal).
Second, note that the matrix is more dense closer to origin, and becomes less dense as it extends outwards.
This is because the hash index of each word was computed by starting from one, and taking the next number counting upwards if an observed word does not already have a hash.
So often-observed words tend to get lower hash indices than less-often-observed words.
Which means that words like 'the', '?', 'PHP', and 'C\#' have low indices, while words like 'herest', 'unmarsharller', and 'initwithtitle' have high indices.
So the likelihood that 'initwithtitle' context and 'herest' tag would be observed together is low. Most likely this pairing is a white spot in Figure \ref{fig:sjiActScatter}.

\subsubsection{Attentional Weighting}

Attentional weighting ($W_{j}$) can be thought of as the amount of attentional resources that are dedicated to contextual element $_{j}$.
Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that our attentional resources are limited.
For the StackOverflow model, the attentional weighting was left unbound, and allowed to be calibrated by the logistic regression model.
This was done for two main reasons.
First, the researcher was curious to see what value the logistic regression model would give for $W$.
Second, a non-ACT-R standard measure of attentional weighting was used ($E_{j}$), so it seemed appropriate to allow $W$ to vary alongside exploring this new weighting measure.

\subsection{Entropy}

During initial model testing, the attentional weighting for all words was equal.
That is initially, for a given post with $n$ title words, the attentional weighting $W_{j\in C}$ for all of those words was $W/n$, where the logistic regression model fit $W$ to range between 1-3.
The problem with this approach was that familiar stop words like 'the' and '?', that were observed so often and for all tags, had the same attentional weighting as highly predictive words like 'PHP'.
Granted the $S_{ji}$'s across all tags for context words like 'the' are much lower (and more evenly distributed) than the few $S_{ji}$'s that spike for 'PHP' (e.g., the tag PHP).
So the amount of adjusted log odds from stop words was low, even with equal attentional weighting (since their $S_{ji}$'s were low).
However, it still seemed problematic for two reasons.
First, intuitively it seemed that regardless of word co-occurrence association strengths, because stop words like 'the' are observed with all tags, they are not highly predictive of any tag.
So one could argue that their presence is simply adding error in the model's tag activation calculation.
Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words like 'the' and '?' are taken away from being used for highly predictive words like 'PHP'.
So this researcher decided to look further in the information retrieval literature to see if any prior researchers had solved this problem in an elegant and intuitive manner.

It turns out that there are many connections between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories.
For example, \cite{Dumais1991} shows that a measure of how likely term ${_j}$ is in document ${_i}$ is often the product of a local weighting $L_{ji}$ and global weighting $G_{j}$ measure.
This idea maps rather nicely to ACT-R's contextual weighting.
That is, terms are context words ${_j}$, documents are tags ${_i}$, local weighting is strength of association $S_{ji}$, and global weighting is attentional weighting $W_{j}$.

\cite{Dumais1991} states that predictions tend to improve when a global weighting scheme is used, so that words that occur frequently or with many documents (tags) are not weighted as high.
He reviews four different commonly-used weighting schemes; one of which is a scaled entropy measure.

The scaled entropy measure $E_{j}$ was chosen as a global weighting measure for this dataset because it has a strong theoretical foundation and it predicts well (e.g., \cite{Dumais1991}).
It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
Entropy measures that are mathematically equivalent to Shannon's formulation are used in data compression applications, cryptography, as well as thermodynamic systems.
The scaled entropy ($E_{j}$) of a context word can be thought of as the amount of predictiveness that that word has to any tag.

For example, a stop word like 'the' will have an even distribution of co-occurrences across all tags.
That is, seeing the word 'the' in context is not highly predictive of any particular tag.
This means that the marginal probabilities of 'the' with all tags will be roughly equal.
It can be shown that maximum entropy $H_{j}$ is reached when all marginal probabilities are equal.
This makes intuitive sense because if all marginal probabilities are equal, there is no information gained in observing the context word.
The scaled entropy measure $E_{j}$ transforms the entropy measure to a scale that ranges from 0, 1, where maximum entropy means lowest scaled entropy.
This means that a stop word like 'the' will have a scaled entropy measure close to 0, while a highly-predictive word like 'MySQL' will have a scaled entropy measure close to 1.
Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate the limited attentional resources to predictive contextual elements.

The distribution of attentional weights $E_{j}$'s is included in Figure \ref{fig:attentionalDis}. 
Note that scaled entropy measures for context words are highly skewed towards zero.
That is, there are many words in the title of a post that are stop words like 'the', that one should essentially ignore when computing tag activations.

Note also that there is an effect of attentional weight on frequency.
That is, more frequently encountered words in the title receive lower attentional weight.
This can be seen in Figure \ref{fig:attentionalN}.
Note however that although attentional weight decreases with number of observations, there is still a spread of attentional weighting at a given observation count.
This makes intuitive sense, as a two often-observed title words like 'the' and 'PHP' should have different attentional weightings, since the distribution of marginal probabilities for 'PHP' has more variance than 'the'.
That is, although PHP is often observed, because it is much-more-often observed for the PHP, PHP5, and PHPUnit tags, it is still highly predictive, and should still get a high level of attentional resources.
This type of result can be seen in the named data points in Figure \ref{fig:attentionalN}.

\subsection{PMI}

ACT-R's declarative memory retrieval equations are also seen in other information retrieval literature that use word co-occurrence counts, although it's usually called the Pointwise Mutual Information Index.
For example, \cite{Fu2007} and \cite{Farahat2004} have shown that ACT-R's retrieval equations are mathematically equivalent to the Pointwise Mutual Information (PMI) index when the sample size of the corpus is large.
\cite{Douglass2007} used the PMI-simplified version of ACT-R's retrieval equations with an $S_{ji}$ corpus of around 1 million.
Looking at the descriptives in Table \ref{tab:descript}, the $S_{ji}$ count for this dataset is on the order of 1 million as well.
So it was decided to use the PMI-simplified version of ACT-R's retrieval equations for the model here.

\cite{Budiu2007} and \cite{Farahat2004} have shown that the PMI index for word co-occurrences is a strong predictor of word similarity and information scent.
And on large corpora \cite{Budiu2007} has shown that PMI is the best predictor of word pair similarity (compared to LSA and generalized LSA).
It is interesting to note that ACT-R's declarative memory retrieval theory not only fits human performance data well, but is also a useful and competitive best predictor of word pair similarity in the information retrieval literature.
Often artificial intelligence approaches (LSA for example) stray from the most cognitively-plausible approach in order to achieve better performance for a particular domain.
However there are times where a cognitively-inspired strategy (PMI for example) may achieve better performance than a system that is free from cognitive constraints.

<<echo=false,results=tex>>=
stdprintxtable(xtable(descriptsFrm, caption="Descriptives for dataset", label="tab:descript"))
@

\subsection{Implementation}

\subsubsection{NLP}

Python's Natural Language Processing toolkit was chosen for word tokenizing and lemmatization because it seemed to be the most widely used tool for this type of natural language processing.
The tool wasn't tweaked from base settings or customized for the particular domain.

\subsubsection{Spreading Activation}

Since the model is based on ACT-R's declarative memory retrieval equations, it could have been implemented in Common Lisp (programming language for ACT-R's implementation).
However, it is known that the current implementation of spreading activation (required for $S_{ji}$) in ACT-R does not scale to an $S_{ji}$ corpus over 100,000 (\cite{Douglass1998}, \cite{Douglass2007}).
So it was decided to use a one-off implementation of ACT-R's retrieval equations for this analysis.
This is much like how \cite{Douglass2007} implemented the equations in Erlang when using the theory to test the fan effect on large-scale declarative memory datasets.

R was chosen as the implementation language for two main reasons.
First, it has sparse matrix support and vectorization routines, so that the model could scale to millions of $S_{jis}$'s.
Second, it has excellent data visualization and analysis capabilities. 
This way data visualization of the $S_{ji}$ sparse matrix could be done within the same R running process. 
The matrix could be accessed directly instead of having to code up a file I/O interface or socket interface to perform the visualization.

\section{Results}

The model was trained on 1 million posts from the dataset.
The descriptives are included in Table \ref{tab:descript}.

\subsection{Example Posts}

Three example posts were selected to qualitatively provide a better visualization of model behavior.
These examples are included in Figure FIXME.

Each figure has the title text of the post as the title of the figure.
The user-chosen tags are included below the title.
The model-chosen tags are on the x-axis, ranked by tag activation (y-axis).
Note also that the attentional weight (scaled entropy) for each word in the title is included below each word.

Note how the base level activation and contextual activation combine to predict tags that are both frequently used and contextually relevant.
For example, Figure \ref{fig:visPost1Sji} shows a high co-occurrence for twoway (not tagged) and databinding (tagged), but not for C\# (tagged) or wpf (tagged).
However, when spreading activation is added to base level activation, only the frequent and relevant tags have high activation (Figure \ref{fig:visPost1Act}).
That is, note that twoway (not tagged) decreases in ranking, while wpf, databinding, and C\# (all tagged) increase in ranking.
The plot isn't perfect though.
Note that XML (tagged) is further down in ranking than twoway (not tagged).

Model accuracy is higher for the example in Figures \ref{fig:visPost2Sji} and \ref{fig:visPost2Act}.
Note that the total activation for magic-quotes (tagged) is above 0, meaning that the model is predicting that it is very likely that the user will choose this tag for this post.

Finally, note the example in Figures \ref{fig:visPost3Sji} and \ref{fig:visPost3Act}, where the model is able to correctly predict a tag that is not mentioned explicitly in the text of the title.
C++ most likely has high word co-occurrences with const, char, and *, so the contextual activation for the C++ tag is relatively high (although not one of the top tags for contextual activation).
But since C++ is one of the most frequent tags, when that bit of contextual activation combines with base level activation, it becomes the highest ranked tag.

\subsection{Logistic Regression}

Two test runs using 1,000 posts were computed.
Model performance was assessed for classifying used tags based on tag activation (spreading and base-level components).
This was done using logistic regression.
The best-fit logistic regression coefficients for the intercept, spreading ($S_{ji}$) and base-level (prior) are included in Table \ref{tab:coeffs}.
These two runs are located in the 'uncompressed' section of the table.

Note that all three coefficients for each run are statistically significant ($p<.01$), and that the absolute value for each coefficient across runs varies, but only slightly.
This indicates that 1,000 aggregated posts for a test run is reasonably large to settle on stable coefficient values for the model.

It is interesting to note that although the attentional weighting term ($W$) was allowed to be calibrated by the regression, it did not stray far from one.
Apparently the scaled entropy measure that globally weights each cue allows for total attentional resources ($W$) to remain relatively small (around one) and not stray far from cognitively-plausible defaults (one).

Plots of model performance for the logistic regression for the two test runs are included in Figures \ref{fig:run1LogReg} and \ref{fig:run2LogReg}.
Not-tagged observed instances are collected on the bottom and tagged instances are on the top.
Looking at the histograms, note that model can not perfectly discriminate between tags that will and will not be used for a single post (overlap in distributions).
Nonetheless note that if model predicted tag activation is above 0, there is a high likelihood that this tag will be used for the post (about 90\% likely).
Refer back to Figure \ref{fig:visPost2Act} as an example of a case where the model's total activation for a tag is above 0.

A set of model performance statistics for the two runs are included in Table \ref{tab:fit}.
These two runs are on the run1 and run2 columns of that table.
The McFadden $R_{pseudo}^{2}$ statistic was chosen as the primary overall measure of fit, mostly because it makes intuitive sense and also because it can discriminate between later variations of the model.
Model fit for these runs was marginal (FIXME), (FIXME) for run2.

Looking at model classification shows why the fit is only marginal.
The model is able to correctly predict when a tag will not be used (CR=FIXME) for run1, CR= for run2).
However, if the model is to ever be fielded as a classifier, it must also be able to predict when a tag will be used.
And in this situation the model does only marginal (Hits=FIXME for run1, Hits= FIXME for run2).
That is, of all of the tags used in the 1,000 post tests, the model is only able to correctly predict those tags in about 10\% of cases.

\section{Discussion}

\subsection{Related Work}
This researcher recently discovered that predicting tags for the StackOverflow dataset has been worked on by at least one prior researcher (\cite{Kuo2011}).
\cite{Kuo2011} only looks at the Stack Overflow dataset briefly at the end of the technical report.
However he does mention that the co-occurrence model he used has about a 47\% classification accuracy for predicting StackOverflow tags, although it is unclear how exactly this accuracy is calculated.
I am considering sending this researcher an email, as I'm curious to know exactly how he measured classification accuracy for his co-occurrence model.

\subsection{ACT-R and Scale}

\subsection{Entropy and Stop Words}

\subsection{Future Work}

\subsubsection{User's Table}

\section{Acknowledgments}

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{bibliography}

\end{document}

### 2013-02-03 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 458f673fa7c97bcf63058a72ce00409123f8b32b [formerly 5bab703a26f33f56ee6cb96a16e027417727cb47]
Former-commit-id: 89ea856b5670c4f233e95fef398a336e316c8ad8

@@
 \usepackage[group-separator={,}]{siunitx}
+
+\graphicspath{{./pictures/}} % Specifies the directory where pictures are stored
 
@@
 Human-based tagging of online content has increased in recent years (\cite{Chang2010}).
-A large portion of the growing number of social media sites support some sort of human-directed tagging of posts
+A large portion of the growing number of social media sites support some sort of human-directed tagging of posts.
 Additionally, an increasing number of individuals are using tags as a form of query-based search to find information
@@
 
-Although much good work has been done to characterize the lifetime and growth characteristics of tags for real-time social networking sites such as Twitter
+Although much good work has been done to characterize the lifetime and growth characteristics of hashtags for real-time social networking sites such as Twitter
 (e.g., \cite{Bauer2012}, \cite{Tsur2012}, \cite{Chang2010})
@@
 
-ACT-R's declarative memory retrieval theory was used to model tag retrievals for StackOverflow posts.
-The hypothesis is that the tags used for posts are based on a tag's prior use and the content of words in the title and body of the post.
-The model was trained on 2/3 of the dataset (1M posts), and cross validated on 10 samples of 1,000 posts not contained within the training dataset.
+A variation of ACT-R's declarative memory retrieval theory was used to model tag retrievals for StackOverflow posts.
+The hypothesis is that the tags used for posts are based on a tag's prior use and the strength of association between the tag and the content of words in the title and body of the post.
+The strength of association between post words and tags was calculated using 2/3 of the dataset (1M posts).
 A logistic regression statistical technique was used to calibrate model parameters and optimize performance.
+The weights for each model component were calibrated using a sample of 1,000 posts not contained within the strength of association dataset.
+Finally, the calibrated model was tested on a new sample of 1,000 posts not contained within either previous training dataset.
 Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).
@@
 The StackOverflow dataset consists of all questions and answers posted on stackoverflow.com.
-The questions are all programming related and can be posted by anyone with a StackOverflow account.
-Each question is tagged by the poster with the tags that are most representative of the post.
+The questions are all programming related and can be posted by anyone with an account, which is free to create.
+Each question is tagged by the poster with (presumably) the tags that are most representative of the post.
 Example tags are programming languages (PHP, Common-Lisp, MySQL, C\#) and also general topics (databases, optimization, arrays).
-An Example question from the stackoverflow.com site is included in Figure FIXME.
-Note that Figure FIXME is tagged with 'PHP' and 'magic-quotes' for example.
+An Example question from the stackoverflow.com site is included in Figure \ref{fig:examplePost}.
+Note that the author tagged this post with java, twitter, twitter4j, and twitter−oauth.
+
+\begin{figure}[ht]
+  \centering
+  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act.pdf}}
+  \caption{Example Post}
+  \label{fig:examplePost}
+\end{figure}
 
 The Sep' 11 dataset (\cite{DataDump2011}) was downloaded and imported into the MySQL relational database system.
-The dataset contains FIXME posts from FIXME unique users.
+The dataset contains \num{4945168} posts from \num{559803} unique users.
 
@@
 Model tag predictions are based on the amount of word co-occurrences between words in the title and tags used for that post.
-To build the co-occurrence matrix between title words and tags, all title text and associated tags were extracted from the MySQL database.
-The Python Natural Language Processing Toolkit (\cite{Bird2009}) was used to chunk (i.e., tokenize and lemmetize) the title text.
+To build the co-occurrence matrix between post words and tags, all title and body text and associated tags were extracted from the relational database.
+The Python Natural Language Processing Toolkit (\cite{Bird2009}) was used to chunk (i.e., tokenize and lemmetize) the text in the post.
 The tags were already chunked so those words were not processed any further.
-Afterwords the processed words were fed back in to the MySQL database.
+Afterwords the processed word chunks were fed back in to the database.
 
-A MySQL query was then used to build the title-word, tag word co-occurrence matrix.
-For a given word found in any title, the query counts the number of times that word appears with each tag.
-Each of those counts is recorded as the word co-occurrence for that title word and each associated tag.
-This process is done across all words found in any title.
-The result is a comma-separated value (csv) file of title-word, tag, count for each title-word, tag combination.
+A query was then used to build the post word, tag word co-occurrence matrix.
+For a given word found in the body or title of a post, the query counts the number of times that that word appears with each tag.
+Each of those counts is recorded as the word co-occurrence for that word and each associated tag.
+This process is done across all words found in any post.
+The result is a comma-separated value (csv) file of post-word, tag, count for each combination.
 
@@
 More often used tags more likely to have higher activation.
-A MySQL query was used to build a similar csv for tag occurrences.
-This csv for example would show that C\# is the most often used tag, where FIXME posts were tagged with C\#.
+A SQL query was used to build a similar csv for tag occurrences.
+This csv for example would show that C\# is the most often used tag, where about 11\% of posts were tagged with C\#.
 
@@
 The tag with the highest activation is the most likely tag to be associated with that post.
-Tag activation is a function of three components:
+Tag activation is a function of four components:
 The tag's base level activation (prior odds of being used),
 contextual activation for words in the title (word co-occurrence weighting),
-and contextual activation for words in the body.
+contextual activation for words in the body,
+and an offset to equalize activation values for tags across all posts.
 
-Contextual activation is computed by comparing the words in the title with each tag's word co-occurrence vector.
+Contextual activation is computed by comparing the words in the post with each tag's word co-occurrence vector.
 If a tag's co-occurrences correlate strongly with the words in the title, then that tag's activation will be high.
@@
 
-1 million of the ~1.5 million posts were fed into the sparse matrix and vector to build the population co-occurrences and tag occurrences.
-An interface to the model was developed so that one could feed in a vector of words, and the model would return a sparse vector of tag activations given that context.
+1 million posts (approximately 2/3 of all posts) were fed into R to build the co-occurrences and tag occurrences.
+An interface to the model was developed so that one could supply a vector of words, and the model would return a sparse vector of tag activations given that context.
 In order to generate a model prediction, tag activations for title and body context were computed separately, and then added to each tag's prior base-level activation.
+An offset constant (equal to the average activation of the top 5 model-predicted tags for the post) was then subtracted from the prior value.
 The resulting tags were rank sorted, and the one with the highest activation would then be the model's prediction of the most likely tag associated with that post.
@@
 
-For a given post, the model produces a sparse vector of tag activations.
-For that post, the poster's chosen tags are known (see Figures \ref{fig:screenTreeview}, \ref{fig:screenMagic}, and \ref{fig:screenConst}).
+For a given post, the model produces a vector of tag activations.
+For that post, the poster's chosen tags are known (see Figure \ref{fig:examplePost}).
 So model performance can be measured by comparing the model's tag activations with chosen tags for a post, and then aggregated across posts to get an overall average performance.
@@
 
-Logistic regression was used in order to operationalize this fit measure.
+A logistic regression technique was used in order to operationalize this fit measure.
 For each post, the model's tag activations for the highest 100 tags were recorded, as well as the tag activations for all tags that were chosen by the poster.
@@
 
-These vectors of tag activation, categorization were collected across 10 sets of 1,000 posts.
-Each of those sets did not have any posts that were present in the 1 million posts used to train the model.
+These vectors of tag activation, categorization were collected across a set of 1,000 posts.
+The set of posts did not include any posts that were used to train the model on word association strength.
 So a form of cross validation was used to measure model fit, although the more robust full k-fold cross validation technique was not used (due mostly to resource and time limitations).
 
-R was used to perform the logistic regression.
-The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from two predictor variables: Tag occurrence activation and word co-occurrence activation.
+The base R glm function (with 'binomial' argument) was used to perform the logistic regression.
+The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from the four predictor variables:
+prior tag occurance, word co-occurrence activation for title and body words, and activation offset for the post.
 Further discussion and operationalization of the predictor variables is included in the Model section.
@@
 The tag-prediction model is a cognitively-inspired Bayesian probabilistic model, based on ACT-R's declarative memory retrieval mechanisms (\cite{Anderson2004}).
-These equations describe the likelihood that a chunk from memory will be retrieved, given the current context.
-In this way the task of retrieving the tag with the highest activation from the words in the title is analogous to performing a declarative memory retrieval request for a tag, given the context of title words.
+These equations describe the likelihood that a chunk from memory will be retrieved, given the prior odds of retrieving the chunk and the current context.
+In this way the task of retrieving the tag with the highest activation from the words in the title is similar to performing a declarative memory retrieval request for a tag, given the context of title and body words.
 
-This model is similar in design to the approach used by SNIF-ACT (\cite{Fu2007}, \cite{Pirolli2003}). 
+This model is also similar in design to the approach used by SNIF-ACT (\cite{Fu2007}, \cite{Pirolli2003}). 
 SNIF-ACT can predict link traversal for search queries.
@@
 
-\subsection{ACT-R}
+\subsection{Equations}
 
-ACT-R's declarative memory retrieval equations are included below:
+A formal description of the model's equations are included below:
 
-$A_{i} = B_{i} + \sum_{j\in C}^{ } W_{j} S_{ji}$
-
-$B_{i} = log \frac{p_{i}}{1-p_{i}}$
-
-$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{p(j,i)}{p(j)p(i)} = log \frac{NN(j,i)}{N_{Row}(j)N_{Col}(i)}$
-
-$W_{j} = WE_{j}$
-
-$E_{j} = 1-\frac{H_{j}}{H_{max}}$
-
-$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$
-
-$H_{max} = maxH$
+\begin{table}[!ht]
+  \begin{center} 
+    \caption{Sample table title.} 
+    \label{sample-table} 
+    \vskip 0.12in
+    \begin{tabular}{ll} 
+      \hline
+      Common Name &  Equation \\
+      \hline
+      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} + O$ \\
+      Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
+      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
+      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{p(j,i)}{p(j)p(i)}$ \\
+      & 			$= log \frac{NN(j,i)}{N_{Row}(j)N_{Col}(i)}$ \\
+      Attentional Weight	& $W_{j} = WE_{j}$ \\
+      Chunk Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
+      Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
+      Max Entropy & 		$H_{max} = maxH$ \\
+      \hline
+    \end{tabular} 
+  \end{center} 
+\end{table}
 
@@
 $A_{i}$ is total activation for tag i.
-That activation is a function of two components: Base-level activation $B_{i}$ and contextual activation $S_{ji}$.
+Note the similarities between the StackOverflow model and ACT-R's retrieval equations (base level and contextual component, log odds of prior being adjusted by association strength with context).
 
@@
 So the tag with the highest frequency (C\#) would have the highest base-level activation.
-The distribution of base level tag activations is included in Figure \ref{fig:tagActDis}.
-Note that the majority of tags have very low base level activations, meaning that many of the tags are not frequently used.
-
-In order to better visualize how often each tag is used, the tags' base-level activations were sorted and plotted in Figure \ref{fig:tagActSorted}.
-The tag with the highest frequency of use (C\#) has the highest activation, and is the data point at the far left of the plot.
-Note in particular how sharply tag activation decreases, meaning that only a small set of tags are the ones used almost entirely to tag posts.
+Tag frequency drops off sharply for this dataset.
+The top 1\% of tags based on frequency of use account for approximately 60\% of total tag instances.
 This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.
@@
 Since PHP already has a high base-level activation (frequently used tag), the model should produce a high total activation for the tag PHP, given the context.
-
-The distribution of contextual activations is included in Figure \ref{fig:sjiActDis}.
-Note the negative skew of activations, indicating that there are fewer word co-occurrences that occur very often than those that occur only once.
-However the frequency of word co-occurrences isn't precisely what is used for the $S_{ji}$ calculation.
-Note that the co-occurrences ($N(j,i)$) are scaled by total occurrences (N) as well as total occurrences for the context chunk j ($N_{Row}(j)$) and tag i ($N_{Col}(i)$).
-
-It is also the case that there are many context, tag pairings that are never observed.
-That is, the sparse matrix of $S_{ji}$'s is very sparse.
-To visualize this, a scatter of the $S_{ji}$'s is included in Figure \ref{fig:sjiActScatter}, where a dot is included if the $S_{ji}$ at that location is non-zero.
-Note that the visualization isn't perfect, since the thickness of the dot can influence how sparse the visualization looks.
-But there are a few interesting things to note in this plot.
-First, the context words that are tags often co-occur with that tag (line across main diagonal).
-Second, note that the matrix is more dense closer to origin, and becomes less dense as it extends outwards.
-This is because the hash index of each word was computed by starting from one, and taking the next number counting upwards if an observed word does not already have a hash.
-So often-observed words tend to get lower hash indices than less-often-observed words.
-Which means that words like 'the', '?', 'PHP', and 'C\#' have low indices, while words like 'herest', 'unmarsharller', and 'initwithtitle' have high indices.
-So the likelihood that 'initwithtitle' context and 'herest' tag would be observed together is low. Most likely this pairing is a white spot in Figure \ref{fig:sjiActScatter}.
 
@@
 Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words like 'the' and '?' are taken away from being used for highly predictive words like 'PHP'.
-So this researcher decided to look further in the information retrieval literature to see if any prior researchers had solved this problem in an elegant and intuitive manner.
 
-It turns out that there are many connections between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories.
+It turns out that there are many connections between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
+and many of those theories use a combination of weighting schemes for contextual relevance and global importance of a particular chunk in context.
 For example, \cite{Dumais1991} shows that a measure of how likely term ${_j}$ is in document ${_i}$ is often the product of a local weighting $L_{ji}$ and global weighting $G_{j}$ measure.
@@
 The scaled entropy measure $E_{j}$ transforms the entropy measure to a scale that ranges from 0, 1, where maximum entropy means lowest scaled entropy.
-This means that a stop word like 'the' will have a scaled entropy measure close to 0, while a highly-predictive word like 'MySQL' will have a scaled entropy measure close to 1.
+This means that a stop word like 'the' will have a scaled entropy measure close to 0, while a highly-predictive word like 'Macro' will have a scaled entropy measure close to 1.
 Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate the limited attentional resources to predictive contextual elements.
-
-The distribution of attentional weights $E_{j}$'s is included in Figure \ref{fig:attentionalDis}. 
-Note that scaled entropy measures for context words are highly skewed towards zero.
-That is, there are many words in the title of a post that are stop words like 'the', that one should essentially ignore when computing tag activations.
-
-Note also that there is an effect of attentional weight on frequency.
-That is, more frequently encountered words in the title receive lower attentional weight.
-This can be seen in Figure \ref{fig:attentionalN}.
-Note however that although attentional weight decreases with number of observations, there is still a spread of attentional weighting at a given observation count.
-This makes intuitive sense, as a two often-observed title words like 'the' and 'PHP' should have different attentional weightings, since the distribution of marginal probabilities for 'PHP' has more variance than 'the'.
-That is, although PHP is often observed, because it is much-more-often observed for the PHP, PHP5, and PHPUnit tags, it is still highly predictive, and should still get a high level of attentional resources.
-This type of result can be seen in the named data points in Figure \ref{fig:attentionalN}.
 
@@
 
-ACT-R's declarative memory retrieval equations are also seen in other information retrieval literature that use word co-occurrence counts, although it's usually called the Pointwise Mutual Information Index.
+ACT-R's declarative memory retrieval equations appear in other information retrieval literature that use word co-occurrence counts, although it's usually called the Pointwise Mutual Information Index.
 For example, \cite{Fu2007} and \cite{Farahat2004} have shown that ACT-R's retrieval equations are mathematically equivalent to the Pointwise Mutual Information (PMI) index when the sample size of the corpus is large.
-\cite{Douglass2007} used the PMI-simplified version of ACT-R's retrieval equations with an $S_{ji}$ corpus of around 1 million.
-Looking at the descriptives in Table \ref{tab:descript}, the $S_{ji}$ count for this dataset is on the order of 1 million as well.
-So it was decided to use the PMI-simplified version of ACT-R's retrieval equations for the model here.
-
 \cite{Budiu2007} and \cite{Farahat2004} have shown that the PMI index for word co-occurrences is a strong predictor of word similarity and information scent.
@@
 It is interesting to note that ACT-R's declarative memory retrieval theory not only fits human performance data well, but is also a useful and competitive best predictor of word pair similarity in the information retrieval literature.
-Often artificial intelligence approaches (LSA for example) stray from the most cognitively-plausible approach in order to achieve better performance for a particular domain.
-However there are times where a cognitively-inspired strategy (PMI for example) may achieve better performance than a system that is free from cognitive constraints.
-
-<<echo=false,results=tex>>=
-stdprintxtable(xtable(descriptsFrm, caption="Descriptives for dataset", label="tab:descript"))
-@
 
@@
 Second, it has excellent data visualization and analysis capabilities. 
-This way data visualization of the $S_{ji}$ sparse matrix could be done within the same R running process. 
-The matrix could be accessed directly instead of having to code up a file I/O interface or socket interface to perform the visualization.
+This way statistical data analysis and visualization could be done within the same R process that was running the model.
 
@@
 
-The model was trained on 1 million posts from the dataset.
-The descriptives are included in Table \ref{tab:descript}.
+After training the model's spreading activation matrix and calibrating model parameters, the model can predict roughly 70\% of tags correctly when asked to generate one tag per post on average.
 
@@
 
-Three example posts were selected to qualitatively provide a better visualization of model behavior.
-These examples are included in Figure FIXME.
+A visualization of the model's tag activation for the example post in Figure \ref{fig:examplePost} is included in Figure \ref{fig:modelPost}.
 
-Each figure has the title text of the post as the title of the figure.
-The user-chosen tags are included below the title.
-The model-chosen tags are on the x-axis, ranked by tag activation (y-axis).
-Note also that the attentional weight (scaled entropy) for each word in the title is included below each word.
+\begin{figure}[ht]
+  \centering
+  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act.pdf}}
+  \caption{
+    Model's prediction of top-ten tags for post in Figure \ref{fig:examplePost}.
+    The title is the title text of the post.
+    The attentional weight (entropy) for each word in the title is included below each word.
+    The user-chosen tags are included below the attentional weights.
+    The stacked bar graphs are the total activation for the top-ten predicted model tags for the post.
+    Each bar graph is partitioned into the 3 model components that vary between tags within a post (not the post offset).
+}
+  \label{fig:modelPost}
+\end{figure}
 
 Note how the base level activation and contextual activation combine to predict tags that are both frequently used and contextually relevant.
-For example, Figure \ref{fig:visPost1Sji} shows a high co-occurrence for twoway (not tagged) and databinding (tagged), but not for C\# (tagged) or wpf (tagged).
-However, when spreading activation is added to base level activation, only the frequent and relevant tags have high activation (Figure \ref{fig:visPost1Act}).
-That is, note that twoway (not tagged) decreases in ranking, while wpf, databinding, and C\# (all tagged) increase in ranking.
-The plot isn't perfect though.
-Note that XML (tagged) is further down in ranking than twoway (not tagged).
-
-Model accuracy is higher for the example in Figures \ref{fig:visPost2Sji} and \ref{fig:visPost2Act}.
-Note that the total activation for magic-quotes (tagged) is above 0, meaning that the model is predicting that it is very likely that the user will choose this tag for this post.
-
-Finally, note the example in Figures \ref{fig:visPost3Sji} and \ref{fig:visPost3Act}, where the model is able to correctly predict a tag that is not mentioned explicitly in the text of the title.
-C++ most likely has high word co-occurrences with const, char, and *, so the contextual activation for the C++ tag is relatively high (although not one of the top tags for contextual activation).
-But since C++ is one of the most frequent tags, when that bit of contextual activation combines with base level activation, it becomes the highest ranked tag.
+For example, note that the 'twitter' tag leverages a combination of contextual activation and prior activation to achieve a high total activation.
+Alternatively, the 'twitter-4j' tag relies almost entirely on context to achieve high activation.
+And note that the model is by no means perfect.
+For example, the 'java' tag was used by the post author, and it does have a high prior activation, but context in this particular post was not associated too highly with java, so the model rank of java suffered.
 
@@
 
-Two test runs using 1,000 posts were computed.
-Model performance was assessed for classifying used tags based on tag activation (spreading and base-level components).
-This was done using logistic regression.
-The best-fit logistic regression coefficients for the intercept, spreading ($S_{ji}$) and base-level (prior) are included in Table \ref{tab:coeffs}.
-These two runs are located in the 'uncompressed' section of the table.
+In order to formally test model performance, and to calibrate model parameters, a calibration run using \num{1000} posts were performed.
+For the calibration run, the category prediction (used/not-used tag for post) was regressed on base-level activation, spreading activation for title and body, and post offset.
+Best-fit coefficients for each model component and formal statistical measurements are included in Table \ref{tab:coeffs}.
 
-Note that all three coefficients for each run are statistically significant ($p<.01$), and that the absolute value for each coefficient across runs varies, but only slightly.
-This indicates that 1,000 aggregated posts for a test run is reasonably large to settle on stable coefficient values for the model.
-
-It is interesting to note that although the attentional weighting term ($W$) was allowed to be calibrated by the regression, it did not stray far from one.
+Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory ($R_{pseudo}^{2}=FIXME$).
+It's also interesting to note that although the attentional weighting term ($W$) was allowed to be calibrated by the regression, it did not stray far from one.
 Apparently the scaled entropy measure that globally weights each cue allows for total attentional resources ($W$) to remain relatively small (around one) and not stray far from cognitively-plausible defaults (one).
 
-Plots of model performance for the logistic regression for the two test runs are included in Figures \ref{fig:run1LogReg} and \ref{fig:run2LogReg}.
+A plot of model performance for the logistic regression for the calibration run is included in Figure \ref{fig:runCalLogReg}.
 Not-tagged observed instances are collected on the bottom and tagged instances are on the top.
 Looking at the histograms, note that model can not perfectly discriminate between tags that will and will not be used for a single post (overlap in distributions).
-Nonetheless note that if model predicted tag activation is above 0, there is a high likelihood that this tag will be used for the post (about 90\% likely).
-Refer back to Figure \ref{fig:visPost2Act} as an example of a case where the model's total activation for a tag is above 0.
+If further improvements to the model were found, these would increase the separation between the two histograms.
 
-A set of model performance statistics for the two runs are included in Table \ref{tab:fit}.
-These two runs are on the run1 and run2 columns of that table.
+\subsection{Classification Accuracy}
+
+In order to assess model performance in a dep
+
 The McFadden $R_{pseudo}^{2}$ statistic was chosen as the primary overall measure of fit, mostly because it makes intuitive sense and also because it can discriminate between later variations of the model.

### 2013-02-04 stackExchange/writeups/iccm-2013/writeup.tex
first draft
Former-commit-id: 75089ab0c359c2d5ae0ae1e5af50d9b01e7f93c1 [formerly adfe518c5d85c560ae374caca1db86439df12f6a]
Former-commit-id: 802a08fbf3be43d0fac87c4c806250dee6d92afa

@@
 \usepackage{apacite}
+
 \usepackage[group-separator={,}]{siunitx}
@@
 
-Human-based tagging of online content has increased in recent years (\cite{Chang2010}).
+Human-based tagging of online content has increased in recent years \cite{Chang2010}.
 A large portion of the growing number of social media sites support some sort of human-directed tagging of posts.
 Additionally, an increasing number of individuals are using tags as a form of query-based search to find information
-(e.g., \cite{Diakopoulos2010}).
+(e.g., \citeNP{Diakopoulos2010}).
 
 Although much good work has been done to characterize the lifetime and growth characteristics of hashtags for real-time social networking sites such as Twitter
-(e.g., \cite{Bauer2012}, \cite{Tsur2012}, \cite{Chang2010})
-less is known about a person's motivation for choosing particular tags during the creation of the tweet, post, etc.
+(e.g., \citeNP{Bauer2012}, \citeNP{Tsur2012}, \citeNP{Chang2010})
+less is known about a person's motivation for choosing particular tags during the creation of the tweet or post.
 Modeling tag creation accurately (as opposed to modeing tag growth and behavior after creation) could allow a system to predict the hashtag that a user should use, given the contextual clues available.
@@
 The StackOverflow dataset was chosen because it is relatively constrained, contains a large number of posts and associated tags, and the data are easy to obtain.
-It is constrained because there is a fully enumerated and limited set of tags that a poster can use, as opposed to a less-constrained domain such as Twitter.
+It is constrained because there is a fully enumerated and limited set of tags that an author can use, as opposed to a less-constrained domain such as Twitter.
 The site also makes public all posting data on a quarterly basis, which can be easily downloaded and loaded into a relational database for further analysis.
@@
 
-A variation of ACT-R's declarative memory retrieval theory was used to model tag retrievals for StackOverflow posts.
+A variation of ACT-R's declarative memory retrieval theory \cite{Anderson2004} was used to model tag retrievals for StackOverflow posts.
 The hypothesis is that the tags used for posts are based on a tag's prior use and the strength of association between the tag and the content of words in the title and body of the post.
@@
 The questions are all programming related and can be posted by anyone with an account, which is free to create.
-Each question is tagged by the poster with (presumably) the tags that are most representative of the post.
+Each question is tagged by the author with (presumably) the tags that are most representative of the post.
 Example tags are programming languages (PHP, Common-Lisp, MySQL, C\#) and also general topics (databases, optimization, arrays).
@@
 
-The Sep' 11 dataset (\cite{DataDump2011}) was downloaded and imported into the MySQL relational database system.
+The Sep' 11 dataset \cite{DataDump2011} was downloaded and imported into the MySQL relational database system.
 The dataset contains \num{4945168} posts from \num{559803} unique users.
@@
 To build the co-occurrence matrix between post words and tags, all title and body text and associated tags were extracted from the relational database.
-The Python Natural Language Processing Toolkit (\cite{Bird2009}) was used to chunk (i.e., tokenize and lemmetize) the text in the post.
-The tags were already chunked so those words were not processed any further.
-Afterwords the processed word chunks were fed back in to the database.
+The Python Natural Language Processing Toolkit \cite{Bird2009} was used to chunk (i.e., tokenize and lemmetize) the text in the post.
+The tags were already chunked.
+However, tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
+Afterwords the processed word chunks and root tags were fed back in to the database.
 
@@
 This process is done across all words found in any post.
-The result is a comma-separated value (csv) file of post-word, tag, count for each combination.
+The result is a comma-separated value (csv) file of context word, tag, count for each combination.
 
@@
 A hash was computed for each unique word, so that indices for the sparse matrix and vector could be numeric.
-The dimensions for the co-occurrence sparse matrix were context-word, tag, and the tag occurrence vector was 1 dimensional (tag).
+The dimensions for the co-occurrence sparse matrix were context word, tag, and the tag occurrence vector was 1 dimensional (tag).
 
@@
 For a given post, the model produces a vector of tag activations.
-For that post, the poster's chosen tags are known (see Figure \ref{fig:examplePost}).
+For that post, the author's chosen tags are known (see Figure \ref{fig:examplePost}).
 So model performance can be measured by comparing the model's tag activations with chosen tags for a post, and then aggregated across posts to get an overall average performance.
@@
 A logistic regression technique was used in order to operationalize this fit measure.
-For each post, the model's tag activations for the highest 100 tags were recorded, as well as the tag activations for all tags that were chosen by the poster.
+For each post, three sets of tag groups were combined to most accurately sample tag activations while managing computer resource constriants:
+The model's tag activations for the highest 200 tags, a weighted sample of 800 tags (weigted towards higher activatino), and the tag activations for all tags that were chosen by the author.
 Each of these observations was assigned a category of 0 or 1 depending on if the tag was a chosen tag or not for that post (0 is not chosen).
@@
 
-The tag-prediction model is a cognitively-inspired Bayesian probabilistic model, based on ACT-R's declarative memory retrieval mechanisms (\cite{Anderson2004}).
+The tag-prediction model is a cognitively-inspired Bayesian probabilistic model, based on ACT-R's declarative memory retrieval mechanisms \cite{Anderson2004}.
 These equations describe the likelihood that a chunk from memory will be retrieved, given the prior odds of retrieving the chunk and the current context.
@@
 
-This model is also similar in design to the approach used by SNIF-ACT (\cite{Fu2007}, \cite{Pirolli2003}). 
+This model is also similar in design to the approach used by SNIF-ACT (\citeNP{Fu2007}, \citeNP{Pirolli2003}). 
 SNIF-ACT can predict link traversal for search queries.
@@
       Attentional Weight	& $W_{j} = WE_{j}$ \\
-      Chunk Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
+      Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
       Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
@@
 
-\subsection{Entropy}
+\subsubsection{Entropy}
 
@@
 and many of those theories use a combination of weighting schemes for contextual relevance and global importance of a particular chunk in context.
-For example, \cite{Dumais1991} shows that a measure of how likely term ${_j}$ is in document ${_i}$ is often the product of a local weighting $L_{ji}$ and global weighting $G_{j}$ measure.
+For example, \citeA{Dumais1991} shows that a measure of how likely term ${_j}$ is in document ${_i}$ is often the product of a local weighting $L_{ji}$ and global weighting $G_{j}$ measure.
 This idea maps rather nicely to ACT-R's contextual weighting.
@@
 
-\cite{Dumais1991} states that predictions tend to improve when a global weighting scheme is used, so that words that occur frequently or with many documents (tags) are not weighted as high.
+\citeA{Dumais1991} states that predictions tend to improve when a global weighting scheme is used, so that words that occur frequently or with many documents (tags) are not weighted as high.
 He reviews four different commonly-used weighting schemes; one of which is a scaled entropy measure.
 
-The scaled entropy measure $E_{j}$ was chosen as a global weighting measure for this dataset because it has a strong theoretical foundation and it predicts well (e.g., \cite{Dumais1991}).
+The scaled entropy measure $E_{j}$ was chosen as a global weighting measure for this dataset because it has a strong theoretical foundation and it predicts well (\citeNP{Dumais1991}).
 It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
@@
 ACT-R's declarative memory retrieval equations appear in other information retrieval literature that use word co-occurrence counts, although it's usually called the Pointwise Mutual Information Index.
-For example, \cite{Fu2007} and \cite{Farahat2004} have shown that ACT-R's retrieval equations are mathematically equivalent to the Pointwise Mutual Information (PMI) index when the sample size of the corpus is large.
-\cite{Budiu2007} and \cite{Farahat2004} have shown that the PMI index for word co-occurrences is a strong predictor of word similarity and information scent.
-And on large corpora \cite{Budiu2007} has shown that PMI is the best predictor of word pair similarity (compared to LSA and generalized LSA).
-It is interesting to note that ACT-R's declarative memory retrieval theory not only fits human performance data well, but is also a useful and competitive best predictor of word pair similarity in the information retrieval literature.
-
-\subsection{Implementation}
-
-\subsubsection{NLP}
-
-Python's Natural Language Processing toolkit was chosen for word tokenizing and lemmatization because it seemed to be the most widely used tool for this type of natural language processing.
-The tool wasn't tweaked from base settings or customized for the particular domain.
-
-\subsubsection{Spreading Activation}
-
-Since the model is based on ACT-R's declarative memory retrieval equations, it could have been implemented in Common Lisp (programming language for ACT-R's implementation).
-However, it is known that the current implementation of spreading activation (required for $S_{ji}$) in ACT-R does not scale to an $S_{ji}$ corpus over 100,000 (\cite{Douglass1998}, \cite{Douglass2007}).
-So it was decided to use a one-off implementation of ACT-R's retrieval equations for this analysis.
-This is much like how \cite{Douglass2007} implemented the equations in Erlang when using the theory to test the fan effect on large-scale declarative memory datasets.
+For example, \citeA{Fu2007} and \citeA{Farahat2004} have shown that ACT-R's retrieval equations are mathematically equivalent to the Pointwise Mutual Information (PMI) index when the sample size of the corpus is large.
+\citeA{Budiu2007} and \citeA{Farahat2004} have shown that the PMI index for word co-occurrences is a strong predictor of word similarity and information scent.
+And on large corpora \citeA{Budiu2007} has shown that PMI is the best predictor of word pair similarity (compared to LSA and generalized LSA).
+It is interesting to note that ACT-R's declarative memory retrieval theory not only fits human performance data well, but is also a useful and competitive best predictor of word pair similarity in the general information retrieval literature that is less bounded by cognitive constraints.
 
@@
 
-After training the model's spreading activation matrix and calibrating model parameters, the model can predict roughly 70\% of tags correctly when asked to generate one tag per post on average.
+After training the model's spreading activation matrix and calibrating model parameters, the model can predict 70\% of tags correctly when asked to generate an average of one tag per post.
+
+\subsection{Logistic Regression}
+
+In order to formally test model performance, and to calibrate model parameters, a calibration run using \num{1000} posts were performed.
+For the calibration run, the category prediction (used/not-used tag for post) was regressed on base-level activation, spreading activation for title and body, and post offset.
+Best-fit coefficients for each model component and formal statistical measurements are included in Table \ref{tab:coeffs}.
+
+Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory ($R_{pseudo}^{2}=FIXME$).
+It's also interesting to note that although the attentional weighting term ($W$) was allowed to be calibrated by the regression, it did not stray far from one.
+Apparently the scaled entropy measure that globally weights each cue allows for total attentional resources ($W$) to remain relatively small (around one) and not stray far from cognitively-plausible defaults (one).
+
+A plot of model performance for the logistic regression for the calibration run is included in Figure \ref{fig:runCalLogReg}.
+Not-tagged observed instances are collected on the bottom and tagged instances are on the top.
+Looking at the histograms, note that model can not perfectly discriminate between tags that will and will not be used for a single post (overlap in distributions).
+If further improvements to the model were found, these would increase the separation between the two histograms.
+
+\begin{figure}[ht]
+  \centering
+  \resizebox{\linewidth}{!}{\includegraphics{foo.pdf}}
+  \caption{
+    Logistic regression of tag used on post (binary) regressed on four components of model tag activation.
+  }
+  \label{fig:logReg}
+\end{figure}
 
@@
 
-\subsection{Logistic Regression}
+\subsection{Generalizeability}
 
-In order to formally test model performance, and to calibrate model parameters, a calibration run using \num{1000} posts were performed.
-For the calibration run, the category prediction (used/not-used tag for post) was regressed on base-level activation, spreading activation for title and body, and post offset.
-Best-fit coefficients for each model component and formal statistical measurements are included in Table \ref{tab:coeffs}.
-
-Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory ($R_{pseudo}^{2}=FIXME$).
-It's also interesting to note that although the attentional weighting term ($W$) was allowed to be calibrated by the regression, it did not stray far from one.
-Apparently the scaled entropy measure that globally weights each cue allows for total attentional resources ($W$) to remain relatively small (around one) and not stray far from cognitively-plausible defaults (one).
-
-A plot of model performance for the logistic regression for the calibration run is included in Figure \ref{fig:runCalLogReg}.
-Not-tagged observed instances are collected on the bottom and tagged instances are on the top.
-Looking at the histograms, note that model can not perfectly discriminate between tags that will and will not be used for a single post (overlap in distributions).
-If further improvements to the model were found, these would increase the separation between the two histograms.
+In order to test how well the model (and calibrated parameters) generalize to the rest of the StackOverflow dataset, the model was tested on an additional set of \num{1000} posts.
+These were fresh posts, and not previously used to either train the strength of association matrix or calibrate model parameters.
+The model parameters from the calibrated dataset were held constant and used on the test dataset.
+Model classification accuracy as a function of number of tags predicted by the model was used as the metric to test if the model generalized.
 
@@
 
-In order to assess model performance in a dep
+A plot of model performance for both the calibration and test dataset is included in Figure \ref{fig:ROC}.
 
-The McFadden $R_{pseudo}^{2}$ statistic was chosen as the primary overall measure of fit, mostly because it makes intuitive sense and also because it can discriminate between later variations of the model.
-Model fit for these runs was marginal (FIXME), (FIXME) for run2.
+\begin{figure}[ht]
+  \centering
+  \resizebox{\linewidth}{!}{\includegraphics{ROC.pdf}}
+  \caption{
+    Model performance across the calibration and test dataset, as well as when specific model components are removed.
+    The vertical dotted line represents the point where the model is asked to tag on average one tag per post.
+  }
+  \label{fig:ROC}
+\end{figure}
 
-Looking at model classification shows why the fit is only marginal.
-The model is able to correctly predict when a tag will not be used (CR=FIXME) for run1, CR= for run2).
-However, if the model is to ever be fielded as a classifier, it must also be able to predict when a tag will be used.
-And in this situation the model does only marginal (Hits=FIXME for run1, Hits= FIXME for run2).
-That is, of all of the tags used in the 1,000 post tests, the model is only able to correctly predict those tags in about 10\% of cases.
+Note how model performance is almost identical for the calibration and test dataset (compare full model calibration to full model test).
+This indicates that the model parameters were not overfit, and generalize well to the rest of the dataset.
+It also provides evidence that \num{1000} posts is an adequate number to estimate model parameters accurately.
+
+\subsection{Necessity Analysis}
+
+Figure \ref{fig:ROC} also shows model performance after specific model components are removed.
+Model parameters were recalibrated for each model, except for the full model test dataset which used the parameters from the calibration dataset.
+These plots provide useful information regarding the relative necessity of each component in the full model.
+That is, which model components contribute most strongly to increased performance, and which model components might be redundant (not necessary).
+Note that it is certainly the case that each model component contributes positive and unique variance to model performance.
+However, removing only one component (particularly the words in the body of the post) does not drastically reduce model performance.
+It's only when groups of components are removed that model performance starts to drastically suffer.
 
@@
 \subsection{Related Work}
-This researcher recently discovered that predicting tags for the StackOverflow dataset has been worked on by at least one prior researcher (\cite{Kuo2011}).
-\cite{Kuo2011} only looks at the Stack Overflow dataset briefly at the end of the technical report.
-However he does mention that the co-occurrence model he used has about a 47\% classification accuracy for predicting StackOverflow tags, although it is unclear how exactly this accuracy is calculated.
-I am considering sending this researcher an email, as I'm curious to know exactly how he measured classification accuracy for his co-occurrence model.
+Predicting tags for the StackOverflow dataset has been worked on by at least one prior researcher \cite{Kuo2011}.
+\citeA{Kuo2011} only looks at the Stack Overflow dataset briefly at the end of the technical report.
+However he does mention that the co-occurrence model he used has about a 47\% classification accuracy for predicting StackOverflow tags, when required to predict one tag per post.
+The full model used here achieves 70\% classiciation accuraction when predicting one tag per post on average (height of full model at vertical dotted line in Figure \ref{fig:ROC}.
+The increase in model performance is most likely due to FIXME.
 
 \subsection{ACT-R and Scale}
+\num{39223968} unique context word, tag combinations were contained within the \num{1000000} posts that were used to build the strength of association matrix.
+Since the model is based on ACT-R's declarative memory retrieval equations, it could have been implemented in Common Lisp (programming language for ACT-R's implementation).
+However, it is known that the current implementation of spreading activation (required for $S_{ji}$) in ACT-R does not scale to an $S_{ji}$ corpus over 100,000 (\citeNP{Douglass2009}, \citeNP{Douglass2010}).
+So it was decided to use a one-off R implementation of ACT-R's retrieval equations for this analysis.
+This is much like how \citeA{Douglass2010} implemented the equations in Erlang when using the theory to test the fan effect on large-scale declarative memory datasets.
+
+However, it is increasingly likely that there is no fundamental reason why vanilla ACT-R (common-lisp version) cannot be optimized to handle spreading activation calculations with datasets this large.
+All that is really needed is a more appropriate data structure to handle the large scale.
+This can be seen by comparing the implementation of spreading activation here to the implementation in \citeA{Douglass2010}.
+Both implementations handle sji counts greater than \num{1000000}, with declarative memory retrieval times (time to calculate $A_i$s) around 1 second at this scale.
+However, their implementations could not be more different.
+The Erlang implementation leverages a semantic network of millions of isolated process chunks and massive concurrency to achieve the scale.
+The R implementation here is single threaded and requires only a sparse matrix implementatino to achive the scale.
+Since the programming languages are different, it's not likely the case that only a specific programming language can achive the scale.
+And since the parallelism is different (single threaded vs multi threaded), it's not likely the case that achieving the scale can only be done with massively concurrent implementations.
+So, it's most likely the case that using an appropriate data structure is all that is needed for ACT-R common-lisp to handle sji counts in the millions and above.
 
@@
 
+The Python NLTK documentation \cite{Bird2009} recommended to remove stop words from the analysis.
+Stop words are low-predictor words, and the documentation listed out a short list that were commonly removed (e.g., 'the', 'and', 'or')
+However it seemed somewhat arbitrary to explicitly list out a few stop words and then simply choose those as the words to remove.
+Looking at the scaled entropy measure, it seems quite plausible that this measure of word predictiveness can also be used to identify stop words.
+That is, words with scaled entropy close to zero are the stop words.
+Looking at the lowest ten scaled-entropy words for this dataset, (does, anyone, 've, found, are, seems, has, been, for, support), using scaled entropy to determine stop words seems to work fairly well.
+
 \subsection{Future Work}
 
-\subsubsection{User's Table}
+Classification accuracy for the model is already fairly high.
+However, there is certainly room for improvement, and unexplained variance left to account for.
+In particularl, the prior activatinos for each tag are not (currently) dependent on the particular user that is writing the post, and it seems plausible that prior activations for tags would change for each user.
+That is, if a user has repeatedly tagged questions with Common Lisp, and never with C\#, it seems unrealistic to use the general prior activations for this case, and give C\# the highest prior, and Common Lisp a relatively low prior.
+Instead, it may be worthwhile to maintain prior tag activations for each user, and combine this prior with the global prior when predicting tags for a particular post.
+And this approach could be fielded as well, since each user has to sign into their account before posting a questino on the site.
 
@@
 
+I would like to thank Dr. Mike Byrne and Dr. Fred Oswald for their excellent guidance and recommendations throughout the model development and data analysis process.
+
 \bibliographystyle{apacite}
-
 \setlength{\bibleftmargin}{.125in}
 \setlength{\bibindent}{-\bibleftmargin}
-
 \bibliography{bibliography}

### 2013-02-10 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 6026de11fa4b5c677a1252080980ce5fe850852a [formerly 890ccfc28319a6a92fb9af040c27e71992fdc0f9]
Former-commit-id: 15bb22acebf8dcb085d2ea3dc4b786dd9bead4a6

@@
 That is, terms are context words ${_j}$, documents are tags ${_i}$, local weighting is strength of association $S_{ji}$, and global weighting is attentional weighting $W_{j}$.
-
 \citeA{Dumais1991} states that predictions tend to improve when a global weighting scheme is used, so that words that occur frequently or with many documents (tags) are not weighted as high.
@@
 Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate the limited attentional resources to predictive contextual elements.
-
-\subsection{PMI}
-
-ACT-R's declarative memory retrieval equations appear in other information retrieval literature that use word co-occurrence counts, although it's usually called the Pointwise Mutual Information Index.
-For example, \citeA{Fu2007} and \citeA{Farahat2004} have shown that ACT-R's retrieval equations are mathematically equivalent to the Pointwise Mutual Information (PMI) index when the sample size of the corpus is large.
-\citeA{Budiu2007} and \citeA{Farahat2004} have shown that the PMI index for word co-occurrences is a strong predictor of word similarity and information scent.
-And on large corpora \citeA{Budiu2007} has shown that PMI is the best predictor of word pair similarity (compared to LSA and generalized LSA).
-It is interesting to note that ACT-R's declarative memory retrieval theory not only fits human performance data well, but is also a useful and competitive best predictor of word pair similarity in the general information retrieval literature that is less bounded by cognitive constraints.
 
@@
 
+\subsection{Entropy and ACT-R's Attentional Weighting}
+
+Model performance improved by using a scaled entropy measure to weight the importance of each contextual element in the body and title of a post.
+However, it's uncommon for an ACT-R model that uses the declarative memory retrieval mechanisms to use anything other than equal attentional weighting for each chunk in context.
+But it is much more common outside of the ACT-R modeling community to use both a local (strength of association) and global (attentional weight) term to predict document retrieval \citeA{Dumais1991}.
+In fact, it's been shown that the Pointwise Mutual Information Index (PMI) (a commonly-used measure to predict document retrieval) is mathematically equivalent to ACT-R's retrieval equations when the sample size of the corpus is large (\citeNP{Budiu2007}, \citeNP{Farahat2004}).
+Both global and local weightings are commonly used when the PMI measurement is used.
+This suggests that using a global weighting term for chunks in context (e.g., the scaled entropy measure used ehre) alongside the more commonly-used local weighting (spreading activation) might be useful for other ACT-R models that are taking advantage of large stores of declarative memory datesets to build up the model's base knowledge.
+
 \subsection{Future Work}

### 2013-02-10 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 14a45947fc03e3818491a9788bd8510b271c1d60 [formerly 4dd0d2a0b772728f982bffbbf758f7754ecb1f82]
Former-commit-id: 723742e46f278a8c71b1fe0c1e88b0c015f4401e

@@
   \centering
-  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act.pdf}}
+  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act-cropped.pdf}}
   \caption{Example Post}
@@
 This process is done across all words found in any post.
-The result is a comma-separated value (csv) file of context word, tag, count for each combination.
+The result is a comma-separated value (csv) file of context word, tag, count for each of the \num{39223968} unique context word, tag combinations.
 
@@
 More often used tags more likely to have higher activation.
-A SQL query was used to build a similar csv for tag occurrences.
+A SQL query was used to build a similar csv of tag use counts for the \num{26176} unique tags.
 This csv for example would show that C\# is the most often used tag, where about 11\% of posts were tagged with C\#.
@@
       \hline
-      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} + O$ \\
+      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} - O$ \\
       Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
-      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
       Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{p(j,i)}{p(j)p(i)}$ \\
@@
       Max Entropy & 		$H_{max} = maxH$ \\
+      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
       \hline
@@
 
-Contextual activation ($S_{ji}$) can be thought of as the amount of odds adjustment to base levels, given the current context $_{j}$.
-For example, if the words in the title contained only 'PHP', 'PHP5', and 'PHPUnit', then the number of $_{j}$'s is three.
-And for the 'PHP' tag, those three words are often in the title of PHP-tagged posts.
-So the sum of the three $S_{ji}$'s for the tag PHP will be high, meaning that the odds adjustment for PHP given the current context will be high.
-Since PHP already has a high base-level activation (frequently used tag), the model should produce a high total activation for the tag PHP, given the context.
+Contextual activation can be thought of as the amount of odds adjustment to base levels, given the current context $_{j}$.
+Contextual activation has two components: strength of association ($s_{sji}$) and attentional weight ($w_{j}$).
+Strength of association measures the relatedness of a contextual element $_{i}$ (e.g., the word PHP in the title of a post) with a tag $_{j}$ (e.g., the tag PHP5).
 
@@
 
-Attentional weighting ($W_{j}$) can be thought of as the amount of attentional resources that are dedicated to contextual element $_{j}$.
+The second component of contextual activation, attentional weighting ($W_{j}$), can be thought of as the amount of attentional resources that are dedicated to contextual element $_{j}$.
 Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that our attentional resources are limited.
@@
 
-R was chosen as the implementation language for two main reasons.
-First, it has sparse matrix support and vectorization routines, so that the model could scale to millions of $S_{jis}$'s.
-Second, it has excellent data visualization and analysis capabilities. 
-This way statistical data analysis and visualization could be done within the same R process that was running the model.
+\subsubsection{Offset}
+
+An offset measurement (mean of the top five tags for the post) was used to equalize the top activated tags for each post.
+This was done primarily to improve model fit.
+To see why this is the case, suppose the following scenario:
+The offset measurement is not used, and the model is asked to provide 10 tags across 10 posts.
+The top ten tags for one of these posts have activations that are higher than activations for all other tags in the other nine posts.
+The model would then select the ten tags from the single post as the most likely tags.
+But this doesn't make intuitive sense, because there is a limit to the number of tags that people tend to use for each post (average number of tags per post is ~ 3).
+So the offset measure aims to place all of these 10 posts on the same level, so that asking the model to choose 10 tags from 10 posts will select roughly a single tag from each post.
 
@@
   \centering
-  \resizebox{\linewidth}{!}{\includegraphics{foo.pdf}}
+  \resizebox{\linewidth}{!}{\includegraphics{foo-cropped.pdf}}
   \caption{
@@
   \centering
-  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act.pdf}}
+  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act-cropped.pdf}}
   \caption{
@@
   \centering
-  \resizebox{\linewidth}{!}{\includegraphics{ROC.pdf}}
+  \resizebox{\linewidth}{!}{\includegraphics{ROC-cropped.pdf}}
   \caption{

### 2013-02-10 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: bb4834d7223f2c68e83546109f5d4f36013baf4f [formerly 68cf953e7ac53a7c25f4fabacb204f2ed619bcc0]
Former-commit-id: 5abd299412cbedfe49171f6dc51234ece143b21a

@@
 An Example question from the stackoverflow.com site is included in Figure \ref{fig:examplePost}.
-Note that the author tagged this post with java, twitter, twitter4j, and twitter−oauth.
+Note that the author tagged this post with javascript, firefox, dom, and svg.
 
@@
   \centering
-  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act-cropped.pdf}}
+  \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-screen-cropped.pdf}}
   \caption{Example Post}
@@
 
-A formal description of the model's equations are included below:
+A formal description of the model's equations are included in Table \ref{sample-table}.
 
+\renewcommand{\arraystretch}{1.5}% Wider
+\renewcommand{\tabcolsep}{.5mm}
 \begin{table}[!ht]
@@
       Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
-      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{p(j,i)}{p(j)p(i)}$ \\
-      & 			$= log \frac{NN(j,i)}{N_{Row}(j)N_{Col}(i)}$ \\
-      Attentional Weight	& $W_{j} = WE_{j}$ \\
+      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN(j,i)}{N_{Row}(j)N_{Col}(i)}$ \\
+      Attention Weight	& 	$W_{j} = WE_{j}$ \\
       Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
@@
   \centering
-  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act-cropped.pdf}}
+  \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-act-cropped.pdf}}
   \caption{
@@
 Note how the base level activation and contextual activation combine to predict tags that are both frequently used and contextually relevant.
-For example, note that the 'twitter' tag leverages a combination of contextual activation and prior activation to achieve a high total activation.
-Alternatively, the 'twitter-4j' tag relies almost entirely on context to achieve high activation.
+For example, note that the 'javascript' tag leverages a combination of contextual activation and prior activation to achieve a high total activation.
+Alternatively, the 'svg' tag relies almost entirely on context to achieve high activation.
 And note that the model is by no means perfect.
-For example, the 'java' tag was used by the post author, and it does have a high prior activation, but context in this particular post was not associated too highly with java, so the model rank of java suffered.
+For example, note that 'html' is not tagged by the author, but the model ranks it as the second highest tag.
+And note that author's tag 'firefox' is only ranked 8th by the model.
 

### 2013-02-10 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 3bf8f6fc373a12392633b58dc8f6463d28622085 [formerly 69f76ce5bc5605a30e383dbd23320994e04e9323]
Former-commit-id: 0acd7d61b0e28cf53cbb33fb58b02d44a8089b4a

@@
 
-\title{FIXME}
+\title{Predicting Tags for StackOverflow Posts}
 
@@
 \begin{abstract}
-The abstract should be one paragraph, indented 1/8~inch on both sides,
-in 9~point font with single spacing. The heading {\bf Abstract} should
-be 10~point, bold, centered, with one line space below it. This
-one-paragraph abstract section is required only for standard spoken
-papers and standard posters (i.e., those presentations that will be
-represented by six page papers in the Proceedings).
+  A model was built for the StackOverflow dataset that can predict the tags that will be used by the author of the post, given the content of the post.
+  The tag-predictor model is a cognitively-inspired Bayesian probabilistic model based on ACT-R's declarative memory retrieval mechanisms.
+  Large-scale relational database techniques and efficient sparse matrix data structures were used to implement the model,
+  so that the data structures require only 3GB of memory and retrieval requests with a DM dataset of \num{556677795} chunks occur in less than one second on a single processor.
+  A logistic regression technique was used to calibrate model parameters and measure performance.
+  Model fit was satisfactory ($R_{pseudo}^{2}=.46$, npv = .997, ppv = .70) and model performance in a real-world setting was fairly strong (67\% accuracy when asked to tag one tag per post on average).
 
@@
 
-Human-based tagging of online content has increased in recent years \cite{Chang2010}.
+Human-based tagging of online content has increased in recent years.
 A large portion of the growing number of social media sites support some sort of human-directed tagging of posts.
@@
 less is known about a person's motivation for choosing particular tags during the creation of the tweet or post.
-Modeling tag creation accurately (as opposed to modeing tag growth and behavior after creation) could allow a system to predict the hashtag that a user should use, given the contextual clues available.
+Modeling tag creation accurately (as opposed to modeling tag growth and behavior after creation) could allow a system to predict the hashtag that a user should use, given the contextual clues available.
 The benefits of this type of prediction are numerous.
-For example, users could be introduced to tags (and potentially comunities surrounding those tags) that are of interest to them.
+For example, users could be introduced to tags (and potentially communities surrounding those tags) that are of interest to them.
 Also tag cleanup systems could be implemented where mistags are identified and corrected.
@@
 The weights for each model component were calibrated using a sample of 1,000 posts not contained within the strength of association dataset.
-Finally, the calibrated model was tested on a new sample of 1,000 posts not contained within either previous training dataset.
+The calibrated model was then tested on a new sample of 1,000 posts not contained within either previous training dataset.
 Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).
@@
 Example tags are programming languages (PHP, Common-Lisp, MySQL, C\#) and also general topics (databases, optimization, arrays).
-An Example question from the stackoverflow.com site is included in Figure \ref{fig:examplePost}.
+An example post is included in Figure \ref{fig:examplePost}.
 Note that the author tagged this post with javascript, firefox, dom, and svg.
@@
 
-The Sep' 11 dataset \cite{DataDump2011} was downloaded and imported into the MySQL relational database system.
+The Apr' 11 dataset \cite{DataDump2011} was downloaded and imported into the MySQL relational database system.
 The dataset contains \num{4945168} posts from \num{559803} unique users.
@@
 
-Model tag predictions are based on the amount of word co-occurrences between words in the title and tags used for that post.
+Model tag predictions are based on the amount of word co-occurrences between words in the title and body of the post and tags used for that post.
 To build the co-occurrence matrix between post words and tags, all title and body text and associated tags were extracted from the relational database.
-The Python Natural Language Processing Toolkit \cite{Bird2009} was used to chunk (i.e., tokenize and lemmetize) the text in the post.
+The Python Natural Language Processing Toolkit \cite{Bird2009} was used to chunk (i.e., tokenize and lemmatize) the text in the post.
 The tags were already chunked.
-However, tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
+However for the tags, tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
 Afterwords the processed word chunks and root tags were fed back in to the database.
@@
 This process is done across all words found in any post.
-The result is a comma-separated value (csv) file of context word, tag, count for each of the \num{39223968} unique context word, tag combinations.
+The result is a set of context word, tag, count for each of the \num{39223968} unique context word, tag combinations.
 
@@
 More often used tags more likely to have higher activation.
-A SQL query was used to build a similar csv of tag use counts for the \num{26176} unique tags.
-This csv for example would show that C\# is the most often used tag, where about 11\% of posts were tagged with C\#.
+A SQL query was used to build a similar set of tag use counts for the \num{26176} unique tags.
+This set for example would show that C\# is the most often used tag, where about 11\% of posts were tagged with C\#.
 
@@
 
-For a given title in a post, the model returns the activations for each possible tag.
+For a given title and body of a post, the model returns the activations for each possible tag.
 The tag with the highest activation is the most likely tag to be associated with that post.
@@
 Contextual activation is computed by comparing the words in the post with each tag's word co-occurrence vector.
-If a tag's co-occurrences correlate strongly with the words in the title, then that tag's activation will be high.
-If there is a low correlation between the observed co-occurrences for a tag and the observed words in the title, than that tag's activation will be low.
+If a tag's co-occurrences correlate strongly with the words in the post, then that tag's activation will be high.
+If there is a low correlation between the observed co-occurrences for a tag and the observed words in the post, than that tag's activation will be low.
 The equations that operationalize this process will be included in the Model section.
@@
 The R statistical programming environment was used to build the model and generate model predictions.
-This was done by feeding the word co-occurrence and tag occurrence csv's into a separate sparse matrix and sparse vector in R.
+This was done by feeding the word co-occurrence and tag occurrence files generated by the database queries into a separate sparse matrix and sparse vector in R.
 A hash was computed for each unique word, so that indices for the sparse matrix and vector could be numeric.
@@
 A logistic regression technique was used in order to operationalize this fit measure.
-For each post, three sets of tag groups were combined to most accurately sample tag activations while managing computer resource constriants:
-The model's tag activations for the highest 200 tags, a weighted sample of 800 tags (weigted towards higher activatino), and the tag activations for all tags that were chosen by the author.
+For each post, three sets of tag groups were combined to most accurately sample tag activations while managing computer resource constraints:
+The model's tag activations for the highest 200 tags, a weighted sample of 800 tags (weighted towards higher activation), and the tag activations for all tags that were chosen by the author of the post.
 Each of these observations was assigned a category of 0 or 1 depending on if the tag was a chosen tag or not for that post (0 is not chosen).
@@
 The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from the four predictor variables:
-prior tag occurance, word co-occurrence activation for title and body words, and activation offset for the post.
+prior tag occurrence, word co-occurrence activation for title and body words, and activation offset for the post.
 Further discussion and operationalization of the predictor variables is included in the Model section.
@@
 These equations describe the likelihood that a chunk from memory will be retrieved, given the prior odds of retrieving the chunk and the current context.
-In this way the task of retrieving the tag with the highest activation from the words in the title is similar to performing a declarative memory retrieval request for a tag, given the context of title and body words.
+In this way the task of retrieving the tag with the highest activation from the words in the post is similar to performing a declarative memory retrieval request for a tag, given the context of title and body words.
 
@@
   \begin{center} 
-    \caption{Sample table title.} 
+    \caption{The StackOverflow Tag-Prediction Model} 
     \label{sample-table} 
@@
 Contextual activation can be thought of as the amount of odds adjustment to base levels, given the current context $_{j}$.
-Contextual activation has two components: strength of association ($s_{sji}$) and attentional weight ($w_{j}$).
+Contextual activation has two components: strength of association ($S_{ji}$) and attentional weight ($W_{j}$).
 Strength of association measures the relatedness of a contextual element $_{i}$ (e.g., the word PHP in the title of a post) with a tag $_{j}$ (e.g., the tag PHP5).
@@
 
-It turns out that there are many connections between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
+It's the case that there are many connections and commonalities between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
 and many of those theories use a combination of weighting schemes for contextual relevance and global importance of a particular chunk in context.
@@
 It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
-Entropy measures that are mathematically equivalent to Shannon's formulation are used in data compression applications, cryptography, as well as thermodynamic systems.
+Entropy measures that are mathematically equivalent to Shannon's formulation are used in a wide range of applications, including data compression, cryptography, and thermodynamics.
 The scaled entropy ($E_{j}$) of a context word can be thought of as the amount of predictiveness that that word has to any tag.
@@
 The scaled entropy measure $E_{j}$ transforms the entropy measure to a scale that ranges from 0, 1, where maximum entropy means lowest scaled entropy.
-This means that a stop word like 'the' will have a scaled entropy measure close to 0, while a highly-predictive word like 'Macro' will have a scaled entropy measure close to 1.
+This means that a stop word like 'the' will have a scaled entropy measure close to 0, while a highly-predictive word like 'macro' will have a scaled entropy measure close to 1.
 Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate the limited attentional resources to predictive contextual elements.
@@
 To see why this is the case, suppose the following scenario:
-The offset measurement is not used, and the model is asked to provide 10 tags across 10 posts.
+The offset measurement is not used, and the model is asked to provide ten tags across ten posts.
 The top ten tags for one of these posts have activations that are higher than activations for all other tags in the other nine posts.
 The model would then select the ten tags from the single post as the most likely tags.
-But this doesn't make intuitive sense, because there is a limit to the number of tags that people tend to use for each post (average number of tags per post is ~ 3).
-So the offset measure aims to place all of these 10 posts on the same level, so that asking the model to choose 10 tags from 10 posts will select roughly a single tag from each post.
+But this doesn't make intuitive sense, because there is a limit to the number of tags that people tend to use for each post (average number of tags per post is approximately 3).
+So the offset measure aims to place all of these ten posts on the same level, so that asking the model to choose ten tags from ten posts will select roughly a single tag from each post.
 
@@
 
-After training the model's spreading activation matrix and calibrating model parameters, the model can predict 70\% of tags correctly when asked to generate an average of one tag per post.
+After training the model's spreading activation matrix and calibrating model parameters, the model can predict 67\% of tags correctly when asked to generate an average of one tag per post.
 
@@
 
-In order to formally test model performance, and to calibrate model parameters, a calibration run using \num{1000} posts were performed.
-For the calibration run, the category prediction (used/not-used tag for post) was regressed on base-level activation, spreading activation for title and body, and post offset.
-Best-fit coefficients for each model component and formal statistical measurements are included in Table \ref{tab:coeffs}.
+In order to formally test model performance and to calibrate model parameters, a calibration run using \num{1000} posts was performed.
+For the calibration run, the category prediction (chosen or not chosen tag for a post) was regressed on base-level activation, spreading activation for title and body, and post offset.
+Best-fit coefficients for each model component and formal statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits} respectively.
 
-Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory ($R_{pseudo}^{2}=FIXME$).
-It's also interesting to note that although the attentional weighting term ($W$) was allowed to be calibrated by the regression, it did not stray far from one.
+\renewcommand{\arraystretch}{1}% Wider
+\renewcommand{\tabcolsep}{1mm}
+\begin{table}[!ht]
+  \begin{center} 
+    \caption{Best-fit coefficient estimates from logistic regression} 
+    \label{tab:coeffs} 
+    \vskip 0.12in
+    \begin{tabular}{lllll} 
+      \hline
+      Coefficient & 	Estimate &	Std. Err. &	z value &	$Pr(>|z|)$  \\
+      \hline
+      (Intercept) &	-1.35 &		0.10 &		-13.59 &	\textless2e-16 *** \\
+      prior &		1.08 & 		0.01 &		88.12 & 	\textless2e-16 *** \\
+      sjiTitle &	1.41 &		0.02 &		57.36 &		\textless2e-16 *** \\
+      sjiBody &		2.36 &		0.04 &		58.48 &		\textless2e-16 *** \\
+      offset &		-0.71 &		0.02 &		-39.05 &	\textless2e-16 *** \\
+      \hline
+    \end{tabular} 
+  \end{center} 
+\end{table}
+
+\renewcommand{\arraystretch}{1}% Wider
+\renewcommand{\tabcolsep}{3mm}
+\begin{table}[!ht]
+  \begin{center} 
+    \caption{Model fit metrics} 
+    \label{tab:fits} 
+    \vskip 0.12in
+    \begin{tabular}{ll} 
+      \hline
+      Common Name &  Value	\\
+      \hline
+      McFadden's $R_{pseudo}^{2}$ &	.46 \\
+      True Positives &			647 \\
+      False Positives &			278 \\
+      True Negatives &			\num{796982} \\
+      False Negatives &			\num{2235} \\
+      Positive Predictive Value &	0.70 \\
+      Negative Predictive Value &	0.997 \\
+      \hline
+    \end{tabular} 
+  \end{center} 
+\end{table}
+
+Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory ($R_{pseudo}^{2}=.46$).
+It's also interesting to note that although the attentional weighting term ($W$) was allowed to be calibrated by the regression, it did not stray far from one
+(see coefficients for title and body in Table \ref{tab:coeffs}).
 Apparently the scaled entropy measure that globally weights each cue allows for total attentional resources ($W$) to remain relatively small (around one) and not stray far from cognitively-plausible defaults (one).
 
-A plot of model performance for the logistic regression for the calibration run is included in Figure \ref{fig:runCalLogReg}.
-Not-tagged observed instances are collected on the bottom and tagged instances are on the top.
+A plot of model performance for the logistic regression for the calibration run is included in Figure \ref{fig:logReg}.
+Tag-not-chosen observed instances are collected on the bottom and tag-chosen instances are on the top.
 Looking at the histograms, note that model can not perfectly discriminate between tags that will and will not be used for a single post (overlap in distributions).
-If further improvements to the model were found, these would increase the separation between the two histograms.
+Once further improvements to the model are found, these will increase the separation between the two histograms.
 
@@
 And note that the model is by no means perfect.
-For example, note that 'html' is not tagged by the author, but the model ranks it as the second highest tag.
-And note that author's tag 'firefox' is only ranked 8th by the model.
+For example, 'html' is not tagged by the author, but the model ranks it as the second highest tag.
+And the author's tag 'firefox' is only ranked 8th by the model.
 
-\subsection{Generalizeability}
+\subsection{Generalizability}
 
@@
 However he does mention that the co-occurrence model he used has about a 47\% classification accuracy for predicting StackOverflow tags, when required to predict one tag per post.
-The full model used here achieves 70\% classiciation accuraction when predicting one tag per post on average (height of full model at vertical dotted line in Figure \ref{fig:ROC}.
+The full model used here achieves 67\% classification accuracy when predicting one tag per post on average (height of full model at vertical dotted line in Figure \ref{fig:ROC}.
 The increase in model performance is most likely due to FIXME.
@@
 
-However, it is increasingly likely that there is no fundamental reason why vanilla ACT-R (common-lisp version) cannot be optimized to handle spreading activation calculations with datasets this large.
+However, it is increasingly likely that there is no fundamental reason why vanilla ACT-R (Common Lisp version) cannot be optimized to handle spreading activation calculations with datasets this large.
 All that is really needed is a more appropriate data structure to handle the large scale.
 This can be seen by comparing the implementation of spreading activation here to the implementation in \citeA{Douglass2010}.
-Both implementations handle sji counts greater than \num{1000000}, with declarative memory retrieval times (time to calculate $A_i$s) around 1 second at this scale.
-However, their implementations could not be more different.
+Both implementations handle sji counts greater than \num{1000000}, with declarative memory retrieval times (time to calculate $A_i$'s) around 1 second at this scale.
+However, the implementations could not be more different.
 The Erlang implementation leverages a semantic network of millions of isolated process chunks and massive concurrency to achieve the scale.
-The R implementation here is single threaded and requires only a sparse matrix implementatino to achive the scale.
-Since the programming languages are different, it's not likely the case that only a specific programming language can achive the scale.
+The R implementation here is single threaded and requires only a sparse matrix implementation to achieve the scale.
+Since the programming languages are different, it's not likely the case that only a specific programming language can achieve the scale.
 And since the parallelism is different (single threaded vs multi threaded), it's not likely the case that achieving the scale can only be done with massively concurrent implementations.
-So, it's most likely the case that using an appropriate data structure is all that is needed for ACT-R common-lisp to handle sji counts in the millions and above.
+So, it's most likely the case that using an appropriate data structure is all that is needed for ACT-R Common Lisp to handle sji counts in the millions and above.
 
@@
 Both global and local weightings are commonly used when the PMI measurement is used.
-This suggests that using a global weighting term for chunks in context (e.g., the scaled entropy measure used ehre) alongside the more commonly-used local weighting (spreading activation) might be useful for other ACT-R models that are taking advantage of large stores of declarative memory datesets to build up the model's base knowledge.
+This suggests that using a global weighting term for chunks in context (e.g., the scaled entropy measure used here) alongside the more commonly-used local weighting (spreading activation) might be useful for other ACT-R models that are taking advantage of large stores of declarative memory datasets to build up the model's base knowledge.
 
@@
 Classification accuracy for the model is already fairly high.
-However, there is certainly room for improvement, and unexplained variance left to account for.
-In particularl, the prior activatinos for each tag are not (currently) dependent on the particular user that is writing the post, and it seems plausible that prior activations for tags would change for each user.
+However, there is certainly room for improvement and unexplained variance left to account for.
+In particular, the prior activations for each tag are not (currently) dependent on the particular user that is writing the post, and it seems plausible that prior activations for tags would change for each user.
 That is, if a user has repeatedly tagged questions with Common Lisp, and never with C\#, it seems unrealistic to use the general prior activations for this case, and give C\# the highest prior, and Common Lisp a relatively low prior.
 Instead, it may be worthwhile to maintain prior tag activations for each user, and combine this prior with the global prior when predicting tags for a particular post.
-And this approach could be fielded as well, since each user has to sign into their account before posting a questino on the site.
+And this approach could be fielded as well, since each user has to sign into their account before posting a question on the site.
 

### 2013-02-10 stackExchange/analysis/model/analyzer.R

@@ -198,6 +198,12 @@ plotROC2 = function(items) {
 	dev.off()
 }
 
+getAccuracyAtNAverageTags = function(frm, N=1) {
+	propTagged = with(frm, length(unique(tagFile))/sum(targetP) * N)
+	res = plotROCColumn(frm, "act")
+	res$y[max(which(res$x<propTagged))]
+}
+
 plotHighest = function(contextIndeces, vals) {
 	vals = lapply(vals, function(val) {val[contextIndeces]})
 	frm = data.frame(vals)

### 2013-02-10 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 9b20364a4ea81ec508218b529fe3b4654342e0b4 [formerly 8b23a2e0941e6297285b818c2b5e4dca03ced9b4]
Former-commit-id: 549049bea4a6e08c598cf7221559a1917bcc77c6

@@
 The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from the four predictor variables:
-prior tag occurrence, word co-occurrence activation for title and body words, and activation offset for the post.
+The tag's prior occurrence frequency, word co-occurrence activation for title and body words, and activation offset for the post.
 Further discussion and operationalization of the predictor variables is included in the Model section.
@@
 
-Here, the $_{i}$ subscript denotes activation for a particular tag i, and the $_{j}$ subscript is for context.
-$A_{i}$ is total activation for tag i.
+Here, the $i$ subscript denotes activation for a particular tag $i$, and the $j$ subscript is for context.
+$A_{i}$ is total activation for tag $i$.
 Note the similarities between the StackOverflow model and ACT-R's retrieval equations (base level and contextual component, log odds of prior being adjusted by association strength with context).
@@
 
-Base-level activation ($B_{i}$) is the log prior odds of observing tag i in the world.
+Base-level activation ($B_{i}$) is the log prior odds of observing tag $i$ in the world.
 So the tag with the highest frequency (C\#) would have the highest base-level activation.
@@
 
-Contextual activation can be thought of as the amount of odds adjustment to base levels, given the current context $_{j}$.
+Contextual activation can be thought of as the amount of odds adjustment to base levels, given the current context $j$.
 Contextual activation has two components: strength of association ($S_{ji}$) and attentional weight ($W_{j}$).
-Strength of association measures the relatedness of a contextual element $_{i}$ (e.g., the word PHP in the title of a post) with a tag $_{j}$ (e.g., the tag PHP5).
+Strength of association measures the relatedness of a contextual element $i$ (e.g., the word PHP in the title of a post) with a tag $j$ (e.g., the tag PHP5).
 
@@
 
-The second component of contextual activation, attentional weighting ($W_{j}$), can be thought of as the amount of attentional resources that are dedicated to contextual element $_{j}$.
+The second component of contextual activation, attentional weighting ($W_{j}$), can be thought of as the amount of attentional resources that are dedicated to contextual element $j$.
 Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that our attentional resources are limited.

### 2013-02-11 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 7421747f5b4edccd4eebe8d1e167c90e0f1c7949 [formerly 59e9d5403ceb470d1edb138840c8900f5a887f9b]
Former-commit-id: f6dc590e4998d77b7d77c3b40e6e3038e9e7e9b7

@@
   A model was built for the StackOverflow dataset that can predict the tags that will be used by the author of the post, given the content of the post.
-  The tag-predictor model is a cognitively-inspired Bayesian probabilistic model based on ACT-R's declarative memory retrieval mechanisms.
-  Large-scale relational database techniques and efficient sparse matrix data structures were used to implement the model,
+  The model is a cognitively-inspired Bayesian probabilistic model based on ACT-R's declarative memory retrieval mechanisms.
+  Large-scale relational database techniques and efficient sparse matrix data structures were used for implementation,
   so that the data structures require only 3GB of memory and retrieval requests with a DM dataset of \num{556677795} chunks occur in less than one second on a single processor.
   A logistic regression technique was used to calibrate model parameters and measure performance.
-  Model fit was satisfactory ($R_{pseudo}^{2}=.46$, npv = .997, ppv = .70) and model performance in a real-world setting was fairly strong (67\% accuracy when asked to tag one tag per post on average).
+  Model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.46$, npv = .997, ppv = .70) and model performance on the test dataset is fairly strong (67\% accuracy when asked to tag one tag per post on average).
 
@@
 It is constrained because there is a fully enumerated and limited set of tags that an author can use, as opposed to a less-constrained domain such as Twitter.
-The site also makes public all posting data on a quarterly basis, which can be easily downloaded and loaded into a relational database for further analysis.
+The site also makes public all posting data on a quarterly basis, which can be downloaded and loaded into a local relational database for further analysis.
 
@@
 The hypothesis is that the tags used for posts are based on a tag's prior use and the strength of association between the tag and the content of words in the title and body of the post.
-The strength of association between post words and tags was calculated using 2/3 of the dataset (1M posts).
+The strength of association between post words and tags was calculated using 2/3 of the dataset (one million posts).
 A logistic regression statistical technique was used to calibrate model parameters and optimize performance.
-The weights for each model component were calibrated using a sample of 1,000 posts not contained within the strength of association dataset.
-The calibrated model was then tested on a new sample of 1,000 posts not contained within either previous training dataset.
+The weights for each model component were calibrated using a sample of \num{1000} posts not contained within the strength of association dataset.
+The calibrated model was then tested on a new sample of \num{1000} posts not contained within either previous training dataset.
 Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).
@@
 A hash was computed for each unique word, so that indices for the sparse matrix and vector could be numeric.
-The dimensions for the co-occurrence sparse matrix were context word, tag, and the tag occurrence vector was 1 dimensional (tag).
+The dimensions for the co-occurrence sparse matrix were context word, tag, and the tag occurrence vector was one dimensional (tag).
 
-1 million posts (approximately 2/3 of all posts) were fed into R to build the co-occurrences and tag occurrences.
+One million posts (approximately 2/3 of all posts) were fed into R to build the co-occurrences and tag occurrences.
 An interface to the model was developed so that one could supply a vector of words, and the model would return a sparse vector of tag activations given that context.
@@
 
-These vectors of tag activation, categorization were collected across a set of 1,000 posts.
+These vectors of tag activation, categorization were collected across a set of \num{1000} posts.
 The set of posts did not include any posts that were used to train the model on word association strength.
@@
   \begin{center} 
-    \caption{The StackOverflow Tag-Prediction Model} 
+    \caption{The StackOverflow tag-prediction model} 
     \label{sample-table} 
@@
       \hline
-      McFadden's $R_{pseudo}^{2}$ &	.46 \\
+      McFadden's Pseudo $R_{}^{2}$ &	.46 \\
       True Positives &			647 \\
@@
 
-Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory ($R_{pseudo}^{2}=.46$).
+Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.46$).
 It's also interesting to note that although the attentional weighting term ($W$) was allowed to be calibrated by the regression, it did not stray far from one
@@
 \citeA{Kuo2011} only looks at the Stack Overflow dataset briefly at the end of the technical report.
-However he does mention that the co-occurrence model he used has about a 47\% classification accuracy for predicting StackOverflow tags, when required to predict one tag per post.
-The full model used here achieves 67\% classification accuracy when predicting one tag per post on average (height of full model at vertical dotted line in Figure \ref{fig:ROC}.
-The increase in model performance is most likely due to FIXME.
+However he does mention that the co-occurrence model he used has about a 47\% classification accuracy for predicting StackOverflow tags when required to predict one tag per post.
+The full model used here achieves 67\% classification accuracy when predicting one tag per post on average
+(height of full model at vertical dotted line in Figure \ref{fig:ROC}).
+The increase in model performance is most likely due to the fact that \citeA{Kuo2011} only used \num{10000} posts to generate the co-occurance matrix ($S_{ji}$).
+One million posts (two orders of magnitude higher) was used here.
+Also the model here has some additional components: The offset in particular and possibly the body term as well,
+as it was unclear from \citeA{Kuo2011} if the author used context from both the body and title of a post or just the title. 
+Nonetheless, there are certainly similarities in the approach:
+Co-occurance count was used as the core measure of similarity, and both local (strength of association) and global (idf or entropy) measures were used to compute tag likelihood, given context.
 
@@
 This can be seen by comparing the implementation of spreading activation here to the implementation in \citeA{Douglass2010}.
-Both implementations handle sji counts greater than \num{1000000}, with declarative memory retrieval times (time to calculate $A_i$'s) around 1 second at this scale.
+Both implementations handle sji counts greater than \num{1000000}, with declarative memory retrieval times (time to calculate $A_i$'s) around one second at this scale.
 However, the implementations could not be more different.

### 2013-02-11 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: d82f8f9fa1cb3a0581a678ad989b764b19aafad4 [formerly ff032473ca3222c7c4bdc540623b737201b95309]
Former-commit-id: 05c9ffbce4b58bfe2edcbc6c295ae7c37ee51b08

@@
 For the calibration run, the category prediction (chosen or not chosen tag for a post) was regressed on base-level activation, spreading activation for title and body, and post offset.
-Best-fit coefficients for each model component and formal statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits} respectively.
+However using a single run of the \num{1000} posts was insufficient because there are two model components that require prior estimates of coefficients:
+The offset measure to determine the top-five tags and associated activation and the sampling technique used to pick a representative set of tags for each post.
+So model calibration was performed four times, where the coefficients from the previous run were used as estimates of coefficients for the next run.
+The coefficients converged on best-fit values quickly (running just two times would have sufficed), and did not show any sign of instability with the iterative technique.
+The final best-fit coefficients for each model component and formal statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits} respectively.
 

### 2013-02-11 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 64de62291a3aea7507c0f9986b96480f92c2f134 [formerly 22bfc5925c170b77149339034d87cdeabbdff118]
Former-commit-id: 00c7a481d23f86e42b23f8d11a20160db361385a

@@
 (height of full model at vertical dotted line in Figure \ref{fig:ROC}).
-The increase in model performance is most likely due to the fact that \citeA{Kuo2011} only used \num{10000} posts to generate the co-occurance matrix ($S_{ji}$).
+The increase in model performance is most likely due to the fact that \citeA{Kuo2011} only used \num{10000} posts to generate the co-occurrence matrix ($S_{ji}$).
 One million posts (two orders of magnitude higher) was used here.
@@
 Nonetheless, there are certainly similarities in the approach:
-Co-occurance count was used as the core measure of similarity, and both local (strength of association) and global (idf or entropy) measures were used to compute tag likelihood, given context.
+Co-occurrence count was used as the core measure of similarity, and both local (strength of association) and global (idf or entropy) measures were used to compute tag likelihood, given context.
 

### 2013-02-11 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: b08c74ff49c3ec4d88a183dee8e3691769160501 [formerly f67e212d40daadf85fd41f049207736da4c556bf]
Former-commit-id: 8bd9bc1b9941f44f28a52296975c8bf65aae02d9

@@
 This StackOverflow model is similar because it leverages ACT-R's declarative memory retrieval mechanisms to generate predictions (tag activations instead of link activations).
-However, it is slightly different because it takes into account prior odds of tag activations (set to zero in SNIF-ACT).
-This was set to zero in SNIF-ACT because each link has multiple words, so looking at word co-occurrences between the search query and link words produces a stable predictor.
-It's also the case that search engines already rank order search results.
-So even if prior odds for link activations were used in the SNIF-ACT model, the prior odds for the top few links would be virtually identical, so model predictions would not change much.
-For the StackOverflow dataset, each tag is a single word, and those tags are not already rank ordered.
-So ACT-R's prior odds component of the declarative memory retrieval equations was included for the StackOverflow model.
+However, it is slightly different because it takes into account prior odds of tag activations which wasn't necessary in SNIF-ACT.
 
@@
 Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that our attentional resources are limited.
-For the StackOverflow model, the attentional weighting was left unbound, and allowed to be calibrated by the logistic regression model.
-This was done for two main reasons.
+For the StackOverflow model, the attentional weighting was left unbound and allowed to be calibrated by the logistic regression model.
+This was done for two main reasons:
 First, the researcher was curious to see what value the logistic regression model would give for $W$.
@@
 During initial model testing, the attentional weighting for all words was equal.
-That is initially, for a given post with $n$ title words, the attentional weighting $W_{j\in C}$ for all of those words was $W/n$, where the logistic regression model fit $W$ to range between 1-3.
-The problem with this approach was that familiar stop words like 'the' and '?', that were observed so often and for all tags, had the same attentional weighting as highly predictive words like 'PHP'.
+That is initially, for a given post with $n$ title words, the attentional weighting $W_{j\in C}$ for all of those words was $W/n$.
+The problem with this approach was that familiar stop words like 'the' and '?', that were often observed across all tags had the same attentional weighting as highly predictive words like 'PHP'.
 Granted the $S_{ji}$'s across all tags for context words like 'the' are much lower (and more evenly distributed) than the few $S_{ji}$'s that spike for 'PHP' (e.g., the tag PHP).
-So the amount of adjusted log odds from stop words was low, even with equal attentional weighting (since their $S_{ji}$'s were low).
-However, it still seemed problematic for two reasons.
+So the amount of adjusted log odds from stop words was low even with equal attentional weighting (since their $S_{ji}$'s were low).
+However, it still seemed problematic for two reasons:
 First, intuitively it seemed that regardless of word co-occurrence association strengths, because stop words like 'the' are observed with all tags, they are not highly predictive of any tag.
@@
 
-It's the case that there are many connections and commonalities between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
+It's also the case that there are many connections and commonalities between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
 and many of those theories use a combination of weighting schemes for contextual relevance and global importance of a particular chunk in context.
-For example, \citeA{Dumais1991} shows that a measure of how likely term ${_j}$ is in document ${_i}$ is often the product of a local weighting $L_{ji}$ and global weighting $G_{j}$ measure.
+For example, \citeA{Dumais1991} shows that a measure of how likely term $j$ is in document $i$ is often the product of a local weighting $L_{ji}$ and global weighting $G_{j}$ measure.
 This idea maps rather nicely to ACT-R's contextual weighting.
-That is, terms are context words ${_j}$, documents are tags ${_i}$, local weighting is strength of association $S_{ji}$, and global weighting is attentional weighting $W_{j}$.
+That is, terms are context words $j$, documents are tags $i$, local weighting is strength of association $S_{ji}$, and global weighting is attentional weighting $W_{j}$.
 \citeA{Dumais1991} states that predictions tend to improve when a global weighting scheme is used, so that words that occur frequently or with many documents (tags) are not weighted as high.
@@
 This can be seen by comparing the implementation of spreading activation here to the implementation in \citeA{Douglass2010}.
-Both implementations handle sji counts greater than \num{1000000}, with declarative memory retrieval times (time to calculate $A_i$'s) around one second at this scale.
+Both implementations handle sji counts greater than \num{1000000}, with declarative memory retrieval times (time to calculate $A_{i}$'s) around one second at this scale.
 However, the implementations could not be more different.

### 2013-02-11 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: dfe23ad5e3cbe1ab68f50fc02835bced374ed478 [formerly d0a204d9bfac40773aca60ca6dc67fae32f7ff51]
Former-commit-id: 075518be9385455baa706b5a1b872c7b7c7f7aaf

@@
 The offset measure to determine the top-five tags and associated activation and the sampling technique used to pick a representative set of tags for each post.
-So model calibration was performed four times, where the coefficients from the previous run were used as estimates of coefficients for the next run.
-The coefficients converged on best-fit values quickly (running just two times would have sufficed), and did not show any sign of instability with the iterative technique.
+So model calibration was iteratively bootstrapped, where the coefficients from the previous run were used as estimates of coefficients for the next run.
+The coefficients converged on best-fit values quickly (only required two sets of four iterations), and did not show any signs of instability.
+
 The final best-fit coefficients for each model component and formal statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits} respectively.

### 2013-02-11 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 0a6e602fc520c091242414787d268b5461a1f5b3 [formerly aa7eaba6543aa5a861e0c29439c4ae58b6e0e8f7]
Former-commit-id: 9531bcd2f5dbe3b47470ae52b369779c7c5ce387

@@
 
-\section{Method}
-
-A variation of ACT-R's declarative memory retrieval theory \cite{Anderson2004} was used to model tag retrievals for StackOverflow posts.
-The hypothesis is that the tags used for posts are based on a tag's prior use and the strength of association between the tag and the content of words in the title and body of the post.
-The strength of association between post words and tags was calculated using 2/3 of the dataset (one million posts).
-A logistic regression statistical technique was used to calibrate model parameters and optimize performance.
-The weights for each model component were calibrated using a sample of \num{1000} posts not contained within the strength of association dataset.
-The calibrated model was then tested on a new sample of \num{1000} posts not contained within either previous training dataset.
-Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).
-
 \subsection{Stack Overflow Dataset}
@@
 
-The Apr' 11 dataset \cite{DataDump2011} was downloaded and imported into the MySQL relational database system.
+\section{Method}
+
+\subsubsection{Summary}
+
+A variation of ACT-R's declarative memory retrieval theory \cite{Anderson2004} was used to model tag retrievals for StackOverflow posts.
+The hypothesis is that the tags used for posts are based on a tag's prior use and the strength of association between the tag and the content of words in the title and body of the post.
+The strength of association between post words and tags was calculated using 2/3 of the dataset (one million posts).
+A logistic regression statistical technique was used to calibrate model parameters and optimize performance.
+The weights for each model component were calibrated using a sample of \num{1000} posts not contained within the strength of association dataset.
+The calibrated model was then tested on a new sample of \num{1000} posts not contained within either previous training dataset.
+Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).
+
+The Apr' 11 StackOverflow dataset \cite{DataDump2011} was downloaded and imported into the MySQL relational database system.
 The dataset contains \num{4945168} posts from \num{559803} unique users.
@@
 The Python Natural Language Processing Toolkit \cite{Bird2009} was used to chunk (i.e., tokenize and lemmatize) the text in the post.
-The tags were already chunked.
-However for the tags, tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
+The tags were already chunked, but tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
 Afterwords the processed word chunks and root tags were fed back in to the database.
@@
 A query was then used to build the post word, tag word co-occurrence matrix.
-For a given word found in the body or title of a post, the query counts the number of times that that word appears with each tag.
-Each of those counts is recorded as the word co-occurrence for that word and each associated tag.
-This process is done across all words found in any post.
 The result is a set of context word, tag, count for each of the \num{39223968} unique context word, tag combinations.
@@
 and an offset to equalize activation values for tags across all posts.
-
-Contextual activation is computed by comparing the words in the post with each tag's word co-occurrence vector.
-If a tag's co-occurrences correlate strongly with the words in the post, then that tag's activation will be high.
-If there is a low correlation between the observed co-occurrences for a tag and the observed words in the post, than that tag's activation will be low.
 The equations that operationalize this process will be included in the Model section.
@@
 The R statistical programming environment was used to build the model and generate model predictions.
-This was done by feeding the word co-occurrence and tag occurrence files generated by the database queries into a separate sparse matrix and sparse vector in R.
-A hash was computed for each unique word, so that indices for the sparse matrix and vector could be numeric.
-The dimensions for the co-occurrence sparse matrix were context word, tag, and the tag occurrence vector was one dimensional (tag).
-
 One million posts (approximately 2/3 of all posts) were fed into R to build the co-occurrences and tag occurrences.
-An interface to the model was developed so that one could supply a vector of words, and the model would return a sparse vector of tag activations given that context.
+These were represented as a sparse matrix of co-occurrences and a sparse vector of tag occurrences.
 In order to generate a model prediction, tag activations for title and body context were computed separately, and then added to each tag's prior base-level activation.
@@
 
-The second component of contextual activation, attentional weighting ($W_{j}$), can be thought of as the amount of attentional resources that are dedicated to contextual element $j$.
+Attentional weighting ($W_{j}$), can be thought of as the amount of attentional resources that are dedicated to contextual element $j$.
 Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that our attentional resources are limited.
 For the StackOverflow model, the attentional weighting was left unbound and allowed to be calibrated by the logistic regression model.
-This was done for two main reasons:
+This was done for two reasons:
 First, the researcher was curious to see what value the logistic regression model would give for $W$.
@@
 That is, terms are context words $j$, documents are tags $i$, local weighting is strength of association $S_{ji}$, and global weighting is attentional weighting $W_{j}$.
-\citeA{Dumais1991} states that predictions tend to improve when a global weighting scheme is used, so that words that occur frequently or with many documents (tags) are not weighted as high.
-He reviews four different commonly-used weighting schemes; one of which is a scaled entropy measure.
 
@@
 It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
-Entropy measures that are mathematically equivalent to Shannon's formulation are used in a wide range of applications, including data compression, cryptography, and thermodynamics.
 The scaled entropy ($E_{j}$) of a context word can be thought of as the amount of predictiveness that that word has to any tag.
@@
 For example, a stop word like 'the' will have an even distribution of co-occurrences across all tags.
-That is, seeing the word 'the' in context is not highly predictive of any particular tag.
 This means that the marginal probabilities of 'the' with all tags will be roughly equal.
@@
 
+\subsubsection{Summary}
+
 After training the model's spreading activation matrix and calibrating model parameters, the model can predict 67\% of tags correctly when asked to generate an average of one tag per post.
@@
 The offset measure to determine the top-five tags and associated activation and the sampling technique used to pick a representative set of tags for each post.
-So model calibration was iteratively bootstrapped, where the coefficients from the previous run were used as estimates of coefficients for the next run.
+So model calibration was bootstrapped, where the coefficients from the previous run were used as estimates of coefficients for the next run.
 The coefficients converged on best-fit values quickly (only required two sets of four iterations), and did not show any signs of instability.
@@
 
+\vspace{-2em}
+
 \renewcommand{\arraystretch}{1}% Wider
@@
 
-Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.46$).
-It's also interesting to note that although the attentional weighting term ($W$) was allowed to be calibrated by the regression, it did not stray far from one
+Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.46$, npv = .997, ppv = .70).
+It's also interesting to note that although the attentional weighting terms ($W_{title}, W_{body}$) were calibrated by the regression, they did not stray far from cognitively-plausible defaults
 (see coefficients for title and body in Table \ref{tab:coeffs}).
-Apparently the scaled entropy measure that globally weights each cue allows for total attentional resources ($W$) to remain relatively small (around one) and not stray far from cognitively-plausible defaults (one).
-
-A plot of model performance for the logistic regression for the calibration run is included in Figure \ref{fig:logReg}.
-Tag-not-chosen observed instances are collected on the bottom and tag-chosen instances are on the top.
-Looking at the histograms, note that model can not perfectly discriminate between tags that will and will not be used for a single post (overlap in distributions).
-Once further improvements to the model are found, these will increase the separation between the two histograms.
-
-\begin{figure}[ht]
-  \centering
-  \resizebox{\linewidth}{!}{\includegraphics{foo-cropped.pdf}}
-  \caption{
-    Logistic regression of tag used on post (binary) regressed on four components of model tag activation.
-  }
-  \label{fig:logReg}
-\end{figure}
 
@@
 The increase in model performance is most likely due to the fact that \citeA{Kuo2011} only used \num{10000} posts to generate the co-occurrence matrix ($S_{ji}$).
-One million posts (two orders of magnitude higher) was used here.
+One million posts (two orders of magnitude higher) were used here.
 Also the model here has some additional components: The offset in particular and possibly the body term as well,
@@
 \subsection{ACT-R and Scale}
-\num{39223968} unique context word, tag combinations were contained within the \num{1000000} posts that were used to build the strength of association matrix.
+\num{39223968} unique context word, tag combinations were contained within the one million posts that were used to build the strength of association matrix.
 Since the model is based on ACT-R's declarative memory retrieval equations, it could have been implemented in Common Lisp (programming language for ACT-R's implementation).
@@
 This can be seen by comparing the implementation of spreading activation here to the implementation in \citeA{Douglass2010}.
-Both implementations handle sji counts greater than \num{1000000}, with declarative memory retrieval times (time to calculate $A_{i}$'s) around one second at this scale.
+Both implementations handle sji counts greater than one million, with declarative memory retrieval times (time to calculate $A_{i}$'s) around one second at this scale.
 However, the implementations could not be more different.
@@
 
-\subsection{Future Work}
-
-Classification accuracy for the model is already fairly high.
-However, there is certainly room for improvement and unexplained variance left to account for.
-In particular, the prior activations for each tag are not (currently) dependent on the particular user that is writing the post, and it seems plausible that prior activations for tags would change for each user.
-That is, if a user has repeatedly tagged questions with Common Lisp, and never with C\#, it seems unrealistic to use the general prior activations for this case, and give C\# the highest prior, and Common Lisp a relatively low prior.
-Instead, it may be worthwhile to maintain prior tag activations for each user, and combine this prior with the global prior when predicting tags for a particular post.
-And this approach could be fielded as well, since each user has to sign into their account before posting a question on the site.
-
 \section{Acknowledgments}

### 2013-02-11 stackExchange/analysis/model/analyzer.R
foo
Former-commit-id: 1173cbdf45ce6c94861efa09285410a693a0b559 [formerly f6ea72cc6ccc3975d85291f10f8926cb473e234a]
Former-commit-id: 56c94aa7ea4e4d99e5ff1f083f7db529bb7847fd

@@ -140,8 +140,7 @@ getAllFrms = function() {
 	frms
 }
 
-plotAllFrms = function() {
-	#frms = getAllFrms()
+plotAllFrms = function(frms=getAllFrms()) {
 	legendText = c("Full Model, Calibration", "Full Model, Test", "Without Body Words", "Without Title Words", 
 		"Without Offset", "Without Entropy Weighting", "Without Offset & Entropy", "Without Offset, Entropy, & Body")
 	items = makeItems(frms[c(1,2,3,5,7,9,10,11)], legendText)
@@ -190,7 +189,7 @@ plotROC2 = function(items) {
 			lines(ROC$x, ROC$y, col=col, typ="p", pch=c(cnt, rep(NA, 32)))
 		} else {
 			plot(ROC$x, ROC$y, xlim=xlim, ylim=c(.2, .9), col=col, pch=c(cnt, rep(NA, 32)),
-				xlab="proportion of model tag count to observed tag count", ylab="model tag proportion correct")
+				xlab="Proportion of model tag count to observed tag count", ylab="Model tag proportion correct")
 		}
 	}	
 	legend("topright", legend=columns, col=colors, pch=1:length(columns))

### 2013-02-12 stackExchange/analysis/model/analyzer.R
foo
Former-commit-id: 61c15199cf153b8196aca1135fed7a2790c74482 [formerly ef1f7930854a5d70a4583ac4f2767d9961b10210]
Former-commit-id: c3a6e1000100bc846732b509a5540569b072644a

@@ -151,7 +151,7 @@ makeItems = function(frms, legendText=as.character(c(1:length(frms)))) {
 	cnt = 0
 	foo = function(frm) {
 		cnt <<- cnt+1
-		list(act=frm$act, targetP=frm$targetP, tagFile=frm$tagFile, col=colors[cnt], legendText=legendText[cnt])
+		list(act=frm$act, targetP=frm$targetP, tagFile=frm$tagFile, col=colors[cnt], legendText=legendText[cnt], shapes=shapes[cnt])
 	}
 	lapply(frms, foo)
 }
@@ -170,11 +170,13 @@ plotROCColumn = function(dFrame, column) {
 }
 
 colors = c("red", "green", "blue", "orange", "yellow", "black", "grey", "purple", "magenta", "cyan", "pink")
+shapes = c(1, 2, 3, 4, 5, 20, 7, 8, 9, 10, 11)
 colors = rep("black", 10)
 
 plotROC2 = function(items) {
 	asFig("ROC")
 	colors = unlist(lapply(items, function(item) {item$col}))
+	shapes = unlist(lapply(items, function(item) {item$shape}))
 	foo = function(item) {
 		data.frame(act=item$act, targetP=item$targetP, tagFile=item$tagFile)
 	}
@@ -186,13 +188,13 @@ plotROC2 = function(items) {
 		cnt = cnt + 1
 		col = colors[cnt]
 		if(cnt != 1) {
-			lines(ROC$x, ROC$y, col=col, typ="p", pch=c(cnt, rep(NA, 32)))
+			lines(ROC$x, ROC$y, col=col, typ="p", pch=c(shapes[cnt], rep(NA, 32)))
 		} else {
-			plot(ROC$x, ROC$y, xlim=xlim, ylim=c(.2, .9), col=col, pch=c(cnt, rep(NA, 32)),
+			plot(ROC$x, ROC$y, xlim=xlim, ylim=c(.2, .9), col=col, pch=c(shapes[cnt], rep(NA, 32)),
 				xlab="Proportion of model tag count to observed tag count", ylab="Model tag proportion correct")
 		}
 	}	
-	legend("topright", legend=columns, col=colors, pch=1:length(columns))
+	legend("topright", legend=columns, col=colors, pch=shapes[1:length(columns)])
 	with(items[[1]], abline(v=length(unique(tagFile))/sum(targetP), lty=3))
 	dev.off()
 }

### 2013-02-12 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 5cdaf4fa5b714ba7a0210ccdf2c4cfd36e9afaf8 [formerly c565ee496474f64f3bbdbadc05fa9f6a5b0bfa44]
Former-commit-id: e82e30f6db4ce48a8722061dcef9cb3c7c861f88

@@
 It is constrained because there is a fully enumerated and limited set of tags that an author can use, as opposed to a less-constrained domain such as Twitter.
-The site also makes public all posting data on a quarterly basis, which can be downloaded and loaded into a local relational database for further analysis.
+The site also makes public all post data on a quarterly basis, which can be downloaded and loaded into a relational database for further analysis.
 
@@
   \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-screen-cropped.pdf}}
-  \caption{Example Post}
+  \caption{Example post on stackoverflow.com}
   \label{fig:examplePost}
@@
 The tags were already chunked, but tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
-Afterwords the processed word chunks and root tags were fed back in to the database.
+Afterwords the processed word chunks and root tags were imported into a PostgreSQL database table.
 
@@
 The R statistical programming environment was used to build the model and generate model predictions.
-One million posts (approximately 2/3 of all posts) were fed into R to build the co-occurrences and tag occurrences.
+One million posts (approximately 2/3 of all posts) were loaded into R to build the co-occurrences and tag occurrences.
 These were represented as a sparse matrix of co-occurrences and a sparse vector of tag occurrences.
@@
 An offset constant (equal to the average activation of the top 5 model-predicted tags for the post) was then subtracted from the prior value.
-The resulting tags were rank sorted, and the one with the highest activation would then be the model's prediction of the most likely tag associated with that post.
+The resulting tags were rank sorted, and the one with the highest activation is the model's prediction of the most likely tag associated with that post.
 
@@
 
-These vectors of tag activation, categorization were collected across a set of \num{1000} posts.
-The set of posts did not include any posts that were used to train the model on word association strength.
+These vectors of tag activation, categorization were collected across a set of \num{1000} fresh posts that were not used to train the model on word association strength.
 So a form of cross validation was used to measure model fit, although the more robust full k-fold cross validation technique was not used (due mostly to resource and time limitations).
@@
 
-The tag-prediction model is a cognitively-inspired Bayesian probabilistic model, based on ACT-R's declarative memory retrieval mechanisms \cite{Anderson2004}.
+The tag-prediction model is a cognitively-inspired Bayesian probabilistic model based on ACT-R's declarative memory retrieval mechanisms \cite{Anderson2004}.
 These equations describe the likelihood that a chunk from memory will be retrieved, given the prior odds of retrieving the chunk and the current context.
@@
 Contextual activation has two components: strength of association ($S_{ji}$) and attentional weight ($W_{j}$).
-Strength of association measures the relatedness of a contextual element $i$ (e.g., the word PHP in the title of a post) with a tag $j$ (e.g., the tag PHP5).
+Strength of association measures the relatedness of a contextual element $j$ (e.g., the word PHP in the title of a post) with a tag $i$ (e.g., the tag PHP5).
 
@@
 That is initially, for a given post with $n$ title words, the attentional weighting $W_{j\in C}$ for all of those words was $W/n$.
-The problem with this approach was that familiar stop words like 'the' and '?', that were often observed across all tags had the same attentional weighting as highly predictive words like 'PHP'.
+The problem with this approach is that familiar stop words like 'the' and '?', that are often observed across all tags have the same attentional weighting as highly predictive words like 'PHP'.
 Granted the $S_{ji}$'s across all tags for context words like 'the' are much lower (and more evenly distributed) than the few $S_{ji}$'s that spike for 'PHP' (e.g., the tag PHP).
-So the amount of adjusted log odds from stop words was low even with equal attentional weighting (since their $S_{ji}$'s were low).
-However, it still seemed problematic for two reasons:
-First, intuitively it seemed that regardless of word co-occurrence association strengths, because stop words like 'the' are observed with all tags, they are not highly predictive of any tag.
+So the amount of adjusted log odds from stop words is low even with equal attentional weighting (since their $S_{ji}$'s are low).
+However, this still seemed problematic for two reasons:
+First, because stop words like 'the' are observed with all tags, they are not highly predictive of any tag.
 So one could argue that their presence is simply adding error in the model's tag activation calculation.
@@
 
-The scaled entropy measure $E_{j}$ was chosen as a global weighting measure for this dataset because it has a strong theoretical foundation and it predicts well (\citeNP{Dumais1991}).
+So the scaled entropy measure $E_{j}$ was chosen as a global weighting measure for this dataset because it has a strong theoretical foundation and it predicts well (\citeNP{Dumais1991}).
 It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
@@
     The title is the title text of the post.
-    The attentional weight (entropy) for each word in the title is included below each word.
+    The attentional weight (scaled entropy) for each word in the title is included below each word.
     The user-chosen tags are included below the attentional weights.
     The stacked bar graphs are the total activation for the top-ten predicted model tags for the post.
-    Each bar graph is partitioned into the 3 model components that vary between tags within a post (not the post offset).
+    Each bar graph is partitioned into the three model components that vary across the tags in a single post (not the offset).
 }
@@
 Figure \ref{fig:ROC} also shows model performance after specific model components are removed.
-Model parameters were recalibrated for each model, except for the full model test dataset which used the parameters from the calibration dataset.
+Model parameters were recalibrated for each model, except for the full model test dataset that used the parameters from the calibration dataset.
 These plots provide useful information regarding the relative necessity of each component in the full model.
@@
 (height of full model at vertical dotted line in Figure \ref{fig:ROC}).
-The increase in model performance is most likely due to the fact that \citeA{Kuo2011} only used \num{10000} posts to generate the co-occurrence matrix ($S_{ji}$).
+The increase in model performance here is most likely due to the fact that \citeA{Kuo2011} only used \num{10000} posts to generate the co-occurrence matrix ($S_{ji}$).
 One million posts (two orders of magnitude higher) were used here.
@@
 But it is much more common outside of the ACT-R modeling community to use both a local (strength of association) and global (attentional weight) term to predict document retrieval \citeA{Dumais1991}.
-In fact, it's been shown that the Pointwise Mutual Information Index (PMI) (a commonly-used measure to predict document retrieval) is mathematically equivalent to ACT-R's retrieval equations when the sample size of the corpus is large (\citeNP{Budiu2007}, \citeNP{Farahat2004}).
-Both global and local weightings are commonly used when the PMI measurement is used.
-This suggests that using a global weighting term for chunks in context (e.g., the scaled entropy measure used here) alongside the more commonly-used local weighting (spreading activation) might be useful for other ACT-R models that are taking advantage of large stores of declarative memory datasets to build up the model's base knowledge.
+In fact, it's been shown that the Pointwise Mutual Information Index (PMI) (a commonly-used measure to predict document retrieval) is mathematically equivalent to ACT-R's retrieval equations when the sample size of the corpus is large (\citeNP{Budiu2007}, \citeNP{Farahat2004}),
+and both global and local weightings are commonly used when the PMI measurement is used.
+This suggests that using a global weighting term for chunks in context (e.g., the scaled entropy measure used here) alongside the more commonly-used local weighting (spreading activation) might be useful for other ACT-R models that are taking advantage of large declarative memory datasets to build up the model's base knowledge.
 

### 2013-02-22 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 5e37e8c2147ccb8c5e97497bc85d6e8b7c1c5bfb [formerly e4471efa906f058b1d29cbfd32fa518bab040b9c]
Former-commit-id: d4b902efd494161c262d9b0021d32f478ead1730
foo
Former-commit-id: aa13e6dc1787986f319569364d21296720a03bc0 [formerly 6df9716c05339a39e3456e76781836687004029e]
Former-commit-id: e5a294bf681ec2bd8fd6f87c965f53f73132e3a1

@@
   Houston, TX 77005 USA 
-  \AND {\large \bf Mike Byrne (byrne@rice.edu)} \\
+  \AND {\large \bf Michael D. Byrne (byrne@rice.edu)} \\
   Department of Psychology and Computer Science, 6100 Main Street \\
@@
 \maketitle
+
+\frenchspacing
 
@@
 Human-based tagging of online content has increased in recent years.
+A few examples of human-based tagging of information are hashtags for Twitter tweets and Facebook posts, as well as keywords for research papers.
 A large portion of the growing number of social media sites support some sort of human-directed tagging of posts.
@@
 
-Although much good work has been done to characterize the lifetime and growth characteristics of hashtags for real-time social networking sites such as Twitter
-(e.g., \citeNP{Bauer2012}, \citeNP{Tsur2012}, \citeNP{Chang2010})
-less is known about a person's motivation for choosing particular tags during the creation of the tweet or post.
+There is a growing body of research that aims to characterize the lifetime and growth characteristics of hashtags for real-time social networking sites such as Twitter
+(e.g., \citeNP{Bauer2012, Tsur2012, Chang2010}).
+However, less is known about the underlying process when choosing particular tags during the creation of the tweet or post.
 Modeling tag creation accurately (as opposed to modeling tag growth and behavior after creation) could allow a system to predict the hashtag that a user should use, given the contextual clues available.
-The benefits of this type of prediction are numerous.
+There are numerous benefits to this type of tag prediction.
 For example, users could be introduced to tags (and potentially communities surrounding those tags) that are of interest to them.
@@
 
-A tag-prediction system was built here for the StackOverflow dataset that can solve precisely these sorts of problems.
+We built a tag prediction system for the StackOverflow dataset that can solve precisely these sorts of problems.
 The StackOverflow dataset was chosen because it is relatively constrained, contains a large number of posts and associated tags, and the data are easy to obtain.
@@
 The StackOverflow dataset consists of all questions and answers posted on stackoverflow.com.
-The questions are all programming related and can be posted by anyone with an account, which is free to create.
+The questions are all related to computer programming and can be posted by anyone with an account, which is free to create.
 Each question is tagged by the author with (presumably) the tags that are most representative of the post.
-Example tags are programming languages (PHP, Common-Lisp, MySQL, C\#) and also general topics (databases, optimization, arrays).
+Example tags are programming languages (\emph{PHP}, \emph{Common-Lisp}, \emph{MySQL}, \emph{C\#}) and also general topics (\emph{databases}, \emph{optimization}, \emph{arrays}).
 An example post is included in Figure \ref{fig:examplePost}.
-Note that the author tagged this post with javascript, firefox, dom, and svg.
+Note that the author tagged this post with \emph{javascript}, \emph{firefox}, \emph{dom}, and \emph{svg}.
 
@@
 
+The Apr' 11 StackOverflow dataset \cite{DataDump2011} was used for this analysis.
+The dataset contains \num{4945168} posts from \num{559803} unique users and \num{26176} unique tags.
+The dataset is the entire collection of posts dated between Jul' 2008 (StackOverflow's creation) and Apr' 11.
+
 \section{Method}
@@
 A variation of ACT-R's declarative memory retrieval theory \cite{Anderson2004} was used to model tag retrievals for StackOverflow posts.
-The hypothesis is that the tags used for posts are based on a tag's prior use and the strength of association between the tag and the content of words in the title and body of the post.
+The hypothesis is that the tags used for posts are based on a tag's history of prior use and the strength of association between the tag and the content of words in the title and body of the post.
 The strength of association between post words and tags was calculated using 2/3 of the dataset (one million posts).
@@
 The Apr' 11 StackOverflow dataset \cite{DataDump2011} was downloaded and imported into the MySQL relational database system.
-The dataset contains \num{4945168} posts from \num{559803} unique users.
 
@@
 A SQL query was used to build a similar set of tag use counts for the \num{26176} unique tags.
-This set for example would show that C\# is the most often used tag, where about 11\% of posts were tagged with C\#.
+This set for example would show that \emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
 
@@
 
-The tag-prediction model is a cognitively-inspired Bayesian probabilistic model based on ACT-R's declarative memory retrieval mechanisms \cite{Anderson2004}.
+The tag prediction model is a cognitively-inspired Bayesian probabilistic model based on ACT-R's declarative memory retrieval mechanisms \cite{Anderson2004}.
 These equations describe the likelihood that a chunk from memory will be retrieved, given the prior odds of retrieving the chunk and the current context.
@@
 
-This model is also similar in design to the approach used by SNIF-ACT (\citeNP{Fu2007}, \citeNP{Pirolli2003}). 
+This model is also similar in design to the approach used by SNIF-ACT \cite{Fu2007, Pirolli2003}. 
 SNIF-ACT can predict link traversal for search queries.
@@
   \begin{center} 
-    \caption{The StackOverflow tag-prediction model} 
+    \caption{The StackOverflow tag prediction model} 
     \label{sample-table} 
@@
 Base-level activation ($B_{i}$) is the log prior odds of observing tag $i$ in the world.
-So the tag with the highest frequency (C\#) would have the highest base-level activation.
+So the tag with the highest frequency (\emph{C\#}) would have the highest base-level activation.
 Tag frequency drops off sharply for this dataset.
@@
 This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.
-Otherwise without context, the likelihood of choosing C\# as a tag would be the same as choosing ora-00997 (one of the many tags with only a single occurrence of use).
-One might argue that with context, it's unlikely that the words associated with ora-00997 will occur again, so pure context would be able to still often predict C\#.
+Otherwise without context, the likelihood of choosing \emph{C\#} as a tag would be the same as choosing \emph{ora-00997} (one of the many tags with only a single occurrence of use).
+One might argue that with context, it's unlikely that the words associated with \emph{ora-00997} will occur again, so pure context would be able to still often predict \emph{C\#}.
 However, the ACT-R theory asserts that pure context is often not enough to overcome high base-level activations (e.g., it can be difficult to break a bad habit).
@@
 
-So the scaled entropy measure $E_{j}$ was chosen as a global weighting measure for this dataset because it has a strong theoretical foundation and it predicts well (\citeNP{Dumais1991}).
+So the scaled entropy measure $E_{j}$ was chosen as a global weighting measure for this dataset because it has a strong theoretical foundation and it predicts well \cite{Dumais1991}.
 It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
@@
 The scaled entropy measure $E_{j}$ transforms the entropy measure to a scale that ranges from 0, 1, where maximum entropy means lowest scaled entropy.
-This means that a stop word like 'the' will have a scaled entropy measure close to 0, while a highly-predictive word like 'macro' will have a scaled entropy measure close to 1.
+This means that a stop word like 'the' will have a scaled entropy measure close to 0, while a highly-predictive word like \emph{macro} will have a scaled entropy measure close to 1.
 Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate the limited attentional resources to predictive contextual elements.
@@
 Since the model is based on ACT-R's declarative memory retrieval equations, it could have been implemented in Common Lisp (programming language for ACT-R's implementation).
-However, it is known that the current implementation of spreading activation (required for $S_{ji}$) in ACT-R does not scale to an $S_{ji}$ corpus over 100,000 (\citeNP{Douglass2009}, \citeNP{Douglass2010}).
+However, it is known that the current implementation of spreading activation (required for $S_{ji}$) in ACT-R does not scale to an $S_{ji}$ corpus over 100,000 \cite{Douglass2010, Douglass2009}.
 So it was decided to use a one-off R implementation of ACT-R's retrieval equations for this analysis.
@@
 But it is much more common outside of the ACT-R modeling community to use both a local (strength of association) and global (attentional weight) term to predict document retrieval \citeA{Dumais1991}.
-In fact, it's been shown that the Pointwise Mutual Information Index (PMI) (a commonly-used measure to predict document retrieval) is mathematically equivalent to ACT-R's retrieval equations when the sample size of the corpus is large (\citeNP{Budiu2007}, \citeNP{Farahat2004}),
+In fact, it's been shown that the Pointwise Mutual Information Index (PMI) (a commonly-used measure to predict document retrieval)
+is mathematically equivalent to ACT-R's retrieval equations when the sample size of the corpus is large \cite{Budiu2007, Farahat2004},
 and both global and local weightings are commonly used when the PMI measurement is used.

### 2013-02-22 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 237823efabeb3e20e84e3b99f6bb51000e21dc33 [formerly d0d7392f5f7c1e2459fdb352bf6373a8f274a7a5]
Former-commit-id: c80f84d1371d97d9a6ceb951bf2d169307611b3f

@@
 \usepackage{apacite}
+\usepackage[none]{hyphenat}
 
@@
 The dataset is the entire collection of posts dated between Jul' 2008 (StackOverflow's creation) and Apr' 11.
-
-\section{Method}
-
-\subsubsection{Summary}
-
-A variation of ACT-R's declarative memory retrieval theory \cite{Anderson2004} was used to model tag retrievals for StackOverflow posts.
-The hypothesis is that the tags used for posts are based on a tag's history of prior use and the strength of association between the tag and the content of words in the title and body of the post.
-The strength of association between post words and tags was calculated using 2/3 of the dataset (one million posts).
-A logistic regression statistical technique was used to calibrate model parameters and optimize performance.
-The weights for each model component were calibrated using a sample of \num{1000} posts not contained within the strength of association dataset.
-The calibrated model was then tested on a new sample of \num{1000} posts not contained within either previous training dataset.
-Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).
-
-The Apr' 11 StackOverflow dataset \cite{DataDump2011} was downloaded and imported into the MySQL relational database system.
-
-\subsection{Measuring Co-occurrences}
-
-Model tag predictions are based on the amount of word co-occurrences between words in the title and body of the post and tags used for that post.
-To build the co-occurrence matrix between post words and tags, all title and body text and associated tags were extracted from the relational database.
-The Python Natural Language Processing Toolkit \cite{Bird2009} was used to chunk (i.e., tokenize and lemmatize) the text in the post.
-The tags were already chunked, but tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
-Afterwords the processed word chunks and root tags were imported into a PostgreSQL database table.
-
-A query was then used to build the post word, tag word co-occurrence matrix.
-The result is a set of context word, tag, count for each of the \num{39223968} unique context word, tag combinations.
-
-\subsection{Measuring Tag Occurrences}
-
-Model tag predictions are also based on the frequency that each tag has been used.
-More often used tags more likely to have higher activation.
-A SQL query was used to build a similar set of tag use counts for the \num{26176} unique tags.
-This set for example would show that \emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
-
-\subsection{Generating Model Predictions}
-
-For a given title and body of a post, the model returns the activations for each possible tag.
-The tag with the highest activation is the most likely tag to be associated with that post.
-Tag activation is a function of four components:
-The tag's base level activation (prior odds of being used),
-contextual activation for words in the title (word co-occurrence weighting),
-contextual activation for words in the body,
-and an offset to equalize activation values for tags across all posts.
-The equations that operationalize this process will be included in the Model section.
-
-The R statistical programming environment was used to build the model and generate model predictions.
-One million posts (approximately 2/3 of all posts) were loaded into R to build the co-occurrences and tag occurrences.
-These were represented as a sparse matrix of co-occurrences and a sparse vector of tag occurrences.
-In order to generate a model prediction, tag activations for title and body context were computed separately, and then added to each tag's prior base-level activation.
-An offset constant (equal to the average activation of the top 5 model-predicted tags for the post) was then subtracted from the prior value.
-The resulting tags were rank sorted, and the one with the highest activation is the model's prediction of the most likely tag associated with that post.
-
-\subsection{Measuring Model Fit}
-
-For a given post, the model produces a vector of tag activations.
-For that post, the author's chosen tags are known (see Figure \ref{fig:examplePost}).
-So model performance can be measured by comparing the model's tag activations with chosen tags for a post, and then aggregated across posts to get an overall average performance.
-If the model consistently produces higher tag activations for chosen tags than non-chosen tags, the model should be able to accurately predict the chosen tags.
-
-A logistic regression technique was used in order to operationalize this fit measure.
-For each post, three sets of tag groups were combined to most accurately sample tag activations while managing computer resource constraints:
-The model's tag activations for the highest 200 tags, a weighted sample of 800 tags (weighted towards higher activation), and the tag activations for all tags that were chosen by the author of the post.
-Each of these observations was assigned a category of 0 or 1 depending on if the tag was a chosen tag or not for that post (0 is not chosen).
-Conceptually, if the model has good categorization power, the observations assigned 1 should have higher tag activations than the observations assigned 0.
-
-These vectors of tag activation, categorization were collected across a set of \num{1000} fresh posts that were not used to train the model on word association strength.
-So a form of cross validation was used to measure model fit, although the more robust full k-fold cross validation technique was not used (due mostly to resource and time limitations).
-
-The base R glm function (with 'binomial' argument) was used to perform the logistic regression.
-The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from the four predictor variables:
-The tag's prior occurrence frequency, word co-occurrence activation for title and body words, and activation offset for the post.
-Further discussion and operationalization of the predictor variables is included in the Model section.
+Post authors used approximately 3 tags on average per post.
 
@@
       Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
-      Max Entropy & 		$H_{max} = maxH$ \\
       Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
@@
 Base-level activation ($B_{i}$) is the log prior odds of observing tag $i$ in the world.
+The log prior odds are based on both frequency and recency of use.
+However we are are assuming that frequency of tag use does not change greatly between Jul' 08 and Apr' 11 (time period of dataset).
+This allows us to compute log odds based on frequency only, as is indicated in Table \ref{sample-table}.
 So the tag with the highest frequency (\emph{C\#}) would have the highest base-level activation.
@@
 Contextual activation has two components: strength of association ($S_{ji}$) and attentional weight ($W_{j}$).
-Strength of association measures the relatedness of a contextual element $j$ (e.g., the word PHP in the title of a post) with a tag $i$ (e.g., the tag PHP5).
+Strength of association measures the relatedness of a contextual element $j$ (e.g., the word PHP in the title of a post) with a tag $i$ (e.g., the tag \emph{PHP5}).
 
@@
 Attentional weighting ($W_{j}$), can be thought of as the amount of attentional resources that are dedicated to contextual element $j$.
-Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that our attentional resources are limited.
+Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that attentional resources are limited.
 For the StackOverflow model, the attentional weighting was left unbound and allowed to be calibrated by the logistic regression model.
 This was done for two reasons:
-First, the researcher was curious to see what value the logistic regression model would give for $W$.
+First, the we were curious to see what value the logistic regression model would give for $W$.
 Second, a non-ACT-R standard measure of attentional weighting was used ($E_{j}$), so it seemed appropriate to allow $W$ to vary alongside exploring this new weighting measure.
@@
 That is initially, for a given post with $n$ title words, the attentional weighting $W_{j\in C}$ for all of those words was $W/n$.
-The problem with this approach is that familiar stop words like 'the' and '?', that are often observed across all tags have the same attentional weighting as highly predictive words like 'PHP'.
-Granted the $S_{ji}$'s across all tags for context words like 'the' are much lower (and more evenly distributed) than the few $S_{ji}$'s that spike for 'PHP' (e.g., the tag PHP).
+The problem with this approach is that familiar stop words like ``the'' and ``?'', that are often observed across all tags have the same attentional weighting as highly predictive words like ``PHP''.
+Granted the $S_{ji}$'s across all tags for context words like ``the'' are much lower (and more evenly distributed) than the few $S_{ji}$'s that spike for ``PHP'' (e.g., the tag \emph{PHP}).
 So the amount of adjusted log odds from stop words is low even with equal attentional weighting (since their $S_{ji}$'s are low).
 However, this still seemed problematic for two reasons:
-First, because stop words like 'the' are observed with all tags, they are not highly predictive of any tag.
+First, because stop words like ``the'' are observed with all tags, they are not highly predictive of any tag.
 So one could argue that their presence is simply adding error in the model's tag activation calculation.
-Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words like 'the' and '?' are taken away from being used for highly predictive words like 'PHP'.
+Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words like ``the'' and ``?'' are taken away from being used for highly predictive words like ``PHP''.
 
-It's also the case that there are many connections and commonalities between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
+It's also the case that there are many commonalities between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
 and many of those theories use a combination of weighting schemes for contextual relevance and global importance of a particular chunk in context.
@@
 
-For example, a stop word like 'the' will have an even distribution of co-occurrences across all tags.
-This means that the marginal probabilities of 'the' with all tags will be roughly equal.
+For example, a stop word like ``the'' will have an even distribution of co-occurrences across all tags.
+This means that the marginal probabilities of ``the'' with all tags will be roughly equal.
 It can be shown that maximum entropy $H_{j}$ is reached when all marginal probabilities are equal.
 This makes intuitive sense because if all marginal probabilities are equal, there is no information gained in observing the context word.
-The scaled entropy measure $E_{j}$ transforms the entropy measure to a scale that ranges from 0, 1, where maximum entropy means lowest scaled entropy.
-This means that a stop word like 'the' will have a scaled entropy measure close to 0, while a highly-predictive word like \emph{macro} will have a scaled entropy measure close to 1.
+The scaled entropy measure $E_{j}$ transforms the entropy measure to a scale that ranges from 0 to 1, where maximum entropy means lowest scaled entropy.
+This means that a stop word like ``the'' will have a scaled entropy measure close to 0, while a highly-predictive word like \emph{macro} will have a scaled entropy measure close to 1.
 Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate the limited attentional resources to predictive contextual elements.
@@
 The model would then select the ten tags from the single post as the most likely tags.
-But this doesn't make intuitive sense, because there is a limit to the number of tags that people tend to use for each post (average number of tags per post is approximately 3).
+But this doesn't make intuitive sense, because there is a limit to the number of tags that people tend to use for each post.
 So the offset measure aims to place all of these ten posts on the same level, so that asking the model to choose ten tags from ten posts will select roughly a single tag from each post.
 
-\section{Results}
+\section{Method}
 
@@
 
-After training the model's spreading activation matrix and calibrating model parameters, the model can predict 67\% of tags correctly when asked to generate an average of one tag per post.
+A variation of ACT-R's declarative memory retrieval theory \cite{Anderson2004} was used to model tag retrievals for StackOverflow posts.
+The hypothesis is that the tags used for posts are based on a tag's history of prior use and the strength of association between the tag and the content of words in the title and body of the post.
+The strength of association between post words and tags was calculated using 2/3 of the dataset (one million posts).
+A logistic regression statistical technique was used to calibrate model parameters and optimize performance.
+The weights for each model component were calibrated using a sample of \num{1000} posts not contained within the strength of association dataset.
+The calibrated model was then tested on a new sample of \num{1000} posts not contained within either previous training dataset.
+Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).
+
+FIXME. The Apr' 11 StackOverflow dataset \cite{DataDump2011} was downloaded and imported into the MySQL relational database system.
+
+\subsection{Generating Model Predictions}
+
+For a given title and body of a post, the model returns the activations for each possible tag.
+The tag with the highest activation is the most likely tag to be associated with that post.
+Tag activation is a function of four components:
+The tag's base level activation (prior odds of being used),
+contextual activation for words in the title (word co-occurrence weighting),
+contextual activation for words in the body,
+and an offset to equalize activation values for tags across all posts.
+
+The R statistical programming environment was used to build the model and generate model predictions.
+One million posts (approximately 2/3 of all posts) were loaded into R to build the co-occurrences and tag occurrences.
+These were represented as a sparse matrix of co-occurrences and a sparse vector of tag occurrences.
+In order to generate a model prediction, tag activations for title and body context were computed separately, and then added to each tag's prior base-level activation.
+An offset constant (equal to the average activation of the top 5 model-predicted tags for the post) was then subtracted from the prior value.
+The resulting tags were rank sorted, and the one with the highest activation is the model's prediction of the most likely tag associated with that post.
+
+\subsection{Measuring Co-occurrences}
+
+Model tag predictions are based on the amount of word co-occurrences between words in the title and body of the post and tags used for that post.
+To build the co-occurrence matrix between post words and tags, all title and body text and associated tags were extracted from the relational database.
+The Python Natural Language Processing Toolkit \cite{Bird2009} was used to chunk (i.e., tokenize and lemmatize) the text in the post.
+The tags were already chunked, but tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
+Afterwords the processed word chunks and root tags were imported into a PostgreSQL database table.
+
+A query was then used to build the post word, tag word co-occurrence matrix.
+There are \num{39223968} unique post word, tag word combinations.
+The result is a table of co-occurence counts for each of the combinations.
+
+\subsection{Measuring Tag Occurrences}
+
+Model tag predictions are also based on the frequency that each tag has been used.
+More often used tags more likely to have higher activation.
+A SQL query was used to build a similar set of tag use counts for the \num{26176} unique tags.
+This set for example would show that \emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
+
+\subsection{Measuring Model Fit}
+
+For a given post, the model produces a vector of tag activations.
+For that post, the author's chosen tags are known (see Figure \ref{fig:examplePost}).
+So model performance can be measured by comparing the model's tag activations with chosen tags for a post, and then aggregated across posts to get an overall average performance.
+If the model consistently produces higher tag activations for chosen tags than non-chosen tags, the model should be able to accurately predict the chosen tags.
+
+A logistic regression technique was used in order to operationalize this fit measure.
+For each post, three sets of tag groups were combined to most accurately sample tag activations while managing computer resource constraints:
+The model's tag activations for the highest 200 tags, a weighted sample of 800 tags (weighted towards higher activation), and the tag activations for all tags that were chosen by the author of the post.
+Each of these observations was assigned a category of 0 or 1 depending on if the tag was a chosen tag or not for that post (0 is not chosen).
+Conceptually, if the model has good categorization power, the observations assigned 1 should have higher tag activations than the observations assigned 0.
+
+These vectors of tag activation, categorization were collected across a set of \num{1000} fresh posts that were not used to train the model on word association strength.
+So a form of cross validation was used to measure model fit, although the more robust full k-fold cross validation technique was not used (due mostly to resource and time limitations).
+
+The base R glm function (with ``binomial'' argument) was used to perform the logistic regression.
+The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from the four predictor variables:
+The tag's prior occurrence frequency, word co-occurrence activation for title and body words, and activation offset for the post.
+Further discussion and operationalization of the predictor variables is included in the Model section.
+
+\section{Results}
 
@@
 Note how the base level activation and contextual activation combine to predict tags that are both frequently used and contextually relevant.
-For example, note that the 'javascript' tag leverages a combination of contextual activation and prior activation to achieve a high total activation.
-Alternatively, the 'svg' tag relies almost entirely on context to achieve high activation.
+For example, note that the \emph{javascript} tag leverages a combination of contextual activation and prior activation to achieve a high total activation.
+Alternatively, the \emph{svg} tag relies almost entirely on context to achieve high activation.
 And note that the model is by no means perfect.
-For example, 'html' is not tagged by the author, but the model ranks it as the second highest tag.
-And the author's tag 'firefox' is only ranked 8th by the model.
+For example, \emph{html} is not tagged by the author, but the model ranks it as the second highest tag.
+And the author's tag \emph{firefox} is only ranked 8th by the model.
 
@@
     Model performance across the calibration and test dataset, as well as when specific model components are removed.
+    The y-axis represents model performance (one is perfect), and the x-axis is the proportion of tags that the model generates compared to the number of tags used by the authors of the posts (one is equal number).
     The vertical dotted line represents the point where the model is asked to tag on average one tag per post.
@@
 The Python NLTK documentation \cite{Bird2009} recommended to remove stop words from the analysis.
-Stop words are low-predictor words, and the documentation listed out a short list that were commonly removed (e.g., 'the', 'and', 'or')
+Stop words are low-predictor words, and the documentation listed out a short list that were commonly removed (e.g., ``the'', ``and'', ``or'')
 However it seemed somewhat arbitrary to explicitly list out a few stop words and then simply choose those as the words to remove.
@@
 That is, words with scaled entropy close to zero are the stop words.
-Looking at the lowest ten scaled-entropy words for this dataset, (does, anyone, 've, found, are, seems, has, been, for, support), using scaled entropy to determine stop words seems to work fairly well.
+Looking at the lowest ten scaled-entropy words for this dataset, (``does'', ``anyone'', ``'ve'', ``found'', ``are'', ``seems'', ``has'', ``been'', ``for'', ``support''),
+using scaled entropy to determine stop words seems to work fairly well.
 

### 2013-02-25 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 7cfb2bbbf1cee125175de82abb5cc6372872bf00 [formerly 3b5ace12e24df8f63c7c516123d027c06e77cafa]
Former-commit-id: 8b9c9b798834fc9ef9019f02daf6b32e7095b0c9
foo
Former-commit-id: 19dea6f301cd68e2817fb67b723bc355f1176daf [formerly 3eb841c51298976182d17127ee4bbf11931fc508]
Former-commit-id: 04e617db72e99640f1824d6d04fcdf441370b4be

@@
 There is a growing body of research that aims to characterize the lifetime and growth characteristics of hashtags for real-time social networking sites such as Twitter
-(e.g., \citeNP{Bauer2012, Tsur2012, Chang2010}).
+(e.g., \citeNP{Tsur2012, Chang2010}).
 However, less is known about the underlying process when choosing particular tags during the creation of the tweet or post.
@@
 
+\subsection{Related Work}
+
+Predicting tags for the StackOverflow dataset has been worked on by at least one prior researcher \cite{Kuo2011}.
+\citeauthor{Kuo2011} only looks at the Stack Overflow dataset briefly at the end of the technical report.
+However he does mention that his co-occurrence model has a 47\% classification accuracy for predicting StackOverflow tags when required to predict one tag per post.
+Our model achieves 67\% classification accuracy when predicting one tag per post on average.
+
 \section{Model}
@@
 In this way the task of retrieving the tag with the highest activation from the words in the post is similar to performing a declarative memory retrieval request for a tag, given the context of title and body words.
+The hypothesis is that the tags used for posts are based on a tag's history of prior use and the strength of association between the tag and the content of words in the title and body of the post.
 
-This model is also similar in design to the approach used by SNIF-ACT \cite{Fu2007, Pirolli2003}. 
+This model is also similar in design to the approach used by SNIF-ACT \cite{Fu2007}.
 SNIF-ACT can predict link traversal for search queries.
@@
 
-A formal description of the model's equations are included in Table \ref{sample-table}.
+Table \ref{sample-table} includes a description of the model's equations.
 
-\renewcommand{\arraystretch}{1.5}% Wider
+\renewcommand{\arraystretch}{1.5}
 \renewcommand{\tabcolsep}{.5mm}
@@
       Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
-      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN(j,i)}{N_{Row}(j)N_{Col}(i)}$ \\
-      Attention Weight	& 	$W_{j} = WE_{j}$ \\
+      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
+      Co-occurrence &		$N_{ji} = \sum_{}^{}{observed_{ji}}$ \\
+      Observations &		$N = \sum_{}^{}{\sum_{}^{}{N_{ji}}}$ \\
+      Attention Weight	& 	$W_{j} = W \frac{E_{j}} {\sum_{}^{} {E_{j}}} $ \\
       Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
@@
 This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.
-Otherwise without context, the likelihood of choosing \emph{C\#} as a tag would be the same as choosing \emph{ora-00997} (one of the many tags with only a single occurrence of use).
-One might argue that with context, it's unlikely that the words associated with \emph{ora-00997} will occur again, so pure context would be able to still often predict \emph{C\#}.
-However, the ACT-R theory asserts that pure context is often not enough to overcome high base-level activations (e.g., it can be difficult to break a bad habit).
 
@@
 
-During initial model testing, the attentional weighting for all words was equal.
-That is initially, for a given post with $n$ title words, the attentional weighting $W_{j\in C}$ for all of those words was $W/n$.
+During initial model testing, the attentional weighting for all words was equal, and set to $W/n$ for $n$ title words (and set similarly for body words).
 The problem with this approach is that familiar stop words like ``the'' and ``?'', that are often observed across all tags have the same attentional weighting as highly predictive words like ``PHP''.
 Granted the $S_{ji}$'s across all tags for context words like ``the'' are much lower (and more evenly distributed) than the few $S_{ji}$'s that spike for ``PHP'' (e.g., the tag \emph{PHP}).
-So the amount of adjusted log odds from stop words is low even with equal attentional weighting (since their $S_{ji}$'s are low).
 However, this still seemed problematic for two reasons:
-First, because stop words like ``the'' are observed with all tags, they are not highly predictive of any tag.
-So one could argue that their presence is simply adding error in the model's tag activation calculation.
+First, because stop words like ``the'' are observed with all tags, and not highly predictive of any tag, one could argue that their presence is simply adding error in the model's tag activation calculation.
 Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words like ``the'' and ``?'' are taken away from being used for highly predictive words like ``PHP''.
@@
 The scaled entropy ($E_{j}$) of a context word can be thought of as the amount of predictiveness that that word has to any tag.
-
-For example, a stop word like ``the'' will have an even distribution of co-occurrences across all tags.
-This means that the marginal probabilities of ``the'' with all tags will be roughly equal.
-It can be shown that maximum entropy $H_{j}$ is reached when all marginal probabilities are equal.
-This makes intuitive sense because if all marginal probabilities are equal, there is no information gained in observing the context word.
-The scaled entropy measure $E_{j}$ transforms the entropy measure to a scale that ranges from 0 to 1, where maximum entropy means lowest scaled entropy.
 This means that a stop word like ``the'' will have a scaled entropy measure close to 0, while a highly-predictive word like \emph{macro} will have a scaled entropy measure close to 1.
 Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate the limited attentional resources to predictive contextual elements.
+
+\subsubsection{Entropy and Stop Words}
+
+It was recommended to remove stop words from the analysis \cite{Bird2009}.
+Stop words are low-predictor words, and \citeauthor listed a short list that were commonly removed (e.g., ``the'', ``and'', ``or'').
+However it seemed somewhat arbitrary to explicitly list out a few stop words and then simply choose those as the words to remove.
+Looking at the scaled entropy measure, it seems quite plausible that this measure of word predictiveness can also be used to identify stop words.
+That is, words with scaled entropy close to zero are the stop words.
+Looking at the lowest ten scaled-entropy words for this dataset, (``does'', ``anyone'', ``'ve'', ``found'', ``are'', ``seems'', ``has'', ``been'', ``for'', ``support''),
+using scaled entropy to determine stop words seems to work fairly well.
+So instead of explicitly removing an enumerated list of stop words, we used a data-driven approach and let the scaled entropy measure attentuate the words that were not predictive of any tag.
 
@@
 A variation of ACT-R's declarative memory retrieval theory \cite{Anderson2004} was used to model tag retrievals for StackOverflow posts.
-The hypothesis is that the tags used for posts are based on a tag's history of prior use and the strength of association between the tag and the content of words in the title and body of the post.
 The strength of association between post words and tags was calculated using 2/3 of the dataset (one million posts).
@@
 
-FIXME. The Apr' 11 StackOverflow dataset \cite{DataDump2011} was downloaded and imported into the MySQL relational database system.
-
 \subsection{Generating Model Predictions}
@@
 The tag with the highest activation is the most likely tag to be associated with that post.
-Tag activation is a function of four components:
-The tag's base level activation (prior odds of being used),
-contextual activation for words in the title (word co-occurrence weighting),
-contextual activation for words in the body,
-and an offset to equalize activation values for tags across all posts.
 
@@
 One million posts (approximately 2/3 of all posts) were loaded into R to build the co-occurrences and tag occurrences.
-These were represented as a sparse matrix of co-occurrences and a sparse vector of tag occurrences.
 In order to generate a model prediction, tag activations for title and body context were computed separately, and then added to each tag's prior base-level activation.
@@
 
-\subsection{Measuring Co-occurrences}
+\subsubsection{Measuring Co-occurrences}
 
@@
 There are \num{39223968} unique post word, tag word combinations.
-The result is a table of co-occurence counts for each of the combinations.
+The result is a table of co-occurence counts for each of the combinations, which was loaded into R.
 
-\subsection{Measuring Tag Occurrences}
+\subsubsection{Measuring Tag Occurrences}
 
@@
 More often used tags more likely to have higher activation.
-A SQL query was used to build a similar set of tag use counts for the \num{26176} unique tags.
+A SQL query was used to build a similar table of tag use counts for the \num{26176} unique tags, which was loaded into R.
 This set for example would show that \emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
@@
 
-The base R glm function (with ``binomial'' argument) was used to perform the logistic regression.
 The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from the four predictor variables:
 The tag's prior occurrence frequency, word co-occurrence activation for title and body words, and activation offset for the post.
-Further discussion and operationalization of the predictor variables is included in the Model section.
 
@@
 
-\renewcommand{\arraystretch}{1}% Wider
+\renewcommand{\arraystretch}{1}
 \renewcommand{\tabcolsep}{1mm}
@@
 
-\renewcommand{\arraystretch}{1}% Wider
+\renewcommand{\arraystretch}{1}
 \renewcommand{\tabcolsep}{3mm}
@@
 
-\subsection{Related Work}
-Predicting tags for the StackOverflow dataset has been worked on by at least one prior researcher \cite{Kuo2011}.
-\citeA{Kuo2011} only looks at the Stack Overflow dataset briefly at the end of the technical report.
-However he does mention that the co-occurrence model he used has about a 47\% classification accuracy for predicting StackOverflow tags when required to predict one tag per post.
-The full model used here achieves 67\% classification accuracy when predicting one tag per post on average
-(height of full model at vertical dotted line in Figure \ref{fig:ROC}).
-The increase in model performance here is most likely due to the fact that \citeA{Kuo2011} only used \num{10000} posts to generate the co-occurrence matrix ($S_{ji}$).
-One million posts (two orders of magnitude higher) were used here.
-Also the model here has some additional components: The offset in particular and possibly the body term as well,
-as it was unclear from \citeA{Kuo2011} if the author used context from both the body and title of a post or just the title. 
-Nonetheless, there are certainly similarities in the approach:
-Co-occurrence count was used as the core measure of similarity, and both local (strength of association) and global (idf or entropy) measures were used to compute tag likelihood, given context.
+Our StackOverflow model is a successful case showing that ACT-Rs declarative memory retrieval mechanisms can be scaled up and applied to retrieval requests occuring across large-scale declarative memory datasets.
+The model can successfully tag 67/100 posts correctly, when tasked to predict one tag per post on average.
+The model is also computationall space efficient:
+Large-scale relational database techniques and efficient sparse matrix data structures were used for implementation,
+so that the data structures require only 3GB of memory and retrieval requests with a DM dataset of \num{556677795} chunks occur in less than one second on a single processor.
+Although certainly not perfect, model accuracy is significant and retrieval times are fast, so we are considering potential applications of the model in real-world settings.
 
-\subsection{ACT-R and Scale}
-\num{39223968} unique context word, tag combinations were contained within the one million posts that were used to build the strength of association matrix.
-Since the model is based on ACT-R's declarative memory retrieval equations, it could have been implemented in Common Lisp (programming language for ACT-R's implementation).
-However, it is known that the current implementation of spreading activation (required for $S_{ji}$) in ACT-R does not scale to an $S_{ji}$ corpus over 100,000 \cite{Douglass2010, Douglass2009}.
-So it was decided to use a one-off R implementation of ACT-R's retrieval equations for this analysis.
-This is much like how \citeA{Douglass2010} implemented the equations in Erlang when using the theory to test the fan effect on large-scale declarative memory datasets.
+\subsection{Potential Applications}
 
-However, it is increasingly likely that there is no fundamental reason why vanilla ACT-R (Common Lisp version) cannot be optimized to handle spreading activation calculations with datasets this large.
-All that is really needed is a more appropriate data structure to handle the large scale.
-This can be seen by comparing the implementation of spreading activation here to the implementation in \citeA{Douglass2010}.
-Both implementations handle sji counts greater than one million, with declarative memory retrieval times (time to calculate $A_{i}$'s) around one second at this scale.
-However, the implementations could not be more different.
-The Erlang implementation leverages a semantic network of millions of isolated process chunks and massive concurrency to achieve the scale.
-The R implementation here is single threaded and requires only a sparse matrix implementation to achieve the scale.
-Since the programming languages are different, it's not likely the case that only a specific programming language can achieve the scale.
-And since the parallelism is different (single threaded vs multi threaded), it's not likely the case that achieving the scale can only be done with massively concurrent implementations.
-So, it's most likely the case that using an appropriate data structure is all that is needed for ACT-R Common Lisp to handle sji counts in the millions and above.
+The model could be deployed as a tag recommender system for stackoverflow.com. 
+If a post author does not include the model's predicted best tag for the post, then the model could suggest the tag to the author.
+This could potentially introduce authors to tags that are contextually relevant to their interests that they may not have known about.
+This could also improve the answers received for the post, since other expert programmers may have only been monitoring the model's predicted best tag for the post.
 
-\subsection{Entropy and Stop Words}
+The model could also be used as a tag cleanup system for stackoverflow.com.
+In this scenario the tags for a previously submitted post would be compared againt model predictions.
+If the model's predicted tag activation for a used tag is substantially low, then that tag could either be removed from the post, or the post could be flagged for further analysis by an operator.
 
-The Python NLTK documentation \cite{Bird2009} recommended to remove stop words from the analysis.
-Stop words are low-predictor words, and the documentation listed out a short list that were commonly removed (e.g., ``the'', ``and'', ``or'')
-However it seemed somewhat arbitrary to explicitly list out a few stop words and then simply choose those as the words to remove.
-Looking at the scaled entropy measure, it seems quite plausible that this measure of word predictiveness can also be used to identify stop words.
-That is, words with scaled entropy close to zero are the stop words.
-Looking at the lowest ten scaled-entropy words for this dataset, (``does'', ``anyone'', ``'ve'', ``found'', ``are'', ``seems'', ``has'', ``been'', ``for'', ``support''),
-using scaled entropy to determine stop words seems to work fairly well.
+\subsection{Comparison to \citeA{Kuo2011}}
+
+Our model achieves 67\% classification accuracy when predicting one tag per post on average
+(height of full model at vertical dotted line in Figure \ref{fig:ROC})
+compared to 47\% accuracy for the co-occurence model presented in \citeA{Kuo2011}.
+Most likely we achieved the increase in model performance by adding some additional model components:
+The offset and possibly the body term as well,
+as it was unclear if \citeauthor{Kuo2011} used context from both the body and title of a post or just the title.
+Also we used two orders of magnitude more posts to train the model co-occurrence matrix ($S_{ji}$).
+Nonetheless, there are certainly model similarities:
+Co-occurrence count was used as the core measure of similarity, and both local (strength of association) and global (entropy or idf) weighting measures were used to compute tag likelihood, given context.
 
@@
 However, it's uncommon for an ACT-R model that uses the declarative memory retrieval mechanisms to use anything other than equal attentional weighting for each chunk in context.
-But it is much more common outside of the ACT-R modeling community to use both a local (strength of association) and global (attentional weight) term to predict document retrieval \citeA{Dumais1991}.
+But it is much more common outside of the ACT-R modeling community to use both a local (strength of association) and global (attentional weight) term to predict document retrieval \cite{Dumais1991}.
 In fact, it's been shown that the Pointwise Mutual Information Index (PMI) (a commonly-used measure to predict document retrieval)
@@
 
+\subsection{Future Planned Research}
+
 \section{Acknowledgments}
 
-I would like to thank Dr. Mike Byrne and Dr. Fred Oswald for their excellent guidance and recommendations throughout the model development and data analysis process.
+We would like to thank Fred Oswald for his excellent guidance and recommendations throughout the model development and data analysis process.
 

### 2013-02-25 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 59d3969f8cba2dadfb6df97a2c39c9cc2a64b8f8 [formerly e61f9ca637996a44a6977e998a4193ab04dd2b1c]
Former-commit-id: 86ed36194bf1e97dd30da1965af9b80d63d04bc2

@@
 using scaled entropy to determine stop words seems to work fairly well.
-So instead of explicitly removing an enumerated list of stop words, we used a data-driven approach and let the scaled entropy measure attentuate the words that were not predictive of any tag.
+So instead of explicitly removing an enumerated list of stop words, we used a data-driven approach and let the scaled entropy measure attenuate the words that were not predictive of any tag.
 
@@
 There are \num{39223968} unique post word, tag word combinations.
-The result is a table of co-occurence counts for each of the combinations, which was loaded into R.
+The result is a table of co-occurrence counts for each of the combinations, which was loaded into R.
 
@@
 
-Our StackOverflow model is a successful case showing that ACT-Rs declarative memory retrieval mechanisms can be scaled up and applied to retrieval requests occuring across large-scale declarative memory datasets.
+Our StackOverflow model is a successful case showing that ACT-Rs declarative memory retrieval mechanisms can be scaled up and applied to retrieval requests occurring across large-scale declarative memory datasets.
 The model can successfully tag 67/100 posts correctly, when tasked to predict one tag per post on average.
-The model is also computationall space efficient:
+The model is also computationally space efficient:
 Large-scale relational database techniques and efficient sparse matrix data structures were used for implementation,
@@
 The model could also be used as a tag cleanup system for stackoverflow.com.
-In this scenario the tags for a previously submitted post would be compared againt model predictions.
+In this scenario the tags for a previously submitted post would be compared against model predictions.
 If the model's predicted tag activation for a used tag is substantially low, then that tag could either be removed from the post, or the post could be flagged for further analysis by an operator.
@@
 (height of full model at vertical dotted line in Figure \ref{fig:ROC})
-compared to 47\% accuracy for the co-occurence model presented in \citeA{Kuo2011}.
+compared to 47\% accuracy for the co-occurrence model presented in \citeA{Kuo2011}.
 Most likely we achieved the increase in model performance by adding some additional model components:

### 2013-02-26 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: e5707b05b43dbc86d1423e2cbaa16fbb6e129fe8 [formerly adfdb3328cc86f77830bb92fb8f0de50188bdf07]
Former-commit-id: b4fd790a82f21e8ee5daa2a8a1dd9a8ead4848eb

@@
 
-Table \ref{sample-table} includes a description of the model's equations.
+Table \ref{sample-table} includes a formal description of the model.
 
@@
 So the tag with the highest frequency (\emph{C\#}) would have the highest base-level activation.
-Tag frequency drops off sharply for this dataset.
+
+\subsubsection{Distribution of Tag Frequency}
+
+\emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
+Tag frequency then drops off sharply.
 The top 1\% of tags based on frequency of use account for approximately 60\% of total tag instances.
@@
 
-Attentional weighting ($W_{j}$), can be thought of as the amount of attentional resources that are dedicated to contextual element $j$.
+Attentional weighting ($W_{j}$) can be thought of as the amount of attentional resources that are dedicated to contextual element $j$.
 Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that attentional resources are limited.
@@
 However, this still seemed problematic for two reasons:
-First, because stop words like ``the'' are observed with all tags, and not highly predictive of any tag, one could argue that their presence is simply adding error in the model's tag activation calculation.
+First, because stop words like ``the'' are observed with all tags and not highly predictive of any tag, one could argue that their presence is simply adding error in the model's tag activation calculation.
 Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words like ``the'' and ``?'' are taken away from being used for highly predictive words like ``PHP''.
@@
 This means that a stop word like ``the'' will have a scaled entropy measure close to 0, while a highly-predictive word like \emph{macro} will have a scaled entropy measure close to 1.
-Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate the limited attentional resources to predictive contextual elements.
+Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate limited attentional resources to predictive contextual elements.
 
@@
 It was recommended to remove stop words from the analysis \cite{Bird2009}.
-Stop words are low-predictor words, and \citeauthor listed a short list that were commonly removed (e.g., ``the'', ``and'', ``or'').
+Stop words are low-predictor words, and \citeauthor{Bird2009} listed a short list that were commonly removed (e.g., ``the'', ``and'', ``or'').
 However it seemed somewhat arbitrary to explicitly list out a few stop words and then simply choose those as the words to remove.
@@
 using scaled entropy to determine stop words seems to work fairly well.
-So instead of explicitly removing an enumerated list of stop words, we used a data-driven approach and let the scaled entropy measure attenuate the words that were not predictive of any tag.
+So instead of explicitly removing an enumerated list of stop words, we used a data-driven approach and let the scaled entropy measure attenuate the words that were not predictive of any tag to near-zero levels.
 
@@
 
-An offset measurement (mean of the top five tags for the post) was used to equalize the top activated tags for each post.
+An offset measurement (mean of the top five tags for the post) was used to equalize the top activated tags across posts.
 This was done primarily to improve model fit.
@@
 There are \num{39223968} unique post word, tag word combinations.
-The result is a table of co-occurrence counts for each of the combinations, which was loaded into R.
+The result was a table of co-occurrence counts for each of the combinations, which was loaded into R.
 
@@
 Model tag predictions are also based on the frequency that each tag has been used.
-More often used tags more likely to have higher activation.
+More often used tags have a higher base-level activation.
 A SQL query was used to build a similar table of tag use counts for the \num{26176} unique tags, which was loaded into R.
-This set for example would show that \emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
 
@@
 So model performance can be measured by comparing the model's tag activations with chosen tags for a post, and then aggregated across posts to get an overall average performance.
-If the model consistently produces higher tag activations for chosen tags than non-chosen tags, the model should be able to accurately predict the chosen tags.
 
@@
 Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.46$, npv = .997, ppv = .70).
-It's also interesting to note that although the attentional weighting terms ($W_{title}, W_{body}$) were calibrated by the regression, they did not stray far from cognitively-plausible defaults
+It's also interesting to note that although the attentional weighting ($W$) terms for the title and body were calibrated by the regression, they did not stray far from cognitively-plausible defaults
 (see coefficients for title and body in Table \ref{tab:coeffs}).
@@
 Our StackOverflow model is a successful case showing that ACT-Rs declarative memory retrieval mechanisms can be scaled up and applied to retrieval requests occurring across large-scale declarative memory datasets.
-The model can successfully tag 67/100 posts correctly, when tasked to predict one tag per post on average.
+The model can successfully tag 67\% of posts correctly, when tasked to predict one tag per post on average.
 The model is also computationally space efficient:
@@
 If a post author does not include the model's predicted best tag for the post, then the model could suggest the tag to the author.
-This could potentially introduce authors to tags that are contextually relevant to their interests that they may not have known about.
+This could potentially introduce authors to tags that are contextually relevant to their interests that they may not have been aware of.
 This could also improve the answers received for the post, since other expert programmers may have only been monitoring the model's predicted best tag for the post.
@@
 In this scenario the tags for a previously submitted post would be compared against model predictions.
-If the model's predicted tag activation for a used tag is substantially low, then that tag could either be removed from the post, or the post could be flagged for further analysis by an operator.
+If the model's predicted tag activation for a used tag is substantially low, then that tag could either be removed from the post, or the post could be flagged for further analysis.
 

### 2013-02-26 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 116c33b684092a62fa118d2803c083a0d275c64f [formerly cfc0a3e7d56f54495f8848cddb814753432602d9]
Former-commit-id: 2685ecfe70f9c2f40ce11e521a6c5fe4777d2c23
foo
Former-commit-id: 59b7539dd1c5d85f229c5c2e4f9d2c4f4ac15d25 [formerly 23eabfa0fd685faaca2f94633c9e4e0fbdf185a9]
Former-commit-id: bb1bd5d62f30f31e88df90e00e119f531810ce1c

@@
 \usepackage[none]{hyphenat}
+\usepackage[compact]{titlesec}
 
@@
 Human-based tagging of online content has increased in recent years.
-A few examples of human-based tagging of information are hashtags for Twitter tweets and Facebook posts, as well as keywords for research papers.
+A few examples are authors creating hashtags for Twitter tweets and Facebook posts, as well as keywords for research papers.
 A large portion of the growing number of social media sites support some sort of human-directed tagging of posts.
@@
 
-There is a growing body of research that aims to characterize the lifetime and growth characteristics of hashtags for real-time social networking sites such as Twitter
+A growing body of research is focusing on the lifetime and growth characteristics of hashtags for real-time social networking sites such as Twitter
 (e.g., \citeNP{Tsur2012, Chang2010}).
@@
 
-The Apr' 11 StackOverflow dataset \cite{DataDump2011} was used for this analysis.
-The dataset contains \num{4945168} posts from \num{559803} unique users and \num{26176} unique tags.
-The dataset is the entire collection of posts dated between Jul' 2008 (StackOverflow's creation) and Apr' 11.
-Post authors used approximately 3 tags on average per post.
-
 \subsection{Related Work}
@@
 Table \ref{sample-table} includes a formal description of the model.
+
+\vspace{-1em}
 
@@
 
-\subsubsection{Distribution of Tag Frequency}
-
-\emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
-Tag frequency then drops off sharply.
-The top 1\% of tags based on frequency of use account for approximately 60\% of total tag instances.
-This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.
-
 \subsubsection{Contextual Activation}
@@
 First, because stop words like ``the'' are observed with all tags and not highly predictive of any tag, one could argue that their presence is simply adding error in the model's tag activation calculation.
-Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words like ``the'' and ``?'' are taken away from being used for highly predictive words like ``PHP''.
+Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words are taken away from being used for highly predictive words.
 
@@
 Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate limited attentional resources to predictive contextual elements.
-
-\subsubsection{Entropy and Stop Words}
-
-It was recommended to remove stop words from the analysis \cite{Bird2009}.
-Stop words are low-predictor words, and \citeauthor{Bird2009} listed a short list that were commonly removed (e.g., ``the'', ``and'', ``or'').
-However it seemed somewhat arbitrary to explicitly list out a few stop words and then simply choose those as the words to remove.
-Looking at the scaled entropy measure, it seems quite plausible that this measure of word predictiveness can also be used to identify stop words.
-That is, words with scaled entropy close to zero are the stop words.
-Looking at the lowest ten scaled-entropy words for this dataset, (``does'', ``anyone'', ``'ve'', ``found'', ``are'', ``seems'', ``has'', ``been'', ``for'', ``support''),
-using scaled entropy to determine stop words seems to work fairly well.
-So instead of explicitly removing an enumerated list of stop words, we used a data-driven approach and let the scaled entropy measure attenuate the words that were not predictive of any tag to near-zero levels.
 
@@
 
+\subsubsection{Dataset}
+
+The Apr' 11 StackOverflow dataset \cite{DataDump2011} was used for this analysis.
+The dataset contains \num{4945168} posts from \num{559803} unique users and \num{26176} unique tags.
+The dataset is the entire collection of posts dated between Jul' 2008 (StackOverflow's creation) and Apr' 11.
+Post authors used approximately 3 tags on average per post.
+
+\subsubsection{Distribution of Tag Frequency}
+\emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
+Tag frequency then drops off sharply.
+The top 1\% of tags based on frequency of use account for approximately 60\% of total tag instances.
+This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.
+
 \subsection{Generating Model Predictions}
@@
 The tags were already chunked, but tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
-Afterwords the processed word chunks and root tags were imported into a PostgreSQL database table.
+Afterwords the processed word chunks and root tags were imported into a PostgreSQL table.
 
@@
 A SQL query was used to build a similar table of tag use counts for the \num{26176} unique tags, which was loaded into R.
+
+\subsubsection{Entropy and Stop Words}
+
+It was recommended to remove stop words from the analysis \cite{Bird2009}.
+Stop words are low-predictor words, and \citeauthor{Bird2009} listed a short list that were commonly removed (e.g., ``the'', ``and'', ``or'').
+However it seemed somewhat arbitrary to explicitly list out a few stop words and then simply choose those as the words to remove.
+Looking at the scaled entropy measure, it seems quite plausible that this measure of word predictiveness can also be used to identify stop words.
+That is, words with scaled entropy close to zero are the stop words.
+Looking at the lowest ten scaled-entropy words for this dataset, (``does'', ``anyone'', ``'ve'', ``found'', ``are'', ``seems'', ``has'', ``been'', ``for'', ``support''),
+using scaled entropy to determine stop words seems to work fairly well.
+So instead of explicitly removing an enumerated list of stop words, we used a data-driven approach and let the scaled entropy measure attenuate the words that were not predictive of any tag to near-zero levels.
 
@@
 and both global and local weightings are commonly used when the PMI measurement is used.
-This suggests that using a global weighting term for chunks in context (e.g., the scaled entropy measure used here) alongside the more commonly-used local weighting (spreading activation) might be useful for other ACT-R models that are taking advantage of large declarative memory datasets to build up the model's base knowledge.
+This suggests that using a global weighting term for chunks in context (e.g., the scaled entropy measure used here)
+alongside the more commonly-used local weighting (spreading activation) might be useful for other ACT-R models that are taking advantage of large declarative memory datasets to build up the model's base knowledge.
+
+\subsection{Current Model Issues}
+
+Including the offset term in the model improved performance (see Figure \ref{fig:ROC}).
+However by normalizing the top activations for all posts, the model cannot differentiate between a post where the top predicted tag is certain, and a post where the top-predicted tag is questionable.
+One could argue that normalizing top activations for all posts should cause model performance to decrease, and the opposite was actually observed.
+One hypothesis is that this normalization is required because the retrieval threshold of a tag is adjusted for each post, and changed so that only a small set of tags are retrieved for all posts.
+Further research will explore this and other theories for the behavior of the offset component in the model.
 
 \subsection{Future Planned Research}
+
+Alongside exploring the offset component of the model, our main planned research is to include the post author's past tagging behavior as an additional model predictor.
+We are exploring a Bayesian updating technique that would adjust prior tag activations from global priors to user's priors as more tagging information about the user is collected.
+For the StackOverflow dataset, adding user's tagging history to the model should drastically improve predictive power for veteran users,
+since tags are tied to programming languages and specific technologies, and programmers tend to specialize on a few set of tags that are of interest to them.
+This technique could also be deployed, since a post author must be signed into StackOverflow before creating a new post.
 

### 2013-03-03 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 5577bc7440b4cdc3a2995c2d99824a2f05f8eca4 [formerly 7045503a4c06b9a3ee9bfeef87f713389a7d1819]
Former-commit-id: 40dda42e1b4ef1baec8497b9fa7b52c6c504df88

@@
 \begin{abstract}
-  A model was built for the StackOverflow dataset that can predict the tags that will be used by the author of the post, given the content of the post.
-  The model is a cognitively-inspired Bayesian probabilistic model based on ACT-R's declarative memory retrieval mechanisms.
+  Large-scale rich behavioral datasets such as Twitter, Facebook, LinkedIn, StackOverflow, and Yelp can be mined and explored to study trends of human behavior and human computer interaction.
+  We focused on hashtag creation as a specific human behavior.
+  We explored the StackOverflow dataset, and developed an ACT-R inspired Bayesian propabilistic model that can predict the hashtags used by the author of the post.
   Large-scale relational database techniques and efficient sparse matrix data structures were used for implementation,
-  so that the data structures require only 3GB of memory and retrieval requests with a DM dataset of \num{556677795} chunks occur in less than one second on a single processor.
-  A logistic regression technique was used to calibrate model parameters and measure performance.
-  Model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.46$, npv = .997, ppv = .70) and model performance on the test dataset is fairly strong (67\% accuracy when asked to tag one tag per post on average).
+  so that the data structures require only 3GB of memory and retrieval requests with a DM dataset of hundreds of millions of chunks occur in less than one second on a single processor.
+  The model is a succesful case showing that ACT-R's declarative memory retrieval equations scale, and are relevant to task domains that require large-scale knowledge stores.
 
   \textbf{Keywords:}
-  Machine Learning; ACT-R; Large-Scale Declarative Memory
+  Machine Learning; Data Mining; ACT-R; Large-Scale Declarative Memory
 \end{abstract}
@@
 
-During initial model testing, the attentional weighting for all words was equal, and set to $W/n$ for $n$ title words (and set similarly for body words).
-The problem with this approach is that familiar stop words like ``the'' and ``?'', that are often observed across all tags have the same attentional weighting as highly predictive words like ``PHP''.
-Granted the $S_{ji}$'s across all tags for context words like ``the'' are much lower (and more evenly distributed) than the few $S_{ji}$'s that spike for ``PHP'' (e.g., the tag \emph{PHP}).
+During initial model testing, the attentional weighting for all words was equal and set to $W/n$ for $n$ title words (set similarly for body words).
+The problem with this approach is that words like ``the'' and ``?'' that are often observed across all tags (i.e., ``stop words'') have the same attentional weighting as highly predictive words like ``PHP''.
+Granted the $S_{ji}$'s across all tags for stop words like ``the'' are much lower (and more evenly distributed) than the few $S_{ji}$'s that spike for ``PHP'' (e.g., the tag \emph{PHP}).
 However, this still seemed problematic for two reasons:
-First, because stop words like ``the'' are observed with all tags and not highly predictive of any tag, one could argue that their presence is simply adding error in the model's tag activation calculation.
-Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words are taken away from being used for highly predictive words.
+First, one could argue that stop words are simply adding error in the model's tag activation calculation because they are observed evenly across all tags.
+Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words are unavailable for highly predictive words.
 
 It's also the case that there are many commonalities between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
-and many of those theories use a combination of weighting schemes for contextual relevance and global importance of a particular chunk in context.
+and many of those theories use a combination of weighting schemes for contextual relevance and global importance of a particular chunk.
 For example, \citeA{Dumais1991} shows that a measure of how likely term $j$ is in document $i$ is often the product of a local weighting $L_{ji}$ and global weighting $G_{j}$ measure.
-This idea maps rather nicely to ACT-R's contextual weighting.
-That is, terms are context words $j$, documents are tags $i$, local weighting is strength of association $S_{ji}$, and global weighting is attentional weighting $W_{j}$.
+This idea maps cleanly to ACT-R's equations: 
+Terms are context words $j$, documents are tags $i$, local weighting is strength of association $S_{ji}$, and global weighting is attentional weight $W_{j}$.
 
@@
 It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
-The scaled entropy ($E_{j}$) of a context word can be thought of as the amount of predictiveness that that word has to any tag.
+So the scaled entropy ($E_{j}$) of a context word can be thought of as the amount of predictiveness that that word has to \emph{any} tag.
 This means that a stop word like ``the'' will have a scaled entropy measure close to 0, while a highly-predictive word like \emph{macro} will have a scaled entropy measure close to 1.
-Going back to ACT-R's attentional weighting $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate limited attentional resources to predictive contextual elements.
+Looking at $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate limited attentional resources to predictive contextual elements.
 
@@
 The Apr' 11 StackOverflow dataset \cite{DataDump2011} was used for this analysis.
-The dataset contains \num{4945168} posts from \num{559803} unique users and \num{26176} unique tags.
-The dataset is the entire collection of posts dated between Jul' 2008 (StackOverflow's creation) and Apr' 11.
+This is the entire collection of posts dated between Jul' 2008 (StackOverflow's creation) and Apr' 11.
+It contains \num{4945168} posts from \num{559803} unique users and \num{26176} unique tags.
 Post authors used approximately 3 tags on average per post.
@@
 
-Model tag predictions are based on the amount of word co-occurrences between words in the title and body of the post and tags used for that post.
+Model tag predictions are based on the amount of word co-occurrences between words in the post and tags used for that post.
 To build the co-occurrence matrix between post words and tags, all title and body text and associated tags were extracted from the relational database.
@@
 
-It was recommended to remove stop words from the analysis \cite{Bird2009}.
-Stop words are low-predictor words, and \citeauthor{Bird2009} listed a short list that were commonly removed (e.g., ``the'', ``and'', ``or'').
-However it seemed somewhat arbitrary to explicitly list out a few stop words and then simply choose those as the words to remove.
+\citeA{Bird2009} recommended to remove the low-predictor stop words from the analysis.
+\citeauthor{Bird2009} provided a short list that were commonly removed (e.g., ``the'', ``and'', ``or'').
+However it seemed somewhat arbitrary to remove an explicitly-enumerated list of stop words.
 Looking at the scaled entropy measure, it seems quite plausible that this measure of word predictiveness can also be used to identify stop words.
@@
 Looking at the lowest ten scaled-entropy words for this dataset, (``does'', ``anyone'', ``'ve'', ``found'', ``are'', ``seems'', ``has'', ``been'', ``for'', ``support''),
-using scaled entropy to determine stop words seems to work fairly well.
+using scaled entropy to determine stop words works fairly well.
 So instead of explicitly removing an enumerated list of stop words, we used a data-driven approach and let the scaled entropy measure attenuate the words that were not predictive of any tag to near-zero levels.
@@
 These vectors of tag activation, categorization were collected across a set of \num{1000} fresh posts that were not used to train the model on word association strength.
-So a form of cross validation was used to measure model fit, although the more robust full k-fold cross validation technique was not used (due mostly to resource and time limitations).
-
 The parameters were optimized to best predict categorization (chosen tag, not chosen tag) from the four predictor variables:
 The tag's prior occurrence frequency, word co-occurrence activation for title and body words, and activation offset for the post.
+
+However using a single run of the \num{1000} posts was insufficient because there are two model components that require prior estimates of coefficients:
+The offset measure to determine the top-five tags and associated activation and the sampling technique used to pick a representative set of tags for each post.
+So model calibration was bootstrapped, where the coefficients from the previous run were used as estimates of coefficients for the next run.
+The coefficients converged on best-fit values quickly (only required two sets of four iterations), and did not show any signs of instability.
 
@@
 \subsection{Logistic Regression}
-
-In order to formally test model performance and to calibrate model parameters, a calibration run using \num{1000} posts was performed.
-For the calibration run, the category prediction (chosen or not chosen tag for a post) was regressed on base-level activation, spreading activation for title and body, and post offset.
-However using a single run of the \num{1000} posts was insufficient because there are two model components that require prior estimates of coefficients:
-The offset measure to determine the top-five tags and associated activation and the sampling technique used to pick a representative set of tags for each post.
-So model calibration was bootstrapped, where the coefficients from the previous run were used as estimates of coefficients for the next run.
-The coefficients converged on best-fit values quickly (only required two sets of four iterations), and did not show any signs of instability.
 
@@
 \end{table}
-
-\vspace{-2em}
 
@@
 Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.46$, npv = .997, ppv = .70).
-It's also interesting to note that although the attentional weighting ($W$) terms for the title and body were calibrated by the regression, they did not stray far from cognitively-plausible defaults
+It's also interesting that although the attentional weighting ($W$) terms for the title and body were calibrated by the regression, they did not stray far from cognitively-plausible defaults
 (see coefficients for title and body in Table \ref{tab:coeffs}).
@@
 
-In order to test how well the model (and calibrated parameters) generalize to the rest of the StackOverflow dataset, the model was tested on an additional set of \num{1000} posts.
-These were fresh posts, and not previously used to either train the strength of association matrix or calibrate model parameters.
-The model parameters from the calibrated dataset were held constant and used on the test dataset.
+In order to test how well the model (and calibrated parameters) generalize to the rest of the StackOverflow dataset,
+model parameters were held constant and the model was tested on a fresh set of \num{1000} posts.
 Model classification accuracy as a function of number of tags predicted by the model was used as the metric to test if the model generalized.
@@
 
-A plot of model performance for both the calibration and test dataset is included in Figure \ref{fig:ROC}.
+A plot of model performance for both the calibration and test runs is included in Figure \ref{fig:ROC}.
 
@@
 
-Our StackOverflow model is a successful case showing that ACT-Rs declarative memory retrieval mechanisms can be scaled up and applied to retrieval requests occurring across large-scale declarative memory datasets.
+Our StackOverflow model is a successful case showing that ACT-Rs declarative memory retrieval mechanisms scale, and can be applied to retrieval requests occurring across large-scale declarative memory datasets.
 The model can successfully tag 67\% of posts correctly, when tasked to predict one tag per post on average.
@@
 If a post author does not include the model's predicted best tag for the post, then the model could suggest the tag to the author.
-This could potentially introduce authors to tags that are contextually relevant to their interests that they may not have been aware of.
+This would provide a way to introduce authors to tags that are contextually relevant to their interests that they may not have been aware of.
 This could also improve the answers received for the post, since other expert programmers may have only been monitoring the model's predicted best tag for the post.

### 2013-03-03 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 3213e7b268d556d9b3d3875735f0ea7e3956a490 [formerly d53aa6d174d1a5acbeb2c9cec4b7209d40fab881]
Former-commit-id: 0bf54a4eb3a6bc0357fd06bfd13d6cce3fd6379a

@@
   We focused on hashtag creation as a specific human behavior.
-  We explored the StackOverflow dataset, and developed an ACT-R inspired Bayesian propabilistic model that can predict the hashtags used by the author of the post.
+  We explored the StackOverflow dataset, and developed an ACT-R inspired Bayesian probabilistic model that can predict the hashtags used by the author of the post.
   Large-scale relational database techniques and efficient sparse matrix data structures were used for implementation,
   so that the data structures require only 3GB of memory and retrieval requests with a DM dataset of hundreds of millions of chunks occur in less than one second on a single processor.
-  The model is a succesful case showing that ACT-R's declarative memory retrieval equations scale, and are relevant to task domains that require large-scale knowledge stores.
+  The model is a successful case showing that ACT-R's declarative memory retrieval equations scale, and are relevant to task domains that require large-scale knowledge stores.
 

### 2013-03-03 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 069494f48b98c0cff6aeeb40d494a48656b3fe73 [formerly 319a1a8924d8dc79d32072444964c0d7197e7332]
Former-commit-id: afd8f2b8763025de1c3cd5eba0e30dc28c7f0f46

@@
 \begin{abstract}
-  Large-scale rich behavioral datasets such as Twitter, Facebook, LinkedIn, StackOverflow, and Yelp can be mined and explored to study trends of human behavior and human computer interaction.
+  Large-scale behavioral datasets such as Twitter, Facebook, LinkedIn, StackOverflow, and Yelp can be mined and explored to study trends of human behavior and human computer interaction.
   We focused on hashtag creation as a specific human behavior.
@@
 Looking at the scaled entropy measure, it seems quite plausible that this measure of word predictiveness can also be used to identify stop words.
-That is, words with scaled entropy close to zero are the stop words.
+That is, words with scaled entropy close to zero \emph{are} the stop words.
 Looking at the lowest ten scaled-entropy words for this dataset, (``does'', ``anyone'', ``'ve'', ``found'', ``are'', ``seems'', ``has'', ``been'', ``for'', ``support''),

### 2013-03-04 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: cda8ce26309171ed784c21c2f7d3395ac480e842 [formerly 38129b4ae21918ab8eb0b3e3d8cec260a87cac1f]
Former-commit-id: 0517f534b694a2dc34e6fc07815156165b610a5f

@@
 
-Our StackOverflow model is a successful case showing that ACT-Rs declarative memory retrieval mechanisms scale, and can be applied to retrieval requests occurring across large-scale declarative memory datasets.
+Our StackOverflow model is a successful case showing that ACT-R's declarative memory retrieval mechanisms scale, and can be applied to retrieval requests occurring across large-scale declarative memory datasets.
 The model can successfully tag 67\% of posts correctly, when tasked to predict one tag per post on average.

### 2013-03-16 stackExchange/analysis/model/analyzer-spring.R
Final code for ICCM writeup
Former-commit-id: a729cd466a87a639b5784b9e9581b7df5c76c39e [formerly 99c3b0af02c9c9f109fc0dec76721b09edb1d669]
Former-commit-id: 90e6a003e201e66a66ffb8da3e2079bd2b100c53

@@ -61,7 +61,12 @@ combinePs = function(psGlobal, psLocal, count) {
 	(psGlobal*(maxCount-count+1) + psLocal*count) / (maxCount + 1)
 }
 
-analyzeForICCM = function(frms=getAllFrms()) {
+analyzeForICCM = function() {
+	runSet(sets=8, id=1)
+	runSet(sets=9, id=1)
+	prevFrm = getFrms(8,1)[[1]]
+	runFromPrevious(prevFrm, 8, 2)
+	frms = getAllFrms()
 	plotAllFrms(frms=frms)
 	tagFiles=getTagFiles(makeTagDir(8))
 	visPost(tagFiles[105], makeTagDir(8), coeffs=frms[[1]]$coeffs)
@@ -74,9 +79,9 @@ makeMultivariateROC = function(baseFrms, figName, legendText) {
 }
 
 analyzeBaseFrmForMultivariate = function(baseFrm) {
-	baseFrms = adjustFrms(list(baseFrm), coeffsGlobal, model=formula(targetP ~ prior + userIdLogPriors + sjiTitle + sjiBody + offset + 0))
-	baseFrms[2] = adjustFrms(list(baseFrms[[1]]$frm), coeffsGlobal, model=formula(targetP ~ prior + sjiTitle + sjiBody + offset + 0))
-	baseFrms[3] = adjustFrms(list(baseFrms[[1]]$frm), coeffsGlobal, model=formula(targetP ~ combinedPrior + sjiTitle + sjiBody + offset + 0))
+	baseFrms = adjustFrms(list(baseFrm), coeffsGlobal, model=formula(targetP ~ prior + userIdLogPriors + sjiTitle + sjiBody + offset))
+	baseFrms[2] = adjustFrms(list(baseFrms[[1]]$frm), coeffsGlobal, model=formula(targetP ~ prior + sjiTitle + sjiBody + offset))
+	baseFrms[3] = adjustFrms(list(baseFrms[[1]]$frm), coeffsGlobal, model=formula(targetP ~ combinedPrior + sjiTitle + sjiBody + offset))
 	#tempFrm = baseFrms[[1]]$frm
 	#tempFrm = subset(tempFrm, userIdTagCount > 500)
 	#tempFrm$userPrior[is.infinite(tempFrm$userPrior)] = -10000
@@ -92,10 +97,31 @@ getCoeffs = function(baseFrms) {
[…]

### 2013-03-16 stackExchange/analysis/model/analyzer.R

@@ -112,9 +112,11 @@ adjustFrms = function(frms, coeffs, model) {
 	for(i in c(1:4)) {
 		res = runLogRegFrm(frms[[1]], coeffs, model)
 		coeffs = res$coeffs
+		classLog = res$classLog
+		logit = summary(res$logit)
 	}
 	frms = lapply(frms, function(frm) recomputeActivation(frm, coeffs))
-	return(lapply(frms, function(frm) list(frm=frm, coeffs=coeffs)))
+	return(lapply(frms, function(frm) list(frm=frm, coeffs=coeffs, classLog=classLog, logit=logit)))
 }
 
 getAllFrms = function() {
@@ -208,15 +210,17 @@ getAccuracyAtNAverageTags = function(frm, N=1) {
 plotHighest = function(contextIndeces, vals) {
 	vals = lapply(vals, function(val) {val[contextIndeces]})
 	frm = data.frame(vals)
-	topNum = 10
+	topNum = 12
 	vals = rowSums(frm)
 	res = sort(vals, decreasing=T, index.return=T)
 	sortedChunkHashes = contextIndeces[res$ix]
 	x = sortedChunkHashes[1:topNum]
 	xnames = getChunks(x, dbPriors)
-	if(sum(colnames(frm) == "prior") > 0) {
-		priorOffset = min(frm$prior[x])
-		frm$prior = frm$prior - priorOffset	
+	for (name in colnames(frm)) {
+		offset = min(frm[[name]][x])
+		if (offset < 0) {
+			frm[[name]] = frm[[name]] - offset
+		}
 	}
 	par(mar=c(10,4.1,4.1,2.1))
 	barplot(t(frm[x,]), legend=colnames(frm), names.arg=xnames, las=3)

### 2013-03-16 stackExchange/analysis/model/driver.R

@@ -20,18 +20,8 @@ library(data.table)
 library(plyr)
 
 # A few helper functions
-rateVals3 = function(priorsIndeces, act, observed, tagFile) {
-	cutoff = 20
-	vals = as.vector(B[priorsIndeces]*coeffsGlobal$prior 
-		+ act$sjiTitle[priorsIndeces]*coeffsGlobal$sjiTitle 
-		+ act$sjiBody[priorsIndeces])*coeffsGlobal$sjiBody
-	res = sort(vals, decreasing=T, index.return=T)
-	sortedChunkHashes = priorsIndeces[res$ix]
-	#sortedChunkHashes = sortedChunkHashes[1:cutoff]
-	#sampleIndeces = union(union(observed, sortedChunkHashes), sample(priorsIndeces, 100, replace=F))
-	#sampleIndeces = union(sortedChunkHashes[1:cutoff], union(observed, sample(priorsIndeces, 500, replace=F, prob=vals)))
-	probs = 1/(1+exp(-vals/.5))
-	sampleIndeces = union(sortedChunkHashes[1:cutoff], union(observed, sample(priorsIndeces, 400, replace=F)))
+
+rateValsInner = function(sampleIndeces, act, observed, tagFile) {
 	targetP = rep(0, length(sampleIndeces))
 	targetP[match(observed, sampleIndeces)] = 1
 	tags = getChunks(sampleIndeces, dbPriors)
@@ -43,6 +33,21 @@ rateVals3 = function(priorsIndeces, act, observed, tagFile) {
 	userIdPriors = userIdPriors[sampleIndeces]
 	return(data.frame(tag=tags, sjiTitle=sjiTitle, sjiBody=sjiBody, prior=priors, targetP=targetP, act=priors + sjiTitle + sjiBody,
 		userIdPriors=userIdPriors, userIdTagCount = userIdTagCount))
+}
+
+rateVals3 = function(priorsIndeces, act, observed, tagFile) {
+	cutoff = 20
+	vals = as.vector(B[priorsIndeces]*coeffsGlobal$prior 
+		+ act$sjiTitle[priorsIndeces]*coeffsGlobal$sjiTitle 
+		+ act$sjiBody[priorsIndeces])*coeffsGlobal$sjiBody
+	res = sort(vals, decreasing=T, index.return=T)
+	sortedChunkHashes = priorsIndeces[res$ix]
[…]

### 2013-03-16 stackExchange/analysis/model/model.R

@@ -73,6 +73,7 @@ getPriorsForPost = function(PostId) {
 # Calculate total activation, given base-level activation, sji associations, and context
 act = function(context, B, sji) {
 	weights = contextWeights[context]
+	#weights = rep(1, length(weights))
 	print(sum(weights))
 	if( sum(weights) > 0) {
 		weights = weights / sum(weights)

### 2013-03-16 stackExchange/analysis/model/driver.R
foo
Former-commit-id: 2767ca1a1615b28519b608f9731d9ff351197bbf [formerly f769f7a787fc6dd45ac1629222352303862f7cda]
Former-commit-id: 7e7c27f7d5fe996f42ab1f18419193dd7eb2ff14

@@ -88,7 +88,7 @@ ratePost = function(tagFile, tagDir) {
 }
 
 ratePost2 = function(tagFile, tagDir) {
-	subsetIndeces = getChunkHashes(dFrameGlobal[dFrameGlobal$tagFile == tagFile,]$tag, dbPriors)
+	subsetIndeces = getChunkHashes(frmGlobal[frmGlobal$tagFile == tagFile,]$tag, dbPriors)
 	ratePostInner(tagFile, tagDir, subsetIndeces, rateValsFun=rateVals4)
 }
 
@@ -138,8 +138,8 @@ runSet = function(sets=c(8:17), id=1) {
 	lapply(sets, function(set) {writePosts(ratePosts(set), set, getSubsetId(sjiCSV), id)})
 }
 
-runFromPrevious = function(previousDFrame, set, id) {
-	dFrameGlobal <<- previousDFrame
+runFromPrevious = function(prevFrm, set, id) {
+	frmGlobal <<- prevFrm
 	writePosts(ratePosts(set, ratePostFun=ratePost2), set, getSubsetId(sjiCSV), id)
 }
 

### 2013-03-16 stackExchange/analysis/model/driver.R
foo
Former-commit-id: 0c5fa71ba9073beaf75f1197d409686830d12a59 [formerly 36e6890f2b3154063d2ac73f13afdf99367ca250]
Former-commit-id: f359185cad33300fe8bafb3f1f5dcc20b1c15e0f

@@ -35,7 +35,7 @@ rateValsInner = function(sampleIndeces, act, observed, tagFile) {
 		userIdPriors=userIdPriors, userIdTagCount = userIdTagCount))
 }
 
-rateVals3 = function(priorsIndeces, act, observed, tagFile) {
+rateValsSample = function(priorsIndeces, act, observed, tagFile) {
 	cutoff = 20
 	vals = as.vector(B[priorsIndeces]*coeffsGlobal$prior 
 		+ act$sjiTitle[priorsIndeces]*coeffsGlobal$sjiTitle 
@@ -46,7 +46,7 @@ rateVals3 = function(priorsIndeces, act, observed, tagFile) {
 	rateValsInner(sampleIndeces, act, observed, tagFile)
 }
 
-rateVals4 = function(priorsIndeces, act, observed, tagFile) {
+rateValsNoSample = function(priorsIndeces, act, observed, tagFile) {
 	rateValsInner(priorsIndeces, act, observed, tagFile)
 }
 
@@ -64,7 +64,7 @@ getObservedContext = function(tagFile, tagDir) {
 	return(ret)
 }
 
-ratePostInner = function(tagFile, tagDir, indeces, rateValsFun=rateVals3) {
+ratePostInner = function(tagFile, tagDir, indeces, rateValsFun=rateValsSample) {
 	res = data.frame()
 	if ( length(lst <- getObservedContext(tagFile, tagDir)) > 0 ) {
 		observed = lst$observed
@@ -89,7 +89,7 @@ ratePost = function(tagFile, tagDir) {
 
 ratePost2 = function(tagFile, tagDir) {
 	subsetIndeces = getChunkHashes(frmGlobal[frmGlobal$tagFile == tagFile,]$tag, dbPriors)
-	ratePostInner(tagFile, tagDir, subsetIndeces, rateValsFun=rateVals4)
+	ratePostInner(tagFile, tagDir, subsetIndeces, rateValsFun=rateValsNoSample)
 }
 
 getTagFiles = function(tagDir) {

### 2013-03-16 stackExchange/analysis/model/driver.R
foo
Former-commit-id: d42b373ee0f94cf25d51b0eac64070524c1d7882 [formerly 2f48d63aca7a560350492e6ef0d5e9d86ee9042b]
Former-commit-id: ea7065a70440d04bc80bed01f4ba5af967cc8b95

@@ -83,11 +83,11 @@ ratePostInner = function(tagFile, tagDir, indeces, rateValsFun=rateValsSample) {
 	return(res)
 }
 
-ratePost = function(tagFile, tagDir) {
+ratePostSample = function(tagFile, tagDir) {
 	ratePostInner(tagFile, tagDir, priorsIndeces)
 }
 
-ratePost2 = function(tagFile, tagDir) {
+ratePostNoSample = function(tagFile, tagDir) {
 	subsetIndeces = getChunkHashes(frmGlobal[frmGlobal$tagFile == tagFile,]$tag, dbPriors)
 	ratePostInner(tagFile, tagDir, subsetIndeces, rateValsFun=rateValsNoSample)
 }
@@ -122,7 +122,7 @@ getPostIdForTagFile = function(tagFile) {
 	as.integer(str_extract(tagFile, "[^/][0-9]+$"))
 }
 
-ratePosts = function(subsetId, ratePostFun=ratePost) {
+ratePosts = function(subsetId, ratePostFun=ratePostSample) {
 	tagDir = makeTagDir(subsetId)
 	tagFiles = getTagFiles(tagDir)
 	#res = rbind.fill(parallel::mclapply(tagFiles, ratePost, mc.cores=3, mc.preschedule=T))
@@ -140,7 +140,7 @@ runSet = function(sets=c(8:17), id=1) {
 
 runFromPrevious = function(prevFrm, set, id) {
 	frmGlobal <<- prevFrm
-	writePosts(ratePosts(set, ratePostFun=ratePost2), set, getSubsetId(sjiCSV), id)
+	writePosts(ratePosts(set, ratePostFun=ratePostNoSample), set, getSubsetId(sjiCSV), id)
 }
 
 #coeffsGlobal=list(sjiTitle=1.40, sjiBody=2.36, prior=1.08)

### 2013-03-16 stackExchange/analysis/model/driver.R
foo
Former-commit-id: f839fb4a6dafb6e91ba81a05fc7906b2b4be5597 [formerly b8aeaf802a86359d59c6ac0817ab98aa1727267c]
Former-commit-id: 25787d1005ec97936f11a5c3296758f2f1c332de

@@ -135,7 +135,7 @@ writePosts = function(res, testSubset, modelSubset, id) {
 }
 
 runSet = function(sets=c(8:17), id=1) {
-	lapply(sets, function(set) {writePosts(ratePosts(set), set, getSubsetId(sjiCSV), id)})
+	lapply(sets, function(set) writePosts(ratePosts(set), set, getSubsetId(sjiCSV), id))
 }
 
 runFromPrevious = function(prevFrm, set, id) {

### 2013-03-16 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: e9b7d7119c57f5ea4b1297348812b4ecbad997b1 [formerly 2d41ae893dd27857a0818dbde814bb541663931b]
Former-commit-id: a8a406c08a2e0e688d052021d0b27ee7c5359b33
Final code for ICCM writeup
Former-commit-id: a729cd466a87a639b5784b9e9581b7df5c76c39e [formerly 99c3b0af02c9c9f109fc0dec76721b09edb1d669]
Former-commit-id: 90e6a003e201e66a66ffb8da3e2079bd2b100c53
r3
Former-commit-id: a86e3b29ee2c88ed1170102cd2ec0f2284bb8d38 [formerly 966bb44cb2c254e3ed0c87113a5e342a02548549]
Former-commit-id: 78abff161c5d0db8c4109845c411432aa42571f8

@@
 \begin{abstract}
-  Large-scale behavioral datasets such as Twitter, Facebook, LinkedIn, StackOverflow, and Yelp can be mined and explored to study trends of human behavior and human computer interaction.
-  We focused on hashtag creation as a specific human behavior.
-  We explored the StackOverflow dataset, and developed an ACT-R inspired Bayesian probabilistic model that can predict the hashtags used by the author of the post.
-  Large-scale relational database techniques and efficient sparse matrix data structures were used for implementation,
-  so that the data structures require only 3GB of memory and retrieval requests with a DM dataset of hundreds of millions of chunks occur in less than one second on a single processor.
+  Large-scale behavioral datasets such as Twitter, Facebook, LinkedIn, StackOverflow, and Yelp can be mined and explored to study human behavior and human computer interaction.
+  We focused on hashtag creation as a specific human behavior, since understanding and modeling hashtag creation can lead to improved human-computer systems that can better identify users' goals and interests.
+  We explored the StackOverflow dataset, and developed an ACT-R inspired Bayesian probabilistic model that can predict the hashtags used by the author of the post---predicting the users' programming interests.
+  The model is 67\% accurate when tasked to predict one tag per post on average.
+  This is achieved by choosing the tag that has the highest log odds of being correct,
+  given the tag's prior log odds of occurrence and adjusting for the log likelihood ratio of the words in the post being associated with the tag.
   The model is a successful case showing that ACT-R's declarative memory retrieval equations scale, and are relevant to task domains that require large-scale knowledge stores.
@@
   \textbf{Keywords:}
-  Machine Learning; Data Mining; ACT-R; Large-Scale Declarative Memory
+  Tagging; Large-Scale Semantic Memory; Data Mining; ACT-R
 \end{abstract}
@@
   \centering
-  \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-screen-cropped.pdf}}
+  \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-screen-crop.pdf}}
   \caption{Example post on stackoverflow.com}
@@
 
+The Apr' 11 dataset \cite{DataDump2011} was used for the analysis.
+It contains all posts dated between Jul' 08 (site creation) and Apr' 11.
+There are \num{1468485} posts from \num{559803} unique users and \num{26176} unique tags.
+Post authors used approximately 3 tags per post.
+
+\emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
+Tag frequency drops off sharply and follows Zipf's law.
+The top 1\% of tags ranked by frequency of use account for approximately 60\% of total tag instances.
+
 \subsection{Related Work}
@@
 Predicting tags for the StackOverflow dataset has been worked on by at least one prior researcher \cite{Kuo2011}.
-\citeauthor{Kuo2011} only looks at the Stack Overflow dataset briefly at the end of the technical report.
-However he does mention that his co-occurrence model has a 47\% classification accuracy for predicting StackOverflow tags when required to predict one tag per post.
-Our model achieves 67\% classification accuracy when predicting one tag per post on average.
+\citeauthor{Kuo2011} uses a co-occurrence model that predicts tags based on the words in the post and their relation (co-occurrence) to tags.
+The model was initially built for next-word prediction in large documents, and then adapted to the StackOverflow dataset by constraining the next word predicted to only tags.
+His co-occurrence model has a 47\% classification accuracy for predicting StackOverflow tags when the model predicts one tag per post.
+
+The SNIF-ACT \cite{Fu2007} model also uses co-occurrences and can predict link traversal for search queries.
+That is, given a search query (goal state), fetched results, and words included in those fetched results, the model predicts the most likely link that a person will click on.
+This StackOverflow model is similar because it leverages ACT-R's declarative memory retrieval mechanisms to generate predictions (tag activations instead of link activations).
+However, it is slightly different because it takes into account prior odds of tag activations which was not necessary in SNIF-ACT.
 
@@
 These equations describe the likelihood that a chunk from memory will be retrieved, given the prior odds of retrieving the chunk and the current context.
-In this way the task of retrieving the tag with the highest activation from the words in the post is similar to performing a declarative memory retrieval request for a tag, given the context of title and body words.
+So the task of retrieving the tag with the highest activation from the words in the post is similar to performing a declarative memory retrieval request for a tag, given the context of title and body words.
 The hypothesis is that the tags used for posts are based on a tag's history of prior use and the strength of association between the tag and the content of words in the title and body of the post.
-
-This model is also similar in design to the approach used by SNIF-ACT \cite{Fu2007}.
-SNIF-ACT can predict link traversal for search queries.
-That is, given a search query (goal state), fetched results, and words included in those fetched results, the model predicts the most likely link that a person will click on.
-This StackOverflow model is similar because it leverages ACT-R's declarative memory retrieval mechanisms to generate predictions (tag activations instead of link activations).
-However, it is slightly different because it takes into account prior odds of tag activations which wasn't necessary in SNIF-ACT.
-
-\subsection{Equations}
-
-Table \ref{sample-table} includes a formal description of the model.
 
@@
 
+Table \ref{sample-table} includes a formal description of the model.
 Here, the $i$ subscript denotes activation for a particular tag $i$, and the $j$ subscript is for context.
 $A_{i}$ is total activation for tag $i$.
-Note the similarities between the StackOverflow model and ACT-R's retrieval equations (base level and contextual component, log odds of prior being adjusted by association strength with context).
+
+Similar to SNIF-ACT \cite{Fu2007}, we interpret ACT-R's activation equation as a Bayesian prediction of the likelihood of retrieving a tag, given the tag's prior and updated by context.
+That is, the log odds of retrieving a particular chunk or tag ($A_{i}$) is determined by adjusting the tag's prior log odds ($B_{i}$) with new information ($\sum_{}^{} W_{j} S_{ji}$).
+The main difference is that an offset term ($O$) was included for the StackOverflow model, which is not present in SNIF-ACT or ACT-R's retrieval equations.
 
@@
 The log prior odds are based on both frequency and recency of use.
-However we are are assuming that frequency of tag use does not change greatly between Jul' 08 and Apr' 11 (time period of dataset).
-This allows us to compute log odds based on frequency only, as is indicated in Table \ref{sample-table}.
+However, we are assuming that frequency of tag use does not change greatly between Jul' 08 and Apr' 11 (time period of dataset).
+This is a reasonable assumption for the StackOverflow dataset, since tags relate to programming languages, and the rank of programming-language popularity has not changed drastically within the past few years.
+This also makes calculating the priors computationally tractable for this large-scale dataset, because the prior log odds can be computed based on frequency only, as is indicated in Table \ref{sample-table}.
 So the tag with the highest frequency (\emph{C\#}) would have the highest base-level activation.
@@
 
-It's also the case that there are many commonalities between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
+It is also the case that there are many commonalities between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
 and many of those theories use a combination of weighting schemes for contextual relevance and global importance of a particular chunk.
@@
 This was done primarily to improve model fit.
-To see why this is the case, suppose the following scenario:
+To see why this is the case, consider the following scenario:
 The offset measurement is not used, and the model is asked to provide ten tags across ten posts.
@@
 The model would then select the ten tags from the single post as the most likely tags.
-But this doesn't make intuitive sense, because there is a limit to the number of tags that people tend to use for each post.
+But this does not make sense, because there is a limit to the number of tags that people tend to use for each post.
 So the offset measure aims to place all of these ten posts on the same level, so that asking the model to choose ten tags from ten posts will select roughly a single tag from each post.
@@
 Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).
-
-\subsubsection{Dataset}
-
-The Apr' 11 StackOverflow dataset \cite{DataDump2011} was used for this analysis.
-This is the entire collection of posts dated between Jul' 2008 (StackOverflow's creation) and Apr' 11.
-It contains \num{4945168} posts from \num{559803} unique users and \num{26176} unique tags.
-Post authors used approximately 3 tags on average per post.
-
-\subsubsection{Distribution of Tag Frequency}
-\emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
-Tag frequency then drops off sharply.
-The top 1\% of tags based on frequency of use account for approximately 60\% of total tag instances.
-This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.
 
@@
 
-\citeA{Bird2009} recommended to remove the low-predictor stop words from the analysis.
+\citeA{Bird2009} recommended the removal of low-predictor stop words from the analysis.
 \citeauthor{Bird2009} provided a short list that were commonly removed (e.g., ``the'', ``and'', ``or'').
@@
 Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.46$, npv = .997, ppv = .70).
-It's also interesting that although the attentional weighting ($W$) terms for the title and body were calibrated by the regression, they did not stray far from cognitively-plausible defaults
+It is also interesting that although the attentional weighting ($W$) terms for the title and body were calibrated by the regression, they did not stray far from cognitively-plausible defaults
 (see coefficients for title and body in Table \ref{tab:coeffs}).
@@
   \centering
-  \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-act-cropped.pdf}}
+  \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-act-crop.pdf}}
   \caption{
@@
   \centering
-  \resizebox{\linewidth}{!}{\includegraphics{ROC-cropped.pdf}}
+  \resizebox{\linewidth}{!}{\includegraphics{ROC-crop.pdf}}
   \caption{
@@
 However, removing only one component (particularly the words in the body of the post) does not drastically reduce model performance.
-It's only when groups of components are removed that model performance starts to drastically suffer.
+It is only when groups of components are removed that model performance starts to drastically suffer.
 
@@
 Model performance improved by using a scaled entropy measure to weight the importance of each contextual element in the body and title of a post.
-However, it's uncommon for an ACT-R model that uses the declarative memory retrieval mechanisms to use anything other than equal attentional weighting for each chunk in context.
+However, it is uncommon for an ACT-R model that uses the declarative memory retrieval mechanisms to use anything other than equal attentional weighting for each chunk in context.
 But it is much more common outside of the ACT-R modeling community to use both a local (strength of association) and global (attentional weight) term to predict document retrieval \cite{Dumais1991}.
-In fact, it's been shown that the Pointwise Mutual Information Index (PMI) (a commonly-used measure to predict document retrieval)
+In fact, it has been shown that the Pointwise Mutual Information Index (PMI) (a commonly-used measure to predict document retrieval)
 is mathematically equivalent to ACT-R's retrieval equations when the sample size of the corpus is large \cite{Budiu2007, Farahat2004},

### 2013-03-16 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 4c341b0b78582a9017bfb54ce784b28e477583f9 [formerly bbf712ad341b3b24cb674f265a3c4c95820a8053]
Former-commit-id: 0280f749441fc54204506a17727de8e6005809aa

@@
 
-A logistic regression technique was used in order to operationalize this fit measure.
+A logistic regression technique was used to operationalize this fit measure.
 For each post, three sets of tag groups were combined to most accurately sample tag activations while managing computer resource constraints:
-The model's tag activations for the highest 200 tags, a weighted sample of 800 tags (weighted towards higher activation), and the tag activations for all tags that were chosen by the author of the post.
+The top 20 tags having the highest model activation, a random sample of 400 tags, and the tags that were chosen by the author of the post.
 Each of these observations was assigned a category of 0 or 1 depending on if the tag was a chosen tag or not for that post (0 is not chosen).
@@
 However using a single run of the \num{1000} posts was insufficient because there are two model components that require prior estimates of coefficients:
-The offset measure to determine the top-five tags and associated activation and the sampling technique used to pick a representative set of tags for each post.
+The offset measure to determine the top-five tags and associated activation and the sampling technique used to pick the top 20 model-predicted tags for each post.
 So model calibration was bootstrapped, where the coefficients from the previous run were used as estimates of coefficients for the next run.
@@
 
-The final best-fit coefficients for each model component and formal statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits} respectively.
+\vspace{-1em}
 
@@
 
+The final best-fit coefficients for each model component and formal statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits} respectively.
 Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.46$, npv = .997, ppv = .70).
@@
 And note that the model is by no means perfect.
-For example, \emph{html} is not tagged by the author, but the model ranks it as the second highest tag.
-And the author's tag \emph{firefox} is only ranked 8th by the model.
+For example, \emph{html} and \emph{jquery} are not tagged by the author, but the model ranks them as the second and third highest tag.
+And the author's tag \emph{firefox} is only ranked 12th by the model.
 
@@
 
-\subsection{Classification Accuracy}
+\subsubsection{Classification Accuracy}
 
@@
 
-\subsection{Potential Applications}
+\subsubsection{Potential Applications}
 
@@
 
-\subsection{Comparison to \citeA{Kuo2011}}
+\subsubsection{Comparison to \citeA{Kuo2011}}
 
@@
 
-\subsection{Entropy and ACT-R's Attentional Weighting}
+\subsubsection{Entropy and ACT-R's Attentional Weighting}
 
@@
 
-\subsection{Current Model Issues}
+\subsubsection{Current Model Issues}
 
@@
 One hypothesis is that this normalization is required because the retrieval threshold of a tag is adjusted for each post, and changed so that only a small set of tags are retrieved for all posts.
-Further research will explore this and other theories for the behavior of the offset component in the model.
+Further research will explore this and other theories for the behavior of the offset component.
 
-\subsection{Future Planned Research}
+\subsubsection{Future Planned Research}
 
@@
 We are exploring a Bayesian updating technique that would adjust prior tag activations from global priors to user's priors as more tagging information about the user is collected.
-For the StackOverflow dataset, adding user's tagging history to the model should drastically improve predictive power for veteran users,
-since tags are tied to programming languages and specific technologies, and programmers tend to specialize on a few set of tags that are of interest to them.
-This technique could also be deployed, since a post author must be signed into StackOverflow before creating a new post.
+For the StackOverflow dataset, adding user's tagging history to the model should improve predictive power for veteran users,
+since programmers tend to specialize on a few set of tags that are of interest to them.
+This technique could also be deployed, since a post author must be signed in before creating a post.
 

### 2013-03-17 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 73f78ef349a2604d285356d5b4a9b81143a2a14f [formerly 6cbdf23c60e57f511e041763dc2489918180d760]
Former-commit-id: baceafc71a4ff08ba0dac57922c15bae12b0d47d

@@
 
-Note how model performance is almost identical for the calibration and test dataset (compare full model calibration to full model test).
-This indicates that the model parameters were not overfit, and generalize well to the rest of the dataset.
-It also provides evidence that \num{1000} posts is an adequate number to estimate model parameters accurately.
+Note how model performance is almost identical for the calibration and test dataset (compare full model calibration to full model test), which indicates that model parameters generalize well.
+There is a small decrease in performance for test compared to calibration to the left of the vertical dotted line, and model performance at the dotted line is slightly lower (65\% compared to 67\%).
+Model performance for the test dataset (cross-validated performance) is the most accurate measurement of real-world performance,
+and model accuracy is 65\% for this dataset when tasked to predict one tag per post on average.
 

### 2013-03-17 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 18eb7813e9e7980a50571dd9b0e4e1f95d774dfb [formerly 7e7eefb767de8c0a764f2d01925aaa97b84717f3]
Former-commit-id: 1a908f099316a698171f7e8b7eefa771a0519cc8

@@
   We explored the StackOverflow dataset, and developed an ACT-R inspired Bayesian probabilistic model that can predict the hashtags used by the author of the post---predicting the users' programming interests.
-  The model is 67\% accurate when tasked to predict one tag per post on average.
+  The model is 65\% accurate when tasked to predict one tag per post on average.
   This is achieved by choosing the tag that has the highest log odds of being correct,
@@
       \hline
-      (Intercept) &	-1.35 &		0.10 &		-13.59 &	\textless2e-16 *** \\
-      prior &		1.08 & 		0.01 &		88.12 & 	\textless2e-16 *** \\
-      sjiTitle &	1.41 &		0.02 &		57.36 &		\textless2e-16 *** \\
-      sjiBody &		2.36 &		0.04 &		58.48 &		\textless2e-16 *** \\
-      offset &		-0.71 &		0.02 &		-39.05 &	\textless2e-16 *** \\
+      (Intercept) &	-3.00 &		0.12 &		-25.97 &	\textless2e-16 *** \\
+      prior &		0.62 & 		0.01 &		66.94 & 	\textless2e-16 *** \\
+      sjiTitle &	0.92 &		0.02 &		39.90 &		\textless2e-16 *** \\
+      sjiBody &		1.76 &		0.04 &		43.21 &		\textless2e-16 *** \\
+      offset &		0.78 &		0.02 &		33.25 &		\textless2e-16 *** \\
       \hline
@@
       \hline
-      McFadden's Pseudo $R_{}^{2}$ &	.46 \\
-      True Positives &			647 \\
-      False Positives &			278 \\
-      True Negatives &			\num{796982} \\
-      False Negatives &			\num{2235} \\
+      McFadden's Pseudo $R_{}^{2}$ &	.52 \\
+      True Positives &			507 \\
+      False Positives &			215 \\
+      True Negatives &			\num{417409} \\
+      False Negatives &			\num{2375} \\
       Positive Predictive Value &	0.70 \\
-      Negative Predictive Value &	0.997 \\
+      Negative Predictive Value &	0.994 \\
       \hline
@@
 Our StackOverflow model is a successful case showing that ACT-R's declarative memory retrieval mechanisms scale, and can be applied to retrieval requests occurring across large-scale declarative memory datasets.
-The model can successfully tag 67\% of posts correctly, when tasked to predict one tag per post on average.
+The model can successfully tag 65\% of posts correctly, when tasked to predict one tag per post on average.
 The model is also computationally space efficient:
@@
 
-Our model achieves 67\% classification accuracy when predicting one tag per post on average
+Our model achieves 65\% classification accuracy when predicting one tag per post on average
 (height of full model at vertical dotted line in Figure \ref{fig:ROC})

### 2013-03-18 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: e33b40fa14b137c80e377a87d94517151961170c [formerly dce291dbc8470f0a803c910627c97f35080b7b3c]
Former-commit-id: d572a56a0a2fb0dbf130932cc2f8a5084f9ebc25

@@
 The final best-fit coefficients for each model component and formal statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits} respectively.
-Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.46$, npv = .997, ppv = .70).
+Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.52$, npv = .994, ppv = .70).
 It is also interesting that although the attentional weighting ($W$) terms for the title and body were calibrated by the regression, they did not stray far from cognitively-plausible defaults

### 2013-04-02 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: feb498b0aa7e9b03dc491f4b598ad89599fca2da [formerly 5088d1898425a4aef3cc53324b0b30391bc4fbaf]
Former-commit-id: 911b9ab9554c58350e7cb574d9b1c1a599de4da0
foo
Former-commit-id: 48cf445c0d303851080d9878459cc13873abd8a3 [formerly 4b4dba6b7835cadbc5b7ec24e6d70f093bcd5a73]
Former-commit-id: d784fec4f253e1adda4f1b48e19b5d9eacc6ab8e

@@
   \AND {\large \bf Michael D. Byrne (byrne@rice.edu)} \\
-  Department of Psychology and Computer Science, 6100 Main Street \\
+  Departments of Psychology and Computer Science, 6100 Main Street \\
   Houston, TX 77005 USA }
@@
 \begin{abstract}
-  Large-scale behavioral datasets such as Twitter, Facebook, LinkedIn, StackOverflow, and Yelp can be mined and explored to study human behavior and human computer interaction.
-  We focused on hashtag creation as a specific human behavior, since understanding and modeling hashtag creation can lead to improved human-computer systems that can better identify users' goals and interests.
-  We explored the StackOverflow dataset, and developed an ACT-R inspired Bayesian probabilistic model that can predict the hashtags used by the author of the post---predicting the users' programming interests.
+  Hashtags created by authors of online content provide a view of a user's goals and interests.
+  Predicting users' interests can lead to improved, more user-centered human-computer systems.
+  Large-scale behavioral datasets such as Twitter, Facebook, LinkedIn, StackOverflow, and Yelp can be mined and explored to study these hashtags and how they relate to users' interests.
+  We explored the StackOverflow dataset, and developed an ACT-R inspired Bayesian probabilistic model that can predict the hashtags used by the author of the post.
   The model is 65\% accurate when tasked to predict one tag per post on average.
@@
 
-We built a tag prediction system for the StackOverflow dataset that can solve precisely these sorts of problems.
+We built a tag prediction system for the StackOverflow dataset that is intended to solve precisely these sorts of problems.
 The StackOverflow dataset was chosen because it is relatively constrained, contains a large number of posts and associated tags, and the data are easy to obtain.
 It is constrained because there is a fully enumerated and limited set of tags that an author can use, as opposed to a less-constrained domain such as Twitter.
-The site also makes public all post data on a quarterly basis, which can be downloaded and loaded into a relational database for further analysis.
+The site also makes public all post data on a quarterly basis, which can be downloaded and loaded into a database for further analysis.
 
@@
 There are \num{1468485} posts from \num{559803} unique users and \num{26176} unique tags.
-Post authors used approximately 3 tags per post.
+Post authors used an average of 2.9 tags per post.
 
 \emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
-Tag frequency drops off sharply and follows Zipf's law.
+Tag frequency drops off sharply and approximately follows Zipf's law.
 The top 1\% of tags ranked by frequency of use account for approximately 60\% of total tag instances.
@@
       prior &		0.62 & 		0.01 &		66.94 & 	\textless2e-16 *** \\
-      sjiTitle &	0.92 &		0.02 &		39.90 &		\textless2e-16 *** \\
-      sjiBody &		1.76 &		0.04 &		43.21 &		\textless2e-16 *** \\
+      Wtitle &		0.92 &		0.02 &		39.90 &		\textless2e-16 *** \\
+      Wbody &		1.76 &		0.04 &		43.21 &		\textless2e-16 *** \\
       offset &		0.78 &		0.02 &		33.25 &		\textless2e-16 *** \\
@@
 since programmers tend to specialize on a few set of tags that are of interest to them.
-This technique could also be deployed, since a post author must be signed in before creating a post.
+This technique is possible, since a post author must be signed in before creating a post.
 

### 2013-06-19 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: badac33fe9e9bccc490f1e6696b25811990b7c37 [formerly 35a9c0ad732c4977e428241535871cd85af3cd34]
Former-commit-id: 625efd9cee1c3704adb831523ccfb508bfb012ac
ICCM revisions
Former-commit-id: e5687596b15d3449f5506054428777fb51b0c944 [formerly f7232924e39102cbc7c1081c9846ad260e866bbd]
Former-commit-id: e8cc0b891f6b2bd7707de7aaf7acb88c2ccb70b2

@@
   Hashtags created by authors of online content provide a view of a user's goals and interests.
-  Predicting users' interests can lead to improved, more user-centered human-computer systems.
+  Predicting users' interests can lead to improved, more user centered human-computer systems.
   Large-scale behavioral datasets such as Twitter, Facebook, LinkedIn, StackOverflow, and Yelp can be mined and explored to study these hashtags and how they relate to users' interests.
@@
 
-Predicting tags for the StackOverflow dataset has been worked on by at least one prior researcher \cite{Kuo2011}.
+At least one prior researcher \cite{Kuo2011} has worked on StackOverflow tag prediction.
 \citeauthor{Kuo2011} uses a co-occurrence model that predicts tags based on the words in the post and their relation (co-occurrence) to tags.
 The model was initially built for next-word prediction in large documents, and then adapted to the StackOverflow dataset by constraining the next word predicted to only tags.
-His co-occurrence model has a 47\% classification accuracy for predicting StackOverflow tags when the model predicts one tag per post.
+His co-occurrence model has a 47\% classification accuracy predicting one tag per post.
 
 The SNIF-ACT \cite{Fu2007} model also uses co-occurrences and can predict link traversal for search queries.
-That is, given a search query (goal state), fetched results, and words included in those fetched results, the model predicts the most likely link that a person will click on.
+That is, given a search query (goal state) and fetched results, the model predicts the most likely link that a person will click on.
 This StackOverflow model is similar because it leverages ACT-R's declarative memory retrieval mechanisms to generate predictions (tag activations instead of link activations).
@@
 The weights for each model component were calibrated using a sample of \num{1000} posts not contained within the strength of association dataset.
-The calibrated model was then tested on a new sample of \num{1000} posts not contained within either previous training dataset.
-Classification accuracy was used as a metric of model performance (i.e., how many of the model's predicted tags are correct).
+The calibrated model was then tested on a fresh sample of \num{1000} posts.
+Classification accuracy was used as a metric of model performance (i.e., number of correctly-predicted tags).
 
@@
 For that post, the author's chosen tags are known (see Figure \ref{fig:examplePost}).
-So model performance can be measured by comparing the model's tag activations with chosen tags for a post, and then aggregated across posts to get an overall average performance.
+So model performance can be measured by comparing the model's tag activations with chosen tags for a post, and then aggregated across posts to get an average.
 
@@
 
-\subsection{Logistic Regression}
+\subsubsection{Logistic Regression}
 
@@
 
-The final best-fit coefficients for each model component and formal statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits} respectively.
-Note that all four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.52$, npv = .994, ppv = .70).
+The final best-fit coefficients for each model component and statistical measurements are included in Tables \ref{tab:coeffs} and \ref{tab:fits}.
+All four coefficients for the run are strong predictors of tag use ($p<.01$), and model fit is satisfactory (McFadden's pseudo $R_{}^{2}=.52$, npv = .994, ppv = .70).
 It is also interesting that although the attentional weighting ($W$) terms for the title and body were calibrated by the regression, they did not stray far from cognitively-plausible defaults
@@
 
-\subsection{Example Posts}
+\subsubsection{Example Posts}
 
-A visualization of the model's tag activation for the example post in Figure \ref{fig:examplePost} is included in Figure \ref{fig:modelPost}.
+A visualization of the model's tag activation for the post in Figure \ref{fig:examplePost} is included in Figure \ref{fig:modelPost}.
 
@@
   \caption{
-    Model's prediction of top-ten tags for post in Figure \ref{fig:examplePost}.
-    The title is the title text of the post.
-    The attentional weight (scaled entropy) for each word in the title is included below each word.
-    The user-chosen tags are included below the attentional weights.
-    The stacked bar graphs are the total activation for the top-ten predicted model tags for the post.
-    Each bar graph is partitioned into the three model components that vary across the tags in a single post (not the offset).
+    Model's prediction of the top tags for post in Figure \ref{fig:examplePost}.
+    At the top is the title text of the post, then the attentional weight (scaled entropy) for each word in the title, then the user-chosen tags.
+    The bar graphs are the total activation for the top predicted model tags for the post.
+    Each bar graph is partitioned into the three model components that vary across the tags in a single post.
 }
@@
 
-Note how the base level activation and contextual activation combine to predict tags that are both frequently used and contextually relevant.
+The base level activation and contextual activation combine to predict tags that are both frequently used and contextually relevant.
 For example, note that the \emph{javascript} tag leverages a combination of contextual activation and prior activation to achieve a high total activation.
@@
 
-\subsection{Generalizability}
+\subsubsection{Generalizability}
 
@@
     Model performance across the calibration and test dataset, as well as when specific model components are removed.
-    The y-axis represents model performance (one is perfect), and the x-axis is the proportion of tags that the model generates compared to the number of tags used by the authors of the posts (one is equal number).
-    The vertical dotted line represents the point where the model is asked to tag on average one tag per post.
+    The y-axis is model performance (one is perfect), and the x-axis is the proportion of tags that the model generates compared to the total number of tags used by the post authors (one is equal number).
+    The vertical dotted line is where the model generates one tag per post on average.
   }
@@
 
-\subsection{Necessity Analysis}
+\subsubsection{Necessity Analysis}
 
-Figure \ref{fig:ROC} also shows model performance after specific model components are removed.
+Figure \ref{fig:ROC} also shows performance after specific model components are removed.
 Model parameters were recalibrated for each model, except for the full model test dataset that used the parameters from the calibration dataset.
@@
 
-The model could also be used as a tag cleanup system for stackoverflow.com.
+The model could also be used as a tag cleanup system.
 In this scenario the tags for a previously submitted post would be compared against model predictions.
@@
 
-Our model achieves 65\% classification accuracy when predicting one tag per post on average
-(height of full model at vertical dotted line in Figure \ref{fig:ROC})
+Our model achieves 65\% classification accuracy when predicting one tag per post on average (see Figure \ref{fig:ROC})
 compared to 47\% accuracy for the co-occurrence model presented in \citeA{Kuo2011}.
-Most likely we achieved the increase in model performance by adding some additional model components:
-The offset and possibly the body term as well,
+Most likely we achieved the increase in model performance by adding the offset model component and possibly the body term as well, 
 as it was unclear if \citeauthor{Kuo2011} used context from both the body and title of a post or just the title.
@@
 
+\subsubsection{Model Plausibility}
+
+ACT-R's declarative memory retrieval process is the core of the StackOverflow model.
+This is a more constrained and cognitively-plausible model than the set of machine learning techniques used by \citeA{Kuo2011} for the same task. 
+Interestingly, when using ACT-R's retrieval mechanism, the model performs significantly better than the less constrained models in \citeA{Kuo2011}.
+So there is no reason to think that people do not engage in a process much like this to select tags.
+However, more research is necessary before making the strong claim that the StackOverflow model \emph{is} the process people go through when selecting tags.
+In particular, the tag errors made by authors should be looked at, and model behavior should be tested against these important counter examples.
+Also, places where the StackOverflow model deviate from ACT-R's default retrieval theory (i.e., the offset term, entropy and attentional weighting) should be looked at in greater detail.
+
 \subsubsection{Current Model Issues}
@@
 
-Alongside exploring the offset component of the model, our main planned research is to include the post author's past tagging behavior as an additional model predictor.
+Alongside exploring the offset component and testing cognitive plausibility, our main planned research is to include the post author's past tagging behavior as an additional model predictor.
 We are exploring a Bayesian updating technique that would adjust prior tag activations from global priors to user's priors as more tagging information about the user is collected.
@@
 since programmers tend to specialize on a few set of tags that are of interest to them.
-This technique is possible, since a post author must be signed in before creating a post.
 
@@
 
-We would like to thank Fred Oswald for his excellent guidance and recommendations throughout the model development and data analysis process.
+We would like to thank Fred Oswald for his excellent guidance and feedback throughout the model development process.
 

### 2013-06-19 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: c782536a5292b72da82c3c108b419f1ed421bc0d [formerly 83c11cfce91a3b642b29d1fcbea1aa2deb7c624b]
Former-commit-id: e422b3b2bb063633fb21499ae73d17e6d4e7fb03

@@
 
-ACT-R's declarative memory retrieval process is the core of the StackOverflow model.
+ACT-R's declarative memory retrieval process makes up the core of the StackOverflow model.
 This is a more constrained and cognitively-plausible model than the set of machine learning techniques used by \citeA{Kuo2011} for the same task. 
@@
 So there is no reason to think that people do not engage in a process much like this to select tags.
-However, more research is necessary before making the strong claim that the StackOverflow model \emph{is} the process people go through when selecting tags.
+However, more research is necessary before making the strong claim that the StackOverflow model \emph{is} the process people go through when selecting tags on the site, or tags in general.
 In particular, the tag errors made by authors should be looked at, and model behavior should be tested against these important counter examples.
-Also, places where the StackOverflow model deviate from ACT-R's default retrieval theory (i.e., the offset term, entropy and attentional weighting) should be looked at in greater detail.
+Also, places where the StackOverflow model deviates from ACT-R's default retrieval theory (i.e., the offset term, entropy and attentional weighting) should be looked at in greater detail.
 

### 2013-06-20 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 5edb59e2e31bf5a4ef92355e1e5f316be7b54786 [formerly 303366c5c379c498a3cbe3535b1f0e437de635eb]
Former-commit-id: 5ab358b5f36ef66d3a970a2dde2c312f06a56ef1

@@
 ACT-R's declarative memory retrieval process makes up the core of the StackOverflow model.
-This is a more constrained and cognitively-plausible model than the set of machine learning techniques used by \citeA{Kuo2011} for the same task. 
+This is a more constrained and cognitively-plausible model than the set of general machine learning techniques used by \citeA{Kuo2011} for the same task. 
 Interestingly, when using ACT-R's retrieval mechanism, the model performs significantly better than the less constrained models in \citeA{Kuo2011}.
@@
 
-Alongside exploring the offset component and testing cognitive plausibility, our main planned research is to include the post author's past tagging behavior as an additional model predictor.
+Alongside exploring the offset component and cognitive plausibility, our main planned research is to include the post author's past tagging behavior as an additional model predictor.
 We are exploring a Bayesian updating technique that would adjust prior tag activations from global priors to user's priors as more tagging information about the user is collected.

### 2013-06-20 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 8a9221c45f421ecd2730032b92d7d9e20ffbbfde [formerly d8f3480fb10c27f4d2603a125c047231ebf142ac]
Former-commit-id: 38ce017e476ea3b48c45d732fb45470970b3b14f

@@
 However, more research is necessary before making the strong claim that the StackOverflow model \emph{is} the process people go through when selecting tags on the site, or tags in general.
-In particular, the tag errors made by authors should be looked at, and model behavior should be tested against these important counter examples.
-Also, places where the StackOverflow model deviates from ACT-R's default retrieval theory (i.e., the offset term, entropy and attentional weighting) should be looked at in greater detail.
+In particular, the tag errors made by authors should be analyzed, and model behavior should be tested against these important counter examples.
+Also, places where the StackOverflow model deviates from ACT-R's default retrieval theory (i.e., the offset term, entropy and attentional weighting) should be examined in greater detail.
 

### 2013-06-27 stackExchange/writeups/iccm-2013/writeup.tex
Final revision
Former-commit-id: 3791c8f9eaa95dc92cb83845716a7056d2c13830 [formerly 2c3cdc4133127b114d44681635b91c19df964e3b]
Former-commit-id: 7cc9c5b3d5cb527dbb20d44040ae427c67ca8aaf

@@
 
-Base-level activation ($B_{i}$) is the log prior odds of observing tag $i$ in the world.
-The log prior odds are based on both frequency and recency of use.
+Base-level activation ($B_{i}$) is the prior log odds of an author using tag $i$.
+The prior log odds are based on both frequency and recency of use.
 However, we are assuming that frequency of tag use does not change greatly between Jul' 08 and Apr' 11 (time period of dataset).

### 2013-06-27 stackExchange/writeups/iccm-2013/writeup.tex
Updated table data
Former-commit-id: 2a71fa469b0776f5058a458baef1052cdc7a6b03 [formerly 6acf0019a9d8e795d2b7c774a4ff1388199ed9ae]
Former-commit-id: 79116cc0f3308822344390d583498f21fa033413

@@
       \hline
-      (Intercept) &	-3.00 &		0.12 &		-25.97 &	\textless2e-16 *** \\
+      (Intercept) &	-2.22 &		0.10 &		-22.95 &	\textless2e-16 *** \\
       prior &		0.62 & 		0.01 &		66.94 & 	\textless2e-16 *** \\
-      Wtitle &		0.92 &		0.02 &		39.90 &		\textless2e-16 *** \\
-      Wbody &		1.76 &		0.04 &		43.21 &		\textless2e-16 *** \\
+      Wtitle &		0.93 &		0.02 &		40.00 &		\textless2e-16 *** \\
+      Wbody &		1.75 &		0.04 &		43.16 &		\textless2e-16 *** \\
       offset &		0.78 &		0.02 &		33.25 &		\textless2e-16 *** \\
@@
       True Positives &			507 \\
-      False Positives &			215 \\
-      True Negatives &			\num{417409} \\
+      False Positives &			216 \\
+      True Negatives &			\num{417408} \\
       False Negatives &			\num{2375} \\

### 2013-06-27 stackExchange/analysis/model/analyzer-spring.R
Updated code to generate ICCM data
Former-commit-id: dc2246c1e13d61701a4d371f3a290963719821c6 [formerly fe2b0aadb3f3c8733290f9eb14677e46d7dbe449]
Former-commit-id: 908dd058066ed0f634f741bf84c080c8ddf9cafc

@@ -72,6 +72,8 @@ analyzeForICCM = function() {
 	plotAllFrms(frms=frms)
 	tagFiles=getTagFiles(makeTagDir(8))
 	visPost(tagFiles[105], makeTagDir(8), coeffs=frms[[1]]$coeffs, topNum=12)
+	getLogitSummary(frms[1])
+	getClassLog(frms[1])
 }
 
 analyzeBaseFrmForMultivariate = function(baseFrm, maxCount) {

### 2013-06-27 stackExchange/analysis/model/analyzer.R

@@ -28,7 +28,10 @@ devOff = function() {
 logReg = function(tag, res, model) {
 	myPrint(str_c("working ", length(tag), " tags"))
 	dfSubset = res[res$tag %in% tag,]
-	weights = dfSubset$weights
+	if( ! "weights" %in% colnames(dfSubset)) {
+		myPrint("Did not find weights column in dataframe; running model w/o weights")
+		dfSubset$weights = 1
+	}
 	mylogit = glm(model, data=dfSubset, family=binomial(link="logit"), weights=weights)
 	myPrint(summary(mylogit))
 	tmp = ClassLog(mylogit, dfSubset$targetP)

### 2013-06-28 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: ee4819176add4ccd3ddba367011a2dc8567e00e7 [formerly 6b0abd22ed2424f49174b0c8f55dbe47e4bb4027]
Former-commit-id: 4f8bd67fa826263cc09ae517e085f99991bc3f9b
Final changes for ICCM paper
Former-commit-id: 7831a32e9283798e678b5be5cabb44feb122950e [formerly 15d5de80ad819e87f2b7d414475e8277b7662e51]
Former-commit-id: 9e646c1cc98b6d5e3f8f72cd60ff108b2245abf1

@@
     Model's prediction of the top tags for post in Figure \ref{fig:examplePost}.
-    At the top is the title text of the post, then the attentional weight (scaled entropy) for each word in the title, then the user-chosen tags.
+    At the top is the title text of the post, then the attentional weight (scaled entropy) for each word in the title, then the author-chosen tags.
     The bar graphs are the total activation for the top predicted model tags for the post.
@@
 Our StackOverflow model is a successful case showing that ACT-R's declarative memory retrieval mechanisms scale, and can be applied to retrieval requests occurring across large-scale declarative memory datasets.
-The model can successfully tag 65\% of posts correctly, when tasked to predict one tag per post on average.
+The model can successfully tag 65\% of StackOverflow posts correctly, when tasked to predict one tag per post on average.
 The model is also computationally space efficient:
@@
 
-Alongside exploring the offset component and cognitive plausibility, our main planned research is to include the post author's past tagging behavior as an additional model predictor.
+Alongside exploring the offset component and cognitive plausibility of the model, our main planned research is to include the post author's past tagging behavior as an additional model predictor.
 We are exploring a Bayesian updating technique that would adjust prior tag activations from global priors to user's priors as more tagging information about the user is collected.
@@
 
-We would like to thank Fred Oswald for his excellent guidance and feedback throughout the model development process.
+We would like to thank Fred Oswald for his excellent guidance and feedback throughout the model development and analysis process.
 

### 2013-06-28 stackExchange/analysis/model/analyzer-spring.R

@@ -74,6 +74,8 @@ analyzeForICCM = function() {
 	visPost(tagFiles[105], makeTagDir(8), coeffs=frms[[1]]$coeffs, topNum=12)
 	getLogitSummary(frms[1])
 	getClassLog(frms[1])
+	getAccuracyAtNAverageTags(frms[[1]]$frm, N=1)
+	getAccuracyAtNAverageTags(frms[[2]]$frm, N=1)
 }
 
 analyzeBaseFrmForMultivariate = function(baseFrm, maxCount) {

### 2013-08-12 writeups/proposal/proposal.tex
Init
Former-commit-id: b818013ce82ee39c6c36f1dd4b5ddd5c9e55a1d5 [formerly d987c172f4b2e831baed4d343d0375967008f1ef]
Former-commit-id: e4cf121eaee05137a356830617f7e71902230ab3

\documentclass[man]{apa6}

\usepackage[american]{babel}
\usepackage[utf8]{inputenc}
%\usepackage[compact]{titlesec}

\usepackage{amssymb,amsmath}

\usepackage{graphicx}
\usepackage[group-separator={,}]{siunitx}
\RequirePackage[l2tabu, orthodox]{nag}
\graphicspath{{./figures/}} % Specifies the directory where pictures are stored

\usepackage{csquotes}
\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{bibliography.bib}

\title{Psyc 601 Multivariate Statistics Project}
\shorttitle{Psyc 601 Project}

\author{Clayton Stanley}
\affiliation{Rice University}

\leftheader{Beitzel}

\abstract{
  The base StackOverflow tag-prediction model from the Fall '12 Psychometrics course was improved by incorporating the words in the body of the post and the post author's specific tagging history.
  We focused on hashtag creation as a specific human behavior, since understanding and modeling hashtag creation can lead to improved human-computer systems that can better identify users' goals and interests.
  We used multivariate logistic regression statistical techniques to guide the model building process, and found that adding the two model terms improved model accuracy from 56\% to 69\%.
  The model is a successful case showing that ACT-R's declarative memory retrieval equations scale, and are relevant to task domains that require large-scale knowledge stores.
}

\keywords{StackOverflow, Machine Learning, ACT-R, Tagging, Large-Scale Semantic Memory}

\begin{document}
\maketitle

\tableofcontents
\newpage

\section{Motivation}

TODO: Motivation

\subsection{Research Questions}

\subsubsection{Comparison Between ACT-R and Vector-based Memory Systems}

\subsubsection{Modeling Lifetime of a User's Hashtag}

\subsubsection{Single-user-focused Hashtag Modeling}

\section{Hashtag Prediction}

\subsection{Recommendation Systems}

\subsubsection{Netflix}

\subsubsection{Facebook}

\subsubsection{Twitter}

\section{ACT-R Declarative Memory Theory}

\subsection{ACT-R DM Model}

\subsection{Pointwise Mutual Information}

\section{Matrix-based Models of Semantics}

\subsection{Latent Semantic Analysis}

\subsubsection{Word Order}

\subsubsection{Comparison to PMI}

TODO: Issue of scale.

\section{Vector-based Models of Semantics}

\subsection{Addressing Word Order}

\subsection{Addressing Scalability}

\subsection{BEAGLE}

\subsection{Random Permutations}

\subsubsection{Connection to LSA}

\section{Methods}

\subsection{StackOverflow Dataset}

\subsection{Twitter Dataset}

\subsection{Analyses}

\subsubsection{Top Hashtag Prediction}

\subsubsection{Lifetime of a User's Hashtag}

\subsubsection{User-customized Hashtag Prediction}

\printbibliography

\end{document}

### 2013-08-12 writeups/proposal/proposal.tex
Mon commit
Former-commit-id: 187bbab021a30daea008f2f7813a3bcf3392fa9d [formerly 6258a37350ba2696ecb1952eca6a9352e6e950b6]
Former-commit-id: 6179e69077c912be9041b6a093c8ecf956e28a2f

@@
 
+TODO: Identifying user's goals when searching social media sites is an open research problem \cite{Rose2004}
+
+TODO: Researchers are currently looking for ways to automate the task \cite{Lee2005}, \cite{Jansen2008}
+
+TODO: Hashtags are an additional view of a user's goals when generating content on a site, and can help in this automation.
+TODO: These hashtags relate to user's goals and interests because they are human-generated content; a form of human-based document tagging \cite{Chang2010}
+TODO: People are also using hashtags as an information source on social media cites \cite{Diakopoulos2010} \cite{Kwak2010}
+
+TODO: Natural question to ask is how can we build human-computer systems that help connect users to these hashtag information sources.
+
+TODO: Co-occurence-based modeling is a potentially useful approach. \cite{Efron2010}
+
+TODO: Challenging domain due to time constraints on information. \cite{Bauer2012}
+
+TODO: Useful problem since real-time search on social media sites is growing \cite{Jansen2011}
+
 \subsection{Research Questions}
@@
 
+TODO: Extending work from \cite{Rutledge2008}
+
+TODO: Can user prior be represented in vector-based memory systems?
+
+TODO: Can word order be represented in ACT-R?
+
 \subsubsection{Modeling Lifetime of a User's Hashtag}
+
+TODO: Comparison to global lifetime of a hashtag \cite{Tsur2012}
 
@@
 
-\section{Hashtag Prediction}
+\subsubsection{Task Domain}
+
+TODO: StackOverflow and Twitter; large scale, hashtag prediction, etc.
+
+\section{Prior Work on Hashtag Prediction}
+
+\subsection{StackOverflow}
+
+TODO: \cite{Kuo2011}
+
+\subsection{Twitter}
+
+FIXME: Find cite for twitter hashtag prediction
 
@@
 
-\subsubsection{Netflix}
-
-\subsubsection{Facebook}
-
-\subsubsection{Twitter}
+FIXME: Find cites for recommendation systems.
 
 \section{ACT-R Declarative Memory Theory}
+
+\cite{Anderson2004}
 
@@
 
+TODO: Introduce PMI \cite{Farahat2004}
+
+TODO: Comparison to ACT-R: Virtually equivalent for large corpora \cite{Farahat2004}
+
+\subsection{Scaling the Equations}
+
+\cite{Douglass2010}
+
+TODO: SNIF-ACT \cite{Fu2007} \cite{Pirolli2003}
+
 \section{Matrix-based Models of Semantics}
@@
 \subsection{Latent Semantic Analysis}
+
+TODO: Summarize \cite{Landauer1997}
 
@@
 
-TODO: Issue of scale.
+TODO: Performance and issue of scale. \cite{Budiu2007}
 
 \section{Vector-based Models of Semantics}
+
+TODO: Summarize \cite{Plate1995}
 
@@
 
+TODO: Sumarize Jones' model \cite{Jones2007}
+
 \subsection{Random Permutations}
 
+TODO: Summarize \cite{Sahlgren2008}
+
+\subsection{Comparison between BEAGLE and Random Permutations}
+
+TODO: Summarize \cite{Recchia2010}
+
 \subsubsection{Connection to LSA}
+
+FIXME: Get citation.
+
+\section{Comparison of Vector-based Models and ACT-R}
+
+TODO: Summarize \cite{Rutledge2007}
+
+TODO: This approach can be used to model the fan effect \cite{Rutledge2008}
+
+TODO: However, there are open issues that need to be addressed:
+TODO: Can user prior be represented in vector-based memory systems?
+TODO: Can word order be represented in ACT-R?
+TODO: Do they both scale?
 
@@
 
+\cite{DataDump2013}
+
 \subsection{Twitter Dataset}
+
+\subsection{Tokenizing and Lemmetizing Text}
+
+TODO: Python NLP toolkit \cite{Bird2009}
+
+\subsection{Stop Words and Entropy Weighting}
+
+TODO: Explore entropy weighting instead of stop word removal \cite{Dumais1991}
 

### 2013-08-14 writeups/proposal/proposal.tex
foo
Former-commit-id: c1f2c4c918e39331eb1670719d8bc1b8b2cb6571 [formerly 321d3ae21488d07e4082b61b8c07d2c44c5a7a2a]
Former-commit-id: 8060f7fb3030f596a2624cc9f8d2a28cbeeefd10
foo
Former-commit-id: a0c904c4c347a5d5e37003cb7c16c09bdaea6888 [formerly 8db452a7d709f06293c8edfee936ac76f3b43f64]
Former-commit-id: 797f913fe18152452138192bdffb0819c2086a32

@@
 \addbibresource{bibliography.bib}
+
+% If the toc is too detailed, try limiting the depth of printed sections in the TOC:
+% \setcounter{tocdepth}{2}
 
@@
 
-\subsubsection{Comparison Between ACT-R and Vector-based Memory Systems}
+\subsubsection{Comparison Between ACT-R and Vector-Based Memory Systems}
 
@@
 
-\subsubsection{Modeling Lifetime of a User's Hashtag}
+\subsubsection{Single-User-Focused Hashtag Modeling}
 
@@
 
-\subsubsection{Single-user-focused Hashtag Modeling}
+TODO: How is recency information about the hashtag included in the model?
 
@@
 
-\section{Prior Work on Hashtag Prediction}
+\section{Prior Research}
 
-\subsection{StackOverflow}
+\subsection{Hashtag Prediction}
+
+\subsubsection{StackOverflow}
 
@@
 
-\subsection{Twitter}
+\subsubsection{Google+}
 
-FIXME: Find cite for twitter hashtag prediction
+TODO: Summarize Google+ hashtag prediction \cite{GoogleKeynote2013}
+
+\subsubsection{Twitter}
+
+\cite{Efron2010} Generated related hashtags from 'query'.
+Showed that users found those hashtags useful for the query.
+Not using the model to predict hashtags from tweets, but a model like this could be used for that purpose.
+They were focused more on making the case that hashtags provide useful information that can be used on sites like Twitter as part of a general recommendation system.
 
 \subsection{Recommendation Systems}
+
+\subsubsection{The Netflix Prize}
+
+TODO: Summarize and \cite{Bennett2007}
+
+\subsubsection{Recommending Twitter Users to Follow}
+
+TODO: Summarize and \cite{Hannon2010}
 
@@
 
-\subsection{Pointwise Mutual Information}
+\subsection{Connection to Pointwise Mutual Information}
 
@@
 
-\section{Matrix-based Models of Semantics}
-
-\subsection{Latent Semantic Analysis}
+\section{Latent Semantic Analysis Memory Theory}
 
@@
 
-\subsubsection{Word Order}
+\subsection{Singular Value Decomposition}
 
-\subsubsection{Comparison to PMI}
+\subsection{Word Order}
 
-TODO: Performance and issue of scale. \cite{Budiu2007}
+\subsection{Scaling Issues}
 
-\section{Vector-based Models of Semantics}
+TODO: Compare with PMI, performance and issue of scale. \cite{Budiu2007}
+
+\section{Vector-Based Memory Systems}
 
@@
 
-\subsubsection{Connection to LSA}
+\subsection{Connection to LSA}
 
-FIXME: Get citation.
+TODO: Summarize \cite{Kanerva2000}
 
@@
 
+\section{Incorporating Hashtag Recency}
+
+\subsection{ACT-R}
+
+TODO: ACT-R has a strong theory here.
+TODO: IR literature for Twitter domain uses a model similar to ACT-R \cite{Efron2011}
+
+\subsection{Vector-based Models}
+
+TODO: Unclear how recency will fit in with random permutation vector-based model.
+
 \section{Methods}
@@
 \subsection{Twitter Dataset}
+
+TODO: common-hashtags dataset
+TODO: popular-users dataset
 
@@
 
+\subsubsection{StackOverflow Dataset Modifications}
+
+TODO: Use newest dataset
+TODO: Chunk tags
+
+\subsubsection{Analyze Recency for User-Generated Hashtags}
+
+TODO: Use stackoverflow and twitter dataset
+
+\subsubsection{Incorporating Recency Into Random Permutations}
+
+TODO: Use stackoverflow dataset only for random permutations
+
+\subsubsection{Incorporating Word Order Into ACT-R}
+
+TODO: Use twitter dataset only for ACT-R
+
 \subsubsection{Top Hashtag Prediction}
 
-\subsubsection{Lifetime of a User's Hashtag}
+TODO: vector-based and act-r head to head on twitter
+TODO: head to head on stackoverflow
 

### 2013-08-14 writeups/proposal/proposal.tex
foo
Former-commit-id: 8ae3fc4d2620b8820a1c7a17b9ac8f0b53caf144 [formerly 487d92856600e556c04fa355aa241cc9b40c21ce]
Former-commit-id: 98822d95e69a2e68a970ac75d6307df0c87dacae

@@
 % If the toc is too detailed, try limiting the depth of printed sections in the TOC:
-% \setcounter{tocdepth}{2}
+\setcounter{tocdepth}{2}
 
@@
 
-\subsubsection{Single-User-Focused Hashtag Modeling}
+\subsubsection{Modeling Hashtag Creation for a Single User}
 
@@
 
-\subsubsection{Task Domain}
+\subsubsection{Adding Word-Order Information When Predicting Hashtags}
+
+TODO: speak generally about how word order might be useful for hashtags
+
+\subsubsection{StackOverflow and Twitter Task Domains}
 
@@
 
-\subsubsection{Recommending Twitter Users to Follow}
+\subsubsection{Recommending Followers on Twitter}
 
@@
 
-\section{ACT-R Declarative Memory Theory}
+\subsection{ACT-R Declarative Memory Theory}
 
@@
 
-\subsection{ACT-R DM Model}
+\subsubsection{ACT-R DM Model}
 
-\subsection{Connection to Pointwise Mutual Information}
+\subsubsection{Connection to Pointwise Mutual Information}
 
@@
 
-\subsection{Scaling the Equations}
+\subsubsection{Scaling the Equations}
 
@@
 
-\section{Latent Semantic Analysis Memory Theory}
+\subsection{Latent Semantic Analysis Memory Theory}
 
@@
 
-\subsection{Singular Value Decomposition}
+\subsubsection{Singular Value Decomposition}
 
-\subsection{Word Order}
+\subsubsection{Word Order}
 
-\subsection{Scaling Issues}
+\subsubsection{Scaling Issues}
 
@@
 
-\section{Vector-Based Memory Systems}
+\subsection{Vector-Based Memory Systems}
 
@@
 
-\subsection{Addressing Word Order}
+\subsubsection{Addressing Word Order}
 
-\subsection{Addressing Scalability}
+\subsubsection{Addressing Scalability}
 
-\subsection{BEAGLE}
+\subsubsection{BEAGLE}
 
@@
 
-\subsection{Random Permutations}
+\subsubsection{Random Permutations}
 
@@
 
-\subsection{Comparison between BEAGLE and Random Permutations}
+\subsubsection{Comparison between BEAGLE and Random Permutations}
 
@@
 
-\subsection{Connection to LSA}
+\subsubsection{Connection to LSA}
 
@@
 
-\section{Comparison of Vector-based Models and ACT-R}
+\subsection{Comparison of Vector-Based Models and ACT-R}
 
@@
 
-\section{Incorporating Hashtag Recency}
+\subsection{Incorporating Hashtag Recency}
 
-\subsection{ACT-R}
+\subsubsection{ACT-R}
 
@@
 
-\subsection{Vector-based Models}
+\subsubsection{Vector-Based Models}
 
@@
 
-\subsection{Analyses}
+\subsection{Experiments}
 
@@
 
-TODO: Use stackoverflow and twitter dataset
+TODO: Use stackoverflow and twitter popular-users dataset
 
@@
 
-TODO: Use twitter dataset only for ACT-R
+TODO: Use twitter common-hashtags dataset only for ACT-R
 
@@
 
-TODO: vector-based and act-r head to head on twitter
+TODO: vector-based and act-r head to head on twitter common-hashtags
+TODO: head to head without word order on twitter common-hashtags
 TODO: head to head on stackoverflow
 
-\subsubsection{User-customized Hashtag Prediction}
+\subsubsection{User-Customized Hashtag Prediction}
+
+TODO: head to head on twitter popular-users dataset
+TODO: head to head on stackoverflow
 

### 2013-08-15 writeups/proposal/proposal.tex
foo
Former-commit-id: 9c9e21f20d37118ce31bac2ecbfa9ed1beba0e48 [formerly 9a81e2b44301d2a888d32abe44338d1860a5e365]
Former-commit-id: 5dc9f89720a27f65430497e47c3af0aa0e41fc50

@@
 \usepackage{amssymb,amsmath}
+
+\usepackage[hyphens]{url}
+%\usepackage{hyperref}
+%\usepackage{breakurl}
 
@@
 
+TODO: Useful problem since real-time search on social media sites is growing \cite{Jansen2011}
+TODO: Users want to be quickly connected to the content that they care about, with minimal amount of searching necessary, since the amount of generated content on these sites is so vast.
+
+TODO: Challenging domain due to time constraints on information. \cite{Bauer2012}
+
+TODO: Social media sites are composed almost entirely of human-generated content.
+Users are searching through this content.
+
 TODO: Identifying user's goals when searching social media sites is an open research problem \cite{Rose2004}
@@
 
-TODO: Natural question to ask is how can we build human-computer systems that help connect users to these hashtag information sources.
+TODO: Natural question to ask is if we can predict a users' interested hashtags, we have a better understanding of that user's goals on the site.
+And with that understanding, we can improve their interaction with the site by tailoring the system to give them content that they care most about.
 
 TODO: Co-occurence-based modeling is a potentially useful approach. \cite{Efron2010}
-
-TODO: Challenging domain due to time constraints on information. \cite{Bauer2012}
-
-TODO: Useful problem since real-time search on social media sites is growing \cite{Jansen2011}
 
@@
 
-TODO: Summarize \cite{Rutledge2007}
+TODO: Summarize MALTA \cite{Rutledge2007}
 

### 2013-08-15 writeups/proposal/proposal.tex
Thu commits
Former-commit-id: 1ce64a521fd9b977949af454a1f7e5f8d7d42be6 [formerly 3ea02c1fff9e15ea9e5c0dadbef51f0fb357c211]
Former-commit-id: 58456f74662c202a5ba643918aef26d7fe486927

@@
 
+\usepackage{textcomp}
 \usepackage[american]{babel}
@@
 \usepackage{amssymb,amsmath}
+\usepackage{setspace}
 
@@
 \usepackage{csquotes}
-\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
+\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber,doi=false,uniquename=false]{biblatex}
 \DeclareLanguageMapping{american}{american-apa}
@@
 % If the toc is too detailed, try limiting the depth of printed sections in the TOC:
-\setcounter{tocdepth}{2}
+%\setcounter{tocdepth}{2}
 
@@
 TODO: Useful problem since real-time search on social media sites is growing \cite{Jansen2011}
+TODO: Social media sites are composed almost entirely of human-generated content.
+Users are searching through this content.
 TODO: Users want to be quickly connected to the content that they care about, with minimal amount of searching necessary, since the amount of generated content on these sites is so vast.
@@
 TODO: Challenging domain due to time constraints on information. \cite{Bauer2012}
-
-TODO: Social media sites are composed almost entirely of human-generated content.
-Users are searching through this content.
 
@@
 
-\subsubsection{Modeling Hashtag Creation for a Single User}
+\subsubsection{Modeling Hashtag Use for a Single User}
 
@@
 
+\subsection{Recommendation Systems}
+
+Grown in need with the growth of the web.
+Users on a site need access to specific pieces of information, and the site contains much more information than they can process / care about.
+Recommendation systems are used to tailor the presented information to what is relevant to the user.
+\cite{Pazzani2007}
+
+\subsubsection{The Netflix Prize}
+
+TODO: Summarize and \cite{Bennett2007}
+
+\subsubsection{Recommending Followers on Twitter}
+
+TODO: Summarize and \cite{Hannon2010}
+
+\subsubsection{Hashtags for Recommendation Systems}
+
+\cite{Efron2010} Generated related hashtags from 'query'.
+Showed that users found those hashtags useful for the query.
+Not using the model to predict hashtags from tweets, but a model like this could be used for that purpose.
+They were focused more on making the case that hashtags provide useful information that can be used on sites like Twitter as part of a general recommendation system.
+
 \subsection{Hashtag Prediction}
+
+Previous research focused on StackOverflow, Google+, and Twitter.
+Google+ has deployed automatic tagging of posts.
 
@@
 
-\cite{Efron2010} Generated related hashtags from 'query'.
-Showed that users found those hashtags useful for the query.
-Not using the model to predict hashtags from tweets, but a model like this could be used for that purpose.
-They were focused more on making the case that hashtags provide useful information that can be used on sites like Twitter as part of a general recommendation system.
-
-\subsection{Recommendation Systems}
-
-\subsubsection{The Netflix Prize}
-
-TODO: Summarize and \cite{Bennett2007}
-
-\subsubsection{Recommending Followers on Twitter}
-
-TODO: Summarize and \cite{Hannon2010}
-
-FIXME: Find cites for recommendation systems.
+TODO: None that we are aware of.
 
@@
 
+\begingroup
+\setstretch{1}
+\setlength\bibitemsep{12pt}
 \printbibliography
+\endgroup
 

### 2013-08-16 writeups/proposal/proposal.tex
foo
Former-commit-id: d25ad90b48b4595a355e158baf4107201f93d065 [formerly 6f0f2e4d379526fd73a8ea12538bfd9d34d7013d]
Former-commit-id: 574e885413b29591d664761c0375847bf3754a53
foo
Former-commit-id: 73f94630abd97f5a6e20461eb6ddedcc9282fdf2 [formerly 1df978033bac6518a474b5472a4c670dd5cbae29]
Former-commit-id: 811f713786e599ca65dcdfcc26fd6f20a8a62288

@@
 \addbibresource{bibliography.bib}
+
+\AtEveryBibitem{\clearfield{doi}}
+
+\AtEveryBibitem{
+  \ifentrytype{misc}{}{
+    \clearfield{url}
+  }
+}
 
@@
 
+FIXME: Fix cite for Bauer2012
 TODO: Challenging domain due to time constraints on information. \cite{Bauer2012}
@@
 
+FIXME: Fix cite for Bennett2007
 TODO: Summarize and \cite{Bennett2007}
@@
 TODO: \cite{Kuo2011}
+FIXME: cite recent paper
+FIXME: cite tex.stackexchange.org as a hashtag suggestion tool
 

### 2013-08-19 writeups/proposal/proposal.tex
foo
Former-commit-id: 2fa3e4eec7a4819a2cb911bff30fa92d681ea34e [formerly 38dfe0aec5a1a85862d9c4baab406f4ebecc042a]
Former-commit-id: 1f4c6bc48be14fcce95a3a0832e7fa998ddaf6f3
mon commits
Former-commit-id: 3ba81b20ade3f054e7d33d87cea937f8c0abae27 [formerly 011565c1ec21344d86c1686f30e2ea8fbc8e3bc3]
Former-commit-id: 2aba7fbce7237a29c570e98e702c08173960ef07

@@
 \usepackage[utf8]{inputenc}
+
 %\usepackage[compact]{titlesec}
@@
 
+% Needed to ensure periods are placed at the end of every bibliography entry in the references section
 \AtEveryBibitem{\clearfield{doi}}
 
+% Make sure only urls that are printed in references section are for webpages
 \AtEveryBibitem{
@@
     \clearfield{url}
+  }
+}
+
+% ref: http://tex.stackexchange.com/questions/128592/remove-backslashes-from-url-fields-in-bib-entry
+\DeclareSourcemap{
+  \maps[datatype=bibtex]{
+    \map{
+      \step[fieldsource=url,
+	match=\regexp{\\},
+      replace=\regexp{}]
+    }
   }
@@
 
-TODO: Useful problem since real-time search on social media sites is growing \cite{Jansen2011}
-TODO: Social media sites are composed almost entirely of human-generated content.
-Users are searching through this content.
-TODO: Users want to be quickly connected to the content that they care about, with minimal amount of searching necessary, since the amount of generated content on these sites is so vast.
+Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp, are composed almost entirely of human-generated content.
+The amount of human-generated content on these sites is staggering, and continues to grow.
+Users of Twitter's microblogging service, for example, generate half a billion tweets per day \parencite{TwitterReport2012}.
+However, a single user of a social media site is most likely interested in only a small fragment of this large amount of information.
+One general question to support the the user in this domain is:
+How can we quickly and effectively connect users to the content that they care about?
 
-FIXME: Fix cite for Bauer2012
-TODO: Challenging domain due to time constraints on information. \cite{Bauer2012}
+% FIXME: Fix cite for Bauer2012
+One traditional technique is to provide the user with a powerful search engine so that they can actively look for content that interests them.
+However, users are not forming the same search queries that work so effectively on search engines like Google.
+Users of social media sites are searching for information that is happening now, in real time \cite{Bauer2012}.
+In other words, users on social-media sites want the fresh \emph{stream} of information about a particular topic.
+This can certainly be seen in the main user-interface views that these social media site provide:
+Twitter's homepage showing your followers' tweets, Facebook's homepage showing your friends' posts, StackOverflow's daily digest of posts for particular tags, etc.
 
-TODO: Identifying user's goals when searching social media sites is an open research problem \cite{Rose2004}
+So how can we ensure that users are connected to the information streams that they care most about?
+One promising approach is to model and predict the user's goals when using and searching these social media sites \parencite{Rose2004}.
+If we have a better understanding of a user's goals, then we can tailor the information streams provided to the ones that are most inline with these goals.
+For example, Twitter might suggest other users and hashtags for the user to follow, Facebook could suggest pages to like, Yelp could recommend businesses to check out, etc.
+Researchers are currently looking for ways to automate the goal-identification task for more general search queries like those generated for Google \parencites{Jansen2008}{Lee2005}.
 
-TODO: Researchers are currently looking for ways to automate the task \cite{Lee2005}, \cite{Jansen2008}
+However, social media sites have an advantage for identifying user goals over traditional search engines:
+Human-generated content.
+Twitter does not have only the tweet that a user is currently composing to try and identify that user's goals.
+Rather, the site has all of the user's previously-generated content, and can use that information to provide a much better prediction of a user's goals.
+Further, sites like Twitter, StackOverflow, Google+, and Facebook support ways for the user to explicitly identify their precise goals when creating content: Hashtags.
+These user-created hashtags are a key indicator to a user's goals on a social media site.
 
-TODO: Hashtags are an additional view of a user's goals when generating content on a site, and can help in this automation.
-TODO: These hashtags relate to user's goals and interests because they are human-generated content; a form of human-based document tagging \cite{Chang2010}
-TODO: People are also using hashtags as an information source on social media cites \cite{Diakopoulos2010} \cite{Kwak2010}
+These hashtags relate to a user's goals and interests because they are human-generated content, and a form of human-based document tagging \cite{Chang2010}.
+Further, people are using hashtags as a way to create information streams on social media sites \parencite{Kwak2010}.
+Twitter users, for example, commonly create hashtags for upcoming political events, such as debates and races \cite{Diakopoulos2010}.
+So if we can predict what hashtags a user is interested in, we have a better understanding of the information streams that they care about.
+With that understanding, we can tailor the system to suggest hashtags that they may not yet know about but are of likely interest to them.
+We can help aid in their discovery of fresh and relevant information that match their goals on the site.
 
-TODO: Natural question to ask is if we can predict a users' interested hashtags, we have a better understanding of that user's goals on the site.
-And with that understanding, we can improve their interaction with the site by tailoring the system to give them content that they care most about.
-
-TODO: Co-occurence-based modeling is a potentially useful approach. \cite{Efron2010}
+FIXME: It is certainly the case that real-time search on these social media sites is growing \parencite{Jansen2011}.
 
@@
 
+The core research question we are interested in is:
+What kinds of cognitively-plausible models can best predict the human-generated tags that a user on a social media site is interested in?
+This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
+This process can be framed as a memory retrieval problem.
+That is, each user on a social media site has created content that contain a set of human-generated tags within the context of each post.
+The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.
+
+Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010}.
+Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
+Two of the current state-of-the-art memory models are based on a co-occurrence methodology, and we'll be comparing these two models for hashtag prediction:
+ACT-R's declarative memory retrieval theory and random permutation vector-based memory systems.
+
 \subsubsection{Comparison Between ACT-R and Vector-Based Memory Systems}
 
-TODO: Extending work from \cite{Rutledge2008}
+ACT-R's declarative memory and vector-based memory systems are substantially different in their implementation.
+Nonetheless, both can successfully model the classic fan effect phenomena found in word pairing experiments \parencite{Rutledge2008}.
+We will extend the work done by \textcite{Rutledge2008} by formally comparing the models on large-scale hashtag prediction tasks.
+We'll also replace the vector-based model used by \textcite{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
+The primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory.
 
-TODO: Can user prior be represented in vector-based memory systems?
+An advantage of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
+Word-order information on sites like Twitter may contain highly-predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
+However, it is unclear and unexplored how word order can be represented in the ACT-R declarative memory theory.
+We will explore how the word-order strength of vector-based models can be incorporated into the ACT-R theory when testing the models on hashtag prediction tasks.
 
-TODO: Can word order be represented in ACT-R?
+ACT-R's declarative memory system has a strong theory on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2004}.
+However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.
+We will explore how a user's prior knowledge influences the likelihood that a particular hashtag is generated for both ACT-R's memory theory and vector-based models.
 
-\subsubsection{Modeling Hashtag Use for a Single User}
+\subsubsection{Lifetime of a Hashtag for a Specific User}
 
-TODO: Comparison to global lifetime of a hashtag \cite{Tsur2012}
+Prior research has examined the growth and decay cycle of hashtag use across users \parencite{Tsur2012}.
+However, much less is known about hashtag lifetime within a single user.
+It may not be the case that a hashtag's lifecycle for a single user matches the lifecycle across users.
+Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into the a specific user's prior likelihood of hashtag use.
+So we will characterize the hashtag growth and decay cycle within users.
+The expectation is that this cycle will match ACT-R's decay rate used in the declarative memory system to compute a particular memory item's prior.
+
+\subsubsection{User-Customized Hashtag Prediction}
+
+The end goal of this work is to identify a memory retrieval model that can suggest relevant hashtags to users when they wish to retrieve one.
+These hashtags should be customized to their specific interests and relevant to the content of the post that they are currently creating.
+This will undoubtably require a combination of two primary model components:
+[1] A user's prior likelihood of choosing a particular hashtag, given their prievously-generated content, [2] the likelihood that a particular hashtag is related to the context of the post being generated.
+
+***
 
 TODO: How is recency information about the hashtag included in the model?
-
-\subsubsection{Adding Word-Order Information When Predicting Hashtags}
-
-TODO: speak generally about how word order might be useful for hashtags
 

### 2013-08-20 writeups/proposal/proposal.tex
tue
Former-commit-id: 855b839db86331a00377b0ffb1ca00a40a03d1b1 [formerly 621dd4678b91b8dee6eea31196500d1c9a6e99ed]
Former-commit-id: e096867905019df4db412ba8a484ca40478fabef

@@
 
+\subsubsection{StackOverflow and Twitter Task Domains}
+
+We will compare these two memory systems across two domains where users generate tags for posts: Twitter and StackOverflow.
+Twitter is a micro-blogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
+StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
+
+These two datasets were chosen because the content of the human-generated content is quite different between them, and we are interested in retrieval models that generalize across different task domains.
+However, they are also quite similar on several accounts:
+both domains have emassed large amounts of user-generated content, users generate tags for posts when creating content on the site, and the user data from the datasets are publicly available for analysis.
+
+Also, studying user-based tag generation on these sites is a relevant task domain.
+Models that can accuratly predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these very popular sites.
+These systems can help newer users by recommending proper tags for their newest content.
+On the StackOverflow site for example, experts often subscribe to specific hashtags that interest them, and receive a daily digest of posts that are tagged with those hashtags.
+Helping the user properly tag posts on StackOverflow ensures that right community sees the post, which greatly increases the chance that the question on the post will be answered quickly and correctly.
+For Twitter, hashtags represent streams of information that are possibly interesting to the user.
+Having a recommendation system that can suggest relevant hashtags to the user provides a way for the user to connect to and discover new information streams.
+
 \subsubsection{Comparison Between ACT-R and Vector-Based Memory Systems}
@@
 The expectation is that this cycle will match ACT-R's decay rate used in the declarative memory system to compute a particular memory item's prior.
+This decay rate equation formalizes how recency and frequency of hashtag use relate to the prior probability that a hashtag will be retrieved.
 
@@
 This will undoubtably require a combination of two primary model components:
-[1] A user's prior likelihood of choosing a particular hashtag, given their prievously-generated content, [2] the likelihood that a particular hashtag is related to the context of the post being generated.
+[1] A user's prior likelihood of choosing a particular hashtag, given their prievously-generated content, and [2] the likelihood that a particular hashtag is related to the context of the post being generated.
 
-***
-
-TODO: How is recency information about the hashtag included in the model?
-
-\subsubsection{StackOverflow and Twitter Task Domains}
-
-TODO: StackOverflow and Twitter; large scale, hashtag prediction, etc.
+A user's prior hashtag use will be an essential model component for domains such as Twitter, where the number of possible hashtags is practically infinite.
+In these domains, the model can utilize a user's prior hashtag use as a way to prune the infinite space of possible hashtags, and generate a much smaller and more manageable set for prediction.
+However, not only should the model properly take into account a user's prior hashtag use, but it should also generate new hashtag predictions when a user is most likely choosing a hashtag they have never used before.
+Calibrating the models to properly balance this exploitation of prior hashtag use vs. exploration of new hashtags will be a primary resaerch focus.
+The end goal is to have a model that can properly balance the user's prior tag with the contextual cues in the post in order to generate an accurate prediciton.
 
@@
 
-Grown in need with the growth of the web.
-Users on a site need access to specific pieces of information, and the site contains much more information than they can process / care about.
-Recommendation systems are used to tailor the presented information to what is relevant to the user.
-\cite{Pazzani2007}
+A model that predicts a user-chosen hashtag is a specific type of the more general recommendatino systems.
+These systems have grown in need with the growth of the web.
+Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
+Recommendation systems are used to tailor the information presented to what is most relevant to the user \cite{Pazzani2007}.
 
@@
 
-FIXME: Fix cite for Bennett2007
-TODO: Summarize and \cite{Bennett2007}
+% FIXME: Fix cite for Bennett2007
+The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
+Netflix is an online-subscription service where users can stream their favorite television episodes and movies.
+Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
+In order to provide a high-quality user experience on the site, Netflix created a content-recommendation service.
+This tool recommends to the user movies and television episodes that most likely match their specific interests.
+Netflix wanted to improve the accuracy of their recommendation system, so they held a competition where teams could try to improve on the accuracy of the service.
+
+The winning team utilized a singular value decomposition (SVD) method to identify particular dimensions of the entire content library (e.g., genre, time period, previously watched) that best predicted user preferences.
+This is a dimensionality-reducing technique that aims to filter out noise by forcing the data to be represented from a smaller number of dimensions than were originally present.
+This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
+By using an SVD technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.
 
@@
 
-TODO: Summarize and \cite{Hannon2010}
+Recommendation systems based on the content of user-generated posts have been successfull in recommending people on Twitter for a user to follow.
+\parencite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
+The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
+The profiles were created by generating a word frequency matrix of words x (user + followers + followees), and then providing that matrix as input into the Lucine text search engine framework.
+When the system needed to recommend a set of individuals for a user to follow, it would take all of the words used in posts from that user to construct a search query to be provided to Lucine.
+That search query would return the most likely users that share similar words in posts within their twitter network with the words provided in the query.
+Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user will follow at 25\% accuracy.
 
@@
 
-\cite{Efron2010} Generated related hashtags from 'query'.
-Showed that users found those hashtags useful for the query.
-Not using the model to predict hashtags from tweets, but a model like this could be used for that purpose.
-They were focused more on making the case that hashtags provide useful information that can be used on sites like Twitter as part of a general recommendation system.
+\cite{Efron2010} showed how hashtags might be used as a type of recommendation system for Twitter.
+Users constructed a query of text that represented a search for a topic that they were interested in.
+The system then retrieved the most likely hashtags that were associated with that search query.
+The model is based on building a word co-occurance matrix of words x hashtags, and then uses that matrix to assign a similarity score for each hashtag when provided with a search query.
+Participants graded how relevant the hashtags returned were to each search query.
+The results suggest that hashtags contain useful information that help classify the gist of a post. 
+Although this model does not predict hashtags directly when composing a tweet, a model like this could certainly be tailored for that purpose.
 
@@
 
+\section{Models}
+
+TODO: Formal description of each model.
+
 \section{Methods}

### 2013-08-22 writeups/proposal/proposal.tex
thu commit
Former-commit-id: b9ac40a5c1a86f94e1d2d0368647ebfac116be07 [formerly 0ba736a575820a079856bd502708d0ccf893f597]
Former-commit-id: 93c2cc2ee770064e45ad39a4b5e162d3a94482da

@@
 However, users are not forming the same search queries that work so effectively on search engines like Google.
-Users of social media sites are searching for information that is happening now, in real time \cite{Bauer2012}.
+Users of social media sites are searching for information that is happening now, in real time \parencite{Bauer2012}.
 In other words, users on social-media sites want the fresh \emph{stream} of information about a particular topic.
@@
 However, social media sites have an advantage for identifying user goals over traditional search engines:
-Human-generated content.
+human-generated content.
 Twitter does not have only the tweet that a user is currently composing to try and identify that user's goals.
@@
 
-These hashtags relate to a user's goals and interests because they are human-generated content, and a form of human-based document tagging \cite{Chang2010}.
+These hashtags relate to a user's goals and interests because they are human-generated content, and a form of human-based document tagging \parencite{Chang2010}.
 Further, people are using hashtags as a way to create information streams on social media sites \parencite{Kwak2010}.
-Twitter users, for example, commonly create hashtags for upcoming political events, such as debates and races \cite{Diakopoulos2010}.
+Twitter users, for example, commonly create hashtags for upcoming political events, such as debates and races \parencite{Diakopoulos2010}.
 So if we can predict what hashtags a user is interested in, we have a better understanding of the information streams that they care about.
@@
 Twitter is a micro-blogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
+User-created hashtags on Twitter are embedded into the tweet, and can be used to connect the tweet to other tweets about the same topic through common hashtag use.
+Some recent and often-used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.
 StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
+User-chosen tags on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
+Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, or \emph{Common Lisp}.
 
@@
 However, they are also quite similar on several accounts:
-both domains have emassed large amounts of user-generated content, users generate tags for posts when creating content on the site, and the user data from the datasets are publicly available for analysis.
+Both domains have emassed large amounts of user-generated content, users generate tags for posts when creating content on the site, and the user data from the datasets are publicly available for analysis.
 
@@
 Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
-Recommendation systems are used to tailor the information presented to what is most relevant to the user \cite{Pazzani2007}.
+Recommendation systems are used to tailor the information presented to what is most relevant to the user \parencite{Pazzani2007}.
 
@@
 Recommendation systems based on the content of user-generated posts have been successfull in recommending people on Twitter for a user to follow.
-\parencite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
+\textcite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
 The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
@@
 
-\cite{Efron2010} showed how hashtags might be used as a type of recommendation system for Twitter.
+\textcite{Efron2010} showed how hashtags might be used as a type of recommendation system for Twitter.
 Users constructed a query of text that represented a search for a topic that they were interested in.
@@
 
-Previous research focused on StackOverflow, Google+, and Twitter.
-Google+ has deployed automatic tagging of posts.
+More specific hashtag-recommendation models that attempt to predict a user's chosen hashtag have also been created and tested for a few social media sites.
+Google has even deployed a hashtag-recommendation model to help users properly tag posts for their Google+ social media site \parencite{GoogleKeynote2013}.
+However, to our knowledge a hashtag-recommendation model for Twitter tweets has not yet been attempted.
 
@@
 
-TODO: \cite{Kuo2011}
-FIXME: cite recent paper
-FIXME: cite tex.stackexchange.org as a hashtag suggestion tool
+\textcite{Kuo2011} was the first to model tag use for StackOverflow posts.
+His set of models were originally designed for next-word prediction for large text document collections.
+In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user generated a post, and restrict the space of possible words to only tags.
+He tested the set of next-word predictino models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
+Interestingly, the Bayesian co-occurrance model outperformed the more complicated and computatinoally expensive Singular Value Decomposition and K-nearest-neighbors models.
+This model can accurately predict a user-chosen tag at 47\% accuracy.
+
+\textcite{Stanley2013} also modeled a user's chosen tags for a post on the StackOverflow site.
+The model was based on ACT-R's declarative memory retrieval theory, which at the core is also a Bayesian co-occurrance model.
+However, this model scaled much better than the co-occurrance model presented in \textcite{Kuo2011}.
+Because of this, 100 times more posts were used to build the co-occurrance matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
+After scaling up the model, this model can successfully predict a user's tag for a post at 65\% accuracy.
+It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrance framework, since that is precisely the framework from which ACT-R's declarative memory system is built.
+This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular task domain:
+predictng the user-chosen tags for human-generated content on social media sites.
+
+Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
+We are curious to see the prediction difference between a vector-based models and current co-occurrance models on the StackOverflow dataset.
+Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
+Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, we are curious to see how a cognitively-plausible vector-based model performs at the task.
+
+A tag suggestion tool has also been deployed for the tex.stackexchange.org site. \parencite{LatexTags2013}
+Both the \LaTeX site and StackOverflow are part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
+Once a user has created a question on the Latex site, the site provides subdued text of suggested tags in the text field where the user fills in the tags to associate with the question.
+The post author is not forced to use the suggested tags, and is free to choose tags from them or tags not represented in the suggested set.
+
+The Latex site is not nearly as popular as StackOverflow currently.
+The site has roughly \num{50000} questions compared to \num{5000000} questions on StackOverflow.
+It may be the case that the StackExchange company decided to try out a tag recommendation system on one of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
+Or it is possible that their current tag recommendation system is not efficient enough to work with the demand of the StackOverflow site.
+Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial enough to the user to start adding them to their question and answer sites.
 
@@
 
-TODO: Summarize Google+ hashtag prediction \cite{GoogleKeynote2013}
+The hashtag recommendation system deployed for the Google+ micro-blogging service is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
+The workflow and user interface provided for recommending hashtags for Google+ is similar to how a hashtag recommendation system for Twitter might look.
+When generating content on Google+, the system automatically associates the most likely relevant hashtags with that post.
+The user is free to remove any or all of the tags that the system has associated with the post, and can also use tags that were not included in the recommended set.
+
+Google is apparently highly confident in the accuracy of their system.
+On the Latex site, the User Interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
+For Google+, the User Interface automatically provides tags for a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
+If tag accuracy is high enough, the Google+ technique is certainly favorable, since the user does not have to spend time figuring out which tags best represent this post.
+However, if the model makes even just a few errors, users may become annoyed and frustrated with the recommendation system, since they will have to explicitly correct it every time they generate a post.
 
@@
 
-TODO: None that we are aware of.
+To our knowledge, no model of user-created hashtags when composing tweets on Twitter has yet been built.
+There are two likely reasons for this:
+[1] Modeling user-chosen hashtags and building hashtag-recommendation systems has only explored over the past few years, and
+[2] Twitter is a large database of human-generated content, spanning a wide range of topics, containing a sizeable set of user-generated hashtags, so hashtag prediction for Twitter is a relatively hard problem.
+We are nonetheless curious if declarative memory retrieval models can be applied to this domain to generate useful hashtag predictions.
+The domain shares many similarities with environments where hashtag-recommendation systems have already been deployed:
+Google+'s microblogging service is essentially Twitter, but focused on a specific subset of users and topics.
+The StackExchange sites are also built from a large database of human generated content, where users have chosen specific tags to associate with posts.
+So we will evaluate how well an ACT-R-inspired Bayesian declarative memory model and a new vector-based model predict a user's chosen hashtag when composing a tweet on Twitter.
 

### 2013-08-23 writeups/proposal/proposal.tex
Fri commits
Former-commit-id: 5728261a08d655cee7c1f1ed3c06b1bc541ba922 [formerly d0e58081e5d89f1524b1f0259fc8d235b12a2809]
Former-commit-id: be46e14b71525979d6101ae9753ca692447f7678

@@
 
+\usepackage{xfrac}
 \usepackage{textcomp}
@@
 
+\usepackage{tabu}
 \usepackage{amssymb,amsmath}
@@
 
-\cite{Anderson2004}
+ACT-R \cite{Anderson2004} is a cognitive architecture that formalizes how each cognitive process of the brain (e.g., memory, learning, visual and motor) interacts to produce behavior.
+The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up long-term declarative memory retrieval.
+The equations that make up this system can be described through a rational analysis of long-term memory retrieval.
+That is, given the task of retrieving a chunk of information from long-term declarative memory,
+the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
+what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
+Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likilihood of being retrieved again, given the prior number of times it has been needed.
+Those prior probabilities for each chunk are then adjusted for the current context, so that the posterior probabilities represent the likilihood that a chunk is needed, given prior odds and current environment state.
 
 \subsubsection{ACT-R DM Model}
+
+\begin{table}[!ht]
+  \caption{ACT-R Declarative Memory Model}
+  \label{tabACT-RModel}
+  {\tabulinesep=1.2mm
+    \begin{tabu}{ll}
+      \hline
+      Common Name &  Equation \\
+      \hline
+      Activation &	 	$A_{i} = B_{i} + \sum_{j \in c}^{} W_{j} S_{ji}$ \\
+      Attentional Weight &	$W_{j} = \frac{W}{n}$ \\
+      Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
+      Constant Base Level &	$B_{i} = log \frac{p(i)}{p(\overline{i})}$ \\
+      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)}$ \\
+      Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
+      \hline
+    \end{tabu}
+  }
+\end{table}
+
+A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACT-RModel}.
+The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base level activation ($B_{i}$) and strength of association ($S_{ji}$).
+The recall probability that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
+
+\subsubsection{Base-Level Activation}
+
+Base-level activation reflects the log prior odds of needing an observed chunk again.
+The primary way to calculate these log prior odds is to use the standard base level equation in Table \ref{tabACT-RModel}.
+This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
+Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
+However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
+In this way the standard base level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its current frequency and recency of use.
+
+For some domains, it is reasonable to assume that the base-level activations of each chunk within a particualar window of interest are stable, and do not change within that window.
+Programming-language popularity over the past few years is a reasonable example of this.
+Although the popularity of various programming languages has certainly changed slightly over the past few years (.e.g, \emph{closure}'s growth), it is not the case that the changes have been drastic.
+\emph{C\#}, \emph{Python}, and \emph{Java} are still a few of the most popular languages, \emph{Common Lisp} has not gained or lost much ground in popularity over the past few years, etc.
+
+Further, if this is a reasonable assumption to make, it greatly simplifies the computation of the base-level activation for chunks, and can turn the computation into a tractable problem for large datasets.
+In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log odds ratio of being exposed to a chunk.
+For example, if the StackOverflow tag \emph{PHP} has been used four times as often as the tag \emph{Common Lisp}, then the prior odds of needing \emph{PHP} again is $\frac{.8}{.2}=4$ times that of \emph{Common Lisp}.
+
+\subsubsection{Strength of Association}
+
+Strength of association ($S_{ji}$) reflects the amount of log odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
+Context for the StackOverflow domain for example, can be represented as each word in the title and body of a post.
+Context for the Twitter domain would be the words in a Tweet.
+Association strenght between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
+The likelihood a chunk occurred with the current context ($p(i|j)$) over the likeihood that the chunk occured in any of the other contexts ($p(i|\overline{j})$).
+For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
+So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
+
+Using this assumption for large datasets, the interpretation of the context-adusted odds ratio becomes much simpler.
+If this ratio for a particular tag and context chunk is greater than one, then the log is positive, which means that this context has been observed more often with this tag than in general.
+If this ratio is less than one, then the log is negative, which provides a negative adjustment of total activation since this context has been observed more often in general than with this particular tag.
 

### 2013-08-26 writeups/proposal/proposal.tex
foo
Former-commit-id: e8e06245abfd0c0b47542c07258ee8dbc5937b43 [formerly 4ce8434aebd15fc50381302c2185c07d74c9e4e8]
Former-commit-id: e20d7ecf4e64580855d3b11817e36ea8b0a7232f
foo
Former-commit-id: ebd82cfcb57fb06af8059c0ca75e6fbbc3bfb82d [formerly b356864c73b6b70f2a7d0e5a6d73f5503ec7438d]
Former-commit-id: c554956320e46828e29582d53edba33dd6a8c4ad

@@
 
+Several hashtag recommendation models for Twitter have been developed recently.
+One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a post as a function of prior hashtag use and context.
+\parencite{Mazzia2009}
+The model present by \textcite{Mazzia2009} is very similar to an ACT-R declarative retrieval model.
+The main potential difference is that the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
+Instead, the global prior is an overall average frequency of hashtag use.
+
+Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used from similar tweets.
+\textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurence-based similarity score between a composed tweet and the rest of the tweets in the database.
+\textcites{Zangerle2011, Kwye2012} use the Apache Lucine infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarites.
+Regardless of the method to compute the similarity between tweets, each method collects the hashtags used from a set of the most similar tweets, then ranks the set, and presents the top 5-10 hashtags to the user.
+
+This tweet-centered approach is quite different than the hashtag-centered approach presented in \textcite{Mazzia2009}.
+With a hashtag-centered approach, a direct association is built between the words in a tweet and hashtags that occur in tweets.
+With a tweet-centered approach, that association between words and hashtags is indirect:
+The relationship is built between words in a tweet and a user, and then similar users are assumed to use similar hashtags.
+One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a vector representatino of the content in each tweet has to be maintained to later compute tweet similarity.
+The storage size for a hashtag-centered approach grows with the number of hashtags, but it seems reasonable that this will grow slower than the number of tweets and asymptote after collecting a large sample of tweets.
+So from the perspective of efficient information storage and retrieval, it seems more likely that the hashtag-centered approach used in \textcite{Mazzia2009} will scale as Twitter grows in size and use.
+
+\textcite{kwye2012} also customized their tag predictino model to the user's past hashtag use, and was one of the first to do so when developing a hashtag-prediction model for Twitter.
+The set of recommended hashtags was the union of hashtags used in similar tweets and hashtags used by similar users.
+This was done by storing a content vector of hshtag use for each user (user-centered approach) alongside the content vectors of hashtag use for each tweet (tweet-centered approach).
+To generate recommended hashtags, the model would return the top-ranked hashtags from the hashtags used in 0-50 similar tweets and 0-4 similar users.
+Prediction accuracy improved when the model generated recommendations based on hashtags in both similar tweets and the top few similar users compared to basing recommendations on similar tweets alone.
+
+FIXME: \cite{Godin2013}
+This paragraph will discuss the topics-based approach.
+
+FIXME: Paragraph that no vector-based models have been attempted for Twitter.
+Bayesian model can be improved by customizing predicitons based on past hashtag use.
+
+FIXME: ditch this paragraph?
 To our knowledge, no model of user-created hashtags when composing tweets on Twitter has yet been built.
@@
 So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
+This assumption has been both mentioned and used when deriving the $S_{ji}$ equation \parencite{Anderson1989}, as well as in recent research working with large-scale datasets \parencite{Stanley2013}.
 

### 2013-08-26 writeups/proposal/proposal.tex
foo
Former-commit-id: cf9f679145fba44530b496828036ab51f634c185 [formerly ff187ee8c940177eebc7d76b4aa289850694082d]
Former-commit-id: 4ba43c415cd639a943f13f5969bee025c0f2908d

@@
 \textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurence-based similarity score between a composed tweet and the rest of the tweets in the database.
-\textcites{Zangerle2011, Kwye2012} use the Apache Lucine infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarites.
+\textcites{Zangerle2011, Kywe2012} use the Apache Lucine infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarites.
 Regardless of the method to compute the similarity between tweets, each method collects the hashtags used from a set of the most similar tweets, then ranks the set, and presents the top 5-10 hashtags to the user.
@@
 
-\textcite{kwye2012} also customized their tag predictino model to the user's past hashtag use, and was one of the first to do so when developing a hashtag-prediction model for Twitter.
+\textcite{Kywe2012} also customized their tag predictino model to the user's past hashtag use, and was one of the first to do so when developing a hashtag-prediction model for Twitter.
 The set of recommended hashtags was the union of hashtags used in similar tweets and hashtags used by similar users.

### 2013-08-26 writeups/proposal/proposal.tex
foo
Former-commit-id: 6a45608d5941ef028562fe183477862d16283368 [formerly 4ed51fc7eb51ff77dd9ea92c926726bf48ec4ae5]
Former-commit-id: 9e21fbe337271d5b5146cd11ebecb7254040ddd2

@@
 With a tweet-centered approach, that association between words and hashtags is indirect:
-The relationship is built between words in a tweet and a user, and then similar users are assumed to use similar hashtags.
+The relationship is built between a tweet and its content, and then similar tweets are assumed to use similar hashtags.
 One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a vector representatino of the content in each tweet has to be maintained to later compute tweet similarity.

### 2013-08-26 writeups/proposal/proposal.tex
foo
Former-commit-id: bbe158597345a1535a7222e188fbdaa106db5e2f [formerly 6124dcd79c54c56ada68c48530b604c1446701ea]
Former-commit-id: c00831fb6c30ad4867601c8d3a7d7e39001bd166

@@
 To generate recommended hashtags, the model would return the top-ranked hashtags from the hashtags used in 0-50 similar tweets and 0-4 similar users.
-Prediction accuracy improved when the model generated recommendations based on hashtags in both similar tweets and the top few similar users compared to basing recommendations on similar tweets alone.
+Prediction accuracy improved when the model combined recommendations based on hashtags in both similar tweets and the top few similar users compared to basing recommendations on similar tweets alone.
 

### 2013-08-28 writeups/proposal/proposal.tex
foo
Former-commit-id: bbd9b39487b08d18524a02dfdfee377259a202fe [formerly 80f2178267c5ba456ae4cc6616ed4022a5aa58e4]
Former-commit-id: f5c86f73a91d0af415f1ec2c8e7bcd79541f55f3

@@
 
-Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
-We are curious to see the prediction difference between a vector-based models and current co-occurrance models on the StackOverflow dataset.
-Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
-Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, we are curious to see how a cognitively-plausible vector-based model performs at the task.
-
 A tag suggestion tool has also been deployed for the tex.stackexchange.org site. \parencite{LatexTags2013}
@@
 
-FIXME: \cite{Godin2013}
-This paragraph will discuss the topics-based approach.
+A topics-based approach was used by \textcite{Godin2013}, where the model predicted the most likely set of topics that were associated with a tweet, and tested if these topics matched the tweet by using human raters.
+The latent topics for the colleciton of tweets were generated using a Latent Dirichlet Allocation statistical methodology.
+This is similar to singular value decomposition in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance.
+Human graders rated how well the models generated topics for a tweet related to the content in the tweet.
+The LDA model performed well, and generated relevant topics.
 
-FIXME: Paragraph that no vector-based models have been attempted for Twitter.
-Bayesian model can be improved by customizing predicitons based on past hashtag use.
+However, a topic-generating model is not performing the same task as a hashtag-prediction model.
+The authors argued that these general topics could be used to categorize the post, in addition to the hashtags already used by the author in the post.
+It is unclear though if a user would want to label their tweet with these general topics.
+Perhaps they would rather use more specific hashtags to be more precise in how they label the tweet and how that tweet is associated with tweets within the rest of the community.
+In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
+Now it is certainly the case that the topics generated by a latent-based model may be similar, if not identical, to the hashtags used in the post.
+However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
+Why not take those topics already present in the data and build a model that produces them?
+In other words, the topics data are already in the tweet, so why not use them?
 
-FIXME: ditch this paragraph?
-To our knowledge, no model of user-created hashtags when composing tweets on Twitter has yet been built.
-There are two likely reasons for this:
-[1] Modeling user-chosen hashtags and building hashtag-recommendation systems has only explored over the past few years, and
-[2] Twitter is a large database of human-generated content, spanning a wide range of topics, containing a sizeable set of user-generated hashtags, so hashtag prediction for Twitter is a relatively hard problem.
-We are nonetheless curious if declarative memory retrieval models can be applied to this domain to generate useful hashtag predictions.
-The domain shares many similarities with environments where hashtag-recommendation systems have already been deployed:
-Google+'s microblogging service is essentially Twitter, but focused on a specific subset of users and topics.
-The StackExchange sites are also built from a large database of human generated content, where users have chosen specific tags to associate with posts.
-So we will evaluate how well an ACT-R-inspired Bayesian declarative memory model and a new vector-based model predict a user's chosen hashtag when composing a tweet on Twitter.
+\subsubsection{State of Research on StackOverflow and Twitter}
+
+Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
+We are curious to see the prediction difference between a vector-based models and current co-occurrance models on the StackOverflow dataset.
+Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
+Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, we are curious to see how a cognitively-plausible vector-based model performs at the task.
+
+The current state of research on Twitter hashtag prediction is that a mix of co-occurrence and latent-based models have been used to predict hashtags, and the results are encouraging.
+However, there has been very little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
+\textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
+The ACT-R declarative memory retrieval model takes a combined approach, and shows how past user behavior and current context combine to produce the most likely retrieval.
+Each component is summed together to generate a total activation, and then the likelihodd of chunks are ranked by that summed activation.
+This technique allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
+This type of weighted additive model that combines past user behavior and current tweet context to predict the most likely hashtags has not yet been explored for Twitter.
+
+Also, a vector-based retrieval model has not yet been tested on Twitter.
+One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
+It seems quite likely that the few words used just before a hashtag embedded in a tweet will be highly predictive of the hashtag chosen.
+To our knowledge, only bag-of-words models have been tested on Twitter to predict hashtags, where co-occurrance between words and hashtags are computed independently from where the word is located in the tweet.
+It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.
 

### 2013-08-28 writeups/proposal/proposal.tex
foo
Former-commit-id: 1fa0c7bc892c2fe90369cfab2a4624f620989511 [formerly a05c9c05efd573c31e26cd5cb55a88144019fd06]
Former-commit-id: 6d9af6b8a25a27c7bc825a5ffa3804b5a381a9a6

@@
 
-The current state of research on Twitter hashtag prediction is that a mix of co-occurrence and latent-based models have been used to predict hashtags, and the results are encouraging.
-However, there has been very little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
-\textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
-The ACT-R declarative memory retrieval model takes a combined approach, and shows how past user behavior and current context combine to produce the most likely retrieval.
-Each component is summed together to generate a total activation, and then the likelihodd of chunks are ranked by that summed activation.
-This technique allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
-This type of weighted additive model that combines past user behavior and current tweet context to predict the most likely hashtags has not yet been explored for Twitter.
-
 Also, a vector-based retrieval model has not yet been tested on Twitter.
@@
 It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.
+
+There has been a mix of co-occurrence and latent-based models created to predict hashtags use on Twitter, and the results are encouraging.
+However, there has been very little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
+\textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
+The ACT-R declarative memory retrieval theory shows how past user behavior and current context combine to produce the most likely retrieval.
+Each component is summed together to generate a total activation, and then the likelihodd of chunks are ranked by that summed activation.
+This technique allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
+This type of weighted additive model that combines past user behavior and current tweet context to predict the most likely hashtags has not yet been explored for Twitter.
 

### 2013-08-28 writeups/proposal/proposal.tex
foo
Former-commit-id: 3c53327b0c4a2563025003c51db18c07c3d89a52 [formerly bc7cd4c8412fc5562d212c56fc79b75fdeac0b87]
Former-commit-id: 9b5a12674b249c6284fb576bbb3102a8c62dfb7f

@@
 
-\title{Psyc 601 Multivariate Statistics Project}
-\shorttitle{Psyc 601 Project}
+\title{FIXME}
+\shorttitle{}
 
@@
 \abstract{
-  The base StackOverflow tag-prediction model from the Fall '12 Psychometrics course was improved by incorporating the words in the body of the post and the post author's specific tagging history.
-  We focused on hashtag creation as a specific human behavior, since understanding and modeling hashtag creation can lead to improved human-computer systems that can better identify users' goals and interests.
-  We used multivariate logistic regression statistical techniques to guide the model building process, and found that adding the two model terms improved model accuracy from 56\% to 69\%.
-  The model is a successful case showing that ACT-R's declarative memory retrieval equations scale, and are relevant to task domains that require large-scale knowledge stores.
+  FIXME
 }
 
-\keywords{StackOverflow, Machine Learning, ACT-R, Tagging, Large-Scale Semantic Memory}
+\keywords{FIXME}
 

### 2013-08-28 writeups/proposal/proposal.tex
foo
Former-commit-id: db40a64f155c5fdbdd37ef6cd30b13ceabba8714 [formerly 2c22df825f77559db5b662b556f138f3320d27c6]
Former-commit-id: cadab7c2f658b38996009f77c804caf809036110

@@
 
-TODO: Introduce PMI \cite{Farahat2004}
+\textcite{Farahat2004} introduced the Pointwise Mutual Information (PMI) index that also measures strength of association between two terms.
+It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
+The PMI index is included in Equation \eqref{eqPMI}.
 
-TODO: Comparison to ACT-R: Virtually equivalent for large corpora \cite{Farahat2004}
+\begin{equation}
+  \label{eqPMI}
+  PMI(y,x) = log \frac{p(y|x)}{p(y)}
+\end{equation}
+
+This PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
+Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.
 
@@
 
-\cite{Douglass2010}
+ACT-R's strength of association ($S_{ji}$) and the equivalent pointwise mutual information index ($PMI_{ji}$) have been shown to scale well for large corpora.
+\textcite{Douglass2010} implemented the large-scale version of $S_{ji}$ in Erlang.
+The wall-clock time required for a declarative memory retrieval request scaled linearly with the size of declarative memory with this implementation.
+Retrieval times were around one second with the largest declarative memory that contained over one million associations (co-occurrence pairs).
 
-TODO: SNIF-ACT \cite{Fu2007} \cite{Pirolli2003}
+The SNIF-ACT framework is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search \parencites{Fu2007,Pirolli2003}.
+The $S_{ji}$ values were based on a 118GB local database of 10 million pages of the Stanford Webbase Project.
+So this implementation did not fit into RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.
+
+The SNIF-ACT model uses the strength of association to measure the relationship between the words in a search query and the words contained in each query result.
+A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.
+The links that user chose in experimental testing were largely predictive of the strength of association measure.
+The predictive power of the SNIF-ACT model went above and beyond what would be predicted by simply assuming that the user always clicks on the first link in the search results page.
 

### 2013-08-28 writeups/proposal/proposal.tex
foo
Former-commit-id: 2d09514b331d087b9d249fe3b6e97d68f2ab2fc3 [formerly 238b1f973cf88ab3f4095687030d817ec640dc1b]
Former-commit-id: 2f4386587034d6f15ca04975f09a530885fb293e

@@
 
-The SNIF-ACT model uses the strength of association to measure the relationship between the words in a search query and the words contained in each query result.
+The SNIF-ACT model uses strength of association to measure the relationship between the words in a search query and the words contained in each query result.
 A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.

### 2013-08-29 writeups/proposal/proposal.tex
foo
Former-commit-id: 3ad48f38a10a778709ff1d1929028e4dc33321d6 [formerly 4dfc8d6f26931730a573dbb98b6ec658b57592ac]
Former-commit-id: c5ea866d6776b7f9c09e4ae42bea42cf7584a2ed

@@
   \label{eqPMI}
-  PMI(y,x) = log \frac{p(y|x)}{p(y)}
+  \mathit{PMI}(y,x) = log \frac{p(y|x)}{p(y)}
 \end{equation}
@@
 
-ACT-R's strength of association ($S_{ji}$) and the equivalent pointwise mutual information index ($PMI_{ji}$) have been shown to scale well for large corpora.
+The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
+It is important that this measure scales well enough to handle the sheer volume of information present in online social media datasets.
+Large-scale social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
+If one wants to capitilize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.
+
+ACT-R's strength of association ($S_{ji}$) and the equivalent pointwise mutual information index ($\mathit{PMI}_{ji}$) have been shown to scale well for large corpora.
 \textcite{Douglass2010} implemented the large-scale version of $S_{ji}$ in Erlang.

### 2013-08-29 writeups/proposal/proposal.tex
foo
Former-commit-id: 09e4866ec8c648dee8a89ac5244d61d99e726879 [formerly b0fef70c4733fdec6b003b04109efb764f6c22ef]
Former-commit-id: 2baa3d0c9b9571dd334bc0415e0a8ee69692193a

@@
 
-\textcite{Farahat2004} introduced the Pointwise Mutual Information (PMI) index that also measures strength of association between two terms.
+Pointwise Mutual Information (PMI) is a another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
 It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
@@
 
-This PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
+Further, this PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
 Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.
@@
 The SNIF-ACT framework is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search \parencites{Fu2007,Pirolli2003}.
-The $S_{ji}$ values were based on a 118GB local database of 10 million pages of the Stanford Webbase Project.
+The $S_{ji}$ values were based on a 118GB local database of 10 million pages of the Stanford Webbase Project \parencite{Farahat2004}.
 So this implementation did not fit into RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.

### 2013-08-29 writeups/proposal/proposal.tex
foo
Former-commit-id: ff6a0b4544b895a30e18ea33f345d23e35933300 [formerly 64c8a8ac64071fa1f832b6d8cd548efdd7faf0a5]
Former-commit-id: 84cf4d092c413f829428a4a537338534429a946f

@@
 
-TODO: Summarize \cite{Landauer1997}
+Latent Semantic Analysis \parencite{Landauer1997} is a technique that has mathematical roots in factor analysis and also measures strength of association between two terms.
+The process starts with a set of N documents.
+A word x document co-occurrence matrix is built by counting the number of times each word appears in each document.
+The dimensionality of the full word frequency co-occurrence matrix is then reduced and forced to be represented on a smaller set of dimensions.
+This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
+The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space, which would be similar to computing the cosine of two correlated eigenvectors.
+
+Usually information is lost when the data are restructered into a smaller set of latent dimensions.
+However it was shown that this process actually improves predictiveness and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
+They tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
+Model results were on par with a large sample of applicants to american universities from non-english speaking countries (64.4\% compared to 64.5\%).
+Model performance also improved when the original dimensionality of the data was reduced.
 
@@
 
+LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word x document co-occurrence matrix.
+The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
+The best way to remove a dimension while maintianing all variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
+For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangably accross documents, so these terms would be collapsed to a single dimension and frequncy counts would be pooled.
+By pooling co-occurrence counts the model is able to generate better predictions with less data.
+
 \subsubsection{Word Order}
+
+Although the LSA model has been shown to perform well as a way to measure similarity between two terms, one of the main criticisms is that it is a ``bag of words'' model.
+This means that the model does not account for or pay attention to word order when building the base word x document matrix or consequently, when performing the SVD dimension reduction.
+Other models that can account for word order have been shown to perform better than ``bag of words'' models on tasks that do not even appear to require grammatical structure such as word order to do well.
+For example, \textcite{Jones2007} compared a vector-based model called BEAGLE directly to LSA, training and testing both models on the same corpus used in \textcite{Landauer1997}.
+The context-only BEAGLE model (no word order) performed slightly better than the LSA model at the task (55.6\% vs. 55.3\%).
+However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to 57.5\%.
+It seems reasonable to expect that task performance will improve much more on tasks where word order is used (e.g., generating a Twitter hashtag in the middle of a sentence).
 
@@
 
-TODO: Compare with PMI, performance and issue of scale. \cite{Budiu2007}
+Another issue with LSA is the SVD matrix decomposition technique is computationally very expensive.
+As the size of the original word x document matrix increases, the SVD computation becomes more and more time consuming.
+If the domains where LSA is being applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
+However, with large-scale human-generated dataesets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
+\textcite{Budui2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size dataset.
+Instead, they compared the measures on the smaller TASA corpus created from \num{60527} samples of text from high school to early college textbooks.
+This is one of the available corpuses on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budui2007}.
+
+When \textcite{Budui2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
+However, once the dataset surpasses in size what is computationally feasable with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
+This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
+\textcite{Budui2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
+This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
+Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to 73.75\%, which is 10\% higher than the TASA-trained LSA.
+This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available informaiton to generate predictions.
+
+\subsubsection{Cost of Incremental Updating}
+
+Another issue with LSA is that adding additional observations to an already-computed reduced representation requires rerunning the SVD \textcite{Farahat2004}.
+Since computing the SVD is the most computationally-expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically being added.
+For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.
 

### 2013-08-29 writeups/proposal/proposal.tex
foo
Former-commit-id: b6b3fea5f8feeb504fae674ed9109d7e8dcd6295 [formerly 758c27e72fef213887a7fb6eaca4c3eb3c09a5aa]
Former-commit-id: 44b3fce806f9d7749471ac8b9d3b44f1d89b2ed5

@@
 However, with large-scale human-generated dataesets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
-\textcite{Budui2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size dataset.
+\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size dataset.
 Instead, they compared the measures on the smaller TASA corpus created from \num{60527} samples of text from high school to early college textbooks.
-This is one of the available corpuses on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budui2007}.
+This is one of the available corpuses on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.
 
-When \textcite{Budui2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
+When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
 However, once the dataset surpasses in size what is computationally feasable with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
 This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
-\textcite{Budui2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
+\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
 This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.

### 2013-08-30 writeups/proposal/proposal.tex
foo
Former-commit-id: 6f64ec51c5a212be9d94e8dcca38b420656a4bf8 [formerly 01f3ffd4837d4cba3d282174f29d3182618f215b]
Former-commit-id: 377d4ec6086f9fef22b6fa8b5c54324a1fae265c

@@
 
-TODO: Summarize \cite{Plate1995}
+``holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., chunk) as a vector.
+The representation of the concept is distributed across all elements of the vector.
+This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and saying that the distribution of co-occurrence counts across the rows in the column is the representation of that concept.
+However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.
+
+For vector-based systems, each word is represented by an environment vector $e_{j}$ that is nearly orthogonal to all other words' environment vectors.
+The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix, which is how vector-based systems represent information in a much more compact space.
+Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occured with that $e_{i}$ environment vector.
+These memory vectors can contain environment vectors from bag-of-words co-occurrence information ($c_{i}$) and word order information ($o_{i}$), which can be combined into a single representation ($m_{i}$).
+Over time, memory vectors accumulate a distributed representation of the most common environment vectors that co-occurred with them.
+
+\subsubsection{Retrieval Process}
+
+Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
+Decoding takes the memory vector for context and decodes it back into an environment vector.
+The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
+For example, assume that a user has written the word ``zend'' in a post.
+``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP}.
+Since the environment vector for \emph{PHP} was added to the memory vector for ``zend'' without any encoding, no decoding is necessary.
+So the cosine between the environment vector \emph{PHP} and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.
+
+Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
+Fo the \emph{PHP} and ``zend'' example, the context memory vector would consist of the unordered environment vector for ``zend''.
+Since the memory vector for \emph{PHP} contains the ``zend'' environment vector, it is likely that the correlation between these two memory vectors will be high.
+
+Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
+\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
+However, decoding does require an additional decoding step that resonance does not require, and depending on the decoding operation, this step may be computatinoally expensive.
 
 \subsubsection{Addressing Word Order}
+
+One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
+Representing word order is an example where the endoding and decoding operations are applied.
+As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
+Simply counting co-occurrences of the two immediately-preceding words to hashtags does not include the order information of the words in the representation.
+With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occured just prior to these 1-back words.
+Order information is lost if a simply co-occurrence counting technique is used.
+
+With vector-based systems, order informatino is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
+This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
+Convolution and inverse convolution (i.e,. circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
+This allows for retrieval requests such as ``playing \#$\Phi$'', ``just read \#$\Phi$'', and ``vector-based $\Phi$ systems''.
+
+For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
+Within this corpus, the phrase ``martin luther'' often occured.
+When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-back location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
+Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
+The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.
+
+Using decoding, the word-order memory vector for martin is deconvolved to the right ($o_{martin} \circledast \Phi$) which returns the environment vector for luther ($e_{luther}$).
+Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which returns the word-order memory vector for luther ($o_{luther}$).
+Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$''.
+
+New vectors generated by circular convolution are uncorrelated with all other environment vectors and have the same length.
+This allows the word order information and unordered context information for a chunk to be aggregated and represented in the same memory vector.
+Quite simply, the different types of memory vectors can be summed to create a single memory vector representation for each chunk (i.e., $m_{j} = c_{j} + o_{j}$).
 

### 2013-09-03 writeups/proposal/proposal.tex
Tue writing
Former-commit-id: 00c9b2afe86bb4b50c9fd490922cd84cd7fb7bc0 [formerly 9b8ae1b422371e3a7c729594e60aef13c546eaa8]
Former-commit-id: d0117cc8ca159d6aa9c1cf6e4176e38474170202

@@
 
+Vector-based represnetations compute chunk likelihoods by performing a correlation operation directly on the memory vectors.
+Each memory vector can be thought of as a column on a context word x tag word co-occurrence matrix used for a Bayesian representation.
+Further, the number of rows for a vector-based representation are fixed and set much smaller than the number of rows in the full word x tag co-occurrence matrix.
+Since Bayesian representations have been shown to scale to over hundreds of millions of word x tag co-occurrences \parencite{Stanley2013}, the compressed vector-based representations should scale as well or better.
+
 \subsubsection{BEAGLE}
 
-TODO: Sumarize Jones' model \cite{Jones2007}
+\textcite{Jones2007} created the BEAGLE vector-based memory representation.
+Each memory vector ($m_{j}$) is a summed representation of unordered and ordered word co-occurrences ($c_{j} + o_{j}$).
+This model uses circular convoluation and inverse correlation as a way to encode and decode word order.
+BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (55.6\% compared to 55.3\%).
+BEAGLE with unordered context and word order was also more correllated to the WordNet synonym database than LSA (-.311 compared to -.165).
+Further, it was shown that incorporating word-order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
+This model provided an efficient algorithm for using a single representation for unordered and ordered informaiton that showed performance similar to if not better than LSA.
 
@@
 
-TODO: Summarize \cite{Sahlgren2008}
+In order to represent word order information, a funciton must be used that converts the memory vectors for a set of words (``stack $\Phi$'') into a new uncorrelated environment vector.
+The BEAGLE model uses circular convolution and inverse correlation as the encoding and decoding operations.
+Howerver, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors.
+\textcite{Sahlgren2008} used a much simpler random permutation method for this operation.
+
+Random permutatinos use a large sparse vector of zeros with a few ones in random locations as an environment vector for each of the words.
+In order to create a new vector for a particular word that occurs one back ($\mathit{stack}^{-1}$), that word's environment vector ($e_{stack}$) is bit shifted to the left one position.
+This produces a new environment vector for the combination of ``stack'' and the one back position that is uncorrelated with the original ``stack'' environment vector and all other environment vectors.
+So if ``stack overflow'' occurred in a sentence, the one-back environment vector for ``stack'' would be added to the memory vector for ``overflow'' ($m_{overflow} = m_{overflow'} + e_{stack^{-1}}$).
+
+A retrieval for a word following ``stack'' (``stack $\Phi$'') is done by creating a memory vector out of all information in context.
+In this case, assume only a single piece of order information is available, that ``stack'' precedes the word to retrieve.
+So the memory vector for context is the environment vector for one-back ``stack'' ($m_{context} = e_{stack^{-1}}$).
+The memory vector for ``overflow'' will be correlated the highest with this context vector since it contains this vector.
 
@@
 
+RP used the simplest permutation possible to create a new environment vector.
+Performance is actually better than BEAGLE.
+Computational efficiency is better as well.
+Strong case for using RP over circular convulation for decoding / encoding.
 TODO: Summarize \cite{Recchia2010}

### 2013-09-03 writeups/proposal/proposal.tex
foo
Former-commit-id: f26fb1b725bef4cd978cc37c1be1cb7a60362071 [formerly dfec040bcb0a881d28c3ac61b3b2bc12060ac413]
Former-commit-id: 54970fa7dec0e5d2e84f5f34be93ff84f9f595aa

@@
 
-In order to represent word order information, a funciton must be used that converts the memory vectors for a set of words (``stack $\Phi$'') into a new uncorrelated environment vector.
+In order to represent word order information, a funciton must be used that converts an environment vector for a set of words and positions (e.g., ``stack $\Phi$'') into a new uncorrelated environment vector.
 The BEAGLE model uses circular convolution and inverse correlation as the encoding and decoding operations.
-Howerver, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors.
+Howerver, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors and positions.
 \textcite{Sahlgren2008} used a much simpler random permutation method for this operation.
 
-Random permutatinos use a large sparse vector of zeros with a few ones in random locations as an environment vector for each of the words.
-In order to create a new vector for a particular word that occurs one back ($\mathit{stack}^{-1}$), that word's environment vector ($e_{stack}$) is bit shifted to the left one position.
-This produces a new environment vector for the combination of ``stack'' and the one back position that is uncorrelated with the original ``stack'' environment vector and all other environment vectors.
-So if ``stack overflow'' occurred in a sentence, the one-back environment vector for ``stack'' would be added to the memory vector for ``overflow'' ($m_{overflow} = m_{overflow'} + e_{stack^{-1}}$).
+With random permutations, each word environment vector is a large sparse vector of zeros with a few $[1,-1]$s in random locations. 
+In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is bit shifted to the left one position.
+This produces a new environment vector for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
+
+For example, suppose that the words ``stack overflow'' appear in a sentence, and we want to represent that ``stack'' preceded the word ``overflow''.
+The ``stack'' environment vector ($e_{stack}$) is bit shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).
+That environment vector is then added to the memory vector for ``overflow'' ($m_{overflow} = m_{overflow'} + e_{stack^{-1}}$), so that ``overflow'' remembers that ``stack'' has preceded it in a sentence.
 
@@
 So the memory vector for context is the environment vector for one-back ``stack'' ($m_{context} = e_{stack^{-1}}$).
-The memory vector for ``overflow'' will be correlated the highest with this context vector since it contains this vector.
+The memory vector for ``overflow'' will be correlated the highest with this context vector since ``overflow'' contains this vector.
 

### 2013-09-04 writeups/proposal/proposal.tex
foo
Former-commit-id: 296f58991ce633975436b826ebf1a182cc812e39 [formerly e00b42e5cc441d28f544568f233e49adb9004d66]
Former-commit-id: d81d8556e5955210777b145e1a967098030a04d9

@@
 
-RP used the simplest permutation possible to create a new environment vector.
-Performance is actually better than BEAGLE.
-Computational efficiency is better as well.
-Strong case for using RP over circular convulation for decoding / encoding.
-TODO: Summarize \cite{Recchia2010}
+Random permutations use one of simplest permutations possible to create a new environment vector.
+This operation is much less costly than circular convoluation used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
+However, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.
+
+\textcite{Recchia2010} found that random permutations could use smaller vectors to produce equivalent performance (better compression), and scale to larger datasets (better computational efficiency).
+The vector x performance tradeoff was tested on a simple paired associative retrieval task. 
+Retrieval accuracy for the vector-based models remained high even after reducing each environment vector's length from \num{2048} to \num{1024}.
+This indicates that vector-based representations using random permutations compress better than those using circular convolution.
+
+To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the TOEFFL (among other tasks).
+The random-permutatino model was trained on \num{2.33} GB corpus.
+The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
+The random permutatino model performed as well or slightly better than teh BEAGLE model when trained on the smaller corpus.
+The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
+These results make a strong case to favor random-permutations over circular convoluation when implementing vector-based memory representations.
 

### 2013-09-05 writeups/proposal/proposal.tex
foo
Former-commit-id: 7922efbd649f4225819b31ce549cfed4622dfa96 [formerly b8d94ba71e576ce2ab2af3307a9d20aa51a1a8a4]
Former-commit-id: 1b77b5f4b1d4a4e3a8ca69f9461c87ac05b436c9
foo
Former-commit-id: f20f4f9e8a179e089561d27875f5c552f0c54b94 [formerly ffc30d1b00c5e671da1a477ce058166b06809490]
Former-commit-id: feb73204e741126f3f8287d2ba5049023c5de704

@@
-\documentclass[man]{apa6}
+\documentclass[man,floatsintext]{apa6}
 
@@
 
-TODO: Summarize \cite{Kanerva2000}
+Both random permutations and LSA calculate word similarity on a reduced matrix.
+For LSA, the original word x document matrix is compressed to a factor x document matrix and the number of factors are much less than the number of words.
+For the StackOverflow and Twitter datasets, the number of documents are the number of different tags.
+
+For random permutations, each tag is represented by a compressed vector with a length much less than the number of words.
+Note that if teh length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular column in the vector.
+If done in this way, building a matrix of all words' environment vectors x tags would recover the original word x document matrix \parencite{Kanerva2000}.
+However, each word's environment vector is compressed by using a length much less than the number of unique words and using a combination of a few [1,-1]s to represent the word.
+
+Information is lost with this compression in both cases: by starting with a compressed representation with random permutaitons or by reducing to a compressed representation with LSA.
+However, the way that the compression works in these two cases is not the same.
+LSA uses SVD to find words that have highly correlated tag occurance distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
+A vector-based approach like BEAGLE or random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
+Nonetheless, random permutations have been shown to perform similarl to if not better than LSA when trained on the TASA and tested on the TOEFFL \parencites{Sahlgren2008,Jones2007}.
+
+This is a conterintuitive result.
+It seems plausible that an SVD technique that takes into account word similarity when reducing teh matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
+This does not seem to be the case in testing however, as vector-based approaches like BEAGLE and random permutations perform just as well as the LSA when trained and tested on similar datasets.
+Nonetheless, it has been suggested to perform an SVD reduction technique on a compressed random-permutation vector \parencite{Kanerva2000}.
+This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrance counts from latent topics together.
 

### 2013-09-07 writeups/proposal/proposal.tex
foo
Former-commit-id: 9f687664e2b3ea19c0cd9caa81abca913524a392 [formerly 4f1d001fb1d33ee4ac474ac602e52fcbc345e3ac]
Former-commit-id: 5ec43722072896e9e7341abd4e9c9032476d948d

@@
 
-ACT-R \cite{Anderson2004} is a cognitive architecture that formalizes how each cognitive process of the brain (e.g., memory, learning, visual and motor) interacts to produce behavior.
+ACT-R \parencite{Anderson2004} is a cognitive architecture that formalizes how each cognitive process of the brain (e.g., memory, learning, visual and motor) interacts to produce behavior.
 The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up long-term declarative memory retrieval.
@@
 
-TODO: Summarize MALTA \cite{Rutledge2007}
+Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
+Each model computes an activation for each word to retrieve, given prior knowledge and context.
+Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
+For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
+\textcite{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect, which is a psychological phenomina that ACT-R's current Bayesian DM system can model as well.
 
-TODO: This approach can be used to model the fan effect \cite{Rutledge2008}
+\subsubsection{Base-Level Activation}
 
-TODO: However, there are open issues that need to be addressed:
-TODO: Can user prior be represented in vector-based memory systems?
-TODO: Can word order be represented in ACT-R?
-TODO: Do they both scale?
+However, vector-based models and ACT-R's current Bayesian DM model are not identical, and there are open issues with both models that need to be addressed before further evaluating the benefits of one over the other.
+For example, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
+More recent tweets on a news site like Twitter should be more relevant, for example.
+\textcite{Efron2011} tested this claim by evaluating a set of query likeihood models on returning relevant Twitter tweets from search queries.
+They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.
 
-\subsection{Incorporating Hashtag Recency}
+ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
+It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
+Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
+ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
+This type of information on a user's prior behavior is not being utilized with vector-based models currently.
+It may certainly be the case that vector-based models can be modified to incorporate base-level activation, and this idea will be pursued with this research.
 
-\subsubsection{ACT-R}
+\subsubsection{Word Order}
 
-TODO: ACT-R has a strong theory here.
-TODO: IR literature for Twitter domain uses a model similar to ACT-R \cite{Efron2011}
+One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occerrence information.
+With the ACT-R model for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
+It may be that multiple co-occurrence matrices must be used to represent word-order co-occurrences and unordered co-occurrences separately.
+However, it would be more parsimonious if these matrices could be combined into a single matrix representation, much like how this informatino is combined for vector-based models.
 
-\subsubsection{Vector-Based Models}
-
-TODO: Unclear how recency will fit in with random permutation vector-based model.
+\subsubsection{Activation Calculation}
 

### 2013-09-07 writeups/proposal/proposal.tex
foo
Former-commit-id: bde61dcc57f748f91014dcfe5058911bc5405c43 [formerly d88d27868bcc5839723ea7a0625a361743a1917b]
Former-commit-id: 9b48575b21bc57dd0fb76765e08c3a9fa1775101

@@
 
+Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
+Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
+These two computations are certainly not equivalent, even though they are most likely highly correlated with one another.
+It may be interesting to explore how model performance changes when one calculation is swapped out for another.
+For example, a vector-based model could be used to generate a compress (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
+Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
+It's unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
+But it may be the case that using Bayesian computation on a compressed vector-based model improves performance.
+
 \section{Models}
+
+Two models will be evaluated for hashtag prediction on both the Twitter and StackOverflow datasets:
+A vector-based model and Bayesian retrieval model.
+The random permutation model described in \textcite{Sahlgren2008} will be used for the vector-based model, since it has shown to perform better than the BEAGLE model \parencite{Recchia2010}.
+The modified ACT-R Bayesian model described in \textcite{Stanley2013} will be used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
+
+\subsubsection{ACT-R Bayesian Model}
+
+\begin{table}[!ht]
+  \caption{The StackOverflow tag prediction model}
+  \label{tabModACTRModel}
+  {\tabulinesep=1.2mm
+    \begin{tabu}{ll}
+      \hline
+      Common Name &  Equation \\
+      \hline
+      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} - O$ \\
+      Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
+      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
+      Co-occurrence &		$N_{ji} = \sum_{}^{}{observed_{ji}}$ \\
+      Observations &		$N = \sum_{}^{}{\sum_{}^{}{N_{ji}}}$ \\
+      Attentional Weight& 	$W_{j} = W \frac{E_{j}} {\sum_{}^{} {E_{j}}} $ \\
+      Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
+      Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
+      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
+      \hline
+    \end{tabu}
+  }
+\end{table}
+
+A formal description of the modified ACT-R Bayesian model from \textcite{Stanley2013} is included in Table \ref{tabModACTRModel}.
+The activation of each tag ($A_{i}$) is a function of four components:
+That tag's base level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
+This model has been modified from the standard ACT-R model (Table \ref{tabACT-RModel}) in three ways:
+Two separate contextual compoennts were used (words in title and body of post), an offset term was added, and an entropy metric was used to compute the attential weight ($W_{j}$).
+
+The offset term was added to normalize the mean activation across retrievals.
+This allowed the use of a logistical reggresion statistical technique to calibrate the weights for each of the terms.
+The title and body contextual components were separated since it seemed plausible that the words in the title would be better predictors of the tag used than the words in the body.
+It actually turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
+The entropy metric was used to compute the attentional weight for two reasons:
+Stop words could be included in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
+Using an entropy metric (as opposed to weighting each word equally in the post) allows the model to allocate limited attentional resources ($W$) to important contextual elements.
+
+\subsubsection{Random Permutation Model}
+
+\begin{table}[!ht]
+  \caption{Random Permutation Model}
+  \label{tabRandPermModel}
+  {\tabulinesep=1.2mm
+    \begin{tabu}{ll}
+      \hline
+      Common Name &  Equation \\
+      \hline
+      Activation &		$A_{i} = r(m_{C},m_{i})$ \\
+      Memory Vector &		$m_{i} = e_{i} + \sum_{j \in all past} c_{j} + \sum_{j \in all past} \sum_{k \in locations} o_{j,k}$ \\
+      Unordered Context &	$c_{i} = e_{i}$ \\
+      Ordered Context &		$o_{i,k} = e_{i^{-k}}$ \\
+      Context Memory Vector &	$m_{C} = \sum_{j \in C} c_{j} + \sum_{j \in C} \sum_{j \in locations} o_{j,k}$ \\
+      Environment Vector & 	$e_{i} = rand$ \\
+      \hline
+    \end{tabu}
+  }
+\end{table}
+
+A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
 

### 2013-09-09 writeups/proposal/proposal.tex
Mon commits
Former-commit-id: 0c1101609063f328d4a56e111b18fd615384cf26 [formerly 993e3e5e6f2061eea359b90ba771ed678c029624]
Former-commit-id: d118d6bf9eb0594f02316cf3bbf709256dfada58

@@
 However, users are not forming the same search queries that work so effectively on search engines like Google.
-Users of social media sites are searching for information that is happening now, in real time \parencite{Bauer2012}.
+A growing number of users on social media sites are searching for information that is happening now, in real time \parencites{Bauer2012,Jansen2011}.
 In other words, users on social-media sites want the fresh \emph{stream} of information about a particular topic.
@@
 We can help aid in their discovery of fresh and relevant information that match their goals on the site.
-
-FIXME: It is certainly the case that real-time search on these social media sites is growing \parencite{Jansen2011}.
 
@@
 
+\subsubsection{\textcite{Stanley2013}'s ACT-R DM Model for StackOverflow}
+
+\begin{table}[!ht]
+  \caption{The StackOverflow tag prediction model}
+  \label{tabModACTRModel}
+  {\tabulinesep=1.2mm
+    \begin{tabu}{ll}
+      \hline
+      Common Name &  Equation \\
+      \hline
+      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} - O$ \\
+      Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
+      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
+      Co-occurrence &		$N_{ji} = \sum_{}^{}{observed_{ji}}$ \\
+      Observations &		$N = \sum_{}^{}{\sum_{}^{}{N_{ji}}}$ \\
+      Attentional Weight& 	$W_{j} = W \frac{E_{j}} {\sum_{}^{} {E_{j}}} $ \\
+      Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
+      Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
+      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
+      \hline
+    \end{tabu}
+  }
+\end{table}
+
+A formal description of the modified ACT-R Bayesian model from \textcite{Stanley2013} is included in Table \ref{tabModACTRModel}.
+The activation of each tag ($A_{i}$) is a function of four components:
+That tag's base level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
+This model has been modified from the standard ACT-R model (Table \ref{tabACT-RModel}) in three ways:
+Two separate contextual compoennts were used (words in title and body of post), an offset term was added, and an entropy metric was used to compute the attential weight ($W_{j}$).
+
+The offset term was added to normalize the mean activation across retrievals.
+This allowed the use of a logistical reggresion statistical technique to calibrate the weights for each of the terms.
+The title and body contextual components were separated since it seemed plausible that the words in the title would be better predictors of the tag used than the words in the body.
+It actually turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
+The entropy metric was used to compute the attentional weight for two reasons:
+Stop words could be included in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
+Using an entropy metric (as opposed to weighting each word equally in the post) allows the model to allocate limited attentional resources ($W$) to important contextual elements.
+
 \subsection{Latent Semantic Analysis Memory Theory}
@@
 
-\subsubsection{Random Permutations}
+\subsubsection{Random Permutation Model}
 
@@
 
-With random permutations, each word environment vector is a large sparse vector of zeros with a few $[1,-1]$s in random locations. 
-In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is bit shifted to the left one position.
-This produces a new environment vector for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
+\begin{table}[!ht]
+  \caption{Random Permutation Model}
+  \label{tabRandPermModel}
+  {\tabulinesep=1.2mm
+    \begin{tabu}{ll}
+      \hline
+      Common Name &  Equation \\
+      \hline
+      Activation &		$A_{i} = r(m_{C},m_{i})$ \\
+      Memory Vector &		$m_{i} = \sum_{i \in all past} c_{i} + \sum_{i \in all past} \sum_{l \in locations} o_{i,l}$ \\
+      Unordered Context &	$c_{i} = e_{i}$ \\
+      Ordered Context &		$o_{i,l} = e_{i^{-l}}$ \\
+      Context Memory Vector &	$m_{C} = \sum_{i \in C} c_{i} + \sum_{i \in C} \sum_{l \in locations} o_{i,l}$ \\
+      Environment Vector & 	$e_{i} = rand$ \\
+      \hline
+    \end{tabu}
+  }
+\end{table}
+
+A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
+With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few $[1,-1]$s in random locations. 
+In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector ($e_{i}$) is bit shifted to the left one position ($^{-1}$).
+This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
 
@@
 In this case, assume only a single piece of order information is available, that ``stack'' precedes the word to retrieve.
-So the memory vector for context is the environment vector for one-back ``stack'' ($m_{context} = e_{stack^{-1}}$).
-The memory vector for ``overflow'' will be correlated the highest with this context vector since ``overflow'' contains this vector.
+So the memory vector for context is the environment vector for one-back ``stack'' ($m_{C} = e_{stack^{-1}}$).
+The memory vector for ``overflow'' ($m_{overflow}$) will be correlated the highest with this context vector ($m_{C}$) since ``overflow'' contains this vector.
 
@@
 A vector-based model and Bayesian retrieval model.
-The random permutation model described in \textcite{Sahlgren2008} will be used for the vector-based model, since it has shown to perform better than the BEAGLE model \parencite{Recchia2010}.
-The modified ACT-R Bayesian model described in \textcite{Stanley2013} will be used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
+The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} will be used for the vector-based model, since it has shown to perform better than the BEAGLE model \parencite{Recchia2010}.
+The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRMOdel} will be used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
 
-\subsubsection{ACT-R Bayesian Model}
-
-\begin{table}[!ht]
-  \caption{The StackOverflow tag prediction model}
-  \label{tabModACTRModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} - O$ \\
-      Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
-      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
-      Co-occurrence &		$N_{ji} = \sum_{}^{}{observed_{ji}}$ \\
-      Observations &		$N = \sum_{}^{}{\sum_{}^{}{N_{ji}}}$ \\
-      Attentional Weight& 	$W_{j} = W \frac{E_{j}} {\sum_{}^{} {E_{j}}} $ \\
-      Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
-      Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
-      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-A formal description of the modified ACT-R Bayesian model from \textcite{Stanley2013} is included in Table \ref{tabModACTRModel}.
-The activation of each tag ($A_{i}$) is a function of four components:
-That tag's base level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
-This model has been modified from the standard ACT-R model (Table \ref{tabACT-RModel}) in three ways:
-Two separate contextual compoennts were used (words in title and body of post), an offset term was added, and an entropy metric was used to compute the attential weight ($W_{j}$).
-
-The offset term was added to normalize the mean activation across retrievals.
-This allowed the use of a logistical reggresion statistical technique to calibrate the weights for each of the terms.
-The title and body contextual components were separated since it seemed plausible that the words in the title would be better predictors of the tag used than the words in the body.
-It actually turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
-The entropy metric was used to compute the attentional weight for two reasons:
-Stop words could be included in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
-Using an entropy metric (as opposed to weighting each word equally in the post) allows the model to allocate limited attentional resources ($W$) to important contextual elements.
-
-\subsubsection{Random Permutation Model}
-
-\begin{table}[!ht]
-  \caption{Random Permutation Model}
-  \label{tabRandPermModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Activation &		$A_{i} = r(m_{C},m_{i})$ \\
-      Memory Vector &		$m_{i} = e_{i} + \sum_{j \in all past} c_{j} + \sum_{j \in all past} \sum_{k \in locations} o_{j,k}$ \\
-      Unordered Context &	$c_{i} = e_{i}$ \\
-      Ordered Context &		$o_{i,k} = e_{i^{-k}}$ \\
-      Context Memory Vector &	$m_{C} = \sum_{j \in C} c_{j} + \sum_{j \in C} \sum_{j \in locations} o_{j,k}$ \\
-      Environment Vector & 	$e_{i} = rand$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
-
-TODO: Formal description of each model.
+These models will be the starting point for modeling tag use on the Twitter and StackOverflow sites.
+The models may be modified during testing if the data indicate that slight modifications are needed to improve model performance.
+Also there are planned modifications to each model as part of the research:
+incorporating user prior information into the vector-based model, and incorporating word order information into the Bayesian model.
 

### 2013-09-10 writeups/proposal/proposal.tex
Tue commits
Former-commit-id: 87d03def9e1f62687c64ecd435baafba1e4b3011 [formerly e7d234e46c49fcbd0d1e3744b699a90f67e4d322]
Former-commit-id: 13370119671ebe0ddde061fe58504b386456aa28

@@
 
+\usepackage{subfigure}
 \usepackage{graphicx}
@@
 \RequirePackage[l2tabu, orthodox]{nag}
-\graphicspath{{./figures/}} % Specifies the directory where pictures are stored
+\graphicspath{{./figures/dissertation/}} % Specifies the directory where pictures are stored
 
@@
 The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} will be used for the vector-based model, since it has shown to perform better than the BEAGLE model \parencite{Recchia2010}.
-The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRMOdel} will be used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
+The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel} will be used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
 
@@
 
+A modified ACT-R declarative memory retrieval model and random-permutation vector-based model will be evaluated on predicting user-chosen tags for Twitter tweets and StackOverflow posts.
+Parameters for each model will be calibrated on a large subset of the Twitter and StackOverflow datasets.
+Each model will be tested by evaluting the accuracy of model-chosen-tags on a fresh test subset of the datasets.
+Accuracy will be assessed by comparing model performance with the human data.
+
 \subsection{StackOverflow Dataset}
 
-\cite{DataDump2013}
+% FIXME: Verify these numbers
+The most recent quarterly-published StackOverflow dataset \parencite{DataDump2013} will be used to test the models on user-chosen tags for posts.
+This dataset contains around 5 million posts and \num{30000} unique tags.
+The site requires that for each post, the user must associate at least one tag with the post.
+The post author is allowed to choose multiple tags for the post.
+The average number of tags per post is around 3.
+
+% FIXME: Use a different example figure
+\begin{figure}[ht]
+  \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-screen-crop.pdf}}
+  \caption{Example post on stackoverflow.com}
+  \label{figSOExample}
+\end{figure}
+
+An example StackOverflow post is included in Figure \ref{figSOExample}.
+The author used 4 tags for this post: \emph{javascript}, \emph{firefox}, \emph{dom}, and \emph{svg}.
 
@@
 
-TODO: common-hashtags dataset
-TODO: popular-users dataset
+% FIXME: Verify these numbers
+Unlike StackOverflow, the entire Twitter dataset is not released to the public.
+Due to this and the sheer size of all of Twitter, subsets of the Twitter dataset must be collected.
+So two Twitter datasets will be collected for model evaluation:
+A common-hashtags and popular-users dataset.
+The common-hashtags dataset contains approximately 2 million tweets containing at least one of the 20 most popular hashtags.
+This dataset was collected over a period of 2 days.
+The purpose of this dataset is to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
 
-\subsection{Tokenizing and Lemmetizing Text}
+The popular-users dataset will contain all of the tweets from the top \num{2000} users on Twitter.
+This dataset will be used to explore the growth and decay characteristics of a hashtag for a single user.
+The temporal dynamics of ACT-R's decay rate model for a chunk will be evaluated against this data.
+As a result, the most appropriate model of user-centered frequency and recency effects (i.e., each hashtag's prior likelihood given a particular user) will be determined.
 
-TODO: Python NLP toolkit \cite{Bird2009}
+% FIXME: Use actual tweet figures
+\begin{figure}[ht]
+  \hfill
+  \subfigure[Title A]{\includegraphics[width=8cm]{tweet-screen-crop.pdf}}
+  \hfill
+  \subfigure[Title B]{\includegraphics[width=8cm]{tweet-screen-crop.pdf}}
+  \hfill
+  \caption{Example tweets on Twitter.com}
+  \label{figTweetExample}
+\end{figure}
+
+% FIXME: Write summary paragraph for twitter tweets figures.
+
+\subsection{Pre-processing the Text}
+
+The text from all dataset will be pre-processed using standard natural-language processing techniques.
+The Python NLP toolkit \parencite{Bird2009} will be used to tokenize (chunk) and lemmetize (convert to root form) each word.
+The standard chunking algorithm will be modified, so that it also ensures that hashtags and tag words (e.g., the word \emph{PHP} in a post) are not separated into multiple chunks.
+With the StackOverflow dataset, it is possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
+So tag words in the post and used tags for the post will be converted to their root tag when applicable.
 
@@
 TODO: Explore entropy weighting instead of stop word removal \cite{Dumais1991}
+
+\subsection{Attentuating Spam Posts}
+
+\cite{Mazzia2009}
 

### 2013-09-11 writeups/proposal/proposal.tex
Wed commits
Former-commit-id: 3c65ccd5d4fc2d121e9a7a35fcface675201cd56 [formerly 920a7eaa885bdd2e6c3fa519f201398a89aceb42]
Former-commit-id: 0eb5cb2299d224b4880586c19d232a5d72718c76

@@
 
-TODO: Explore entropy weighting instead of stop word removal \cite{Dumais1991}
+``Stop words'' are commonly-used words that co-occur with all tags (i.e., have little to no discriminating power).
+It is typically recommended \parencite{Bird2009} to remove stop words such as ``the'', ``a'', ``or'' when building the co-occurrence matrix.
+Otherwise the matrix may not scale as well since there are more words to keep track of. 
+Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occuring stop words.
+\textcite{Sahlgren2008} removed stop words for the random-permutation vector-based model, for example.
+\textcite{Jones2007} did as well for the circular-correlation model, but also mentioned that an entrop weighting technique has also been used previously to attentuate (rather than remove) stop words.
 
-\subsection{Attentuating Spam Posts}
+\textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset to reduce the weight of low-predictor stop words instead of removing them.
+This data-driven technique identifies stop words by calculating the discriminating power (i.e., entropy) of each word to predict any tag.
+The result maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
+It also seems less arbitrary than explicitly identifying a list of stop words to remove from the analysis.
+This technique worked well fo rthe StackOverflow model, and will be further explored for Twitter hashtag prediction.
+However, if it becomes computationally infeasable to compute this weighting for Twitter, then teh stop words will simply be removed, as was done for previous research.
 
-\cite{Mazzia2009}
+\subsection{Attentuating Twitter Spam Tweets}
+
+Although Twitter has an infrastructure to easily report and remove spam accounts, the spam that does get through can contain tweets with numerous hashtags.
+For example, some spam accounts generate tweets where all words in the tweet are hashtags, or generate multiple tweets with repeating content and hashtags.
+Feeding these spam tweets into statistical models that generate hashtag predictions can bais the models' chosen hashtags.
+
+\textcite{Mazzia2009} developed a novel approach to dealing with hashtag spam on Twitter.
+When training his model, it attentuates the importance of multiple tweets from teh same user that contain the same hashtags.
+As a user generates more than 10 tweets that contain the same hashtag, all tweets afterwards that contain this hashtag will be attentuated so that the overall weight is no more than that for 10 tweets.
+When testing the model, only at most 1 post from a user is included in the test set, since it is much easier to suggest relevant hashtags to a spammer than it is for an actual user.
+This exact technique may not be used for this analysis, but certainly some form of attentuating or identifying twitter spam tweets will be explored.
 

### 2013-09-16 writeups/proposal/proposal.tex
mon commits
Former-commit-id: a0ac5f98f983837c4a661dbbeaec4380d6f2e895 [formerly 21e712ea589220cefcf8b3bcb5b981b0abc55407]
Former-commit-id: e17be59415fb329f96100f8f2acfb1cdc287b40a

@@
 
-TODO: Use newest dataset
-TODO: Chunk tags
+\textcite{Stanley2013} developed a model for the StackOverflow dataset that was based on all posts up to April, 2011.
+This model will be used for this research, but the dataset will be updated to the most recent release.
+When moving from the older dataset to the newest dataset, several performance assessments will be made.
+
+First, model performance will be compared between the newest dataset and the Apr' 11 dataset to ensure that performance is consistent.
+Also this is a way to verify that the newest dataset was correctly pre-processed and loaded.
+
+Next, incremental improvements will be made to the pre-processing stages and performance will be compared after each change.
+For example, in \textcite{Stanley2013} the technique used to chunk the words does not ensure that tags appearing in the post remain as a single word unit.
+So a word like \emph{C\#} in a post was chunked as two separate words (\emph{C} and \emph{\#}).
+This reduces predictive power of the model since it muddles the highly-predictive co-occurrence observations where a tag appears in the post.
+So the chunking technique will be changed to take into account tags in the post, and its effect on performance will be evaluated.
 
@@
 
-TODO: Use stackoverflow and twitter popular-users dataset
+How does a user's prior hashtag use influence their likelihood of choosing that hashtag in the future?
+The ACT-R theory places strong constraints on the lifetime and growth characteristics of chunks in declarative memory.
+The more recently and frequently used chunks have higher prior activations, and are more likely to be needed in the future.
+It is an open question if this ACT-R theory of chunk growth and decay accurately reflects hashtag growth and decay for each user on Twitter and StackOverflow.
+
+The full StackOverflow dataset and Twitter popular-users dataset will be used to characterize tag growth and decay for a user on each site.
+Afterwards, the predicted results from the ACT-R theory will be compared against the observed data and evaluated.
+If the ACT-R theory of decay rate and frequency/recency affects fits the data reasonably well, then the prior activation components for the models developed in this research will be based on this theory.
+However, if the theory does not fit hashtag growth and decay for each user, then an alternative model will be proposed that is a better fit, and that model will be used instead.
 
@@
 
-TODO: Use stackoverflow dataset only for random permutations
+How can prior likelihood of a user choosing a particular tag be incorporated in the random permutations vector-based model?
+Currently vector-based models generate activations for each tag that are based purely on context.
+That is, if a particular tag has been used 10 times more often than the other tags by a specific user, that information is not included in a vector-based model.
+It seems plausible that performance should improve for vector-based models if this prior tag activation is included.
+Adding a prior activation term for the Bayesian StackOverflow model certainly improved performance \parencite{Stanley2013}.
+So a method to incorporate a user's prior tag history into a random permutation model will be developed and then tested on the full StackOverflow dataset and Twitter popular-users dataset.
 
@@
 
-TODO: Use twitter common-hashtags dataset only for ACT-R
+How can word-order information be incorporated into the ACT-R Bayesian model?
+Currently ACT-R uses a word co-occurrence matrix that does not include word position.
+A co-occurrence is counted if a word and a tag co-occur anywhere in a tweet or post.
+However for Twitter in particular, it may be the case that the few words immediately preceding hashtags are highly predictive, and the word position for those words should be included in the model.
+So a method to include word order into a Bayesian model will be developed.
+This updated model will then be evaluated on only teh Twitter common-hashtags dataset, since this dataset contains hashtags that occur in positions throughout the tweet (e.g., embedded within or at the end of a tweet).
+This model will not be evaluated on the StackOverflow dataset since chosen tags for a StackOverflow post are not embedded and mixed within a post, so word-order information is unavailable and most likely unused.
 

### 2013-09-17 writeups/proposal/proposal.tex
finished first draft
Former-commit-id: 0a609af09e5fbe0f9599cfde80a42e14214f1a5c [formerly 0e1d88f08804768b596525c983820f3d1c2d0ce4]
Former-commit-id: e0ea74523114417ca1d0541080dd6ed9353be3ca

@@
 
-\subsubsection{StackOverflow Dataset Modifications}
+Each proposed experiment aims to answer a specific research question in this study.
+The results from Experiment 1 will show how a user's prior hashtag use influences their likelihood of using that hashtag in the future, and test whether the observed data fit the ACT-R frequency/recency decay rate theory.
+The output from Experiment 1 will be a model of how a user's prior hashtag use influences the probability of using that hashtag in the future for both the Twitter and StackOverflow datasets.
+Experiment 2 will use the results from Experiment 1 and develop the best method to incorporate user's past tagging history in the random-permutations model, where user history is currently not being used.
+Experiment 3 will explore various methods for incorporating word order information into the ACT-R Bayesian model, where this information is currently not being used.
+The outputs of Experiment 2 and 3 are improved ACT-R and random-permutation models that have properly incorporated the strengths of each other: word-order strength from random permutations and user history from ACT-R.
+Finally, Experiments 4 and 5 will directly compare model accuracy between the two models and evaluate overall performance.
+Experiment 4 is focused primarily on hashtag prediction when no user prior history is available, while Experiment 5 will test how much performance improves when that history is available.
+
+\subsubsection{Preparation Experiment: StackOverflow Dataset Modifications}
 
@@
 
-\subsubsection{Analyze Recency for User-Generated Hashtags}
+\subsubsection{Experiment 1: Analyze Recency for User-Generated Hashtags}
 
@@
 
-\subsubsection{Incorporating Recency Into Random Permutations}
+\subsubsection{Experiment 2: Incorporating Recency Into Random Permutations}
 
@@
 
-\subsubsection{Incorporating Word Order Into ACT-R}
+\subsubsection{Experiment 3: Incorporating Word Order Into ACT-R}
 
@@
 
-\subsubsection{Top Hashtag Prediction}
+\subsubsection{Experiment 4: Top Hashtag Prediction}
 
-TODO: vector-based and act-r head to head on twitter common-hashtags
-TODO: head to head without word order on twitter common-hashtags
-TODO: head to head on stackoverflow
+After incorporating the relative strengths of each model into the other, the accuracy of the two models will be compared directly.
+Both models will be evaluated on predicting hashtags for tweets in the popular-hashtags dataset and predicting tags for posts on StackOverflow.
+However, the user prior hashtag use will not be available for the popular-hashtags dataset, since this subset was collected across the entire twitter user population.
 
-\subsubsection{User-Customized Hashtag Prediction}
+\subsubsection{Experiment 5: User-Customized Hashtag Prediction}
 
-TODO: head to head on twitter popular-users dataset
-TODO: head to head on stackoverflow
+In order to directly cmopare performance differences for each model when generating user-customized tag predictions, the models will be tested on the StackOverflow dataset and the Twitter popular-users dataset.
+Each dataset contains the entire tag history of either all users (StackOverflow) or the most popular users (Twitter).
+The main evaluation will be the amount of performance gain achieved when the models generate tags that are based on past user behavior compared to models that generate tags that are bases purely on context.
+
+\subsection{Summary}
+
+The end result of this research will be a better understanding of how a user's past tagging history influences future tag use,
+as well as two cognitively-plausible models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
+Overall performance differences between an ACT-R's Bayesian retrieval model and a random-permutation vector-based model will be directly compared.
+Improvements made in each model in order to increase performance at this tagging task may suggest model modifications that could be made to improve model performance for other tasks.
+Further, these improvements may uncover model modifications that could improve model fit across all tasks (i.e., suggest cognitive architectural modifications).
+In addition, if model accuracy is high enough, then these developed models may be applied to general tag recommendation systems for either Twitter or StackOverflow,
+and possibly for other sites where users generate and tag content.
 

### 2013-09-18 writeups/proposal/proposal.tex
wed commits
Former-commit-id: 69ccd125a21dd55ae812e80eb8210437fc9d3fdd [formerly 4b50a974c0765bd1ca1b1063e90a5c0f60079b99]
Former-commit-id: ce9bbad2fc4c772f86bf079581cf724015b0fdc9

@@
 Twitter is a micro-blogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
+A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#''.
 User-created hashtags on Twitter are embedded into the tweet, and can be used to connect the tweet to other tweets about the same topic through common hashtag use.
+After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
 Some recent and often-used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.
+
+% FIXME: Use better tweet figures
+\begin{figure}[!htbp]
+  \hfill
+  \subfigure[]{\includegraphics[width=8cm]{twitterTweetExample1.pdf}}
+  \hfill
+  \subfigure[]{\includegraphics[width=8cm]{twitterTweetExample2.pdf}}
+  \hfill
+  \caption{Example tweets on Twitter.com}
+  \label{figTweetExample}
+\end{figure}
+
+FIXME: Write summary paragraph for twitter tweets figures.
+
 StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
+A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
 User-chosen tags on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
 Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, or \emph{Common Lisp}.
+
+% FIXME: Use a different example figure
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-screen-crop.pdf}}}
+  \caption{Example post on stackoverflow.com}
+  \label{figSOExample}
+\end{figure}
+
+An example StackOverflow post is included in Figure \ref{figSOExample}.
+The site requires that for each post, the user must associate at least one tag with the post.
+The author used 4 tags for this post: \emph{javascript}, \emph{firefox}, \emph{dom}, and \emph{svg}, and the average number of tags per post is around 3.
 
@@
 The hashtag recommendation system deployed for the Google+ micro-blogging service is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
+The user interface for this recommenation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.
+
+\begin{figure}[!htbp]
+  \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample.pdf}}}
+  \caption{Example post on google+. The system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-generated post}
+  \label{figGoogle+Hashtags}
+\end{figure}
+
 The workflow and user interface provided for recommending hashtags for Google+ is similar to how a hashtag recommendation system for Twitter might look.
@@
 
-% FIXME: Verify these numbers
 The most recent quarterly-published StackOverflow dataset \parencite{DataDump2013} will be used to test the models on user-chosen tags for posts.
-This dataset contains around 5 million posts and \num{30000} unique tags.
-The site requires that for each post, the user must associate at least one tag with the post.
-The post author is allowed to choose multiple tags for the post.
-The average number of tags per post is around 3.
-
-% FIXME: Use a different example figure
-\begin{figure}[ht]
-  \resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-screen-crop.pdf}}
-  \caption{Example post on stackoverflow.com}
-  \label{figSOExample}
-\end{figure}
-
-An example StackOverflow post is included in Figure \ref{figSOExample}.
-The author used 4 tags for this post: \emph{javascript}, \emph{firefox}, \emph{dom}, and \emph{svg}.
+This dataset contains around \num{5.7} million posts and \num{34377} unique tags.
 
@@
 
-% FIXME: Verify these numbers
 Unlike StackOverflow, the entire Twitter dataset is not released to the public.
-Due to this and the sheer size of all of Twitter, subsets of the Twitter dataset must be collected.
-So two Twitter datasets will be collected for model evaluation:
-A common-hashtags and popular-users dataset.
-The common-hashtags dataset contains approximately 2 million tweets containing at least one of the 20 most popular hashtags.
-This dataset was collected over a period of 2 days.
+Due to this and the sheer size of Twitter, subsets of the Twitter dataset must be collected.
+So two Twitter datasets will be collected for model evaluation: a common-hashtags and popular-users dataset.
+The common-hashtags dataset will contain approximately 2 million tweets where at least one of the 100 most popular hashtags are used.
+This dataset will be collected in real time until the tweet quota is reached, which should take just a few days.
 The purpose of this dataset is to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
@@
 The popular-users dataset will contain all of the tweets from the top \num{2000} users on Twitter.
+This dataset will be collected by fetching all tweet history for each of the top users.
+Due to Twitter's query limitation for their API, it is estimated that this process will take about a week.
 This dataset will be used to explore the growth and decay characteristics of a hashtag for a single user.
@@
 As a result, the most appropriate model of user-centered frequency and recency effects (i.e., each hashtag's prior likelihood given a particular user) will be determined.
-
-% FIXME: Use actual tweet figures
-\begin{figure}[ht]
-  \hfill
-  \subfigure[Title A]{\includegraphics[width=8cm]{tweet-screen-crop.pdf}}
-  \hfill
-  \subfigure[Title B]{\includegraphics[width=8cm]{tweet-screen-crop.pdf}}
-  \hfill
-  \caption{Example tweets on Twitter.com}
-  \label{figTweetExample}
-\end{figure}
-
-% FIXME: Write summary paragraph for twitter tweets figures.
 

### 2013-09-20 writeups/proposal/proposal.tex
removed a few fixmes
Former-commit-id: 53bd91cd0417f301e2adc52aa359b82ae5b61af5 [formerly df127ee6acfa18f0e42b2a60b97489977f562984]
Former-commit-id: ecc91681c821569ec865b3689754fe4ca0616868

@@
 %\usepackage[compact]{titlesec}
-
 \usepackage{tabu}
@@
 
-% FIXME: Use better tweet figures
 \begin{figure}[!htbp]
+  {%
+    \setlength{\fboxsep}{0pt}%
+    \setlength{\fboxrule}{1pt}%
     \hfill
-  \subfigure[]{\includegraphics[width=8cm]{twitterTweetExample1.pdf}}
+    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample1-crop.pdf}}}
     \hfill
-  \subfigure[]{\includegraphics[width=8cm]{twitterTweetExample2.pdf}}
+    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample3-crop.pdf}}}
+    \hfill
+    \vfill
+    \hfill
+    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample4-crop.pdf}}}
+    \hfill
+    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample5-crop.pdf}}}
     \hfill
@@
     \label{figTweetExample}
+  }%
 \end{figure}
 
-FIXME: Write summary paragraph for twitter tweets figures.
+Some example tweets on Twitter are included in Figure \ref{figTweetExample}.
+Note how users both add hashtags to the end of a tweet as a summary of the message and emphasize important concepts words in the tweet by turning them into hashtags.
+That is, hashtags are both intermixed within the words of the tweet and added to the end of the message.
 
@@
 \begin{figure}[!htbp]
-  \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample.pdf}}}
+  \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample-crop.pdf}}}
   \caption{Example post on google+. The system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-generated post}

### 2013-09-23 writeups/proposal/proposal.tex
foo
Former-commit-id: dbae64aa0af5e242146aafca447fe3df2b3e5513 [formerly f40ec5ac8e5c9873ed2c8709838e6775dcd27d43]
Former-commit-id: 0ae74501bf620ef9dc4c1e44718b0fc82c4946a5
spell checked
Former-commit-id: bc27cf4a681892758c1b954e6710c74111fab7d9 [formerly 617e8fbab83471cee84a4d90efcc24eead05360f]
Former-commit-id: de5a1e16e71560fe047dc10948233971d2df65bc

@@
 
-TODO: Motivation
-
 Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp, are composed almost entirely of human-generated content.
@@
 However, a single user of a social media site is most likely interested in only a small fragment of this large amount of information.
-One general question to support the the user in this domain is:
+One general question to support the user in this domain is:
 How can we quickly and effectively connect users to the content that they care about?
@@
 One promising approach is to model and predict the user's goals when using and searching these social media sites \parencite{Rose2004}.
-If we have a better understanding of a user's goals, then we can tailor the information streams provided to the ones that are most inline with these goals.
+If we have a better understanding of a user's goals, then we can tailor the information streams provided to the ones that are most in line with these goals.
 For example, Twitter might suggest other users and hashtags for the user to follow, Facebook could suggest pages to like, Yelp could recommend businesses to check out, etc.
@@
 We will compare these two memory systems across two domains where users generate tags for posts: Twitter and StackOverflow.
-Twitter is a micro-blogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
+Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
 A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#''.
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visPost-2977094-screen-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{stackoverflowPostExample1-crop.pdf}}}
   \caption{Example post on stackoverflow.com}
@@
 However, they are also quite similar on several accounts:
-Both domains have emassed large amounts of user-generated content, users generate tags for posts when creating content on the site, and the user data from the datasets are publicly available for analysis.
+Both domains have amassed large amounts of user-generated content, users generate tags for posts when creating content on the site, and the user data from the datasets are publicly available for analysis.
 
 Also, studying user-based tag generation on these sites is a relevant task domain.
-Models that can accuratly predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these very popular sites.
+Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these very popular sites.
 These systems can help newer users by recommending proper tags for their newest content.
@@
 However, much less is known about hashtag lifetime within a single user.
-It may not be the case that a hashtag's lifecycle for a single user matches the lifecycle across users.
-Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into the a specific user's prior likelihood of hashtag use.
+It may not be the case that a hashtag's life cycle for a single user matches the life cycle across users.
+Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into a specific user's prior likelihood of hashtag use.
 So we will characterize the hashtag growth and decay cycle within users.
@@
 These hashtags should be customized to their specific interests and relevant to the content of the post that they are currently creating.
-This will undoubtably require a combination of two primary model components:
-[1] A user's prior likelihood of choosing a particular hashtag, given their prievously-generated content, and [2] the likelihood that a particular hashtag is related to the context of the post being generated.
+This will undoubtedly require a combination of two primary model components:
+[1] A user's prior likelihood of choosing a particular hashtag, given their previously-generated content, and [2] the likelihood that a particular hashtag is related to the context of the post being generated.
 
@@
 However, not only should the model properly take into account a user's prior hashtag use, but it should also generate new hashtag predictions when a user is most likely choosing a hashtag they have never used before.
-Calibrating the models to properly balance this exploitation of prior hashtag use vs. exploration of new hashtags will be a primary resaerch focus.
-The end goal is to have a model that can properly balance the user's prior tag with the contextual cues in the post in order to generate an accurate prediciton.
+Calibrating the models to properly balance this exploitation of prior hashtag use versus exploration of new hashtags will be a primary research focus.
+The end goal is to have a model that can properly balance the user's prior tag with the contextual cues in the post in order to generate an accurate prediction.
 
@@
 
-A model that predicts a user-chosen hashtag is a specific type of the more general recommendatino systems.
+A model that predicts a user-chosen hashtag is a specific type of the more general recommendation systems.
 These systems have grown in need with the growth of the web.
@@
 
-Recommendation systems based on the content of user-generated posts have been successfull in recommending people on Twitter for a user to follow.
+Recommendation systems based on the content of user-generated posts have been successful in recommending people on Twitter for a user to follow.
 \textcite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
@@
 The system then retrieved the most likely hashtags that were associated with that search query.
-The model is based on building a word co-occurance matrix of words x hashtags, and then uses that matrix to assign a similarity score for each hashtag when provided with a search query.
+The model is based on building a word co-occurrence matrix of words x hashtags, and then uses that matrix to assign a similarity score for each hashtag when provided with a search query.
 Participants graded how relevant the hashtags returned were to each search query.
@@
 In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user generated a post, and restrict the space of possible words to only tags.
-He tested the set of next-word predictino models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
-Interestingly, the Bayesian co-occurrance model outperformed the more complicated and computatinoally expensive Singular Value Decomposition and K-nearest-neighbors models.
+He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
+Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive Singular Value Decomposition and K-nearest-neighbors models.
 This model can accurately predict a user-chosen tag at 47\% accuracy.
@@
 \textcite{Stanley2013} also modeled a user's chosen tags for a post on the StackOverflow site.
-The model was based on ACT-R's declarative memory retrieval theory, which at the core is also a Bayesian co-occurrance model.
-However, this model scaled much better than the co-occurrance model presented in \textcite{Kuo2011}.
-Because of this, 100 times more posts were used to build the co-occurrance matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
+The model was based on ACT-R's declarative memory retrieval theory, which at the core is also a Bayesian co-occurrence model.
+However, this model scaled much better than the co-occurrence model presented in \textcite{Kuo2011}.
+Because of this, 100 times more posts were used to build the co-occurrence matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
 After scaling up the model, this model can successfully predict a user's tag for a post at 65\% accuracy.
-It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrance framework, since that is precisely the framework from which ACT-R's declarative memory system is built.
+It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is built.
 This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular task domain:
-predictng the user-chosen tags for human-generated content on social media sites.
+predicting the user-chosen tags for human-generated content on social media sites.
 
-A tag suggestion tool has also been deployed for the tex.stackexchange.org site. \parencite{LatexTags2013}
+A tag suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} site. \parencite{LatexTags2013}
 Both the \LaTeX site and StackOverflow are part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
@@
 
-The hashtag recommendation system deployed for the Google+ micro-blogging service is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
-The user interface for this recommenation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.
+The hashtag recommendation system deployed for the Google+ microblogging service is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
+The user interface for this recommendation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.
 
@@
   \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample-crop.pdf}}}
-  \caption{Example post on google+. The system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-generated post}
+  \caption{Example post on Google+. The system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-generated post}
   \label{figGoogle+Hashtags}
@@
 Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used from similar tweets.
-\textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurence-based similarity score between a composed tweet and the rest of the tweets in the database.
-\textcites{Zangerle2011, Kywe2012} use the Apache Lucine infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarites.
+\textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
+\textcites{Zangerle2011, Kywe2012} use the Apache Lucine infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarities.
 Regardless of the method to compute the similarity between tweets, each method collects the hashtags used from a set of the most similar tweets, then ranks the set, and presents the top 5-10 hashtags to the user.
@@
 The relationship is built between a tweet and its content, and then similar tweets are assumed to use similar hashtags.
-One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a vector representatino of the content in each tweet has to be maintained to later compute tweet similarity.
+One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a vector representation of the content in each tweet has to be maintained to later compute tweet similarity.
 The storage size for a hashtag-centered approach grows with the number of hashtags, but it seems reasonable that this will grow slower than the number of tweets and asymptote after collecting a large sample of tweets.
@@
 
-\textcite{Kywe2012} also customized their tag predictino model to the user's past hashtag use, and was one of the first to do so when developing a hashtag-prediction model for Twitter.
+\textcite{Kywe2012} also customized their tag prediction model to the user's past hashtag use, and was one of the first to do so when developing a hashtag-prediction model for Twitter.
 The set of recommended hashtags was the union of hashtags used in similar tweets and hashtags used by similar users.
-This was done by storing a content vector of hshtag use for each user (user-centered approach) alongside the content vectors of hashtag use for each tweet (tweet-centered approach).
+This was done by storing a content vector of hashtag use for each user (user-centered approach) alongside the content vectors of hashtag use for each tweet (tweet-centered approach).
 To generate recommended hashtags, the model would return the top-ranked hashtags from the hashtags used in 0-50 similar tweets and 0-4 similar users.
@@
 A topics-based approach was used by \textcite{Godin2013}, where the model predicted the most likely set of topics that were associated with a tweet, and tested if these topics matched the tweet by using human raters.
-The latent topics for the colleciton of tweets were generated using a Latent Dirichlet Allocation statistical methodology.
+The latent topics for the collection of tweets were generated using a Latent Dirichlet Allocation (LDA) statistical methodology.
 This is similar to singular value decomposition in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance.
@@
 Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
-We are curious to see the prediction difference between a vector-based models and current co-occurrance models on the StackOverflow dataset.
+We are curious to see the prediction difference between a vector-based models and current co-occurrence models on the StackOverflow dataset.
 Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
@@
 It seems quite likely that the few words used just before a hashtag embedded in a tweet will be highly predictive of the hashtag chosen.
-To our knowledge, only bag-of-words models have been tested on Twitter to predict hashtags, where co-occurrance between words and hashtags are computed independently from where the word is located in the tweet.
+To our knowledge, only bag-of-words models have been tested on Twitter to predict hashtags, where co-occurrence between words and hashtags are computed independently from where the word is located in the tweet.
 It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.
@@
 The ACT-R declarative memory retrieval theory shows how past user behavior and current context combine to produce the most likely retrieval.
-Each component is summed together to generate a total activation, and then the likelihodd of chunks are ranked by that summed activation.
+Each component is summed together to generate a total activation, and then the likelihood of chunks are ranked by that summed activation.
 This technique allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
@@
 what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
-Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likilihood of being retrieved again, given the prior number of times it has been needed.
-Those prior probabilities for each chunk are then adjusted for the current context, so that the posterior probabilities represent the likilihood that a chunk is needed, given prior odds and current environment state.
+Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of being retrieved again, given the prior number of times it has been needed.
+Those prior probabilities for each chunk are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and current environment state.
 
@@
 
-For some domains, it is reasonable to assume that the base-level activations of each chunk within a particualar window of interest are stable, and do not change within that window.
+For some domains, it is reasonable to assume that the base-level activations of each chunk within a particular window of interest are stable, and do not change within that window.
 Programming-language popularity over the past few years is a reasonable example of this.
-Although the popularity of various programming languages has certainly changed slightly over the past few years (.e.g, \emph{closure}'s growth), it is not the case that the changes have been drastic.
+Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., \emph{closure}'s growth), it is not the case that the changes have been drastic.
 \emph{C\#}, \emph{Python}, and \emph{Java} are still a few of the most popular languages, \emph{Common Lisp} has not gained or lost much ground in popularity over the past few years, etc.
@@
 Context for the Twitter domain would be the words in a Tweet.
-Association strenght between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
-The likelihood a chunk occurred with the current context ($p(i|j)$) over the likeihood that the chunk occured in any of the other contexts ($p(i|\overline{j})$).
+Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
+The likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
 For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
@@
 
-Using this assumption for large datasets, the interpretation of the context-adusted odds ratio becomes much simpler.
+Using this assumption for large datasets, the interpretation of the context-adjusted odds ratio becomes much simpler.
 If this ratio for a particular tag and context chunk is greater than one, then the log is positive, which means that this context has been observed more often with this tag than in general.
@@
 Large-scale social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
-If one wants to capitilize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.
+If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.
 
@@
 The SNIF-ACT framework is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search \parencites{Fu2007,Pirolli2003}.
-The $S_{ji}$ values were based on a 118GB local database of 10 million pages of the Stanford Webbase Project \parencite{Farahat2004}.
+The $S_{ji}$ values were based on a 118GB local database of 10 million pages of the Stanford WebBase Project \parencite{Farahat2004}.
 So this implementation did not fit into RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.
@@
 This model has been modified from the standard ACT-R model (Table \ref{tabACT-RModel}) in three ways:
-Two separate contextual compoennts were used (words in title and body of post), an offset term was added, and an entropy metric was used to compute the attential weight ($W_{j}$).
+Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy metric was used to compute the attentional weight ($W_{j}$).
 
 The offset term was added to normalize the mean activation across retrievals.
-This allowed the use of a logistical reggresion statistical technique to calibrate the weights for each of the terms.
+This allowed the use of a logistical regression statistical technique to calibrate the weights for each of the terms.
 The title and body contextual components were separated since it seemed plausible that the words in the title would be better predictors of the tag used than the words in the body.
@@
 
-Latent Semantic Analysis \parencite{Landauer1997} is a technique that has mathematical roots in factor analysis and also measures strength of association between two terms.
+Latent Semantic Analysis \parencite{Landauer1997} (LSA) is a technique that has mathematical roots in factor analysis and also measures strength of association between two terms.
 The process starts with a set of N documents.
@@
 
-Usually information is lost when the data are restructered into a smaller set of latent dimensions.
-However it was shown that this process actually improves predictiveness and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
+Usually information is lost when the data are restructured into a smaller set of latent dimensions.
+However it was shown that this process actually improves productiveness and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
 They tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
-Model results were on par with a large sample of applicants to american universities from non-english speaking countries (64.4\% compared to 64.5\%).
+Model results were on par with a large sample of applicants to american universities from non-English speaking countries (64.4\% compared to 64.5\%).
 Model performance also improved when the original dimensionality of the data was reduced.
@@
 The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
-The best way to remove a dimension while maintianing all variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
-For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangably accross documents, so these terms would be collapsed to a single dimension and frequncy counts would be pooled.
+The best way to remove a dimension while maintaining all variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
+For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
 By pooling co-occurrence counts the model is able to generate better predictions with less data.
@@
 If the domains where LSA is being applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
-However, with large-scale human-generated dataesets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
+However, with large-scale human-generated datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
 \textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size dataset.
-Instead, they compared the measures on the smaller TASA corpus created from \num{60527} samples of text from high school to early college textbooks.
-This is one of the available corpuses on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.
+Instead, they compared the measures on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
+This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.
 
 When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
-However, once the dataset surpasses in size what is computationally feasable with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
+However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
 This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
@@
 Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to 73.75\%, which is 10\% higher than the TASA-trained LSA.
-This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available informaiton to generate predictions.
+This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.
 
@@
 The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix, which is how vector-based systems represent information in a much more compact space.
-Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occured with that $e_{i}$ environment vector.
+Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occurred with that $e_{i}$ environment vector.
 These memory vectors can contain environment vectors from bag-of-words co-occurrence information ($c_{i}$) and word order information ($o_{i}$), which can be combined into a single representation ($m_{i}$).
@@
 Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
-Fo the \emph{PHP} and ``zend'' example, the context memory vector would consist of the unordered environment vector for ``zend''.
+For the \emph{PHP} and ``zend'' example, the context memory vector would consist of the unordered environment vector for ``zend''.
 Since the memory vector for \emph{PHP} contains the ``zend'' environment vector, it is likely that the correlation between these two memory vectors will be high.
@@
 \textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
-However, decoding does require an additional decoding step that resonance does not require, and depending on the decoding operation, this step may be computatinoally expensive.
+However, decoding does require an additional decoding step that resonance does not require, and depending on the decoding operation, this step may be computationally expensive.
 
@@
 One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
-Representing word order is an example where the endoding and decoding operations are applied.
+Representing word order is an example where the encoding and decoding operations are applied.
 As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
 Simply counting co-occurrences of the two immediately-preceding words to hashtags does not include the order information of the words in the representation.
-With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occured just prior to these 1-back words.
+With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
 Order information is lost if a simply co-occurrence counting technique is used.
 
-With vector-based systems, order informatino is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
+With vector-based systems, order information is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
 This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
-Convolution and inverse convolution (i.e,. circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
+Convolution and inverse convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
 This allows for retrieval requests such as ``playing \#$\Phi$'', ``just read \#$\Phi$'', and ``vector-based $\Phi$ systems''.
@@
 For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
-Within this corpus, the phrase ``martin luther'' often occured.
+Within this corpus, the phrase ``martin luther'' often occurred.
 When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-back location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
@@
 
-Vector-based represnetations compute chunk likelihoods by performing a correlation operation directly on the memory vectors.
+Vector-based representations compute chunk likelihoods by performing a correlation operation directly on the memory vectors.
 Each memory vector can be thought of as a column on a context word x tag word co-occurrence matrix used for a Bayesian representation.
@@
 Each memory vector ($m_{j}$) is a summed representation of unordered and ordered word co-occurrences ($c_{j} + o_{j}$).
-This model uses circular convoluation and inverse correlation as a way to encode and decode word order.
+This model uses circular convolution and inverse correlation as a way to encode and decode word order.
 BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (55.6\% compared to 55.3\%).
-BEAGLE with unordered context and word order was also more correllated to the WordNet synonym database than LSA (-.311 compared to -.165).
+BEAGLE with unordered context and word order was also more correlated to the WordNet synonym database than LSA (-.311 compared to -.165).
 Further, it was shown that incorporating word-order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
-This model provided an efficient algorithm for using a single representation for unordered and ordered informaiton that showed performance similar to if not better than LSA.
+This model provided an efficient algorithm for using a single representation for unordered and ordered information that showed performance similar to if not better than LSA.
 
@@
 
-In order to represent word order information, a funciton must be used that converts an environment vector for a set of words and positions (e.g., ``stack $\Phi$'') into a new uncorrelated environment vector.
+In order to represent word order information, a function must be used that converts an environment vector for a set of words and positions (e.g., ``stack $\Phi$'') into a new uncorrelated environment vector.
 The BEAGLE model uses circular convolution and inverse correlation as the encoding and decoding operations.
-Howerver, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors and positions.
+However, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors and positions.
 \textcite{Sahlgren2008} used a much simpler random permutation method for this operation.
@@
 Random permutations use one of simplest permutations possible to create a new environment vector.
-This operation is much less costly than circular convoluation used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
+This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
 However, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.
@@
 
-To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the TOEFFL (among other tasks).
-The random-permutatino model was trained on \num{2.33} GB corpus.
+To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFFL) assessment (among other tasks).
+The random-permutation model was trained on \num{2.33} GB corpus.
 The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
-The random permutatino model performed as well or slightly better than teh BEAGLE model when trained on the smaller corpus.
+The random permutation model performed as well or slightly better than the BEAGLE model when trained on the smaller corpus.
 The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
-These results make a strong case to favor random-permutations over circular convoluation when implementing vector-based memory representations.
+These results make a strong case to favor random-permutations over circular convolution when implementing vector-based memory representations.
 
@@
 For random permutations, each tag is represented by a compressed vector with a length much less than the number of words.
-Note that if teh length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular column in the vector.
+Note that if the length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular column in the vector.
 If done in this way, building a matrix of all words' environment vectors x tags would recover the original word x document matrix \parencite{Kanerva2000}.
@@
 
-Information is lost with this compression in both cases: by starting with a compressed representation with random permutaitons or by reducing to a compressed representation with LSA.
+Information is lost with this compression in both cases: by starting with a compressed representation with random permutations or by reducing to a compressed representation with LSA.
 However, the way that the compression works in these two cases is not the same.
-LSA uses SVD to find words that have highly correlated tag occurance distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
+LSA uses SVD to find words that have highly correlated tag occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
 A vector-based approach like BEAGLE or random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
-Nonetheless, random permutations have been shown to perform similarl to if not better than LSA when trained on the TASA and tested on the TOEFFL \parencites{Sahlgren2008,Jones2007}.
+Nonetheless, random permutations have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFFL \parencites{Sahlgren2008,Jones2007}.
 
-This is a conterintuitive result.
-It seems plausible that an SVD technique that takes into account word similarity when reducing teh matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
+This is a counterintuitive result.
+It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
 This does not seem to be the case in testing however, as vector-based approaches like BEAGLE and random permutations perform just as well as the LSA when trained and tested on similar datasets.
 Nonetheless, it has been suggested to perform an SVD reduction technique on a compressed random-permutation vector \parencite{Kanerva2000}.
-This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrance counts from latent topics together.
+This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.
 
@@
 For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
-\textcite{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect, which is a psychological phenomina that ACT-R's current Bayesian DM system can model as well.
+\textcite{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect, which is a psychological phenomena that ACT-R's current Bayesian DM system can model as well.
 
@@
 More recent tweets on a news site like Twitter should be more relevant, for example.
-\textcite{Efron2011} tested this claim by evaluating a set of query likeihood models on returning relevant Twitter tweets from search queries.
+\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
 They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.
@@
 
-One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occerrence information.
+One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
 With the ACT-R model for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
 It may be that multiple co-occurrence matrices must be used to represent word-order co-occurrences and unordered co-occurrences separately.
-However, it would be more parsimonious if these matrices could be combined into a single matrix representation, much like how this informatino is combined for vector-based models.
+However, it would be more parsimonious if these matrices could be combined into a single matrix representation, much like how this information is combined for vector-based models.
 
@@
 Parameters for each model will be calibrated on a large subset of the Twitter and StackOverflow datasets.
-Each model will be tested by evaluting the accuracy of model-chosen-tags on a fresh test subset of the datasets.
+Each model will be tested by evaluating the accuracy of model-chosen-tags on a fresh test subset of the datasets.
 Accuracy will be assessed by comparing model performance with the human data.
@@
 The text from all dataset will be pre-processed using standard natural-language processing techniques.
-The Python NLP toolkit \parencite{Bird2009} will be used to tokenize (chunk) and lemmetize (convert to root form) each word.
+The Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} will be used to tokenize (chunk) and lemmatize (convert to root form) each word.
 The standard chunking algorithm will be modified, so that it also ensures that hashtags and tag words (e.g., the word \emph{PHP} in a post) are not separated into multiple chunks.
@@
 Otherwise the matrix may not scale as well since there are more words to keep track of. 
-Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occuring stop words.
+Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.
 \textcite{Sahlgren2008} removed stop words for the random-permutation vector-based model, for example.
-\textcite{Jones2007} did as well for the circular-correlation model, but also mentioned that an entrop weighting technique has also been used previously to attentuate (rather than remove) stop words.
+\textcite{Jones2007} did as well for the circular-correlation model, but also mentioned that an entropy weighting technique has also been used previously to attenuate (rather than remove) stop words.
 
@@
 It also seems less arbitrary than explicitly identifying a list of stop words to remove from the analysis.
-This technique worked well fo rthe StackOverflow model, and will be further explored for Twitter hashtag prediction.
-However, if it becomes computationally infeasable to compute this weighting for Twitter, then teh stop words will simply be removed, as was done for previous research.
+This technique worked well for the StackOverflow model, and will be further explored for Twitter hashtag prediction.
+However, if it becomes computationally infeasible to compute this weighting for Twitter, then the stop words will simply be removed, as was done for previous research.
 
-\subsection{Attentuating Twitter Spam Tweets}
+\subsection{Attenuating Twitter Spam Tweets}
 
@@
 For example, some spam accounts generate tweets where all words in the tweet are hashtags, or generate multiple tweets with repeating content and hashtags.
-Feeding these spam tweets into statistical models that generate hashtag predictions can bais the models' chosen hashtags.
+Feeding these spam tweets into statistical models that generate hashtag predictions can bias the models' chosen hashtags.
 
 \textcite{Mazzia2009} developed a novel approach to dealing with hashtag spam on Twitter.
-When training his model, it attentuates the importance of multiple tweets from teh same user that contain the same hashtags.
-As a user generates more than 10 tweets that contain the same hashtag, all tweets afterwards that contain this hashtag will be attentuated so that the overall weight is no more than that for 10 tweets.
+When training his model, it attenuates the importance of multiple tweets from the same user that contain the same hashtags.
+As a user generates more than 10 tweets that contain the same hashtag, all tweets afterwards that contain this hashtag will be attenuated so that the overall weight is no more than that for 10 tweets.
 When testing the model, only at most 1 post from a user is included in the test set, since it is much easier to suggest relevant hashtags to a spammer than it is for an actual user.
-This exact technique may not be used for this analysis, but certainly some form of attentuating or identifying twitter spam tweets will be explored.
+This exact technique may not be used for this analysis, but certainly some form of attenuating or identifying twitter spam tweets will be explored.
 
@@
 So a method to include word order into a Bayesian model will be developed.
-This updated model will then be evaluated on only teh Twitter common-hashtags dataset, since this dataset contains hashtags that occur in positions throughout the tweet (e.g., embedded within or at the end of a tweet).
+This updated model will then be evaluated on only the Twitter common-hashtags dataset, since this dataset contains hashtags that occur in positions throughout the tweet (e.g., embedded within or at the end of a tweet).
 This model will not be evaluated on the StackOverflow dataset since chosen tags for a StackOverflow post are not embedded and mixed within a post, so word-order information is unavailable and most likely unused.
@@
 
-In order to directly cmopare performance differences for each model when generating user-customized tag predictions, the models will be tested on the StackOverflow dataset and the Twitter popular-users dataset.
+In order to directly compare performance differences for each model when generating user-customized tag predictions, the models will be tested on the StackOverflow dataset and the Twitter popular-users dataset.
 Each dataset contains the entire tag history of either all users (StackOverflow) or the most popular users (Twitter).

### 2013-09-23 writeups/proposal/proposal.tex
foo
Former-commit-id: b8777ce7e056bd9c551392f0f23cb74fad518c13 [formerly 5f55ea69ee36ef944bf81d579621e80903e73c70]
Former-commit-id: 7a2d377bd9bea6cfa6ccab4d68d644e201e429b4

@@
 
-% FIXME: Use a different example figure
+% FIXME: Use vector-graphics for this figure 
 \begin{figure}[!htbp]
@@
 The site requires that for each post, the user must associate at least one tag with the post.
-The author used 4 tags for this post: \emph{javascript}, \emph{firefox}, \emph{dom}, and \emph{svg}, and the average number of tags per post is around 3.
+The author used 4 tags for this post: \emph{javascript}, \emph{html}, \emph{internet-explorer}, and \emph{xhtml}, and the average number of tags per post is around 3.
 

### 2013-09-24 writeups/proposal/proposal.tex
tue commits
Former-commit-id: fe806d39bbb2a9c9a68c17390c8ae985b379a181 [formerly 7635790e5bd2845c18b5c1475cc80cef87f2a3b9]
Former-commit-id: fefed1923c006ff746adaa401698ddab0b43babf

@@
 
-Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp, are composed almost entirely of human-generated content.
-The amount of human-generated content on these sites is staggering, and continues to grow.
-Users of Twitter's microblogging service, for example, generate half a billion tweets per day \parencite{TwitterReport2012}.
+Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp, are composed almost entirely of human-created content.
+The amount of user content on these sites is staggering, and continues to grow.
+Users of Twitter's microblogging service for example, generate half a billion tweets per day \parencite{TwitterReport2012}.
 However, a single user of a social media site is most likely interested in only a small fragment of this large amount of information.
@@
 This can certainly be seen in the main user-interface views that these social media site provide:
-Twitter's homepage showing your followers' tweets, Facebook's homepage showing your friends' posts, StackOverflow's daily digest of posts for particular tags, etc.
+Twitter's homepage showing your followers' tweets, Facebook's homepage showing your friends' posts, StackOverflow's daily digest of posts for particular tags.
 
@@
 If we have a better understanding of a user's goals, then we can tailor the information streams provided to the ones that are most in line with these goals.
-For example, Twitter might suggest other users and hashtags for the user to follow, Facebook could suggest pages to like, Yelp could recommend businesses to check out, etc.
+For example, Twitter might suggest other users and hashtags for the user to follow, Facebook could suggest pages to like, Yelp could recommend businesses to check out.
 Researchers are currently looking for ways to automate the goal-identification task for more general search queries like those generated for Google \parencites{Jansen2008}{Lee2005}.
 
-However, social media sites have an advantage for identifying user goals over traditional search engines:
-human-generated content.
-Twitter does not have only the tweet that a user is currently composing to try and identify that user's goals.
-Rather, the site has all of the user's previously-generated content, and can use that information to provide a much better prediction of a user's goals.
+However, social media sites have an advantage for identifying user goals over traditional search engines: human-created content.
+Twitter has more than just the tweet that a user is currently composing to try and identify that user's goals.
+The site also has all of the user's previous content, and can use that information to provide a much better prediction of a user's goals.
 Further, sites like Twitter, StackOverflow, Google+, and Facebook support ways for the user to explicitly identify their precise goals when creating content: Hashtags.
@@
 
-These hashtags relate to a user's goals and interests because they are human-generated content, and a form of human-based document tagging \parencite{Chang2010}.
+These hashtags relate to a user's goals and interests because they are a form of human-based document tagging \parencite{Chang2010}.
 Further, people are using hashtags as a way to create information streams on social media sites \parencite{Kwak2010}.
-Twitter users, for example, commonly create hashtags for upcoming political events, such as debates and races \parencite{Diakopoulos2010}.
+Twitter users for example, commonly create and use hashtags for upcoming political events that interst them, such as debates and races \parencite{Diakopoulos2010}.
 So if we can predict what hashtags a user is interested in, we have a better understanding of the information streams that they care about.
@@
 
-The core research question we are interested in is:
-What kinds of cognitively-plausible models can best predict the human-generated tags that a user on a social media site is interested in?
+The core research question I am interested in is:
+What kinds of cognitively-plausible models can best predict the human-created tags that a user on a social media site creates and uses?
 This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
 This process can be framed as a memory retrieval problem.
-That is, each user on a social media site has created content that contain a set of human-generated tags within the context of each post.
+That is, each user on a social media site has created content that contain a set of human-created tags within the context of each post.
 The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.
@@
 Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
-Two of the current state-of-the-art memory models are based on a co-occurrence methodology, and we'll be comparing these two models for hashtag prediction:
+Two of the current state-of-the-art memory models are based on a co-occurrence methodology, and I'll be comparing these two models for hashtag prediction:
 ACT-R's declarative memory retrieval theory and random permutation vector-based memory systems.
@@
 
-We will compare these two memory systems across two domains where users generate tags for posts: Twitter and StackOverflow.
+I will compare these two memory systems across two domains where users generate tags for posts: Twitter and StackOverflow.
 Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
 A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#''.
-User-created hashtags on Twitter are embedded into the tweet, and can be used to connect the tweet to other tweets about the same topic through common hashtag use.
 After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
+So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
 Some recent and often-used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.
@@
 
+Alongside testing these declarative memory retrieval models on Twitter hashtag use, they will be tested on tag use for StackOverflow posts.
 StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
@@
 An example StackOverflow post is included in Figure \ref{figSOExample}.
-The site requires that for each post, the user must associate at least one tag with the post.
-The author used 4 tags for this post: \emph{javascript}, \emph{html}, \emph{internet-explorer}, and \emph{xhtml}, and the average number of tags per post is around 3.
+The site requires that for each post the user must associate at least one tag with the post, and the average number of tags per post is around 3.
+The author used 4 tags for this post: \emph{javascript}, \emph{html}, \emph{internet-explorer}, and \emph{xhtml}.
 
-These two datasets were chosen because the content of the human-generated content is quite different between them, and we are interested in retrieval models that generalize across different task domains.
+The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and I am interested in retrieval models that generalize across different task domains.
 However, they are also quite similar on several accounts:
-Both domains have amassed large amounts of user-generated content, users generate tags for posts when creating content on the site, and the user data from the datasets are publicly available for analysis.
+Both domains have amassed large amounts of user-created content, users generate tags for posts when creating content on the site (hashtags and tags for Twitter and StackOverflow respectively),
+and the user data from the datasets are publicly available for analysis.
 
@@
 Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these very popular sites.
-These systems can help newer users by recommending proper tags for their newest content.
+These systems can also help newer users by recommending proper tags for their newest content.
 On the StackOverflow site for example, experts often subscribe to specific hashtags that interest them, and receive a daily digest of posts that are tagged with those hashtags.
@@
 ACT-R's declarative memory and vector-based memory systems are substantially different in their implementation.
-Nonetheless, both can successfully model the classic fan effect phenomena found in word pairing experiments \parencite{Rutledge2008}.
-We will extend the work done by \textcite{Rutledge2008} by formally comparing the models on large-scale hashtag prediction tasks.
+Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., the ``fan effect'' \parencite{Rutledge2008}).
+I will extend the work done by \textcite{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
 We'll also replace the vector-based model used by \textcite{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
@@
 However, it is unclear and unexplored how word order can be represented in the ACT-R declarative memory theory.
-We will explore how the word-order strength of vector-based models can be incorporated into the ACT-R theory when testing the models on hashtag prediction tasks.
+I will explore how the word-order strength of vector-based models can be incorporated into the ACT-R theory when testing the models on hashtag prediction tasks.
 
-ACT-R's declarative memory system has a strong theory on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2004}.
+ACT-R's declarative memory system places strong constraints on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2004}.
+This decay rate equation formalizes how recency and frequency of hashtag use relate to the prior probability that a hashtag will be retrieved.
 However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.
-We will explore how a user's prior knowledge influences the likelihood that a particular hashtag is generated for both ACT-R's memory theory and vector-based models.
+I will explore how a user's prior knowledge influences the likelihood that a particular hashtag is chosen for both ACT-R's memory theory and vector-based models.
 
@@
 However, much less is known about hashtag lifetime within a single user.
-It may not be the case that a hashtag's life cycle for a single user matches the life cycle across users.
+It may not necessarily follow that a hashtag's life cycle for a single user matches the life cycle across users.
 Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into a specific user's prior likelihood of hashtag use.
-So we will characterize the hashtag growth and decay cycle within users.
+So I will characterize the hashtag growth and decay cycle for a single user.
 The expectation is that this cycle will match ACT-R's decay rate used in the declarative memory system to compute a particular memory item's prior.
-This decay rate equation formalizes how recency and frequency of hashtag use relate to the prior probability that a hashtag will be retrieved.
 
@@
 This will undoubtedly require a combination of two primary model components:
-[1] A user's prior likelihood of choosing a particular hashtag, given their previously-generated content, and [2] the likelihood that a particular hashtag is related to the context of the post being generated.
+[1] A user's prior likelihood of choosing a particular hashtag, given their previous content, and [2] the likelihood that a particular hashtag is related to the context of the post being created.
 
@@
 
-A model that predicts a user-chosen hashtag is a specific type of the more general recommendation systems.
-These systems have grown in need with the growth of the web.
+A model that predicts a user-chosen hashtag is a specific type of a recommendation system.
+These general recommendation systems have grown in need with the growth of the web.
 Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
@@
 This tool recommends to the user movies and television episodes that most likely match their specific interests.
+
 Netflix wanted to improve the accuracy of their recommendation system, so they held a competition where teams could try to improve on the accuracy of the service.
-
-The winning team utilized a singular value decomposition (SVD) method to identify particular dimensions of the entire content library (e.g., genre, time period, previously watched) that best predicted user preferences.
-This is a dimensionality-reducing technique that aims to filter out noise by forcing the data to be represented from a smaller number of dimensions than were originally present.
+The winning team utilized a statistical model that performed a type of factor analysis to identify important dimensions in the data (e.g., drama, action, cult classic) and place each user along each dimension.
 This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
-By using an SVD technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.
+By using a factor-analysis technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.
 
@@
 
-Recommendation systems based on the content of user-generated posts have been successful in recommending people on Twitter for a user to follow.
+Recommendation systems based on the content of user-created posts have been successful in recommending people on Twitter for a user to follow.
 \textcite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
@@
 Google has even deployed a hashtag-recommendation model to help users properly tag posts for their Google+ social media site \parencite{GoogleKeynote2013}.
-However, to our knowledge a hashtag-recommendation model for Twitter tweets has not yet been attempted.
+A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} site.
+Models for Twitter and StackOverflow are also being proposed and tested, but none have yet been deployed.
 
@@
 His set of models were originally designed for next-word prediction for large text document collections.
-In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user generated a post, and restrict the space of possible words to only tags.
+In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user created a post, and restrict the space of possible words to only tags.
 He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
@@
 This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular task domain:
-predicting the user-chosen tags for human-generated content on social media sites.
+predicting the user-chosen tags for human-created content on social media sites.
 
-A tag suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} site. \parencite{LatexTags2013}
+% FIXME: Add figure for LaTeX tag suggestion example
+
+A tag suggestion tool has been deployed for the \url{http://tex.stackexchange.org} site \parencite{LatexTags2013}.
 Both the \LaTeX site and StackOverflow are part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
@@
   \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample-crop.pdf}}}
-  \caption{Example post on Google+. The system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-generated post}
+  \caption{Example post on Google+. The system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-created post}
   \label{figGoogle+Hashtags}
@@
 Google is apparently highly confident in the accuracy of their system.
-On the Latex site, the User Interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
-For Google+, the User Interface automatically provides tags for a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
-If tag accuracy is high enough, the Google+ technique is certainly favorable, since the user does not have to spend time figuring out which tags best represent this post.
-However, if the model makes even just a few errors, users may become annoyed and frustrated with the recommendation system, since they will have to explicitly correct it every time they generate a post.
+On the Latex site, the user interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
+For Google+, the user interface automatically provides tags for a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
+If tag accuracy is high enough, the Google+ technique is certainly favorable since the user does not have to spend time figuring out which tags best represent this post.
+However, if the model makes even just a few errors, users may become annoyed and frustrated with the recommendation system since they will have to explicitly correct it every time they generate a post.
 
@@
 \parencite{Mazzia2009}
-The model present by \textcite{Mazzia2009} is very similar to an ACT-R declarative retrieval model.
+The model presented by \textcite{Mazzia2009} is very similar to an ACT-R declarative retrieval model.
 The main potential difference is that the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
@@
 Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
-We are curious to see the prediction difference between a vector-based models and current co-occurrence models on the StackOverflow dataset.
+I am curious to see the prediction difference between a vector-based models and current co-occurrence models on the StackOverflow dataset.
 Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
-Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, we are curious to see how a cognitively-plausible vector-based model performs at the task.
+Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, I am curious to see how a cognitively-plausible vector-based model performs at the task.
 
@@
 If the domains where LSA is being applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
-However, with large-scale human-generated datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
+However, with large-scale human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
 \textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size dataset.
@@
 
-For example, suppose that the words ``stack overflow'' appear in a sentence, and we want to represent that ``stack'' preceded the word ``overflow''.
+For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow''.
 The ``stack'' environment vector ($e_{stack}$) is bit shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).

### 2013-09-25 writeups/proposal/proposal.tex
Wed commits
Former-commit-id: 8d0a7b06f11f1ef3ca5f2f6b6033b416509c1333 [formerly 44f691c9e6f2e8a528aed1ceccf3a197dfb55ee3]
Former-commit-id: 433372b72146985b9a273f575872e5f401eb5d5c

@@
 The latent topics for the collection of tweets were generated using a Latent Dirichlet Allocation (LDA) statistical methodology.
-This is similar to singular value decomposition in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance.
+This is similar to a factor analysis approach in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance.
 Human graders rated how well the models generated topics for a tweet related to the content in the tweet.
 The LDA model performed well, and generated relevant topics.
+When the model produced five most-likely hashtags for each tweet, 80\% of the time at least one of those hashtags was suitable (as determined by human raters).
 
@@
 In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
-Now it is certainly the case that the topics generated by a latent-based model may be similar, if not identical, to the hashtags used in the post.
+Now it is certainly the case that the topics generated by a latent-based model may be similar to the hashtags used in the post.
 However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
@@
 Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
-I am curious to see the prediction difference between a vector-based models and current co-occurrence models on the StackOverflow dataset.
+I am curious to see the prediction difference between a vector-based model and current co-occurrence models on the StackOverflow dataset.
 Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
-Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, I am curious to see how a cognitively-plausible vector-based model performs at the task.
+Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, it will be interesting to see how a cognitively-plausible vector-based model performs at the task.
 
@@
 One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
-It seems quite likely that the few words used just before a hashtag embedded in a tweet will be highly predictive of the hashtag chosen.
-To our knowledge, only bag-of-words models have been tested on Twitter to predict hashtags, where co-occurrence between words and hashtags are computed independently from where the word is located in the tweet.
+It seems quite likely that words used just before a hashtag in a tweet (combined with their position) will be highly predictive of the hashtag chosen.
+To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
 It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.
@@
 The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up long-term declarative memory retrieval.
-The equations that make up this system can be described through a rational analysis of long-term memory retrieval.
+The equations that make up this system can be derived through a rational analysis of long-term memory retrieval.
 That is, given the task of retrieving a chunk of information from long-term declarative memory,
@@
 what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
-Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of being retrieved again, given the prior number of times it has been needed.
-Those prior probabilities for each chunk are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and current environment state.
+Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of being retrieved again, given the prior history of exposure to the chunk.
+These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adusted by current environment state.
 
@@
 The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base level activation ($B_{i}$) and strength of association ($S_{ji}$).
-The recall probability that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
+The recall probability ($P_{i}$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
 
@@
 
-For some domains, it is reasonable to assume that the base-level activations of each chunk within a particular window of interest are stable, and do not change within that window.
+For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable, and do not change within that window.
 Programming-language popularity over the past few years is a reasonable example of this.
 Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., \emph{closure}'s growth), it is not the case that the changes have been drastic.
-\emph{C\#}, \emph{Python}, and \emph{Java} are still a few of the most popular languages, \emph{Common Lisp} has not gained or lost much ground in popularity over the past few years, etc.
+\emph{C\#}, \emph{Python}, and \emph{Java} are still a few of the most popular languages, \emph{Common Lisp} has not gained or lost much ground in popularity over the past few years.
 
 Further, if this is a reasonable assumption to make, it greatly simplifies the computation of the base-level activation for chunks, and can turn the computation into a tractable problem for large datasets.
-In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log odds ratio of being exposed to a chunk.
+In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log odds ratio of being exposed to that chunk compared to all other chunk exposures.
 For example, if the StackOverflow tag \emph{PHP} has been used four times as often as the tag \emph{Common Lisp}, then the prior odds of needing \emph{PHP} again is $\frac{.8}{.2}=4$ times that of \emph{Common Lisp}.
@@
 
-Using this assumption for large datasets, the interpretation of the context-adjusted odds ratio becomes much simpler.
+Using this assumption for large datasets ($S_{ji} = log \frac{p(i|j)}{p(i)}$), the interpretation of the context-adjusted odds ratio becomes much simpler.
 If this ratio for a particular tag and context chunk is greater than one, then the log is positive, which means that this context has been observed more often with this tag than in general.
@@
 
-Further, this PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
+Further, this PMI equation is actually identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
 Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.
+That is, the PMI equation \emph{is} the simplified form of ACT-R's strength of association index ($S_{ji}$).
 
@@
 The $S_{ji}$ values were based on a 118GB local database of 10 million pages of the Stanford WebBase Project \parencite{Farahat2004}.
-So this implementation did not fit into RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.
+So this implementation did not fit completely in RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.
 
-The SNIF-ACT model uses strength of association to measure the relationship between the words in a search query and the words contained in each query result.
-A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.
-The links that user chose in experimental testing were largely predictive of the strength of association measure.
-The predictive power of the SNIF-ACT model went above and beyond what would be predicted by simply assuming that the user always clicks on the first link in the search results page.
+%TODO: This paragraph doesn't fit well after the previous one, and in the scaling section
+%The SNIF-ACT model uses strength of association to measure the relationship between the words in a search query and the words contained in each query result.
+%A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.
+%The links that user chose in experimental testing were largely predictive of the strength of association measure.
+%The predictive power of the SNIF-ACT model went above and beyond what would be predicted by simply assuming that the user always clicks on the first link in the search results page.
 
@@
 This model has been modified from the standard ACT-R model (Table \ref{tabACT-RModel}) in three ways:
-Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy metric was used to compute the attentional weight ($W_{j}$).
+Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy measure was used to compute the attentional weight ($W_{j}$).
 
@@
 It actually turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
-The entropy metric was used to compute the attentional weight for two reasons:
-Stop words could be included in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
-Using an entropy metric (as opposed to weighting each word equally in the post) allows the model to allocate limited attentional resources ($W$) to important contextual elements.
+The entropy measure was used to compute the attentional weight for two reasons:
+Stop words could be statistically derived and attentuated in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
+Using an entropy measure (as opposed to weighting each word equally in the post) allows the model to allocate limited attentional resources ($W$) to important contextual elements.
 
@@
 This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
-The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space, which would be similar to computing the cosine of two correlated eigenvectors.
+The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
 
@@
 Model results were on par with a large sample of applicants to american universities from non-English speaking countries (64.4\% compared to 64.5\%).
-Model performance also improved when the original dimensionality of the data was reduced.
+Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
+Afterwards further reductions in the dimensionality of the data began to degrade performance.
 
@@
 When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
+This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
 However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
-This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
 \textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
@@
 For example, assume that a user has written the word ``zend'' in a post.
-``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP}.
+``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP} ($e_{PHP}$).
 Since the environment vector for \emph{PHP} was added to the memory vector for ``zend'' without any encoding, no decoding is necessary.
-So the cosine between the environment vector \emph{PHP} and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.
+So the cosine between the environment vector \emph{PHP} ($e_{PHP}$) and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.
 
@@
 
+The resonance and decoding processes are inversions of each other.
 Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
@@
 One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
-Representing word order is an example where the encoding and decoding operations are applied.
+Representing word order is an example where both the encoding and decoding operations are applied.
 As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
@@
 Within this corpus, the phrase ``martin luther'' often occurred.
-When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-back location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
-Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
+When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-forward location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
+Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-back location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
 The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.
@@
 A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
-With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few $[1,-1]$s in random locations. 
+With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few $<1,-1>$s in random locations. 
 In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector ($e_{i}$) is bit shifted to the left one position ($^{-1}$).
@@
 
-Random permutations use one of simplest permutations possible to create a new environment vector.
+Random permutations usees one of simplest permutations possible to create a new environment vector.
 This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
@@
 For random permutations, each tag is represented by a compressed vector with a length much less than the number of words.
-Note that if the length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular column in the vector.
+Note that if the length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular row in the column vector.
 If done in this way, building a matrix of all words' environment vectors x tags would recover the original word x document matrix \parencite{Kanerva2000}.
-However, each word's environment vector is compressed by using a length much less than the number of unique words and using a combination of a few [1,-1]s to represent the word.
+However, each word's environment vector is compressed using a length much less than the number of unique words and using a combination of a few $<1,-1>$s to represent the word.
 
@@
 It may be interesting to explore how model performance changes when one calculation is swapped out for another.
-For example, a vector-based model could be used to generate a compress (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
+For example, a vector-based model could be used to generate a compressed (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
 Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.

### 2013-10-01 writeups/proposal/proposal.tex
read through entire proposal
Former-commit-id: 805f270e8d8c5293aa2aebaa5846f77107ff90c0 [formerly 5b074f791cf991108423ca5a0053bda49abc6922]
Former-commit-id: ca9c66816dd9f145501e78530f95d9b7667b4754

@@
 When the system needed to recommend a set of individuals for a user to follow, it would take all of the words used in posts from that user to construct a search query to be provided to Lucine.
-That search query would return the most likely users that share similar words in posts within their twitter network with the words provided in the query.
+That search query would return the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
 Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user will follow at 25\% accuracy.
@@
 These models will be the starting point for modeling tag use on the Twitter and StackOverflow sites.
-The models may be modified during testing if the data indicate that slight modifications are needed to improve model performance.
+The models may be modified during testing if the data indicate that modifications are needed to improve model performance.
 Also there are planned modifications to each model as part of the research:
@@
 Parameters for each model will be calibrated on a large subset of the Twitter and StackOverflow datasets.
-Each model will be tested by evaluating the accuracy of model-chosen-tags on a fresh test subset of the datasets.
+Each model will be tested by evaluating the accuracy of model-chosen tags on a fresh test subset of the datasets.
 Accuracy will be assessed by comparing model performance with the human data.
@@
 The temporal dynamics of ACT-R's decay rate model for a chunk will be evaluated against this data.
-As a result, the most appropriate model of user-centered frequency and recency effects (i.e., each hashtag's prior likelihood given a particular user) will be determined.
+Afterwards, the most appropriate model of user-centered frequency and recency effects (i.e., each hashtag's prior likelihood given a particular user) will be determined.
 
@@
 
-The text from all dataset will be pre-processed using standard natural-language processing techniques.
+The text from each dataset will be pre-processed using standard natural language processing techniques.
 The Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} will be used to tokenize (chunk) and lemmatize (convert to root form) each word.
-The standard chunking algorithm will be modified, so that it also ensures that hashtags and tag words (e.g., the word \emph{PHP} in a post) are not separated into multiple chunks.
+The standard chunking algorithm will be modified, so that it also ensures that Twitter hashtags and StackOverflow tag words (e.g., the word \emph{PHP} in a post) are not separated into multiple chunks.
 With the StackOverflow dataset, it is possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
@@
 
-``Stop words'' are commonly-used words that co-occur with all tags (i.e., have little to no discriminating power).
+``Stop words'' are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
 It is typically recommended \parencite{Bird2009} to remove stop words such as ``the'', ``a'', ``or'' when building the co-occurrence matrix.
-Otherwise the matrix may not scale as well since there are more words to keep track of. 
+Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in text.
 Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.
-\textcite{Sahlgren2008} removed stop words for the random-permutation vector-based model, for example.
+\textcite{Sahlgren2008} removed stop words for the random-permutation vector-based model to handle this issue.
 \textcite{Jones2007} did as well for the circular-correlation model, but also mentioned that an entropy weighting technique has also been used previously to attenuate (rather than remove) stop words.
@@
 When testing the model, only at most 1 post from a user is included in the test set, since it is much easier to suggest relevant hashtags to a spammer than it is for an actual user.
-This exact technique may not be used for this analysis, but certainly some form of attenuating or identifying twitter spam tweets will be explored.
+This exact technique may not be used for this analysis, but certainly some form of attenuating or identifying Twitter spam tweets will be explored.
 
@@
 First, model performance will be compared between the newest dataset and the Apr' 11 dataset to ensure that performance is consistent.
-Also this is a way to verify that the newest dataset was correctly pre-processed and loaded.
+This is also a way to verify that the newest dataset was correctly pre-processed and loaded.
 
@@
 After incorporating the relative strengths of each model into the other, the accuracy of the two models will be compared directly.
-Both models will be evaluated on predicting hashtags for tweets in the popular-hashtags dataset and predicting tags for posts on StackOverflow.
-However, the user prior hashtag use will not be available for the popular-hashtags dataset, since this subset was collected across the entire twitter user population.
+First both models will be evaluated on predicting tags based purely on context (i.e., no prior user history), using the popular-hashtags, popular-users, and StackOverflow datasets.
+The results will show the performance differences for the contextual component of each model.
 
@@
 
-In order to directly compare performance differences for each model when generating user-customized tag predictions, the models will be tested on the StackOverflow dataset and the Twitter popular-users dataset.
-Each dataset contains the entire tag history of either all users (StackOverflow) or the most popular users (Twitter).
-The main evaluation will be the amount of performance gain achieved when the models generate tags that are based on past user behavior compared to models that generate tags that are bases purely on context.
+Next the full models (prior and context components) will be tested on the Twitter popular-users dataset and the StackOverflow dataset.
+The popular-hashtags dataset will not be included in this analysis, since this subset was collected across the entire Twitter user population, so prior hashtag use is not available.
+This test will directly compare performance differences for each model when generating user-customized tag predictions.
+The main evaluation will be the amount of performance gain achieved when the models generate tags that are based on past user behavior compared to models that generate tags that are based purely on context.
 
@@
 as well as two cognitively-plausible models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
-Overall performance differences between an ACT-R's Bayesian retrieval model and a random-permutation vector-based model will be directly compared.
-Improvements made in each model in order to increase performance at this tagging task may suggest model modifications that could be made to improve model performance for other tasks.
-Further, these improvements may uncover model modifications that could improve model fit across all tasks (i.e., suggest cognitive architectural modifications).
-In addition, if model accuracy is high enough, then these developed models may be applied to general tag recommendation systems for either Twitter or StackOverflow,
+Overall performance differences between an ACT-R's Bayesian retrieval model and a random-permutation vector-based model will be directly compared on a real-world tagging task, which has never been done previously.
+The strengths of each model (word order and user prior likelihood) will be incorporated into the other, and performance differences will be evaluated.
+The main two proposed model modifications (incorporating word order into ACT-R and user history into vector-based representations) have not been tested before, and seem likely to significantly improve performance.
+Further, these two modifications in particular are aimed at the architectural level of each model, may very well be task independent, and may generalize and improve model fit across other tasks.
+Finally, if model accuracy is high enough, then the developed models may be applied to general tag recommendation systems for either Twitter or StackOverflow,
 and possibly for other sites where users generate and tag content.

### 2013-10-02 writeups/proposal/proposal.tex
foo
Former-commit-id: 496906991ccc1838e6d8ce6e58f21a590d6e5c80 [formerly 4055675559ae36636d274c3b8730857928e72e31]
Former-commit-id: 5deef66a6f58b7738b3c495df534f741a44cc3b6
foo
Former-commit-id: 099d5e79dc25877fdd408c2362508b0c4d91658d [formerly 9b750697e007311bee146ed5a16ba135990ae585]
Former-commit-id: 51c0a167072d907de440011535b6cb38dfb57840

@@
 \DeclareLanguageMapping{american}{american-apa}
+
 \addbibresource{bibliography.bib}
@@
 }
+
+% Don't let biblatex change any of the casing in the bibliography file
+\DeclareRobustCommand{\MakeSentenceCase*}[1]{{#1}}
 
@@
 
-% FIXME: Fix cite for Bauer2012
 One traditional technique is to provide the user with a powerful search engine so that they can actively look for content that interests them.
 However, users are not forming the same search queries that work so effectively on search engines like Google.
-A growing number of users on social media sites are searching for information that is happening now, in real time \parencites{Bauer2012,Jansen2011}.
+A growing number of users on social media sites are searching for information that is happening now, in real time \parencite{Jansen2011}.
 In other words, users on social-media sites want the fresh \emph{stream} of information about a particular topic.
@@
 
-% FIXME: Use vector-graphics for this figure 
 \begin{figure}[!htbp]
@@
 
-% FIXME: Fix cite for Bennett2007
 The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
@@
 
-% FIXME: Add figure for LaTeX tag suggestion example
+A tag suggestion tool has been deployed for the \url{http://tex.stackexchange.org} \parencite{LatexTags2013} and \url{http://superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
+\LaTeX, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
+For the two sites where a tag suggestion tool has been deployed, once a user has created a question, suggested tags appear just below the text field where the user types the tags to associate with the question.
+The post author is not forced to use the suggested tags, and is free to choose tags from them or tags not represented in the suggested set.
+An example of the interface for a created post on the SuperUser site is included in Figure \ref{figSuperUserSuggestion}.
 
-A tag suggestion tool has been deployed for the \url{http://tex.stackexchange.org} site \parencite{LatexTags2013}.
-Both the \LaTeX site and StackOverflow are part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
-Once a user has created a question on the Latex site, the site provides subdued text of suggested tags in the text field where the user fills in the tags to associate with the question.
-The post author is not forced to use the suggested tags, and is free to choose tags from them or tags not represented in the suggested set.
+\begin{figure}[!htbp]
+  \scalebox{.9}{\resizebox{\linewidth}{!}{\includegraphics{superUserTagSuggestion.pdf}}}
+  \caption{Example of tag suggestions when creating a post on the SuperUser site}
+  \label{figSuperUserSuggestion}
+\end{figure}
 
-The Latex site is not nearly as popular as StackOverflow currently.
-The site has roughly \num{50000} questions compared to \num{5000000} questions on StackOverflow.
-It may be the case that the StackExchange company decided to try out a tag recommendation system on one of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
+The Latex and SuperUser sites are not nearly as popular as StackOverflow currently.
+The sites have roughly \num{50000} and \num{189000} questions respectively compared to \num{5000000} questions on StackOverflow.
+It may be the case that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
 Or it is possible that their current tag recommendation system is not efficient enough to work with the demand of the StackOverflow site.

### 2013-10-02 writeups/proposal/proposal.tex
foo
Former-commit-id: c2f89f8acc16bb383dfe88e3a9be0a4724c5efd4 [formerly c7a5861399874443fca5e63557e70427dd2ebfe5]
Former-commit-id: 5a45f6fc83eb226c8976d7c0714befe524b81103

@@
 Further, people are using hashtags as a way to create information streams on social media sites \parencite{Kwak2010}.
-Twitter users for example, commonly create and use hashtags for upcoming political events that interst them, such as debates and races \parencite{Diakopoulos2010}.
+Twitter users for example, commonly create and use hashtags for upcoming political events that interest them, such as debates and races \parencite{Diakopoulos2010}.
 So if we can predict what hashtags a user is interested in, we have a better understanding of the information streams that they care about.
@@
 Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of being retrieved again, given the prior history of exposure to the chunk.
-These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adusted by current environment state.
+These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.
 
@@
 The entropy measure was used to compute the attentional weight for two reasons:
-Stop words could be statistically derived and attentuated in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
+Stop words could be statistically derived and attenuated in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
 Using an entropy measure (as opposed to weighting each word equally in the post) allows the model to allocate limited attentional resources ($W$) to important contextual elements.
@@
 
-Random permutations usees one of simplest permutations possible to create a new environment vector.
+Random permutations uses one of simplest permutations possible to create a new environment vector.
 This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.

### 2013-10-02 writeups/proposal/proposal.tex
foo
Former-commit-id: ee9bd784e6616ba2049508e5c06dfb566ee9917d [formerly c6f962873f33d5ee75847827c25798da710f6ce2]
Former-commit-id: e5cf7845ecd171f0c6a1785f7f2a6d56f3064e71
foo
Former-commit-id: 941fac6cfa638da1c86adf4375e78db2d713ccba [formerly 2b9ea25b5467859f16af5847215affdac94198c3]
Former-commit-id: 7af1cf806d7237a1d09cf5046caca536d2bfcaa7

@@
-\documentclass[man,floatsintext]{apa6}
+\documentclass[man,floatsintext,donotrepeattitle]{apa6}
 
@@
 
-\title{FIXME}
-\shorttitle{}
+\title{User-customized hashtag/tag prediction on Twitter and StackOverflow:\protect \\ Comparing vector-based and Bayesian models of tag retrieval}
+\shorttitle{Tag prediction on Twitter and StackOverflow}
 
@@
 This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
-However, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.
+Further, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.
 

### 2013-10-04 writeups/proposal/proposal.tex
fri edits
Former-commit-id: 02a889d7adecce4a3b9a2ff20e4b8b96500fff4d [formerly c76f970ce84da57b898e3515a4a65f58eb81326e]
Former-commit-id: 4b78dc47bb90e32f87b115fb273f1bc5e3e08826

@@
 
-Alongside testing these declarative memory retrieval models on Twitter hashtag use, they will be tested on tag use for StackOverflow posts.
+Alongside testing these declarative memory retrieval models on Twitter hashtag use, they will also be tested on tag use for StackOverflow posts.
 StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
@@
 This will undoubtedly require a combination of two primary model components:
-[1] A user's prior likelihood of choosing a particular hashtag, given their previous content, and [2] the likelihood that a particular hashtag is related to the context of the post being created.
+[1] A user's prior likelihood of choosing a particular hashtag, given their previous hashtag use, and [2] the likelihood that a particular hashtag is related to the context of the post being created.
 
@@
 Netflix wanted to improve the accuracy of their recommendation system, so they held a competition where teams could try to improve on the accuracy of the service.
-The winning team utilized a statistical model that performed a type of factor analysis to identify important dimensions in the data (e.g., drama, action, cult classic) and place each user along each dimension.
+The winning team utilized a statistical model that performed a type of factor analysis to identify important dimensions in the data (e.g., drama, action, cult classic) and place values for each user along each dimension.
 This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
@@
 Google has even deployed a hashtag-recommendation model to help users properly tag posts for their Google+ social media site \parencite{GoogleKeynote2013}.
-A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} site.
+A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
 Models for Twitter and StackOverflow are also being proposed and tested, but none have yet been deployed.
@@
 Or it is possible that their current tag recommendation system is not efficient enough to work with the demand of the StackOverflow site.
-Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial enough to the user to start adding them to their question and answer sites.
+Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial to the user and have started to add them to their question and answer sites.
 
@@
 The relationship is built between a tweet and its content, and then similar tweets are assumed to use similar hashtags.
-One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a vector representation of the content in each tweet has to be maintained to later compute tweet similarity.
+One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a representation of the content in each tweet has to be maintained to later compute tweet similarity.
 The storage size for a hashtag-centered approach grows with the number of hashtags, but it seems reasonable that this will grow slower than the number of tweets and asymptote after collecting a large sample of tweets.
@@
 Stop words could be statistically derived and attenuated in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
-Using an entropy measure (as opposed to weighting each word equally in the post) allows the model to allocate limited attentional resources ($W$) to important contextual elements.
+Using an entropy measure (as opposed to weighting each word equally in the post) for the attentional weight term allows the model to allocate limited attentional resources ($W$) to important contextual elements.
 
@@
 A word x document co-occurrence matrix is built by counting the number of times each word appears in each document.
-The dimensionality of the full word frequency co-occurrence matrix is then reduced and forced to be represented on a smaller set of dimensions.
+The dimensionality of the full word frequency co-occurrence matrix is then forced to be represented on a smaller set of dimensions.
 This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
@@
 For example, \textcite{Jones2007} compared a vector-based model called BEAGLE directly to LSA, training and testing both models on the same corpus used in \textcite{Landauer1997}.
-The context-only BEAGLE model (no word order) performed slightly better than the LSA model at the task (55.6\% vs. 55.3\%).
+The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the task (55.6\% vs. 55.3\%).
 However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to 57.5\%.
@@
 
-Another issue with LSA is the SVD matrix decomposition technique is computationally very expensive.
+Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
 As the size of the original word x document matrix increases, the SVD computation becomes more and more time consuming.
@@
 \textcite{Recchia2010} found that random permutations could use smaller vectors to produce equivalent performance (better compression), and scale to larger datasets (better computational efficiency).
-The vector x performance tradeoff was tested on a simple paired associative retrieval task. 
+The vector length x performance tradeoff was tested on a simple paired associative retrieval task. 
 Retrieval accuracy for the vector-based models remained high even after reducing each environment vector's length from \num{2048} to \num{1024}.
@@
 Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
-Each model computes an activation for each word to retrieve, given prior knowledge and context.
+Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
 Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.

### 2013-10-04 writeups/proposal/proposal.tex
apa6 formatting changes
Former-commit-id: 6d929863507f25c4c04103472b91b51cd879182c [formerly 31da54f90694d1585ce4ece60138c0ce3be3bfd1]
Former-commit-id: bdf25a16e7b821b2b555cccc3c0ba41174ba1743

@@
 
+% ref: http://tex.stackexchange.com/questions/28516/how-to-change-the-title-of-toc
+\addto\captionsamerican{% Replace ``english'' with the language you use
+  \renewcommand{\contentsname}%
+  {Table of Contents}%
+}
+
 %\usepackage[compact]{titlesec}
@@
 \usepackage{setspace}
+
+% ref: http://tex.stackexchange.com/questions/48509/insert-list-of-figures-in-the-table-of-contents
+\usepackage[nottoc,numbib]{tocbibind}
 
@@
 
-\title{User-customized hashtag/tag prediction on Twitter and StackOverflow:\protect \\ Comparing vector-based and Bayesian models of tag retrieval}
-\shorttitle{Tag prediction on Twitter and StackOverflow}
+\title{User-customized hashtag/tag prediction on Twitter and StackOverflow:\protect \\ Comparing vector-based and ACT-R Bayesian models of tag retrieval}
+%\shorttitle{Tag prediction on Twitter and StackOverflow}
+\shorttitle{}
 
@@
 
+\keywords{FIXME}
+
+\begin{document}
+
 \abstract{
@@
 
-\keywords{FIXME}
-
-\begin{document}
 \maketitle
@@
 \tableofcontents
+\newpage
+
+\listoftables
+\newpage
+
+\listoffigures
 \newpage
@@
     \hfill
-    \caption{Example tweets on Twitter.com}
+    \caption{Example tweets on twitter.com}
     \label{figTweetExample}
@@
   \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample-crop.pdf}}}
-  \caption{Example post on Google+. The system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-created post}
+  \caption{Example post on Google+}
   \label{figGoogle+Hashtags}
@@
 When generating content on Google+, the system automatically associates the most likely relevant hashtags with that post.
+For the example in Figure \ref{figGoogle+Hashtags}, the system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-created post.
 The user is free to remove any or all of the tags that the system has associated with the post, and can also use tags that were not included in the recommended set.
@@
 \begin{table}[!ht]
-  \caption{ACT-R Declarative Memory Model}
+  \caption{ACT-R declarative memory model}
   \label{tabACT-RModel}
@@
 \begin{table}[!ht]
-  \caption{Random Permutation Model}
+  \caption{Random permutation model}
   \label{tabRandPermModel}
@@
 \setlength\bibitemsep{12pt}
-\printbibliography
+\clearpage
+\printbibliography[heading=bibintoc]
 \endgroup

### 2013-10-04 writeups/proposal/proposal.tex
ran through style checker
Former-commit-id: 9afad637c7bdbdd16c01fce95600b9318e08e0cd [formerly 6041cf9def12b0b97b35d184794533280b64b025]
Former-commit-id: 0deebb8a2887354695f9d390255c29b0dbb98854

@@
 The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and I am interested in retrieval models that generalize across different task domains.
-However, they are also quite similar on several accounts:
+However, they are also similar on several accounts:
 Both domains have amassed large amounts of user-created content, users generate tags for posts when creating content on the site (hashtags and tags for Twitter and StackOverflow respectively),
@@
 
-Pointwise Mutual Information (PMI) is a another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
+Pointwise Mutual Information (PMI) is another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
 It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
@@
 BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (55.6\% compared to 55.3\%).
-BEAGLE with unordered context and word order was also more correlated to the WordNet synonym database than LSA (-.311 compared to -.165).
+Also, BEAGLE with unordered context and word order was more correlated to the WordNet synonym database than LSA (-.311 compared to -.165).
 Further, it was shown that incorporating word-order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).

### 2013-10-06 writeups/proposal/proposal.tex
foo
Former-commit-id: 5d39450145900210f2d0354e3ade31cca54726a0 [formerly 3b51d3b610e8e700a938b7552e75905b1a9775a5]
Former-commit-id: 1fec6f77c529e831d6c424e45b944a7134a80c15

@@
 
-\subsubsection{StackOverflow and Twitter Task Domains}
+\subsubsection{StackOverflow and Twitter}
 
@@
 
-The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and I am interested in retrieval models that generalize across different task domains.
+The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and I am interested in retrieval models that generalize across tasks.
 However, they are also similar on several accounts:
@@
 
-Also, studying user-based tag generation on these sites is a relevant task domain.
+Also, studying user-based tag generation on these sites has relevant real-world application.
 Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these very popular sites.
@@
 After scaling up the model, this model can successfully predict a user's tag for a post at 65\% accuracy.
-It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is built.
-This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular task domain:
-predicting the user-chosen tags for human-created content on social media sites.
+It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
+This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular environment
+(i.e., when predicting the user-chosen tags for human-created content on social media sites).
 
@@
 as well as two cognitively-plausible models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
-Overall performance differences between an ACT-R's Bayesian retrieval model and a random-permutation vector-based model will be directly compared on a real-world tagging task, which has never been done previously.
+Overall performance differences between an ACT-R's Bayesian retrieval model and a random-permutation vector-based model will be directly compared on two real-world tagging tasks, which has never been done previously.
 The strengths of each model (word order and user prior likelihood) will be incorporated into the other, and performance differences will be evaluated.

### 2013-10-06 writeups/proposal/proposal.tex
foo
Former-commit-id: 35c0f6e625a76a0b78e2092b5e39f7ebba16dd29 [formerly ff85ea3711fd6879f7dfd3c7a9d923c4e3fc58d3]
Former-commit-id: f7fb7da7afc3a81998729bad1bb9d5317fa40b39

@@
 Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
-Two of the current state-of-the-art memory models are based on a co-occurrence methodology, and I'll be comparing these two models for hashtag prediction:
+Two of the current state-of-the-art memory retrieval models are based on a co-occurrence methodology, and I'll be comparing these two models for hashtag prediction:
 ACT-R's declarative memory retrieval theory and random permutation vector-based memory systems.
@@
 
-Alongside testing these declarative memory retrieval models on Twitter hashtag use, they will also be tested on tag use for StackOverflow posts.
+Alongside testing these memory retrieval models on Twitter hashtag use, they will also be tested on tag use for StackOverflow posts.
 StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
@@
 User-chosen tags on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
-Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, or \emph{Common Lisp}.
+Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.
 
@@
 
-ACT-R's declarative memory and vector-based memory systems are substantially different in their implementation.
+ACT-R's declarative memory and vector-based memory systems are substantially different in their mathematical structure.
 Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., the ``fan effect'' \parencite{Rutledge2008}).
 I will extend the work done by \textcite{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
-We'll also replace the vector-based model used by \textcite{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
+I'll also replace the vector-based model used by \textcite{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
 The primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory.
@@
 A user's prior hashtag use will be an essential model component for domains such as Twitter, where the number of possible hashtags is practically infinite.
-In these domains, the model can utilize a user's prior hashtag use as a way to prune the infinite space of possible hashtags, and generate a much smaller and more manageable set for prediction.
+In these domains, the model can utilize a user's hashtag history as a way to prune the infinite space of possible hashtags, and generate a much smaller and more manageable set for prediction.
 However, not only should the model properly take into account a user's prior hashtag use, but it should also generate new hashtag predictions when a user is most likely choosing a hashtag they have never used before.
@@
 That search query would return the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
-Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user will follow at 25\% accuracy.
+Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user follows at 25\% accuracy.
 
@@
 Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive Singular Value Decomposition and K-nearest-neighbors models.
-This model can accurately predict a user-chosen tag at 47\% accuracy.
+This Bayesian model can accurately predict a user-chosen tag at 47\% accuracy.
 
@@
 On the Latex site, the user interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
-For Google+, the user interface automatically provides tags for a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
+For Google+, the user interface automatically tags a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
 If tag accuracy is high enough, the Google+ technique is certainly favorable since the user does not have to spend time figuring out which tags best represent this post.
@@
 Programming-language popularity over the past few years is a reasonable example of this.
-Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., \emph{closure}'s growth), it is not the case that the changes have been drastic.
-\emph{C\#}, \emph{Python}, and \emph{Java} are still a few of the most popular languages, \emph{Common Lisp} has not gained or lost much ground in popularity over the past few years.
+Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., Clojure's growth), it is not the case that the changes have been drastic.
+C\#, Python, and Java are still a few of the most popular languages, Common Lisp has not gained or lost much ground in popularity over the past few years.
 
@@
 Strength of association ($S_{ji}$) reflects the amount of log odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
-Context for the StackOverflow domain for example, can be represented as each word in the title and body of a post.
-Context for the Twitter domain would be the words in a Tweet.
+Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
+Context for the Twitter domain are the words in a Tweet.
 Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
@@
 
-The SNIF-ACT framework is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search \parencites{Fu2007,Pirolli2003}.
-The $S_{ji}$ values were based on a 118GB local database of 10 million pages of the Stanford WebBase Project \parencite{Farahat2004}.
+The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search.
+\textcite{Farahat2004} used $S_{ji}$ values based on a 118GB local database of 10 million pages of the Stanford WebBase Project.
 So this implementation did not fit completely in RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.
@@
 
+A formal description of the modified ACT-R Bayesian model from \textcite{Stanley2013} is included in Table \ref{tabModACTRModel}.
+
 \begin{table}[!ht]
-  \caption{The StackOverflow tag prediction model}
+  \caption{StackOverflow tag prediction model}
   \label{tabModACTRModel}
@@
 
-A formal description of the modified ACT-R Bayesian model from \textcite{Stanley2013} is included in Table \ref{tabModACTRModel}.
 The activation of each tag ($A_{i}$) is a function of four components:
@@
 However it was shown that this process actually improves productiveness and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
-They tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
+\textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
 Model results were on par with a large sample of applicants to american universities from non-English speaking countries (64.4\% compared to 64.5\%).
@@
 The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
-The best way to remove a dimension while maintaining all variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
+The best way to remove a dimension while maintaining the most variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
 For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
@@
 However, with large-scale human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
-\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size dataset.
-Instead, they compared the measures on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
+\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
+Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
 This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.
@@
 
-Another issue with LSA is that adding additional observations to an already-computed reduced representation requires rerunning the SVD \textcite{Farahat2004}.
+Another issue with LSA is that adding additional observations to an already-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
 Since computing the SVD is the most computationally-expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically being added.
@@
 
-``holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., chunk) as a vector.
+``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., chunk) as a vector.
 The representation of the concept is distributed across all elements of the vector.
-This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and saying that the distribution of co-occurrence counts across the rows in the column is the representation of that concept.
+This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and viewing the distribution of co-occurrence counts across the rows in the column as the representation of that concept.
 However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.
@@
 For example, assume that a user has written the word ``zend'' in a post.
-``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP} ($e_{PHP}$).
+``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP} ($e_{\mathit{PHP}}$).
 Since the environment vector for \emph{PHP} was added to the memory vector for ``zend'' without any encoding, no decoding is necessary.
-So the cosine between the environment vector \emph{PHP} ($e_{PHP}$) and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.
+So the cosine between the environment vector \emph{PHP} ($e_{\mathit{PHP}}$) and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.
 
@@
 As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
-Simply counting co-occurrences of the two immediately-preceding words to hashtags does not include the order information of the words in the representation.
+Simply counting co-occurrences of the two immediately-preceding words and hashtags does not include the order information of the words in the representation.
 With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
@@
 Within this corpus, the phrase ``martin luther'' often occurred.
-When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-forward location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
-Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-back location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
+When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-back location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
+Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
 The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.
@@
 The vector length x performance tradeoff was tested on a simple paired associative retrieval task. 
-Retrieval accuracy for the vector-based models remained high even after reducing each environment vector's length from \num{2048} to \num{1024}.
+Retrieval accuracy for the random-permutation vector-based models remained high even after reducing each environment vector's length from \num{2048} to \num{1024}.
 This indicates that vector-based representations using random permutations compress better than those using circular convolution.
 
-To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFFL) assessment (among other tasks).
+To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFL) assessment (among other tasks).
 The random-permutation model was trained on \num{2.33} GB corpus.
@@
 A vector-based approach like BEAGLE or random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
-Nonetheless, random permutations have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFFL \parencites{Sahlgren2008,Jones2007}.
+Nonetheless, random permutations have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFL \parencites{Sahlgren2008,Jones2007}.
 
@@
 
-However, vector-based models and ACT-R's current Bayesian DM model are not identical, and there are open issues with both models that need to be addressed before further evaluating the benefits of one over the other.
+However, vector-based models and ACT-R's current Bayesian DM model are not identical, and these differences need to be addressed in order to better evaluate the benefits of one over the other.
 For example, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.

### 2013-10-06 writeups/proposal/proposal.tex
foo
Former-commit-id: 1035d81dbdb83ec5aca8ecb96527ad59d02d68de [formerly 8f193866504e24548fb080d349aae7c628cd9d2a]
Former-commit-id: 8f0cab42523b43281cedffc639098e5efcf1b8cd

@@
 what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
-Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of being retrieved again, given the prior history of exposure to the chunk.
+Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of needing retrieval again, given the prior history of exposure to the chunk.
 These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.
@@
 Further, if this is a reasonable assumption to make, it greatly simplifies the computation of the base-level activation for chunks, and can turn the computation into a tractable problem for large datasets.
-In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log odds ratio of being exposed to that chunk compared to all other chunk exposures.
+In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log odds ratio of exposure to that chunk compared to exposure to all other chunks.
 For example, if the StackOverflow tag \emph{PHP} has been used four times as often as the tag \emph{Common Lisp}, then the prior odds of needing \emph{PHP} again is $\frac{.8}{.2}=4$ times that of \emph{Common Lisp}.
@@
 As the size of the original word x document matrix increases, the SVD computation becomes more and more time consuming.
-If the domains where LSA is being applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
+If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
 However, with large-scale human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
@@
 Another issue with LSA is that adding additional observations to an already-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
-Since computing the SVD is the most computationally-expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically being added.
+Since computing the SVD is the most computationally-expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
 For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.

### 2013-10-06 writeups/proposal/proposal.tex
foo
Former-commit-id: 5927992fe44d7b08ef715a895d9bc27c8afffc6e [formerly 74e943b3e3c789f66d4ca9f6b0ad8d83cc2d1373]
Former-commit-id: 44898ea3d19ceb94c55b1c3edb6b7010c81c07d0

@@
 A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
-With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few $<1,-1>$s in random locations. 
+With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few $\{1,-1\}$s in random locations. 
 In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector ($e_{i}$) is bit shifted to the left one position ($^{-1}$).
@@
 If done in this way, building a matrix of all words' environment vectors x tags would recover the original word x document matrix \parencite{Kanerva2000}.
-However, each word's environment vector is compressed using a length much less than the number of unique words and using a combination of a few $<1,-1>$s to represent the word.
+However, each word's environment vector is compressed using a length much less than the number of unique words and using a combination of a few $\{1,-1\}$s to represent the word.
 

### 2013-10-06 writeups/proposal/proposal.tex
foo
Former-commit-id: bb734b37947a27134813331ead399a81c3b05171 [formerly 2df2d3ca55a21eec736d7e9ef461b4410f43c5f3]
Former-commit-id: 6d705b974bceb3abc79263708d5279a3c16542b1
foo
Former-commit-id: a9b6c08cae7c17671510e2b87bb1f6d4374f2e47 [formerly be0b1d3e61a374ee2c3e24b7b0b70f17883c4e87]
Former-commit-id: f572afdc6b45953edab071c76f43861b8aa0e270
foo
Former-commit-id: 2cfae7bd10bf1fd2598644a84faedbb84e6d8373 [formerly e32e5425aa8bb49b9725c16faa6d04bc5e1fb0a0]
Former-commit-id: f28e9b7304d115dc74c63c80df8af5796a40cebd

@@
 \usepackage[hyphens]{url}
-%\usepackage{hyperref}
+
+% ref: http://tex.stackexchange.com/questions/73862/how-can-i-make-a-clickable-table-of-contents
+\usepackage{hyperref}
+\hypersetup{
+  colorlinks,
+  citecolor=black,
+  filecolor=black,
+  linkcolor=black,
+  urlcolor=black
+}
+%\hypersetup{linktocpage}
+
 %\usepackage{breakurl}
-
 \usepackage{subfigure}
@@
 \title{User-customized hashtag/tag prediction on Twitter and StackOverflow:\protect \\ Comparing vector-based and ACT-R Bayesian models of tag retrieval}
-%\shorttitle{Tag prediction on Twitter and StackOverflow}
-\shorttitle{}
+\shorttitle{Tag prediction on Twitter and StackOverflow}
+%\shorttitle{}
 
@@
 }
+
+\note{\today}
 
@@
 
+\subsubsection{StackExchange}
+
 A tag suggestion tool has been deployed for the \url{http://tex.stackexchange.org} \parencite{LatexTags2013} and \url{http://superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
@@
 \begin{figure}[!htbp]
-  \scalebox{.9}{\resizebox{\linewidth}{!}{\includegraphics{superUserTagSuggestion.pdf}}}
+  \scalebox{.9}{\resizebox{\linewidth}{!}{\includegraphics{superUserTagSuggestion-crop.pdf}}}
   \caption{Example of tag suggestions when creating a post on the SuperUser site}
@@
 
-\subsubsection{\textcite{Stanley2013}'s ACT-R DM Model for StackOverflow}
+% ref: http://tex.stackexchange.com/questions/53513/hyperref-token-not-allowed
+\subsubsection{\texorpdfstring{\textcite{Stanley2013}}{Stanley2013}'s ACT-R DM Model for StackOverflow}
 

### 2013-10-07 writeups/proposal/proposal.tex
foo
Former-commit-id: dd8791e696ca44c9bef4610edbe4bdd293b787bd [formerly 5470f505c320d2c274b3667c076a852bd99a31f7]
Former-commit-id: 67e059cae110397abb43e58de060cc0e88c6643d
foo
Former-commit-id: b9f0649be30e5088b4d085468f6c6d3e8fba9c4f [formerly f9dd904595b2ca77ddd073f0784ea6c430992386]
Former-commit-id: 17385a4df455d10d1caa2636fe1ea949b76bcaf3

@@
 
-\title{User-customized hashtag/tag prediction on Twitter and StackOverflow:\protect \\ Comparing vector-based and ACT-R Bayesian models of tag retrieval}
+\title{User-customized hashtag/tag prediction on Twitter and StackOverflow:\protect \\ Comparing vector-based and ACT-R memory models of tag retrieval}
 \shorttitle{Tag prediction on Twitter and StackOverflow}
@@
 
-\keywords{FIXME}
+%\keywords{}
 
@@
 \abstract{
-  FIXME
+  The amount of user-created content on social media sites continues to grow.
+  To help organize and structure this information, many users create tags for documents to help connect their content to other posts on similar topics and other users that might have interest.
+  If an accurate model of the tags users choose when creating content is available, then the system can recommend tags for a user to use, in order to connect the user to similar topics and people as quickly as possible.
+  This works aims to develop and compare tag-recommendation models for users choosing hashtags when composing a tweet on Twitter and choosing tags when asking a question on the StackOverflow site.
+  Two cognitively-plausible memory models will be evaluated: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
+  Experiments will be aimed to [1] discover the best way to leverage a user's prior hashtag use as a future predictor, [2] improve the models so that each model's strengths are incorporated into the other,
+  and [3] formally evaluate and compare each model's performance when recommending hashtags on Twitter and tags on StackOverflow.
+  This research will focus on developing user-customized models of hashtag prediction, that best balance a user's past hashtag use and the current content in a post being created to predict the chosen hashtags.
+  Planned modifications to the two memory models are aimed at the architectural level, and may suggest architectural modifications that generalize to other tasks.
+  Finally, the user-customized tag-recommendation models developed with this research may be immediately useful for tag-recommendation systems on the popular Twitter and StackOverflow social media sites.
 }
@@
 However, they are also similar on several accounts:
-Both domains have amassed large amounts of user-created content, users generate tags for posts when creating content on the site (hashtags and tags for Twitter and StackOverflow respectively),
+Both domains have amassed large amounts of user-created content, users generate tags for posts when creating content on the site (hashtags for Twitter and tags for StackOverflow),
 and the user data from the datasets are publicly available for analysis.
@@
 For Twitter, hashtags represent streams of information that are possibly interesting to the user.
-Having a recommendation system that can suggest relevant hashtags to the user provides a way for the user to connect to and discover new information streams.
+Having a recommendation system that can suggest relevant hashtags to the user provides a way for the user to connect to and discover new information streams that interest them.
 
@@
 The results suggest that hashtags contain useful information that help classify the gist of a post. 
-Although this model does not predict hashtags directly when composing a tweet, a model like this could certainly be tailored for that purpose.
+Although this model does not predict the specific hashtags used when composing a tweet, a model like this could certainly be tailored for that purpose.
 
@@
 More specific hashtag-recommendation models that attempt to predict a user's chosen hashtag have also been created and tested for a few social media sites.
-Google has even deployed a hashtag-recommendation model to help users properly tag posts for their Google+ social media site \parencite{GoogleKeynote2013}.
+Google has even deployed a hashtag-recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
 A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
@@
 The main potential difference is that the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
-Instead, the global prior is an overall average frequency of hashtag use.
+Instead, the global prior is an overall average frequency of hashtag use across users.
 
@@
 
-There has been a mix of co-occurrence and latent-based models created to predict hashtags use on Twitter, and the results are encouraging.
+There has been a mix of co-occurrence and latent-based models created to predict hashtag use on Twitter, and the results are encouraging.
 However, there has been very little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
@@
 This type of weighted additive model that combines past user behavior and current tweet context to predict the most likely hashtags has not yet been explored for Twitter.
+
+Both StackOverflow models \parencites{Stanley2013,Kuo2011} did not take into account the specific user's prior history when generating tag predictions.
+Rather their models, much like \textcite{Mazzia2009}'s Twitter model, uses the aggregate tag history across all users as the model's prior activation for each tag.
+One of the main purposes of this research is to improve on these user-agnostic Bayesian models and include each specific user's past tagging history as a model component.
+Using customized priors that are tailored to a user's specific tagging history should be crucial for sites like Twitter where the possible tag space is infinite,
+but should also improve performance for more constrained sites like StackOverflow since each user is not interested in all of the possible tags on the site.
 
@@
 \subsubsection{ACT-R DM Model}
+
+A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACT-RModel}.
 
@@
 
-A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACT-RModel}.
 The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base level activation ($B_{i}$) and strength of association ($S_{ji}$).
-The recall probability ($P_{i}$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
+The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
 
@@
 Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
-Context for the Twitter domain are the words in a Tweet.
+Context for the Twitter domain are the words in a tweet.
 Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
@@
 
+As one of the largest tests, \textcite{Farahat2004} used the PMI index based on a 118GB local database of 10 million pages of the Stanford WebBase Project.
+So this implementation did not fit completely in RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.
+
 The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search.
-\textcite{Farahat2004} used $S_{ji}$ values based on a 118GB local database of 10 million pages of the Stanford WebBase Project.
-So this implementation did not fit completely in RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.
+The SNIF-ACT model uses strength of association to measure the relationship between the words in a search query and the words contained in each query result.
+A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.
+The links that user chose in experimental testing were largely predictive of the strength of association measure.
 
 %TODO: This paragraph doesn't fit well after the previous one, and in the scaling section
-%The SNIF-ACT model uses strength of association to measure the relationship between the words in a search query and the words contained in each query result.
-%A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.
-%The links that user chose in experimental testing were largely predictive of the strength of association measure.
 %The predictive power of the SNIF-ACT model went above and beyond what would be predicted by simply assuming that the user always clicks on the first link in the search results page.
@@
 A word x document co-occurrence matrix is built by counting the number of times each word appears in each document.
-The dimensionality of the full word frequency co-occurrence matrix is then forced to be represented on a smaller set of dimensions.
+The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
 This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
@@
 Usually information is lost when the data are restructured into a smaller set of latent dimensions.
-However it was shown that this process actually improves productiveness and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
+However it was shown that this process actually improves prediction and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
 \textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
@@
 However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to 57.5\%.
-It seems reasonable to expect that task performance will improve much more on tasks where word order is used (e.g., generating a Twitter hashtag in the middle of a sentence).
+It seems reasonable to expect that task performance will improve much more on tasks where word order is used (e.g., generating a hashtag in the middle of a sentence when composing a tweet).
 
@@
 \textcite{Sahlgren2008} used a much simpler random permutation method for this operation.
+
+A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
 
@@
 
-A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
 With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few $\{1,-1\}$s in random locations. 
-In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector ($e_{i}$) is bit shifted to the left one position ($^{-1}$).
+In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector ($e_{i}$) is shifted to the left one position ($^{-1}$).
 This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
@@
 For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow''.
-The ``stack'' environment vector ($e_{stack}$) is bit shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).
+The ``stack'' environment vector ($e_{stack}$) is shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).
 That environment vector is then added to the memory vector for ``overflow'' ($m_{overflow} = m_{overflow'} + e_{stack^{-1}}$), so that ``overflow'' remembers that ``stack'' has preceded it in a sentence.

### 2013-10-10 writeups/proposal/proposal.tex
foo
Former-commit-id: 06eae111baf45fc24e06e33dd76dc0350e40a979 [formerly 20845f47b219eebde832c4f94b845606e10998f9]
Former-commit-id: 09fc4e14a5078522ca8d9f80d8da939bcd0d84f0

@@
   The amount of user-created content on social media sites continues to grow.
-  To help organize and structure this information, many users create tags for documents to help connect their content to other posts on similar topics and other users that might have interest.
+  To help organize and structure this information, many users create tags for documents to help connect their content to other posts on similar topics and other users that might have similar interests.
   If an accurate model of the tags users choose when creating content is available, then the system can recommend tags for a user to use, in order to connect the user to similar topics and people as quickly as possible.
   This works aims to develop and compare tag-recommendation models for users choosing hashtags when composing a tweet on Twitter and choosing tags when asking a question on the StackOverflow site.
-  Two cognitively-plausible memory models will be evaluated: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
+  Two state-of-the-art memory models will be evaluated: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
   Experiments will be aimed to [1] discover the best way to leverage a user's prior hashtag use as a future predictor, [2] improve the models so that each model's strengths are incorporated into the other,
@@
 This can certainly be seen in the main user-interface views that these social media site provide:
-Twitter's homepage showing your followers' tweets, Facebook's homepage showing your friends' posts, StackOverflow's daily digest of posts for particular tags.
+Twitter's homepage showing your followees' tweets, Facebook's homepage showing your friends' posts, StackOverflow's daily digest of posts for particular tags.
 
@@
 Further, sites like Twitter, StackOverflow, Google+, and Facebook support ways for the user to explicitly identify their precise goals when creating content: Hashtags.
-These user-created hashtags are a key indicator to a user's goals on a social media site.
+These user-created hashtags are a key indicator of a user's goals on a social media site.
 
@@
 The core research question I am interested in is:
-What kinds of cognitively-plausible models can best predict the human-created tags that a user on a social media site creates and uses?
+What kinds of cognitively-plausible models can best predict the human-created tags that a user on a social media site uses?
 This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
@@
 Calibrating the models to properly balance this exploitation of prior hashtag use versus exploration of new hashtags will be a primary research focus.
-The end goal is to have a model that can properly balance the user's prior tag with the contextual cues in the post in order to generate an accurate prediction.
+The end goal is to have a model that can properly balance the user's tag history with the contextual cues in the post in order to generate an accurate prediction.
 
@@
 Or it is possible that their current tag recommendation system is not efficient enough to work with the demand of the StackOverflow site.
-Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial to the user and have started to add them to their question and answer sites.
+Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial to the user and has started to add them to their question and answer sites.
 
@@
 Several hashtag recommendation models for Twitter have been developed recently.
-One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a post as a function of prior hashtag use and context.
-\parencite{Mazzia2009}
+One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a post as a function of prior hashtag use and context \parencite{Mazzia2009}.
 The model presented by \textcite{Mazzia2009} is very similar to an ACT-R declarative retrieval model.
@@
 Each component is summed together to generate a total activation, and then the likelihood of chunks are ranked by that summed activation.
-This technique allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
+This technique is compensatory and allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
 This type of weighted additive model that combines past user behavior and current tweet context to predict the most likely hashtags has not yet been explored for Twitter.
@@
 
-Further, this PMI equation is actually identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
+Further, this PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
 Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.
@@
 
+The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search.
+
 As one of the largest tests, \textcite{Farahat2004} used the PMI index based on a 118GB local database of 10 million pages of the Stanford WebBase Project.
@@
 
-The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search.
-The SNIF-ACT model uses strength of association to measure the relationship between the words in a search query and the words contained in each query result.
-A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.
-The links that user chose in experimental testing were largely predictive of the strength of association measure.
+%The SNIF-ACT model uses strength of association to measure the relationship between the words in a search query and the words contained in each query result.
+%A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.
+%The links that user chose in experimental testing were largely predictive of the strength of association measure.
 
@@
 
-A formal description of the modified ACT-R Bayesian model from \textcite{Stanley2013} is included in Table \ref{tabModACTRModel}.
+\textcite{Stanley2013} modified the standard ACT-R model in Table \ref{tabACT-RModel} in order to more accurately predict tag use on StackOverflow.
+A formal description of the modified ACT-R Bayesian model is included in Table \ref{tabModACTRModel}.
 
@@
 The title and body contextual components were separated since it seemed plausible that the words in the title would be better predictors of the tag used than the words in the body.
-It actually turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
+It turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
 The entropy measure was used to compute the attentional weight for two reasons:
@@
 One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
-Representing word order is an example where both the encoding and decoding operations are applied.
+Representing word order is a case where both the encoding and decoding operations are applied.
 As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
@@
 Within this corpus, the phrase ``martin luther'' often occurred.
-When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-back location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
-Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
+When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated by convolving a 1-back location vector with luther ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
+Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin with and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
 The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.
@@
 Using decoding, the word-order memory vector for martin is deconvolved to the right ($o_{martin} \circledast \Phi$) which returns the environment vector for luther ($e_{luther}$).
-Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which returns the word-order memory vector for luther ($o_{luther}$).
+Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which correlates highest with the word-order memory vector for luther ($o_{luther}$).
 Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$''.
@@
 Vector-based representations compute chunk likelihoods by performing a correlation operation directly on the memory vectors.
-Each memory vector can be thought of as a column on a context word x tag word co-occurrence matrix used for a Bayesian representation.
+Each memory vector can be thought of as a column on a context word x tag co-occurrence matrix used for a Bayesian representation.
 Further, the number of rows for a vector-based representation are fixed and set much smaller than the number of rows in the full word x tag co-occurrence matrix.
@@
 Each memory vector ($m_{j}$) is a summed representation of unordered and ordered word co-occurrences ($c_{j} + o_{j}$).
-This model uses circular convolution and inverse correlation as a way to encode and decode word order.
+This model uses circular convolution and circular correlation as a way to encode and decode word order.
 BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (55.6\% compared to 55.3\%).
@@
 In order to represent word order information, a function must be used that converts an environment vector for a set of words and positions (e.g., ``stack $\Phi$'') into a new uncorrelated environment vector.
-The BEAGLE model uses circular convolution and inverse correlation as the encoding and decoding operations.
+The BEAGLE model uses circular convolution and circular correlation as the encoding and decoding operations.
 However, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors and positions.
@@
 
-\textcite{Recchia2010} found that random permutations could use smaller vectors to produce equivalent performance (better compression), and scale to larger datasets (better computational efficiency).
+\textcite{Recchia2010} found that random permutations could use smaller vectors than circular convolution approaches (e.g., BEAGLE) to produce equivalent performance (better compression),
+and also scale to larger datasets (better computational efficiency).
 The vector length x performance tradeoff was tested on a simple paired associative retrieval task. 
-Retrieval accuracy for the random-permutation vector-based models remained high even after reducing each environment vector's length from \num{2048} to \num{1024}.
+Retrieval accuracy for the random-permutation vector-based models remained high relative to the circular convolution model even after reducing each environment vector's length from \num{2048} to \num{1024}.
 This indicates that vector-based representations using random permutations compress better than those using circular convolution.
@@
 It may be interesting to explore how model performance changes when one calculation is swapped out for another.
-For example, a vector-based model could be used to generate a compressed (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
+For example, a vector-based model could generate a compressed (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
 Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
@@
 Also there are planned modifications to each model as part of the research:
-incorporating user prior information into the vector-based model, and incorporating word order information into the Bayesian model.
+incorporating user prior information into the vector-based model and incorporating word order information into the Bayesian model.
 
@@
 So two Twitter datasets will be collected for model evaluation: a common-hashtags and popular-users dataset.
-The common-hashtags dataset will contain approximately 2 million tweets where at least one of the 100 most popular hashtags are used.
+The common-hashtags dataset will contain approximately 2 million tweets where at least one of the 500 most popular hashtags are used.
 This dataset will be collected in real time until the tweet quota is reached, which should take just a few days.
@@
 With the StackOverflow dataset, it is possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
-So tag words in the post and used tags for the post will be converted to their root tag when applicable.
+So tag words in the post and used tags for the post will be converted to their root tag when possible.
 
@@
 \textcite{Sahlgren2008} removed stop words for the random-permutation vector-based model to handle this issue.
-\textcite{Jones2007} did as well for the circular-correlation model, but also mentioned that an entropy weighting technique has also been used previously to attenuate (rather than remove) stop words.
+\textcite{Jones2007} did as well for the circular convolution model, but also mentioned that an entropy weighting technique has also been used previously to attenuate (rather than remove) stop words.
 

### 2013-10-10 writeups/proposal/proposal.tex
foo
Former-commit-id: 1697063aef3d7eb3a3409f079809c3963ecc6c03 [formerly 600eef6d1b8808e2ca402441cd51361bb7b6c565]
Former-commit-id: be434b9b216f72183faed8e600ff25830845ce23

@@
 Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
-It's unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
+It is unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
 But it may be the case that using Bayesian computation on a compressed vector-based model improves performance.

### 2013-10-10 writeups/proposal/proposal.tex
foo
Former-commit-id: 2834c40ccf88f5633adb5c1937c780322c1e7385 [formerly ae44f790ffc678356de513b4e2be8662a2396fd4]
Former-commit-id: b06ef2714e97384db55acbba4240a22dcf2facf1

@@
 This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
-Convolution and inverse convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
+Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
 This allows for retrieval requests such as ``playing \#$\Phi$'', ``just read \#$\Phi$'', and ``vector-based $\Phi$ systems''.

### 2013-10-10 writeups/proposal/proposal.tex
foo
Former-commit-id: 013458e253effc27d3755ca5652999cf647b33db [formerly 7b5d530318df6d9595d199e76e2db41178fbe230]
Former-commit-id: 7c6308d01a1a2632ef50c56624a4eaf26c39b6db

@@
 The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
-For example, assume that a user has written the word ``zend'' in a post.
-``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP} ($e_{\mathit{PHP}}$).
-Since the environment vector for \emph{PHP} was added to the memory vector for ``zend'' without any encoding, no decoding is necessary.
-So the cosine between the environment vector \emph{PHP} ($e_{\mathit{PHP}}$) and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.
 
 Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
-For the \emph{PHP} and ``zend'' example, the context memory vector would consist of the unordered environment vector for ``zend''.
-Since the memory vector for \emph{PHP} contains the ``zend'' environment vector, it is likely that the correlation between these two memory vectors will be high.
+For example, assume that a user has written the word ``zend'' in a post, and she wants to assign a tag for the post.
+``zend'' has co-occurred with the tag \emph{PHP} many times previously, so the unordered memory vector for ``PHP'' ($c_{\mathit{PHP}}$) contains the unordered environment vector for \emph{zend} ($e_{zend}$).
+To determine the most correlated memory vector with context, a memory vector is created from the current context ($e_{zend}$).
+The cosine between that context memory vector and the memory vector for \emph{PHP} will be high, since the \emph{PHP} memory vector ($c_{\mathit{PHP}}$) often contains the context environment vector ($e_{zend}$).
+Since the cosine is high, the tag \emph{PHP} will most likely be returned as the vector that resonates highest with the current context.
 

### 2013-10-10 writeups/proposal/proposal.tex
foo
Former-commit-id: 82bf16200aea676d7925092fc1395cbcdbf500c0 [formerly 17c8f537cde8542e193cf718c8131282192ff596]
Former-commit-id: c28162db69a154223350f3b6c8b4309603c65940

@@
 Two of the current state-of-the-art memory retrieval models are based on a co-occurrence methodology, and I'll be comparing these two models for hashtag prediction:
-ACT-R's declarative memory retrieval theory and random permutation vector-based memory systems.
+ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.
 
@@
 
-The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search.
+The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} is also based on the large-scale version of $S_{ji}$, and scaled well when working with a large-scale search query database for internet search.
 
@@
 
-A retrieval for a word following ``stack'' (``stack $\Phi$'') is done by creating a memory vector out of all information in context.
+A resonance retrieval for a word following ``stack'' (``stack $\Phi$'') is done by creating a memory vector out of all information in context.
 In this case, assume only a single piece of order information is available, that ``stack'' precedes the word to retrieve.

### 2013-10-31 writeups/proposal/proposal.tex
just before working on second revision
Former-commit-id: 43865551d9abd53036e8ea1b68130393341268a2 [formerly 3056d53a16e74011590b202e5fbfa0839531c551]
Former-commit-id: 9f7af75ee164b2508d3e39da6c058b74d2c013e4
revised abstract
Former-commit-id: 80dc2558aaf000fd97164d722ac0415f93594453 [formerly 222bb70d73bec02165de827b76b443be2a34bd60]
Former-commit-id: 22db7d8ba8c890415ff0d679ec82682a750e26a2

@@
 \abstract{
-  The amount of user-created content on social media sites continues to grow.
-  To help organize and structure this information, many users create tags for documents to help connect their content to other posts on similar topics and other users that might have similar interests.
-  If an accurate model of the tags users choose when creating content is available, then the system can recommend tags for a user to use, in order to connect the user to similar topics and people as quickly as possible.
-  This works aims to develop and compare tag-recommendation models for users choosing hashtags when composing a tweet on Twitter and choosing tags when asking a question on the StackOverflow site.
-  Two state-of-the-art memory models will be evaluated: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
-  Experiments will be aimed to [1] discover the best way to leverage a user's prior hashtag use as a future predictor, [2] improve the models so that each model's strengths are incorporated into the other,
+  This research aims to develop and compare tag-recommendation models for users choosing hashtags when composing a tweet on Twitter and choosing tags when asking a question on the StackOverflow site.
+  Two state-of-the-art cognitively-plausible memory models will be evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
+  Experiments are designed to [1] discover the best way to leverage a user's prior hashtag use as a future predictor, [2] improve the models so that each model's strengths are incorporated into the other,
   and [3] formally evaluate and compare each model's performance when recommending hashtags on Twitter and tags on StackOverflow.
-  This research will focus on developing user-customized models of hashtag prediction, that best balance a user's past hashtag use and the current content in a post being created to predict the chosen hashtags.
-  Planned modifications to the two memory models are aimed at the architectural level, and may suggest architectural modifications that generalize to other tasks.
-  Finally, the user-customized tag-recommendation models developed with this research may be immediately useful for tag-recommendation systems on the popular Twitter and StackOverflow social media sites.
+  The user-customized tag-recommendation models developed with this research should be immediately useful for tag-recommendation systems on the popular Twitter and StackOverflow social media sites.
+  Further, the planned performance modifications to the two memory models are aimed at the architectural level of each model.
+  If these architectural modifications to two cognitively-plausible memory models significantly improve performance,
+  they may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
 }

### 2013-11-05 writeups/proposal/proposal.tex
foo
Former-commit-id: 27fd11485aa5822e8f449e3ae41f43914ca3ea11 [formerly e4890c68f65f48386999ec46b9d5b5a228cd04a1]
Former-commit-id: 588000624b447413b7624b0f66fd1dd6da270a9f
foo
Former-commit-id: 121441ae8e2b8f8a9d07e29eadbd4f7fbeba92fe [formerly bac466661c1542099fcc2ea310ac08748c62b11b]
Former-commit-id: 126f285cb73d70b1a67ed1f820e08a7605979564

@@
 \usepackage[utf8]{inputenc}
+
+% http://tex.stackexchange.com/questions/36423/random-unwanted-space-between-paragraphs
+\raggedbottom
 
@@
 ACT-R's declarative memory and vector-based memory systems are substantially different in their mathematical structure.
-Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., the ``fan effect'' \parencite{Rutledge2008}).
-I will extend the work done by \textcite{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
-I'll also replace the vector-based model used by \textcite{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
+Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., \cite{Rutledge2008} modeling the ``fan effect'').
+I will extend the work done by \citeauthor{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
+I'll also replace the vector-based model used by \citeauthor{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
 The primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory.

### 2013-11-05 writeups/proposal/proposal.tex
foo
Former-commit-id: 85c979b768ca4a3d89f834ed36c4d3b5b547688e [formerly a390c1a60cd2b273c472fc4091d70357340bd1f8]
Former-commit-id: bd94d28181a4045c7e4ddfe6f05887b26d59ccf9
foo
Former-commit-id: 91cfe1bafe79849daf16f30274ebf85e7f6cf60c [formerly 0df9a17c84c16a4657dd23fbc83907a198e087e7]
Former-commit-id: 12e02085a6c98b2479892d572705c1f1443b2cd7

@@
 \usepackage[utf8]{inputenc}
+
+% Single space after period (Mike prefers)
+\frenchspacing
 
@@
 
-\title{User-customized hashtag/tag prediction on Twitter and StackOverflow:\protect \\ Comparing vector-based and ACT-R memory models of tag retrieval}
+\title{Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow}
 \shorttitle{Tag prediction on Twitter and StackOverflow}
@@
 
-Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp, are composed almost entirely of human-created content.
+Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp are composed almost entirely of human-created content.
 The amount of user content on these sites is staggering, and continues to grow.
-Users of Twitter's microblogging service for example, generate half a billion tweets per day \parencite{TwitterReport2012}.
+Users of Twitter's microblogging service, for example, generate half a billion tweets per day \parencite{TwitterReport2012}.
 However, a single user of a social media site is most likely interested in only a small fragment of this large amount of information.
@@
 
-One traditional technique is to provide the user with a powerful search engine so that they can actively look for content that interests them.
-However, users are not forming the same search queries that work so effectively on search engines like Google.
-A growing number of users on social media sites are searching for information that is happening now, in real time \parencite{Jansen2011}.
+A growing number of users on social media sites are searching for information that is happening in real time \parencite{Jansen2011}.
 In other words, users on social-media sites want the fresh \emph{stream} of information about a particular topic.
-This can certainly be seen in the main user-interface views that these social media site provide:
+This can certainly be seen in the main user-interface views that these social media sites provide:
 Twitter's homepage showing your followees' tweets, Facebook's homepage showing your friends' posts, StackOverflow's daily digest of posts for particular tags.
@@
 
-However, social media sites have an advantage for identifying user goals over traditional search engines: human-created content.
+However, social media sites have an advantage over traditional search engines for identifying user goals: human-created content.
 Twitter has more than just the tweet that a user is currently composing to try and identify that user's goals.
 The site also has all of the user's previous content, and can use that information to provide a much better prediction of a user's goals.
-Further, sites like Twitter, StackOverflow, Google+, and Facebook support ways for the user to explicitly identify their precise goals when creating content: Hashtags.
+Further, sites like Twitter, StackOverflow, Google+, and Facebook support ways for the user to explicitly identify their precise goals when creating content: hashtags.
 These user-created hashtags are a key indicator of a user's goals on a social media site.
@@
 Further, people are using hashtags as a way to create information streams on social media sites \parencite{Kwak2010}.
-Twitter users for example, commonly create and use hashtags for upcoming political events that interest them, such as debates and races \parencite{Diakopoulos2010}.
+Twitter users, for example, commonly create and use hashtags for upcoming political events that interest them, such as debates and races \parencite{Diakopoulos2010}.
 So if we can predict what hashtags a user is interested in, we have a better understanding of the information streams that they care about.
@@
 
-The core research question I am interested in is:
+The core research question motivating this proposal is:
 What kinds of cognitively-plausible models can best predict the human-created tags that a user on a social media site uses?
@@
 Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
-Two of the current state-of-the-art memory retrieval models are based on a co-occurrence methodology, and I'll be comparing these two models for hashtag prediction:
+Two of the current state-of-the-art memory retrieval models are based on a co-occurrence methodology, and I will be comparing these two models for hashtag prediction:
 ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.
@@
 These systems can also help newer users by recommending proper tags for their newest content.
-On the StackOverflow site for example, experts often subscribe to specific hashtags that interest them, and receive a daily digest of posts that are tagged with those hashtags.
+On the StackOverflow site, for example, experts often subscribe to specific hashtags that interest them, and receive a daily digest of posts that are tagged with those hashtags.
 Helping the user properly tag posts on StackOverflow ensures that right community sees the post, which greatly increases the chance that the question on the post will be answered quickly and correctly.
@@
 I will extend the work done by \citeauthor{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
-I'll also replace the vector-based model used by \citeauthor{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
+I will also replace the vector-based model used by \citeauthor{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
 The primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory.
@@
 However, not only should the model properly take into account a user's prior hashtag use, but it should also generate new hashtag predictions when a user is most likely choosing a hashtag they have never used before.
-Calibrating the models to properly balance this exploitation of prior hashtag use versus exploration of new hashtags will be a primary research focus.
+Calibrating the models to properly balance this exploitation of prior hashtag use versus exploration of new hashtags will be a primary research question.
 The end goal is to have a model that can properly balance the user's tag history with the contextual cues in the post in order to generate an accurate prediction.
@@
 One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
-With the ACT-R model for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
+With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
 It may be that multiple co-occurrence matrices must be used to represent word-order co-occurrences and unordered co-occurrences separately.

### 2013-11-05 writeups/proposal/proposal.tex
foo
Former-commit-id: a7c7e6b1de28862dcb4db0f4db8d345f80b00c1e [formerly 8fe6b1259de6a39d5348b7a74a83bc853f627f19]
Former-commit-id: 9dcc86dd3b5db8798478d5be7365c1981676e562

@@
 The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
-The profiles were created by generating a word frequency matrix of words x (user + followers + followees), and then providing that matrix as input into the Lucine text search engine framework.
+The profiles were created by generating a word frequency matrix of words by (user + followers + followees), and then providing that matrix as input into the Lucine text search engine framework.
 When the system needed to recommend a set of individuals for a user to follow, it would take all of the words used in posts from that user to construct a search query to be provided to Lucine.
@@
 The system then retrieved the most likely hashtags that were associated with that search query.
-The model is based on building a word co-occurrence matrix of words x hashtags, and then uses that matrix to assign a similarity score for each hashtag when provided with a search query.
+The model is based on building a word co-occurrence matrix of words by hashtags, and then uses that matrix to assign a similarity score for each hashtag when provided with a search query.
 Participants graded how relevant the hashtags returned were to each search query.
@@
 The process starts with a set of N documents.
-A word x document co-occurrence matrix is built by counting the number of times each word appears in each document.
+A word by document co-occurrence matrix is built by counting the number of times each word appears in each document.
 The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
@@
 
-LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word x document co-occurrence matrix.
+LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix.
 The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
@@
 Although the LSA model has been shown to perform well as a way to measure similarity between two terms, one of the main criticisms is that it is a ``bag of words'' model.
-This means that the model does not account for or pay attention to word order when building the base word x document matrix or consequently, when performing the SVD dimension reduction.
+This means that the model does not account for or pay attention to word order when building the base word by document matrix or consequently, when performing the SVD dimension reduction.
 Other models that can account for word order have been shown to perform better than ``bag of words'' models on tasks that do not even appear to require grammatical structure such as word order to do well.
@@
 Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
-As the size of the original word x document matrix increases, the SVD computation becomes more and more time consuming.
+As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
 If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
@@
 Vector-based representations compute chunk likelihoods by performing a correlation operation directly on the memory vectors.
-Each memory vector can be thought of as a column on a context word x tag co-occurrence matrix used for a Bayesian representation.
-Further, the number of rows for a vector-based representation are fixed and set much smaller than the number of rows in the full word x tag co-occurrence matrix.
-Since Bayesian representations have been shown to scale to over hundreds of millions of word x tag co-occurrences \parencite{Stanley2013}, the compressed vector-based representations should scale as well or better.
+Each memory vector can be thought of as a column on a context word by tag co-occurrence matrix used for a Bayesian representation.
+Further, the number of rows for a vector-based representation are fixed and set much smaller than the number of rows in the full word by tag co-occurrence matrix.
+Since Bayesian representations have been shown to scale to over hundreds of millions of word by tag co-occurrences \parencite{Stanley2013}, the compressed vector-based representations should scale as well or better.
 
@@
 and also scale to larger datasets (better computational efficiency).
-The vector length x performance tradeoff was tested on a simple paired associative retrieval task. 
+The vector length by performance tradeoff was tested on a simple paired associative retrieval task. 
 Retrieval accuracy for the random-permutation vector-based models remained high relative to the circular convolution model even after reducing each environment vector's length from \num{2048} to \num{1024}.
@@
 Both random permutations and LSA calculate word similarity on a reduced matrix.
-For LSA, the original word x document matrix is compressed to a factor x document matrix and the number of factors are much less than the number of words.
+For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors are much less than the number of words.
 For the StackOverflow and Twitter datasets, the number of documents are the number of different tags.
@@
 Note that if the length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular row in the column vector.
-If done in this way, building a matrix of all words' environment vectors x tags would recover the original word x document matrix \parencite{Kanerva2000}.
+If done in this way, building a matrix of all words' environment vectors by tags would recover the original word by document matrix \parencite{Kanerva2000}.
 However, each word's environment vector is compressed using a length much less than the number of unique words and using a combination of a few $\{1,-1\}$s to represent the word.

### 2013-11-05 writeups/proposal/proposal.tex
foo
Former-commit-id: 5d31e8c29019c95ea3e991bbf18329725c1bd47b [formerly d0c8fd955214b5a76f4b51ec62814e95dfc2fe73]
Former-commit-id: 36ce7fbc39478e224d4754d001482d25d271f0bb

@@
 The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
-The profiles were created by generating a word frequency matrix of words by (user + followers + followees), and then providing that matrix as input into the Lucine text search engine framework.
-When the system needed to recommend a set of individuals for a user to follow, it would take all of the words used in posts from that user to construct a search query to be provided to Lucine.
+The profiles were created by generating a word frequency matrix of words by (user + followers + followees), and then providing that matrix as input into the Lucene text search engine framework.
+
+Lucene \parencite{McCandless2010} is a commonly-used open-source software tool that measures the similarity between a query and candidate documents in a database.
+It uses the term frequency--inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
+
+When the system needed to recommend a set of individuals for a user to follow, it would take all of the words used in posts from that user to construct a search query to be provided to Lucene.
 That search query would return the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
@@
 A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
-Models for Twitter and StackOverflow are also being proposed and tested, but none have yet been deployed.
+Models for Twitter and StackOverflow have also been proposed and tested, but none have yet been deployed.
 
@@
 \textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
-\textcites{Zangerle2011, Kywe2012} use the Apache Lucine infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarities.
+\textcites{Zangerle2011, Kywe2012} use the Apache Lucene infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarities.
 Regardless of the method to compute the similarity between tweets, each method collects the hashtags used from a set of the most similar tweets, then ranks the set, and presents the top 5-10 hashtags to the user.

### 2013-11-05 writeups/proposal/proposal.tex
foo
Former-commit-id: 0edfad65b9b09b071d96f9418c3d2ef1e6ef8598 [formerly c0ade82b37f5679c6a97f9fc65950aa1fc35dee3]
Former-commit-id: 5225c957013eb076a827ed9bbb4c25954f1e8a7c

@@
 The model was based on ACT-R's declarative memory retrieval theory, which at the core is also a Bayesian co-occurrence model.
-However, this model scaled much better than the co-occurrence model presented in \textcite{Kuo2011}.
+However several modifications to the model were made from the standard Bayesian ACT-R declarative memory theory, so it was more sophisticated than the standard ACT-R model and the model used by \textcite{Kuo2011}.
+\citeauthor{Stanley2013} used a data-driven entropy weighting measure to attenuate low-predictor words and increase the weighting of high-predictor words in the body and title of a post.
+An offset term was also included which leveled the model's overall prediction for each post and enabled the use of a logistic regression statistical technique to calibrate the model's parameters.
+A separate model component was used for the words in the body and title of the post, and it was unclear if \citeauthor{Kuo2011} used the words in the body of the post for his models.
+
+\citeauthor{Stanley2013}'s model also scaled much better than \citeauthor{Kuo2011}'s co-occurrence model. 
 Because of this, 100 times more posts were used to build the co-occurrence matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
-After scaling up the model, this model can successfully predict a user's tag for a post at 65\% accuracy.
+After including the modifications and scaling up the model, the model can successfully predict a user's tag for a post at 65\% accuracy.
+
 It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
@@
 A tag suggestion tool has been deployed for the \url{http://tex.stackexchange.org} \parencite{LatexTags2013} and \url{http://superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
-\LaTeX, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
+LaTeX, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
 For the two sites where a tag suggestion tool has been deployed, once a user has created a question, suggested tags appear just below the text field where the user types the tags to associate with the question.

### 2013-11-05 writeups/proposal/proposal.tex
foo
Former-commit-id: c8b9b1ca53b26262082d90559715ad843dc7ca63 [formerly 0782f334feafc62beedfd0e7527d5e1d079bbe51]
Former-commit-id: 58595f5bc64a148f7486546d4f587a0ed36b9ea8

@@
 
+The exact tag-recommendation algorithm used for the LaTeX and SuperUser sites is not publicized.
+StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags, as the algorithm has not yet been deployed for the busiest StackOverflow site.
 The Latex and SuperUser sites are not nearly as popular as StackOverflow currently.
 The sites have roughly \num{50000} and \num{189000} questions respectively compared to \num{5000000} questions on StackOverflow.
-It may be the case that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
-Or it is possible that their current tag recommendation system is not efficient enough to work with the demand of the StackOverflow site.
+
+So it stands to reason that the algorithm is relatively complex, and there are possible scaling and efficiency concerns with the algorithm's implementation.
+Or it is also possible that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
 Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial to the user and has started to add them to their question and answer sites.
@@
 The user is free to remove any or all of the tags that the system has associated with the post, and can also use tags that were not included in the recommended set.
+
+Google+'s tag-recommendation algorithm is not publicized, but it appears to be rather sophisticated.
+Looking at the example post in Figure \ref{figGoogle+Hashtags}, \#BusterPosey is suggested without having that specific word in the post.
+Also \textcite{GoogleKeynote2013} mentioned that Google uses its sophisticated image recognition software to suggest hashtags for images that are included in the post (e.g., \#EiffelTower),
+even when the post does not have any keywords to provide additional context.
 

### 2013-11-05 writeups/proposal/proposal.tex
foo
Former-commit-id: 63995900765b0c801ce041aaddd6fae53d96882f [formerly a5e8f67726fe5f9f5383f8af8d2701cb5c7d92a3]
Former-commit-id: 95bbbb5a52d52a836c99c1d93201aeac69e4464b

@@
 Google+'s tag-recommendation algorithm is not publicized, but it appears to be rather sophisticated.
-Looking at the example post in Figure \ref{figGoogle+Hashtags}, \#BusterPosey is suggested without having that specific word in the post.
-Also \textcite{GoogleKeynote2013} mentioned that Google uses its sophisticated image recognition software to suggest hashtags for images that are included in the post (e.g., \#EiffelTower),
+Looking at the example post in Figure \ref{figGoogle+Hashtags}, \emph{\#BusterPosey} is suggested without having that specific word in the post.
+Also \textcite{GoogleKeynote2013} mentioned that the algorithm uses Google's sophisticated image recognition software to suggest hashtags for images that are included in the post (e.g., \emph{\#EiffelTower}),
 even when the post does not have any keywords to provide additional context.

### 2013-11-05 writeups/proposal/proposal.tex
foo
Former-commit-id: 77c747083cb5c1981f8f873667c6df5bca5ce545 [formerly 1a3099eb56bb2d31dd595fab9960aeb5a6aacb23]
Former-commit-id: b553b8cc5e91fc0a4be2366279ab8aabb6797d38

@@
 
-With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few $\{1,-1\}$s in random locations. 
-In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector ($e_{i}$) is shifted to the left one position ($^{-1}$).
+With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few one and negative one values in random locations. 
+In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is shifted to the left one position.
 This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
@@
 If done in this way, building a matrix of all words' environment vectors by tags would recover the original word by document matrix \parencite{Kanerva2000}.
-However, each word's environment vector is compressed using a length much less than the number of unique words and using a combination of a few $\{1,-1\}$s to represent the word.
+However, each word's environment vector is compressed using a length much less than the number of unique words and represented by using a combination of a few one and negative one values at a unique combination of positions.
 
@@
 
-However, vector-based models and ACT-R's current Bayesian DM model are not identical, and these differences need to be addressed in order to better evaluate the benefits of one over the other.
+Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the benefits of one over the other.
 For example, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
@@
 These models will be the starting point for modeling tag use on the Twitter and StackOverflow sites.
-The models may be modified during testing if the data indicate that modifications are needed to improve model performance.
-Also there are planned modifications to each model as part of the research:
+The models may be modified during testing if the data indicate that modifications will increase the accuracy in which the model's recommended tags agree with the post author's chosen tags.
+Also there are planned modifications to each model as part of the research, as these modifications seem likely to improve performance: 
 incorporating user prior information into the vector-based model and incorporating word order information into the Bayesian model.

### 2013-11-05 writeups/proposal/proposal.tex
foo
Former-commit-id: bdcde8597a5b4028f9712084df8c3602105541d6 [formerly 366a9c0de9cee936fa2f954bdf4491c3fd3ca3d0]
Former-commit-id: 4bb34949b10e00addefec65f2f72ce37b440617c

@@
 
+One straightforward way to include a user's prior tag history into a random permutation model is to include ACT-R's base-level activation term ($B_{i}$) as an additional term in the model.
+However this may not be the most mathematically appropriate thing to do, since the base-level activation is represented as a log odds value,
+and the usual single strength of association term in a random permutation model is a correlation metric.
+It may be more appropriate to somehow attenuate the memory vectors over time,
+so that an association between a word in a post and a tag (i.e., a memory vector) that has long passed is no longer included in that memory vector's representation.
+I will try these ideas and any others that are discovered during the model development process, and the one that is the most cognitively plausible and accurate will be chosen.
+
 \subsubsection{Experiment 3: Incorporating Word Order Into ACT-R}
@@
 So a method to include word order into a Bayesian model will be developed.
+
+One method to include word order in a Bayesian co-occurrence matrix is to represent word position on a third, orthogonal dimension, and turn the matrix into a 3-dimensional array.
+So a co-occurrence is represented by a triplet: context word by word position relative to tag by tag.
+When a retrieval request for a tag is made, the strength of association between each word in context (and its position) and the tags in the co-occurrence array would be computed.
+That computation could be made tractable by storing only a small set of orders on the dimension axis and binning.
+
+For example, one could store only three: near, medium, far, which is equivalent to storing three separate co-occurrence matrices.
+So when a retrieval request is made (e.g., when a user writes the \# character when composing a tweet), the strength of association for each word would be computed using the appropriate matrix.
+That is, the ``far'' matrix would be used for words far away from the cursor position, ``near'' for words just written, and ``medium'' for words mid distance from the current focus point.
+This method seems straightforward to implement, but it is slightly unsatisfying since information is lost when binning is used.
+So I will try to discover a better technique to incorporate word order information into the ACT-R Bayesian model than the one just described, but will fall back to this technique if a better method is not found.
+
 This updated model will then be evaluated on only the Twitter common-hashtags dataset, since this dataset contains hashtags that occur in positions throughout the tweet (e.g., embedded within or at the end of a tweet).

### 2013-11-05 writeups/proposal/proposal.tex
foo
Former-commit-id: 6dd3aef4bf858f8aa203d221a32a175cce575071 [formerly be7e18366dbc8d2350cfdd87fb1eb65ae616d993]
Former-commit-id: 8d00c6a2f5d99e965343f471fff62e29a16f25d9

@@
 When a retrieval request for a tag is made, the strength of association between each word in context (and its position) and the tags in the co-occurrence array would be computed.
-That computation could be made tractable by storing only a small set of orders on the dimension axis and binning.
+That computation could be made tractable by storing only a small set of values on the order axis and binning.
 
@@
 This method seems straightforward to implement, but it is slightly unsatisfying since information is lost when binning is used.
+It also does not scale, since an additional 2-dimensional sparse matrix must be added for every additional value stored on the word order dimension.
 So I will try to discover a better technique to incorporate word order information into the ACT-R Bayesian model than the one just described, but will fall back to this technique if a better method is not found.

### 2013-11-06 writeups/proposal/proposal.tex
foo
Former-commit-id: 56d9faa6b57b2df84d0a3ab1b780bbec5a1a38d6 [formerly 4466f8b240312acf099b75bfa6ccd8e48c4d1574]
Former-commit-id: 37cfe94a0c68e3f129cbbf5aa12a4170098ce331
foo
Former-commit-id: 1fc41395d0c9b22d23fc56ac8b564be753d3f17f [formerly 177837ab8abf7673ca430b113106812d2d421152]
Former-commit-id: 45be167e1909f3342f3cca1c2236f8c94da8a0b9

@@
 \title{Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow}
-\shorttitle{Tag prediction on Twitter and StackOverflow}
+\shorttitle{Comparing memory models of tag retrieval}
 %\shorttitle{}
@@
 
-The end result of this research will be a better understanding of how a user's past tagging history influences future tag use,
-as well as two cognitively-plausible models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
-Overall performance differences between an ACT-R's Bayesian retrieval model and a random-permutation vector-based model will be directly compared on two real-world tagging tasks, which has never been done previously.
-The strengths of each model (word order and user prior likelihood) will be incorporated into the other, and performance differences will be evaluated.
+The end result of this research will be [1] a better understanding of how a user's past tagging history influences future tag use,
+[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on a large-scale real-world retrieval task,
+[3] improvements to each memory model where the strengths of each model (word order and user prior likelihood) are incorporated into the other,
+and [4] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets.
+
+The tag-recommendation models developed through this research should be immediately relevant for real-world tasks.
+If model accuracy is high enough, then the developed models may be directly deployed as tag recommendation systems for either Twitter or StackOverflow.
+It is also likely that the developed models will scale if deployed to these two sites, since during development they will be tested on a large portion of each site's currently available data.
+
+This research also aims to find task-agnostic memory models that can be applied across a wide range of domains.
+As much as possible, the two developed models that are evaluated on the StackOverflow and Twitter datasets will not differ across the datasets.
+This should help ensure that the developed models are accurate in general, and can potentially be used as tag recommendation systems for other domains.
+Examples of similar tagging tasks are hashtag use on Facebook, flagging (i.e., tagging) Wikipedia articles for content issues or overly-opinionated material,
+auto labeling incoming mail and placing it into appropriate folders, and spam filtering.
+Having an accurate, efficient co-occurrence memory model that correctly retrieves relevant tags can be applied to a wide variety of real-world tasks.
+
+This research also has numerous components that are relevant to improving our understanding of declarative memory.
+It should be emphasized that it has only recently become possible to evaluate psychological theories of declarative memory (e.g., ACT-R Bayesian, random-permutation vector based)
+on large-scale real-world tasks where the theory can be tested on hundreds of millions of data points.
+These information-rich large-scale environments provide a unique opportunity to really stress test and explore the impacts of the psychological constraints of each theory on a much larger scale than was previously possible.
+This environment will allow, for example, overall performance differences between an ACT-R Bayesian retrieval model and a random-permutation vector-based model to be directly compared on two real-world tagging tasks,
+which has never been done previously.
+
+This data-rich environment will also enable the proposed modifications to each model to be quickly evaluated, so that further model improvements can be iteratively proposed and tested.
 The main two proposed model modifications (incorporating word order into ACT-R and user history into vector-based representations) have not been tested before, and seem likely to significantly improve performance.
-Further, these two modifications in particular are aimed at the architectural level of each model, may very well be task independent, and may generalize and improve model fit across other tasks.
-Finally, if model accuracy is high enough, then the developed models may be applied to general tag recommendation systems for either Twitter or StackOverflow,
-and possibly for other sites where users generate and tag content.
+
+Further, the proposed model modifications are aimed at the architectural level, may very well be task independent, and may generalize and improve model fit across other tasks.
+Discovering an effective way to represent the time decay of information in a vector-based model could improve model fit on other tasks where the memory model has been tested. 
+Finding a task-agnostic way to represent word order information in a Bayesian ACT-R model could improve model accuracy when applied to other domains where it seems plausible that participants are utilizing word order.
+Introducing, exploring, and evaluating reasonable architectural modifications to two commonly-used memory models is worth the effort:
+It will increase our understanding of how the effects of imposing specific architectural constraints on memory models influence what information is retrieved.
 

### 2013-11-06 writeups/proposal/proposal.tex
foo
Former-commit-id: 6f33a13acf79593b4d3fbd36583210f58a52dc85 [formerly 3af3fba524bb1edaacd3fcfb4e12fd972b2cabb1]
Former-commit-id: a8fa7817fe0fb7a3741fbdde67e5d9f91bd038d5

@@
 
-\subsection{Recommendation Systems}
-
-A model that predicts a user-chosen hashtag is a specific type of a recommendation system.
-These general recommendation systems have grown in need with the growth of the web.
-Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
-Recommendation systems are used to tailor the information presented to what is most relevant to the user \parencite{Pazzani2007}.
-
-\subsubsection{The Netflix Prize}
-
-The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
-Netflix is an online-subscription service where users can stream their favorite television episodes and movies.
-Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
-In order to provide a high-quality user experience on the site, Netflix created a content-recommendation service.
-This tool recommends to the user movies and television episodes that most likely match their specific interests.
-
-Netflix wanted to improve the accuracy of their recommendation system, so they held a competition where teams could try to improve on the accuracy of the service.
-The winning team utilized a statistical model that performed a type of factor analysis to identify important dimensions in the data (e.g., drama, action, cult classic) and place values for each user along each dimension.
-This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
-By using a factor-analysis technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.
-
-\subsubsection{Recommending Followers on Twitter}
-
-Recommendation systems based on the content of user-created posts have been successful in recommending people on Twitter for a user to follow.
-\textcite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
-The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
-The profiles were created by generating a word frequency matrix of words by (user + followers + followees), and then providing that matrix as input into the Lucene text search engine framework.
-
-Lucene \parencite{McCandless2010} is a commonly-used open-source software tool that measures the similarity between a query and candidate documents in a database.
-It uses the term frequency--inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
-
-When the system needed to recommend a set of individuals for a user to follow, it would take all of the words used in posts from that user to construct a search query to be provided to Lucene.
-That search query would return the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
-Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user follows at 25\% accuracy.
-
-\subsubsection{Hashtags for Recommendation Systems}
-
-\textcite{Efron2010} showed how hashtags might be used as a type of recommendation system for Twitter.
-Users constructed a query of text that represented a search for a topic that they were interested in.
-The system then retrieved the most likely hashtags that were associated with that search query.
-The model is based on building a word co-occurrence matrix of words by hashtags, and then uses that matrix to assign a similarity score for each hashtag when provided with a search query.
-Participants graded how relevant the hashtags returned were to each search query.
-The results suggest that hashtags contain useful information that help classify the gist of a post. 
-Although this model does not predict the specific hashtags used when composing a tweet, a model like this could certainly be tailored for that purpose.
-
-\subsection{Hashtag Prediction}
-
-More specific hashtag-recommendation models that attempt to predict a user's chosen hashtag have also been created and tested for a few social media sites.
-Google has even deployed a hashtag-recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
-A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
-Models for Twitter and StackOverflow have also been proposed and tested, but none have yet been deployed.
-
-\subsubsection{StackOverflow}
-
-\textcite{Kuo2011} was the first to model tag use for StackOverflow posts.
-His set of models were originally designed for next-word prediction for large text document collections.
-In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user created a post, and restrict the space of possible words to only tags.
-He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
-Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive Singular Value Decomposition and K-nearest-neighbors models.
-This Bayesian model can accurately predict a user-chosen tag at 47\% accuracy.
-
-\textcite{Stanley2013} also modeled a user's chosen tags for a post on the StackOverflow site.
-The model was based on ACT-R's declarative memory retrieval theory, which at the core is also a Bayesian co-occurrence model.
-However several modifications to the model were made from the standard Bayesian ACT-R declarative memory theory, so it was more sophisticated than the standard ACT-R model and the model used by \textcite{Kuo2011}.
-\citeauthor{Stanley2013} used a data-driven entropy weighting measure to attenuate low-predictor words and increase the weighting of high-predictor words in the body and title of a post.
-An offset term was also included which leveled the model's overall prediction for each post and enabled the use of a logistic regression statistical technique to calibrate the model's parameters.
-A separate model component was used for the words in the body and title of the post, and it was unclear if \citeauthor{Kuo2011} used the words in the body of the post for his models.
-
-\citeauthor{Stanley2013}'s model also scaled much better than \citeauthor{Kuo2011}'s co-occurrence model. 
-Because of this, 100 times more posts were used to build the co-occurrence matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
-After including the modifications and scaling up the model, the model can successfully predict a user's tag for a post at 65\% accuracy.
-
-It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
-This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular environment
-(i.e., when predicting the user-chosen tags for human-created content on social media sites).
-
-\subsubsection{StackExchange}
-
-A tag suggestion tool has been deployed for the \url{http://tex.stackexchange.org} \parencite{LatexTags2013} and \url{http://superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
-LaTeX, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
-For the two sites where a tag suggestion tool has been deployed, once a user has created a question, suggested tags appear just below the text field where the user types the tags to associate with the question.
-The post author is not forced to use the suggested tags, and is free to choose tags from them or tags not represented in the suggested set.
-An example of the interface for a created post on the SuperUser site is included in Figure \ref{figSuperUserSuggestion}.
-
-\begin{figure}[!htbp]
-  \scalebox{.9}{\resizebox{\linewidth}{!}{\includegraphics{superUserTagSuggestion-crop.pdf}}}
-  \caption{Example of tag suggestions when creating a post on the SuperUser site}
-  \label{figSuperUserSuggestion}
-\end{figure}
-
-The exact tag-recommendation algorithm used for the LaTeX and SuperUser sites is not publicized.
-StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags, as the algorithm has not yet been deployed for the busiest StackOverflow site.
-The Latex and SuperUser sites are not nearly as popular as StackOverflow currently.
-The sites have roughly \num{50000} and \num{189000} questions respectively compared to \num{5000000} questions on StackOverflow.
-
-So it stands to reason that the algorithm is relatively complex, and there are possible scaling and efficiency concerns with the algorithm's implementation.
-Or it is also possible that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
-Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial to the user and has started to add them to their question and answer sites.
-
-\subsubsection{Google+}
-
-The hashtag recommendation system deployed for the Google+ microblogging service is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
-The user interface for this recommendation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.
-
-\begin{figure}[!htbp]
-  \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample-crop.pdf}}}
-  \caption{Example post on Google+}
-  \label{figGoogle+Hashtags}
-\end{figure}
-
-The workflow and user interface provided for recommending hashtags for Google+ is similar to how a hashtag recommendation system for Twitter might look.
-When generating content on Google+, the system automatically associates the most likely relevant hashtags with that post.
-For the example in Figure \ref{figGoogle+Hashtags}, the system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-created post.
-The user is free to remove any or all of the tags that the system has associated with the post, and can also use tags that were not included in the recommended set.
-
-Google+'s tag-recommendation algorithm is not publicized, but it appears to be rather sophisticated.
-Looking at the example post in Figure \ref{figGoogle+Hashtags}, \emph{\#BusterPosey} is suggested without having that specific word in the post.
-Also \textcite{GoogleKeynote2013} mentioned that the algorithm uses Google's sophisticated image recognition software to suggest hashtags for images that are included in the post (e.g., \emph{\#EiffelTower}),
-even when the post does not have any keywords to provide additional context.
-
-Google is apparently highly confident in the accuracy of their system.
-On the Latex site, the user interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
-For Google+, the user interface automatically tags a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
-If tag accuracy is high enough, the Google+ technique is certainly favorable since the user does not have to spend time figuring out which tags best represent this post.
-However, if the model makes even just a few errors, users may become annoyed and frustrated with the recommendation system since they will have to explicitly correct it every time they generate a post.
-
-\subsubsection{Twitter}
-
-Several hashtag recommendation models for Twitter have been developed recently.
-One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a post as a function of prior hashtag use and context \parencite{Mazzia2009}.
-The model presented by \textcite{Mazzia2009} is very similar to an ACT-R declarative retrieval model.
-The main potential difference is that the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
-Instead, the global prior is an overall average frequency of hashtag use across users.
-
-Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used from similar tweets.
-\textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
-\textcites{Zangerle2011, Kywe2012} use the Apache Lucene infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarities.
-Regardless of the method to compute the similarity between tweets, each method collects the hashtags used from a set of the most similar tweets, then ranks the set, and presents the top 5-10 hashtags to the user.
-
-This tweet-centered approach is quite different than the hashtag-centered approach presented in \textcite{Mazzia2009}.
-With a hashtag-centered approach, a direct association is built between the words in a tweet and hashtags that occur in tweets.
-With a tweet-centered approach, that association between words and hashtags is indirect:
-The relationship is built between a tweet and its content, and then similar tweets are assumed to use similar hashtags.
-One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a representation of the content in each tweet has to be maintained to later compute tweet similarity.
-The storage size for a hashtag-centered approach grows with the number of hashtags, but it seems reasonable that this will grow slower than the number of tweets and asymptote after collecting a large sample of tweets.
-So from the perspective of efficient information storage and retrieval, it seems more likely that the hashtag-centered approach used in \textcite{Mazzia2009} will scale as Twitter grows in size and use.
-
-\textcite{Kywe2012} also customized their tag prediction model to the user's past hashtag use, and was one of the first to do so when developing a hashtag-prediction model for Twitter.
-The set of recommended hashtags was the union of hashtags used in similar tweets and hashtags used by similar users.
-This was done by storing a content vector of hashtag use for each user (user-centered approach) alongside the content vectors of hashtag use for each tweet (tweet-centered approach).
-To generate recommended hashtags, the model would return the top-ranked hashtags from the hashtags used in 0-50 similar tweets and 0-4 similar users.
-Prediction accuracy improved when the model combined recommendations based on hashtags in both similar tweets and the top few similar users compared to basing recommendations on similar tweets alone.
-
-A topics-based approach was used by \textcite{Godin2013}, where the model predicted the most likely set of topics that were associated with a tweet, and tested if these topics matched the tweet by using human raters.
-The latent topics for the collection of tweets were generated using a Latent Dirichlet Allocation (LDA) statistical methodology.
-This is similar to a factor analysis approach in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance.
-Human graders rated how well the models generated topics for a tweet related to the content in the tweet.
-The LDA model performed well, and generated relevant topics.
-When the model produced five most-likely hashtags for each tweet, 80\% of the time at least one of those hashtags was suitable (as determined by human raters).
-
-However, a topic-generating model is not performing the same task as a hashtag-prediction model.
-The authors argued that these general topics could be used to categorize the post, in addition to the hashtags already used by the author in the post.
-It is unclear though if a user would want to label their tweet with these general topics.
-Perhaps they would rather use more specific hashtags to be more precise in how they label the tweet and how that tweet is associated with tweets within the rest of the community.
-In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
-Now it is certainly the case that the topics generated by a latent-based model may be similar to the hashtags used in the post.
-However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
-Why not take those topics already present in the data and build a model that produces them?
-In other words, the topics data are already in the tweet, so why not use them?
-
-\subsubsection{State of Research on StackOverflow and Twitter}
-
-Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
-I am curious to see the prediction difference between a vector-based model and current co-occurrence models on the StackOverflow dataset.
-Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
-Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, it will be interesting to see how a cognitively-plausible vector-based model performs at the task.
-
-Also, a vector-based retrieval model has not yet been tested on Twitter.
-One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
-It seems quite likely that words used just before a hashtag in a tweet (combined with their position) will be highly predictive of the hashtag chosen.
-To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
-It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.
-
-There has been a mix of co-occurrence and latent-based models created to predict hashtag use on Twitter, and the results are encouraging.
-However, there has been very little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
-\textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
-The ACT-R declarative memory retrieval theory shows how past user behavior and current context combine to produce the most likely retrieval.
-Each component is summed together to generate a total activation, and then the likelihood of chunks are ranked by that summed activation.
-This technique is compensatory and allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
-This type of weighted additive model that combines past user behavior and current tweet context to predict the most likely hashtags has not yet been explored for Twitter.
-
-Both StackOverflow models \parencites{Stanley2013,Kuo2011} did not take into account the specific user's prior history when generating tag predictions.
-Rather their models, much like \textcite{Mazzia2009}'s Twitter model, uses the aggregate tag history across all users as the model's prior activation for each tag.
-One of the main purposes of this research is to improve on these user-agnostic Bayesian models and include each specific user's past tagging history as a model component.
-Using customized priors that are tailored to a user's specific tagging history should be crucial for sites like Twitter where the possible tag space is infinite,
-but should also improve performance for more constrained sites like StackOverflow since each user is not interested in all of the possible tags on the site.
-
 \subsection{ACT-R Declarative Memory Theory}
@@
 
-%The SNIF-ACT model uses strength of association to measure the relationship between the words in a search query and the words contained in each query result.
-%A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.
-%The links that user chose in experimental testing were largely predictive of the strength of association measure.
-
-%TODO: This paragraph doesn't fit well after the previous one, and in the scaling section
-%The predictive power of the SNIF-ACT model went above and beyond what would be predicted by simply assuming that the user always clicks on the first link in the search results page.
+\subsection{Latent Semantic Analysis Memory Theory}
+
+Latent Semantic Analysis \parencite{Landauer1997} (LSA) is a technique that has mathematical roots in factor analysis and also measures strength of association between two terms.
+The process starts with a set of N documents.
+A word by document co-occurrence matrix is built by counting the number of times each word appears in each document.
+The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
+This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
+The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
+
+Usually information is lost when the data are restructured into a smaller set of latent dimensions.
+However it was shown that this process actually improves prediction and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
+\textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
+Model results were on par with a large sample of applicants to american universities from non-English speaking countries (64.4\% compared to 64.5\%).
+Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
+Afterwards further reductions in the dimensionality of the data began to degrade performance.
+
+\subsubsection{Singular Value Decomposition}
+
+LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix.
+The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
+The best way to remove a dimension while maintaining the most variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
+For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
+By pooling co-occurrence counts the model is able to generate better predictions with less data.
+
+\subsubsection{Word Order}
+
+Although the LSA model has been shown to perform well as a way to measure similarity between two terms, one of the main criticisms is that it is a ``bag of words'' model.
+This means that the model does not account for or pay attention to word order when building the base word by document matrix or consequently, when performing the SVD dimension reduction.
+Other models that can account for word order have been shown to perform better than ``bag of words'' models on tasks that do not even appear to require grammatical structure such as word order to do well.
+For example, \textcite{Jones2007} compared a vector-based model called BEAGLE directly to LSA, training and testing both models on the same corpus used in \textcite{Landauer1997}.
+The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the task (55.6\% vs. 55.3\%).
+However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to 57.5\%.
+It seems reasonable to expect that task performance will improve much more on tasks where word order is used (e.g., generating a hashtag in the middle of a sentence when composing a tweet).
+
+\subsubsection{Scaling Issues}
+
+Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
+As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
+If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
+However, with large-scale human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
+\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
+Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
+This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.
+
+When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
+This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
+However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
+\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
+This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
+Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to 73.75\%, which is 10\% higher than the TASA-trained LSA.
+This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.
+
+\subsubsection{Cost of Incremental Updating}
+
+Another issue with LSA is that adding additional observations to an already-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
+Since computing the SVD is the most computationally-expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
+For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.
+
+\subsection{Vector-Based Memory Systems}
+
+``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., chunk) as a vector.
+The representation of the concept is distributed across all elements of the vector.
+This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and viewing the distribution of co-occurrence counts across the rows in the column as the representation of that concept.
+However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.
+
+For vector-based systems, each word is represented by an environment vector $e_{j}$ that is nearly orthogonal to all other words' environment vectors.
+The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix, which is how vector-based systems represent information in a much more compact space.
+Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occurred with that $e_{i}$ environment vector.
+These memory vectors can contain environment vectors from bag-of-words co-occurrence information ($c_{i}$) and word order information ($o_{i}$), which can be combined into a single representation ($m_{i}$).
+Over time, memory vectors accumulate a distributed representation of the most common environment vectors that co-occurred with them.
+
+\subsubsection{Retrieval Process}
+
+Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
+Decoding takes the memory vector for context and decodes it back into an environment vector.
+The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
+
+Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
+For example, assume that a user has written the word ``zend'' in a post, and she wants to assign a tag for the post.
+``zend'' has co-occurred with the tag \emph{PHP} many times previously, so the unordered memory vector for ``PHP'' ($c_{\mathit{PHP}}$) contains the unordered environment vector for \emph{zend} ($e_{zend}$).
+To determine the most correlated memory vector with context, a memory vector is created from the current context ($e_{zend}$).
+The cosine between that context memory vector and the memory vector for \emph{PHP} will be high, since the \emph{PHP} memory vector ($c_{\mathit{PHP}}$) often contains the context environment vector ($e_{zend}$).
+Since the cosine is high, the tag \emph{PHP} will most likely be returned as the vector that resonates highest with the current context.
+
+The resonance and decoding processes are inversions of each other.
+Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
+\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
+However, decoding does require an additional decoding step that resonance does not require, and depending on the decoding operation, this step may be computationally expensive.
+
+\subsubsection{Addressing Word Order}
+
+One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
+Representing word order is a case where both the encoding and decoding operations are applied.
+As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
+Simply counting co-occurrences of the two immediately-preceding words and hashtags does not include the order information of the words in the representation.
+With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
+Order information is lost if a simply co-occurrence counting technique is used.
+
+With vector-based systems, order information is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
+This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
+Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
+This allows for retrieval requests such as ``playing \#$\Phi$'', ``just read \#$\Phi$'', and ``vector-based $\Phi$ systems''.
+
+For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
+Within this corpus, the phrase ``martin luther'' often occurred.
+When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated by convolving a 1-back location vector with luther ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
+Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin with and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
+The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.
+
+Using decoding, the word-order memory vector for martin is deconvolved to the right ($o_{martin} \circledast \Phi$) which returns the environment vector for luther ($e_{luther}$).
+Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which correlates highest with the word-order memory vector for luther ($o_{luther}$).
+Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$''.
+
+New vectors generated by circular convolution are uncorrelated with all other environment vectors and have the same length.
+This allows the word order information and unordered context information for a chunk to be aggregated and represented in the same memory vector.
+Quite simply, the different types of memory vectors can be summed to create a single memory vector representation for each chunk (i.e., $m_{j} = c_{j} + o_{j}$).
+
+\subsubsection{Addressing Scalability}
+
+Vector-based representations compute chunk likelihoods by performing a correlation operation directly on the memory vectors.
+Each memory vector can be thought of as a column on a context word by tag co-occurrence matrix used for a Bayesian representation.
+Further, the number of rows for a vector-based representation are fixed and set much smaller than the number of rows in the full word by tag co-occurrence matrix.
+Since Bayesian representations have been shown to scale to over hundreds of millions of word by tag co-occurrences \parencite{Stanley2013}, the compressed vector-based representations should scale as well or better.
+
+\subsubsection{BEAGLE}
+
+\textcite{Jones2007} created the BEAGLE vector-based memory representation.
+Each memory vector ($m_{j}$) is a summed representation of unordered and ordered word co-occurrences ($c_{j} + o_{j}$).
+This model uses circular convolution and circular correlation as a way to encode and decode word order.
+BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (55.6\% compared to 55.3\%).
+Also, BEAGLE with unordered context and word order was more correlated to the WordNet synonym database than LSA (-.311 compared to -.165).
+Further, it was shown that incorporating word-order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
+This model provided an efficient algorithm for using a single representation for unordered and ordered information that showed performance similar to if not better than LSA.
+
+\subsubsection{Random Permutation Model}
+
+In order to represent word order information, a function must be used that converts an environment vector for a set of words and positions (e.g., ``stack $\Phi$'') into a new uncorrelated environment vector.
+The BEAGLE model uses circular convolution and circular correlation as the encoding and decoding operations.
+However, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors and positions.
+\textcite{Sahlgren2008} used a much simpler random permutation method for this operation.
+
+A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
+
+\begin{table}[!ht]
+  \caption{Random permutation model}
+  \label{tabRandPermModel}
+  {\tabulinesep=1.2mm
+    \begin{tabu}{ll}
+      \hline
+      Common Name &  Equation \\
+      \hline
+      Activation &		$A_{i} = r(m_{C},m_{i})$ \\
+      Memory Vector &		$m_{i} = \sum_{i \in all past} c_{i} + \sum_{i \in all past} \sum_{l \in locations} o_{i,l}$ \\
+      Unordered Context &	$c_{i} = e_{i}$ \\
+      Ordered Context &		$o_{i,l} = e_{i^{-l}}$ \\
+      Context Memory Vector &	$m_{C} = \sum_{i \in C} c_{i} + \sum_{i \in C} \sum_{l \in locations} o_{i,l}$ \\
+      Environment Vector & 	$e_{i} = rand$ \\
+      \hline
+    \end{tabu}
+  }
+\end{table}
+
+With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few one and negative one values in random locations. 
+In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is shifted to the left one position.
+This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
+
+For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow''.
+The ``stack'' environment vector ($e_{stack}$) is shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).
+That environment vector is then added to the memory vector for ``overflow'' ($m_{overflow} = m_{overflow'} + e_{stack^{-1}}$), so that ``overflow'' remembers that ``stack'' has preceded it in a sentence.
+
+A resonance retrieval for a word following ``stack'' (``stack $\Phi$'') is done by creating a memory vector out of all information in context.
+In this case, assume only a single piece of order information is available, that ``stack'' precedes the word to retrieve.
+So the memory vector for context is the environment vector for one-back ``stack'' ($m_{C} = e_{stack^{-1}}$).
+The memory vector for ``overflow'' ($m_{overflow}$) will be correlated the highest with this context vector ($m_{C}$) since ``overflow'' contains this vector.
+
+\subsubsection{Comparison between BEAGLE and Random Permutations}
+
+Random permutations uses one of simplest permutations possible to create a new environment vector.
+This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
+Further, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.
+
+\textcite{Recchia2010} found that random permutations could use smaller vectors than circular convolution approaches (e.g., BEAGLE) to produce equivalent performance (better compression),
+and also scale to larger datasets (better computational efficiency).
+The vector length by performance tradeoff was tested on a simple paired associative retrieval task. 
+Retrieval accuracy for the random-permutation vector-based models remained high relative to the circular convolution model even after reducing each environment vector's length from \num{2048} to \num{1024}.
+This indicates that vector-based representations using random permutations compress better than those using circular convolution.
+
+To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFL) assessment (among other tasks).
+The random-permutation model was trained on \num{2.33} GB corpus.
+The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
+The random permutation model performed as well or slightly better than the BEAGLE model when trained on the smaller corpus.
+The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
+These results make a strong case to favor random-permutations over circular convolution when implementing vector-based memory representations.
+
+\subsubsection{Connection to LSA}
+
+Both random permutations and LSA calculate word similarity on a reduced matrix.
+For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors are much less than the number of words.
+For the StackOverflow and Twitter datasets, the number of documents are the number of different tags.
+
+For random permutations, each tag is represented by a compressed vector with a length much less than the number of words.
+Note that if the length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular row in the column vector.
+If done in this way, building a matrix of all words' environment vectors by tags would recover the original word by document matrix \parencite{Kanerva2000}.
+However, each word's environment vector is compressed using a length much less than the number of unique words and represented by using a combination of a few one and negative one values at a unique combination of positions.
+
+Information is lost with this compression in both cases: by starting with a compressed representation with random permutations or by reducing to a compressed representation with LSA.
+However, the way that the compression works in these two cases is not the same.
+LSA uses SVD to find words that have highly correlated tag occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
+A vector-based approach like BEAGLE or random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
+Nonetheless, random permutations have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFL \parencites{Sahlgren2008,Jones2007}.
+
+This is a counterintuitive result.
+It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
+This does not seem to be the case in testing however, as vector-based approaches like BEAGLE and random permutations perform just as well as the LSA when trained and tested on similar datasets.
+Nonetheless, it has been suggested to perform an SVD reduction technique on a compressed random-permutation vector \parencite{Kanerva2000}.
+This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.
+
+\subsection{Comparison of Vector-Based Models and ACT-R}
+
+Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
+Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
+Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
+For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
+\textcite{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect, which is a psychological phenomena that ACT-R's current Bayesian DM system can model as well.
+
+\subsubsection{Base-Level Activation}
+
+Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the benefits of one over the other.
+For example, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
+More recent tweets on a news site like Twitter should be more relevant, for example.
+\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
+They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.
+
+ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
+It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
+Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
+ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
+This type of information on a user's prior behavior is not being utilized with vector-based models currently.
+It may certainly be the case that vector-based models can be modified to incorporate base-level activation, and this idea will be pursued with this research.
+
+\subsubsection{Word Order}
+
+One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
+With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
+It may be that multiple co-occurrence matrices must be used to represent word-order co-occurrences and unordered co-occurrences separately.
+However, it would be more parsimonious if these matrices could be combined into a single matrix representation, much like how this information is combined for vector-based models.
+
+\subsubsection{Activation Calculation}
+
+Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
+Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
+These two computations are certainly not equivalent, even though they are most likely highly correlated with one another.
+It may be interesting to explore how model performance changes when one calculation is swapped out for another.
+For example, a vector-based model could generate a compressed (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
+Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
+It is unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
+But it may be the case that using Bayesian computation on a compressed vector-based model improves performance.
+
+\subsection{Recommendation Systems}
+
+A model that predicts a user-chosen hashtag is a specific type of a recommendation system.
+These general recommendation systems have grown in need with the growth of the web.
+Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
+Recommendation systems are used to tailor the information presented to what is most relevant to the user \parencite{Pazzani2007}.
+
+\subsubsection{The Netflix Prize}
+
+The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
+Netflix is an online-subscription service where users can stream their favorite television episodes and movies.
+Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
+In order to provide a high-quality user experience on the site, Netflix created a content-recommendation service.
+This tool recommends to the user movies and television episodes that most likely match their specific interests.
+
+Netflix wanted to improve the accuracy of their recommendation system, so they held a competition where teams could try to improve on the accuracy of the service.
+The winning team utilized a statistical model that performed a type of factor analysis to identify important dimensions in the data (e.g., drama, action, cult classic) and place values for each user along each dimension.
+This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
+By using a factor-analysis technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.
+
+\subsubsection{Recommending Followers on Twitter}
+
+Recommendation systems based on the content of user-created posts have been successful in recommending people on Twitter for a user to follow.
+\textcite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
+The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
+The profiles were created by generating a word frequency matrix of words by (user + followers + followees), and then providing that matrix as input into the Lucene text search engine framework.
+
+Lucene \parencite{McCandless2010} is a commonly-used open-source software tool that measures the similarity between a query and candidate documents in a database.
+It uses the term frequency--inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
+
+When the system needed to recommend a set of individuals for a user to follow, it would take all of the words used in posts from that user to construct a search query to be provided to Lucene.
+That search query would return the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
+Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user follows at 25\% accuracy.
+
+\subsubsection{Hashtags for Recommendation Systems}
+
+\textcite{Efron2010} showed how hashtags might be used as a type of recommendation system for Twitter.
+Users constructed a query of text that represented a search for a topic that they were interested in.
+The system then retrieved the most likely hashtags that were associated with that search query.
+The model is based on building a word co-occurrence matrix of words by hashtags, and then uses that matrix to assign a similarity score for each hashtag when provided with a search query.
+Participants graded how relevant the hashtags returned were to each search query.
+The results suggest that hashtags contain useful information that help classify the gist of a post. 
+Although this model does not predict the specific hashtags used when composing a tweet, a model like this could certainly be tailored for that purpose.
+
+\subsection{Hashtag Prediction}
+
+More specific hashtag-recommendation models that attempt to predict a user's chosen hashtag have also been created and tested for a few social media sites.
+Google has even deployed a hashtag-recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
+A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
+Models for Twitter and StackOverflow have also been proposed and tested, but none have yet been deployed.
+
+\subsubsection{StackOverflow}
+
+\textcite{Kuo2011} was the first to model tag use for StackOverflow posts.
+His set of models were originally designed for next-word prediction for large text document collections.
+In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user created a post, and restrict the space of possible words to only tags.
+He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
+Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive Singular Value Decomposition and K-nearest-neighbors models.
+This Bayesian model can accurately predict a user-chosen tag at 47\% accuracy.
 
 % ref: http://tex.stackexchange.com/questions/53513/hyperref-token-not-allowed
-\subsubsection{\texorpdfstring{\textcite{Stanley2013}}{Stanley2013}'s ACT-R DM Model for StackOverflow}
-
-\textcite{Stanley2013} modified the standard ACT-R model in Table \ref{tabACT-RModel} in order to more accurately predict tag use on StackOverflow.
+\paragraph{\texorpdfstring{\textcite{Stanley2013}}{Stanley2013}'s ACT-R DM Model for StackOverflow}
+
+\textcite{Stanley2013} also modeled a user's chosen tags for a post on the StackOverflow site.
+They modified the standard ACT-R model in Table \ref{tabACT-RModel} in order to more accurately predict tag use on StackOverflow.
 A formal description of the modified ACT-R Bayesian model is included in Table \ref{tabModACTRModel}.
@@
 
-\subsection{Latent Semantic Analysis Memory Theory}
-
-Latent Semantic Analysis \parencite{Landauer1997} (LSA) is a technique that has mathematical roots in factor analysis and also measures strength of association between two terms.
-The process starts with a set of N documents.
-A word by document co-occurrence matrix is built by counting the number of times each word appears in each document.
-The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
-This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
-The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
-
-Usually information is lost when the data are restructured into a smaller set of latent dimensions.
-However it was shown that this process actually improves prediction and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
-\textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
-Model results were on par with a large sample of applicants to american universities from non-English speaking countries (64.4\% compared to 64.5\%).
-Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
-Afterwards further reductions in the dimensionality of the data began to degrade performance.
-
-\subsubsection{Singular Value Decomposition}
-
-LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix.
-The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
-The best way to remove a dimension while maintaining the most variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
-For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
-By pooling co-occurrence counts the model is able to generate better predictions with less data.
-
-\subsubsection{Word Order}
-
-Although the LSA model has been shown to perform well as a way to measure similarity between two terms, one of the main criticisms is that it is a ``bag of words'' model.
-This means that the model does not account for or pay attention to word order when building the base word by document matrix or consequently, when performing the SVD dimension reduction.
-Other models that can account for word order have been shown to perform better than ``bag of words'' models on tasks that do not even appear to require grammatical structure such as word order to do well.
-For example, \textcite{Jones2007} compared a vector-based model called BEAGLE directly to LSA, training and testing both models on the same corpus used in \textcite{Landauer1997}.
-The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the task (55.6\% vs. 55.3\%).
-However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to 57.5\%.
-It seems reasonable to expect that task performance will improve much more on tasks where word order is used (e.g., generating a hashtag in the middle of a sentence when composing a tweet).
-
-\subsubsection{Scaling Issues}
-
-Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
-As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
-If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
-However, with large-scale human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
-\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
-Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
-This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.
-
-When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
-This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
-However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
-\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
-This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
-Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to 73.75\%, which is 10\% higher than the TASA-trained LSA.
-This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.
-
-\subsubsection{Cost of Incremental Updating}
-
-Another issue with LSA is that adding additional observations to an already-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
-Since computing the SVD is the most computationally-expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
-For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.
-
-\subsection{Vector-Based Memory Systems}
-
-``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., chunk) as a vector.
-The representation of the concept is distributed across all elements of the vector.
-This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and viewing the distribution of co-occurrence counts across the rows in the column as the representation of that concept.
-However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.
-
-For vector-based systems, each word is represented by an environment vector $e_{j}$ that is nearly orthogonal to all other words' environment vectors.
-The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix, which is how vector-based systems represent information in a much more compact space.
-Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occurred with that $e_{i}$ environment vector.
-These memory vectors can contain environment vectors from bag-of-words co-occurrence information ($c_{i}$) and word order information ($o_{i}$), which can be combined into a single representation ($m_{i}$).
-Over time, memory vectors accumulate a distributed representation of the most common environment vectors that co-occurred with them.
-
-\subsubsection{Retrieval Process}
-
-Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
-Decoding takes the memory vector for context and decodes it back into an environment vector.
-The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
-
-Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
-For example, assume that a user has written the word ``zend'' in a post, and she wants to assign a tag for the post.
-``zend'' has co-occurred with the tag \emph{PHP} many times previously, so the unordered memory vector for ``PHP'' ($c_{\mathit{PHP}}$) contains the unordered environment vector for \emph{zend} ($e_{zend}$).
-To determine the most correlated memory vector with context, a memory vector is created from the current context ($e_{zend}$).
-The cosine between that context memory vector and the memory vector for \emph{PHP} will be high, since the \emph{PHP} memory vector ($c_{\mathit{PHP}}$) often contains the context environment vector ($e_{zend}$).
-Since the cosine is high, the tag \emph{PHP} will most likely be returned as the vector that resonates highest with the current context.
-
-The resonance and decoding processes are inversions of each other.
-Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
-\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
-However, decoding does require an additional decoding step that resonance does not require, and depending on the decoding operation, this step may be computationally expensive.
-
-\subsubsection{Addressing Word Order}
-
-One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
-Representing word order is a case where both the encoding and decoding operations are applied.
-As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
-Simply counting co-occurrences of the two immediately-preceding words and hashtags does not include the order information of the words in the representation.
-With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
-Order information is lost if a simply co-occurrence counting technique is used.
-
-With vector-based systems, order information is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
-This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
-Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
-This allows for retrieval requests such as ``playing \#$\Phi$'', ``just read \#$\Phi$'', and ``vector-based $\Phi$ systems''.
-
-For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
-Within this corpus, the phrase ``martin luther'' often occurred.
-When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated by convolving a 1-back location vector with luther ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
-Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin with and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
-The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.
-
-Using decoding, the word-order memory vector for martin is deconvolved to the right ($o_{martin} \circledast \Phi$) which returns the environment vector for luther ($e_{luther}$).
-Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which correlates highest with the word-order memory vector for luther ($o_{luther}$).
-Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$''.
-
-New vectors generated by circular convolution are uncorrelated with all other environment vectors and have the same length.
-This allows the word order information and unordered context information for a chunk to be aggregated and represented in the same memory vector.
-Quite simply, the different types of memory vectors can be summed to create a single memory vector representation for each chunk (i.e., $m_{j} = c_{j} + o_{j}$).
-
-\subsubsection{Addressing Scalability}
-
-Vector-based representations compute chunk likelihoods by performing a correlation operation directly on the memory vectors.
-Each memory vector can be thought of as a column on a context word by tag co-occurrence matrix used for a Bayesian representation.
-Further, the number of rows for a vector-based representation are fixed and set much smaller than the number of rows in the full word by tag co-occurrence matrix.
-Since Bayesian representations have been shown to scale to over hundreds of millions of word by tag co-occurrences \parencite{Stanley2013}, the compressed vector-based representations should scale as well or better.
-
-\subsubsection{BEAGLE}
-
-\textcite{Jones2007} created the BEAGLE vector-based memory representation.
-Each memory vector ($m_{j}$) is a summed representation of unordered and ordered word co-occurrences ($c_{j} + o_{j}$).
-This model uses circular convolution and circular correlation as a way to encode and decode word order.
-BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (55.6\% compared to 55.3\%).
-Also, BEAGLE with unordered context and word order was more correlated to the WordNet synonym database than LSA (-.311 compared to -.165).
-Further, it was shown that incorporating word-order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
-This model provided an efficient algorithm for using a single representation for unordered and ordered information that showed performance similar to if not better than LSA.
-
-\subsubsection{Random Permutation Model}
-
-In order to represent word order information, a function must be used that converts an environment vector for a set of words and positions (e.g., ``stack $\Phi$'') into a new uncorrelated environment vector.
-The BEAGLE model uses circular convolution and circular correlation as the encoding and decoding operations.
-However, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors and positions.
-\textcite{Sahlgren2008} used a much simpler random permutation method for this operation.
-
-A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
-
-\begin{table}[!ht]
-  \caption{Random permutation model}
-  \label{tabRandPermModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Activation &		$A_{i} = r(m_{C},m_{i})$ \\
-      Memory Vector &		$m_{i} = \sum_{i \in all past} c_{i} + \sum_{i \in all past} \sum_{l \in locations} o_{i,l}$ \\
-      Unordered Context &	$c_{i} = e_{i}$ \\
-      Ordered Context &		$o_{i,l} = e_{i^{-l}}$ \\
-      Context Memory Vector &	$m_{C} = \sum_{i \in C} c_{i} + \sum_{i \in C} \sum_{l \in locations} o_{i,l}$ \\
-      Environment Vector & 	$e_{i} = rand$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few one and negative one values in random locations. 
-In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is shifted to the left one position.
-This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
-
-For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow''.
-The ``stack'' environment vector ($e_{stack}$) is shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).
-That environment vector is then added to the memory vector for ``overflow'' ($m_{overflow} = m_{overflow'} + e_{stack^{-1}}$), so that ``overflow'' remembers that ``stack'' has preceded it in a sentence.
-
-A resonance retrieval for a word following ``stack'' (``stack $\Phi$'') is done by creating a memory vector out of all information in context.
-In this case, assume only a single piece of order information is available, that ``stack'' precedes the word to retrieve.
-So the memory vector for context is the environment vector for one-back ``stack'' ($m_{C} = e_{stack^{-1}}$).
-The memory vector for ``overflow'' ($m_{overflow}$) will be correlated the highest with this context vector ($m_{C}$) since ``overflow'' contains this vector.
-
-\subsubsection{Comparison between BEAGLE and Random Permutations}
-
-Random permutations uses one of simplest permutations possible to create a new environment vector.
-This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
-Further, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.
-
-\textcite{Recchia2010} found that random permutations could use smaller vectors than circular convolution approaches (e.g., BEAGLE) to produce equivalent performance (better compression),
-and also scale to larger datasets (better computational efficiency).
-The vector length by performance tradeoff was tested on a simple paired associative retrieval task. 
-Retrieval accuracy for the random-permutation vector-based models remained high relative to the circular convolution model even after reducing each environment vector's length from \num{2048} to \num{1024}.
-This indicates that vector-based representations using random permutations compress better than those using circular convolution.
-
-To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFL) assessment (among other tasks).
-The random-permutation model was trained on \num{2.33} GB corpus.
-The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
-The random permutation model performed as well or slightly better than the BEAGLE model when trained on the smaller corpus.
-The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
-These results make a strong case to favor random-permutations over circular convolution when implementing vector-based memory representations.
-
-\subsubsection{Connection to LSA}
-
-Both random permutations and LSA calculate word similarity on a reduced matrix.
-For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors are much less than the number of words.
-For the StackOverflow and Twitter datasets, the number of documents are the number of different tags.
-
-For random permutations, each tag is represented by a compressed vector with a length much less than the number of words.
-Note that if the length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular row in the column vector.
-If done in this way, building a matrix of all words' environment vectors by tags would recover the original word by document matrix \parencite{Kanerva2000}.
-However, each word's environment vector is compressed using a length much less than the number of unique words and represented by using a combination of a few one and negative one values at a unique combination of positions.
-
-Information is lost with this compression in both cases: by starting with a compressed representation with random permutations or by reducing to a compressed representation with LSA.
-However, the way that the compression works in these two cases is not the same.
-LSA uses SVD to find words that have highly correlated tag occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
-A vector-based approach like BEAGLE or random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
-Nonetheless, random permutations have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFL \parencites{Sahlgren2008,Jones2007}.
-
-This is a counterintuitive result.
-It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
-This does not seem to be the case in testing however, as vector-based approaches like BEAGLE and random permutations perform just as well as the LSA when trained and tested on similar datasets.
-Nonetheless, it has been suggested to perform an SVD reduction technique on a compressed random-permutation vector \parencite{Kanerva2000}.
-This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.
-
-\subsection{Comparison of Vector-Based Models and ACT-R}
-
-Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
-Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
-Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
-For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
-\textcite{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect, which is a psychological phenomena that ACT-R's current Bayesian DM system can model as well.
-
-\subsubsection{Base-Level Activation}
-
-Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the benefits of one over the other.
-For example, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
-More recent tweets on a news site like Twitter should be more relevant, for example.
-\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
-They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.
-
-ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
-It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
-Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
-ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
-This type of information on a user's prior behavior is not being utilized with vector-based models currently.
-It may certainly be the case that vector-based models can be modified to incorporate base-level activation, and this idea will be pursued with this research.
-
-\subsubsection{Word Order}
-
-One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
-With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
-It may be that multiple co-occurrence matrices must be used to represent word-order co-occurrences and unordered co-occurrences separately.
-However, it would be more parsimonious if these matrices could be combined into a single matrix representation, much like how this information is combined for vector-based models.
-
-\subsubsection{Activation Calculation}
-
-Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
-Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
-These two computations are certainly not equivalent, even though they are most likely highly correlated with one another.
-It may be interesting to explore how model performance changes when one calculation is swapped out for another.
-For example, a vector-based model could generate a compressed (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
-Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
-It is unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
-But it may be the case that using Bayesian computation on a compressed vector-based model improves performance.
+\paragraph{Comparison Between \citeauthor{Kuo2011}'s and \citeauthor{Stanley2013}'s Models}
+
+\textcite{Stanley2013}'s model was based on ACT-R's declarative memory retrieval theory, which at the core is also a Bayesian co-occurrence model.
+However several modifications to the model were made from the standard Bayesian ACT-R declarative memory theory, so it was more sophisticated than the standard ACT-R model and the model used by \textcite{Kuo2011}.
+\citeauthor{Stanley2013} used a data-driven entropy weighting measure to attenuate low-predictor words and increase the weighting of high-predictor words in the body and title of a post.
+An offset term was also included which leveled the model's overall prediction for each post and enabled the use of a logistic regression statistical technique to calibrate the model's parameters.
+A separate model component was used for the words in the body and title of the post, and it was unclear if \citeauthor{Kuo2011} used the words in the body of the post for his models.
+
+\citeauthor{Stanley2013}'s model also scaled much better than \citeauthor{Kuo2011}'s co-occurrence model. 
+Because of this, 100 times more posts were used to build the co-occurrence matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
+After including the modifications and scaling up the model, the model can successfully predict a user's tag for a post at 65\% accuracy.
+
+It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
+This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular environment
+(i.e., when predicting the user-chosen tags for human-created content on social media sites).
+
+\subsubsection{StackExchange}
+
+A tag suggestion tool has been deployed for the \url{http://tex.stackexchange.org} \parencite{LatexTags2013} and \url{http://superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
+LaTeX, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
+For the two sites where a tag suggestion tool has been deployed, once a user has created a question, suggested tags appear just below the text field where the user types the tags to associate with the question.
+The post author is not forced to use the suggested tags, and is free to choose tags from them or tags not represented in the suggested set.
+An example of the interface for a created post on the SuperUser site is included in Figure \ref{figSuperUserSuggestion}.
+
+\begin{figure}[!htbp]
+  \scalebox{.9}{\resizebox{\linewidth}{!}{\includegraphics{superUserTagSuggestion-crop.pdf}}}
+  \caption{Example of tag suggestions when creating a post on the SuperUser site}
+  \label{figSuperUserSuggestion}
+\end{figure}
+
+The exact tag-recommendation algorithm used for the LaTeX and SuperUser sites is not publicized.
+StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags, as the algorithm has not yet been deployed for the busiest StackOverflow site.
+The Latex and SuperUser sites are not nearly as popular as StackOverflow currently.
+The sites have roughly \num{50000} and \num{189000} questions respectively compared to \num{5000000} questions on StackOverflow.
+
+So it stands to reason that the algorithm is relatively complex, and there are possible scaling and efficiency concerns with the algorithm's implementation.
+Or it is also possible that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
+Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial to the user and has started to add them to their question and answer sites.
+
+\subsubsection{Google+}
+
+The hashtag recommendation system deployed for the Google+ microblogging service is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
+The user interface for this recommendation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.
+
+\begin{figure}[!htbp]
+  \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample-crop.pdf}}}
+  \caption{Example post on Google+}
+  \label{figGoogle+Hashtags}
+\end{figure}
+
+The workflow and user interface provided for recommending hashtags for Google+ is similar to how a hashtag recommendation system for Twitter might look.
+When generating content on Google+, the system automatically associates the most likely relevant hashtags with that post.
+For the example in Figure \ref{figGoogle+Hashtags}, the system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-created post.
+The user is free to remove any or all of the tags that the system has associated with the post, and can also use tags that were not included in the recommended set.
+
+Google+'s tag-recommendation algorithm is not publicized, but it appears to be rather sophisticated.
+Looking at the example post in Figure \ref{figGoogle+Hashtags}, \emph{\#BusterPosey} is suggested without having that specific word in the post.
+Also \textcite{GoogleKeynote2013} mentioned that the algorithm uses Google's sophisticated image recognition software to suggest hashtags for images that are included in the post (e.g., \emph{\#EiffelTower}),
+even when the post does not have any keywords to provide additional context.
+
+Google is apparently highly confident in the accuracy of their system.
+On the Latex site, the user interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
+For Google+, the user interface automatically tags a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
+If tag accuracy is high enough, the Google+ technique is certainly favorable since the user does not have to spend time figuring out which tags best represent this post.
+However, if the model makes even just a few errors, users may become annoyed and frustrated with the recommendation system since they will have to explicitly correct it every time they generate a post.
+
+\subsubsection{Twitter}
+
+Several hashtag recommendation models for Twitter have been developed recently.
+One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a post as a function of prior hashtag use and context \parencite{Mazzia2009}.
+The model presented by \textcite{Mazzia2009} is very similar to an ACT-R declarative retrieval model.
+The main potential difference is that the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
+Instead, the global prior is an overall average frequency of hashtag use across users.
+
+Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used from similar tweets.
+\textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
+\textcites{Zangerle2011, Kywe2012} use the Apache Lucene infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarities.
+Regardless of the method to compute the similarity between tweets, each method collects the hashtags used from a set of the most similar tweets, then ranks the set, and presents the top 5-10 hashtags to the user.
+
+This tweet-centered approach is quite different than the hashtag-centered approach presented in \textcite{Mazzia2009}.
+With a hashtag-centered approach, a direct association is built between the words in a tweet and hashtags that occur in tweets.
+With a tweet-centered approach, that association between words and hashtags is indirect:
+The relationship is built between a tweet and its content, and then similar tweets are assumed to use similar hashtags.
+One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a representation of the content in each tweet has to be maintained to later compute tweet similarity.
+The storage size for a hashtag-centered approach grows with the number of hashtags, but it seems reasonable that this will grow slower than the number of tweets and asymptote after collecting a large sample of tweets.
+So from the perspective of efficient information storage and retrieval, it seems more likely that the hashtag-centered approach used in \textcite{Mazzia2009} will scale as Twitter grows in size and use.
+
+\textcite{Kywe2012} also customized their tag prediction model to the user's past hashtag use, and was one of the first to do so when developing a hashtag-prediction model for Twitter.
+The set of recommended hashtags was the union of hashtags used in similar tweets and hashtags used by similar users.
+This was done by storing a content vector of hashtag use for each user (user-centered approach) alongside the content vectors of hashtag use for each tweet (tweet-centered approach).
+To generate recommended hashtags, the model would return the top-ranked hashtags from the hashtags used in 0-50 similar tweets and 0-4 similar users.
+Prediction accuracy improved when the model combined recommendations based on hashtags in both similar tweets and the top few similar users compared to basing recommendations on similar tweets alone.
+
+A topics-based approach was used by \textcite{Godin2013}, where the model predicted the most likely set of topics that were associated with a tweet, and tested if these topics matched the tweet by using human raters.
+The latent topics for the collection of tweets were generated using a Latent Dirichlet Allocation (LDA) statistical methodology.
+This is similar to a factor analysis approach in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance.
+Human graders rated how well the models generated topics for a tweet related to the content in the tweet.
+The LDA model performed well, and generated relevant topics.
+When the model produced five most-likely hashtags for each tweet, 80\% of the time at least one of those hashtags was suitable (as determined by human raters).
+
+However, a topic-generating model is not performing the same task as a hashtag-prediction model.
+The authors argued that these general topics could be used to categorize the post, in addition to the hashtags already used by the author in the post.
+It is unclear though if a user would want to label their tweet with these general topics.
+Perhaps they would rather use more specific hashtags to be more precise in how they label the tweet and how that tweet is associated with tweets within the rest of the community.
+In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
+Now it is certainly the case that the topics generated by a latent-based model may be similar to the hashtags used in the post.
+However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
+Why not take those topics already present in the data and build a model that produces them?
+In other words, the topics data are already in the tweet, so why not use them?
+
+\subsubsection{State of Research on StackOverflow and Twitter}
+
+Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
+I am curious to see the prediction difference between a vector-based model and current co-occurrence models on the StackOverflow dataset.
+Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
+Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, it will be interesting to see how a cognitively-plausible vector-based model performs at the task.
+
+Also, a vector-based retrieval model has not yet been tested on Twitter.
+One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
+It seems quite likely that words used just before a hashtag in a tweet (combined with their position) will be highly predictive of the hashtag chosen.
+To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
+It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.
+
+There has been a mix of co-occurrence and latent-based models created to predict hashtag use on Twitter, and the results are encouraging.
+However, there has been very little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
+\textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
+The ACT-R declarative memory retrieval theory shows how past user behavior and current context combine to produce the most likely retrieval.
+Each component is summed together to generate a total activation, and then the likelihood of chunks are ranked by that summed activation.
+This technique is compensatory and allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
+This type of weighted additive model that combines past user behavior and current tweet context to predict the most likely hashtags has not yet been explored for Twitter.
+
+Both StackOverflow models \parencites{Stanley2013,Kuo2011} did not take into account the specific user's prior history when generating tag predictions.
+Rather their models, much like \textcite{Mazzia2009}'s Twitter model, uses the aggregate tag history across all users as the model's prior activation for each tag.
+One of the main purposes of this research is to improve on these user-agnostic Bayesian models and include each specific user's past tagging history as a model component.
+Using customized priors that are tailored to a user's specific tagging history should be crucial for sites like Twitter where the possible tag space is infinite,
+but should also improve performance for more constrained sites like StackOverflow since each user is not interested in all of the possible tags on the site.
 

### 2013-11-07 writeups/proposal/proposal.tex
foo
Former-commit-id: 7c62ef3e1e6f9d7e2e6cc1a7c11f735dd5085b66 [formerly 8533fbcd72ce94cd80eff8d4d80e250643917fa1]
Former-commit-id: ab5ea7fca9e78c793fe5efc54fa600cc906ccc57

@@
 One straightforward way to include a user's prior tag history into a random permutation model is to include ACT-R's base-level activation term ($B_{i}$) as an additional term in the model.
-However this may not be the most mathematically appropriate thing to do, since the base-level activation is represented as a log odds value,
-and the usual single strength of association term in a random permutation model is a correlation metric.
+However this may not be the most mathematically appropriate thing to do, since base-level activation is represented as a log odds value,
+and the single strength of association term in a random permutation model is computed by a correlation.
 It may be more appropriate to somehow attenuate the memory vectors over time,
@@
 When a retrieval request for a tag is made, the strength of association between each word in context (and its position) and the tags in the co-occurrence array would be computed.
-That computation could be made tractable by storing only a small set of values on the order axis and binning.
-
-For example, one could store only three: near, medium, far, which is equivalent to storing three separate co-occurrence matrices.
+That computation could be made tractable by storing only a small set of values on the order axis and binning over that dimension.
+
+For example, one could store only three word order values: near, medium, far, which is equivalent to storing three separate co-occurrence matrices.
 So when a retrieval request is made (e.g., when a user writes the \# character when composing a tweet), the strength of association for each word would be computed using the appropriate matrix.
@@
 
-The tag-recommendation models developed through this research should be immediately relevant for real-world tasks.
-If model accuracy is high enough, then the developed models may be directly deployed as tag recommendation systems for either Twitter or StackOverflow.
+The tag-recommendation models developed through this research should be immediately relevant to real-world tasks.
+If model accuracy is high enough, then the developed models may be used as a foundation for tag recommendation systems that are deployed on Twitter or StackOverflow.
 It is also likely that the developed models will scale if deployed to these two sites, since during development they will be tested on a large portion of each site's currently available data.
@@
 This research also has numerous components that are relevant to improving our understanding of declarative memory.
-It should be emphasized that it has only recently become possible to evaluate psychological theories of declarative memory (e.g., ACT-R Bayesian, random-permutation vector based)
-on large-scale real-world tasks where the theory can be tested on hundreds of millions of data points.
-These information-rich large-scale environments provide a unique opportunity to really stress test and explore the impacts of the psychological constraints of each theory on a much larger scale than was previously possible.
+It should be emphasized that it has only recently become possible to evaluate psychological theories of declarative memory on large-scale real-world tasks where the theory can be tested on hundreds of millions of data points.
+These information-rich large-scale environments provide a unique opportunity to really stress test and explore the impacts of the psychological constraints of each theory
+on a much larger scale than has previously been possible.
 This environment will allow, for example, overall performance differences between an ACT-R Bayesian retrieval model and a random-permutation vector-based model to be directly compared on two real-world tagging tasks,
@@
 Further, the proposed model modifications are aimed at the architectural level, may very well be task independent, and may generalize and improve model fit across other tasks.
-Discovering an effective way to represent the time decay of information in a vector-based model could improve model fit on other tasks where the memory model has been tested. 
+Discovering an effective way to represent the time decay of information in a vector-based model could improve model fit on other tasks where the memory model has been used. 
 Finding a task-agnostic way to represent word order information in a Bayesian ACT-R model could improve model accuracy when applied to other domains where it seems plausible that participants are utilizing word order.

### 2013-11-07 writeups/proposal/proposal.tex
foo
Former-commit-id: 436ee89efb7c5e4719da38f8f64d99f9dd982524 [formerly 0976639a46d10fc7273de09316fe5c8aac3bbeb7]
Former-commit-id: 35277e3edcf4bc36f851eaf51ad1bced624fb82a

@@
 Lucene \parencite{McCandless2010} is a commonly-used open-source software tool that measures the similarity between a query and candidate documents in a database.
-It uses the term frequency--inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
+It uses the term frequency---inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
 

### 2013-11-07 writeups/proposal/proposal.tex
foo
Former-commit-id: 5918b46e49a1c152c0b77d4cb8325a1cdbaf4557 [formerly db83c52687909c3ec9764ec7d0b0b68a4774ac2f]
Former-commit-id: 02f679de7c067fc27e5a0db99ce52902406b8d76

@@
   and [3] formally evaluate and compare each model's performance when recommending hashtags on Twitter and tags on StackOverflow.
-  The user-customized tag-recommendation models developed with this research should be immediately useful for tag-recommendation systems on the popular Twitter and StackOverflow social media sites.
+  The user-customized tag-recommendation models developed with this research should be immediately useful as foundations for tag-recommendation systems on the popular Twitter and StackOverflow social media sites.
   Further, the planned performance modifications to the two memory models are aimed at the architectural level of each model.
@@
 Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
-These two computations are certainly not equivalent, even though they are most likely highly correlated with one another.
+These two computations are certainly not equivalent, even though they are most likely correlated with one another.
 It may be interesting to explore how model performance changes when one calculation is swapped out for another.
@@
 The end result of this research will be [1] a better understanding of how a user's past tagging history influences future tag use,
-[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on a large-scale real-world retrieval task,
+[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on large-scale real-world retrieval tasks,
 [3] improvements to each memory model where the strengths of each model (word order and user prior likelihood) are incorporated into the other,

### 2013-11-13 writeups/proposal/proposal.tex
foo
Former-commit-id: f73f200dab3af212181c1c8e534273f6399f6c05 [formerly 465c0ae6729318d32df39f4ce24a53041f4762fd]
Former-commit-id: ebd5c634fdc98e57727ae3ef8bb2a2dcae74d260

@@
 The offset term was added to normalize the mean activation across retrievals.
-This allowed the use of a logistical regression statistical technique to calibrate the weights for each of the terms.
+This allowed the use of a logistic regression statistical technique to calibrate the weights for each of the terms.
 The title and body contextual components were separated since it seemed plausible that the words in the title would be better predictors of the tag used than the words in the body.
@@
 A tag suggestion tool has been deployed for the \url{http://tex.stackexchange.org} \parencite{LatexTags2013} and \url{http://superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
-LaTeX, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
+Latex, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
 For the two sites where a tag suggestion tool has been deployed, once a user has created a question, suggested tags appear just below the text field where the user types the tags to associate with the question.
@@
 
-The exact tag-recommendation algorithm used for the LaTeX and SuperUser sites is not publicized.
+The exact tag-recommendation algorithm used for the Latex and SuperUser sites is not publicized.
 StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags, as the algorithm has not yet been deployed for the busiest StackOverflow site.

### 2013-11-13 writeups/proposal/proposal.tex
foo
Former-commit-id: d7960e5ef0a4e90274b47e281f0b87e006bc7571 [formerly 02dc4cdc4062f890d3bb65e75fb7d248a5d1cb15]
Former-commit-id: 22dc94742a137166bc0d1f34ab241d5ebb7b2129

@@
 
-A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACT-RModel}.
+A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACTRModel}.
 
@@
   \caption{ACT-R declarative memory model}
-  \label{tabACT-RModel}
+  \label{tabACTRModel}
   {\tabulinesep=1.2mm
@@
 Base-level activation reflects the log prior odds of needing an observed chunk again.
-The primary way to calculate these log prior odds is to use the standard base level equation in Table \ref{tabACT-RModel}.
+The primary way to calculate these log prior odds is to use the standard base level equation in Table \ref{tabACTRModel}.
 This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
@@
 \textcite{Stanley2013} also modeled a user's chosen tags for a post on the StackOverflow site.
-They modified the standard ACT-R model in Table \ref{tabACT-RModel} in order to more accurately predict tag use on StackOverflow.
+They modified the standard ACT-R model in Table \ref{tabACTRModel} in order to more accurately predict tag use on StackOverflow.
 A formal description of the modified ACT-R Bayesian model is included in Table \ref{tabModACTRModel}.
@@
 That tag's base level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
-This model has been modified from the standard ACT-R model (Table \ref{tabACT-RModel}) in three ways:
+This model has been modified from the standard ACT-R model (Table \ref{tabACTRModel}) in three ways:
 Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy measure was used to compute the attentional weight ($W_{j}$).
@@
 Rather their models, much like \textcite{Mazzia2009}'s Twitter model, uses the aggregate tag history across all users as the model's prior activation for each tag.
+
+As some preliminary research, I have developed and tested a user-customized model for StackOverflow by modifying the model in \textcite{Stanley2013}.
+To incorporate the user's prior tag history into the model, the base-level activation term was modified from the original equation in ACT-R.
+$B_{i}$ is based on log odds occurrence, which in turn is based on frequency of tag use, or probability of tag use $p_{i}$.
+The change to incorporate user's priors was to compute a weighted average for $p_{i}$,
+where the probability of using tag $i$ would range from the global probability ($pGlobal_{i}$) to the user's specific probability ($pUser_{i}$) as the user's total tag count ($count$) increases.
+Overall model accuracy improved by 5\% after incorporating the user's prior tag history into the model.
+
+This preliminary research is encouraging, and suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
+It also seems plausible that model accuracy can improve an even greater amount by
+[1] figuring out the optimal way to combine global tag history with a user's past tagging history, and
+[2] using the time-dependent form of the base level component in Table \ref{tabACTRModel}, in order to account for the fact that a user's tagging preferences change over time.
+So these proposed modifications will be explored further when incorporating a user's past tagging history into the models for this research.
+
 One of the main purposes of this research is to improve on these user-agnostic Bayesian models and include each specific user's past tagging history as a model component.

### 2013-11-13 writeups/proposal/proposal.tex
foo
Former-commit-id: e24613b98debada3d6ec4daf9468063c1a8d2e5f [formerly 54972aa7434a4c503467d45afebb0209b60da825]
Former-commit-id: 12468cc3313ac3db8cff29479dcaae705b727571

@@
 $B_{i}$ is based on log odds occurrence, which in turn is based on frequency of tag use, or probability of tag use $p_{i}$.
-The change to incorporate user's priors was to compute a weighted average for $p_{i}$,
-where the probability of using tag $i$ would range from the global probability ($pGlobal_{i}$) to the user's specific probability ($pUser_{i}$) as the user's total tag count ($count$) increases.
+To incorporate a user's prior, a weighted average for $p_{i}$ was used,
+where the probability of using tag $i$ ranges from the global probability ($pGlobal_{i}$) to the user's specific probability ($pUser_{i}$) as the user's total tag count increases.
 Overall model accuracy improved by 5\% after incorporating the user's prior tag history into the model.

### 2013-11-13 writeups/proposal/proposal.tex
foo
Former-commit-id: 8b9532f7544884a7bbf7075940f87b014c3a3159 [formerly d977002407e4e8bdc41f3174cb98a7d63082b353]
Former-commit-id: fef35ede069e88f298d7b286bbbf3fbaa973f4dd

@@
 
-For vector-based systems, each word is represented by an environment vector $e_{j}$ that is nearly orthogonal to all other words' environment vectors.
+For vector-based systems, each word is represented by an environment vector $e_{i}$ that is nearly orthogonal to all other words' environment vectors.
 The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix, which is how vector-based systems represent information in a much more compact space.
@@
 This allows the word order information and unordered context information for a chunk to be aggregated and represented in the same memory vector.
-Quite simply, the different types of memory vectors can be summed to create a single memory vector representation for each chunk (i.e., $m_{j} = c_{j} + o_{j}$).
+Quite simply, the different types of memory vectors can be summed to create a single memory vector representation for each chunk (i.e., $m_{i} = c_{i} + o_{i}$).
 
@@
 \textcite{Jones2007} created the BEAGLE vector-based memory representation.
-Each memory vector ($m_{j}$) is a summed representation of unordered and ordered word co-occurrences ($c_{j} + o_{j}$).
+Each memory vector ($m_{i}$) is a summed representation of unordered and ordered word co-occurrences ($c_{i} + o_{i}$).
 This model uses circular convolution and circular correlation as a way to encode and decode word order.

### 2013-11-13 writeups/proposal/proposal.tex
foo
Former-commit-id: 79beac62d9806c9ff092335339feec666e23fcfd [formerly d20e6e661561a3fa10520a9754093737d6ca6296]
Former-commit-id: 5c97b1e6eaa2b0cb9dc48f53056575489be90be1

@@
 
-When the system needed to recommend a set of individuals for a user to follow, it would take all of the words used in posts from that user to construct a search query to be provided to Lucene.
-That search query would return the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
+When \citeauthor{Hannon2010}'s system recommended a set of individuals for a user to follow, it collected all of the words used in posts from that user, constructed a search query, and provided that query to Lucene.
+Lucene then returned the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
 Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user follows at 25\% accuracy.

### 2014-01-03 writeups/proposal/proposal.tex
migrated changes
Former-commit-id: ba2375da7972de2be09a8dae34f53ec1d358cf51 [formerly 62375ee355721490f16c28d7f41a554cf03a27d2]
Former-commit-id: 52591fd917144caba99a01e9dfdb2ffd7769d3a3

@@
 
-A growing number of users on social media sites are searching for information that is happening in real time \parencite{Jansen2011}.
-In other words, users on social-media sites want the fresh \emph{stream} of information about a particular topic.
+A growing number of users on social media sites are focusing on and looking for recently created information when performing a search \parencite{Jansen2011}.
+In other words, users on social media sites want the fresh \emph{stream} of information about a particular topic.
 This can certainly be seen in the main user-interface views that these social media sites provide:
-Twitter's homepage showing your followees' tweets, Facebook's homepage showing your friends' posts, StackOverflow's daily digest of posts for particular tags.
+Twitter's homepage showing followees' tweets, Facebook's homepage showing friends' posts, and StackOverflow's daily digest of posts for particular tags.
 
@@
 Prior research has examined the growth and decay cycle of hashtag use across users \parencite{Tsur2012}.
-However, much less is known about hashtag lifetime within a single user.
-It may not necessarily follow that a hashtag's life cycle for a single user matches the life cycle across users.
+However, much less is known about hashtag lifetime for individual users.
+It may not necessarily follow that a hashtag's life cycle for individual users matches the life cycle across users.
 Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into a specific user's prior likelihood of hashtag use.
-So I will characterize the hashtag growth and decay cycle for a single user.
+So I will characterize the hashtag growth and decay cycle for individual users.
 The expectation is that this cycle will match ACT-R's decay rate used in the declarative memory system to compute a particular memory item's prior.
@@
 The main potential difference is that the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
-Instead, the global prior is an overall average frequency of hashtag use across users.
+Instead, the global prior is an overall average frequency of hashtag use across all users.
 
@@
 \textcites{Zangerle2011, Kywe2012} use the Apache Lucene infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarities.
-Regardless of the method to compute the similarity between tweets, each method collects the hashtags used from a set of the most similar tweets, then ranks the set, and presents the top 5-10 hashtags to the user.
+Regardless of the method to compute the similarity between tweets,
+each method collects the hashtags used from a set of the most similar tweets, then ranks the hashtags in the set, and presents the top 5-10 hashtags to the user.
 
@@
 Due to Twitter's query limitation for their API, it is estimated that this process will take about a week.
-This dataset will be used to explore the growth and decay characteristics of a hashtag for a single user.
+This dataset will be used to explore the growth and decay characteristics of a hashtag for individual users.
 The temporal dynamics of ACT-R's decay rate model for a chunk will be evaluated against this data.
@@
 This environment will allow, for example, overall performance differences between an ACT-R Bayesian retrieval model and a random-permutation vector-based model to be directly compared on two real-world tagging tasks,
-which has never been done previously.
+which has not been done previously.
 

### 2014-01-03 twitter/topUsers/scrapeUsers.py
code to grab top 1000 users
Former-commit-id: 48cddd86cb73bfdee4923788fa10d8f6dc3e2cb4 [formerly 78415c648333d3b008b38e72aecdc934331588f8]
Former-commit-id: 3c1d97c6652860499fb51a589c304f8e35c0ae89

@@ -1,0 +1,22 @@
+import urllib2
+from bs4 import BeautifulSoup
+import itertools
+
+
+def scrape_twitaholic(url):
+    soup = BeautifulSoup(urllib2.urlopen(url).read())
+    res = []
+    for tr in soup.findAll('tr', {'style': 'border-top:1px solid black;'}):
+        temp = tr.find('td', {'class': 'statcol_name'})
+        res.append(temp.a['title'].split('(')[1][4:-1])
+    return res
+
+res = []
+for i in range(10):
+    i = i + 1
+    url = 'http://twitaholic.com/top' + str(i) + '00/followers/'
+    res.append(scrape_twitaholic(url))
+    res
+
+for item in itertools.chain(*res):
+    print item

### 2014-01-04 twitter/topUsers/scrapeUsers.py
foo
Former-commit-id: bc1579681916e75036aad3c3c7a958788cb19fef [formerly a5727d05be20acb373c4cf5fe21e9b273c8b85a5]
Former-commit-id: 550bc92985813dae86511b720e0a44d0012b7bc4

@@ -1,6 +1,16 @@
 import urllib2
 from bs4 import BeautifulSoup
 import itertools
+import csv
+import os, errno
+import os.path
+import string
+import pg
+from collections import OrderedDict
+
+
+_dir = os.path.dirname(os.path.abspath(__file__))
+cur = pg.connect(host="127.0.0.1")
 
 
 def scrape_twitaholic(url):
@@ -11,12 +21,29 @@ def scrape_twitaholic(url):
         res.append(temp.a['title'].split('(')[1][4:-1])
     return res
 
-res = []
-for i in range(10):
-    i = i + 1
-    url = 'http://twitaholic.com/top' + str(i) + '00/followers/'
-    res.append(scrape_twitaholic(url))
-    res
+file = 'top1000.csv'
+def generateTopUsersCSV():
+    res = []
+    for i in range(10):
+        i = i + 1
+        url = 'http://twitaholic.com/top' + str(i) + '00/followers/'
+        res.append(scrape_twitaholic(url))
+    res = itertools.chain(*res)
+    res = list(OrderedDict.fromkeys(res))
+    res = filter(None, res)
+    with open(file, 'wb') as csvfile:
+        csvWriter = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
+        rank = 0
+        for item in res:
+            rank = rank + 1
+            csvWriter.writerow([rank, item])
 
-for item in itertools.chain(*res):
-    print item
+
+def storeTopUsers(dir, file):
+    cmd = "copy topUsers (rank, userName) from '${dir}/${file}' delimiters ',' csv"
+    cmd = string.Template(cmd).substitute(locals())
+    cur.query(cmd)
+
+generateTopUsersCSV()
+storeTopUsers(_dir, file)
+

### 2014-01-07 twitter/topUsers/preProcess.R
foo
Former-commit-id: 95c718097f39f0d12508d397c871c9306935fa0b [formerly 4db1b783f6219694e6c6cad823c544a730102e06]
Former-commit-id: 476d548395333820c527d57c5002364c7210e9b1

@@ -9,6 +9,8 @@ options("scipen"=100, "digits"=4)
 
 sqldf("select * from tweets where user_screen_name = 'claytonstanley1'")
 sqldf('select count(*) from tweets')
+sqldf('select count(*) from topUsers')
+sqldf('delete from topUsers')
 sqldf('select t.user_screen_name,rank from tweets as t join topusers as u on t.user_screen_name = u.user_screen_name group by t.user_screen_name,u.rank order by rank')
 data.table(sqldf('select * from topusers order by rank'))
 sqldf("select * from tweets where user_screen_name='jlo'")

### 2014-01-07 twitter/topUsers/scrapeUsers.py

@@ -143,7 +143,7 @@ def userAlreadyCollected (user_screen_name):
     return len(res) > 0
 
 def getAllTweetsForTopUsers ():
-    res = _cur.query('select (user_screen_name) from topUsers order by rank asc').getresult()
+    res = _cur.query('select (user_screen_name) from topUsers order by rank asc limit 10000').getresult()
     screenNames = [[user[0]] for user in res]
     screenNames = itertools.chain(*screenNames)
     for screenName in screenNames:
@@ -160,7 +160,8 @@ def getAllTweetsForTopUsers ():
             time.sleep(60)
             pass
 
-#generateTopUsers(scrapeFun=generateTopUsersSocialBakers, topUsersFile='top10000SocialBakers.csv')
+generateTopUsers(scrapeFun=generateTopUsersSocialBakers, topUsersFile='top10000SocialBakers.csv')
+#generateTopUsers(scrapeFun=lambda : generateTopUsersSocialBakers(numUsers=100000), topUsersFile='top100000SocialBakers.csv')
 #getAllTweets('claytonstanley1')
 #getAllTweetsForTopUsers()
 

### 2014-01-07 twitter/topUsers/preProcess.R
foo
Former-commit-id: d86c184902e76ca89b8dea93dbc99b7ba632cdec [formerly cf699263ff7a43a0aa2695c5d8fae6af5fd4674e]
Former-commit-id: 985395ce5871896827093e4cb0a62d9c5e4a71a4

@@ -11,7 +11,7 @@ sqldf("select * from tweets where user_screen_name = 'claytonstanley1'")
 sqldf('select count(*) from tweets')
 sqldf('select count(*) from topUsers')
 sqldf('delete from topUsers')
-sqldf('select t.user_screen_name,rank from tweets as t join topusers as u on t.user_screen_name = u.user_screen_name group by t.user_screen_name,u.rank order by rank')
+data.table(sqldf('select t.user_screen_name,rank from tweets as t join topusers as u on t.user_screen_name = u.user_screen_name group by t.user_screen_name,u.rank order by rank'))
 data.table(sqldf('select * from topusers order by rank'))
 sqldf("select * from tweets where user_screen_name='jlo'")
 sqldf('select user_screen_name,count(*) from tweets group by user_screen_name')

### 2014-01-07 writeups/proposal/proposal.tex
merge
Former-commit-id: 43f6f39b37cf89e2cbbc0086dcf04447c68815d2 [formerly bacc4126246ca39f15aaeb9ced534ca8ab51f73f]
Former-commit-id: 45fc34c6172708076973a8e31fc9978e6aa22488

@@
 
-Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
-I am curious to see the prediction difference between a vector-based model and current co-occurrence models on the StackOverflow dataset.
+\textcite{Stanley2013} used an ACT-R inspired Bayesian memory model to predict human tagging on the StackOverflow site, but to our knowledge no vector-based holographic memory model has been tested or used.
 Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
 Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, it will be interesting to see how a cognitively-plausible vector-based model performs at the task.
+Prediction differences between the two models may help reveal ways where the constraints of each model can be adjusted in order to improve accuracy.
 

### 2014-01-07 writeups/proposal/proposal.tex
merge
Former-commit-id: 83c281e059e8ae39158f65201aed9e4ab8b2e054 [formerly 0bfc9115dd3c3a151aa40aaa7afcc129920e6c83]
Former-commit-id: b396325413ad09575644838d019f84cc81e87c04

@@
 \section{Motivation}
+
+The ACT-R cognitive architecture \parencite{Anderson2004} contains one of the leading theories of human declarative memory.
+The declarative memory component in ACT-R is a Bayesian statistical model.
+This is the same type of co-occurrence Naive Bayes statistical model that is prominent in the more general machine-learning field.
+When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
+The primary reason for this is that the stability of statistical models increase as sample size increases.
+
+However it is much less common in the cognitive modeling community to use a large number of observations to build up declarative memory when creating ACT-R models.
+If a cognitive modeler is using the declarative memory component of ACT-R,
+it is common to either construct the co-occurrences by hand, or use a small training set of data to initially build up the declarative memory system.
+This approach can and has worked well.
+Essentially the modeler is only including the observations in declarative memory that are relevant to the specific task that they are studying.
+
+This approach is problematic for two main reasons.
+First, it means that the declarative memory system for every developed model must be hand crafted and customized for that model, which takes enormous time and effort.
+Second, it means that modelers are continuing to use small sample sizes for the declarative memory components in their models.
+This has important implications for understanding declarative memory:
+It means that we do yet fully understand how well the ACT-R declarative memory theory scales when the number of chunks begins to approach the number of chunks that are actually stored in human declarative memory.
+
+There is, however, an opportunity to change this approach.
+With the massive growth of human-created content on social media sites (e.g., Twitter, StackOverflow, Wikipedia), and public access to many of those databases,
+it is now possible to evaluate psychological theories of declarative memory on large-scale real-world tasks where the theory can be tested on hundreds of millions to billions of data points.
+This will provide a way to test if these models scale, and evaluate the psychological validity of these models as the number of observations used increases by orders of magnitude.
+
+There have already been a few researchers that have used these large databases to evaluate how well declarative memory models scale,
+for both ACT-R Bayesian models \parencites{Fu2007,Pirolli2003,Stanley2013,Douglass2010} and vector-based memory models \parencites{Jones2007,Rutledge2008,Recchia2010,Sahlgren2008}.
+
+The work for this proposal will expand on the knowledge gained from previous work that has tested retrieval models on large-scale datasets.
+It will focus specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutations) can be improved further, by testing these models on two real-world tagging tasks.
+One primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory when the datasets used for each model contain hundreds of millions to billions of data points.
+Another objective is to test how much improvement is gained by incorporating the strengths of each model into the other (i.e., modifying model components).
+For example, it is currently unexplored how a vector-based model can be modified to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
+Since large datasets will be used, a rich volume of data will be available to test the models.
+This enables constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.
+
+\subsection{Motivation for Task}
 
@@
 I will also replace the vector-based model used by \citeauthor{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
-The primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory.
 
@@
 However, not only should the model properly take into account a user's prior hashtag use, but it should also generate new hashtag predictions when a user is most likely choosing a hashtag they have never used before.
-Calibrating the models to properly balance this exploitation of prior hashtag use versus exploration of new hashtags will be a primary research question.
 The end goal is to have a model that can properly balance the user's tag history with the contextual cues in the post in order to generate an accurate prediction.

### 2014-01-07 twitter/topUsers/scrapeUsers.py
foo
Former-commit-id: 3a8cd81834d80f7b811abde9c0ff33ca299e2dcf [formerly cc0f7a9bcdbd4d3ba6c5b640954c88b8d1af5a06]
Former-commit-id: 5b5bc5c16fdff7dad075b23c965508f64b363bed

@@ -94,6 +94,21 @@ def getTweets(screen_name, **kwargs):
 def isRetweet(tweet):
     return hasattr(tweet, 'retweeted_status')
 
+def write2csv(res, file):
+    with open(file, 'wb') as f:
+        writer = csv.writer(f, quoting=csv.QUOTE_ALL)
+        writer.writerows(res)
+
+def getInfoForUser(screen_name):
+    user = _api.get_user(screen_name)
+    res = [[user.id, user.created_at, user.description.encode('utf-8'), user.followers_count, user.friends_count,
+            user.lang, user.location.encode('utf-8'), user.name.encode('utf-8'), user.screen_name.lower(), user.verified]]
+    file = '/tmp/%s_user.csv' % screen_name
+    print "writing %s results to file: %s" % (len(res), file)
+    write2csv(res, file)
+    print "updating database with results in temp file"
+    _cur.query("copy twitter_users (id,created_at,description,followers_count,friends_count,lang,location,name,user_screen_name,verified) from '%s' delimiters ',' csv" % (file))
+
 def getAllTweets(screen_name):
     def getTweetsBetween(greaterThanID, lessThanID):
         alltweets = []
@@ -127,41 +142,46 @@ def getAllTweets(screen_name):
     alltweets.extend(getTweetsBetween(0, lessThanID))
     outTweets = [[tweet.id_str, tweet.user.id, tweet.user.screen_name.lower(), tweet.created_at, isRetweet(tweet), tweet.in_reply_to_status_id_str,
                   tweet.lang, tweet.truncated, tweet.text.encode("utf-8")] for tweet in alltweets]
-    #print(outTweets)
     file = '/tmp/%s_tweets.csv' % screen_name
-    print "writing %s results to temp file: %s" % (len(outTweets), file)
-    with open(file, 'wb') as f:
-        writer = csv.writer(f, quoting=csv.QUOTE_ALL)
-        writer.writerows(outTweets)
+    print "writing %s results to file: %s" % (len(outTweets), file)
+    write2csv(outTweets, file)
     print "updating database with results in temp file"
-    cmd = "copy tweets (id, user_id, user_screen_name, created_at, retweeted, in_reply_to_status_id, lang, truncated,text) from '${file}' delimiters ',' csv"
-    cmd = string.Template(cmd).substitute(locals())
-    _cur.query(cmd)
[…]

### 2014-01-08 writeups/proposal/proposal.tex
foo
Former-commit-id: 6cadf8f2e340bb41bd7482385cefdd863f8a9d1c [formerly d808a94d26cdc1fdc0711db0a77de4991b33a12f]
Former-commit-id: 4ffbf034b1f7fdb3e260663be11589ea4ce236e1

@@
 
-This approach is problematic for two main reasons.
+This approach is problematic for two main reasons:
 First, it means that the declarative memory system for every developed model must be hand crafted and customized for that model, which takes enormous time and effort.
 Second, it means that modelers are continuing to use small sample sizes for the declarative memory components in their models.
+
 This has important implications for understanding declarative memory:
-It means that we do yet fully understand how well the ACT-R declarative memory theory scales when the number of chunks begins to approach the number of chunks that are actually stored in human declarative memory.
+First, it is difficult to verify and explore modifications to the declarative memory architecture because the dataset is not large enough or rich enough to rigorously test modifications to the theory.
+It also means that we do yet fully understand how well the ACT-R declarative memory theory scales when the number of chunks begins to approach the number of chunks that are actually stored in human declarative memory.
 
@@
 
-There have already been a few researchers that have used these large databases to evaluate how well declarative memory models scale,
+A few researchers have started to use these large databases to evaluate how well declarative memory models scale,
 for both ACT-R Bayesian models \parencites{Fu2007,Pirolli2003,Stanley2013,Douglass2010} and vector-based memory models \parencites{Jones2007,Rutledge2008,Recchia2010,Sahlgren2008}.

### 2014-01-08 twitter/topUsers/scrapeUsers.py

@@ -206,5 +206,5 @@ def getUserInfoForTopUsers ():
 #generateTopUsers(scrapeFun=generateTopUsersSocialBakers, topUsersFile='top10000SocialBakers.csv')
 #generateTopUsers(scrapeFun=lambda : generateTopUsersSocialBakers(numUsers=100000), topUsersFile='top100000SocialBakers.csv')
 #getAllTweets('claytonstanley1')
-#getAllTweetsForTopUsers()
-getUserInfoForTopUsers()
+getAllTweetsForTopUsers()
+#getUserInfoForTopUsers()

### 2014-01-08 twitter/topUsers/preProcess.R
foo
Former-commit-id: ebadcb8544cca92399fa1e96e54317fb1fd80f5d [formerly f5edcda16e345f6911518ca4ad78188893984719]
Former-commit-id: 2ffa302911fc35bfac69d6bf514f56064fe72912

@@ -19,6 +19,7 @@ sqldf('select lang,count(*) as count from tweets group by lang order by count de
 sqldf('select retweeted,count(*) as count from tweets group by retweeted order by count desc')
 sqldf('select truncated,count(*) as count from tweets group by truncated order by count desc')
 sqldf("select * from topUsers")[1:100,]
+data.table(sqldf("select * from twitter_users"))
 
 tweetsTbl
 tweetsTbl = data.table(sqldf("select * from tweets"))

### 2014-01-08 twitter/topUsers/preProcess.R
foo
Former-commit-id: e4f6f04dbade2bc8297a00f3f4556c0026ce454d [formerly 5f5929f0304aaf75755e90de6d88ff4d2b24474e]
Former-commit-id: 9c4721d351b35ac9ac2adde1088f68b2e1ff37e4

@@ -1,4 +1,5 @@
 library(RPostgreSQL)
+library(stringr)
 library(sqldf)
 library(data.table)
 PATH = getPathToThisFile()
@@ -20,6 +21,12 @@ sqldf('select retweeted,count(*) as count from tweets group by retweeted order b
 sqldf('select truncated,count(*) as count from tweets group by truncated order by count desc')
 sqldf("select * from topUsers")[1:100,]
 data.table(sqldf("select * from twitter_users"))
+twitter_users = data.table(read.csv(str_c(PATH, "/data/tables/twitter_users.csv")))
+topUsers = data.table(read.csv(str_c(PATH, "/data/tables/topUsers.csv")))
+twitter_users[created_at=='2010-10-29 19:05:25',]
+?fread
+fread('col1,col2\n5,"4\n3"')
+
 
 tweetsTbl
 tweetsTbl = data.table(sqldf("select * from tweets"))

### 2014-01-08 twitter/topUsers/scrapeUsers.py

@@ -72,6 +72,13 @@ def generateTopUsers (scrapeFun=generateTopUsersTwitaholic, topUsersFile='top100
     generateTopUsersCSV(scrapeFun=scrapeFun, topUsersFile=topUsersFile)
     storeTopUsers(topUsersFile=topUsersFile)
 
+def backupTables ():
+    tableNames = ['topUsers', 'twitter_users']
+    for tableName in tableNames:
+        file = "%s/data/tables/%s.csv" % (_dir, tableName)
+        cmd = string.Template("copy ${tableName} to '${file}' delimiter ',' csv header").substitute(locals())
+        _cur.query(cmd)
+
 def getTweepyAPI ():
     consumer_key="vKbz24SqytZnYO33FNkR7w"
     consumer_secret="jjobro8Chy9aKMzo8szYMz9tHftONLRkjNnrxk0"
@@ -206,5 +213,6 @@ def getUserInfoForTopUsers ():
 #generateTopUsers(scrapeFun=generateTopUsersSocialBakers, topUsersFile='top10000SocialBakers.csv')
 #generateTopUsers(scrapeFun=lambda : generateTopUsersSocialBakers(numUsers=100000), topUsersFile='top100000SocialBakers.csv')
 #getAllTweets('claytonstanley1')
-getAllTweetsForTopUsers()
+#getAllTweetsForTopUsers()
 #getUserInfoForTopUsers()
+backupTables()

### 2014-01-09 writeups/proposal/proposal.tex
merged changes
Former-commit-id: 571f10916e36405721af978482c2dbb41408a869 [formerly fa1d510982dcddb6544c9241203b7620cd799f38]
Former-commit-id: f907412a9fbbebe7aead3d47296ff7120f42b598

@@
 \begin{table}[!ht]
-  \caption{StackOverflow tag prediction model}
+  \caption{\citeauthor{Stanley2013}'s StackOverflow tag prediction model}
   \label{tabModACTRModel}
@@
 The model presented by \textcite{Mazzia2009} is very similar to an ACT-R declarative retrieval model.
-The main potential difference is that the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
+The main potential difference is that --- like \textcite{Kuo2011} and \textcite{Stanley2013} --- the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
 Instead, the global prior is an overall average frequency of hashtag use across all users.
@@
 
-One of the main purposes of this research is to improve on these user-agnostic Bayesian models and include each specific user's past tagging history as a model component.
+One of the main purposes of this research is to improve on these user-agnostic statistical models and include each specific user's past tagging history as a model component.
 Using customized priors that are tailored to a user's specific tagging history should be crucial for sites like Twitter where the possible tag space is infinite,
@@
 
-\subsection{StackOverflow Dataset}
+\subsubsection{StackOverflow Dataset}
 
@@
 
-\subsection{Twitter Dataset}
+\subsubsection{Twitter Dataset}
 
@@
 Due to this and the sheer size of Twitter, subsets of the Twitter dataset must be collected.
-So two Twitter datasets will be collected for model evaluation: a common-hashtags and popular-users dataset.
-The common-hashtags dataset will contain approximately 2 million tweets where at least one of the 500 most popular hashtags are used.
-This dataset will be collected in real time until the tweet quota is reached, which should take just a few days.
+So two Twitter datasets will be collected for model evaluation: a popular-hashtags and popular-users dataset.
+The popular-hashtags dataset will contain approximately 2 million tweets where at least one of the 500 most popular hashtags are used.
+This dataset will be collected in real time until the tweet quota is reached.
 The purpose of this dataset is to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
@@
 This dataset will be collected by fetching all tweet history for each of the top users.
-Due to Twitter's query limitation for their API, it is estimated that this process will take about a week.
-This dataset will be used to explore the growth and decay characteristics of a hashtag for individual users.
+It will be used to explore the growth and decay characteristics of a hashtag for individual users.
 The temporal dynamics of ACT-R's decay rate model for a chunk will be evaluated against this data.
-Afterwards, the most appropriate model of user-centered frequency and recency effects (i.e., each hashtag's prior likelihood given a particular user) will be determined.
 
@@
 
-This updated model will then be evaluated on only the Twitter common-hashtags dataset, since this dataset contains hashtags that occur in positions throughout the tweet (e.g., embedded within or at the end of a tweet).
+This updated model will then be evaluated on only the Twitter popular-hashtags dataset, since this dataset contains a large (yet not infinite) set of hashtags that occur in various positions throughout the tweet.
 This model will not be evaluated on the StackOverflow dataset since chosen tags for a StackOverflow post are not embedded and mixed within a post, so word-order information is unavailable and most likely unused.
@@
 auto labeling incoming mail and placing it into appropriate folders, and spam filtering.
-Having an accurate, efficient co-occurrence memory model that correctly retrieves relevant tags can be applied to a wide variety of real-world tasks.
+Having an accurate and efficient co-occurrence memory model that correctly retrieves relevant tags can be applied to a wide variety of real-world tasks.
 

### 2014-01-09 twitter/topUsers/scrapeUsers.py
foo
Former-commit-id: ca52bd73807bf29b852fa6f90e60e78cd881a6b8 [formerly 6710ef8072336681adb109a25ea7f8578c660bdc]
Former-commit-id: a625e115187097e49d12bf6190e68acfaa96f4fd

@@ -75,7 +75,7 @@ def generateTopUsers (scrapeFun=generateTopUsersTwitaholic, topUsersFile='top100
 def backupTables ():
     tableNames = ['topUsers', 'twitter_users']
     for tableName in tableNames:
-        file = "%s/data/tables/%s.csv" % (_dir, tableName)
+        file = "%s/dissertationData/tables/%s.csv" % (_dir, tableName)
         cmd = string.Template("copy ${tableName} to '${file}' delimiter ',' csv header").substitute(locals())
         _cur.query(cmd)
 

### 2014-01-09 twitter/topUsers/scrapeUsers.py
foo
Former-commit-id: afb7b8c1e147954423b2d5bf7029556123ab4e7d [formerly 938b0fc01cbca458f29b0f32f49aa11f5fb0240d]
Former-commit-id: 50089f3eb4b5704d3e0eab8b19f106a1453fd020

@@ -50,21 +50,24 @@ def generateTopUsersSocialBakers (numUsers=10000):
         res.append(scrape_socialbakers(url))
     return res
 
+_topUsersDir = "%s/dissertationData/topRankedUsers" % (_dir)
+
 def generateTopUsersCSV (scrapeFun, topUsersFile):
     res = scrapeFun()
-    res = itertools.chain(*res)
-    res = list(OrderedDict.fromkeys(res))
+    res = list(itertools.chain(*res))
+    res = [x.lower() for x in res]
+    res = OrderedDict.fromkeys(res)
     res = filter(None, res)
-    with open(topUsersFile, 'wb') as csvfile:
+    with open("%s/%s" % (_topUsersDir, topUsersFile), 'wb') as csvfile:
         csvWriter = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
         rank = 0
         for item in res:
             rank = rank + 1
-            csvWriter.writerow([item.lower(), rank])
+            csvWriter.writerow([item, rank])
 
 def storeTopUsers(topUsersFile):
-    dir = _dir
-    cmd = "copy topUsers (user_screen_name, rank) from '${dir}/${topUsersFile}' delimiters ',' csv"
+    topUsersDir = _topUsersDir
+    cmd = "copy topUsers (user_screen_name, rank) from '${topUsersDir}/${topUsersFile}' delimiters ',' csv"
     cmd = string.Template(cmd).substitute(locals())
     _cur.query(cmd)
 
@@ -211,7 +214,9 @@ def getUserInfoForTopUsers ():
     getForTopUsers(alreadyCollectedFun=userInfoAlreadyCollected, getForUserFun=getInfoForUser, getRemainingHitsFun=getRemainingHitsGetUser, hitsAlwaysGreaterThan=30, groupFun=lambda x: chunker(x, 100))
 
 #generateTopUsers(scrapeFun=generateTopUsersSocialBakers, topUsersFile='top10000SocialBakers.csv')
+#generateTopUsers(scrapeFun=generateTopUsersTwitaholic, topUsersFile='top1000Twitaholic.csv')
 #generateTopUsers(scrapeFun=lambda : generateTopUsersSocialBakers(numUsers=100000), topUsersFile='top100000SocialBakers.csv')
+#storeTopUsers(topUsersFile='top100000SocialBakers.csv')
 #getAllTweets('claytonstanley1')
 #getAllTweetsForTopUsers()
 #getUserInfoForTopUsers()

### 2014-01-09 twitter/topUsers/preProcess.R
foo
Former-commit-id: 5f159ec9a2974002b4ad0a3aa4f8f85a70de9144 [formerly 6eeb4f887e5378aed438ad26cdf6c98ea8596f05]
Former-commit-id: 4a5a60dccd68b8285bd54bb4b7f82cf11f52cd94

@@ -13,7 +13,7 @@ options("scipen"=100, "digits"=4)
 sqldf("select * from tweets where user_screen_name = 'claytonstanley1'")
 sqldf('select count(*) from tweets')
 sqldf('select count(*) from topUsers')
-sqldf('delete from topUsers')
+sqldf('select * from topUsers')
 data.table(sqldf('select t.user_screen_name,rank from tweets as t join topusers as u on t.user_screen_name = u.user_screen_name group by t.user_screen_name,u.rank order by rank'))
 data.table(sqldf('select * from topusers order by rank'))
 sqldf("select * from tweets where user_screen_name='jlo'")

### 2014-01-09 twitter/topUsers/preProcess.R
foo
Former-commit-id: bfe3fd99799797bd5a52a9c52785ef92d22d884c [formerly 9bf2e76c2e20eec9db1859c9e823f5044ad7877e]
Former-commit-id: b2b3949f2d01d21ae8629cdf8710bf2fc7a74b44

@@ -4,6 +4,8 @@ library(sqldf)
 library(data.table)
 library(reshape2)
 library(assertthat)
+library(tm)
+library(tau)
 PATH = getPathToThisFile()
 
 options(sqldf.RPostgreSQL.user = 'claytonstanley',
@@ -28,28 +30,35 @@ topUsers = data.table(read.csv(str_c(PATH, "/data/tables/topUsers.csv")))
 twitter_users[created_at=='2010-10-29 19:05:25',]
 #fread('col1,col2\n5,"4\n3"')
 
-findHashtags = function(str) {
-	assert_that(length(str) == 1) # code isn't yet vectorized
-	regex = "#(\\d|\\w)+"
-	matches = gregexpr(regex, str)
-	hashtags = regmatches(str,matches)
-	list(hashtag=hashtags[[1]], start=unlist(matches[[1]]))
+getTokenizedTbl = function(tweetsTbl) {
+	regex = '\\S+'
+	matches = with(tweetsTbl, regmatches(text, gregexpr(regex, text, perl=T)))
+	wideTbl = data.table(id=tweetsTbl$id, matches=matches)[length(unlist(matches)) > 0,]
+	extractMatches = function (m) {
+		if (length(m) == 0) {
+			list(chunk=NULL, pos=NULL)
+		} else {
+			list(chunk=m, pos=1:length(m))
+		}
+	}
+	wideTbl[,extractMatches(unlist(matches)), by=id]
 }
 
-findHashtags('foo #bar #baz')
+foo = getTokenizedTbl(data.table(id=c(1,2,3), text=c('kfkf idid','','ie #2')))
 
-addHashtagsToTweets = function(tweetsTbl) {
-	hashtags = tweetsTbl[, findHashtags(text), by=id]
-	setkey(hashtags,id)
-	setkey(tweetsTbl,id)
-	tweetsTbl[hashtags, `:=`(hashtag=hashtag, start=start)]
-	invisible()
-}
+tweetsTbl = data.table(sqldf("select * from tweets limit 1000000"))
+setkey(tweetsTbl, id)
+tweetsTbl
+tokenizedTbl = getTokenizedTbl(tweetsTbl)
+tokenizedTbl
+htOfTokenizedTbl= tokenizedTbl[grepl('^#', chunk),]
+setkey(htOfTokenizedTbl,id)
[…]

### 2014-01-16 writeups/proposal/proposal.tex
foo
Former-commit-id: c04fca3ad06f7d84f695a57829666dd6f6798401 [formerly 4332a89b0b3acf3ed9541be8039aae759fce0d6e]
Former-commit-id: 3a19506cf90af2263411a2d7e63f6667c2feb50d

@@
 
-The ACT-R cognitive architecture \parencite{Anderson2004} contains one of the leading theories of human declarative memory.
+The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human declarative memory.
 The declarative memory component in ACT-R is a Bayesian statistical model.
@@
 
-However it is much less common in the cognitive modeling community to use a large number of observations to build up declarative memory when creating ACT-R models.
+However, it is much less common in the cognitive modeling community to use a large number of observations to build up declarative memory when creating ACT-R models.
 If a cognitive modeler is using the declarative memory component of ACT-R,
@@
 
-The work for this proposal will expand on the knowledge gained from previous work that has tested retrieval models on large-scale datasets.
-It will focus specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutations) can be improved further, by testing these models on two real-world tagging tasks.
-One primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory when the datasets used for each model contain hundreds of millions to billions of data points.
-Another objective is to test how much improvement is gained by incorporating the strengths of each model into the other (i.e., modifying model components).
+The work for this proposal will expand on previous work that has tested retrieval models on large-scale datasets.
+It will focus specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutations) can be improved, by testing these models on two real-world tagging tasks.
+One primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory when large datasets are used for each model. 
+Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
 For example, it is currently unexplored how a vector-based model can be modified to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
-Since large datasets will be used, a rich volume of data will be available to test the models.
-This enables constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.
+Large-scale datasets will be used throughout the model development and evaluation process.
+This provides an ideal testing environment for constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.
 
@@
 With that understanding, we can tailor the system to suggest hashtags that they may not yet know about but are of likely interest to them.
-We can help aid in their discovery of fresh and relevant information that match their goals on the site.
+We can help aid in their discovery of fresh and relevant information that matches their goals on the site.
 
@@
 
-Alongside testing these memory retrieval models on Twitter hashtag use, they will also be tested on tag use for StackOverflow posts.
+These memory retrieval models will also be tested on tag use for StackOverflow posts.
 StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
@@
 
-ACT-R's declarative memory system places strong constraints on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2004}.
+ACT-R's declarative memory system places strong constraints on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2007}.
 This decay rate equation formalizes how recency and frequency of hashtag use relate to the prior probability that a hashtag will be retrieved.
@@
 
-A user's prior hashtag use will be an essential model component for domains such as Twitter, where the number of possible hashtags is practically infinite.
+A user's prior hashtag use will be an essential model component for domains such as Twitter, where the number of possible hashtags is effectively infinite.
 In these domains, the model can utilize a user's hashtag history as a way to prune the infinite space of possible hashtags, and generate a much smaller and more manageable set for prediction.
@@
 
-ACT-R \parencite{Anderson2004} is a cognitive architecture that formalizes how each cognitive process of the brain (e.g., memory, learning, visual and motor) interacts to produce behavior.
+ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the brain (e.g., memory, learning, visual and motor) interacts to produce behavior.
 The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up long-term declarative memory retrieval.
@@
 Usually information is lost when the data are restructured into a smaller set of latent dimensions.
-However it was shown that this process actually improves prediction and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
+However, it was shown that this process actually improves prediction and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
 \textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
@@
 Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occurred with that $e_{i}$ environment vector.
-These memory vectors can contain environment vectors from bag-of-words co-occurrence information ($c_{i}$) and word order information ($o_{i}$), which can be combined into a single representation ($m_{i}$).
+These memory vectors can contain environment vectors from position-independent word co-occurrence information ($c_{i}$) and word order information ($o_{i}$), which can be combined into a single representation ($m_{i}$).
 Over time, memory vectors accumulate a distributed representation of the most common environment vectors that co-occurred with them.
@@
 
-More specific hashtag-recommendation models that attempt to predict a user's chosen hashtag have also been created and tested for a few social media sites.
+Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
 Google has even deployed a hashtag-recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
 A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
-Models for Twitter and StackOverflow have also been proposed and tested, but none have yet been deployed.
+Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
 
@@
 
-\textcite{Stanley2013}'s model was based on ACT-R's declarative memory retrieval theory, which at the core is also a Bayesian co-occurrence model.
-However several modifications to the model were made from the standard Bayesian ACT-R declarative memory theory, so it was more sophisticated than the standard ACT-R model and the model used by \textcite{Kuo2011}.
+Several modifications to \textcite{Stanley2013}'s model were made from the standard Bayesian ACT-R declarative memory theory,
+so it was more sophisticated than the standard ACT-R model and the model used by \textcite{Kuo2011}.
 \citeauthor{Stanley2013} used a data-driven entropy weighting measure to attenuate low-predictor words and increase the weighting of high-predictor words in the body and title of a post.
@@
 
-\textcite{Stanley2013} used an ACT-R inspired Bayesian memory model to predict human tagging on the StackOverflow site, but to our knowledge no vector-based holographic memory model has been tested or used.
+\textcite{Stanley2013} used an ACT-R inspired Bayesian memory model to predict human tagging on the StackOverflow site, but to our knowledge no vector-based memory model has been tested or used.
 Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
@@
 
-There has been a mix of co-occurrence and latent-based models created to predict hashtag use on Twitter, and the results are encouraging.
+There has been a mix of co-occurrence models created to predict hashtag use on Twitter, and the results are encouraging.
 However, there has been very little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
@@
 
-A modified ACT-R declarative memory retrieval model and random-permutation vector-based model will be evaluated on predicting user-chosen tags for Twitter tweets and StackOverflow posts.
+The two memory models will be evaluated on predicting user-chosen tags for Twitter tweets and StackOverflow posts.
 Parameters for each model will be calibrated on a large subset of the Twitter and StackOverflow datasets.
@@
 One straightforward way to include a user's prior tag history into a random permutation model is to include ACT-R's base-level activation term ($B_{i}$) as an additional term in the model.
-However this may not be the most mathematically appropriate thing to do, since base-level activation is represented as a log odds value,
+However, this may not be the most mathematically appropriate thing to do, since base-level activation is represented as a log odds value,
 and the single strength of association term in a random permutation model is computed by a correlation.
@@
 A co-occurrence is counted if a word and a tag co-occur anywhere in a tweet or post.
-However for Twitter in particular, it may be the case that the few words immediately preceding hashtags are highly predictive, and the word position for those words should be included in the model.
+However, for Twitter in particular, it may be the case that the few words immediately preceding hashtags are highly predictive, and the word position for those words should be included in the model.
 So a method to include word order into a Bayesian model will be developed.
@@
 
-The end result of this research will be [1] a better understanding of how a user's past tagging history influences future tag use,
-[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on large-scale real-world retrieval tasks,
-[3] improvements to each memory model where the strengths of each model (word order and user prior likelihood) are incorporated into the other,
-and [4] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets.
-
-The tag-recommendation models developed through this research should be immediately relevant to real-world tasks.
-If model accuracy is high enough, then the developed models may be used as a foundation for tag recommendation systems that are deployed on Twitter or StackOverflow.
-It is also likely that the developed models will scale if deployed to these two sites, since during development they will be tested on a large portion of each site's currently available data.
-
-This research also aims to find task-agnostic memory models that can be applied across a wide range of domains.
-As much as possible, the two developed models that are evaluated on the StackOverflow and Twitter datasets will not differ across the datasets.
-This should help ensure that the developed models are accurate in general, and can potentially be used as tag recommendation systems for other domains.
-Examples of similar tagging tasks are hashtag use on Facebook, flagging (i.e., tagging) Wikipedia articles for content issues or overly-opinionated material,
-auto labeling incoming mail and placing it into appropriate folders, and spam filtering.
-Having an accurate and efficient co-occurrence memory model that correctly retrieves relevant tags can be applied to a wide variety of real-world tasks.
-
-This research also has numerous components that are relevant to improving our understanding of declarative memory.
+This research has numerous components that are relevant to improving our understanding of declarative memory.
 It should be emphasized that it has only recently become possible to evaluate psychological theories of declarative memory on large-scale real-world tasks where the theory can be tested on hundreds of millions of data points.
@@
 Finding a task-agnostic way to represent word order information in a Bayesian ACT-R model could improve model accuracy when applied to other domains where it seems plausible that participants are utilizing word order.
+
+Additionally, the tag-recommendation models developed through this research should be immediately relevant to real-world tasks.
+If model accuracy is high enough, then the developed models may be used as a foundation for tag recommendation systems that are deployed on Twitter or StackOverflow.
+
+Also since the developed models will be task agnostic, they can potentially be used as tag recommendation systems for other domains.
+Examples of similar tagging tasks are hashtag use on Facebook, flagging (i.e., tagging) Wikipedia articles for content issues or overly-opinionated material,
+auto labeling incoming mail and placing it into appropriate folders, and spam filtering.
+Having an accurate and efficient co-occurrence memory model that correctly retrieves relevant tags can be applied to a wide variety of real-world tasks.
+
+The end result of this research will be [1] a better understanding of how a user's past tagging history influences future tag use,
+[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on large-scale real-world retrieval tasks,
+[3] improvements to each memory model where the strengths of each model (word order and user prior likelihood) are incorporated into the other,
+and [4] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
 Introducing, exploring, and evaluating reasonable architectural modifications to two commonly-used memory models is worth the effort:

### 2014-01-17 twitter/topUsers/preProcess.R
foo
Former-commit-id: 254577ddf28a5f48856102d8c33b3ca52e676685 [formerly 9ed1ebc49cf6f2013a3a5027b94878603279620a]
Former-commit-id: db2834ead282378063ea432731a204320b5e618e

@@ -129,34 +129,34 @@ compareHashtagTbls <- function() {
 	hashtagTblText[hashtagTblTokenText]
 }
 
-computeActs <- function(hashtags, dt, d) {
+computeActs <- function(hashtags, dt, cTime, d) {
 	myPrint(hashtags)
 	myPrint(dt)
 	myPrint(d)
-	cTime = dt[length(dt)]
+	#cTime = dt[length(dt)]
 	myPrint(cTime)
 	dt = cTime - dt
 	myPrint(dt)
 	indeces = dt>0
 	hashtagsSub = hashtags[indeces] 
 	myPrint(hashtagsSub)
+	cTimeSub = cTime[indeces]
 	dtSub = dt[indeces]
 	dtSubRep = rep(dtSub, times=length(d))
 	dRep = rep(d, each=length(dtSub))
 	hashtagsSubRep = rep(hashtagsSub)
 	myPrint(dtSubRep)
 	myPrint(dRep)
-	list(hashtag=hashtagsSubRep, partialAct=dtSubRep^(-dRep), dt=rep(cTime, length(hashtagsSubRep)), d=dRep)
+	list(hashtag=hashtagsSubRep, partialAct=dtSubRep^(-dRep), dt=cTimeSub, d=dRep)
 }
 
 computeActsForUser <- function(hashtag, dt, ds) {
-	myPrint(hashtag)
 	retIndeces = which(!duplicated(dt))[-1]
 	expect_true(length(retIndeces) > 0)
 	partialRes = data.table(i=retIndeces)
 	myPrint(partialRes)
-	partialRes = partialRes[, computeActs(hashtag[1:i], dt[1:i], d=ds), by=i]
-	partialRes[, i := NULL]
+	partialRes = partialRes[, list(hashtag=hashtag[1:i], dt=dt[1:i], cTime=dt[i]), by=i]
+	partialRes = with(partialRes, as.data.table(computeActs(hashtag, dt, cTime, d=ds)))
 	partialRes
 }
 
@@ -167,7 +167,7 @@ computeActsByUser <- function(hashtagsTbl, ds) {
 	res = partialRes[, list(N=.N, act=log(sum(partialAct))), by=list(dt, hashtag, d, user_screen_name)]
 	with(res, expect_that(any(is.infinite(act)), is_false()))
 	Rprof(NULL)
-	myPrint(summaryRprof())
+	print(summaryRprof())
 	res
 }
 

### 2014-01-17 twitter/topUsers/preProcess.R
foo
Former-commit-id: 11e592972085e03c9c4464ac53925ec6ccd3d899 [formerly bd922e25af41a41431c58a2525bfb1e848fc5d42]
Former-commit-id: 41b863699b72cc855dd1a4bc2b8a45fa889c2d69

@@ -141,16 +141,18 @@ computeActs <- function(hashtags, dt, cTime, d) {
 	hashtagsSub = hashtags[indeces] 
 	myPrint(hashtagsSub)
 	cTimeSub = cTime[indeces]
+	cTimeSubRep = rep(cTimeSub, times=length(d))
 	dtSub = dt[indeces]
 	dtSubRep = rep(dtSub, times=length(d))
 	dRep = rep(d, each=length(dtSub))
 	hashtagsSubRep = rep(hashtagsSub)
 	myPrint(dtSubRep)
 	myPrint(dRep)
-	list(hashtag=hashtagsSubRep, partialAct=dtSubRep^(-dRep), dt=cTimeSub, d=dRep)
+	list(hashtag=hashtagsSubRep, partialAct=dtSubRep^(-dRep), dt=cTimeSubRep, d=dRep)
 }
 
-computeActsForUser <- function(hashtag, dt, ds) {
+computeActsForUser <- function(hashtag, dt, ds, user_screen_name) {
+	print(sprintf('computing partial activation for user %s', user_screen_name))
 	retIndeces = which(!duplicated(dt))[-1]
 	expect_true(length(retIndeces) > 0)
 	partialRes = data.table(i=retIndeces)
@@ -162,8 +164,9 @@ computeActsForUser <- function(hashtag, dt, ds) {
 
 computeActsByUser <- function(hashtagsTbl, ds) {
 	Rprof()
-	partialRes = hashtagsTbl[, computeActsForUser(hashtag, dt, ds), by=user_screen_name]
+	partialRes = hashtagsTbl[, computeActsForUser(hashtag, dt, ds, user_screen_name), by=user_screen_name]
 	myPrint(partialRes)
+	print('computing activations across table')
 	res = partialRes[, list(N=.N, act=log(sum(partialAct))), by=list(dt, hashtag, d, user_screen_name)]
 	with(res, expect_that(any(is.infinite(act)), is_false()))
 	Rprof(NULL)
@@ -210,6 +213,17 @@ testPriorActivations <- function() {
 	expectedActTbl
 	expect_that(actTbl, is_equivalent_to(expectedActTbl))
 
+	testHashtagsTbl = data.table(user_screen_name=c(1,1,1), dt=c(0,2,3), hashtag=c('a','b','c'))
[…]

### 2014-01-17 twitter/topUsers/preProcess.R
foo
Former-commit-id: f7d54fad36b7668efd0b4e5577838ce19653a823 [formerly 73854693d9b519ddec50e6b8510d10806175fcb7]
Former-commit-id: cf0900fee113e2d70571b1421c8943a572e7ed10

@@ -324,7 +324,7 @@ curWS <- function() {
 	unique(hashtagsTbl$user_screen_name)
 	visCompare(hashtagsTbl[user_screen_name=='icarly'], modelHashtagsTbl[topHashtag==T & user_screen_name=='icarly',], db)
 	summarizeExtremes(modelHashtagsTbl)
-	modelVsPredTbl = genAggModelVsPredTbl(hashtagsTbl[user_screen_name=='icarly' | user_screen_name=='icarly'])
+	modelVsPredTbl = genAggModelVsPredTbl(hashtagsTbl[user_screen_name=='billcosby' | user_screen_name=='icarly'])
 	modelVsPredTbl = genAggModelVsPredTbl(hashtagsTbl)
 	modelVsPredTblBig = modelVsPredTbl 
 	modelVsPredTbl

### 2014-02-18 twitter/topUsers/preProcess.R
foo
Former-commit-id: 578d4c87223d94c04a524eb39a341d11ddad32d2 [formerly 765899e7e229229c0aa5c5c693cc8f7dc64cd9c9]
Former-commit-id: ed0313a6932fceba39261e2fb207081b96da99e5

@@ -407,11 +407,11 @@ genAggModelVsPredTbl <- function(hashtagsTbl, ds=c(0,.1,.2,.3,.4,.5,.6,.7,.8,.9,
 }
 
 visModelVsPredTbl <- function(modelVsPredTbl) {
-	assign('p1', ggplot(modelVsPredTbl[maxNP==T & topHashtag & hashtagUsedP], aes(totN, d)) +
+	assign('p1', ggplot(modelVsPredTbl[predUsedBest == T], aes(totN, d)) +
 	       geom_point() +
 	       xlab('total number of hashtags for user'))
-	modelVsPredTbl[topHashtag & hashtagUsedP, meanPC := mean(NCell/totN), by=user_screen_name]
-	modelVsPredTbl[topHashtag & hashtagUsedP, relPC := NCell/totN - mean(NCell/totN), by=user_screen_name]
+	modelVsPredTbl[topHashtag & hashtagUsedP, meanPC := mean(acc), by=user_screen_name]
+	modelVsPredTbl[topHashtag & hashtagUsedP, relPC := acc - mean(acc), by=user_screen_name]
 	modelVsPredTbl[topHashtag & hashtagUsedP, meanRelPC := mean(relPC), by=d]
 	assign('p2', ggplot(modelVsPredTbl[topHashtag & hashtagUsedP], aes(log(d),relPC)) +
 	       geom_point() +
@@ -419,9 +419,9 @@ visModelVsPredTbl <- function(modelVsPredTbl) {
 	       xlab('log(d)') +
 	       ylab('normalized mean for user'))
 	assign('p3', ggplot(modelVsPredTbl[topHashtag & hashtagUsedP & user_screen_name %in% sample(unique(user_screen_name), size=min(20, length(unique(user_screen_name))))],
-			    aes(log(d),NCell/totN, group=as.factor(user_screen_name))) + geom_line() +
+			    aes(log(d),acc, group=as.factor(user_screen_name))) + geom_line() +
 	       ylab('proportion correct'))
-	assign('p4', ggplot(modelVsPredTbl[topHashtag & hashtagUsedP & maxNP], aes(d)) +
+	assign('p4', ggplot(modelVsPredTbl[predUsedBest == T], aes(d)) +
 	       geom_histogram(aes(y = ..density..)) +
 	       geom_density())
 	dev.new()
[…]

### 2014-02-18 twitter/topUsers/preProcess.R
foo
Former-commit-id: 0fba1353a8e3e30513bf5750c53cde27ca88b228 [formerly a6fa67da5941b67461526935d06a14cea7a78c8f]
Former-commit-id: 84544f012d6850afc4ae010334e22749ae2ff436

@@ -672,6 +672,7 @@ compareMeanDV <- function(modelVsPredTbl, DV) {
 	DV = substitute(DV)
 	expr = bquote(tableModelVsPredTbl(modelVsPredTbl)[, withCI(.(DV)), keyby=list(DVName, datasetGroup)])
 	sumTbl = eval(expr)
+	dev.new()
 	myLog(ggplot(sumTbl, aes(x=factor(datasetGroup), y=meanVal, fill=DVName)) +
 	      geom_bar(position=position_dodge(), stat='identity') +
 	      geom_errorbar(aes(ymin=minCI, ymax=maxCI), position=position_dodge(width=0.9), width=0.1, size=0.3))

### 2014-02-18 twitter/topUsers/preProcess.R
foo
Former-commit-id: 1689e78174f1a805f49a6a5ebb759ea86c4b26d6 [formerly 2f5ce9b2c25f15270ed20235dfdb42bd7c874c26]
Former-commit-id: 2f5b6ec3820a6c4d59d86088af30ada860f4729e

@@ -668,15 +668,29 @@ compare2Runs <- function(modelVsPredTbl, runNums) {
 	sumTbl[, list(diff=acc[2]-acc[1], direction=paste(runNums[2], '-', runNums[1])), by=list(datasetNameRoot, DVName, user_screen_name)][, genComparisonTbl(.SD)]
 }
 
+plotBarSumTbl <- function(sumTbl, fillCol) {
+	fillCol = substitute(fillCol)
+	dev.new()
+	expr = bquote(myLog(ggplot(sumTbl, aes(x=factor(datasetGroup), y=meanVal, fill=.(fillCol))) +
+			    geom_bar(position=position_dodge(), stat='identity') +
+			    geom_errorbar(aes(ymin=minCI, ymax=maxCI), position=position_dodge(width=0.9), width=0.1, size=0.3)))
+	eval(expr)
+	sumTbl
+}
+
 compareMeanDV <- function(modelVsPredTbl, DV) {
 	DV = substitute(DV)
 	expr = bquote(tableModelVsPredTbl(modelVsPredTbl)[, withCI(.(DV)), keyby=list(DVName, datasetGroup)])
 	sumTbl = eval(expr)
-	dev.new()
-	myLog(ggplot(sumTbl, aes(x=factor(datasetGroup), y=meanVal, fill=DVName)) +
-	      geom_bar(position=position_dodge(), stat='identity') +
-	      geom_errorbar(aes(ymin=minCI, ymax=maxCI), position=position_dodge(width=0.9), width=0.1, size=0.3))
-	sumTbl
+	plotBarSumTbl(sumTbl, DVName)
+}
+
+plotDVDiffs <- function(sumTbl) {
+	plotBarSumTbl(sumTbl, direction)
+}
+
+plotRunDiffs <- function(sumTbl) {
+	plotBarSumTbl(sumTbl, DVName)
 }
 
 compareOptimalDs <- function(modelVsPredTbl) compareMeanDV(modelVsPredTbl, median)
@@ -713,9 +727,9 @@ analyzeModelVsPredTbl <- function(modelVsPredTbl) {
 	modelVsPredTbl[topHashtag == T & d==0][, list(totN, sum(NCell)), by=list(user_screen_name,d,DVName, datasetName)][!is.na(totN)][,list(res=totN-V2)][, withCI(res)]
 	# Check that the Ns for each dataset look right	
[…]

### 2014-02-18 twitter/topUsers/preProcess.R
foo
Former-commit-id: 4cc2ce47fbd83cc30e4418b52ed4589f5483b461 [formerly 06300a90fe1f86bd07d62898ba07f4f065c53abf]
Former-commit-id: 9787ae1c628e2c5f519ca00c3631f25309694d81

@@ -425,13 +425,13 @@ visModelVsPredTbl <- function(modelVsPredTbl) {
 	       geom_histogram(aes(y = ..density..)) +
 	       geom_density())
 	dev.new()
-	myLog(p1)
+	myPlotPrint(p1)
 	dev.new()
-	myLog(p2)
+	myPlotPrint(p2)
 	dev.new()
-	myLog(p3)
+	myPlotPrint(p3)
 	dev.new()
-	myLog(p4)
+	myPlotPrint(p4)
 }
 
 tableModelVsPredTbl <- function(modelVsPredTbl) {
@@ -671,9 +671,9 @@ compare2Runs <- function(modelVsPredTbl, runNums) {
 plotBarSumTbl <- function(sumTbl, fillCol) {
 	fillCol = substitute(fillCol)
 	dev.new()
-	expr = bquote(myLog(ggplot(sumTbl, aes(x=factor(datasetGroup), y=meanVal, fill=.(fillCol))) +
-			    geom_bar(position=position_dodge(), stat='identity') +
-			    geom_errorbar(aes(ymin=minCI, ymax=maxCI), position=position_dodge(width=0.9), width=0.1, size=0.3)))
+	expr = bquote(myPlotPrint(ggplot(sumTbl, aes(x=factor(datasetGroup), y=meanVal, fill=.(fillCol))) +
+				  geom_bar(position=position_dodge(), stat='identity') +
+				  geom_errorbar(aes(ymin=minCI, ymax=maxCI), position=position_dodge(width=0.9), width=0.1, size=0.3)))
 	eval(expr)
 	sumTbl
 }

### 2014-02-18 twitter/topUsers/preProcess.R
foo
Former-commit-id: c0dba434a460d8ef015d75f7d1708e10a26e2a70 [formerly 830d9ac46314b3ce09d1ead13180e73db9c27f42]
Former-commit-id: a8a2d70dfbe84c132bd38b3b5f32ff49ca1a011d

@@ -437,10 +437,10 @@ visModelVsPredTbl <- function(modelVsPredTbl) {
 tableModelVsPredTbl <- function(modelVsPredTbl) {
 	# Summary table of optimal d values and sample variance
 	modelVsPredTbl[predUsedBest == T][, list(mean=mean(d), median=median(d), totN=mean(totN), NCell=mean(NCell), acc=mean(NCell/totN),
-					    datasetName=datasetName[1],
-					    datasetGroup=datasetGroup[1],
-					    #sdCI=sqrt(CIVar(d)), sdCI1=sqrt(CIVar2(d)), meanCI=CI(d))
-					    sd=sd(d)), by=list(datasetNameRoot, runNum, DVName)]
+						 datasetName=datasetName[1],
+						 datasetGroup=datasetGroup[1],
+						 #sdCI=sqrt(CIVar(d)), sdCI1=sqrt(CIVar2(d)), meanCI=CI(d))
+						 sd=sd(d)), by=list(datasetNameRoot, runNum, DVName)]
 }
 
 modelVsPredDir <- function() {
@@ -507,7 +507,7 @@ getQuerySOQ <- function(val) {
 		 where N > %d
 		 order by N asc
 		 limit 100)
-		 ", val)
+		", val)
 }
 
 combineFilters <- function(f1, f2) {
@@ -702,9 +702,9 @@ wrapQuotes <- function(charVect) {
 }
 
 plotTemporal <- function(modelHashtagsTbl, hashtagsTbl) {
-	db = makeDB(do.call(function(x) sample(x, length(x)), list(unique(res$hashtagsTbl$hashtag))))
-	visHashtags(res$hashtagsTbl, db)
-	visCompare(res$hashtagsTbl, res$modelHashtagsTbl[topHashtagPost==T & d %in% c(0,20,.5,.8)], db)
+	db = makeDB(do.call(function(x) sample(x, length(x)), list(unique(hashtagsTbl$hashtag))))
+	visHashtags(hashtagsTbl, db)
+	visCompare(hashtagsTbl, modelHashtagsTbl[topHashtagPost==T & d %in% c(0,20,.5,.8)], db)
 }
 
 analyzeTemporal <- function(modelVsPredTbl) {
@@ -713,11 +713,11 @@ analyzeTemporal <- function(modelVsPredTbl) {
 	#user_screen_names = screenTbl[, paste(user_screen_name, sep='', collapse=','), by=datasetType]
 	#user_screen_names = user_screen_names[, V1]
[…]

### 2014-02-18 twitter/topUsers/preProcess.R
foo
Former-commit-id: 32927e83007bc409aba54cf39c74aedffa7685e2 [formerly b0c9a2e1c27577f21f7a7fc03bf15e4d5291e4d4]
Former-commit-id: 87b60088895b4fca8bd667740958d9aeaed20757

@@ -61,6 +61,11 @@ myLog <- function(str, forLogLevel=1) {
 	if (logLevel >= forLogLevel) print(str)
 }
 
+myPlotPrint <- function(fig) {
+	dev.new()
+	myLog(fig)
+}
+
 # ref: http://stackoverflow.com/questions/5060076/convert-html-character-entity-encoding-in-r
 html2txt <- function(str) {
 	xpathApply(htmlParse(str, asText=TRUE),
@@ -276,23 +281,21 @@ computeActsByUser <- function(hashtagsTbl, ds) {
 }
 
 visHashtags <- function(hashtagsTbl, db) {
-	hashtagsTbl[, {dev.new(); plot(dt, main=user_screen_name)}, by=user_screen_name]
-	hashtagsTbl[, {dev.new(); plot(getHashes(hashtag, db), dt, main=user_screen_name)}, by=user_screen_name]
+	plots = hashtagsTbl[, list(resPlots=list(ggplot(.SD, aes(x=hashtag, y=dt)) + geom_point())), by=user_screen_name]
+	plots[, myPlotPrint(resPlots), by=user_screen_name]
+	plots
 }
 
 visCompare <- function(hashtagsTbl, modelHashtagsTbl, db) {
 	expect_that(sort(unique(hashtagsTbl$user_screen_name)), is_equivalent_to(sort(unique(modelHashtagsTbl$user_screen_name))))
-	plotDFun <- function(hashtagsTbl, modelHashtagsTbl, userScreenName, d) {
-		dev.new()
-		with(hashtagsTbl, plot(getHashes(hashtag, db), dt, main=sprintf('%s, %f', userScreenName, d)))
-		with(modelHashtagsTbl, lines(getHashes(hashtag, db), dt, col='red', typ='p', pch=4, cex=.1))
+	plotBuildFun <- function(modelHashtagsTbl) {
+		list(resPlots=list(ggplot(hashtagsTbl, aes(x=hashtag, y=dt)) +
+				   geom_point() + 
+				   geom_point(data=modelHashtagsTbl, aes(x=hashtag, y=dt), colour="red", size=1)))
 	}
-	plotFun <- function(hashtagsTbl, modelHashtagsTbl, userScreenName) {
-		modelHashtagsTbl[, plotDFun(hashtagsTbl, .SD, userScreenName, d), by=d]
-
-	}
[…]

### 2014-02-18 twitter/topUsers/preProcess.R
foo
Former-commit-id: db215e352f03e6064dc3a3dcfe6b0b67728583ce [formerly 5f8628d52ad412f5989bb8c15fefbe73feb511bd]
Former-commit-id: c3b1a502c58e0a6de6ac4ed385f5575a99360ef3

@@ -61,9 +61,13 @@ myLog <- function(str, forLogLevel=1) {
 	if (logLevel >= forLogLevel) print(str)
 }
 
-myPlotPrint <- function(fig) {
+savePlotsP = T
+
+myPlotPrint <- function(fig, name) {
 	dev.new()
 	myLog(fig)
+	if (savePlotsP) ggsave(filename=sprintf('%s/figures/%s.pdf', PATH, name), plot=fig)
+	fig
 }
 
 # ref: http://stackoverflow.com/questions/5060076/convert-html-character-entity-encoding-in-r
@@ -280,13 +284,13 @@ computeActsByUser <- function(hashtagsTbl, ds) {
 	res
 }
 
-visHashtags <- function(hashtagsTbl, db) {
+visHashtags <- function(hashtagsTbl) {
 	plots = hashtagsTbl[, list(resPlots=list(ggplot(.SD, aes(x=hashtag, y=dt)) + geom_point())), by=user_screen_name]
-	plots[, myPlotPrint(resPlots), by=user_screen_name]
+	plots[, list(myPlotPrint(resPlots[[1]], sprintf('HTByTime-%s', .BY[1]))), by=user_screen_name]
 	plots
 }
 
-visCompare <- function(hashtagsTbl, modelHashtagsTbl, db) {
+visCompare <- function(hashtagsTbl, modelHashtagsTbl) {
 	expect_that(sort(unique(hashtagsTbl$user_screen_name)), is_equivalent_to(sort(unique(modelHashtagsTbl$user_screen_name))))
 	plotBuildFun <- function(modelHashtagsTbl) {
 		list(resPlots=list(ggplot(hashtagsTbl, aes(x=hashtag, y=dt)) +
@@ -294,7 +298,7 @@ visCompare <- function(hashtagsTbl, modelHashtagsTbl, db) {
 				   geom_point(data=modelHashtagsTbl, aes(x=hashtag, y=dt), colour="red", size=1)))
 	}
 	plots = modelHashtagsTbl[, plotBuildFun(.SD), by=list(user_screen_name, d)]
-	plots[, myPlotPrint(resPlots), by=list(user_screen_name, d)]
+	plots[, list(myPlotPrint(resPlots[[1]], sprintf('HTMByTime-%s-%s', .BY[1], .BY[2]))), by=list(user_screen_name, d)]
 	plots
 }
 
[…]

### 2014-02-18 twitter/topUsers/preProcess.R
foo
Former-commit-id: 83d2022cc471afe179dfa52b46700ef97cba5878 [formerly 9354ed205eff0b8e9c0da499242fdbb3c4d1074d]
Former-commit-id: edae448c172d620876abdac6a87f92e0428d2874

@@ -431,10 +431,11 @@ visModelVsPredTbl <- function(modelVsPredTbl) {
 	assign('p4', ggplot(modelVsPredTbl[predUsedBest == T], aes(d)) +
 	       geom_histogram(aes(y = ..density..)) +
 	       geom_density())
-	myPlotPrint(p1)
-	myPlotPrint(p2)
-	myPlotPrint(p3)
-	myPlotPrint(p4)
+	ext = sprintf('%s-%s', p1$data$datasetName[1], p1$data$DVName[1])
+	myPlotPrint(p1, sprintf('visDByN-%s', ext)) 
+	myPlotPrint(p2, sprintf('visNormMean-%s', ext)) 
+	myPlotPrint(p3, sprintf('visAcc-%s', ext)) 
+	myPlotPrint(p4, sprintf('visHistD-%s', ext)) 
 }
 
 tableModelVsPredTbl <- function(modelVsPredTbl) {

### 2014-02-18 twitter/topUsers/preProcess.R
foo
Former-commit-id: 14965a1879c44723dbefed23e8b9f1cc6b62f20d [formerly f5fe823f15d947037a67b2a7658b2dbd14d6b3a0]
Former-commit-id: 22ec95b3c4525a6c2c2d0f5b6fa285d19a40f11b

@@ -672,11 +672,12 @@ compare2Runs <- function(modelVsPredTbl, runNums) {
 	sumTbl[, list(diff=acc[2]-acc[1], direction=paste(runNums[2], '-', runNums[1])), by=list(datasetNameRoot, DVName, user_screen_name)][, genComparisonTbl(.SD)]
 }
 
-plotBarSumTbl <- function(sumTbl, fillCol) {
+plotBarSumTbl <- function(sumTbl, fillCol, figName) {
 	fillCol = substitute(fillCol)
 	expr = bquote(myPlotPrint(ggplot(sumTbl, aes(x=factor(datasetGroup), y=meanVal, fill=.(fillCol))) +
 				  geom_bar(position=position_dodge(), stat='identity') +
-				  geom_errorbar(aes(ymin=minCI, ymax=maxCI), position=position_dodge(width=0.9), width=0.1, size=0.3)))
+				  geom_errorbar(aes(ymin=minCI, ymax=maxCI), position=position_dodge(width=0.9), width=0.1, size=0.3),
+				  figName))
 	eval(expr)
 	sumTbl
 }
@@ -685,15 +686,15 @@ compareMeanDV <- function(modelVsPredTbl, DV) {
 	DV = substitute(DV)
 	expr = bquote(tableModelVsPredTbl(modelVsPredTbl)[, withCI(.(DV)), keyby=list(DVName, datasetGroup)])
 	sumTbl = eval(expr)
-	plotBarSumTbl(sumTbl, DVName)
+	plotBarSumTbl(sumTbl, DVName, sprintf('compareMeanDV-%s', deparse(DV)))
 }
 
 plotDVDiffs <- function(sumTbl) {
-	plotBarSumTbl(sumTbl, direction)
+	plotBarSumTbl(sumTbl, direction, sprintf('compareDVDiffs'))
 }
 
 plotRunDiffs <- function(sumTbl) {
-	plotBarSumTbl(sumTbl, DVName)
+	plotBarSumTbl(sumTbl, DVName, sprintf('compareRunDiffs'))
 }
 
 compareOptimalDs <- function(modelVsPredTbl) compareMeanDV(modelVsPredTbl, median)

### 2014-02-18 twitter/scratch/scratch.R
foo
Former-commit-id: cb0850ca19afd8e11c1143978129ec4bd12b94b9 [formerly 05eb8a8f923fa09f0065990f701d01da9c4df69e]
Former-commit-id: 68ae329e092ff38fd29f005331f04eb3b60000ba

@@ -1,12 +1,0 @@
-consumerKey = "vKbz24SqytZnYO33FNkR7w"
-consumerSecret = "jjobro8Chy9aKMzo8szYMz9tHftONLRkjNnrxk0"
-twitCred = getTwitterOAuth(consumerKey, consumerSecret)
-
-save(list="twitCred", file="~/twitteR_credentials")
-
-load("~/twitteR_credentials")
-
-registerTwitterOAuth(twitCred)
-
-
-

### 2014-02-18 twitter/scratch/scratch.py

@@ -1,7 +1,0 @@
-
-import tweetstream
-
-stream = tweetstream.SampleStream("claytonstanley1", "0987)(*&ujmUJM")
-
-for tweet in stream:
-	print tweet

### 2014-02-18 twitter/scratch/streamScratch.R

@@ -1,20 +1,0 @@
-PATH = getPathToThisFile()
-
-library(streamR)
-library(RCurl)
-library(ROAuth)
-library(stringr)
-
-library(twitteR)
-
-load("~/twitteR_credentials")
-
-registerTwitterOAuth(twitCred)
-
-sampleStream(file="tweetsScratch.json", oauth=twitCred, timeout=60)
-
-library(rjson)
-
-fromJSON(file=str_c(PATH, '/jsonTmp.json'))
-
-parseTweets(str_c(PATH, '/jsonTmp.json'))

### 2014-02-18 twitter/scratch2.py

@@ -1,50 +1,0 @@
-import sys
-import tweepy
-from bson import json_util
-import json
-import datetime
-import codecs
-import os
-
-consumer_key="vKbz24SqytZnYO33FNkR7w"
-consumer_secret="jjobro8Chy9aKMzo8szYMz9tHftONLRkjNnrxk0"
-access_key = "363361813-FKSdmwSbzuUzHWg326fTGJM7Bu2hTviqEetjMgu8"
-access_secret = "VKgzDnTvDUWR1csliUR3BiMOI2oqO9NzocNKX1jPd4" 
-
-auth = tweepy.OAuthHandler(consumer_key, consumer_secret)
-auth.set_access_token(access_key, access_secret)
-api = tweepy.API(auth)
-
-@classmethod
-def parse(cls, api, raw):
-	status = cls.first_parse(api, raw)
-	setattr(status, 'json', json.dumps(raw))
-	return status
-
-tweepy.models.Status.first_parse = tweepy.models.Status.parse
-tweepy.models.Status.parse = parse
-
-_dir = os.path.dirname(os.path.abspath(__file__))
-
-class CustomStreamListener(tweepy.StreamListener):
-	def on_status(self, status):
-		print status.user.screen_name
-		with codecs.open(os.path.join(_dir, 'data/tweets.json'), "a", 'utf-8') as textFile:
-			textFile.write(status.json + '\n')
-
-	def on_error(self, status_code):
-		print >> sys.stderr, 'Encountered error with status code:', status_code
-		return True # Don't kill the stream
-
-	def on_timeout(self):
-		print >> sys.stderr, 'Timeout...'
-		return True # Don't kill the stream
-
-sapi = tweepy.streaming.Stream(auth, CustomStreamListener())
-# sapi.filter(track=['curiosity'])
-# sapi.filter(track=['believetour'])
-# sapi.filter(locations=[-122.75,36.8,-121.75,37.8])
-
[…]

### 2014-02-18 dissProject/preProcess.R
cleaned up ...s
Former-commit-id: 0a6c08bfca5e129df0b57393d51ec3bb314c3183 [formerly ec9a2ee712b0619472dba8b43e77caa700ccfaa5]
Former-commit-id: d54bb919518ff9d5567990a00a1536d817ac3dc1

@@ -192,7 +192,7 @@ getPostsTbl <- function(sqlStr) {
 	postsTbl
 }
 
-getHashtagsTbl <- function(tweetsTbl, config, ...) {
+getHashtagsTbl <- function(tweetsTbl, config) {
 	from = config$from
 	tokenizedTbl = getTokenizedTbl(tweetsTbl, from=from, regex='\\S+')
 	htOfTokenizedTbl = tokenizedTbl[grepl('^#', chunk),]
@@ -202,7 +202,7 @@ getHashtagsTbl <- function(tweetsTbl, config, ...) {
 	hashtagsTbl
 }
 
-getTagsTbl <- function(postsTbl, config, ...) {
+getTagsTbl <- function(postsTbl, config) {
 	tokenizedTbl = getTokenizedTbl(postsTbl, from='tagsNoHtml', regex='(?<=<)[^>]+(?=>)')
 	if (config$convertTagSynonymsP) {
 		withKey(tokenizedTbl, chunk,
@@ -455,26 +455,33 @@ modelVsPredOutFile <- function(name) {
 	sprintf('%s/%s.csv', modelVsPredDir(), name)
 }
 
-runPrior <- function(query, config, getPostsFun, getHashtagsFun, ...) {
+runPrior <- function(config) {
 	withProf({
-		postsTbl = getPostsFun(query)
-		hashtagsTbl = getHashtagsFun(postsTbl, config=config, ...)
-		res = genAggModelVsPredTbl(hashtagsTbl, config=config, ...)
+		postsTbl = config$getPostsFun(config$query)
+		hashtagsTbl = config$getHashtagsFun(postsTbl, config=config)
+		res = genAggModelVsPredTbl(hashtagsTbl, config=config)
 		modelVsPredTbl = res$modelVsPredTbl
 		modelHashtagsTbl = res$modelHashtagsTbl
 		list(modelVsPredTbl=modelVsPredTbl, modelHashtagsTbl=modelHashtagsTbl, hashtagsTbl=hashtagsTbl)
 	})
 }
 
-defaultTConfig = list(from='tokenText', accumModelHashtagsTbl=F)
-defaultSOConfig = list(convertTagSynonymsP=F, accumModelHashtagsTbl=F)
+defaultTConfig = list(from='tokenText',
+		      accumModelHashtagsTbl=F,
+		      getPostsFun=getTweetsTbl,
+		      getHashtagsFun=getHashtagsTbl)
 
[…]

### 2014-02-18 dissProject/preProcess.R
foo
Former-commit-id: 62c1cb621d7aa7a56379935d0cd98e293da75890 [formerly 0161435cefcc57d5b2fe7206f5609ed23565890a]
Former-commit-id: 96f91fe5a39073c8a55aafdbb09c6b9488943263

@@ -168,8 +168,10 @@ getDiffTimeSinceFirst <- function(ts) {
 	as.numeric(difftime(ts, ts[1], units='secs'))
 }
 
-getTweetsTbl <- function(sqlStr="select * from tweets limit 10000") {
+getTweetsTbl <- function(sqlStr="select * from tweets limit 10000", config) {
 	tweetsTbl = data.table(sqldf(sqlStr))
+	stopifnot(sort(unique(tweetsTbl$retweeted)) == c('False', 'True'))
+	if (!config$includeRetweetsP) tweetsTbl = tweetsTbl[retweeted == 'False']
 	addTokenText(tweetsTbl, from='text')
 	setkey(tweetsTbl, id)
 	tweetsTbl[, dt := getDiffTimeSinceFirst(created_at), by=user_screen_name]
@@ -182,7 +184,7 @@ getTagSynonymsTbl <- function(sqlStr='select * from tag_synonyms') {
 	tagSynonymsTbl
 }
 
-getPostsTbl <- function(sqlStr) {
+getPostsTbl <- function(sqlStr, config) {
 	postsTbl = data.table(sqldf(sqlStr))
 	setkey(postsTbl, id)
 	stopifnot(!duplicated(postsTbl$id))
@@ -457,7 +459,7 @@ modelVsPredOutFile <- function(name) {
 
 runPrior <- function(config) {
 	withProf({
-		postsTbl = config$getPostsFun(config$query)
+		postsTbl = config$getPostsFun(config$query, config=config)
 		hashtagsTbl = config$getHashtagsFun(postsTbl, config=config)
 		res = genAggModelVsPredTbl(hashtagsTbl, config=config)
 		modelVsPredTbl = res$modelVsPredTbl
@@ -469,9 +471,10 @@ runPrior <- function(config) {
 defaultTConfig = list(from='tokenText',
 		      accumModelHashtagsTbl=F,
 		      getPostsFun=getTweetsTbl,
-		      getHashtagsFun=getHashtagsTbl)
+		      getHashtagsFun=getHashtagsTbl,
+		      includeRetweetsP=F)
 
-defaultSOConfig = list(convertTagSynonymsP=F,
+defaultSOConfig = list(convertTagSynonymsP=T,
 		       accumModelHashtagsTbl=F,
 		       getPostsFun=getPostsTbl,
 		       getHashtagsFun=getTagsTbl)
[…]

### 2014-02-19 dissProject/preProcess.R
foo
Former-commit-id: f39f0a25f3eb2042b6d5709084eb9b80283f486c [formerly 7123204ec2349174d3bb849349a03be04b800ad7]
Former-commit-id: 9a5122c01c2511df7e9a2d7488a5b5c303481696

@@ -457,11 +457,11 @@ modelVsPredOutFile <- function(name) {
 	sprintf('%s/%s.csv', modelVsPredDir(), name)
 }
 
-runPrior <- function(config) {
+runPrior <- function(config, ...) {
 	withProf({
 		postsTbl = config$getPostsFun(config$query, config=config)
 		hashtagsTbl = config$getHashtagsFun(postsTbl, config=config)
-		res = genAggModelVsPredTbl(hashtagsTbl, config=config)
+		res = genAggModelVsPredTbl(hashtagsTbl, config=config, ...)
 		modelVsPredTbl = res$modelVsPredTbl
 		modelHashtagsTbl = res$modelHashtagsTbl
 		list(modelVsPredTbl=modelVsPredTbl, modelHashtagsTbl=modelHashtagsTbl, hashtagsTbl=hashtagsTbl)
@@ -479,12 +479,12 @@ defaultSOConfig = list(convertTagSynonymsP=T,
 		       getPostsFun=getPostsTbl,
 		       getHashtagsFun=getTagsTbl)
 
-runPriorT <- function(config=defaultTConfig) {
-	runPrior(config)
+runPriorT <- function(config=defaultTConfig, ...) {
+	runPrior(config, ...)
 }
 
-runPriorSO <- function(config=defaultSOConfig) {
-	runPrior(config)
+runPriorSO <- function(config=defaultSOConfig, ...) {
+	runPrior(config, ...)
 }
 
 modConfig <- function(config, mods) {
@@ -529,7 +529,7 @@ combineFilters <- function(f1, f2='1=1') {
 }
 
 makeTRun <- function(val, outFileName, config) {
-	function() runPriorT(outFile=modelVsPredOutFile(outFileName), config=modConfig(config, query=config$query(val)))
+	function() runPriorT(config=modConfig(config, list(query=config$query(val))), outFile=modelVsPredOutFile(outFileName))
 }
 
 makeTRunr1 <- function(val, outFileName, ...) {
@@ -574,7 +574,7 @@ runTTweets5e4 <- makeTRunr3(50000, 'TTweetsgt5e4', filters="user_screen_name != 
 runTTweets5e4r2 <- makeTRunr4(50000, 'TTweetsgt5e4r2', filters="user_screen_name != 'stanhjerleid'")
 
 makeSORun <- function(val, outFileName, config) {
[…]

### 2014-02-19 dissProject/tests/test-R.R
foo
Former-commit-id: c5eeafd84f0976a4787ea6c193d1e147560bbb98 [formerly dbb555369f212522d07eb036d7179c7ff0c834bc]
Former-commit-id: 02dbe28a5547ccdc343d48c556eca3113d5d8ede

@@ -70,22 +70,22 @@ test_that("testOnlyFirstT", {
 
 test_that("testModelVsPred", {
 	expectedTbl = myReadCSV(modelVsPredOutFile('testing1'))
-	resTbl = runPriorT("select * from tweets where user_screen_name = 'ap'", outFile='/tmp/modelVsPred.csv')
-	expect_equivalent(expectedTbl, resTbl)
+	resTbl = runPriorT(config=modConfig(defaultTConfig, list(query="select * from tweets where user_screen_name = 'ap'")), outFile='/tmp/modelVsPred.csv')
+	expect_equivalent(expectedTbl, resTbl$modelVsPredTbl)
 	expectedTbl = myReadCSV(modelVsPredOutFile('testing2'))
 	expectedTbl[, totN := as.integer(totN)]
-	resTbl = runPriorT("select * from tweets where user_screen_name = 'thebucktlist'", outFile='/tmp/modelVsPred.csv')
-	expect_equivalent(expectedTbl, resTbl)
+	resTbl = runPriorT(config=modConfig(defaultTConfig, list(query="select * from tweets where user_screen_name = 'thebucktlist'")), outFile='/tmp/modelVsPred.csv')
+	expect_equivalent(expectedTbl, resTbl$modelVsPredTbl)
 	expectedTbl = myReadCSV(modelVsPredOutFile('twitter_ru'))
-	resTbl = runPriorT("select * from tweets where user_screen_name = 'twitter_ru'", outFile='/tmp/modelVsPred.csv')
-	expect_equivalent(expectedTbl, resTbl)
+	resTbl = runPriorT(config=modConfig(defaultTConfig, list(query="select * from tweets where user_screen_name = 'twitter_ru'")), outFile='/tmp/modelVsPred.csv')
+	expect_equivalent(expectedTbl, resTbl$modelVsPredTbl)
 })
 
 test_that("testModelVsPredSO", {
 	  expectedTbl = myReadCSV(modelVsPredOutFile('testingSO1'))
 	  expectedTbl[, user_screen_name := as.character(user_screen_name)]
-	  resTbl = runPriorSO("select * from posts where post_type_id = 1 and owner_user_id = 20")
-	  expect_equivalent(expectedTbl, resTbl)
+	  resTbl = runPriorSO(config=modConfig(defaultSOConfig, list(query="select * from posts where post_type_id = 1 and owner_user_id = 20")))
[…]

### 2014-02-19 dissProject/preProcess.R
foo
Former-commit-id: eae6bdf543230471612dd1cd60196dc512c9346e [formerly 715fce78e381cd98a8ef03ce84a2d41fb91cc618]
Former-commit-id: dd65144534f851dd72b65b368f64795c301b6574

@@ -130,7 +130,6 @@ sqlScratch <- function() {
 	data.table(sqldf('select * from topusers order by rank'))
 	sqldf("select * from tweets where user_screen_name='jlo'")
 	data.table(sqldf('select * from (select user_screen_name,count(*) from tweets group by user_screen_name) as t join topUsers as u on t.user_screen_name = u.user_screen_name order by rank'))
-	sqldf("select * from topUsers")[1:100,]
 	data.table(sqldf("select retweeted, count(*) from tweets group by retweeted"))[, min(count) / (max(count) + min(count))]
 	sqldf("select * from tweets where 1=1 and user_screen_name='katyperry' limit 10")
 	foo = data.table(sqldf("select * from twitter_users"))
@@ -358,6 +357,11 @@ onlyFirstT <- function(bool) {
 	ret
 }
 
+guardAllEqualP <- function(vect) {
+	stopifnot(length(unique(vect)) <= 1)
+	vect
+}
+
 modelVsPredForDV <- function(modelHashtagsTbl, DVName) {
 	tempTbl = modelHashtagsTbl[, list(NCell=.N, DVName=DVName), by=c('user_screen_name', DVName, 'hashtagUsedP', 'd')]
 	setnames(tempTbl, DVName, 'topHashtag')
@@ -433,7 +437,7 @@ visModelVsPredTbl <- function(modelVsPredTbl) {
 	assign('p4', ggplot(modelVsPredTbl[predUsedBest == T], aes(d)) +
 	       geom_histogram(aes(y = ..density..)) +
 	       geom_density())
-	ext = sprintf('%s-%s', p1$data$datasetName[1], p1$data$DVName[1])
+	ext = sprintf('%s-%s', guardAllEqualP(p1$data$datasetName)[1], guardAllEqualP(p1$data$DVName)[1])
 	myPlotPrint(p1, sprintf('visDByN-%s', ext)) 
 	myPlotPrint(p2, sprintf('visNormMean-%s', ext)) 
 	myPlotPrint(p3, sprintf('visAcc-%s', ext)) 
@@ -443,8 +447,8 @@ visModelVsPredTbl <- function(modelVsPredTbl) {
 tableModelVsPredTbl <- function(modelVsPredTbl) {
 	# Summary table of optimal d values and sample variance
[…]

### 2014-02-19 dissProject/preProcess.R
foo
Former-commit-id: 0de2ec086ba55ff6eb47b67993afb4c66f7dad1c [formerly dbcc2b9eb76cec237849d05324786a26a24c6235]
Former-commit-id: 8b9890a0726ed22d86da1552d74d685a4e1af944

@@ -130,7 +130,7 @@ sqlScratch <- function() {
 	data.table(sqldf('select * from topusers order by rank'))
 	sqldf("select * from tweets where user_screen_name='jlo'")
 	data.table(sqldf('select * from (select user_screen_name,count(*) from tweets group by user_screen_name) as t join topUsers as u on t.user_screen_name = u.user_screen_name order by rank'))
-	data.table(sqldf("select retweeted, count(*) from tweets group by retweeted"))[, min(count) / (max(count) + min(count))]
+	data.table(sqldf("select retweeted, count(*) from tweets group by retweeted"))
 	sqldf("select * from tweets where 1=1 and user_screen_name='katyperry' limit 10")
 	foo = data.table(sqldf("select * from twitter_users"))
 	twitter_users = myReadCSV(str_c(PATH, "/dissertationData/tables/twitter_users.csv"))

### 2014-03-06 writeups/midwestCogsci2014/abstract.tex
foo
Former-commit-id: 111a9e8abd8f578eb2bab86ca4a0a00e8a743fe3 [formerly 7de7033ce6feb90537902a118303eb24e752a30c]
Former-commit-id: a773a01b2b3db4f4b715be204a8cd46527bb62fd

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage[group-separator={,}]{siunitx}

\title{Abstract Submission for Fourth Annual Midwestern Cognitive Science Conference}

\author{Clayton Stanley (clayton.stanley@rice.edu) \\
  Department of Psychology, 6100 Main Street \\
  Houston, TX 77005 USA 
  \and Michael D. Byrne (byrne@rice.edu) \\
  Departments of Psychology and Computer Science, 6100 Main Street \\
  Houston, TX 77005 USA \\
}

\begin{document}

\maketitle

%TC:break Abstract
\begin{abstract}
  This research explores and further validates ACT-R's base-level learning architectural component on two large-scale real-world datasets.
  We framed the process where an author chooses tags and hashtags on StackOverflow posts and Twitter tweets as a declarative memory retrieval problem.
  Over \num{265000} posts and 2.4 million tweets from \num{3900} total authors were collected. 
  ACT-R's base-level learning equation and the simplified version of that equation (which assumes equally-spaced retrievals for each chunk) were used to predict each author's chosen hashtags, given their previous hashtag use.
  The model performance space was explored from over 70 million total declarative memory retrievals, at 22 different levels of the decay rate parameter, and 48 different subsets of the two datasets.
  The results show that for these datasets, the optimal value for the decay rate parameter for the simplified form of the equation is 0.4, slightly less than but close to the standard 0.5 value. 
  However, the results also suggest that a higher (0.7) value is optimal when the standard form of the equation is used.
  Further, the standard form can be implemented in a computationally efficient manner (20 ms per retrieval), and using the standard equation significantly improves model accuracy within subjects by 5\%.
  Taken together, this research suggests that it may be advantageous to set the standard form of the base-level learning equation as the default in ACT-R,
  and also increase the default value for the decay rate parameter to 0.7.
  The results also highlight the importance of customizing model retrieval predictions based on each user's prior history, as model accuracy is still significant when a retrieval is based purely on history (i.e., no context).
\end{abstract}
%TC:break _main_

\end{document}

### 2014-03-06 dissProject/preProcess.R

@@ -562,8 +562,13 @@ modConfig <- function(config, mods) {
 	newConfig
 }
 
+getQueryUsersSubset <- function(val, from) {
+	sprintf('select user_screen_name from twitter_users where %s > %d order by %s asc limit 100', from, val, from)
+}
+
+
 getQueryGeneralT <- function(val, from, filters) {
-	sprintf('select * from tweets where %s and user_screen_name in (select user_screen_name from twitter_users where %s > %d order by %s asc limit 100)', filters, from, val, from)
+	sprintf('select * from tweets where %s and user_screen_name in (%s)', filters, getQueryUsersSubset(val, from))
 }
 
 getQueryT <- function(val, filters='1=1') {
@@ -580,16 +585,8 @@ getQuerySO <- function(val) {
 }
 
 getQuerySOQ <- function(val) {
-	sprintf("select id, owner_user_id, creation_date, creation_epoch, title, tags from posts where post_type_id = 1 and owner_user_id in 
-		(select Owner_User_Id from
-		 (select owner_user_id, Post_Type_Id, count(*) as N from Posts 
-		  where Post_Type_Id = 1
-		  group by Owner_User_Id,Post_Type_Id) as foo2
-		 join Users on Users.Id = foo2.Owner_User_Id
-		 where N > %d
-		 order by N asc
-		 limit 100)
-		", val)
+	sprintf("select id, owner_user_id, creation_date, creation_epoch, title, tags from posts
+		where post_type_id = 1 and owner_user_id in (select id from users where num_questions > %d order by num_questions asc limit 100)", val)
 }
 
 combineFilters <- function(f1, f2='1=1') {
@@ -600,46 +597,109 @@ makeTRun <- function(val, outFileName, config) {
 	function() runPriorT(config=modConfig(config, list(query=config$query(val), modelVsPredOutFile=getModelVsPredOutFile(outFileName))))
 }
 
-makeTRunr1 <- function(val, outFileName, ...) {
[…]

### 2014-03-06 dissProject/preProcess.R
foo
Former-commit-id: bd44d4b26a3a88db6d75b15761baeae002476905 [formerly e2c535443beef18ad366d1de343e946dd6830e66]
Former-commit-id: 2fc4cc6cecd9e0b7afdce441c07a64b9a9b83900

@@ -668,9 +668,11 @@ getSummaryStats <- function() {
 	tPostsCntFromRunsTbl = rbindlist(lapply(runTQueries, getTPostsFromRunQuery))
 	SOUsersFromRunsTbl = rbindlist(lapply(runSOQueries, getSOUsersFromRunQuery))
 	SOPostsCntFromRunsTbl = rbindlist(lapply(runSOQueries, getSOPostsFromRunQuery))
-	usersFromRunsTbl = rbind(tUsersFromRunsTbl, SOUsersFrmRunsTbl)
+	usersFromRunsTbl = rbind(tUsersFromRunsTbl, SOUsersFromRunsTbl, use.names=F)
+	unique(usersFromRunsTbl)
 	postsCntFromRunsTbl = rbind(tPostsCntFromRunsTbl, SOPostsCntFromRunsTbl)
-	modelVsPredTbl[predUsedBest == T][runNum == 2][, list(acc=median(acc), d=median(d), N=.N), by=DVName]
+	postsCntFromRunsTbl[, sum(count), by=dataset]
+	modelVsPredTbl[predUsedBest == T][runNum == 2][, list(acc=median(acc), d=median(d), N=.N), by=list(DVName, datasetType)]
 	modelVsPredTbl[hashtagUsedP == T][topHashtag == T][DVName == 'topHashtagAct'][, sum(totN), by=datasetType]
 	modelVsPredTbl[, .N, by=d]
 }

### 2014-03-06 dissProject/preProcess.R
foo
Former-commit-id: 470f319b19756d00fc1193334c48738e99e9a4a3 [formerly 5f4db3eeb62f4de299459d48dd453730e1a53d82]
Former-commit-id: 26d11d87266f9026a0587c69d70a57be7c5f2b78

@@ -972,8 +972,8 @@ genTokenizedTblSO <- function(filters='1=1', bundleSize=10000) {
 			      maxRowId = sqldf(sprintf("select max(row_id) from post_tokenized"))$max
 			      maxRowId = if (is.na(maxRowId)) 0 else maxRowId
 			      myLog(sprintf('Writing %s rows to post_tokenized table', nrow(tbl)))
-			      tbl[, pos_as_char := as.character(pos)][, row_id := seq(maxRowId+1, length.out=length(id))]
-			      setcolorder(tbl, c('row_id', 'id', 'chunk', 'pos', 'pos_as_char', 'type'))
+			      tbl[, row_id := seq(maxRowId+1, length.out=length(id))]
+			      setcolorder(tbl, c('row_id', 'id', 'chunk', 'pos', 'type'))
 			      withDBConnect(dbCon, dbWriteTable(dbCon, "post_tokenized", tbl, append=T, row.names=0))
 		      }
 		      while (T) {

### 2014-03-07 dissProject/preProcess.R
foo
Former-commit-id: dc69ec78a880c9b9350effb23543828c28d952dc [formerly 52f959aba595bcfa9dc05ce926d26ab45db65799]
Former-commit-id: b4c7e592ea44d09ba19f564e0d505dd93206faaf

@@ -121,6 +121,7 @@ myReadCSV <- function(file, ...) {
 }
 
 myWriteCSV <- function(data, file, ...) {
+	myLog(sprintf('Writing results to "%s"', file))
 	write.csv(data, file=file, row.names=F, ...)
 }
 
@@ -914,20 +915,27 @@ analyzeModelVsPredTbl <- function(modelVsPredTbl) {
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostOL2' & datasetName=='SOQgt500r2' & d < 1])
 }
 
-getNcoocTbl <- function(type, idsQuery) {
-	myLog(sprintf('Getting Ncooc table for type %s and idsQuery %s', type, idsQuery))
-	resTbl = as.data.table(sqldf(sprintf("select title_tbl.chunk as chunk, tag_tbl.chunk as tag, title_tbl.pos - tag_tbl.pos as pos_from_tag, count(tag_tbl.id) as partial_N from
-					     (select * from post_tokenized
-					      where id in (%s)
-					      and type = 'tag') as tag_tbl
-					     join
-					     (select * from post_tokenized
-					      where id in (%s)
-					      and type = '%s') as title_tbl
-					     on tag_tbl.id = title_tbl.id
-					     group by title_tbl.chunk, tag_tbl.chunk, pos_from_tag
-					     order by title_tbl.chunk, tag_tbl.chunk, pos_from_tag",
-					     idsQuery, idsQuery, type)))
+getNcoocTbl <- function(type, chunkTableQuery) {
+	myLog(sprintf('Getting Ncooc table for type %s', type))
+	sqldf('truncate table temp_cooc')
+	sqldf(sprintf("insert into temp_cooc (tag_chunk_id, context_chunk_id, pos_from_tag, partial_N)
+		      select L.chunk_id as tag_chunk_id, R.chunk_id as context_chunk_id, L.pos - R.pos as pos_from_tag, count(L.post_id) as partial_N
+		      from temp_post_tokenized as L
+		      join temp_post_tokenized as R
+		      on L.post_id = R.post_id
+		      where L.post_type_id = (select tokenized_types.id from tokenized_types where tokenized_types.type_name = 'tag')
+		      and R.post_type_id = (select tokenized_types.id from tokenized_types where tokenized_types.type_name = '%s')
[…]

### 2014-03-07 dissProject/preProcess.R
foo
Former-commit-id: 7e04a0705d1730f7d4a01f751a60932dca68c852 [formerly 29861232d44958f4e6e6751b711e1c24568d3334]
Former-commit-id: cd5176c362d2786b02529ea0cea45b9958829247

@@ -1076,7 +1076,7 @@ computeAct <- function(context, sjiTbl) {
 
 curWS <- function() {
 	runGenNcoocTblSO1thru3000000()
-	runGenNcoocTblSO1thru100()
+	runGenNcoocTblSO1thru10000()
 	fooTbl = myReadCSV('~/src/dissertation/firstSOProject/analysis/model/title-chunks-subset-4.csv')
 	userPTbl = withProf(getUserPTbl(defaultSOConfig))
 	userPTbl

### 2014-03-07 dissProject/preProcess.R
foo
Former-commit-id: e1f18f39a030e7e8e95934b60a96cb7c1abacc96 [formerly f99c4c5fade9ef9702fac363e0fe15ecc7e02c82]
Former-commit-id: 256bd846d96e4dac7d97c5cb18d3a48e4549baa0

@@ -936,6 +936,7 @@ getNcoocTbl <- function(type, chunkTableQuery) {
 				     on c.id = temp_cooc.context_chunk_id
 				     order by tag, chunk, pos_from_tag'
 				     ))
+	sqldf('truncate table temp_cooc')
 	resTbl[, pos_from_tag := NULL][, pos_from_tag := 'NaN'] # FIXME: Now that this has been regression tested against the prior R implementation, is this needed anymore?
 	NcoocTbl = resTbl[, list(posFromTag=pos_from_tag, partialN=partial_n, NChunkTag=sum(partial_n)), by=list(chunk, tag)]
 	setkey(NcoocTbl, chunk, tag, posFromTag)
@@ -962,6 +963,8 @@ genNcoocTblSO <- function(subsetName, startId, endId)  {
 	myWriteCSV(NcoocTblTitle, file=outFile)
 	outFile = sprintf('%s/NcoocTblBody-%s.csv', coocDir(), fullSubsetName)
 	myWriteCSV(NcoocTblBody, file=outFile) 
+	sqldf('truncate table temp_post_tokenized')
+	return()
 }
 
 genTokenizedTblSO <- function(filters='1=1', bundleSize=10000) {

### 2014-03-07 writeups/midwestCogsci2014/abstract.tex
foo
Former-commit-id: 19d7521d273acf99fa0e63130434f1ef9d574b1a [formerly 89684cc975529cbd1ca66cb17868792f749f4320]
Former-commit-id: 4bbb1f192dcb9dbab060dda25096cc8526862811

@@
 \begin{abstract}
-  This research explores and further validates ACT-R's base-level learning architectural component on two large-scale real-world datasets.
-  We framed the process where an author chooses tags and hashtags on StackOverflow posts and Twitter tweets as a declarative memory retrieval problem.
+  This research explores and further validates ACT-R's base-level learning mechanism on two large-scale real-world datasets.
+  We framed the process of choosing tags and hashtags on StackOverflow posts and Twitter tweets as a declarative memory retrieval problem.
   Over \num{265000} posts and 2.4 million tweets from \num{3900} total authors were collected. 
-  ACT-R's base-level learning equation and the simplified version of that equation (which assumes equally-spaced retrievals for each chunk) were used to predict each author's chosen hashtags, given their previous hashtag use.
-  The model performance space was explored from over 70 million total declarative memory retrievals, at 22 different levels of the decay rate parameter, and 48 different subsets of the two datasets.
-  The results show that for these datasets, the optimal value for the decay rate parameter for the simplified form of the equation is 0.4, slightly less than but close to the standard 0.5 value. 
-  However, the results also suggest that a higher (0.7) value is optimal when the standard form of the equation is used.
-  Further, the standard form can be implemented in a computationally efficient manner (20 ms per retrieval), and using the standard equation significantly improves model accuracy within subjects by 5\%.
-  Taken together, this research suggests that it may be advantageous to set the standard form of the base-level learning equation as the default in ACT-R,
-  and also increase the default value for the decay rate parameter to 0.7.
-  The results also highlight the importance of customizing model retrieval predictions based on each user's prior history, as model accuracy is still significant when a retrieval is based purely on history (i.e., no context).
+  ACT-R's base-level learning equation and the simplified version (which assumes equally-spaced retrievals for each chunk) were used to predict each author's chosen hashtags, given their previous hashtag use.
+  The model performance space was explored across 22 different levels of the decay rate parameter and 48 different dataset subsets.
+  The results show that for these datasets, the optimal value for the decay rate parameter for the simplified equation is 0.4, close to the standard 0.5 value. 
+  When using the standard equation, the optimal decay rate increases to 0.7 and model accuracy improves from 29\% to 34\%. 
+  Further, the standard equation can be implemented in a computationally efficient manner (20 ms per retrieval).
+  Taken together, this research suggests that when modeling retrievals on a broader range of tasks --- or at least similar tasks with a large amount of retrievals over an extended period of time --
+  it may be worthwhile to test if model accuracy improves after using the standard base-level learning equation and slightly increasing the decay rate.
+  The results also highlight the importance of customizing model retrieval predictions based on each user's prior history,
+  as tag retrieval accuracy is 34\% for these datasets when using only base-level learning and no contextual information.
 \end{abstract}

### 2014-03-07 dissProject/preProcess.R
foo
Former-commit-id: e8b43d4ddd175c5a4f484af3611b940f2ee79017 [formerly 411778d00dd78c7d42a0aa5c2eb6a9ce3a1a197c]
Former-commit-id: 21f554b5c6d76dc8c56b9140059f6ce60cee53cf

@@ -673,7 +673,7 @@ getSummaryStats <- function() {
 	unique(usersFromRunsTbl)
 	postsCntFromRunsTbl = rbind(tPostsCntFromRunsTbl, SOPostsCntFromRunsTbl)
 	postsCntFromRunsTbl[, sum(count), by=dataset]
-	modelVsPredTbl[predUsedBest == T][runNum == 2][, list(acc=median(acc), d=median(d), N=.N), by=list(DVName, datasetType)]
+	modelVsPredTbl[predUsedBest == T][runNum == 2][, list(acc=mean(acc), d=median(d), N=.N), by=list(DVName, datasetType)][DVName %in% c('topHashtagPost', 'topHashtagPostOL2')][, mean(acc), by=DVName]
 	modelVsPredTbl[hashtagUsedP == T][topHashtag == T][DVName == 'topHashtagAct'][, sum(totN), by=datasetType]
 	modelVsPredTbl[, .N, by=d]
 }

### 2014-03-07 writeups/midwestCogsci2014/abstract.tex
foo
Former-commit-id: d4c274ffaad9c5c9936a849c5296cf00f6d23ad7 [formerly 3ae3d27130b91d5e28fc7ec616979d40e36114ca]
Former-commit-id: 69c9f40551508ae7c6e968b237174a429fd0b05e

@@
   The results also highlight the importance of customizing model retrieval predictions based on each user's prior history,
-  as tag retrieval accuracy is 34\% for these datasets when using only base-level learning and no contextual information.
+  as tag retrieval accuracy is 34\% over the extensive tag space in these datasets when using only base-level learning and no contextual information.
 \end{abstract}

### 2014-03-07 writeups/midwestCogsci2014/abstract.tex
foo
Former-commit-id: d34a4741abe087636c52ca0f83cb98afcb9dfb5a [formerly 2972edc66d9938026b82c5c50804bb1768bb5d2a]
Former-commit-id: 4f53af8dda773cd749410f838271a8c25f4e2a0e

--- 2014-03-07 writeups/midwestCogsci2014/abstract.tex
@@
   We framed the process of choosing tags and hashtags on StackOverflow posts and Twitter tweets as a declarative memory retrieval problem.
-  Over \num{265000} posts and 2.4 million tweets from \num{3900} total authors were collected. 
-  ACT-R's base-level learning equation and the simplified version (which assumes equally-spaced retrievals for each chunk) were used to predict each author's chosen hashtags, given their previous hashtag use.
-  The model performance space was explored across 22 different levels of the decay rate parameter and 48 different dataset subsets.
-  The results show that for these datasets, the optimal value for the decay rate parameter for the simplified equation is 0.4, close to the standard 0.5 value. 
+  \num{265000} posts and 2.4 million tweets from \num{3900} total authors were collected. 
+  ACT-R's base-level learning equation and the simplified version (assumes equally-spaced retrievals for each chunk) were used to predict each author's chosen hashtags, given their previous hashtag use.
+  The model performance space was explored across 22 levels of the decay rate parameter and 48 dataset subsets.
+  The results show that for these datasets, the optimal decay rate parameter value for the simplified equation is 0.4, close to the standard 0.5 value. 
   When using the standard equation, the optimal decay rate increases to 0.7 and model accuracy improves from 29\% to 34\%. 
@@
   The results also highlight the importance of customizing model retrieval predictions based on each user's prior history,
-  as tag retrieval accuracy is 34\% for these datasets when using only base-level learning and no contextual information.
+  as tag retrieval accuracy over the extensive tag space in these datasets is 34\% when using only base-level learning and no contextual information.
 \end{abstract}

### 2014-03-07 writeups/midwestCogsci2014/abstract.tex
foo
Former-commit-id: 89b5e4c81f9ddec801f8766004d5a37ae5395730 [formerly 35c24e184933c7a42a087ce0a4c565c5b98717b7]
Former-commit-id: a16e2a7a380ccce7f55d6ae7777330988bdc4986

@@
   Further, the standard equation can be implemented in a computationally efficient manner (20 ms per retrieval).
-  Taken together, this research suggests that when modeling retrievals on a broader range of tasks --- or at least similar tasks with a large amount of retrievals over an extended period of time --
+  Taken together, this research suggests that when modeling retrievals on a broader range of tasks --- or at least similar tasks consisting of a large amount of retrievals over an extended period of time --
   it may be worthwhile to test if model accuracy improves after using the standard base-level learning equation and slightly increasing the decay rate.

### 2014-06-17 writeups/fb2014/abstract.tex
foo

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage[group-separator={,}]{siunitx}
\frenchspacing
\raggedright

\title{Abstract Submission for Fourth Annual Midwestern Cognitive Science Conference}

\author{Clayton Stanley (clayton.stanley@rice.edu) \\
  Department of Psychology, 6100 Main Street \\
  Houston, TX 77005 USA 
  \and Michael D. Byrne (byrne@rice.edu) \\
  Departments of Psychology and Computer Science, 6100 Main Street \\
  Houston, TX 77005 USA \\
}

\begin{document}

\maketitle

%TC:break Abstract
\begin{abstract}
  This research explores how large-scale behavioral datasets can be used to evaluate and improve models in declarative memory.
  An author's task of choosing hashtags when composing a tweet on Twitter and choosing tags when asking a question on StackOverflow can be framed as a declarative memory retrieval problem.
  Two state-of-the-art declarative memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random permutation vector-based model.
  Experiments were designed to
  [1] discover the best way to leverage a user's prior hashtag use as a future predictor,
  [2] improve the models so that each model's strengths are incorporated into the other,
  and [3] formally evaluate and compare each model's performance when recommending hashtags on Twitter and tags on StackOverflow.
  The results show that each user's own past behavior is a strong predictor of future behavior, and past behavior can be combined with context in a straightforward way for both the Bayesian and vector-based context models.
  Also both models improve after dealing with high frequency words in some way, but they differ in how best to handle these stop words.
  This shows that there is no single best way to deal with stop words when using a co-occurrence model, and the most appropriate method is dependent on the type of co-occurrence model that is used.
  Overall, since the modifications to the two memory models were aimed at the architectural level, and these modifications improved performance,
  they may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
\end{abstract}
%TC:break _main_

\end{document}

### 2014-06-17 writeups/fb2014/abstract.tex
foo

@@
 
-\title{Abstract Submission for Fourth Annual Midwestern Cognitive Science Conference}
+\title{Using Large Datasets to Evaluate and Improve Models of Declarative Memory}
 
@@
   Houston, TX 77005 USA 
-  \and Michael D. Byrne (byrne@rice.edu) \\
-  Departments of Psychology and Computer Science, 6100 Main Street \\
-  Houston, TX 77005 USA \\
 }
@@
 \begin{abstract}
-  This research explores how large-scale behavioral datasets can be used to evaluate and improve models in declarative memory.
+  This research explores how large-scale behavioral datasets can be used to evaluate and improve models of declarative memory.
   An author's task of choosing hashtags when composing a tweet on Twitter and choosing tags when asking a question on StackOverflow can be framed as a declarative memory retrieval problem.

### 2014-06-17 writeups/fb2014/bio/bio.tex
foo

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage[group-separator={,}]{siunitx}
\frenchspacing
\raggedright

\title{Bio}

\author{Clayton Stanley (clayton.stanley@rice.edu) \\
  Department of Psychology, 6100 Main Street \\
  Houston, TX 77005 USA 
}

\begin{document}

\maketitle

%TC:break Abstract
\begin{abstract}
  I am currently a PhD candidate in the Dr. Mike Byrne's Computer-Human Interaction Laboratory (CHIL) at Rice University.
  I have a mixed training background in mathematics, modeling, statistics, and human-computer interaction,
  after completing a BS in Physics at the Air Force Academy, an MA in Psychology at Rice, and 3 years in industry as a software engineer and cognitive scientist.
  Broadly I am interested in leveraging computational modeling and statistical techniques to better understand and model human behavior in human-machine systems.
  More specifically I am interested in domains in which large sets of user performance data for a particular task have been amassed.
  For my dissertation research, I am exploring how large-scale behavioral datasets like StackOverflow and Twitter can be used as a testbed to evaluate and improve models of declarative memory, and consequently, 
  improve how well the tagging models can predict each user's behavior. 
\end{abstract}
%TC:break _main_

\end{document}

### 2014-07-14 dissProject/preProcess.R
foo

@@ -3574,7 +3574,6 @@ runGenAndSaveCurWorkspaceg3s6 <- function() genAndSaveCurWorkspace(groupConfigG3
 runGenAndSaveCurWorkspaceg4s6 <- function() genAndSaveCurWorkspace(groupConfigG4S6)
 
 curWS <- function() {
-	# FIXME: Rerun context (s3 change, logregTbl change,  and name change from Hyman to Enthyman)
 	withProf(runContext20g1s6(regen='useAlreadyLoaded'))
 	runContext500g1s6(regen='useAlreadyLoaded', numRunsT=1, numRunsSO=1)
 	runContext500g1s6(regen=F, numRunsT=1, numRunsSO=1)

### 2014-07-14 writeups/dissertation/dissertation.tex
foo

--- 2014-01-16 writeups/proposal/proposal.tex
@@
 % If the toc is too detailed, try limiting the depth of printed sections in the TOC:
-%\setcounter{tocdepth}{2}
+\setcounter{tocdepth}{2}
 
@@
 
-\subsubsection{StackOverflow Dataset}
+\subsection{StackOverflow Dataset}
 
@@
 
-\subsubsection{Twitter Dataset}
+\subsection{Twitter Dataset}
 
@@
 
-\subsection{Pre-processing the Text}
+\subsection{Aquiring Datasets}
+
+\subsection{Tokenization}
 
@@
 
-\subsection{Stop Words and Entropy Weighting}
+\subsection{Handling Stop Words}
 
@@
 
-\subsection{Attenuating Twitter Spam Tweets}
-
-Although Twitter has an infrastructure to easily report and remove spam accounts, the spam that does get through can contain tweets with numerous hashtags.
-For example, some spam accounts generate tweets where all words in the tweet are hashtags, or generate multiple tweets with repeating content and hashtags.
-Feeding these spam tweets into statistical models that generate hashtag predictions can bias the models' chosen hashtags.
-
-\textcite{Mazzia2009} developed a novel approach to dealing with hashtag spam on Twitter.
-When training his model, it attenuates the importance of multiple tweets from the same user that contain the same hashtags.
-As a user generates more than 10 tweets that contain the same hashtag, all tweets afterwards that contain this hashtag will be attenuated so that the overall weight is no more than that for 10 tweets.
-When testing the model, only at most 1 post from a user is included in the test set, since it is much easier to suggest relevant hashtags to a spammer than it is for an actual user.
-This exact technique may not be used for this analysis, but certainly some form of attenuating or identifying Twitter spam tweets will be explored.
-
-\subsection{Experiments}
-
-Each proposed experiment aims to answer a specific research question in this study.
-The results from Experiment 1 will show how a user's prior hashtag use influences their likelihood of using that hashtag in the future, and test whether the observed data fit the ACT-R frequency/recency decay rate theory.
-The output from Experiment 1 will be a model of how a user's prior hashtag use influences the probability of using that hashtag in the future for both the Twitter and StackOverflow datasets.
-Experiment 2 will use the results from Experiment 1 and develop the best method to incorporate user's past tagging history in the random-permutations model, where user history is currently not being used.
-Experiment 3 will explore various methods for incorporating word order information into the ACT-R Bayesian model, where this information is currently not being used.
-The outputs of Experiment 2 and 3 are improved ACT-R and random-permutation models that have properly incorporated the strengths of each other: word-order strength from random permutations and user history from ACT-R.
-Finally, Experiments 4 and 5 will directly compare model accuracy between the two models and evaluate overall performance.
-Experiment 4 is focused primarily on hashtag prediction when no user prior history is available, while Experiment 5 will test how much performance improves when that history is available.
-
-\subsubsection{Preparation Experiment: StackOverflow Dataset Modifications}
-
-\textcite{Stanley2013} developed a model for the StackOverflow dataset that was based on all posts up to April, 2011.
-This model will be used for this research, but the dataset will be updated to the most recent release.
-When moving from the older dataset to the newest dataset, several performance assessments will be made.
-
-First, model performance will be compared between the newest dataset and the Apr' 11 dataset to ensure that performance is consistent.
-This is also a way to verify that the newest dataset was correctly pre-processed and loaded.
-
-Next, incremental improvements will be made to the pre-processing stages and performance will be compared after each change.
-For example, in \textcite{Stanley2013} the technique used to chunk the words does not ensure that tags appearing in the post remain as a single word unit.
-So a word like \emph{C\#} in a post was chunked as two separate words (\emph{C} and \emph{\#}).
-This reduces predictive power of the model since it muddles the highly-predictive co-occurrence observations where a tag appears in the post.
-So the chunking technique will be changed to take into account tags in the post, and its effect on performance will be evaluated.
-
-\subsubsection{Experiment 1: Analyze Recency for User-Generated Hashtags}
+%FIXME: Add paragraph for freq weighting
+
+\section{Results}
+
+\subsection{Past User Behavior}
 
@@
 
-\subsubsection{Experiment 2: Incorporating Recency Into Random Permutations}
+\subsubsection{Method}
+
+\subsubsection{Results}
+
+\subsection{Adding Offset Term}
+
+\subsubsection{Method}
+\subsubsection{Results}
+
+\subsection{Twitter Subsets}
+
+\subsubsection{Method}
+\subsubsection{Results}
+
+\subsection{Frequency vs Entropy vs Stoplist Weighting}
+
+\subsubsection{Method}
+\subsubsection{Results}
+
+\subsection{Dataset Size}
+
+\subsubsection{Method}
+\subsubsection{Results}
+
+\subsection{Amount of Compression for Random Permutation}
+
+\subsubsection{Method}
+\subsubsection{Results}
+
+\subsection{Method to Combine Prior and Context}
 
@@
 
-\subsubsection{Experiment 3: Incorporating Word Order Into ACT-R}
-
-How can word-order information be incorporated into the ACT-R Bayesian model?
-Currently ACT-R uses a word co-occurrence matrix that does not include word position.
-A co-occurrence is counted if a word and a tag co-occur anywhere in a tweet or post.
-However, for Twitter in particular, it may be the case that the few words immediately preceding hashtags are highly predictive, and the word position for those words should be included in the model.
-So a method to include word order into a Bayesian model will be developed.
-
-One method to include word order in a Bayesian co-occurrence matrix is to represent word position on a third, orthogonal dimension, and turn the matrix into a 3-dimensional array.
-So a co-occurrence is represented by a triplet: context word by word position relative to tag by tag.
-When a retrieval request for a tag is made, the strength of association between each word in context (and its position) and the tags in the co-occurrence array would be computed.
-That computation could be made tractable by storing only a small set of values on the order axis and binning over that dimension.
-
-For example, one could store only three word order values: near, medium, far, which is equivalent to storing three separate co-occurrence matrices.
-So when a retrieval request is made (e.g., when a user writes the \# character when composing a tweet), the strength of association for each word would be computed using the appropriate matrix.
-That is, the ``far'' matrix would be used for words far away from the cursor position, ``near'' for words just written, and ``medium'' for words mid distance from the current focus point.
-This method seems straightforward to implement, but it is slightly unsatisfying since information is lost when binning is used.
-It also does not scale, since an additional 2-dimensional sparse matrix must be added for every additional value stored on the word order dimension.
-So I will try to discover a better technique to incorporate word order information into the ACT-R Bayesian model than the one just described, but will fall back to this technique if a better method is not found.
-
-This updated model will then be evaluated on only the Twitter popular-hashtags dataset, since this dataset contains a large (yet not infinite) set of hashtags that occur in various positions throughout the tweet.
-This model will not be evaluated on the StackOverflow dataset since chosen tags for a StackOverflow post are not embedded and mixed within a post, so word-order information is unavailable and most likely unused.
-
-\subsubsection{Experiment 4: Top Hashtag Prediction}
-
-After incorporating the relative strengths of each model into the other, the accuracy of the two models will be compared directly.
-First both models will be evaluated on predicting tags based purely on context (i.e., no prior user history), using the popular-hashtags, popular-users, and StackOverflow datasets.
-The results will show the performance differences for the contextual component of each model.
-
-\subsubsection{Experiment 5: User-Customized Hashtag Prediction}
-
-Next the full models (prior and context components) will be tested on the Twitter popular-users dataset and the StackOverflow dataset.
-The popular-hashtags dataset will not be included in this analysis, since this subset was collected across the entire Twitter user population, so prior hashtag use is not available.
-This test will directly compare performance differences for each model when generating user-customized tag predictions.
-The main evaluation will be the amount of performance gain achieved when the models generate tags that are based on past user behavior compared to models that generate tags that are based purely on context.
-
-\subsection{Summary}
+\subsubsection{Method}
+\subsubsection{Results}
+
+\subsection{Word Order}
+
+\subsubsection{Method}
+\subsubsection{Results}
+
+\subsection{Impact of Prior vs Context}
+
+\subsubsection{Method}
+\subsubsection{Results}
+
+\subsection{User-customized Co-occurrence Matrix}
+
+\subsubsection{Method}
+\subsubsection{Results}
+
+\section{General Discussion}
+
+\section{Summary}
 

### 2014-07-14 writeups/dissertation/dissertation.tex
foo

@@
 
-\subsection{StackOverflow Dataset}
-
-The most recent quarterly-published StackOverflow dataset \parencite{DataDump2013} will be used to test the models on user-chosen tags for posts.
+% Started re-reading and re-writing here.
+
+\subsection{Datasets}
+
+The most recent quarterly published StackOverflow dataset \parencite{DataDump2013} was used to test the models on user chosen tags for posts.
 This dataset contains around \num{5.7} million posts and \num{34377} unique tags.
 
-\subsection{Twitter Dataset}
-
 Unlike StackOverflow, the entire Twitter dataset is not released to the public.
-Due to this and the sheer size of Twitter, subsets of the Twitter dataset must be collected.
-So two Twitter datasets will be collected for model evaluation: a popular-hashtags and popular-users dataset.
-The popular-hashtags dataset will contain approximately 2 million tweets where at least one of the 500 most popular hashtags are used.
-This dataset will be collected in real time until the tweet quota is reached.
+Due to this and the sheer size of Twitter, subsets of the Twitter dataset were collected.
+
+\subsubsection{Popular Users}
+
+A popular-users dataset was collected for both StackOverflow and Twitter.
+This dataset consists of all of the tweets from 24 different samples of the top users for each site.
+This dataset was collected by fetching all tweet history for each of the top users.
+It was used to explore the growth and decay characteristics of a hashtag for individual users.
+The temporal dynamics of ACT-R's decay rate model for a chunk were evaluated against this data.
+
+\subsubsection{Popular Hashtags}
+
+A additional popular-hashtags dataset was collected for Twitter.
 The purpose of this dataset is to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
-
-The popular-users dataset will contain all of the tweets from the top \num{2000} users on Twitter.
-This dataset will be collected by fetching all tweet history for each of the top users.
-It will be used to explore the growth and decay characteristics of a hashtag for individual users.
-The temporal dynamics of ACT-R's decay rate model for a chunk will be evaluated against this data.
+The dataset contains approximately three million tweets where at least one of the 400 most popular hashtags are used, and was collected in real time until the tweet quota was reached.
+
+Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
+It took around four days to collect a single dataset, and one collection was done each week.
+Multiple datasets were collected to ensure that the results weren't dependent on a particular sample of hashtags, and would generalize across different popular-hashtag datasets and different time periods.
 
@@
 
+\subsubsection{StackOverflow}
+
+The entire StackOverflow dataset is packaged and released quarterly to the public.
+So this package was downloaded and the database tables were imported into PostgreSQL using \textcite{so2db} importing tool.
+
+\subsubsection{Twitter}
+
+Twitter provides two types of APIs for aquiring data: A query-based and streaming interface.
+The \textcite{Tweepy} Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
+The query-based interface was used to collect all tweets for specific popular users.
+At most \num{3000} of the most recent tweets were collected for each user, since Twitter only allows access to this number of tweets for each user.
+
+The streaming interface was used to collect a sample of all of the tweets that included one of the top 400 hashtags at that time.
+Twitter's streaming API only allows at most 400 hashtags to be monitored at a time.
+Otherwise a larger number of hashtags would have been collected.
+Twitter allows the streaming interface to randomly sample about 1\% of all of the real-time data that is generated on Twitter.
+This sampling rate was adequate to collect around three million tweets within a few days.
+
 \subsection{Tokenization}
 
-The text from each dataset will be pre-processed using standard natural language processing techniques.
-The Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} will be used to tokenize (chunk) and lemmatize (convert to root form) each word.
-The standard chunking algorithm will be modified, so that it also ensures that Twitter hashtags and StackOverflow tag words (e.g., the word \emph{PHP} in a post) are not separated into multiple chunks.
-With the StackOverflow dataset, it is possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
-So tag words in the post and used tags for the post will be converted to their root tag when possible.
+A custom tokenization algorithm created by \textcite{Owoputi2013} was used to chunk the words in the text.
+This algorithm was specifically created for tokenizing Twitter content.
+However during testing it was found to be robust and well suited for StackOverflow content.
+The initial model results with the new tokenizer were qualitatively very similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
+This is most likely because the language for the samples taken for both sites is informal English,
+and it seams reasonable to think that tokenizing all of the informal slang, jargon, and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
+So the same tokenization algorithm was used for both StackOverflow and Twitter, since performance was similar for both algorithms on StackOverflow,
+and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.
+
+\subsection{Stemming}
+
+\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing 'ed' and 'ing' endings), but
+with the StackOverflow dataset, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
+This can be thought of as semantic stemming, where two words with identical meaning are converted to an agreed-upon canonical word.
+So tag words in the post and used tags for the post were converted to their root tag when possible.
 
@@
 \subsubsection{Method}
-
 \subsubsection{Results}
@@
 
-\subsection{Frequency vs Entropy vs Stoplist Weighting}
+\subsection{Stop Words}
 

### 2014-07-15 writeups/dissertation/dissertation.tex
foo

@@
 Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.
-\textcite{Sahlgren2008} removed stop words for the random-permutation vector-based model to handle this issue.
-\textcite{Jones2007} did as well for the circular convolution model, but also mentioned that an entropy weighting technique has also been used previously to attenuate (rather than remove) stop words.
-
-\textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset to reduce the weight of low-predictor stop words instead of removing them.
+
+Stop words are commonly removed by either filtering based on a pre-determined list or filtering based on the high-frequency words that occur in the dataset.
+\textcite{Sahlgren2008} tried both methods to remove stop words for the random-permutation vector-based model, and found that the data-driven approach
+(i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a pre-determined list.
+
+\textcite{Jones2007} also filtered the stop words for the circular convolution model,
+but also mentioned that an entropy weighting technique \parencite{Dumais1991} has been used previously to attenuate (rather than remove) stop words.
 This data-driven technique identifies stop words by calculating the discriminating power (i.e., entropy) of each word to predict any tag.
+Each word is then weighted (not removed) based on their discriminating power.
+
+\textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset.
 The result maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
 It also seems less arbitrary than explicitly identifying a list of stop words to remove from the analysis.
-This technique worked well for the StackOverflow model, and will be further explored for Twitter hashtag prediction.
-However, if it becomes computationally infeasible to compute this weighting for Twitter, then the stop words will simply be removed, as was done for previous research.
-
-%FIXME: Add paragraph for freq weighting
+
+All three stop word techniques (i.e., pre-determined list, high-frequency filtering, and entropy weighting) were explored for this analysis.
 
@@
 How does a user's prior hashtag use influence their likelihood of choosing that hashtag in the future?
-The ACT-R theory places strong constraints on the lifetime and growth characteristics of chunks in declarative memory.
-The more recently and frequently used chunks have higher prior activations, and are more likely to be needed in the future.
-It is an open question if this ACT-R theory of chunk growth and decay accurately reflects hashtag growth and decay for each user on Twitter and StackOverflow.
-
-The full StackOverflow dataset and Twitter popular-users dataset will be used to characterize tag growth and decay for a user on each site.
-Afterwards, the predicted results from the ACT-R theory will be compared against the observed data and evaluated.
-If the ACT-R theory of decay rate and frequency/recency affects fits the data reasonably well, then the prior activation components for the models developed in this research will be based on this theory.
-However, if the theory does not fit hashtag growth and decay for each user, then an alternative model will be proposed that is a better fit, and that model will be used instead.
+This question was tested by examing how well ACT-R's base-level activation component of declarative memory predicts the tags that authors choose, given each author's past tagging behavior.
+The model predicts that the more recently and frequently used chunks have higher prior activations, and are more likely to be needed in the future.
+
+\subsubsection{Base-level ACT-R model}
+
+The base-level activation component from the full ACT-R declarative memory model in Table \ref{tabACTRModel} is included in Table \ref{tabACTRBLLModel}.
+
+\begin{table}[!ht]
+  \caption{Base-level component of ACT-R declarative memory}
+  \label{tabACTRBLLModel}
+  {\tabulinesep=1.2mm
+    \begin{tabu}{ll}
+      \hline
+      Common Name &  Equation \\
+      \hline
+      Base Level Activation &	$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
+      Optimized Learning &	$B_{i} = log \frac{n}{1-d} - d * log L$
+      \hline
+    \end{tabu}
+  }
+\end{table}
+
+Both terms are methods for computing the base-level activation ($B_{i}$) of a chunk in declarative memory.
+The standard form has a time dependence and accounts for the fact that the rate of presentations for a given object can change over time.
+The optimized learning form assumes that this rate is constant.
+
+For both equatons, $i$ represents each chunk type (e.g., a word), $j$ is an observation for a chunk (e.g,.
+the 2nd time that ``a'' appears), and $n$ is the number of times each chunk type has been observed.
+Since the optimized learning form assumes a constant rate for each chunk type, it only has to keep track of the time since the first presentation of a chunk type $L$ and the total number of chunks observed for that type.
+It does not have to keep track of the time each specific chunk was observed.
+Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R common-lisp implementation.
+
+The decay rate parameter $d$ represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
+Higher values mean that chunk activation decays faster, which means that the most recently-presented chunks will have the highest activation, and the total number of times that a chunk type has been presented matters less.
+Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks types that have occured most frequently will have the highest activation,
+and activation depends less on the specific chunk types that have been observed in the recent past.
+
+The base-level ($B_{i}$) component in Table \ref{ACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{ModACTRModel}).
+\authorcite{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log odds computation of the probability that each chunk type is likely to be presented again, given past frequency.
+If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log n_{i}$,
+which is a nonlinear transformation of the $B_{i}$ used by \authorcite{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
+Both forms maintain the same rank ordering, so they should produce similar results.
+So the $B_{i}$ component used by \authorcite{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).
 
 \subsubsection{Method}
+
+The full StackOverflow dataset and Twitter popular-users dataset were used to characterize tag growth and decay for a user on each site.
+Each user's tag behavior from the popular-users dataset for each site was analyzed.
+
+Multiple samples were taken across a broad range of user types for each dataset to ensure that the model results weren't dependent on specific attributes in a single sample.
+For the StackOverflow popular-users dataset, a total of 12 slices of users were taken at different levels of reputation and total number of questions asked.
+500 users were sampled that had a reputation higher than \num{500}, \num{1000}, \num{5000}, \num{10000}, \num{50000}, and \num{100000}.
+100 users were sampled that had asked a total number of questions higher than \num{50}, \num{100}, \num{200}, \num{300}, \num{400}, \num{500}.
+
+For the Twitter popular-users dataset, 12 slices were also taken, this time at different levels of number of followers and total number of tweets.
+100 users were sampled that had a total number of followers higher than \num{1000}, \num{5000}, \num{10000}, \num{100000}, \num{1000000}, and \num{10000000},
+and a total number of tweets higher than \num{100}, \num{500}, \num{1000}, \num{5000}, \num{10000}, and \num{50000}.
+
 \subsubsection{Results}

### 2014-07-15 writeups/dissertation/dissertation.tex
foo

@@
 
-For both equatons, $i$ represents each chunk type (e.g., a word), $j$ is an observation for a chunk (e.g,.
-the 2nd time that ``a'' appears), and $n$ is the number of times each chunk type has been observed.
-Since the optimized learning form assumes a constant rate for each chunk type, it only has to keep track of the time since the first presentation of a chunk type $L$ and the total number of chunks observed for that type.
-It does not have to keep track of the time each specific chunk was observed.
+For both equatons, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g,.
+the 2nd time that ``a'' appears), and $n$ is the number of times each chunk has been observed.
+Since the optimized learning form assumes a constant presentation rate for each chunk,
+it only has to keep track of the time since the first presentation of a chunk $L$ and the total number of instances observed for that chunk. 
+It does not have to keep track of the time each specific chunk instance was observed.
 Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R common-lisp implementation.
@@
 The decay rate parameter $d$ represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
-Higher values mean that chunk activation decays faster, which means that the most recently-presented chunks will have the highest activation, and the total number of times that a chunk type has been presented matters less.
-Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks types that have occured most frequently will have the highest activation,
-and activation depends less on the specific chunk types that have been observed in the recent past.
+Higher values mean that chunk activation decays faster, which means that the most recently-presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
+Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks that have occured most frequently will have the highest activation,
+and activation depends less on the specific chunk that have been observed in the recent past.
 
 The base-level ($B_{i}$) component in Table \ref{ACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{ModACTRModel}).
-\authorcite{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log odds computation of the probability that each chunk type is likely to be presented again, given past frequency.
+\authorcite{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log odds computation of the probability that each chunk is likely to be presented again, given past frequency.
 If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log n_{i}$,
@@
 
+20 different values of the decay rate parameter were explored for each slice
+(0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.8, 2, 5, 10, 20).
+The values were chosen in order to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d=.5$), to a pure recency model ($d=20$).
+The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
+However, only values equal to or below 1 were examined for the optimized learning form since values above 1 for this form cause a bifurcation and are not psychologically interesting.
+
+Both the standard form and optimized learning form of $B_{i}$ were used to predict each user's specific tag history.
+To do this, a user's tag history was collected, and then each tag instance from oldest to newest was examined.
+For each of these tagging instances, the model generated an activation value for each tag based on that tag's frequency and recency of use. 
+This means, for example, that the model always missed characterizing the first observed tag for each user, since at that point the model does not have any history of tag use for that user.
+This also means that the tag space for each user is constrained and limited to the tags that each user has used in the past.
+
+At each tagging instance, the model has activation values for a set of tags.
+Model accurracy was assessed by rank ordering these activations and choosing the model's highest $N$ tags, where $N$ is the number of tags used for that specific post.
+The model's chosen $N$ tags were compared to the user's chosen $N$ tags, and if a model's chosen tag was in the set of user's chosen tags, then the prediction was marked as correct.
+Otherwise, it was marked as incorrect.
+
+This process was performed across all users in each of the 12 dataset slices for both the Twitter and StackOverflow popular-users datasets.
+It was repeated for each of the decay rate values and the two different forms of $B_{i}$.
+
 \subsubsection{Results}
+
+\paragraph{Visualizing Tag Use}
+
+To better illustrate the dynamics of each user's tagging behavior,
+profile plots for a single user for both StackOverflow and Twitter are included in Figures \ref{figHTByTime-520957} and \ref{figHTByTime-fashionista_com} respectively.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTByTime-520957-crop.pdf}}}
+  \caption{Tagging profile of a single user on StackOverflow}
+  \label{figHTByTime-520957}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTByTime-fashionista_com-crop.pdf}}}
+  \caption{Tagging profile of a single user on Twitter}
+  \label{figHTByTime-fashionista_com}
+\end{figure}
+
+The specific value of the hashtag used for both plots is nominal and not important for this analysis.
+What is interesting is how large the tagging space is for these users.
+There are bands of commonly-used tags for each user, but there are also instances where new tags are used, and a large percentage of tags used are not in the most-common bands.
+This helps illustrate how challenging modeling this task can be.
+The model's goal is to predict at each instance what tag will be chosen, but since the tag space is so large and varied, this task is quite challenging.
+
+Although the tag space for the Twitter user is smaller than the StackOverflow user in these plots, this is not a consistent trend.
+These plots were chosen from the set to also show that the amount of tag variation for individual users also varies: some use many tags and some use just a few.
+To accurately model human performance in this domain, the model must be able to make accurate predictions across all variations of user tagging behavior.
+
+\paragraph{Visualizing Model Performance}
+
+As a first method for understanding model performance, profile plots of model's predicted tags for individual users were overlaid on that user's actual tag use.
+These plots for StackOverflow and Twitter are included in Figures \ref{figHTMByTime-520957} and \ref{figHTMByTime-fashionista_com} respectively.
+These are the same plots shown in Figures \ref{figHTByTime-520957} and \ref{figHTByTime-fashionista_com} with model predictions overlaid in red.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
+  \caption{Model tagging profile of a single user on StackOverflow}
+  \label{figHTMByTime-520957}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
+  \caption{Model tagging profile of a single user on Twitter}
+  \label{figHTMByTime-fashionista_com}
+\end{figure}
+
+Also shown in these figures is how model predictions change as the decay rate parameter changes.
+The model changes less from tag to tag with lower decay rate values.
+This makes sense because lower decay rate values lead to the model weighting frequency more than recency in the activation calculation.
+These plots also suggest that model accuracy is best when a blend of frequency and recency is used (i.e., a decay rate between 0 and 20).
+This will be more formally tested when model accuracy across decay rate values is analyzed across all users in each dataset slice and all dataset slices.
+
+\paragraph{Model Performance Across Decay Rate Values}
+
+Moving up one level of analysis, model accuracy for all users in a single dataset slice was analyzed.
+A plot of the results for one of the 12 slices for both Twitter and StackOverflow are included in Figures \ref{figPriorSOQSliceDs} and \ref{figPriorTwitterSliceDs} respectively.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}}
+  \caption{Model performance for a single dataset slice for StackOverflow}
+  \label{figPriorSOQSliceDsStd}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}}
+  \caption{Model performance for a single dataset slice for StackOverflow}
+  \label{figPriorSOQSliceDsOL}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}}
+  \caption{Model performance for a single dataset slice for Twitter}
+  \label{figPriorTwitterSliceDsStd}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}}
+  \caption{Model performance for a single dataset slice for Twitter}
+  \label{figPriorTwitterSliceDsOL}
+\end{figure}
+
+Model accuracy at each decay rate value for each user in the dataset slice are included in the plots.
+Each user's mean accuracy was subtracted off in order to compute a normalized mean for that user.
+This was done so that the relative accuracy change for different decay rate values could be more easily visualized.
+The technique is a visualization to what is done when testing the main effect in a repeated measures design.
+
+One main effect in all of the plots is that there is an optimal decay rate value between 0 and 20 that produces the highest model accuracy.
+For the standard $B_{i}$ equation, model accuracy for a pure frequency model ($d=0$) is also worse than using a pure recency model ($d=20$), particularly for Twitter.
+It makes sense that a pure recency model for Twitter is relatively more accurate than for StackOverflow, as the hashtag lifetime for Twitter is much less than a programming-language tag on StackOverflow.
+
+When comparing the standard and optimized learning forms of the equation, it is apparent that a pure recency model does not work very well for optimized learning.
+The best-fit decay rate value for the optimized learning form is also not as clear and pronounced as it is for the standard form (less of a peak).
+And the relative accuracy gained from using a pure frequency based model to a blend of frequency and recency is less for the optimized learning form than the standard form of $B_{i}$.
+
+\paragraph{Best-fit Decay Rate Values}
+
+To identify the best decay rate value for the dataset slices,
+each user's best decay rate value was taken and the results are plotted in histogram form in Figures \ref{figPriorSOQHistDs} and \ref{figPriorTwitterSliceDs} respectively.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}}
+  \caption{Best-fit decay rate for a single dataset slice for StackOverflow}
+  \label{figPriorSOQHistDs}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}}
+  \caption{Best-fit decay rate for a single dataset slice for StackOverflow}
+  \label{figPriorSOQHistDsOL}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}}
+  \caption{Best-fit decay rate for a single dataset slice for Twitter}
+  \label{figPriorTwitterHistDs}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}}
+  \caption{Best-fit decay rate for a single dataset slice for Twitter}
+  \label{figPriorTwitterHistDsOL}
+\end{figure}
+
+First, the best-fit values are close to the 0.5 default value used in ACT-R when the optimized learning form is used.
+Comparing across optimized learning and the standard form, the best-fit values for the standard form are slightly higher for both the StackOverflow and Twitter dataset slices.
+Comparing best-fit values will be more thoroughly tested when looking at the values across all dataset slices.
+
+\paragraph{Aggregate Decay Rate and Accuracy}
+
+jjj
 

### 2014-07-15 dissProject/preProcess.R

@@ -1700,12 +1700,12 @@ analyzeTemporal <- function(modelVsPredTbl) {
 	user_screen_names = wrapQuotes(c('rickeysmiley','fashionista_com','laurenpope','mtvindia','officialrcti'))
 	user_screen_names = wrapQuotes(c('fashionista_com'))
 	runTbls = runPriorT(config=modConfig(defaultTSjiPConfig, list(accumModelHashtagsTbl=T,
-								      query=sprintf("user_screen_name in (%s)", user_screen_names))))
+								      query=sprintf("select user_screen_name from twitter_users where user_screen_name in (%s)", user_screen_names))))
 	plotTemporal(runTbls)
 	user_screen_names = wrapQuotes(c('520957','238260','413225','807325','521180'))
 	user_screen_names = wrapQuotes(c('520957','238260'))
 	runTbls = runPriorSO(config=modConfig(defaultSOSjiPConfig, list(accumModelHashtagsTbl=T,
-									query=sprintf("user_screen_name in (%s)", user_screen_names))))
+									query=sprintf("select id as owner_user_id from users where id in (%s)", user_screen_names))))
 	plotTemporal(runTbls)
 }
 

### 2014-07-15 writeups/dissertation/dissertation.tex
foo

@@
 Moving up one level of analysis, model accuracy for all users in a single dataset slice was analyzed.
-A plot of the results for one of the 12 slices for both Twitter and StackOverflow are included in Figures \ref{figPriorSOQSliceDs} and \ref{figPriorTwitterSliceDs} respectively.
+A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDs} and \ref{figPriorTwitterSliceDs} respectively.
 
@@
 
-\paragraph{Aggregate Decay Rate and Accuracy}
-
-jjj
+\paragraph{Aggregate Model Performance}
+
+Finally, model accuraccy was analyzed across all dataset slices for StackOverflow and Twitter.
+Mean accuracy for both sites is included in Figure \ref{figPriorAcc}. 
+The mean was computed by taking the mean of the mean model accuracy for each dataset slice.
+All confidence intervals are 95\% bootstrapped for the mean of the means for each slice.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc--crop.pdf}}}
+  \caption{Model accuracy}
+  \label{figPriorAcc}
+\end{figure}
+
+Accuracy for the standard form (0.35) is higher than the optimized learning form (0.30).
+This makes sense given that the standard form does not assume equal presentation rate for each chunk, and hashtags on Twitter in particular show trends of high rate of use and then shift to low rate of use.
+The standard form of $B_{i}$ is able to more accurately account for these trends.
+
+The best-fit decay rate values were also analyzed across all dataset slices.
+These results are included in Figure \ref{figPriorDecay}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
+  \caption{Best-fit decay rate}
+  \label{figPriorDecay}
+\end{figure}
+
+Decay rate values that produced the most accurate model performance are lower for the optimized learning form (0.43) of $B_{i}$ compared to the standard form (0.80).
+However, the optimal values for optimized learning are near 0.5, which lines up with default value used in ACT-R.
+So the decay rate values for the standard form are slightly higher than the default values used in ACT-R, but the values found when optimized learning is used line up nicely with the ACT-R defaults.
+
+\subsubsection{Discussion}
+
+Given the large tag space used by post authors on StackOverflow and Twitter, it was surprising that model performance is respectable even when only past tag history and no context is used.
+This speaks to the importance of not only taking into account overall past tagging history, but also to ensuring that the prior component of activation for each tag is customized to each user's specific past tagging history.
+
+Also apparent for these datasets is that the optimized learning form of $B_{i}$ is not as accurate as the standard form.
+Further, for the optimized learning form there is not much difference between using a pure frequency-based model and a model where frequency and recency are optimally blended.
+This makes sense given that the optimized learning equation in Table \ref{tabACTRBLLModel} collapses to a pure frequency model as the amount of time since the presentation of the first recorded chunk instance increases,
+and also since the timespan of recorded tag use for StackOverflow and Twitter is large (years).
+
+So it may be worthwhile to use the standard form instead of the optimized learning form when declarative memory is modeled across a broader range of tasks.
+However, this shouldn't be taken as a strong recommendation,
+since the performance benefits seen here could simply be due to the fact that the declarative memory retrievals span across a large period of time, which isn't the case for all tasks.
+Nonetheless, the only real reason not to use the standard form is due to computational resource limitations.
+Since this becomes less of an issue as time progresses, it may be worthwhile to reconsider the default value used in ACT-R.
+It may be the case that the optimized learning form of the equation is no longer computationally necessary.
+And if the standard form is used as the default (or simply tried in another experiment),
+the results from this study suggest that the optimal decay rate value should be set slightly higher (0.8) than the default when optimized learning is used (0.5).
 

### 2014-07-15 writeups/dissertation/dissertation.tex
foo

@@
       Base Level Activation &	$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
-      Optimized Learning &	$B_{i} = log \frac{n}{1-d} - d * log L$
+      Optimized Learning &	$B_{i} = log \frac{n}{1-d} - d * log L$ \\
       \hline
@@
 The base-level ($B_{i}$) component in Table \ref{ACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{ModACTRModel}).
-\authorcite{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log odds computation of the probability that each chunk is likely to be presented again, given past frequency.
+\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log odds computation of the probability that each chunk is likely to be presented again, given past frequency.
 If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log n_{i}$,
-which is a nonlinear transformation of the $B_{i}$ used by \authorcite{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
+which is a nonlinear transformation of the $B_{i}$ used by \citeauthor{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
 Both forms maintain the same rank ordering, so they should produce similar results.
-So the $B_{i}$ component used by \authorcite{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).
+So the $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).
 

### 2014-07-16 writeups/dissertation/dissertation.tex
foo

@@
 
-\section{Results}
-
-\subsection{Past User Behavior}
+\section{Past User Behavior}
 
@@
 
-\subsubsection{Base-level ACT-R model}
+\subsection{Base-level ACT-R model}
 
@@
 
-\subsubsection{Method}
+\subsection{Method}
 
@@
 
-\subsubsection{Results}
-
-\paragraph{Visualizing Tag Use}
+\subsection{Results}
+
+\subsubsection{Visualizing Tag Use}
 
@@
 
-\paragraph{Visualizing Model Performance}
+\subsubsection{Visualizing Model Performance}
 
@@
 
-\paragraph{Model Performance Across Decay Rate Values}
+\subsubsection{Model Performance Across Decay Rate Values}
 
@@
 
-\paragraph{Best-fit Decay Rate Values}
+\subsubsection{Best-fit Decay Rate Values}
 
@@
 
-\paragraph{Aggregate Model Performance}
+\subsubsection{Aggregate Model Performance}
 
@@
 
-\subsubsection{Discussion}
+\subsection{Discussion}
 
@@
 
-\subsection{Adding Offset Term}
+\section{Combining Predictors}
+
+The declarative memory models analyzed in this study consist of two main components: past behavior and context.
+Past tag use was isolated in the first study to see how well ACT-R's prior component of declarative memory (i.e., base-level learning) fits each user's tag use over time.
+However for the rest of the studies the past behavior and context declarative memory components were combined as well as looked at in isolation.
+
+\subsection{Methods}
+
+\subsubsection{Logistic Regression}
+
+In order to best combine these predictors, a logistic regression statistical technique similar to \textcite{Stanley2013} was used to determine the most optimal weights for each term.
+Logistic regression is used on these datasets in the following way.
+When a user creates a tag for a post, a retrieval request for the model is made.
+For each request, the model returns a set of tags (all tags that have been observed for that user in the past) and activations associated with each tag.
+An activation value is computed for each model component (e.g., prior tag use, context) for each tag.
+These values are recorded, and tag instances that match what the author actually tagged are marked as a 1, while all others are marked with a 0.
+This process is repeated across all posts where a user chooses a tag, and across all users in the dataset.
+Once all of the results are aggregated, a logistic regression is run where the optimal weights for each model component are found that maximize the model's ability to correctly label tags as 0 or 1 based on total activation.
+
+Model accuracy is evaluated by using the optimal weights, rank ordering all recorded tags by model activation using those weights,
+asking the model to tag the top N posts with a 1 (where N is the total number of recorded instances where an author chose a tag),
+and then comparing the model's chosen tags with the used tags for each post (i.e., looking at the ``hits'').
+This is equivalent to throttling the threshold in the logistic regression such that the total number of observations labeled as a 1 match the total number of recorded instances of tagging in the dataset.
+
+For this form of model accuracy, the model labels N posts with a 1 where the labeling is not constrained in any way within each post.
+That is, the labeling is relaxed across all posts, and there are no constraints where the model must label a specific number of 1s for each post based on the number of tags that the author used for that post.
+So model accuracy values found with the relaxed method are actually slightly conservative, since the model may be labeling more 1s for a post than the number of tags the author used for that post.
+
+However, this relaxed metric was used for the rest of the experiments that involve combining terms because the optimal weights found by the regression are based on using the relaxed metric.
+If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression wasn't optimized to the constrained form, and weights may change if that was done.
+The constrained form of model accuracy can not be used for the regression because there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation.
+The constrained form could be used in a regression if a custom form of the regression was implemented.
+However this was not persued since around \num{2500} total regressions were run on these datasets for two reasons:
+A custom implementation in R would not likely run in compiled C, and even if it was implemented in a compiled language it would be computationally intractible without a closed-form solution.
+
+\subsubsection{Adding Offset Term}
+
+This logistic regression method is a computationally efficient way to find the optimal weights for the predictors.
+However when the regression can not be constrained to label a specific number of 1s for each observation (as is the case here), to be used properly an offset term must be added as an additional model predictor.
+
+The offset term ensures that the activation for the small number of top tags for each post are equal across posts.
+This ensures that the logistic regression only labels a few 1s for each post.
+\textcite{Stanley2013} did this by subtracting off the mean total activation for the top 5 tags for each post.
+However this method is complicated since computing the toal activation for a tag requires already knowing the weights for each model component that are used to compute that activation.
+\citeauthor{Stanley2013} used an iterative approach to determine the weights used for the offset term, and kept iterating until those weights matched the weights returned after running the logistic regression.
+
+A simplified version of \citeauthor{Stanley2013}'s method was used for the results in this study.
+Instead of subtracting off the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
+So for the prior component, the activation value used in the logistic regression is the prior activation with the mean of the top 5 prior activations subtracted off of this value.
+The same process is used for the context component.
+Decoupling the offset term and computing an offset for each model component in isolation means that an iterative approach is no longer necessary,
+since the optimal weights of the terms are no longer needed when computing the offset.
+
+\subsubsection{Datasets}
+
+Model accuracy was computed on the popular-hashtags dataset for Twitter and a random sample of posts across the entire StackOverflow dataset.
+For each of the four Twitter popular-hashtags datasets, model retrievals were done across 15 runs and 500 posts within each run.
+For the StackOverflow dataset, 5 runs of 100 posts were used, since the results for StackOverflow are more stable and retrievals take more wall-clock time for this dataset due to the size of the co-occurrence matrix.
+All runs contained separate posts, and all posts used for runs were not used to generate the co-occurrence matrix used for the two retrieval models.
+
+\subsubsection{Computing Prior}
+
+For each randomly-sampled post for StackOverflow, the prior activation is based on that user's prior tagging history, and not on any global prior of specific tag frequency across users.
+This was done so that the prior component used for StackOverflow for both the popular-users and randomly-sampled subsets was the same.
+
+However, a custom user prior for the Twitter popular-hashtags dataset cannot be used, since the entire Twitter dataset cannot be downloaded.
+This dataset contains samples of a large set of various users that choose a specific set of hashtags being monitored.
+So in order to generate a custom user prior for each sample in the dataset, that user's entire tagging history must be accessed through the API.
+Since access rates are limited across that interface, it is infeasable to do this process for every user that appears in the Twitter popular-users dataset.
+So instead, the global prior tag history for all users in the Twitter popular-users dataset was used to compute prior activation values for each tag.
+
+\subsubsection{Co-occurrence Matrices}
+
+The context component for both the Random Permutation and Bayesian model requires a co-occurrence matrix of words and observed tags.
+Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
+The matrix for \num{3000000} posts was only used for Twitter, since the number of words in a StackOverflow post is much higher than Twitter on average,
+and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high-performance desktop, even with highly optimized representations.
+Already FIXME total co-occuring word tag observations are used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
+It is also the case that \num{1000000} posts for StackOverflow is already around $1/7$ of the total posts created on the site,
+so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrent matrix instead,
+especially since FIXME co-occurrences are already used with \num{1000000} posts.
+
+\section{Results for Combining Terms}
+
+\subsection{Adding Offset}
+
+Model accuracy relaxed across posts with and without the offset model component was evaluated for both the Bayesian and random-permutation model.
 
 \subsubsection{Method}
+
+The four Twitter popular-users datasets and the StackOverflow randomly-sampled dataset were used for the analysis.
+Logistic regressions were run both with and without the offset component added to each model.
+The results for each run for each dataset were collected and averaged.
+The results across the four Twitter datasets were collected together, and the average was done across all runs in all four datasets.
+
 \subsubsection{Results}
 
+The impact of adding the offset to each of the models for StackOverflow and Twitter are included in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetTwitter}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
+  \caption{Impact of adding model offset for StackOverflow}
+  \label{figContextOffsetSO}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
+  \caption{Impact of adding model offset for Twitter}
+  \label{figContextOffsetT}
+\end{figure}
+
+In both plots, if the model name includes ``full'' then all model components are included.
+When they are enumerated, then components are left out.
+The random-permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
+So, for example, the ``standard Prior model relaxed across posts w/o offset'' means that only the prior model component was used, and the offset for that component was not subtracted off when determining activation.
+As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.
+
+Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites.
+Accuracy is improved even when only a single term is used.
+Although in this case the weight of the term is irrelevant (only one term, weighting for that term doesn't matter),
+using an offset still improves accuracy since the offset term ensures that only a small number of tags per posts are predicted by the model to be the chosen tags.
+
+The random-permutation model in particular is greatly affected by the offset term, for both the StackOverflow and Twitter.
+This suggests that the mean activation for the random-permutation model for the top few tags in each post can vary greatly across posts.
+Only if this mean activation is subtracted off will the logistic regression model generate optimal weights and predictions that are comparative to the Bayesian model when model accuracy relaxed across posts is examined.
+
+\subsubsection{Discussion}
+
+It is actually somewhat dissapointing that the offset term is even needed.
+If it were the case that model activation for a tag in a specific post could be interpreted as an exact activation value that did not depend on the specific post,
+then the model would be able to say when, for example, the third top tag in a post is more likely to be a chosen tag than the second top tag in another post.
+That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of retrievals to do for a specific post.
+If this were the case, then adding the offset term would actually decrease accuracy, since this absolute value of the activation values would be subtracted off.
+However, since accuracy improves when adding the offset term for both models, it does not seem that activation values have much meaning across posts (and only within a single post) for these datasets.
+
+A likely and simple explaination for this is that possibly there are posts where a large set of tags have high activation,
+but the author is throttling their retrieval threshold for each post such that only a small set of tags are actually used.
+This is probably not the retrieval threshold used in the core declarative memory system that determines if a retrieval for the term actually happens.
+It is unlikely that this retrieval threshold can be directly controlled and manipulated.
+Rather, it could be a high-level task retrieval threshold that the author has control over, where he/she actively sets it based on competing goals of tagging posts an not overly saturating a post with tags.
+Adding the offset term into the model is a way to implement a simple version of this high-level retrieval threshold.
+
 \subsection{Twitter Subsets}
 
+Four subsets of the popular-users dataset were collected for Twitter.
+This was to ensure that model results could be interpreted as effects that were not specific to a certain time period or set of popular hashtags.
+
 \subsubsection{Method}
+
+The four subsets were collected across a period of one month.
+The hashtags used in each subset were chosen by scraping a site that published the current trending hashtags for each city.
+Hashtags for each US city were collected, and then 400 were randomly sampled from this set.
+Only 400 were selected since this is the maximum that the Twitter API can monitor at one time.
+
 \subsubsection{Results}
+
+A plot of model accuracy for components of both the Random Permutation and Bayesian models and isolated by subset are included in Figure \ref{figContextSubsets}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-contextStandardByGroupT-crop.pdf}}}
+  \caption{Model accuracy for each of the four Twitter popular-users subsets}
+  \label{figContextOffsetT}
+\end{figure}
+
+Although there is a main effect of subset, there are no practically significant interactions.
+Relative model performance remains consistent across the four subsets.
+This means that model behavior for the popular-users subset for Twitter can be interpreted as behavior that will persist across different time periods,
+if the same method of choosing popular hashtags is done at a later date.
+It may be more general that model behavior also does not depend on the specific method of choosing the popular hashtags, but since only a single method was used for this analysis, this is still an open question.
 

### 2014-07-16 dissProject/preProcess.R

@@ -1703,7 +1703,7 @@ analyzeTemporal <- function(modelVsPredTbl) {
 								      query=sprintf("select user_screen_name from twitter_users where user_screen_name in (%s)", user_screen_names))))
 	plotTemporal(runTbls)
 	user_screen_names = wrapQuotes(c('520957','238260','413225','807325','521180'))
-	user_screen_names = wrapQuotes(c('520957','238260'))
+	user_screen_names = wrapQuotes(c('520957'))
 	runTbls = runPriorSO(config=modConfig(defaultSOSjiPConfig, list(accumModelHashtagsTbl=T,
 									query=sprintf("select id as owner_user_id from users where id in (%s)", user_screen_names))))
 	plotTemporal(runTbls)
@@ -1736,8 +1736,9 @@ analyzePrior <- function(modelVsPredTbl) {
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorStd' & datasetName=='SOQgt500r2'])
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorOL2' & datasetName=='TFollowgt10Mr2' & d < 1])
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorOL2' & datasetName=='SOQgt500r2' & d < 1])
-	compareOptimalDs(bestTbl[DVName %in% c('topHashtagPostPriorStd', 'topHashtagPostPriorOL2') & runNum == 2])
-	compareOptimalAcc(bestTbl[DVName %in% c('topHashtagPostPriorStd', 'topHashtagPostPriorOL2') & runNum == 2])
+	tbl = compareOptimalDs(bestTbl[DVName %in% c('topHashtagPostPriorStd', 'topHashtagPostPriorOL2') & runNum == 2])
+	tbl = compareOptimalAcc(bestTbl[DVName %in% c('topHashtagPostPriorStd', 'topHashtagPostPriorOL2') & runNum == 2])
+	tbl[, mean(meanVal), by=DVName]
 }
 
 asTopHashtagAcross <- function(vect) {
@@ -1901,6 +1902,10 @@ analyzeContext <- function(modelHashtagTbls, modelVsPredTbl) {
 				       'PriorStdTweetNentropy', 'PriorStdNoffsetTweetNoffset',
 				       'PriorStdNoffsetTweetOrderNoffsetTweetOrderlessNoffset', 'PriorStdTweetOrderTweetOrderless'))
[…]

### 2014-07-16 writeups/dissertation/dissertation.tex
foo

@@
 
-A plot of model accuracy for components of both the Random Permutation and Bayesian models and isolated by subset are included in Figure \ref{figContextSubsets}.
+A plot of model accuracy for components of both the Random Permutation and Bayesian models and isolated by Twitter popular-users subset are included in Figure \ref{figContextSubsets}.
 
@@
   \caption{Model accuracy for each of the four Twitter popular-users subsets}
-  \label{figContextOffsetT}
-\end{figure}
-
-Although there is a main effect of subset, there are no practically significant interactions.
+  \label{figContextSubsets}
+\end{figure}
+
+Although there is a main effect of subset, there are minimal to no practically significant interactions.
 Relative model performance remains consistent across the four subsets.
-This means that model behavior for the popular-users subset for Twitter can be interpreted as behavior that will persist across different time periods,
+
+\subsubsection{Discussion}
+
+Since the interactions across popular-users datasets are minimal, model behavior for the popular-users subset for Twitter can be interpreted as behavior that will persist across different time periods,
 if the same method of choosing popular hashtags is done at a later date.
-It may be more general that model behavior also does not depend on the specific method of choosing the popular hashtags, but since only a single method was used for this analysis, this is still an open question.
+It may be more general that model behavior also does not depend on the specific method of choosing the popular hashtags, but since only a single method was used for this analysis,
+the results from this study cannot by themselves be used to answer this question.
 
@@
 
+\textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random-permutation model:
+A data-driven frequency-weighted approach was tried, along with using a previously-derived stoplist, and no stop word removal at all.
+Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and uing a previously-derived stoplist hardly increased performance.
+However, \citeauthor{Sahlgren2008} did not try to weight the stop words, and only tried methods to remove them.
+So alongside looking at frequency weighting, previously derived, and no stop word removal, \textcite{Stanley2013}'s method of attentuating stop words based on Entropy was also explored.
+
 \subsubsection{Method}
+
+The runs from the Twitter popular-users and StackOverflow randomly-sampled datasets were used for the analysis.
+Both the random permutation and Bayesian models were tested.
+Activations for context were computed by using the two methods for stop word removal, the entropy weighting method, and no method (i.e., no stop word processing).
+
+\paragraph{Determing Frequency Cutoff}
+
+The same previously-derived stoplist from FIXME that \textcite{Sahlgren2008} used was used for this analysis,
+The same entropy weighting method using by \textcite{Stanley2013} was examined,
+and the same method of removing the high-frequency words from \citeauthor{Sahlgren2008} was used.
+
+However a cutoff point must be identified and used in order to remove the high-frequency words.
+\textcite{Sahlgren2008} used a cutoff where words that occurred more than \num{15000} times were removed.
+Since the StackOverflow and Twitter datasets are much larger than the datasets used by \citeauthor{Sahlgren2008}, the same cutoff could not be used.
+The ideal cutoff was determined by choosing the value where the standard deviation of the counts for each word for each of the tags were all relatively small.
+That is, the spikes were suppressed for standard deviations for particular tags that still contained words with a very large number of observations.
+The cutoff was also chosen such that the size of the spikes was about as large as the spikes generated when the entropy method was used.
+
+The plot of the standard deviations across tags that was used to identify the cutoff for StackOverflow and Twitter is included in Figures \ref{figContextCutoffSO} and \ref{figContextCuttoffT}.
+
+FIXME:
+
+The no-removal plot shows the size of the spikes when high-frequency words are removed.
+The entropy plot shows how the size decreases by more than an order of magnitude when the entropy weighting measure is used.
+The cutoff was chosen such that the frequency plot looks similar to the entropy plot, with spikes around the same size.
+Note that using a previously-defined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
+The cuttoff that produced the frequency plot is where a word represents more than \num{.0004} occurrences in the dataset.
+This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset were not given.
+
+Also, no strong claims are being made that the cutoff identified is the ideal frequency percentage cutoff for all datasets when a frequency cutoff is used.
+The method used to identify this cutoff was simply aimed to find a reasonable value where comparisons between a frequency cutoff and entropy weighting techniques would be equitable.
+
 \subsubsection{Results}

### 2014-07-16 dissProject/preProcess.R

@@ -1902,9 +1902,9 @@ analyzeContext <- function(modelHashtagTbls, modelVsPredTbl) {
 				       'PriorStdTweetNentropy', 'PriorStdNoffsetTweetNoffset',
 				       'PriorStdNoffsetTweetOrderNoffsetTweetOrderlessNoffset', 'PriorStdTweetOrderTweetOrderless'))
 	compareMeanDVDefault(tbl[sizeNum == 2 & dsetType == 'twitter' & DVName %in% DVNames], acc, figName='OffsetT')
-	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweetNentroy', 'PriorStdTweetOrderTweetOrderless'))
 	# RESULT: 4 different subsets for twitter popular hashtags show same effects
-	# T standard all groups
+	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweetNentropy', 'PriorStdTweetOrderTweetOrderless',
+				       'TweetNentropy', 'TweetOrderTweetOrderless'))
 	compareMeanDVDefault(tbl[sizeNum == 2 & dsetType == 'twitter' & DVName %in% DVNames], acc, figName='contextStandardByGroupT', groupCol='groupNum')
 	# RESULT: Show base performance for both models
 	# SO standard

### 2014-07-18 writeups/dissertation/dissertation.tex
foo

@@
 
-The four Twitter popular-users datasets and the StackOverflow randomly-sampled dataset were used for the analysis.
+The four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset were used for the analysis.
 Logistic regressions were run both with and without the offset component added to each model.
@@
 
-The runs from the Twitter popular-users and StackOverflow randomly-sampled datasets were used for the analysis.
+The runs from the Twitter popular-hashtags and StackOverflow randomly-sampled datasets were used for the analysis.
 Both the random permutation and Bayesian models were tested.
@@
 
-The plot of the standard deviations across tags that was used to identify the cutoff for StackOverflow and Twitter is included in Figures \ref{figContextCutoffSO} and \ref{figContextCuttoffT}.
+The plots for the standard deviations across tags that was used to identify the cutoff for StackOverflow and Twitter are included in Figures \ref{figContextCutoffSO} and \ref{figContextCuttoffT}.
 
@@
 
+\subsubsection{Results for the Random Permutation Model}
+
+Plots for the three methods of handling stop words for the random-permutation model on the StackOverflow dataset are included in Figure \ref{figContextStopRPSO}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsSOPerm-crop.pdf}}}
+  \caption{Stop-word techniques for the random-permutation model for StackOverflow}
+  \label{figContextStopRPSO}
+\end{figure}
+
+The results show that model accuracy improves if stop words are handled in any of the three ways (pre-defined stoplist filtering, entropy weighting, and frequency filtering).
+Using a pre-defined stoplist makes the smallest improvment, while the entropy and frequency techniques improve performance the most, and frequency shows a slight edge over entropy.
+However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
+
+The three stop word techniques were then tested for the random-permutation model on the Twitter dataset, and the results are included in Figure \ref{figContextStopRPT}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTPerm-crop.pdf}}}
+  \caption{Stop-word techniques for the random-permutation model for Twitter}
+  \label{figContextStopRPSO}
+\end{figure}
+
+For the random-permutation model for Twitter, the entropy weighting technique is the best method for handling stop words.
+Removing stop-words based on a pre-determined list actually decreases accuracy, and using data-driven frequency filtering has only a small to no effect.
+
+\subsubsection{Discussion}
+
+Comparing across Figures \ref{figContextStopRPSO} and \ref{figContextStopRPT}, entropy and frequency techniques are superior to filtering based on a pre-determined list.
+However it is more difficult to choose the best approach when comparing entropy weighting and frequency filtering.
+Frequency filtering is slightly better for StackOverflow, while entropy weighting is slightly better for Twitter.
+However frequency filtering does have a few drawbacks compared to entropy weighting:
+It produces accuracy measurments that have higher levels of noise (see Figure \ref{figContextStopRPSO} and frequency models for the title component in StackOverflow).
+Also, the ideal cutoff threshold must be identified in order to filter based on frequency,
+and identifying this cutoff requires a search, and this search must be done for each dataset that frequency filtering is used on.
+This threshold is a free parameter in frequency filtering that is not required when weighting stop words by entropy.
+Therefore, favoring parsimony, for the random-permutation model, attentuating stop words based on entropy produces more accurate results when compared to filtering the stop words based on frequency of use.
+This result is interesting since filtering stop words based on frequency is the default method for random permutation models \parencite{Sahlgren2008}.
+This suggests that a simpler method (entropy weighting) can be used instead of the more common frequency-filtering technique to achieve the same (or better) accuracy measurements.
+
+\subsubsection{Results for the Bayesian Model}
+
+Since accuracy measurements when using a pre-determined list to remove stop words for the random-permutation model were much lower than the frequency and entropy methods, this method was not examined for the Bayesian model.
+The results for the frequency and entropy techniques for the Bayesian model on the StackOverflow dataset are include in Figure \ref{figContextStopBSO}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsSOSji-crop.pdf}}}
+  \caption{Stop-word techniques for the Bayesian model for StackOverflow}
+  \label{figContextStopBSO}
+\end{figure}
+
+The results show that entropy weighting produces the most accurate results.
+Also, accuracy actually decreases when using frequency filtering (compared to no filtering) for the Bayesian model.
+So entropy weighting is the clear winner in this case.
+
+Similar results are found when the Bayesian model is tested on the Twitter dataset.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTSji-crop.pdf}}}
+  \caption{Stop-word techniques for the Bayesian model for Twitter}
+  \label{figContextStopBSO}
+\end{figure}
+
+Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer, and frequency filtering decreases performance relative to no filtering.
+
+\subsubsection{Discussion}
+
+One result that should not be overlooked is how poorly using a pre-determined stoplist performed when compare to the two data-driven frequency and entropy approaches.
+This is most likely because the StackOverflow and Twitter datasets are large and contain high-frequency words that are domain specific (e.g., emoticons) and are not included in the pre-determined list.
+
+Another interesting result was how the Bayesian model actually performed worse when stop words were removed based on frequency.
+This was not the case for the random-permutation model, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random-permutation model, and is most likely never the case for the random-permutation model.
+Model accuracy most likely decreases for the Bayesian model because this model already has a normalization component built in that the random-permutation model does not have.
+When computing activaiton for the random-permutation model, the corrollation is calculated directly on a matrix of counts of co-occurrences for words and tags.
+However for the Bayesian model this count matrix is converted to a log odds ($S_{ji}$) matrix,
+where both the number of observations for the word and the tag are normalized when computing the value for each cell (see the $S_{ji}$ computation in Table \ref{tabModACTRModel}).
+So it is likely the case that a form of frequency \emph{attentuation} is already built into the Bayesian model.
+And therefore adding a frequency filter on top of this does nothing to improve accuracy, and only removes predictive information.
+
+\textcite{Stanley2013} found that the entropy-weighting technique to attentuate stop words for the Bayesian model improved model accuracy.
+The same effect is found for the StackOverflow and Twitter datasets in this study.
+Further, when this entropy-weighting technique is compared directly to the more common frequency-filtering technique, weighting by entropy produces more accurate results.
+This was clearly the case for the Bayesian model.
+Both frequency and entropy techniques performed equally for the random-permutation model.
+However since the entropy-weighting technique is parameter free and does not require a cutoff to tune, it is also a better choice for the random-permutation model for these datasets.
+
+The effect sizes when comparing no stop-word removal technique to either the entropy or frequency techniques are quite large.
+This shows the importance of correctly identifying and handling stop words when building co-occurrence matrices for large-scale natural-language datasets.
+For further research, it may be worthwhile to explore an even larger range of stop-word handling techniques,
+since model accuracy seems to be quite sensitive to the different techniques. 
+However, one conclusion that can be made from this research is that the entropy-weighting technique is a contender, and should be explored much more often than it has been previously.
+
+\subsection{Dataset Size}
+
+The default number of documents used to build the co-occurrence matrix was \num{1000000} for StackOverflow and \num{3000000} for Twitter.
+In order to see how accuracy changed as function of document size, 
+co-occurence matrices for a range of smaller document sizes were built.
+Both the random-permutation and Bayesian models were tested on all document sizes.
+
+\subsubsection{Method}
+
+For StackOverflow, the following number of posts were used to build separate co-occurrence matrices: \num{1000}, \num{10000}, \num{100000}, and the default \num{1000000}.
+For Twitter, these matrix sizes were tested alongside the default \num{3000000}.
+The four Twitter popular-hashtags datasets and the randomly-sampled StackOverflow dataset were used to test model accuracy.
+
+\subsubsection{Results for StackOverflow}
+
+Model accuracy as a function of co-occurrence matrix size for StackOverflow is included in Figure \ref{figContextDSizeSO}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
+  \caption{Effect of size of co-occurrence matrix for StackOverflow}
+  \label{figContextDSizeSO}
+\end{figure}
+
+Accuracy improves for both the Bayesian and random-permutation models as the size of the documents used to generate the co-occurrence matrix increases.
+However model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random-permutation model, while the frequency technique only begins to plateau.
+The Bayesian model does not plateau and accuracy continues to improveas document size increases.
+
+\subsubsection{Discussion}
+
+It is likely that the random-permutation model is saturating when entropy weighting is used for the largest document size for StackOverflow.
+Even if the non-predictive words based on entropy are attenuated,
+when the high-frequency words are included random-permutatoin model,
+this attentuation is not enough to overcome the range effects that grow with document size.
+These effects end up saturating all of the correlations near one, and consequently decrease discriminability and accuracy.
+However, the effect of saturation is small compared to the large effect from using either frequency filtering or entropy weighting to handle the stop words.
+The Bayesian model does not saturate when using entropy weighting because each cell in the co-occurrence matrix is normalized by word frequency and tag frequency.
+
+\subsubsection{Results for Twitter}
+
+The overall trend of increasing accuracy with document size is present for Twitter as well.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
+  \caption{Effect of size of co-occurrence matrix for Twitter}
+  \label{figContextDSizeSO}
+\end{figure}
+
+Both entropy weighting and frequency filtering in Figure \ref{figContextDSizeSO} plateu for the random-permutation model with increasing document size.
+Once again, the Bayesian model continues to improve with document size.
+
+\subsubsection{Discussion}
+
+It is interesting to compare model accuracy betwen the Bayesian and random-permutation model at different document sizes for both StackOverflow and Twitter.
+At smaller document sizes, the random-permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as document size increases.
+In other words, the Bayesian model needs larger co-occurrence matrices than the random-permutation model to work properly.
+As the size of the dataset increases, the compression may start to lose crucial information, and the uncompressed Bayesian model should be used.
+As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random-permutation model is actually beneficial, and the compressed random-permutation model should be used.
+This crosstalk and noise is beneficial because it helps soften spurious $S_{ji}$ values when the values for each individual cell in the co-occurrence matrix are still volitle, which happens at smaller document sizes.
+
+This result has important implications.
+The document size is rarely chosen, and more likely constrained by the number of documents in the dataset being studied.
+If that dataset is relatively small, it may be better to use the compressed random-permutation model than the Bayesian model, and vice-versa if the dataset is large.
+
+\subsection{Amount of Compression for Random Permutation}
+
+Another parameter in the random-permutation model is the number of rows used to represent the all of the words.
+Each word is represented by two randomly placed ones and two negative ones across the rows.
+All that is required is that the four positions and signs of each of these environment vectors is unique.
+But this means multiple environment vectors can overlap when looking at a single row.
+This is a lossy form of compression, and allows the number of rows to represent the words to be much less than the total number of words.
+As the number of rows decreases, the amount of cross talk between the environment vectors increases, compression increases, and the amount of information stored decreases.
+
+\subsubsection{Method}
+
+Three different row-size values were tested for the random-permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
+A low value (\num{100}), a more standard value (\num{2048}), and a high value (\num{10000}) of rows were tested.
+A separate co-occurrence representation was built for each of these row values.
+Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
+Accuracy measurments were the same as were used for previous runs of these datasets: mean of the mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.
+
 \subsubsection{Results}
 
-\subsection{Dataset Size}
+Results for using different levels of compression for the random-permutation model for the StackOverflow dataset are included in Figure \ref{figContextDimSO}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
+  \caption{Effect of compression on random-permutation model for StackOverflow}
+  \label{figContextDimSO}
+\end{figure}
+
+In the figure, the full random permutation model that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
+This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random-permutation model for all other results.
+The plot is also broken out for different size datasets used to build the co-occurrence matrix.
+Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy.
+However \num{2048} rows is sufficient for these dataests, as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy.
+
+Results are similar for Twitter.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeT-crop.pdf}}}
+  \caption{Effect of compression on random-permutation model for Twitter}
+  \label{figContextDimT}
+\end{figure}
+
+Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100}, and little to no impact when increasing from \num{2048} to \num{10000}.
+This result is consistent across all document sizes used to build the co-occurrence matrix.
+
+\subsubsection{Discussion}
+
+The fact that the random-permutation model only needs around \num{2048} rows to represent all of the words in the Twitter and StackOverflow datasets is impressive.
+Twitter has around FIXME unique words for the \num{3000000} tweets in each of the popular-users datasets.
+Stackoverflow has FIXME unique words for the \num{1000000} posts used to generate the co-occurrence matrix.
+So it is impressive that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
+This is likely because word use frequency drops off sharply, and there are many words in both datasets which are used only once or a few times.
+Since these words have little to no ability to influence predictions, it makes sense that they can be safely removed from the dataset.
+
+However it should be pointed out that the way the random-permutation model compresses the words is completly random.
+It is not an LSA approach where the few ones and negative ones for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
+Everything is randomly chosen with the random-permutaton model.
+The environment vector for each word is randomly selected, and no SVD is performed to choose these in a more discriminating manner.
+So it is that more impressive that even with a randomly-chosen form of compression, only \num{2048} rows are needed to represent all of the words in these datasets.
+
+\subsection{Method to Combine Prior and Context}
+
+The activation value for the context component of the vector-based model is a correlation, ranging from zero to one.
+This is not a log odds value.
+Currently, prior and context for the random-permutation model are combined by simply adding the terms.
+However, this means that ACT-R's prior base-level learning component (which is a log odds value) is added to a correlation value, which may not be mathematically appropriate.
+It would be more appropriate to convert the correlation value to a log-odds value, and then add this context term to the log-odds prior term.
 
 \subsubsection{Method}
+
+For a given retrieval, the random-permutation model returns a set of activations that are correlation values for each of the tags.
+So a method was identified to convert a distribution of correlations to a distribution of log-odds values.
+To do this, the one-tailed cumulative distribution function value for each correlation is computed.
+That probability is then converted to a log odds value in the usual way.
+
+This technique has a few nice mathematical properties:
+The conversion technique is simple and computationally cheap.
+Also, it behaves properly at specific points in the distribution.
+For example, if the conversion is performed one the mean correlation value in the set, the result is a log odds value of zero.
+This makes sense given that log odds can be interpreted as the amount of odds information (positive or negative) that should be adjusted to the prior log odds.
+Since the mean correlation value contains no information regarding if it is better or worse than the other correlation values, it is appropriate to assign this correlation value a zero log odds adjustment value.
+
+The log-odds transformation technique was tested for the random-permutaiton model on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.
+
 \subsubsection{Results}
 
-\subsection{Amount of Compression for Random Permutation}
-
-\subsubsection{Method}
-\subsubsection{Results}
-
-\subsection{Method to Combine Prior and Context}
-
-How can prior likelihood of a user choosing a particular tag be incorporated in the random permutations vector-based model?
-Currently vector-based models generate activations for each tag that are based purely on context.
-That is, if a particular tag has been used 10 times more often than the other tags by a specific user, that information is not included in a vector-based model.
-It seems plausible that performance should improve for vector-based models if this prior tag activation is included.
-Adding a prior activation term for the Bayesian StackOverflow model certainly improved performance \parencite{Stanley2013}.
-So a method to incorporate a user's prior tag history into a random permutation model will be developed and then tested on the full StackOverflow dataset and Twitter popular-users dataset.
-
-One straightforward way to include a user's prior tag history into a random permutation model is to include ACT-R's base-level activation term ($B_{i}$) as an additional term in the model.
-However, this may not be the most mathematically appropriate thing to do, since base-level activation is represented as a log odds value,
-and the single strength of association term in a random permutation model is computed by a correlation.
-It may be more appropriate to somehow attenuate the memory vectors over time,
-so that an association between a word in a post and a tag (i.e., a memory vector) that has long passed is no longer included in that memory vector's representation.
-I will try these ideas and any others that are discovered during the model development process, and the one that is the most cognitively plausible and accurate will be chosen.
-
-\subsubsection{Method}
-\subsubsection{Results}
+Model accuracy results after using the log-odds transformation are included in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
+  \caption{Random-permutation model accuracy when using log-odds trasformation on context for StackOverflow}
+  \label{figContextLogoddsSO}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
+  \caption{Random-permutation model accuracy when using log-odds trasformation on context for Twitter}
+  \label{figContextLogoddsT}
+\end{figure}
+
+Using the log-odds transformation technique has little to no effect on model accuracy for both StackOverflow and Twitter.
+When the technique is used and accuracy is examined for a model with only a single predictor, the transformation does not change accuracy.
+When predictors are combined and the full model is examined, the transformation is in the direction of slightly increasing accuracy, at least for the StackOverflow dataset.
+However any improvement when using the transformation is minimal.
+
+\subsubsection{Discussion}
+
+Vector-based models such as BEAGLE \parencite{Jones2007} and the random-permutation model \parencite{Sahlgren2008} have included only a context component in the past (i.e., no prior component).
+When the prior component is included for the random-permutation model, accuracy improves significantly (compare accuracy for full models to context models in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}).
+However, it is somewhat surprising that it does not matter much how the context and prior terms are added for the random-permutation model.
+Model performance is similar when the context component used for the model is based on a correlation compared to a log odds value.
+
+Although there may be a slight edge in increased accuracy when a log-odds transformation is used for the StackOverflow dataset.
+But that result is not statistically reliable for the sample size used in this analysis, and even if present the effect size is small.
+Nonetheless, it does seem appropriate to convert the context to a log odds adjustment value when context is added as an additional term to the random-permutation model.
+Since performance does not decrease when the transformation is made, may slightly increase,
+and leads to a more natural interpretation of each model term as a log odds value, it is reasonable to include this step for the random-permutation model.
+Future research could explore more improved methods to properly combine the terms.
 

### 2014-07-18 dissProject/preProcess.R

@@ -1906,6 +1906,45 @@ analyzeContext <- function(modelHashtagTbls, modelVsPredTbl) {
 	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweetNentropy', 'PriorStdTweetOrderTweetOrderless',
 				       'TweetNentropy', 'TweetOrderTweetOrderless'))
 	compareMeanDVDefault(tbl[sizeNum == 2 & dsetType == 'twitter' & DVName %in% DVNames], acc, figName='contextStandardByGroupT', groupCol='groupNum')
+	# RESULT: Freq weighting is better than entropy weighting for RP, and standard stoplist doesn't work well at all
+	# SO compare entropy to stoplist to freq to frentropy
+	DVNames = asTopHashtagAcross(c('PriorStd', 'TitleOrderlessEntropy', 'TitleOrderless', 'TitleOrderlessStoplist', 'TitleOrderlessFreq', 'TitleOrderlessFrentropy',
+				       'PriorStdTitleOrderlessEntropyBodyOrderlessEntropy', 'PriorStdTitleOrderlessBodyOrderless',
+				       'PriorStdTitleOrderlessStoplistBodyOrderlessStoplist',
+				       'PriorStdTitleOrderlessFreqBodyOrderlessFreq',
+				       'PriorStdTitleOrderlessFrentropyBodyOrderlessFrentropy'
+				       ))
+	compareMeanDVDefault(tbl[sizeNum == 2 & dsetType == 'stackoverflow' & DVName %in% DVNames], acc, figName='allWeightingsSOPerm')
+	DVNames = asTopHashtagAcross(c('PriorStd', 'TweetOrderlessEntropy', 'TweetOrderless', 'TweetOrderlessStoplist', 'TweetOrderlessFreq', 'TweetOrderlessFrentropy',
+				       'PriorStdTweetOrderEntropyTweetOrderlessEntropy', 'PriorStdTweetOrderTweetOrderless',
+				       'PriorStdTweetOrderStoplistTweetOrderlessStoplist',
+				       'PriorStdTweetOrderFreqTweetOrderlessFreq',
+				       'PriorStdTweetOrderFrentropyTweetOrderlessFrentropy'
+				       ))
+	compareMeanDVDefault(tbl[sizeNum == 2 & dsetType == 'twitter' & DVName %in% DVNames], acc, figName='allWeightingsTPerm', groupCol='dsetGroup')
[…]

### 2014-07-18 dissProject/preProcess.R
foo

@@ -1938,13 +1938,37 @@ analyzeContext <- function(modelHashtagTbls, modelVsPredTbl) {
 				       'PriorStdTitleOrderlessLgdimBodyOrderlessLgdim'))
 	compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames & sizeNum != 2], acc, figName='dimBySizeSO', groupCol='sizeNum')
 	# T Entropy all sizes
-	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweet',
-				       'PriorStdTweetOrderTweetOrderless', 'PriorStdTweetOrderEntropyTweetOrderlessEntropy',
-				       'PriorStdTweetOrderFreqTweetOrderlessFreq',
+	DVNames = asTopHashtagAcross(c('PriorStd',
+				       'PriorStdTweetOrderEntropyTweetOrderlessEntropy',
 				       'PriorStdTweetOrderSmdimTweetOrderlessSmdim',
 				       'PriorStdTweetOrderLgdimTweetOrderlessLgdim'))
 	compareMeanDVDefault(tbl[dsetType == 'twitter' & DVName %in% DVNames], acc, figName='dimBySizeT', groupCol='sizeNum')
-
+	# RESULT: Bayes and RP with frequency cutoff scales; RP with entropy saturates for SO
+	# SO Entropy all sizes
+	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTitleBody', 
+				       'PriorStdTitleOrderlessBodyOrderless',
+				       'PriorStdTitleOrderlessEntropyBodyOrderlessEntropy',
+				       'PriorStdTitleOrderlessFreqBodyOrderlessFreq'))
+	compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames & sizeNum != 2], acc, figName='freqVsEntropyBySizeSO', groupCol='sizeNum')
+	# T Entropy all sizes
+	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweet',
+				       'PriorStdTweetOrderEntropyTweetOrderlessEntropy',
+				       'PriorStdTweetOrderFreqTweetOrderlessFreq'))
+	compareMeanDVDefault(tbl[dsetType == 'twitter' & DVName %in% DVNames], acc, figName='freqVsEntropyBySizeT', groupCol='sizeNum')
+	# RESULT: Logodds technique works
[…]

### 2014-07-18 writeups/dissertation/dissertation.tex
foo

@@
 
+Vector-based models like the random-permutation model can easily incorporate word order when making predictions.
+This is one of their primary strenghts that has been discussed \parencites{Jones2007, Sahlgren2008}.
+In order to thouroughly test the impact of adding word order into the model, various methods for incorporating word order were tested on the StackOverflow and Twitter datasets.
+
+\subsubsection{Method}
+
+Three different methods for incorporating word order in the random-permutation model were tested:
+Taking just the sign of the word's position relative to the tag (direction method), using the position (standard method), and using the position but only including words that were used near the tag (window method).
+For the standard method, all words and their respective position are used to build the representation.
+For the direction method, essentially two representations are combined: all co-occurrences with words to the left of the tag, and all co-occurrences with words to the right of the tag.
+The window method is similar to the standard method, where each word's relative position is maintained, but it only includes words near the tag to build the representation.
+\textcite{Sahlgren2008} tested similar methods, and found that accuracy was highest when a narrow window was used for the window method (only including words at most two positions to the left and right of the tag).
+Teh same $+2$ $-2$ range was used to test the window method on the Twitter dataset.
+
+Only the Twitter dataset was tested, since the StackOverflow dataset contains no word order information between words in the title and body of the post and the tags used for hte post.
+The author chooses tags after filling out the title and body forms in the post, so the tags are never intermixed with words in the post like they are for Twitter.
+
+The same Twitter popular-users datasets and StackOverflow popular-hashtags datasets were used for testing.
+Both an ordered and orderless context component were created for the random-permutation models.
+The orderless context component uses the same co-occurrence matrix used by the Bayesian model to create the representation, since the Bayesian model does not contain any order information.
+Models were tested that used only one of the three word-order methods for the ordered context component, used only the orderless context component, used a combination of the two components, and combined context with prior.
+
+\subsubsection{Results}
+
+Model accuracy results after incorporating word order into the random-permutation model are included in Figure \ref{figContextOrder}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
+  \caption{Model accuracy for various word order methods for the random-permutation model for Twitter}
+  \label{figContextLogoddsT}
+\end{figure}
+
+All bar plots that include an order component where ``window'' or ``direction'' is not included in the title use the standard ordering method (no direction or windowing, all position information is included)
+This is the default ordering method used, so the names for these bar plots were kept consistent to the names used in other figures so that comparisons could be made across figures if necessary
+(similar to what was done for the default row-size for the random-permutation model).
+Base models that do not use entropy weighting were also included so that effect sizes could be compared between entropy weighting and adding word-order information.
+
+Looking first at models where no prior is added, the random-permutation model using only orderless context performs similar to the model using only order context with direction.
+All other order-only models perform worse than the direction order model and worse than the orderless model.
+
+Full models were also plotted against the model where only the prior and orderless context was included (i.e., no order).
+This was to test how much of the variance in model accuracy is shared between the orderless and order models.
+The results show that there is very little unique variance in the order-only models, so order information contributes very little to overall model performance.
+This can be seen by comparing the two-component model with only prior and orderless context to the three-component model with prior, orderless, and any of the three methods for the order context term.
+
+\subsubsection{Discussion}
+
+This was a surprising result.
+One of the main benefits discussed for the random-permutation model is that it can easily incorporate word-order information into the representation.
+But including word-order information does not dramatically improve performance. 
+For the random-permutation model, it is much more important to properly handle high-frequency stop words and include a prior component than it is to include word order information,
+at least for the Twitter dataset tested in this study.
+
+The original BEAGLE model \parencite{Jones2007} actually showed only minimal improvement when incorporating word order as well.
+In absolute terms, this model only improved by 2 percentage points in accuracy when trained on the TASA and tested on the TOEFL.
+\textcite{Sahlgren2008}'s results for word order for the random-permutation model are similar.
+At a row size of around \num{2000}, model accuracy improves by less than a percentage point when order information is incorporated.
+The results for larger row sizes are somewhat inconsistent, since the full model (order and orderless) sometimes underperforms the order-only model, which does not make sense.
+\textcite{Sahlgren2008} most likely did not use a regression technique to opimize the weights for each of the order and orderless model terms.
+So although incorporating word order into the random-permutation model does seem to slightly improve performance,
+the strenght of the model is more likely to be in how effeciently it can represent the co-occurrence matrix, than its ability to include word order in the representation.
+
+Since it was shown that including word-order information has little to no sigificant improvement on model accuracy, an attempt to incorporate word order information into the Bayesian model was not made.
+
+\subsection{Impact of Prior vs Context}
+
 \subsubsection{Method}
@@
 
-\subsection{Impact of Prior vs Context}
+\subsection{User-customized Co-occurrence Matrix}
 
@@
 
-\subsection{User-customized Co-occurrence Matrix}
-
-\subsubsection{Method}
-\subsubsection{Results}
-
 \section{General Discussion}
+
+FIXME: Include a paragraph about how the random-permutaiton model works so well due to the compression abilities, and not because of word order.
 

### 2014-07-18 dissProject/preProcess.R

@@ -1571,7 +1571,6 @@ renameColDVName <- function(tbl) {
 		    makeStandardMapping('Entropy', 'w/ entropy'),
 		    makeStandardMapping('Stoplist', 'w/ stoplist'),
 		    makeStandardMapping('Enthyman', 'w/ entropy and log odds'),
-		    makeStandardMapping('Hyman', 'w/ entropy and log odds'),
 		    makeStandardMapping('Direction', 'w/ entropy and direction'),
 		    makeStandardMapping('Window', 'w/ entropy and window'),
 		    makeStandardMapping('Freq', 'w/ freq'),
@@ -1612,9 +1611,9 @@ makeLogregMapping <- function() {
 	  'actTitleOrderlessEnthyman', 'RP title w/ entropy and log odds',
 	  'actBodyOrderlessEnthyman', 'RP body w/ entropy and log odds',
 	  'actTweetOrderlessEnthyman', 'RP orderless context w/ entropy and log odds',
-	  'actTitleOrderlessHyman', 'RP title w/ entropy and log odds',
-	  'actBodyOrderlessHyman', 'RP body w/ entropy and log odds',
-	  'actTweetOrderlessHyman', 'RP orderless context w/ entropy and log odds',
+	  'actTitleOrderlessEnthyman', 'RP title w/ entropy and log odds',
+	  'actBodyOrderlessEnthyman', 'RP body w/ entropy and log odds',
+	  'actTweetOrderlessEnthyman', 'RP orderless context w/ entropy and log odds',
 	  'actTweetUsercontext', 'Bayes orderless context w/ entropy and user sji',
 	  'actTweetOrderlessEnthyser', 'RP orderless context w/ entropy, log odds, and user sji')
 }
@@ -1969,6 +1968,24 @@ analyzeContext <- function(modelHashtagTbls, modelVsPredTbl) {
 				       'PriorStdTweetOrderNenthymanTweetOrderlessNenthyman',
 				       'TweetOrderless', 'TweetOrderlessNenthyman'))
 	compareMeanDVDefault(tbl[sizeNum == 2 & dsetType == 'twitter' & DVName %in% DVNames], acc, figName='logoddsT', groupCol='dsetGroup')
+	# RESULT: Not much information in order for RP
[…]

### 2014-07-18 writeups/dissertation/dissertation.tex
foo

@@
 
+Adding a user's prior tag history as an additional component has been shown to improve performance, but is this always the case?
+If enough context information has been collected, maybe a context-only model can perform just as well as a combined context and prior model.
+If this were the case, then it may be less important that the random-permutation model does not include a prior term by default.
+
 \subsubsection{Method}
+
+Full models were tested against context-only models for both sites and both the Bayesian and random-permutation model types.
+The four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled datasets were tested first.
+
 \subsubsection{Results}
+
+The results are included in Figures \ref{figContextPriorneededSO} and \ref{figContextPriorneededT}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
+  \caption{Model accuracy for various word order methods for the random-permutation model for Twitter}
+  \label{figContextPriorneededSO}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddT-crop.pdf}}}
+  \caption{Model accuracy for various word order methods for the random-permutation model for Twitter}
+  \label{figContextPriorneededT}
+\end{figure}
+
+For both the Bayesian and random-permutation models, accuracy remains the same when the prior compoennt is ermoved for StackOverflow.
+Accuracy does dramatically decrease for Twitter when the prior is removed, however.
+
+\subsubsection{Discussion}
+
+These results show that prior informatino is not always needed when a large amount of context is avaialble and not much information has been collected about each user's past history.
+The StackOverflow randomly-sampled dataset selects posts across the entire dataset, regardless of each user's popularity or tag history.
+Across the dataset, each user only asks FIXME questions on average, so there is not a large amount of past tag history available for the prior component to use,
+which is why this component is not a strong predictor in this case.
+
+However, can a context-only model perform just as well as a combined context and prior model when users with the largest amount of StackOverflow history are tested?
+If each user's tagging history is a useful predictor for StackOverflow,
+then this component should at least generate unique predictive variance in the case where users with large amounts of history are tested.
+
+\subsubsection{Method}
+
+The same StackOverflow popular-users datasets used to test the prior component in isolation were used to test the prior component when context was added.
+However, since computing activations for the context component take an order of magnitude more wall-clock time than computing the prior comonent, not all observations could be included for the test.
+To reduce the observations, the last FIXME posts from each user in the datasets were sampled.
+Using the last observations for each user ensures that the observations tested have aquired the greatest amount of past tagging history in the datasets.
+So if a context-only model performs just as well as the full model in this case, it would be surprising.
+
+\subsubsection{Results}
+
+The results comparing the full model to a context-only only model for the StackOverflow popular-users dataset are included in Figure \ref{figPUAddPrior}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddPUserSO-crop.pdf}}}
+  \caption{Unique variance in prior component for popular StackOverflow users}
+  \label{figPUAddPrior}
+\end{figure}
+
+In this case, past tag history does contain unique variance for accurately predicting tag use.
+Accuracy improves for both the Bayesian and random-permutation models when the prior model component is combined with context compared to context alone.
+
+\subsubsection{Discussion}
+
+This shows that incorporating past tagging history is important for even a site like StackOverflow where a large amount of context information has been collected.
+However, for this type of site it only becomes important when a reasonably large amount of a user's past tagging history has been collected.
+For a randomly-sampled user on StackOverflow, a context-only model captures the majority of the predicitve variance.
 

### 2014-07-22 writeups/dissertation/dissertation.tex
foo

@@
 However, since computing activations for the context component take an order of magnitude more wall-clock time than computing the prior comonent, not all observations could be included for the test.
-To reduce the observations, the last FIXME posts from each user in the datasets were sampled.
-Using the last observations for each user ensures that the observations tested have aquired the greatest amount of past tagging history in the datasets.
+To reduce the observations, the most recent FIXME posts from each user in the datasets were sampled.
+Using the most recent observations for each user ensures that the observations tested have aquired the greatest amount of past tagging history in the datasets.
 So if a context-only model performs just as well as the full model in this case, it would be surprising.
@@
 
+The co-occurrence matrix for the context component of both models has been built by aggregating across users and collecting all instances of context words and associated tags.
+That is, the default co-occurrence matrix is not customized to each user, so the same matrix is used for all users.
+This is commonly done mainly because there is never enough context by tag co-occurrences for a single user to generate a stable co-occurrence matrix.
+If there was enough information, acuracy would most likely improve since it is not likely that every user's word by tag co-occurrence profile is exactly the same.
+However, since several hundred posts for each user in the popular-users datasets are available, an analysis was tried where user-custom co-occurence matrices were used.
+
 \subsubsection{Method}
-\subsubsection{Results}
+
+Separate co-occurrence matrices were built for each user in the sampled StackOverflow and Twitter popular-users dataset.
+The same sampling method used to test the impact of the prior component compared to the context component for the popular-users datasets was used,
+but the oldest posts that were not included in the testing set were used to build the co-occurrence matrices.
+So for each user, the co-occurrence matrix was created from their oldest posts, up to (and not including) their FIXME newer posts that were used to test model accuracy.
+
+\subsubsection{Result}
+
+Model accuracy when using a user-customized co-occurrence matrix for StackOverflow and Twitter is shown in Figures \ref{figPUserCustomSO} and \ref{figPUserCustomT}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiSO-crop.pdf}}}
+  \caption{Model accuracy with user-customized context compoenent for StackOverflow}
+  \label{figPUCustomSO}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiT-crop.pdf}}}
+  \caption{Model accuracy with user-customized context component for Twitter}
+  \label{figPUCustomT}
+\end{figure}
+
+The results compare accuracy for models that use a global co-occurrence and user-specific co-occurence matrix.
+For StackOverflow, model accuracy for the user-specific (labeled as ``user sji'') can be compared against the default global co-occurrence matrix.
+Accuracy decreases dramatically when a user-specific co-occurrence matrix is used.
+
+The results are similar for Twitter.
+However since the entire dataset is not available for Twitter, a global co-occurrence matrix condition can not be tested,
+since the popular-users dataset samples across a wide range of hashtags, and due to Twitter's API restrictions it would be infeasable to collect all posts for each of the hashtags that are used in the dataset.
+But user-specific co-occurence matrices can be built, since the popular-users dataset does contain all context and hashtag observations for each user.
+
+Model accuracy when only the single user-customized context predictor is used is relatively poor, and the prior term is a much better predictor.
+Further, when this component is combined with the prior component, model performance improves only slightly over performance when only the prior is used.
+
+\subsubsection{Discussion}
+
+Both models require large numbers of context co-occurrence observations to make stable predictinos from context information.
+Otherwise, the co-occurrence matrices do not have enough data to properly stabilize.
+Millions of posts were used to generate the global co-occurrence matrices for Twitter and StackOverflow.
+If this number of posts had been collected for a single user, then the user-customized co-occurrence matrix for that user should lead to more accurate predictions compared to using the global matrix.
+However, only around 100 posts were used to build the user-specific matrices for each user.
+
+This is most likly why model accuracy was so poor when user-specific matrices were used:
+Even for the large datasets analyzed in this study, there is still not enough history collected for each user to build customized co-occurrence matrices.
+The prior model component can be customized for each user, since the prior is only concerned with tag use, and there are a reasonbly small set of tags that the model is choosing from.
+However, the context component is built from the the set of all of the prior tags crossed with all of the possible words, which is a much larger space than just the set of tags.
+So building a stable context component requires a much larger set of data than the prior component.
+
+That said, if a user only chooses a small set of tags (so the tag space is small), and computing a global co-occurrence matrix is infeasable (e.g., with the Twitter popular-users dataset), 
+then it may be worthwhiel to add the user-custom context model compoennt as an additional predictor.
+As was shown for Twitter, model accuracy can slightly improve over a model that just uses a custom prior, even when the co-occurrence matrix only contains around 100 posts.
 
@@
 
-FIXME: Include a paragraph about how the random-permutaiton model works so well due to the compression abilities, and not because of word order.
-
-\section{Summary}
-
-This research has numerous components that are relevant to improving our understanding of declarative memory.
-It should be emphasized that it has only recently become possible to evaluate psychological theories of declarative memory on large-scale real-world tasks where the theory can be tested on hundreds of millions of data points.
-These information-rich large-scale environments provide a unique opportunity to really stress test and explore the impacts of the psychological constraints of each theory
-on a much larger scale than has previously been possible.
-This environment will allow, for example, overall performance differences between an ACT-R Bayesian retrieval model and a random-permutation vector-based model to be directly compared on two real-world tagging tasks,
-which has not been done previously.
-
-This data-rich environment will also enable the proposed modifications to each model to be quickly evaluated, so that further model improvements can be iteratively proposed and tested.
-The main two proposed model modifications (incorporating word order into ACT-R and user history into vector-based representations) have not been tested before, and seem likely to significantly improve performance.
-
-Further, the proposed model modifications are aimed at the architectural level, may very well be task independent, and may generalize and improve model fit across other tasks.
-Discovering an effective way to represent the time decay of information in a vector-based model could improve model fit on other tasks where the memory model has been used. 
-Finding a task-agnostic way to represent word order information in a Bayesian ACT-R model could improve model accuracy when applied to other domains where it seems plausible that participants are utilizing word order.
-
-Additionally, the tag-recommendation models developed through this research should be immediately relevant to real-world tasks.
-If model accuracy is high enough, then the developed models may be used as a foundation for tag recommendation systems that are deployed on Twitter or StackOverflow.
-
-Also since the developed models will be task agnostic, they can potentially be used as tag recommendation systems for other domains.
+\subsubsection{Representation}
+
+Overall, both the random-permutaiton and Bayesian models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
+Given that the storage space required to represent the co-occurrences for the random-permutation is dramatically smaller than the uncompressed Bayesian model,
+and that the compression is lossy and essentially random (i.e., no SVD),
+it is quite interesting that the random-permutation model performs nearly as well as the full Bayesian model for this task.
+
+These results pose an interesting and important fundamental question about declarative memory:
+How, precisely, is the neural system configured to represent co-occurrences for long-term memory?
+Since, at the behavioral level, both models perform similarly, it is possible that the neural system is using some sort of vector-based compressed storage (e.g., random permutations) to store co-occurrences.
+Vector-based models are simple enough that they can easily be implemented in a neural network.
+Perhaps a form of vector-based compressed storage is being used at the neural level,
+since it can be implemented in neurons and it is a way for the machinery to generate predictions that closely match the statistical structure of the environment (i.e., that closely match the Bayesian representation).
+
+As neural imaging systems improve, this question of neural representation can be best answered by looking directly at the neural activity.
+However, there are possible behavioral studies that could be done to provide evidence that may favor one representation over the other:
+The largest differences between the Bayesian and vector-based models happen with small numbers of co-occurrence observations.
+The vector-based model begins to generate more accurate predictions sooner than the Bayesian models,
+since the noise introduced from the compression helps smooth out activation values until a larger number of co-occurrences are observed and the co-occurrence values stabilize.
+For a behavioral study, novel words could be introduced alongside the corpus of known words, and the participant could be asked to make retrievals of the novel words and known words as more novel words are introduced.
+Both models would make different predictions about the rate at which activation values for the novel words stabilize relative to the known words, and such make different predictions about retrieval error rates over time.
+Learning a new language might be a domain where large amounts of these sorts of retrievals could be collected.
+For instance, the online site Duolingo might contain enough collective data about language learning and word translation to better tease apart the two context models.
+
+\subsubsection{Compression}
+
+Also, the relatively strong performance of the random-permutation model is not due to the model's ability to represent word order.
+Performance is similar for the random-permutation and Bayesian models when the co-occurrence matrix for both models is the same, and not based on word order.
+The random-permutation model seems to work so well primarily due to the way that the full co-occurrence representation is compressed.
+
+Compression for the random-permutation model is essentiallly random.
+That is, no SVD is done to identify the most predictive components,
+so the compressed representation used by the random-permutation model does not contain nearly as much of the variance from the original co-occurrence matrix that it could have.
+However, performing a full SVD on a co-occurrence matrix as large as the ones generated for StackOverflow and Twitter is computationally infeasable.
+So, it also seems plausible that this SVD computation is infeasable at the neural level.
+\textcite{Landauer1997} explicitly state that they make no claims that the brain is computing an SVD at the neural level;
+only that the neural machinery is set up so that the representation makes predictions that are similar to what would have been generated if the matrix decomposition for an SVD would have been computed.
+Perhaps the randomly-compressed representation used for the random-permutaiton model is the way in which the brain tries to achive the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
+
+\subsubsection{Attentional Weight}
+
+The declarative memory model for ACT-R has an attentional weight term $W_{j}$ that can be used to implement how stop words are handled.
+The random-permutaiton model was also easily modified to have an attentional weight term by attentuating each word's set of ones and negative ones values by that word's attentional weight.
+Both a method for stop-word removal (frequency filtering) and stop-word attentuating (entropy weighting) performed well for both models.
+Also, using a proper method for handling the low-predictor words produced some of the largest improvements in model accuracy compared to the other explorations
+(e.g., compared to word order, properly adding terms, compression amount for random-permutation representation).
+Since the ACT-R declarative memory theory already has a component for attentional weight, it may be the case that the neural system performs some form of stop-word weighting when cmoputing activations,
+and this is reflected in the attentional weight term.
+That is, $W_{j}$ can be defined mathematically, has the same general form across tasks, and both the frequency-filtering and entropy-weighting methods for stop-word removal are likely first candidates for the definition.
+This claim, however, would require that no top-down activatino is used to compute the attentional weight, which would mean that the $W_{j}$ is not under conscious control, which is not likely to be correct.
+Perhaps it is a combination of top-down and bottom-up systems that determine $W_{j}$, and the entropy-weighting and frequency-filtering techniques are methods of formalizing the bottom-up component.
+
+\subsubsection{Generalizing Results}
+
+It seems likely that the entropy method for computing the attentional weight term $W_{j}$ may be applicable to other declarative memory tasks.
+The entropy method is parameter free (unlike the frequency-filtering method), so it completely data driven (much like the co-occurrence matrix) and does not require tuning.
+It is common to handle stop words in some manner when working with NLP and large text documents.
+So, if these large natural-language documents are used to test declarative memory, then the entropy method may be a more cognitively plausible way of dealing with stop words rather than 
+filtering based on a pre-determined list or based on the frequency count in the dataset, particulary since the entropy method substantially increases model accuracy for these datasets and since it is parameter free.
+
+Adding a prior component to the random-permutatino model also increased model accuracy for both StackOverflow and Twitter.
+It is unsurprising that this occurred.
+However, it is quite surprising that this is the first time that the prior component has been included in a random-permutation model.
+If the random-permutatino model is explored as a declarative memory theory for other datasets, it should be worthwhile to include the prior component for those explorations.
+Although it did not seem to matter much how the prior component is combined with the context component in this study, 
+there may be better ways than the log-odds transformation to combine the terms.
+Further research could focus on exploring a larger set of methods for combining the context component of the random-permutation model with the prior component from the ACT-R theory.
+
+\subsubsection{Future Applications}
+
+The tag-recommendation models developed through this research should be immediately relevant to real-world tasks.
+Model accuracy is high enough (particularly for StackOverflow) for the developed models to be used as a foundation for tag recommendation systems that are deployed on Twitter or StackOverflow.
+Also since the developed models are task agnostic, they can potentially be used as tag recommendation systems for other domains.
 Examples of similar tagging tasks are hashtag use on Facebook, flagging (i.e., tagging) Wikipedia articles for content issues or overly-opinionated material,
@@
 
-The end result of this research will be [1] a better understanding of how a user's past tagging history influences future tag use,
+\subsubsection{Summary}
+
+The end result of this research is [1] a better understanding of how a user's past tagging history influences future tag use,
 [2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on large-scale real-world retrieval tasks,
-[3] improvements to each memory model where the strengths of each model (word order and user prior likelihood) are incorporated into the other,
-and [4] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
-Introducing, exploring, and evaluating reasonable architectural modifications to two commonly-used memory models is worth the effort:
-It will increase our understanding of how the effects of imposing specific architectural constraints on memory models influence what information is retrieved.
+and [3] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
+This work has shown how past user history is important, and how a context-only model like the random-permutation model can be modified to include the prior component.
+The work has also shown how the strength of the random-permutation model is not in its ability to represent word order,
+and rather in its ability to efficiently compress the full co-occurrence representation used for the Bayesian model.
+Several different formalisms for ACT-R's attentional weighting term were also explored, and an entropy-weighting technique was found to perform well without requiring any parameter tuning.
+Finally, this work has helped raise the question about how co-occurrence observations are represented at the neural level, since two models with very different representations performed similarly well.
+The results from the study and the new research questions raised will help to better understand how imposing specific architectural constraints on memory models influence what information is retrieved.
+
+\subsubsection{Acknowledgements}
+
+The majority of the modeling and analysis was done with the R statistical computing environment \parencite{RCoreTeam2014}, and the data.table \parencite{Dowle2014} and ggplot2 \parencite{Wickham2009} packages.
 

### 2014-07-22 writeups/dissertation/dissertation.tex
foo

@@
 
-\subsection{Aquiring Datasets}
+\subsection{Acquiring Datasets}
 
@@
 
-Twitter provides two types of APIs for aquiring data: A query-based and streaming interface.
+Twitter provides two types of APIs for acquiring data: A query-based and streaming interface.
 The \textcite{Tweepy} Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
@@
 How does a user's prior hashtag use influence their likelihood of choosing that hashtag in the future?
-This question was tested by examing how well ACT-R's base-level activation component of declarative memory predicts the tags that authors choose, given each author's past tagging behavior.
+This question was tested by examining how well ACT-R's base-level activation component of declarative memory predicts the tags that authors choose, given each author's past tagging behavior.
 The model predicts that the more recently and frequently used chunks have higher prior activations, and are more likely to be needed in the future.
@@
 
-For both equatons, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g,.
-the 2nd time that ``a'' appears), and $n$ is the number of times each chunk has been observed.
+For both equations, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``a'' appears), and $n$ is the number of times each chunk has been observed.
 Since the optimized learning form assumes a constant presentation rate for each chunk,
@@
 Higher values mean that chunk activation decays faster, which means that the most recently-presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
-Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks that have occured most frequently will have the highest activation,
+Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks that have occurred most frequently will have the highest activation,
 and activation depends less on the specific chunk that have been observed in the recent past.
@@
 At each tagging instance, the model has activation values for a set of tags.
-Model accurracy was assessed by rank ordering these activations and choosing the model's highest $N$ tags, where $N$ is the number of tags used for that specific post.
+Model accuracy was assessed by rank ordering these activations and choosing the model's highest $N$ tags, where $N$ is the number of tags used for that specific post.
 The model's chosen $N$ tags were compared to the user's chosen $N$ tags, and if a model's chosen tag was in the set of user's chosen tags, then the prediction was marked as correct.
@@
 
-Finally, model accuraccy was analyzed across all dataset slices for StackOverflow and Twitter.
+Finally, model accuracy was analyzed across all dataset slices for StackOverflow and Twitter.
 Mean accuracy for both sites is included in Figure \ref{figPriorAcc}. 
@@
 The constrained form could be used in a regression if a custom form of the regression was implemented.
-However this was not persued since around \num{2500} total regressions were run on these datasets for two reasons:
-A custom implementation in R would not likely run in compiled C, and even if it was implemented in a compiled language it would be computationally intractible without a closed-form solution.
+However this was not pursued since around \num{2500} total regressions were run on these datasets for two reasons:
+A custom implementation in R would not likely run in compiled C, and even if it was implemented in a compiled language it would be computationally intractable without a closed-form solution.
 
@@
 \textcite{Stanley2013} did this by subtracting off the mean total activation for the top 5 tags for each post.
-However this method is complicated since computing the toal activation for a tag requires already knowing the weights for each model component that are used to compute that activation.
+However this method is complicated since computing the total activation for a tag requires already knowing the weights for each model component that are used to compute that activation.
 \citeauthor{Stanley2013} used an iterative approach to determine the weights used for the offset term, and kept iterating until those weights matched the weights returned after running the logistic regression.
@@
 So in order to generate a custom user prior for each sample in the dataset, that user's entire tagging history must be accessed through the API.
-Since access rates are limited across that interface, it is infeasable to do this process for every user that appears in the Twitter popular-users dataset.
+Since access rates are limited across that interface, it is infeasible to do this process for every user that appears in the Twitter popular-users dataset.
 So instead, the global prior tag history for all users in the Twitter popular-users dataset was used to compute prior activation values for each tag.
@@
 and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high-performance desktop, even with highly optimized representations.
-Already FIXME total co-occuring word tag observations are used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
+Already FIXME total co-occurring word tag observations are used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
 It is also the case that \num{1000000} posts for StackOverflow is already around $1/7$ of the total posts created on the site,
-so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrent matrix instead,
+so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrence matrix instead,
 especially since FIXME co-occurrences are already used with \num{1000000} posts.
@@
 The random-permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
-So, for example, the ``standard Prior model relaxed across posts w/o offset'' means that only the prior model component was used, and the offset for that component was not subtracted off when determining activation.
+So, for example, the ``standard Prior model relaxed across posts without offset'' means that only the prior model component was used, and the offset for that component was not subtracted off when determining activation.
 As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.
@@
 
-It is actually somewhat dissapointing that the offset term is even needed.
+It is actually somewhat disappointing that the offset term is even needed.
 If it were the case that model activation for a tag in a specific post could be interpreted as an exact activation value that did not depend on the specific post,
@@
 
-A likely and simple explaination for this is that possibly there are posts where a large set of tags have high activation,
+A likely and simple explanation for this is that possibly there are posts where a large set of tags have high activation,
 but the author is throttling their retrieval threshold for each post such that only a small set of tags are actually used.
@@
 A data-driven frequency-weighted approach was tried, along with using a previously-derived stoplist, and no stop word removal at all.
-Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and uing a previously-derived stoplist hardly increased performance.
+Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and using a previously-derived stoplist hardly increased performance.
 However, \citeauthor{Sahlgren2008} did not try to weight the stop words, and only tried methods to remove them.
-So alongside looking at frequency weighting, previously derived, and no stop word removal, \textcite{Stanley2013}'s method of attentuating stop words based on Entropy was also explored.
+So alongside looking at frequency weighting, previously derived, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on Entropy was also explored.
 
@@
 
-\paragraph{Determing Frequency Cutoff}
+\paragraph{Determining Frequency Cutoff}
 
@@
 Note that using a previously-defined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
-The cuttoff that produced the frequency plot is where a word represents more than \num{.0004} occurrences in the dataset.
+The cutoff that produced the frequency plot is where a word represents more than \num{.0004} occurrences in the dataset.
 This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset were not given.
@@
 The results show that model accuracy improves if stop words are handled in any of the three ways (pre-defined stoplist filtering, entropy weighting, and frequency filtering).
-Using a pre-defined stoplist makes the smallest improvment, while the entropy and frequency techniques improve performance the most, and frequency shows a slight edge over entropy.
+Using a pre-defined stoplist makes the smallest improvement, while the entropy and frequency techniques improve performance the most, and frequency shows a slight edge over entropy.
 However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
@@
 However frequency filtering does have a few drawbacks compared to entropy weighting:
-It produces accuracy measurments that have higher levels of noise (see Figure \ref{figContextStopRPSO} and frequency models for the title component in StackOverflow).
+It produces accuracy measurements that have higher levels of noise (see Figure \ref{figContextStopRPSO} and frequency models for the title component in StackOverflow).
 Also, the ideal cutoff threshold must be identified in order to filter based on frequency,
@@
 This threshold is a free parameter in frequency filtering that is not required when weighting stop words by entropy.
-Therefore, favoring parsimony, for the random-permutation model, attentuating stop words based on entropy produces more accurate results when compared to filtering the stop words based on frequency of use.
+Therefore, favoring parsimony, for the random-permutation model, attenuating stop words based on entropy produces more accurate results when compared to filtering the stop words based on frequency of use.
 This result is interesting since filtering stop words based on frequency is the default method for random permutation models \parencite{Sahlgren2008}.
@@
 Model accuracy most likely decreases for the Bayesian model because this model already has a normalization component built in that the random-permutation model does not have.
-When computing activaiton for the random-permutation model, the corrollation is calculated directly on a matrix of counts of co-occurrences for words and tags.
+When computing activation for the random-permutation model, the correlation is calculated directly on a matrix of counts of co-occurrences for words and tags.
 However for the Bayesian model this count matrix is converted to a log odds ($S_{ji}$) matrix,
 where both the number of observations for the word and the tag are normalized when computing the value for each cell (see the $S_{ji}$ computation in Table \ref{tabModACTRModel}).
-So it is likely the case that a form of frequency \emph{attentuation} is already built into the Bayesian model.
+So it is likely the case that a form of frequency \emph{attenuation} is already built into the Bayesian model.
 And therefore adding a frequency filter on top of this does nothing to improve accuracy, and only removes predictive information.
 
-\textcite{Stanley2013} found that the entropy-weighting technique to attentuate stop words for the Bayesian model improved model accuracy.
+\textcite{Stanley2013} found that the entropy-weighting technique to attenuate stop words for the Bayesian model improved model accuracy.
 The same effect is found for the StackOverflow and Twitter datasets in this study.
@@
 In order to see how accuracy changed as function of document size, 
-co-occurence matrices for a range of smaller document sizes were built.
+co-occurrence matrices for a range of smaller document sizes were built.
 Both the random-permutation and Bayesian models were tested on all document sizes.
@@
 However model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random-permutation model, while the frequency technique only begins to plateau.
-The Bayesian model does not plateau and accuracy continues to improveas document size increases.
+The Bayesian model does not plateau and accuracy continues to improves document size increases.
 
@@
 Even if the non-predictive words based on entropy are attenuated,
-when the high-frequency words are included random-permutatoin model,
-this attentuation is not enough to overcome the range effects that grow with document size.
-These effects end up saturating all of the correlations near one, and consequently decrease discriminability and accuracy.
+when the high-frequency words are included random-permutation model,
+this attenuation is not enough to overcome the range effects that grow with document size.
+These effects end up saturating all of the correlations near one, and consequently decrease discrimination and accuracy.
 However, the effect of saturation is small compared to the large effect from using either frequency filtering or entropy weighting to handle the stop words.
@@
 
-Both entropy weighting and frequency filtering in Figure \ref{figContextDSizeSO} plateu for the random-permutation model with increasing document size.
+Both entropy weighting and frequency filtering in Figure \ref{figContextDSizeSO} plateau for the random-permutation model with increasing document size.
 Once again, the Bayesian model continues to improve with document size.
@@
 
-It is interesting to compare model accuracy betwen the Bayesian and random-permutation model at different document sizes for both StackOverflow and Twitter.
+It is interesting to compare model accuracy between the Bayesian and random-permutation model at different document sizes for both StackOverflow and Twitter.
 At smaller document sizes, the random-permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as document size increases.
@@
 As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random-permutation model is actually beneficial, and the compressed random-permutation model should be used.
-This crosstalk and noise is beneficial because it helps soften spurious $S_{ji}$ values when the values for each individual cell in the co-occurrence matrix are still volitle, which happens at smaller document sizes.
+This crosstalk and noise is beneficial because it helps soften spurious $S_{ji}$ values when the values for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller document sizes.
 
@@
 Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
-Accuracy measurments were the same as were used for previous runs of these datasets: mean of the mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.
+Accuracy measurements were the same as were used for previous runs of these datasets: mean of the mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.
 
@@
 Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy.
-However \num{2048} rows is sufficient for these dataests, as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy.
+However \num{2048} rows is sufficient for these datasets, as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy.
 
@@
 
-However it should be pointed out that the way the random-permutation model compresses the words is completly random.
+However it should be pointed out that the way the random-permutation model compresses the words is completely random.
 It is not an LSA approach where the few ones and negative ones for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
-Everything is randomly chosen with the random-permutaton model.
+Everything is randomly chosen with the random-permutation model.
 The environment vector for each word is randomly selected, and no SVD is performed to choose these in a more discriminating manner.
@@
 
-The log-odds transformation technique was tested for the random-permutaiton model on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.
+The log-odds transformation technique was tested for the random-permutation model on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
-  \caption{Random-permutation model accuracy when using log-odds trasformation on context for StackOverflow}
+  \caption{Random-permutation model accuracy when using log-odds transformation on context for StackOverflow}
   \label{figContextLogoddsSO}
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
-  \caption{Random-permutation model accuracy when using log-odds trasformation on context for Twitter}
+  \caption{Random-permutation model accuracy when using log-odds transformation on context for Twitter}
   \label{figContextLogoddsT}
@@
 Vector-based models like the random-permutation model can easily incorporate word order when making predictions.
-This is one of their primary strenghts that has been discussed \parencites{Jones2007, Sahlgren2008}.
-In order to thouroughly test the impact of adding word order into the model, various methods for incorporating word order were tested on the StackOverflow and Twitter datasets.
+This is one of their primary strengths that has been discussed \parencites{Jones2007, Sahlgren2008}.
+In order to thoroughly test the impact of adding word order into the model, various methods for incorporating word order were tested on the StackOverflow and Twitter datasets.
 
@@
 \textcite{Sahlgren2008} tested similar methods, and found that accuracy was highest when a narrow window was used for the window method (only including words at most two positions to the left and right of the tag).
-Teh same $+2$ $-2$ range was used to test the window method on the Twitter dataset.
-
-Only the Twitter dataset was tested, since the StackOverflow dataset contains no word order information between words in the title and body of the post and the tags used for hte post.
+The same $+2$ $-2$ range was used to test the window method on the Twitter dataset.
+
+Only the Twitter dataset was tested, since the StackOverflow dataset contains no word order information between words in the title and body of the post and the tags used for the post.
 The author chooses tags after filling out the title and body forms in the post, so the tags are never intermixed with words in the post like they are for Twitter.
@@
 The results for larger row sizes are somewhat inconsistent, since the full model (order and orderless) sometimes underperforms the order-only model, which does not make sense.
-\textcite{Sahlgren2008} most likely did not use a regression technique to opimize the weights for each of the order and orderless model terms.
+\textcite{Sahlgren2008} most likely did not use a regression technique to optimize the weights for each of the order and orderless model terms.
 So although incorporating word order into the random-permutation model does seem to slightly improve performance,
-the strenght of the model is more likely to be in how effeciently it can represent the co-occurrence matrix, than its ability to include word order in the representation.
-
-Since it was shown that including word-order information has little to no sigificant improvement on model accuracy, an attempt to incorporate word order information into the Bayesian model was not made.
+the strength of the model is more likely to be in how efficiently it can represent the co-occurrence matrix, than its ability to include word order in the representation.
+
+Since it was shown that including word-order information has little to no significant improvement on model accuracy, an attempt to incorporate word order information into the Bayesian model was not made.
 
@@
 
-For both the Bayesian and random-permutation models, accuracy remains the same when the prior compoennt is ermoved for StackOverflow.
+For both the Bayesian and random-permutation models, accuracy remains the same when the prior component is removed for StackOverflow.
 Accuracy does dramatically decrease for Twitter when the prior is removed, however.
@@
 
-These results show that prior informatino is not always needed when a large amount of context is avaialble and not much information has been collected about each user's past history.
+These results show that prior information is not always needed when a large amount of context is available and not much information has been collected about each user's past history.
 The StackOverflow randomly-sampled dataset selects posts across the entire dataset, regardless of each user's popularity or tag history.
@@
 The same StackOverflow popular-users datasets used to test the prior component in isolation were used to test the prior component when context was added.
-However, since computing activations for the context component take an order of magnitude more wall-clock time than computing the prior comonent, not all observations could be included for the test.
+However, since computing activations for the context component take an order of magnitude more wall-clock time than computing the prior component, not all observations could be included for the test.
 To reduce the observations, the most recent FIXME posts from each user in the datasets were sampled.
-Using the most recent observations for each user ensures that the observations tested have aquired the greatest amount of past tagging history in the datasets.
+Using the most recent observations for each user ensures that the observations tested have acquired the greatest amount of past tagging history in the datasets.
 So if a context-only model performs just as well as the full model in this case, it would be surprising.
@@
 However, for this type of site it only becomes important when a reasonably large amount of a user's past tagging history has been collected.
-For a randomly-sampled user on StackOverflow, a context-only model captures the majority of the predicitve variance.
+For a randomly-sampled user on StackOverflow, a context-only model captures the majority of the predictive variance.
 
@@
 This is commonly done mainly because there is never enough context by tag co-occurrences for a single user to generate a stable co-occurrence matrix.
-If there was enough information, acuracy would most likely improve since it is not likely that every user's word by tag co-occurrence profile is exactly the same.
-However, since several hundred posts for each user in the popular-users datasets are available, an analysis was tried where user-custom co-occurence matrices were used.
+If there was enough information, accuracy would most likely improve since it is not likely that every user's word by tag co-occurrence profile is exactly the same.
+However, since several hundred posts for each user in the popular-users datasets are available, an analysis was tried where user-custom co-occurrence matrices were used.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiSO-crop.pdf}}}
-  \caption{Model accuracy with user-customized context compoenent for StackOverflow}
+  \caption{Model accuracy with user-customized context component for StackOverflow}
   \label{figPUCustomSO}
@@
 
-The results compare accuracy for models that use a global co-occurrence and user-specific co-occurence matrix.
+The results compare accuracy for models that use a global co-occurrence and user-specific co-occurrence matrix.
 For StackOverflow, model accuracy for the user-specific (labeled as ``user sji'') can be compared against the default global co-occurrence matrix.
@@
 However since the entire dataset is not available for Twitter, a global co-occurrence matrix condition can not be tested,
-since the popular-users dataset samples across a wide range of hashtags, and due to Twitter's API restrictions it would be infeasable to collect all posts for each of the hashtags that are used in the dataset.
-But user-specific co-occurence matrices can be built, since the popular-users dataset does contain all context and hashtag observations for each user.
+since the popular-users dataset samples across a wide range of hashtags, and due to Twitter's API restrictions it would be infeasible to collect all posts for each of the hashtags that are used in the dataset.
+But user-specific co-occurrence matrices can be built, since the popular-users dataset does contain all context and hashtag observations for each user.
 
@@
 
-Both models require large numbers of context co-occurrence observations to make stable predictinos from context information.
+Both models require large numbers of context co-occurrence observations to make stable predictions from context information.
 Otherwise, the co-occurrence matrices do not have enough data to properly stabilize.
@@
 
-This is most likly why model accuracy was so poor when user-specific matrices were used:
+This is most likely why model accuracy was so poor when user-specific matrices were used:
 Even for the large datasets analyzed in this study, there is still not enough history collected for each user to build customized co-occurrence matrices.
-The prior model component can be customized for each user, since the prior is only concerned with tag use, and there are a reasonbly small set of tags that the model is choosing from.
-However, the context component is built from the the set of all of the prior tags crossed with all of the possible words, which is a much larger space than just the set of tags.
+The prior model component can be customized for each user, since the prior is only concerned with tag use, and there are a reasonably small set of tags that the model is choosing from.
+However, the context component is built from the set of all of the prior tags crossed with all of the possible words, which is a much larger space than just the set of tags.
 So building a stable context component requires a much larger set of data than the prior component.
 
-That said, if a user only chooses a small set of tags (so the tag space is small), and computing a global co-occurrence matrix is infeasable (e.g., with the Twitter popular-users dataset), 
-then it may be worthwhiel to add the user-custom context model compoennt as an additional predictor.
+That said, if a user only chooses a small set of tags (so the tag space is small), and computing a global co-occurrence matrix is infeasible (e.g., with the Twitter popular-users dataset), 
+then it may be worthwhile to add the user-custom context model component as an additional predictor.
 As was shown for Twitter, model accuracy can slightly improve over a model that just uses a custom prior, even when the co-occurrence matrix only contains around 100 posts.
@@
 
-Overall, both the random-permutaiton and Bayesian models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
+Overall, both the random-permutation and Bayesian models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
 Given that the storage space required to represent the co-occurrences for the random-permutation is dramatically smaller than the uncompressed Bayesian model,
@@
 
-Compression for the random-permutation model is essentiallly random.
+Compression for the random-permutation model is essentially random.
 That is, no SVD is done to identify the most predictive components,
 so the compressed representation used by the random-permutation model does not contain nearly as much of the variance from the original co-occurrence matrix that it could have.
-However, performing a full SVD on a co-occurrence matrix as large as the ones generated for StackOverflow and Twitter is computationally infeasable.
-So, it also seems plausible that this SVD computation is infeasable at the neural level.
+However, performing a full SVD on a co-occurrence matrix as large as the ones generated for StackOverflow and Twitter is computationally infeasible.
+So, it also seems plausible that this SVD computation is infeasible at the neural level.
 \textcite{Landauer1997} explicitly state that they make no claims that the brain is computing an SVD at the neural level;
 only that the neural machinery is set up so that the representation makes predictions that are similar to what would have been generated if the matrix decomposition for an SVD would have been computed.
-Perhaps the randomly-compressed representation used for the random-permutaiton model is the way in which the brain tries to achive the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
+Perhaps the randomly-compressed representation used for the random-permutation model is the way in which the brain tries to achieve the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
 
@@
 The declarative memory model for ACT-R has an attentional weight term $W_{j}$ that can be used to implement how stop words are handled.
-The random-permutaiton model was also easily modified to have an attentional weight term by attentuating each word's set of ones and negative ones values by that word's attentional weight.
-Both a method for stop-word removal (frequency filtering) and stop-word attentuating (entropy weighting) performed well for both models.
+The random-permutation model was also easily modified to have an attentional weight term by attenuating each word's set of ones and negative ones values by that word's attentional weight.
+Both a method for stop-word removal (frequency filtering) and stop-word attenuating (entropy weighting) performed well for both models.
 Also, using a proper method for handling the low-predictor words produced some of the largest improvements in model accuracy compared to the other explorations
 (e.g., compared to word order, properly adding terms, compression amount for random-permutation representation).
-Since the ACT-R declarative memory theory already has a component for attentional weight, it may be the case that the neural system performs some form of stop-word weighting when cmoputing activations,
+Since the ACT-R declarative memory theory already has a component for attentional weight, it may be the case that the neural system performs some form of stop-word weighting when computing activations,
 and this is reflected in the attentional weight term.
 That is, $W_{j}$ can be defined mathematically, has the same general form across tasks, and both the frequency-filtering and entropy-weighting methods for stop-word removal are likely first candidates for the definition.
-This claim, however, would require that no top-down activatino is used to compute the attentional weight, which would mean that the $W_{j}$ is not under conscious control, which is not likely to be correct.
+This claim, however, would require that no top-down activation is used to compute the attentional weight, which would mean that the $W_{j}$ is not under conscious control, which is not likely to be correct.
 Perhaps it is a combination of top-down and bottom-up systems that determine $W_{j}$, and the entropy-weighting and frequency-filtering techniques are methods of formalizing the bottom-up component.
@@
 So, if these large natural-language documents are used to test declarative memory, then the entropy method may be a more cognitively plausible way of dealing with stop words rather than 
-filtering based on a pre-determined list or based on the frequency count in the dataset, particulary since the entropy method substantially increases model accuracy for these datasets and since it is parameter free.
-
-Adding a prior component to the random-permutatino model also increased model accuracy for both StackOverflow and Twitter.
+filtering based on a pre-determined list or based on the frequency count in the dataset, particularly since the entropy method substantially increases model accuracy for these datasets and since it is parameter free.
+
+Adding a prior component to the random-permutation model also increased model accuracy for both StackOverflow and Twitter.
 It is unsurprising that this occurred.
 However, it is quite surprising that this is the first time that the prior component has been included in a random-permutation model.
-If the random-permutatino model is explored as a declarative memory theory for other datasets, it should be worthwhile to include the prior component for those explorations.
+If the random-permutation model is explored as a declarative memory theory for other datasets, it should be worthwhile to include the prior component for those explorations.
 Although it did not seem to matter much how the prior component is combined with the context component in this study, 

### 2014-07-22 writeups/dissertation/dissertation.tex
foo

@@
 
-The majority of the modeling and analysis was done with the R statistical computing environment \parencite{RCoreTeam2014}, and the data.table \parencite{Dowle2014} and ggplot2 \parencite{Wickham2009} packages.
+The majority of the modeling and analysis was done with the R statistical computing environment \parencite{RCoreTeam2014} and two important R packages:
+data.table \parencite{Dowle2014} was leveraged to implement the models efficiently, and ggplot2 \parencite{Wickham2009} was used to produce the visualizations.
 

### 2014-07-23 writeups/dissertation/dissertation.tex
foo

@@
 
+Overall, both the random-permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
+These results provide support that choosing a hashtag when composing a tweet and tagging a StackOverflow post is a long-term memory retrieval process.
+However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
+So is most likely the case that the task is not done solely with declarative memory processes.
+
+Or, perhaps authors are taking advantage of context information that is not currently represented in the model.
+For example, time of day might be a useful predictor of tag use.
+\textcite{Cohn2012} built a tag-prediction model that was based primarily on temporal cycles of tag use.
+This technique could be naturally incorporated into a co-occurrence framework by including time of day as an additional context by tag co-occurrence for each post.
+This would allow time of day information to be used alongside the rest of the context information in the words of the post.
+A similar technique could be used to add additional context predictors to the model.
+These co-occurrences could either be added on as additional columns to the existing co-occurrence matrix,
+or a separate co-occurrence matrix could be used for each additional context component, similar to how separate matrices and context predictors were created for the words in the title and body of a post.
+Including additional context predictors such as time of day may help improve model accuracy,
+and perhaps it is the case that the rest of the variance left to predict is from these context components that the author is using but are not yet included in the model.
+That is, perhaps this tagging process is completely declarative, and the remaining cues that the author is using are yet to be discovered.
+
+\tectcite{Anderson1991} also used tasks similar to tag use on StackOverflow and Twitter when deriving the prior component of the declarative memory theory.
+\citeauthor{Anderson1991} examined, for example, the frequency of word choice over time used for the title of articles for the New York Times,
+and used the results to derive the power law equation for the prior term in the declarative memory model.
+ACT-R's declarative memory theory claims that the system has evolved to make predictions that closely match the statistical structure of the environment.
+Using a power-law equation to model future tag use based on past tag use is a reasonable statistical approach to the problem.
+It is also the case that this same equation is an accurate model for the prior component of declarative memory, as was shown by \citeauthor{Anderson1991} in the original derivation, 
+and as is supported in the results from this study.
+
 \subsubsection{Representation}
 
-Overall, both the random-permutation and Bayesian models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
 Given that the storage space required to represent the co-occurrences for the random-permutation is dramatically smaller than the uncompressed Bayesian model,

### 2014-07-23 writeups/dissertation/dissertation.tex
foo

@@
 
-\tectcite{Anderson1991} also used tasks similar to tag use on StackOverflow and Twitter when deriving the prior component of the declarative memory theory.
+\textcite{Anderson1991} also used tasks similar to tag use on StackOverflow and Twitter when deriving the prior component of the declarative memory theory.
 \citeauthor{Anderson1991} examined, for example, the frequency of word choice over time used for the title of articles for the New York Times,

### 2014-07-23 writeups/dissertation/dissertation.tex
foo

@@
 
-\textcite{Anderson1991} also used tasks similar to tag use on StackOverflow and Twitter when deriving the prior component of the declarative memory theory.
+\textcite{Anderson1991} also looked at tasks similar to tag use on StackOverflow and Twitter when deriving the prior component of the declarative memory theory.
 \citeauthor{Anderson1991} examined, for example, the frequency of word choice over time used for the title of articles for the New York Times,

### 2014-07-23 writeups/dissertation/dissertation.tex
foo

@@
 The entire StackOverflow dataset is packaged and released quarterly to the public.
-So this package was downloaded and the database tables were imported into PostgreSQL using \textcite{so2db} importing tool.
+So this package was downloaded and the database tables were imported into PostgreSQL.
 
@@
 Twitter provides two types of APIs for acquiring data: A query-based and streaming interface.
-The \textcite{Tweepy} Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
+The Tweepy Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
 The query-based interface was used to collect all tweets for specific popular users.
@@
 
-The base-level ($B_{i}$) component in Table \ref{ACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{ModACTRModel}).
+The base-level ($B_{i}$) component in Table \ref{tabACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{tabModACTRModel}).
 \citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log odds computation of the probability that each chunk is likely to be presented again, given past frequency.
@@
 To better illustrate the dynamics of each user's tagging behavior,
-profile plots for a single user for both StackOverflow and Twitter are included in Figures \ref{figHTByTime-520957} and \ref{figHTByTime-fashionista_com} respectively.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTByTime-520957-crop.pdf}}}
-  \caption{Tagging profile of a single user on StackOverflow}
-  \label{figHTByTime-520957}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTByTime-fashionista_com-crop.pdf}}}
-  \caption{Tagging profile of a single user on Twitter}
-  \label{figHTByTime-fashionista_com}
+profile plots for a single user for both StackOverflow and Twitter are included in Figure \ref{figHTByTime}.
+
+\begin{figure}[!htbp]
+  {%
+    \setlength{\fboxsep}{0pt}%
+    \setlength{\fboxrule}{1pt}%
+    \hfill
+    \subfigure[][StackOverflow]{\includegraphics[width=8cm]{HTByTime-520957-crop.pdf}}
+    \hfill
+    \subfigure[][Twitter]{\includegraphics[width=8cm]{HTByTime-fashionista_com-crop.pdf}}
+    \caption{Tagging profile for a single user}
+    \label{figHTByTime}
+  }%
 \end{figure}
@@
 These plots for StackOverflow and Twitter are included in Figures \ref{figHTMByTime-520957} and \ref{figHTMByTime-fashionista_com} respectively.
-These are the same plots shown in Figures \ref{figHTByTime-520957} and \ref{figHTByTime-fashionista_com} with model predictions overlaid in red.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
+These are the same plots shown in Figure \ref{figHTByTime} with model predictions overlaid in red.
+
+\begin{figure}[!htbp]
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
   \caption{Model tagging profile of a single user on StackOverflow}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
   \caption{Model tagging profile of a single user on Twitter}
@@
 Moving up one level of analysis, model accuracy for all users in a single dataset slice was analyzed.
-A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDs} and \ref{figPriorTwitterSliceDs} respectively.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}}
+A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDsStd} and \ref{figPriorTwitterSliceDsStd} respectively.
+
+\begin{figure}[!htbp]
+  {%
+    \setlength{\fboxsep}{0pt}%
+    \setlength{\fboxrule}{1pt}%
+    \hfill
+    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
+    \hfill
+    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
     \caption{Model performance for a single dataset slice for StackOverflow}
     \label{figPriorSOQSliceDsStd}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}}
-  \caption{Model performance for a single dataset slice for StackOverflow}
-  \label{figPriorSOQSliceDsOL}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}}
+  }%
+\end{figure}
+
+\begin{figure}[!htbp]
+  {%
+    \setlength{\fboxsep}{0pt}%
+    \setlength{\fboxrule}{1pt}%
+    \hfill
+    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}
+    \hfill
+    \subfigure[][Optmized learning form]{\includegraphics[width=8cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
     \caption{Model performance for a single dataset slice for Twitter}
     \label{figPriorTwitterSliceDsStd}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}}
-  \caption{Model performance for a single dataset slice for Twitter}
-  \label{figPriorTwitterSliceDsOL}
+  }%
 \end{figure}
@@
 To identify the best decay rate value for the dataset slices,
-each user's best decay rate value was taken and the results are plotted in histogram form in Figures \ref{figPriorSOQHistDs} and \ref{figPriorTwitterSliceDs} respectively.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}}
-  \caption{Best-fit decay rate for a single dataset slice for StackOverflow}
-  \label{figPriorSOQHistDs}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}}
+each user's best decay rate value was taken and the results are plotted in histogram form in Figures \ref{figPriorSOQHistDsOL} and \ref{figPriorTwitterHistDsOL} respectively.
+
+\begin{figure}[!htbp]
+  {%
+    \setlength{\fboxsep}{0pt}%
+    \setlength{\fboxrule}{1pt}%
+    \hfill
+    \subfigure[][Standard form]{\includegraphics[width=8cm]{visHistD-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
+    \hfill
+    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visHistD-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
     \caption{Best-fit decay rate for a single dataset slice for StackOverflow}
     \label{figPriorSOQHistDsOL}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}}
-  \caption{Best-fit decay rate for a single dataset slice for Twitter}
-  \label{figPriorTwitterHistDs}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}}
+  }%
+\end{figure}
+
+\begin{figure}[!htbp]
+  {%
+    \setlength{\fboxsep}{0pt}%
+    \setlength{\fboxrule}{1pt}%
+    \hfill
+    \subfigure[][Standard form]{\includegraphics[width=8cm]{visHistD-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}
+    \hfill
+    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visHistD-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
     \caption{Best-fit decay rate for a single dataset slice for Twitter}
     \label{figPriorTwitterHistDsOL}
+  }%
 \end{figure}
@@
 
-The impact of adding the offset to each of the models for StackOverflow and Twitter are included in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetTwitter}.
+The impact of adding the offset to each of the models for StackOverflow and Twitter are included in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.
 
@@
   \caption{Stop-word techniques for the random-permutation model for Twitter}
-  \label{figContextStopRPSO}
+  \label{figContextStopRPT}
 \end{figure}
@@
   \caption{Stop-word techniques for the Bayesian model for Twitter}
-  \label{figContextStopBSO}
+  \label{figContextStopBT}
 \end{figure}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-orderT-crop.pdf}}}
   \caption{Model accuracy for various word order methods for the random-permutation model for Twitter}
-  \label{figContextLogoddsT}
+  \label{figContextOrder}
 \end{figure}
@@
 
-Model accuracy when using a user-customized co-occurrence matrix for StackOverflow and Twitter is shown in Figures \ref{figPUserCustomSO} and \ref{figPUserCustomT}.
+Model accuracy when using a user-customized co-occurrence matrix for StackOverflow and Twitter is shown in Figures \ref{figPUCustomSO} and \ref{figPUCustomT}.
 

### 2014-07-23 writeups/dissertation/dissertation.tex
foo

@@
 and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high-performance desktop, even with highly optimized representations.
-Already FIXME total co-occurring word tag observations are used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
+Already \num{835500000} total co-occurring word tag observations are used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
 It is also the case that \num{1000000} posts for StackOverflow is already around $1/7$ of the total posts created on the site,
 so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrence matrix instead,
-especially since FIXME co-occurrences are already used with \num{1000000} posts.
+especially since \num{835500000} co-occurrences are already used with \num{1000000} posts.
 
@@
 The fact that the random-permutation model only needs around \num{2048} rows to represent all of the words in the Twitter and StackOverflow datasets is impressive.
-Twitter has around FIXME unique words for the \num{3000000} tweets in each of the popular-users datasets.
-Stackoverflow has FIXME unique words for the \num{1000000} posts used to generate the co-occurrence matrix.
+Twitter has around \num{2383000} unique words for the \num{3000000} tweets in each of the popular-users datasets.
+Stackoverflow has \num{5848000} unique words for the \num{1000000} posts used to generate the co-occurrence matrix.
 So it is impressive that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
@@
 The StackOverflow randomly-sampled dataset selects posts across the entire dataset, regardless of each user's popularity or tag history.
-Across the dataset, each user only asks FIXME questions on average, so there is not a large amount of past tag history available for the prior component to use,
+Across the dataset, each user only asks around five questions on average, so there is not a large amount of past tag history available for the prior component to use,
 which is why this component is not a strong predictor in this case.
@@
 However, since computing activations for the context component take an order of magnitude more wall-clock time than computing the prior component, not all observations could be included for the test.
-To reduce the observations, the most recent FIXME posts from each user in the datasets were sampled.
+To reduce the observations, the most recent 50 posts from each user in the datasets were sampled.
+The total number of users sampled was also reduced to 40 users for StackOverflow and 80 users for Twitter.
 Using the most recent observations for each user ensures that the observations tested have acquired the greatest amount of past tagging history in the datasets.
@@
 but the oldest posts that were not included in the testing set were used to build the co-occurrence matrices.
-So for each user, the co-occurrence matrix was created from their oldest posts, up to (and not including) their FIXME newer posts that were used to test model accuracy.
+So for each user, the co-occurrence matrix was created from their oldest posts, up to (and not including) their 50 newer posts that were used to test model accuracy.
 

### 2014-07-23 writeups/dissertation/dissertation.tex
foo

@@
 To better illustrate the dynamics of each user's tagging behavior,
-profile plots for a single user for both StackOverflow and Twitter are included in Figure \ref{figHTByTime}.
+profile plots for a single user for both StackOverflow and Twitter are included in Figure \ref{figPriorHTByTime}.
 
@@
     \caption{Tagging profile for a single user}
-    \label{figHTByTime}
+    \label{figPriorHTByTime}
   }%
@@
 As a first method for understanding model performance, profile plots of model's predicted tags for individual users were overlaid on that user's actual tag use.
-These plots for StackOverflow and Twitter are included in Figures \ref{figHTMByTime-520957} and \ref{figHTMByTime-fashionista_com} respectively.
-These are the same plots shown in Figure \ref{figHTByTime} with model predictions overlaid in red.
+These plots for StackOverflow and Twitter are included in Figures \ref{figPriorHTMByTimeSO} and \ref{figPriorHTMByTimeT} respectively.
+These are the same plots shown in Figure \ref{figPriorHTByTime} with model predictions overlaid in red.
 
@@
   \caption{Model tagging profile of a single user on StackOverflow}
-  \label{figHTMByTime-520957}
+  \label{figPriorHTMByTimeSO}
 \end{figure}
@@
   \caption{Model tagging profile of a single user on Twitter}
-  \label{figHTMByTime-fashionista_com}
+  \label{figPriorHTMByTimeT}
 \end{figure}
@@
 
-The same previously-derived stoplist from FIXME that \textcite{Sahlgren2008} used was used for this analysis,
+The same previously-derived stoplist from \textcite{Salton2014} that \textcite{Sahlgren2008} used was used for this analysis,
 The same entropy weighting method using by \textcite{Stanley2013} was examined,
@@
 
-The plots for the standard deviations across tags that was used to identify the cutoff for StackOverflow and Twitter are included in Figures \ref{figContextCutoffSO} and \ref{figContextCuttoffT}.
-
-FIXME:
-
-The no-removal plot shows the size of the spikes when high-frequency words are removed.
+The plots for the standard deviations across tags that were used to identify the cutoff for StackOverflow and Twitter are included in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT}.
+
+\begin{figure}[!htbp]
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{memMatSO-crop.pdf}}}
+  \caption{Variance in observation counts of words for each tag for StackOverflow}
+  \label{figContextCutoffSO}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{memMatT-crop.pdf}}}
+  \caption{Variance in observation counts of words for each tag for Twitter}
+  \label{figContextCutoffT}
+\end{figure}
+
+The original plot shows the size of the spikes when high-frequency words are removed.
 The entropy plot shows how the size decreases by more than an order of magnitude when the entropy weighting measure is used.
@@
 Note that using a previously-defined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
-The cutoff that produced the frequency plot is where a word represents more than \num{.0004} occurrences in the dataset.
-This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset were not given.
+The cutoff that produced the frequency plot is where a word represents more than \num{.04}\% of total occurrences in the dataset.
+This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset was not given.
 
@@
 
-Model accuracy as a function of co-occurrence matrix size for StackOverflow is included in Figure \ref{figContextDSizeSO}.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
+Model accuracy as a function of co-occurrence matrix size for StackOverflow is included in Figure \ref{figContextDocumentSizeSO}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-freqVsEntropyBySizeSO-crop.pdf}}}
   \caption{Effect of size of co-occurrence matrix for StackOverflow}
-  \label{figContextDSizeSO}
+  \label{figContextDocumentSizeSO}
 \end{figure}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-freqVsEntropyBySizeT-crop.pdf}}}
   \caption{Effect of size of co-occurrence matrix for Twitter}
-  \label{figContextDSizeSO}
-\end{figure}
-
-Both entropy weighting and frequency filtering in Figure \ref{figContextDSizeSO} plateau for the random-permutation model with increasing document size.
+  \label{figContextDocumentSizeT}
+\end{figure}
+
+Both entropy weighting and frequency filtering in Figure \ref{figContextDocumentSizeT} plateau for the random-permutation model with increasing document size.
 Once again, the Bayesian model continues to improve with document size.
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
   \caption{Random-permutation model accuracy when using log-odds transformation on context for Twitter}

### 2014-07-23 dissProject/preProcess.R

@@ -2619,27 +2619,31 @@ makeCombinedEnvironmentTbl <- function(sjiTbl, config) {
 }
 
 plotMemMat <- function() {
+	plotTbl <- function(tbl, name) {
+		tbl[, id := 1:nrow(.SD)]
+		tbl = melt(tbl,
+			   measure=c('sdStoplist', 'sdFreq', 'sdEntropy', 'sdOrig'),
+			   variable.name='type',
+			   value.name='sd')
+		assign('p1', ggplot(tbl, aes(id, sd)) +
+		       geom_point() +
+		       defaultGGPlotOpts + 
+		       facet_grid(.~type) + 
+		       theme(axis.text.x = element_blank()))
+		myPlotPrint(p1, name)
+	}
 	sdStoplist = apply(permMemMatSOOrderless$stoplist, 2, sd)
 	sdFreq = apply(permMemMatSOOrderless$freq, 2, sd)
 	sdEntropy = apply(permMemMatSOOrderless$entropy, 2, sd)
 	sdOrig = apply(permMemMatSOOrderless$orig, 2, sd)
+	tbl = data.table(sdStoplist, sdFreq, sdEntropy, sdOrig)
+	plotTbl(tbl, 'memMatSO')
 	sdStoplist = apply(permMemMatTOrderless$stoplist, 2, sd)
 	sdFreq = apply(permMemMatTOrderless$freq, 2, sd)
 	sdEntropy = apply(permMemMatTOrderless$entropy, 2, sd)
 	sdOrig = apply(permMemMatTOrderless$orig, 2, sd)
-	permEnvTblSO[, .N, by=stopWordWeight]
-	permEnvTblSO[, .N, by=freqWeight]
-	permEnvTblT[, .N, by=stopWordWeight]
-	permEnvTblT[, .N, by=freqWeight]
-	permEnvTblSO
-	sd(sdOrig)
-	sd(sdEntropy)
-	sd(sdFreq)
-	sd(sdStoplist)
-	plot(sdOrig)
-	plot(sdStoplist)
-	plot(sdEntropy)
-	plot(sdFreq)
+	tbl = data.table(sdStoplist, sdFreq, sdEntropy, sdOrig)
+	plotTbl(tbl, 'memMatT')
 }
 
 getCurWorkspace <- function(groupConfig) {
@@ -3570,6 +3574,9 @@ runGenAndSaveCurWorkspaceg3s6 <- function() genAndSaveCurWorkspace(groupConfigG3
[…]

### 2014-07-24 dissProject/preProcess.R
foo

@@ -2632,18 +2632,16 @@ plotMemMat <- function() {
 		       theme(axis.text.x = element_blank()))
 		myPlotPrint(p1, name)
 	}
-	sdStoplist = apply(permMemMatSOOrderless$stoplist, 2, sd)
-	sdFreq = apply(permMemMatSOOrderless$freq, 2, sd)
-	sdEntropy = apply(permMemMatSOOrderless$entropy, 2, sd)
-	sdOrig = apply(permMemMatSOOrderless$orig, 2, sd)
-	tbl = data.table(sdStoplist, sdFreq, sdEntropy, sdOrig)
-	plotTbl(tbl, 'memMatSO')
-	sdStoplist = apply(permMemMatTOrderless$stoplist, 2, sd)
-	sdFreq = apply(permMemMatTOrderless$freq, 2, sd)
-	sdEntropy = apply(permMemMatTOrderless$entropy, 2, sd)
-	sdOrig = apply(permMemMatTOrderless$orig, 2, sd)
-	tbl = data.table(sdStoplist, sdFreq, sdEntropy, sdOrig)
-	plotTbl(tbl, 'memMatT')
+	buildTbl <- function(mat) {
+		sdStoplist = apply(mat$stoplist, 2, sd)
+		sdFreq = apply(mat$freq, 2, sd)
+		sdEntropy = apply(mat$entropy, 2, sd)
+		sdOrig = apply(mat$orig, 2, sd)
+		tbl = data.table(sdStoplist, sdFreq, sdEntropy, sdOrig)
+		tbl
+	}
+	plotTbl(buildTbl(permMemMatSOOrderless), 'memMatSO')
+	plotTbl(buildTbl(permMemMatTOrderless), 'memMatT')
 }
 
 getCurWorkspace <- function(groupConfig) {

### 2014-07-24 dissProject/preProcess.R
foo

@@ -3572,9 +3572,6 @@ runGenAndSaveCurWorkspaceg3s6 <- function() genAndSaveCurWorkspace(groupConfigG3
 runGenAndSaveCurWorkspaceg4s6 <- function() genAndSaveCurWorkspace(groupConfigG4S6)
 
 curWS <- function() {
-	sqldf('select owner_user_id, count(id) from posts where post_type_id = 1 group by owner_user_id')
-	tbl = sqldt('select id, num_questions from users')
-	tbl[!is.na(num_questions)][, median(num_questions)]
 	withProf(runContext20g1s6(regen='useAlreadyLoaded'))
 	runContext500g1s6(regen='useAlreadyLoaded', numRunsT=1, numRunsSO=1)
 	runContext500g1s6(regen=F, numRunsT=1, numRunsSO=1)

### 2014-07-24 dissProject/preProcess.R
foo

@@ -3582,7 +3582,6 @@ curWS <- function() {
 	withProf(runContext20g1s1(regen='useAlreadyLoaded'))
 	setLogLevel(2)
 	modelVsPredTbl = buildTables(file_path_sans_ext(Filter(isContextRun, list.files(path=getDirModelVsPred()))))
-	modelVsPredTbl.new = copy(modelVsPredTbl)
 	modelVsPredTbl = buildTables(file_path_sans_ext(Filter(isPriorRun, list.files(path=getDirModelVsPred()))))
 	modelVsPredTbl = buildTables(file_path_sans_ext(Filter(isPUserRun, list.files(path=getDirModelVsPred()))))
 	logregTbl = buildTablesLogreg(file_path_sans_ext(Filter(function(r) isPUserRun(r) | isContextRun(r), list.files(path=getDirLogreg()))))
@@ -3600,10 +3599,7 @@ curWS <- function() {
 	withProf(myLoadImage(groupConfigG1S3))
 	withProf(myLoadImage(groupConfigPUserS6))
 	test_dir(sprintf("%s/%s", PATH, 'tests'), reporter='summary')
-	tables()
 	.ls.objects(order.by='Size')
-	lapply(permMemMatSOOrderless, dim)
-	as.sparseMatrix(permMemMatSOOrderless[['orig']])
 	# Checking that tweets for twitter users from each followers_count,statuses_count scale are being collected properly
 	usersWithTweetsTbl = sqldt("select distinct on (t.user_screen_name) t.user_screen_name,u.followers_count,u.statuses_count
 				   from tweets as t join twitter_users as u on t.user_screen_name = u.user_screen_name"

### 2014-07-24 writeups/dissertation/dissertation.tex
foo

@@
 \abstract{
-  This research aims to develop and compare tag-recommendation models for users choosing hashtags when composing a tweet on Twitter and choosing tags when asking a question on the StackOverflow site.
-  Two state-of-the-art cognitively-plausible memory models will be evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
-  Experiments are designed to [1] discover the best way to leverage a user's prior hashtag use as a future predictor, [2] improve the models so that each model's strengths are incorporated into the other,
-  and [3] formally evaluate and compare each model's performance when recommending hashtags on Twitter and tags on StackOverflow.
-  The user-customized tag-recommendation models developed with this research should be immediately useful as foundations for tag-recommendation systems on the popular Twitter and StackOverflow social media sites.
-  Further, the planned performance modifications to the two memory models are aimed at the architectural level of each model.
-  If these architectural modifications to two cognitively-plausible memory models significantly improve performance,
-  they may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
+  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
+  The task of choosing a hashtag for a tweet and tagging a post on StackOverflow was framed as a declarative memory retrieval problem.
+  Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
+  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
+  The results show that past user behavior of tag use is a strong predictor of future performance, and this past behavior was successfully incorporated into the random-permutation model that previously used only context.
+  Also, the common natural-language processing method to attenuate low-predictor words in the dataset was linked to ACT-R's attentional weight term,
+  and an entropy-weighting technique was developed that improved model accuracy by more efficiently allocating limited attentional resources to the predictive words in the posts and tweets.
+  Word order was not found to be strong predictor of tag use, and the random-permutation model performed comparable to the Bayesian model without including word order,
+  which shows that the strength of the random-permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
+  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
+  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
 }
@@
 The entropy plot shows how the size decreases by more than an order of magnitude when the entropy weighting measure is used.
-The cutoff was chosen such that the frequency plot looks similar to the entropy plot, with spikes around the same size.
+The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
 Note that using a previously-defined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.

### 2014-07-28 writeups/dissertation/dissertation.tex
foo

@@
   Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
-  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
-  The results show that past user behavior of tag use is a strong predictor of future performance, and this past behavior was successfully incorporated into the random-permutation model that previously used only context.
+  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random permutation vector-based model.
+  The results show that past user behavior of tag use is a strong predictor of future performance, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
   Also, the common natural-language processing method to attenuate low-predictor words in the dataset was linked to ACT-R's attentional weight term,
   and an entropy-weighting technique was developed that improved model accuracy by more efficiently allocating limited attentional resources to the predictive words in the posts and tweets.
-  Word order was not found to be strong predictor of tag use, and the random-permutation model performed comparable to the Bayesian model without including word order,
-  which shows that the strength of the random-permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
+  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparable to the Bayesian model without including word order,
+  which shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
   The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
@@
 It will focus specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutations) can be improved, by testing these models on two real-world tagging tasks.
-One primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory when large datasets are used for each model. 
+One primary objective is to compare the random permutation vector-based model to ACT-R's declarative memory theory when large datasets are used for each model. 
 Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
@@
 The core research question motivating this proposal is:
-What kinds of cognitively-plausible models can best predict the human-created tags that a user on a social media site uses?
+What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
 This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
 This process can be framed as a memory retrieval problem.
-That is, each user on a social media site has created content that contain a set of human-created tags within the context of each post.
+That is, each user on a social media site has created content that contain a set of tags within the context of each post.
 The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.
@@
 So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
-Some recent and often-used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.
+Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.
 
@@
 I will extend the work done by \citeauthor{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
-I will also replace the vector-based model used by \citeauthor{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
+I will also replace the vector-based model used by \citeauthor{Rutledge2008} with an improved and simplified random permutation model \parencite{Sahlgren2008}.
 
 An advantage of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
-Word-order information on sites like Twitter may contain highly-predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
+Word order information on sites like Twitter may contain highly predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
 However, it is unclear and unexplored how word order can be represented in the ACT-R declarative memory theory.
@@
 
-The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base level activation ($B_{i}$) and strength of association ($S_{ji}$).
+The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base-level activation ($B_{i}$) and strength of association ($S_{ji}$).
 The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
@@
 Base-level activation reflects the log prior odds of needing an observed chunk again.
-The primary way to calculate these log prior odds is to use the standard base level equation in Table \ref{tabACTRModel}.
+The primary way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
 This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
@@
 However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
-In this way the standard base level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its current frequency and recency of use.
+In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its current frequency and recency of use.
 
 For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable, and do not change within that window.
-Programming-language popularity over the past few years is a reasonable example of this.
+Programming language popularity over the past few years is a reasonable example of this.
 Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., Clojure's growth), it is not the case that the changes have been drastic.
@@
 Further, if this is a reasonable assumption to make, it greatly simplifies the computation of the base-level activation for chunks, and can turn the computation into a tractable problem for large datasets.
-In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log odds ratio of exposure to that chunk compared to exposure to all other chunks.
+In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log-odds ratio of exposure to that chunk compared to exposure to all other chunks.
 For example, if the StackOverflow tag \emph{PHP} has been used four times as often as the tag \emph{Common Lisp}, then the prior odds of needing \emph{PHP} again is $\frac{.8}{.2}=4$ times that of \emph{Common Lisp}.
@@
 
-Strength of association ($S_{ji}$) reflects the amount of log odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
+Strength of association ($S_{ji}$) reflects the amount of log-odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
 Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
@@
 The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
-The best way to remove a dimension while maintaining the most variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
+The best way to remove a dimension while maintaining the most variability is to look for words with highly similar co-occurrence patterns and deduce that these words represent the same concept.
 For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
@@
 
-When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
-This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
+When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-sized datasets, LSA performed much better (23\% vs. 60\%).
+This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model outperformed the full dimensionality model on TASA-sized datasets.
 However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
@@
 As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
-Simply counting co-occurrences of the two immediately-preceding words and hashtags does not include the order information of the words in the representation.
+Simply counting co-occurrences of the two immediately preceding words and hashtags does not include the order information of the words in the representation.
 With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
@@
 Also, BEAGLE with unordered context and word order was more correlated to the WordNet synonym database than LSA (-.311 compared to -.165).
-Further, it was shown that incorporating word-order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
+Further, it was shown that incorporating word order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
 This model provided an efficient algorithm for using a single representation for unordered and ordered information that showed performance similar to if not better than LSA.
@@
 The vector length by performance tradeoff was tested on a simple paired associative retrieval task. 
-Retrieval accuracy for the random-permutation vector-based models remained high relative to the circular convolution model even after reducing each environment vector's length from \num{2048} to \num{1024}.
+Retrieval accuracy for the random permutation vector-based models remained high relative to the circular convolution model even after reducing each environment vector's length from \num{2048} to \num{1024}.
 This indicates that vector-based representations using random permutations compress better than those using circular convolution.
@@
 To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFL) assessment (among other tasks).
-The random-permutation model was trained on \num{2.33} GB corpus.
+The random permutation model was trained on \num{2.33} GB corpus.
 The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
@@
 The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
-These results make a strong case to favor random-permutations over circular convolution when implementing vector-based memory representations.
+These results make a strong case to favor random permutations over circular convolution when implementing vector-based memory representations.
 
@@
 This does not seem to be the case in testing however, as vector-based approaches like BEAGLE and random permutations perform just as well as the LSA when trained and tested on similar datasets.
-Nonetheless, it has been suggested to perform an SVD reduction technique on a compressed random-permutation vector \parencite{Kanerva2000}.
+Nonetheless, it has been suggested to perform an SVD reduction technique on a compressed random permutation vector \parencite{Kanerva2000}.
 This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.
@@
 With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
-It may be that multiple co-occurrence matrices must be used to represent word-order co-occurrences and unordered co-occurrences separately.
+It may be that multiple co-occurrence matrices must be used to represent ordered and unordered co-occurrences separately.
 However, it would be more parsimonious if these matrices could be combined into a single matrix representation, much like how this information is combined for vector-based models.
@@
 The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
-Netflix is an online-subscription service where users can stream their favorite television episodes and movies.
+Netflix is an online subscription service where users can stream their favorite television episodes and movies.
 Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
-In order to provide a high-quality user experience on the site, Netflix created a content-recommendation service.
+In order to provide a high quality user experience on the site, Netflix created a content-recommendation service.
 This tool recommends to the user movies and television episodes that most likely match their specific interests.
@@
 This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
-By using a factor-analysis technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.
+By using a factor analysis technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.
 
@@
 
-Lucene \parencite{McCandless2010} is a commonly-used open-source software tool that measures the similarity between a query and candidate documents in a database.
+Lucene \parencite{McCandless2010} is a commonly-used open source software tool that measures the similarity between a query and candidate documents in a database.
 It uses the term frequency---inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
@@
 The activation of each tag ($A_{i}$) is a function of four components:
-That tag's base level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
+That tag's base-level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
 This model has been modified from the standard ACT-R model (Table \ref{tabACTRModel}) in three ways:
@@
 To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
-It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.
+It will be interesting to see how much (if at all) accuracy improves when incorporating word order information into the model.
 
@@
 To incorporate the user's prior tag history into the model, the base-level activation term was modified from the original equation in ACT-R.
-$B_{i}$ is based on log odds occurrence, which in turn is based on frequency of tag use, or probability of tag use $p_{i}$.
+$B_{i}$ is based on log-odds occurrence, which in turn is based on frequency of tag use, or probability of tag use $p_{i}$.
 To incorporate a user's prior, a weighted average for $p_{i}$ was used,
@@
 [1] figuring out the optimal way to combine global tag history with a user's past tagging history, and
-[2] using the time-dependent form of the base level component in Table \ref{tabACTRModel}, in order to account for the fact that a user's tagging preferences change over time.
+[2] using the time-dependent form of the base-level component in Table \ref{tabACTRModel}, in order to account for the fact that a user's tagging preferences change over time.
 So these proposed modifications will be explored further when incorporating a user's past tagging history into the models for this research.
@@
 with the StackOverflow dataset, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
-This can be thought of as semantic stemming, where two words with identical meaning are converted to an agreed-upon canonical word.
+This can be thought of as semantic stemming, where two words with identical meaning are converted to the same canonical word.
 So tag words in the post and used tags for the post were converted to their root tag when possible.
@@
 Stop words are commonly removed by either filtering based on a pre-determined list or filtering based on the high-frequency words that occur in the dataset.
-\textcite{Sahlgren2008} tried both methods to remove stop words for the random-permutation vector-based model, and found that the data-driven approach
+\textcite{Sahlgren2008} tried both methods to remove stop words for the random permutation vector-based model, and found that the data-driven approach
 (i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a pre-determined list.
@@
       \hline
-      Base Level Activation &	$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
+      Base-Level Activation &	$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
       Optimized Learning &	$B_{i} = log \frac{n}{1-d} - d * log L$ \\
@@
 It does not have to keep track of the time each specific chunk instance was observed.
-Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R common-lisp implementation.
+Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R Common Lisp implementation.
 
@@
 The base-level ($B_{i}$) component in Table \ref{tabACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{tabModACTRModel}).
-\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log odds computation of the probability that each chunk is likely to be presented again, given past frequency.
+\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log-odds computation of the probability that each chunk is likely to be presented again, given past frequency.
 If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log n_{i}$,
@@
 What is interesting is how large the tagging space is for these users.
-There are bands of commonly-used tags for each user, but there are also instances where new tags are used, and a large percentage of tags used are not in the most-common bands.
+There are bands of commonly-used tags for each user, but there are also instances where new tags are used, and a large percentage of tags used are not in the most common bands.
 This helps illustrate how challenging modeling this task can be.
@@
 For the standard $B_{i}$ equation, model accuracy for a pure frequency model ($d=0$) is also worse than using a pure recency model ($d=20$), particularly for Twitter.
-It makes sense that a pure recency model for Twitter is relatively more accurate than for StackOverflow, as the hashtag lifetime for Twitter is much less than a programming-language tag on StackOverflow.
+It makes sense that a pure recency model for Twitter is relatively more accurate than for StackOverflow, as the hashtag lifetime for Twitter is much less than a tag for a programming language on StackOverflow.
 
@@
 
-The context component for both the Random Permutation and Bayesian model requires a co-occurrence matrix of words and observed tags.
+The context component for both the random permutation and Bayesian model requires a co-occurrence matrix of words and observed tags.
 Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
 The matrix for \num{3000000} posts was only used for Twitter, since the number of words in a StackOverflow post is much higher than Twitter on average,
-and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high-performance desktop, even with highly optimized representations.
+and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high performance desktop, even with highly optimized representations.
 Already \num{835500000} total co-occurring word tag observations are used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
@@
 
-Model accuracy relaxed across posts with and without the offset model component was evaluated for both the Bayesian and random-permutation model.
+Model accuracy relaxed across posts with and without the offset model component was evaluated for both the Bayesian and random permutation model.
 
@@
 When they are enumerated, then components are left out.
-The random-permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
+The random permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
 So, for example, the ``standard Prior model relaxed across posts without offset'' means that only the prior model component was used, and the offset for that component was not subtracted off when determining activation.
@@
 
-The random-permutation model in particular is greatly affected by the offset term, for both the StackOverflow and Twitter.
-This suggests that the mean activation for the random-permutation model for the top few tags in each post can vary greatly across posts.
+The random permutation model in particular is greatly affected by the offset term, for both the StackOverflow and Twitter.
+This suggests that the mean activation for the random permutation model for the top few tags in each post can vary greatly across posts.
 Only if this mean activation is subtracted off will the logistic regression model generate optimal weights and predictions that are comparative to the Bayesian model when model accuracy relaxed across posts is examined.
@@
 
-A plot of model accuracy for components of both the Random Permutation and Bayesian models and isolated by Twitter popular-users subset are included in Figure \ref{figContextSubsets}.
+A plot of model accuracy for components of both the random permutation and Bayesian models and isolated by Twitter popular-users subset are included in Figure \ref{figContextSubsets}.
 
@@
 
-\textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random-permutation model:
+\textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
 A data-driven frequency-weighted approach was tried, along with using a previously-derived stoplist, and no stop word removal at all.
@@
 The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
-Note that using a previously-defined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
+Note that using a predefined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
 The cutoff that produced the frequency plot is where a word represents more than \num{.04}\% of total occurrences in the dataset.
@@
 
-Plots for the three methods of handling stop words for the random-permutation model on the StackOverflow dataset are included in Figure \ref{figContextStopRPSO}.
+Plots for the three methods of handling stop words for the random permutation model on the StackOverflow dataset are included in Figure \ref{figContextStopRPSO}.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsSOPerm-crop.pdf}}}
-  \caption{Stop-word techniques for the random-permutation model for StackOverflow}
+  \caption{Stop-word techniques for the random permutation model for StackOverflow}
   \label{figContextStopRPSO}
@@
 
-The results show that model accuracy improves if stop words are handled in any of the three ways (pre-defined stoplist filtering, entropy weighting, and frequency filtering).
+The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
 Using a pre-defined stoplist makes the smallest improvement, while the entropy and frequency techniques improve performance the most, and frequency shows a slight edge over entropy.
@@
 
-The three stop word techniques were then tested for the random-permutation model on the Twitter dataset, and the results are included in Figure \ref{figContextStopRPT}.
+The three stop word techniques were then tested for the random permutation model on the Twitter dataset, and the results are included in Figure \ref{figContextStopRPT}.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTPerm-crop.pdf}}}
-  \caption{Stop-word techniques for the random-permutation model for Twitter}
+  \caption{Stop-word techniques for the random permutation model for Twitter}
   \label{figContextStopRPT}
@@
 
-For the random-permutation model for Twitter, the entropy weighting technique is the best method for handling stop words.
+For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words.
 Removing stop-words based on a pre-determined list actually decreases accuracy, and using data-driven frequency filtering has only a small to no effect.
@@
 This threshold is a free parameter in frequency filtering that is not required when weighting stop words by entropy.
-Therefore, favoring parsimony, for the random-permutation model, attenuating stop words based on entropy produces more accurate results when compared to filtering the stop words based on frequency of use.
+Therefore, favoring parsimony, for the random permutation model, attenuating stop words based on entropy produces more accurate results when compared to filtering the stop words based on frequency of use.
 This result is interesting since filtering stop words based on frequency is the default method for random permutation models \parencite{Sahlgren2008}.
@@
 
-Since accuracy measurements when using a pre-determined list to remove stop words for the random-permutation model were much lower than the frequency and entropy methods, this method was not examined for the Bayesian model.
+Since accuracy measurements when using a pre-determined list to remove stop words for the random permutation model were much lower than the frequency and entropy methods, this method was not examined for the Bayesian model.
 The results for the frequency and entropy techniques for the Bayesian model on the StackOverflow dataset are include in Figure \ref{figContextStopBSO}.
@@
 Another interesting result was how the Bayesian model actually performed worse when stop words were removed based on frequency.
-This was not the case for the random-permutation model, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random-permutation model, and is most likely never the case for the random-permutation model.
-Model accuracy most likely decreases for the Bayesian model because this model already has a normalization component built in that the random-permutation model does not have.
-When computing activation for the random-permutation model, the correlation is calculated directly on a matrix of counts of co-occurrences for words and tags.
-However for the Bayesian model this count matrix is converted to a log odds ($S_{ji}$) matrix,
+This was not the case for the random permutation model, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random permutation model, and is most likely never the case for the random permutation model.
+Model accuracy most likely decreases for the Bayesian model because this model already has a normalization component built in that the random permutation model does not have.
+When computing activation for the random permutation model, the correlation is calculated directly on a matrix of counts of co-occurrences for words and tags.
+However for the Bayesian model this count matrix is converted to a log-odds ($S_{ji}$) matrix,
 where both the number of observations for the word and the tag are normalized when computing the value for each cell (see the $S_{ji}$ computation in Table \ref{tabModACTRModel}).
@@
 This was clearly the case for the Bayesian model.
-Both frequency and entropy techniques performed equally for the random-permutation model.
-However since the entropy-weighting technique is parameter free and does not require a cutoff to tune, it is also a better choice for the random-permutation model for these datasets.
+Both frequency and entropy techniques performed equally for the random permutation model.
+However since the entropy-weighting technique is parameter free and does not require a cutoff to tune, it is also a better choice for the random permutation model for these datasets.
 
@@
 co-occurrence matrices for a range of smaller document sizes were built.
-Both the random-permutation and Bayesian models were tested on all document sizes.
+Both the random permutation and Bayesian models were tested on all document sizes.
 
@@
 
-Accuracy improves for both the Bayesian and random-permutation models as the size of the documents used to generate the co-occurrence matrix increases.
-However model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random-permutation model, while the frequency technique only begins to plateau.
+Accuracy improves for both the Bayesian and random permutation models as the size of the documents used to generate the co-occurrence matrix increases.
+However model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random permutation model, while the frequency technique only begins to plateau.
 The Bayesian model does not plateau and accuracy continues to improves document size increases.
@@
 
-It is likely that the random-permutation model is saturating when entropy weighting is used for the largest document size for StackOverflow.
+It is likely that the random permutation model is saturating when entropy weighting is used for the largest document size for StackOverflow.
 Even if the non-predictive words based on entropy are attenuated,
-when the high-frequency words are included random-permutation model,
+when the high-frequency words are included random permutation model,
 this attenuation is not enough to overcome the range effects that grow with document size.
@@
 
-Both entropy weighting and frequency filtering in Figure \ref{figContextDocumentSizeT} plateau for the random-permutation model with increasing document size.
+Both entropy weighting and frequency filtering in Figure \ref{figContextDocumentSizeT} plateau for the random permutation model with increasing document size.
 Once again, the Bayesian model continues to improve with document size.
@@
 
-It is interesting to compare model accuracy between the Bayesian and random-permutation model at different document sizes for both StackOverflow and Twitter.
-At smaller document sizes, the random-permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as document size increases.
-In other words, the Bayesian model needs larger co-occurrence matrices than the random-permutation model to work properly.
+It is interesting to compare model accuracy between the Bayesian and random permutation model at different document sizes for both StackOverflow and Twitter.
+At smaller document sizes, the random permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as document size increases.
+In other words, the Bayesian model needs larger co-occurrence matrices than the random permutation model to work properly.
 As the size of the dataset increases, the compression may start to lose crucial information, and the uncompressed Bayesian model should be used.
-As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random-permutation model is actually beneficial, and the compressed random-permutation model should be used.
+As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random permutation model is actually beneficial, and the compressed random permutation model should be used.
 This crosstalk and noise is beneficial because it helps soften spurious $S_{ji}$ values when the values for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller document sizes.
@@
 The document size is rarely chosen, and more likely constrained by the number of documents in the dataset being studied.
-If that dataset is relatively small, it may be better to use the compressed random-permutation model than the Bayesian model, and vice-versa if the dataset is large.
+If that dataset is relatively small, it may be better to use the compressed random permutation model than the Bayesian model, and vice-versa if the dataset is large.
 
@@
 
-Another parameter in the random-permutation model is the number of rows used to represent the all of the words.
+Another parameter in the random permutation model is the number of rows used to represent the all of the words.
 Each word is represented by two randomly placed ones and two negative ones across the rows.
@@
 
-Three different row-size values were tested for the random-permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
+Three different row-size values were tested for the random permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
 A low value (\num{100}), a more standard value (\num{2048}), and a high value (\num{10000}) of rows were tested.
@@
 
-Results for using different levels of compression for the random-permutation model for the StackOverflow dataset are included in Figure \ref{figContextDimSO}.
+Results for using different levels of compression for the random permutation model for the StackOverflow dataset are included in Figure \ref{figContextDimSO}.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
-  \caption{Effect of compression on random-permutation model for StackOverflow}
+  \caption{Effect of compression on random permutation model for StackOverflow}
   \label{figContextDimSO}
@@
 In the figure, the full random permutation model that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
-This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random-permutation model for all other results.
+This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random permutation model for all other results.
 The plot is also broken out for different size datasets used to build the co-occurrence matrix.
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeT-crop.pdf}}}
-  \caption{Effect of compression on random-permutation model for Twitter}
+  \caption{Effect of compression on random permutation model for Twitter}
   \label{figContextDimT}
@@
 
-The fact that the random-permutation model only needs around \num{2048} rows to represent all of the words in the Twitter and StackOverflow datasets is impressive.
+The fact that the random permutation model only needs around \num{2048} rows to represent all of the words in the Twitter and StackOverflow datasets is impressive.
 Twitter has around \num{2383000} unique words for the \num{3000000} tweets in each of the popular-users datasets.
@@
 
-However it should be pointed out that the way the random-permutation model compresses the words is completely random.
+However it should be pointed out that the way the random permutation model compresses the words is completely random.
 It is not an LSA approach where the few ones and negative ones for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
-Everything is randomly chosen with the random-permutation model.
+Everything is randomly chosen with the random permutation model.
 The environment vector for each word is randomly selected, and no SVD is performed to choose these in a more discriminating manner.
@@
 The activation value for the context component of the vector-based model is a correlation, ranging from zero to one.
-This is not a log odds value.
-Currently, prior and context for the random-permutation model are combined by simply adding the terms.
-However, this means that ACT-R's prior base-level learning component (which is a log odds value) is added to a correlation value, which may not be mathematically appropriate.
+This is not a log-odds value.
+Currently, prior and context for the random permutation model are combined by simply adding the terms.
+However, this means that ACT-R's prior base-level learning component (which is a log-odds value) is added to a correlation value, which may not be mathematically appropriate.
 It would be more appropriate to convert the correlation value to a log-odds value, and then add this context term to the log-odds prior term.
@@
 
-For a given retrieval, the random-permutation model returns a set of activations that are correlation values for each of the tags.
+For a given retrieval, the random permutation model returns a set of activations that are correlation values for each of the tags.
 So a method was identified to convert a distribution of correlations to a distribution of log-odds values.
 To do this, the one-tailed cumulative distribution function value for each correlation is computed.
-That probability is then converted to a log odds value in the usual way.
+That probability is then converted to a log-odds value in the usual way.
 
@@
 Also, it behaves properly at specific points in the distribution.
-For example, if the conversion is performed one the mean correlation value in the set, the result is a log odds value of zero.
+For example, if the conversion is performed one the mean correlation value in the set, the result is a log-odds value of zero.
 This makes sense given that log odds can be interpreted as the amount of odds information (positive or negative) that should be adjusted to the prior log odds.
-Since the mean correlation value contains no information regarding if it is better or worse than the other correlation values, it is appropriate to assign this correlation value a zero log odds adjustment value.
-
-The log-odds transformation technique was tested for the random-permutation model on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.
+Since the mean correlation value contains no information regarding if it is better or worse than the other correlation values, it is appropriate to assign this correlation value a zero log-odds adjustment value.
+
+The log-odds transformation technique was tested for the random permutation model on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
-  \caption{Random-permutation model accuracy when using log-odds transformation on context for StackOverflow}
+  \caption{Random permutation model accuracy when using log-odds transformation on context for StackOverflow}
   \label{figContextLogoddsSO}
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
-  \caption{Random-permutation model accuracy when using log-odds transformation on context for Twitter}
+  \caption{Random permutation model accuracy when using log-odds transformation on context for Twitter}
   \label{figContextLogoddsT}
@@
 
-Vector-based models such as BEAGLE \parencite{Jones2007} and the random-permutation model \parencite{Sahlgren2008} have included only a context component in the past (i.e., no prior component).
-When the prior component is included for the random-permutation model, accuracy improves significantly (compare accuracy for full models to context models in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}).
-However, it is somewhat surprising that it does not matter much how the context and prior terms are added for the random-permutation model.
-Model performance is similar when the context component used for the model is based on a correlation compared to a log odds value.
+Vector-based models such as BEAGLE \parencite{Jones2007} and the random permutation model \parencite{Sahlgren2008} have included only a context component in the past (i.e., no prior component).
+When the prior component is included for the random permutation model, accuracy improves significantly (compare accuracy for full models to context models in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}).
+However, it is somewhat surprising that it does not matter much how the context and prior terms are added for the random permutation model.
+Model performance is similar when the context component used for the model is based on a correlation compared to a log-odds value.
 
@@
 But that result is not statistically reliable for the sample size used in this analysis, and even if present the effect size is small.
-Nonetheless, it does seem appropriate to convert the context to a log odds adjustment value when context is added as an additional term to the random-permutation model.
+Nonetheless, it does seem appropriate to convert the context to a log-odds adjustment value when context is added as an additional term to the random permutation model.
 Since performance does not decrease when the transformation is made, may slightly increase,
-and leads to a more natural interpretation of each model term as a log odds value, it is reasonable to include this step for the random-permutation model.
+and leads to a more natural interpretation of each model term as a log-odds value, it is reasonable to include this step for the random permutation model.
 Future research could explore more improved methods to properly combine the terms.
@@
 
-Vector-based models like the random-permutation model can easily incorporate word order when making predictions.
+Vector-based models like the random permutation model can easily incorporate word order when making predictions.
 This is one of their primary strengths that has been discussed \parencites{Jones2007, Sahlgren2008}.
@@
 
-Three different methods for incorporating word order in the random-permutation model were tested:
+Three different methods for incorporating word order in the random permutation model were tested:
 Taking just the sign of the word's position relative to the tag (direction method), using the position (standard method), and using the position but only including words that were used near the tag (window method).
@@
 The same Twitter popular-users datasets and StackOverflow popular-hashtags datasets were used for testing.
-Both an ordered and orderless context component were created for the random-permutation models.
+Both an ordered and orderless context component were created for the random permutation models.
 The orderless context component uses the same co-occurrence matrix used by the Bayesian model to create the representation, since the Bayesian model does not contain any order information.
-Models were tested that used only one of the three word-order methods for the ordered context component, used only the orderless context component, used a combination of the two components, and combined context with prior.
+Models were tested that used only one of the three word order methods for the ordered context component, used only the orderless context component, used a combination of the two components, and combined context with prior.
 
@@
 
-Model accuracy results after incorporating word order into the random-permutation model are included in Figure \ref{figContextOrder}
+Model accuracy results after incorporating word order into the random permutation model are included in Figure \ref{figContextOrder}
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-orderT-crop.pdf}}}
-  \caption{Model accuracy for various word order methods for the random-permutation model for Twitter}
+  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
   \label{figContextOrder}
@@
 This is the default ordering method used, so the names for these bar plots were kept consistent to the names used in other figures so that comparisons could be made across figures if necessary
-(similar to what was done for the default row-size for the random-permutation model).
-Base models that do not use entropy weighting were also included so that effect sizes could be compared between entropy weighting and adding word-order information.
-
-Looking first at models where no prior is added, the random-permutation model using only orderless context performs similar to the model using only order context with direction.
+(similar to what was done for the default row-size for the random permutation model).
+Base models that do not use entropy weighting were also included so that effect sizes could be compared between entropy weighting and adding word order information.
+
+Looking first at models where no prior is added, the random permutation model using only orderless context performs similar to the model using only order context with direction.
 All other order-only models perform worse than the direction order model and worse than the orderless model.
@@
 This was a surprising result.
-One of the main benefits discussed for the random-permutation model is that it can easily incorporate word-order information into the representation.
-But including word-order information does not dramatically improve performance. 
-For the random-permutation model, it is much more important to properly handle high-frequency stop words and include a prior component than it is to include word order information,
+One of the main benefits discussed for the random permutation model is that it can easily incorporate word order information into the representation.
+But including word order information does not dramatically improve performance. 
+For the random permutation model, it is much more important to properly handle high-frequency stop words and include a prior component than it is to include word order information,
 at least for the Twitter dataset tested in this study.
@@
 In absolute terms, this model only improved by 2 percentage points in accuracy when trained on the TASA and tested on the TOEFL.
-\textcite{Sahlgren2008}'s results for word order for the random-permutation model are similar.
+\textcite{Sahlgren2008}'s results for word order for the random permutation model are similar.
 At a row size of around \num{2000}, model accuracy improves by less than a percentage point when order information is incorporated.
@@
 \textcite{Sahlgren2008} most likely did not use a regression technique to optimize the weights for each of the order and orderless model terms.
-So although incorporating word order into the random-permutation model does seem to slightly improve performance,
+So although incorporating word order into the random permutation model does seem to slightly improve performance,
 the strength of the model is more likely to be in how efficiently it can represent the co-occurrence matrix, than its ability to include word order in the representation.
 
-Since it was shown that including word-order information has little to no significant improvement on model accuracy, an attempt to incorporate word order information into the Bayesian model was not made.
+Since it was shown that including word order information has little to no significant improvement on model accuracy, an attempt to incorporate word order information into the Bayesian model was not made.
 
@@
 If enough context information has been collected, maybe a context-only model can perform just as well as a combined context and prior model.
-If this were the case, then it may be less important that the random-permutation model does not include a prior term by default.
+If this were the case, then it may be less important that the random permutation model does not include a prior term by default.
 
@@
 
-Full models were tested against context-only models for both sites and both the Bayesian and random-permutation model types.
+Full models were tested against context-only models for both sites and both the Bayesian and random permutation model types.
 The four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled datasets were tested first.
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
-  \caption{Model accuracy for various word order methods for the random-permutation model for Twitter}
+  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
   \label{figContextPriorneededSO}
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddT-crop.pdf}}}
-  \caption{Model accuracy for various word order methods for the random-permutation model for Twitter}
+  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
   \label{figContextPriorneededT}
@@
 
-For both the Bayesian and random-permutation models, accuracy remains the same when the prior component is removed for StackOverflow.
+For both the Bayesian and random permutation models, accuracy remains the same when the prior component is removed for StackOverflow.
 Accuracy does dramatically decrease for Twitter when the prior is removed, however.
@@
 In this case, past tag history does contain unique variance for accurately predicting tag use.
-Accuracy improves for both the Bayesian and random-permutation models when the prior model component is combined with context compared to context alone.
+Accuracy improves for both the Bayesian and random permutation models when the prior model component is combined with context compared to context alone.
 
@@
 
-Overall, both the random-permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
+Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
 These results provide support that choosing a hashtag when composing a tweet and tagging a StackOverflow post is a long-term memory retrieval process.
@@
 
-Given that the storage space required to represent the co-occurrences for the random-permutation is dramatically smaller than the uncompressed Bayesian model,
+Given that the storage space required to represent the co-occurrences for the random permutation is dramatically smaller than the uncompressed Bayesian model,
 and that the compression is lossy and essentially random (i.e., no SVD),
-it is quite interesting that the random-permutation model performs nearly as well as the full Bayesian model for this task.
+it is quite interesting that the random permutation model performs nearly as well as the full Bayesian model for this task.
 
@@
 
-Also, the relatively strong performance of the random-permutation model is not due to the model's ability to represent word order.
-Performance is similar for the random-permutation and Bayesian models when the co-occurrence matrix for both models is the same, and not based on word order.
-The random-permutation model seems to work so well primarily due to the way that the full co-occurrence representation is compressed.
-
-Compression for the random-permutation model is essentially random.
+Also, the relatively strong performance of the random permutation model is not due to the model's ability to represent word order.
+Performance is similar for the random permutation and Bayesian models when the co-occurrence matrix for both models is the same, and not based on word order.
+The random permutation model seems to work so well primarily due to the way that the full co-occurrence representation is compressed.
+
+Compression for the random permutation model is essentially random.
 That is, no SVD is done to identify the most predictive components,
-so the compressed representation used by the random-permutation model does not contain nearly as much of the variance from the original co-occurrence matrix that it could have.
+so the compressed representation used by the random permutation model does not contain nearly as much of the variance from the original co-occurrence matrix that it could have.
 However, performing a full SVD on a co-occurrence matrix as large as the ones generated for StackOverflow and Twitter is computationally infeasible.
@@
 only that the neural machinery is set up so that the representation makes predictions that are similar to what would have been generated if the matrix decomposition for an SVD would have been computed.
-Perhaps the randomly-compressed representation used for the random-permutation model is the way in which the brain tries to achieve the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
+Perhaps the randomly-compressed representation used for the random permutation model is the way in which the brain tries to achieve the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
 
@@
 The declarative memory model for ACT-R has an attentional weight term $W_{j}$ that can be used to implement how stop words are handled.
-The random-permutation model was also easily modified to have an attentional weight term by attenuating each word's set of ones and negative ones values by that word's attentional weight.
+The random permutation model was also easily modified to have an attentional weight term by attenuating each word's set of ones and negative ones values by that word's attentional weight.
 Both a method for stop-word removal (frequency filtering) and stop-word attenuating (entropy weighting) performed well for both models.
 Also, using a proper method for handling the low-predictor words produced some of the largest improvements in model accuracy compared to the other explorations
-(e.g., compared to word order, properly adding terms, compression amount for random-permutation representation).
+(e.g., compared to word order, properly adding terms, compression amount for random permutation representation).
 Since the ACT-R declarative memory theory already has a component for attentional weight, it may be the case that the neural system performs some form of stop-word weighting when computing activations,
@@
 
-Adding a prior component to the random-permutation model also increased model accuracy for both StackOverflow and Twitter.
+Adding a prior component to the random permutation model also increased model accuracy for both StackOverflow and Twitter.
 It is unsurprising that this occurred.
-However, it is quite surprising that this is the first time that the prior component has been included in a random-permutation model.
-If the random-permutation model is explored as a declarative memory theory for other datasets, it should be worthwhile to include the prior component for those explorations.
+However, it is quite surprising that this is the first time that the prior component has been included in a random permutation model.
+If the random permutation model is explored as a declarative memory theory for other datasets, it should be worthwhile to include the prior component for those explorations.
 Although it did not seem to matter much how the prior component is combined with the context component in this study, 
 there may be better ways than the log-odds transformation to combine the terms.
-Further research could focus on exploring a larger set of methods for combining the context component of the random-permutation model with the prior component from the ACT-R theory.
+Further research could focus on exploring a larger set of methods for combining the context component of the random permutation model with the prior component from the ACT-R theory.
 
@@
 auto labeling incoming mail and placing it into appropriate folders, and spam filtering.
-Having an accurate and efficient co-occurrence memory model that correctly retrieves relevant tags can be applied to a wide variety of real-world tasks.
+Having an accurate and efficient co-occurrence memory model that correctly retrieves relevant tags can be applied to a wide variety of tasks.
 
@@
 and [3] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
-This work has shown how past user history is important, and how a context-only model like the random-permutation model can be modified to include the prior component.
-The work has also shown how the strength of the random-permutation model is not in its ability to represent word order,
+This work has shown how past user history is important, and how a context-only model like the random permutation model can be modified to include the prior component.
+The work has also shown how the strength of the random permutation model is not in its ability to represent word order,
 and rather in its ability to efficiently compress the full co-occurrence representation used for the Bayesian model.

### 2014-07-28 writeups/dissertation/dissertation.tex
foo

@@
   Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random permutation vector-based model.
-  The results show that past user behavior of tag use is a strong predictor of future performance, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
-  Also, the common natural-language processing method to attenuate low-predictor words in the dataset was linked to ACT-R's attentional weight term,
+  The results show that past user behavior of tag use is a strong predictor of future behavior, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
+  Also, the common natural language processing method to attenuate low-predictor words in the dataset was linked to ACT-R's attentional weight term,
   and an entropy-weighting technique was developed that improved model accuracy by more efficiently allocating limited attentional resources to the predictive words in the posts and tweets.
@@
 The declarative memory component in ACT-R is a Bayesian statistical model.
-This is the same type of co-occurrence Naive Bayes statistical model that is prominent in the more general machine-learning field.
+This is the same type of co-occurrence Naive Bayes statistical model that is prominent in the more general machine learning field.
 When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
@@
 
-This approach is problematic for two main reasons:
+But this approach is problematic for two main reasons:
 First, it means that the declarative memory system for every developed model must be hand crafted and customized for that model, which takes enormous time and effort.
@@
 In other words, users on social media sites want the fresh \emph{stream} of information about a particular topic.
-This can certainly be seen in the main user-interface views that these social media sites provide:
+This can certainly be seen in the main user interface views that these social media sites provide:
 Twitter's homepage showing followees' tweets, Facebook's homepage showing friends' posts, and StackOverflow's daily digest of posts for particular tags.
@@
 Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
-Two of the current state-of-the-art memory retrieval models are based on a co-occurrence methodology, and I will be comparing these two models for hashtag prediction:
+At least two of the current memory retrieval models are based on a co-occurrence methodology, and I will be comparing these two models for hashtag prediction:
 ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.
@@
 
-I will compare these two memory systems across two domains where users generate tags for posts: Twitter and StackOverflow.
+These two memory systems were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
 Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
@@
 
-These memory retrieval models will also be tested on tag use for StackOverflow posts.
+These memory retrieval models were also tested on tag use for StackOverflow posts.
 StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
 A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
-User-chosen tags on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
+The tags chosen by a user on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
 Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.
@@
 Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., \cite{Rutledge2008} modeling the ``fan effect'').
-I will extend the work done by \citeauthor{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
-I will also replace the vector-based model used by \citeauthor{Rutledge2008} with an improved and simplified random permutation model \parencite{Sahlgren2008}.
 
@@
 However, it is unclear and unexplored how word order can be represented in the ACT-R declarative memory theory.
-I will explore how the word-order strength of vector-based models can be incorporated into the ACT-R theory when testing the models on hashtag prediction tasks.
 
@@
 However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.
-I will explore how a user's prior knowledge influences the likelihood that a particular hashtag is chosen for both ACT-R's memory theory and vector-based models.
 
@@
 Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into a specific user's prior likelihood of hashtag use.
-So I will characterize the hashtag growth and decay cycle for individual users.
-The expectation is that this cycle will match ACT-R's decay rate used in the declarative memory system to compute a particular memory item's prior.
 
@@
 Base-level activation reflects the log prior odds of needing an observed chunk again.
-The primary way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
+The default way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
 This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
@@
 It is important that this measure scales well enough to handle the sheer volume of information present in online social media datasets.
-Large-scale social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
+Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
 If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.
@@
 If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
-However, with large-scale human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
+However, with large human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
 \textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
@@
 
-Another issue with LSA is that adding additional observations to an already-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
-Since computing the SVD is the most computationally-expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
+Another issue with LSA is that adding additional observations to an previously-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
+Since computing the SVD is the most computationally expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
 For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.
@@
 This type of information on a user's prior behavior is not being utilized with vector-based models currently.
-It may certainly be the case that vector-based models can be modified to incorporate base-level activation, and this idea will be pursued with this research.
 
@@
 These two computations are certainly not equivalent, even though they are most likely correlated with one another.
-It may be interesting to explore how model performance changes when one calculation is swapped out for another.
-For example, a vector-based model could generate a compressed (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
-Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
-It is unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
-But it may be the case that using Bayesian computation on a compressed vector-based model improves performance.
 
@@
 Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
-In order to provide a high quality user experience on the site, Netflix created a content-recommendation service.
+In order to provide a high quality user experience on the site, Netflix created a content recommendation service.
 This tool recommends to the user movies and television episodes that most likely match their specific interests.
@@
 Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
-Google has even deployed a hashtag-recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
+Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
 A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
@@
 It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
-This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular environment
+This suggests that more cognitively-plausible memory retrieval models may outperform more general machine learning approaches in this particular environment
 (i.e., when predicting the user-chosen tags for human-created content on social media sites).
@@
 Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}.
-Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, it will be interesting to see how a cognitively-plausible vector-based model performs at the task.
-Prediction differences between the two models may help reveal ways where the constraints of each model can be adjusted in order to improve accuracy.
 
@@
 To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
-It will be interesting to see how much (if at all) accuracy improves when incorporating word order information into the model.
 
@@
 This preliminary research is encouraging, and suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
-It also seems plausible that model accuracy can improve an even greater amount by
-[1] figuring out the optimal way to combine global tag history with a user's past tagging history, and
-[2] using the time-dependent form of the base-level component in Table \ref{tabACTRModel}, in order to account for the fact that a user's tagging preferences change over time.
-So these proposed modifications will be explored further when incorporating a user's past tagging history into the models for this research.
+It also seems plausible that model accuracy can improve an even greater amount by using the time-dependent form of the base-level component in Table \ref{tabACTRModel},
+in order to account for the fact that a user's tagging preferences change over time.
 
@@
 
-Two models will be evaluated for hashtag prediction on both the Twitter and StackOverflow datasets:
+Two models were evaluated for hashtag prediction on both the Twitter and StackOverflow datasets:
 A vector-based model and Bayesian retrieval model.
-The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} will be used for the vector-based model, since it has shown to perform better than the BEAGLE model \parencite{Recchia2010}.
-The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel} will be used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
-
-These models will be the starting point for modeling tag use on the Twitter and StackOverflow sites.
-The models may be modified during testing if the data indicate that modifications will increase the accuracy in which the model's recommended tags agree with the post author's chosen tags.
-Also there are planned modifications to each model as part of the research, as these modifications seem likely to improve performance: 
-incorporating user prior information into the vector-based model and incorporating word order information into the Bayesian model.
+The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} was used for the vector-based model, since it has shown to perform better than the BEAGLE model \parencite{Recchia2010}.
+The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel} was used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
 
@@
 
-The two memory models will be evaluated on predicting user-chosen tags for Twitter tweets and StackOverflow posts.
-Parameters for each model will be calibrated on a large subset of the Twitter and StackOverflow datasets.
-Each model will be tested by evaluating the accuracy of model-chosen tags on a fresh test subset of the datasets.
-Accuracy will be assessed by comparing model performance with the human data.
-
-% Started re-reading and re-writing here.
+The two memory models were evaluated on predicting user-chosen tags for Twitter tweets and StackOverflow posts.
+The co-occurrence matrices used for the context component of each model were generated from a large subset of the Twitter and StackOverflow datasets.
+Each model was tested by evaluating the accuracy of model-chosen tags on a fresh test subset of the datasets.
+Accuracy was assessed by comparing model performance with the human data.
+
+% Started rereading and rewriting here.
 
@@
 A popular-users dataset was collected for both StackOverflow and Twitter.
-This dataset consists of all of the tweets from 24 different samples of the top users for each site.
-This dataset was collected by fetching all tweet history for each of the top users.
+This dataset consists of all of the tweets from various samples of the top users for each site.
+This dataset was collected by fetching all past tagging history for each of the top users.
 It was used to explore the growth and decay characteristics of a hashtag for individual users.
@@
 Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
-It took around four days to collect a single dataset, and one collection was done each week.
 Multiple datasets were collected to ensure that the results weren't dependent on a particular sample of hashtags, and would generalize across different popular-hashtag datasets and different time periods.
+Each dataset required around four days to collect enough tweets to reach the quota, and one collection was done each week.
 
@@
 The entire StackOverflow dataset is packaged and released quarterly to the public.
-So this package was downloaded and the database tables were imported into PostgreSQL.
+So the packaged dataset was downloaded and the database tables were imported into PostgreSQL.
 
@@
 This is most likely because the language for the samples taken for both sites is informal English,
-and it seams reasonable to think that tokenizing all of the informal slang, jargon, and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
+and it seams reasonable to think that tokenizing all of the informal language and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
 So the same tokenization algorithm was used for both StackOverflow and Twitter, since performance was similar for both algorithms on StackOverflow,
@@
 This can be thought of as semantic stemming, where two words with identical meaning are converted to the same canonical word.
-So tag words in the post and used tags for the post were converted to their root tag when possible.
+So tags used for the post were converted to their root tag when possible.
 
@@
 
-``Stop words'' are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
-It is typically recommended \parencite{Bird2009} to remove stop words such as ``the'', ``a'', ``or'' when building the co-occurrence matrix.
+``Stop words'' (e.g., ``the'', ``a'', ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
+It is typically recommended \parencite{Bird2009} to remove stop words when building the co-occurrence matrix.
 Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in text.
@@
 
-Stop words are commonly removed by either filtering based on a pre-determined list or filtering based on the high-frequency words that occur in the dataset.
+Stop words are commonly removed by either filtering based on a predetermined list or filtering based on the high-frequency words that occur in the dataset.
 \textcite{Sahlgren2008} tried both methods to remove stop words for the random permutation vector-based model, and found that the data-driven approach
-(i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a pre-determined list.
+(i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a predetermined list.
 
@@
 
-All three stop word techniques (i.e., pre-determined list, high-frequency filtering, and entropy weighting) were explored for this analysis.
+All three stop word techniques (i.e., predetermined list, high-frequency filtering, and entropy weighting) were explored for this analysis.
 
@@
 The decay rate parameter $d$ represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
-Higher values mean that chunk activation decays faster, which means that the most recently-presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
+Higher values mean that chunk activation decays faster, which means that the most recently presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
 Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks that have occurred most frequently will have the highest activation,
@@
 
-The full StackOverflow dataset and Twitter popular-users dataset were used to characterize tag growth and decay for a user on each site.
+The StackOverflow and Twitter popular-users datasets were used to characterize tag growth and decay for a user on each site.
 Each user's tag behavior from the popular-users dataset for each site was analyzed.
@@
 (0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.8, 2, 5, 10, 20).
-The values were chosen in order to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d=.5$), to a pure recency model ($d=20$).
+The values were chosen to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d=.5$), to a pure recency model ($d=20$).
 The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
@@
 Both the standard form and optimized learning form of $B_{i}$ were used to predict each user's specific tag history.
-To do this, a user's tag history was collected, and then each tag instance from oldest to newest was examined.
+To do this, a user's entire tag history was collected, and then each tag instance from oldest to newest was examined.
 For each of these tagging instances, the model generated an activation value for each tag based on that tag's frequency and recency of use. 
@@
 
-The specific value of the hashtag used for both plots is nominal and not important for this analysis.
-What is interesting is how large the tagging space is for these users.
-There are bands of commonly-used tags for each user, but there are also instances where new tags are used, and a large percentage of tags used are not in the most common bands.
+The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
+What is interesting is the large tag space for these users.
+There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
 This helps illustrate how challenging modeling this task can be.
-The model's goal is to predict at each instance what tag will be chosen, but since the tag space is so large and varied, this task is quite challenging.
+The model's goal is to predict at each instance what tag will be chosen, but since the tag space is so large and varied, the model must predict the correct few tags that are used from a much larger set of possible tags.
 
@@
 These plots were chosen from the set to also show that the amount of tag variation for individual users also varies: some use many tags and some use just a few.
-To accurately model human performance in this domain, the model must be able to make accurate predictions across all variations of user tagging behavior.
+To accurately model human performance in this domain, the model must be able to make correct predictions across all variations of user tagging behavior.
 
@@
 Also shown in these figures is how model predictions change as the decay rate parameter changes.
-The model changes less from tag to tag with lower decay rate values.
+With lower decay rate values, the model is less likely to switch from tag to tag for each post.
 This makes sense because lower decay rate values lead to the model weighting frequency more than recency in the activation calculation.
@@
 
-One main effect in all of the plots is that there is an optimal decay rate value between 0 and 20 that produces the highest model accuracy.
+One large effect in all of the plots is that there is an optimal decay rate value between 0 and 20 that produces the highest model accuracy.
 For the standard $B_{i}$ equation, model accuracy for a pure frequency model ($d=0$) is also worse than using a pure recency model ($d=20$), particularly for Twitter.
@@
 
-First, the best-fit values are close to the 0.5 default value used in ACT-R when the optimized learning form is used.
+The best-fit values are close to the 0.5 default value used in ACT-R when the optimized learning form is used.
 Comparing across optimized learning and the standard form, the best-fit values for the standard form are slightly higher for both the StackOverflow and Twitter dataset slices.
@@
 
+The mean was computed by taking the mean of the median best-fit decay rate value for each dataset slice.
+The median was used since the distribution of best-fit decay rate values for each dataset slice is somewhat positively skewed (e.g., the histograms in Figures \ref{figPriorSOQHistDsOL} and \ref{figPriorTwitterHistDsOL}).
+Also, there were users in the popular-users dataset that did not have enough observations to generate stable predictions, and best-fit decay rate values for these users could be as high as 20.
+Using the median effectively trims these unstable values from the sample, without having to define a cut point and go through an outlier removal process.
+
 Decay rate values that produced the most accurate model performance are lower for the optimized learning form (0.43) of $B_{i}$ compared to the standard form (0.80).
@@
 Given the large tag space used by post authors on StackOverflow and Twitter, it was surprising that model performance is respectable even when only past tag history and no context is used.
-This speaks to the importance of not only taking into account overall past tagging history, but also to ensuring that the prior component of activation for each tag is customized to each user's specific past tagging history.
+This speaks to the importance of taking into account overall past tagging history, and ensuring that the prior component of activation for each tag is customized to each user's specific past tagging history.
 
@@
 
+That said, comparing the effect size between changing the decay rate value and using the standard form,
+model accuracy improves much more when the standard form replaces the optimized learning form than when the decay rate value is increased from 0.5 to 0.8.
+For these datasets, it is much more important to use the standard form of the equation than it is to adjust the decay rate, although model accuracy does slightly improve if a higher decay rate is chosen with the standard form.
+
 \section{Combining Predictors}
@@
 Past tag use was isolated in the first study to see how well ACT-R's prior component of declarative memory (i.e., base-level learning) fits each user's tag use over time.
-However for the rest of the studies the past behavior and context declarative memory components were combined as well as looked at in isolation.
+However, for the rest of the studies the past behavior and context declarative memory components were combined as well as looked at in isolation.
 
@@
 In order to best combine these predictors, a logistic regression statistical technique similar to \textcite{Stanley2013} was used to determine the most optimal weights for each term.
-Logistic regression is used on these datasets in the following way.
+Logistic regression is used on these datasets in the following way:
 When a user creates a tag for a post, a retrieval request for the model is made.
@@
 
-Model accuracy is evaluated by using the optimal weights, rank ordering all recorded tags by model activation using those weights,
-asking the model to tag the top N posts with a 1 (where N is the total number of recorded instances where an author chose a tag),
+Model accuracy is evaluated by rank ordering all recorded tags by model activation using the optimal weights,
+asking the model to tag the top N posts with a 1 (where N is the total number of recorded tagging instances for an author),
 and then comparing the model's chosen tags with the used tags for each post (i.e., looking at the ``hits'').
@@
 If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression wasn't optimized to the constrained form, and weights may change if that was done.
-The constrained form of model accuracy can not be used for the regression because there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation.
-The constrained form could be used in a regression if a custom form of the regression was implemented.
-However this was not pursued since around \num{2500} total regressions were run on these datasets for two reasons:
-A custom implementation in R would not likely run in compiled C, and even if it was implemented in a compiled language it would be computationally intractable without a closed-form solution.
+Also, there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation. 
+To optimize parameters based on the constrained form,
+a full-enumeration numerical approach would have to be developed that systematically searches the space of coefficient weights and uses a custom objective fitting function to find the best fit.
+However, this was not pursued for three reasons:
+Around \num{2500} total regressions were run on these datasets,
+a custom implementation in R would not likely run in compiled C,
+and even if it was implemented in a compiled language it would be computationally intractable without a closed-form solution.
 
@@
 
-The offset term ensures that the activation for the small number of top tags for each post are equal across posts.
+The offset term keeps the activation for the small number of top tags for each post equal across posts.
 This ensures that the logistic regression only labels a few 1s for each post.
@@
 For the StackOverflow dataset, 5 runs of 100 posts were used, since the results for StackOverflow are more stable and retrievals take more wall-clock time for this dataset due to the size of the co-occurrence matrix.
-All runs contained separate posts, and all posts used for runs were not used to generate the co-occurrence matrix used for the two retrieval models.
+All runs contained different posts, and all posts used for runs were not used to generate the co-occurrence matrix for the two retrieval models.
 
@@
 However, a custom user prior for the Twitter popular-hashtags dataset cannot be used, since the entire Twitter dataset cannot be downloaded.
-This dataset contains samples of a large set of various users that choose a specific set of hashtags being monitored.
+The popular-hashtags dataset contains samples of a large set of various users that happened to use one of a specific set of hashtags being monitored.
 So in order to generate a custom user prior for each sample in the dataset, that user's entire tagging history must be accessed through the API.
-Since access rates are limited across that interface, it is infeasible to do this process for every user that appears in the Twitter popular-users dataset.
-So instead, the global prior tag history for all users in the Twitter popular-users dataset was used to compute prior activation values for each tag.
+Since access rates are limited across that interface, it is infeasible to do this process for every user that appears in the Twitter popular-hashtags dataset.
+So instead, the global prior tag history for all users in the Twitter popular-hashtags dataset was used to compute prior activation values for each tag.
 
@@
 In both plots, if the model name includes ``full'' then all model components are included.
-When they are enumerated, then components are left out.
+When they are enumerated, then at least one of the components is left out.
 The random permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
@@
 then the model would be able to say when, for example, the third top tag in a post is more likely to be a chosen tag than the second top tag in another post.
-That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of retrievals to do for a specific post.
+That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of retrievals to use for a specific post.
 If this were the case, then adding the offset term would actually decrease accuracy, since this absolute value of the activation values would be subtracted off.
@@
 
-A plot of model accuracy for components of both the random permutation and Bayesian models and isolated by Twitter popular-users subset are included in Figure \ref{figContextSubsets}.
+Model accuracy results for components of both the random permutation and Bayesian models grouped by Twitter popular-users subset are included in Figure \ref{figContextSubsets}.
 
@@
 
-However a cutoff point must be identified and used in order to remove the high-frequency words.
+However, a cutoff point must be identified and used in order to remove the high-frequency words.
 \textcite{Sahlgren2008} used a cutoff where words that occurred more than \num{15000} times were removed.
 Since the StackOverflow and Twitter datasets are much larger than the datasets used by \citeauthor{Sahlgren2008}, the same cutoff could not be used.
+
 The ideal cutoff was determined by choosing the value where the standard deviation of the counts for each word for each of the tags were all relatively small.
-That is, the spikes were suppressed for standard deviations for particular tags that still contained words with a very large number of observations.
-The cutoff was also chosen such that the size of the spikes was about as large as the spikes generated when the entropy method was used.
-
 The plots for the standard deviations across tags that were used to identify the cutoff for StackOverflow and Twitter are included in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT}.
@@
 The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
-Using a pre-defined stoplist makes the smallest improvement, while the entropy and frequency techniques improve performance the most, and frequency shows a slight edge over entropy.
+Using a predefined stoplist makes the smallest improvement, while the entropy and frequency techniques improve performance the most, and frequency shows a slight edge over entropy.
 However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
@@
 For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words.
-Removing stop-words based on a pre-determined list actually decreases accuracy, and using data-driven frequency filtering has only a small to no effect.
+Removing stop-words based on a predetermined list actually decreases accuracy, and using data-driven frequency filtering has only a small to no effect.
 
@@
 
-Comparing across Figures \ref{figContextStopRPSO} and \ref{figContextStopRPT}, entropy and frequency techniques are superior to filtering based on a pre-determined list.
+Comparing across Figures \ref{figContextStopRPSO} and \ref{figContextStopRPT}, entropy and frequency techniques are superior to filtering based on a predetermined list.
 However it is more difficult to choose the best approach when comparing entropy weighting and frequency filtering.
 Frequency filtering is slightly better for StackOverflow, while entropy weighting is slightly better for Twitter.
-However frequency filtering does have a few drawbacks compared to entropy weighting:
+However, frequency filtering does have a few drawbacks compared to entropy weighting:
 It produces accuracy measurements that have higher levels of noise (see Figure \ref{figContextStopRPSO} and frequency models for the title component in StackOverflow).
@@
 Therefore, favoring parsimony, for the random permutation model, attenuating stop words based on entropy produces more accurate results when compared to filtering the stop words based on frequency of use.
-This result is interesting since filtering stop words based on frequency is the default method for random permutation models \parencite{Sahlgren2008}.
+This result is interesting since filtering stop words based on frequency is the more common default method for random permutation models \parencite{Sahlgren2008}.
 This suggests that a simpler method (entropy weighting) can be used instead of the more common frequency-filtering technique to achieve the same (or better) accuracy measurements.
@@
 
-Since accuracy measurements when using a pre-determined list to remove stop words for the random permutation model were much lower than the frequency and entropy methods, this method was not examined for the Bayesian model.
+Since accuracy measurements when using a predetermined list to remove stop words for the random permutation model were much lower than the frequency and entropy methods, this method was not examined for the Bayesian model.
 The results for the frequency and entropy techniques for the Bayesian model on the StackOverflow dataset are include in Figure \ref{figContextStopBSO}.
@@
 
-One result that should not be overlooked is how poorly using a pre-determined stoplist performed when compare to the two data-driven frequency and entropy approaches.
-This is most likely because the StackOverflow and Twitter datasets are large and contain high-frequency words that are domain specific (e.g., emoticons) and are not included in the pre-determined list.
+One result that should not be overlooked is how poorly using a predetermined stoplist performed when compared to the two data-driven frequency and entropy approaches.
+This is most likely because the StackOverflow and Twitter datasets are large and contain high-frequency words that are domain specific (e.g., emoticons) and are not included in the predetermined list.
 
 Another interesting result was how the Bayesian model actually performed worse when stop words were removed based on frequency.
-This was not the case for the random permutation model, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random permutation model, and is most likely never the case for the random permutation model.
+This was not the case for the random permutation model in this study, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random permutation model,
+and is most likely never the case for the random permutation model.
 Model accuracy most likely decreases for the Bayesian model because this model already has a normalization component built in that the random permutation model does not have.
@@
 Both frequency and entropy techniques performed equally for the random permutation model.
-However since the entropy-weighting technique is parameter free and does not require a cutoff to tune, it is also a better choice for the random permutation model for these datasets.
+However, since the entropy-weighting technique is parameter free and does not require a cutoff to tune, it is also a better choice for the random permutation model for these datasets.
 
 The effect sizes when comparing no stop-word removal technique to either the entropy or frequency techniques are quite large.
-This shows the importance of correctly identifying and handling stop words when building co-occurrence matrices for large-scale natural-language datasets.
+This shows the importance of correctly identifying and handling stop words when building co-occurrence matrices for large-scale natural language datasets.
 For further research, it may be worthwhile to explore an even larger range of stop-word handling techniques,
-since model accuracy seems to be quite sensitive to the different techniques. 
-However, one conclusion that can be made from this research is that the entropy-weighting technique is a contender, and should be explored much more often than it has been previously.
+since model accuracy seems to be quite sensitive to the specific method used. 
+However, one conclusion that can be made from this research is that the entropy-weighting method is a contender, and should be explored much more often than it has been previously.
 
@@
 
-It is likely that the random permutation model is saturating when entropy weighting is used for the largest document size for StackOverflow.
+It is likely that the random permutation model is starting to saturate when entropy weighting is used for the largest document size for StackOverflow.
 Even if the non-predictive words based on entropy are attenuated,
-when the high-frequency words are included random permutation model,
+when the high-frequency words are included in the random permutation model,
 this attenuation is not enough to overcome the range effects that grow with document size.
@@
 As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random permutation model is actually beneficial, and the compressed random permutation model should be used.
-This crosstalk and noise is beneficial because it helps soften spurious $S_{ji}$ values when the values for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller document sizes.
+This noise is beneficial because it helps soften spurious $S_{ji}$ values when the values for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller document sizes.
 
@@
 It is common to handle stop words in some manner when working with NLP and large text documents.
-So, if these large natural-language documents are used to test declarative memory, then the entropy method may be a more cognitively plausible way of dealing with stop words rather than 
-filtering based on a pre-determined list or based on the frequency count in the dataset, particularly since the entropy method substantially increases model accuracy for these datasets and since it is parameter free.
+So, if these large natural language documents are used to test declarative memory, then the entropy method may be a more cognitively plausible way of dealing with stop words rather than 
+filtering based on a predetermined list or based on the frequency count in the dataset, particularly since the entropy method substantially increases model accuracy for these datasets and since it is parameter free.
 
@@
 Also since the developed models are task agnostic, they can potentially be used as tag recommendation systems for other domains.
-Examples of similar tagging tasks are hashtag use on Facebook, flagging (i.e., tagging) Wikipedia articles for content issues or overly-opinionated material,
+Examples of similar tagging tasks are hashtag use on Facebook, flagging (i.e., tagging) Wikipedia articles for content issues or overly opinionated material,
 auto labeling incoming mail and placing it into appropriate folders, and spam filtering.

### 2014-07-28 writeups/dissertation/dissertation.tex
foo

@@
 
-Another parameter in the random permutation model is the number of rows used to represent the all of the words.
+Another parameter in the random permutation model is the number of rows used to represent the all of the different words.
 Each word is represented by two randomly placed ones and two negative ones across the rows.
@@
 It is not an LSA approach where the few ones and negative ones for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
-Everything is randomly chosen with the random permutation model.
 The environment vector for each word is randomly selected, and no SVD is performed to choose these in a more discriminating manner.
@@
 This is not a log-odds value.
-Currently, prior and context for the random permutation model are combined by simply adding the terms.
+For the previous results, prior and context for the random permutation model are combined by simply adding the terms.
 However, this means that ACT-R's prior base-level learning component (which is a log-odds value) is added to a correlation value, which may not be mathematically appropriate.
@@
 For example, if the conversion is performed one the mean correlation value in the set, the result is a log-odds value of zero.
-This makes sense given that log odds can be interpreted as the amount of odds information (positive or negative) that should be adjusted to the prior log odds.
+This makes sense given that log odds can be interpreted as the amount of odds information (positive or negative) that should be adjusted to the prior log odds, given that new piece of information.
 Since the mean correlation value contains no information regarding if it is better or worse than the other correlation values, it is appropriate to assign this correlation value a zero log-odds adjustment value.
@@
 For the standard method, all words and their respective position are used to build the representation.
-For the direction method, essentially two representations are combined: all co-occurrences with words to the left of the tag, and all co-occurrences with words to the right of the tag.
+For the direction method, essentially two representations are created: all co-occurrences with words to the left of the tag, and all co-occurrences with words to the right of the tag.
 The window method is similar to the standard method, where each word's relative position is maintained, but it only includes words near the tag to build the representation.
@@
 
-The same Twitter popular-users datasets and StackOverflow popular-hashtags datasets were used for testing.
+The same Twitter popular-users datasets and StackOverflow randomly-sampled datasets were used for testing.
 Both an ordered and orderless context component were created for the random permutation models.
@@
 This is the default ordering method used, so the names for these bar plots were kept consistent to the names used in other figures so that comparisons could be made across figures if necessary
-(similar to what was done for the default row-size for the random permutation model).
+(similar to what was done for the default row size for the random permutation model).
 Base models that do not use entropy weighting were also included so that effect sizes could be compared between entropy weighting and adding word order information.
@@
 Looking first at models where no prior is added, the random permutation model using only orderless context performs similar to the model using only order context with direction.
-All other order-only models perform worse than the direction order model and worse than the orderless model.
+All other order-only models perform worse than the order model using direction and worse than the orderless model.
 
@@
 
-Full models were tested against context-only models for both sites and both the Bayesian and random permutation model types.
-The four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled datasets were tested first.
+Full models were tested against context-only models for both sites. 
+The four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset were used for the test.
 
@@
 The results compare accuracy for models that use a global co-occurrence and user-specific co-occurrence matrix.
-For StackOverflow, model accuracy for the user-specific (labeled as ``user sji'') can be compared against the default global co-occurrence matrix.
+For StackOverflow, model accuracy for the user-specific co-occurrence matrix (labeled as ``user sji'') can be compared against the default global co-occurrence matrix.
 Accuracy decreases dramatically when a user-specific co-occurrence matrix is used.
@@
 The results are similar for Twitter.
-However since the entire dataset is not available for Twitter, a global co-occurrence matrix condition can not be tested,
-since the popular-users dataset samples across a wide range of hashtags, and due to Twitter's API restrictions it would be infeasible to collect all posts for each of the hashtags that are used in the dataset.
+However, the entire dataset is not available for Twitter, so a global co-occurrence matrix condition can not be tested,
+since the popular-users dataset samples across a wide range of hashtags, and due to Twitter's API restrictions it would be infeasible to collect all posts that use each of the hashtags in the dataset.
 But user-specific co-occurrence matrices can be built, since the popular-users dataset does contain all context and hashtag observations for each user.
@@
 
-Both models require large numbers of context co-occurrence observations to make stable predictions from context information.
+Both models require large numbers of context co-occurrence observations to make accurate predictions from context information.
 Otherwise, the co-occurrence matrices do not have enough data to properly stabilize.
@@
 How, precisely, is the neural system configured to represent co-occurrences for long-term memory?
-Since, at the behavioral level, both models perform similarly, it is possible that the neural system is using some sort of vector-based compressed storage (e.g., random permutations) to store co-occurrences.
+Since at the behavioral level both models perform similarly, it is possible that the neural system is using either representation to store co-occurrences.
 Vector-based models are simple enough that they can easily be implemented in a neural network.
 Perhaps a form of vector-based compressed storage is being used at the neural level,
-since it can be implemented in neurons and it is a way for the machinery to generate predictions that closely match the statistical structure of the environment (i.e., that closely match the Bayesian representation).
+since it can be implemented in neurons and is an efficient way for the machinery to generate predictions that closely match the statistical structure of the environment
+(i.e., that closely match the uncompressed Bayesian representation).
 
@@
 Performance is similar for the random permutation and Bayesian models when the co-occurrence matrix for both models is the same, and not based on word order.
+Performance also does not improve by much when word order is added as the last predictor in the random permutation model.
 The random permutation model seems to work so well primarily due to the way that the full co-occurrence representation is compressed.
 
-Compression for the random permutation model is essentially random.
+And compression for the random permutation model is essentially random.
 That is, no SVD is done to identify the most predictive components,
@@
 So, it also seems plausible that this SVD computation is infeasible at the neural level.
-\textcite{Landauer1997} explicitly state that they make no claims that the brain is computing an SVD at the neural level;
+When deriving the LSA approach, \textcite{Landauer1997} explicitly state that they make no claims that the brain is computing an SVD at the neural level;
 only that the neural machinery is set up so that the representation makes predictions that are similar to what would have been generated if the matrix decomposition for an SVD would have been computed.
-Perhaps the randomly-compressed representation used for the random permutation model is the way in which the brain tries to achieve the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
+Perhaps the randomly-compressed representation used for the random permutation model is the way in which the brain tries to efficiently achieve the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
 
@@
 It seems likely that the entropy method for computing the attentional weight term $W_{j}$ may be applicable to other declarative memory tasks.
-The entropy method is parameter free (unlike the frequency-filtering method), so it completely data driven (much like the co-occurrence matrix) and does not require tuning.
+The entropy method is parameter free (unlike the frequency-filtering method), completely data driven (much like the co-occurrence matrix), and does not require tuning.
 It is common to handle stop words in some manner when working with NLP and large text documents.
@@
 It is unsurprising that this occurred.
-However, it is quite surprising that this is the first time that the prior component has been included in a random permutation model.
+However, it is quite surprising that this is the first time that the prior component has been included and tested in a random permutation model.
 If the random permutation model is explored as a declarative memory theory for other datasets, it should be worthwhile to include the prior component for those explorations.

### 2014-07-28 writeups/dissertation/dissertation.tex
foo

@@
 This algorithm was specifically created for tokenizing Twitter content.
-However during testing it was found to be robust and well suited for StackOverflow content.
+However, during testing it was also found to be robust and well suited for StackOverflow content.
 The initial model results with the new tokenizer were qualitatively very similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
@@
 This logistic regression method is a computationally efficient way to find the optimal weights for the predictors.
-However when the regression can not be constrained to label a specific number of 1s for each observation (as is the case here), to be used properly an offset term must be added as an additional model predictor.
+However, when the regression can not be constrained to label a specific number of 1s for each observation (as is the case here), to be used properly an offset term must be added as an additional model predictor.
 
@@
 \textcite{Stanley2013} did this by subtracting off the mean total activation for the top 5 tags for each post.
-However this method is complicated since computing the total activation for a tag requires already knowing the weights for each model component that are used to compute that activation.
+However, this method is complicated since computing the total activation for a tag requires already knowing the weights for each model component that are used to compute that activation.
 \citeauthor{Stanley2013} used an iterative approach to determine the weights used for the offset term, and kept iterating until those weights matched the weights returned after running the logistic regression.
@@
 Comparing across Figures \ref{figContextStopRPSO} and \ref{figContextStopRPT}, entropy and frequency techniques are superior to filtering based on a predetermined list.
-However it is more difficult to choose the best approach when comparing entropy weighting and frequency filtering.
+However, it is more difficult to choose the best approach when comparing entropy weighting and frequency filtering.
 Frequency filtering is slightly better for StackOverflow, while entropy weighting is slightly better for Twitter.
@@
 When computing activation for the random permutation model, the correlation is calculated directly on a matrix of counts of co-occurrences for words and tags.
-However for the Bayesian model this count matrix is converted to a log-odds ($S_{ji}$) matrix,
+For the Bayesian model this count matrix is converted to a log-odds ($S_{ji}$) matrix,
 where both the number of observations for the word and the tag are normalized when computing the value for each cell (see the $S_{ji}$ computation in Table \ref{tabModACTRModel}).
@@
 Accuracy improves for both the Bayesian and random permutation models as the size of the documents used to generate the co-occurrence matrix increases.
-However model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random permutation model, while the frequency technique only begins to plateau.
+However, model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random permutation model, while the frequency technique only begins to plateau.
 The Bayesian model does not plateau and accuracy continues to improves document size increases.
@@
 Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy.
-However \num{2048} rows is sufficient for these datasets, as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy.
+But \num{2048} rows is sufficient for these datasets, as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy.
 
@@
 
-However it should be pointed out that the way the random permutation model compresses the words is completely random.
+However, it should be pointed out that the way the random permutation model compresses the words is completely random.
 It is not an LSA approach where the few ones and negative ones for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
@@
 When predictors are combined and the full model is examined, the transformation is in the direction of slightly increasing accuracy, at least for the StackOverflow dataset.
-However any improvement when using the transformation is minimal.
+However, any improvement when using the transformation is minimal.
 

### 2014-07-29 writeups/dissertation/dissertation.tex
foo

@@
 
-% Don't let biblatex change any of the casing in the bibliography file
+% Do not let biblatex change any of the casing in the bibliography file
 \DeclareRobustCommand{\MakeSentenceCase*}[1]{{#1}}
@@
 
-\section{Methods}
+\section{Method}
 
@@
 Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
-Multiple datasets were collected to ensure that the results weren't dependent on a particular sample of hashtags, and would generalize across different popular-hashtag datasets and different time periods.
+Multiple datasets were collected to ensure that the results were not dependent on a particular sample of hashtags, and would generalize across different popular-hashtag datasets and different time periods.
 Each dataset required around four days to collect enough tweets to reach the quota, and one collection was done each week.
@@
 
-Multiple samples were taken across a broad range of user types for each dataset to ensure that the model results weren't dependent on specific attributes in a single sample.
+Multiple samples were taken across a broad range of user types for each dataset to ensure that the model results were not dependent on specific attributes in a single sample.
 For the StackOverflow popular-users dataset, a total of 12 slices of users were taken at different levels of reputation and total number of questions asked.
@@
 So it may be worthwhile to use the standard form instead of the optimized learning form when declarative memory is modeled across a broader range of tasks.
-However, this shouldn't be taken as a strong recommendation,
-since the performance benefits seen here could simply be due to the fact that the declarative memory retrievals span across a large period of time, which isn't the case for all tasks.
+However, this should not be taken as a strong recommendation,
+since the performance benefits seen here could simply be due to the fact that the declarative memory retrievals span across a large period of time, which is not the case for all tasks.
 Nonetheless, the only real reason not to use the standard form is due to computational resource limitations.
@@
 
-\subsection{Methods}
+\subsection{Method}
 
@@
 However, this relaxed metric was used for the rest of the experiments that involve combining terms because the optimal weights found by the regression are based on using the relaxed metric.
-If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression wasn't optimized to the constrained form, and weights may change if that was done.
+If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression was not optimized to the constrained form, and weights may change if that was done.
 Also, there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation. 
@@
 
-\section{Results for Combining Terms}
+\section{Results for Combining Predictors}
 
@@
 Accuracy is improved even when only a single term is used.
-Although in this case the weight of the term is irrelevant (only one term, weighting for that term doesn't matter),
+Although in this case the weight of the term is irrelevant (only one term, weighting for that term does not matter),
 using an offset still improves accuracy since the offset term ensures that only a small number of tags per posts are predicted by the model to be the chosen tags.
@@
 
+\subsubsection{Activation Calculation}
+
+The Bayesian and vector-based memory models differ in how the co-occurrence matrix is represented (lossy compression for vector-based models).
+They also differ in how the activation of each tag is calculated, given the different representations.
+The vector-based memory models, like LSA, compute a correlation between the context and each tag, while the Bayesian models sum the log likelihoods that each word in the context occurs with each tag.
+These two methods for computing the activation are correlated but not identical.
+Further research could try to tease apart how model accuracy is independently influenced by the representation (compressed or not) and the activation calculation (correlation or log likelihood).
+To do this, four models could be constructed (uncompressed and compressed representation crossed with correlation and log likelihood activation) compared to the two in this study,
+so that the effects of representation and activation calculation could be partialed out.
+This would allow, for example, the method of compression for the random permutation vector-based model to be analyzed independent of activation calculation.
+The results would provide additional evidence to help determine the specific representation that the declarative memory system uses to represent co-occurrences,
+and also the specific activation calculation that is carried out by the system.
+
 \subsubsection{Future Applications}

### 2014-07-29 writeups/dissertation/dissertation.tex
foo

@@
 
-\subsection{Impact of Prior vs Context}
+\subsection{Coefficients for Predictors}
+
+The logistic regression is used to determine the optimal weight for each model component.
+A regression is done for each run in the dataset, and since multiple runs were used for the StackOverflow and Twitter datasets, multiple samples of coefficient weights for each model component were computed.
+The variance of the weights for each component show how likely it is that the regression is overfitting the model,
+and the mean shows the best estimate weight of each component relative to the other.
+
+\subsubsection{Method}
+
+Coefficients for all of the runs for the Twitter popular-hashtags and StackOverflow randomly-sampled datasets were collected.
+This totaled 5 sets of coefficient weights for StackOverflow (5 runs) and 60 for Twitter (15 runs for each of the 4 dataset samples).
+
+\subsubsection{Results}
+
+Coefficients for each model component of the full models are included in Figures \ref{figContextCoeffSO} and \ref{figContextCoeffT}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
+  \caption{Coefficient values for full models for StackOverflow}
+  \label{figContextCoeffSO}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
+  \caption{Coefficient values for full models for Twitter}
+  \label{figContextCoeffT}
+\end{figure}
+
+The range of the confidence intervals is much more narrow than the range of the coefficient values between the components of each model.
+This shows that the model is not being overfit, as the weights for each component are relatively stable across the different runs.
+The weight of the prior relative to the context terms is also fairly consistent across models and across the two sites.
+Across the sites, context is a much stronger predictor than past behavior for StackOverflow, but past behavior is more important for Twitter. 
+Within a site, both models are consistent in how they favor context or past behavior.
+Also, looking at StackOverflow, both models are consistent in weighting the body more than the title, and the title more than the prior term.
+
+\subsubsection{Discussion}
+
+It was not surprising that the model was not overfit in this scenario, since the number of observed tags is much greater than the number of coefficients being fit.
+Each model has at most three terms, those weights are being fit to 500 posts within each run for Twitter and 100 for StackOverflow, and multiple tags are used for each post (particularly for StackOverflow).
+It is interesting that although the values of the weights are stable within a site (StackOverflow or Twitter), the weights change dramatically when looking across the sites.
+This illustrates that the impact of a user's past history relative to the current context can be domain specific, which means that for best results the weights for each of the terms should be recalibrated in new scenarios.
+However, the underlying reason that the weights change between the sites cannot be answered from this study alone.
+This relative impact may be changing from either a true alteration in the underlying weight of the prior compared to context for different domains,
+or simply due to the fact that not enough information on one of the two terms is available.
+
+\subsection{Impact of Prior vs. Context}
 

### 2014-07-29 dissProject/preProcess.R

@@ -1382,7 +1382,7 @@ buildTablesLogreg <- function(outFileNames) {
 	addModelType(logregTbl)
 	logregTbl[, DVName := predName]
 	# FIXME: Remove once runs generating coeffStd have completed 
-	logregTbl[, coeffStd := logregTbl[['coeff']] / logregTbl[['Std. Error']]]
+	#logregTbl[, coeffStd := logregTbl[['coeff']] / logregTbl[['Std. Error']]]
 	logregTbl
 }
 
@@ -1604,7 +1604,9 @@ makeRemapForMapping <- function(keyword, text) {
 }
 
 makeLogregMapping <- function() {
-	c('actPriorStd', 'Standard prior model',
+	c('actPriorStd', 'Standard prior model relaxed across posts',
+	  'actPriorStdBayes', 'Bayes prior relaxed across posts',
+	  'actPriorStdRP', 'RP prior relaxed across posts',
 	  'actTitle', 'Bayes title w/ entropy',
 	  'actBody', 'Bayes body w/ entropy',
 	  'actTweet', 'Bayes orderless context w/ entropy',
@@ -2046,10 +2048,16 @@ analyzeContext <- function(modelHashtagTbls, modelVsPredTbl) {
 }
 
 analyzeLogreg <- function() {
-	bestFitNames = c('actPriorStd_actTitle_actBody', 'actPriorStd_actTitleOrderlessEnthyman_actBodyOrderlessEnthyman', 'actPriorStd_actTitleOrderlessEnthyman_actBodyOrderlessEnthyman')
-	compareMeanDVLogreg(logregTbl[(sizeNum == 2 | is.na(sizeNum)) & dsetType == 'stackoverflow' & bestFitName %in% bestFitNames], coeffStd, figName='foo')
+	dTbl = modelVsPredTbl[predUsedBest == T][, list(d,  DVName)]
+	dTbl
+	logregTbl
+	bestFitNames = c('actPriorStd_actTitle_actBody', 'actPriorStd_actTitleOrderlessEnthyman_actBodyOrderlessEnthyman')
+	logregTbl
+	logregTbl[modelType == 'perm' & DVName == 'actPriorStd', DVName := 'actPriorStdRP']
+	logregTbl[modelType == 'sji' & DVName == 'actPriorStd', DVName := 'actPriorStdBayes']
[…]

### 2014-07-30 writeups/dissertation/dissertation.tex
foo

@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffSO-crop.pdf}}}
   \caption{Coefficient values for full models for StackOverflow}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffT-crop.pdf}}}
   \caption{Coefficient values for full models for Twitter}
@@
 
-The range of the confidence intervals is much more narrow than the range of the coefficient values between the components of each model.
+Even at these small sample sizes, the range of the confidence intervals is much more narrow than the range of the coefficient values between the components of each model.
 This shows that the model is not being overfit, as the weights for each component are relatively stable across the different runs.

### 2014-07-30 dissProject/preProcess.R

@@ -2047,17 +2047,37 @@ analyzeContext <- function(modelHashtagTbls, modelVsPredTbl) {
 	compareMeanDVDefault(dWideTbl, dDiff, figName='dContext', groupCol='dsetType')
 }
 
+bestFitNameToDVName <- function(vect, pre) {
+	vect = gsub('^act', pre, vect)	
+	vect = gsub('_act', '', vect)
+}
+
 analyzeLogreg <- function() {
-	dTbl = modelVsPredTbl[predUsedBest == T][, list(d,  DVName)]
+	dTbl = modelVsPredTbl[predUsedBest == T][, list(d,  runNum, groupNum, sizeNum, DVName)][grepl('^topHashtagAcross', DVName)]
+	setkey(dTbl, DVName, runNum, groupNum, sizeNum, d)
+	logregTbl
 	dTbl
-	logregTbl
-	bestFitNames = c('actPriorStd_actTitle_actBody', 'actPriorStd_actTitleOrderlessEnthyman_actBodyOrderlessEnthyman')
-	logregTbl
+	logregTbl[, bestFitNameAsDVName := bestFitNameToDVName(bestFitName, 'topHashtagAcross')]
+	setkey(logregTbl, bestFitNameAsDVName, runNum, groupNum, sizeNum, d)
+	logregTbl[, predUsedBest := F]
+	logregTbl[dTbl, predUsedBest := T, nomatch=0]
 	logregTbl[modelType == 'perm' & DVName == 'actPriorStd', DVName := 'actPriorStdRP']
 	logregTbl[modelType == 'sji' & DVName == 'actPriorStd', DVName := 'actPriorStdBayes']
-	compareMeanDVLogreg(logregTbl[(sizeNum == 2) & dsetType == 'stackoverflow' & DVName != '(Intercept)' & bestFitName %in% bestFitNames], coeffStd, figName='coeffSO')
+	bestFitNames = c('actPriorStd_actTitle_actBody', 'actPriorStd_actTitleOrderlessEnthyman_actBodyOrderlessEnthyman')
+	assign('p1', ggplot(
+			    logregTbl[predUsedBest & (sizeNum == 2) & dsetType == 'stackoverflow' & DVName != '(Intercept)' & bestFitName %in% bestFitNames],
+			    aes(x=dsetGroup, y=coeffStd, fill=DVName)) +
+	       #stat_summary(fun.y=mean, geom="bar", width=0.7, position=position_dodge()) +
[…]

### 2014-07-30 writeups/dissertation/dissertation.tex
foo

@@
 Finally, model accuracy was analyzed across all dataset slices for StackOverflow and Twitter.
-Mean accuracy for both sites is included in Figure \ref{figPriorAcc}. 
-The mean was computed by taking the mean of the mean model accuracy for each dataset slice.
-All confidence intervals are 95\% bootstrapped for the mean of the means for each slice.
+Accuracy for both sites is included in Figure \ref{figPriorAcc}. 
+Aggregate accuracy was computed by taking the mean of every user's model accuracy across all users in all dataset slices.
+All confidence intervals are 95\% bootstrapped for the mean.
 
@@
 The best-fit decay rate values were also analyzed across all dataset slices.
+Aggregate best-fit decay rates were computed by taking the median of all user's best-fit decay rate value across all dataset slices.
+The median was used since there were users in the popular-users dataset that did not have enough observations to generate stable predictions, and best-fit decay rate values for these users could be as high as 20.
+Using the median effectively trims these unstable values from the sample, without having to define a cut point and go through an outlier removal process.
+
 These results are included in Figure \ref{figPriorDecay}.
+Error bars represent the 95\% bootstrapped confidence interval of the median.
 
@@
 \end{figure}
-
-The mean was computed by taking the mean of the median best-fit decay rate value for each dataset slice.
-The median was used since the distribution of best-fit decay rate values for each dataset slice is somewhat positively skewed (e.g., the histograms in Figures \ref{figPriorSOQHistDsOL} and \ref{figPriorTwitterHistDsOL}).
-Also, there were users in the popular-users dataset that did not have enough observations to generate stable predictions, and best-fit decay rate values for these users could be as high as 20.
-Using the median effectively trims these unstable values from the sample, without having to define a cut point and go through an outlier removal process.
 
@@
 The impact of adding the offset to each of the models for StackOverflow and Twitter are included in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.
+The error bars represent the 95\% bootstrapped confidence interval of the mean model accuracy across all runs in dataset.
 

### 2014-07-30 dissProject/preProcess.R

@@ -160,17 +160,21 @@ sqldt <- function(...) {
 
 # Computing CIs around statistics
 
-CIMedian <- function(x) {
-	res = boot(x,function(x,i) median(x[i]), R=1000)
+CIBoot <- function(x, fun) {
+	res = boot(x,function(x,i) fun(x[i]), R=1000)
 	res
 	res = boot.ci(res, type="norm")
 	lower = unlist(res)[['normal2']]
 	upper = unlist(res)[['normal3']]
-	mid = median(x)
+	mid = fun(x)
 	if (is.null(lower)) lower=mid
 	if (is.null(upper)) upper=mid
 	c(upper, mid, lower)
 }
+
+CIMedian <- function(x) CIBoot(x, median)
+
+CIMean <- function(x) CIBoot(x, mean)
 
 getTokenizedTbl <- function(tweetsTbl, from, regex) {
 	matches = regmatches(tweetsTbl[[from]], gregexpr(regex, tweetsTbl[[from]], perl=T))
@@ -1408,7 +1412,7 @@ buildTables <- function(outFileNames) {
 	modelVsPredTbl
 }
 
-withCI <- function(dat, CIFun=CI) {
+withCI <- function(dat, CIFun=CIMean) {
 	res = CIFun(dat)
 	list(N=length(dat), meanVal=res[2], minCI=res[3], maxCI=res[1])
 }
@@ -1463,7 +1467,7 @@ plotBarSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol='dsetG
 	sumTbl
 }
 
-compareMeanDV <- function(modelVsPredTbl, DV, extras=NULL, figName='', groupCol='dsetGroup', CIFun=CI, title='Model Name') {
+compareMeanDV <- function(modelVsPredTbl, DV, extras=NULL, figName='', groupCol='dsetGroup', CIFun=CIMean, title='Model Name') {
 	DV = substitute(DV)
 	modelVsPredTbl
 	expr = bquote(modelVsPredTbl[, withCI(.(DV), CIFun=CIFun), keyby=list(DVName, .(as.symbol(groupCol)))])
@@ -1739,7 +1743,7 @@ analyzePrior <- function(modelVsPredTbl) {
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorOL2' & datasetName=='SOQgt500r2' & d < 1])
[…]

### 2014-07-30 writeups/dissertation/dissertation.tex
foo

@@
 Coefficients for each model component of the full models are included in Figures \ref{figContextCoeffSO} and \ref{figContextCoeffT}
+Error bars represent the standard deviation of coefficient values for each component.
 
@@
 
-Even at these small sample sizes, the range of the confidence intervals is much more narrow than the range of the coefficient values between the components of each model.
+The standard deviation is used instead of the confidence interval of the mean so that the plots show the overall stability of each coefficient across the different runs.
+The range of the distribution of coefficient values is much more narrow than the range of the mean coefficient values between the components of each model.
 This shows that the model is not being overfit, as the weights for each component are relatively stable across the different runs.
+
 The weight of the prior relative to the context terms is also fairly consistent across models and across the two sites.

### 2014-07-30 dissProject/preProcess.R

@@ -176,6 +176,10 @@ CIMedian <- function(x) CIBoot(x, median)
 
 CIMean <- function(x) CIBoot(x, mean)
 
+CIsd <- function(x) {
+	c(mean(x) + sd(x), mean(x), mean(x) - sd(x))
+}
+
 getTokenizedTbl <- function(tweetsTbl, from, regex) {
 	matches = regmatches(tweetsTbl[[from]], gregexpr(regex, tweetsTbl[[from]], perl=T))
 	wideTbl = data.table(id=tweetsTbl$id, matches=matches)
@@ -1759,7 +1763,7 @@ compareMeanDVDefault <- function(...) {
 }
 
 compareMeanDVLogreg <- function(...) {
-	compareMeanDV(..., groupCol='dsetGroup', extras=list(ylab('Coefficient Value')), title='Cofficient Name')
+	compareMeanDV(..., groupCol='dsetGroup', extras=list(ylab('Coefficient Value')), title='Cofficient Name', CIFun=CIsd)
 }
 
 analyzePUser <- function(modelVsPredTbl) {

### 2014-07-30 writeups/dissertation/dissertation.tex
foo

@@
 
-% Started rereading and rewriting here.
-
 \subsection{Datasets}
 
-The most recent quarterly published StackOverflow dataset \parencite{DataDump2013} was used to test the models on user chosen tags for posts.
+The most recent quarterly published StackOverflow dataset \parencite{DataDump2014} was used to test the models on user chosen tags for posts.
 This dataset contains around \num{5.7} million posts and \num{34377} unique tags.
@@
 
-\subsubsection{Popular Users}
+\subsubsection{Popular Users Subset}
 
 A popular-users dataset was collected for both StackOverflow and Twitter.
-This dataset consists of all of the tweets from various samples of the top users for each site.
+This dataset consists of all of the tweets and posts from various samples of the top users for each site.
 This dataset was collected by fetching all past tagging history for each of the top users.
@@
 
-\subsubsection{Popular Hashtags}
+\subsubsection{Popular Hashtags Subset}
 
@@
 
+\subsubsection{Randomly Sampled Subset}
+
+Since the entire dataset is available for StackOverflow, randomly-sampled posts were used to test the context models rather than subsetting and creating a popular-tags dataset (analogous to Twitter).
+If all Twitter data were available, then the models would have been tested on a randomly-sampled dataset for Twitter as well.
+However, since this is not the case for Twitter, the popular-hashtags subset was used instead of a randomly-sampled subset for that site,
+while models for StackOverflow were tested on the more challenging dataset where posts were randomly sampled across all posts created on the site.
+
 \subsection{Acquiring Datasets}
@@
 The entire StackOverflow dataset is packaged and released quarterly to the public.
-So the packaged dataset was downloaded and the database tables were imported into PostgreSQL.
+So the most recent packaged dataset was downloaded and the database tables were imported into PostgreSQL.
 
@@
 It is typically recommended \parencite{Bird2009} to remove stop words when building the co-occurrence matrix.
-Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in text.
+Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in the text.
 Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.
 
-Stop words are commonly removed by either filtering based on a predetermined list or filtering based on the high-frequency words that occur in the dataset.
+Stop words are usually removed by either filtering based on a predetermined list or filtering based on the high-frequency words that occur in the dataset.
 \textcite{Sahlgren2008} tried both methods to remove stop words for the random permutation vector-based model, and found that the data-driven approach
@@
 Since the optimized learning form assumes a constant presentation rate for each chunk,
-it only has to keep track of the time since the first presentation of a chunk $L$ and the total number of instances observed for that chunk. 
+it only has to keep track of the time since the first presentation of a chunk ($L$) and the total number of instances observed for that chunk ($n$). 
 It does not have to keep track of the time each specific chunk instance was observed.
@@
 
-The decay rate parameter $d$ represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
+The decay rate parameter ($d$) represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
 Higher values mean that chunk activation decays faster, which means that the most recently presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
 Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks that have occurred most frequently will have the highest activation,
-and activation depends less on the specific chunk that have been observed in the recent past.
+and activation depends less on the specific chunks that have been observed in the recent past.
 
@@
 The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
-However, only values equal to or below 1 were examined for the optimized learning form since values above 1 for this form cause a bifurcation and are not psychologically interesting.
+Only values equal to or below 1 were examined for the optimized learning form, since values above 1 for this form cause a bifurcation and are not psychologically interesting.
 
@@
 To do this, a user's entire tag history was collected, and then each tag instance from oldest to newest was examined.
-For each of these tagging instances, the model generated an activation value for each tag based on that tag's frequency and recency of use. 
+For each of these tagging instances, the model generated an activation value for each possible tag based on that tag's frequency and recency of use. 
 This means, for example, that the model always missed characterizing the first observed tag for each user, since at that point the model does not have any history of tag use for that user.
@@
 
-Model accuracy at each decay rate value for each user in the dataset slice are included in the plots.
+Model accuracies at each decay rate value for each user in the dataset slice are included in the plots.
 Each user's mean accuracy was subtracted off in order to compute a normalized mean for that user.
 This was done so that the relative accuracy change for different decay rate values could be more easily visualized.
-The technique is a visualization to what is done when testing the main effect in a repeated measures design.
+The plotting technique is a analogous to what is done when testing the main effect in a repeated measures design (i.e., each subject's mean score is subtracted off).
 
@@
 
-These results are included in Figure \ref{figPriorDecay}.
+Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
 Error bars represent the 95\% bootstrapped confidence interval of the median.
@@
 Decay rate values that produced the most accurate model performance are lower for the optimized learning form (0.43) of $B_{i}$ compared to the standard form (0.80).
-However, the optimal values for optimized learning are near 0.5, which lines up with default value used in ACT-R.
+The optimal values for optimized learning are near 0.5, which lines up with default value used in ACT-R.
 So the decay rate values for the standard form are slightly higher than the default values used in ACT-R, but the values found when optimized learning is used line up nicely with the ACT-R defaults.
@@
 However, this should not be taken as a strong recommendation,
-since the performance benefits seen here could simply be due to the fact that the declarative memory retrievals span across a large period of time, which is not the case for all tasks.
+since the performance benefits seen in this study could simply be due to the fact that the declarative memory retrievals span across a large period of time, which is not the case for all tasks.
 Nonetheless, the only real reason not to use the standard form is due to computational resource limitations.
-Since this becomes less of an issue as time progresses, it may be worthwhile to reconsider the default value used in ACT-R.
+Since this becomes less of an issue as time progresses and hardware speed increases, it may be time to reconsider the default value used in ACT-R.
 It may be the case that the optimized learning form of the equation is no longer computationally necessary.
@@
 Past tag use was isolated in the first study to see how well ACT-R's prior component of declarative memory (i.e., base-level learning) fits each user's tag use over time.
-However, for the rest of the studies the past behavior and context declarative memory components were combined as well as looked at in isolation.
+For the rest of the studies the past behavior and context declarative memory components were combined as well as looked at in isolation.
 
@@
 Model accuracy is evaluated by rank ordering all recorded tags by model activation using the optimal weights,
-asking the model to tag the top N posts with a 1 (where N is the total number of recorded tagging instances for an author),
+asking the model to tag the top N posts with a 1 (where N is the total number of recorded tagging instances in the dataset sample),
 and then comparing the model's chosen tags with the used tags for each post (i.e., looking at the ``hits'').
@@
 If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression was not optimized to the constrained form, and weights may change if that was done.
-Also, there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation. 
+Also, there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation where a small set of 1s occurred. 
 To optimize parameters based on the constrained form,
@@
 This ensures that the logistic regression only labels a few 1s for each post.
-\textcite{Stanley2013} did this by subtracting off the mean total activation for the top 5 tags for each post.
+\textcite{Stanley2013} did this by subtracting off from each post the mean total activation for the top 5 tags for that post.
 However, this method is complicated since computing the total activation for a tag requires already knowing the weights for each model component that are used to compute that activation.
@@
 
-The context component for both the random permutation and Bayesian model requires a co-occurrence matrix of words and observed tags.
+The context component for both the random permutation and Bayesian model requires the same co-occurrence matrix of words and observed tags.
 Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
@@
 The results for each run for each dataset were collected and averaged.
-The results across the four Twitter datasets were collected together, and the average was done across all runs in all four datasets.
+The results across the four Twitter datasets were collected together, and the average was taken across all runs in all four datasets.
 
@@
 
-It is actually somewhat disappointing that the offset term is even needed.
+It is somewhat disappointing that the offset term is even needed.
 If it were the case that model activation for a tag in a specific post could be interpreted as an exact activation value that did not depend on the specific post,
@@
 but the author is throttling their retrieval threshold for each post such that only a small set of tags are actually used.
-This is probably not the retrieval threshold used in the core declarative memory system that determines if a retrieval for the term actually happens.
+This is probably not the retrieval threshold used in the core declarative memory system that determines if a retrieval for the term is successful or not.
 It is unlikely that this retrieval threshold can be directly controlled and manipulated.
@@
 Note that using a predefined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
-The cutoff that produced the frequency plot is where a word represents more than \num{.04}\% of total occurrences in the dataset.
+The cutoff that produced the frequency plot in Figure \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
 This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset was not given.
@@
 However, frequency filtering does have a few drawbacks compared to entropy weighting:
-It produces accuracy measurements that have higher levels of noise (see Figure \ref{figContextStopRPSO} and frequency models for the title component in StackOverflow).
+It produces accuracy measurements that have higher levels of noise (see Figure \ref{figContextStopRPSO} and the frequency model for the title component in StackOverflow).
 Also, the ideal cutoff threshold must be identified in order to filter based on frequency,
 and identifying this cutoff requires a search, and this search must be done for each dataset that frequency filtering is used on.
-This threshold is a free parameter in frequency filtering that is not required when weighting stop words by entropy.
+This cutoff value is a free parameter in frequency filtering that is not required when weighting stop words by entropy.
 Therefore, favoring parsimony, for the random permutation model, attenuating stop words based on entropy produces more accurate results when compared to filtering the stop words based on frequency of use.
@@
 This is most likely because the StackOverflow and Twitter datasets are large and contain high-frequency words that are domain specific (e.g., emoticons) and are not included in the predetermined list.
+This shows that data-driven techniques to identify stop words can be much more effective than using a predetermined list.
 
@@
 where both the number of observations for the word and the tag are normalized when computing the value for each cell (see the $S_{ji}$ computation in Table \ref{tabModACTRModel}).
-So it is likely the case that a form of frequency \emph{attenuation} is already built into the Bayesian model.
+So it is likely the case that a form of frequency attenuation is already built into the Bayesian model.
 And therefore adding a frequency filter on top of this does nothing to improve accuracy, and only removes predictive information.
 
-\textcite{Stanley2013} found that the entropy-weighting technique to attenuate stop words for the Bayesian model improved model accuracy.
+\textcite{Stanley2013} found that the entropy-weighting technique to attenuate stop words for the Bayesian model significantly improved model accuracy.
 The same effect is found for the StackOverflow and Twitter datasets in this study.
@@
 For Twitter, these matrix sizes were tested alongside the default \num{3000000}.
-The four Twitter popular-hashtags datasets and the randomly-sampled StackOverflow dataset were used to test model accuracy.
+All runs from the four Twitter popular-hashtags datasets and the randomly-sampled StackOverflow dataset were used to test model accuracy.
 
@@
 However, model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random permutation model, while the frequency technique only begins to plateau.
-The Bayesian model does not plateau and accuracy continues to improves document size increases.
+The Bayesian model does not plateau and accuracy continues to improve as document size increases.
 
@@
 
-It is interesting to compare model accuracy between the Bayesian and random permutation model at different document sizes for both StackOverflow and Twitter.
+It is interesting to compare model accuracy between the Bayesian and random permutation model at the smallest and largest document sizes for both StackOverflow and Twitter.
 At smaller document sizes, the random permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as document size increases.
@@
 Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
-Accuracy measurements were the same as were used for previous runs of these datasets: mean of the mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.
+Accuracy measurements were the same as were used for previous runs of these datasets: mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.
 
@@
 However, it should be pointed out that the way the random permutation model compresses the words is completely random.
-It is not an LSA approach where the few ones and negative ones for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
+It is not an LSA approach where the few ones and negative ones (i.e., the factors) for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
 The environment vector for each word is randomly selected, and no SVD is performed to choose these in a more discriminating manner.
@@
 For a given retrieval, the random permutation model returns a set of activations that are correlation values for each of the tags.
-So a method was identified to convert a distribution of correlations to a distribution of log-odds values.
-To do this, the one-tailed cumulative distribution function value for each correlation is computed.
+So a method was created to convert a distribution of correlations to a distribution of log-odds values.
+To do this, the one-tailed cumulative distribution function value for each correlation in the distribution is computed.
 That probability is then converted to a log-odds value in the usual way.
@@
 This technique has a few nice mathematical properties:
-The conversion technique is simple and computationally cheap.
+It is simple and computationally cheap.
 Also, it behaves properly at specific points in the distribution.
-For example, if the conversion is performed one the mean correlation value in the set, the result is a log-odds value of zero.
+For example, if the conversion is performed on the mean correlation value in the set, the result is a log-odds value of zero.
 This makes sense given that log odds can be interpreted as the amount of odds information (positive or negative) that should be adjusted to the prior log odds, given that new piece of information.
@@
 This was to test how much of the variance in model accuracy is shared between the orderless and order models.
-The results show that there is very little unique variance in the order-only models, so order information contributes very little to overall model performance.
+The results show that there is very little unique variance in the order-only models, so order information contributes very little to overall model performance above and beyond what an orderless model contributes.
 This can be seen by comparing the two-component model with only prior and orderless context to the three-component model with prior, orderless, and any of the three methods for the order context term.
@@
 
+Also, the reason why the direction version of ordering is the best predictor of the three ordered components is most likely because order information does not have a large effect,
+so the two representations for words to the left and right of a tag are highly correlated.
+This means that this ordered component is essentially composed of the all co-occurrence observations in the orderless component after being randomly sampled and partitioned into two separate representations.
+The co-occurrences in these representations are highly correlated with each other, and highly correlated with the orderless component,
+which is why model accuracy is similar for the orderless component and direction ordered component, and why accuracy does not improve when adding the direction ordered component to the orderless component.
+
 Since it was shown that including word order information has little to no significant improvement on model accuracy, an attempt to incorporate word order information into the Bayesian model was not made.
@@
 The standard deviation is used instead of the confidence interval of the mean so that the plots show the overall stability of each coefficient across the different runs.
-The range of the distribution of coefficient values is much more narrow than the range of the mean coefficient values between the components of each model.
+The range of the distribution of coefficient values within a component is much more narrow than the range of the mean coefficient values between the components of each model.
 This shows that the model is not being overfit, as the weights for each component are relatively stable across the different runs.
@@
 This relative impact may be changing from either a true alteration in the underlying weight of the prior compared to context for different domains,
-or simply due to the fact that not enough information on one of the two terms is available.
+or simply due to the fact that the amount of information collected for each of the terms changes between the domains (less information leads to lower weighting in general).
 
@@
 If each user's tagging history is a useful predictor for StackOverflow,
-then this component should at least generate unique predictive variance in the case where users with large amounts of history are tested.
+then the prior component should at least generate unique predictive variance in the case where users with large amounts of history are tested.
 
@@
 However, for this type of site it only becomes important when a reasonably large amount of a user's past tagging history has been collected.
-For a randomly-sampled user on StackOverflow, a context-only model captures the majority of the predictive variance.
+For a randomly sampled user on StackOverflow, a context-only model captures the majority of the predictive variance.
 
@@
 If this number of posts had been collected for a single user, then the user-customized co-occurrence matrix for that user should lead to more accurate predictions compared to using the global matrix.
-However, only around 100 posts were used to build the user-specific matrices for each user.
+However, in this study only around 100 posts were used to build the user-specific matrices for each user.
 
@@
 However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
-So is most likely the case that the task is not done solely with declarative memory processes.
+So it is most likely the case that the task is not done solely with declarative memory processes.
 
@@
 When deriving the LSA approach, \textcite{Landauer1997} explicitly state that they make no claims that the brain is computing an SVD at the neural level;
-only that the neural machinery is set up so that the representation makes predictions that are similar to what would have been generated if the matrix decomposition for an SVD would have been computed.
+only that the neural machinery is set up such that the representation makes predictions that are similar to what would have been generated if the matrix decomposition for an SVD would have been computed.
 Perhaps the randomly-compressed representation used for the random permutation model is the way in which the brain tries to efficiently achieve the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
@@
 This work has shown how past user history is important, and how a context-only model like the random permutation model can be modified to include the prior component.
-The work has also shown how the strength of the random permutation model is not in its ability to represent word order,
+The work has also shown how the strength of the random permutation model may not be in its ability to represent word order,
 and rather in its ability to efficiently compress the full co-occurrence representation used for the Bayesian model.

### 2014-07-31 writeups/dissertation/dissertation.tex
foo

@@
 
+The best-fit decay rate values were analyzed across all dataset slices.
+Aggregate best-fit decay rates were computed by taking the median of all user's best-fit decay rate values across all dataset slices.
+The median was used since there were users in the popular-users dataset that did not have enough observations to generate stable predictions, and best-fit decay rate values for these users could be as high as 20.
+Using the median effectively trims these unstable values from the sample, without having to define a cut point for an outlier removal process.
+
+Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
+Error bars represent the 95\% bootstrapped confidence interval of the median.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
+  \caption{Best-fit decay rate}
+  \label{figPriorDecay}
+\end{figure}
+
+Decay rate values that produced the most accurate model performance are lower for the optimized learning form (0.43) of $B_{i}$ compared to the standard form (0.80).
+The optimal values for optimized learning are near 0.5, which lines up with default value used in ACT-R.
+So the decay rate values for the standard form are slightly higher than the default values used in ACT-R, but the values found when optimized learning is used line up nicely with the ACT-R defaults.
+
 Finally, model accuracy was analyzed across all dataset slices for StackOverflow and Twitter.
@@
 
-The best-fit decay rate values were also analyzed across all dataset slices.
-Aggregate best-fit decay rates were computed by taking the median of all user's best-fit decay rate value across all dataset slices.
-The median was used since there were users in the popular-users dataset that did not have enough observations to generate stable predictions, and best-fit decay rate values for these users could be as high as 20.
-Using the median effectively trims these unstable values from the sample, without having to define a cut point and go through an outlier removal process.
-
-Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
-Error bars represent the 95\% bootstrapped confidence interval of the median.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
-  \caption{Best-fit decay rate}
-  \label{figPriorDecay}
-\end{figure}
-
-Decay rate values that produced the most accurate model performance are lower for the optimized learning form (0.43) of $B_{i}$ compared to the standard form (0.80).
-The optimal values for optimized learning are near 0.5, which lines up with default value used in ACT-R.
-So the decay rate values for the standard form are slightly higher than the default values used in ACT-R, but the values found when optimized learning is used line up nicely with the ACT-R defaults.
+ACT-R sets the default decay rate value to 0.5, and the best-fit value for the standard form is higher (0.8) for these datasets.
+The figure shows that although model accuracy does improve for the standard form when increasing the decay rate parameter from 0.5 to the optimal value,
+that percentage improvement is not as large as the improvement gained from using the standard form compared to the optimized learning form.
+Nonetheless, model accuracy is highest when both the standard form is used and the decay rate value is increased.
 
@@
 the results from this study suggest that the optimal decay rate value should be set slightly higher (0.8) than the default when optimized learning is used (0.5).
-
-That said, comparing the effect size between changing the decay rate value and using the standard form,
-model accuracy improves much more when the standard form replaces the optimized learning form than when the decay rate value is increased from 0.5 to 0.8.
-For these datasets, it is much more important to use the standard form of the equation than it is to adjust the decay rate, although model accuracy does slightly improve if a higher decay rate is chosen with the standard form.
 

### 2014-07-31 dissProject/preProcess.R

@@ -1566,6 +1566,7 @@ renameColDVName <- function(tbl) {
 	setkey(tbl, DVName)
 	mapping = c('topHashtagAcrossPriorStd', 'Standard Prior Model Relaxed Across Posts',
 		    'topHashtagPostPriorStd', 'Standard Prior Model',
+		    'topHashtagPostPriorStdDefaultD', 'Standard Prior Model at Default d=0.5',
 		    'topHashtagPostPriorOL2', 'Optimized Learning Model',
 		    'topHashtagAcrossPriorStdNoffset', 'Standard Prior Model Relaxed Across Posts w/o offset',
 		    makeLogregMapping(),
@@ -1738,6 +1739,10 @@ analyzePrior <- function(modelVsPredTbl) {
 	# Check that the Ns for each dataset look right	
 	modelVsPredTbl[, list(N=.N, names=list(unique(datasetName))), by=list(dsetType, dsetGroup, runNum,datasetNameRoot)]
 	bestTbl = modelVsPredTbl[predUsedBest == T]
+	tbl = modelVsPredTbl[topHashtag & hashtagUsedP] 
+	tbl[d == 0.5 & !maxNP, DVName := paste0(DVName, 'DefaultD')]
+	tbl = rbind(tbl, tbl[d == 0.5 & maxNP][, DVName := paste0(DVName, 'DefaultD')])
+	tbl = tbl[maxNP | d == 0.5]
 	#dvDiffsTbl = plotDVDiffs(rbind(bestTbl[runNum==2, compare2DVs(.SD, c('topHashtagPostPriorStd', 'topHashtagPostPriorOL2'), sortedOrder=c(1,2)), by=list(dsetType, dsetGroup), .SDcols=colnames(bestTbl)],
 	#			       bestTbl[runNum==2, compare2DVs(.SD, c('topHashtagPostPriorStd', 'topHashtagAcrossPriorStd')), by=list(dsetType, dsetGroup), .SDcols=colnames(bestTbl)]
 	#			       ))
@@ -1745,8 +1750,10 @@ analyzePrior <- function(modelVsPredTbl) {
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorStd' & datasetName=='SOQgt500r2'])
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorOL2' & datasetName=='TFollowgt10Mr2' & d < 1])
[…]

### 2014-07-31 dissProject/preProcess.R
foo

@@ -1564,11 +1564,11 @@ groupN <- function(n, d) {
 
 renameColDVName <- function(tbl) {
 	setkey(tbl, DVName)
-	mapping = c('topHashtagAcrossPriorStd', 'Standard Prior Model Relaxed Across Posts',
-		    'topHashtagPostPriorStd', 'Standard Prior Model',
-		    'topHashtagPostPriorStdDefaultD', 'Standard Prior Model at Default d=0.5',
-		    'topHashtagPostPriorOL2', 'Optimized Learning Model',
-		    'topHashtagAcrossPriorStdNoffset', 'Standard Prior Model Relaxed Across Posts w/o offset',
+	mapping = c('topHashtagAcrossPriorStd', 'Standard prior model relaxed across posts',
+		    'topHashtagPostPriorStd', 'Standard prior model',
+		    'topHashtagPostPriorStdDefaultD', 'Standard prior model at default d=0.5',
+		    'topHashtagPostPriorOL2', 'Optimized learning model',
+		    'topHashtagAcrossPriorStdNoffset', 'Standard prior model relaxed across posts w/o offset',
 		    makeLogregMapping(),
 		    makeStandardMappingBayes('', 'w/ entropy'),
 		    makeStandardMappingBayes('Freq', 'w/ freq'),
@@ -1751,7 +1751,8 @@ analyzePrior <- function(modelVsPredTbl) {
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorOL2' & datasetName=='TFollowgt10Mr2' & d < 1])
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorOL2' & datasetName=='SOQgt500r2' & d < 1])
 	resTbl = compareOptimalDs(tbl[DVName %in% c(paste0('topHashtagPostPriorStd', c('')), paste0('topHashtagPostPriorOL2', c(''))) & runNum == 2])
-	resTbl = compareOptimalAcc(tbl[DVName %in% c(paste0('topHashtagPostPriorStd', c('', 'DefaultD')), paste0('topHashtagPostPriorOL2', c(''))) & runNum == 2])
+	DVNames = c(paste0('topHashtagPostPriorStd', c('', 'DefaultD')), paste0('topHashtagPostPriorOL2', c('')))
+	resTbl = compareOptimalAcc(tbl[DVName %in% DVNames & runNum == 2])
[…]

### 2014-07-31 writeups/dissertation/dissertation.tex
foo

@@
 The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
-It is important that this measure scales well enough to handle the sheer volume of information present in online social media datasets.
+It is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
 Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
@@
 Unlike StackOverflow, the entire Twitter dataset is not released to the public.
-Due to this and the sheer size of Twitter, subsets of the Twitter dataset were collected.
+Due to this and the scale of Twitter, subsets of the Twitter dataset were collected.
 
@@
 Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
-Multiple datasets were collected to ensure that the results were not dependent on a particular sample of hashtags, and would generalize across different popular-hashtag datasets and different time periods.
+Multiple datasets were collected to ensure that the results were not dependent on a particular sample of hashtags and would generalize across different popular-hashtag datasets and different time periods.
 Each dataset required around four days to collect enough tweets to reach the quota, and one collection was done each week.
@@
 
-\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing 'ed' and 'ing' endings), but
+\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing ``ed'' and ``ing'' endings), but
 with the StackOverflow dataset, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
@@
 
-How does a user's prior hashtag use influence their likelihood of choosing that hashtag in the future?
+How does a user's prior tag use influence their likelihood of choosing that tag in the future?
 This question was tested by examining how well ACT-R's base-level activation component of declarative memory predicts the tags that authors choose, given each author's past tagging behavior.
-The model predicts that the more recently and frequently used chunks have higher prior activations, and are more likely to be needed in the future.
+The model predicts that the more recently and frequently used tags have higher prior activations, and are more likely to be needed in the future.
 
@@
 
-20 different values of the decay rate parameter were explored for each slice
+22 different values of the decay rate parameter were explored for each slice
 (0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.8, 2, 5, 10, 20).
-The values were chosen to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d=.5$), to a pure recency model ($d=20$).
+The values were chosen to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d \approx .5$), to a pure recency model ($d=20$).
 The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
@@
 
-The best-fit decay rate values were analyzed across all dataset slices.
 Aggregate best-fit decay rates were computed by taking the median of all user's best-fit decay rate values across all dataset slices.
@@
 Nonetheless, the only real reason not to use the standard form is due to computational resource limitations.
-Since this becomes less of an issue as time progresses and hardware speed increases, it may be time to reconsider the default value used in ACT-R.
+Since this is becoming less of an issue as time progresses and hardware speed increases, it may be time to reconsider the default form of base-level learning used in ACT-R.
 It may be the case that the optimized learning form of the equation is no longer computationally necessary.
@@
 However, this relaxed metric was used for the rest of the experiments that involve combining terms because the optimal weights found by the regression are based on using the relaxed metric.
-If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression was not optimized to the constrained form, and weights may change if that was done.
+If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression was not optimized to the constrained form, and weights may change as a consequence.
 Also, there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation where a small set of 1s occurred. 
@@
 
-Model accuracy was computed on the popular-hashtags dataset for Twitter and a random sample of posts across the entire StackOverflow dataset.
+Model accuracy was tested on the popular-hashtags dataset for Twitter and a random sample of posts across the entire StackOverflow dataset.
 For each of the four Twitter popular-hashtags datasets, model retrievals were done across 15 runs and 500 posts within each run.
@@
 
-The context component for both the random permutation and Bayesian model requires the same co-occurrence matrix of words and observed tags.
+The context component for both the random permutation and Bayesian model is based on the same co-occurrence matrix of words and observed tags.
 Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
   \caption{Impact of adding model offset for StackOverflow}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
   \caption{Impact of adding model offset for Twitter}
@@
 The random permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
-So, for example, the ``standard Prior model relaxed across posts without offset'' means that only the prior model component was used, and the offset for that component was not subtracted off when determining activation.
+So, for example, the ``standard prior model relaxed across posts without offset'' means that only the prior model component was used, and the offset for that component was not subtracted off when determining activation.
 As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.
@@
 However, \citeauthor{Sahlgren2008} did not try to weight the stop words, and only tried methods to remove them.
-So alongside looking at frequency weighting, previously derived, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on Entropy was also explored.
+So alongside looking at frequency weighting, previously derived, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on entropy was also explored.
 
@@
 
+% FIXME: Cite is too recent
 The same previously-derived stoplist from \textcite{Salton2014} that \textcite{Sahlgren2008} used was used for this analysis,
@@
 
-The original plot shows the size of the spikes when high-frequency words are removed.
+The original plot shows the size of the spikes when high-frequency words are not removed.
 The entropy plot shows how the size decreases by more than an order of magnitude when the entropy weighting measure is used.
@@
 Since accuracy measurements when using a predetermined list to remove stop words for the random permutation model were much lower than the frequency and entropy methods, this method was not examined for the Bayesian model.
-The results for the frequency and entropy techniques for the Bayesian model on the StackOverflow dataset are include in Figure \ref{figContextStopBSO}.
+The results for the frequency and entropy techniques for the Bayesian model on the StackOverflow dataset are included in Figure \ref{figContextStopBSO}.
 
@@
 As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random permutation model is actually beneficial, and the compressed random permutation model should be used.
-This noise is beneficial because it helps soften spurious $S_{ji}$ values when the values for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller document sizes.
+This noise is beneficial because it helps soften spurious $S_{ji}$ values when the counts for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller document sizes.
 
@@
 The plot is also broken out for different size datasets used to build the co-occurrence matrix.
+
 Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy.
@@
 Results are similar for Twitter.
+Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100}, and little to no impact when increasing from \num{2048} to \num{10000}.
+This result is consistent across all document sizes used to build the co-occurrence matrix.
 
@@
 \end{figure}
-
-Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100}, and little to no impact when increasing from \num{2048} to \num{10000}.
-This result is consistent across all document sizes used to build the co-occurrence matrix.
 
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
   \caption{Random permutation model accuracy when using log-odds transformation on context for StackOverflow}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
   \caption{Random permutation model accuracy when using log-odds transformation on context for Twitter}
@@
 
-The same Twitter popular-users datasets and StackOverflow randomly-sampled datasets were used for testing.
+The same four Twitter popular-users datasets were used for testing.
 Both an ordered and orderless context component were created for the random permutation models.
@@
 But including word order information does not dramatically improve performance. 
-For the random permutation model, it is much more important to properly handle high-frequency stop words and include a prior component than it is to include word order information,
+For the random permutation model, it is much more important to properly handle low-predictor stop words and include a prior component than it is to include word order information,
 at least for the Twitter dataset tested in this study.
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffSO-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffSO-crop.pdf}}}
   \caption{Coefficient values for full models for StackOverflow}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffT-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffT-crop.pdf}}}
   \caption{Coefficient values for full models for Twitter}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
   \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddT-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddT-crop.pdf}}}
   \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
@@
 However, since computing activations for the context component take an order of magnitude more wall-clock time than computing the prior component, not all observations could be included for the test.
-To reduce the observations, the most recent 50 posts from each user in the datasets were sampled.
+To reduce the observations, the most recent 50 posts from each user in the datasets were selected.
 The total number of users sampled was also reduced to 40 users for StackOverflow and 80 users for Twitter.
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiSO-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiSO-crop.pdf}}}
   \caption{Model accuracy with user-customized context component for StackOverflow}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiT-crop.pdf}}}
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiT-crop.pdf}}}
   \caption{Model accuracy with user-customized context component for Twitter}
@@
 since the noise introduced from the compression helps smooth out activation values until a larger number of co-occurrences are observed and the co-occurrence values stabilize.
-For a behavioral study, novel words could be introduced alongside the corpus of known words, and the participant could be asked to make retrievals of the novel words and known words as more novel words are introduced.
+For a behavioral study, novel words could be introduced alongside a corpus of known words, and the participant could be asked to make retrievals of the novel words and known words as more novel words are introduced.
 Both models would make different predictions about the rate at which activation values for the novel words stabilize relative to the known words, and such make different predictions about retrieval error rates over time.

### 2014-08-01 writeups/dissertation/dissertation.tex
foo

@@
 \usepackage[utf8]{inputenc}
+
+% Default apa6 formatting for a CI
+\newcommand{\CI}[2]{95\% CI [#1, #2]}
 
@@
 \textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
-Model results were on par with a large sample of applicants to american universities from non-English speaking countries (64.4\% compared to 64.5\%).
+Model results were on par with a large sample of applicants to american universities from non-English speaking countries (\num{64.4}\% compared to \num{64.5}\%).
 Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
@@
 For example, \textcite{Jones2007} compared a vector-based model called BEAGLE directly to LSA, training and testing both models on the same corpus used in \textcite{Landauer1997}.
-The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the task (55.6\% vs. 55.3\%).
-However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to 57.5\%.
+The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the task (\num{55.6}\% vs. \num{55.3}\%).
+However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to \num{57.5}\%.
 It seems reasonable to expect that task performance will improve much more on tasks where word order is used (e.g., generating a hashtag in the middle of a sentence when composing a tweet).
@@
 This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
-Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to 73.75\%, which is 10\% higher than the TASA-trained LSA.
+Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to \num{73.75}\%, which is 10\% higher than the TASA-trained LSA.
 This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.
@@
 This model uses circular convolution and circular correlation as a way to encode and decode word order.
-BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (55.6\% compared to 55.3\%).
-Also, BEAGLE with unordered context and word order was more correlated to the WordNet synonym database than LSA (-.311 compared to -.165).
+BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (\num{55.6}\% compared to \num{55.3}\%).
+Also, BEAGLE with unordered context and word order was more correlated to the WordNet synonym database than LSA (\num{-.311} compared to \num{-.165}).
 Further, it was shown that incorporating word order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
@@
 
-The best-fit values are close to the 0.5 default value used in ACT-R when the optimized learning form is used.
+The best-fit values are close to the \num{0.5} default value used in ACT-R when the optimized learning form is used.
 Comparing across optimized learning and the standard form, the best-fit values for the standard form are slightly higher for both the StackOverflow and Twitter dataset slices.
@@
 
-Decay rate values that produced the most accurate model performance are lower for the optimized learning form (0.43) of $B_{i}$ compared to the standard form (0.80).
-The optimal values for optimized learning are near 0.5, which lines up with default value used in ACT-R.
+Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{0.43}) of $B_{i}$ compared to the standard form (\num{0.80}).
+The optimal values for optimized learning are near \num{0.5}, which lines up with default value used in ACT-R.
 So the decay rate values for the standard form are slightly higher than the default values used in ACT-R, but the values found when optimized learning is used line up nicely with the ACT-R defaults.
@@
 
-Accuracy for the standard form (0.35) is higher than the optimized learning form (0.30).
+Accuracy for the standard form (\num{0.35}) is higher than the optimized learning form (\num{0.30}).
 This makes sense given that the standard form does not assume equal presentation rate for each chunk, and hashtags on Twitter in particular show trends of high rate of use and then shift to low rate of use.
@@
 
-ACT-R sets the default decay rate value to 0.5, and the best-fit value for the standard form is higher (0.8) for these datasets.
-The figure shows that although model accuracy does improve for the standard form when increasing the decay rate parameter from 0.5 to the optimal value,
+ACT-R sets the default decay rate value to \num{0.5}, and the best-fit value for the standard form is higher (\num{0.8}) for these datasets.
+The figure shows that although model accuracy does improve for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
 that percentage improvement is not as large as the improvement gained from using the standard form compared to the optimized learning form.
@@
 And if the standard form is used as the default (or simply tried in another experiment),
-the results from this study suggest that the optimal decay rate value should be set slightly higher (0.8) than the default when optimized learning is used (0.5).
+the results from this study suggest that the optimal decay rate value should be set slightly higher (\num{0.8}) than the default when optimized learning is used (\num{0.5}).
 
@@
 The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
-Using a predefined stoplist makes the smallest improvement, while the entropy and frequency techniques improve performance the most, and frequency shows a slight edge over entropy.
+Using a predefined stoplist makes the smallest improvement (\num{.45} to \num{.49}), while the entropy (\num{.60}) and frequency (\num{.63}) techniques improve performance the most,
+and frequency shows a slight edge over entropy.
 However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
@@
 
-For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words.
-Removing stop-words based on a predetermined list actually decreases accuracy, and using data-driven frequency filtering has only a small to no effect.
+For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\num{.25} to \num{.26}).
+Removing stop-words based on a predetermined list actually decreases accuracy (\num{.25} to \num{.23}), and using data-driven frequency filtering has only a small to no effect.
 
@@
 
-The results show that entropy weighting produces the most accurate results.
-Also, accuracy actually decreases when using frequency filtering (compared to no filtering) for the Bayesian model.
+The results show that entropy weighting produces the most accurate results (\num{.61} to \num{.63}).
+Also, accuracy actually decreases when using frequency filtering (compared to no filtering) (\num{.61} to \num{.56}) for the Bayesian model.
 So entropy weighting is the clear winner in this case.
@@
 
-Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer, and frequency filtering decreases performance relative to no filtering.
+Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer (\num{.29} to \num{.32}), and frequency filtering decreases performance relative to no filtering (\num{.29} to \num{.28}).
 
@@
 
-Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy.
-But \num{2048} rows is sufficient for these datasets, as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy.
+Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy (\num{.60} to \num{.55} for the \num{1000000} document dataset size).
+But \num{2048} rows is sufficient for these datasets,
+as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy (\num{.60} to \num{.60} for the \num{1000000} document size).
 
 Results are similar for Twitter.
-Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100}, and little to no impact when increasing from \num{2048} to \num{10000}.
-This result is consistent across all document sizes used to build the co-occurrence matrix.
+Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100} (\num{.26} to \num{.24} for the \num{3000000} document size),
+and little to no impact when increasing from \num{2048} to \num{10000} (\num{.26} to \num{.26} at \num{3000000}).
+This results are consistent across all document sizes used to build the co-occurrence matrix.
 
@@
 
-Using the log-odds transformation technique has little to no effect on model accuracy for both StackOverflow and Twitter.
+Using the log-odds transformation technique has a small effect on model accuracy for StackOverflow (\num{.45} to \num{.47}) and minimal to no effect on Twitter (\num{.25} to \num{.25}).
 When the technique is used and accuracy is examined for a model with only a single predictor, the transformation does not change accuracy.
@@
 
-For both the Bayesian and random permutation models, accuracy remains the same when the prior component is removed for StackOverflow.
-Accuracy does dramatically decrease for Twitter when the prior is removed, however.
+For StackOverflow, accuracy remains the same when the prior component is removed for both the Bayesian (\num{.63} to \num{.63}) and random permutation (\num{.63} to \num{.63}) models.
+Accuracy does dramatically decrease for Twitter when the prior is removed, however (\num{.32} to \num{.14} for Bayesian, \num{.25} to \num{.12} for random permutation).
 
@@
 In this case, past tag history does contain unique variance for accurately predicting tag use.
-Accuracy improves for both the Bayesian and random permutation models when the prior model component is combined with context compared to context alone.
+Accuracy improves for both the Bayesian (\num{.51} to \num{.63}) and random permutation (\num{.47} to \num{.58}) models when the prior model component is combined with context compared to context alone.
 

### 2014-08-01 dissProject/preProcess.R

@@ -1947,6 +1947,18 @@ analyzeContext <- function(modelVsPredTbl) {
 	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweet', 'PriorStdTweetFrentropy', 'Tweet',
 				       'TweetNentropy', 'TweetFreq', 'TweetFrentropy', 'PriorStdTweetNentropy', 'PriorStdTweetFreq'))
 	compareMeanDVDefault(tbl[sizeNum == 2 & dsetType == 'twitter' & DVName %in% DVNames], acc, figName='allWeightingsTSji', groupCol='dsetGroup')
+	# RESULT: Bayes and RP with frequency cutoff scales; RP with entropy saturates for SO
+	# SO Entropy all sizes
+	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTitleBody', 
+				       'PriorStdTitleOrderlessBodyOrderless',
+				       'PriorStdTitleOrderlessEntropyBodyOrderlessEntropy',
+				       'PriorStdTitleOrderlessFreqBodyOrderlessFreq'))
+	compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames & sizeNum != 2], acc, figName='freqVsEntropyBySizeSO', groupCol='sizeNum')
+	# T Entropy all sizes
+	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweet',
+				       'PriorStdTweetOrderEntropyTweetOrderlessEntropy',
+				       'PriorStdTweetOrderFreqTweetOrderlessFreq'))
+	compareMeanDVDefault(tbl[dsetType == 'twitter' & DVName %in% DVNames], acc, figName='freqVsEntropyBySizeT', groupCol='sizeNum')
 	# RESULT: Increasing rows in matrix does not dramatically improve performance, even at different size datasets; entropy and freq cutoff is much more important for RP
 	# SO Entropy all sizes
 	DVNames = asTopHashtagAcross(c('PriorStd',
@@ -1960,18 +1972,6 @@ analyzeContext <- function(modelVsPredTbl) {
 				       'PriorStdTweetOrderSmdimTweetOrderlessSmdim',
 				       'PriorStdTweetOrderLgdimTweetOrderlessLgdim'))
[…]

### 2014-08-01 writeups/dissertation/dissertation.tex
foo

@@
 
-Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites.
+Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites, except that the full Bayesian model without an offset was slightly more accurate on Twitter
+(\num{.53} to \num{.60} and \num{.22} to \num{.45} for Bayesian and random permutation for StackOverflow, \num{.31} to \num{.29} and \num{.24} to \num{.25} for Bayesian and random permutation for Twitter).
 Accuracy is improved even when only a single term is used.
@@
 For StackOverflow, model accuracy for the user-specific co-occurrence matrix (labeled as ``user sji'') can be compared against the default global co-occurrence matrix.
-Accuracy decreases dramatically when a user-specific co-occurrence matrix is used.
+Accuracy decreases dramatically when a user-specific co-occurrence matrix is used for both the Bayesian (\num{.64} to \num{.35}) and random permutation (\num{.58} to \num{.36}) models.
 
@@
 
-Model accuracy when only the single user-customized context predictor is used is relatively poor, and the prior term is a much better predictor.
-Further, when this component is combined with the prior component, model performance improves only slightly over performance when only the prior is used.
+Model accuracy when only the single user-customized context predictor is used is relatively poor (\num{.17} and \num{.21} for Bayesian and random permutation),
+and the prior term is a much better predictor (\num{.37}).
+Further, when this component is combined with the prior component, model performance improves only slightly over performance when only the prior is used (\num{.37} to \num{.39} for both the Bayesian and random permutation).
 
@@
 As was shown for Twitter, model accuracy can slightly improve over a model that just uses a custom prior, even when the co-occurrence matrix only contains around 100 posts.
+
+Further, if the number of co-occurrences is small, and a user-custom co-occurrence matrix is used, it may be advantageous to use the compressed random permutation model instead of the uncompressed Bayesian model.
+For the Twitter popular-users dataset, only a user-custom co-occurrence matrix could be built, and when this matrix is used the compressed random permutation model performed better than the uncompressed Bayesian model.
+When the global co-occurrence matrix was removed from the StackOverflow dataset, and a user-custom matrix was used instead, the accuracy for the random permutation model decreased less than the Bayesian model.
+This provides further evidence showing that there is an interaction between dataset size and optimal model:
+At lower numbers of co-occurrences the compressed random permutation model is able to perform better than the Bayesian model,
+and the Bayesian model surpasses the compressed model at higher co-occurrence counts once the cells in the full co-occurrence matrix stabilize.
 

### 2014-08-01 writeups/dissertation/dissertation.tex
foo

@@
 
+\usepackage[nodayofweek]{datetime}% http://ctan.org/pkg/datetime
 \usepackage{xfrac}
@@
 \leftheader{Beitzel}
+
+\def\committee{
+  Joseph R. Cavallaro, Chair \\
+  Professor of Electrical and Computer Engineering and Computer Science \and
+  Behnaam Aazhang \\
+  J.S. Abercrombie Professor of Electrical and Computer Engineering \and
+  Scott Rixner\\
+  Associate Professor of Electrical and Computer Engineering and Computer Science \and
+  Ashutosh Sabharwal\\
+  Assistant Professor of Electrical and Computer Engineering \\
+}
+
+% Replace the title page in the apa6 style sheet with the title page from the Rice template
+% The title page style was taken from Rice University LaTeX PhD template: http://scholarship.rice.edu/handle/1911/21747
+% The following code for the abstract page was taken from the apa6 \maketitle command.
+\makeatletter
+\def\maketitle{
+  \begin{titlepage}
+    \singlespacing
+    \null
+    \setcounter{page}{1}
+    \vfill
+    \vbox to 0pt{\vss\vbox to 8.75in{\parskip 0pt \parindent 0pt\centering
+      {\large \uppercase\expandafter{Rice University} \par} \vskip 0pt plus 3fil
+      {\Large \bf \@title \par} \vskip 0pt plus 1.3fil
+      {\large by \par} \vskip 0pt plus 1fil
+      {\large \bf \@author \par} \vskip 0pt plus 3fil
+      {\large \sc A Thesis Submitted \\
+	in Partial Fulfillment of the \\
+      Requirements for the Degree \par} \vskip 0pt plus 1.1fil
+      {\large \bf Doctor of Philosophy \par} \vskip 0pt plus 4.5fil
+      \begingroup
+      \leftskip .5\textwidth \rightskip 0pt plus 1fil
+      {\sc Approved, Thesis Committee: \par} \vskip 2.5em
+      \moveright\leftskip\vbox{\hrule width \leftskip}
+      \def\and{\par \vskip 2.25em minus 1em\moveright\leftskip\vbox{\hrule width \leftskip}}
+      \committee
+      \par \endgroup
+      \vskip 0pt plus 2fil
+    {\large Houston, TX \par} \vskip 0pt plus .8fil
+    {\large \monthnameenglish, \number\year \par}
+  }}
+  \doublespacing
+  \newpage
+  %BDB\hyphenpenalty 10000
+  \fussy
+  \@ifundefined{@abstract}{}{%
+    \section{\normalfont\normalsize\abstractname}% BDB
+    \noindent\@abstract\par% BDB
+    \@ifundefined{@keywords}{}{%
+      \setlength{\parindent}{0.4in}% BDB
+      \indent\textit{\keywordname:} \@keywords%
+    }%
+    \newpage
+  }
+  \@ifundefined{def@donotrepeattitle}{
+    \section{\protect\normalfont{\@title}}
+  }{}%
+  \raggedright%
+  \setlength{\parindent}{0.4in}%
+\end{titlepage}}
+\makeatother
 

### 2014-08-01 dissProject/preProcess.R
foo

@@ -1784,15 +1784,16 @@ analyzePUser <- function(modelVsPredTbl) {
 	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTitleBody', 'TitleBody', 'Body', 'Title',
 				       'TitleOrderlessEnthymanBodyOrderlessEnthyman',
 				       'PriorStdTitleOrderlessEnthymanBodyOrderlessEnthyman'))
-	compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames], acc, figName='priorAddPUserSO')
-	# RESULT: User sji is not very helpful for Twitter, and prior is useful again
-	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweetUsercontext', 'PriorStdTweetOrderlessEnthyser', 'TweetUsercontext', 'TweetOrderlessEnthyser'))
-	compareMeanDVDefault(tbl[dsetType == 'twitter' & DVName %in% DVNames], acc, figName='userSjiSO')
+	resTbl = compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames], acc, figName='priorAddPUserSO')
+	resTbl[, mean(meanVal), by=DVName]
 	# RESULT: User sji is much less helpful than standard sji for StackOverflow
 	DVNames = asTopHashtagAcross(c('TitleOrderlessEnthyser', 'TitleUsercontext', 'TitleOrderlessEnthyman', 'TitleOrderlessEnthyser', 'Title',
 				       'PriorStdTitleOrderlessEnthyserBodyOrderlessEnthyser', 'PriorStdTitleBody', 'PriorStdTitleOrderlessEnthymanBodyOrderlessEnthyman',
 				       'PriorStdTitleUsercontextBodyUsercontext'))
-	compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames], acc, figName='userSjiT')
+	resTbl = compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames], acc, figName='userSjiT')
+	# RESULT: User sji is not very helpful for Twitter, and prior is useful again
+	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweetUsercontext', 'PriorStdTweetOrderlessEnthyser', 'TweetUsercontext', 'TweetOrderlessEnthyser'))
[…]

### 2014-08-11 writeups/dissertation/dissertation.tex
foo

--- 2014-08-01 writeups/dissertation/dissertation.tex
@@
 
-\title{Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow}
+\title{Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow}
 \shorttitle{Comparing memory models of tag retrieval}
@@
 \leftheader{Beitzel}
+
+\def\committee{
+  Michael D. Byrne, Chair \\
+  Professor, Psychology and Computer Science \and
+  Phillip Kortum \\
+  Assistant Professor, Psychology \and
+  David Lane \\
+  Associate Professor, Psychology, Statistics, and Management \and
+  Devika Subramanian \\
+  Professor, Computer Science and Electrical Engineering \\
+}
+
+% Replace the title page in the apa6 style sheet with the title page from the Rice template
+% The title page style was taken from Rice University LaTeX PhD template: http://scholarship.rice.edu/handle/1911/21747
+% The following code for the abstract page was taken from the apa6 \maketitle command.
+\makeatletter
+\def\makeDisCover{
+  \def\maketitle{
+    \begin{titlepage}
+      \singlespacing
+      \null
+      \vfill
+      \vbox to 0pt{\vss\vbox to 8.75in{\parskip 0pt \parindent 0pt\centering
+	{\uppercase\expandafter{Rice University} \par} \vskip 0pt plus 3fil
+	{\bf \@title \par} \vskip 0pt plus 1.3fil
+	{by \par} \vskip 0pt plus 1fil
+	{\bf \@author \par} \vskip 0pt plus 3fil
+	{\sc A Thesis Submitted \\
+	  in Partial Fulfillment of the \\
+	Requirements for the Degree \par} \vskip 0pt plus 1.1fil
+	{\bf Doctor of Philosophy \par} \vskip 0pt plus 4.5fil
+	\begingroup
+	\leftskip .45\textwidth \rightskip 0pt plus 1fil
+	{\sc Approved, Thesis Committee: \par} \vskip 2.5em
+	\moveright\leftskip\vbox{\hrule width \leftskip}
+	\def\and{\par \vskip 2.25em minus 1em\moveright\leftskip\vbox{\hrule width \leftskip}}
+	\committee
+	\par \endgroup
+	\vskip 0pt plus 2fil
+	{Houston, TEXAS \par} \vskip 0pt 
+	{\monthnameenglish{} \number\year \par}
+      }}
+      \doublespacing
+      \newpage
+  %BDB\hyphenpenalty 10000
+      \fussy
+      \@ifundefined{@abstract}{}{%
+	\section{\normalfont\normalsize\abstractname}% BDB
+	\noindent\@abstract\par% BDB
+	\@ifundefined{@keywords}{}{%
+	  \setlength{\parindent}{0.4in}% BDB
+	  \indent\textit{\keywordname:} \@keywords%
+	}%
+	\newpage
+      }
+      \@ifundefined{def@donotrepeattitle}{
+	\section{\protect\normalfont{\@title}}
+      }{}%
+      \raggedright%
+      \setlength{\parindent}{0.4in}%
+    \end{titlepage}}
+  }
+\makeatother
+
+%\makeDisCover
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
-  \caption{Best-fit decay rate}
+  \caption{Overall best-fit decay rate for StackOverflow and Twitter}
   \label{figPriorDecay}
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc--crop.pdf}}}
-  \caption{Model accuracy}
+  \caption{Overall model accuracy for StackOverflow and Twitter}
   \label{figPriorAcc}
@@
   \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
-  \caption{Impact of adding model offset for StackOverflow}
+  \caption{Impact of adding offset component for StackOverflow}
   \label{figContextOffsetSO}
@@
   \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
-  \caption{Impact of adding model offset for Twitter}
+  \caption{Impact of adding offset component for Twitter}
   \label{figContextOffsetT}
@@
 
-% FIXME: Cite is too recent
-The same previously-derived stoplist from \textcite{Salton2014} that \textcite{Sahlgren2008} used was used for this analysis,
+The same 571-word Cornell SMART stoplist used by \textcite{Sahlgren2008} was also used for this analysis,
 The same entropy weighting method using by \textcite{Stanley2013} was examined,
@@
   \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
-  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
+  \caption{Model accuracy when prior is removed for StackOverflow randomly-sampled dataset}
   \label{figContextPriorneededSO}
@@
   \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddT-crop.pdf}}}
-  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
+  \caption{Model accuracy when prior is removed for Twitter popular-hashtags dataset}
   \label{figContextPriorneededT}
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddPUserSO-crop.pdf}}}
-  \caption{Unique variance in prior component for popular StackOverflow users}
+  \caption{Model accuracy when prior is removed for StackOverflow popular-users dataset}
   \label{figPUAddPrior}
@@
 
-\subsubsection{Result}
+\subsubsection{Results}
 
@@
 
+\subsection{General Results}
+
+Table \ref{tabOverallResults} summarizes model accuracies for full models (all components) on all of the datasets tested in the study.
+
+\begin{table}[!ht]
+  \caption{Summary of accuracy for each model and dataset}
+  \label{tabOverallResults}
+  {\tabulinesep=1.2mm
+    \begin{tabu}{llll}
+      \hline
+	Site 		& Dataset		& Model						& Accuracy \\
+	\hline
+	StackOverflow	& popular-users		& Bayesian with entropy				& \num{.63} \\
+			& 			& RP with entropy and log odds 			& \num{.58} \\
+			& randomly-sampled	& Bayesian with entropy 			& \num{.63} \\
+			& 			& RP with entropy and log odds			& \num{.63} \\
+	Twitter		& popular-users		& Bayesian with entropy and user sji		& \num{.39} \\
+			& 			& RP with entropy, log odds, and user sji	& \num{.39} \\
+			& popular-hashtags	& Bayesian with entropy				& \num{.32} \\
+			& 			& RP with entropy and log odds			& \num{.25} \\
+	\hline
+    \end{tabu}
+  }
+\end{table}
+
+After modifying the models by adding an entropy weighting mechanism to both models and a prior term to the random permutation model,
+accuracy for the compressed random permutation model was nearly equal to the Bayesian model.
+The largest difference between the two models was on the Twitter popular-hashtags dataset (7 percent).
+Both models performed nearly equal for the StackOverflow datasets tested,
+and both models were much more accurate when tested on StackOverflow compared to Twitter.
+
 \section{General Discussion}
@@
 Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
+The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
 These results provide support that choosing a hashtag when composing a tweet and tagging a StackOverflow post is a long-term memory retrieval process.
@@
 and perhaps it is the case that the rest of the variance left to predict is from these context components that the author is using but are not yet included in the model.
-That is, perhaps this tagging process is completely declarative, and the remaining cues that the author is using are yet to be discovered.
 

### 2014-08-11 dissProject/preProcess.R

@@ -1790,10 +1790,10 @@ analyzePUser <- function(modelVsPredTbl) {
 	DVNames = asTopHashtagAcross(c('TitleOrderlessEnthyser', 'TitleUsercontext', 'TitleOrderlessEnthyman', 'TitleOrderlessEnthyser', 'Title',
 				       'PriorStdTitleOrderlessEnthyserBodyOrderlessEnthyser', 'PriorStdTitleBody', 'PriorStdTitleOrderlessEnthymanBodyOrderlessEnthyman',
 				       'PriorStdTitleUsercontextBodyUsercontext'))
-	resTbl = compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames], acc, figName='userSjiT')
+	resTbl = compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames], acc, figName='userSjiSO')
 	# RESULT: User sji is not very helpful for Twitter, and prior is useful again
 	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweetUsercontext', 'PriorStdTweetOrderlessEnthyser', 'TweetUsercontext', 'TweetOrderlessEnthyser'))
-	resTbl = compareMeanDVDefault(tbl[dsetType == 'twitter' & DVName %in% DVNames], acc, figName='userSjiSO')
+	resTbl = compareMeanDVDefault(tbl[dsetType == 'twitter' & DVName %in% DVNames], acc, figName='userSjiT')
 	# RESULT: .7 d is better for T than SO, but effect is small.
 	# Also once additional predictors are added to the model, the effect diminishes.
 	# SO and T compare two d values	

### 2014-08-13 writeups/dissertation/dissertation.tex
foo

@@
 
-Another issue with LSA is that adding additional observations to an previously-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
+Another issue with LSA is that adding additional observations to a previously-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
 Since computing the SVD is the most computationally expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
@@
 
-Random permutations uses one of simplest permutations possible to create a new environment vector.
+Random permutations uses one of simplest operations possible to create a new environment vector.
 This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
@@
 Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
-A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
+A tag-suggestion tool has also been deployed for the \url{tex.stackexchange.org} and \url{superuser.com} StackExchange sites.
 Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
@@
 
-It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
+It is interesting that in both studies the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
 This suggests that more cognitively-plausible memory retrieval models may outperform more general machine learning approaches in this particular environment
@@
 
-A tag suggestion tool has been deployed for the \url{http://tex.stackexchange.org} \parencite{LatexTags2013} and \url{http://superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
+A tag suggestion tool has been deployed for the \url{tex.stackexchange.org} \parencite{LatexTags2013} and \url{superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
 Latex, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
@@
 and it seams reasonable to think that tokenizing all of the informal language and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
-So the same tokenization algorithm was used for both StackOverflow and Twitter, since performance was similar for both algorithms on StackOverflow,
+So the same tokenization algorithm was used for both StackOverflow and Twitter, since performance was similar to \textcite{Stanley2013} for StackOverflow,
 and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.
@@
 This helps illustrate how challenging modeling this task can be.
-The model's goal is to predict at each instance what tag will be chosen, but since the tag space is so large and varied, the model must predict the correct few tags that are used from a much larger set of possible tags.
+The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is dynamically changing over time.
 
@@
 
-The best-fit values are close to the \num{0.5} default value used in ACT-R when the optimized learning form is used.
+The best-fit values are close to the \num{0.5} default value set in ACT-R when the optimized learning form is used.
 Comparing across optimized learning and the standard form, the best-fit values for the standard form are slightly higher for both the StackOverflow and Twitter dataset slices.

### 2014-08-14 writeups/dissertation/dissertation.tex
foo

@@
 Nonetheless, the only real reason not to use the standard form is due to computational resource limitations.
-Since this is becoming less of an issue as time progresses and hardware speed increases, it may be time to reconsider the default form of base-level learning used in ACT-R.
+Since this is becoming less of an issue as hardware speed increases, it may be time to reconsider the default form of base-level learning used in ACT-R.
 It may be the case that the optimized learning form of the equation is no longer computationally necessary.
@@
 For each of the four Twitter popular-hashtags datasets, model retrievals were done across 15 runs and 500 posts within each run.
-For the StackOverflow dataset, 5 runs of 100 posts were used, since the results for StackOverflow are more stable and retrievals take more wall-clock time for this dataset due to the size of the co-occurrence matrix.
+For the StackOverflow randomly-sampled dataset, 5 runs of 100 posts were used,
+since the results for StackOverflow are more stable and retrievals take more wall-clock time for this dataset due to the size of the co-occurrence matrix.
 All runs contained different posts, and all posts used for runs were not used to generate the co-occurrence matrix for the two retrieval models.
@@
 
-The impact of adding the offset to each of the models for StackOverflow and Twitter are included in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.
+The impact of adding the offset to each of the models for StackOverflow and Twitter is shown in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.
 The error bars represent the 95\% bootstrapped confidence interval of the mean model accuracy across all runs in dataset.
@@
 The random permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
-So, for example, the ``standard prior model relaxed across posts without offset'' means that only the prior model component was used, and the offset for that component was not subtracted off when determining activation.
+So, for example, the ``standard prior model relaxed across posts without offset'' means that only the prior model component was used and the offset for that component was not subtracted off when determining activation.
 As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.
 
-Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites, except that the full Bayesian model without an offset was slightly more accurate on Twitter
-(\num{.53} to \num{.60} and \num{.22} to \num{.45} for Bayesian and random permutation for StackOverflow, \num{.31} to \num{.29} and \num{.24} to \num{.25} for Bayesian and random permutation for Twitter).
+Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites
+(\num{.53} to \num{.60} and \num{.22} to \num{.45} for Bayesian and random permutation for StackOverflow, \num{.31} to \num{.29} and \num{.24} to \num{.25} for Bayesian and random permutation for Twitter),
+except that the full Bayesian model without an offset was slightly more accurate on Twitter.
 Accuracy is improved even when only a single term is used.
@@
 The four subsets were collected across a period of one month.
+Each was collected sequentially, and about one subset was collected per week.
 The hashtags used in each subset were chosen by scraping a site that published the current trending hashtags for each city.
-Hashtags for each US city were collected, and then 400 were randomly sampled from this set.
+Before starting to collect for each subset, the current trending popular hashtags for each US city were collected, and then 400 were randomly sampled from this set.
 Only 400 were selected since this is the maximum that the Twitter API can monitor at one time.
@@
 Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and using a previously-derived stoplist hardly increased performance.
-However, \citeauthor{Sahlgren2008} did not try to weight the stop words, and only tried methods to remove them.
+However, \citeauthor{Sahlgren2008} did not try to weight the stop words and only tried methods to remove them.
 So alongside looking at frequency weighting, previously derived, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on entropy was also explored.
@@
 
-\paragraph{Determining Frequency Cutoff}
-
 The same 571-word Cornell SMART stoplist used by \textcite{Sahlgren2008} was also used for this analysis,
@@
 and the same method of removing the high-frequency words from \citeauthor{Sahlgren2008} was used.
+
+\paragraph{Determining Frequency Cutoff}
 
@@
 Note that using a predefined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
-The cutoff that produced the frequency plot in Figure \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
+The cutoff that produced the frequency plots in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
 This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset was not given.
@@
 This cutoff value is a free parameter in frequency filtering that is not required when weighting stop words by entropy.
-Therefore, favoring parsimony, for the random permutation model, attenuating stop words based on entropy produces more accurate results when compared to filtering the stop words based on frequency of use.
+Therefore, favoring parsimony, for the random permutation model, attenuating stop words based on entropy is a slightly superior method compared to removing stop words based on frequency of use.
 This result is interesting since filtering stop words based on frequency is the more common default method for random permutation models \parencite{Sahlgren2008}.
@@
 
-Another interesting result was how the Bayesian model actually performed worse when stop words were removed based on frequency.
+Another interesting result was how the Bayesian model actually performed worse when stop words were removed based on frequency compared to no stop word removal.
 This was not the case for the random permutation model in this study, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random permutation model,
@@
 Another parameter in the random permutation model is the number of rows used to represent the all of the different words.
-Each word is represented by two randomly placed ones and two negative ones across the rows.
+Each word's environment vector is represented by two randomly placed ones and two negative ones across the rows.
 All that is required is that the four positions and signs of each of these environment vectors is unique.
@@
 
-In the figure, the full random permutation model that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
+In the figure, the model name that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
 This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random permutation model for all other results.
@@
 This is not a log-odds value.
-For the previous results, prior and context for the random permutation model are combined by simply adding the terms.
+For the previous results, prior and context for the random permutation model were combined by simply adding the terms.
 However, this means that ACT-R's prior base-level learning component (which is a log-odds value) is added to a correlation value, which may not be mathematically appropriate.
@@
 So although incorporating word order into the random permutation model does seem to slightly improve performance,
-the strength of the model is more likely to be in how efficiently it can represent the co-occurrence matrix, than its ability to include word order in the representation.
+the strength of the model is more likely to be in how efficiently it can represent the co-occurrence matrix rather than its ability to include word order in the representation.
 
@@
 The same sampling method used to test the impact of the prior component compared to the context component for the popular-users datasets was used,
-but the oldest posts that were not included in the testing set were used to build the co-occurrence matrices.
+and the co-occurrence matrices were built from the oldest posts that were not included in the testing set.
 So for each user, the co-occurrence matrix was created from their oldest posts, up to (and not including) their 50 newer posts that were used to test model accuracy.
@@
 
-The declarative memory model for ACT-R has an attentional weight term $W_{j}$ that can be used to implement how stop words are handled.
+The Bayesian declarative memory model for ACT-R has an attentional weight term $W_{j}$ that can be used to implement how stop words are handled.
 The random permutation model was also easily modified to have an attentional weight term by attenuating each word's set of ones and negative ones values by that word's attentional weight.

### 2014-09-17 writeups/dissertation/dissertation.tex
foo

@@
   The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
-  The task of choosing a hashtag for a tweet and tagging a post on StackOverflow was framed as a declarative memory retrieval problem.
+  The tasks of choosing a hashtag for a tweet and tagging a post on StackOverflow were framed as declarative memory retrieval problems.
+  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R based Bayesian model and a random permutation vector-based model.
   Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
-  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random permutation vector-based model.
   The results show that past user behavior of tag use is a strong predictor of future behavior, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
-  Also, the common natural language processing method to attenuate low-predictor words in the dataset was linked to ACT-R's attentional weight term,
-  and an entropy-weighting technique was developed that improved model accuracy by more efficiently allocating limited attentional resources to the predictive words in the posts and tweets.
-  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparable to the Bayesian model without including word order,
-  which shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
+  Also, ACT-R's attentional weight term was linked to a common entropy-weighting natural language processing method used to attenuate low-predictor words.
+  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparable to the Bayesian model without including word order.
+  This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
+  Finally, model accuracy was moderate to high for the tasks, which supports the theory that choosing tags on StackOverflow and Twitter is primarily a declarative memory retrieval process. 
   The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
@@
 
-\section{Motivation}
-
-The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human declarative memory.
+\section{Introduction}
+
+The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human cognitive processing.
 The declarative memory component in ACT-R is a Bayesian statistical model.
@@
 
-But this approach is problematic for two main reasons:
+However, this approach is problematic for two main reasons:
 First, it means that the declarative memory system for every developed model must be hand crafted and customized for that model, which takes enormous time and effort.
@@
 
-The work for this proposal will expand on previous work that has tested retrieval models on large-scale datasets.
-It will focus specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutations) can be improved, by testing these models on two real-world tagging tasks.
+The work in this study expands on previous work that has tested retrieval models on large-scale datasets.
+It focuses specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and vector-based) can be improved by testing these models on two real-world tagging tasks.
 One primary objective is to compare the random permutation vector-based model to ACT-R's declarative memory theory when large datasets are used for each model. 
@@
 For example, it is currently unexplored how a vector-based model can be modified to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
-Large-scale datasets will be used throughout the model development and evaluation process.
+Large-scale datasets were used throughout the model development and evaluation process.
 This provides an ideal testing environment for constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.
 
-\subsection{Motivation for Task}
+\subsection{The Task: Predicting User-Generated Tags}
 
@@
 One promising approach is to model and predict the user's goals when using and searching these social media sites \parencite{Rose2004}.
-If we have a better understanding of a user's goals, then we can tailor the information streams provided to the ones that are most in line with these goals.
+If we have a better understanding of a user's goals, then we can tailor the information streams to the content that is most in line with these goals.
 For example, Twitter might suggest other users and hashtags for the user to follow, Facebook could suggest pages to like, Yelp could recommend businesses to check out.
@@
 So if we can predict what hashtags a user is interested in, we have a better understanding of the information streams that they care about.
-With that understanding, we can tailor the system to suggest hashtags that they may not yet know about but are of likely interest to them.
+We can use that information to correctly suggest hashtags that they have previously used when they are generating new content.
+
+Further, if a particular information stream has multiple hashtags associated with it (common for Twitter), these semantically similar hashtags can also be recommended to users.
+So we can tailor the system to suggest hashtags that the user may not yet know about but are of likely interest to them and growing in popularity across the site.
 We can help aid in their discovery of fresh and relevant information that matches their goals on the site.
@@
 This process can be framed as a memory retrieval problem.
-That is, each user on a social media site has created content that contain a set of tags within the context of each post.
 The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.
+That is, each user on a social media site has created content that contains a set of tags within the context of each post.
 
@@
 Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
-At least two of the current memory retrieval models are based on a co-occurrence methodology, and I will be comparing these two models for hashtag prediction:
+At least two of the current memory retrieval models are based on a co-occurrence methodology:
 ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.
@@
 Word order information on sites like Twitter may contain highly predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
-However, it is unclear and unexplored how word order can be represented in the ACT-R declarative memory theory.
+Word order has been incorporated into ACT-R sentence-parsing models previously, but this has not been done using declarative memory.
+Rather, production rules were defined that ensured that the model followed a set of rules (e.g., left-to-right reading) when parsing sentences \parencite{Lewis2006}.
+It is certainly true that at least some production rules are needed when choosing words and creating content. 
+However, it is unexplored if some of the process of sentence construction and word selection can be represented in the declarative memory system.
 
 ACT-R's declarative memory system places strong constraints on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2007}.
-This decay rate equation formalizes how recency and frequency of hashtag use relate to the prior probability that a hashtag will be retrieved.
+This decay rate equation formalizes how recency and frequency of memory use relate to the prior probability that a memory will be retrieved.
 However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.
@@
 
-\section{Prior Research}
+\section{Theoretical Background}
 
@@
 
-ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the brain (e.g., memory, learning, visual and motor) interacts to produce behavior.
-The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up long-term declarative memory retrieval.
-The equations that make up this system can be derived through a rational analysis of long-term memory retrieval.
+ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
+The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
+The equations that make up this system were derived from a rational analysis of long-term memory retrieval.
 That is, given the task of retrieving a chunk of information from long-term declarative memory,
@@
       Constant Base Level &	$B_{i} = log \frac{p(i)}{p(\overline{i})}$ \\
-      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)}$ \\
+      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
       Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
@@
 However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
-In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its current frequency and recency of use.
+In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its frequency and recency of use.
 
@@
 So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
-This assumption has been both mentioned and used when deriving the $S_{ji}$ equation \parencite{Anderson1989}, as well as in recent research working with large-scale datasets \parencite{Stanley2013}.
+This assumption was described when deriving the $S_{ji}$ equation \parencite{Anderson1989}, and used in recent research working with large-scale datasets \parencites{Stanley2013,Farahat2004,Douglass2010}.
 
@@
 It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
-The PMI index is included in Equation \eqref{eqPMI}.
+The PMI index is presented as Equation \eqref{eqPMI}.
 
@@
 
-Latent Semantic Analysis \parencite{Landauer1997} (LSA) is a technique that has mathematical roots in factor analysis and also measures strength of association between two terms.
+Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
+Both ACT-R's Bayesian declarative memory model and LSA compute activations based on a co-occurrence matrix.
+However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
 The process starts with a set of N documents.
@@
 \textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
-Model results were on par with a large sample of applicants to american universities from non-English speaking countries (\num{64.4}\% compared to \num{64.5}\%).
+Model results were on par with a large sample of applicants to American universities from non-English speaking countries (\num{64.4}\% compared to \num{64.5}\%).
 Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
@@
 If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
-However, with large human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
+However, with large human-created datasets such as Twitter, StackOverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
 \textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
@@
 
-``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., chunk) as a vector.
+``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., a chunk) as a vector.
 The representation of the concept is distributed across all elements of the vector.
@@
 \textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
-However, decoding does require an additional decoding step that resonance does not require, and depending on the decoding operation, this step may be computationally expensive.
 
@@
 BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (\num{55.6}\% compared to \num{55.3}\%).
-Also, BEAGLE with unordered context and word order was more correlated to the WordNet synonym database than LSA (\num{-.311} compared to \num{-.165}).
 Further, it was shown that incorporating word order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
@@
 A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
+With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few one and negative one values in random locations. 
+In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is shifted to the left one position.
+This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
 
@@
 
-With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few one and negative one values in random locations. 
-In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is shifted to the left one position.
-This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
-
 For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow''.
@@
 
+The fan effect describes a behavior of human memory where memory items are more difficult to retrieve if they have occurred in a larger number of different contexts.
+\textcite{Rutledge2008} used \textcite{Anderson1974} classic experimental paradigm to test the fan effect.
+In the task, participants are exposed to a set of person and place pairs (e.g., ``the hippie is in the park'').
+They are later asked to recall if a particular person and place pair was presented to them before.
+The main manipulation is the ``fan'' of the person and place for each pair.
+The fan is defined as the number of different contexts that each item in the pair is presented in.
+For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store'', then ``hippie'' would have a fan of two.
+Recall accuracy decrease and time to recall increase as total fan of the person and pair increases.
+
+Both the ACT-R declarative memory theory and vector-based models produce the fan effect due to the effect that fan has on the co-occurrence matrix.
+As a particular word (e.g., ``hippie'') is seen in more and more contexts, the co-occurrence count of the word across all contexts is spread across multiple contexts.
+This means that when a retrieval for a context given the word occurs, multiple contexts have relatively high activation.
+This makes it more difficult to test if a signal (e.g., if a word appears in a specific context) is higher than the background noise (e.g., all other contexts that the word has appeared in),
+which makes the retrieval more uncertain.
+
+Both vector-based and Bayesian models produce the fan effect since they both are based on a co-occurrence matrix.
+However, vector-based models do further processing on that matrix, reduce its dimensionality, and use a correlation instead of a log odds value to compute the activation.
+Nonetheless, vector-based models still produce the fan effect because at their core they are still computing activations based on co-occurrences.
+
 \subsubsection{Base-Level Activation}
 
-Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the benefits of one over the other.
-For example, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
-More recent tweets on a news site like Twitter should be more relevant, for example.
+Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the relative benefits of the two models. 
+In particular, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
+
+The frequency and recency of how content has been used in the past can be a strong predictor of the future importance of that information. 
+The recent content in tweets on a news site like Twitter should be more relevant, for example.
 \textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
@@
 ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
-This type of information on a user's prior behavior is not being utilized with vector-based models currently.
+Vector-based models have focused on computing the associative strength of a chunk and specific contextual elements.
+It is unexplored how accuracy changes for these models once the information about a user's prior behavior (e.g., ACT-R's base-level activation component) is added.
 
@@
 With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
-It may be that multiple co-occurrence matrices must be used to represent ordered and unordered co-occurrences separately.
-However, it would be more parsimonious if these matrices could be combined into a single matrix representation, much like how this information is combined for vector-based models.
+Although there are certainly ways to incorporate word order into the ACT-R DM model.
+For example, an additional dimension could be added to the co-occurrence matrix representing word position.
+But adding word-order information to ACT-R's default Bayesian model is not common.
 
@@
 
+Using a correlation can be problematic due to range effects in observed co-occurrences.
+If there are particular words for a tag that co-occur much more often than all of the other words with that tag, that single large value diminishes the power of all other words to contribute to the correlation calculation.
+This does not happen with Bayesian retrieval models because each pair of word and tag values is normalized by the total number of words that appeared with that tag and total number of tags that appeared with that word
+(see the strength of association equation in Table \ref{tabACTRModel}).
+The high-frequency words are often removed from co-occurrence matrix for vector-based models to reduce this effect.
+
 \subsection{Recommendation Systems}
@@
 A model that predicts a user-chosen hashtag is a specific type of a recommendation system.
-These general recommendation systems have grown in need with the growth of the web.
+General recommendation systems have grown in need with the growth of the web.
 Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
@@
 It uses the term frequency---inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
+This measure is the product of two terms: the frequency count of the word in the document and the log inverse of the ratio of the number of documents that the word appeared in to the total number of documents.
+The term frequency component measures the strength of association between a word and a document.
+The inverse document frequency component measures how much information is contained within a specific word, independent of document.
+ACT-R also uses two similar components to compute the activation of a word with a specific document.
+The strength of association component of the ACT-R Bayesian co-occurrence model is a more sophisticated method for computing the term frequency component,
+and ACT-R's attentional weight component of the declarative memory model can be used to measure the information content within each word.
 
@@
 Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
+Recommendation models for StackOverflow, Google+, and Twitter will be described in more detail to follow.
 
@@
 His set of models were originally designed for next-word prediction for large text document collections.
+\citeauthor{Kuo2011} examined the accuracy of three different commonly-used models for next-word prediction: a Bayesian model, a K-nearest-neighbors model, and a Latent Semantic Analysis (LSA) model.
 In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user created a post, and restrict the space of possible words to only tags.
 He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
-Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive Singular Value Decomposition and K-nearest-neighbors models.
+Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive LSA (since it requires Singular Value Decomposition) and K-nearest-neighbors models.
 This Bayesian model can accurately predict a user-chosen tag at 47\% accuracy.
@@
 It turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
-The entropy measure was used to compute the attentional weight for two reasons:
-Stop words could be statistically derived and attenuated in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
-Using an entropy measure (as opposed to weighting each word equally in the post) for the attentional weight term allows the model to allocate limited attentional resources ($W$) to important contextual elements.
+
+The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation and it predicts well \parencite{Dumais1991}.
+It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
+So the scaled entropy ($E_{j}$) of a context word can be thought of as the amount of predictiveness that that word has to \emph{any} tag.
+This means that a stop word like ``the'' will have a scaled entropy measure close to 0, while a highly-predictive word like \emph{macro} will have a scaled entropy measure close to 1.
+Looking at $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate limited attentional resources to predictive contextual elements.
 
@@
 
-Several modifications to \textcite{Stanley2013}'s model were made from the standard Bayesian ACT-R declarative memory theory,
-so it was more sophisticated than the standard ACT-R model and the model used by \textcite{Kuo2011}.
-\citeauthor{Stanley2013} used a data-driven entropy weighting measure to attenuate low-predictor words and increase the weighting of high-predictor words in the body and title of a post.
-An offset term was also included which leveled the model's overall prediction for each post and enabled the use of a logistic regression statistical technique to calibrate the model's parameters.
-A separate model component was used for the words in the body and title of the post, and it was unclear if \citeauthor{Kuo2011} used the words in the body of the post for his models.
-
-\citeauthor{Stanley2013}'s model also scaled much better than \citeauthor{Kuo2011}'s co-occurrence model. 
+Since \textcite{Stanley2013} made several modifications to standard ACT-R model, it was more sophisticated than both the standard ACT-R model and the Bayesian model used by \textcite{Kuo2011}.
+Also, it was unclear if \citeauthor{Kuo2011} used both the words in the title and the body of the post as separate components for his models, or if only the words in the title were used.
+
+\citeauthor{Stanley2013}'s model also scaled much better than \citeauthor{Kuo2011}'s Bayesian co-occurrence model. 
 Because of this, 100 times more posts were used to build the co-occurrence matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
@@
 
-The exact tag-recommendation algorithm used for the Latex and SuperUser sites is not publicized.
-StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags, as the algorithm has not yet been deployed for the busiest StackOverflow site.
+The exact tag-recommendation algorithm used for the Latex and SuperUser sites is not public.
+After exploratory testing, StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags.
 The Latex and SuperUser sites are not nearly as popular as StackOverflow currently.
@@
 
-So it stands to reason that the algorithm is relatively complex, and there are possible scaling and efficiency concerns with the algorithm's implementation.
+So it stands to reason that the algorithm is relatively complex, and there could be possible scaling and efficiency concerns with the algorithm's implementation, so it has not yet been deployed to StackOverflow.
 Or it is also possible that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
@@
 
-Google+'s tag-recommendation algorithm is not publicized, but it appears to be rather sophisticated.
+Google+'s tag-recommendation algorithm is not public, but it appears to be rather sophisticated.
 Looking at the example post in Figure \ref{figGoogle+Hashtags}, \emph{\#BusterPosey} is suggested without having that specific word in the post.
@@
 \textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
-\textcites{Zangerle2011, Kywe2012} use the Apache Lucene infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarities.
+\textcites{Zangerle2011, Kywe2012} use the Apache Lucene infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score to compute similarities.
 Regardless of the method to compute the similarity between tweets,
@@
 In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
-Now it is certainly the case that the topics generated by a latent-based model may be similar to the hashtags used in the post.
+Now it is certainly the case that the topics generated by an LDA model may be similar to the hashtags used in the post.
 However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
-Why not take those topics already present in the data and build a model that produces them?
-In other words, the topics data are already in the tweet, so why not use them?
+On the other hand, a topics-based approach is more general than a hashtags-based approach, since topics can be suggested for posts that do not use hashtags (which is quite common across Twitter).
 
@@
 
-\textcite{Stanley2013} used an ACT-R inspired Bayesian memory model to predict human tagging on the StackOverflow site, but to our knowledge no vector-based memory model has been tested or used.
-Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}.
+\textcite{Stanley2013} used an ACT-R based Bayesian memory model to predict human tagging on the StackOverflow site, but to our knowledge no vector-based memory model has been tested on the site.
+Both types of models have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007},
+so it should be straightforward to test a vector-based model on the StackOverflow site.
 
@@
 
-As some preliminary research, I have developed and tested a user-customized model for StackOverflow by modifying the model in \textcite{Stanley2013}.
-To incorporate the user's prior tag history into the model, the base-level activation term was modified from the original equation in ACT-R.
+As some initial research, I have developed and tested a user-customized model for StackOverflow by modifying the model in \textcite{Stanley2013}.
+To incorporate the user's prior tag history into the model, the base-level activation term ($B_{i}$) was modified from the original equation in ACT-R.
 $B_{i}$ is based on log-odds occurrence, which in turn is based on frequency of tag use, or probability of tag use $p_{i}$.
@@
 where the probability of using tag $i$ ranges from the global probability ($pGlobal_{i}$) to the user's specific probability ($pUser_{i}$) as the user's total tag count increases.
-Overall model accuracy improved by 5\% after incorporating the user's prior tag history into the model.
-
-This preliminary research is encouraging, and suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
+
+The global probability is the same $p_{i}$ used in \textcite{Stanley2013}'s study.
+The user's specific probability is computed the same way that the global probability is computed (i.e., the log odds that the tag occurred relative to the occurrence of all other tags).
+However, each user's probability is constrained to the set of tags that the specific user has used in the past. 
+When a user has no past tag history, the global prior probability is used.
+Once a user has provided at least five tags, that user's specific prior probability is used.
+Between zero and five tags, both terms are used but the weight of the global prior relative to the user prior decreases as the number of tags increases.
+
+Both linear decay and exponential decay transition functions were tried.
+Using either transition function, overall model accuracy improved by 5\% after incorporating the user's prior tag history into the model.
+
+This preliminary research suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
 It also seems plausible that model accuracy can improve an even greater amount by using the time-dependent form of the base-level component in Table \ref{tabACTRModel},
@@
 Twitter has around \num{2383000} unique words for the \num{3000000} tweets in each of the popular-users datasets.
-Stackoverflow has \num{5848000} unique words for the \num{1000000} posts used to generate the co-occurrence matrix.
+StackOverflow has \num{5848000} unique words for the \num{1000000} posts used to generate the co-occurrence matrix.
 So it is impressive that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
@@
 
+\subsubsection{Software Repository}
+
+All software code created for this research is publicly available on GitHub \parencite{StanleyRepo2014}.
+
 \begingroup

### 2014-09-21 writeups/dissertation/dissertation.tex
foo

@@
 \usepackage[utf8]{inputenc}
-
-% Default apa6 formatting for a CI
-\newcommand{\CI}[2]{95\% CI [#1, #2]}
 
@@
 Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
-A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#''.
+A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#.''
 After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
@@
 Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
-This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.
+This is one of the available corpora on the LSA website and is intended to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.
 
@@
 Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
-This allows for retrieval requests such as ``playing \#$\Phi$'', ``just read \#$\Phi$'', and ``vector-based $\Phi$ systems''.
+This allows for retrieval requests such as ``playing \#$\Phi$,'' ``just read \#$\Phi$,'' and ``vector-based $\Phi$ systems.''
 
@@
 Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which correlates highest with the word-order memory vector for luther ($o_{luther}$).
-Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$''.
+Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$.''
 
@@
 
-For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow''.
+For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow.''
 The ``stack'' environment vector ($e_{stack}$) is shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).
@@
 The fan is defined as the number of different contexts that each item in the pair is presented in.
-For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store'', then ``hippie'' would have a fan of two.
+For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store,'' then ``hippie'' would have a fan of two.
 Recall accuracy decrease and time to recall increase as total fan of the person and pair increases.
@@
 
-This preliminary research suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
+This research suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
 It also seems plausible that model accuracy can improve an even greater amount by using the time-dependent form of the base-level component in Table \ref{tabACTRModel},
@@
 The purpose of this dataset is to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
-The dataset contains approximately three million tweets where at least one of the 400 most popular hashtags are used, and was collected in real time until the tweet quota was reached.
+The dataset contains approximately three million tweets where at least one of the 400 most popular hashtags are used.
+A quota of three million tweets was selected because previously research testing Bayesian co-occurrence models on large-scale datasets has used on the order of one million documents
+\parencites{Stanley2013,Douglass2010,Budiu2007,Kuo2011}.
+Tweets were collected in real time until the tweet quota was reached.
 
@@
 
-``Stop words'' (e.g., ``the'', ``a'', ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
+``Stop words'' (e.g., ``the,'' ``a,'' ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
 It is typically recommended \parencite{Bird2009} to remove stop words when building the co-occurrence matrix.
@@
 
-For both equations, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``a'' appears), and $n$ is the number of times each chunk has been observed.
+For both equations, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``loop'' appears), and $n$ is the number of times each chunk has been observed.
 Since the optimized learning form assumes a constant presentation rate for each chunk,
@@
 \citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log-odds computation of the probability that each chunk is likely to be presented again, given past frequency.
-If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log n_{i}$,
+If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log \left ( n_{i} \right )$,
 which is a nonlinear transformation of the $B_{i}$ used by \citeauthor{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
@@
     \subfigure[][Twitter]{\includegraphics[width=8cm]{HTByTime-fashionista_com-crop.pdf}}
-    \caption{Tagging profile for a single user}
+    \caption{
+      Tagging profile for a single user.
+      The x-axis represents the specific hashtag a user chooses for a post and time progresses vertically.
+      A point in the plot represents when a user chooses a specific hashtag for a specific post.
+      A user's chosen hashtag for their oldest post that contains a hashtag is the lowest point on the plot.
+    }
     \label{figPriorHTByTime}
@@
 There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
+For example, the StackOverflow user in Figure \ref{figPriorHTByTime} has a pronounced band where the same hashtag is chosen repeatedly over time. 
+However there are many instances of other tags being used in parallel to the tag that is producing the band (that is repeatedly being used). 
+The Twitter user has a few clusters where the same hashtag is chosen repeatedly but only for a brief period of time.
+As time progresses, a band of choosing another hashtag is observed, and then the user switches to a new hashtag after that.
+More recent posts for the Twitter user are more varied, and there is no longer a single hashtag that the user chooses repeatedly.
 This helps illustrate how challenging modeling this task can be.
-The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is dynamically changing over time.
+The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is changing over time.
 
@@
     \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
-    \caption{Model performance for a single dataset slice for StackOverflow}
+    \caption{
+    Model performance for a single dataset slice for StackOverflow.
+  }
     \label{figPriorSOQSliceDsStd}
@@
 Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
-Error bars represent the 95\% bootstrapped confidence interval of the median.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
-  \caption{Overall best-fit decay rate for StackOverflow and Twitter}
+  \caption{
+  Overall best-fit decay rate for StackOverflow and Twitter.
+  Results are shown for the two Twitter and StackOverflow (SO) popular-users datasets.
+  Error bars represent the 95\% bootstrapped confidence interval for the median.
+}
   \label{figPriorDecay}
@@
 Aggregate accuracy was computed by taking the mean of every user's model accuracy across all users in all dataset slices.
-All confidence intervals are 95\% bootstrapped for the mean.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc--crop.pdf}}}
-  \caption{Overall model accuracy for StackOverflow and Twitter}
+  \caption{
+    Overall model accuracy for StackOverflow and Twitter at optimal decay-rate values.
+    Best-fit decay-rate values for the standard and optimized learning model are \num{.8} and \num{.43}.
+    All error bars are the 95\% bootstrapped confidence interval for the mean. 
+  }
   \label{figPriorAcc}
@@
 
+The decay rate parameter in $B_{i}$ determines how fast a presentation of a chunk in the past decays over time.
+Higher values of $d$ mean that activation decays more quickly, which means that the chunks used more recently are the most active and the total frequency count of a chunk matters less.
+Looking back at Figure \ref{figPriorDecay}, the optimal decay rate parameter when the standard form of $B_{i}$ is used is higher for Twitter than StackOverflow.
+It makes sense that the lifetime of a specific hashtag on Twitter is shorter overall than the lifetime of a tag (e.g., \emph{C#}) on StackOverflow.
+This means that the decay rate for Twitter should be higher than StackOverflow, which is what is observed.
+
 \section{Combining Predictors}
 
-The declarative memory models analyzed in this study consist of two main components: past behavior and context.
+The declarative memory models analyzed for this research consist of two main components: past behavior and context.
 Past tag use was isolated in the first study to see how well ACT-R's prior component of declarative memory (i.e., base-level learning) fits each user's tag use over time.
@@
 In order to best combine these predictors, a logistic regression statistical technique similar to \textcite{Stanley2013} was used to determine the most optimal weights for each term.
-Logistic regression is used on these datasets in the following way:
+Each term in the full model (see the total activation equation Table \ref{tabModACTRModel}) is weighted so that model accuracy is maximized.  
+\textcite{Stanley2013} found that the optimal weights for the StackOverflow model for the prior term, title context term, and body context term were \num{.62}, \num{.93}, and \num{1.75} respectively.
+
+The optimal weights for each model component are computed in the following way:
 When a user creates a tag for a post, a retrieval request for the model is made.
@@
 
-A simplified version of \citeauthor{Stanley2013}'s method was used for the results in this study.
+A simplified version of \citeauthor{Stanley2013}'s method was used for the results for this research.
 Instead of subtracting off the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
@@
 
-\subsubsection{Computing Prior}
+\subsubsection{Computing Priors}
 
@@
 and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high performance desktop, even with highly optimized representations.
-Already \num{835500000} total co-occurring word tag observations are used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
-It is also the case that \num{1000000} posts for StackOverflow is already around $1/7$ of the total posts created on the site,
+Already \num{835500000} total co-occurring word tag observations were used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
+It was also the case that \num{1000000} posts for StackOverflow was already around $1/7$ of the total posts created on the site,
 so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrence matrix instead,
-especially since \num{835500000} co-occurrences are already used with \num{1000000} posts.
+especially since \num{835500000} co-occurrences were already used with \num{1000000} posts.
 
@@
 
-Model accuracy relaxed across posts with and without the offset model component was evaluated for both the Bayesian and random permutation model.
+The offset component was added to both models, and model accuracy was examined before and after the addition.
 
@@
 The impact of adding the offset to each of the models for StackOverflow and Twitter is shown in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.
-The error bars represent the 95\% bootstrapped confidence interval of the mean model accuracy across all runs in dataset.
 
@@
   \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
-  \caption{Impact of adding offset component for StackOverflow}
+  \caption{
+    Impact of adding offset component for StackOverflow.
+    Error bars represent the 95\% bootstrapped confidence interval for the mean model accuracy across all runs in dataset.
+    The random permutation models are represented as ``RP'' and Bayesian models as ``Bayes.''
+}
   \label{figContextOffsetSO}
@@
   \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
-  \caption{Impact of adding offset component for Twitter}
+  \caption{
+    Impact of adding offset component for Twitter.
+    All error bars are the 95\% bootstrapped confidence interval for the mean.
+  }
   \label{figContextOffsetT}
@@
 When they are enumerated, then at least one of the components is left out.
-The random permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
+The random permutation and Bayesian models are plotted alongside each other so that comparisons across the two model types can be done easily.
 So, for example, the ``standard prior model relaxed across posts without offset'' means that only the prior model component was used and the offset for that component was not subtracted off when determining activation.
@@
 Accuracy is improved even when only a single term is used.
-Although in this case the weight of the term is irrelevant (only one term, weighting for that term does not matter),
-using an offset still improves accuracy since the offset term ensures that only a small number of tags per posts are predicted by the model to be the chosen tags.
+Although in this case the term is not weighted,
+but using an offset still improves accuracy since the offset term ensures that only a small number of tags per posts are predicted by the model to be the chosen tags.
 
@@
 then the model would be able to say when, for example, the third top tag in a post is more likely to be a chosen tag than the second top tag in another post.
-That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of retrievals to use for a specific post.
-If this were the case, then adding the offset term would actually decrease accuracy, since this absolute value of the activation values would be subtracted off.
-However, since accuracy improves when adding the offset term for both models, it does not seem that activation values have much meaning across posts (and only within a single post) for these datasets.
-
-A likely and simple explanation for this is that possibly there are posts where a large set of tags have high activation,
-but the author is throttling their retrieval threshold for each post such that only a small set of tags are actually used.
-This is probably not the retrieval threshold used in the core declarative memory system that determines if a retrieval for the term is successful or not.
-It is unlikely that this retrieval threshold can be directly controlled and manipulated.
-Rather, it could be a high-level task retrieval threshold that the author has control over, where he/she actively sets it based on competing goals of tagging posts an not overly saturating a post with tags.
-Adding the offset term into the model is a way to implement a simple version of this high-level retrieval threshold.
+That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of tags to use for a specific post.
+If this were the case, then adding the offset term would actually decrease accuracy, since this absolute value of the activation values would be removed.
+However, since accuracy improves when adding the offset term for both models, it does not seem that absolute activation values have much meaning across posts for these datasets,
+and that the relative activation values for each tag within a post is what is important.
+
+One simple and likely explanation is that there are posts where a large set of tags have high activation,
+but the author is choosing to use only a fixed small set of the top tags for each post, regardless of the activation values of the top tags.
+The author determines the number of tags to use for a post based on competing goals of tagging posts an not overly saturating a post with tags.
+For example, the author may be writing a question on StackOverflow about \emph{javascript}, and a large set of related tags are retrieved from declarative memory that are all above the retrieval threshold
+(e.g., \emph{javascript}, \emph{dom}, \emph{svg}, \emph{html}, \emph{jquery}, \emph{css}),
+but the author only chooses to use the tags \emph{javascript}, \emph{dom}, and \emph{svg}, since he has decided to tag this post with three tags, and those are the three tags with the highest activation.
+Adding the offset term into the model is a way to implement a simple version of this strategy where the author is actively choosing to use only a small set of tags for each post.
 
@@
 Since the interactions across popular-users datasets are minimal, model behavior for the popular-users subset for Twitter can be interpreted as behavior that will persist across different time periods,
-if the same method of choosing popular hashtags is done at a later date.
-It may be more general that model behavior also does not depend on the specific method of choosing the popular hashtags, but since only a single method was used for this analysis,
+if the same method of choosing popular hashtags is conducted at a later date.
+It may also be the case that that model behavior does not depend on the specific method of choosing the popular hashtags, but since only a single method was used for this analysis,
 the results from this study cannot by themselves be used to answer this question.
@@
 Both the random permutation and Bayesian models were tested.
-Activations for context were computed by using the two methods for stop word removal, the entropy weighting method, and no method (i.e., no stop word processing).
-
-The same 571-word Cornell SMART stoplist used by \textcite{Sahlgren2008} was also used for this analysis,
-The same entropy weighting method using by \textcite{Stanley2013} was examined,
-and the same method of removing the high-frequency words from \citeauthor{Sahlgren2008} was used.
+Activations for context were computed by using five different methods to handle stop words:
+The two removal methods used in \textcite{Sahlgren2008} (i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist),
+the entropy-weighting method used in \textcite{Stanley2013},
+combining the entropy-weighting and frequency-filtering methods,
+and no removal or weighting of stop words.
 
@@
 
-However, a cutoff point must be identified and used in order to remove the high-frequency words.
+A cutoff point must be identified and used in order to remove the high-frequency words.
 \textcite{Sahlgren2008} used a cutoff where words that occurred more than \num{15000} times were removed.
@@
 The original plot shows the size of the spikes when high-frequency words are not removed.
-The entropy plot shows how the size decreases by more than an order of magnitude when the entropy weighting measure is used.
+The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used (compare the difference between the original subplot and the entropy subplot).
 The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
-Note that using a predefined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
 The cutoff that produced the frequency plots in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
@@
 
+Note that using a predefined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
 Also, no strong claims are being made that the cutoff identified is the ideal frequency percentage cutoff for all datasets when a frequency cutoff is used.
-The method used to identify this cutoff was simply aimed to find a reasonable value where comparisons between a frequency cutoff and entropy weighting techniques would be equitable.
+The method used to identify this cutoff was simply intended to find a reasonable value where comparisons between a frequency cutoff and entropy weighting techniques would be equitable.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsSOPerm-crop.pdf}}}
-  \caption{Stop-word techniques for the random permutation model for StackOverflow}
+  \caption{
+    Stop-word techniques for the random permutation model for StackOverflow.
+    The five methods for handling stop words are shown both for the title model component in isolation and the full model.
+    ``Stoplist'', ``freq'', ``entropy'' are the stoplist removal method, removal based on frequency count, and weighting based entropy metric.
+}
   \label{figContextStopRPSO}
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTPerm-crop.pdf}}}
-  \caption{Stop-word techniques for the random permutation model for Twitter}
+  \caption{
+  Stop-word techniques for the random permutation model for Twitter.
+  The five methods for handling stop words are shown both for the context model component (i.e., no prior term) and the full model.
+}
   \label{figContextStopRPT}
@@
 For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\num{.25} to \num{.26}).
-Removing stop-words based on a predetermined list actually decreases accuracy (\num{.25} to \num{.23}), and using data-driven frequency filtering has only a small to no effect.
+Removing stop-words based on a predetermined list actually decreases accuracy (\num{.25} to \num{.23}), and using data-driven frequency filtering has only a small to no effect (\num{.25} to \num{.25}).
 
@@
 
-Since accuracy measurements when using a predetermined list to remove stop words for the random permutation model were much lower than the frequency and entropy methods, this method was not examined for the Bayesian model.
+Since accuracy measurements when using a predetermined list to remove stop words for the random permutation model were much worse than the frequency and entropy methods, this method was not examined for the Bayesian model.
 The results for the frequency and entropy techniques for the Bayesian model on the StackOverflow dataset are included in Figure \ref{figContextStopBSO}.
@@
 So it is likely the case that a form of frequency attenuation is already built into the Bayesian model.
-And therefore adding a frequency filter on top of this does nothing to improve accuracy, and only removes predictive information.
+Therefore, adding a frequency filter on top of this does nothing to improve accuracy and only removes predictive information.
 
@@
 since model accuracy seems to be quite sensitive to the specific method used. 
-However, one conclusion that can be made from this research is that the entropy-weighting method is a contender, and should be explored much more often than it has been previously.
+However, one conclusion that can be made from this research is that the entropy-weighting method is a contender, and should be considered much more often than it has been previously.
 
@@
 It is likely that the random permutation model is starting to saturate when entropy weighting is used for the largest document size for StackOverflow.
-Even if the non-predictive words based on entropy are attenuated,
-when the high-frequency words are included in the random permutation model,
-this attenuation is not enough to overcome the range effects that grow with document size.
-These effects end up saturating all of the correlations near one, and consequently decrease discrimination and accuracy.
+Saturation happens when certain words co-occur with a tag much more often than all of the other words.
+This introduces range effects when the correlation is computed, which ends up driving all of the correlations near one, and consequently decreases discrimination and accuracy.
+Saturation happens when entropy weighting is used because the non-predictive words are only attenuated and not removed, and this attenuation is not enough to overcome the range effects that grow with document size.
+
 However, the effect of saturation is small compared to the large effect from using either frequency filtering or entropy weighting to handle the stop words.

### 2014-09-22 writeups/dissertation/dissertation.tex
foo

@@
 This will undoubtedly require a combination of two primary model components:
-[1] A user's prior likelihood of choosing a particular hashtag, given their previous hashtag use, and [2] the likelihood that a particular hashtag is related to the context of the post being created.
+[1] a user's prior likelihood of choosing a particular hashtag, given their previous hashtag use, and [2] the likelihood that a particular hashtag is related to the context of the post being created.
 
@@
 Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
-The likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
+the likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
 For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
@@
 
-Two models were evaluated for hashtag prediction on both the Twitter and StackOverflow datasets:
-A vector-based model and Bayesian retrieval model.
+Two models were evaluated for hashtag prediction on both the Twitter and StackOverflow datasets: a vector-based model and Bayesian retrieval model.
 The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} was used for the vector-based model, since it has shown to perform better than the BEAGLE model \parencite{Recchia2010}.
@@
 
-Twitter provides two types of APIs for acquiring data: A query-based and streaming interface.
+Twitter provides two types of APIs for acquiring data: a query-based and streaming interface.
 The Tweepy Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
@@
 Higher values mean that chunk activation decays faster, which means that the most recently presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
-Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks that have occurred most frequently will have the highest activation,
+Lower decay rate values flip the importance of frequency and recency:
+Lower values mean that chunks that have occurred most frequently will have the highest activation,
 and activation depends less on the specific chunks that have been observed in the recent past.
@@
 Looking back at Figure \ref{figPriorDecay}, the optimal decay rate parameter when the standard form of $B_{i}$ is used is higher for Twitter than StackOverflow.
-It makes sense that the lifetime of a specific hashtag on Twitter is shorter overall than the lifetime of a tag (e.g., \emph{C#}) on StackOverflow.
+It makes sense that the lifetime of a specific hashtag on Twitter is shorter overall than the lifetime of a tag (e.g., \emph{C\#}) on StackOverflow.
 This means that the decay rate for Twitter should be higher than StackOverflow, which is what is observed.
@@
 
+However, document size is not the only factor that determines whether the Bayesian model or random permutation model will perform better.
+For example, the document size where the Bayesian model begins to outperform the random permutation model is quite different for StackOverflow and Twitter.
+This is because the number of observations used to generate the co-occurrence matrix is a function of four components:
+document size, number of words in each document, total number of unique tags, and number of tags used per document.
+The average number of tags used per post for StackOverflow and Twitter were the same order of magnitude (around 3 for StackOverflow, slightly less for Twitter).
+
+However, the average number of words in a StackOverflow post is much greater than the 140 character limit used for tweets.
+More words per post means that the number of observations per cell in the co-occurrence matrix grows more quickly with document size.
+The Bayesian model is more likely to perform better with datasets like StackOverflow that have these larger document sizes.
+
+Also, the popular-hashtags Twitter dataset only contains 400 unique tags while the randomly-sampled StackOverflow dataset has \num{34377} unique tags.
+More unique tags means that the co-occurrence counts are spread across a wider range of tag columns,
+which means that it takes a larger number of total documents before the Bayesian model performs better than the random permutation model.
+
+The transition point where accuracy for the Bayesian model outperformed the random permutation model was much higher for StackOverflow than Twitter.
+This is most likely because StackOverflow has a almost \num{100} times more unique tags than the Twitter popular-hashtags dataset.
+The number of words in a StackOverflow post is also much higher than Twitter, which should have lowered the transition point.
+However, the effect of words in a post must be less than the effect of number of unique tags, as the transition point was still much higher for StackOverflow given that it should have lowered due to post size.
+
+The compressed random permutation model should perform better with smaller document sizes, less words per document, more unique tags, and fewer number of tags per post.
+However, since there are so many factors that determine whether the Bayesian or random permutation model will perform better,
+and only two different datasets were examined in this research,
+it is difficult to make any specific recommendations as far as exact transition cut points (e.g., document size, words per document) where either the Bayesian or random permutation model should be used.
+Also, it is likely that the specific domain being studied interacts with the factors already mentioned to determine which model is better for that domain.
+Nevertheless, if a researcher is working in a domain with limited data (e.g., small document size, large tag space) that produces a sparse co-occurrence matrix,
+then it may be worthwhile to experiment with using a compressed random permutation model instead of the full Bayesian model.
+
 \subsection{Amount of Compression for Random Permutation}
@@
 So it is impressive that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
-This is likely because word use frequency drops off sharply, and there are many words in both datasets which are used only once or a few times.
+This is likely because word use frequency approximately follows Zipf's law \parencite{Stanley2013}.
+This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply when moving to words that are used fewer times. 
 Since these words have little to no ability to influence predictions, it makes sense that they can be safely removed from the dataset.
@@
 So a method was created to convert a distribution of correlations to a distribution of log-odds values.
-To do this, the one-tailed cumulative distribution function value for each correlation in the distribution is computed.
-That probability is then converted to a log-odds value in the usual way.
-
-This technique has a few nice mathematical properties:
+To do this, the one-tailed cumulative distribution function value for each correlation in the distribution was computed.
+That probability was then converted to a log-odds value in the usual way.
+
+This technique has a few advantages: 
 It is simple and computationally cheap.
@@
 
-Although there may be a slight edge in increased accuracy when a log-odds transformation is used for the StackOverflow dataset.
-But that result is not statistically reliable for the sample size used in this analysis, and even if present the effect size is small.
+There may be a slight edge in increased accuracy when a log-odds transformation is used for the StackOverflow dataset.
+However, that result is not statistically reliable for the sample size used in this analysis (confidence intervals overlap in Figure \ref{figContextLogoddsSO}),
+and even if present the overall change in accuracy is small (\num{.45} to \num{.47}).
 Nonetheless, it does seem appropriate to convert the context to a log-odds adjustment value when context is added as an additional term to the random permutation model.
@@
 Three different methods for incorporating word order in the random permutation model were tested:
-Taking just the sign of the word's position relative to the tag (direction method), using the position (standard method), and using the position but only including words that were used near the tag (window method).
+taking just the sign of the word's position relative to the tag (direction method), using the position (standard method), and using the position but only including words that were used near the tag (window method).
 For the standard method, all words and their respective position are used to build the representation.
@@
 
-Model accuracy results after incorporating word order into the random permutation model are included in Figure \ref{figContextOrder}
+Model accuracy results after incorporating word order into the random permutation model are included in Figure \ref{figContextOrder}.
 
@@
 At a row size of around \num{2000}, model accuracy improves by less than a percentage point when order information is incorporated.
-The results for larger row sizes are somewhat inconsistent, since the full model (order and orderless) sometimes underperforms the order-only model, which does not make sense.
+The results for larger row sizes are somewhat inconsistent, since the full model (order and orderless) sometimes underperforms the order-only model,
+which should never happen since the order-only model is a subset of the full model.
 \textcite{Sahlgren2008} most likely did not use a regression technique to optimize the weights for each of the order and orderless model terms.
@@
 
-Since it was shown that including word order information has little to no significant improvement on model accuracy, an attempt to incorporate word order information into the Bayesian model was not made.
+Since it was shown that including word order information has little to no significant improvement on model accuracy, no attempt to incorporate word order information into the Bayesian model was made.
 
@@
 A regression is done for each run in the dataset, and since multiple runs were used for the StackOverflow and Twitter datasets, multiple samples of coefficient weights for each model component were computed.
-The variance of the weights for each component show how likely it is that the regression is overfitting the model,
+The variance of the weights for each component show how likely it is that the regression is overfitting the data,
 and the mean shows the best estimate weight of each component relative to the other.
@@
   \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffSO-crop.pdf}}}
-  \caption{Coefficient values for full models for StackOverflow}
+  \caption{
+  Coefficient values for each component of the full models for StackOverflow.
+  Models used are the best-performing Bayesian and random permutation models
+  (i.e., Bayesian and RP have entropy weighting, RP converts context activation from correlation to log odds).
+}
   \label{figContextCoeffSO}
@@
   \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffT-crop.pdf}}}
-  \caption{Coefficient values for full models for Twitter}
+  \caption{Coefficient values for each component of the full models for Twitter}
   \label{figContextCoeffT}
@@
 
-This shows that incorporating past tagging history is important for even a site like StackOverflow where a large amount of context information has been collected.
+This shows that incorporating past tagging history is important even for a site like StackOverflow where a large amount of context information has been collected.
 However, for this type of site it only becomes important when a reasonably large amount of a user's past tagging history has been collected.
@@
 \begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiSO-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiSO-crop.pdf}}}
   \caption{Model accuracy with user-customized context component for StackOverflow}
@@
 \begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiT-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiT-crop.pdf}}}
   \caption{Model accuracy with user-customized context component for Twitter}
@@
 The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
-These results provide support that choosing a hashtag when composing a tweet and tagging a StackOverflow post is a long-term memory retrieval process.
+These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is a long-term memory retrieval process.
 However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
@@
 
-Or, perhaps authors are taking advantage of context information that is not currently represented in the model.
-For example, time of day might be a useful predictor of tag use.
-\textcite{Cohn2012} built a tag-prediction model that was based primarily on temporal cycles of tag use.
+Alternatively, perhaps authors are taking advantage of context information that is not currently represented in the model.
+For example, time of day might be a useful predictor of tag use on Twitter.
+\textcite{Cohn2012} built a tag-prediction model for Twitter that was based primarily on temporal cycles of hashtag use.
 This technique could be naturally incorporated into a co-occurrence framework by including time of day as an additional context by tag co-occurrence for each post.
@@
 
-\textcite{Anderson1991} also looked at tasks similar to tag use on StackOverflow and Twitter when deriving the prior component of the declarative memory theory.
+\textcite{Anderson1991} also looked at similar tagging tasks with large datasets when deriving the prior component of the declarative memory theory.
 \citeauthor{Anderson1991} examined, for example, the frequency of word choice over time used for the title of articles for the New York Times,
@@
 Model accuracy is high enough (particularly for StackOverflow) for the developed models to be used as a foundation for tag recommendation systems that are deployed on Twitter or StackOverflow.
-Also since the developed models are task agnostic, they can potentially be used as tag recommendation systems for other domains.
+Also, since the developed models are task agnostic, they can potentially be used as tag recommendation systems for other domains.
 Examples of similar tagging tasks are hashtag use on Facebook, flagging (i.e., tagging) Wikipedia articles for content issues or overly opinionated material,

### 2014-09-22 dissProject/preProcess.R
foo

@@ -468,7 +468,7 @@ visModelVsPredTbl <- function(modelVsPredTbl) {
 	       geom_line(aes(log(d), meanRelPC, group=user_screen_name[1])) +
 	       defaultGGPlotOpts + 
 	       xlab('log(d)') +
-	       ylab('Normalized Mean'))
+	       ylab('Normalized Mean Model Accuracy'))
 	assign('p3', ggplot(modelVsPredTbl[topHashtag & hashtagUsedP & user_screen_name %in% sample(unique(user_screen_name), size=min(20, length(unique(user_screen_name))))],
 			    aes(log(d),acc, group=as.factor(user_screen_name))) +
 	       geom_line() +
@@ -478,7 +478,7 @@ visModelVsPredTbl <- function(modelVsPredTbl) {
 	modelVsPredTbl[predUsedBest == T & d <= .4]
 	assign('p4', ggplot(modelVsPredTbl[predUsedBest == T & d <= 2], aes(d)) +
 	       geom_histogram(breaks=defaultBaseConfig$ds[defaultBaseConfig$ds <= min(2, max(modelVsPredTbl[,d+.1]))], aes(y = ..density..)) +
-	       geom_density() +
+	       #geom_density() +
 	       defaultGGPlotOpts + 
 	       xlab('Best Fit Decay Rate Parameter (d) by Subject') +
 	       ylab('Density'))
@@ -1520,7 +1520,7 @@ renameColGroupNum <- function(tbl) {
 
 renameColSizeNum <- function(tbl) {
 	mapTbl = data.table(sizeNum=as.character(c(1,6,5,4,3,2)),
-			    newName=c('Testing 1e5 SO, 3e6 Twitter', '1e3 Documents', '1e4 Documents', '1e5 Documents', '1e6 Documents', '3e6 Documents'))
+			    newName=c('Testing 1e5 SO, 3e6 Twitter', '1*10^3 Documents', '1*10^4 Documents', '1*10^5 Documents', '1*10^6 Documents', '3*10^6 Documents'))
 	tbl[, sizeNum := as.character(sizeNum)]
 	setkey(tbl, sizeNum)
 	tbl[mapTbl, sizeNum := newName]
@@ -1613,20 +1613,20 @@ makeRemapForMapping <- function(keyword, text) {
 }
 
 makeLogregMapping <- function() {
-	c('actPriorStd', 'Standard prior model relaxed across posts',
-	  'actPriorStdBayes', 'Bayes prior relaxed across posts',
[…]

### 2014-09-22 writeups/dissertation/dissertation.tex
foo

@@
 \AtEveryBibitem{\clearfield{doi}}
+
+% http://tex.stackexchange.com/questions/55780/disable-month-in-biblatex-bibliography
+\AtEveryBibitem{\clearfield{month}}
+\AtEveryCitekey{\clearfield{month}}
 
@@
 
+\section{Acknowledgements}
+
+I would first like to thank my advisor, Mike Byrne, for all of his feedback and suggestions for this dissertation research
+and also his excellent guidance throughout my time as a graduate student.
+I would also like to thank my parents, Mike and Patsy, for the positive influence that they have provided me for my entire life.
+Finally, I would like to thank my brothers Brendan and Justin, sister Kellen, extended family, and friends, for all of their support during my time in the Rice Psychology graduate program.
+
+This work was supported in part by the Gertrude Maurin Fund.
+The resources were used to purchase high-performance computing hardware and the research would not have been possible without this crucial equipment.
+
+I would also like to mention a few open-sourced software packages that were heavily leveraged to do this research:
+The majority of the modeling and analysis was done with the R statistical computing environment \parencite{RCoreTeam2014} and two important R packages:
+data.table \parencite{Dowle2014} was leveraged to implement the models efficiently, and ggplot2 \parencite{Wickham2009} was used to produce the visualizations.
+
+\newpage
+
 \tableofcontents
@@
 
-\subsubsection{Acknowledgements}
-
-The majority of the modeling and analysis was done with the R statistical computing environment \parencite{RCoreTeam2014} and two important R packages:
-data.table \parencite{Dowle2014} was leveraged to implement the models efficiently, and ggplot2 \parencite{Wickham2009} was used to produce the visualizations.
-
-\subsubsection{Software Repository}
-
-All software code created for this research is publicly available on GitHub \parencite{StanleyRepo2014}.
-
 \begingroup
@@
 
+\clearpage
+\section{Appendix 1: Support Code}
+
+All software code created for this research is publicly available on GitHub \parencite{StanleyRepo2014}.
+
 \end{document}

### 2014-09-22 writeups/dissertation/dissertation.tex
foo

@@
 
+\subsubsection{Directory Structure}
+
+The repository is separated into three main folders:
+dissProject, firstSOProject, and writeups.
+The firstSOProject folder contains the support code for the \textcite{Stanley2013} research.
+The dissProject folder contains support code for this dissertation research.
+The writeups folder contains all writeup projects created during the dissertation process (e.g., the LaTeX for this dissertation, Keynote presentations).
+
+Within the dissProject folder, there are two main files that contain the majority of the code for the dissertation research: scrapeUsers.py and preProcess.R.
+The python file contains the necessary code to scrape the Twitter site for tweets, import results into the Postgres database, and perform various cleanup functions that were easiest to do in Python.
+The preProcess.R file contains all modeling and analysis code for the project.
+Within the file, there are implementations for the Bayesian and random permutation models.
+Also there is code to test the models by running each model through the various StackOverflow and Twitter test datasets.
+
+\subsubsection{Model Implementations}
+
+The Bayesian model is implemented using the data.table package, so that fast joins can be performed on the table when computing the strength of association values.
+The random permutation model is represented as a matrix, so that correlations between the model and context vectors are fast.
+The matrix is not sparse, since there are currently no sparse matrix packages for R that have a vectorized function to compute correlations between a vector and a matrix.
+
+The Bayesian model requires slightly more time to compute activation values and also more memory space to store the representation.
+This is because the random permutation model is compressed and only needs to store around \num{2048} rows in the co-occurrence matrix.
+However, both models are implemented efficiently enough so that all analysis runs and activation computations can be kept in memory.
+64 GB of memory was required to perform the analysis, particularly on the largest Twitter (3 million) and StackOverflow (1 million) document sizes.
+
+\subsubsection{Design Decisions}
+
+All model manipulations (e.g., using an entropy weighting metric, using a model with only the prior term) were expressed as part of a configuration object.
+That object is a list of configuration name, value pairs.
+The configuration object is passed to the analysis code, which determines the specific type of run that is performed.
+
+This configuration design enables a large amount of code reuse and overlap.
+This helps minimize the likelihood of a programming mistake, since functions are reused multiple times and overall static code length is kept to a minimum.
+
+\subsubsection{Testing}
+
+Each analysis run produced a set of result files.
+Those result files were checked into the repository so that changes could be tracked.
+Code changes were tested by looking at differences between previously checked in output files and the output files produced after a code change was made.
+
+A small amount of unit testing was also performed.
+The core behavior of each model implementation (Bayesian and random permutation) was tested in isolation to ensure that the activation values returned by the models were mathematically correct.
+Unit test code is located within dissProject/test.
+
+\subsubsection{Parallelization}
+
+None of the analysis code is finely parallelized.
+Most of the inner loop in R ends up running in compiled C, and execution within those function calls occurs on a single core.
+However, since multiple models were explored, each different model configuration was executed in parallel.
+This type of course-grained parallelization was achieved using R's built-in Parralel package.
+Also some parallelization was done at the OS level when various make targets would run in parallel.
+
+\subsubsection{Database}
+
+All raw Tweet and post data were stored in Postgres tables.
+Also, after the documents were tokenized, the tokenized versions of the raw data were stored in Postgres tables.
+When R ran an analysis, it would query the tables to retrieve the correct set of tokenized words and tags.
+So there is a lot of R code that uses the RPostgreSQL package to communicate with Postgres as the analysis is running.
+
+\subsubsection{Running the Analysis}
+
+All analysis runs were scripted and can be run by using the appropriate target in dissProject/Makefile.
+The three main targets to run the models through the prior-only runs on the popular users dataset,
+the full models on the randomly-sampled and popular-hashtags datasets,
+and the full models on the popular users dataset are ``parrunPriors'', ``parrunContext'', and ``parrunPUsers'' respectively.
+Each of these targets takes tens of hours to complete.
+``parrunContext'' is the longest and takes around 72 hours on a desktop machine with 64 GB of memory and 24 virtual cores.
+
+Smaller test runs were also created so that model changes could be more rapidly developed.
+These typically use smaller Twitter and StackOverflow document sizes (e.g., only \num{10000} StackOverflow posts) to create the co-occurrence matrix
+and a smaller number of posts to test the models on (e.g., 20 instead of 500).
+Running these functions takes around a minute to complete.
+To run these, the R function is called directly (i.e., not a make target) within an already-running R session.
+As an example, the ``runContext20g1s6'' function takes only a few minutes to complete, runs the models on a document size of \num{1000}, with only a single run per model, and only 20 sampled posts for each run.
+
 \end{document}

### 2014-09-23 writeups/dissertation/dissertation.tex
foo

@@
 The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
-It is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
+As a applied concern, it is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
 Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
 If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.
+
+However, scale is also a theoretical concern.
+The declarative memory system is tasked to quickly return a single chunk of information that is stored alongside millions to billions of other unique chunks given only the current context and past history of use.
+To do this task, it seems plausible that a model of declarative memory must use a representation that produces accurate retrievals even when large amounts of observations are included.
+Further, it seems plausible that the time to compute an activation value by a computer is correlated with the time to compute an activation value by the brain.
+So any model of declarative memory should also be reasonably computationally efficient even with large co-occurrence datasets.
 
@@
 As one of the largest tests, \textcite{Farahat2004} used the PMI index based on a 118GB local database of 10 million pages of the Stanford WebBase Project.
-So this implementation did not fit completely in RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.
+Even at this large dataset size, the ACT-R declarative memory theory still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.
 
@@
 Unlike StackOverflow, the entire Twitter dataset is not released to the public.
-Due to this and the scale of Twitter, subsets of the Twitter dataset were collected.
-
-\subsubsection{Popular Users Subset}
+Due to this and the scale of Twitter, only the part of the Twitter dataset relevant to this research was collected.
+
+\subsubsection{Popular Users Dataset}
 
@@
 
-\subsubsection{Popular Hashtags Subset}
+\subsubsection{Popular Hashtags Dataset}
 
@@
 
-\subsubsection{Randomly Sampled Subset}
-
-Since the entire dataset is available for StackOverflow, randomly-sampled posts were used to test the context models rather than subsetting and creating a popular-tags dataset (analogous to Twitter).
+\subsubsection{Randomly Sampled Dataset}
+
+Since the entire dataset is available for StackOverflow, randomly-sampled posts were used to test the context models rather than creating a popular-tags dataset (analogous to Twitter).
 If all Twitter data were available, then the models would have been tested on a randomly-sampled dataset for Twitter as well.
-However, since this is not the case for Twitter, the popular-hashtags subset was used instead of a randomly-sampled subset for that site,
+However, since this is not the case for Twitter, the popular-hashtags dataset was used instead of a randomly-sampled dataset for that site,
 while models for StackOverflow were tested on the more challenging dataset where posts were randomly sampled across all posts created on the site.
@@
 For each randomly-sampled post for StackOverflow, the prior activation is based on that user's prior tagging history, and not on any global prior of specific tag frequency across users.
-This was done so that the prior component used for StackOverflow for both the popular-users and randomly-sampled subsets was the same.
+This was done so that the prior component used for StackOverflow for both the popular-users and randomly-sampled datasets was the same.
 

### 2014-09-24 dissProject/preProcess.R
foo

@@ -1456,18 +1456,39 @@ compareDBestVsMax <- function(modelVsPredTbl) {
 	sumTbl[, list(diff=acc[2]-acc[1], direction=paste('best d', '-', 'max d')), by=list(datasetName, user_screen_name, DVName)][, getComparisonTbl(.SD)]
 }
 
+plotTbl <- function(plot, extras, figName) {
+	lapply(extras, function(extra) plot <<- plot + extra)
+	myPlotPrint(plot, figName)
+}
+
 plotBarSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol='dsetGroup') {
 	groupCol = as.symbol(groupCol)
 	fillCol = substitute(fillCol)
 	renameCols(sumTbl)
-	sumTbl
 	expr = bquote(ggplot(sumTbl, aes(x=factor(.(groupCol)), y=meanVal, fill=.(fillCol))) +
 		      geom_bar(aes(y=meanVal), width=0.7, position=position_dodge(), stat='identity') +
 		      geom_errorbar(aes(ymin=minCI, ymax=maxCI), position=position_dodge(width=0.7), width=0.1, size=0.3) + 
 		      defaultGGPlotOpts)
 	plot = eval(expr)
-	lapply(extras, function(extra) plot <<- plot + extra)
-	myPlotPrint(plot, figName)
+	plotTbl(plot, extras, figName)
+	sumTbl
+}
+
+plotLineSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol) {
+	groupCol = as.symbol(groupCol)
+	fillCol = substitute(fillCol)
+	renameCols(sumTbl)
+	eval(bquote(setkey(sumTbl, .(groupCol), .(fillCol))))
+	sizeNums = eval(bquote(sumTbl[!duplicated(.(groupCol))][, .(groupCol)]))
+	sizeNums = parse(text=sizeNums)
+	expr = bquote(ggplot(sumTbl, aes(x=factor(.(groupCol)), y=meanVal, colour=.(fillCol))) + 
+		      geom_line(aes(group=.(fillCol))) + 
+		      geom_point() + 
+		      scale_x_discrete(breaks = unique(sumTbl[, .(groupCol)]), labels = sizeNums) +
+		      geom_errorbar(aes(ymin=minCI, ymax=maxCI), width=0.1, size=0.3) + 
+		      defaultGGPlotOpts)
+	plot = eval(expr)
[…]

### 2014-09-24 writeups/dissertation/dissertation.tex
foo

@@
 profile plots for a single user for both StackOverflow and Twitter are included in Figure \ref{figPriorHTByTime}.
+The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
+What is interesting is the large tag space for these users.
+There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
+For example, the StackOverflow user in Figure \ref{figPriorHTByTime} has a pronounced band where the same hashtag is chosen repeatedly over time. 
+However there are many instances of other tags being used in parallel to the tag that is producing the band (that is repeatedly being used). 
+The Twitter user has a few clusters where the same hashtag is chosen repeatedly but only for a brief period of time.
+As time progresses, a band of choosing another hashtag is observed, and then the user switches to a new hashtag after that.
+More recent posts for the Twitter user are more varied, and there is no longer a single hashtag that the user chooses repeatedly.
+This helps illustrate how challenging modeling this task can be.
+The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is changing over time.
 
@@
 
-The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
-What is interesting is the large tag space for these users.
-There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
-For example, the StackOverflow user in Figure \ref{figPriorHTByTime} has a pronounced band where the same hashtag is chosen repeatedly over time. 
-However there are many instances of other tags being used in parallel to the tag that is producing the band (that is repeatedly being used). 
-The Twitter user has a few clusters where the same hashtag is chosen repeatedly but only for a brief period of time.
-As time progresses, a band of choosing another hashtag is observed, and then the user switches to a new hashtag after that.
-More recent posts for the Twitter user are more varied, and there is no longer a single hashtag that the user chooses repeatedly.
-This helps illustrate how challenging modeling this task can be.
-The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is changing over time.
-
 Although the tag space for the Twitter user is smaller than the StackOverflow user in these plots, this is not a consistent trend.
@@
 each user's best decay rate value was taken and the results are plotted in histogram form in Figures \ref{figPriorSOQHistDsOL} and \ref{figPriorTwitterHistDsOL} respectively.
+The best-fit values are close to the \num{0.5} default value set in ACT-R when the optimized learning form is used.
+Comparing across optimized learning and the standard form, the best-fit values for the standard form are slightly higher for both the StackOverflow and Twitter dataset slices.
+Comparing best-fit values will be more thoroughly tested when looking at the values across all dataset slices.
 
@@
 
-The best-fit values are close to the \num{0.5} default value set in ACT-R when the optimized learning form is used.
-Comparing across optimized learning and the standard form, the best-fit values for the standard form are slightly higher for both the StackOverflow and Twitter dataset slices.
-Comparing best-fit values will be more thoroughly tested when looking at the values across all dataset slices.
-
 \subsubsection{Aggregate Model Performance}
@@
 Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
+Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{0.43}) of $B_{i}$ compared to the standard form (\num{0.80}).
+The optimal values for optimized learning are near \num{0.5}, which lines up with default value used in ACT-R.
+So the decay rate values for the standard form are slightly higher than the default values used in ACT-R, but the values found when optimized learning is used line up nicely with the ACT-R defaults.
 
@@
 \end{figure}
-
-Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{0.43}) of $B_{i}$ compared to the standard form (\num{0.80}).
-The optimal values for optimized learning are near \num{0.5}, which lines up with default value used in ACT-R.
-So the decay rate values for the standard form are slightly higher than the default values used in ACT-R, but the values found when optimized learning is used line up nicely with the ACT-R defaults.
 
@@
 The impact of adding the offset to each of the models for StackOverflow and Twitter is shown in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
+In both plots, if the model name includes ``full'' then all model components are included.
+When they are enumerated, then at least one of the components is left out.
+The random permutation and Bayesian models are plotted alongside each other so that comparisons across the two model types can be done easily.
+So, for example, the ``standard prior model relaxed across posts without offset'' means that only the prior model component was used and the offset for that component was not subtracted off when determining activation.
+As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
   \caption{
@@
 \begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
   \caption{
@@
 \end{figure}
-
-In both plots, if the model name includes ``full'' then all model components are included.
-When they are enumerated, then at least one of the components is left out.
-The random permutation and Bayesian models are plotted alongside each other so that comparisons across the two model types can be done easily.
-So, for example, the ``standard prior model relaxed across posts without offset'' means that only the prior model component was used and the offset for that component was not subtracted off when determining activation.
-As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.
 
@@
 
+In order to ensure that model results are not dependent on the specific Twitter subset collected,
+it is important that the overall accuracy differences between the different models do not drastically shift from subset to subset.
+That is, it is important that the effect size of model is much greater than the effect size of the interaction between Twitter subset and model.
+A small interaction relative to the main effect means that the rank ordering of the different models remains consistent across the different Twitter subsets.
+Further, a small effect size of subset relative to the effect of model means that the exact accuracy for each model remains consistent across the subsets.
+To measure the effect sizes of subset, model type, and the interaction, a two-way between-subjects ANOVA was performed.
+
 \subsubsection{Results}
@@
 Model accuracy results for components of both the random permutation and Bayesian models grouped by Twitter popular-users subset are included in Figure \ref{figContextSubsets}.
+Accuracy is affected significantly by subset ($F(3, 340) = 764, p < .01$)), model ($F(4, 340) = 3617, p < .01$), and the interaction between model and subset ($F(12, 340) = 142, p < .01$).
+However, the effect size of model type on accuracy is \num{8.6} times larger than the interaction term ($\eta ^ {2} = 0.770$ versus $\eta ^ {2} = 0.090$).
+This shows that the performance of a model relative to the other models remains consistent across the four subsets.
+Further, the effect size of model type is \num{6.3} times larger than the effect of Twitter subset ($\eta ^ {2} = 0.770$ versus $\eta ^ {2} = 0.122$).
+This shows that not only is relative model performance maintained across Twitter subsets, but the exact value of model accuracy for each model remains reasonably consistent across subsets.
 
@@
 
-Although there is a main effect of subset, there are minimal to no practically significant interactions.
-Relative model performance remains consistent across the four subsets.
-
 \subsubsection{Discussion}
 
-Since the interactions across popular-users datasets are minimal, model behavior for the popular-users subset for Twitter can be interpreted as behavior that will persist across different time periods,
+Since the effect sizes of subset and the interaction between model type and subset are minimal relative to the effect size of model,
+model behavior for the popular-users subset for Twitter can be interpreted as behavior that will persist across different time periods,
 if the same method of choosing popular hashtags is conducted at a later date.
@@
 Plots for the three methods of handling stop words for the random permutation model on the StackOverflow dataset are included in Figure \ref{figContextStopRPSO}.
+The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
+Using a predefined stoplist makes the smallest improvement (\num{.45} to \num{.49}), while the entropy (\num{.60}) and frequency (\num{.63}) techniques improve performance the most,
+and frequency shows a slight edge over entropy.
+However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
 
@@
 
-The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
-Using a predefined stoplist makes the smallest improvement (\num{.45} to \num{.49}), while the entropy (\num{.60}) and frequency (\num{.63}) techniques improve performance the most,
-and frequency shows a slight edge over entropy.
-However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
-
 The three stop word techniques were then tested for the random permutation model on the Twitter dataset, and the results are included in Figure \ref{figContextStopRPT}.
+For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\num{.25} to \num{.26}).
+Removing stop-words based on a predetermined list actually decreases accuracy (\num{.25} to \num{.23}), and using data-driven frequency filtering has only a small to no effect (\num{.25} to \num{.25}).
 
@@
 \end{figure}
-
-For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\num{.25} to \num{.26}).
-Removing stop-words based on a predetermined list actually decreases accuracy (\num{.25} to \num{.23}), and using data-driven frequency filtering has only a small to no effect (\num{.25} to \num{.25}).
 
@@
 Similar results are found when the Bayesian model is tested on the Twitter dataset.
+Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer (\num{.29} to \num{.32}), and frequency filtering decreases performance relative to no filtering (\num{.29} to \num{.28}).
 
@@
 \end{figure}
-
-Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer (\num{.29} to \num{.32}), and frequency filtering decreases performance relative to no filtering (\num{.29} to \num{.28}).
 
@@
 Model accuracy as a function of co-occurrence matrix size for StackOverflow is included in Figure \ref{figContextDocumentSizeSO}.
+Accuracy improves for both the Bayesian and random permutation models as the size of the documents used to generate the co-occurrence matrix increases.
+However, model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random permutation model, while the frequency technique only begins to plateau.
+The Bayesian model does not plateau and accuracy continues to improve as document size increases.
 
@@
 \end{figure}
-
-Accuracy improves for both the Bayesian and random permutation models as the size of the documents used to generate the co-occurrence matrix increases.
-However, model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random permutation model, while the frequency technique only begins to plateau.
-The Bayesian model does not plateau and accuracy continues to improve as document size increases.
 
@@
 The overall trend of increasing accuracy with document size is present for Twitter as well.
+Both entropy weighting and frequency filtering in Figure \ref{figContextDocumentSizeT} plateau for the random permutation model with increasing document size.
+Once again, the Bayesian model continues to improve with document size.
 
@@
 \end{figure}
-
-Both entropy weighting and frequency filtering in Figure \ref{figContextDocumentSizeT} plateau for the random permutation model with increasing document size.
-Once again, the Bayesian model continues to improve with document size.
 
@@
 Results for using different levels of compression for the random permutation model for the StackOverflow dataset are included in Figure \ref{figContextDimSO}.
+In the figure, the model name that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
+This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random permutation model for all other results.
+The plot is also broken out for different size datasets used to build the co-occurrence matrix.
 
@@
 \end{figure}
-
-In the figure, the model name that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
-This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random permutation model for all other results.
-The plot is also broken out for different size datasets used to build the co-occurrence matrix.
 
@@
 Model accuracy results after using the log-odds transformation are included in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}.
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
-  \caption{Random permutation model accuracy when using log-odds transformation on context for StackOverflow}
-  \label{figContextLogoddsSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
-  \caption{Random permutation model accuracy when using log-odds transformation on context for Twitter}
-  \label{figContextLogoddsT}
-\end{figure}
-
 Using the log-odds transformation technique has a small effect on model accuracy for StackOverflow (\num{.45} to \num{.47}) and minimal to no effect on Twitter (\num{.25} to \num{.25}).
@@
 However, any improvement when using the transformation is minimal.
+
+\begin{figure}[!htbp]
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
+  \caption{Random permutation model accuracy when using log-odds transformation on context for StackOverflow}
+  \label{figContextLogoddsSO}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
+  \caption{Random permutation model accuracy when using log-odds transformation on context for Twitter}
+  \label{figContextLogoddsT}
+\end{figure}
 
@@
 Model accuracy results after incorporating word order into the random permutation model are included in Figure \ref{figContextOrder}.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-orderT-crop.pdf}}}
-  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
-  \label{figContextOrder}
-\end{figure}
-
 All bar plots that include an order component where ``window'' or ``direction'' is not included in the title use the standard ordering method (no direction or windowing, all position information is included)
@@
 Base models that do not use entropy weighting were also included so that effect sizes could be compared between entropy weighting and adding word order information.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-orderT-crop.pdf}}}
+  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
+  \label{figContextOrder}
+\end{figure}
 
@@
 Error bars represent the standard deviation of coefficient values for each component.
+The standard deviation is used instead of the confidence interval of the mean so that the plots show the overall stability of each coefficient across the different runs.
+The range of the distribution of coefficient values within a component is much more narrow than the range of the mean coefficient values between the components of each model.
+This shows that the model is not being overfit, as the weights for each component are relatively stable across the different runs.
 
@@
 
-The standard deviation is used instead of the confidence interval of the mean so that the plots show the overall stability of each coefficient across the different runs.
-The range of the distribution of coefficient values within a component is much more narrow than the range of the mean coefficient values between the components of each model.
-This shows that the model is not being overfit, as the weights for each component are relatively stable across the different runs.
-
 The weight of the prior relative to the context terms is also fairly consistent across models and across the two sites.
@@
 The results are included in Figures \ref{figContextPriorneededSO} and \ref{figContextPriorneededT}.
+For StackOverflow, accuracy remains the same when the prior component is removed for both the Bayesian (\num{.63} to \num{.63}) and random permutation (\num{.63} to \num{.63}) models.
+Accuracy does dramatically decrease for Twitter when the prior is removed, however (\num{.32} to \num{.14} for Bayesian, \num{.25} to \num{.12} for random permutation).
 
@@
 \end{figure}
-
-For StackOverflow, accuracy remains the same when the prior component is removed for both the Bayesian (\num{.63} to \num{.63}) and random permutation (\num{.63} to \num{.63}) models.
-Accuracy does dramatically decrease for Twitter when the prior is removed, however (\num{.32} to \num{.14} for Bayesian, \num{.25} to \num{.12} for random permutation).
 
@@
 The results comparing the full model to a context-only only model for the StackOverflow popular-users dataset are included in Figure \ref{figPUAddPrior}.
+In this case, past tag history does contain unique variance for accurately predicting tag use.
+Accuracy improves for both the Bayesian (\num{.51} to \num{.63}) and random permutation (\num{.47} to \num{.58}) models when the prior model component is combined with context compared to context alone.
 
@@
 \end{figure}
-
-In this case, past tag history does contain unique variance for accurately predicting tag use.
-Accuracy improves for both the Bayesian (\num{.51} to \num{.63}) and random permutation (\num{.47} to \num{.58}) models when the prior model component is combined with context compared to context alone.
 
@@
 Model accuracy when using a user-customized co-occurrence matrix for StackOverflow and Twitter is shown in Figures \ref{figPUCustomSO} and \ref{figPUCustomT}.
+The results compare accuracy for models that use a global co-occurrence and user-specific co-occurrence matrix.
+For StackOverflow, model accuracy for the user-specific co-occurrence matrix (labeled as ``user sji'') can be compared against the default global co-occurrence matrix.
+Accuracy decreases dramatically when a user-specific co-occurrence matrix is used for both the Bayesian (\num{.64} to \num{.35}) and random permutation (\num{.58} to \num{.36}) models.
 
@@
 \end{figure}
-
-The results compare accuracy for models that use a global co-occurrence and user-specific co-occurrence matrix.
-For StackOverflow, model accuracy for the user-specific co-occurrence matrix (labeled as ``user sji'') can be compared against the default global co-occurrence matrix.
-Accuracy decreases dramatically when a user-specific co-occurrence matrix is used for both the Bayesian (\num{.64} to \num{.35}) and random permutation (\num{.58} to \num{.36}) models.
 

### 2014-09-24 dissProject/preProcess.R

@@ -27,6 +27,7 @@ library(reshape2)
 library(boot)
 library(digest)
 library(parallel)
+library(heplots)
 
 PATH = getPathToThisFile()
 FILE = getNameOfThisFile()
@@ -1957,6 +1958,13 @@ analyzeContext <- function(modelVsPredTbl) {
 	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTweetNentropy', 'PriorStdTweetOrderTweetOrderless',
 				       'TweetNentropy', 'TweetOrderTweetOrderless'))
 	compareMeanDVDefault(tbl[sizeNum == 2 & dsetType == 'twitter' & DVName %in% DVNames], acc, figName='contextStandardByGroupT', groupCol='groupNum')
+	
+	fooTbl = copy(tbl[sizeNum == 2 & dsetType == 'twitter' & DVName %in% DVNames])
+	fooTbl[, groupNum := factor(groupNum)]
+	fit = aov(acc ~ DVName * groupNum, data=fooTbl)
+	summary(fit)
+	etasq(fit, anova=T, partial=F)
+	
 	# RESULT: Freq weighting is better than entropy weighting for RP, and standard stoplist doesn't work well at all
 	# SO compare entropy to stoplist to freq to frentropy
 	DVNames = asTopHashtagAcross(c('PriorStd', 'TitleOrderlessEntropy', 'TitleOrderless', 'TitleOrderlessStoplist', 'TitleOrderlessFreq', 'TitleOrderlessFrentropy',

### 2014-09-25 writeups/dissertation/dissertation.tex
foo

@@
 
-\section{Introduction}
+\section{Chapter 1: Introduction}
 
@@
 
-\section{Theoretical Background}
+\subsection{Roadmap}
+
+This document is separated into 7 chapters.
+Chapter 1 introduced the declarative memory research and motivated why it is interesting both from a theoretical and applied perspective to evaluate declarative memory models on large-scale datasets.
+The task to test the models was also introduced: tagging of user-generated content on StackOverflow and Twitter.
+Chapter 2 describes the vector-based and Bayesian declarative memory models that will be evaluated in this research, and shows how these models are related to the other common declarative memory models that are studied.
+Chapter 3 presents the overall method for testing the models on the StackOverflow and Twitter datasets.
+Chapter 4 shows the results for the first test of the models, where the model term for past user behavior is examined in isolation, and tags are predicted based purely on a user's past tagging history.
+Chapter 5 adds the context component to the models, and shows how model accuracy is influenced by changing various architectural components of each model.
+Chapter 6 describes model results when prior and context are combined and tested on both of the dataset subsets used for Chapter 4 and Chapter 5.
+Finally, Chapter 7 provides a general discussion of the overall findings.
+
+\clearpage
+\section{Chapter 2: Theoretical Background}
+
+The ACT-R declarative memory theory and the vector-based retrieval models will be described.
+The Latent Semantic Analysis (LSA) theory is closely related to vector-based retrieval models, so an overview of LSA will be included.
+The theory of the ACT-R and vector-based models will be compared.
+Also, different tagging models have already been applied to various recommendation systems to tag user-generated content,
+so these systems will be described and the models used will be related back to the Bayesian and vector-based declarative memory models.
 
@@
 
-\section{Models}
+\clearpage
+\section{Chapter 3: Overall Method}
+
+Each specific result of the research (to be presented in Chapters 4-6) requires a sightly different method to study.
+However, there is a common general methodology that applies across all results that will be described first.
+Different subsets of the StackOverflow and Twitter datasets were used evaluate past user behavior in isolation and to test the context component of each memory model.
+These datasets will be named and described in general terms.
+Also, the general methods used to tokenize and stem the words in all of the datasets will be described.
+
+\subsection{Models}
 
@@
 The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel} was used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
-
-\section{Method}
 
@@
 
-\section{Past User Behavior}
+\clearpage
+\section{Chapter 4: Past User Behavior}
 
@@
 
-\section{Combining Predictors}
+\clearpage
+\section{Chapter 5: Combining Predictors}
 
@@
 
-\subsection{Method}
+Chapter 5 will examine how model accuracy is influenced by two methodological concerns (adding an offset term and examining the effect of dataset subset for Twitter)
+and five architectural concerns
+(method of handling stop words, influence of dataset size, amount of compression used for the vector-based models, method to combine the prior and context for the vector-based models, and word order).
+The two methodological concerns were examined to ensure that the technique to calibrate model parameters was accurate and also that results for Twitter were not specific to the subset of data collected.
+
+The five architectural concerns were examined for the following reasons:
+During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.
+Due to research from \textcite{Budiu2007}, model accuracy for both the Bayesian and LSA models is influenced by the number of documents in the dataset,
+so the effect of dataset size on accuracy was examined in more detail.
+Specifically,
+since both the Bayesian and random-permutation models are being evaluated, analyzing the effect of dataset size on accuracy in this research will allow direct comparisons between the two models on a common dataset.
+This has not yet been possible since the two models have never been compared on the same dataset within the same study.
+\textcite{Sahlgren2008} showed how different levels of compression in the random-permutation model influence accuracy, so this was examined for three reasons:
+to replicate the findings,
+as an additional check to ensure that the random-permutation model was implemented correctly, and also to see if the same levels of compression are valid for the much larger datasets used in this research.
+One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random-permutation model,
+so two different ways to add the model terms were tested.
+Finally, previous research has shown that including word order in the vector-based models improves performance \parencites{Jones2007, Sahlgren2008},
+so word order was also examined to see if the effect could be replicated and also to measure the overall importance of word order on model accuracy relative to other architectural manipulations.
+
+\subsection{Overall Method}
 
@@
 especially since \num{835500000} co-occurrences were already used with \num{1000000} posts.
-
-\section{Results for Combining Predictors}
 
@@
 
+\clearpage
+\section{Chapter 6: Combining Predictors on the Popular-Users Dataset}
+
 \subsection{Impact of Prior vs. Context}
@@
 
-\subsection{General Results}
+\clearpage
+\section{Chapter 7: General Discussion}
+
+\subsection{Overall Results}
 
 Table \ref{tabOverallResults} summarizes model accuracies for full models (all components) on all of the datasets tested in the study.
+After modifying the models by adding an entropy weighting mechanism to both models and a prior term to the random permutation model,
+accuracy for the compressed random permutation model was nearly equal to the Bayesian model.
+The largest difference between the two models was on the Twitter popular-hashtags dataset (7 percent).
+Both models performed nearly equal for the StackOverflow datasets tested,
+and both models were much more accurate when tested on StackOverflow compared to Twitter.
 
@@
 
-After modifying the models by adding an entropy weighting mechanism to both models and a prior term to the random permutation model,
-accuracy for the compressed random permutation model was nearly equal to the Bayesian model.
-The largest difference between the two models was on the Twitter popular-hashtags dataset (7 percent).
-Both models performed nearly equal for the StackOverflow datasets tested,
-and both models were much more accurate when tested on StackOverflow compared to Twitter.
-
-\section{General Discussion}
-
 Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
@@
 
+\subsection{Implications and Future Directions}
+
 \subsubsection{Generalizing Results}

### 2014-09-25 writeups/dissertation/dissertation.tex
foo

@@
 During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.
+
 Due to research from \textcite{Budiu2007}, model accuracy for both the Bayesian and LSA models is influenced by the number of documents in the dataset,
@@
 This has not yet been possible since the two models have never been compared on the same dataset within the same study.
+
 \textcite{Sahlgren2008} showed how different levels of compression in the random-permutation model influence accuracy, so this was examined for three reasons:
@@
 as an additional check to ensure that the random-permutation model was implemented correctly, and also to see if the same levels of compression are valid for the much larger datasets used in this research.
+
 One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random-permutation model,
 so two different ways to add the model terms were tested.
+
 Finally, previous research has shown that including word order in the vector-based models improves performance \parencites{Jones2007, Sahlgren2008},
@@
 
+Chapter 4 isolated the prior component of the model and analyzed accuracy on the popular-users dataset.
+Chapter 5 added the context components to the models and analyzed accuracy on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.
+Chapter 6 shows the results when the combined prior and context models are tested across all of the datasets.
+Looking at model behavior across the datasets provides a way to show how the importance of the prior component relative to the context component can change depending on specific properties in the dataset that is studied.
+Also, an experiment to a user-custom context co-occurrence matrix was attempted on the popular-users dataset.
+This was tried because it seemed theoretically interesting to see if model performance improves when the context co-occurrence matrix is customized to each user.
+
 \subsection{Impact of Prior vs. Context}
@@
 \section{Chapter 7: General Discussion}
+
+The general discussion is separated into two parts.
+The first section will describe the overall results and reemphasize key findings.
+The final section will discuss general theoretical implications of the results and describe future applications and directions.
 

### 2014-09-26 writeups/dissertation/dissertation.tex
foo

@@
 \documentclass[man,floatsintext,donotrepeattitle]{apa6}
+
+% http://tex.stackexchange.com/questions/279/how-do-i-ensure-that-figures-appear-in-the-section-theyre-associated-with
+\usepackage{placeins}
 
@@
 and also his excellent guidance throughout my time as a graduate student.
-I would also like to thank my parents, Mike and Patsy, for the positive influence that they have provided me for my entire life.
+I would also like to thank my parents, Mike and Patsy, for the positive influence that they have provided me throughout my life.
 Finally, I would like to thank my brothers Brendan and Justin, sister Kellen, extended family, and friends, for all of their support during my time in the Rice Psychology graduate program.
@@
 It is certainly true that at least some production rules are needed when choosing words and creating content. 
-However, it is unexplored if some of the process of sentence construction and word selection can be represented in the declarative memory system.
+However, it is unexplored if some of the process of sentence construction and word selection can be represented in the ACT-R declarative memory system.
 
@@
 
-For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable, and do not change within that window.
+For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable and do not change within that window.
 Programming language popularity over the past few years is a reasonable example of this.
@@
 The fan effect describes a behavior of human memory where memory items are more difficult to retrieve if they have occurred in a larger number of different contexts.
-\textcite{Rutledge2008} used \textcite{Anderson1974} classic experimental paradigm to test the fan effect.
+\textcite{Rutledge2008} used \textcite{Anderson1974}'s classic experimental paradigm to test the fan effect.
 In the task, participants are exposed to a set of person and place pairs (e.g., ``the hippie is in the park'').
@@
 For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store,'' then ``hippie'' would have a fan of two.
-Recall accuracy decrease and time to recall increase as total fan of the person and pair increases.
+Recall accuracy decrease and time to recall increase as the sum total fan of the person and pair increases.
 
@@
 Although there are certainly ways to incorporate word order into the ACT-R DM model.
-For example, an additional dimension could be added to the co-occurrence matrix representing word position.
+For example, an additional dimension could be added to the co-occurrence matrix that represents word position.
 But adding word-order information to ACT-R's default Bayesian model is not common.
@@
 The strength of association component of the ACT-R Bayesian co-occurrence model is a more sophisticated method for computing the term frequency component,
-and ACT-R's attentional weight component of the declarative memory model can be used to measure the information content within each word.
+and ACT-R's attentional weight component of the declarative memory model can be used as a measure of the information content within each word.
 
@@
 
-A additional popular-hashtags dataset was collected for Twitter.
+An additional popular-hashtags dataset was collected for Twitter.
 The purpose of this dataset is to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
@@
 Both terms are methods for computing the base-level activation ($B_{i}$) of a chunk in declarative memory.
-The standard form has a time dependence and accounts for the fact that the rate of presentations for a given object can change over time.
+The standard form has a time dependence and accounts for the fact that the rate of presentations for a given object (e.g., the word ``loop'') can change over time.
 The optimized learning form assumes that this rate is constant.
 
-For both equations, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``loop'' appears), and $n$ is the number of times each chunk has been observed.
+For both equations, $i$ represents each chunk (e.g., the word ``loop''), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``loop'' appears), and $n$ is the number of times each chunk has been observed.
 Since the optimized learning form assumes a constant presentation rate for each chunk,
@@
 
+\FloatBarrier
+
 Finally, model accuracy was analyzed across all dataset slices for StackOverflow and Twitter.
@@
 Looking back at Figure \ref{figPriorDecay}, the optimal decay rate parameter when the standard form of $B_{i}$ is used is higher for Twitter than StackOverflow.
-It makes sense that the lifetime of a specific hashtag on Twitter is shorter overall than the lifetime of a tag (e.g., \emph{C\#}) on StackOverflow.
+It makes sense that the lifetime of a specific hashtag on Twitter is shorter overall than the lifetime of a tag (e.g., \emph{C\#}) on StackOverflow,
+since hashtag topics are generally more transient on Twitter than StackOverflow (e.g., hashtags for specific political events or upcoming conferences).
 This means that the decay rate for Twitter should be higher than StackOverflow, which is what is observed.
@@
 
-Due to research from \textcite{Budiu2007}, model accuracy for both the Bayesian and LSA models is influenced by the number of documents in the dataset,
+Due to research from \textcite{Budiu2007} and \textcite{Recchia2010}, model accuracy for the Bayesian and vector-based models is influenced by the number of documents in the dataset,
 so the effect of dataset size on accuracy was examined in more detail.
@@
 
-The offset component was added to both models, and model accuracy was examined before and after the addition.
+The first methodological concern was the offset component.
+To test the effect of adding the component on model accuracy, the offset component was added to both models, and model accuracy was examined before and after the addition.
 
@@
 
-Four subsets of the popular-users dataset were collected for Twitter.
+As a second methodological concern, model accuracy was examined across each of the dataset subsets collected for Twitter.
+Four popular-hashtags subsets were collected for Twitter at different time periods.
 This was to ensure that model results could be interpreted as effects that were not specific to a certain time period or set of popular hashtags.
@@
 
-Model accuracy results for components of both the random permutation and Bayesian models grouped by Twitter popular-users subset are included in Figure \ref{figContextSubsets}.
+Model accuracy results for components of both the random permutation and Bayesian models grouped by Twitter popular-hashtags subset are included in Figure \ref{figContextSubsets}.
 Accuracy is affected significantly by subset ($F(3, 340) = 764, p < .01$)), model ($F(4, 340) = 3617, p < .01$), and the interaction between model and subset ($F(12, 340) = 142, p < .01$).
-However, the effect size of model type on accuracy is \num{8.6} times larger than the interaction term ($\eta ^ {2} = 0.770$ versus $\eta ^ {2} = 0.090$).
+However, the effect size of model type on accuracy is \num{8.6} times larger than the interaction term ($\eta ^ {2} = 0.77$ versus $\eta ^ {2} = 0.090$).
 This shows that the performance of a model relative to the other models remains consistent across the four subsets.
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-contextStandardByGroupT-crop.pdf}}}
-  \caption{Model accuracy for each of the four Twitter popular-users subsets}
+  \caption{Model accuracy for each of the four Twitter popular-hashtags subsets}
   \label{figContextSubsets}
@@
 Since the effect sizes of subset and the interaction between model type and subset are minimal relative to the effect size of model,
-model behavior for the popular-users subset for Twitter can be interpreted as behavior that will persist across different time periods,
+model behavior for the popular-hashtags subset for Twitter can be interpreted as behavior that will persist across different time periods,
 if the same method of choosing popular hashtags is conducted at a later date.
@@
 
+The first architectural concern examined how the method of handling stop words influenced model accuracy. 
 \textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
@@
 
-The original plot shows the size of the spikes when high-frequency words are not removed.
-The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used (compare the difference between the original subplot and the entropy subplot).
+The no processing plot shows the size of the spikes when high-frequency words are not removed.
+The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used (compare the difference between the no processing subplot and the entropy subplot).
 The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
@@
 
+The second architectural concern examined how the number of documents used to build the co-occurrence matrix influenced model accuracy.
 The default number of documents used to build the co-occurrence matrix was \num{1000000} for StackOverflow and \num{3000000} for Twitter.
@@
 
+As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.
 Another parameter in the random permutation model is the number of rows used to represent the all of the different words.
@@
 So it is impressive that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
-This is likely because word use frequency approximately follows Zipf's law \parencite{Stanley2013}.
+This is likely because word use frequency for StackOverflow and Twitter approximately follows Zipf's law \parencite{Stanley2013}.
 This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply when moving to words that are used fewer times. 
@@
 
+As a fourth architectural concern, two different methods to combine the prior and context components for the random-permutation model were explored.
 The activation value for the context component of the vector-based model is a correlation, ranging from zero to one.
@@
 
+As a final architectural concern, vector-based models with and without word order were tested to examine the effect of word order on model accuracy.
 Vector-based models like the random permutation model can easily incorporate word order when making predictions.
@@
 
-The same four Twitter popular-users datasets were used for testing.
+The same four Twitter popular-hashtags datasets were used for testing.
 Both an ordered and orderless context component were created for the random permutation models.
@@
 so the two representations for words to the left and right of a tag are highly correlated.
-This means that this ordered component is essentially composed of the all co-occurrence observations in the orderless component after being randomly sampled and partitioned into two separate representations.
+This means that this direction ordered component is essentially composed of all the co-occurrence observations in the orderless component after being randomly sampled and partitioned into two separate representations.
 The co-occurrences in these representations are highly correlated with each other, and highly correlated with the orderless component,
@@
 
-The logistic regression is used to determine the optimal weight for each model component.
+All of the methodological and architectural tests were done on the Twitter popular-hashtags dataset and the StackOverflow randomly-sampled dataset.
+For these tests, the logistic regression is used to determine the optimal weight for each model component.
 A regression is done for each run in the dataset, and since multiple runs were used for the StackOverflow and Twitter datasets, multiple samples of coefficient weights for each model component were computed.
@@
 
-Coefficients for each model component of the full models are included in Figures \ref{figContextCoeffSO} and \ref{figContextCoeffT}
+Coefficients for each model component of the full models are included in Figures \ref{figContextCoeffSO} and \ref{figContextCoeffT}.
 Error bars represent the standard deviation of coefficient values for each component.
@@
 
-The same StackOverflow popular-users datasets used to test the prior component in isolation were used to test the prior component when context was added.
+The same StackOverflow popular-users dataset used to test the prior component in isolation was used to test the prior component when context was added.
 However, since computing activations for the context component take an order of magnitude more wall-clock time than computing the prior component, not all observations could be included for the test.
@@
 Also, since the developed models are task agnostic, they can potentially be used as tag recommendation systems for other domains.
-Examples of similar tagging tasks are hashtag use on Facebook, flagging (i.e., tagging) Wikipedia articles for content issues or overly opinionated material,
+Examples of similar tagging tasks are hashtag use on Facebook, flagging (i.e., tagging), Wikipedia articles for content issues or overly opinionated material,
 auto labeling incoming mail and placing it into appropriate folders, and spam filtering.
@@
 and rather in its ability to efficiently compress the full co-occurrence representation used for the Bayesian model.
+Also from a methodological perspective,
+this work has shown how large collections of publicly-available behavioral datasets can be used to explore and test different theories of declarative memory and their individual architectural components.
 Several different formalisms for ACT-R's attentional weighting term were also explored, and an entropy-weighting technique was found to perform well without requiring any parameter tuning.

### 2014-09-26 writeups/dissertation/dissertation.tex
foo

@@
 
-As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.
 Another parameter in the random permutation model is the number of rows used to represent the all of the different words.
@@
 As the number of rows decreases, the amount of cross talk between the environment vectors increases, compression increases, and the amount of information stored decreases.
+As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.
 
@@
 
-As a fourth architectural concern, two different methods to combine the prior and context components for the random-permutation model were explored.
 The activation value for the context component of the vector-based model is a correlation, ranging from zero to one.
@@
 However, this means that ACT-R's prior base-level learning component (which is a log-odds value) is added to a correlation value, which may not be mathematically appropriate.
-It would be more appropriate to convert the correlation value to a log-odds value, and then add this context term to the log-odds prior term.
+As a fourth architectural concern, another method to combine the prior and context components for the random-permutation model was explored.
 
@@
 For a given retrieval, the random permutation model returns a set of activations that are correlation values for each of the tags.
+Instead of simply adding this correlation value to the prior model term, it may be more appropriate to convert the correlation value to a log-odds value first, and then add this context term to the log-odds prior term.
 So a method was created to convert a distribution of correlations to a distribution of log-odds values.
@@
 All of the methodological and architectural tests were done on the Twitter popular-hashtags dataset and the StackOverflow randomly-sampled dataset.
-For these tests, the logistic regression is used to determine the optimal weight for each model component.
+For these tests, a logistic regression statistical technique is used to determine the optimal weight for each model component.
 A regression is done for each run in the dataset, and since multiple runs were used for the StackOverflow and Twitter datasets, multiple samples of coefficient weights for each model component were computed.

### 2014-10-02 writeups/dissertation/dissertation.tex
foo

@@
 \usepackage{graphicx}
+
 \usepackage[group-separator={,}]{siunitx}
+\newcommand{\numNoZero}[1]{{\sisetup{add-integer-zero=false}\num{#1}}}
+
 \RequirePackage[l2tabu, orthodox]{nag}
@@
 Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
-Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{0.43}) of $B_{i}$ compared to the standard form (\num{0.80}).
+Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{.43}) of $B_{i}$ compared to the standard form (\num{.80}).
 The optimal values for optimized learning are near \num{0.5}, which lines up with default value used in ACT-R.
@@
 
-Accuracy for the standard form (\num{0.35}) is higher than the optimized learning form (\num{0.30}).
+Accuracy for the standard form (\numNoZero{.35}) is higher than the optimized learning form (\numNoZero{.30}).
 This makes sense given that the standard form does not assume equal presentation rate for each chunk, and hashtags on Twitter in particular show trends of high rate of use and then shift to low rate of use.
@@
 Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites
-(\num{.53} to \num{.60} and \num{.22} to \num{.45} for Bayesian and random permutation for StackOverflow, \num{.31} to \num{.29} and \num{.24} to \num{.25} for Bayesian and random permutation for Twitter),
+(\numNoZero{.53} to \numNoZero{.60} and \numNoZero{.22} to \numNoZero{.45} for Bayesian and random permutation for StackOverflow,
+\numNoZero{.31} to \numNoZero{.29} and \numNoZero{.24} to \numNoZero{.25} for Bayesian and random permutation for Twitter),
 except that the full Bayesian model without an offset was slightly more accurate on Twitter.
@@
 Model accuracy results for components of both the random permutation and Bayesian models grouped by Twitter popular-hashtags subset are included in Figure \ref{figContextSubsets}.
-Accuracy is affected significantly by subset ($F(3, 340) = 764, p < .01$)), model ($F(4, 340) = 3617, p < .01$), and the interaction between model and subset ($F(12, 340) = 142, p < .01$).
+Accuracy is affected significantly by subset ($F(3, 340) = 764, p < .001$)), model ($F(4, 340) = 3617, p < .001$), and the interaction between model and subset ($F(12, 340) = 142, p < .001$).
 However, the effect size of model type on accuracy is \num{8.6} times larger than the interaction term ($\eta ^ {2} = 0.77$ versus $\eta ^ {2} = 0.090$).
@@
 The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
-Using a predefined stoplist makes the smallest improvement (\num{.45} to \num{.49}), while the entropy (\num{.60}) and frequency (\num{.63}) techniques improve performance the most,
+Using a predefined stoplist makes the smallest improvement (\numNoZero{.45} to \numNoZero{.49}), while the entropy (\numNoZero{.60}) and frequency (\numNoZero{.63}) techniques improve performance the most,
 and frequency shows a slight edge over entropy.
@@
 The three stop word techniques were then tested for the random permutation model on the Twitter dataset, and the results are included in Figure \ref{figContextStopRPT}.
-For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\num{.25} to \num{.26}).
-Removing stop-words based on a predetermined list actually decreases accuracy (\num{.25} to \num{.23}), and using data-driven frequency filtering has only a small to no effect (\num{.25} to \num{.25}).
+For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\numNoZero{.25} to \numNoZero{.26}).
+Removing stop-words based on a predetermined list actually decreases accuracy (\numNoZero{.25} to \numNoZero{.23}), and using data-driven frequency filtering has only a small to no effect (\numNoZero{.25} to \numNoZero{.25}).
 
@@
 
-The results show that entropy weighting produces the most accurate results (\num{.61} to \num{.63}).
-Also, accuracy actually decreases when using frequency filtering (compared to no filtering) (\num{.61} to \num{.56}) for the Bayesian model.
+The results show that entropy weighting produces the most accurate results (\numNoZero{.61} to \numNoZero{.63}).
+Also, accuracy actually decreases when using frequency filtering (compared to no filtering) (\numNoZero{.61} to \numNoZero{.56}) for the Bayesian model.
 So entropy weighting is the clear winner in this case.
@@
 Similar results are found when the Bayesian model is tested on the Twitter dataset.
-Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer (\num{.29} to \num{.32}), and frequency filtering decreases performance relative to no filtering (\num{.29} to \num{.28}).
+Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer (\numNoZero{.29} to \numNoZero{.32}),
+and frequency filtering decreases performance relative to no filtering (\numNoZero{.29} to \numNoZero{.28}).
 
@@
 
-Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy (\num{.60} to \num{.55} for the \num{1000000} document dataset size).
+Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy (\numNoZero{.60} to \numNoZero{.55} for the \num{1000000} document dataset size).
 But \num{2048} rows is sufficient for these datasets,
-as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy (\num{.60} to \num{.60} for the \num{1000000} document size).
+as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy (\numNoZero{.60} to \numNoZero{.60} for the \num{1000000} document size).
 
 Results are similar for Twitter.
-Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100} (\num{.26} to \num{.24} for the \num{3000000} document size),
-and little to no impact when increasing from \num{2048} to \num{10000} (\num{.26} to \num{.26} at \num{3000000}).
+Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100} (\numNoZero{.26} to \numNoZero{.24} for the \num{3000000} document size),
+and little to no impact when increasing from \num{2048} to \num{10000} (\numNoZero{.26} to \numNoZero{.26} at \num{3000000}).
 This results are consistent across all document sizes used to build the co-occurrence matrix.
@@
 Model accuracy results after using the log-odds transformation are included in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}.
-Using the log-odds transformation technique has a small effect on model accuracy for StackOverflow (\num{.45} to \num{.47}) and minimal to no effect on Twitter (\num{.25} to \num{.25}).
+Using the log-odds transformation technique has a small effect on model accuracy for StackOverflow (\numNoZero{.45} to \numNoZero{.47}) and minimal to no effect on Twitter (\numNoZero{.25} to \numNoZero{.25}).
 When the technique is used and accuracy is examined for a model with only a single predictor, the transformation does not change accuracy.
@@
 However, that result is not statistically reliable for the sample size used in this analysis (confidence intervals overlap in Figure \ref{figContextLogoddsSO}),
-and even if present the overall change in accuracy is small (\num{.45} to \num{.47}).
+and even if present the overall change in accuracy is small (\numNoZero{.45} to \numNoZero{.47}).
 Nonetheless, it does seem appropriate to convert the context to a log-odds adjustment value when context is added as an additional term to the random permutation model.
@@
 The results are included in Figures \ref{figContextPriorneededSO} and \ref{figContextPriorneededT}.
-For StackOverflow, accuracy remains the same when the prior component is removed for both the Bayesian (\num{.63} to \num{.63}) and random permutation (\num{.63} to \num{.63}) models.
-Accuracy does dramatically decrease for Twitter when the prior is removed, however (\num{.32} to \num{.14} for Bayesian, \num{.25} to \num{.12} for random permutation).
+For StackOverflow, accuracy remains the same when the prior component is removed for both the Bayesian (\numNoZero{.63} to \numNoZero{.63}) and random permutation (\numNoZero{.63} to \numNoZero{.63}) models.
+Accuracy does dramatically decrease for Twitter when the prior is removed, however (\numNoZero{.32} to \numNoZero{.14} for Bayesian, \numNoZero{.25} to \numNoZero{.12} for random permutation).
 
@@
 In this case, past tag history does contain unique variance for accurately predicting tag use.
-Accuracy improves for both the Bayesian (\num{.51} to \num{.63}) and random permutation (\num{.47} to \num{.58}) models when the prior model component is combined with context compared to context alone.
+Accuracy improves for both the Bayesian (\numNoZero{.51} to \numNoZero{.63}) and random permutation (\numNoZero{.47} to \numNoZero{.58}) models
+when the prior model component is combined with context compared to context alone.
 
@@
 For StackOverflow, model accuracy for the user-specific co-occurrence matrix (labeled as ``user sji'') can be compared against the default global co-occurrence matrix.
-Accuracy decreases dramatically when a user-specific co-occurrence matrix is used for both the Bayesian (\num{.64} to \num{.35}) and random permutation (\num{.58} to \num{.36}) models.
+Accuracy decreases dramatically when a user-specific co-occurrence matrix is used for both the Bayesian (\numNoZero{.64} to \numNoZero{.35}) and random permutation (\numNoZero{.58} to \numNoZero{.36}) models.
 
@@
 
-Model accuracy when only the single user-customized context predictor is used is relatively poor (\num{.17} and \num{.21} for Bayesian and random permutation),
-and the prior term is a much better predictor (\num{.37}).
-Further, when this component is combined with the prior component, model performance improves only slightly over performance when only the prior is used (\num{.37} to \num{.39} for both the Bayesian and random permutation).
+Model accuracy when only the single user-customized context predictor is used is relatively poor (\numNoZero{.17} and \numNoZero{.21} for Bayesian and random permutation),
+and the prior term is a much better predictor (\numNoZero{.37}).
+Further, when this component is combined with the prior component,
+model performance improves only slightly over performance when only the prior is used (\numNoZero{.37} to \numNoZero{.39} for both the Bayesian and random permutation).
 
@@
 	\hline
-	StackOverflow	& popular-users		& Bayesian with entropy				& \num{.63} \\
-			& 			& RP with entropy and log odds 			& \num{.58} \\
-			& randomly-sampled	& Bayesian with entropy 			& \num{.63} \\
-			& 			& RP with entropy and log odds			& \num{.63} \\
-	Twitter		& popular-users		& Bayesian with entropy and user sji		& \num{.39} \\
-			& 			& RP with entropy, log odds, and user sji	& \num{.39} \\
-			& popular-hashtags	& Bayesian with entropy				& \num{.32} \\
-			& 			& RP with entropy and log odds			& \num{.25} \\
+	StackOverflow	& popular-users		& Bayesian with entropy				& \numNoZero{.63} \\
+			& 			& RP with entropy and log odds 			& \numNoZero{.58} \\
+			& randomly-sampled	& Bayesian with entropy 			& \numNoZero{.63} \\
+			& 			& RP with entropy and log odds			& \numNoZero{.63} \\
+	Twitter		& popular-users		& Bayesian with entropy and user sji		& \numNoZero{.39} \\
+			& 			& RP with entropy, log odds, and user sji	& \numNoZero{.39} \\
+			& popular-hashtags	& Bayesian with entropy				& \numNoZero{.32} \\
+			& 			& RP with entropy and log odds			& \numNoZero{.25} \\
 	\hline

### 2014-10-15 writeups/dissertation/dissertation.tex
foo

@@
   The tasks of choosing a hashtag for a tweet and tagging a post on StackOverflow were framed as declarative memory retrieval problems.
-  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R based Bayesian model and a random permutation vector-based model.
-  Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
-  The results show that past user behavior of tag use is a strong predictor of future behavior, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
+  Two state-of-the-art cognitively-plausible declarative memory models were evaluated on how accurately they predict a user's chosen tags: an ACT-R based Bayesian model and a random permutation vector-based model.
+  Millions of posts and tweets were collected, and both declarative memory models were used to predict Twitter hashtags and StackOverflow tags.
+  The results show that past user behavior of tag use is a strong predictor of future behavior.
+  Furthermore, past behavior was successfully incorporated into the random permutation model that previously used only context.
   Also, ACT-R's attentional weight term was linked to a common entropy-weighting natural language processing method used to attenuate low-predictor words.
-  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparable to the Bayesian model without including word order.
+  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparably to the Bayesian model without including word order.
   This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
@@
 The declarative memory component in ACT-R is a Bayesian statistical model.
-This is the same type of co-occurrence Naive Bayes statistical model that is prominent in the more general machine learning field.
+This is the same type of co-occurrence-based Naive Bayes statistical model that is prominent in the more general machine learning field.
 When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
@@
 This approach can and has worked well.
-Essentially the modeler is only including the observations in declarative memory that are relevant to the specific task that they are studying.
+Essentially, the modeler is only including the observations in declarative memory that are relevant to the specific task that they are studying.
 
@@
 This has important implications for understanding declarative memory:
-First, it is difficult to verify and explore modifications to the declarative memory architecture because the dataset is not large enough or rich enough to rigorously test modifications to the theory.
+First, it is difficult to verify and explore modifications to the declarative memory architecture because the datasets are not large enough or rich enough to rigorously test modifications to the theory.
 It also means that we do yet fully understand how well the ACT-R declarative memory theory scales when the number of chunks begins to approach the number of chunks that are actually stored in human declarative memory.
@@
 Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp are composed almost entirely of human-created content.
-The amount of user content on these sites is staggering, and continues to grow.
+The amount of user content on these sites is quite large, and continues to grow.
 Users of Twitter's microblogging service, for example, generate half a billion tweets per day \parencite{TwitterReport2012}.
@@
 
-A growing number of users on social media sites are focusing on and looking for recently created information when performing a search \parencite{Jansen2011}.
+A growing number of users on social media sites are focusing on and looking for recently-created information when performing a search \parencite{Jansen2011}.
 In other words, users on social media sites want the fresh \emph{stream} of information about a particular topic.
@@
 These two memory systems were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
-Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
-A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#.''
+Twitter is a microblogging service where users create 140 character ``tweets'' and broadcast those messages out to the users who follow their posts.
+A user is free to annotate important concepts or words in the tweet by preceding words with ``\#,'' thereby making the word into a hashtag.
 After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
@@
 Note how users both add hashtags to the end of a tweet as a summary of the message and emphasize important concepts words in the tweet by turning them into hashtags.
-That is, hashtags are both intermixed within the words of the tweet and added to the end of the message.
-
-These memory retrieval models were also tested on tag use for StackOverflow posts.
+That is, hashtags can be both intermixed within the words of the tweet and added to the end of the message.
+
+The two memory retrieval models were also tested on tag use for StackOverflow posts.
 StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
@@
 Also, studying user-based tag generation on these sites has relevant real-world application.
-Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these very popular sites.
+Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these popular sites.
 These systems can also help newer users by recommending proper tags for their newest content.
@@
 Having a recommendation system that can suggest relevant hashtags to the user provides a way for the user to connect to and discover new information streams that interest them.
+Also, by suggesting relevant hashtags the system can help make sure a user's content is seen by other users who are interested in the same type of information.
 
@@
 The ACT-R declarative memory theory and the vector-based retrieval models will be described.
-The Latent Semantic Analysis (LSA) theory is closely related to vector-based retrieval models, so an overview of LSA will be included.
-The theory of the ACT-R and vector-based models will be compared.
-Also, different tagging models have already been applied to various recommendation systems to tag user-generated content,
-so these systems will be described and the models used will be related back to the Bayesian and vector-based declarative memory models.
+The Latent Semantic Analysis (LSA) theory is closely related to vector-based retrieval models, so an overview of LSA will also be included.
+The theories underlying the ACT-R and vector-based models will be compared.
+Also, different tagging models have already been applied to various recommendation systems to tag user-generated content.
+Many of these models are based on the same Bayesian statistical framework that is the foundation of the ACT-R declarative memory theory.
+So these systems will be described and the models used will be discussed within the context of Bayesian statistical models, LSA approaches, and the ACT-R declarative memory theory.
 
@@
 The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
-The equations that make up this system were derived from a rational analysis of long-term memory retrieval.
-That is, given the task of retrieving a chunk of information from long-term declarative memory,
+The equations that make up this system were derived from a rational analysis of declarative memory retrieval.
+That is, given the task of retrieving a chunk of information from declarative memory,
 the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
@@
       Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
-      Constant Base Level &	$B_{i} = log \frac{p(i)}{p(\overline{i})}$ \\
+      Constant Base Level &	$B_{i} = \frac{p_{i}}{1-p_{i}}$ \\
       Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
@@
 The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
+Each term in Table \ref{tabACTRModel} will be discussed in greater detail to follow.
 
@@
 It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
+The index has been used to accurately and efficiently measure strength of association information for large-scale datasets \parencite{Budiu2007,Farahat2004}.
 The PMI index is presented as Equation \eqref{eqPMI}.
@@
 
+\subsubsection{Attentional Weight}
+
+The attentional weight term allows for some contextual components to be weighted more than others when the total $S_{ji}$ activation is computed.
+This reflects the fact that the declarative memory system has limited attentional resources ($W$) that must be spread across all of the items in context.
+Most ACT-R models simply weight each contextual element equally (e.g., all words in a StackOverflow post have equal weight), so each word's attentional weight is simply normalized by the number of items in context.
+
 \subsubsection{Scaling the Equations}
@@
 The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
-As a applied concern, it is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
+As an applied concern, it is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
 Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
@@
 
-The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} is also based on the large-scale version of $S_{ji}$, and scaled well when working with a large-scale search query database for internet search.
-
-As one of the largest tests, \textcite{Farahat2004} used the PMI index based on a 118GB local database of 10 million pages of the Stanford WebBase Project.
+The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} was developed to predict user navigation of web-based hyperlinks (e.g., choosing between a set of results from a search query).
+The theory uses the large-scale version of $S_{ji}$ to measure the relatedness between a user's current goals and the current context.
+It was shown to scale well when working with a large-scale search query database for internet search.
+
+As one of the largest tests for the PMI index, \textcite{Farahat2004} generated the co-occurrence matrix from a 118GB local database of 10 million pages of the Stanford WebBase Project.
+\citeauthor{Farahat2004} compared strength of association values using the PMI index and similarity ratings derived from human raters.
 Even at this large dataset size, the ACT-R declarative memory theory still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.
@@
 The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
-This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
+This process, described in more detail to follow, is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
 The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
@@
 For example, \textcite{Jones2007} compared a vector-based model called BEAGLE directly to LSA, training and testing both models on the same corpus used in \textcite{Landauer1997}.
-The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the task (\num{55.6}\% vs. \num{55.3}\%).
+The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the TOEFL task (\num{55.6}\% vs. \num{55.3}\%).
 However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to \num{57.5}\%.
@@
 
-Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
+Another issue with LSA is that the SVD matrix decomposition technique is computationally expensive.
 As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
@@
 This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
-Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to \num{73.75}\%, which is 10\% higher than the TASA-trained LSA.
+Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to \num{73.8}\%, which is 10\% higher than the TASA-trained LSA.
 This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.
@@
 The resonance and decoding processes are inversions of each other.
-Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
+Usually both resonance and decoding will return similar rank orderings of most likely chunks given context.
 \textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
@@
 Both random permutations and LSA calculate word similarity on a reduced matrix.
-For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors are much less than the number of words.
+For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors is much less than the number of words.
 For the StackOverflow and Twitter datasets, the number of documents are the number of different tags.
@@
 It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
-This does not seem to be the case in testing however, as vector-based approaches like BEAGLE and random permutations perform just as well as the LSA when trained and tested on similar datasets.
-Nonetheless, it has been suggested to perform an SVD reduction technique on a compressed random permutation vector \parencite{Kanerva2000}.
+This does not seem to be the case in testing however \parencites{Sahlgren2008,Jones2007}.
+Nonetheless, \textcite{Kanerva2000} suggested to perform an SVD reduction technique on a compressed random permutation vector.
 This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.
@@
 For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
-\textcite{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect, which is a psychological phenomena that ACT-R's current Bayesian DM system can model as well.
+\citeAuthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect.
 
@@
 
-Both vector-based and Bayesian models produce the fan effect since they both are based on a co-occurrence matrix.
+Both vector-based and Bayesian models produce the fan effect since they both are based on co-occurrences.
 However, vector-based models do further processing on that matrix, reduce its dimensionality, and use a correlation instead of a log odds value to compute the activation.
@@
 
-One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
+One strength of a vector-based approach over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
 With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
-Although there are certainly ways to incorporate word order into the ACT-R DM model.
-For example, an additional dimension could be added to the co-occurrence matrix that represents word position.
-But adding word-order information to ACT-R's default Bayesian model is not common.
+To represent word order in a Bayesian model, an additional dimension could be added to the co-occurrence matrix that represents word position.
+However, adding word-order information to ACT-R's default Bayesian declarative memory model is not common.
+Instead, when word order information is used to model reading for example, production rules are defined so that words are read from left to right (e.g., \textcite{Lewis2006}).
 
@@
 The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
-Netflix is an online subscription service where users can stream their favorite television episodes and movies.
+Netflix is an online subscription service where users can stream television episodes and movies.
 Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
@@
 Lucene \parencite{McCandless2010} is a commonly-used open source software tool that measures the similarity between a query and candidate documents in a database.
-It uses the term frequency---inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
+It uses the term frequency-inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
 This measure is the product of two terms: the frequency count of the word in the document and the log inverse of the ratio of the number of documents that the word appeared in to the total number of documents.
@@
 The inverse document frequency component measures how much information is contained within a specific word, independent of document.
-ACT-R also uses two similar components to compute the activation of a word with a specific document.
-The strength of association component of the ACT-R Bayesian co-occurrence model is a more sophisticated method for computing the term frequency component,
-and ACT-R's attentional weight component of the declarative memory model can be used as a measure of the information content within each word.
 
@@
 He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
-Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive LSA (since it requires Singular Value Decomposition) and K-nearest-neighbors models.
+Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive LSA and K-nearest-neighbors models.
 This Bayesian model can accurately predict a user-chosen tag at 47\% accuracy.
@@
 So it stands to reason that the algorithm is relatively complex, and there could be possible scaling and efficiency concerns with the algorithm's implementation, so it has not yet been deployed to StackOverflow.
-Or it is also possible that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
+Alternatively, it is also possible that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
 Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial to the user and has started to add them to their question and answer sites.
@@
 
-The hashtag recommendation system deployed for the Google+ microblogging service is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
+The hashtag recommendation system deployed for the Google+ social media site is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
 The user interface for this recommendation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.
@@
 One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a post as a function of prior hashtag use and context \parencite{Mazzia2009}.
-The model presented by \textcite{Mazzia2009} is very similar to an ACT-R declarative retrieval model.
+The model presented by \textcite{Mazzia2009} is quite similar to an ACT-R declarative retrieval model.
 The main potential difference is that --- like \textcite{Kuo2011} and \textcite{Stanley2013} --- the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
 Instead, the global prior is an overall average frequency of hashtag use across all users.
+Further, no recency information was used to compute the global prior, so hashtags used in the recent past were not weighted more heavily. 
+This was most likely not much of an issue for \citeAuthor{Mazzia2009}'s study, since all hashtags were collected within a relatively short period of time (about 3 months).
 
@@
 \textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
-\textcites{Zangerle2011, Kywe2012} use the Apache Lucene infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score to compute similarities.
+\textcites{Zangerle2011, Kywe2012} used the Apache Lucene infrastructure to compute the similarity score, while \textcite{Li2011} used a custom similarity score to compute similarities.
 Regardless of the method to compute the similarity between tweets,
@@
 There has been a mix of co-occurrence models created to predict hashtag use on Twitter, and the results are encouraging.
-However, there has been very little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
+However, there has been little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
 \textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
@@
 \clearpage
-\section{Chapter 3: Overall Method}
+\section{Chapter 3: General Methodology}
 
@@
 Unlike StackOverflow, the entire Twitter dataset is not released to the public.
-Due to this and the scale of Twitter, only the part of the Twitter dataset relevant to this research was collected.
+Due to this and the scale of Twitter, only a part of the Twitter dataset relevant to this research was collected.
 
@@
 It was used to explore the growth and decay characteristics of a hashtag for individual users.
-The temporal dynamics of ACT-R's decay rate model for a chunk were evaluated against this data.
+The temporal dynamics of ACT-R's decay rate model for a chunk were evaluated against these data.
+Multiple samples were taken across a broad range of user types for each dataset to ensure that the model results were not dependent on specific attributes in a single sample.
+
+For StackOverflow, top users were identified by their reputation score on the site (based on their upvoted questions and answers) and the total number of questions asked on the site.
+To create the StackOverflow popular-users dataset, a total of 12 slices of users were taken at different levels of reputation and total number of questions asked.
+500 users were sampled that had a reputation higher than \num{500}, \num{1000}, \num{5000}, \num{10000}, \num{50000}, and \num{100000}.
+100 users were sampled that had asked a total number of questions higher than \num{50}, \num{100}, \num{200}, \num{300}, \num{400}, \num{500}.
+
+For Twitter, top users were those that had the largest number of followers and had authored the largest number of tweets.
+To create the Twitter popular-users dataset, 12 slices were also taken, this time at different levels of number of followers and total number of tweets.
+100 users were sampled that had a total number of followers higher than \num{1000}, \num{5000}, \num{10000}, \num{100000}, \num{1000000}, and \num{10000000},
+and a total number of tweets higher than \num{100}, \num{500}, \num{1000}, \num{5000}, \num{10000}, and \num{50000}.
 
@@
 
-Twitter provides two types of APIs for acquiring data: a query-based and streaming interface.
+Twitter provides two types of APIs for acquiring data: one based on streaming and a direct query-based interface.
 The Tweepy Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
@@
 However, during testing it was also found to be robust and well suited for StackOverflow content.
-The initial model results with the new tokenizer were qualitatively very similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
+The initial model results with the new tokenizer were qualitatively similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
 This is most likely because the language for the samples taken for both sites is informal English,
@@
 The StackOverflow and Twitter popular-users datasets were used to characterize tag growth and decay for a user on each site.
+The popular-users dataset contains 12 slices of popular users for both StackOverflow and Twitter.
 Each user's tag behavior from the popular-users dataset for each site was analyzed.
-
-Multiple samples were taken across a broad range of user types for each dataset to ensure that the model results were not dependent on specific attributes in a single sample.
-For the StackOverflow popular-users dataset, a total of 12 slices of users were taken at different levels of reputation and total number of questions asked.
-500 users were sampled that had a reputation higher than \num{500}, \num{1000}, \num{5000}, \num{10000}, \num{50000}, and \num{100000}.
-100 users were sampled that had asked a total number of questions higher than \num{50}, \num{100}, \num{200}, \num{300}, \num{400}, \num{500}.
-
-For the Twitter popular-users dataset, 12 slices were also taken, this time at different levels of number of followers and total number of tweets.
-100 users were sampled that had a total number of followers higher than \num{1000}, \num{5000}, \num{10000}, \num{100000}, \num{1000000}, and \num{10000000},
-and a total number of tweets higher than \num{100}, \num{500}, \num{1000}, \num{5000}, \num{10000}, and \num{50000}.
 
@@
 Only values equal to or below 1 were examined for the optimized learning form, since values above 1 for this form cause a bifurcation and are not psychologically interesting.
+The bifurcation occurs because at a decay rate of 1 the denominator of the first term to compute $B_{i}$ goes to zero (see Table \ref{tabACTRBLLModel}).
 
@@
 \begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
   \caption{Model tagging profile of a single user on StackOverflow}
@@
 \begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
   \caption{Model tagging profile of a single user on Twitter}
@@
     Model performance for a single dataset slice for StackOverflow.
+    This includes model accuracy as a function of decay rate value for each user in the dataset slice.
+    Model accuracy is shown both when the standard from of base-level activation ($B_{i}$) and the optimized learning form of the equation are used. 
   }
@@
     \subfigure[][Optmized learning form]{\includegraphics[width=8cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
-    \caption{Model performance for a single dataset slice for Twitter}
+    \caption{
+      Model performance for a single dataset slice for Twitter.
+      This includes model accuracy as a function of decay rate value for each user in the dataset slice.
+    }
     \label{figPriorTwitterSliceDsStd}
@@
 
-When comparing the standard and optimized learning forms of the equation, it is apparent that a pure recency model does not work very well for optimized learning.
+When comparing the standard and optimized learning forms of the equation, it is apparent that a pure recency model does not work well for optimized learning.
 The best-fit decay rate value for the optimized learning form is also not as clear and pronounced as it is for the standard form (less of a peak).
@@
     \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visHistD-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
-    \caption{Best-fit decay rate for a single dataset slice for StackOverflow}
+    \caption{
+      Best-fit decay rate for a single dataset slice for StackOverflow.
+      The histogram represents the count of users for each best-fit decay rate value.
+      Each bar represents a discrete decay rate value that was explored.
+  }
     \label{figPriorSOQHistDsOL}
@@
     \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visHistD-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
-    \caption{Best-fit decay rate for a single dataset slice for Twitter}
+    \caption{
+      Best-fit decay rate for a single dataset slice for Twitter.
+      Best-fit values were slightly higher for Twitter, so the range of the histogram is higher for Twitter than StackOverflow.
+  }
     \label{figPriorTwitterHistDsOL}
@@
 Finally, model accuracy was analyzed across all dataset slices for StackOverflow and Twitter.
+This was done by examining model accuracy at a specific decay rate value for each model.
+Accuracy was assessed at each model's previously-determined best-fit values (\num{.43} and \num{.80} for the optimized learning and standard form), and the standard form's default value (\num{0.5}).
+Since decay rate values were explored at \num{0.1} increments between the 0 and 1 range, the decay rate values used to examine model accuracy were rounded to the nearest decay rate value that was included in the run.
+So a best-fit value of \num{.4} was used for the optimized learning model instead of the \num{.43} average found from the dataset slices.
 Accuracy for both sites is included in Figure \ref{figPriorAcc}. 
@@
   \caption{
-    Overall model accuracy for StackOverflow and Twitter at optimal decay-rate values.
-    Best-fit decay-rate values for the standard and optimized learning model are \num{.8} and \num{.43}.
+    Overall model accuracy for StackOverflow and Twitter at optimal decay rate values.
     All error bars are the 95\% bootstrapped confidence interval for the mean. 
@@
 
-Accuracy for the standard form (\numNoZero{.35}) is higher than the optimized learning form (\numNoZero{.30}).
+Accuracy for the standard form (\numNoZero{.33}) is higher than the optimized learning form (\numNoZero{.28}).
 This makes sense given that the standard form does not assume equal presentation rate for each chunk, and hashtags on Twitter in particular show trends of high rate of use and then shift to low rate of use.
@@
 This was to test how much of the variance in model accuracy is shared between the orderless and order models.
-The results show that there is very little unique variance in the order-only models, so order information contributes very little to overall model performance above and beyond what an orderless model contributes.
+The results show that there is little unique variance in the order-only models, so order information contributes little to overall model performance above and beyond what an orderless model contributes.
 This can be seen by comparing the two-component model with only prior and orderless context to the three-component model with prior, orderless, and any of the three methods for the order context term.
@@
 The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
-These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is a long-term memory retrieval process.
+These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is a declarative memory retrieval process.
 However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
@@
 These results pose an interesting and important fundamental question about declarative memory:
-How, precisely, is the neural system configured to represent co-occurrences for long-term memory?
+How, precisely, is the neural system configured to represent co-occurrences for declarative memory?
 Since at the behavioral level both models perform similarly, it is possible that the neural system is using either representation to store co-occurrences.
@@
 Several different formalisms for ACT-R's attentional weighting term were also explored, and an entropy-weighting technique was found to perform well without requiring any parameter tuning.
-Finally, this work has helped raise the question about how co-occurrence observations are represented at the neural level, since two models with very different representations performed similarly well.
+Finally, this work has helped raise the question about how co-occurrence observations are represented at the neural level, since two models with quite different representations performed similarly well.
 The results from the study and the new research questions raised will help to better understand how imposing specific architectural constraints on memory models influence what information is retrieved.

### 2014-10-15 dissProject/preProcess.R

@@ -1601,6 +1601,8 @@ renameColDVName <- function(tbl) {
 	mapping = c('topHashtagAcrossPriorStd', 'Standard prior model relaxed across posts',
 		    'topHashtagPostPriorStd', 'Standard prior model',
 		    'topHashtagPostPriorStdDefaultD', 'Standard prior model at default d=0.5',
+		    'topHashtagPostPriorStdBestDStd', 'Standard prior model at best-fit d=0.8',
+		    'topHashtagPostPriorOL2BestDOL', 'Optimized learning model at best-fit d=0.4',
 		    'topHashtagPostPriorOL2', 'Optimized learning model',
 		    'topHashtagAcrossPriorStdNoffset', 'Standard prior model relaxed across posts w/o offset',
 		    makeLogregMapping(),
@@ -1774,9 +1776,6 @@ analyzePrior <- function(modelVsPredTbl) {
 	modelVsPredTbl[, list(N=.N, names=list(unique(datasetName))), by=list(dsetType, dsetGroup, runNum,datasetNameRoot)]
 	bestTbl = modelVsPredTbl[predUsedBest == T]
 	tbl = modelVsPredTbl[topHashtag & hashtagUsedP] 
-	tbl[d == 0.5 & !maxNP, DVName := paste0(DVName, 'DefaultD')]
-	tbl = rbind(tbl, tbl[d == 0.5 & maxNP][, DVName := paste0(DVName, 'DefaultD')])
-	tbl = tbl[maxNP | d == 0.5]
 	#dvDiffsTbl = plotDVDiffs(rbind(bestTbl[runNum==2, compare2DVs(.SD, c('topHashtagPostPriorStd', 'topHashtagPostPriorOL2'), sortedOrder=c(1,2)), by=list(dsetType, dsetGroup), .SDcols=colnames(bestTbl)],
 	#			       bestTbl[runNum==2, compare2DVs(.SD, c('topHashtagPostPriorStd', 'topHashtagAcrossPriorStd')), by=list(dsetType, dsetGroup), .SDcols=colnames(bestTbl)]
 	#			       ))
@@ -1785,8 +1784,15 @@ analyzePrior <- function(modelVsPredTbl) {
[…]

### 2014-10-16 writeups/dissertation/dissertation.tex
foo

@@
 In order to best combine these predictors, a logistic regression statistical technique similar to \textcite{Stanley2013} was used to determine the most optimal weights for each term.
-Each term in the full model (see the total activation equation Table \ref{tabModACTRModel}) is weighted so that model accuracy is maximized.  
-\textcite{Stanley2013} found that the optimal weights for the StackOverflow model for the prior term, title context term, and body context term were \num{.62}, \num{.93}, and \num{1.75} respectively.
-
-The optimal weights for each model component are computed in the following way:
+\textcite{Stanley2013} weighted each term in the total activation equation Table \ref{tabModACTRModel}) to maximize model accuracy.
+\citeAuthor{Stanley2013} found that the optimal weights for the StackOverflow model for the prior term, title context term, and body context term were \num{.62}, \num{.93}, and \num{1.75} respectively.
+
+The optimal weights for each model component are computed in the same manner as \textcite{Stanley2013}:
 When a user creates a tag for a post, a retrieval request for the model is made.
@@
 Instead of subtracting off the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
-So for the prior component, the activation value used in the logistic regression is the prior activation with the mean of the top 5 prior activations subtracted off of this value.
+So for the prior component, the activation value used in the logistic regression is the prior activation with the mean of the top 5 prior activations subtracted from this value.
 The same process is used for the context component.
@@
 The matrix for \num{3000000} posts was only used for Twitter, since the number of words in a StackOverflow post is much higher than Twitter on average,
-and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high performance desktop, even with highly optimized representations.
+and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too large for a current high-performance desktop, even with highly optimized representations.
 Already \num{835500000} total co-occurring word tag observations were used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
@@
 In both plots, if the model name includes ``full'' then all model components are included.
-When they are enumerated, then at least one of the components is left out.
+When the model components are enumerated, then at least one of the components is left out.
 The random permutation and Bayesian models are plotted alongside each other so that comparisons across the two model types can be done easily.
@@
 This suggests that the mean activation for the random permutation model for the top few tags in each post can vary greatly across posts.
-Only if this mean activation is subtracted off will the logistic regression model generate optimal weights and predictions that are comparative to the Bayesian model when model accuracy relaxed across posts is examined.
+Only if this mean activation is subtracted will the logistic regression model generate optimal weights and predictions that are comparable to the Bayesian model when model accuracy relaxed across posts is examined.
 
@@
 Adding the offset term into the model is a way to implement a simple version of this strategy where the author is actively choosing to use only a small set of tags for each post.
+Also, this type of strategic behavior of limiting the total number of chosen tags could be easily implemented in the production system of ACT-R.
 
@@
 The hashtags used in each subset were chosen by scraping a site that published the current trending hashtags for each city.
-Before starting to collect for each subset, the current trending popular hashtags for each US city were collected, and then 400 were randomly sampled from this set.
+Before starting to collect for each subset, the current trending popular hashtags for the 400 most populous US cities were collected, and then 400 popular hashtags were randomly sampled from this set.
 Only 400 were selected since this is the maximum that the Twitter API can monitor at one time.
@@
 
-Model accuracy results for components of both the random permutation and Bayesian models grouped by Twitter popular-hashtags subset are included in Figure \ref{figContextSubsets}.
-Accuracy is affected significantly by subset ($F(3, 340) = 764, p < .001$)), model ($F(4, 340) = 3617, p < .001$), and the interaction between model and subset ($F(12, 340) = 142, p < .001$).
-However, the effect size of model type on accuracy is \num{8.6} times larger than the interaction term ($\eta ^ {2} = 0.77$ versus $\eta ^ {2} = 0.090$).
+Model accuracy results for components of both the random permutation and Bayesian models grouped by Twitter popular-hashtags subset are depicted in Figure \ref{figContextSubsets}.
+Accuracy is affected significantly by subset ($F(3, 340) = 764, p < .001, \eta ^ {2} = .12$), model ($F(4, 340) = 3617, p < .001, \eta ^ {2} = .77$),
+and the interaction between model and subset ($F(12, 340) = 142, p < .001, \eta ^ {2} = .09$).
+However, the effect size of model type on accuracy is \num{8.6} times larger than the interaction term ($\eta ^ {2} = .77$ versus $\eta ^ {2} = .09$).
 This shows that the performance of a model relative to the other models remains consistent across the four subsets.
-Further, the effect size of model type is \num{6.3} times larger than the effect of Twitter subset ($\eta ^ {2} = 0.770$ versus $\eta ^ {2} = 0.122$).
+Further, the effect size of model type is \num{6.3} times larger than the effect of Twitter subset ($\eta ^ {2} = .77$ versus $\eta ^ {2} = .12$).
 This shows that not only is relative model performance maintained across Twitter subsets, but the exact value of model accuracy for each model remains reasonably consistent across subsets.
@@
 
-The first architectural concern examined how the method of handling stop words influenced model accuracy. 
+The method of handling stop words has been shown to influence model accuracy \parencite{Sahlgren2008,Stanley2013}, so this was the first architectural concern that was examined.
 \textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
-A data-driven frequency-weighted approach was tried, along with using a previously-derived stoplist, and no stop word removal at all.
-Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and using a previously-derived stoplist hardly increased performance.
+A data-driven frequency-weighted approach was tried, along with using a previously-generated stoplist, and no stop word removal at all.
+Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and using a previously-generated stoplist hardly increased performance.
 However, \citeauthor{Sahlgren2008} did not try to weight the stop words and only tried methods to remove them.
-So alongside looking at frequency weighting, previously derived, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on entropy was also explored.
+So alongside looking at frequency weighting, previously generated, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on entropy was also explored.
 
@@
   \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{memMatSO-crop.pdf}}}
-  \caption{Variance in observation counts of words for each tag for StackOverflow}
+  \caption{
+    Variance in observation counts of words for each tag for StackOverflow.
+    Plots are shown for the four different methods for handling stop words.
+    Each plot contains the standard deviation of the total counts for each word in context as a function of hashtag.
+    High values mean that there are context words for a hashtag that have a high number of counts relative to the counts for all other context words for that hashtag.
+  }
   \label{figContextCutoffSO}
@@
   \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{memMatT-crop.pdf}}}
-  \caption{Variance in observation counts of words for each tag for Twitter}
+  \caption{
+    Variance in observation counts of words for each tag for Twitter.
+  }
   \label{figContextCutoffT}
@@
 
-The three stop word techniques were then tested for the random permutation model on the Twitter dataset, and the results are included in Figure \ref{figContextStopRPT}.
+The three stop word techniques were then tested for the random permutation model on the Twitter dataset, and the results are depicted in Figure \ref{figContextStopRPT}.
 For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\numNoZero{.25} to \numNoZero{.26}).
@@
 Since accuracy measurements when using a predetermined list to remove stop words for the random permutation model were much worse than the frequency and entropy methods, this method was not examined for the Bayesian model.
-The results for the frequency and entropy techniques for the Bayesian model on the StackOverflow dataset are included in Figure \ref{figContextStopBSO}.
+The results for the frequency and entropy techniques for the Bayesian model on the StackOverflow dataset are depicted in Figure \ref{figContextStopBSO}.
 
@@
 
-\subsection{Dataset Size}
-
-The second architectural concern examined how the number of documents used to build the co-occurrence matrix influenced model accuracy.
+\subsection{Corpus Size}
+
+The number of documents used to build the co-occurrence matrix has been shown to influence model accuracy \parencite{Sahlgren2008}, so this was the second architectural concern that was examined.
 The default number of documents used to build the co-occurrence matrix was \num{1000000} for StackOverflow and \num{3000000} for Twitter.
-In order to see how accuracy changed as function of document size, 
-co-occurrence matrices for a range of smaller document sizes were built.
-Both the random permutation and Bayesian models were tested on all document sizes.
+In order to see how accuracy changed as function of corpus size, 
+co-occurrence matrices for a range of smaller corpus sizes were built.
+Both the random permutation and Bayesian models were tested on all corpus sizes.
 
@@
 
-Model accuracy as a function of co-occurrence matrix size for StackOverflow is included in Figure \ref{figContextDocumentSizeSO}.
+Model accuracy as a function of co-occurrence matrix size for StackOverflow is depicted in Figure \ref{figContextDocumentSizeSO}.
 Accuracy improves for both the Bayesian and random permutation models as the size of the documents used to generate the co-occurrence matrix increases.
-However, model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random permutation model, while the frequency technique only begins to plateau.
-The Bayesian model does not plateau and accuracy continues to improve as document size increases.
+However, model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest corpus size for the random permutation model, while the frequency technique only begins to plateau.
+The Bayesian model does not plateau and accuracy continues to improve as corpus size increases.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-freqVsEntropyBySizeSO-crop.pdf}}}
-  \caption{Effect of size of co-occurrence matrix for StackOverflow}
+  \caption{
+    Effect of size of co-occurrence matrix for StackOverflow.
+    Error bars represent the 95\% bootstrapped confidence interval for the mean model accuracy across all runs in dataset.
+  }
   \label{figContextDocumentSizeSO}
@@
 
-It is likely that the random permutation model is starting to saturate when entropy weighting is used for the largest document size for StackOverflow.
+It is likely that the random permutation model is starting to saturate when entropy weighting is used for the largest corpus size for StackOverflow.
 Saturation happens when certain words co-occur with a tag much more often than all of the other words.
 This introduces range effects when the correlation is computed, which ends up driving all of the correlations near one, and consequently decreases discrimination and accuracy.
-Saturation happens when entropy weighting is used because the non-predictive words are only attenuated and not removed, and this attenuation is not enough to overcome the range effects that grow with document size.
+Saturation happens when entropy weighting is used because the non-predictive words are only attenuated and not removed, and this attenuation is not enough to overcome the range effects that grow with corpus size.
 
@@
 
-The overall trend of increasing accuracy with document size is present for Twitter as well.
-Both entropy weighting and frequency filtering in Figure \ref{figContextDocumentSizeT} plateau for the random permutation model with increasing document size.
-Once again, the Bayesian model continues to improve with document size.
+The overall trend of increasing accuracy with corpus size is present for Twitter as well.
+Both entropy weighting and frequency filtering in Figure \ref{figContextDocumentSizeT} plateau for the random permutation model with increasing corpus size.
+Once again, the Bayesian model continues to improve with corpus size.
 
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-freqVsEntropyBySizeT-crop.pdf}}}
-  \caption{Effect of size of co-occurrence matrix for Twitter}
+  \caption{
+    Effect of size of co-occurrence matrix for Twitter.
+    Error bars represent the 95\% bootstrapped confidence interval for the mean.
+  }
   \label{figContextDocumentSizeT}
@@
 
-It is interesting to compare model accuracy between the Bayesian and random permutation model at the smallest and largest document sizes for both StackOverflow and Twitter.
-At smaller document sizes, the random permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as document size increases.
+It is interesting to compare model accuracy between the Bayesian and random permutation model at the smallest and largest corpus sizes for both StackOverflow and Twitter.
+At smaller corpus sizes, the random permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as corpus size increases.
 In other words, the Bayesian model needs larger co-occurrence matrices than the random permutation model to work properly.
@@
 As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random permutation model is actually beneficial, and the compressed random permutation model should be used.
-This noise is beneficial because it helps soften spurious $S_{ji}$ values when the counts for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller document sizes.
+This noise is beneficial because it helps soften spurious $S_{ji}$ values when the counts for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller corpus sizes.
 
 This result has important implications.
-The document size is rarely chosen, and more likely constrained by the number of documents in the dataset being studied.
+The corpus size is rarely chosen, and more likely constrained by the number of documents in the dataset being studied.
 If that dataset is relatively small, it may be better to use the compressed random permutation model than the Bayesian model, and vice-versa if the dataset is large.
 
-However, document size is not the only factor that determines whether the Bayesian model or random permutation model will perform better.
-For example, the document size where the Bayesian model begins to outperform the random permutation model is quite different for StackOverflow and Twitter.
+However, corpus size is not the only factor that determines whether the Bayesian model or random permutation model will perform better.
+For example, the corpus size where the Bayesian model begins to outperform the random permutation model is quite different for StackOverflow and Twitter.
 This is because the number of observations used to generate the co-occurrence matrix is a function of four components:
-document size, number of words in each document, total number of unique tags, and number of tags used per document.
+corpus size, number of words in each document, total number of unique tags, and number of tags used per document.
 The average number of tags used per post for StackOverflow and Twitter were the same order of magnitude (around 3 for StackOverflow, slightly less for Twitter).
@@
 However, the average number of words in a StackOverflow post is much greater than the 140 character limit used for tweets.
-More words per post means that the number of observations per cell in the co-occurrence matrix grows more quickly with document size.
-The Bayesian model is more likely to perform better with datasets like StackOverflow that have these larger document sizes.
+More words per post means that the number of observations per cell in the co-occurrence matrix grows more quickly with corpus size.
+The Bayesian model is more likely to perform better with datasets like StackOverflow that have these larger corpus sizes.
 
@@
 
-The compressed random permutation model should perform better with smaller document sizes, less words per document, more unique tags, and fewer number of tags per post.
+The compressed random permutation model should perform better with smaller corpus sizes, less words per document, more unique tags, and fewer number of tags per post.
 However, since there are so many factors that determine whether the Bayesian or random permutation model will perform better,
 and only two different datasets were examined in this research,
-it is difficult to make any specific recommendations as far as exact transition cut points (e.g., document size, words per document) where either the Bayesian or random permutation model should be used.
+it is difficult to make any specific recommendations as far as exact transition cut points (e.g., corpus size, words per document) where either the Bayesian or random permutation model should be used.
 Also, it is likely that the specific domain being studied interacts with the factors already mentioned to determine which model is better for that domain.
-Nevertheless, if a researcher is working in a domain with limited data (e.g., small document size, large tag space) that produces a sparse co-occurrence matrix,
+Nevertheless, if a researcher is working in a domain with limited data (e.g., small corpus size, large tag space) that produces a sparse co-occurrence matrix,
 then it may be worthwhile to experiment with using a compressed random permutation model instead of the full Bayesian model.
@@
 
-Results for using different levels of compression for the random permutation model for the StackOverflow dataset are included in Figure \ref{figContextDimSO}.
+Results for using different levels of compression for the random permutation model for the StackOverflow dataset are depicted in Figure \ref{figContextDimSO}.
 In the figure, the model name that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
@@
 
-Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy (\numNoZero{.60} to \numNoZero{.55} for the \num{1000000} document dataset size).
+Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy (\numNoZero{.60} to \numNoZero{.55} for the \num{1000000} document corpus size).
 But \num{2048} rows is sufficient for these datasets,
-as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy (\numNoZero{.60} to \numNoZero{.60} for the \num{1000000} document size).
+as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy (\numNoZero{.60} to \numNoZero{.60} for the \num{1000000} corpus size).
 
 Results are similar for Twitter.
-Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100} (\numNoZero{.26} to \numNoZero{.24} for the \num{3000000} document size),
+Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100} (\numNoZero{.26} to \numNoZero{.24} for the \num{3000000} corpus size),
 and little to no impact when increasing from \num{2048} to \num{10000} (\numNoZero{.26} to \numNoZero{.26} at \num{3000000}).
-This results are consistent across all document sizes used to build the co-occurrence matrix.
+This results are consistent across all corpus sizes used to build the co-occurrence matrix.
 
@@
 So it is impressive that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
-This is likely because word use frequency for StackOverflow and Twitter approximately follows Zipf's law \parencite{Stanley2013}.
+This is likely because word use frequency for StackOverflow and Twitter approximately follows Zipf's law.
+\textcite{Stanley2013} showed that frequency of tag use for StackOverflow approximately follows a Zipf's law distribution.
 This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply when moving to words that are used fewer times. 
@@
 That probability was then converted to a log-odds value in the usual way.
+
+For example, say that the model is choosing between a set of tags, and those tags have a mean correlation of .2 with context and a standard deviation of 0.1.
+If a tag in the set has correlation of .2 with context, then approximately half of the observed tags have a correlation lower than this tag.
+So the one-tailed cumulative distribution value for this tag is .5.
+If a tag in the set has a correlation of .3, then approximately half plus .34 (cumulative distribution of normal curve from mean to 1 standard deviation outward) of the observed tags have a lower correlation.
+So the one-tailed cumulative distribution value for this tag is .84.
 
@@
 
-Model accuracy results after using the log-odds transformation are included in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}.
+Model accuracy results after using the log-odds transformation are depicted in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}.
 Using the log-odds transformation technique has a small effect on model accuracy for StackOverflow (\numNoZero{.45} to \numNoZero{.47}) and minimal to no effect on Twitter (\numNoZero{.25} to \numNoZero{.25}).
@@
 and leads to a more natural interpretation of each model term as a log-odds value, it is reasonable to include this step for the random permutation model.
-Future research could explore more improved methods to properly combine the terms.
+Future research could explore additional methods to properly combine the terms.
 
@@
 
-Model accuracy results after incorporating word order into the random permutation model are included in Figure \ref{figContextOrder}.
+Model accuracy results after incorporating word order into the random permutation model are depicted in Figure \ref{figContextOrder}.
 All bar plots that include an order component where ``window'' or ``direction'' is not included in the title use the standard ordering method (no direction or windowing, all position information is included)
@@
 
-Also, the reason why the direction version of ordering is the best predictor of the three ordered components is most likely because order information does not have a large effect,
+Also, the reason why the directional version of ordering is the best predictor of the three ordered components is most likely because order information does not have a large effect,
 so the two representations for words to the left and right of a tag are highly correlated.
-This means that this direction ordered component is essentially composed of all the co-occurrence observations in the orderless component after being randomly sampled and partitioned into two separate representations.
+This means that this directionally-ordered component is essentially composed of all the co-occurrence observations in the orderless component after being randomly sampled and partitioned into two separate representations.
 The co-occurrences in these representations are highly correlated with each other, and highly correlated with the orderless component,
@@
 All of the methodological and architectural tests were done on the Twitter popular-hashtags dataset and the StackOverflow randomly-sampled dataset.
-For these tests, a logistic regression statistical technique is used to determine the optimal weight for each model component.
-A regression is done for each run in the dataset, and since multiple runs were used for the StackOverflow and Twitter datasets, multiple samples of coefficient weights for each model component were computed.
+For these tests, a logistic regression statistical technique was used to determine the optimal weight for each model component.
+A regression was done for each run in the dataset, and since multiple runs were used for the StackOverflow and Twitter datasets, multiple samples of coefficient weights for each model component were computed.
 The variance of the weights for each component show how likely it is that the regression is overfitting the data,
@@
 
-Coefficients for each model component of the full models are included in Figures \ref{figContextCoeffSO} and \ref{figContextCoeffT}.
-Error bars represent the standard deviation of coefficient values for each component.
-The standard deviation is used instead of the confidence interval of the mean so that the plots show the overall stability of each coefficient across the different runs.
+Coefficients for each model component of the full models are depicted in Figures \ref{figContextCoeffSO} and \ref{figContextCoeffT}.
 The range of the distribution of coefficient values within a component is much more narrow than the range of the mean coefficient values between the components of each model.
@@
     (i.e., Bayesian and RP have entropy weighting, RP converts context activation from correlation to log odds).
+    Error bars represent the standard deviation of coefficient values for each component.
+    The standard deviation is used instead of the confidence interval of the mean so that the plots show the overall stability of each coefficient across the different runs.
   }
@@
 Also, an experiment to a user-custom context co-occurrence matrix was attempted on the popular-users dataset.
-This was tried because it seemed theoretically interesting to see if model performance improves when the context co-occurrence matrix is customized to each user.
+This was tried because it could be theoretically interesting to see if model performance improves when the context co-occurrence matrix is customized to each user.
 
@@
 
-The results are included in Figures \ref{figContextPriorneededSO} and \ref{figContextPriorneededT}.
+The results are depicted in Figures \ref{figContextPriorneededSO} and \ref{figContextPriorneededT}.
 For StackOverflow, accuracy remains the same when the prior component is removed for both the Bayesian (\numNoZero{.63} to \numNoZero{.63}) and random permutation (\numNoZero{.63} to \numNoZero{.63}) models.
@@
 
+This raises the question about how much past user behavior is enough for a site like StackOverflow to benefit from adding past user behavior as an additional predictor.
+When the model predicts tags from a random sample of StackOverflow users, adding the prior component to the model does not significantly increase model accuracy compared to using context alone.
+However, when the model predicts tags from a set of users that ask the most questions on the site, adding the prior does significantly increase accuracy.
+For the StackOverflow dataset, a user asks on average 5 questions, and adds 3 tags to a post.
+So the model uses approximately 15 prior observations when computing the prior component from a randomly-sampled on the site.
+For the popular-users dataset used to generate the results in Figure \ref{figPUAddPrior}, a user asked 180 questions on average.
+So for this dataset, the model uses approximately 540 prior observations when computing the prior activation.
+So for the StackOverflow site, the prior component begins to generate unique predictive variance above and beyond the context component somewhere between when a user has generated 15 and 540 prior observations.
+This result will be different for other sites that have collected a different amount of contextual information.
+For sites like StackOverflow where the contextual component is already a strong predictor, it will take a larger number of prior observations before adding the prior component becomes beneficial.
+Alternatively for sites like Twitter where contextual information is much more noisy and not as strong of a predictor, it will take fewer observations before adding the prior becomes beneficial.
+
 \subsection{User-customized Co-occurrence Matrix}
@@
 That is, the default co-occurrence matrix is not customized to each user, so the same matrix is used for all users.
-This is commonly done mainly because there is never enough context by tag co-occurrences for a single user to generate a stable co-occurrence matrix.
+This is commonly done because there is usually not enough context by tag co-occurrences for a single user to generate a stable co-occurrence matrix.
 If there was enough information, accuracy would most likely improve since it is not likely that every user's word by tag co-occurrence profile is exactly the same.
@@
 However, both models are implemented efficiently enough so that all analysis runs and activation computations can be kept in memory.
-64 GB of memory was required to perform the analysis, particularly on the largest Twitter (3 million) and StackOverflow (1 million) document sizes.
+64 GB of memory was required to perform the analysis, particularly on the largest Twitter (3 million) and StackOverflow (1 million) corpus sizes.
 
@@
 Smaller test runs were also created so that model changes could be more rapidly developed.
-These typically use smaller Twitter and StackOverflow document sizes (e.g., only \num{10000} StackOverflow posts) to create the co-occurrence matrix
+These typically use smaller Twitter and StackOverflow corpus sizes (e.g., only \num{10000} StackOverflow posts) to create the co-occurrence matrix
 and a smaller number of posts to test the models on (e.g., 20 instead of 500).
@@
 To run these, the R function is called directly (i.e., not a make target) within an already-running R session.
-As an example, the ``runContext20g1s6'' function takes only a few minutes to complete, runs the models on a document size of \num{1000}, with only a single run per model, and only 20 sampled posts for each run.
+As an example, the ``runContext20g1s6'' function takes only a few minutes to complete, runs the models on a corpus size of \num{1000}, with only a single run per model, and only 20 sampled posts for each run.
 

### 2014-10-16 dissProject/preProcess.R

@@ -1516,7 +1516,7 @@ compareMeanDVBySize <- function(modelVsPredTbl, DV, extras=NULL, figName='', gro
 	plotLineSumTbl(sumTbl, DVName, sprintf('compareMeanDV-%s-%s', deparse(DV), figName), groupCol=groupCol,
 		       extras=c(list(theme(legend.position='top', legend.direction='vertical'),
 				     ylab('Proportion Correct'),
-				     xlab('Dataset Size')),
+				     xlab('Corpus Size')),
 				extras))
 }
 
@@ -1604,30 +1604,30 @@ renameColDVName <- function(tbl) {
 		    'topHashtagPostPriorStdBestDStd', 'Standard prior model at best-fit d=0.8',
 		    'topHashtagPostPriorOL2BestDOL', 'Optimized learning model at best-fit d=0.4',
 		    'topHashtagPostPriorOL2', 'Optimized learning model',
-		    'topHashtagAcrossPriorStdNoffset', 'Standard prior model relaxed across posts w/o offset',
+		    'topHashtagAcrossPriorStdNoffset', 'Standard prior model relaxed across posts without offset',
 		    makeLogregMapping(),
-		    makeStandardMappingBayes('', 'w/ entropy'),
-		    makeStandardMappingBayes('Freq', 'w/ freq'),
-		    makeStandardMappingBayes('Frentropy', 'w/ entropy and freq'),
+		    makeStandardMappingBayes('', 'with entropy'),
+		    makeStandardMappingBayes('Freq', 'with freq'),
+		    makeStandardMappingBayes('Frentropy', 'with entropy and freq'),
 		    makeStandardMappingBayes('Nentropy', ''),
-		    makeStandardMappingBayes('Usercontext', 'w/ entropy and user sji'),
-		    makeStandardMappingBayes('Noffset', 'w/o offset', 'PriorStdNoffset'),
+		    makeStandardMappingBayes('Usercontext', 'with entropy and user sji'),
+		    makeStandardMappingBayes('Noffset', 'without offset', 'PriorStdNoffset'),
 		    makeStandardMapping('', ''),
-		    makeStandardMapping('Entropy', 'w/ entropy'),
-		    makeStandardMapping('Stoplist', 'w/ stoplist'),
-		    makeStandardMapping('Enthyman', 'w/ entropy and log odds'),
-		    makeStandardMapping('Direction', 'w/ entropy and direction'),
[…]

### 2014-10-17 dissProject/preProcess.R
foo

@@ -3691,9 +3691,9 @@ curWS <- function() {
 	runPUserTFollowSji1kTest(regen=F)
 	withProf(runContext20g1s1(regen='useAlreadyLoaded'))
 	setLogLevel(2)
+	modelVsPredTbl = buildTables(file_path_sans_ext(Filter(isPriorRun, list.files(path=getDirModelVsPred()))))
 	modelVsPredTbl = buildTables(file_path_sans_ext(Filter(isContextRun, list.files(path=getDirModelVsPred()))))
 	logregTbl = buildTablesLogreg(file_path_sans_ext(Filter(function(r) isContextRun(r), list.files(path=getDirLogreg()))))
-	modelVsPredTbl = buildTables(file_path_sans_ext(Filter(isPriorRun, list.files(path=getDirModelVsPred()))))
 	modelVsPredTbl = buildTables(file_path_sans_ext(Filter(isPUserRun, list.files(path=getDirModelVsPred()))))
 	modelHashtagsTbls = buildTablesModelHashtags(file_path_sans_ext(Filter(isContextRun, list.files(path=getDirModelHashtags()))))
 

### 2014-10-17 writeups/dissertation/dissertation.tex
foo

@@
 \raggedbottom
+
+% Changing horizontal space between subsubsection title and text, without loading titlesec package.
+% Not loading titlesec because it clashes with apa6 package.
+% Took base \@startsection command from apa6 .cls file, and tweaked horizontal spacing. 
+% http://tex.stackexchange.com/questions/22576/redefining-sectioning-commands
+% http://www.tex.ac.uk/cgi-bin/texfaq2html?label=atsigns
+% http://tex.stackexchange.com/questions/60127/subsubsection-description-and-title-not-separated
+\makeatletter
+\renewcommand{\subsubsection}{%
+  \@startsection
+  {subsubsection}%
+  {3}%
+  {\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-.4em}%
+  {\normalfont\normalsize\bfseries\addperi}}
+\renewcommand{\paragraph}{%
+  \@startsection
+  {paragraph}%
+  {4}%
+  {\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-.4em}%
+  {\normalfont\normalsize\bfseries\itshape\addperi}}
+\makeatother
 
@@
 For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
-\citeAuthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect.
+\citeauthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect.
 
@@
 Further, no recency information was used to compute the global prior, so hashtags used in the recent past were not weighted more heavily. 
-This was most likely not much of an issue for \citeAuthor{Mazzia2009}'s study, since all hashtags were collected within a relatively short period of time (about 3 months).
+This was most likely not much of an issue for \citeauthor{Mazzia2009}'s study, since all hashtags were collected within a relatively short period of time (about 3 months).
 
@@
 \textcite{Stanley2013} weighted each term in the total activation equation Table \ref{tabModACTRModel}) to maximize model accuracy.
-\citeAuthor{Stanley2013} found that the optimal weights for the StackOverflow model for the prior term, title context term, and body context term were \num{.62}, \num{.93}, and \num{1.75} respectively.
+\citeauthor{Stanley2013} found that the optimal weights for the StackOverflow model for the prior term, title context term, and body context term were \num{.62}, \num{.93}, and \num{1.75} respectively.
 
@@
 Activations for context were computed by using five different methods to handle stop words:
-The two removal methods used in \textcite{Sahlgren2008} (i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist),
+The two removal methods used in \textcite{Sahlgren2008}
+(i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist \parencite{Salton1988}),
 the entropy-weighting method used in \textcite{Stanley2013},
@@
 The ideal cutoff was determined by choosing the value where the standard deviation of the counts for each word for each of the tags were all relatively small.
-The plots for the standard deviations across tags that were used to identify the cutoff for StackOverflow and Twitter are included in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT}.
+The plots for the standard deviations across tags that were used to identify the cutoff for StackOverflow and Twitter are shown in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT}.
 
@@
 
-Plots for the three methods of handling stop words for the random permutation model on the StackOverflow dataset are included in Figure \ref{figContextStopRPSO}.
+Plots for the three methods of handling stop words for the random permutation model on the StackOverflow dataset are shown in Figure \ref{figContextStopRPSO}.
 The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
@@
 
-The results comparing the full model to a context-only only model for the StackOverflow popular-users dataset are included in Figure \ref{figPUAddPrior}.
+The results comparing the full model to a context-only only model for the StackOverflow popular-users dataset are depicted in Figure \ref{figPUAddPrior}.
 In this case, past tag history does contain unique variance for accurately predicting tag use.
@@
 
-This raises the question about how much past user behavior is enough for a site like StackOverflow to benefit from adding past user behavior as an additional predictor.
+This raises an important applied question: Exactly how much past user behavior is enough for a site like StackOverflow to benefit from adding past user behavior as an additional predictor.
 When the model predicts tags from a random sample of StackOverflow users, adding the prior component to the model does not significantly increase model accuracy compared to using context alone.
 However, when the model predicts tags from a set of users that ask the most questions on the site, adding the prior does significantly increase accuracy.
-For the StackOverflow dataset, a user asks on average 5 questions, and adds 3 tags to a post.
-So the model uses approximately 15 prior observations when computing the prior component from a randomly-sampled on the site.
+For the StackOverflow dataset, a user asks on average 5 questions and adds 3 tags to a post.
+So the model uses approximately 15 prior observations when computing the prior component for a randomly-sampled user on the site.
 For the popular-users dataset used to generate the results in Figure \ref{figPUAddPrior}, a user asked 180 questions on average.
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiSO-crop.pdf}}}
-  \caption{Model accuracy with user-customized context component for StackOverflow}
+  \caption{
+    Model accuracy with user-customized context component for StackOverflow.
+    A model name with ``user sji'' uses a co-occurrence matrix customized for each user instead of a single global co-occurrence matrix.
+  }
   \label{figPUCustomSO}
@@
 \begin{table}[!ht]
-  \caption{Summary of accuracy for each model and dataset}
+  \caption{
+    Summary of accuracy for each model and dataset.
+    The Bayesian model includes entropy weighting for the context words,
+    the RP model includes entropy weighting and log odds transformation when combining the context and prior components.
+    Both models use a user-customized co-occurrence matrix for the Twitter popular-users dataset, since a global co-occurrence
+    matrix is not available for this dataset.
+  }
   \label{tabOverallResults}
@@
 The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
-These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is a declarative memory retrieval process.
+These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
 However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
@@
 
-Alternatively, perhaps authors are taking advantage of context information that is not currently represented in the model.
+Alternatively, perhaps authors of posts are taking advantage of context information that is not currently represented in the model.
 For example, time of day might be a useful predictor of tag use on Twitter.
@@
 
-And compression for the random permutation model is essentially random.
+As previously noted, compression for the random permutation model is essentially random.
 That is, no SVD is done to identify the most predictive components,
@@
 However, performing a full SVD on a co-occurrence matrix as large as the ones generated for StackOverflow and Twitter is computationally infeasible.
-So, it also seems plausible that this SVD computation is infeasible at the neural level.
-When deriving the LSA approach, \textcite{Landauer1997} explicitly state that they make no claims that the brain is computing an SVD at the neural level;
+Also, when deriving the LSA approach, \textcite{Landauer1997} explicitly state that they make no claims that the brain is computing an SVD at the neural level;
 only that the neural machinery is set up such that the representation makes predictions that are similar to what would have been generated if the matrix decomposition for an SVD would have been computed.
@@
 
-The Bayesian declarative memory model for ACT-R has an attentional weight term $W_{j}$ that can be used to implement how stop words are handled.
+The Bayesian declarative memory model for ACT-R has an attentional weight term $W_{j}$ that can be used to attenuate or remove stop words.
 The random permutation model was also easily modified to have an attentional weight term by attenuating each word's set of ones and negative ones values by that word's attentional weight.
@@
 This would allow, for example, the method of compression for the random permutation vector-based model to be analyzed independent of activation calculation.
-The results would provide additional evidence to help determine the specific representation that the declarative memory system uses to represent co-occurrences,
+The results would help determine the specific representation that the declarative memory system uses to represent co-occurrences,
 and also the specific activation calculation that is carried out by the system.
@@
 
-The tag-recommendation models developed through this research should be immediately relevant to real-world tasks.
+The tag-recommendation models developed through this research should be relevant to real-world tasks.
 Model accuracy is high enough (particularly for StackOverflow) for the developed models to be used as a foundation for tag recommendation systems that are deployed on Twitter or StackOverflow.
-Also, since the developed models are task agnostic, they can potentially be used as tag recommendation systems for other domains.
-Examples of similar tagging tasks are hashtag use on Facebook, flagging (i.e., tagging), Wikipedia articles for content issues or overly opinionated material,
-auto labeling incoming mail and placing it into appropriate folders, and spam filtering.
-Having an accurate and efficient co-occurrence memory model that correctly retrieves relevant tags can be applied to a wide variety of tasks.
+However, the focus for this research was to evaluate the accuracy of two cognitively-plausible declarative memory models, so the models were more constrained than general machine learning models of tagging.
+So it is unlikely that these general models of declarative memory will outperform specific machine learning models that have been optimized for a specific tagging task. 
+Nevertheless, one of the advantages of the general declarative memory models is that they are task agnostic, so it would take minimal time to deploy these models on a new tagging task.
 

### 2014-10-22 writeups/dissertation/dissertation.tex
foo

@@
 
-I would also like to mention a few open-sourced software packages that were heavily leveraged to do this research:
+I would also like to the creators of a few open-sourced software packages that were heavily leveraged to do this research:
 The majority of the modeling and analysis was done with the R statistical computing environment \parencite{RCoreTeam2014} and two important R packages:

### 2014-10-23 writeups/dissertation/dissertation.tex
foo

--- 2014-10-17 writeups/dissertation/dissertation.tex
@@
 
-I would also like to mention a few open-sourced software packages that were heavily leveraged to do this research:
+I would also like to thank the creators of a few open-sourced software packages that were heavily leveraged to do this research:
 The majority of the modeling and analysis was done with the R statistical computing environment \parencite{RCoreTeam2014} and two important R packages:
@@
 These systems can also help newer users by recommending proper tags for their newest content.
-On the StackOverflow site, for example, experts often subscribe to specific hashtags that interest them, and receive a daily digest of posts that are tagged with those hashtags.
-Helping the user properly tag posts on StackOverflow ensures that right community sees the post, which greatly increases the chance that the question on the post will be answered quickly and correctly.
+On the StackOverflow site, for example, experts often subscribe to specific tags that interest them, and receive a daily digest of posts that tagged with those tags.
+Helping the user properly tag posts on StackOverflow ensures that right community sees the post, which greatly increases the chance that the question on the site will be answered quickly and correctly.
 For Twitter, hashtags represent streams of information that are possibly interesting to the user.
@@
 For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store,'' then ``hippie'' would have a fan of two.
-Recall accuracy decrease and time to recall increase as the sum total fan of the person and pair increases.
+Recall accuracy decrease and time to recall increase as the sum total fan of the person and place increases.
 
@@
 There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
-For example, the StackOverflow user in Figure \ref{figPriorHTByTime} has a pronounced band where the same hashtag is chosen repeatedly over time. 
-However there are many instances of other tags being used in parallel to the tag that is producing the band (that is repeatedly being used). 
-The Twitter user has a few clusters where the same hashtag is chosen repeatedly but only for a brief period of time.
-As time progresses, a band of choosing another hashtag is observed, and then the user switches to a new hashtag after that.
-More recent posts for the Twitter user are more varied, and there is no longer a single hashtag that the user chooses repeatedly.
-This helps illustrate how challenging modeling this task can be.
-The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is changing over time.
 
@@
 
+For example, the StackOverflow user in Figure \ref{figPriorHTByTime} has a pronounced band where the same hashtag is chosen repeatedly over time. 
+However there are many instances of other tags being used in parallel to the tag that is producing the band (that is repeatedly being used). 
+The Twitter user has a few clusters where the same hashtag is chosen repeatedly but only for a brief period of time.
+As time progresses, a band of choosing another hashtag is observed, and then the user switches to a new hashtag after that.
+More recent posts for the Twitter user are more varied, and there is no longer a single hashtag that the user chooses repeatedly.
+This helps illustrate how challenging modeling this task can be.
+The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is changing over time.
+
 Although the tag space for the Twitter user is smaller than the StackOverflow user in these plots, this is not a consistent trend.
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
+  \scalebox{.7}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
   \caption{Model tagging profile of a single user on StackOverflow}
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
+  \scalebox{.7}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
   \caption{Model tagging profile of a single user on Twitter}
@@
 This was done by examining model accuracy at a specific decay rate value for each model.
-Accuracy was assessed at each model's previously-determined best-fit values (\num{.43} and \num{.80} for the optimized learning and standard form), and the standard form's default value (\num{0.5}).
+Accuracy was assessed at each model's previously-determined best-fit values (\num{.43} and \num{.80} for the optimized learning and standard form) and the standard form's default value (\num{0.5}).
 Since decay rate values were explored at \num{0.1} increments between the 0 and 1 range, the decay rate values used to examine model accuracy were rounded to the nearest decay rate value that was included in the run.
@@
 ACT-R sets the default decay rate value to \num{0.5}, and the best-fit value for the standard form is higher (\num{0.8}) for these datasets.
-The figure shows that although model accuracy does improve for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
+The figure shows that although model accuracy does slightly improve for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
 that percentage improvement is not as large as the improvement gained from using the standard form compared to the optimized learning form.
@@
 In order to best combine these predictors, a logistic regression statistical technique similar to \textcite{Stanley2013} was used to determine the most optimal weights for each term.
-\textcite{Stanley2013} weighted each term in the total activation equation Table \ref{tabModACTRModel}) to maximize model accuracy.
+\textcite{Stanley2013} weighted each term in the total activation equation (Table \ref{tabModACTRModel}) to maximize model accuracy.
 \citeauthor{Stanley2013} found that the optimal weights for the StackOverflow model for the prior term, title context term, and body context term were \num{.62}, \num{.93}, and \num{1.75} respectively.
 
-The optimal weights for each model component are computed in the same manner as \textcite{Stanley2013}:
+The optimal weights for each model component were computed in the same manner as \textcite{Stanley2013}:
 When a user creates a tag for a post, a retrieval request for the model is made.
@@
 
-Model accuracy is evaluated by rank ordering all recorded tags by model activation using the optimal weights,
+Model accuracy was evaluated by rank ordering all recorded tags by model activation using the optimal weights,
 asking the model to tag the top N posts with a 1 (where N is the total number of recorded tagging instances in the dataset sample),
@@
 but the author is choosing to use only a fixed small set of the top tags for each post, regardless of the activation values of the top tags.
-The author determines the number of tags to use for a post based on competing goals of tagging posts an not overly saturating a post with tags.
+The author determines the number of tags to use for a post based on competing goals of tagging posts and not overly saturating a post with tags.
 For example, the author may be writing a question on StackOverflow about \emph{javascript}, and a large set of related tags are retrieved from declarative memory that are all above the retrieval threshold
@@
 
-Another interesting result was how the Bayesian model actually performed worse when stop words were removed based on frequency compared to no stop word removal.
+Another interesting result is how the Bayesian model actually performed worse when stop words were removed based on frequency compared to no stop word removal.
 This was not the case for the random permutation model in this study, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random permutation model,
@@
 The transition point where accuracy for the Bayesian model outperformed the random permutation model was much higher for StackOverflow than Twitter.
-This is most likely because StackOverflow has a almost \num{100} times more unique tags than the Twitter popular-hashtags dataset.
+This is most likely because StackOverflow has almost \num{100} times more unique tags than the Twitter popular-hashtags dataset.
 The number of words in a StackOverflow post is also much higher than Twitter, which should have lowered the transition point.
@@
 
-Another parameter in the random permutation model is the number of rows used to represent the all of the different words.
+Another parameter in the random permutation model is the number of rows used to represent all of the different words.
 Each word's environment vector is represented by two randomly placed ones and two negative ones across the rows.
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
-  \caption{Effect of compression on random permutation model for StackOverflow}
+  \caption{
+    Effect of compression on random permutation model for StackOverflow.
+    Error bars represent the 95\% bootstrapped confidence interval for the mean.
+  }
   \label{figContextDimSO}
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeT-crop.pdf}}}
-  \caption{Effect of compression on random permutation model for Twitter}
+  \caption{
+    Effect of compression on random permutation model for Twitter.
+    Error bars represent the 95\% bootstrapped confidence interval for the mean.
+  }
   \label{figContextDimT}
@@
 the strength of the model is more likely to be in how efficiently it can represent the co-occurrence matrix rather than its ability to include word order in the representation.
+That is, the strength of the model is in its compressed unordered context component term and not the ordered component, at least for the Twitter dataset studied.
 
@@
 
-This raises an important applied question: Exactly how much past user behavior is enough for a site like StackOverflow to benefit from adding past user behavior as an additional predictor.
+This raises an important applied question: Exactly how much past user behavior is enough for a site like StackOverflow to benefit from adding past user behavior as an additional predictor?
 When the model predicts tags from a random sample of StackOverflow users, adding the prior component to the model does not significantly increase model accuracy compared to using context alone.
@@
 model performance improves only slightly over performance when only the prior is used (\numNoZero{.37} to \numNoZero{.39} for both the Bayesian and random permutation).
+
+\FloatBarrier
 
@@
 
-Given that the storage space required to represent the co-occurrences for the random permutation is dramatically smaller than the uncompressed Bayesian model,
+Given that the storage space required to represent the co-occurrences for the random permutation model is dramatically smaller than the uncompressed Bayesian model,
 and that the compression is lossy and essentially random (i.e., no SVD),

### 2014-11-14 writeups/psycMethods/psycMethods.tex
foo

--- 2014-10-23 writeups/dissertation/dissertation.tex
@@
 \title{Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow}
-\shorttitle{Comparing memory models of tag retrieval}
-%\shorttitle{}
+%\shorttitle{Comparing memory models of tag retrieval}
+\shorttitle{}
 
@@
 	{Houston, TEXAS \par} \vskip 0pt 
-	{\monthnameenglish{} \number\year \par}
+	{December \number\year \par}
       }}
@@
       \@ifundefined{@abstract}{}{%
-	\section{\normalfont\normalsize\abstractname}% BDB
-	\noindent\@abstract\par% BDB
+	\thispagestyle{empty}
+	\section{\normalfont\normalsize \abstractname}% BDB
+	\begingroup
+	\parskip \z@ \parindent \z@ \par
+	\null \vskip .1in \centering
+	\@title
+	\vskip .3in
+	{by \par}
+	\@author \par
+	\vskip .5in
+      \endgroup
+      \@abstract\par% BDB
       \@ifundefined{@keywords}{}{%
@@
       }%
+      \thispagestyle{empty}
       \newpage
@@
     \raggedright%
+  \end{titlepage}
   \setlength{\parindent}{0.4in}%
-    \end{titlepage}}
-  }
+}}
 \makeatother
 
-%\makeDisCover
+\makeDisCover
 
@@
 \begin{document}
+
+\pagenumbering{roman}
 
@@
 
-\section{Acknowledgements}
-
-I would first like to thank my advisor, Mike Byrne, for all of his feedback and suggestions for this dissertation research
-and also his excellent guidance throughout my time as a graduate student.
-I would also like to thank my parents, Mike and Patsy, for the positive influence that they have provided me throughout my life.
-Finally, I would like to thank my brothers Brendan and Justin, sister Kellen, extended family, and friends, for all of their support during my time in the Rice Psychology graduate program.
-
-This work was supported in part by the Gertrude Maurin Fund.
-The resources were used to purchase high-performance computing hardware and the research would not have been possible without this crucial equipment.
-
-I would also like to thank the creators of a few open-sourced software packages that were heavily leveraged to do this research:
-The majority of the modeling and analysis was done with the R statistical computing environment \parencite{RCoreTeam2014} and two important R packages:
-data.table \parencite{Dowle2014} was leveraged to implement the models efficiently, and ggplot2 \parencite{Wickham2009} was used to produce the visualizations.
-
 \newpage
 
-\tableofcontents
-\newpage
-
-\listoftables
-\newpage
-
-\listoffigures
-\newpage
+\pagenumbering{arabic}
 
@@
 
-The core research question motivating this proposal is:
+The core question motivating this research is:
 What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
@@
 
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample1-crop.pdf}}}
-    \hfill
-    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample3-crop.pdf}}}
-    \hfill
-    \vfill
-    \hfill
-    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample4-crop.pdf}}}
-    \hfill
-    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample5-crop.pdf}}}
-    \hfill
-    \caption{Example tweets on twitter.com}
-    \label{figTweetExample}
-  }%
-\end{figure}
-
-Some example tweets on Twitter are included in Figure \ref{figTweetExample}.
-Note how users both add hashtags to the end of a tweet as a summary of the message and emphasize important concepts words in the tweet by turning them into hashtags.
-That is, hashtags can be both intermixed within the words of the tweet and added to the end of the message.
-
 The two memory retrieval models were also tested on tag use for StackOverflow posts.
@@
 Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{stackoverflowPostExample1-crop.pdf}}}
-  \caption{Example post on stackoverflow.com}
-  \label{figSOExample}
-\end{figure}
-
-An example StackOverflow post is included in Figure \ref{figSOExample}.
-The site requires that for each post the user must associate at least one tag with the post, and the average number of tags per post is around 3.
-The author used 4 tags for this post: \emph{javascript}, \emph{html}, \emph{internet-explorer}, and \emph{xhtml}.
 
@@
 However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.
-
-\subsubsection{Lifetime of a Hashtag for a Specific User}
-
-Prior research has examined the growth and decay cycle of hashtag use across users \parencite{Tsur2012}.
-However, much less is known about hashtag lifetime for individual users.
-It may not necessarily follow that a hashtag's life cycle for individual users matches the life cycle across users.
-Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into a specific user's prior likelihood of hashtag use.
-
-\subsubsection{User-Customized Hashtag Prediction}
-
-The end goal of this work is to identify a memory retrieval model that can suggest relevant hashtags to users when they wish to retrieve one.
-These hashtags should be customized to their specific interests and relevant to the content of the post that they are currently creating.
-This will undoubtedly require a combination of two primary model components:
-[1] a user's prior likelihood of choosing a particular hashtag, given their previous hashtag use, and [2] the likelihood that a particular hashtag is related to the context of the post being created.
-
-A user's prior hashtag use will be an essential model component for domains such as Twitter, where the number of possible hashtags is effectively infinite.
-In these domains, the model can utilize a user's hashtag history as a way to prune the infinite space of possible hashtags, and generate a much smaller and more manageable set for prediction.
-However, not only should the model properly take into account a user's prior hashtag use, but it should also generate new hashtag predictions when a user is most likely choosing a hashtag they have never used before.
-The end goal is to have a model that can properly balance the user's tag history with the contextual cues in the post in order to generate an accurate prediction.
 
@@
 
-Using this assumption for large datasets ($S_{ji} = log \frac{p(i|j)}{p(i)}$), the interpretation of the context-adjusted odds ratio becomes much simpler.
-If this ratio for a particular tag and context chunk is greater than one, then the log is positive, which means that this context has been observed more often with this tag than in general.
-If this ratio is less than one, then the log is negative, which provides a negative adjustment of total activation since this context has been observed more often in general than with this particular tag.
-
 \subsubsection{Connection to Pointwise Mutual Information}
@@
 Most ACT-R models simply weight each contextual element equally (e.g., all words in a StackOverflow post have equal weight), so each word's attentional weight is simply normalized by the number of items in context.
-
-\subsubsection{Scaling the Equations}
-
-The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
-As an applied concern, it is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
-Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
-If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.
-
-However, scale is also a theoretical concern.
-The declarative memory system is tasked to quickly return a single chunk of information that is stored alongside millions to billions of other unique chunks given only the current context and past history of use.
-To do this task, it seems plausible that a model of declarative memory must use a representation that produces accurate retrievals even when large amounts of observations are included.
-Further, it seems plausible that the time to compute an activation value by a computer is correlated with the time to compute an activation value by the brain.
-So any model of declarative memory should also be reasonably computationally efficient even with large co-occurrence datasets.
-
-ACT-R's strength of association ($S_{ji}$) and the equivalent pointwise mutual information index ($\mathit{PMI}_{ji}$) have been shown to scale well for large corpora.
-\textcite{Douglass2010} implemented the large-scale version of $S_{ji}$ in Erlang.
-The wall-clock time required for a declarative memory retrieval request scaled linearly with the size of declarative memory with this implementation.
-Retrieval times were around one second with the largest declarative memory that contained over one million associations (co-occurrence pairs).
-
-The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} was developed to predict user navigation of web-based hyperlinks (e.g., choosing between a set of results from a search query).
-The theory uses the large-scale version of $S_{ji}$ to measure the relatedness between a user's current goals and the current context.
-It was shown to scale well when working with a large-scale search query database for internet search.
-
-As one of the largest tests for the PMI index, \textcite{Farahat2004} generated the co-occurrence matrix from a 118GB local database of 10 million pages of the Stanford WebBase Project.
-\citeauthor{Farahat2004} compared strength of association values using the PMI index and similarity ratings derived from human raters.
-Even at this large dataset size, the ACT-R declarative memory theory still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.
 
@@
 
-\subsubsection{Retrieval Process}
-
-Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
-Decoding takes the memory vector for context and decodes it back into an environment vector.
-The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
-
-Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
-For example, assume that a user has written the word ``zend'' in a post, and she wants to assign a tag for the post.
-``zend'' has co-occurred with the tag \emph{PHP} many times previously, so the unordered memory vector for ``PHP'' ($c_{\mathit{PHP}}$) contains the unordered environment vector for \emph{zend} ($e_{zend}$).
-To determine the most correlated memory vector with context, a memory vector is created from the current context ($e_{zend}$).
-The cosine between that context memory vector and the memory vector for \emph{PHP} will be high, since the \emph{PHP} memory vector ($c_{\mathit{PHP}}$) often contains the context environment vector ($e_{zend}$).
-Since the cosine is high, the tag \emph{PHP} will most likely be returned as the vector that resonates highest with the current context.
-
-The resonance and decoding processes are inversions of each other.
-Usually both resonance and decoding will return similar rank orderings of most likely chunks given context.
-\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
-
 \subsubsection{Addressing Word Order}
@@
 This allows for retrieval requests such as ``playing \#$\Phi$,'' ``just read \#$\Phi$,'' and ``vector-based $\Phi$ systems.''
-
-For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
-Within this corpus, the phrase ``martin luther'' often occurred.
-When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated by convolving a 1-back location vector with luther ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
-Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin with and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
-The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.
-
-Using decoding, the word-order memory vector for martin is deconvolved to the right ($o_{martin} \circledast \Phi$) which returns the environment vector for luther ($e_{luther}$).
-Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which correlates highest with the word-order memory vector for luther ($o_{luther}$).
-Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$.''
-
-New vectors generated by circular convolution are uncorrelated with all other environment vectors and have the same length.
-This allows the word order information and unordered context information for a chunk to be aggregated and represented in the same memory vector.
-Quite simply, the different types of memory vectors can be summed to create a single memory vector representation for each chunk (i.e., $m_{i} = c_{i} + o_{i}$).
 
@@
 
-This is a counterintuitive result.
-It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
-This does not seem to be the case in testing however \parencites{Sahlgren2008,Jones2007}.
-Nonetheless, \textcite{Kanerva2000} suggested to perform an SVD reduction technique on a compressed random permutation vector.
-This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.
-
 \subsection{Comparison of Vector-Based Models and ACT-R}
@@
 For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
-\citeauthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect.
-
-The fan effect describes a behavior of human memory where memory items are more difficult to retrieve if they have occurred in a larger number of different contexts.
-\textcite{Rutledge2008} used \textcite{Anderson1974}'s classic experimental paradigm to test the fan effect.
-In the task, participants are exposed to a set of person and place pairs (e.g., ``the hippie is in the park'').
-They are later asked to recall if a particular person and place pair was presented to them before.
-The main manipulation is the ``fan'' of the person and place for each pair.
-The fan is defined as the number of different contexts that each item in the pair is presented in.
-For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store,'' then ``hippie'' would have a fan of two.
-Recall accuracy decrease and time to recall increase as the sum total fan of the person and place increases.
-
-Both the ACT-R declarative memory theory and vector-based models produce the fan effect due to the effect that fan has on the co-occurrence matrix.
-As a particular word (e.g., ``hippie'') is seen in more and more contexts, the co-occurrence count of the word across all contexts is spread across multiple contexts.
-This means that when a retrieval for a context given the word occurs, multiple contexts have relatively high activation.
-This makes it more difficult to test if a signal (e.g., if a word appears in a specific context) is higher than the background noise (e.g., all other contexts that the word has appeared in),
-which makes the retrieval more uncertain.
-
-Both vector-based and Bayesian models produce the fan effect since they both are based on co-occurrences.
-However, vector-based models do further processing on that matrix, reduce its dimensionality, and use a correlation instead of a log odds value to compute the activation.
-Nonetheless, vector-based models still produce the fan effect because at their core they are still computing activations based on co-occurrences.
 
@@
 
-\subsection{Recommendation Systems}
-
-A model that predicts a user-chosen hashtag is a specific type of a recommendation system.
-General recommendation systems have grown in need with the growth of the web.
-Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
-Recommendation systems are used to tailor the information presented to what is most relevant to the user \parencite{Pazzani2007}.
-
-\subsubsection{The Netflix Prize}
-
-The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
-Netflix is an online subscription service where users can stream television episodes and movies.
-Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
-In order to provide a high quality user experience on the site, Netflix created a content recommendation service.
-This tool recommends to the user movies and television episodes that most likely match their specific interests.
-
-Netflix wanted to improve the accuracy of their recommendation system, so they held a competition where teams could try to improve on the accuracy of the service.
-The winning team utilized a statistical model that performed a type of factor analysis to identify important dimensions in the data (e.g., drama, action, cult classic) and place values for each user along each dimension.
-This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
-By using a factor analysis technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.
-
-\subsubsection{Recommending Followers on Twitter}
-
-Recommendation systems based on the content of user-created posts have been successful in recommending people on Twitter for a user to follow.
-\textcite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
-The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
-The profiles were created by generating a word frequency matrix of words by (user + followers + followees), and then providing that matrix as input into the Lucene text search engine framework.
-
-Lucene \parencite{McCandless2010} is a commonly-used open source software tool that measures the similarity between a query and candidate documents in a database.
-It uses the term frequency-inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
-This measure is the product of two terms: the frequency count of the word in the document and the log inverse of the ratio of the number of documents that the word appeared in to the total number of documents.
-The term frequency component measures the strength of association between a word and a document.
-The inverse document frequency component measures how much information is contained within a specific word, independent of document.
-
-When \citeauthor{Hannon2010}'s system recommended a set of individuals for a user to follow, it collected all of the words used in posts from that user, constructed a search query, and provided that query to Lucene.
-Lucene then returned the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
-Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user follows at 25\% accuracy.
-
-\subsubsection{Hashtags for Recommendation Systems}
-
-\textcite{Efron2010} showed how hashtags might be used as a type of recommendation system for Twitter.
-Users constructed a query of text that represented a search for a topic that they were interested in.
-The system then retrieved the most likely hashtags that were associated with that search query.
-The model is based on building a word co-occurrence matrix of words by hashtags, and then uses that matrix to assign a similarity score for each hashtag when provided with a search query.
-Participants graded how relevant the hashtags returned were to each search query.
-The results suggest that hashtags contain useful information that help classify the gist of a post. 
-Although this model does not predict the specific hashtags used when composing a tweet, a model like this could certainly be tailored for that purpose.
-
 \subsection{Hashtag Prediction}
 
-Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
+Models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
 Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
@@
 Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
-Recommendation models for StackOverflow, Google+, and Twitter will be described in more detail to follow.
+Recommendation models for StackOverflow and Twitter will be described in more detail to follow.
 
@@
 
-It is interesting that in both studies the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
-This suggests that more cognitively-plausible memory retrieval models may outperform more general machine learning approaches in this particular environment
-(i.e., when predicting the user-chosen tags for human-created content on social media sites).
-
-\subsubsection{StackExchange}
-
-A tag suggestion tool has been deployed for the \url{tex.stackexchange.org} \parencite{LatexTags2013} and \url{superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
-Latex, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
-For the two sites where a tag suggestion tool has been deployed, once a user has created a question, suggested tags appear just below the text field where the user types the tags to associate with the question.
-The post author is not forced to use the suggested tags, and is free to choose tags from them or tags not represented in the suggested set.
-An example of the interface for a created post on the SuperUser site is included in Figure \ref{figSuperUserSuggestion}.
-
-\begin{figure}[!htbp]
-  \scalebox{.9}{\resizebox{\linewidth}{!}{\includegraphics{superUserTagSuggestion-crop.pdf}}}
-  \caption{Example of tag suggestions when creating a post on the SuperUser site}
-  \label{figSuperUserSuggestion}
-\end{figure}
-
-The exact tag-recommendation algorithm used for the Latex and SuperUser sites is not public.
-After exploratory testing, StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags.
-The Latex and SuperUser sites are not nearly as popular as StackOverflow currently.
-The sites have roughly \num{50000} and \num{189000} questions respectively compared to \num{5000000} questions on StackOverflow.
-
-So it stands to reason that the algorithm is relatively complex, and there could be possible scaling and efficiency concerns with the algorithm's implementation, so it has not yet been deployed to StackOverflow.
-Alternatively, it is also possible that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
-Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial to the user and has started to add them to their question and answer sites.
-
-\subsubsection{Google+}
-
-The hashtag recommendation system deployed for the Google+ social media site is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
-The user interface for this recommendation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.
-
-\begin{figure}[!htbp]
-  \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample-crop.pdf}}}
-  \caption{Example post on Google+}
-  \label{figGoogle+Hashtags}
-\end{figure}
-
-The workflow and user interface provided for recommending hashtags for Google+ is similar to how a hashtag recommendation system for Twitter might look.
-When generating content on Google+, the system automatically associates the most likely relevant hashtags with that post.
-For the example in Figure \ref{figGoogle+Hashtags}, the system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-created post.
-The user is free to remove any or all of the tags that the system has associated with the post, and can also use tags that were not included in the recommended set.
-
-Google+'s tag-recommendation algorithm is not public, but it appears to be rather sophisticated.
-Looking at the example post in Figure \ref{figGoogle+Hashtags}, \emph{\#BusterPosey} is suggested without having that specific word in the post.
-Also \textcite{GoogleKeynote2013} mentioned that the algorithm uses Google's sophisticated image recognition software to suggest hashtags for images that are included in the post (e.g., \emph{\#EiffelTower}),
-even when the post does not have any keywords to provide additional context.
-
-Google is apparently highly confident in the accuracy of their system.
-On the Latex site, the user interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
-For Google+, the user interface automatically tags a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
-If tag accuracy is high enough, the Google+ technique is certainly favorable since the user does not have to spend time figuring out which tags best represent this post.
-However, if the model makes even just a few errors, users may become annoyed and frustrated with the recommendation system since they will have to explicitly correct it every time they generate a post.
-
 \subsubsection{Twitter}
@@
 
-A topics-based approach was used by \textcite{Godin2013}, where the model predicted the most likely set of topics that were associated with a tweet, and tested if these topics matched the tweet by using human raters.
-The latent topics for the collection of tweets were generated using a Latent Dirichlet Allocation (LDA) statistical methodology.
-This is similar to a factor analysis approach in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance.
-Human graders rated how well the models generated topics for a tweet related to the content in the tweet.
-The LDA model performed well, and generated relevant topics.
-When the model produced five most-likely hashtags for each tweet, 80\% of the time at least one of those hashtags was suitable (as determined by human raters).
-
-However, a topic-generating model is not performing the same task as a hashtag-prediction model.
-The authors argued that these general topics could be used to categorize the post, in addition to the hashtags already used by the author in the post.
-It is unclear though if a user would want to label their tweet with these general topics.
-Perhaps they would rather use more specific hashtags to be more precise in how they label the tweet and how that tweet is associated with tweets within the rest of the community.
-In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
-Now it is certainly the case that the topics generated by an LDA model may be similar to the hashtags used in the post.
-However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
-On the other hand, a topics-based approach is more general than a hashtags-based approach, since topics can be suggested for posts that do not use hashtags (which is quite common across Twitter).
-
 \subsubsection{State of Research on StackOverflow and Twitter}
@@
 Rather their models, much like \textcite{Mazzia2009}'s Twitter model, uses the aggregate tag history across all users as the model's prior activation for each tag.
-
-As some initial research, I have developed and tested a user-customized model for StackOverflow by modifying the model in \textcite{Stanley2013}.
-To incorporate the user's prior tag history into the model, the base-level activation term ($B_{i}$) was modified from the original equation in ACT-R.
-$B_{i}$ is based on log-odds occurrence, which in turn is based on frequency of tag use, or probability of tag use $p_{i}$.
-To incorporate a user's prior, a weighted average for $p_{i}$ was used,
-where the probability of using tag $i$ ranges from the global probability ($pGlobal_{i}$) to the user's specific probability ($pUser_{i}$) as the user's total tag count increases.
-
-The global probability is the same $p_{i}$ used in \textcite{Stanley2013}'s study.
-The user's specific probability is computed the same way that the global probability is computed (i.e., the log odds that the tag occurred relative to the occurrence of all other tags).
-However, each user's probability is constrained to the set of tags that the specific user has used in the past. 
-When a user has no past tag history, the global prior probability is used.
-Once a user has provided at least five tags, that user's specific prior probability is used.
-Between zero and five tags, both terms are used but the weight of the global prior relative to the user prior decreases as the number of tags increases.
-
-Both linear decay and exponential decay transition functions were tried.
-Using either transition function, overall model accuracy improved by 5\% after incorporating the user's prior tag history into the model.
-
-This research suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
-It also seems plausible that model accuracy can improve an even greater amount by using the time-dependent form of the base-level component in Table \ref{tabACTRModel},
-in order to account for the fact that a user's tagging preferences change over time.
 
@@
 So tags used for the post were converted to their root tag when possible.
-
-\subsection{Handling Stop Words}
-
-``Stop words'' (e.g., ``the,'' ``a,'' ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
-It is typically recommended \parencite{Bird2009} to remove stop words when building the co-occurrence matrix.
-Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in the text.
-Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.
-
-Stop words are usually removed by either filtering based on a predetermined list or filtering based on the high-frequency words that occur in the dataset.
-\textcite{Sahlgren2008} tried both methods to remove stop words for the random permutation vector-based model, and found that the data-driven approach
-(i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a predetermined list.
-
-\textcite{Jones2007} also filtered the stop words for the circular convolution model,
-but also mentioned that an entropy weighting technique \parencite{Dumais1991} has been used previously to attenuate (rather than remove) stop words.
-This data-driven technique identifies stop words by calculating the discriminating power (i.e., entropy) of each word to predict any tag.
-Each word is then weighted (not removed) based on their discriminating power.
-
-\textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset.
-The result maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
-It also seems less arbitrary than explicitly identifying a list of stop words to remove from the analysis.
-
-All three stop word techniques (i.e., predetermined list, high-frequency filtering, and entropy weighting) were explored for this analysis.
 
@@
 
-\subsubsection{Visualizing Tag Use}
-
-To better illustrate the dynamics of each user's tagging behavior,
-profile plots for a single user for both StackOverflow and Twitter are included in Figure \ref{figPriorHTByTime}.
-The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
-What is interesting is the large tag space for these users.
-There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[][StackOverflow]{\includegraphics[width=8cm]{HTByTime-520957-crop.pdf}}
-    \hfill
-    \subfigure[][Twitter]{\includegraphics[width=8cm]{HTByTime-fashionista_com-crop.pdf}}
-    \caption{
-      Tagging profile for a single user.
-      The x-axis represents the specific hashtag a user chooses for a post and time progresses vertically.
-      A point in the plot represents when a user chooses a specific hashtag for a specific post.
-      A user's chosen hashtag for their oldest post that contains a hashtag is the lowest point on the plot.
-    }
-    \label{figPriorHTByTime}
-  }%
-\end{figure}
-
-For example, the StackOverflow user in Figure \ref{figPriorHTByTime} has a pronounced band where the same hashtag is chosen repeatedly over time. 
-However there are many instances of other tags being used in parallel to the tag that is producing the band (that is repeatedly being used). 
-The Twitter user has a few clusters where the same hashtag is chosen repeatedly but only for a brief period of time.
-As time progresses, a band of choosing another hashtag is observed, and then the user switches to a new hashtag after that.
-More recent posts for the Twitter user are more varied, and there is no longer a single hashtag that the user chooses repeatedly.
-This helps illustrate how challenging modeling this task can be.
-The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is changing over time.
-
-Although the tag space for the Twitter user is smaller than the StackOverflow user in these plots, this is not a consistent trend.
-These plots were chosen from the set to also show that the amount of tag variation for individual users also varies: some use many tags and some use just a few.
-To accurately model human performance in this domain, the model must be able to make correct predictions across all variations of user tagging behavior.
-
-\subsubsection{Visualizing Model Performance}
-
-As a first method for understanding model performance, profile plots of model's predicted tags for individual users were overlaid on that user's actual tag use.
-These plots for StackOverflow and Twitter are included in Figures \ref{figPriorHTMByTimeSO} and \ref{figPriorHTMByTimeT} respectively.
-These are the same plots shown in Figure \ref{figPriorHTByTime} with model predictions overlaid in red.
-
-\begin{figure}[!htbp]
-  \scalebox{.7}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
-  \caption{Model tagging profile of a single user on StackOverflow}
-  \label{figPriorHTMByTimeSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.7}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
-  \caption{Model tagging profile of a single user on Twitter}
-  \label{figPriorHTMByTimeT}
-\end{figure}
-
-Also shown in these figures is how model predictions change as the decay rate parameter changes.
-With lower decay rate values, the model is less likely to switch from tag to tag for each post.
-This makes sense because lower decay rate values lead to the model weighting frequency more than recency in the activation calculation.
-These plots also suggest that model accuracy is best when a blend of frequency and recency is used (i.e., a decay rate between 0 and 20).
-This will be more formally tested when model accuracy across decay rate values is analyzed across all users in each dataset slice and all dataset slices.
-
 \subsubsection{Model Performance Across Decay Rate Values}
 
-Moving up one level of analysis, model accuracy for all users in a single dataset slice was analyzed.
+Looking first at only a single dataset slice, model accuracy for all users in that slice was analyzed.
 A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDsStd} and \ref{figPriorTwitterSliceDsStd} respectively.
@@
   }%
+
 \end{figure}
@@
 Chapter 5 will examine how model accuracy is influenced by two methodological concerns (adding an offset term and examining the effect of dataset subset for Twitter)
-and five architectural concerns
-(method of handling stop words, influence of dataset size, amount of compression used for the vector-based models, method to combine the prior and context for the vector-based models, and word order).
+and four architectural concerns
+(method of handling stop words, influence of dataset size, method to combine prior and context, and word order).
 The two methodological concerns were examined to ensure that the technique to calibrate model parameters was accurate and also that results for Twitter were not specific to the subset of data collected.
 
-The five architectural concerns were examined for the following reasons:
+The four architectural concerns were examined for the following reasons:
 During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.
@@
 This has not yet been possible since the two models have never been compared on the same dataset within the same study.
-
-\textcite{Sahlgren2008} showed how different levels of compression in the random-permutation model influence accuracy, so this was examined for three reasons:
-to replicate the findings,
-as an additional check to ensure that the random-permutation model was implemented correctly, and also to see if the same levels of compression are valid for the much larger datasets used in this research.
 
@@
 
-The method of handling stop words has been shown to influence model accuracy \parencite{Sahlgren2008,Stanley2013}, so this was the first architectural concern that was examined.
+``Stop words'' (e.g., ``the,'' ``a,'' ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
+It is typically recommended \parencite{Bird2009} to remove stop words when building the co-occurrence matrix.
+Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in the text.
+Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.
+
+The method of handling stop words has been shown to influence model accuracy \parencite{Sahlgren2008,Stanley2013,Jones2007}, so this was the first architectural concern that was examined.
 \textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
@@
 However, \citeauthor{Sahlgren2008} did not try to weight the stop words and only tried methods to remove them.
-So alongside looking at frequency weighting, previously generated, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on entropy was also explored.
+
+\textcite{Jones2007} also filtered the stop words for the circular convolution model,
+but also mentioned that an entropy weighting technique \parencite{Dumais1991} has been used previously to attenuate (rather than remove) stop words.
+This data-driven technique identifies stop words by calculating the discriminating power (i.e., entropy) of each word to predict any tag.
+Each word is then weighted (not removed) based on their discriminating power.
+
+\textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset.
+The result maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
+It also seems less arbitrary than explicitly identifying a list of stop words to remove from the analysis.
+
+All three stop word techniques (i.e., predetermined list, high-frequency filtering, and \textcite{Stanley2013}'s entropy weighting) were explored for this analysis.
 
@@
 
-\subsection{Amount of Compression for Random Permutation}
-
-Another parameter in the random permutation model is the number of rows used to represent all of the different words.
-Each word's environment vector is represented by two randomly placed ones and two negative ones across the rows.
-All that is required is that the four positions and signs of each of these environment vectors is unique.
-But this means multiple environment vectors can overlap when looking at a single row.
-This is a lossy form of compression, and allows the number of rows to represent the words to be much less than the total number of words.
-As the number of rows decreases, the amount of cross talk between the environment vectors increases, compression increases, and the amount of information stored decreases.
-As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.
-
-\subsubsection{Method}
-
-Three different row-size values were tested for the random permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
-A low value (\num{100}), a more standard value (\num{2048}), and a high value (\num{10000}) of rows were tested.
-A separate co-occurrence representation was built for each of these row values.
-Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
-Accuracy measurements were the same as were used for previous runs of these datasets: mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.
-
-\subsubsection{Results}
-
-Results for using different levels of compression for the random permutation model for the StackOverflow dataset are depicted in Figure \ref{figContextDimSO}.
-In the figure, the model name that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
-This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random permutation model for all other results.
-The plot is also broken out for different size datasets used to build the co-occurrence matrix.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
-  \caption{
-    Effect of compression on random permutation model for StackOverflow.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean.
-  }
-  \label{figContextDimSO}
-\end{figure}
-
-Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy (\numNoZero{.60} to \numNoZero{.55} for the \num{1000000} document corpus size).
-But \num{2048} rows is sufficient for these datasets,
-as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy (\numNoZero{.60} to \numNoZero{.60} for the \num{1000000} corpus size).
-
-Results are similar for Twitter.
-Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100} (\numNoZero{.26} to \numNoZero{.24} for the \num{3000000} corpus size),
-and little to no impact when increasing from \num{2048} to \num{10000} (\numNoZero{.26} to \numNoZero{.26} at \num{3000000}).
-This results are consistent across all corpus sizes used to build the co-occurrence matrix.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeT-crop.pdf}}}
-  \caption{
-    Effect of compression on random permutation model for Twitter.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean.
-  }
-  \label{figContextDimT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-The fact that the random permutation model only needs around \num{2048} rows to represent all of the words in the Twitter and StackOverflow datasets is impressive.
-Twitter has around \num{2383000} unique words for the \num{3000000} tweets in each of the popular-users datasets.
-StackOverflow has \num{5848000} unique words for the \num{1000000} posts used to generate the co-occurrence matrix.
-So it is impressive that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
-This is likely because word use frequency for StackOverflow and Twitter approximately follows Zipf's law.
-\textcite{Stanley2013} showed that frequency of tag use for StackOverflow approximately follows a Zipf's law distribution.
-This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply when moving to words that are used fewer times. 
-Since these words have little to no ability to influence predictions, it makes sense that they can be safely removed from the dataset.
-
-However, it should be pointed out that the way the random permutation model compresses the words is completely random.
-It is not an LSA approach where the few ones and negative ones (i.e., the factors) for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
-The environment vector for each word is randomly selected, and no SVD is performed to choose these in a more discriminating manner.
-So it is that more impressive that even with a randomly-chosen form of compression, only \num{2048} rows are needed to represent all of the words in these datasets.
-
 \subsection{Method to Combine Prior and Context}
@@
 or simply due to the fact that the amount of information collected for each of the terms changes between the domains (less information leads to lower weighting in general).
-
-\clearpage
-\section{Chapter 6: Combining Predictors on the Popular-Users Dataset}
-
-Chapter 4 isolated the prior component of the model and analyzed accuracy on the popular-users dataset.
-Chapter 5 added the context components to the models and analyzed accuracy on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.
-Chapter 6 shows the results when the combined prior and context models are tested across all of the datasets.
-Looking at model behavior across the datasets provides a way to show how the importance of the prior component relative to the context component can change depending on specific properties in the dataset that is studied.
-Also, an experiment to a user-custom context co-occurrence matrix was attempted on the popular-users dataset.
-This was tried because it could be theoretically interesting to see if model performance improves when the context co-occurrence matrix is customized to each user.
-
-\subsection{Impact of Prior vs. Context}
-
-Adding a user's prior tag history as an additional component has been shown to improve performance, but is this always the case?
-If enough context information has been collected, maybe a context-only model can perform just as well as a combined context and prior model.
-If this were the case, then it may be less important that the random permutation model does not include a prior term by default.
-
-\subsubsection{Method}
-
-Full models were tested against context-only models for both sites. 
-The four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset were used for the test.
-
-\subsubsection{Results}
-
-The results are depicted in Figures \ref{figContextPriorneededSO} and \ref{figContextPriorneededT}.
-For StackOverflow, accuracy remains the same when the prior component is removed for both the Bayesian (\numNoZero{.63} to \numNoZero{.63}) and random permutation (\numNoZero{.63} to \numNoZero{.63}) models.
-Accuracy does dramatically decrease for Twitter when the prior is removed, however (\numNoZero{.32} to \numNoZero{.14} for Bayesian, \numNoZero{.25} to \numNoZero{.12} for random permutation).
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
-  \caption{Model accuracy when prior is removed for StackOverflow randomly-sampled dataset}
-  \label{figContextPriorneededSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddT-crop.pdf}}}
-  \caption{Model accuracy when prior is removed for Twitter popular-hashtags dataset}
-  \label{figContextPriorneededT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-These results show that prior information is not always needed when a large amount of context is available and not much information has been collected about each user's past history.
-The StackOverflow randomly-sampled dataset selects posts across the entire dataset, regardless of each user's popularity or tag history.
-Across the dataset, each user only asks around five questions on average, so there is not a large amount of past tag history available for the prior component to use,
-which is why this component is not a strong predictor in this case.
-
-However, can a context-only model perform just as well as a combined context and prior model when users with the largest amount of StackOverflow history are tested?
-If each user's tagging history is a useful predictor for StackOverflow,
-then the prior component should at least generate unique predictive variance in the case where users with large amounts of history are tested.
-
-\subsubsection{Method}
-
-The same StackOverflow popular-users dataset used to test the prior component in isolation was used to test the prior component when context was added.
-However, since computing activations for the context component take an order of magnitude more wall-clock time than computing the prior component, not all observations could be included for the test.
-To reduce the observations, the most recent 50 posts from each user in the datasets were selected.
-The total number of users sampled was also reduced to 40 users for StackOverflow and 80 users for Twitter.
-Using the most recent observations for each user ensures that the observations tested have acquired the greatest amount of past tagging history in the datasets.
-So if a context-only model performs just as well as the full model in this case, it would be surprising.
-
-\subsubsection{Results}
-
-The results comparing the full model to a context-only only model for the StackOverflow popular-users dataset are depicted in Figure \ref{figPUAddPrior}.
-In this case, past tag history does contain unique variance for accurately predicting tag use.
-Accuracy improves for both the Bayesian (\numNoZero{.51} to \numNoZero{.63}) and random permutation (\numNoZero{.47} to \numNoZero{.58}) models
-when the prior model component is combined with context compared to context alone.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddPUserSO-crop.pdf}}}
-  \caption{Model accuracy when prior is removed for StackOverflow popular-users dataset}
-  \label{figPUAddPrior}
-\end{figure}
-
-\subsubsection{Discussion}
-
-This shows that incorporating past tagging history is important even for a site like StackOverflow where a large amount of context information has been collected.
-However, for this type of site it only becomes important when a reasonably large amount of a user's past tagging history has been collected.
-For a randomly sampled user on StackOverflow, a context-only model captures the majority of the predictive variance.
-
-This raises an important applied question: Exactly how much past user behavior is enough for a site like StackOverflow to benefit from adding past user behavior as an additional predictor?
-When the model predicts tags from a random sample of StackOverflow users, adding the prior component to the model does not significantly increase model accuracy compared to using context alone.
-However, when the model predicts tags from a set of users that ask the most questions on the site, adding the prior does significantly increase accuracy.
-For the StackOverflow dataset, a user asks on average 5 questions and adds 3 tags to a post.
-So the model uses approximately 15 prior observations when computing the prior component for a randomly-sampled user on the site.
-For the popular-users dataset used to generate the results in Figure \ref{figPUAddPrior}, a user asked 180 questions on average.
-So for this dataset, the model uses approximately 540 prior observations when computing the prior activation.
-So for the StackOverflow site, the prior component begins to generate unique predictive variance above and beyond the context component somewhere between when a user has generated 15 and 540 prior observations.
-This result will be different for other sites that have collected a different amount of contextual information.
-For sites like StackOverflow where the contextual component is already a strong predictor, it will take a larger number of prior observations before adding the prior component becomes beneficial.
-Alternatively for sites like Twitter where contextual information is much more noisy and not as strong of a predictor, it will take fewer observations before adding the prior becomes beneficial.
-
-\subsection{User-customized Co-occurrence Matrix}
-
-The co-occurrence matrix for the context component of both models has been built by aggregating across users and collecting all instances of context words and associated tags.
-That is, the default co-occurrence matrix is not customized to each user, so the same matrix is used for all users.
-This is commonly done because there is usually not enough context by tag co-occurrences for a single user to generate a stable co-occurrence matrix.
-If there was enough information, accuracy would most likely improve since it is not likely that every user's word by tag co-occurrence profile is exactly the same.
-However, since several hundred posts for each user in the popular-users datasets are available, an analysis was tried where user-custom co-occurrence matrices were used.
-
-\subsubsection{Method}
-
-Separate co-occurrence matrices were built for each user in the sampled StackOverflow and Twitter popular-users dataset.
-The same sampling method used to test the impact of the prior component compared to the context component for the popular-users datasets was used,
-and the co-occurrence matrices were built from the oldest posts that were not included in the testing set.
-So for each user, the co-occurrence matrix was created from their oldest posts, up to (and not including) their 50 newer posts that were used to test model accuracy.
-
-\subsubsection{Results}
-
-Model accuracy when using a user-customized co-occurrence matrix for StackOverflow and Twitter is shown in Figures \ref{figPUCustomSO} and \ref{figPUCustomT}.
-The results compare accuracy for models that use a global co-occurrence and user-specific co-occurrence matrix.
-For StackOverflow, model accuracy for the user-specific co-occurrence matrix (labeled as ``user sji'') can be compared against the default global co-occurrence matrix.
-Accuracy decreases dramatically when a user-specific co-occurrence matrix is used for both the Bayesian (\numNoZero{.64} to \numNoZero{.35}) and random permutation (\numNoZero{.58} to \numNoZero{.36}) models.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiSO-crop.pdf}}}
-  \caption{
-    Model accuracy with user-customized context component for StackOverflow.
-    A model name with ``user sji'' uses a co-occurrence matrix customized for each user instead of a single global co-occurrence matrix.
-  }
-  \label{figPUCustomSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiT-crop.pdf}}}
-  \caption{Model accuracy with user-customized context component for Twitter}
-  \label{figPUCustomT}
-\end{figure}
-
-The results are similar for Twitter.
-However, the entire dataset is not available for Twitter, so a global co-occurrence matrix condition can not be tested,
-since the popular-users dataset samples across a wide range of hashtags, and due to Twitter's API restrictions it would be infeasible to collect all posts that use each of the hashtags in the dataset.
-But user-specific co-occurrence matrices can be built, since the popular-users dataset does contain all context and hashtag observations for each user.
-
-Model accuracy when only the single user-customized context predictor is used is relatively poor (\numNoZero{.17} and \numNoZero{.21} for Bayesian and random permutation),
-and the prior term is a much better predictor (\numNoZero{.37}).
-Further, when this component is combined with the prior component,
-model performance improves only slightly over performance when only the prior is used (\numNoZero{.37} to \numNoZero{.39} for both the Bayesian and random permutation).
-
-\FloatBarrier
-
-\subsubsection{Discussion}
-
-Both models require large numbers of context co-occurrence observations to make accurate predictions from context information.
-Otherwise, the co-occurrence matrices do not have enough data to properly stabilize.
-Millions of posts were used to generate the global co-occurrence matrices for Twitter and StackOverflow.
-If this number of posts had been collected for a single user, then the user-customized co-occurrence matrix for that user should lead to more accurate predictions compared to using the global matrix.
-However, in this study only around 100 posts were used to build the user-specific matrices for each user.
-
-This is most likely why model accuracy was so poor when user-specific matrices were used:
-Even for the large datasets analyzed in this study, there is still not enough history collected for each user to build customized co-occurrence matrices.
-The prior model component can be customized for each user, since the prior is only concerned with tag use, and there are a reasonably small set of tags that the model is choosing from.
-However, the context component is built from the set of all of the prior tags crossed with all of the possible words, which is a much larger space than just the set of tags.
-So building a stable context component requires a much larger set of data than the prior component.
-
-That said, if a user only chooses a small set of tags (so the tag space is small), and computing a global co-occurrence matrix is infeasible (e.g., with the Twitter popular-users dataset), 
-then it may be worthwhile to add the user-custom context model component as an additional predictor.
-As was shown for Twitter, model accuracy can slightly improve over a model that just uses a custom prior, even when the co-occurrence matrix only contains around 100 posts.
-
-Further, if the number of co-occurrences is small, and a user-custom co-occurrence matrix is used, it may be advantageous to use the compressed random permutation model instead of the uncompressed Bayesian model.
-For the Twitter popular-users dataset, only a user-custom co-occurrence matrix could be built, and when this matrix is used the compressed random permutation model performed better than the uncompressed Bayesian model.
-When the global co-occurrence matrix was removed from the StackOverflow dataset, and a user-custom matrix was used instead, the accuracy for the random permutation model decreased less than the Bayesian model.
-This provides further evidence showing that there is an interaction between dataset size and optimal model:
-At lower numbers of co-occurrences the compressed random permutation model is able to perform better than the Bayesian model,
-and the Bayesian model surpasses the compressed model at higher co-occurrence counts once the cells in the full co-occurrence matrix stabilize.
 
@@
 
-\clearpage
-\section{Appendix 1: Support Code}
-
-All software code created for this research is publicly available on GitHub \parencite{StanleyRepo2014}.
-
-\subsubsection{Directory Structure}
-
-The repository is separated into three main folders:
-dissProject, firstSOProject, and writeups.
-The firstSOProject folder contains the support code for the \textcite{Stanley2013} research.
-The dissProject folder contains support code for this dissertation research.
-The writeups folder contains all writeup projects created during the dissertation process (e.g., the LaTeX for this dissertation, Keynote presentations).
-
-Within the dissProject folder, there are two main files that contain the majority of the code for the dissertation research: scrapeUsers.py and preProcess.R.
-The python file contains the necessary code to scrape the Twitter site for tweets, import results into the Postgres database, and perform various cleanup functions that were easiest to do in Python.
-The preProcess.R file contains all modeling and analysis code for the project.
-Within the file, there are implementations for the Bayesian and random permutation models.
-Also there is code to test the models by running each model through the various StackOverflow and Twitter test datasets.
-
-\subsubsection{Model Implementations}
-
-The Bayesian model is implemented using the data.table package, so that fast joins can be performed on the table when computing the strength of association values.
-The random permutation model is represented as a matrix, so that correlations between the model and context vectors are fast.
-The matrix is not sparse, since there are currently no sparse matrix packages for R that have a vectorized function to compute correlations between a vector and a matrix.
-
-The Bayesian model requires slightly more time to compute activation values and also more memory space to store the representation.
-This is because the random permutation model is compressed and only needs to store around \num{2048} rows in the co-occurrence matrix.
-However, both models are implemented efficiently enough so that all analysis runs and activation computations can be kept in memory.
-64 GB of memory was required to perform the analysis, particularly on the largest Twitter (3 million) and StackOverflow (1 million) corpus sizes.
-
-\subsubsection{Design Decisions}
-
-All model manipulations (e.g., using an entropy weighting metric, using a model with only the prior term) were expressed as part of a configuration object.
-That object is a list of configuration name, value pairs.
-The configuration object is passed to the analysis code, which determines the specific type of run that is performed.
-
-This configuration design enables a large amount of code reuse and overlap.
-This helps minimize the likelihood of a programming mistake, since functions are reused multiple times and overall static code length is kept to a minimum.
-
-\subsubsection{Testing}
-
-Each analysis run produced a set of result files.
-Those result files were checked into the repository so that changes could be tracked.
-Code changes were tested by looking at differences between previously checked in output files and the output files produced after a code change was made.
-
-A small amount of unit testing was also performed.
-The core behavior of each model implementation (Bayesian and random permutation) was tested in isolation to ensure that the activation values returned by the models were mathematically correct.
-Unit test code is located within dissProject/test.
-
-\subsubsection{Parallelization}
-
-None of the analysis code is finely parallelized.
-Most of the inner loop in R ends up running in compiled C, and execution within those function calls occurs on a single core.
-However, since multiple models were explored, each different model configuration was executed in parallel.
-This type of course-grained parallelization was achieved using R's built-in Parralel package.
-Also some parallelization was done at the OS level when various make targets would run in parallel.
-
-\subsubsection{Database}
-
-All raw Tweet and post data were stored in Postgres tables.
-Also, after the documents were tokenized, the tokenized versions of the raw data were stored in Postgres tables.
-When R ran an analysis, it would query the tables to retrieve the correct set of tokenized words and tags.
-So there is a lot of R code that uses the RPostgreSQL package to communicate with Postgres as the analysis is running.
-
-\subsubsection{Running the Analysis}
-
-All analysis runs were scripted and can be run by using the appropriate target in dissProject/Makefile.
-The three main targets to run the models through the prior-only runs on the popular users dataset,
-the full models on the randomly-sampled and popular-hashtags datasets,
-and the full models on the popular users dataset are ``parrunPriors'', ``parrunContext'', and ``parrunPUsers'' respectively.
-Each of these targets takes tens of hours to complete.
-``parrunContext'' is the longest and takes around 72 hours on a desktop machine with 64 GB of memory and 24 virtual cores.
-
-Smaller test runs were also created so that model changes could be more rapidly developed.
-These typically use smaller Twitter and StackOverflow corpus sizes (e.g., only \num{10000} StackOverflow posts) to create the co-occurrence matrix
-and a smaller number of posts to test the models on (e.g., 20 instead of 500).
-Running these functions takes around a minute to complete.
-To run these, the R function is called directly (i.e., not a make target) within an already-running R session.
-As an example, the ``runContext20g1s6'' function takes only a few minutes to complete, runs the models on a corpus size of \num{1000}, with only a single run per model, and only 20 sampled posts for each run.
-
 \end{document}

### 2014-11-14 writeups/psycMethods/psycMethods.tex
foo

@@
 
-\subsection{The Task: Predicting User-Generated Tags}
-
-Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp are composed almost entirely of human-created content.
-The amount of user content on these sites is quite large, and continues to grow.
-Users of Twitter's microblogging service, for example, generate half a billion tweets per day \parencite{TwitterReport2012}.
-However, a single user of a social media site is most likely interested in only a small fragment of this large amount of information.
-One general question to support the user in this domain is:
-How can we quickly and effectively connect users to the content that they care about?
-
-A growing number of users on social media sites are focusing on and looking for recently-created information when performing a search \parencite{Jansen2011}.
-In other words, users on social media sites want the fresh \emph{stream} of information about a particular topic.
-This can certainly be seen in the main user interface views that these social media sites provide:
-Twitter's homepage showing followees' tweets, Facebook's homepage showing friends' posts, and StackOverflow's daily digest of posts for particular tags.
-
-So how can we ensure that users are connected to the information streams that they care most about?
-One promising approach is to model and predict the user's goals when using and searching these social media sites \parencite{Rose2004}.
-If we have a better understanding of a user's goals, then we can tailor the information streams to the content that is most in line with these goals.
-For example, Twitter might suggest other users and hashtags for the user to follow, Facebook could suggest pages to like, Yelp could recommend businesses to check out.
-Researchers are currently looking for ways to automate the goal-identification task for more general search queries like those generated for Google \parencites{Jansen2008}{Lee2005}.
-
-However, social media sites have an advantage over traditional search engines for identifying user goals: human-created content.
-Twitter has more than just the tweet that a user is currently composing to try and identify that user's goals.
-The site also has all of the user's previous content, and can use that information to provide a much better prediction of a user's goals.
-Further, sites like Twitter, StackOverflow, Google+, and Facebook support ways for the user to explicitly identify their precise goals when creating content: hashtags.
-These user-created hashtags are a key indicator of a user's goals on a social media site.
-
-These hashtags relate to a user's goals and interests because they are a form of human-based document tagging \parencite{Chang2010}.
-Further, people are using hashtags as a way to create information streams on social media sites \parencite{Kwak2010}.
-Twitter users, for example, commonly create and use hashtags for upcoming political events that interest them, such as debates and races \parencite{Diakopoulos2010}.
-So if we can predict what hashtags a user is interested in, we have a better understanding of the information streams that they care about.
-We can use that information to correctly suggest hashtags that they have previously used when they are generating new content.
-
-Further, if a particular information stream has multiple hashtags associated with it (common for Twitter), these semantically similar hashtags can also be recommended to users.
-So we can tailor the system to suggest hashtags that the user may not yet know about but are of likely interest to them and growing in popularity across the site.
-We can help aid in their discovery of fresh and relevant information that matches their goals on the site.
-
 \subsection{Research Questions}
@@
 
-Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010}.
+Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010} and StackOverflow tag use \parencites{Kuo2011,Stanley2013}.
 Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
@@
 After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
-So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
 Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.
@@
 and the user data from the datasets are publicly available for analysis.
-
-Also, studying user-based tag generation on these sites has relevant real-world application.
-Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these popular sites.
-These systems can also help newer users by recommending proper tags for their newest content.
-On the StackOverflow site, for example, experts often subscribe to specific tags that interest them, and receive a daily digest of posts that tagged with those tags.
-Helping the user properly tag posts on StackOverflow ensures that right community sees the post, which greatly increases the chance that the question on the site will be answered quickly and correctly.
-For Twitter, hashtags represent streams of information that are possibly interesting to the user.
-Having a recommendation system that can suggest relevant hashtags to the user provides a way for the user to connect to and discover new information streams that interest them.
-Also, by suggesting relevant hashtags the system can help make sure a user's content is seen by other users who are interested in the same type of information.
 
@@
 
-\clearpage
 \section{Chapter 2: Theoretical Background}
@@
 
-\subsubsection{ACT-R DM Model}
-
-A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACTRModel}.
-
-\begin{table}[!ht]
-  \caption{ACT-R declarative memory model}
-  \label{tabACTRModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Activation &	 	$A_{i} = B_{i} + \sum_{j \in c}^{} W_{j} S_{ji}$ \\
-      Attentional Weight &	$W_{j} = \frac{W}{n}$ \\
-      Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
-      Constant Base Level &	$B_{i} = \frac{p_{i}}{1-p_{i}}$ \\
-      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
-      Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base-level activation ($B_{i}$) and strength of association ($S_{ji}$).
-The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
-Each term in Table \ref{tabACTRModel} will be discussed in greater detail to follow.
-
-\subsubsection{Base-Level Activation}
-
-Base-level activation reflects the log prior odds of needing an observed chunk again.
-The default way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
-This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
-Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
-However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
-In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its frequency and recency of use.
-
-For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable and do not change within that window.
-Programming language popularity over the past few years is a reasonable example of this.
-Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., Clojure's growth), it is not the case that the changes have been drastic.
-C\#, Python, and Java are still a few of the most popular languages, Common Lisp has not gained or lost much ground in popularity over the past few years.
-
-Further, if this is a reasonable assumption to make, it greatly simplifies the computation of the base-level activation for chunks, and can turn the computation into a tractable problem for large datasets.
-In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log-odds ratio of exposure to that chunk compared to exposure to all other chunks.
-For example, if the StackOverflow tag \emph{PHP} has been used four times as often as the tag \emph{Common Lisp}, then the prior odds of needing \emph{PHP} again is $\frac{.8}{.2}=4$ times that of \emph{Common Lisp}.
-
-\subsubsection{Strength of Association}
-
-Strength of association ($S_{ji}$) reflects the amount of log-odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
-Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
-Context for the Twitter domain are the words in a tweet.
-Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
-the likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
-For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
-So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
-This assumption was described when deriving the $S_{ji}$ equation \parencite{Anderson1989}, and used in recent research working with large-scale datasets \parencites{Stanley2013,Farahat2004,Douglass2010}.
-
-\subsubsection{Connection to Pointwise Mutual Information}
-
-Pointwise Mutual Information (PMI) is another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
-It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
-The index has been used to accurately and efficiently measure strength of association information for large-scale datasets \parencite{Budiu2007,Farahat2004}.
-The PMI index is presented as Equation \eqref{eqPMI}.
-
-\begin{equation}
-  \label{eqPMI}
-  \mathit{PMI}(y,x) = log \frac{p(y|x)}{p(y)}
-\end{equation}
-
-Further, this PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
-Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.
-That is, the PMI equation \emph{is} the simplified form of ACT-R's strength of association index ($S_{ji}$).
-
-\subsubsection{Attentional Weight}
-
-The attentional weight term allows for some contextual components to be weighted more than others when the total $S_{ji}$ activation is computed.
-This reflects the fact that the declarative memory system has limited attentional resources ($W$) that must be spread across all of the items in context.
-Most ACT-R models simply weight each contextual element equally (e.g., all words in a StackOverflow post have equal weight), so each word's attentional weight is simply normalized by the number of items in context.
-
 \subsection{Latent Semantic Analysis Memory Theory}
@@
 The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
-This process, described in more detail to follow, is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
+LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix.
 The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
-
-Usually information is lost when the data are restructured into a smaller set of latent dimensions.
-However, it was shown that this process actually improves prediction and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
-\textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
-Model results were on par with a large sample of applicants to American universities from non-English speaking countries (\num{64.4}\% compared to \num{64.5}\%).
-Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
-Afterwards further reductions in the dimensionality of the data began to degrade performance.
-
-\subsubsection{Singular Value Decomposition}
-
-LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix.
-The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
-The best way to remove a dimension while maintaining the most variability is to look for words with highly similar co-occurrence patterns and deduce that these words represent the same concept.
-For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
-By pooling co-occurrence counts the model is able to generate better predictions with less data.
-
-\subsubsection{Word Order}
-
-Although the LSA model has been shown to perform well as a way to measure similarity between two terms, one of the main criticisms is that it is a ``bag of words'' model.
-This means that the model does not account for or pay attention to word order when building the base word by document matrix or consequently, when performing the SVD dimension reduction.
-Other models that can account for word order have been shown to perform better than ``bag of words'' models on tasks that do not even appear to require grammatical structure such as word order to do well.
-For example, \textcite{Jones2007} compared a vector-based model called BEAGLE directly to LSA, training and testing both models on the same corpus used in \textcite{Landauer1997}.
-The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the TOEFL task (\num{55.6}\% vs. \num{55.3}\%).
-However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to \num{57.5}\%.
-It seems reasonable to expect that task performance will improve much more on tasks where word order is used (e.g., generating a hashtag in the middle of a sentence when composing a tweet).
-
-\subsubsection{Scaling Issues}
-
-Another issue with LSA is that the SVD matrix decomposition technique is computationally expensive.
-As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
-If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
-However, with large human-created datasets such as Twitter, StackOverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
-\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
-Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
-This is one of the available corpora on the LSA website and is intended to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.
-
-When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-sized datasets, LSA performed much better (23\% vs. 60\%).
-This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model outperformed the full dimensionality model on TASA-sized datasets.
-However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
-\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
-This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
-Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to \num{73.8}\%, which is 10\% higher than the TASA-trained LSA.
-This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.
-
-\subsubsection{Cost of Incremental Updating}
-
-Another issue with LSA is that adding additional observations to a previously-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
-Since computing the SVD is the most computationally expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
-For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.
 
@@
 This allows for retrieval requests such as ``playing \#$\Phi$,'' ``just read \#$\Phi$,'' and ``vector-based $\Phi$ systems.''
-
-\subsubsection{Addressing Scalability}
-
-Vector-based representations compute chunk likelihoods by performing a correlation operation directly on the memory vectors.
-Each memory vector can be thought of as a column on a context word by tag co-occurrence matrix used for a Bayesian representation.
-Further, the number of rows for a vector-based representation are fixed and set much smaller than the number of rows in the full word by tag co-occurrence matrix.
-Since Bayesian representations have been shown to scale to over hundreds of millions of word by tag co-occurrences \parencite{Stanley2013}, the compressed vector-based representations should scale as well or better.
 
@@
 and also scale to larger datasets (better computational efficiency).
-The vector length by performance tradeoff was tested on a simple paired associative retrieval task. 
-Retrieval accuracy for the random permutation vector-based models remained high relative to the circular convolution model even after reducing each environment vector's length from \num{2048} to \num{1024}.
-This indicates that vector-based representations using random permutations compress better than those using circular convolution.
-
-To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFL) assessment (among other tasks).
-The random permutation model was trained on \num{2.33} GB corpus.
-The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
-The random permutation model performed as well or slightly better than the BEAGLE model when trained on the smaller corpus.
-The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
 These results make a strong case to favor random permutations over circular convolution when implementing vector-based memory representations.
@@
 For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors is much less than the number of words.
-For the StackOverflow and Twitter datasets, the number of documents are the number of different tags.
-
 For random permutations, each tag is represented by a compressed vector with a length much less than the number of words.
-Note that if the length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular row in the column vector.
-If done in this way, building a matrix of all words' environment vectors by tags would recover the original word by document matrix \parencite{Kanerva2000}.
-However, each word's environment vector is compressed using a length much less than the number of unique words and represented by using a combination of a few one and negative one values at a unique combination of positions.
 
@@
 Nonetheless, random permutations have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFL \parencites{Sahlgren2008,Jones2007}.
-
-\subsection{Comparison of Vector-Based Models and ACT-R}
-
-Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
-Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
-Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
-For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
-
-\subsubsection{Base-Level Activation}
-
-Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the relative benefits of the two models. 
-In particular, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
-
-The frequency and recency of how content has been used in the past can be a strong predictor of the future importance of that information. 
-The recent content in tweets on a news site like Twitter should be more relevant, for example.
-\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
-They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.
-
-ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
-It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
-Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
-ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
-Vector-based models have focused on computing the associative strength of a chunk and specific contextual elements.
-It is unexplored how accuracy changes for these models once the information about a user's prior behavior (e.g., ACT-R's base-level activation component) is added.
-
-\subsubsection{Word Order}
-
-One strength of a vector-based approach over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
-With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
-To represent word order in a Bayesian model, an additional dimension could be added to the co-occurrence matrix that represents word position.
-However, adding word-order information to ACT-R's default Bayesian declarative memory model is not common.
-Instead, when word order information is used to model reading for example, production rules are defined so that words are read from left to right (e.g., \textcite{Lewis2006}).
-
-\subsubsection{Activation Calculation}
-
-Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
-Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
-These two computations are certainly not equivalent, even though they are most likely correlated with one another.
-
-Using a correlation can be problematic due to range effects in observed co-occurrences.
-If there are particular words for a tag that co-occur much more often than all of the other words with that tag, that single large value diminishes the power of all other words to contribute to the correlation calculation.
-This does not happen with Bayesian retrieval models because each pair of word and tag values is normalized by the total number of words that appeared with that tag and total number of tags that appeared with that word
-(see the strength of association equation in Table \ref{tabACTRModel}).
-The high-frequency words are often removed from co-occurrence matrix for vector-based models to reduce this effect.
 
@@
 \textcite{Stanley2013} also modeled a user's chosen tags for a post on the StackOverflow site.
-They modified the standard ACT-R model in Table \ref{tabACTRModel} in order to more accurately predict tag use on StackOverflow.
+They modified the standard ACT-R model in order to more accurately predict tag use on StackOverflow.
 A formal description of the modified ACT-R Bayesian model is included in Table \ref{tabModACTRModel}.
@@
 That tag's base-level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
-This model has been modified from the standard ACT-R model (Table \ref{tabACTRModel}) in three ways:
+This model has been modified from the standard ACT-R model in three ways:
+% FIXME: four ways; also base level term
 Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy measure was used to compute the attentional weight ($W_{j}$).
@@
 Looking at $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate limited attentional resources to predictive contextual elements.
-
-\paragraph{Comparison Between \citeauthor{Kuo2011}'s and \citeauthor{Stanley2013}'s Models}
-
-Since \textcite{Stanley2013} made several modifications to standard ACT-R model, it was more sophisticated than both the standard ACT-R model and the Bayesian model used by \textcite{Kuo2011}.
-Also, it was unclear if \citeauthor{Kuo2011} used both the words in the title and the body of the post as separate components for his models, or if only the words in the title were used.
-
-\citeauthor{Stanley2013}'s model also scaled much better than \citeauthor{Kuo2011}'s Bayesian co-occurrence model. 
-Because of this, 100 times more posts were used to build the co-occurrence matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
-After including the modifications and scaling up the model, the model can successfully predict a user's tag for a post at 65\% accuracy.
 
@@
 \textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
-\textcites{Zangerle2011, Kywe2012} used the Apache Lucene infrastructure to compute the similarity score, while \textcite{Li2011} used a custom similarity score to compute similarities.
-Regardless of the method to compute the similarity between tweets,
-each method collects the hashtags used from a set of the most similar tweets, then ranks the hashtags in the set, and presents the top 5-10 hashtags to the user.
-
-This tweet-centered approach is quite different than the hashtag-centered approach presented in \textcite{Mazzia2009}.
-With a hashtag-centered approach, a direct association is built between the words in a tweet and hashtags that occur in tweets.
-With a tweet-centered approach, that association between words and hashtags is indirect:
-The relationship is built between a tweet and its content, and then similar tweets are assumed to use similar hashtags.
-One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a representation of the content in each tweet has to be maintained to later compute tweet similarity.
-The storage size for a hashtag-centered approach grows with the number of hashtags, but it seems reasonable that this will grow slower than the number of tweets and asymptote after collecting a large sample of tweets.
-So from the perspective of efficient information storage and retrieval, it seems more likely that the hashtag-centered approach used in \textcite{Mazzia2009} will scale as Twitter grows in size and use.
-
-\textcite{Kywe2012} also customized their tag prediction model to the user's past hashtag use, and was one of the first to do so when developing a hashtag-prediction model for Twitter.
-The set of recommended hashtags was the union of hashtags used in similar tweets and hashtags used by similar users.
-This was done by storing a content vector of hashtag use for each user (user-centered approach) alongside the content vectors of hashtag use for each tweet (tweet-centered approach).
-To generate recommended hashtags, the model would return the top-ranked hashtags from the hashtags used in 0-50 similar tweets and 0-4 similar users.
-Prediction accuracy improved when the model combined recommendations based on hashtags in both similar tweets and the top few similar users compared to basing recommendations on similar tweets alone.
+Each method collects the hashtags used from a set of the most similar tweets, then ranks the hashtags in the set, and presents the top 5-10 hashtags to the user.
 
@@
 This technique is compensatory and allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
-This type of weighted additive model that combines past user behavior and current tweet context to predict the most likely hashtags has not yet been explored for Twitter.
-
-Both StackOverflow models \parencites{Stanley2013,Kuo2011} did not take into account the specific user's prior history when generating tag predictions.
-Rather their models, much like \textcite{Mazzia2009}'s Twitter model, uses the aggregate tag history across all users as the model's prior activation for each tag.
 
@@
 
-\clearpage
 \section{Chapter 3: General Methodology}
@@
 
-\clearpage
 \section{Chapter 4: Past User Behavior}
@@
 
-The base-level activation component from the full ACT-R declarative memory model in Table \ref{tabACTRModel} is included in Table \ref{tabACTRBLLModel}.
+The base-level activation component from the full ACT-R declarative memory model is included in Table \ref{tabACTRBLLModel}.
 
@@
 
-\clearpage
 \section{Chapter 5: Combining Predictors}
@@
 
-\subsection{Method to Combine Prior and Context}
+\subsection{Combining Prior and Context}
 
@@
 
-\clearpage
 \section{Chapter 7: General Discussion}

### 2014-11-17 writeups/psycMethods/psycMethods.tex
foo

@@
-\documentclass[man,floatsintext,donotrepeattitle]{apa6}
+%\documentclass[man,floatsintext,donotrepeattitle]{apa6}
+\documentclass[man,donotrepeattitle]{apa6}
 
@@
 \title{Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow}
-%\shorttitle{Comparing memory models of tag retrieval}
-\shorttitle{}
-
-\author{Clayton Stanley}
+\shorttitle{Comparing memory models of tag retrieval}
+%\shorttitle{}
+
+\author{Clayton Stanley and Michael D. Byrne}
 \affiliation{Rice University}
@@
 
-\def\committee{
-  Michael D. Byrne, Chair \\
-  Professor, Psychology and Computer Science \and
-  Phillip Kortum \\
-  Assistant Professor, Psychology \and
-  David Lane \\
-  Associate Professor, Psychology, Statistics, and Management \and
-  Devika Subramanian \\
-  Professor, Computer Science and Electrical Engineering \\
-}
-
-% Replace the title page in the apa6 style sheet with the title page from the Rice template
-% The title page style was taken from Rice University LaTeX PhD template: http://scholarship.rice.edu/handle/1911/21747
-% The following code for the abstract page was taken from the apa6 \maketitle command.
-\makeatletter
-\def\makeDisCover{
-  \def\maketitle{
-    \begin{titlepage}
-      \singlespacing
-      \null
-      \vfill
-      \vbox to 0pt{\vss\vbox to 8.75in{\parskip 0pt \parindent 0pt\centering
-	{\uppercase\expandafter{Rice University} \par} \vskip 0pt plus 3fil
-	{\bf \@title \par} \vskip 0pt plus 1.3fil
-	{by \par} \vskip 0pt plus 1fil
-	{\bf \@author \par} \vskip 0pt plus 3fil
-	{\sc A Thesis Submitted \\
-	  in Partial Fulfillment of the \\
-	Requirements for the Degree \par} \vskip 0pt plus 1.1fil
-	{\bf Doctor of Philosophy \par} \vskip 0pt plus 4.5fil
-	\begingroup
-	\leftskip .45\textwidth \rightskip 0pt plus 1fil
-	{\sc Approved, Thesis Committee: \par} \vskip 2.5em
-	\moveright\leftskip\vbox{\hrule width \leftskip}
-	\def\and{\par \vskip 2.25em minus 1em\moveright\leftskip\vbox{\hrule width \leftskip}}
-	\committee
-	\par \endgroup
-	\vskip 0pt plus 2fil
-	{Houston, TEXAS \par} \vskip 0pt 
-	{December \number\year \par}
-      }}
-      \doublespacing
-      \newpage
-  %BDB\hyphenpenalty 10000
-      \fussy
-      \@ifundefined{@abstract}{}{%
-	\thispagestyle{empty}
-	\section{\normalfont\normalsize \abstractname}% BDB
-	\begingroup
-	\parskip \z@ \parindent \z@ \par
-	\null \vskip .1in \centering
-	\@title
-	\vskip .3in
-	{by \par}
-	\@author \par
-	\vskip .5in
-      \endgroup
-      \@abstract\par% BDB
-      \@ifundefined{@keywords}{}{%
-	\setlength{\parindent}{0.4in}% BDB
-	\indent\textit{\keywordname:} \@keywords%
-      }%
-      \thispagestyle{empty}
-      \newpage
-    }
-    \@ifundefined{def@donotrepeattitle}{
-      \section{\protect\normalfont{\@title}}
-    }{}%
-    \raggedright%
-  \end{titlepage}
-  \setlength{\parindent}{0.4in}%
-}}
-\makeatother
-
-\makeDisCover
-
-%\keywords{}
+\keywords{ACT-R declarative memory theory, vector-based models, LSA, machine learning}
 
 \begin{document}
-
-\pagenumbering{roman}
 
@@
 
-\note{\today}
+%\note{\today}
 
@@
 
-\newpage
-
-\pagenumbering{arabic}
+%\newpage
 
@@
 The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.
-That is, each user on a social media site has created content that contains a set of tags within the context of each post.
-
-Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010} and StackOverflow tag use \parencites{Kuo2011,Stanley2013}.
+
+Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010} and StackOverflow tag use \parencites{Stanley2013}.
 Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
@@
 and the user data from the datasets are publicly available for analysis.
-
-\subsubsection{Comparison Between ACT-R and Vector-Based Memory Systems}
-
-ACT-R's declarative memory and vector-based memory systems are substantially different in their mathematical structure.
-Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., \cite{Rutledge2008} modeling the ``fan effect'').
-
-An advantage of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
-Word order information on sites like Twitter may contain highly predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
-Word order has been incorporated into ACT-R sentence-parsing models previously, but this has not been done using declarative memory.
-Rather, production rules were defined that ensured that the model followed a set of rules (e.g., left-to-right reading) when parsing sentences \parencite{Lewis2006}.
-It is certainly true that at least some production rules are needed when choosing words and creating content. 
-However, it is unexplored if some of the process of sentence construction and word selection can be represented in the ACT-R declarative memory system.
-
-ACT-R's declarative memory system places strong constraints on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2007}.
-This decay rate equation formalizes how recency and frequency of memory use relate to the prior probability that a memory will be retrieved.
-However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.
 
@@
 However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
-The process starts with a set of N documents.
-A word by document co-occurrence matrix is built by counting the number of times each word appears in each document.
-The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
-LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix.
+LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix, while maintaining as much of the variance in the original full matrix as possible.
 The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
@@
 One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
-Representing word order is a case where both the encoding and decoding operations are applied.
-As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
-Simply counting co-occurrences of the two immediately preceding words and hashtags does not include the order information of the words in the representation.
-With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
-Order information is lost if a simply co-occurrence counting technique is used.
-
 With vector-based systems, order information is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
 This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
-Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
+Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007}'s BEAGLE model to implement these operations.
 This allows for retrieval requests such as ``playing \#$\Phi$,'' ``just read \#$\Phi$,'' and ``vector-based $\Phi$ systems.''
 
-\subsubsection{BEAGLE}
-
-\textcite{Jones2007} created the BEAGLE vector-based memory representation.
-Each memory vector ($m_{i}$) is a summed representation of unordered and ordered word co-occurrences ($c_{i} + o_{i}$).
-This model uses circular convolution and circular correlation as a way to encode and decode word order.
-BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (\num{55.6}\% compared to \num{55.3}\%).
-Further, it was shown that incorporating word order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
-This model provided an efficient algorithm for using a single representation for unordered and ordered information that showed performance similar to if not better than LSA.
-
 \subsubsection{Random Permutation Model}
@@
 In order to represent word order information, a function must be used that converts an environment vector for a set of words and positions (e.g., ``stack $\Phi$'') into a new uncorrelated environment vector.
-The BEAGLE model uses circular convolution and circular correlation as the encoding and decoding operations.
-However, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors and positions.
-\textcite{Sahlgren2008} used a much simpler random permutation method for this operation.
+\textcite{Sahlgren2008} used a simple random permutation method for this operation.
 
@@
 
-For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow.''
-The ``stack'' environment vector ($e_{stack}$) is shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).
-That environment vector is then added to the memory vector for ``overflow'' ($m_{overflow} = m_{overflow'} + e_{stack^{-1}}$), so that ``overflow'' remembers that ``stack'' has preceded it in a sentence.
-
-A resonance retrieval for a word following ``stack'' (``stack $\Phi$'') is done by creating a memory vector out of all information in context.
-In this case, assume only a single piece of order information is available, that ``stack'' precedes the word to retrieve.
-So the memory vector for context is the environment vector for one-back ``stack'' ($m_{C} = e_{stack^{-1}}$).
-The memory vector for ``overflow'' ($m_{overflow}$) will be correlated the highest with this context vector ($m_{C}$) since ``overflow'' contains this vector.
-
-\subsubsection{Comparison between BEAGLE and Random Permutations}
-
-Random permutations uses one of simplest operations possible to create a new environment vector.
-This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
-Further, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.
-
-\textcite{Recchia2010} found that random permutations could use smaller vectors than circular convolution approaches (e.g., BEAGLE) to produce equivalent performance (better compression),
-and also scale to larger datasets (better computational efficiency).
-These results make a strong case to favor random permutations over circular convolution when implementing vector-based memory representations.
-
 \subsubsection{Connection to LSA}
 
-Both random permutations and LSA calculate word similarity on a reduced matrix.
+Both vector-based models and LSA calculate word similarity on a reduced matrix.
 For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors is much less than the number of words.
-For random permutations, each tag is represented by a compressed vector with a length much less than the number of words.
-
-Information is lost with this compression in both cases: by starting with a compressed representation with random permutations or by reducing to a compressed representation with LSA.
+For vector-based models, each tag is represented by a compressed vector with a length much less than the number of words.
+
+Information is lost with this compression in both cases: by starting with a compressed representation with vector-based models or by reducing to a compressed representation with LSA.
 However, the way that the compression works in these two cases is not the same.
 LSA uses SVD to find words that have highly correlated tag occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
-A vector-based approach like BEAGLE or random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
-Nonetheless, random permutations have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFL \parencites{Sahlgren2008,Jones2007}.
+A vector-based approach like random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
+Nonetheless, vector-based models have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFL \parencites{Sahlgren2008,Jones2007}.
 
@@
 
-\textcite{Kuo2011} was the first to model tag use for StackOverflow posts.
-His set of models were originally designed for next-word prediction for large text document collections.
-\citeauthor{Kuo2011} examined the accuracy of three different commonly-used models for next-word prediction: a Bayesian model, a K-nearest-neighbors model, and a Latent Semantic Analysis (LSA) model.
-In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user created a post, and restrict the space of possible words to only tags.
-He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
-Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive LSA and K-nearest-neighbors models.
-This Bayesian model can accurately predict a user-chosen tag at 47\% accuracy.
-
-% ref: http://tex.stackexchange.com/questions/53513/hyperref-token-not-allowed
-\paragraph{\texorpdfstring{\textcite{Stanley2013}}{Stanley2013}'s ACT-R DM Model for StackOverflow}
-
-\textcite{Stanley2013} also modeled a user's chosen tags for a post on the StackOverflow site.
+\textcite{Stanley2013} used a declarative memory model to predict a user's chosen tags for a post on the StackOverflow site.
 They modified the standard ACT-R model in order to more accurately predict tag use on StackOverflow.
@@
       Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
-      Co-occurrence &		$N_{ji} = \sum_{}^{}{observed_{ji}}$ \\
-      Observations &		$N = \sum_{}^{}{\sum_{}^{}{N_{ji}}}$ \\
-      Attentional Weight& 	$W_{j} = W \frac{E_{j}} {\sum_{}^{} {E_{j}}} $ \\
+      Attentional Weight & 	$W_{j} = W \frac{E_{j}} {\sum_{}^{} {E_{j}}} $ \\
       Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
       Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
-      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
       \hline
@@
 That tag's base-level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
-This model has been modified from the standard ACT-R model in three ways:
-% FIXME: four ways; also base level term
-Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy measure was used to compute the attentional weight ($W_{j}$).
+This model has been modified from the standard ACT-R model in four ways:
+Two separate contextual components were used (words in title and body of post), an offset term was added, an entropy measure was used to compute the attentional weight ($W_{j}$),
+and the base-level activation of a tag was based on global past tag use and not customized to a user's past tagging history.
 
@@
 This allowed the use of a logistic regression statistical technique to calibrate the weights for each of the terms.
-The title and body contextual components were separated since it seemed plausible that the words in the title would be better predictors of the tag used than the words in the body.
-It turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
-
-The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation and it predicts well \parencite{Dumais1991}.
-It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
-So the scaled entropy ($E_{j}$) of a context word can be thought of as the amount of predictiveness that that word has to \emph{any} tag.
-This means that a stop word like ``the'' will have a scaled entropy measure close to 0, while a highly-predictive word like \emph{macro} will have a scaled entropy measure close to 1.
+The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation \parencite{Dumais1991} and is parameter free.
 Looking at $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate limited attentional resources to predictive contextual elements.
@@
 The model presented by \textcite{Mazzia2009} is quite similar to an ACT-R declarative retrieval model.
-The main potential difference is that --- like \textcite{Kuo2011} and \textcite{Stanley2013} --- the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
-Instead, the global prior is an overall average frequency of hashtag use across all users.
-Further, no recency information was used to compute the global prior, so hashtags used in the recent past were not weighted more heavily. 
-This was most likely not much of an issue for \citeauthor{Mazzia2009}'s study, since all hashtags were collected within a relatively short period of time (about 3 months).
+The main potential difference is that --- like \textcite{Stanley2013} --- the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
 
@@
 \textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
-Each method collects the hashtags used from a set of the most similar tweets, then ranks the hashtags in the set, and presents the top 5-10 hashtags to the user.
-
-\subsubsection{State of Research on StackOverflow and Twitter}
-
-\textcite{Stanley2013} used an ACT-R based Bayesian memory model to predict human tagging on the StackOverflow site, but to our knowledge no vector-based memory model has been tested on the site.
-Both types of models have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007},
-so it should be straightforward to test a vector-based model on the StackOverflow site.
-
-Also, a vector-based retrieval model has not yet been tested on Twitter.
-One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
-It seems quite likely that words used just before a hashtag in a tweet (combined with their position) will be highly predictive of the hashtag chosen.
-To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
 
@@
 
-One of the main purposes of this research is to improve on these user-agnostic statistical models and include each specific user's past tagging history as a model component.
-Using customized priors that are tailored to a user's specific tagging history should be crucial for sites like Twitter where the possible tag space is infinite,
-but should also improve performance for more constrained sites like StackOverflow since each user is not interested in all of the possible tags on the site.
-
 \section{Chapter 3: General Methodology}
@@
 Two models were evaluated for hashtag prediction on both the Twitter and StackOverflow datasets: a vector-based model and Bayesian retrieval model.
-The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} was used for the vector-based model, since it has shown to perform better than the BEAGLE model \parencite{Recchia2010}.
+The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} was used for the vector-based model,
+since it has shown to perform better than the original BEAGLE vector-based model \parencite{Recchia2010}.
 The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel} was used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
@@
 A quota of three million tweets was selected because previously research testing Bayesian co-occurrence models on large-scale datasets has used on the order of one million documents
-\parencites{Stanley2013,Douglass2010,Budiu2007,Kuo2011}.
+\parencites{Stanley2013,Douglass2010,Budiu2007}.
 Tweets were collected in real time until the tweet quota was reached.
@@
 The initial model results with the new tokenizer were qualitatively similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
-This is most likely because the language for the samples taken for both sites is informal English,
-and it seams reasonable to think that tokenizing all of the informal language and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
-So the same tokenization algorithm was used for both StackOverflow and Twitter, since performance was similar to \textcite{Stanley2013} for StackOverflow,
+So the same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter, since performance was similar to \textcite{Stanley2013} for StackOverflow,
 and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.
@@
 Higher values mean that chunk activation decays faster, which means that the most recently presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
-Lower decay rate values flip the importance of frequency and recency:
-Lower values mean that chunks that have occurred most frequently will have the highest activation,
-and activation depends less on the specific chunks that have been observed in the recent past.
+Lower decay rate values flip the importance of frequency and recency.
 
@@
 To do this, a user's entire tag history was collected, and then each tag instance from oldest to newest was examined.
-For each of these tagging instances, the model generated an activation value for each possible tag based on that tag's frequency and recency of use. 
-This means, for example, that the model always missed characterizing the first observed tag for each user, since at that point the model does not have any history of tag use for that user.
-This also means that the tag space for each user is constrained and limited to the tags that each user has used in the past.
+For each of these tagging instances, the model generated an activation value for each possible tag based on that tag's frequency and recency of past use.
 
@@
 
-\subsubsection{Best-fit Decay Rate Values}
-
-To identify the best decay rate value for the dataset slices,
-each user's best decay rate value was taken and the results are plotted in histogram form in Figures \ref{figPriorSOQHistDsOL} and \ref{figPriorTwitterHistDsOL} respectively.
-The best-fit values are close to the \num{0.5} default value set in ACT-R when the optimized learning form is used.
-Comparing across optimized learning and the standard form, the best-fit values for the standard form are slightly higher for both the StackOverflow and Twitter dataset slices.
-Comparing best-fit values will be more thoroughly tested when looking at the values across all dataset slices.
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[][Standard form]{\includegraphics[width=8cm]{visHistD-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
-    \hfill
-    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visHistD-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
-    \caption{
-      Best-fit decay rate for a single dataset slice for StackOverflow.
-      The histogram represents the count of users for each best-fit decay rate value.
-      Each bar represents a discrete decay rate value that was explored.
-  }
-    \label{figPriorSOQHistDsOL}
-  }%
-\end{figure}
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[][Standard form]{\includegraphics[width=8cm]{visHistD-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}
-    \hfill
-    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visHistD-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
-    \caption{
-      Best-fit decay rate for a single dataset slice for Twitter.
-      Best-fit values were slightly higher for Twitter, so the range of the histogram is higher for Twitter than StackOverflow.
-  }
-    \label{figPriorTwitterHistDsOL}
-  }%
-\end{figure}
-
 \subsubsection{Aggregate Model Performance}
@@
 
-The decay rate parameter in $B_{i}$ determines how fast a presentation of a chunk in the past decays over time.
-Higher values of $d$ mean that activation decays more quickly, which means that the chunks used more recently are the most active and the total frequency count of a chunk matters less.
-Looking back at Figure \ref{figPriorDecay}, the optimal decay rate parameter when the standard form of $B_{i}$ is used is higher for Twitter than StackOverflow.
-It makes sense that the lifetime of a specific hashtag on Twitter is shorter overall than the lifetime of a tag (e.g., \emph{C\#}) on StackOverflow,
-since hashtag topics are generally more transient on Twitter than StackOverflow (e.g., hashtags for specific political events or upcoming conferences).
-This means that the decay rate for Twitter should be higher than StackOverflow, which is what is observed.
-
 \section{Chapter 5: Combining Predictors}
@@
 
-Chapter 5 will examine how model accuracy is influenced by two methodological concerns (adding an offset term and examining the effect of dataset subset for Twitter)
-and four architectural concerns
+Chapter 5 will examine how model accuracy is influenced by four architectural concerns
 (method of handling stop words, influence of dataset size, method to combine prior and context, and word order).
-The two methodological concerns were examined to ensure that the technique to calibrate model parameters was accurate and also that results for Twitter were not specific to the subset of data collected.
-
 The four architectural concerns were examined for the following reasons:
 During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.
-
 Due to research from \textcite{Budiu2007} and \textcite{Recchia2010}, model accuracy for the Bayesian and vector-based models is influenced by the number of documents in the dataset,
 so the effect of dataset size on accuracy was examined in more detail.
-Specifically,
-since both the Bayesian and random-permutation models are being evaluated, analyzing the effect of dataset size on accuracy in this research will allow direct comparisons between the two models on a common dataset.
-This has not yet been possible since the two models have never been compared on the same dataset within the same study.
-
 One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random-permutation model,
 so two different ways to add the model terms were tested.
-
 Finally, previous research has shown that including word order in the vector-based models improves performance \parencites{Jones2007, Sahlgren2008},
@@
 In order to best combine these predictors, a logistic regression statistical technique similar to \textcite{Stanley2013} was used to determine the most optimal weights for each term.
-\textcite{Stanley2013} weighted each term in the total activation equation (Table \ref{tabModACTRModel}) to maximize model accuracy.
-\citeauthor{Stanley2013} found that the optimal weights for the StackOverflow model for the prior term, title context term, and body context term were \num{.62}, \num{.93}, and \num{1.75} respectively.
-
-The optimal weights for each model component were computed in the same manner as \textcite{Stanley2013}:
 When a user creates a tag for a post, a retrieval request for the model is made.
@@
 
-For this form of model accuracy, the model labels N posts with a 1 where the labeling is not constrained in any way within each post.
-That is, the labeling is relaxed across all posts, and there are no constraints where the model must label a specific number of 1s for each post based on the number of tags that the author used for that post.
-So model accuracy values found with the relaxed method are actually slightly conservative, since the model may be labeling more 1s for a post than the number of tags the author used for that post.
-
-However, this relaxed metric was used for the rest of the experiments that involve combining terms because the optimal weights found by the regression are based on using the relaxed metric.
-If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression was not optimized to the constrained form, and weights may change as a consequence.
-Also, there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation where a small set of 1s occurred. 
-To optimize parameters based on the constrained form,
-a full-enumeration numerical approach would have to be developed that systematically searches the space of coefficient weights and uses a custom objective fitting function to find the best fit.
-However, this was not pursued for three reasons:
-Around \num{2500} total regressions were run on these datasets,
-a custom implementation in R would not likely run in compiled C,
-and even if it was implemented in a compiled language it would be computationally intractable without a closed-form solution.
-
 \subsubsection{Adding Offset Term}
@@
 However, when the regression can not be constrained to label a specific number of 1s for each observation (as is the case here), to be used properly an offset term must be added as an additional model predictor.
-
 The offset term keeps the activation for the small number of top tags for each post equal across posts.
 This ensures that the logistic regression only labels a few 1s for each post.
-\textcite{Stanley2013} did this by subtracting off from each post the mean total activation for the top 5 tags for that post.
-However, this method is complicated since computing the total activation for a tag requires already knowing the weights for each model component that are used to compute that activation.
-\citeauthor{Stanley2013} used an iterative approach to determine the weights used for the offset term, and kept iterating until those weights matched the weights returned after running the logistic regression.
 
@@
 
-However, a custom user prior for the Twitter popular-hashtags dataset cannot be used, since the entire Twitter dataset cannot be downloaded.
-The popular-hashtags dataset contains samples of a large set of various users that happened to use one of a specific set of hashtags being monitored.
-So in order to generate a custom user prior for each sample in the dataset, that user's entire tagging history must be accessed through the API.
-Since access rates are limited across that interface, it is infeasible to do this process for every user that appears in the Twitter popular-hashtags dataset.
+However, a custom user prior for the Twitter popular-hashtags dataset cannot be used, since the entire Twitter dataset cannot be downloaded, so the history of every user in the popular-hashtags dataset cannot be collected.
 So instead, the global prior tag history for all users in the Twitter popular-hashtags dataset was used to compute prior activation values for each tag.
@@
 and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too large for a current high-performance desktop, even with highly optimized representations.
-Already \num{835500000} total co-occurring word tag observations were used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
-It was also the case that \num{1000000} posts for StackOverflow was already around $1/7$ of the total posts created on the site,
-so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrence matrix instead,
-especially since \num{835500000} co-occurrences were already used with \num{1000000} posts.
-
-\subsection{Adding Offset}
-
-The first methodological concern was the offset component.
-To test the effect of adding the component on model accuracy, the offset component was added to both models, and model accuracy was examined before and after the addition.
-
-\subsubsection{Method}
-
-The four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset were used for the analysis.
-Logistic regressions were run both with and without the offset component added to each model.
-The results for each run for each dataset were collected and averaged.
-The results across the four Twitter datasets were collected together, and the average was taken across all runs in all four datasets.
-
-\subsubsection{Results}
-
-The impact of adding the offset to each of the models for StackOverflow and Twitter is shown in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.
-In both plots, if the model name includes ``full'' then all model components are included.
-When the model components are enumerated, then at least one of the components is left out.
-The random permutation and Bayesian models are plotted alongside each other so that comparisons across the two model types can be done easily.
-So, for example, the ``standard prior model relaxed across posts without offset'' means that only the prior model component was used and the offset for that component was not subtracted off when determining activation.
-As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
-  \caption{
-    Impact of adding offset component for StackOverflow.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean model accuracy across all runs in dataset.
-    The random permutation models are represented as ``RP'' and Bayesian models as ``Bayes.''
-}
-  \label{figContextOffsetSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
-  \caption{
-    Impact of adding offset component for Twitter.
-    All error bars are the 95\% bootstrapped confidence interval for the mean.
-  }
-  \label{figContextOffsetT}
-\end{figure}
-
-Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites
-(\numNoZero{.53} to \numNoZero{.60} and \numNoZero{.22} to \numNoZero{.45} for Bayesian and random permutation for StackOverflow,
-\numNoZero{.31} to \numNoZero{.29} and \numNoZero{.24} to \numNoZero{.25} for Bayesian and random permutation for Twitter),
-except that the full Bayesian model without an offset was slightly more accurate on Twitter.
-Accuracy is improved even when only a single term is used.
-Although in this case the term is not weighted,
-but using an offset still improves accuracy since the offset term ensures that only a small number of tags per posts are predicted by the model to be the chosen tags.
-
-The random permutation model in particular is greatly affected by the offset term, for both the StackOverflow and Twitter.
-This suggests that the mean activation for the random permutation model for the top few tags in each post can vary greatly across posts.
-Only if this mean activation is subtracted will the logistic regression model generate optimal weights and predictions that are comparable to the Bayesian model when model accuracy relaxed across posts is examined.
-
-\subsubsection{Discussion}
-
-It is somewhat disappointing that the offset term is even needed.
-If it were the case that model activation for a tag in a specific post could be interpreted as an exact activation value that did not depend on the specific post,
-then the model would be able to say when, for example, the third top tag in a post is more likely to be a chosen tag than the second top tag in another post.
-That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of tags to use for a specific post.
-If this were the case, then adding the offset term would actually decrease accuracy, since this absolute value of the activation values would be removed.
-However, since accuracy improves when adding the offset term for both models, it does not seem that absolute activation values have much meaning across posts for these datasets,
-and that the relative activation values for each tag within a post is what is important.
-
-One simple and likely explanation is that there are posts where a large set of tags have high activation,
-but the author is choosing to use only a fixed small set of the top tags for each post, regardless of the activation values of the top tags.
-The author determines the number of tags to use for a post based on competing goals of tagging posts and not overly saturating a post with tags.
-For example, the author may be writing a question on StackOverflow about \emph{javascript}, and a large set of related tags are retrieved from declarative memory that are all above the retrieval threshold
-(e.g., \emph{javascript}, \emph{dom}, \emph{svg}, \emph{html}, \emph{jquery}, \emph{css}),
-but the author only chooses to use the tags \emph{javascript}, \emph{dom}, and \emph{svg}, since he has decided to tag this post with three tags, and those are the three tags with the highest activation.
-Adding the offset term into the model is a way to implement a simple version of this strategy where the author is actively choosing to use only a small set of tags for each post.
-Also, this type of strategic behavior of limiting the total number of chosen tags could be easily implemented in the production system of ACT-R.
-
-\subsection{Twitter Subsets}
-
-As a second methodological concern, model accuracy was examined across each of the dataset subsets collected for Twitter.
-Four popular-hashtags subsets were collected for Twitter at different time periods.
-This was to ensure that model results could be interpreted as effects that were not specific to a certain time period or set of popular hashtags.
-
-\subsubsection{Method}
-
-The four subsets were collected across a period of one month.
-Each was collected sequentially, and about one subset was collected per week.
-The hashtags used in each subset were chosen by scraping a site that published the current trending hashtags for each city.
-Before starting to collect for each subset, the current trending popular hashtags for the 400 most populous US cities were collected, and then 400 popular hashtags were randomly sampled from this set.
-Only 400 were selected since this is the maximum that the Twitter API can monitor at one time.
-
-In order to ensure that model results are not dependent on the specific Twitter subset collected,
-it is important that the overall accuracy differences between the different models do not drastically shift from subset to subset.
-That is, it is important that the effect size of model is much greater than the effect size of the interaction between Twitter subset and model.
-A small interaction relative to the main effect means that the rank ordering of the different models remains consistent across the different Twitter subsets.
-Further, a small effect size of subset relative to the effect of model means that the exact accuracy for each model remains consistent across the subsets.
-To measure the effect sizes of subset, model type, and the interaction, a two-way between-subjects ANOVA was performed.
-
-\subsubsection{Results}
-
-Model accuracy results for components of both the random permutation and Bayesian models grouped by Twitter popular-hashtags subset are depicted in Figure \ref{figContextSubsets}.
-Accuracy is affected significantly by subset ($F(3, 340) = 764, p < .001, \eta ^ {2} = .12$), model ($F(4, 340) = 3617, p < .001, \eta ^ {2} = .77$),
-and the interaction between model and subset ($F(12, 340) = 142, p < .001, \eta ^ {2} = .09$).
-However, the effect size of model type on accuracy is \num{8.6} times larger than the interaction term ($\eta ^ {2} = .77$ versus $\eta ^ {2} = .09$).
-This shows that the performance of a model relative to the other models remains consistent across the four subsets.
-Further, the effect size of model type is \num{6.3} times larger than the effect of Twitter subset ($\eta ^ {2} = .77$ versus $\eta ^ {2} = .12$).
-This shows that not only is relative model performance maintained across Twitter subsets, but the exact value of model accuracy for each model remains reasonably consistent across subsets.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-contextStandardByGroupT-crop.pdf}}}
-  \caption{Model accuracy for each of the four Twitter popular-hashtags subsets}
-  \label{figContextSubsets}
-\end{figure}
-
-\subsubsection{Discussion}
-
-Since the effect sizes of subset and the interaction between model type and subset are minimal relative to the effect size of model,
-model behavior for the popular-hashtags subset for Twitter can be interpreted as behavior that will persist across different time periods,
-if the same method of choosing popular hashtags is conducted at a later date.
-It may also be the case that that model behavior does not depend on the specific method of choosing the popular hashtags, but since only a single method was used for this analysis,
-the results from this study cannot by themselves be used to answer this question.
 
@@
 ``Stop words'' (e.g., ``the,'' ``a,'' ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
-It is typically recommended \parencite{Bird2009} to remove stop words when building the co-occurrence matrix.
-Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in the text.
-Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.
-
 The method of handling stop words has been shown to influence model accuracy \parencite{Sahlgren2008,Stanley2013,Jones2007}, so this was the first architectural concern that was examined.
@@
 A data-driven frequency-weighted approach was tried, along with using a previously-generated stoplist, and no stop word removal at all.
-Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and using a previously-generated stoplist hardly increased performance.
 However, \citeauthor{Sahlgren2008} did not try to weight the stop words and only tried methods to remove them.
-
-\textcite{Jones2007} also filtered the stop words for the circular convolution model,
-but also mentioned that an entropy weighting technique \parencite{Dumais1991} has been used previously to attenuate (rather than remove) stop words.
-This data-driven technique identifies stop words by calculating the discriminating power (i.e., entropy) of each word to predict any tag.
-Each word is then weighted (not removed) based on their discriminating power.
 
@@
 
-Note that using a predefined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
-Also, no strong claims are being made that the cutoff identified is the ideal frequency percentage cutoff for all datasets when a frequency cutoff is used.
-The method used to identify this cutoff was simply intended to find a reasonable value where comparisons between a frequency cutoff and entropy weighting techniques would be equitable.
-
 \subsubsection{Results for the Random Permutation Model}
@@
 Plots for the three methods of handling stop words for the random permutation model on the StackOverflow dataset are shown in Figure \ref{figContextStopRPSO}.
-The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
-Using a predefined stoplist makes the smallest improvement (\numNoZero{.45} to \numNoZero{.49}), while the entropy (\numNoZero{.60}) and frequency (\numNoZero{.63}) techniques improve performance the most,
-and frequency shows a slight edge over entropy.
-However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
+In the plot, if the model name includes ``full'' then all model components are included.
+When the model components are enumerated, then at least one of the components is left out.
+So, for example, the ``standard prior model relaxed across posts'' means that only the prior model component was used when determining activation. 
+As another example, ``RP only title with freq'' means that only the title context component was used, and that stop words were handled using the frequency filtering technique.
 
@@
 
+The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
+Using a predefined stoplist makes the smallest improvement (\numNoZero{.45} to \numNoZero{.49}), while the entropy (\numNoZero{.60}) and frequency (\numNoZero{.63}) techniques improve performance the most,
+and frequency shows a slight edge over entropy.
+However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
+
 The three stop word techniques were then tested for the random permutation model on the Twitter dataset, and the results are depicted in Figure \ref{figContextStopRPT}.
@@
 \end{figure}
-
-\subsubsection{Discussion}
-
-Comparing across Figures \ref{figContextStopRPSO} and \ref{figContextStopRPT}, entropy and frequency techniques are superior to filtering based on a predetermined list.
-However, it is more difficult to choose the best approach when comparing entropy weighting and frequency filtering.
-Frequency filtering is slightly better for StackOverflow, while entropy weighting is slightly better for Twitter.
-However, frequency filtering does have a few drawbacks compared to entropy weighting:
-It produces accuracy measurements that have higher levels of noise (see Figure \ref{figContextStopRPSO} and the frequency model for the title component in StackOverflow).
-Also, the ideal cutoff threshold must be identified in order to filter based on frequency,
-and identifying this cutoff requires a search, and this search must be done for each dataset that frequency filtering is used on.
-This cutoff value is a free parameter in frequency filtering that is not required when weighting stop words by entropy.
-Therefore, favoring parsimony, for the random permutation model, attenuating stop words based on entropy is a slightly superior method compared to removing stop words based on frequency of use.
-This result is interesting since filtering stop words based on frequency is the more common default method for random permutation models \parencite{Sahlgren2008}.
-This suggests that a simpler method (entropy weighting) can be used instead of the more common frequency-filtering technique to achieve the same (or better) accuracy measurements.
 
@@
 corpus size, number of words in each document, total number of unique tags, and number of tags used per document.
-The average number of tags used per post for StackOverflow and Twitter were the same order of magnitude (around 3 for StackOverflow, slightly less for Twitter).
-
-However, the average number of words in a StackOverflow post is much greater than the 140 character limit used for tweets.
-More words per post means that the number of observations per cell in the co-occurrence matrix grows more quickly with corpus size.
-The Bayesian model is more likely to perform better with datasets like StackOverflow that have these larger corpus sizes.
-
-Also, the popular-hashtags Twitter dataset only contains 400 unique tags while the randomly-sampled StackOverflow dataset has \num{34377} unique tags.
-More unique tags means that the co-occurrence counts are spread across a wider range of tag columns,
-which means that it takes a larger number of total documents before the Bayesian model performs better than the random permutation model.
 
@@
 
-For example, say that the model is choosing between a set of tags, and those tags have a mean correlation of .2 with context and a standard deviation of 0.1.
-If a tag in the set has correlation of .2 with context, then approximately half of the observed tags have a correlation lower than this tag.
-So the one-tailed cumulative distribution value for this tag is .5.
-If a tag in the set has a correlation of .3, then approximately half plus .34 (cumulative distribution of normal curve from mean to 1 standard deviation outward) of the observed tags have a lower correlation.
-So the one-tailed cumulative distribution value for this tag is .84.
-
 This technique has a few advantages: 
@@
 
-Vector-based models such as BEAGLE \parencite{Jones2007} and the random permutation model \parencite{Sahlgren2008} have included only a context component in the past (i.e., no prior component).
+Vector-based models such as the random permutation model \parencite{Sahlgren2008} have included only a context component in the past (i.e., no prior component).
 When the prior component is included for the random permutation model, accuracy improves significantly (compare accuracy for full models to context models in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}).
@@
 At a row size of around \num{2000}, model accuracy improves by less than a percentage point when order information is incorporated.
-The results for larger row sizes are somewhat inconsistent, since the full model (order and orderless) sometimes underperforms the order-only model,
-which should never happen since the order-only model is a subset of the full model.
-\textcite{Sahlgren2008} most likely did not use a regression technique to optimize the weights for each of the order and orderless model terms.
+
 So although incorporating word order into the random permutation model does seem to slightly improve performance,
@@
 which is why model accuracy is similar for the orderless component and direction ordered component, and why accuracy does not improve when adding the direction ordered component to the orderless component.
-
-Since it was shown that including word order information has little to no significant improvement on model accuracy, no attempt to incorporate word order information into the Bayesian model was made.
-
-\subsection{Coefficients for Predictors}
-
-All of the methodological and architectural tests were done on the Twitter popular-hashtags dataset and the StackOverflow randomly-sampled dataset.
-For these tests, a logistic regression statistical technique was used to determine the optimal weight for each model component.
-A regression was done for each run in the dataset, and since multiple runs were used for the StackOverflow and Twitter datasets, multiple samples of coefficient weights for each model component were computed.
-The variance of the weights for each component show how likely it is that the regression is overfitting the data,
-and the mean shows the best estimate weight of each component relative to the other.
-
-\subsubsection{Method}
-
-Coefficients for all of the runs for the Twitter popular-hashtags and StackOverflow randomly-sampled datasets were collected.
-This totaled 5 sets of coefficient weights for StackOverflow (5 runs) and 60 for Twitter (15 runs for each of the 4 dataset samples).
-
-\subsubsection{Results}
-
-Coefficients for each model component of the full models are depicted in Figures \ref{figContextCoeffSO} and \ref{figContextCoeffT}.
-The range of the distribution of coefficient values within a component is much more narrow than the range of the mean coefficient values between the components of each model.
-This shows that the model is not being overfit, as the weights for each component are relatively stable across the different runs.
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffSO-crop.pdf}}}
-  \caption{
-    Coefficient values for each component of the full models for StackOverflow.
-    Models used are the best-performing Bayesian and random permutation models
-    (i.e., Bayesian and RP have entropy weighting, RP converts context activation from correlation to log odds).
-    Error bars represent the standard deviation of coefficient values for each component.
-    The standard deviation is used instead of the confidence interval of the mean so that the plots show the overall stability of each coefficient across the different runs.
-  }
-  \label{figContextCoeffSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffT-crop.pdf}}}
-  \caption{Coefficient values for each component of the full models for Twitter}
-  \label{figContextCoeffT}
-\end{figure}
-
-The weight of the prior relative to the context terms is also fairly consistent across models and across the two sites.
-Across the sites, context is a much stronger predictor than past behavior for StackOverflow, but past behavior is more important for Twitter. 
-Within a site, both models are consistent in how they favor context or past behavior.
-Also, looking at StackOverflow, both models are consistent in weighting the body more than the title, and the title more than the prior term.
-
-\subsubsection{Discussion}
-
-It was not surprising that the model was not overfit in this scenario, since the number of observed tags is much greater than the number of coefficients being fit.
-Each model has at most three terms, those weights are being fit to 500 posts within each run for Twitter and 100 for StackOverflow, and multiple tags are used for each post (particularly for StackOverflow).
-It is interesting that although the values of the weights are stable within a site (StackOverflow or Twitter), the weights change dramatically when looking across the sites.
-This illustrates that the impact of a user's past history relative to the current context can be domain specific, which means that for best results the weights for each of the terms should be recalibrated in new scenarios.
-However, the underlying reason that the weights change between the sites cannot be answered from this study alone.
-This relative impact may be changing from either a true alteration in the underlying weight of the prior compared to context for different domains,
-or simply due to the fact that the amount of information collected for each of the terms changes between the domains (less information leads to lower weighting in general).
 
@@
 
-Alternatively, perhaps authors of posts are taking advantage of context information that is not currently represented in the model.
-For example, time of day might be a useful predictor of tag use on Twitter.
-\textcite{Cohn2012} built a tag-prediction model for Twitter that was based primarily on temporal cycles of hashtag use.
-This technique could be naturally incorporated into a co-occurrence framework by including time of day as an additional context by tag co-occurrence for each post.
-This would allow time of day information to be used alongside the rest of the context information in the words of the post.
-A similar technique could be used to add additional context predictors to the model.
-These co-occurrences could either be added on as additional columns to the existing co-occurrence matrix,
-or a separate co-occurrence matrix could be used for each additional context component, similar to how separate matrices and context predictors were created for the words in the title and body of a post.
-Including additional context predictors such as time of day may help improve model accuracy,
-and perhaps it is the case that the rest of the variance left to predict is from these context components that the author is using but are not yet included in the model.
-
 \textcite{Anderson1991} also looked at similar tagging tasks with large datasets when deriving the prior component of the declarative memory theory.
@@
 
-\begingroup
-\setstretch{1}
-\setlength\bibitemsep{12pt}
 \clearpage
 \printbibliography[heading=bibintoc]
-\endgroup
 

### 2014-11-18 writeups/psycMethods/psycMethods.tex
foo

@@
 
-\section{Chapter 1: Introduction}
+\section{Introduction}
 
@@
 What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
-This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
+This question can be tested by generating a prediction for the most likely used tag whenever a user is about to generate a tag, and then comparing the user's chosen tag to the model's prediction.
 This process can be framed as a memory retrieval problem.
-The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.
+The process of suggesting a relevant tag to a user can be thought of as a memory retrieval request for a tag, given prior tag use and current context.
 
@@
 
-The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and I am interested in retrieval models that generalize across tasks.
+The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and we are interested in retrieval models that generalize across tasks.
 However, they are also similar on several accounts:
@@
 
-This document is separated into 7 chapters.
-Chapter 1 introduced the declarative memory research and motivated why it is interesting both from a theoretical and applied perspective to evaluate declarative memory models on large-scale datasets.
-The task to test the models was also introduced: tagging of user-generated content on StackOverflow and Twitter.
-Chapter 2 describes the vector-based and Bayesian declarative memory models that will be evaluated in this research, and shows how these models are related to the other common declarative memory models that are studied.
-Chapter 3 presents the overall method for testing the models on the StackOverflow and Twitter datasets.
-Chapter 4 shows the results for the first test of the models, where the model term for past user behavior is examined in isolation, and tags are predicted based purely on a user's past tagging history.
-Chapter 5 adds the context component to the models, and shows how model accuracy is influenced by changing various architectural components of each model.
-Chapter 6 describes model results when prior and context are combined and tested on both of the dataset subsets used for Chapter 4 and Chapter 5.
-Finally, Chapter 7 provides a general discussion of the overall findings.
-
-\section{Chapter 2: Theoretical Background}
+This rest of this document is separated into 5 sections.
+``Theoretical Background'' describes the vector-based and Bayesian declarative memory models that will be evaluated in this research,
+and shows how these models are related to the other common declarative memory models that are studied.
+``General Methodology'' presents the overall method for testing the models on the StackOverflow and Twitter datasets.
+``Past User Behavior'' shows the results for the first test of the models, where the model term for past user behavior is examined in isolation, and tags are predicted based purely on a user's past tagging history.
+Combining Predictors`` adds the context component to the models, and shows how model accuracy is influenced by changing various architectural components of each model.
+The final section provides a general discussion of the overall findings.
+
+\section{Theoretical Background}
 
@@
 A vector-based approach like random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
-Nonetheless, vector-based models have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFL \parencites{Sahlgren2008,Jones2007}.
+Nonetheless, vector-based models have been shown to perform similar to if not better than LSA \parencites{Sahlgren2008,Jones2007}.
 
@@
 
-\section{Chapter 3: General Methodology}
-
-Each specific result of the research (to be presented in Chapters 4-6) requires a sightly different method to study.
+\section{General Methodology}
+
+Each specific result of the research requires a sightly different method to study.
 However, there is a common general methodology that applies across all results that will be described first.
@@
 500 users were sampled that had a reputation higher than \num{500}, \num{1000}, \num{5000}, \num{10000}, \num{50000}, and \num{100000}.
-100 users were sampled that had asked a total number of questions higher than \num{50}, \num{100}, \num{200}, \num{300}, \num{400}, \num{500}.
+100 users were sampled that had asked a total number of questions higher than \num{50}, \num{100}, \num{200}, \num{300}, \num{400}, and \num{500}.
 
@@
 An additional popular-hashtags dataset was collected for Twitter.
-The purpose of this dataset is to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
+The purpose of this dataset was to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
 The dataset contains approximately three million tweets where at least one of the 400 most popular hashtags are used.
@@
 
-\section{Chapter 4: Past User Behavior}
+\section{Past User Behavior}
 
@@
 The optimal values for optimized learning are near \num{0.5}, which lines up with default value used in ACT-R.
-So the decay rate values for the standard form are slightly higher than the default values used in ACT-R, but the values found when optimized learning is used line up nicely with the ACT-R defaults.
 
@@
 
-\section{Chapter 5: Combining Predictors}
+\section{Combining Predictors}
 
 The declarative memory models analyzed for this research consist of two main components: past behavior and context.
-Past tag use was isolated in the first study to see how well ACT-R's prior component of declarative memory (i.e., base-level learning) fits each user's tag use over time.
-For the rest of the studies the past behavior and context declarative memory components were combined as well as looked at in isolation.
-
-Chapter 5 will examine how model accuracy is influenced by four architectural concerns
-(method of handling stop words, influence of dataset size, method to combine prior and context, and word order).
+Past tag use was isolated in the first section to see how well ACT-R's prior component of declarative memory (i.e., base-level learning) fits each user's tag use over time.
+For the next section the past behavior and context declarative memory components were combined as well as looked at in isolation.
+
+This section will examine how model accuracy is influenced by four architectural concerns:
+method of handling stop words, influence of dataset size, method to combine prior and context, and word order.
 The four architectural concerns were examined for the following reasons:
@@
 so the effect of dataset size on accuracy was examined in more detail.
-One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random-permutation model,
+One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random permutation model,
 so two different ways to add the model terms were tested.
@@
 
-A simplified version of \citeauthor{Stanley2013}'s method was used for the results for this research.
+A simplified version of \textcite{Stanley2013}'s method was used for the results for this research.
 Instead of subtracting off the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
@@
 \textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
-A data-driven frequency-weighted approach was tried, along with using a previously-generated stoplist, and no stop word removal at all.
+A data-driven frequency-filtering approach was tried, along with using a previously-generated stoplist, and no stop word removal at all.
 However, \citeauthor{Sahlgren2008} did not try to weight the stop words and only tried methods to remove them.
@@
 The no processing plot shows the size of the spikes when high-frequency words are not removed.
-The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used (compare the difference between the no processing subplot and the entropy subplot).
+The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used.
 The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
@@
 So, for example, the ``standard prior model relaxed across posts'' means that only the prior model component was used when determining activation. 
-As another example, ``RP only title with freq'' means that only the title context component was used, and that stop words were handled using the frequency filtering technique.
+As another example, ``RP only title with freq'' means that only the title context component of the random permutation model was used, and that stop words were handled using the frequency filtering technique.
 
@@
 and frequency shows a slight edge over entropy.
-However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
+However, using the frequency-filtering technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
 
@@
 
-\subsubsection{Discussion}
-
-It is likely that the random permutation model is starting to saturate when entropy weighting is used for the largest corpus size for StackOverflow.
-Saturation happens when certain words co-occur with a tag much more often than all of the other words.
-This introduces range effects when the correlation is computed, which ends up driving all of the correlations near one, and consequently decreases discrimination and accuracy.
-Saturation happens when entropy weighting is used because the non-predictive words are only attenuated and not removed, and this attenuation is not enough to overcome the range effects that grow with corpus size.
-
-However, the effect of saturation is small compared to the large effect from using either frequency filtering or entropy weighting to handle the stop words.
-The Bayesian model does not saturate when using entropy weighting because each cell in the co-occurrence matrix is normalized by word frequency and tag frequency.
-
 \subsubsection{Results for Twitter}
@@
 However, corpus size is not the only factor that determines whether the Bayesian model or random permutation model will perform better.
-For example, the corpus size where the Bayesian model begins to outperform the random permutation model is quite different for StackOverflow and Twitter.
 This is because the number of observations used to generate the co-occurrence matrix is a function of four components:
@@
 However, this means that ACT-R's prior base-level learning component (which is a log-odds value) is added to a correlation value, which may not be mathematically appropriate.
-As a fourth architectural concern, another method to combine the prior and context components for the random-permutation model was explored.
+As a fourth architectural concern, another method to combine the prior and context components for the random permutation model was explored.
 
@@
 Using the log-odds transformation technique has a small effect on model accuracy for StackOverflow (\numNoZero{.45} to \numNoZero{.47}) and minimal to no effect on Twitter (\numNoZero{.25} to \numNoZero{.25}).
-When the technique is used and accuracy is examined for a model with only a single predictor, the transformation does not change accuracy.
-When predictors are combined and the full model is examined, the transformation is in the direction of slightly increasing accuracy, at least for the StackOverflow dataset.
-However, any improvement when using the transformation is minimal.
 
@@
 Model accuracy results after incorporating word order into the random permutation model are depicted in Figure \ref{figContextOrder}.
-All bar plots that include an order component where ``window'' or ``direction'' is not included in the title use the standard ordering method (no direction or windowing, all position information is included)
-This is the default ordering method used, so the names for these bar plots were kept consistent to the names used in other figures so that comparisons could be made across figures if necessary
-(similar to what was done for the default row size for the random permutation model).
+All bar plots that include an order component where ``window'' or ``direction'' is not included in the title use the standard ordering method (no direction or windowing, all position information is included).
+This is the default ordering method used, so the names for these bar plots were kept consistent to the names used in other figures so that comparisons could be made across figures if necessary.
 Base models that do not use entropy weighting were also included so that effect sizes could be compared between entropy weighting and adding word order information.
@@
 
-\section{Chapter 7: General Discussion}
+\section{General Discussion}
 
@@
 
-Table \ref{tabOverallResults} summarizes model accuracies for full models (all components) on all of the datasets tested in the study.
+Table \ref{tabOverallResults} summarizes model accuracies for all models and datasets tested in the study.
+This includes the prior-only model tested on the popular-users dataset and the full models (context and prior) tested on the popular-hashtags and randomly-sampled datasets.
 After modifying the models by adding an entropy weighting mechanism to both models and a prior term to the random permutation model,
@@
     the RP model includes entropy weighting and log odds transformation when combining the context and prior components.
-    Both models use a user-customized co-occurrence matrix for the Twitter popular-users dataset, since a global co-occurrence
-    matrix is not available for this dataset.
   }
@@
 	\hline
-	StackOverflow	& popular-users		& Bayesian with entropy				& \numNoZero{.63} \\
-			& 			& RP with entropy and log odds 			& \numNoZero{.58} \\
+	StackOverflow	& popular-users		& Standard prior model				& \numNoZero{.28} \\
 			& randomly-sampled	& Bayesian with entropy 			& \numNoZero{.63} \\
 			& 			& RP with entropy and log odds			& \numNoZero{.63} \\
-	Twitter		& popular-users		& Bayesian with entropy and user sji		& \numNoZero{.39} \\
-			& 			& RP with entropy, log odds, and user sji	& \numNoZero{.39} \\
+	Twitter		& popular-users		& Standard prior model				& \numNoZero{.37} \\
 			& popular-hashtags	& Bayesian with entropy				& \numNoZero{.32} \\
@@
 Also, using a proper method for handling the low-predictor words produced some of the largest improvements in model accuracy compared to the other explorations
-(e.g., compared to word order, properly adding terms, compression amount for random permutation representation).
+(e.g., compared to word order or properly adding terms).
 Since the ACT-R declarative memory theory already has a component for attentional weight, it may be the case that the neural system performs some form of stop-word weighting when computing activations,

### 2014-11-18 dissProject/preProcess.R
foo

@@ -1798,7 +1798,10 @@ analyzePrior <- function(modelVsPredTbl) {
 	tbl2[d == .8, DVName := paste0(DVName, 'BestDStd')]
 	DVNames = c('topHashtagPostPriorStdDefaultD', 'topHashtagPostPriorOL2BestDOL', 'topHashtagPostPriorStdBestDStd')
 	resTbl = compareOptimalAcc(tbl2[DVName %in% DVNames & runNum == 2])
-	resTbl[, mean(meanVal), by=list(DVName)]
+	resTbl[, dsetType := 'twitter']
+	resTbl
+	resTbl[grepl('^SO', dsetGroup), dsetType := 'stackoverflow']
+	resTbl[, mean(meanVal), by=list(DVName, dsetType)]
 	resTbl
 	tbl
 }

### 2014-11-18 dissProject/preProcess.R
foo

@@ -1468,16 +1468,23 @@ plotBarSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol='dsetG
 	groupCol = as.symbol(groupCol)
 	fillCol = substitute(fillCol)
 	renameCols(sumTbl)
-	expr = bquote(ggplot(sumTbl, aes(x=factor(.(groupCol)), y=meanVal, fill=.(fillCol))) +
-		      geom_bar(aes(y=meanVal), width=0.7, position=position_dodge(width=.8), stat='identity') +
-		      #geom_text(aes(y=0, label=.(fillCol), hjust=0, vjust=-1.25), position=position_dodge(width=1.4)) +
-		      geom_errorbar(aes(ymin=minCI, ymax=maxCI), position=position_dodge(width=.8), width=0.1, size=0.3) + 
-		      #scale_x_discrete(expand=c(.04,0)) + 
-		      defaultGGPlotOpts)
-	plot = eval(expr)
-	plotTbl(plot, extras, figName)
+        if (eval(bquote(nrow(sumTbl[, .N, by=.(groupCol)]))) == 1) {
+		eval(bquote(sumTbl[, .(fillCol) := paste(strwrap(.(fillCol), width=25), collapse='\n'), by=.(fillCol)]))
+		plot = eval(bquote(ggplot(sumTbl, aes(x=factor(.(fillCol)), y=meanVal)) +
+				   geom_bar(aes(y=meanVal), width=0.7, fill='white', colour='black', position=position_dodge(width=.8), stat='identity')))
+	} else {
+		plot = eval(bquote(ggplot(sumTbl, aes(x=factor(.(groupCol)), y=meanVal, fill=.(fillCol))) + 
+				   geom_bar(aes(y=meanVal), width=0.7, position=position_dodge(width=.8), stat='identity')))
+	}
+	args = list(#geom_text(aes(y=0, label=.(fillCol), hjust=0, vjust=-1.25), position=position_dodge(width=1.4)),
+		    geom_errorbar(aes(ymin=minCI, ymax=maxCI), position=position_dodge(width=.8), width=0.2, size=0.6),
+		    #scale_x_discrete(expand=c(.04,0)),
+		    defaultGGPlotOpts)
+	plotTbl(plot, c(args, extras), figName)
 	sumTbl
 }
+
+
 
[…]

### 2014-11-19 writeups/dissertation/dissertation.tex
foo

@@
   Assistant Professor, Psychology \and
-  David Lane \\
-  Associate Professor, Psychology, Statistics, and Management \and
   Devika Subramanian \\
@@
 
+\thispagestyle{empty}
 \section{Acknowledgements}
@@
 
+\thispagestyle{empty}
 \tableofcontents
@@
 
+\thispagestyle{empty}
 \listoftables
@@
 
+\thispagestyle{empty}
 \listoffigures
 \newpage
+
+\pagenumbering{arabic}
 

### 2014-11-24 writeups/dissertation/dissertation.tex
foo

@@
 
+\subsubsection{Dataset Descriptives}
+
+The StackOverflow datasets had a much larger document size than the Twitter datasets overall.
+There were on average 9 and 158 words in the title and body of a post for StackOverflow when sampled across the entire site. 
+For Twitter, there were 13 words per tweet for the popular-users dataset and 18 for the popular-hashtags dataset.
+
+The average number of tags per dataset was lower overall for Twitter.
+For Twitter, authors used on average 1.1 hashtags per post for the popular-users dataset and 2.6 for the popular-hashtags dataset.
+For StackOverflow, 2.8 tags were used on average across all posts on the site.
+
+Since the document size and tags per document were higher for StackOverflow, the co-occurrence matrix for StackOverflow contained many more observations compared to Twitter.
+Taking 1 million posts across the site, the StackOverflow co-occurrence matrix contained \num{835000000} total observations and \num{55000000} unique combinations of context words and tags.
+For Twitter, when building a co-occurrence matrix for the popular-hashtags dataset from 3 million posts, the matrix contained \num{39000000} total observations and \num{4000000} unique combinations.
+
+Hashtags for Twitter were both placed at the very end of the Tweet and embedded within the middle of the tweet.
+For the popular-hashtags dataset, 11\% of hashtags were the last word in the Tweet and the rest were embedded within the tweet (not at the end).
+Tags for StackOverflow were always placed at the very end of the post, since the author chose tags by placing them in a separate text field before submitting the post.
+
+Each of the four Twitter popular-hashtags datasets contained the 400 most-popular hashtags at that time.
+There was only a small amount of overlap between the hashtags chosen for each of the datasets.
+Only 20\% of hashtags occurred two or more times across the four popular-hashtags datasets and the majority of those (66\%) occurred in only two of the four datasets.
+
+The StackOverflow dataset contains many orders of magnitude fewer unique users than Twitter.
+This means that the odds that the same user is sampled multiple times is less for the Twitter popular-hashtags dataset compared to all posts across the StackOverflow site.
+For Twitter, an author contributed an average of 2 tweets to the popular-hashtags dataset.
+Across the StackOverflow site, an author contributed 5 posts on average.
+
+\paragraph{Tag Frequency Distribution}
+
+\textcite{Stanley2013} found that the tag frequency for StackOverflow drops off sharply and approximately follows Zipf's Law.
+The same effect was found for the 2014 StackOverflow dataset in this research and the two Twitter dataset subsets.
+The results are depicted in Figure \ref{figDescTagFreq}.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{descr-tagFrequencyCount.pdf}}}
+  \caption{
+    Log-log plot of tag frequency versus rank for the StackOverflow dataset and Twitter popular-users and popular-hashtags dataset subsets.
+    Rank was assigned to each unique tag after ordering by tag frequency (most frequent given smallest rank).
+}
+  \label{figDescTagFreq}
+\end{figure}
+
+Tag frequency for StackOverflow and Twitter follow a similar Zipf's Law pattern, where tag use drops off sharply after moving away from the most popular tags.
+However, none of the three distributions follow Zipf's Law perfectly, since none of the plots show perfectly linear lines.
+Nonetheless, using three separate linear models to fit the data for the three distributions captures 96\% of the variance.
+So these log-log plots are mostly linear, and both StackOverflow and Twitter closely follow Zipf's Law for tag use frequency.
+
 \subsection{Acquiring Datasets}

### 2014-11-24 dissProject/preProcess.R

@@ -1843,14 +1843,13 @@ analyzePUser <- function(modelVsPredTbl) {
 	tbl
 	tbl[, .N, by=DVName]
 	# RESULT: Adding prior is useful when a lot of data about the user has been collected
-	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTitleBody', 'TitleBody', 'Body', 'Title',
+	DVNames = asTopHashtagAcross(c('PriorStd', 'PriorStdTitleBody', 'TitleBody',
 				       'TitleOrderlessEnthymanBodyOrderlessEnthyman',
 				       'PriorStdTitleOrderlessEnthymanBodyOrderlessEnthyman'))
 	resTbl = compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames], acc, figName='priorAddPUserSO')
 	resTbl[, mean(meanVal), by=DVName]
 	# RESULT: User sji is much less helpful than standard sji for StackOverflow
-	DVNames = asTopHashtagAcross(c('TitleOrderlessEnthyser', 'TitleUsercontext', 'TitleOrderlessEnthyman', 'TitleOrderlessEnthyser', 'Title',
-				       'PriorStdTitleOrderlessEnthyserBodyOrderlessEnthyser', 'PriorStdTitleBody', 'PriorStdTitleOrderlessEnthymanBodyOrderlessEnthyman',
+	DVNames = asTopHashtagAcross(c('PriorStdTitleOrderlessEnthyserBodyOrderlessEnthyser', 'PriorStdTitleBody', 'PriorStdTitleOrderlessEnthymanBodyOrderlessEnthyman',
 				       'PriorStdTitleUsercontextBodyUsercontext'))
 	resTbl = compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames], acc, figName='userSjiSO')
 	# RESULT: User sji is not very helpful for Twitter, and prior is useful again
@@ -3703,14 +3702,17 @@ countNumUniqueTHashtags <- function() {
 	fooTbl = setDT(sqldf('select * from top_hashtag_hashtags'))
 	fooTbl[rank < 400][, .N, by=hashtag]
 	fooTbl[rank < 400][, .N, by=hashtag][N>1]
+	fooTbl[rank < 400][, .N, by=hashtag][N==2]
 	307/1565
+	203/307
 	fooTbl[, .N, by=hashtag]
 	fooTbl[, .N, by=hashtag][N>1]
 	583/2267
 }
 
[…]

### 2014-11-24 writeups/dissertation/dissertation.tex
foo

@@
 
-Hashtags for Twitter were both placed at the very end of the Tweet and embedded within the middle of the tweet.
+Hashtags for Twitter were both placed at the end of the Tweet and embedded within the middle of the tweet.
 For the popular-hashtags dataset, 11\% of hashtags were the last word in the Tweet and the rest were embedded within the tweet (not at the end).
-Tags for StackOverflow were always placed at the very end of the post, since the author chose tags by placing them in a separate text field before submitting the post.
+Tags for StackOverflow were always placed at the end of the post, since the author chose tags by placing them in a separate text field before submitting the post.
 

### 2014-11-24 dissProject/preProcess.R
foo

@@ -1486,7 +1486,7 @@ plotBarSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol='dsetG
 	sumTbl
 }
 
-plotLineSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol) {
+plotLineSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol, title='Model Name') {
 	groupCol = as.symbol(groupCol)
 	fillCol = substitute(fillCol)
 	renameCols(sumTbl)
@@ -1500,8 +1500,9 @@ plotLineSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol) {
 		      geom_line(aes(group=.(fillCol), linetype=.(fillCol)), size=1.4) + 
 		      geom_point() +
 		      scale_x_discrete(breaks = unique(sumTbl[, .(groupCol)]), labels = sizeNums) +
+		      scale_colour_discrete(name=title) +
 		      geom_errorbar(aes(ymin=minCI, ymax=maxCI), width=0.1, size=0.3) + 
-		      scale_linetype_manual(values=lineTypes) + 
+		      scale_linetype_manual(name=title, values=lineTypes) + 
 		      guides(fill = guide_legend(keywidth = 1, keyheight = 1),
 			     linetype=guide_legend(keywidth = 5, keyheight = 1),
 			     colour=guide_legend(keywidth = 5, keyheight = 1)) +

### 2014-11-24 dissProject/preProcess.R
foo

@@ -3762,6 +3762,8 @@ fooDesc <- function() {
 	       xlab('Log Rank') + 
 	       ylab('Log Frequency Count') + 
 	       defaultGGPlotOpts + 
+	       scale_colour_discrete(name='Dataset Type') +
+	       scale_shape_discrete(name='Dataset Type') +
 	       theme(legend.position='top', legend.direction='vertical'))
 	plotTbl(p1, list(), 'descr-tagFrequencyCount')
 }

### 2014-11-24 dissProject/preProcess.R
foo

@@ -3762,8 +3762,8 @@ fooDesc <- function() {
 	       xlab('Log Rank') + 
 	       ylab('Log Frequency Count') + 
 	       defaultGGPlotOpts + 
-	       scale_colour_discrete(name='Dataset Type') +
-	       scale_shape_discrete(name='Dataset Type') +
+	       guides(shape=guide_legend(title='Dataset Type')) +
+	       guides(colour=guide_legend(title='Dataset Type')) +
 	       theme(legend.position='top', legend.direction='vertical'))
 	plotTbl(p1, list(), 'descr-tagFrequencyCount')
 }

### 2014-11-24 writeups/dissertation/dissertation.tex
foo

@@
 
+\subsubsection{Model Accuracy Differences Across Sites}
+
+Model accuracy for StackOverflow was much higher overall than for Twitter.
+The Twitter popular-hashtags dataset contained 400 unique hashtags for each of the four dataset subsets.
+The StackOverflow randomly-sampled dataset contained \num{34377} unique tags, since it sampled posts across the entire site.
+As the model is tasked to predict fewer and fewer hashtags, it should become more accurate (less chance of making an incorrect prediction).
+However, even with a much smaller tag set to predict, model accuracy for Twitter was still not nearly as high as StackOverflow.
+This is likely due to the fact that tag frequency use drops off sharply for both StackOverflow and Twitter.
+This means that the models on both sites are generally switching between the top few hundred tags when making predictions.
+So, increasing the total unique number of tags past that point makes only a small negative impact on model accuracy.
+
+However, there was a large difference in total number of users for StackOverflow and Twitter.
+A user on StackOverflow contributed 2.5 times more posts to the StackOverflow randomly-sampled dataset than the Twitter popular-hashtags dataset.
+Also note that the large co-occurrence matrices for StackOverflow and Twitter that contained millions of documents were not user specific.
+This means that the co-occurrence matrix for Twitter was built from a much larger set of unique users compared to StackOverflow.
+It is likely that co-occurrence matrices for individual users are correlated but not equivalent.
+So, as the co-occurrence matrix is built from a larger number of unique users, a greater amount of cross-user noise is introduced into the matrix.
+This noise would decrease model accuracy since the contextual component of the model would lose predictive information by having a noisier co-occurrence matrix.
+This is most likely one of the main reasons that model accuracy for Twitter is much lower than StackOverflow:
+Tag prediction becomes much more challenging as the number of users on a site increases.
+
+The second most-likely reason that model accuracy is much higher for StackOverflow is that the co-occurrence matrix for this site contains much more data due to the large document size for the site.
+The co-occurrence matrix for the StackOverflow randomly-sampled dataset contained 20 times more observations than the matrix for one of the Twitter popular-hashtags datasets.
+Model accuracy generally increases as more observations are used to generate the co-occurrence matrices, since this allows each cell in the matrix to stabilize to its true value.
+
 \subsubsection{Representation}

### 2014-11-24 dissProject/preProcess.R

@@ -3741,10 +3741,10 @@ fooDesc <- function() {
 	sjiTblTOrderless[, sum(partialN)]
 
 	fooTbl = setDT(sqldf('select user_screen_name, count(id) from posts where post_type_id = 1 group by user_screen_name'))
-	fooTbl[, list(mean(count), median(count), max(count))]
+	fooTbl[, list(mean(count), median(count), max(count), trMean=mean(count, trim=.1))]
 
 	fooTbl = setDT(sqldf('select user_screen_name, id from top_hashtag_tweets limit 1000000'))
-	fooTbl[, .N, by=user_screen_name][, list(mean(N), median(N), max(N))]
+	fooTbl[, .N, by=user_screen_name][, list(mean(N), median(N), max(N), trMean=mean(N, trim=.1))]
 
 	fooTbl = setDT(sqldf("select id, chunk from top_hashtag_tokenized where type = 'hashtag' limit 1000000"))
 	curplotTbl = fooTbl[, .N, by=chunk][, type := 'Twitter Popular Hashtags']

### 2014-11-24 writeups/dissertation/dissertation.tex
foo

@@
 
-\subsubsection{Compression}
+\subsubsection{Word Order}
 
@@
 The random permutation model seems to work so well primarily due to the way that the full co-occurrence representation is compressed.
+
+\subsubsection{Compression}
 
@@
 Perhaps the randomly-compressed representation used for the random permutation model is the way in which the brain tries to efficiently achieve the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
+
+\subsubsection{Corpus Size}
+
+The random permutation model was more accurate than the Bayesian model at smaller corpus sizes, while the results were flipped for larger document sizes.
+This is interesting from a practical standpoint because the corpus size is rarely chosen in declarative memory research, and is constrained primarily by the amount of data available in the domain being studied.
+If a researcher is working within a domain that has a relatively small corpus size (generally near or less than \num{1000} documents),
+it may be worthwhile to test the random permutation model in this domain, as this compressed model may be more accurate than the uncompressed Bayesian model.
+
+Further, it is interesting to study the effect of model accuracy on corpus size because people require exposure to only a small number of documents before tagging documents appropriately.
+A user asked around five questions on average for StackOverflow, model accuracy was high for StackOverflow, but the model required training on 1 million documents before achieving that accuracy.
+The user only needed to ``train'' on five questions and the model needed to train on 1 million questions.
+
+This is a large discrepancy, but this does not necessarily invalidate the declarative memory models used for this research.
+First, the random permutation model was still accurate at smaller document sizes, so perhaps this result provides evidence to favor the random permutation model over the Bayesian model.
+However, the smallest corpus size tested for these models was \num{1000} documents, which is much greater than the average number of questions or tweets that a user contributes to the site.
+
+Second, the models may require large amounts of training because the co-occurrence matrices are being built from scratch.
+That is, the models are being trained on the StackOverflow and Twitter domains without any prior general co-occurrence information to start with.
+Presumably, when a person starts using a site, they can leverage all of their past exposure to different word co-occurrences and then simply add on top of that base the domain-specific co-occurrences.
+In that since, one could argue that the models are undertrained on generic co-occurrences (people are exposed to many more than one million documents in their lifetime) and overtrained on specific co-occurrences.
+It may be interesting to test these memory models on a hybrid co-occurrence matrix,
+where the majority of the observations are taken from a generic domain (e.g., american literature) and a minority are domain-specific (e.g., StackOverflow word, tag pairs).
 

### 2014-11-25 writeups/dissertation/dissertation.tex
foo

@@
-%\documentclass[man,floatsintext,donotrepeattitle]{apa6}
-\documentclass[man,donotrepeattitle]{apa6}
+\documentclass[man,floatsintext,donotrepeattitle]{apa6}
+%\documentclass[man,donotrepeattitle]{apa6}
 
@@
 
-Hashtags for Twitter were both placed at the end of the Tweet and embedded within the middle of the tweet.
-For the popular-hashtags dataset, 11\% of hashtags were the last word in the Tweet and the rest were embedded within the tweet (not at the end).
+Hashtags for Twitter were both placed at the end of the tweet and embedded within the middle of the tweet.
+For the popular-hashtags dataset, 11\% of hashtags were the last word in the tweet and the rest were embedded within the tweet (not at the end).
 Tags for StackOverflow were always placed at the end of the post, since the author chose tags by placing them in a separate text field before submitting the post.
@@
 
-All raw Tweet and post data were stored in Postgres tables.
+All raw tweet and post data were stored in Postgres tables.
 Also, after the documents were tokenized, the tokenized versions of the raw data were stored in Postgres tables.

### 2014-11-25 writeups/dissertation/dissertation.tex
foo

@@
 Specifically,
-since both the Bayesian and random-permutation models are being evaluated, analyzing the effect of dataset size on accuracy in this research will allow direct comparisons between the two models on a common dataset.
+since both the Bayesian and random permutation models are being evaluated, analyzing the effect of dataset size on accuracy in this research will allow direct comparisons between the two models on a common dataset.
 This has not yet been possible since the two models have never been compared on the same dataset within the same study.
 
-\textcite{Sahlgren2008} showed how different levels of compression in the random-permutation model influence accuracy, so this was examined for three reasons:
+\textcite{Sahlgren2008} showed how different levels of compression in the random permutation model influence accuracy, so this was examined for three reasons:
 to replicate the findings,
-as an additional check to ensure that the random-permutation model was implemented correctly, and also to see if the same levels of compression are valid for the much larger datasets used in this research.
-
-One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random-permutation model,
+as an additional check to ensure that the random permutation model was implemented correctly, and also to see if the same levels of compression are valid for the much larger datasets used in this research.
+
+One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random permutation model,
 so two different ways to add the model terms were tested.

### 2014-11-25 writeups/dissertation/dissertation.tex
foo

@@
 What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
-This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
+This question can be tested by generating a prediction for the most likely used tag whenever a user is about to generate a tag, and then comparing the user's chosen tag to the model's prediction.
 This process can be framed as a memory retrieval problem.
-The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.
+The process of suggesting a relevant tag to a user can be thought of as a memory retrieval request for a tag, given prior tag use and current context.
 That is, each user on a social media site has created content that contains a set of tags within the context of each post.
 
-Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010}.
+Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010} and StackOverflow tag use \parencites{Stanley2013}
 Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
@@
 However, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors and positions.
-\textcite{Sahlgren2008} used a much simpler random permutation method for this operation.
+\textcite{Sahlgren2008} used a simple random permutation method for this operation.
 
@@
 
-Both random permutations and LSA calculate word similarity on a reduced matrix.
+Both vector-based models and LSA calculate word similarity on a reduced matrix.
 For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors is much less than the number of words.
@@
 
-Information is lost with this compression in both cases: by starting with a compressed representation with random permutations or by reducing to a compressed representation with LSA.
+Information is lost with this compression in both cases: by starting with a compressed representation with vector-based models or by reducing to a compressed representation with LSA.
 However, the way that the compression works in these two cases is not the same.
@@
 That tag's base-level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
-This model has been modified from the standard ACT-R model (Table \ref{tabACTRModel}) in three ways:
-Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy measure was used to compute the attentional weight ($W_{j}$).
+This model has been modified from the standard ACT-R model (Table \ref{tabACTRModel}) in four ways:
+Two separate contextual components were used (words in title and body of post), an offset term was added, an entropy measure was used to compute the attentional weight ($W_{j}$),
+and the base-level activation of a tag was based on global past tag use and not customized to a user's past tagging history.
 
@@
 
-The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation and it predicts well \parencite{Dumais1991}.
+The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation \parencite{Dumais1991} and is parameter free.
 It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
@@
 and it seams reasonable to think that tokenizing all of the informal language and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
-So the same tokenization algorithm was used for both StackOverflow and Twitter, since performance was similar to \textcite{Stanley2013} for StackOverflow,
+So the same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter, since performance was similar to \textcite{Stanley2013} for StackOverflow,
 and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.
@@
 \textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
-A data-driven frequency-weighted approach was tried, along with using a previously-generated stoplist, and no stop word removal at all.
+A data-driven frequency-filtering approach was tried, along with using a previously-generated stoplist, and no stop word removal at all.
 Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and using a previously-generated stoplist hardly increased performance.
 However, \citeauthor{Sahlgren2008} did not try to weight the stop words and only tried methods to remove them.
-So alongside looking at frequency weighting, previously generated, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on entropy was also explored.
+So alongside looking at frequency filtering, previously generated, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on entropy was also explored.
 
@@
 Model accuracy results after incorporating word order into the random permutation model are depicted in Figure \ref{figContextOrder}.
-All bar plots that include an order component where ``window'' or ``direction'' is not included in the title use the standard ordering method (no direction or windowing, all position information is included)
+All bar plots that include an order component where ``window'' or ``direction'' is not included in the title use the standard ordering method (no direction or windowing, all position information is included).
 This is the default ordering method used, so the names for these bar plots were kept consistent to the names used in other figures so that comparisons could be made across figures if necessary

### 2014-12-02 writeups/dissertation/dissertation.tex
foo

@@
   Professor, Psychology and Computer Science \and
-  Phillip Kortum \\
+  Philip Kortum \\
   Assistant Professor, Psychology \and
@@
 \begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{descr-tagFrequencyCount.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{descr-tagFrequencyCount-crop.pdf}}}
   \caption{

### 2014-12-03 writeups/dissertation/dissertation.tex
foo

@@
 
-Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010} and StackOverflow tag use \parencites{Stanley2013}
+Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010} and StackOverflow tag use \parencites{Stanley2013}.
 Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.

### 2014-12-03 writeups/psycMethods/psycMethods.tex
foo

@@
 \begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{memMatSO-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{memMatSO-crop.pdf}}}
   \caption{
@@
 \begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{memMatT-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{memMatT-crop.pdf}}}
   \caption{
@@
 \begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
   \caption{Random permutation model accuracy when using log-odds transformation on context for StackOverflow}
@@
 \begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
   \caption{Random permutation model accuracy when using log-odds transformation on context for Twitter}
@@
 The general discussion is separated into two parts.
-The first section will describe the overall results and reemphasize key findings.
-The final section will discuss general theoretical implications of the results and describe future applications and directions.
+The first section will describe the overall results, reemphasize key findings, and discuss the theoretical implications of each key finding.
+The final section will describe future applications and directions for the research.
 
@@
 
-\textcite{Anderson1991} also looked at similar tagging tasks with large datasets when deriving the prior component of the declarative memory theory.
-\citeauthor{Anderson1991} examined, for example, the frequency of word choice over time used for the title of articles for the New York Times,
-and used the results to derive the power law equation for the prior term in the declarative memory model.
-ACT-R's declarative memory theory claims that the system has evolved to make predictions that closely match the statistical structure of the environment.
-Using a power-law equation to model future tag use based on past tag use is a reasonable statistical approach to the problem.
-It is also the case that this same equation is an accurate model for the prior component of declarative memory, as was shown by \citeauthor{Anderson1991} in the original derivation, 
-and as is supported in the results from this study.
-
 \subsubsection{Representation}
@@
 
-As neural imaging systems improve, this question of neural representation can be best answered by looking directly at the neural activity.
-However, there are possible behavioral studies that could be done to provide evidence that may favor one representation over the other:
-The largest differences between the Bayesian and vector-based models happen with small numbers of co-occurrence observations.
-The vector-based model begins to generate more accurate predictions sooner than the Bayesian models,
-since the noise introduced from the compression helps smooth out activation values until a larger number of co-occurrences are observed and the co-occurrence values stabilize.
-For a behavioral study, novel words could be introduced alongside a corpus of known words, and the participant could be asked to make retrievals of the novel words and known words as more novel words are introduced.
-Both models would make different predictions about the rate at which activation values for the novel words stabilize relative to the known words, and such make different predictions about retrieval error rates over time.
-Learning a new language might be a domain where large amounts of these sorts of retrievals could be collected.
-For instance, the online site Duolingo might contain enough collective data about language learning and word translation to better tease apart the two context models.
-
 \subsubsection{Compression}
-
-Also, the relatively strong performance of the random permutation model is not due to the model's ability to represent word order.
-Performance is similar for the random permutation and Bayesian models when the co-occurrence matrix for both models is the same, and not based on word order.
-Performance also does not improve by much when word order is added as the last predictor in the random permutation model.
-The random permutation model seems to work so well primarily due to the way that the full co-occurrence representation is compressed.
 
@@
 Perhaps the randomly-compressed representation used for the random permutation model is the way in which the brain tries to efficiently achieve the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
+
+\subsubsection{Word Order}
+
+The relatively strong performance of the random permutation model is not due to the model's ability to represent word order.
+Performance is similar for the random permutation and Bayesian models when the co-occurrence matrix for both models is the same, and not based on word order.
+Performance also does not improve by much when word order is added as the last predictor in the random permutation model.
+The random permutation model seems to work so well primarily due to the way that the full co-occurrence representation is compressed.
+
+\subsubsection{Corpus Size}
+
+The random permutation model was more accurate than the Bayesian model at smaller corpus sizes, while the results were flipped for larger document sizes.
+This is interesting from a practical standpoint because the corpus size is rarely chosen in declarative memory research, and is constrained primarily by the amount of data available in the domain being studied.
+If a researcher is working within a domain that has a relatively small corpus size (generally near or less than \num{1000} documents),
+it may be worthwhile to test the random permutation model in this domain, as this compressed model may be more accurate than the uncompressed Bayesian model.
+
+Further, it is interesting to study the effect of model accuracy on corpus size because people require exposure to only a small number of documents before tagging documents appropriately.
+A user asked around five questions on average for StackOverflow, model accuracy was high for StackOverflow, but the model required training on 1 million documents before achieving that accuracy.
+The user only needed to ``train'' on five questions and the model needed to train on 1 million questions.
+
+This is a large discrepancy, but this does not necessarily invalidate the declarative memory models used for this research.
+First, the random permutation model was still accurate at smaller document sizes, so perhaps this result provides evidence to favor the random permutation model over the Bayesian model.
+However, the smallest corpus size tested for these models was \num{1000} documents, which is much greater than the average number of questions or tweets that a user contributes to the site.
+
+Second, the models may require large amounts of training because the co-occurrence matrices are being built from scratch.
+That is, the models are being trained on the StackOverflow and Twitter domains without any prior general co-occurrence information to start with.
+Presumably, when a person starts using a site, they can leverage all of their past exposure to different word co-occurrences and then simply add on top of that base the domain-specific co-occurrences.
+In that since, one could argue that the models are undertrained on generic co-occurrences (people are exposed to many more than one million documents in their lifetime) and overtrained on specific co-occurrences.
+It may be interesting to test these memory models on a hybrid co-occurrence matrix,
+where the majority of the observations are taken from a generic domain (e.g., american literature) and a minority are domain-specific (e.g., StackOverflow word, tag pairs).
 
@@
 
-\subsubsection{Summary}
-
-The end result of this research is [1] a better understanding of how a user's past tagging history influences future tag use,
-[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on large-scale real-world retrieval tasks,
-and [3] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
-This work has shown how past user history is important, and how a context-only model like the random permutation model can be modified to include the prior component.
-The work has also shown how the strength of the random permutation model may not be in its ability to represent word order,
-and rather in its ability to efficiently compress the full co-occurrence representation used for the Bayesian model.
-Also from a methodological perspective,
-this work has shown how large collections of publicly-available behavioral datasets can be used to explore and test different theories of declarative memory and their individual architectural components.
-Several different formalisms for ACT-R's attentional weighting term were also explored, and an entropy-weighting technique was found to perform well without requiring any parameter tuning.
-Finally, this work has helped raise the question about how co-occurrence observations are represented at the neural level, since two models with quite different representations performed similarly well.
-The results from the study and the new research questions raised will help to better understand how imposing specific architectural constraints on memory models influence what information is retrieved.
-
 \clearpage

### 2014-12-03 writeups/psycMethods/psycMethods.tex
foo

@@
 ``Past User Behavior'' shows the results for the first test of the models, where the model term for past user behavior is examined in isolation, and tags are predicted based purely on a user's past tagging history.
-Combining Predictors`` adds the context component to the models, and shows how model accuracy is influenced by changing various architectural components of each model.
+``Combining Predictors'' adds the context component to the models, and shows how model accuracy is influenced by changing various architectural components of each model.
 The final section provides a general discussion of the overall findings.
@@
 The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation \parencite{Dumais1991} and is parameter free.
-Looking at $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate limited attentional resources to predictive contextual elements.
 
@@
 Several hashtag recommendation models for Twitter have been developed recently.
-One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a post as a function of prior hashtag use and context \parencite{Mazzia2009}.
+One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a tweet as a function of prior hashtag use and context \parencite{Mazzia2009}.
 The model presented by \textcite{Mazzia2009} is quite similar to an ACT-R declarative retrieval model.
@@
 An additional popular-hashtags dataset was collected for Twitter.
-The purpose of this dataset was to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
+This dataset was used to test the context component of the models for Twitter.
 The dataset contains approximately three million tweets where at least one of the 400 most popular hashtags are used.
@@
 Otherwise a larger number of hashtags would have been collected.
-Twitter allows the streaming interface to randomly sample about 1\% of all of the real-time data that is generated on Twitter.
+Twitter limits the streaming interface to randomly sample about 1\% of all of the real-time data that is generated on Twitter.
 This sampling rate was adequate to collect around three million tweets within a few days.
@@
 
-Looking first at only a single dataset slice, model accuracy for all users in that slice was analyzed.
 A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDsStd} and \ref{figPriorTwitterSliceDsStd} respectively.
@@
 So a best-fit value of \num{.4} was used for the optimized learning model instead of the \num{.43} average found from the dataset slices.
+Aggregate accuracy was computed by taking the mean of every user's model accuracy across all users in all dataset slices.
+
 Accuracy for both sites is included in Figure \ref{figPriorAcc}. 
-Aggregate accuracy was computed by taking the mean of every user's model accuracy across all users in all dataset slices.
+Accuracy for the standard form (\numNoZero{.33}) is higher than the optimized learning form (\numNoZero{.28}).
+This makes sense given that the standard form does not assume equal presentation rate for each chunk, and hashtags on Twitter in particular show trends of high rate of use and then shift to low rate of use.
+The standard form of $B_{i}$ is able to more accurately account for these trends.
 
@@
 
-Accuracy for the standard form (\numNoZero{.33}) is higher than the optimized learning form (\numNoZero{.28}).
-This makes sense given that the standard form does not assume equal presentation rate for each chunk, and hashtags on Twitter in particular show trends of high rate of use and then shift to low rate of use.
-The standard form of $B_{i}$ is able to more accurately account for these trends.
-
 ACT-R sets the default decay rate value to \num{0.5}, and the best-fit value for the standard form is higher (\num{0.8}) for these datasets.
-The figure shows that although model accuracy does slightly improve for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
-that percentage improvement is not as large as the improvement gained from using the standard form compared to the optimized learning form.
+The figure shows that although model accuracy does slightly improve (\num{0.325} to \num{0.335}) for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
+that percentage improvement is not as large (\num{0.28} to \num{0.33}) as the improvement gained from using the standard form compared to the optimized learning form.
 Nonetheless, model accuracy is highest when both the standard form is used and the decay rate value is increased.
@@
 For the StackOverflow randomly-sampled dataset, 5 runs of 100 posts were used,
-since the results for StackOverflow are more stable and retrievals take more wall-clock time for this dataset due to the size of the co-occurrence matrix.
+since the results for StackOverflow were more stable and retrievals took more wall-clock time for this dataset due to the size of the co-occurrence matrix.
 All runs contained different posts, and all posts used for runs were not used to generate the co-occurrence matrix for the two retrieval models.
@@
 
+The method of handling stop words has been shown to influence model accuracy \parencite{Sahlgren2008,Stanley2013,Jones2007}, so this was the first architectural concern that was examined.
 ``Stop words'' (e.g., ``the,'' ``a,'' ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
-The method of handling stop words has been shown to influence model accuracy \parencite{Sahlgren2008,Stanley2013,Jones2007}, so this was the first architectural concern that was examined.
 \textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
@@
 
-The effect sizes when comparing no stop-word removal technique to either the entropy or frequency techniques are quite large.
+There is a large improvement in model accuracy when either an entropy or frequency technique is used to handle stop words.
 This shows the importance of correctly identifying and handling stop words when building co-occurrence matrices for large-scale natural language datasets.
@@
 co-occurrence matrices for a range of smaller corpus sizes were built.
-Both the random permutation and Bayesian models were tested on all corpus sizes.
 
@@
 However, this means that ACT-R's prior base-level learning component (which is a log-odds value) is added to a correlation value, which may not be mathematically appropriate.
-As a fourth architectural concern, another method to combine the prior and context components for the random permutation model was explored.
+As a third architectural concern, another method to combine the prior and context components for the random permutation model was explored.
 

### 2014-12-05 docs/src/public.tex
foo

--- 2014-11-25 writeups/dissertation/dissertation.tex
@@
-\documentclass[man,floatsintext,donotrepeattitle]{apa6}
-%\documentclass[man,donotrepeattitle]{apa6}
-
-% http://tex.stackexchange.com/questions/279/how-do-i-ensure-that-figures-appear-in-the-section-theyre-associated-with
-\usepackage{placeins}
-
-\usepackage[nodayofweek]{datetime}% http://ctan.org/pkg/datetime
-\usepackage{xfrac}
-\usepackage{textcomp}
-\usepackage[american]{babel}
-\usepackage[utf8]{inputenc}
-
-% Single space after period (Mike prefers)
-\frenchspacing
-
-% http://tex.stackexchange.com/questions/36423/random-unwanted-space-between-paragraphs
-\raggedbottom
-
-% Changing horizontal space between subsubsection title and text, without loading titlesec package.
-% Not loading titlesec because it clashes with apa6 package.
-% Took base \@startsection command from apa6 .cls file, and tweaked horizontal spacing. 
-% http://tex.stackexchange.com/questions/22576/redefining-sectioning-commands
-% http://www.tex.ac.uk/cgi-bin/texfaq2html?label=atsigns
-% http://tex.stackexchange.com/questions/60127/subsubsection-description-and-title-not-separated
-\makeatletter
-\renewcommand{\subsubsection}{%
-  \@startsection
-  {subsubsection}%
-  {3}%
-  {\parindent}%
-  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-  {-.4em}%
-  {\normalfont\normalsize\bfseries\addperi}}
-\renewcommand{\paragraph}{%
-  \@startsection
-  {paragraph}%
-  {4}%
-  {\parindent}%
-  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-  {-.4em}%
-  {\normalfont\normalsize\bfseries\itshape\addperi}}
-\makeatother
-
-% ref: http://tex.stackexchange.com/questions/28516/how-to-change-the-title-of-toc
-\addto\captionsamerican{% Replace ``english'' with the language you use
-  \renewcommand{\contentsname}%
-  {Table of Contents}%
-}
-
-%\usepackage[compact]{titlesec}
-\usepackage{tabu}
-\usepackage{amssymb,amsmath}
-\usepackage{setspace}
-
-% ref: http://tex.stackexchange.com/questions/48509/insert-list-of-figures-in-the-table-of-contents
-\usepackage[nottoc,numbib]{tocbibind}
-
-\usepackage[hyphens]{url}
-
-% ref: http://tex.stackexchange.com/questions/73862/how-can-i-make-a-clickable-table-of-contents
-\usepackage{hyperref}
-\hypersetup{
-  colorlinks,
-  citecolor=black,
-  filecolor=black,
-  linkcolor=black,
-  urlcolor=black
-}
-%\hypersetup{linktocpage}
-
-%\usepackage{breakurl}
-\usepackage{subfigure}
-\usepackage{graphicx}
-
-\usepackage[group-separator={,}]{siunitx}
-\newcommand{\numNoZero}[1]{{\sisetup{add-integer-zero=false}\num{#1}}}
-
-\RequirePackage[l2tabu, orthodox]{nag}
-\graphicspath{{./figures/dissertation/}} % Specifies the directory where pictures are stored
-
-\usepackage{csquotes}
-\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber,doi=false,uniquename=false]{biblatex}
-\DeclareLanguageMapping{american}{american-apa}
-
-\addbibresource{bibliography.bib}
-
-% Needed to ensure periods are placed at the end of every bibliography entry in the references section
-\AtEveryBibitem{\clearfield{doi}}
-
-% http://tex.stackexchange.com/questions/55780/disable-month-in-biblatex-bibliography
-\AtEveryBibitem{\clearfield{month}}
-\AtEveryCitekey{\clearfield{month}}
-
-% Make sure only urls that are printed in references section are for webpages
-\AtEveryBibitem{
-  \ifentrytype{misc}{}{
-    \clearfield{url}
-  }
-}
-
-% ref: http://tex.stackexchange.com/questions/128592/remove-backslashes-from-url-fields-in-bib-entry
-\DeclareSourcemap{
-  \maps[datatype=bibtex]{
-    \map{
-      \step[fieldsource=url,
-	match=\regexp{\\},
-      replace=\regexp{}]
-    }
-  }
-}
-
-% Do not let biblatex change any of the casing in the bibliography file
-\DeclareRobustCommand{\MakeSentenceCase*}[1]{{#1}}
-
-% If the toc is too detailed, try limiting the depth of printed sections in the TOC:
-\setcounter{tocdepth}{2}
-
-\title{Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow}
-%\shorttitle{Comparing memory models of tag retrieval}
-\shorttitle{}
-
-\author{Clayton Stanley}
-\affiliation{Rice University}
-
-\leftheader{Beitzel}
-
-\def\committee{
-  Michael D. Byrne, Chair \\
-  Professor, Psychology and Computer Science \and
-  Phillip Kortum \\
-  Assistant Professor, Psychology \and
-  Devika Subramanian \\
-  Professor, Computer Science and Electrical Engineering \\
-}
-
-% Replace the title page in the apa6 style sheet with the title page from the Rice template
-% The title page style was taken from Rice University LaTeX PhD template: http://scholarship.rice.edu/handle/1911/21747
-% The following code for the abstract page was taken from the apa6 \maketitle command.
-\makeatletter
-\def\makeDisCover{
-  \def\maketitle{
-    \begin{titlepage}
-      \singlespacing
-      \null
-      \vfill
-      \vbox to 0pt{\vss\vbox to 8.75in{\parskip 0pt \parindent 0pt\centering
-	{\uppercase\expandafter{Rice University} \par} \vskip 0pt plus 3fil
-	{\bf \@title \par} \vskip 0pt plus 1.3fil
-	{by \par} \vskip 0pt plus 1fil
-	{\bf \@author \par} \vskip 0pt plus 3fil
-	{\sc A Thesis Submitted \\
-	  in Partial Fulfillment of the \\
-	Requirements for the Degree \par} \vskip 0pt plus 1.1fil
-	{\bf Doctor of Philosophy \par} \vskip 0pt plus 4.5fil
-	\begingroup
-	\leftskip .45\textwidth \rightskip 0pt plus 1fil
-	{\sc Approved, Thesis Committee: \par} \vskip 2.5em
-	\moveright\leftskip\vbox{\hrule width \leftskip}
-	\def\and{\par \vskip 2.25em minus 1em\moveright\leftskip\vbox{\hrule width \leftskip}}
-	\committee
-	\par \endgroup
-	\vskip 0pt plus 2fil
-	{Houston, TEXAS \par} \vskip 0pt 
-	{December \number\year \par}
-      }}
-      \doublespacing
-      \newpage
-  %BDB\hyphenpenalty 10000
-      \fussy
-      \@ifundefined{@abstract}{}{%
-	\thispagestyle{empty}
-	\section{\normalfont\normalsize \abstractname}% BDB
-	\begingroup
-	\parskip \z@ \parindent \z@ \par
-	\null \vskip .1in \centering
-	\@title
-	\vskip .3in
-	{by \par}
-	\@author \par
-	\vskip .5in
-      \endgroup
-      \@abstract\par% BDB
-      \@ifundefined{@keywords}{}{%
-	\setlength{\parindent}{0.4in}% BDB
-	\indent\textit{\keywordname:} \@keywords%
-      }%
-      \thispagestyle{empty}
-      \newpage
-    }
-    \@ifundefined{def@donotrepeattitle}{
-      \section{\protect\normalfont{\@title}}
-    }{}%
-    \raggedright%
-  \end{titlepage}
-  \setlength{\parindent}{0.4in}%
-}}
-\makeatother
-
-\makeDisCover
-
-%\keywords{}
-
-\begin{document}
-
-\pagenumbering{roman}
-
-\abstract{
-  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
-  The tasks of choosing a hashtag for a tweet and tagging a post on StackOverflow were framed as declarative memory retrieval problems.
-  Two state-of-the-art cognitively-plausible declarative memory models were evaluated on how accurately they predict a user's chosen tags: an ACT-R based Bayesian model and a random permutation vector-based model.
-  Millions of posts and tweets were collected, and both declarative memory models were used to predict Twitter hashtags and StackOverflow tags.
-  The results show that past user behavior of tag use is a strong predictor of future behavior.
-  Furthermore, past behavior was successfully incorporated into the random permutation model that previously used only context.
-  Also, ACT-R's attentional weight term was linked to a common entropy-weighting natural language processing method used to attenuate low-predictor words.
-  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparably to the Bayesian model without including word order.
-  This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
-  Finally, model accuracy was moderate to high for the tasks, which supports the theory that choosing tags on StackOverflow and Twitter is primarily a declarative memory retrieval process. 
-  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
-  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
-}
-
-\note{\today}
-
-\maketitle
-
-\thispagestyle{empty}
-\section{Acknowledgements}
-
-I would first like to thank my advisor, Mike Byrne, for all of his feedback and suggestions for this dissertation research
-and also his excellent guidance throughout my time as a graduate student.
-I would also like to thank my parents, Mike and Patsy, for the positive influence that they have provided me throughout my life.
-Finally, I would like to thank my brothers Brendan and Justin, sister Kellen, extended family, and friends, for all of their support during my time in the Rice Psychology graduate program.
-
-This work was supported in part by the Gertrude Maurin Fund.
-The resources were used to purchase high-performance computing hardware and the research would not have been possible without this crucial equipment.
-
-I would also like to thank the creators of a few open-sourced software packages that were heavily leveraged to do this research:
-The majority of the modeling and analysis was done with the R statistical computing environment \parencite{RCoreTeam2014} and two important R packages:
-data.table \parencite{Dowle2014} was leveraged to implement the models efficiently, and ggplot2 \parencite{Wickham2009} was used to produce the visualizations.
-
-\newpage
-
-\thispagestyle{empty}
-\tableofcontents
-\newpage
-
-\thispagestyle{empty}
-\listoftables
-\newpage
-
-\thispagestyle{empty}
-\listoffigures
-\newpage
-
-\pagenumbering{arabic}
-
-\section{Chapter 1: Introduction}
-
-The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human cognitive processing.
-The declarative memory component in ACT-R is a Bayesian statistical model.
-This is the same type of co-occurrence-based Naive Bayes statistical model that is prominent in the more general machine learning field.
-When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
-The primary reason for this is that the stability of statistical models increase as sample size increases.
-
-However, it is much less common in the cognitive modeling community to use a large number of observations to build up declarative memory when creating ACT-R models.
-If a cognitive modeler is using the declarative memory component of ACT-R,
-it is common to either construct the co-occurrences by hand, or use a small training set of data to initially build up the declarative memory system.
-This approach can and has worked well.
-Essentially, the modeler is only including the observations in declarative memory that are relevant to the specific task that they are studying.
-
-However, this approach is problematic for two main reasons:
-First, it means that the declarative memory system for every developed model must be hand crafted and customized for that model, which takes enormous time and effort.
-Second, it means that modelers are continuing to use small sample sizes for the declarative memory components in their models.
-
-This has important implications for understanding declarative memory:
-First, it is difficult to verify and explore modifications to the declarative memory architecture because the datasets are not large enough or rich enough to rigorously test modifications to the theory.
-It also means that we do yet fully understand how well the ACT-R declarative memory theory scales when the number of chunks begins to approach the number of chunks that are actually stored in human declarative memory.
-
-There is, however, an opportunity to change this approach.
-With the massive growth of human-created content on social media sites (e.g., Twitter, StackOverflow, Wikipedia), and public access to many of those databases,
-it is now possible to evaluate psychological theories of declarative memory on large-scale real-world tasks where the theory can be tested on hundreds of millions to billions of data points.
-This will provide a way to test if these models scale, and evaluate the psychological validity of these models as the number of observations used increases by orders of magnitude.
-
-A few researchers have started to use these large databases to evaluate how well declarative memory models scale,
-for both ACT-R Bayesian models \parencites{Fu2007,Pirolli2003,Stanley2013,Douglass2010} and vector-based memory models \parencites{Jones2007,Rutledge2008,Recchia2010,Sahlgren2008}.
-
-The work in this study expands on previous work that has tested retrieval models on large-scale datasets.
-It focuses specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and vector-based) can be improved by testing these models on two real-world tagging tasks.
-One primary objective is to compare the random permutation vector-based model to ACT-R's declarative memory theory when large datasets are used for each model. 
-Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
-For example, it is currently unexplored how a vector-based model can be modified to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
-Large-scale datasets were used throughout the model development and evaluation process.
-This provides an ideal testing environment for constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.
-
-\subsection{The Task: Predicting User-Generated Tags}
-
-Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp are composed almost entirely of human-created content.
-The amount of user content on these sites is quite large, and continues to grow.
-Users of Twitter's microblogging service, for example, generate half a billion tweets per day \parencite{TwitterReport2012}.
-However, a single user of a social media site is most likely interested in only a small fragment of this large amount of information.
-One general question to support the user in this domain is:
-How can we quickly and effectively connect users to the content that they care about?
-
-A growing number of users on social media sites are focusing on and looking for recently-created information when performing a search \parencite{Jansen2011}.
-In other words, users on social media sites want the fresh \emph{stream} of information about a particular topic.
-This can certainly be seen in the main user interface views that these social media sites provide:
-Twitter's homepage showing followees' tweets, Facebook's homepage showing friends' posts, and StackOverflow's daily digest of posts for particular tags.
-
-So how can we ensure that users are connected to the information streams that they care most about?
-One promising approach is to model and predict the user's goals when using and searching these social media sites \parencite{Rose2004}.
-If we have a better understanding of a user's goals, then we can tailor the information streams to the content that is most in line with these goals.
-For example, Twitter might suggest other users and hashtags for the user to follow, Facebook could suggest pages to like, Yelp could recommend businesses to check out.
-Researchers are currently looking for ways to automate the goal-identification task for more general search queries like those generated for Google \parencites{Jansen2008}{Lee2005}.
-
-However, social media sites have an advantage over traditional search engines for identifying user goals: human-created content.
-Twitter has more than just the tweet that a user is currently composing to try and identify that user's goals.
-The site also has all of the user's previous content, and can use that information to provide a much better prediction of a user's goals.
-Further, sites like Twitter, StackOverflow, Google+, and Facebook support ways for the user to explicitly identify their precise goals when creating content: hashtags.
-These user-created hashtags are a key indicator of a user's goals on a social media site.
-
-These hashtags relate to a user's goals and interests because they are a form of human-based document tagging \parencite{Chang2010}.
-Further, people are using hashtags as a way to create information streams on social media sites \parencite{Kwak2010}.
-Twitter users, for example, commonly create and use hashtags for upcoming political events that interest them, such as debates and races \parencite{Diakopoulos2010}.
-So if we can predict what hashtags a user is interested in, we have a better understanding of the information streams that they care about.
-We can use that information to correctly suggest hashtags that they have previously used when they are generating new content.
-
-Further, if a particular information stream has multiple hashtags associated with it (common for Twitter), these semantically similar hashtags can also be recommended to users.
-So we can tailor the system to suggest hashtags that the user may not yet know about but are of likely interest to them and growing in popularity across the site.
-We can help aid in their discovery of fresh and relevant information that matches their goals on the site.
-
-\subsection{Research Questions}
-
-The core research question motivating this proposal is:
-What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
-This question can be tested by generating a prediction for the most likely used tag whenever a user is about to generate a tag, and then comparing the user's chosen tag to the model's prediction.
-This process can be framed as a memory retrieval problem.
-The process of suggesting a relevant tag to a user can be thought of as a memory retrieval request for a tag, given prior tag use and current context.
-That is, each user on a social media site has created content that contains a set of tags within the context of each post.
-
-Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010} and StackOverflow tag use \parencites{Stanley2013}
-Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
-At least two of the current memory retrieval models are based on a co-occurrence methodology:
-ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.
-
-\subsubsection{StackOverflow and Twitter}
-
-These two memory systems were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
-Twitter is a microblogging service where users create 140 character ``tweets'' and broadcast those messages out to the users who follow their posts.
-A user is free to annotate important concepts or words in the tweet by preceding words with ``\#,'' thereby making the word into a hashtag.
-After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
-So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
-Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample1-crop.pdf}}}
-    \hfill
-    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample3-crop.pdf}}}
-    \hfill
-    \vfill
-    \hfill
-    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample4-crop.pdf}}}
-    \hfill
-    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample5-crop.pdf}}}
-    \hfill
-    \caption{Example tweets on twitter.com}
-    \label{figTweetExample}
-  }%
-\end{figure}
-
-Some example tweets on Twitter are included in Figure \ref{figTweetExample}.
-Note how users both add hashtags to the end of a tweet as a summary of the message and emphasize important concepts words in the tweet by turning them into hashtags.
-That is, hashtags can be both intermixed within the words of the tweet and added to the end of the message.
-
-The two memory retrieval models were also tested on tag use for StackOverflow posts.
-StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
-A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
-The tags chosen by a user on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
-Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{stackoverflowPostExample1-crop.pdf}}}
-  \caption{Example post on stackoverflow.com}
-  \label{figSOExample}
-\end{figure}
-
-An example StackOverflow post is included in Figure \ref{figSOExample}.
-The site requires that for each post the user must associate at least one tag with the post, and the average number of tags per post is around 3.
-The author used 4 tags for this post: \emph{javascript}, \emph{html}, \emph{internet-explorer}, and \emph{xhtml}.
-
-The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and I am interested in retrieval models that generalize across tasks.
-However, they are also similar on several accounts:
-Both domains have amassed large amounts of user-created content, users generate tags for posts when creating content on the site (hashtags for Twitter and tags for StackOverflow),
-and the user data from the datasets are publicly available for analysis.
-
-Also, studying user-based tag generation on these sites has relevant real-world application.
-Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these popular sites.
-These systems can also help newer users by recommending proper tags for their newest content.
-On the StackOverflow site, for example, experts often subscribe to specific tags that interest them, and receive a daily digest of posts that tagged with those tags.
-Helping the user properly tag posts on StackOverflow ensures that right community sees the post, which greatly increases the chance that the question on the site will be answered quickly and correctly.
-For Twitter, hashtags represent streams of information that are possibly interesting to the user.
-Having a recommendation system that can suggest relevant hashtags to the user provides a way for the user to connect to and discover new information streams that interest them.
-Also, by suggesting relevant hashtags the system can help make sure a user's content is seen by other users who are interested in the same type of information.
-
-\subsubsection{Comparison Between ACT-R and Vector-Based Memory Systems}
-
-ACT-R's declarative memory and vector-based memory systems are substantially different in their mathematical structure.
-Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., \cite{Rutledge2008} modeling the ``fan effect'').
-
-An advantage of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
-Word order information on sites like Twitter may contain highly predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
-Word order has been incorporated into ACT-R sentence-parsing models previously, but this has not been done using declarative memory.
-Rather, production rules were defined that ensured that the model followed a set of rules (e.g., left-to-right reading) when parsing sentences \parencite{Lewis2006}.
-It is certainly true that at least some production rules are needed when choosing words and creating content. 
-However, it is unexplored if some of the process of sentence construction and word selection can be represented in the ACT-R declarative memory system.
-
-ACT-R's declarative memory system places strong constraints on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2007}.
-This decay rate equation formalizes how recency and frequency of memory use relate to the prior probability that a memory will be retrieved.
-However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.
-
-\subsubsection{Lifetime of a Hashtag for a Specific User}
-
-Prior research has examined the growth and decay cycle of hashtag use across users \parencite{Tsur2012}.
-However, much less is known about hashtag lifetime for individual users.
-It may not necessarily follow that a hashtag's life cycle for individual users matches the life cycle across users.
-Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into a specific user's prior likelihood of hashtag use.
-
-\subsubsection{User-Customized Hashtag Prediction}
-
-The end goal of this work is to identify a memory retrieval model that can suggest relevant hashtags to users when they wish to retrieve one.
-These hashtags should be customized to their specific interests and relevant to the content of the post that they are currently creating.
-This will undoubtedly require a combination of two primary model components:
-[1] a user's prior likelihood of choosing a particular hashtag, given their previous hashtag use, and [2] the likelihood that a particular hashtag is related to the context of the post being created.
-
-A user's prior hashtag use will be an essential model component for domains such as Twitter, where the number of possible hashtags is effectively infinite.
-In these domains, the model can utilize a user's hashtag history as a way to prune the infinite space of possible hashtags, and generate a much smaller and more manageable set for prediction.
-However, not only should the model properly take into account a user's prior hashtag use, but it should also generate new hashtag predictions when a user is most likely choosing a hashtag they have never used before.
-The end goal is to have a model that can properly balance the user's tag history with the contextual cues in the post in order to generate an accurate prediction.
-
-\subsection{Roadmap}
-
-This document is separated into 7 chapters.
-Chapter 1 introduced the declarative memory research and motivated why it is interesting both from a theoretical and applied perspective to evaluate declarative memory models on large-scale datasets.
-The task to test the models was also introduced: tagging of user-generated content on StackOverflow and Twitter.
-Chapter 2 describes the vector-based and Bayesian declarative memory models that will be evaluated in this research, and shows how these models are related to the other common declarative memory models that are studied.
-Chapter 3 presents the overall method for testing the models on the StackOverflow and Twitter datasets.
-Chapter 4 shows the results for the first test of the models, where the model term for past user behavior is examined in isolation, and tags are predicted based purely on a user's past tagging history.
-Chapter 5 adds the context component to the models, and shows how model accuracy is influenced by changing various architectural components of each model.
-Chapter 6 describes model results when prior and context are combined and tested on both of the dataset subsets used for Chapter 4 and Chapter 5.
-Finally, Chapter 7 provides a general discussion of the overall findings.
-
-\clearpage
-\section{Chapter 2: Theoretical Background}
-
-The ACT-R declarative memory theory and the vector-based retrieval models will be described.
-The Latent Semantic Analysis (LSA) theory is closely related to vector-based retrieval models, so an overview of LSA will also be included.
-The theories underlying the ACT-R and vector-based models will be compared.
-Also, different tagging models have already been applied to various recommendation systems to tag user-generated content.
-Many of these models are based on the same Bayesian statistical framework that is the foundation of the ACT-R declarative memory theory.
-So these systems will be described and the models used will be discussed within the context of Bayesian statistical models, LSA approaches, and the ACT-R declarative memory theory.
-
-\subsection{ACT-R Declarative Memory Theory}
-
-ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
-The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
-The equations that make up this system were derived from a rational analysis of declarative memory retrieval.
-That is, given the task of retrieving a chunk of information from declarative memory,
-the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
-what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
-Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of needing retrieval again, given the prior history of exposure to the chunk.
-These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.
-
-\subsubsection{ACT-R DM Model}
-
-A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACTRModel}.
-
-\begin{table}[!ht]
-  \caption{ACT-R declarative memory model}
-  \label{tabACTRModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Activation &	 	$A_{i} = B_{i} + \sum_{j \in c}^{} W_{j} S_{ji}$ \\
-      Attentional Weight &	$W_{j} = \frac{W}{n}$ \\
-      Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
-      Constant Base Level &	$B_{i} = \frac{p_{i}}{1-p_{i}}$ \\
-      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
-      Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base-level activation ($B_{i}$) and strength of association ($S_{ji}$).
-The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
-Each term in Table \ref{tabACTRModel} will be discussed in greater detail to follow.
-
-\subsubsection{Base-Level Activation}
-
-Base-level activation reflects the log prior odds of needing an observed chunk again.
-The default way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
-This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
-Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
-However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
-In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its frequency and recency of use.
-
-For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable and do not change within that window.
-Programming language popularity over the past few years is a reasonable example of this.
-Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., Clojure's growth), it is not the case that the changes have been drastic.
-C\#, Python, and Java are still a few of the most popular languages, Common Lisp has not gained or lost much ground in popularity over the past few years.
-
-Further, if this is a reasonable assumption to make, it greatly simplifies the computation of the base-level activation for chunks, and can turn the computation into a tractable problem for large datasets.
-In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log-odds ratio of exposure to that chunk compared to exposure to all other chunks.
-For example, if the StackOverflow tag \emph{PHP} has been used four times as often as the tag \emph{Common Lisp}, then the prior odds of needing \emph{PHP} again is $\frac{.8}{.2}=4$ times that of \emph{Common Lisp}.
-
-\subsubsection{Strength of Association}
-
-Strength of association ($S_{ji}$) reflects the amount of log-odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
-Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
-Context for the Twitter domain are the words in a tweet.
-Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
-the likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
-For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
-So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
-This assumption was described when deriving the $S_{ji}$ equation \parencite{Anderson1989}, and used in recent research working with large-scale datasets \parencites{Stanley2013,Farahat2004,Douglass2010}.
-
-Using this assumption for large datasets ($S_{ji} = log \frac{p(i|j)}{p(i)}$), the interpretation of the context-adjusted odds ratio becomes much simpler.
-If this ratio for a particular tag and context chunk is greater than one, then the log is positive, which means that this context has been observed more often with this tag than in general.
-If this ratio is less than one, then the log is negative, which provides a negative adjustment of total activation since this context has been observed more often in general than with this particular tag.
-
-\subsubsection{Connection to Pointwise Mutual Information}
-
-Pointwise Mutual Information (PMI) is another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
-It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
-The index has been used to accurately and efficiently measure strength of association information for large-scale datasets \parencite{Budiu2007,Farahat2004}.
-The PMI index is presented as Equation \eqref{eqPMI}.
-
-\begin{equation}
-  \label{eqPMI}
-  \mathit{PMI}(y,x) = log \frac{p(y|x)}{p(y)}
-\end{equation}
-
-Further, this PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
-Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.
-That is, the PMI equation \emph{is} the simplified form of ACT-R's strength of association index ($S_{ji}$).
-
-\subsubsection{Attentional Weight}
-
-The attentional weight term allows for some contextual components to be weighted more than others when the total $S_{ji}$ activation is computed.
-This reflects the fact that the declarative memory system has limited attentional resources ($W$) that must be spread across all of the items in context.
-Most ACT-R models simply weight each contextual element equally (e.g., all words in a StackOverflow post have equal weight), so each word's attentional weight is simply normalized by the number of items in context.
-
-\subsubsection{Scaling the Equations}
-
-The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
-As an applied concern, it is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
-Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
-If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.
-
-However, scale is also a theoretical concern.
-The declarative memory system is tasked to quickly return a single chunk of information that is stored alongside millions to billions of other unique chunks given only the current context and past history of use.
-To do this task, it seems plausible that a model of declarative memory must use a representation that produces accurate retrievals even when large amounts of observations are included.
-Further, it seems plausible that the time to compute an activation value by a computer is correlated with the time to compute an activation value by the brain.
-So any model of declarative memory should also be reasonably computationally efficient even with large co-occurrence datasets.
-
-ACT-R's strength of association ($S_{ji}$) and the equivalent pointwise mutual information index ($\mathit{PMI}_{ji}$) have been shown to scale well for large corpora.
-\textcite{Douglass2010} implemented the large-scale version of $S_{ji}$ in Erlang.
-The wall-clock time required for a declarative memory retrieval request scaled linearly with the size of declarative memory with this implementation.
-Retrieval times were around one second with the largest declarative memory that contained over one million associations (co-occurrence pairs).
-
-The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} was developed to predict user navigation of web-based hyperlinks (e.g., choosing between a set of results from a search query).
-The theory uses the large-scale version of $S_{ji}$ to measure the relatedness between a user's current goals and the current context.
-It was shown to scale well when working with a large-scale search query database for internet search.
-
-As one of the largest tests for the PMI index, \textcite{Farahat2004} generated the co-occurrence matrix from a 118GB local database of 10 million pages of the Stanford WebBase Project.
-\citeauthor{Farahat2004} compared strength of association values using the PMI index and similarity ratings derived from human raters.
-Even at this large dataset size, the ACT-R declarative memory theory still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.
-
-\subsection{Latent Semantic Analysis Memory Theory}
-
-Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
-Both ACT-R's Bayesian declarative memory model and LSA compute activations based on a co-occurrence matrix.
-However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
-The process starts with a set of N documents.
-A word by document co-occurrence matrix is built by counting the number of times each word appears in each document.
-The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
-This process, described in more detail to follow, is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
-The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
-
-Usually information is lost when the data are restructured into a smaller set of latent dimensions.
-However, it was shown that this process actually improves prediction and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
-\textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
-Model results were on par with a large sample of applicants to American universities from non-English speaking countries (\num{64.4}\% compared to \num{64.5}\%).
-Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
-Afterwards further reductions in the dimensionality of the data began to degrade performance.
-
-\subsubsection{Singular Value Decomposition}
-
-LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix.
-The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
-The best way to remove a dimension while maintaining the most variability is to look for words with highly similar co-occurrence patterns and deduce that these words represent the same concept.
-For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
-By pooling co-occurrence counts the model is able to generate better predictions with less data.
-
-\subsubsection{Word Order}
-
-Although the LSA model has been shown to perform well as a way to measure similarity between two terms, one of the main criticisms is that it is a ``bag of words'' model.
-This means that the model does not account for or pay attention to word order when building the base word by document matrix or consequently, when performing the SVD dimension reduction.
-Other models that can account for word order have been shown to perform better than ``bag of words'' models on tasks that do not even appear to require grammatical structure such as word order to do well.
-For example, \textcite{Jones2007} compared a vector-based model called BEAGLE directly to LSA, training and testing both models on the same corpus used in \textcite{Landauer1997}.
-The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the TOEFL task (\num{55.6}\% vs. \num{55.3}\%).
-However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to \num{57.5}\%.
-It seems reasonable to expect that task performance will improve much more on tasks where word order is used (e.g., generating a hashtag in the middle of a sentence when composing a tweet).
-
-\subsubsection{Scaling Issues}
-
-Another issue with LSA is that the SVD matrix decomposition technique is computationally expensive.
-As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
-If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
-However, with large human-created datasets such as Twitter, StackOverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
-\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
-Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
-This is one of the available corpora on the LSA website and is intended to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.
-
-When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-sized datasets, LSA performed much better (23\% vs. 60\%).
-This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model outperformed the full dimensionality model on TASA-sized datasets.
-However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
-\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
-This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
-Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to \num{73.8}\%, which is 10\% higher than the TASA-trained LSA.
-This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.
-
-\subsubsection{Cost of Incremental Updating}
-
-Another issue with LSA is that adding additional observations to a previously-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
-Since computing the SVD is the most computationally expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
-For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.
-
-\subsection{Vector-Based Memory Systems}
-
-``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., a chunk) as a vector.
-The representation of the concept is distributed across all elements of the vector.
-This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and viewing the distribution of co-occurrence counts across the rows in the column as the representation of that concept.
-However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.
-
-For vector-based systems, each word is represented by an environment vector $e_{i}$ that is nearly orthogonal to all other words' environment vectors.
-The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix, which is how vector-based systems represent information in a much more compact space.
-Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occurred with that $e_{i}$ environment vector.
-These memory vectors can contain environment vectors from position-independent word co-occurrence information ($c_{i}$) and word order information ($o_{i}$), which can be combined into a single representation ($m_{i}$).
-Over time, memory vectors accumulate a distributed representation of the most common environment vectors that co-occurred with them.
-
-\subsubsection{Retrieval Process}
-
-Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
-Decoding takes the memory vector for context and decodes it back into an environment vector.
-The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
-
-Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
-For example, assume that a user has written the word ``zend'' in a post, and she wants to assign a tag for the post.
-``zend'' has co-occurred with the tag \emph{PHP} many times previously, so the unordered memory vector for ``PHP'' ($c_{\mathit{PHP}}$) contains the unordered environment vector for \emph{zend} ($e_{zend}$).
-To determine the most correlated memory vector with context, a memory vector is created from the current context ($e_{zend}$).
-The cosine between that context memory vector and the memory vector for \emph{PHP} will be high, since the \emph{PHP} memory vector ($c_{\mathit{PHP}}$) often contains the context environment vector ($e_{zend}$).
-Since the cosine is high, the tag \emph{PHP} will most likely be returned as the vector that resonates highest with the current context.
-
-The resonance and decoding processes are inversions of each other.
-Usually both resonance and decoding will return similar rank orderings of most likely chunks given context.
-\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
-
-\subsubsection{Addressing Word Order}
-
-One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
-Representing word order is a case where both the encoding and decoding operations are applied.
-As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
-Simply counting co-occurrences of the two immediately preceding words and hashtags does not include the order information of the words in the representation.
-With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
-Order information is lost if a simply co-occurrence counting technique is used.
-
-With vector-based systems, order information is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
-This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
-Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
-This allows for retrieval requests such as ``playing \#$\Phi$,'' ``just read \#$\Phi$,'' and ``vector-based $\Phi$ systems.''
-
-For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
-Within this corpus, the phrase ``martin luther'' often occurred.
-When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated by convolving a 1-back location vector with luther ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
-Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin with and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
-The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.
-
-Using decoding, the word-order memory vector for martin is deconvolved to the right ($o_{martin} \circledast \Phi$) which returns the environment vector for luther ($e_{luther}$).
-Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which correlates highest with the word-order memory vector for luther ($o_{luther}$).
-Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$.''
-
-New vectors generated by circular convolution are uncorrelated with all other environment vectors and have the same length.
-This allows the word order information and unordered context information for a chunk to be aggregated and represented in the same memory vector.
-Quite simply, the different types of memory vectors can be summed to create a single memory vector representation for each chunk (i.e., $m_{i} = c_{i} + o_{i}$).
-
-\subsubsection{Addressing Scalability}
-
-Vector-based representations compute chunk likelihoods by performing a correlation operation directly on the memory vectors.
-Each memory vector can be thought of as a column on a context word by tag co-occurrence matrix used for a Bayesian representation.
-Further, the number of rows for a vector-based representation are fixed and set much smaller than the number of rows in the full word by tag co-occurrence matrix.
-Since Bayesian representations have been shown to scale to over hundreds of millions of word by tag co-occurrences \parencite{Stanley2013}, the compressed vector-based representations should scale as well or better.
-
-\subsubsection{BEAGLE}
-
-\textcite{Jones2007} created the BEAGLE vector-based memory representation.
-Each memory vector ($m_{i}$) is a summed representation of unordered and ordered word co-occurrences ($c_{i} + o_{i}$).
-This model uses circular convolution and circular correlation as a way to encode and decode word order.
-BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (\num{55.6}\% compared to \num{55.3}\%).
-Further, it was shown that incorporating word order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
-This model provided an efficient algorithm for using a single representation for unordered and ordered information that showed performance similar to if not better than LSA.
-
-\subsubsection{Random Permutation Model}
-
-In order to represent word order information, a function must be used that converts an environment vector for a set of words and positions (e.g., ``stack $\Phi$'') into a new uncorrelated environment vector.
-The BEAGLE model uses circular convolution and circular correlation as the encoding and decoding operations.
-However, that is certainly not the only operation that can generate new uncorrelated vectors from a set of memory vectors and positions.
-\textcite{Sahlgren2008} used a simple random permutation method for this operation.
-
-A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
-With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few one and negative one values in random locations. 
-In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is shifted to the left one position.
-This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
-
-\begin{table}[!ht]
-  \caption{Random permutation model}
-  \label{tabRandPermModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Activation &		$A_{i} = r(m_{C},m_{i})$ \\
-      Memory Vector &		$m_{i} = \sum_{i \in all past} c_{i} + \sum_{i \in all past} \sum_{l \in locations} o_{i,l}$ \\
-      Unordered Context &	$c_{i} = e_{i}$ \\
-      Ordered Context &		$o_{i,l} = e_{i^{-l}}$ \\
-      Context Memory Vector &	$m_{C} = \sum_{i \in C} c_{i} + \sum_{i \in C} \sum_{l \in locations} o_{i,l}$ \\
-      Environment Vector & 	$e_{i} = rand$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow.''
-The ``stack'' environment vector ($e_{stack}$) is shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).
-That environment vector is then added to the memory vector for ``overflow'' ($m_{overflow} = m_{overflow'} + e_{stack^{-1}}$), so that ``overflow'' remembers that ``stack'' has preceded it in a sentence.
-
-A resonance retrieval for a word following ``stack'' (``stack $\Phi$'') is done by creating a memory vector out of all information in context.
-In this case, assume only a single piece of order information is available, that ``stack'' precedes the word to retrieve.
-So the memory vector for context is the environment vector for one-back ``stack'' ($m_{C} = e_{stack^{-1}}$).
-The memory vector for ``overflow'' ($m_{overflow}$) will be correlated the highest with this context vector ($m_{C}$) since ``overflow'' contains this vector.
-
-\subsubsection{Comparison between BEAGLE and Random Permutations}
-
-Random permutations uses one of simplest operations possible to create a new environment vector.
-This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
-Further, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.
-
-\textcite{Recchia2010} found that random permutations could use smaller vectors than circular convolution approaches (e.g., BEAGLE) to produce equivalent performance (better compression),
-and also scale to larger datasets (better computational efficiency).
-The vector length by performance tradeoff was tested on a simple paired associative retrieval task. 
-Retrieval accuracy for the random permutation vector-based models remained high relative to the circular convolution model even after reducing each environment vector's length from \num{2048} to \num{1024}.
-This indicates that vector-based representations using random permutations compress better than those using circular convolution.
-
-To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFL) assessment (among other tasks).
-The random permutation model was trained on \num{2.33} GB corpus.
-The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
-The random permutation model performed as well or slightly better than the BEAGLE model when trained on the smaller corpus.
-The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
-These results make a strong case to favor random permutations over circular convolution when implementing vector-based memory representations.
-
-\subsubsection{Connection to LSA}
-
-Both vector-based models and LSA calculate word similarity on a reduced matrix.
-For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors is much less than the number of words.
-For the StackOverflow and Twitter datasets, the number of documents are the number of different tags.
-
-For random permutations, each tag is represented by a compressed vector with a length much less than the number of words.
-Note that if the length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular row in the column vector.
-If done in this way, building a matrix of all words' environment vectors by tags would recover the original word by document matrix \parencite{Kanerva2000}.
-However, each word's environment vector is compressed using a length much less than the number of unique words and represented by using a combination of a few one and negative one values at a unique combination of positions.
-
-Information is lost with this compression in both cases: by starting with a compressed representation with vector-based models or by reducing to a compressed representation with LSA.
-However, the way that the compression works in these two cases is not the same.
-LSA uses SVD to find words that have highly correlated tag occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
-A vector-based approach like BEAGLE or random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
-Nonetheless, random permutations have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFL \parencites{Sahlgren2008,Jones2007}.
-
-This is a counterintuitive result.
-It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
-This does not seem to be the case in testing however \parencites{Sahlgren2008,Jones2007}.
-Nonetheless, \textcite{Kanerva2000} suggested to perform an SVD reduction technique on a compressed random permutation vector.
-This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.
-
-\subsection{Comparison of Vector-Based Models and ACT-R}
-
-Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
-Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
-Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
-For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
-\citeauthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect.
-
-The fan effect describes a behavior of human memory where memory items are more difficult to retrieve if they have occurred in a larger number of different contexts.
-\textcite{Rutledge2008} used \textcite{Anderson1974}'s classic experimental paradigm to test the fan effect.
-In the task, participants are exposed to a set of person and place pairs (e.g., ``the hippie is in the park'').
-They are later asked to recall if a particular person and place pair was presented to them before.
-The main manipulation is the ``fan'' of the person and place for each pair.
-The fan is defined as the number of different contexts that each item in the pair is presented in.
-For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store,'' then ``hippie'' would have a fan of two.
-Recall accuracy decrease and time to recall increase as the sum total fan of the person and place increases.
-
-Both the ACT-R declarative memory theory and vector-based models produce the fan effect due to the effect that fan has on the co-occurrence matrix.
-As a particular word (e.g., ``hippie'') is seen in more and more contexts, the co-occurrence count of the word across all contexts is spread across multiple contexts.
-This means that when a retrieval for a context given the word occurs, multiple contexts have relatively high activation.
-This makes it more difficult to test if a signal (e.g., if a word appears in a specific context) is higher than the background noise (e.g., all other contexts that the word has appeared in),
-which makes the retrieval more uncertain.
-
-Both vector-based and Bayesian models produce the fan effect since they both are based on co-occurrences.
-However, vector-based models do further processing on that matrix, reduce its dimensionality, and use a correlation instead of a log odds value to compute the activation.
-Nonetheless, vector-based models still produce the fan effect because at their core they are still computing activations based on co-occurrences.
-
-\subsubsection{Base-Level Activation}
-
-Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the relative benefits of the two models. 
-In particular, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
-
-The frequency and recency of how content has been used in the past can be a strong predictor of the future importance of that information. 
-The recent content in tweets on a news site like Twitter should be more relevant, for example.
-\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
-They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.
-
-ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
-It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
-Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
-ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
-Vector-based models have focused on computing the associative strength of a chunk and specific contextual elements.
-It is unexplored how accuracy changes for these models once the information about a user's prior behavior (e.g., ACT-R's base-level activation component) is added.
-
-\subsubsection{Word Order}
-
-One strength of a vector-based approach over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
-With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
-To represent word order in a Bayesian model, an additional dimension could be added to the co-occurrence matrix that represents word position.
-However, adding word-order information to ACT-R's default Bayesian declarative memory model is not common.
-Instead, when word order information is used to model reading for example, production rules are defined so that words are read from left to right (e.g., \textcite{Lewis2006}).
-
-\subsubsection{Activation Calculation}
-
-Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
-Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
-These two computations are certainly not equivalent, even though they are most likely correlated with one another.
-
-Using a correlation can be problematic due to range effects in observed co-occurrences.
-If there are particular words for a tag that co-occur much more often than all of the other words with that tag, that single large value diminishes the power of all other words to contribute to the correlation calculation.
-This does not happen with Bayesian retrieval models because each pair of word and tag values is normalized by the total number of words that appeared with that tag and total number of tags that appeared with that word
-(see the strength of association equation in Table \ref{tabACTRModel}).
-The high-frequency words are often removed from co-occurrence matrix for vector-based models to reduce this effect.
-
-\subsection{Recommendation Systems}
-
-A model that predicts a user-chosen hashtag is a specific type of a recommendation system.
-General recommendation systems have grown in need with the growth of the web.
-Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
-Recommendation systems are used to tailor the information presented to what is most relevant to the user \parencite{Pazzani2007}.
-
-\subsubsection{The Netflix Prize}
-
-The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
-Netflix is an online subscription service where users can stream television episodes and movies.
-Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
-In order to provide a high quality user experience on the site, Netflix created a content recommendation service.
-This tool recommends to the user movies and television episodes that most likely match their specific interests.
-
-Netflix wanted to improve the accuracy of their recommendation system, so they held a competition where teams could try to improve on the accuracy of the service.
-The winning team utilized a statistical model that performed a type of factor analysis to identify important dimensions in the data (e.g., drama, action, cult classic) and place values for each user along each dimension.
-This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
-By using a factor analysis technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.
-
-\subsubsection{Recommending Followers on Twitter}
-
-Recommendation systems based on the content of user-created posts have been successful in recommending people on Twitter for a user to follow.
-\textcite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
-The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
-The profiles were created by generating a word frequency matrix of words by (user + followers + followees), and then providing that matrix as input into the Lucene text search engine framework.
-
-Lucene \parencite{McCandless2010} is a commonly-used open source software tool that measures the similarity between a query and candidate documents in a database.
-It uses the term frequency-inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
-This measure is the product of two terms: the frequency count of the word in the document and the log inverse of the ratio of the number of documents that the word appeared in to the total number of documents.
-The term frequency component measures the strength of association between a word and a document.
-The inverse document frequency component measures how much information is contained within a specific word, independent of document.
-
-When \citeauthor{Hannon2010}'s system recommended a set of individuals for a user to follow, it collected all of the words used in posts from that user, constructed a search query, and provided that query to Lucene.
-Lucene then returned the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
-Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user follows at 25\% accuracy.
-
-\subsubsection{Hashtags for Recommendation Systems}
-
-\textcite{Efron2010} showed how hashtags might be used as a type of recommendation system for Twitter.
-Users constructed a query of text that represented a search for a topic that they were interested in.
-The system then retrieved the most likely hashtags that were associated with that search query.
-The model is based on building a word co-occurrence matrix of words by hashtags, and then uses that matrix to assign a similarity score for each hashtag when provided with a search query.
-Participants graded how relevant the hashtags returned were to each search query.
-The results suggest that hashtags contain useful information that help classify the gist of a post. 
-Although this model does not predict the specific hashtags used when composing a tweet, a model like this could certainly be tailored for that purpose.
-
-\subsection{Hashtag Prediction}
-
-Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
-Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
-A tag-suggestion tool has also been deployed for the \url{tex.stackexchange.org} and \url{superuser.com} StackExchange sites.
-Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
-Recommendation models for StackOverflow, Google+, and Twitter will be described in more detail to follow.
-
-\subsubsection{StackOverflow}
-
-\textcite{Kuo2011} was the first to model tag use for StackOverflow posts.
-His set of models were originally designed for next-word prediction for large text document collections.
-\citeauthor{Kuo2011} examined the accuracy of three different commonly-used models for next-word prediction: a Bayesian model, a K-nearest-neighbors model, and a Latent Semantic Analysis (LSA) model.
-In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user created a post, and restrict the space of possible words to only tags.
-He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
-Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive LSA and K-nearest-neighbors models.
-This Bayesian model can accurately predict a user-chosen tag at 47\% accuracy.
-
-% ref: http://tex.stackexchange.com/questions/53513/hyperref-token-not-allowed
-\paragraph{\texorpdfstring{\textcite{Stanley2013}}{Stanley2013}'s ACT-R DM Model for StackOverflow}
-
-\textcite{Stanley2013} also modeled a user's chosen tags for a post on the StackOverflow site.
-They modified the standard ACT-R model in Table \ref{tabACTRModel} in order to more accurately predict tag use on StackOverflow.
-A formal description of the modified ACT-R Bayesian model is included in Table \ref{tabModACTRModel}.
-
-\begin{table}[!ht]
-  \caption{\citeauthor{Stanley2013}'s StackOverflow tag prediction model}
-  \label{tabModACTRModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} - O$ \\
-      Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
-      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
-      Co-occurrence &		$N_{ji} = \sum_{}^{}{observed_{ji}}$ \\
-      Observations &		$N = \sum_{}^{}{\sum_{}^{}{N_{ji}}}$ \\
-      Attentional Weight& 	$W_{j} = W \frac{E_{j}} {\sum_{}^{} {E_{j}}} $ \\
-      Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
-      Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
-      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-The activation of each tag ($A_{i}$) is a function of four components:
-That tag's base-level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
-This model has been modified from the standard ACT-R model (Table \ref{tabACTRModel}) in four ways:
-Two separate contextual components were used (words in title and body of post), an offset term was added, an entropy measure was used to compute the attentional weight ($W_{j}$),
-and the base-level activation of a tag was based on global past tag use and not customized to a user's past tagging history.
-
-The offset term was added to normalize the mean activation across retrievals.
-This allowed the use of a logistic regression statistical technique to calibrate the weights for each of the terms.
-The title and body contextual components were separated since it seemed plausible that the words in the title would be better predictors of the tag used than the words in the body.
-It turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
-
-The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation \parencite{Dumais1991} and is parameter free.
-It is based on Shannon's Information Theory, which measures the amount of information content (predictiveness) of a word.
-So the scaled entropy ($E_{j}$) of a context word can be thought of as the amount of predictiveness that that word has to \emph{any} tag.
-This means that a stop word like ``the'' will have a scaled entropy measure close to 0, while a highly-predictive word like \emph{macro} will have a scaled entropy measure close to 1.
-Looking at $W_{j}$, if $E_{j}$ is low then $W_{j}$ is low, meaning that the scaled entropy measure helps allocate limited attentional resources to predictive contextual elements.
-
-\paragraph{Comparison Between \citeauthor{Kuo2011}'s and \citeauthor{Stanley2013}'s Models}
-
-Since \textcite{Stanley2013} made several modifications to standard ACT-R model, it was more sophisticated than both the standard ACT-R model and the Bayesian model used by \textcite{Kuo2011}.
-Also, it was unclear if \citeauthor{Kuo2011} used both the words in the title and the body of the post as separate components for his models, or if only the words in the title were used.
-
-\citeauthor{Stanley2013}'s model also scaled much better than \citeauthor{Kuo2011}'s Bayesian co-occurrence model. 
-Because of this, 100 times more posts were used to build the co-occurrence matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
-After including the modifications and scaling up the model, the model can successfully predict a user's tag for a post at 65\% accuracy.
-
-It is interesting that in both studies the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
-This suggests that more cognitively-plausible memory retrieval models may outperform more general machine learning approaches in this particular environment
-(i.e., when predicting the user-chosen tags for human-created content on social media sites).
-
-\subsubsection{StackExchange}
-
-A tag suggestion tool has been deployed for the \url{tex.stackexchange.org} \parencite{LatexTags2013} and \url{superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
-Latex, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
-For the two sites where a tag suggestion tool has been deployed, once a user has created a question, suggested tags appear just below the text field where the user types the tags to associate with the question.
-The post author is not forced to use the suggested tags, and is free to choose tags from them or tags not represented in the suggested set.
-An example of the interface for a created post on the SuperUser site is included in Figure \ref{figSuperUserSuggestion}.
-
-\begin{figure}[!htbp]
-  \scalebox{.9}{\resizebox{\linewidth}{!}{\includegraphics{superUserTagSuggestion-crop.pdf}}}
-  \caption{Example of tag suggestions when creating a post on the SuperUser site}
-  \label{figSuperUserSuggestion}
-\end{figure}
-
-The exact tag-recommendation algorithm used for the Latex and SuperUser sites is not public.
-After exploratory testing, StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags.
-The Latex and SuperUser sites are not nearly as popular as StackOverflow currently.
-The sites have roughly \num{50000} and \num{189000} questions respectively compared to \num{5000000} questions on StackOverflow.
-
-So it stands to reason that the algorithm is relatively complex, and there could be possible scaling and efficiency concerns with the algorithm's implementation, so it has not yet been deployed to StackOverflow.
-Alternatively, it is also possible that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
-Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial to the user and has started to add them to their question and answer sites.
-
-\subsubsection{Google+}
-
-The hashtag recommendation system deployed for the Google+ social media site is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
-The user interface for this recommendation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.
-
-\begin{figure}[!htbp]
-  \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample-crop.pdf}}}
-  \caption{Example post on Google+}
-  \label{figGoogle+Hashtags}
-\end{figure}
-
-The workflow and user interface provided for recommending hashtags for Google+ is similar to how a hashtag recommendation system for Twitter might look.
-When generating content on Google+, the system automatically associates the most likely relevant hashtags with that post.
-For the example in Figure \ref{figGoogle+Hashtags}, the system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-created post.
-The user is free to remove any or all of the tags that the system has associated with the post, and can also use tags that were not included in the recommended set.
-
-Google+'s tag-recommendation algorithm is not public, but it appears to be rather sophisticated.
-Looking at the example post in Figure \ref{figGoogle+Hashtags}, \emph{\#BusterPosey} is suggested without having that specific word in the post.
-Also \textcite{GoogleKeynote2013} mentioned that the algorithm uses Google's sophisticated image recognition software to suggest hashtags for images that are included in the post (e.g., \emph{\#EiffelTower}),
-even when the post does not have any keywords to provide additional context.
-
-Google is apparently highly confident in the accuracy of their system.
-On the Latex site, the user interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
-For Google+, the user interface automatically tags a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
-If tag accuracy is high enough, the Google+ technique is certainly favorable since the user does not have to spend time figuring out which tags best represent this post.
-However, if the model makes even just a few errors, users may become annoyed and frustrated with the recommendation system since they will have to explicitly correct it every time they generate a post.
-
-\subsubsection{Twitter}
-
-Several hashtag recommendation models for Twitter have been developed recently.
-One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a post as a function of prior hashtag use and context \parencite{Mazzia2009}.
-The model presented by \textcite{Mazzia2009} is quite similar to an ACT-R declarative retrieval model.
-The main potential difference is that --- like \textcite{Kuo2011} and \textcite{Stanley2013} --- the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
-Instead, the global prior is an overall average frequency of hashtag use across all users.
-Further, no recency information was used to compute the global prior, so hashtags used in the recent past were not weighted more heavily. 
-This was most likely not much of an issue for \citeauthor{Mazzia2009}'s study, since all hashtags were collected within a relatively short period of time (about 3 months).
-
-Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used from similar tweets.
-\textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
-\textcites{Zangerle2011, Kywe2012} used the Apache Lucene infrastructure to compute the similarity score, while \textcite{Li2011} used a custom similarity score to compute similarities.
-Regardless of the method to compute the similarity between tweets,
-each method collects the hashtags used from a set of the most similar tweets, then ranks the hashtags in the set, and presents the top 5-10 hashtags to the user.
-
-This tweet-centered approach is quite different than the hashtag-centered approach presented in \textcite{Mazzia2009}.
-With a hashtag-centered approach, a direct association is built between the words in a tweet and hashtags that occur in tweets.
-With a tweet-centered approach, that association between words and hashtags is indirect:
-The relationship is built between a tweet and its content, and then similar tweets are assumed to use similar hashtags.
-One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a representation of the content in each tweet has to be maintained to later compute tweet similarity.
-The storage size for a hashtag-centered approach grows with the number of hashtags, but it seems reasonable that this will grow slower than the number of tweets and asymptote after collecting a large sample of tweets.
-So from the perspective of efficient information storage and retrieval, it seems more likely that the hashtag-centered approach used in \textcite{Mazzia2009} will scale as Twitter grows in size and use.
-
-\textcite{Kywe2012} also customized their tag prediction model to the user's past hashtag use, and was one of the first to do so when developing a hashtag-prediction model for Twitter.
-The set of recommended hashtags was the union of hashtags used in similar tweets and hashtags used by similar users.
-This was done by storing a content vector of hashtag use for each user (user-centered approach) alongside the content vectors of hashtag use for each tweet (tweet-centered approach).
-To generate recommended hashtags, the model would return the top-ranked hashtags from the hashtags used in 0-50 similar tweets and 0-4 similar users.
-Prediction accuracy improved when the model combined recommendations based on hashtags in both similar tweets and the top few similar users compared to basing recommendations on similar tweets alone.
-
-A topics-based approach was used by \textcite{Godin2013}, where the model predicted the most likely set of topics that were associated with a tweet, and tested if these topics matched the tweet by using human raters.
-The latent topics for the collection of tweets were generated using a Latent Dirichlet Allocation (LDA) statistical methodology.
-This is similar to a factor analysis approach in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance.
-Human graders rated how well the models generated topics for a tweet related to the content in the tweet.
-The LDA model performed well, and generated relevant topics.
-When the model produced five most-likely hashtags for each tweet, 80\% of the time at least one of those hashtags was suitable (as determined by human raters).
-
-However, a topic-generating model is not performing the same task as a hashtag-prediction model.
-The authors argued that these general topics could be used to categorize the post, in addition to the hashtags already used by the author in the post.
-It is unclear though if a user would want to label their tweet with these general topics.
-Perhaps they would rather use more specific hashtags to be more precise in how they label the tweet and how that tweet is associated with tweets within the rest of the community.
-In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
-Now it is certainly the case that the topics generated by an LDA model may be similar to the hashtags used in the post.
-However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
-On the other hand, a topics-based approach is more general than a hashtags-based approach, since topics can be suggested for posts that do not use hashtags (which is quite common across Twitter).
-
-\subsubsection{State of Research on StackOverflow and Twitter}
-
-\textcite{Stanley2013} used an ACT-R based Bayesian memory model to predict human tagging on the StackOverflow site, but to our knowledge no vector-based memory model has been tested on the site.
-Both types of models have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007},
-so it should be straightforward to test a vector-based model on the StackOverflow site.
-
-Also, a vector-based retrieval model has not yet been tested on Twitter.
-One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
-It seems quite likely that words used just before a hashtag in a tweet (combined with their position) will be highly predictive of the hashtag chosen.
-To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
-
-There has been a mix of co-occurrence models created to predict hashtag use on Twitter, and the results are encouraging.
-However, there has been little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
-\textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
-The ACT-R declarative memory retrieval theory shows how past user behavior and current context combine to produce the most likely retrieval.
-Each component is summed together to generate a total activation, and then the likelihood of chunks are ranked by that summed activation.
-This technique is compensatory and allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
-This type of weighted additive model that combines past user behavior and current tweet context to predict the most likely hashtags has not yet been explored for Twitter.
-
-Both StackOverflow models \parencites{Stanley2013,Kuo2011} did not take into account the specific user's prior history when generating tag predictions.
-Rather their models, much like \textcite{Mazzia2009}'s Twitter model, uses the aggregate tag history across all users as the model's prior activation for each tag.
-
-As some initial research, I have developed and tested a user-customized model for StackOverflow by modifying the model in \textcite{Stanley2013}.
-To incorporate the user's prior tag history into the model, the base-level activation term ($B_{i}$) was modified from the original equation in ACT-R.
-$B_{i}$ is based on log-odds occurrence, which in turn is based on frequency of tag use, or probability of tag use $p_{i}$.
-To incorporate a user's prior, a weighted average for $p_{i}$ was used,
-where the probability of using tag $i$ ranges from the global probability ($pGlobal_{i}$) to the user's specific probability ($pUser_{i}$) as the user's total tag count increases.
-
-The global probability is the same $p_{i}$ used in \textcite{Stanley2013}'s study.
-The user's specific probability is computed the same way that the global probability is computed (i.e., the log odds that the tag occurred relative to the occurrence of all other tags).
-However, each user's probability is constrained to the set of tags that the specific user has used in the past. 
-When a user has no past tag history, the global prior probability is used.
-Once a user has provided at least five tags, that user's specific prior probability is used.
-Between zero and five tags, both terms are used but the weight of the global prior relative to the user prior decreases as the number of tags increases.
-
-Both linear decay and exponential decay transition functions were tried.
-Using either transition function, overall model accuracy improved by 5\% after incorporating the user's prior tag history into the model.
-
-This research suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
-It also seems plausible that model accuracy can improve an even greater amount by using the time-dependent form of the base-level component in Table \ref{tabACTRModel},
-in order to account for the fact that a user's tagging preferences change over time.
-
-One of the main purposes of this research is to improve on these user-agnostic statistical models and include each specific user's past tagging history as a model component.
-Using customized priors that are tailored to a user's specific tagging history should be crucial for sites like Twitter where the possible tag space is infinite,
-but should also improve performance for more constrained sites like StackOverflow since each user is not interested in all of the possible tags on the site.
-
-\clearpage
-\section{Chapter 3: General Methodology}
-
-Each specific result of the research (to be presented in Chapters 4-6) requires a sightly different method to study.
-However, there is a common general methodology that applies across all results that will be described first.
-Different subsets of the StackOverflow and Twitter datasets were used evaluate past user behavior in isolation and to test the context component of each memory model.
-These datasets will be named and described in general terms.
-Also, the general methods used to tokenize and stem the words in all of the datasets will be described.
-
-\subsection{Models}
-
-Two models were evaluated for hashtag prediction on both the Twitter and StackOverflow datasets: a vector-based model and Bayesian retrieval model.
-The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} was used for the vector-based model,
-since it has shown to perform better than the original BEAGLE vector-based model \parencite{Recchia2010}.
-The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel} was used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
-
-The two memory models were evaluated on predicting user-chosen tags for Twitter tweets and StackOverflow posts.
-The co-occurrence matrices used for the context component of each model were generated from a large subset of the Twitter and StackOverflow datasets.
-Each model was tested by evaluating the accuracy of model-chosen tags on a fresh test subset of the datasets.
-Accuracy was assessed by comparing model performance with the human data.
-
-\subsection{Datasets}
-
-The most recent quarterly published StackOverflow dataset \parencite{DataDump2014} was used to test the models on user chosen tags for posts.
-This dataset contains around \num{5.7} million posts and \num{34377} unique tags.
-
-Unlike StackOverflow, the entire Twitter dataset is not released to the public.
-Due to this and the scale of Twitter, only a part of the Twitter dataset relevant to this research was collected.
-
-\subsubsection{Popular Users Dataset}
-
-A popular-users dataset was collected for both StackOverflow and Twitter.
-This dataset consists of all of the tweets and posts from various samples of the top users for each site.
-This dataset was collected by fetching all past tagging history for each of the top users.
-It was used to explore the growth and decay characteristics of a hashtag for individual users.
-The temporal dynamics of ACT-R's decay rate model for a chunk were evaluated against these data.
-Multiple samples were taken across a broad range of user types for each dataset to ensure that the model results were not dependent on specific attributes in a single sample.
-
-For StackOverflow, top users were identified by their reputation score on the site (based on their upvoted questions and answers) and the total number of questions asked on the site.
-To create the StackOverflow popular-users dataset, a total of 12 slices of users were taken at different levels of reputation and total number of questions asked.
-500 users were sampled that had a reputation higher than \num{500}, \num{1000}, \num{5000}, \num{10000}, \num{50000}, and \num{100000}.
-100 users were sampled that had asked a total number of questions higher than \num{50}, \num{100}, \num{200}, \num{300}, \num{400}, and \num{500}.
-
-For Twitter, top users were those that had the largest number of followers and had authored the largest number of tweets.
-To create the Twitter popular-users dataset, 12 slices were also taken, this time at different levels of number of followers and total number of tweets.
-100 users were sampled that had a total number of followers higher than \num{1000}, \num{5000}, \num{10000}, \num{100000}, \num{1000000}, and \num{10000000},
-and a total number of tweets higher than \num{100}, \num{500}, \num{1000}, \num{5000}, \num{10000}, and \num{50000}.
-
-\subsubsection{Popular Hashtags Dataset}
-
-An additional popular-hashtags dataset was collected for Twitter.
-The purpose of this dataset was to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
-The dataset contains approximately three million tweets where at least one of the 400 most popular hashtags are used.
-A quota of three million tweets was selected because previously research testing Bayesian co-occurrence models on large-scale datasets has used on the order of one million documents
-\parencites{Stanley2013,Douglass2010,Budiu2007,Kuo2011}.
-Tweets were collected in real time until the tweet quota was reached.
-
-Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
-Multiple datasets were collected to ensure that the results were not dependent on a particular sample of hashtags and would generalize across different popular-hashtag datasets and different time periods.
-Each dataset required around four days to collect enough tweets to reach the quota, and one collection was done each week.
-
-\subsubsection{Randomly Sampled Dataset}
-
-Since the entire dataset is available for StackOverflow, randomly-sampled posts were used to test the context models rather than creating a popular-tags dataset (analogous to Twitter).
-If all Twitter data were available, then the models would have been tested on a randomly-sampled dataset for Twitter as well.
-However, since this is not the case for Twitter, the popular-hashtags dataset was used instead of a randomly-sampled dataset for that site,
-while models for StackOverflow were tested on the more challenging dataset where posts were randomly sampled across all posts created on the site.
-
-\subsubsection{Dataset Descriptives}
-
-The StackOverflow datasets had a much larger document size than the Twitter datasets overall.
-There were on average 9 and 158 words in the title and body of a post for StackOverflow when sampled across the entire site. 
-For Twitter, there were 13 words per tweet for the popular-users dataset and 18 for the popular-hashtags dataset.
-
-The average number of tags per dataset was lower overall for Twitter.
-For Twitter, authors used on average 1.1 hashtags per post for the popular-users dataset and 2.6 for the popular-hashtags dataset.
-For StackOverflow, 2.8 tags were used on average across all posts on the site.
-
-Since the document size and tags per document were higher for StackOverflow, the co-occurrence matrix for StackOverflow contained many more observations compared to Twitter.
-Taking 1 million posts across the site, the StackOverflow co-occurrence matrix contained \num{835000000} total observations and \num{55000000} unique combinations of context words and tags.
-For Twitter, when building a co-occurrence matrix for the popular-hashtags dataset from 3 million posts, the matrix contained \num{39000000} total observations and \num{4000000} unique combinations.
-
-Hashtags for Twitter were both placed at the end of the tweet and embedded within the middle of the tweet.
-For the popular-hashtags dataset, 11\% of hashtags were the last word in the tweet and the rest were embedded within the tweet (not at the end).
-Tags for StackOverflow were always placed at the end of the post, since the author chose tags by placing them in a separate text field before submitting the post.
-
-Each of the four Twitter popular-hashtags datasets contained the 400 most-popular hashtags at that time.
-There was only a small amount of overlap between the hashtags chosen for each of the datasets.
-Only 20\% of hashtags occurred two or more times across the four popular-hashtags datasets and the majority of those (66\%) occurred in only two of the four datasets.
-
-The StackOverflow dataset contains many orders of magnitude fewer unique users than Twitter.
-This means that the odds that the same user is sampled multiple times is less for the Twitter popular-hashtags dataset compared to all posts across the StackOverflow site.
-For Twitter, an author contributed an average of 2 tweets to the popular-hashtags dataset.
-Across the StackOverflow site, an author contributed 5 posts on average.
-
-\paragraph{Tag Frequency Distribution}
-
-\textcite{Stanley2013} found that the tag frequency for StackOverflow drops off sharply and approximately follows Zipf's Law.
-The same effect was found for the 2014 StackOverflow dataset in this research and the two Twitter dataset subsets.
-The results are depicted in Figure \ref{figDescTagFreq}.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{descr-tagFrequencyCount.pdf}}}
-  \caption{
-    Log-log plot of tag frequency versus rank for the StackOverflow dataset and Twitter popular-users and popular-hashtags dataset subsets.
-    Rank was assigned to each unique tag after ordering by tag frequency (most frequent given smallest rank).
-}
-  \label{figDescTagFreq}
-\end{figure}
-
-Tag frequency for StackOverflow and Twitter follow a similar Zipf's Law pattern, where tag use drops off sharply after moving away from the most popular tags.
-However, none of the three distributions follow Zipf's Law perfectly, since none of the plots show perfectly linear lines.
-Nonetheless, using three separate linear models to fit the data for the three distributions captures 96\% of the variance.
-So these log-log plots are mostly linear, and both StackOverflow and Twitter closely follow Zipf's Law for tag use frequency.
-
-\subsection{Acquiring Datasets}
-
-\subsubsection{StackOverflow}
-
-The entire StackOverflow dataset is packaged and released quarterly to the public.
-So the most recent packaged dataset was downloaded and the database tables were imported into PostgreSQL.
-
-\subsubsection{Twitter}
-
-Twitter provides two types of APIs for acquiring data: one based on streaming and a direct query-based interface.
-The Tweepy Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
-The query-based interface was used to collect all tweets for specific popular users.
-At most \num{3000} of the most recent tweets were collected for each user, since Twitter only allows access to this number of tweets for each user.
-
-The streaming interface was used to collect a sample of all of the tweets that included one of the top 400 hashtags at that time.
-Twitter's streaming API only allows at most 400 hashtags to be monitored at a time.
-Otherwise a larger number of hashtags would have been collected.
-Twitter allows the streaming interface to randomly sample about 1\% of all of the real-time data that is generated on Twitter.
-This sampling rate was adequate to collect around three million tweets within a few days.
-
-\subsection{Tokenization}
-
-A custom tokenization algorithm created by \textcite{Owoputi2013} was used to chunk the words in the text.
-This algorithm was specifically created for tokenizing Twitter content.
-However, during testing it was also found to be robust and well suited for StackOverflow content.
-The initial model results with the new tokenizer were qualitatively similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
-This is most likely because the language for the samples taken for both sites is informal English,
-and it seams reasonable to think that tokenizing all of the informal language and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
-So the same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter, since performance was similar to \textcite{Stanley2013} for StackOverflow,
-and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.
-
-\subsection{Stemming}
-
-\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing ``ed'' and ``ing'' endings), but
-with the StackOverflow dataset, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
-This can be thought of as semantic stemming, where two words with identical meaning are converted to the same canonical word.
-So tags used for the post were converted to their root tag when possible.
-
-\subsection{Handling Stop Words}
-
-``Stop words'' (e.g., ``the,'' ``a,'' ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
-It is typically recommended \parencite{Bird2009} to remove stop words when building the co-occurrence matrix.
-Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in the text.
-Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.
-
-Stop words are usually removed by either filtering based on a predetermined list or filtering based on the high-frequency words that occur in the dataset.
-\textcite{Sahlgren2008} tried both methods to remove stop words for the random permutation vector-based model, and found that the data-driven approach
-(i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a predetermined list.
-
-\textcite{Jones2007} also filtered the stop words for the circular convolution model,
-but also mentioned that an entropy weighting technique \parencite{Dumais1991} has been used previously to attenuate (rather than remove) stop words.
-This data-driven technique identifies stop words by calculating the discriminating power (i.e., entropy) of each word to predict any tag.
-Each word is then weighted (not removed) based on their discriminating power.
-
-\textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset.
-The result maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
-It also seems less arbitrary than explicitly identifying a list of stop words to remove from the analysis.
-
-All three stop word techniques (i.e., predetermined list, high-frequency filtering, and entropy weighting) were explored for this analysis.
-
-\clearpage
-\section{Chapter 4: Past User Behavior}
-
-How does a user's prior tag use influence their likelihood of choosing that tag in the future?
-This question was tested by examining how well ACT-R's base-level activation component of declarative memory predicts the tags that authors choose, given each author's past tagging behavior.
-The model predicts that the more recently and frequently used tags have higher prior activations, and are more likely to be needed in the future.
-
-\subsection{Base-level ACT-R model}
-
-The base-level activation component from the full ACT-R declarative memory model in Table \ref{tabACTRModel} is included in Table \ref{tabACTRBLLModel}.
-
-\begin{table}[!ht]
-  \caption{Base-level component of ACT-R declarative memory}
-  \label{tabACTRBLLModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Base-Level Activation &	$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
-      Optimized Learning &	$B_{i} = log \frac{n}{1-d} - d * log L$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-Both terms are methods for computing the base-level activation ($B_{i}$) of a chunk in declarative memory.
-The standard form has a time dependence and accounts for the fact that the rate of presentations for a given object (e.g., the word ``loop'') can change over time.
-The optimized learning form assumes that this rate is constant.
-
-For both equations, $i$ represents each chunk (e.g., the word ``loop''), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``loop'' appears), and $n$ is the number of times each chunk has been observed.
-Since the optimized learning form assumes a constant presentation rate for each chunk,
-it only has to keep track of the time since the first presentation of a chunk ($L$) and the total number of instances observed for that chunk ($n$). 
-It does not have to keep track of the time each specific chunk instance was observed.
-Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R Common Lisp implementation.
-
-The decay rate parameter ($d$) represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
-Higher values mean that chunk activation decays faster, which means that the most recently presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
-Lower decay rate values flip the importance of frequency and recency:
-Lower values mean that chunks that have occurred most frequently will have the highest activation,
-and activation depends less on the specific chunks that have been observed in the recent past.
-
-The base-level ($B_{i}$) component in Table \ref{tabACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{tabModACTRModel}).
-\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log-odds computation of the probability that each chunk is likely to be presented again, given past frequency.
-If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log \left ( n_{i} \right )$,
-which is a nonlinear transformation of the $B_{i}$ used by \citeauthor{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
-Both forms maintain the same rank ordering, so they should produce similar results.
-So the $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).
-
-\subsection{Method}
-
-The StackOverflow and Twitter popular-users datasets were used to characterize tag growth and decay for a user on each site.
-The popular-users dataset contains 12 slices of popular users for both StackOverflow and Twitter.
-Each user's tag behavior from the popular-users dataset for each site was analyzed.
-
-22 different values of the decay rate parameter were explored for each slice
-(0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.8, 2, 5, 10, 20).
-The values were chosen to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d \approx .5$), to a pure recency model ($d=20$).
-The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
-Only values equal to or below 1 were examined for the optimized learning form, since values above 1 for this form cause a bifurcation and are not psychologically interesting.
-The bifurcation occurs because at a decay rate of 1 the denominator of the first term to compute $B_{i}$ goes to zero (see Table \ref{tabACTRBLLModel}).
-
-Both the standard form and optimized learning form of $B_{i}$ were used to predict each user's specific tag history.
-To do this, a user's entire tag history was collected, and then each tag instance from oldest to newest was examined.
-For each of these tagging instances, the model generated an activation value for each possible tag based on that tag's frequency and recency of use. 
-This means, for example, that the model always missed characterizing the first observed tag for each user, since at that point the model does not have any history of tag use for that user.
-This also means that the tag space for each user is constrained and limited to the tags that each user has used in the past.
-
-At each tagging instance, the model has activation values for a set of tags.
-Model accuracy was assessed by rank ordering these activations and choosing the model's highest $N$ tags, where $N$ is the number of tags used for that specific post.
-The model's chosen $N$ tags were compared to the user's chosen $N$ tags, and if a model's chosen tag was in the set of user's chosen tags, then the prediction was marked as correct.
-Otherwise, it was marked as incorrect.
-
-This process was performed across all users in each of the 12 dataset slices for both the Twitter and StackOverflow popular-users datasets.
-It was repeated for each of the decay rate values and the two different forms of $B_{i}$.
-
-\subsection{Results}
-
-\subsubsection{Visualizing Tag Use}
-
-To better illustrate the dynamics of each user's tagging behavior,
-profile plots for a single user for both StackOverflow and Twitter are included in Figure \ref{figPriorHTByTime}.
-The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
-What is interesting is the large tag space for these users.
-There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[][StackOverflow]{\includegraphics[width=8cm]{HTByTime-520957-crop.pdf}}
-    \hfill
-    \subfigure[][Twitter]{\includegraphics[width=8cm]{HTByTime-fashionista_com-crop.pdf}}
-    \caption{
-      Tagging profile for a single user.
-      The x-axis represents the specific hashtag a user chooses for a post and time progresses vertically.
-      A point in the plot represents when a user chooses a specific hashtag for a specific post.
-      A user's chosen hashtag for their oldest post that contains a hashtag is the lowest point on the plot.
-    }
-    \label{figPriorHTByTime}
-  }%
-\end{figure}
-
-For example, the StackOverflow user in Figure \ref{figPriorHTByTime} has a pronounced band where the same hashtag is chosen repeatedly over time. 
-However there are many instances of other tags being used in parallel to the tag that is producing the band (that is repeatedly being used). 
-The Twitter user has a few clusters where the same hashtag is chosen repeatedly but only for a brief period of time.
-As time progresses, a band of choosing another hashtag is observed, and then the user switches to a new hashtag after that.
-More recent posts for the Twitter user are more varied, and there is no longer a single hashtag that the user chooses repeatedly.
-This helps illustrate how challenging modeling this task can be.
-The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is changing over time.
-
-Although the tag space for the Twitter user is smaller than the StackOverflow user in these plots, this is not a consistent trend.
-These plots were chosen from the set to also show that the amount of tag variation for individual users also varies: some use many tags and some use just a few.
-To accurately model human performance in this domain, the model must be able to make correct predictions across all variations of user tagging behavior.
-
-\subsubsection{Visualizing Model Performance}
-
-As a first method for understanding model performance, profile plots of model's predicted tags for individual users were overlaid on that user's actual tag use.
-These plots for StackOverflow and Twitter are included in Figures \ref{figPriorHTMByTimeSO} and \ref{figPriorHTMByTimeT} respectively.
-These are the same plots shown in Figure \ref{figPriorHTByTime} with model predictions overlaid in red.
-
-\begin{figure}[!htbp]
-  \scalebox{.7}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
-  \caption{Model tagging profile of a single user on StackOverflow}
-  \label{figPriorHTMByTimeSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.7}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
-  \caption{Model tagging profile of a single user on Twitter}
-  \label{figPriorHTMByTimeT}
-\end{figure}
-
-Also shown in these figures is how model predictions change as the decay rate parameter changes.
-With lower decay rate values, the model is less likely to switch from tag to tag for each post.
-This makes sense because lower decay rate values lead to the model weighting frequency more than recency in the activation calculation.
-These plots also suggest that model accuracy is best when a blend of frequency and recency is used (i.e., a decay rate between 0 and 20).
-This will be more formally tested when model accuracy across decay rate values is analyzed across all users in each dataset slice and all dataset slices.
-
-\subsubsection{Model Performance Across Decay Rate Values}
-
-Moving up one level of analysis, model accuracy for all users in a single dataset slice was analyzed.
-A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDsStd} and \ref{figPriorTwitterSliceDsStd} respectively.
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
-    \hfill
-    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
-    \caption{
-    Model performance for a single dataset slice for StackOverflow.
-    This includes model accuracy as a function of decay rate value for each user in the dataset slice.
-    Model accuracy is shown both when the standard from of base-level activation ($B_{i}$) and the optimized learning form of the equation are used. 
-  }
-    \label{figPriorSOQSliceDsStd}
-  }%
-\end{figure}
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}
-    \hfill
-    \subfigure[][Optmized learning form]{\includegraphics[width=8cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
-    \caption{
-      Model performance for a single dataset slice for Twitter.
-      This includes model accuracy as a function of decay rate value for each user in the dataset slice.
-    }
-    \label{figPriorTwitterSliceDsStd}
-  }%
-\end{figure}
-
-Model accuracies at each decay rate value for each user in the dataset slice are included in the plots.
-Each user's mean accuracy was subtracted off in order to compute a normalized mean for that user.
-This was done so that the relative accuracy change for different decay rate values could be more easily visualized.
-The plotting technique is a analogous to what is done when testing the main effect in a repeated measures design (i.e., each subject's mean score is subtracted off).
-
-One large effect in all of the plots is that there is an optimal decay rate value between 0 and 20 that produces the highest model accuracy.
-For the standard $B_{i}$ equation, model accuracy for a pure frequency model ($d=0$) is also worse than using a pure recency model ($d=20$), particularly for Twitter.
-It makes sense that a pure recency model for Twitter is relatively more accurate than for StackOverflow, as the hashtag lifetime for Twitter is much less than a tag for a programming language on StackOverflow.
-
-When comparing the standard and optimized learning forms of the equation, it is apparent that a pure recency model does not work well for optimized learning.
-The best-fit decay rate value for the optimized learning form is also not as clear and pronounced as it is for the standard form (less of a peak).
-And the relative accuracy gained from using a pure frequency based model to a blend of frequency and recency is less for the optimized learning form than the standard form of $B_{i}$.
-
-\subsubsection{Best-fit Decay Rate Values}
-
-To identify the best decay rate value for the dataset slices,
-each user's best decay rate value was taken and the results are plotted in histogram form in Figures \ref{figPriorSOQHistDsOL} and \ref{figPriorTwitterHistDsOL} respectively.
-The best-fit values are close to the \num{0.5} default value set in ACT-R when the optimized learning form is used.
-Comparing across optimized learning and the standard form, the best-fit values for the standard form are slightly higher for both the StackOverflow and Twitter dataset slices.
-Comparing best-fit values will be more thoroughly tested when looking at the values across all dataset slices.
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[][Standard form]{\includegraphics[width=8cm]{visHistD-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
-    \hfill
-    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visHistD-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
-    \caption{
-      Best-fit decay rate for a single dataset slice for StackOverflow.
-      The histogram represents the count of users for each best-fit decay rate value.
-      Each bar represents a discrete decay rate value that was explored.
-  }
-    \label{figPriorSOQHistDsOL}
-  }%
-\end{figure}
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[][Standard form]{\includegraphics[width=8cm]{visHistD-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}
-    \hfill
-    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visHistD-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
-    \caption{
-      Best-fit decay rate for a single dataset slice for Twitter.
-      Best-fit values were slightly higher for Twitter, so the range of the histogram is higher for Twitter than StackOverflow.
-  }
-    \label{figPriorTwitterHistDsOL}
-  }%
-\end{figure}
-
-\subsubsection{Aggregate Model Performance}
-
-Aggregate best-fit decay rates were computed by taking the median of all user's best-fit decay rate values across all dataset slices.
-The median was used since there were users in the popular-users dataset that did not have enough observations to generate stable predictions, and best-fit decay rate values for these users could be as high as 20.
-Using the median effectively trims these unstable values from the sample, without having to define a cut point for an outlier removal process.
-
-Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
-Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{.43}) of $B_{i}$ compared to the standard form (\num{.80}).
-The optimal values for optimized learning are near \num{0.5}, which lines up with default value used in ACT-R.
-So the decay rate values for the standard form are slightly higher than the default values used in ACT-R, but the values found when optimized learning is used line up nicely with the ACT-R defaults.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
-  \caption{
-  Overall best-fit decay rate for StackOverflow and Twitter.
-  Results are shown for the two Twitter and StackOverflow (SO) popular-users datasets.
-  Error bars represent the 95\% bootstrapped confidence interval for the median.
-}
-  \label{figPriorDecay}
-\end{figure}
-
-\FloatBarrier
-
-Finally, model accuracy was analyzed across all dataset slices for StackOverflow and Twitter.
-This was done by examining model accuracy at a specific decay rate value for each model.
-Accuracy was assessed at each model's previously-determined best-fit values (\num{.43} and \num{.80} for the optimized learning and standard form) and the standard form's default value (\num{0.5}).
-Since decay rate values were explored at \num{0.1} increments between the 0 and 1 range, the decay rate values used to examine model accuracy were rounded to the nearest decay rate value that was included in the run.
-So a best-fit value of \num{.4} was used for the optimized learning model instead of the \num{.43} average found from the dataset slices.
-Accuracy for both sites is included in Figure \ref{figPriorAcc}. 
-Aggregate accuracy was computed by taking the mean of every user's model accuracy across all users in all dataset slices.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc--crop.pdf}}}
-  \caption{
-    Overall model accuracy for StackOverflow and Twitter at optimal decay rate values.
-    All error bars are the 95\% bootstrapped confidence interval for the mean. 
-  }
-  \label{figPriorAcc}
-\end{figure}
-
-Accuracy for the standard form (\numNoZero{.33}) is higher than the optimized learning form (\numNoZero{.28}).
-This makes sense given that the standard form does not assume equal presentation rate for each chunk, and hashtags on Twitter in particular show trends of high rate of use and then shift to low rate of use.
-The standard form of $B_{i}$ is able to more accurately account for these trends.
-
-ACT-R sets the default decay rate value to \num{0.5}, and the best-fit value for the standard form is higher (\num{0.8}) for these datasets.
-The figure shows that although model accuracy does slightly improve for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
-that percentage improvement is not as large as the improvement gained from using the standard form compared to the optimized learning form.
-Nonetheless, model accuracy is highest when both the standard form is used and the decay rate value is increased.
-
-\subsection{Discussion}
-
-Given the large tag space used by post authors on StackOverflow and Twitter, it was surprising that model performance is respectable even when only past tag history and no context is used.
-This speaks to the importance of taking into account overall past tagging history, and ensuring that the prior component of activation for each tag is customized to each user's specific past tagging history.
-
-Also apparent for these datasets is that the optimized learning form of $B_{i}$ is not as accurate as the standard form.
-Further, for the optimized learning form there is not much difference between using a pure frequency-based model and a model where frequency and recency are optimally blended.
-This makes sense given that the optimized learning equation in Table \ref{tabACTRBLLModel} collapses to a pure frequency model as the amount of time since the presentation of the first recorded chunk instance increases,
-and also since the timespan of recorded tag use for StackOverflow and Twitter is large (years).
-
-So it may be worthwhile to use the standard form instead of the optimized learning form when declarative memory is modeled across a broader range of tasks.
-However, this should not be taken as a strong recommendation,
-since the performance benefits seen in this study could simply be due to the fact that the declarative memory retrievals span across a large period of time, which is not the case for all tasks.
-Nonetheless, the only real reason not to use the standard form is due to computational resource limitations.
-Since this is becoming less of an issue as hardware speed increases, it may be time to reconsider the default form of base-level learning used in ACT-R.
-It may be the case that the optimized learning form of the equation is no longer computationally necessary.
-And if the standard form is used as the default (or simply tried in another experiment),
-the results from this study suggest that the optimal decay rate value should be set slightly higher (\num{0.8}) than the default when optimized learning is used (\num{0.5}).
-
-The decay rate parameter in $B_{i}$ determines how fast a presentation of a chunk in the past decays over time.
-Higher values of $d$ mean that activation decays more quickly, which means that the chunks used more recently are the most active and the total frequency count of a chunk matters less.
-Looking back at Figure \ref{figPriorDecay}, the optimal decay rate parameter when the standard form of $B_{i}$ is used is higher for Twitter than StackOverflow.
-It makes sense that the lifetime of a specific hashtag on Twitter is shorter overall than the lifetime of a tag (e.g., \emph{C\#}) on StackOverflow,
-since hashtag topics are generally more transient on Twitter than StackOverflow (e.g., hashtags for specific political events or upcoming conferences).
-This means that the decay rate for Twitter should be higher than StackOverflow, which is what is observed.
-
-\clearpage
-\section{Chapter 5: Combining Predictors}
-
-The declarative memory models analyzed for this research consist of two main components: past behavior and context.
-Past tag use was isolated in the first study to see how well ACT-R's prior component of declarative memory (i.e., base-level learning) fits each user's tag use over time.
-For the rest of the studies the past behavior and context declarative memory components were combined as well as looked at in isolation.
-
-Chapter 5 will examine how model accuracy is influenced by two methodological concerns (adding an offset term and examining the effect of dataset subset for Twitter)
-and five architectural concerns
-(method of handling stop words, influence of dataset size, amount of compression used for the vector-based models, method to combine the prior and context for the vector-based models, and word order).
-The two methodological concerns were examined to ensure that the technique to calibrate model parameters was accurate and also that results for Twitter were not specific to the subset of data collected.
-
-The five architectural concerns were examined for the following reasons:
-During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.
-
-Due to research from \textcite{Budiu2007} and \textcite{Recchia2010}, model accuracy for the Bayesian and vector-based models is influenced by the number of documents in the dataset,
-so the effect of dataset size on accuracy was examined in more detail.
-Specifically,
-since both the Bayesian and random permutation models are being evaluated, analyzing the effect of dataset size on accuracy in this research will allow direct comparisons between the two models on a common dataset.
-This has not yet been possible since the two models have never been compared on the same dataset within the same study.
-
-\textcite{Sahlgren2008} showed how different levels of compression in the random permutation model influence accuracy, so this was examined for three reasons:
-to replicate the findings,
-as an additional check to ensure that the random permutation model was implemented correctly, and also to see if the same levels of compression are valid for the much larger datasets used in this research.
-
-One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random permutation model,
-so two different ways to add the model terms were tested.
-
-Finally, previous research has shown that including word order in the vector-based models improves performance \parencites{Jones2007, Sahlgren2008},
-so word order was also examined to see if the effect could be replicated and also to measure the overall importance of word order on model accuracy relative to other architectural manipulations.
-
-\subsection{Overall Method}
-
-\subsubsection{Logistic Regression}
-
-In order to best combine these predictors, a logistic regression statistical technique similar to \textcite{Stanley2013} was used to determine the most optimal weights for each term.
-\textcite{Stanley2013} weighted each term in the total activation equation (Table \ref{tabModACTRModel}) to maximize model accuracy.
-\citeauthor{Stanley2013} found that the optimal weights for the StackOverflow model for the prior term, title context term, and body context term were \num{.62}, \num{.93}, and \num{1.75} respectively.
-
-The optimal weights for each model component were computed in the same manner as \textcite{Stanley2013}:
-When a user creates a tag for a post, a retrieval request for the model is made.
-For each request, the model returns a set of tags (all tags that have been observed for that user in the past) and activations associated with each tag.
-An activation value is computed for each model component (e.g., prior tag use, context) for each tag.
-These values are recorded, and tag instances that match what the author actually tagged are marked as a 1, while all others are marked with a 0.
-This process is repeated across all posts where a user chooses a tag, and across all users in the dataset.
-Once all of the results are aggregated, a logistic regression is run where the optimal weights for each model component are found that maximize the model's ability to correctly label tags as 0 or 1 based on total activation.
-
-Model accuracy was evaluated by rank ordering all recorded tags by model activation using the optimal weights,
-asking the model to tag the top N posts with a 1 (where N is the total number of recorded tagging instances in the dataset sample),
-and then comparing the model's chosen tags with the used tags for each post (i.e., looking at the ``hits'').
-This is equivalent to throttling the threshold in the logistic regression such that the total number of observations labeled as a 1 match the total number of recorded instances of tagging in the dataset.
-
-For this form of model accuracy, the model labels N posts with a 1 where the labeling is not constrained in any way within each post.
-That is, the labeling is relaxed across all posts, and there are no constraints where the model must label a specific number of 1s for each post based on the number of tags that the author used for that post.
-So model accuracy values found with the relaxed method are actually slightly conservative, since the model may be labeling more 1s for a post than the number of tags the author used for that post.
-
-However, this relaxed metric was used for the rest of the experiments that involve combining terms because the optimal weights found by the regression are based on using the relaxed metric.
-If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression was not optimized to the constrained form, and weights may change as a consequence.
-Also, there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation where a small set of 1s occurred. 
-To optimize parameters based on the constrained form,
-a full-enumeration numerical approach would have to be developed that systematically searches the space of coefficient weights and uses a custom objective fitting function to find the best fit.
-However, this was not pursued for three reasons:
-Around \num{2500} total regressions were run on these datasets,
-a custom implementation in R would not likely run in compiled C,
-and even if it was implemented in a compiled language it would be computationally intractable without a closed-form solution.
-
-\subsubsection{Adding Offset Term}
-
-This logistic regression method is a computationally efficient way to find the optimal weights for the predictors.
-However, when the regression can not be constrained to label a specific number of 1s for each observation (as is the case here), to be used properly an offset term must be added as an additional model predictor.
-
-The offset term keeps the activation for the small number of top tags for each post equal across posts.
-This ensures that the logistic regression only labels a few 1s for each post.
-\textcite{Stanley2013} did this by subtracting off from each post the mean total activation for the top 5 tags for that post.
-However, this method is complicated since computing the total activation for a tag requires already knowing the weights for each model component that are used to compute that activation.
-\citeauthor{Stanley2013} used an iterative approach to determine the weights used for the offset term, and kept iterating until those weights matched the weights returned after running the logistic regression.
-
-A simplified version of \textcite{Stanley2013}'s method was used for the results for this research.
-Instead of subtracting off the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
-So for the prior component, the activation value used in the logistic regression is the prior activation with the mean of the top 5 prior activations subtracted from this value.
-The same process is used for the context component.
-Decoupling the offset term and computing an offset for each model component in isolation means that an iterative approach is no longer necessary,
-since the optimal weights of the terms are no longer needed when computing the offset.
-
-\subsubsection{Datasets}
-
-Model accuracy was tested on the popular-hashtags dataset for Twitter and a random sample of posts across the entire StackOverflow dataset.
-For each of the four Twitter popular-hashtags datasets, model retrievals were done across 15 runs and 500 posts within each run.
-For the StackOverflow randomly-sampled dataset, 5 runs of 100 posts were used,
-since the results for StackOverflow are more stable and retrievals take more wall-clock time for this dataset due to the size of the co-occurrence matrix.
-All runs contained different posts, and all posts used for runs were not used to generate the co-occurrence matrix for the two retrieval models.
-
-\subsubsection{Computing Priors}
-
-For each randomly-sampled post for StackOverflow, the prior activation is based on that user's prior tagging history, and not on any global prior of specific tag frequency across users.
-This was done so that the prior component used for StackOverflow for both the popular-users and randomly-sampled datasets was the same.
-
-However, a custom user prior for the Twitter popular-hashtags dataset cannot be used, since the entire Twitter dataset cannot be downloaded.
-The popular-hashtags dataset contains samples of a large set of various users that happened to use one of a specific set of hashtags being monitored.
-So in order to generate a custom user prior for each sample in the dataset, that user's entire tagging history must be accessed through the API.
-Since access rates are limited across that interface, it is infeasible to do this process for every user that appears in the Twitter popular-hashtags dataset.
-So instead, the global prior tag history for all users in the Twitter popular-hashtags dataset was used to compute prior activation values for each tag.
-
-\subsubsection{Co-occurrence Matrices}
-
-The context component for both the random permutation and Bayesian model is based on the same co-occurrence matrix of words and observed tags.
-Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
-The matrix for \num{3000000} posts was only used for Twitter, since the number of words in a StackOverflow post is much higher than Twitter on average,
-and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too large for a current high-performance desktop, even with highly optimized representations.
-Already \num{835500000} total co-occurring word tag observations were used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
-It was also the case that \num{1000000} posts for StackOverflow was already around $1/7$ of the total posts created on the site,
-so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrence matrix instead,
-especially since \num{835500000} co-occurrences were already used with \num{1000000} posts.
-
-\subsection{Adding Offset}
-
-The first methodological concern was the offset component.
-To test the effect of adding the component on model accuracy, the offset component was added to both models, and model accuracy was examined before and after the addition.
-
-\subsubsection{Method}
-
-The four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset were used for the analysis.
-Logistic regressions were run both with and without the offset component added to each model.
-The results for each run for each dataset were collected and averaged.
-The results across the four Twitter datasets were collected together, and the average was taken across all runs in all four datasets.
-
-\subsubsection{Results}
-
-The impact of adding the offset to each of the models for StackOverflow and Twitter is shown in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.
-In both plots, if the model name includes ``full'' then all model components are included.
-When the model components are enumerated, then at least one of the components is left out.
-The random permutation and Bayesian models are plotted alongside each other so that comparisons across the two model types can be done easily.
-So, for example, the ``standard prior model relaxed across posts without offset'' means that only the prior model component was used and the offset for that component was not subtracted off when determining activation.
-As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
-  \caption{
-    Impact of adding offset component for StackOverflow.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean model accuracy across all runs in dataset.
-    The random permutation models are represented as ``RP'' and Bayesian models as ``Bayes.''
-}
-  \label{figContextOffsetSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
-  \caption{
-    Impact of adding offset component for Twitter.
-    All error bars are the 95\% bootstrapped confidence interval for the mean.
-  }
-  \label{figContextOffsetT}
-\end{figure}
-
-Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites
-(\numNoZero{.53} to \numNoZero{.60} and \numNoZero{.22} to \numNoZero{.45} for Bayesian and random permutation for StackOverflow,
-\numNoZero{.31} to \numNoZero{.29} and \numNoZero{.24} to \numNoZero{.25} for Bayesian and random permutation for Twitter),
-except that the full Bayesian model without an offset was slightly more accurate on Twitter.
-Accuracy is improved even when only a single term is used.
-Although in this case the term is not weighted,
-but using an offset still improves accuracy since the offset term ensures that only a small number of tags per posts are predicted by the model to be the chosen tags.
-
-The random permutation model in particular is greatly affected by the offset term, for both the StackOverflow and Twitter.
-This suggests that the mean activation for the random permutation model for the top few tags in each post can vary greatly across posts.
-Only if this mean activation is subtracted will the logistic regression model generate optimal weights and predictions that are comparable to the Bayesian model when model accuracy relaxed across posts is examined.
-
-\subsubsection{Discussion}
-
-It is somewhat disappointing that the offset term is even needed.
-If it were the case that model activation for a tag in a specific post could be interpreted as an exact activation value that did not depend on the specific post,
-then the model would be able to say when, for example, the third top tag in a post is more likely to be a chosen tag than the second top tag in another post.
-That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of tags to use for a specific post.
-If this were the case, then adding the offset term would actually decrease accuracy, since this absolute value of the activation values would be removed.
-However, since accuracy improves when adding the offset term for both models, it does not seem that absolute activation values have much meaning across posts for these datasets,
-and that the relative activation values for each tag within a post is what is important.
-
-One simple and likely explanation is that there are posts where a large set of tags have high activation,
-but the author is choosing to use only a fixed small set of the top tags for each post, regardless of the activation values of the top tags.
-The author determines the number of tags to use for a post based on competing goals of tagging posts and not overly saturating a post with tags.
-For example, the author may be writing a question on StackOverflow about \emph{javascript}, and a large set of related tags are retrieved from declarative memory that are all above the retrieval threshold
-(e.g., \emph{javascript}, \emph{dom}, \emph{svg}, \emph{html}, \emph{jquery}, \emph{css}),
-but the author only chooses to use the tags \emph{javascript}, \emph{dom}, and \emph{svg}, since he has decided to tag this post with three tags, and those are the three tags with the highest activation.
-Adding the offset term into the model is a way to implement a simple version of this strategy where the author is actively choosing to use only a small set of tags for each post.
-Also, this type of strategic behavior of limiting the total number of chosen tags could be easily implemented in the production system of ACT-R.
-
-\subsection{Twitter Subsets}
-
-As a second methodological concern, model accuracy was examined across each of the dataset subsets collected for Twitter.
-Four popular-hashtags subsets were collected for Twitter at different time periods.
-This was to ensure that model results could be interpreted as effects that were not specific to a certain time period or set of popular hashtags.
-
-\subsubsection{Method}
-
-The four subsets were collected across a period of one month.
-Each was collected sequentially, and about one subset was collected per week.
-The hashtags used in each subset were chosen by scraping a site that published the current trending hashtags for each city.
-Before starting to collect for each subset, the current trending popular hashtags for the 400 most populous US cities were collected, and then 400 popular hashtags were randomly sampled from this set.
-Only 400 were selected since this is the maximum that the Twitter API can monitor at one time.
-
-In order to ensure that model results are not dependent on the specific Twitter subset collected,
-it is important that the overall accuracy differences between the different models do not drastically shift from subset to subset.
-That is, it is important that the effect size of model is much greater than the effect size of the interaction between Twitter subset and model.
-A small interaction relative to the main effect means that the rank ordering of the different models remains consistent across the different Twitter subsets.
-Further, a small effect size of subset relative to the effect of model means that the exact accuracy for each model remains consistent across the subsets.
-To measure the effect sizes of subset, model type, and the interaction, a two-way between-subjects ANOVA was performed.
-
-\subsubsection{Results}
-
-Model accuracy results for components of both the random permutation and Bayesian models grouped by Twitter popular-hashtags subset are depicted in Figure \ref{figContextSubsets}.
-Accuracy is affected significantly by subset ($F(3, 340) = 764, p < .001, \eta ^ {2} = .12$), model ($F(4, 340) = 3617, p < .001, \eta ^ {2} = .77$),
-and the interaction between model and subset ($F(12, 340) = 142, p < .001, \eta ^ {2} = .09$).
-However, the effect size of model type on accuracy is \num{8.6} times larger than the interaction term ($\eta ^ {2} = .77$ versus $\eta ^ {2} = .09$).
-This shows that the performance of a model relative to the other models remains consistent across the four subsets.
-Further, the effect size of model type is \num{6.3} times larger than the effect of Twitter subset ($\eta ^ {2} = .77$ versus $\eta ^ {2} = .12$).
-This shows that not only is relative model performance maintained across Twitter subsets, but the exact value of model accuracy for each model remains reasonably consistent across subsets.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-contextStandardByGroupT-crop.pdf}}}
-  \caption{Model accuracy for each of the four Twitter popular-hashtags subsets}
-  \label{figContextSubsets}
-\end{figure}
-
-\subsubsection{Discussion}
-
-Since the effect sizes of subset and the interaction between model type and subset are minimal relative to the effect size of model,
-model behavior for the popular-hashtags subset for Twitter can be interpreted as behavior that will persist across different time periods,
-if the same method of choosing popular hashtags is conducted at a later date.
-It may also be the case that that model behavior does not depend on the specific method of choosing the popular hashtags, but since only a single method was used for this analysis,
-the results from this study cannot by themselves be used to answer this question.
-
-\subsection{Stop Words}
-
-The method of handling stop words has been shown to influence model accuracy \parencite{Sahlgren2008,Stanley2013}, so this was the first architectural concern that was examined.
-\textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
-A data-driven frequency-filtering approach was tried, along with using a previously-generated stoplist, and no stop word removal at all.
-Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and using a previously-generated stoplist hardly increased performance.
-However, \citeauthor{Sahlgren2008} did not try to weight the stop words and only tried methods to remove them.
-So alongside looking at frequency filtering, previously generated, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on entropy was also explored.
-
-\subsubsection{Method}
-
-The runs from the Twitter popular-hashtags and StackOverflow randomly-sampled datasets were used for the analysis.
-Both the random permutation and Bayesian models were tested.
-Activations for context were computed by using five different methods to handle stop words:
-The two removal methods used in \textcite{Sahlgren2008}
-(i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist \parencite{Salton1988}),
-the entropy-weighting method used in \textcite{Stanley2013},
-combining the entropy-weighting and frequency-filtering methods,
-and no removal or weighting of stop words.
-
-\paragraph{Determining Frequency Cutoff}
-
-A cutoff point must be identified and used in order to remove the high-frequency words.
-\textcite{Sahlgren2008} used a cutoff where words that occurred more than \num{15000} times were removed.
-Since the StackOverflow and Twitter datasets are much larger than the datasets used by \citeauthor{Sahlgren2008}, the same cutoff could not be used.
-
-The ideal cutoff was determined by choosing the value where the standard deviation of the counts for each word for each of the tags were all relatively small.
-The plots for the standard deviations across tags that were used to identify the cutoff for StackOverflow and Twitter are shown in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT}.
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{memMatSO-crop.pdf}}}
-  \caption{
-    Variance in observation counts of words for each tag for StackOverflow.
-    Plots are shown for the four different methods for handling stop words.
-    Each plot contains the standard deviation of the total counts for each word in context as a function of hashtag.
-    High values mean that there are context words for a hashtag that have a high number of counts relative to the counts for all other context words for that hashtag.
-  }
-  \label{figContextCutoffSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{memMatT-crop.pdf}}}
-  \caption{
-    Variance in observation counts of words for each tag for Twitter.
-  }
-  \label{figContextCutoffT}
-\end{figure}
-
-The no processing plot shows the size of the spikes when high-frequency words are not removed.
-The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used (compare the difference between the no processing subplot and the entropy subplot).
-The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
-The cutoff that produced the frequency plots in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
-This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset was not given.
-
-Note that using a predefined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
-Also, no strong claims are being made that the cutoff identified is the ideal frequency percentage cutoff for all datasets when a frequency cutoff is used.
-The method used to identify this cutoff was simply intended to find a reasonable value where comparisons between a frequency cutoff and entropy weighting techniques would be equitable.
-
-\subsubsection{Results for the Random Permutation Model}
-
-Plots for the three methods of handling stop words for the random permutation model on the StackOverflow dataset are shown in Figure \ref{figContextStopRPSO}.
-The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
-Using a predefined stoplist makes the smallest improvement (\numNoZero{.45} to \numNoZero{.49}), while the entropy (\numNoZero{.60}) and frequency (\numNoZero{.63}) techniques improve performance the most,
-and frequency shows a slight edge over entropy.
-However, using the frequency-filtering technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsSOPerm-crop.pdf}}}
-  \caption{
-    Stop-word techniques for the random permutation model for StackOverflow.
-    The five methods for handling stop words are shown both for the title model component in isolation and the full model.
-    ``Stoplist'', ``freq'', ``entropy'' are the stoplist removal method, removal based on frequency count, and weighting based entropy metric.
-}
-  \label{figContextStopRPSO}
-\end{figure}
-
-The three stop word techniques were then tested for the random permutation model on the Twitter dataset, and the results are depicted in Figure \ref{figContextStopRPT}.
-For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\numNoZero{.25} to \numNoZero{.26}).
-Removing stop-words based on a predetermined list actually decreases accuracy (\numNoZero{.25} to \numNoZero{.23}), and using data-driven frequency filtering has only a small to no effect (\numNoZero{.25} to \numNoZero{.25}).
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTPerm-crop.pdf}}}
-  \caption{
-  Stop-word techniques for the random permutation model for Twitter.
-  The five methods for handling stop words are shown both for the context model component (i.e., no prior term) and the full model.
-}
-  \label{figContextStopRPT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-Comparing across Figures \ref{figContextStopRPSO} and \ref{figContextStopRPT}, entropy and frequency techniques are superior to filtering based on a predetermined list.
-However, it is more difficult to choose the best approach when comparing entropy weighting and frequency filtering.
-Frequency filtering is slightly better for StackOverflow, while entropy weighting is slightly better for Twitter.
-However, frequency filtering does have a few drawbacks compared to entropy weighting:
-It produces accuracy measurements that have higher levels of noise (see Figure \ref{figContextStopRPSO} and the frequency model for the title component in StackOverflow).
-Also, the ideal cutoff threshold must be identified in order to filter based on frequency,
-and identifying this cutoff requires a search, and this search must be done for each dataset that frequency filtering is used on.
-This cutoff value is a free parameter in frequency filtering that is not required when weighting stop words by entropy.
-Therefore, favoring parsimony, for the random permutation model, attenuating stop words based on entropy is a slightly superior method compared to removing stop words based on frequency of use.
-This result is interesting since filtering stop words based on frequency is the more common default method for random permutation models \parencite{Sahlgren2008}.
-This suggests that a simpler method (entropy weighting) can be used instead of the more common frequency-filtering technique to achieve the same (or better) accuracy measurements.
-
-\subsubsection{Results for the Bayesian Model}
-
-Since accuracy measurements when using a predetermined list to remove stop words for the random permutation model were much worse than the frequency and entropy methods, this method was not examined for the Bayesian model.
-The results for the frequency and entropy techniques for the Bayesian model on the StackOverflow dataset are depicted in Figure \ref{figContextStopBSO}.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsSOSji-crop.pdf}}}
-  \caption{Stop-word techniques for the Bayesian model for StackOverflow}
-  \label{figContextStopBSO}
-\end{figure}
-
-The results show that entropy weighting produces the most accurate results (\numNoZero{.61} to \numNoZero{.63}).
-Also, accuracy actually decreases when using frequency filtering (compared to no filtering) (\numNoZero{.61} to \numNoZero{.56}) for the Bayesian model.
-So entropy weighting is the clear winner in this case.
-
-Similar results are found when the Bayesian model is tested on the Twitter dataset.
-Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer (\numNoZero{.29} to \numNoZero{.32}),
-and frequency filtering decreases performance relative to no filtering (\numNoZero{.29} to \numNoZero{.28}).
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTSji-crop.pdf}}}
-  \caption{Stop-word techniques for the Bayesian model for Twitter}
-  \label{figContextStopBT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-One result that should not be overlooked is how poorly using a predetermined stoplist performed when compared to the two data-driven frequency and entropy approaches.
-This is most likely because the StackOverflow and Twitter datasets are large and contain high-frequency words that are domain specific (e.g., emoticons) and are not included in the predetermined list.
-This shows that data-driven techniques to identify stop words can be much more effective than using a predetermined list.
-
-Another interesting result is how the Bayesian model actually performed worse when stop words were removed based on frequency compared to no stop word removal.
-This was not the case for the random permutation model in this study, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random permutation model,
-and is most likely never the case for the random permutation model.
-Model accuracy most likely decreases for the Bayesian model because this model already has a normalization component built in that the random permutation model does not have.
-When computing activation for the random permutation model, the correlation is calculated directly on a matrix of counts of co-occurrences for words and tags.
-For the Bayesian model this count matrix is converted to a log-odds ($S_{ji}$) matrix,
-where both the number of observations for the word and the tag are normalized when computing the value for each cell (see the $S_{ji}$ computation in Table \ref{tabModACTRModel}).
-So it is likely the case that a form of frequency attenuation is already built into the Bayesian model.
-Therefore, adding a frequency filter on top of this does nothing to improve accuracy and only removes predictive information.
-
-\textcite{Stanley2013} found that the entropy-weighting technique to attenuate stop words for the Bayesian model significantly improved model accuracy.
-The same effect is found for the StackOverflow and Twitter datasets in this study.
-Further, when this entropy-weighting technique is compared directly to the more common frequency-filtering technique, weighting by entropy produces more accurate results.
-This was clearly the case for the Bayesian model.
-Both frequency and entropy techniques performed equally for the random permutation model.
-However, since the entropy-weighting technique is parameter free and does not require a cutoff to tune, it is also a better choice for the random permutation model for these datasets.
-
-The effect sizes when comparing no stop-word removal technique to either the entropy or frequency techniques are quite large.
-This shows the importance of correctly identifying and handling stop words when building co-occurrence matrices for large-scale natural language datasets.
-For further research, it may be worthwhile to explore an even larger range of stop-word handling techniques,
-since model accuracy seems to be quite sensitive to the specific method used. 
-However, one conclusion that can be made from this research is that the entropy-weighting method is a contender, and should be considered much more often than it has been previously.
-
-\subsection{Corpus Size}
-
-The number of documents used to build the co-occurrence matrix has been shown to influence model accuracy \parencite{Sahlgren2008}, so this was the second architectural concern that was examined.
-The default number of documents used to build the co-occurrence matrix was \num{1000000} for StackOverflow and \num{3000000} for Twitter.
-In order to see how accuracy changed as function of corpus size, 
-co-occurrence matrices for a range of smaller corpus sizes were built.
-Both the random permutation and Bayesian models were tested on all corpus sizes.
-
-\subsubsection{Method}
-
-For StackOverflow, the following number of posts were used to build separate co-occurrence matrices: \num{1000}, \num{10000}, \num{100000}, and the default \num{1000000}.
-For Twitter, these matrix sizes were tested alongside the default \num{3000000}.
-All runs from the four Twitter popular-hashtags datasets and the randomly-sampled StackOverflow dataset were used to test model accuracy.
-
-\subsubsection{Results for StackOverflow}
-
-Model accuracy as a function of co-occurrence matrix size for StackOverflow is depicted in Figure \ref{figContextDocumentSizeSO}.
-Accuracy improves for both the Bayesian and random permutation models as the size of the documents used to generate the co-occurrence matrix increases.
-However, model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest corpus size for the random permutation model, while the frequency technique only begins to plateau.
-The Bayesian model does not plateau and accuracy continues to improve as corpus size increases.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-freqVsEntropyBySizeSO-crop.pdf}}}
-  \caption{
-    Effect of size of co-occurrence matrix for StackOverflow.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean model accuracy across all runs in dataset.
-  }
-  \label{figContextDocumentSizeSO}
-\end{figure}
-
-\subsubsection{Discussion}
-
-It is likely that the random permutation model is starting to saturate when entropy weighting is used for the largest corpus size for StackOverflow.
-Saturation happens when certain words co-occur with a tag much more often than all of the other words.
-This introduces range effects when the correlation is computed, which ends up driving all of the correlations near one, and consequently decreases discrimination and accuracy.
-Saturation happens when entropy weighting is used because the non-predictive words are only attenuated and not removed, and this attenuation is not enough to overcome the range effects that grow with corpus size.
-
-However, the effect of saturation is small compared to the large effect from using either frequency filtering or entropy weighting to handle the stop words.
-The Bayesian model does not saturate when using entropy weighting because each cell in the co-occurrence matrix is normalized by word frequency and tag frequency.
-
-\subsubsection{Results for Twitter}
-
-The overall trend of increasing accuracy with corpus size is present for Twitter as well.
-Both entropy weighting and frequency filtering in Figure \ref{figContextDocumentSizeT} plateau for the random permutation model with increasing corpus size.
-Once again, the Bayesian model continues to improve with corpus size.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-freqVsEntropyBySizeT-crop.pdf}}}
-  \caption{
-    Effect of size of co-occurrence matrix for Twitter.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean.
-  }
-  \label{figContextDocumentSizeT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-It is interesting to compare model accuracy between the Bayesian and random permutation model at the smallest and largest corpus sizes for both StackOverflow and Twitter.
-At smaller corpus sizes, the random permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as corpus size increases.
-In other words, the Bayesian model needs larger co-occurrence matrices than the random permutation model to work properly.
-As the size of the dataset increases, the compression may start to lose crucial information, and the uncompressed Bayesian model should be used.
-As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random permutation model is actually beneficial, and the compressed random permutation model should be used.
-This noise is beneficial because it helps soften spurious $S_{ji}$ values when the counts for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller corpus sizes.
-
-This result has important implications.
-The corpus size is rarely chosen, and more likely constrained by the number of documents in the dataset being studied.
-If that dataset is relatively small, it may be better to use the compressed random permutation model than the Bayesian model, and vice-versa if the dataset is large.
-
-However, corpus size is not the only factor that determines whether the Bayesian model or random permutation model will perform better.
-For example, the corpus size where the Bayesian model begins to outperform the random permutation model is quite different for StackOverflow and Twitter.
-This is because the number of observations used to generate the co-occurrence matrix is a function of four components:
-corpus size, number of words in each document, total number of unique tags, and number of tags used per document.
-The average number of tags used per post for StackOverflow and Twitter were the same order of magnitude (around 3 for StackOverflow, slightly less for Twitter).
-
-However, the average number of words in a StackOverflow post is much greater than the 140 character limit used for tweets.
-More words per post means that the number of observations per cell in the co-occurrence matrix grows more quickly with corpus size.
-The Bayesian model is more likely to perform better with datasets like StackOverflow that have these larger corpus sizes.
-
-Also, the popular-hashtags Twitter dataset only contains 400 unique tags while the randomly-sampled StackOverflow dataset has \num{34377} unique tags.
-More unique tags means that the co-occurrence counts are spread across a wider range of tag columns,
-which means that it takes a larger number of total documents before the Bayesian model performs better than the random permutation model.
-
-The transition point where accuracy for the Bayesian model outperformed the random permutation model was much higher for StackOverflow than Twitter.
-This is most likely because StackOverflow has almost \num{100} times more unique tags than the Twitter popular-hashtags dataset.
-The number of words in a StackOverflow post is also much higher than Twitter, which should have lowered the transition point.
-However, the effect of words in a post must be less than the effect of number of unique tags, as the transition point was still much higher for StackOverflow given that it should have lowered due to post size.
-
-The compressed random permutation model should perform better with smaller corpus sizes, less words per document, more unique tags, and fewer number of tags per post.
-However, since there are so many factors that determine whether the Bayesian or random permutation model will perform better,
-and only two different datasets were examined in this research,
-it is difficult to make any specific recommendations as far as exact transition cut points (e.g., corpus size, words per document) where either the Bayesian or random permutation model should be used.
-Also, it is likely that the specific domain being studied interacts with the factors already mentioned to determine which model is better for that domain.
-Nevertheless, if a researcher is working in a domain with limited data (e.g., small corpus size, large tag space) that produces a sparse co-occurrence matrix,
-then it may be worthwhile to experiment with using a compressed random permutation model instead of the full Bayesian model.
-
-\subsection{Amount of Compression for Random Permutation}
-
-Another parameter in the random permutation model is the number of rows used to represent all of the different words.
-Each word's environment vector is represented by two randomly placed ones and two negative ones across the rows.
-All that is required is that the four positions and signs of each of these environment vectors is unique.
-But this means multiple environment vectors can overlap when looking at a single row.
-This is a lossy form of compression, and allows the number of rows to represent the words to be much less than the total number of words.
-As the number of rows decreases, the amount of cross talk between the environment vectors increases, compression increases, and the amount of information stored decreases.
-As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.
-
-\subsubsection{Method}
-
-Three different row-size values were tested for the random permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
-A low value (\num{100}), a more standard value (\num{2048}), and a high value (\num{10000}) of rows were tested.
-A separate co-occurrence representation was built for each of these row values.
-Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
-Accuracy measurements were the same as were used for previous runs of these datasets: mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.
-
-\subsubsection{Results}
-
-Results for using different levels of compression for the random permutation model for the StackOverflow dataset are depicted in Figure \ref{figContextDimSO}.
-In the figure, the model name that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
-This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random permutation model for all other results.
-The plot is also broken out for different size datasets used to build the co-occurrence matrix.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
-  \caption{
-    Effect of compression on random permutation model for StackOverflow.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean.
-  }
-  \label{figContextDimSO}
-\end{figure}
-
-Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy (\numNoZero{.60} to \numNoZero{.55} for the \num{1000000} document corpus size).
-But \num{2048} rows is sufficient for these datasets,
-as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy (\numNoZero{.60} to \numNoZero{.60} for the \num{1000000} corpus size).
-
-Results are similar for Twitter.
-Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100} (\numNoZero{.26} to \numNoZero{.24} for the \num{3000000} corpus size),
-and little to no impact when increasing from \num{2048} to \num{10000} (\numNoZero{.26} to \numNoZero{.26} at \num{3000000}).
-This results are consistent across all corpus sizes used to build the co-occurrence matrix.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeT-crop.pdf}}}
-  \caption{
-    Effect of compression on random permutation model for Twitter.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean.
-  }
-  \label{figContextDimT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-The fact that the random permutation model only needs around \num{2048} rows to represent all of the words in the Twitter and StackOverflow datasets is impressive.
-Twitter has around \num{2383000} unique words for the \num{3000000} tweets in each of the popular-users datasets.
-StackOverflow has \num{5848000} unique words for the \num{1000000} posts used to generate the co-occurrence matrix.
-So it is impressive that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
-This is likely because word use frequency for StackOverflow and Twitter approximately follows Zipf's law.
-\textcite{Stanley2013} showed that frequency of tag use for StackOverflow approximately follows a Zipf's law distribution.
-This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply when moving to words that are used fewer times. 
-Since these words have little to no ability to influence predictions, it makes sense that they can be safely removed from the dataset.
-
-However, it should be pointed out that the way the random permutation model compresses the words is completely random.
-It is not an LSA approach where the few ones and negative ones (i.e., the factors) for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
-The environment vector for each word is randomly selected, and no SVD is performed to choose these in a more discriminating manner.
-So it is that more impressive that even with a randomly-chosen form of compression, only \num{2048} rows are needed to represent all of the words in these datasets.
-
-\subsection{Combining Prior and Context}
-
-The activation value for the context component of the vector-based model is a correlation, ranging from zero to one.
-This is not a log-odds value.
-For the previous results, prior and context for the random permutation model were combined by simply adding the terms.
-However, this means that ACT-R's prior base-level learning component (which is a log-odds value) is added to a correlation value, which may not be mathematically appropriate.
-As a fourth architectural concern, another method to combine the prior and context components for the random permutation model was explored.
-
-\subsubsection{Method}
-
-For a given retrieval, the random permutation model returns a set of activations that are correlation values for each of the tags.
-Instead of simply adding this correlation value to the prior model term, it may be more appropriate to convert the correlation value to a log-odds value first, and then add this context term to the log-odds prior term.
-So a method was created to convert a distribution of correlations to a distribution of log-odds values.
-To do this, the one-tailed cumulative distribution function value for each correlation in the distribution was computed.
-That probability was then converted to a log-odds value in the usual way.
-
-For example, say that the model is choosing between a set of tags, and those tags have a mean correlation of .2 with context and a standard deviation of 0.1.
-If a tag in the set has correlation of .2 with context, then approximately half of the observed tags have a correlation lower than this tag.
-So the one-tailed cumulative distribution value for this tag is .5.
-If a tag in the set has a correlation of .3, then approximately half plus .34 (cumulative distribution of normal curve from mean to 1 standard deviation outward) of the observed tags have a lower correlation.
-So the one-tailed cumulative distribution value for this tag is .84.
-
-This technique has a few advantages: 
-It is simple and computationally cheap.
-Also, it behaves properly at specific points in the distribution.
-For example, if the conversion is performed on the mean correlation value in the set, the result is a log-odds value of zero.
-This makes sense given that log odds can be interpreted as the amount of odds information (positive or negative) that should be adjusted to the prior log odds, given that new piece of information.
-Since the mean correlation value contains no information regarding if it is better or worse than the other correlation values, it is appropriate to assign this correlation value a zero log-odds adjustment value.
-
-The log-odds transformation technique was tested for the random permutation model on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.
-
-\subsubsection{Results}
-
-Model accuracy results after using the log-odds transformation are depicted in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}.
-Using the log-odds transformation technique has a small effect on model accuracy for StackOverflow (\numNoZero{.45} to \numNoZero{.47}) and minimal to no effect on Twitter (\numNoZero{.25} to \numNoZero{.25}).
-When the technique is used and accuracy is examined for a model with only a single predictor, the transformation does not change accuracy.
-When predictors are combined and the full model is examined, the transformation is in the direction of slightly increasing accuracy, at least for the StackOverflow dataset.
-However, any improvement when using the transformation is minimal.
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
-  \caption{Random permutation model accuracy when using log-odds transformation on context for StackOverflow}
-  \label{figContextLogoddsSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
-  \caption{Random permutation model accuracy when using log-odds transformation on context for Twitter}
-  \label{figContextLogoddsT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-Vector-based models such as BEAGLE \parencite{Jones2007} and the random permutation model \parencite{Sahlgren2008} have included only a context component in the past (i.e., no prior component).
-When the prior component is included for the random permutation model, accuracy improves significantly (compare accuracy for full models to context models in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}).
-However, it is somewhat surprising that it does not matter much how the context and prior terms are added for the random permutation model.
-Model performance is similar when the context component used for the model is based on a correlation compared to a log-odds value.
-
-There may be a slight edge in increased accuracy when a log-odds transformation is used for the StackOverflow dataset.
-However, that result is not statistically reliable for the sample size used in this analysis (confidence intervals overlap in Figure \ref{figContextLogoddsSO}),
-and even if present the overall change in accuracy is small (\numNoZero{.45} to \numNoZero{.47}).
-Nonetheless, it does seem appropriate to convert the context to a log-odds adjustment value when context is added as an additional term to the random permutation model.
-Since performance does not decrease when the transformation is made, may slightly increase,
-and leads to a more natural interpretation of each model term as a log-odds value, it is reasonable to include this step for the random permutation model.
-Future research could explore additional methods to properly combine the terms.
-
-\subsection{Word Order}
-
-As a final architectural concern, vector-based models with and without word order were tested to examine the effect of word order on model accuracy.
-Vector-based models like the random permutation model can easily incorporate word order when making predictions.
-This is one of their primary strengths that has been discussed \parencites{Jones2007, Sahlgren2008}.
-In order to thoroughly test the impact of adding word order into the model, various methods for incorporating word order were tested on the StackOverflow and Twitter datasets.
-
-\subsubsection{Method}
-
-Three different methods for incorporating word order in the random permutation model were tested:
-taking just the sign of the word's position relative to the tag (direction method), using the position (standard method), and using the position but only including words that were used near the tag (window method).
-For the standard method, all words and their respective position are used to build the representation.
-For the direction method, essentially two representations are created: all co-occurrences with words to the left of the tag, and all co-occurrences with words to the right of the tag.
-The window method is similar to the standard method, where each word's relative position is maintained, but it only includes words near the tag to build the representation.
-\textcite{Sahlgren2008} tested similar methods, and found that accuracy was highest when a narrow window was used for the window method (only including words at most two positions to the left and right of the tag).
-The same $+2$ $-2$ range was used to test the window method on the Twitter dataset.
-
-Only the Twitter dataset was tested, since the StackOverflow dataset contains no word order information between words in the title and body of the post and the tags used for the post.
-The author chooses tags after filling out the title and body forms in the post, so the tags are never intermixed with words in the post like they are for Twitter.
-
-The same four Twitter popular-hashtags datasets were used for testing.
-Both an ordered and orderless context component were created for the random permutation models.
-The orderless context component uses the same co-occurrence matrix used by the Bayesian model to create the representation, since the Bayesian model does not contain any order information.
-Models were tested that used only one of the three word order methods for the ordered context component, used only the orderless context component, used a combination of the two components, and combined context with prior.
-
-\subsubsection{Results}
-
-Model accuracy results after incorporating word order into the random permutation model are depicted in Figure \ref{figContextOrder}.
-All bar plots that include an order component where ``window'' or ``direction'' is not included in the title use the standard ordering method (no direction or windowing, all position information is included).
-This is the default ordering method used, so the names for these bar plots were kept consistent to the names used in other figures so that comparisons could be made across figures if necessary
-(similar to what was done for the default row size for the random permutation model).
-Base models that do not use entropy weighting were also included so that effect sizes could be compared between entropy weighting and adding word order information.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-orderT-crop.pdf}}}
-  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
-  \label{figContextOrder}
-\end{figure}
-
-Looking first at models where no prior is added, the random permutation model using only orderless context performs similar to the model using only order context with direction.
-All other order-only models perform worse than the order model using direction and worse than the orderless model.
-
-Full models were also plotted against the model where only the prior and orderless context was included (i.e., no order).
-This was to test how much of the variance in model accuracy is shared between the orderless and order models.
-The results show that there is little unique variance in the order-only models, so order information contributes little to overall model performance above and beyond what an orderless model contributes.
-This can be seen by comparing the two-component model with only prior and orderless context to the three-component model with prior, orderless, and any of the three methods for the order context term.
-
-\subsubsection{Discussion}
-
-This was a surprising result.
-One of the main benefits discussed for the random permutation model is that it can easily incorporate word order information into the representation.
-But including word order information does not dramatically improve performance. 
-For the random permutation model, it is much more important to properly handle low-predictor stop words and include a prior component than it is to include word order information,
-at least for the Twitter dataset tested in this study.
-
-The original BEAGLE model \parencite{Jones2007} actually showed only minimal improvement when incorporating word order as well.
-In absolute terms, this model only improved by 2 percentage points in accuracy when trained on the TASA and tested on the TOEFL.
-\textcite{Sahlgren2008}'s results for word order for the random permutation model are similar.
-At a row size of around \num{2000}, model accuracy improves by less than a percentage point when order information is incorporated.
-The results for larger row sizes are somewhat inconsistent, since the full model (order and orderless) sometimes underperforms the order-only model,
-which should never happen since the order-only model is a subset of the full model.
-\textcite{Sahlgren2008} most likely did not use a regression technique to optimize the weights for each of the order and orderless model terms.
-So although incorporating word order into the random permutation model does seem to slightly improve performance,
-the strength of the model is more likely to be in how efficiently it can represent the co-occurrence matrix rather than its ability to include word order in the representation.
-That is, the strength of the model is in its compressed unordered context component term and not the ordered component, at least for the Twitter dataset studied.
-
-Also, the reason why the directional version of ordering is the best predictor of the three ordered components is most likely because order information does not have a large effect,
-so the two representations for words to the left and right of a tag are highly correlated.
-This means that this directionally-ordered component is essentially composed of all the co-occurrence observations in the orderless component after being randomly sampled and partitioned into two separate representations.
-The co-occurrences in these representations are highly correlated with each other, and highly correlated with the orderless component,
-which is why model accuracy is similar for the orderless component and direction ordered component, and why accuracy does not improve when adding the direction ordered component to the orderless component.
-
-Since it was shown that including word order information has little to no significant improvement on model accuracy, no attempt to incorporate word order information into the Bayesian model was made.
-
-\subsection{Coefficients for Predictors}
-
-All of the methodological and architectural tests were done on the Twitter popular-hashtags dataset and the StackOverflow randomly-sampled dataset.
-For these tests, a logistic regression statistical technique was used to determine the optimal weight for each model component.
-A regression was done for each run in the dataset, and since multiple runs were used for the StackOverflow and Twitter datasets, multiple samples of coefficient weights for each model component were computed.
-The variance of the weights for each component show how likely it is that the regression is overfitting the data,
-and the mean shows the best estimate weight of each component relative to the other.
-
-\subsubsection{Method}
-
-Coefficients for all of the runs for the Twitter popular-hashtags and StackOverflow randomly-sampled datasets were collected.
-This totaled 5 sets of coefficient weights for StackOverflow (5 runs) and 60 for Twitter (15 runs for each of the 4 dataset samples).
-
-\subsubsection{Results}
-
-Coefficients for each model component of the full models are depicted in Figures \ref{figContextCoeffSO} and \ref{figContextCoeffT}.
-The range of the distribution of coefficient values within a component is much more narrow than the range of the mean coefficient values between the components of each model.
-This shows that the model is not being overfit, as the weights for each component are relatively stable across the different runs.
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffSO-crop.pdf}}}
-  \caption{
-    Coefficient values for each component of the full models for StackOverflow.
-    Models used are the best-performing Bayesian and random permutation models
-    (i.e., Bayesian and RP have entropy weighting, RP converts context activation from correlation to log odds).
-    Error bars represent the standard deviation of coefficient values for each component.
-    The standard deviation is used instead of the confidence interval of the mean so that the plots show the overall stability of each coefficient across the different runs.
-  }
-  \label{figContextCoeffSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-coeffStd-coeffT-crop.pdf}}}
-  \caption{Coefficient values for each component of the full models for Twitter}
-  \label{figContextCoeffT}
-\end{figure}
-
-The weight of the prior relative to the context terms is also fairly consistent across models and across the two sites.
-Across the sites, context is a much stronger predictor than past behavior for StackOverflow, but past behavior is more important for Twitter. 
-Within a site, both models are consistent in how they favor context or past behavior.
-Also, looking at StackOverflow, both models are consistent in weighting the body more than the title, and the title more than the prior term.
-
-\subsubsection{Discussion}
-
-It was not surprising that the model was not overfit in this scenario, since the number of observed tags is much greater than the number of coefficients being fit.
-Each model has at most three terms, those weights are being fit to 500 posts within each run for Twitter and 100 for StackOverflow, and multiple tags are used for each post (particularly for StackOverflow).
-It is interesting that although the values of the weights are stable within a site (StackOverflow or Twitter), the weights change dramatically when looking across the sites.
-This illustrates that the impact of a user's past history relative to the current context can be domain specific, which means that for best results the weights for each of the terms should be recalibrated in new scenarios.
-However, the underlying reason that the weights change between the sites cannot be answered from this study alone.
-This relative impact may be changing from either a true alteration in the underlying weight of the prior compared to context for different domains,
-or simply due to the fact that the amount of information collected for each of the terms changes between the domains (less information leads to lower weighting in general).
-
-\clearpage
-\section{Chapter 6: Combining Predictors on the Popular-Users Dataset}
-
-Chapter 4 isolated the prior component of the model and analyzed accuracy on the popular-users dataset.
-Chapter 5 added the context components to the models and analyzed accuracy on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.
-Chapter 6 shows the results when the combined prior and context models are tested across all of the datasets.
-Looking at model behavior across the datasets provides a way to show how the importance of the prior component relative to the context component can change depending on specific properties in the dataset that is studied.
-Also, an experiment to a user-custom context co-occurrence matrix was attempted on the popular-users dataset.
-This was tried because it could be theoretically interesting to see if model performance improves when the context co-occurrence matrix is customized to each user.
-
-\subsection{Impact of Prior vs. Context}
-
-Adding a user's prior tag history as an additional component has been shown to improve performance, but is this always the case?
-If enough context information has been collected, maybe a context-only model can perform just as well as a combined context and prior model.
-If this were the case, then it may be less important that the random permutation model does not include a prior term by default.
-
-\subsubsection{Method}
-
-Full models were tested against context-only models for both sites. 
-The four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset were used for the test.
-
-\subsubsection{Results}
-
-The results are depicted in Figures \ref{figContextPriorneededSO} and \ref{figContextPriorneededT}.
-For StackOverflow, accuracy remains the same when the prior component is removed for both the Bayesian (\numNoZero{.63} to \numNoZero{.63}) and random permutation (\numNoZero{.63} to \numNoZero{.63}) models.
-Accuracy does dramatically decrease for Twitter when the prior is removed, however (\numNoZero{.32} to \numNoZero{.14} for Bayesian, \numNoZero{.25} to \numNoZero{.12} for random permutation).
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
-  \caption{Model accuracy when prior is removed for StackOverflow randomly-sampled dataset}
-  \label{figContextPriorneededSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddT-crop.pdf}}}
-  \caption{Model accuracy when prior is removed for Twitter popular-hashtags dataset}
-  \label{figContextPriorneededT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-These results show that prior information is not always needed when a large amount of context is available and not much information has been collected about each user's past history.
-The StackOverflow randomly-sampled dataset selects posts across the entire dataset, regardless of each user's popularity or tag history.
-Across the dataset, each user only asks around five questions on average, so there is not a large amount of past tag history available for the prior component to use,
-which is why this component is not a strong predictor in this case.
-
-However, can a context-only model perform just as well as a combined context and prior model when users with the largest amount of StackOverflow history are tested?
-If each user's tagging history is a useful predictor for StackOverflow,
-then the prior component should at least generate unique predictive variance in the case where users with large amounts of history are tested.
-
-\subsubsection{Method}
-
-The same StackOverflow popular-users dataset used to test the prior component in isolation was used to test the prior component when context was added.
-However, since computing activations for the context component take an order of magnitude more wall-clock time than computing the prior component, not all observations could be included for the test.
-To reduce the observations, the most recent 50 posts from each user in the datasets were selected.
-The total number of users sampled was also reduced to 40 users for StackOverflow and 80 users for Twitter.
-Using the most recent observations for each user ensures that the observations tested have acquired the greatest amount of past tagging history in the datasets.
-So if a context-only model performs just as well as the full model in this case, it would be surprising.
-
-\subsubsection{Results}
-
-The results comparing the full model to a context-only only model for the StackOverflow popular-users dataset are depicted in Figure \ref{figPUAddPrior}.
-In this case, past tag history does contain unique variance for accurately predicting tag use.
-Accuracy improves for both the Bayesian (\numNoZero{.51} to \numNoZero{.63}) and random permutation (\numNoZero{.47} to \numNoZero{.58}) models
-when the prior model component is combined with context compared to context alone.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddPUserSO-crop.pdf}}}
-  \caption{Model accuracy when prior is removed for StackOverflow popular-users dataset}
-  \label{figPUAddPrior}
-\end{figure}
-
-\subsubsection{Discussion}
-
-This shows that incorporating past tagging history is important even for a site like StackOverflow where a large amount of context information has been collected.
-However, for this type of site it only becomes important when a reasonably large amount of a user's past tagging history has been collected.
-For a randomly sampled user on StackOverflow, a context-only model captures the majority of the predictive variance.
-
-This raises an important applied question: Exactly how much past user behavior is enough for a site like StackOverflow to benefit from adding past user behavior as an additional predictor?
-When the model predicts tags from a random sample of StackOverflow users, adding the prior component to the model does not significantly increase model accuracy compared to using context alone.
-However, when the model predicts tags from a set of users that ask the most questions on the site, adding the prior does significantly increase accuracy.
-For the StackOverflow dataset, a user asks on average 5 questions and adds 3 tags to a post.
-So the model uses approximately 15 prior observations when computing the prior component for a randomly-sampled user on the site.
-For the popular-users dataset used to generate the results in Figure \ref{figPUAddPrior}, a user asked 180 questions on average.
-So for this dataset, the model uses approximately 540 prior observations when computing the prior activation.
-So for the StackOverflow site, the prior component begins to generate unique predictive variance above and beyond the context component somewhere between when a user has generated 15 and 540 prior observations.
-This result will be different for other sites that have collected a different amount of contextual information.
-For sites like StackOverflow where the contextual component is already a strong predictor, it will take a larger number of prior observations before adding the prior component becomes beneficial.
-Alternatively for sites like Twitter where contextual information is much more noisy and not as strong of a predictor, it will take fewer observations before adding the prior becomes beneficial.
-
-\subsection{User-customized Co-occurrence Matrix}
-
-The co-occurrence matrix for the context component of both models has been built by aggregating across users and collecting all instances of context words and associated tags.
-That is, the default co-occurrence matrix is not customized to each user, so the same matrix is used for all users.
-This is commonly done because there is usually not enough context by tag co-occurrences for a single user to generate a stable co-occurrence matrix.
-If there was enough information, accuracy would most likely improve since it is not likely that every user's word by tag co-occurrence profile is exactly the same.
-However, since several hundred posts for each user in the popular-users datasets are available, an analysis was tried where user-custom co-occurrence matrices were used.
-
-\subsubsection{Method}
-
-Separate co-occurrence matrices were built for each user in the sampled StackOverflow and Twitter popular-users dataset.
-The same sampling method used to test the impact of the prior component compared to the context component for the popular-users datasets was used,
-and the co-occurrence matrices were built from the oldest posts that were not included in the testing set.
-So for each user, the co-occurrence matrix was created from their oldest posts, up to (and not including) their 50 newer posts that were used to test model accuracy.
-
-\subsubsection{Results}
-
-Model accuracy when using a user-customized co-occurrence matrix for StackOverflow and Twitter is shown in Figures \ref{figPUCustomSO} and \ref{figPUCustomT}.
-The results compare accuracy for models that use a global co-occurrence and user-specific co-occurrence matrix.
-For StackOverflow, model accuracy for the user-specific co-occurrence matrix (labeled as ``user sji'') can be compared against the default global co-occurrence matrix.
-Accuracy decreases dramatically when a user-specific co-occurrence matrix is used for both the Bayesian (\numNoZero{.64} to \numNoZero{.35}) and random permutation (\numNoZero{.58} to \numNoZero{.36}) models.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiSO-crop.pdf}}}
-  \caption{
-    Model accuracy with user-customized context component for StackOverflow.
-    A model name with ``user sji'' uses a co-occurrence matrix customized for each user instead of a single global co-occurrence matrix.
-  }
-  \label{figPUCustomSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-userSjiT-crop.pdf}}}
-  \caption{Model accuracy with user-customized context component for Twitter}
-  \label{figPUCustomT}
-\end{figure}
-
-The results are similar for Twitter.
-However, the entire dataset is not available for Twitter, so a global co-occurrence matrix condition can not be tested,
-since the popular-users dataset samples across a wide range of hashtags, and due to Twitter's API restrictions it would be infeasible to collect all posts that use each of the hashtags in the dataset.
-But user-specific co-occurrence matrices can be built, since the popular-users dataset does contain all context and hashtag observations for each user.
-
-Model accuracy when only the single user-customized context predictor is used is relatively poor (\numNoZero{.17} and \numNoZero{.21} for Bayesian and random permutation),
-and the prior term is a much better predictor (\numNoZero{.37}).
-Further, when this component is combined with the prior component,
-model performance improves only slightly over performance when only the prior is used (\numNoZero{.37} to \numNoZero{.39} for both the Bayesian and random permutation).
-
-\FloatBarrier
-
-\subsubsection{Discussion}
-
-Both models require large numbers of context co-occurrence observations to make accurate predictions from context information.
-Otherwise, the co-occurrence matrices do not have enough data to properly stabilize.
-Millions of posts were used to generate the global co-occurrence matrices for Twitter and StackOverflow.
-If this number of posts had been collected for a single user, then the user-customized co-occurrence matrix for that user should lead to more accurate predictions compared to using the global matrix.
-However, in this study only around 100 posts were used to build the user-specific matrices for each user.
-
-This is most likely why model accuracy was so poor when user-specific matrices were used:
-Even for the large datasets analyzed in this study, there is still not enough history collected for each user to build customized co-occurrence matrices.
-The prior model component can be customized for each user, since the prior is only concerned with tag use, and there are a reasonably small set of tags that the model is choosing from.
-However, the context component is built from the set of all of the prior tags crossed with all of the possible words, which is a much larger space than just the set of tags.
-So building a stable context component requires a much larger set of data than the prior component.
-
-That said, if a user only chooses a small set of tags (so the tag space is small), and computing a global co-occurrence matrix is infeasible (e.g., with the Twitter popular-users dataset), 
-then it may be worthwhile to add the user-custom context model component as an additional predictor.
-As was shown for Twitter, model accuracy can slightly improve over a model that just uses a custom prior, even when the co-occurrence matrix only contains around 100 posts.
-
-Further, if the number of co-occurrences is small, and a user-custom co-occurrence matrix is used, it may be advantageous to use the compressed random permutation model instead of the uncompressed Bayesian model.
-For the Twitter popular-users dataset, only a user-custom co-occurrence matrix could be built, and when this matrix is used the compressed random permutation model performed better than the uncompressed Bayesian model.
-When the global co-occurrence matrix was removed from the StackOverflow dataset, and a user-custom matrix was used instead, the accuracy for the random permutation model decreased less than the Bayesian model.
-This provides further evidence showing that there is an interaction between dataset size and optimal model:
-At lower numbers of co-occurrences the compressed random permutation model is able to perform better than the Bayesian model,
-and the Bayesian model surpasses the compressed model at higher co-occurrence counts once the cells in the full co-occurrence matrix stabilize.
-
-\clearpage
-\section{Chapter 7: General Discussion}
-
-The general discussion is separated into two parts.
-The first section will describe the overall results and reemphasize key findings.
-The final section will discuss general theoretical implications of the results and describe future applications and directions.
-
-\subsection{Overall Results}
-
-Table \ref{tabOverallResults} summarizes model accuracies for full models (all components) on all of the datasets tested in the study.
-After modifying the models by adding an entropy weighting mechanism to both models and a prior term to the random permutation model,
-accuracy for the compressed random permutation model was nearly equal to the Bayesian model.
-The largest difference between the two models was on the Twitter popular-hashtags dataset (7 percent).
-Both models performed nearly equal for the StackOverflow datasets tested,
-and both models were much more accurate when tested on StackOverflow compared to Twitter.
-
-\begin{table}[!ht]
-  \caption{
-    Summary of accuracy for each model and dataset.
-    The Bayesian model includes entropy weighting for the context words,
-    the RP model includes entropy weighting and log odds transformation when combining the context and prior components.
-    Both models use a user-customized co-occurrence matrix for the Twitter popular-users dataset, since a global co-occurrence
-    matrix is not available for this dataset.
-  }
-  \label{tabOverallResults}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{llll}
-      \hline
-	Site 		& Dataset		& Model						& Accuracy \\
-	\hline
-	StackOverflow	& popular-users		& Bayesian with entropy				& \numNoZero{.63} \\
-			& 			& RP with entropy and log odds 			& \numNoZero{.58} \\
-			& randomly-sampled	& Bayesian with entropy 			& \numNoZero{.63} \\
-			& 			& RP with entropy and log odds			& \numNoZero{.63} \\
-	Twitter		& popular-users		& Bayesian with entropy and user sji		& \numNoZero{.39} \\
-			& 			& RP with entropy, log odds, and user sji	& \numNoZero{.39} \\
-			& popular-hashtags	& Bayesian with entropy				& \numNoZero{.32} \\
-			& 			& RP with entropy and log odds			& \numNoZero{.25} \\
-	\hline
-    \end{tabu}
-  }
-\end{table}
-
-Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
-The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
-These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
-However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
-So it is most likely the case that the task is not done solely with declarative memory processes.
-
-Alternatively, perhaps authors of posts are taking advantage of context information that is not currently represented in the model.
-For example, time of day might be a useful predictor of tag use on Twitter.
-\textcite{Cohn2012} built a tag-prediction model for Twitter that was based primarily on temporal cycles of hashtag use.
-This technique could be naturally incorporated into a co-occurrence framework by including time of day as an additional context by tag co-occurrence for each post.
-This would allow time of day information to be used alongside the rest of the context information in the words of the post.
-A similar technique could be used to add additional context predictors to the model.
-These co-occurrences could either be added on as additional columns to the existing co-occurrence matrix,
-or a separate co-occurrence matrix could be used for each additional context component, similar to how separate matrices and context predictors were created for the words in the title and body of a post.
-Including additional context predictors such as time of day may help improve model accuracy,
-and perhaps it is the case that the rest of the variance left to predict is from these context components that the author is using but are not yet included in the model.
-
-\textcite{Anderson1991} also looked at similar tagging tasks with large datasets when deriving the prior component of the declarative memory theory.
-\citeauthor{Anderson1991} examined, for example, the frequency of word choice over time used for the title of articles for the New York Times,
-and used the results to derive the power law equation for the prior term in the declarative memory model.
-ACT-R's declarative memory theory claims that the system has evolved to make predictions that closely match the statistical structure of the environment.
-Using a power-law equation to model future tag use based on past tag use is a reasonable statistical approach to the problem.
-It is also the case that this same equation is an accurate model for the prior component of declarative memory, as was shown by \citeauthor{Anderson1991} in the original derivation, 
-and as is supported in the results from this study.
-
-\subsubsection{Model Accuracy Differences Across Sites}
-
-Model accuracy for StackOverflow was much higher overall than for Twitter.
-The Twitter popular-hashtags dataset contained 400 unique hashtags for each of the four dataset subsets.
-The StackOverflow randomly-sampled dataset contained \num{34377} unique tags, since it sampled posts across the entire site.
-As the model is tasked to predict fewer and fewer hashtags, it should become more accurate (less chance of making an incorrect prediction).
-However, even with a much smaller tag set to predict, model accuracy for Twitter was still not nearly as high as StackOverflow.
-This is likely due to the fact that tag frequency use drops off sharply for both StackOverflow and Twitter.
-This means that the models on both sites are generally switching between the top few hundred tags when making predictions.
-So, increasing the total unique number of tags past that point makes only a small negative impact on model accuracy.
-
-However, there was a large difference in total number of users for StackOverflow and Twitter.
-A user on StackOverflow contributed 2.5 times more posts to the StackOverflow randomly-sampled dataset than the Twitter popular-hashtags dataset.
-Also note that the large co-occurrence matrices for StackOverflow and Twitter that contained millions of documents were not user specific.
-This means that the co-occurrence matrix for Twitter was built from a much larger set of unique users compared to StackOverflow.
-It is likely that co-occurrence matrices for individual users are correlated but not equivalent.
-So, as the co-occurrence matrix is built from a larger number of unique users, a greater amount of cross-user noise is introduced into the matrix.
-This noise would decrease model accuracy since the contextual component of the model would lose predictive information by having a noisier co-occurrence matrix.
-This is most likely one of the main reasons that model accuracy for Twitter is much lower than StackOverflow:
-Tag prediction becomes much more challenging as the number of users on a site increases.
-
-The second most-likely reason that model accuracy is much higher for StackOverflow is that the co-occurrence matrix for this site contains much more data due to the large document size for the site.
-The co-occurrence matrix for the StackOverflow randomly-sampled dataset contained 20 times more observations than the matrix for one of the Twitter popular-hashtags datasets.
-Model accuracy generally increases as more observations are used to generate the co-occurrence matrices, since this allows each cell in the matrix to stabilize to its true value.
-
-\subsubsection{Representation}
-
-Given that the storage space required to represent the co-occurrences for the random permutation model is dramatically smaller than the uncompressed Bayesian model,
-and that the compression is lossy and essentially random (i.e., no SVD),
-it is quite interesting that the random permutation model performs nearly as well as the full Bayesian model for this task.
-
-These results pose an interesting and important fundamental question about declarative memory:
-How, precisely, is the neural system configured to represent co-occurrences for declarative memory?
-Since at the behavioral level both models perform similarly, it is possible that the neural system is using either representation to store co-occurrences.
-Vector-based models are simple enough that they can easily be implemented in a neural network.
-Perhaps a form of vector-based compressed storage is being used at the neural level,
-since it can be implemented in neurons and is an efficient way for the machinery to generate predictions that closely match the statistical structure of the environment
-(i.e., that closely match the uncompressed Bayesian representation).
-
-As neural imaging systems improve, this question of neural representation can be best answered by looking directly at the neural activity.
-However, there are possible behavioral studies that could be done to provide evidence that may favor one representation over the other:
-The largest differences between the Bayesian and vector-based models happen with small numbers of co-occurrence observations.
-The vector-based model begins to generate more accurate predictions sooner than the Bayesian models,
-since the noise introduced from the compression helps smooth out activation values until a larger number of co-occurrences are observed and the co-occurrence values stabilize.
-For a behavioral study, novel words could be introduced alongside a corpus of known words, and the participant could be asked to make retrievals of the novel words and known words as more novel words are introduced.
-Both models would make different predictions about the rate at which activation values for the novel words stabilize relative to the known words, and such make different predictions about retrieval error rates over time.
-Learning a new language might be a domain where large amounts of these sorts of retrievals could be collected.
-For instance, the online site Duolingo might contain enough collective data about language learning and word translation to better tease apart the two context models.
-
-\subsubsection{Word Order}
-
-Also, the relatively strong performance of the random permutation model is not due to the model's ability to represent word order.
-Performance is similar for the random permutation and Bayesian models when the co-occurrence matrix for both models is the same, and not based on word order.
-Performance also does not improve by much when word order is added as the last predictor in the random permutation model.
-The random permutation model seems to work so well primarily due to the way that the full co-occurrence representation is compressed.
-
-\subsubsection{Compression}
-
-As previously noted, compression for the random permutation model is essentially random.
-That is, no SVD is done to identify the most predictive components,
-so the compressed representation used by the random permutation model does not contain nearly as much of the variance from the original co-occurrence matrix that it could have.
-However, performing a full SVD on a co-occurrence matrix as large as the ones generated for StackOverflow and Twitter is computationally infeasible.
-Also, when deriving the LSA approach, \textcite{Landauer1997} explicitly state that they make no claims that the brain is computing an SVD at the neural level;
-only that the neural machinery is set up such that the representation makes predictions that are similar to what would have been generated if the matrix decomposition for an SVD would have been computed.
-Perhaps the randomly-compressed representation used for the random permutation model is the way in which the brain tries to efficiently achieve the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
-
-\subsubsection{Corpus Size}
-
-The random permutation model was more accurate than the Bayesian model at smaller corpus sizes, while the results were flipped for larger document sizes.
-This is interesting from a practical standpoint because the corpus size is rarely chosen in declarative memory research, and is constrained primarily by the amount of data available in the domain being studied.
-If a researcher is working within a domain that has a relatively small corpus size (generally near or less than \num{1000} documents),
-it may be worthwhile to test the random permutation model in this domain, as this compressed model may be more accurate than the uncompressed Bayesian model.
-
-Further, it is interesting to study the effect of model accuracy on corpus size because people require exposure to only a small number of documents before tagging documents appropriately.
-A user asked around five questions on average for StackOverflow, model accuracy was high for StackOverflow, but the model required training on 1 million documents before achieving that accuracy.
-The user only needed to ``train'' on five questions and the model needed to train on 1 million questions.
-
-This is a large discrepancy, but this does not necessarily invalidate the declarative memory models used for this research.
-First, the random permutation model was still accurate at smaller document sizes, so perhaps this result provides evidence to favor the random permutation model over the Bayesian model.
-However, the smallest corpus size tested for these models was \num{1000} documents, which is much greater than the average number of questions or tweets that a user contributes to the site.
-
-Second, the models may require large amounts of training because the co-occurrence matrices are being built from scratch.
-That is, the models are being trained on the StackOverflow and Twitter domains without any prior general co-occurrence information to start with.
-Presumably, when a person starts using a site, they can leverage all of their past exposure to different word co-occurrences and then simply add on top of that base the domain-specific co-occurrences.
-In that since, one could argue that the models are undertrained on generic co-occurrences (people are exposed to many more than one million documents in their lifetime) and overtrained on specific co-occurrences.
-It may be interesting to test these memory models on a hybrid co-occurrence matrix,
-where the majority of the observations are taken from a generic domain (e.g., american literature) and a minority are domain-specific (e.g., StackOverflow word, tag pairs).
-
-\subsubsection{Attentional Weight}
-
-The Bayesian declarative memory model for ACT-R has an attentional weight term $W_{j}$ that can be used to attenuate or remove stop words.
-The random permutation model was also easily modified to have an attentional weight term by attenuating each word's set of ones and negative ones values by that word's attentional weight.
-Both a method for stop-word removal (frequency filtering) and stop-word attenuating (entropy weighting) performed well for both models.
-Also, using a proper method for handling the low-predictor words produced some of the largest improvements in model accuracy compared to the other explorations
-(e.g., compared to word order, properly adding terms, compression amount for random permutation representation).
-Since the ACT-R declarative memory theory already has a component for attentional weight, it may be the case that the neural system performs some form of stop-word weighting when computing activations,
-and this is reflected in the attentional weight term.
-That is, $W_{j}$ can be defined mathematically, has the same general form across tasks, and both the frequency-filtering and entropy-weighting methods for stop-word removal are likely first candidates for the definition.
-This claim, however, would require that no top-down activation is used to compute the attentional weight, which would mean that the $W_{j}$ is not under conscious control, which is not likely to be correct.
-Perhaps it is a combination of top-down and bottom-up systems that determine $W_{j}$, and the entropy-weighting and frequency-filtering techniques are methods of formalizing the bottom-up component.
-
-\subsection{Implications and Future Directions}
-
-\subsubsection{Generalizing Results}
-
-It seems likely that the entropy method for computing the attentional weight term $W_{j}$ may be applicable to other declarative memory tasks.
-The entropy method is parameter free (unlike the frequency-filtering method), completely data driven (much like the co-occurrence matrix), and does not require tuning.
-It is common to handle stop words in some manner when working with NLP and large text documents.
-So, if these large natural language documents are used to test declarative memory, then the entropy method may be a more cognitively plausible way of dealing with stop words rather than 
-filtering based on a predetermined list or based on the frequency count in the dataset, particularly since the entropy method substantially increases model accuracy for these datasets and since it is parameter free.
-
-Adding a prior component to the random permutation model also increased model accuracy for both StackOverflow and Twitter.
-It is unsurprising that this occurred.
-However, it is quite surprising that this is the first time that the prior component has been included and tested in a random permutation model.
-If the random permutation model is explored as a declarative memory theory for other datasets, it should be worthwhile to include the prior component for those explorations.
-Although it did not seem to matter much how the prior component is combined with the context component in this study, 
-there may be better ways than the log-odds transformation to combine the terms.
-Further research could focus on exploring a larger set of methods for combining the context component of the random permutation model with the prior component from the ACT-R theory.
-
-\subsubsection{Activation Calculation}
-
-The Bayesian and vector-based memory models differ in how the co-occurrence matrix is represented (lossy compression for vector-based models).
-They also differ in how the activation of each tag is calculated, given the different representations.
-The vector-based memory models, like LSA, compute a correlation between the context and each tag, while the Bayesian models sum the log likelihoods that each word in the context occurs with each tag.
-These two methods for computing the activation are correlated but not identical.
-Further research could try to tease apart how model accuracy is independently influenced by the representation (compressed or not) and the activation calculation (correlation or log likelihood).
-To do this, four models could be constructed (uncompressed and compressed representation crossed with correlation and log likelihood activation) compared to the two in this study,
-so that the effects of representation and activation calculation could be partialed out.
-This would allow, for example, the method of compression for the random permutation vector-based model to be analyzed independent of activation calculation.
-The results would help determine the specific representation that the declarative memory system uses to represent co-occurrences,
-and also the specific activation calculation that is carried out by the system.
-
-\subsubsection{Future Applications}
-
-The tag-recommendation models developed through this research should be relevant to real-world tasks.
-Model accuracy is high enough (particularly for StackOverflow) for the developed models to be used as a foundation for tag recommendation systems that are deployed on Twitter or StackOverflow.
-However, the focus for this research was to evaluate the accuracy of two cognitively-plausible declarative memory models, so the models were more constrained than general machine learning models of tagging.
-So it is unlikely that these general models of declarative memory will outperform specific machine learning models that have been optimized for a specific tagging task. 
-Nevertheless, one of the advantages of the general declarative memory models is that they are task agnostic, so it would take minimal time to deploy these models on a new tagging task.
-
-\subsubsection{Summary}
-
-The end result of this research is [1] a better understanding of how a user's past tagging history influences future tag use,
-[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on large-scale real-world retrieval tasks,
-and [3] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
-This work has shown how past user history is important, and how a context-only model like the random permutation model can be modified to include the prior component.
-The work has also shown how the strength of the random permutation model may not be in its ability to represent word order,
-and rather in its ability to efficiently compress the full co-occurrence representation used for the Bayesian model.
-Also from a methodological perspective,
-this work has shown how large collections of publicly-available behavioral datasets can be used to explore and test different theories of declarative memory and their individual architectural components.
-Several different formalisms for ACT-R's attentional weighting term were also explored, and an entropy-weighting technique was found to perform well without requiring any parameter tuning.
-Finally, this work has helped raise the question about how co-occurrence observations are represented at the neural level, since two models with quite different representations performed similarly well.
-The results from the study and the new research questions raised will help to better understand how imposing specific architectural constraints on memory models influence what information is retrieved.
-
-\begingroup
-\setstretch{1}
-\setlength\bibitemsep{12pt}
-\clearpage
-\printbibliography[heading=bibintoc]
-\endgroup
-
-\clearpage
-\section{Appendix 1: Support Code}
-
-All software code created for this research is publicly available on GitHub \parencite{StanleyRepo2014}.
-
 \subsubsection{Directory Structure}
@@
 As an example, the ``runContext20g1s6'' function takes only a few minutes to complete, runs the models on a corpus size of \num{1000}, with only a single run per model, and only 20 sampled posts for each run.
-
-\end{document}

### 2014-12-05 writeups/dissertation/dissertation.tex
foo

@@
 
-\subsubsection{Directory Structure}
-
-The repository is separated into three main folders:
-dissProject, firstSOProject, and writeups.
-The firstSOProject folder contains the support code for the \textcite{Stanley2013} research.
-The dissProject folder contains support code for this dissertation research.
-The writeups folder contains all writeup projects created during the dissertation process (e.g., the LaTeX for this dissertation, Keynote presentations).
-
-Within the dissProject folder, there are two main files that contain the majority of the code for the dissertation research: scrapeUsers.py and preProcess.R.
-The python file contains the necessary code to scrape the Twitter site for tweets, import results into the Postgres database, and perform various cleanup functions that were easiest to do in Python.
-The preProcess.R file contains all modeling and analysis code for the project.
-Within the file, there are implementations for the Bayesian and random permutation models.
-Also there is code to test the models by running each model through the various StackOverflow and Twitter test datasets.
-
-\subsubsection{Model Implementations}
-
-The Bayesian model is implemented using the data.table package, so that fast joins can be performed on the table when computing the strength of association values.
-The random permutation model is represented as a matrix, so that correlations between the model and context vectors are fast.
-The matrix is not sparse, since there are currently no sparse matrix packages for R that have a vectorized function to compute correlations between a vector and a matrix.
-
-The Bayesian model requires slightly more time to compute activation values and also more memory space to store the representation.
-This is because the random permutation model is compressed and only needs to store around \num{2048} rows in the co-occurrence matrix.
-However, both models are implemented efficiently enough so that all analysis runs and activation computations can be kept in memory.
-64 GB of memory was required to perform the analysis, particularly on the largest Twitter (3 million) and StackOverflow (1 million) corpus sizes.
-
-\subsubsection{Design Decisions}
-
-All model manipulations (e.g., using an entropy weighting metric, using a model with only the prior term) were expressed as part of a configuration object.
-That object is a list of configuration name, value pairs.
-The configuration object is passed to the analysis code, which determines the specific type of run that is performed.
-
-This configuration design enables a large amount of code reuse and overlap.
-This helps minimize the likelihood of a programming mistake, since functions are reused multiple times and overall static code length is kept to a minimum.
-
-\subsubsection{Testing}
-
-Each analysis run produced a set of result files.
-Those result files were checked into the repository so that changes could be tracked.
-Code changes were tested by looking at differences between previously checked in output files and the output files produced after a code change was made.
-
-A small amount of unit testing was also performed.
-The core behavior of each model implementation (Bayesian and random permutation) was tested in isolation to ensure that the activation values returned by the models were mathematically correct.
-Unit test code is located within dissProject/test.
-
-\subsubsection{Parallelization}
-
-None of the analysis code is finely parallelized.
-Most of the inner loop in R ends up running in compiled C, and execution within those function calls occurs on a single core.
-However, since multiple models were explored, each different model configuration was executed in parallel.
-This type of course-grained parallelization was achieved using R's built-in Parralel package.
-Also some parallelization was done at the OS level when various make targets would run in parallel.
-
-\subsubsection{Database}
-
-All raw tweet and post data were stored in Postgres tables.
-Also, after the documents were tokenized, the tokenized versions of the raw data were stored in Postgres tables.
-When R ran an analysis, it would query the tables to retrieve the correct set of tokenized words and tags.
-So there is a lot of R code that uses the RPostgreSQL package to communicate with Postgres as the analysis is running.
-
-\subsubsection{Running the Analysis}
-
-All analysis runs were scripted and can be run by using the appropriate target in dissProject/Makefile.
-The three main targets to run the models through the prior-only runs on the popular users dataset,
-the full models on the randomly-sampled and popular-hashtags datasets,
-and the full models on the popular users dataset are ``parrunPriors'', ``parrunContext'', and ``parrunPUsers'' respectively.
-Each of these targets takes tens of hours to complete.
-``parrunContext'' is the longest and takes around 72 hours on a desktop machine with 64 GB of memory and 24 virtual cores.
-
-Smaller test runs were also created so that model changes could be more rapidly developed.
-These typically use smaller Twitter and StackOverflow corpus sizes (e.g., only \num{10000} StackOverflow posts) to create the co-occurrence matrix
-and a smaller number of posts to test the models on (e.g., 20 instead of 500).
-Running these functions takes around a minute to complete.
-To run these, the R function is called directly (i.e., not a make target) within an already-running R session.
-As an example, the ``runContext20g1s6'' function takes only a few minutes to complete, runs the models on a corpus size of \num{1000}, with only a single run per model, and only 20 sampled posts for each run.
+\input{../../docs/src/public.tex}
 

### 2014-12-05 docs/README/readme.tex
foo

\documentclass[man,floatsintext,donotrepeattitle]{apa6}
\frenchspacing
\raggedbottom
\title{README for Clayton Stanley's dissertation}
%\shorttitle{Comparing memory models of tag retrieval}
\shorttitle{README}
\author{Clayton Stanley}
\affiliation{Rice University}
\usepackage[group-separator={,}]{siunitx}
\usepackage{csquotes}
\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber,doi=false,uniquename=false]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{bibliography.bib}

\begin{document}
\maketitle

\section{Description}

\input{../src/public.tex}

\section{Configuring environment}

\subsubsection{Git Repositories}

There are two main git repositories for this work.
The main repository contains all of the source code but very little source data.
This repo is located on chil under ``srv/git/stanley-dissertation.git''.
The source data was separated from the source code for two reasons:
First, the size of the source data is much larger than the code, so if someone wants to look at just the source, it takes much less time to clone the source repo with this method.
Second, the source data for Twitter should not be made public, due to Twitter's terms of service,
and to encourage replication and transparency of research, the source code is publicized and available on GitHub.
So the source data cannot be included in this public repository.

The source repo is included as a submodule of the main repository, but in order to access it, you have to have credentials on the chil server.
This way access to the repository can be limited, but the repo is still properly linked to the main repository.

\subsubsection{Setting Up Environment}

Run the following commands to configure the environment:

\begin{itemize}
  \item ``git clone ssh://[chil uname]@chil.rice.edu/srv/git/stanley-dissertation'':
    Clone the main repository.
  \item ``git submodule update --init --recursive'':
    Initialize and update all submodules.
    This will pull in the source data submodule.
  \item ``make -C submodules/dissertationData pull'':
    Get additional compressed source data for the source data repo.
    This includes large .csv text data extracted from the postgres tables.
    These .csv files are too large to be tracked with git.
  \item ``make -C submodules/dissertationData extract'':
    Uncompress the compressed source data.
\end{itemize}

\end{document}

### 2014-12-05 check-dev-env.sh
foo

@@ -1,0 +1,47 @@
+#!/bin/bash
+
+set -e
+
+pathToThisDir="$( cd "$( dirname "${BASH_SOURCE[0]}" )" && pwd )"
+
+checkDependency() {
+	if eval "$1"; then 
+		echo "Passed test: $1"
+	else
+		printf "Failed. $2\n"
+		exit 1
+	fi
+}
+
+checkNoPassword() {
+	ip=$1
+	status=$(ssh -o BatchMode=yes -o ConnectTimeout=5 $ip echo ok 2>&1)
+	[[ $status == ok ]]
+}
+
+checkSSHHostname() {
+	egrep -iq "^Host[[:space:]]+$1$" ~/.ssh/config
+}
+
+checkFont() {
+	system_profiler SPFontsDataType | grep -iq "$1:"
+}
+
+
+hostnameTemplate=$(cat <<'EOF'
+Host chil
+	User		clayton	
+	HostName 	chil.rice.edu
+EOF
+)
+
+
+checkDependency "[[ -n '$(which make)' ]]" "Install make (suggestion: Install XCode and check the box to install command line tools)"
+checkDependency "[[ -n '$(which ssh-copy-id)' ]]" "Install ssh-copy-id (suggestion: sudo port install openssh)"
+checkDependency "checkSSHHostname chil" "chil hostname not configured: add something like this (replace User field with your chil login name) to ~/.ssh/config: \n$hostnameTemplate"
+checkDependency "checkNoPassword chil" "chil not good: use ssh-copy-id to enable passwordless login to chil using public/private keys; otherwise submodule updating gets tedious"
+
+echo "Initializing and updating submodules"
+( cd "$pathToThisDir" && git submodule update --init --recursive )
+
+echo "Success! Developer environment correctly configured"

### 2014-12-05 docs/README/readme.tex
foo

@@
 \addbibresource{bibliography.bib}
+%\usepackage{enumitem}
+\usepackage{paralist}
+\usepackage[doublespacing]{setspace}
+\usepackage{tocloft}
 
@@
 
-Run the following commands to configure the environment:
+First, clone the main repository:
 
-\begin{itemize}
-  \item ``git clone ssh://[chil uname]@chil.rice.edu/srv/git/stanley-dissertation'':
+\begin{singlespace}
+\begin{compactitem}
+  \item \emph{git clone ssh://[chil uname]@chil.rice.edu/srv/git/stanley-dissertation}:
     Clone the main repository.
-  \item ``git submodule update --init --recursive'':
+\end{compactitem}
+\end{singlespace}
+
+Then, cd to the top level of the repository and run the following commands to configure the environment:
+
+\begin{singlespace}
+\begin{compactitem} 
+  \item \emph{./check-dev-env.sh}:
+    Ensures that ssh keys, hostname alias, binary dependencies, etc. are all configured properly.
+    If this isn't run first, none of the following commands will work.
+  \item \emph{git submodule update --init --recursive}:
     Initialize and update all submodules.
     This will pull in the source data submodule.
-  \item ``make -C submodules/dissertationData pull'':
+  \item \emph{make -C submodules/dissertationData pull}:
     Get additional compressed source data for the source data repo.
@@
     These .csv files are too large to be tracked with git.
-  \item ``make -C submodules/dissertationData extract'':
+  \item \emph{make -C submodules/dissertationData extract}:
     Uncompress the compressed source data.
-\end{itemize}
+\end{compactitem}
+\end{singlespace}
+
+If you want this done all in one go, run the \emph{make post-clone} command at the top level of the repository:
+You'll need a version of make installed to do this.
+\emph{check-dev-env} checks for this binary dependency and includes a suggestion on how to install make if you don't have it.
 

### 2014-12-05 docs/README/readme.tex
foo

@@
 \usepackage[group-separator={,}]{siunitx}
+
 \usepackage{csquotes}
@@
 \DeclareLanguageMapping{american}{american-apa}
+\usepackage[american]{babel}
 \addbibresource{bibliography.bib}
+
 %\usepackage{enumitem}
@@
 
+\subsubection{Manual Environment Setup}
+
+Setting up the postgres tables is currently not automated.
+This involves creating the tables, all indexes, importing the .csv tables from the source data repository, and then updating the tables with a few additional columns.
+This process \emph{was} documented in the form of a script, but that script hasn't been tested.
+It contains code for both creating the tables and updating columns in the tables.
+Those steps must be done separately, and the tables should be imported into postgres before the updates are called.
+Currently that script is located in ``dissProject/createTables.sql''.
+I figure that no one other than myself will want to get the database configured,
+since the results of all the model runs are saved in separate .csv files and the db is not required to analyze or visualize them.
+If I ever have to confiure the database on another machine for this research, I'll separate the ``createTables.sql'' file into two separate files, 
+automate and test the whole process, and create an additional make target to configure the db.
+
+\printbibliography[heading=bibintoc]
+
 \end{document}

### 2014-12-05 docs/README/readme.tex
foo

@@
 
-\subsubection{Manual Environment Setup}
+\subsubsection{Manual Environment Setup}
 
@@
 since the results of all the model runs are saved in separate .csv files and the db is not required to analyze or visualize them.
-If I ever have to confiure the database on another machine for this research, I'll separate the ``createTables.sql'' file into two separate files, 
+If I ever have to configure the database on another machine for this research, I'll separate the ``createTables.sql'' file into two separate files, 
 automate and test the whole process, and create an additional make target to configure the db.

### 2015-05-27 writeups/psycMethods/psycMethods.tex
foo

@@
 \abstract{
-  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
+  The growth of social media and user-created content on online sites provides unique opportunities to study models of human declarative memory.
   The tasks of choosing a hashtag for a tweet and tagging a post on StackOverflow were framed as declarative memory retrieval problems.
@@
   Furthermore, past behavior was successfully incorporated into the random permutation model that previously used only context.
-  Also, ACT-R's attentional weight term was linked to a common entropy-weighting natural language processing method used to attenuate low-predictor words.
+  Also, ACT-R's attentional weight term was linked to an entropy-weighting natural language processing method used to attenuate high-frequency words (e.g., articles and prepositions).
   Word order was not found to be strong predictor of tag use, and the random permutation model performed comparably to the Bayesian model without including word order.
   This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
-  Finally, model accuracy was moderate to high for the tasks, which supports the theory that choosing tags on StackOverflow and Twitter is primarily a declarative memory retrieval process. 
+  Finally, model accuracy was moderate to high for the tasks, which supports the hypothesis that choosing tags on StackOverflow and Twitter is essentially a declarative memory retrieval process. 
   The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
@@
 The declarative memory component in ACT-R is a Bayesian statistical model.
-This is the same type of co-occurrence-based Naive Bayes statistical model that is prominent in the more general machine learning field.
+This is the same type of co-occurrence-based Naive Bayes statistical model that is prominent in the field of machine learning.
 When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
@@
 If a cognitive modeler is using the declarative memory component of ACT-R,
-it is common to either construct the co-occurrences by hand, or use a small training set of data to initially build up the declarative memory system.
+it is common to either construct the co-occurrences by hand, or use a small training set of data to initially parameterize the declarative memory system.
 This approach can and has worked well.
@@
 
-A few researchers have started to use these large databases to evaluate how well declarative memory models scale,
+A few researchers have started to use large-scale databases to evaluate how well declarative memory models scale,
 for both ACT-R Bayesian models \parencites{Fu2007,Pirolli2003,Stanley2013,Douglass2010} and vector-based memory models \parencites{Jones2007,Rutledge2008,Recchia2010,Sahlgren2008}.
@@
 The core question motivating this research is:
-What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
-This question can be tested by generating a prediction for the most likely used tag whenever a user is about to generate a tag, and then comparing the user's chosen tag to the model's prediction.
-This process can be framed as a memory retrieval problem.
+What kinds of cognitively-plausible models can best predict the tags that a user on a social media site selects?
+This question can be tested by generating a prediction for the most likely selected tag whenever a user is about to generate a tag, and then comparing the user's chosen tag to the model's prediction.
 The process of suggesting a relevant tag to a user can be thought of as a memory retrieval request for a tag, given prior tag use and current context.
 
-Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010} and StackOverflow tag use \parencites{Stanley2013}.
+Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag selection \parencite{Efron2010} and StackOverflow tag selection \parencites{Stanley2013}.
 Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
@@
 
-These two memory systems were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
+These two memory models were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
 Twitter is a microblogging service where users create 140 character ``tweets'' and broadcast those messages out to the users who follow their posts.
@@
 
-The two memory retrieval models were also tested on tag use for StackOverflow posts.
+The two memory retrieval models were also tested on tag selection for StackOverflow posts.
 StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
@@
 
-The ACT-R declarative memory theory and the vector-based retrieval models will be described.
-The Latent Semantic Analysis (LSA) theory is closely related to vector-based retrieval models, so an overview of LSA will also be included.
+The ACT-R declarative memory model and vector-based retrieval model will be described.
+Latent Semantic Analysis (LSA) is closely related to vector-based retrieval models, so an overview of LSA will also be included.
 The theories underlying the ACT-R and vector-based models will be compared.
@@
 ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
-The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
+The declarative memory system is a component of the architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
 The equations that make up this system were derived from a rational analysis of declarative memory retrieval.
@@
 However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
-LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix, while maintaining as much of the variance in the original full matrix as possible.
+LSA uses the Singular Value Decomposition (SVD) process to reduce the dimensionality of the word by document co-occurrence matrix, while maintaining as much of the variance in the original full matrix as possible.
 The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
@@
 
-Models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
+Models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites,
+though most of these models are not based on theories of human declarative memory.
 Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
@@
 This allowed the use of a logistic regression statistical technique to calibrate the weights for each of the terms.
-The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation \parencite{Dumais1991} and is parameter free.
+The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation \parencite{Dumais1991} and is parameter-free.
 
@@
 One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a tweet as a function of prior hashtag use and context \parencite{Mazzia2009}.
-The model presented by \textcite{Mazzia2009} is quite similar to an ACT-R declarative retrieval model.
-The main potential difference is that --- like \textcite{Stanley2013} --- the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
-
-Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used from similar tweets.
+This is quite similar to an ACT-R declarative retrieval model.
+The main potential difference is that---like \textcite{Stanley2013}---the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
+
+Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used in similar tweets.
 \textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
@@
 \textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
-The ACT-R declarative memory retrieval theory shows how past user behavior and current context combine to produce the most likely retrieval.
+The ACT-R declarative memory retrieval theory shows how past experience and current context combine to produce the most likely retrieval.
 Each component is summed together to generate a total activation, and then the likelihood of chunks are ranked by that summed activation.
@@
 Unlike StackOverflow, the entire Twitter dataset is not released to the public.
-Due to this and the scale of Twitter, only a part of the Twitter dataset relevant to this research was collected.
+Due to this and the scale of Twitter, only subsets of the Twitter dataset were examined.
+The relevant subsets are described below.
 
@@
 
-Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
+Four popular-hashtags datasets were collected at four different time periods, spanning one month.
 Multiple datasets were collected to ensure that the results were not dependent on a particular sample of hashtags and would generalize across different popular-hashtag datasets and different time periods.
@@
 The entire StackOverflow dataset is packaged and released quarterly to the public.
-So the most recent packaged dataset was downloaded and the database tables were imported into PostgreSQL.
+So the most recent packaged dataset was downloaded and the tables were imported into a relational database.
 
@@
 Twitter provides two types of APIs for acquiring data: one based on streaming and a direct query-based interface.
-The Tweepy Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
 The query-based interface was used to collect all tweets for specific popular users.
@@
 This algorithm was specifically created for tokenizing Twitter content.
-However, during testing it was also found to be robust and well suited for StackOverflow content.
+However, during testing it was also found to be robust and well-suited for StackOverflow content.
 The initial model results with the new tokenizer were qualitatively similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
-So the same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter, since performance was similar to \textcite{Stanley2013} for StackOverflow,
-and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.
+So the same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter, since it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.
 
@@
 The StackOverflow and Twitter popular-users datasets were used to characterize tag growth and decay for a user on each site.
-The popular-users dataset contains 12 slices of popular users for both StackOverflow and Twitter.
-Each user's tag behavior from the popular-users dataset for each site was analyzed.
-
-22 different values of the decay rate parameter were explored for each slice
+Each user's tag behavior from each site was analyzed.
+
+22 different values of the decay rate parameter were explored
 (0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.8, 2, 5, 10, 20).
@@
 Model accuracies at each decay rate value for each user in the dataset slice are included in the plots.
-Each user's mean accuracy was subtracted off in order to compute a normalized mean for that user.
+Each user's mean accuracy was subtracted in order to compute a normalized mean for that user.
 This was done so that the relative accuracy change for different decay rate values could be more easily visualized.
-The plotting technique is a analogous to what is done when testing the main effect in a repeated measures design (i.e., each subject's mean score is subtracted off).
+The plotting technique is a analogous to what is done when testing the main effect in a repeated measures design (i.e., each subject's mean score is subtracted).
 
@@
 method of handling stop words, influence of dataset size, method to combine prior and context, and word order.
-The four architectural concerns were examined for the following reasons:
+These particular architectural concerns were examined for the following reasons:
 During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.
@@
 \textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset.
-The result maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
+This method maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
 It also seems less arbitrary than explicitly identifying a list of stop words to remove from the analysis.
@@
 The two removal methods used in \textcite{Sahlgren2008}
-(i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist \parencite{Salton1988}),
+(i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist: \textcite{Salton1988}),
 the entropy-weighting method used in \textcite{Stanley2013},
@@
 
-The no processing plot shows the size of the spikes when high-frequency words are not removed.
+The ``no processing'' plot shows the size of the spikes when high-frequency words are not removed.
 The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used.
@@
 Both frequency and entropy techniques performed equally for the random permutation model.
-However, since the entropy-weighting technique is parameter free and does not require a cutoff to tune, it is also a better choice for the random permutation model for these datasets.
+However, since the entropy-weighting technique is parameter-free and does not require a cutoff to tune, it is also a better choice for the random permutation model for these datasets.
 
@@
 It is interesting to compare model accuracy between the Bayesian and random permutation model at the smallest and largest corpus sizes for both StackOverflow and Twitter.
-At smaller corpus sizes, the random permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as corpus size increases.
+At smaller corpus sizes, the random permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference diminishes as corpus size increases.
 In other words, the Bayesian model needs larger co-occurrence matrices than the random permutation model to work properly.
@@
 The number of words in a StackOverflow post is also much higher than Twitter, which should have lowered the transition point.
-However, the effect of words in a post must be less than the effect of number of unique tags, as the transition point was still much higher for StackOverflow given that it should have lowered due to post size.
+However, the effect of the number of words in a post must be less than the effect of number of unique tags, as the transition point was still much higher for StackOverflow given that it should have lowered due to post size.
 
@@
 Vector-based models like the random permutation model can easily incorporate word order when making predictions.
-This is one of their primary strengths that has been discussed \parencites{Jones2007, Sahlgren2008}.
+This has been touted as one of their primary strengths \parencites{Jones2007, Sahlgren2008}.
 In order to thoroughly test the impact of adding word order into the model, various methods for incorporating word order were tested on the StackOverflow and Twitter datasets.
@@
 The first section will describe the overall results, reemphasize key findings, and discuss the theoretical implications of each key finding.
-The final section will describe future applications and directions for the research.
+The final section will describe potential applications and directions for the research.
 
@@
 
-Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
+Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter reasonably well.
 The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
@@
 However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
-So it is most likely the case that the task is not done solely with declarative memory processes.
+One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval
+(e.g., strategic processes, determining the number of tags to employ, attempts to add humor).
 
@@
 How, precisely, is the neural system configured to represent co-occurrences for declarative memory?
-Since at the behavioral level both models perform similarly, it is possible that the neural system is using either representation to store co-occurrences.
+Since at the behavioral level both models perform similarly, it is possible that the neural system is using something similar to either representation to store co-occurrences.
 Vector-based models are simple enough that they can easily be implemented in a neural network.
@@
 Performance also does not improve by much when word order is added as the last predictor in the random permutation model.
-The random permutation model seems to work so well primarily due to the way that the full co-occurrence representation is compressed.
+The random permutation model seems to work so well mainly due to the way that the full co-occurrence representation is compressed.
 
@@
 The random permutation model was more accurate than the Bayesian model at smaller corpus sizes, while the results were flipped for larger document sizes.
-This is interesting from a practical standpoint because the corpus size is rarely chosen in declarative memory research, and is constrained primarily by the amount of data available in the domain being studied.
+This is interesting from a practical standpoint because the corpus size is rarely chosen in declarative memory research, and is constrained mainly by the amount of data available in the domain being studied.
 If a researcher is working within a domain that has a relatively small corpus size (generally near or less than \num{1000} documents),
@@
 Further, it is interesting to study the effect of model accuracy on corpus size because people require exposure to only a small number of documents before tagging documents appropriately.
-A user asked around five questions on average for StackOverflow, model accuracy was high for StackOverflow, but the model required training on 1 million documents before achieving that accuracy.
+On StackOverflow, each user asked an average of five questions.
+Model accuracy was high for StackOverflow, but the model required training on 1 million documents before achieving that accuracy.
 The user only needed to ``train'' on five questions and the model needed to train on 1 million questions.
@@
 Presumably, when a person starts using a site, they can leverage all of their past exposure to different word co-occurrences and then simply add on top of that base the domain-specific co-occurrences.
-In that since, one could argue that the models are undertrained on generic co-occurrences (people are exposed to many more than one million documents in their lifetime) and overtrained on specific co-occurrences.
+In that sense, one could argue that the models are undertrained on generic co-occurrences (people are exposed to many more than one million documents in their lifetime) and overtrained on specific co-occurrences.
 It may be interesting to test these memory models on a hybrid co-occurrence matrix,
@@
 and this is reflected in the attentional weight term.
-That is, $W_{j}$ can be defined mathematically, has the same general form across tasks, and both the frequency-filtering and entropy-weighting methods for stop-word removal are likely first candidates for the definition.
-This claim, however, would require that no top-down activation is used to compute the attentional weight, which would mean that the $W_{j}$ is not under conscious control, which is not likely to be correct.
-Perhaps it is a combination of top-down and bottom-up systems that determine $W_{j}$, and the entropy-weighting and frequency-filtering techniques are methods of formalizing the bottom-up component.
+That is, $W_{j}$ can be defined mathematically, has the same general form across tasks, and both the frequency-filtering and entropy-weighting methods for stop-word removal are possible first candidates for the definition.
 
@@
 It seems likely that the entropy method for computing the attentional weight term $W_{j}$ may be applicable to other declarative memory tasks.
-The entropy method is parameter free (unlike the frequency-filtering method), completely data driven (much like the co-occurrence matrix), and does not require tuning.
-It is common to handle stop words in some manner when working with NLP and large text documents.
-So, if these large natural language documents are used to test declarative memory, then the entropy method may be a more cognitively plausible way of dealing with stop words rather than 
-filtering based on a predetermined list or based on the frequency count in the dataset, particularly since the entropy method substantially increases model accuracy for these datasets and since it is parameter free.
+The entropy method is parameter-free (unlike the frequency-filtering method), completely data driven (much like the co-occurrence matrix), does not require tuning,
+and substantially increases model accuracy for these datasets.
+It is common to handle stop words in some manner when working with Natural Language Processing (NLP) and large text corpuses.
+So, if these large corpuses are used to test declarative memory,
+then the entropy method may be a more cognitively plausible way of dealing with stop words rather than filtering based on a predetermined list or the frequency count in the dataset.
 
@@
 
-\subsubsection{Future Applications}
+\subsubsection{Potential Applications}
 

### 2015-05-27 dissProject/preProcess.R
foo

@@ -1496,17 +1496,18 @@ plotLineSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol, titl
 	lineTypes = c(1,2,3,5)
 	lineTypes = lineTypes[1:length(sumTbl[, unique(DVName)])]
 	lineTypes
-	expr = bquote(ggplot(sumTbl, aes(x=factor(.(groupCol)), y=meanVal, colour=.(fillCol))) + 
-		      geom_line(aes(group=.(fillCol), linetype=.(fillCol)), size=1.4) + 
-		      geom_point() +
+	shapeTypes = c(15, 16, 17, 18, 12)
+	shapeTypes = shapeTypes[1:length(sumTbl[, unique(DVName)])]
+	expr = bquote(ggplot(sumTbl, aes(x=factor(.(groupCol)), y=meanVal)) +
+		      geom_line(aes(group=.(fillCol), linetype=.(fillCol), colour=.(fillCol)), size=.6) + 
+		      geom_point(aes(shape=.(fillCol)), size=3) +
 		      scale_x_discrete(breaks = unique(sumTbl[, .(groupCol)]), labels = sizeNums) +
 		      scale_colour_discrete(name=title) +
+		      scale_shape_manual(name=title, values=shapeTypes) +
 		      geom_errorbar(aes(ymin=minCI, ymax=maxCI), width=0.1, size=0.3) + 
 		      scale_linetype_manual(name=title, values=lineTypes) + 
-		      guides(fill = guide_legend(keywidth = 1, keyheight = 1),
-			     linetype=guide_legend(keywidth = 5, keyheight = 1),
-			     colour=guide_legend(keywidth = 5, keyheight = 1)) +
-		      defaultGGPlotOpts)
+		      defaultGGPlotOpts +
+		      theme(legend.key.width=unit(3, 'cm'), legend.key.height=unit(.5, 'cm')))
 	plot = eval(expr)
 	plotTbl(plot, extras, figName)
 	sumTbl
@@ -1569,7 +1570,7 @@ renameColGroupNum <- function(tbl) {
 			    newName=c('Dataset 1', 'Dataset 2', 'Dataset 3', 'Dataset 4'), key='groupNum')
 	tbl[, groupNum := as.character(groupNum)]
 	setkey(tbl, groupNum)
-	tbl[mapTbl, groupNum := newName]
+	tbl[mapTbl, groupNum := newName, nomatch=0]
 }
 
 renameColSizeNum <- function(tbl) {
[…]

### 2015-05-28 writeups/psycMethods/psycMethods.tex
foo

@@
 
-\title{Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow}
+\title{Comparing vector-based and Bayesian memory models using large-scale datasets:\protect\\User-generated hashtag and tag prediction on Twitter and StackOverflow}
 \shorttitle{Comparing memory models of tag retrieval}
@@
   The growth of social media and user-created content on online sites provides unique opportunities to study models of human declarative memory.
-  The tasks of choosing a hashtag for a tweet and tagging a post on StackOverflow were framed as declarative memory retrieval problems.
-  Two state-of-the-art cognitively-plausible declarative memory models were evaluated on how accurately they predict a user's chosen tags: an ACT-R based Bayesian model and a random permutation vector-based model.
-  Millions of posts and tweets were collected, and both declarative memory models were used to predict Twitter hashtags and StackOverflow tags.
+  By framing the task of choosing a hashtag for a tweet and tagging a post on StackOverflow as a declarative memory retrieval problem,
+  two cognitively-plausible declarative memory models were applied to millions of posts and tweets and evaluated on how accurately they predict a user's chosen tags.
+  An ACT-R based Bayesian model and a random permutation vector-based model were tested on the large datasets. 
   The results show that past user behavior of tag use is a strong predictor of future behavior.
@@
   Also, ACT-R's attentional weight term was linked to an entropy-weighting natural language processing method used to attenuate high-frequency words (e.g., articles and prepositions).
-  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparably to the Bayesian model without including word order.
+  Word order was not found to be a strong predictor of tag use, and the random permutation model performed comparably to the Bayesian model without including word order.
   This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
-  Finally, model accuracy was moderate to high for the tasks, which supports the hypothesis that choosing tags on StackOverflow and Twitter is essentially a declarative memory retrieval process. 
   The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
@@
 
-The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human cognitive processing.
-The declarative memory component in ACT-R is a Bayesian statistical model.
-This is the same type of co-occurrence-based Naive Bayes statistical model that is prominent in the field of machine learning.
-When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
-The primary reason for this is that the stability of statistical models increase as sample size increases.
-
-However, it is much less common in the cognitive modeling community to use a large number of observations to build up declarative memory when creating ACT-R models.
-If a cognitive modeler is using the declarative memory component of ACT-R,
-it is common to either construct the co-occurrences by hand, or use a small training set of data to initially parameterize the declarative memory system.
-This approach can and has worked well.
-Essentially, the modeler is only including the observations in declarative memory that are relevant to the specific task that they are studying.
-
-However, this approach is problematic for two main reasons:
-First, it means that the declarative memory system for every developed model must be hand crafted and customized for that model, which takes enormous time and effort.
-Second, it means that modelers are continuing to use small sample sizes for the declarative memory components in their models.
-
-This has important implications for understanding declarative memory:
-First, it is difficult to verify and explore modifications to the declarative memory architecture because the datasets are not large enough or rich enough to rigorously test modifications to the theory.
-It also means that we do yet fully understand how well the ACT-R declarative memory theory scales when the number of chunks begins to approach the number of chunks that are actually stored in human declarative memory.
-
-There is, however, an opportunity to change this approach.
-With the massive growth of human-created content on social media sites (e.g., Twitter, StackOverflow, Wikipedia), and public access to many of those databases,
+With the massive growth of human-created content on social media sites (e.g., Twitter, StackOverflow, Wikipedia) and public access to many of those databases,
 it is now possible to evaluate psychological theories of declarative memory on large-scale real-world tasks where the theory can be tested on hundreds of millions to billions of data points.
-This will provide a way to test if these models scale, and evaluate the psychological validity of these models as the number of observations used increases by orders of magnitude.
+Without using large datasets it is difficult to verify and explore modifications to the declarative memory architecture because the datasets are not large enough or rich enough to rigorously test modifications to the theory.
 
@@
 It focuses specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and vector-based) can be improved by testing these models on two real-world tagging tasks.
-One primary objective is to compare the random permutation vector-based model to ACT-R's declarative memory theory when large datasets are used for each model. 
+One primary objective is to compare the random permutation vector-based model to ACT-R's Bayesian declarative memory theory when large datasets are used for each model. 
 Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
@@
 ``Theoretical Background'' describes the vector-based and Bayesian declarative memory models that will be evaluated in this research,
-and shows how these models are related to the other common declarative memory models that are studied.
+shows how these models are related to the other common declarative memory models,
+and identifies examples of similar tagging models that have already been applied to various tag recommendation systems.
 ``General Methodology'' presents the overall method for testing the models on the StackOverflow and Twitter datasets.
@@
 
-The ACT-R declarative memory model and vector-based retrieval model will be described.
-Latent Semantic Analysis (LSA) is closely related to vector-based retrieval models, so an overview of LSA will also be included.
-The theories underlying the ACT-R and vector-based models will be compared.
-Also, different tagging models have already been applied to various recommendation systems to tag user-generated content.
-Many of these models are based on the same Bayesian statistical framework that is the foundation of the ACT-R declarative memory theory.
-So these systems will be described and the models used will be discussed within the context of Bayesian statistical models, LSA approaches, and the ACT-R declarative memory theory.
-
 \subsection{ACT-R Declarative Memory Theory}
@@
 The declarative memory system is a component of the architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
-The equations that make up this system were derived from a rational analysis of declarative memory retrieval.
+The equations that make up this system were derived from a rational analysis of declarative memory retrieval \parencite{Anderson1989}.
 That is, given the task of retrieving a chunk of information from declarative memory,
@@
 These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.
-
-\subsection{Latent Semantic Analysis Memory Theory}
-
-Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
-Both ACT-R's Bayesian declarative memory model and LSA compute activations based on a co-occurrence matrix.
-However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
-LSA uses the Singular Value Decomposition (SVD) process to reduce the dimensionality of the word by document co-occurrence matrix, while maintaining as much of the variance in the original full matrix as possible.
-The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
 
@@
 
+Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
+Both ACT-R's Bayesian declarative memory model and LSA compute activations based on a co-occurrence matrix.
+However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
+LSA uses the Singular Value Decomposition (SVD) process to reduce the dimensionality of the word by document co-occurrence matrix, while maintaining as much of the variance in the original full matrix as possible.
+The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
+
 Both vector-based models and LSA calculate word similarity on a reduced matrix.
@@
 
-Each specific result of the research requires a sightly different method to study.
-However, there is a common general methodology that applies across all results that will be described first.
 Different subsets of the StackOverflow and Twitter datasets were used evaluate past user behavior in isolation and to test the context component of each memory model.
@@
 
-Two models were evaluated for hashtag prediction on both the Twitter and StackOverflow datasets: a vector-based model and Bayesian retrieval model.
-The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} was used for the vector-based model,
-since it has shown to perform better than the original BEAGLE vector-based model \parencite{Recchia2010}.
-The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel} was used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.
-
+Two models were evaluated for hashtag prediction on both the Twitter and StackOverflow datasets:
+the random permutation vector-based model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} and 
+the modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel}. 
 The two memory models were evaluated on predicting user-chosen tags for Twitter tweets and StackOverflow posts.
@@
 \textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset.
+The technique is formally described in Table \ref{tabModACTRModel} and uses information theory to determine the amount of predictive information contained in each word.
+Words that are often used with a small number of tags have high information content while words used often across all tags carry low information content.
 This method maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
-It also seems less arbitrary than explicitly identifying a list of stop words to remove from the analysis.
-
 All three stop word techniques (i.e., predetermined list, high-frequency filtering, and \textcite{Stanley2013}'s entropy weighting) were explored for this analysis.
@@
 
-There is a large improvement in model accuracy when either an entropy or frequency technique is used to handle stop words.
-This shows the importance of correctly identifying and handling stop words when building co-occurrence matrices for large-scale natural language datasets.
-For further research, it may be worthwhile to explore an even larger range of stop-word handling techniques,
-since model accuracy seems to be quite sensitive to the specific method used. 
-However, one conclusion that can be made from this research is that the entropy-weighting method is a contender, and should be considered much more often than it has been previously.
-
 \subsection{Corpus Size}
@@
 The first section will describe the overall results, reemphasize key findings, and discuss the theoretical implications of each key finding.
-The final section will describe potential applications and directions for the research.
+The final section will describe future directions for the research.
 
@@
 
-\subsubsection{Potential Applications}
-
-The tag-recommendation models developed through this research should be relevant to real-world tasks.
-Model accuracy is high enough (particularly for StackOverflow) for the developed models to be used as a foundation for tag recommendation systems that are deployed on Twitter or StackOverflow.
-However, the focus for this research was to evaluate the accuracy of two cognitively-plausible declarative memory models, so the models were more constrained than general machine learning models of tagging.
-So it is unlikely that these general models of declarative memory will outperform specific machine learning models that have been optimized for a specific tagging task. 
-Nevertheless, one of the advantages of the general declarative memory models is that they are task agnostic, so it would take minimal time to deploy these models on a new tagging task.
+\subsection{Summary}
+
+This work has shown how large collections of publicly-available behavioral datasets can be used to explore and test different theories of declarative memory and their individual architectural components.
+The end result of this research is [1] a better understanding of how a user's past tagging history influences future tag use,
+[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on large-scale real-world retrieval tasks,
+and [3] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets.
+
+This work has shown how past user history is important, and how a context-only model like the random permutation model can be modified to include the prior component.
+The work has also shown how the strength of the random permutation model may not be in its ability to represent word order,
+but rather in its ability to efficiently compress the full co-occurrence representation used for the Bayesian model.
+Several different formalisms for ACT-R's attentional weighting term were also explored, and an entropy-weighting technique was found to perform well without requiring any parameter tuning.
+Finally, this work has helped raise the question about how co-occurrence observations are represented at the neural level, since two models with quite different representations performed similarly well.
+The results from the study and the new research questions raised will help to better understand how imposing specific architectural constraints on memory models influence what information is retrieved.
 

### 2015-05-28 writeups/psycMethods/psycMethods.tex
foo

@@
 The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
-Only values equal to or below 1 were examined for the optimized learning form, since values above 1 for this form cause a bifurcation and are not psychologically interesting.
-The bifurcation occurs because at a decay rate of 1 the denominator of the first term to compute $B_{i}$ goes to zero (see Table \ref{tabACTRBLLModel}).
+Only values below 1 were examined for the optimized learning form, since values above 1 for this form are not mathematically defined. 
+They are not mathematically defined at a decay rate greater than 1 because for these values the first term to compute $B_{i}$ is the logarithm of a negative number (see Table \ref{tabACTRBLLModel}).
+At a decay rate equal to 1, every base-level activation $B_{i}$ is infinite, which is not psychologically interesting.
 

### 2015-05-28 writeups/psycMethods/psycMethods.tex
foo

@@
 
-The work in this study expands on previous work that has tested retrieval models on large-scale datasets.
+This study expands on previous work that has tested retrieval models on large-scale datasets.
 It focuses specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and vector-based) can be improved by testing these models on two real-world tagging tasks.
@@
 A simplified version of \textcite{Stanley2013}'s method was used for the results for this research.
-Instead of subtracting off the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
+Instead of subtracting the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
 So for the prior component, the activation value used in the logistic regression is the prior activation with the mean of the top 5 prior activations subtracted from this value.
@@
 So a method was created to convert a distribution of correlations to a distribution of log-odds values.
-To do this, the one-tailed cumulative distribution function value for each correlation in the distribution was computed.
+To do this, the one-tailed cumulative distribution function value for each correlation was computed.
 That probability was then converted to a log-odds value in the usual way.
@@
 All bar plots that include an order component where ``window'' or ``direction'' is not included in the title use the standard ordering method (no direction or windowing, all position information is included).
-This is the default ordering method used, so the names for these bar plots were kept consistent to the names used in other figures so that comparisons could be made across figures if necessary.
+This is the default ordering method used, so the names for these bar plots were kept consistent to the names used in other figures so that comparisons could be made across figures.
 Base models that do not use entropy weighting were also included so that effect sizes could be compared between entropy weighting and adding word order information.

### 2015-05-28 writeups/psycMethods/psycMethods.tex
foo

@@
 
-Different subsets of the StackOverflow and Twitter datasets were used evaluate past user behavior in isolation and to test the context component of each memory model.
+Different subsets of the StackOverflow and Twitter datasets were used to evaluate past user behavior in isolation and to test the context component of each memory model.
 These datasets will be named and described in general terms.
@@
 the modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel}. 
-The two memory models were evaluated on predicting user-chosen tags for Twitter tweets and StackOverflow posts.
 The co-occurrence matrices used for the context component of each model were generated from a large subset of the Twitter and StackOverflow datasets.
@@
 This method maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
-All three stop word techniques (i.e., predetermined list, high-frequency filtering, and \textcite{Stanley2013}'s entropy weighting) were explored for this analysis.
 
@@
 That is, no SVD is done to identify the most predictive components,
-so the compressed representation used by the random permutation model does not contain nearly as much of the variance from the original co-occurrence matrix that it could have.
+so the compressed representation used by the random permutation model does not contain nearly as much of the variance from the original co-occurrence matrix than it could have.
 However, performing a full SVD on a co-occurrence matrix as large as the ones generated for StackOverflow and Twitter is computationally infeasible.

### 2016-01-24 writeups/psycMethods/psycMethods.tex
foo

@@
 
+\subsubsection{ACT-R DM Model}
+
+A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACTRModel}.
+
+\begin{table}[!ht]
+  \caption{ACT-R declarative memory model}
+  \label{tabACTRModel}
+  {\tabulinesep=1.2mm
+    \begin{tabu}{ll}
+      \hline
+      Common Name &  Equation \\
+      \hline
+      Activation &	 	$A_{i} = B_{i} + \sum_{j \in c}^{} W_{j} S_{ji}$ \\
+      Attentional Weight &	$W_{j} = \frac{W}{n}$ \\
+      Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
+      Constant Base Level &	$B_{i} = \frac{p_{i}}{1-p_{i}}$ \\
+      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
+      Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
+      \hline
+    \end{tabu}
+  }
+\end{table}
+
+The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base-level activation ($B_{i}$) and strength of association ($S_{ji}$).
+The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
+Each term in Table \ref{tabACTRModel} will be discussed in greater detail to follow.
+
+\subsubsection{Base-Level Activation}
+
+Base-level activation reflects the log prior odds of needing an observed chunk again.
+The default way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
+This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
+Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
+However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
+In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its frequency and recency of use.
+
+\subsubsection{Strength of Association}
+
+Strength of association ($S_{ji}$) reflects the amount of log-odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
+Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
+Context for the Twitter domain are the words in a tweet.
+Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
+the likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
+For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
+So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
+This assumption was described when deriving the $S_{ji}$ equation \parencite{Anderson1989}, and used in recent research working with large-scale datasets \parencites{Stanley2013,Farahat2004,Douglass2010}.
+
+\subsubsection{Connection to Pointwise Mutual Information}
+
+Pointwise Mutual Information (PMI) is another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
+It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
+The index has been used to accurately and efficiently measure strength of association information for large-scale datasets \parencite{Budiu2007,Farahat2004}.
+The PMI index is presented as Equation \eqref{eqPMI}.
+
+\begin{equation}
+  \label{eqPMI}
+  \mathit{PMI}(y,x) = log \frac{p(y|x)}{p(y)}
+\end{equation}
+
+Further, this PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
+Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.
+
+\subsubsection{Attentional Weight}
+
+The attentional weight term allows for some contextual components to be weighted more than others when the total $S_{ji}$ activation is computed.
+This reflects the fact that the declarative memory system has limited attentional resources ($W$) that must be spread across all of the items in context.
+Most ACT-R models simply weight each contextual element equally (e.g., all words in a StackOverflow post have equal weight), so each word's attentional weight is simply normalized by the number of items in context.
+
 \subsection{Vector-Based Memory Systems}
@@
 Over time, memory vectors accumulate a distributed representation of the most common environment vectors that co-occurred with them.
+
+\subsubsection{Retrieval Process}
+
+Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
+Decoding takes the memory vector for context and decodes it back into an environment vector.
+The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
+
+Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
+For example, assume that a user has written the word ``zend'' in a post, and she wants to assign a tag for the post.
+``zend'' has co-occurred with the tag \emph{PHP} many times previously, so the unordered memory vector for ``PHP'' ($c_{\mathit{PHP}}$) contains the unordered environment vector for \emph{zend} ($e_{zend}$).
+To determine the most correlated memory vector with context, a memory vector is created from the current context ($e_{zend}$).
+The cosine between that context memory vector and the memory vector for \emph{PHP} will be high, since the \emph{PHP} memory vector ($c_{\mathit{PHP}}$) often contains the context environment vector ($e_{zend}$).
+Since the cosine is high, the tag \emph{PHP} will most likely be returned as the vector that resonates highest with the current context.
+
+The resonance and decoding processes are inversions of each other.
+Usually both resonance and decoding will return similar rank orderings of most likely chunks given context.
+\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
 
@@
 Nonetheless, vector-based models have been shown to perform similar to if not better than LSA \parencites{Sahlgren2008,Jones2007}.
+
+\subsection{Comparison of Vector-Based Models and ACT-R}
+
+Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
+Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
+Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
+For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
+\citeauthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect.
+
+The fan effect describes a behavior of human memory where memory items are more difficult to retrieve if they have occurred in a larger number of different contexts.
+\textcite{Rutledge2008} used \textcite{Anderson1974}'s classic experimental paradigm to test the fan effect.
+In the task, participants are exposed to a set of person and place pairs (e.g., ``the hippie is in the park'').
+They are later asked to recall if a particular person and place pair was presented to them before.
+The main manipulation is the ``fan'' of the person and place for each pair.
+The fan is defined as the number of different contexts that each item in the pair is presented in.
+For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store,'' then ``hippie'' would have a fan of two.
+Recall accuracy decrease and time to recall increase as the sum total fan of the person and place increases.
+
+Both the ACT-R declarative memory theory and vector-based models produce the fan effect due to the effect that fan has on the co-occurrence matrix.
+As a particular word (e.g., ``hippie'') is seen in more and more contexts, the co-occurrence count of the word across all contexts is spread across multiple contexts.
+This means that when a retrieval for a context given the word occurs, multiple contexts have relatively high activation.
+This makes it more difficult to test if a signal (e.g., if a word appears in a specific context) is higher than the background noise (e.g., all other contexts that the word has appeared in),
+which makes the retrieval more uncertain.
+
+Both vector-based and Bayesian models produce the fan effect since they both are based on co-occurrences.
+However, vector-based models do further processing on that matrix, reduce its dimensionality, and use a correlation instead of a log odds value to compute the activation.
+Nonetheless, vector-based models still produce the fan effect because at their core they are still computing activations based on co-occurrences.
+
+\subsubsection{Base-Level Activation}
+
+Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the relative benefits of the two models. 
+In particular, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
+
+The frequency and recency of how content has been used in the past can be a strong predictor of the future importance of that information. 
+The recent content in tweets on a news site like Twitter should be more relevant, for example.
+\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
+They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.
+
+ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
+It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
+Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
+ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
+Vector-based models have focused on computing the associative strength of a chunk and specific contextual elements.
+It is unexplored how accuracy changes for these models once the information about a user's prior behavior (e.g., ACT-R's base-level activation component) is added.
+
+\subsubsection{Word Order}
+
+One strength of a vector-based approach over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
+With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
+To represent word order in a Bayesian model, an additional dimension could be added to the co-occurrence matrix that represents word position.
+However, adding word-order information to ACT-R's default Bayesian declarative memory model is not common.
+Instead, when word order information is used to model reading for example, production rules are defined so that words are read from left to right (e.g., \textcite{Lewis2006}).
+
+\subsubsection{Activation Calculation}
+
+Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
+Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
+These two computations are certainly not equivalent, even though they are most likely correlated with one another.
+
+Using a correlation can be problematic due to range effects in observed co-occurrences.
+If there are particular words for a tag that co-occur much more often than all of the other words with that tag, that single large value diminishes the power of all other words to contribute to the correlation calculation.
+This does not happen with Bayesian retrieval models because each pair of word and tag values is normalized by the total number of words that appeared with that tag and total number of tags that appeared with that word
+(see the strength of association equation in Table \ref{tabACTRModel}).
+The high-frequency words are often removed from co-occurrence matrix for vector-based models to reduce this effect.
 

### 2016-01-24 writeups/psycMethods/psycMethods.tex
foo

@@
 
+As a general methodology for this research, we are using a cognitively-constrained exploratory model development process.
+Since we are working with large-scale datasets, we have enough data to explore many different modifications simultaneously without the risk of overfitting.
+However, we are not interested in bringing to bear the entire collection of machine learning algorithms to this prediction task in order to completely optimize a model.
+Rather, we are interested in evaluating how well two cognitive models can explain the results, and then making selective modifications to them to see how performance changes.
+The modifications that we explored were picked such that they influenced model accuracy, could be easily added to (or removed from) the cognitive models without massive modifications, and could be considered cognitively plausible.
+
 \subsection{Research Questions}
@@
 
-\subsubsection{Addressing Word Order}
+\subsubsection{Word Order for Vector-Based Systems}
 

### 2016-01-24 writeups/psycMethods/psycMethods.tex
foo

@@
 
+\subsection{Testing for Reliable Differences}
+
+Confidence intervals were used throughout for statistical inference instead of explicit p-values.
+This was done because it is the most straightforward way to visualize all of the mean differences in accuracy (i.e., bar and line plots with a visual indication of the error) for each of the models, which eases interpretation
+(similar to the argument made by FIXME:textcite).
+It is also a conservative estimate of reliability, as using confidence intervals on the means essentially ignores the fact that this is a within-subjects design, where accuracy for each model is measured for each subject.
+However, we would rather err on the side of making the results easier to interpret at the cost of using conservative estimates for reliability
+(i.e., use error bars overlaid on a visualization of the means instead of statistical tests on the difference scores).
+In fact, using large datasets for analysis is what empowers us to do this:
+With this much data, we can use straightforward (although more conservative) inferentials and still see reliable results.
+This allows us to concentrate on proper data visualization of the main findings.
+
 \section{Past User Behavior}
@@
 Accuracy improves for both the Bayesian and random permutation models as the size of the documents used to generate the co-occurrence matrix increases.
-However, model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest corpus size for the random permutation model, while the frequency technique only begins to plateau.
-The Bayesian model does not plateau and accuracy continues to improve as corpus size increases.
+However for the random permutation model, accuracy begins to plateau for the largest corpus size for all three tested versions. 
+The Bayesian model does not plateau nearly as much and accuracy continues to improve even at the largest measured corpus size.
 

### 2016-01-24 dissProject/preProcess.R
foo

@@ -1806,7 +1806,11 @@ analyzePrior <- function(modelVsPredTbl) {
 	# Check that the Ns for each dataset look right	
 	modelVsPredTbl[, list(N=.N, names=list(unique(datasetName))), by=list(dsetType, dsetGroup, runNum,datasetNameRoot)]
 	bestTbl = modelVsPredTbl[predUsedBest == T]
+	bestTbl[runNum == 2, median(d), .(DVName, dsetGroup)]
 	tbl = modelVsPredTbl[topHashtag & hashtagUsedP] 
+	tbl
+	bestTbl[runNum == 2, .(.N, uniqueN(user_screen_name)), .(DVName, dsetGroup)]
+	bestTbl
 	#dvDiffsTbl = plotDVDiffs(rbind(bestTbl[runNum==2, compare2DVs(.SD, c('topHashtagPostPriorStd', 'topHashtagPostPriorOL2'), sortedOrder=c(1,2)), by=list(dsetType, dsetGroup), .SDcols=colnames(bestTbl)],
 	#			       bestTbl[runNum==2, compare2DVs(.SD, c('topHashtagPostPriorStd', 'topHashtagAcrossPriorStd')), by=list(dsetType, dsetGroup), .SDcols=colnames(bestTbl)]
 	#			       ))
@@ -1814,7 +1818,8 @@ analyzePrior <- function(modelVsPredTbl) {
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorStd' & datasetName=='SOQgt500r2'])
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorOL2' & datasetName=='TFollowgt10Mr2' & d < 1])
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorOL2' & datasetName=='SOQgt500r2' & d < 1])
-	resTbl = compareOptimalDs(tbl[DVName %in% c(paste0('topHashtagPostPriorStd', c('')), paste0('topHashtagPostPriorOL2', c(''))) & runNum == 2])
+	resTbl = compareOptimalDs(tbl[DVName %in% c(paste0('topHashtagPostPriorStd', c('')), paste0('topHashtagPostPriorOL2', c(''))) & runNum == 2 & predUsedBest == T])
+	resTbl[, mean(meanVal), DVName]
[…]

### 2016-01-28 writeups/psycMethods/psycMethods.tex
foo

@@
 
+\newcommand{\myCI}[4]{{95 \% CI (\num{#1}, \num{#2}, \num{#3}, N=\num{#4})}}
+
 \RequirePackage[l2tabu, orthodox]{nag}
@@
 The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
-Using a predefined stoplist makes the smallest improvement (\numNoZero{.45} to \numNoZero{.49}), while the entropy (\numNoZero{.60}) and frequency (\numNoZero{.63}) techniques improve performance the most,
+Using a predefined stoplist makes the smallest improvement (\numNoZero{.45} to \numNoZero{.49}, \myCI{.032}{.018}{.047}{5})
+while the entropy (\numNoZero{.45} to \numNoZero{.60}, \myCI{.143}{.123}{.163}{5}) and frequency (\numNoZero{.45} to \numNoZero{.63}, \myCI{.178}{.164}{.192}{5}) techniques improve performance the most,
 and frequency shows a slight edge over entropy.
@@
 The three stop word techniques were then tested for the random permutation model on the Twitter dataset, and the results are depicted in Figure \ref{figContextStopRPT}.
-For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\numNoZero{.25} to \numNoZero{.26}).
-Removing stop-words based on a predetermined list actually decreases accuracy (\numNoZero{.25} to \numNoZero{.23}), and using data-driven frequency filtering has only a small to no effect (\numNoZero{.25} to \numNoZero{.25}).
+For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\numNoZero{.25} to \numNoZero{.26}, \myCI{.013}{.011}{.015}{60}).
+Removing stop-words based on a predetermined list actually decreases accuracy (\numNoZero{.25} to \numNoZero{.23}, \myCI{-.017}{-.015}{-.019}{60}),
+and using data-driven frequency filtering has only a small to no effect (\numNoZero{.25} to \numNoZero{.25}, \myCI{.00}{-.003}{.002}{60}).
 
@@
 
-The results show that entropy weighting produces the most accurate results (\numNoZero{.61} to \numNoZero{.63}).
-Also, accuracy actually decreases when using frequency filtering (compared to no filtering) (\numNoZero{.61} to \numNoZero{.56}) for the Bayesian model.
+The results show that entropy weighting produces the most accurate results (\numNoZero{.61} to \numNoZero{.63}, \myCI{.024}{.011}{.037}{5}).
+Also, accuracy actually decreases when using frequency filtering compared to no filtering (\numNoZero{.61} to \numNoZero{.56}, \myCI{-.043}{-.030}{-.057}{5}) for the Bayesian model.
 So entropy weighting is the clear winner in this case.
@@
 Similar results are found when the Bayesian model is tested on the Twitter dataset.
-Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer (\numNoZero{.29} to \numNoZero{.32}),
-and frequency filtering decreases performance relative to no filtering (\numNoZero{.29} to \numNoZero{.28}).
+Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer (\numNoZero{.29} to \numNoZero{.32}, \myCI{.028}{.024}{.031}{60}),
+and frequency filtering decreases performance relative to no filtering (\numNoZero{.29} to \numNoZero{.28}, \myCI{-.017}{-.015}{-.019}{60}).
 
@@
 Model accuracy results after using the log-odds transformation are depicted in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}.
-Using the log-odds transformation technique has a small effect on model accuracy for StackOverflow (\numNoZero{.45} to \numNoZero{.47}) and minimal to no effect on Twitter (\numNoZero{.25} to \numNoZero{.25}).
+Using the log-odds transformation technique has a small effect on model accuracy for StackOverflow (\numNoZero{.45} to \numNoZero{.47}, \myCI{.015}{.009}{.021}{5})
+and minimal to no effect on Twitter (\numNoZero{.25} to \numNoZero{.25}, \myCI{-.002}{-.005}{.001}{60}).
 
@@
 
-There may be a slight edge in increased accuracy when a log-odds transformation is used for the StackOverflow dataset.
-However, that result is not statistically reliable for the sample size used in this analysis (confidence intervals overlap in Figure \ref{figContextLogoddsSO}),
-and even if present the overall change in accuracy is small (\numNoZero{.45} to \numNoZero{.47}).
+There is a slight edge in increased accuracy when a log-odds transformation is used for the StackOverflow dataset, although the magnitude of the improvement in accuracy is small (\numNoZero{.45} to \numNoZero{.47}).
 Nonetheless, it does seem appropriate to convert the context to a log-odds adjustment value when context is added as an additional term to the random permutation model.

### 2016-01-28 dissProject/preProcess.R

@@ -1829,10 +1829,11 @@ analyzePrior <- function(modelVsPredTbl) {
 	tbl2[d == .8, DVName := paste0(DVName, 'BestDStd')]
 	DVNames = c('topHashtagPostPriorStdDefaultD', 'topHashtagPostPriorOL2BestDOL', 'topHashtagPostPriorStdBestDStd')
 	resTbl = compareOptimalAcc(tbl2[DVName %in% DVNames & runNum == 2])
+	resTbl
 	resTbl[, dsetType := 'twitter']
 	resTbl
 	resTbl[grepl('^SO', dsetGroup), dsetType := 'stackoverflow']
-	resTbl[, mean(meanVal), by=list(DVName, dsetType)]
+	resTbl[, mean(meanVal), by=list(DVName)]
 	resTbl
 	tbl
 }
@@ -1968,6 +1969,20 @@ analyzeFB <- function() {
 	compareMeanDVDefault(tbl[dsetType == 'stackoverflow' & DVName %in% DVNames], acc, figName='foo')
 }
 
+getWithinCI <- function(cTbl) {
+	dvTbl = cTbl[, .N, DVName][, data.table(t(combn(DVName, 2, simplify=T)))]
+	dvTbl
+	wTbl = cTbl[, dcast.data.table(.SD, datasetName + runNum + groupNum ~ DVName, value.var='acc')]
+	dvTbl[, rowID := 1:.N]
+	wTbl[, runNum]
+	dvTbl[, name := sprintf('%s - %s', V1, V2)]
+	dvTbl
+	dvTbl[, .(wTbl[, V1, with=F][[1]]), rowID]
+	dvTbl[, .(delta=wTbl[, V1, with=F][[1]] - wTbl[, V2, with=F][[1]]), name
+	      ][, as.list(c(CI(delta), N=.N)), name
+	      ]
+}
+
 analyzeContext <- function(modelVsPredTbl) {
 	#modelHashtagsTbls = Filter(x = modelHashtagsTbls, f = function(tbl) !grepl('-500', tbl[, datasetName[1]]))
 	#ppvTbl = rbindlist(lapply(modelHashtagsTbls, getPPVTblAll))
@@ -2017,22 +2032,50 @@ analyzeContext <- function(modelVsPredTbl) {
 				       'PriorStdTitleOrderlessFreqBodyOrderlessFreq',
 				       'PriorStdTitleOrderlessFrentropyBodyOrderlessFrentropy'
 				       ))
[…]

### 2016-01-31 writeups/psycMethods/psycMethods.tex
foo

@@
 
-\newcommand{\myCI}[4]{{95 \% CI (\num{#1}, \num{#2}, \num{#3}, N=\num{#4})}}
+\newcommand{\myCI}[4]{{mean difference = \num{#1}, 95\% CI [\num{#2}, \num{#3}]}}
 
@@
 
-The entire StackOverflow dataset is packaged and released quarterly to the public.
-So the most recent packaged dataset was downloaded and the tables were imported into a relational database.
+The entire StackOverflow dataset is packaged and released quarterly to the public,
+so the most recent packaged dataset was downloaded and the tables were imported into a relational database.
 
@@
 Confidence intervals were used throughout for statistical inference instead of explicit p-values.
-This was done because it is the most straightforward way to visualize all of the mean differences in accuracy (i.e., bar and line plots with a visual indication of the error) for each of the models, which eases interpretation
+This was done because it is a more straightforward way to visualize all of the mean differences in accuracy (i.e., bar and line plots with a visual indication of the error) for each of the models, which eases interpretation
 (similar to the argument made by FIXME:textcite).
-It is also a conservative estimate of reliability, as using confidence intervals on the means essentially ignores the fact that this is a within-subjects design, where accuracy for each model is measured for each subject.
+It is also a conservative estimate of reliability, as using confidence intervals on the means essentially ignores the fact that this is a within-subjects design, where accuracy for each model is measured for each subject in each dataset.
 However, we would rather err on the side of making the results easier to interpret at the cost of using conservative estimates for reliability
-(i.e., use error bars overlaid on a visualization of the means instead of statistical tests on the difference scores).
+(i.e., use error bars overlaid on a visualization of the means instead of statistical tests on every combination of difference scores).
 In fact, using large datasets for analysis is what empowers us to do this:
 With this much data, we can use straightforward (although more conservative) inferentials and still see reliable results.
-This allows us to concentrate on proper data visualization of the main findings.
+This allows us to concentrate more on proper data visualization of the main findings.
+
+Nonetheless, when accuracy differences between two models are called out explicitly in the text, the confidence interval for the mean difference in accuracy is also included.
+Since this interval is drawn across the observed difference scores, it will be smaller than the intervals included in the plots, since these are drawn independently for each model's overall accuracy.
 
@@
 Both forms maintain the same rank ordering, so they should produce similar results.
-So the $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).
 
@@
 Since decay rate values were explored at \num{0.1} increments between the 0 and 1 range, the decay rate values used to examine model accuracy were rounded to the nearest decay rate value that was included in the run.
-So a best-fit value of \num{.4} was used for the optimized learning model instead of the \num{.43} average found from the dataset slices.
+Consequently, a best-fit value of \num{.4} was used for the optimized learning model instead of the \num{.43} average found from the dataset slices.
 Aggregate accuracy was computed by taking the mean of every user's model accuracy across all users in all dataset slices.
@@
 
-So it may be worthwhile to use the standard form instead of the optimized learning form when declarative memory is modeled across a broader range of tasks.
+Given this, it may be worthwhile to use the standard form instead of the optimized learning form when declarative memory is modeled across a broader range of tasks.
 However, this should not be taken as a strong recommendation,
@@
 However, a custom user prior for the Twitter popular-hashtags dataset cannot be used, since the entire Twitter dataset cannot be downloaded, so the history of every user in the popular-hashtags dataset cannot be collected.
-So instead, the global prior tag history for all users in the Twitter popular-hashtags dataset was used to compute prior activation values for each tag.
+Instead, the global prior tag history for all users in the Twitter popular-hashtags dataset was used to compute prior activation values for each tag.
 
@@
 where both the number of observations for the word and the tag are normalized when computing the value for each cell (see the $S_{ji}$ computation in Table \ref{tabModACTRModel}).
-So it is likely the case that a form of frequency attenuation is already built into the Bayesian model.
-Therefore, adding a frequency filter on top of this does nothing to improve accuracy and only removes predictive information.
+Therefore, it is likely the case that a form of frequency attenuation is already built into the Bayesian model,
+and adding a frequency filter on top of this does nothing to improve accuracy and only removes predictive information.
 
@@
 
-So although incorporating word order into the random permutation model does seem to slightly improve performance,
+Although incorporating word order into the random permutation model does seem to slightly improve performance,
 the strength of the model is more likely to be in how efficiently it can represent the co-occurrence matrix rather than its ability to include word order in the representation.
@@
 
+\subsection{Implications and Future Directions}
+
 \subsubsection{Representation}
@@
 Vector-based models are simple enough that they can easily be implemented in a neural network.
+w
 Perhaps a form of vector-based compressed storage is being used at the neural level,
@@
 where the majority of the observations are taken from a generic domain (e.g., american literature) and a minority are domain-specific (e.g., StackOverflow word, tag pairs).
+
+\subsubsection{Individual vs. Group Representation of Memory}
+
+Even with the large datasets used in this research, there was not enough data on each individual user to build up separate co-occurrence matrices for them.
+That is, the context component of the declarative memory models used in this resaerch was formed as a collection of observed co-occurrences across all individuals.
+Nonetheless, the prior component for the models was customized for each individual (except in the Twitter popular hashtags dataset when it was not possible).
+Ideally, both components would be customized to the individual, and it remains an open question how much accuracy is influenced by using an aggregate or custom context component in the models.
+One practical approach that might be considered for future research:
+Use a large corpus to stand up an initial aggregate co-occurrence matrix (e.g., the Wikipedia corpus),
+then overlay the specific co-occurrences observed for each user on top of this base to create a customized matrix for each user,
+and then test model accuracy with and without the custom overlay.
 
@@
 
-\subsection{Implications and Future Directions}
-
-\subsubsection{Generalizing Results}
-
-It seems likely that the entropy method for computing the attentional weight term $W_{j}$ may be applicable to other declarative memory tasks.
-The entropy method is parameter-free (unlike the frequency-filtering method), completely data driven (much like the co-occurrence matrix), does not require tuning,
-and substantially increases model accuracy for these datasets.
-It is common to handle stop words in some manner when working with Natural Language Processing (NLP) and large text corpuses.
-So, if these large corpuses are used to test declarative memory,
-then the entropy method may be a more cognitively plausible way of dealing with stop words rather than filtering based on a predetermined list or the frequency count in the dataset.
+However, there may be other ways besides adjusting the attentional weight that these models can be modified to achieve the same performance improvement.
+The entropy weighting term is likely solving a problem that resides only in contextual activation and not prior activation,
+since the activation for the prior component for each word is independent of the presentation of other words, so stop words have no way to interact with tag prior activation in this case.
+
+One alternative explanation is that the entropy weighting measure was needed because the contextual co-occurrence matrix for this research was not yet stable, even with these large dataset sizes.
+If the matrix was not yet stable, then by chance certain stop words would contribute more net activation to specific tags and introduce noise into the retrieval process for that tag.
+So further weighting the contribution of these stop words by leveraging the attentional weight term ensured that this noise was not introduced into the system.
+If this is the case, it would be intersting to see how large a dataset is required to not need to modify the attentional weight term.
+However even if this is the case, the random permutation model would not be fixed by larger dataset sizes,
+since this model simply adds environment vectors to produce the current context vector and has no mechanism (unlike Bayesian normalization) for weighting the contribution of an environment vector.
+
+Another explanation for the Bayesian model is that additional attential weighting was needed because the contextual matrix was derived only from word by tag co-occurrences and not (word or tag) by (word or tag) co-occurences.
+This was done for performance reasons as the computational complexity otherwise is simply too high for current hardware.
+It is possible that by adding all pairwise combinations of both words and tags in the posts may soften and stabilize the contribution of stop words when computing contextual activation for a tag, given current context.
+
+\subsubsection{Prior Component}
 
@@
 
+\subsection{Support Code}
+
+All software code created for this research is publicly available on GitHub \parencite{StanleyRepo2014}.
+
 \clearpage

### 2016-01-31 writeups/psycMethods/psycMethods.tex
foo

@@
 the modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel}.
-The co-occurrence matrices used for the context component of each model were generated from a large subset of the Twitter and StackOverflow datasets.
-Each model was tested by evaluating the accuracy of model-chosen tags on a fresh test subset of the datasets.
-Accuracy was assessed by comparing model performance with the human data.
 
@@
 
+\subsection{Implications and Future Directions}
+
+\subsubsection{Tagging as a Declarative Memory Retrieval Process}
+
 Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter reasonably well.
 The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
+
 These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
-However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
+However, both models were not 100 percent accurate, and the amount of variance left to predict was is larger than what would be observed if retrieval noise is included in the model.
 One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval
@@
 
-\subsection{Implications and Future Directions}
+Another possibility is that this is not a declarative memory process at all.
+In particular, it is important to note that tagging on these sites is widely used as a method to communicate specific messages to specific groups of people in the community.
+It is unclear how much this environment overlaps with verbal communication between two people, and the declarative memory requests required to successfully communicate in this setting.
+Nonetheless, these models are reasonably accurate in this setting, particularly for StackOverflow.
+Given the results in this research, it at least seems worthwhile to use these publicly-available tagging environments as a resonable proxy and testbed to explore and validate modifications to the current theory.
+Any insights gained could be followed up with a more precisely controlled experimental design.
 
@@
 Vector-based models are simple enough that they can easily be implemented in a neural network.
-w
 Perhaps a form of vector-based compressed storage is being used at the neural level,
@@
 
+\subsubsection{Observational vs. Experimental Design}
+
+All data used for this study were collected from publicly-available datasets and were not subject to an experimental design (i.e., observational data).
+There are advantages and disadvantes to both observational and experimental designs.
+One primary advantage with the observational approach here is size and availability: data for millions of posts and tweets are immediately available for analysis.
+However one disadvantage is that there is no way to probe a user at a particular point in time, which could be quite advantageous if possible.
+One hybrid approach worth considering for future research is to use the publicly available datasets to build up a representation of declarative memory for each user.
+Afterwards, bring a sample of these specific users into a laboratory setting and probe them with specific retrieval requests.
+Multiple users could be asked to tag the same message, and then models could be tested to see if they reliably predict differences in accuracy between the subjects.
+
+\subsubsection{Which Model is Better?}
+
+After including the prior and entropy components for both models, overall accuracy was similar for the Bayesian and random permutation models for this task.
+So it is unclear with this research precisely which model is more cognitively plausible.
+However, there were two areas in particular where model performance showed relatively large differences: accuracy change as a function of corpus size and reliance on entropy weighting.
+These areas could be used in future research to more directly challenge the models.
+For example, a domain could be picked where enough data is collected to build a custom co-occurrence matrix for each user.
+Then one could look at model accuracy over time for each user as the co-occurrence matrix is increased to see if the trajectory is more accurately modeled by a random permutation (more accurate with lower N) or Bayesian (more accurate with higher N) model.
+Further, if a large-scale domain is picked and it is shown that the Bayesian model does not require entropy weighting for the co-occurrence matrix after the representations are fully stabilized,
+then this would provide a case where the two models are dramatically different: the random permutation model would require explicitly weighting the stop words (entropy weighting) while the Bayesian model would not.
+
 \subsection{Summary}

### 2016-01-31 writeups/psycMethods/psycMethods.tex
foo

@@
 This study expands on previous work that has tested retrieval models on large-scale datasets.
-It focuses specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and vector-based) can be improved by testing these models on two real-world tagging tasks.
+It focuses specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutation vector-based) can be improved by testing these models on two real-world tagging tasks.
 One primary objective is to compare the random permutation vector-based model to ACT-R's Bayesian declarative memory theory when large datasets are used for each model. 
@@
 The declarative memory system is a component of the architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
+At the core of the system is the concept of a ``chunk'': the representation of a piece of knowledge (e.g., a word).
 The equations that make up this system were derived from a rational analysis of declarative memory retrieval \parencite{Anderson1989}.
@@
 For vector-based systems, each word is represented by an environment vector $e_{i}$ that is nearly orthogonal to all other words' environment vectors.
-The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix, which is how vector-based systems represent information in a much more compact space.
+The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix (typically around 2,000 - 4,000), which is how vector-based systems represent information in a much more compact space.
 Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occurred with that $e_{i}$ environment vector.
@@
 The initial model results with the new tokenizer were qualitatively similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
-So the same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter, since it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.
+So the same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter to make comparisons across datasets more equitable.
 
@@
 
-\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing ``ed'' and ``ing'' endings), but
-with the StackOverflow dataset, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
+\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing ``ed'' and ``ing'' endings).
+With the StackOverflow dataset however, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
 This can be thought of as semantic stemming, where two words with identical meaning are converted to the same canonical word.
@@
 Both forms maintain the same rank ordering, so they should produce similar results.
+The $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).
 
@@
 It may be the case that the optimized learning form of the equation is no longer computationally necessary.
-And if the standard form is used as the default (or simply tried in another experiment),
+That said, one may still run up against performance limitations when trying to run large-scale models in real-time settings.
+Nonetheless, if the standard form is used as the default (or simply tried in another experiment),
 the results from this study suggest that the optimal decay rate value should be set slightly higher (\num{0.8}) than the default when optimized learning is used (\num{0.5}).
@@
 The cutoff that produced the frequency plots in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
-This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset was not given.
 
@@
 When the model components are enumerated, then at least one of the components is left out.
-So, for example, the ``standard prior model relaxed across posts'' means that only the prior model component was used when determining activation. 
+For example, the ``standard prior model relaxed across posts'' means that only the prior model component was used when determining activation. 
 As another example, ``RP only title with freq'' means that only the title context component of the random permutation model was used, and that stop words were handled using the frequency filtering technique.
@@
     ``Stoplist'', ``freq'', ``entropy'' are the stoplist removal method, removal based on frequency count, and weighting based entropy metric.
+    All error bars are the 95\% bootstrapped confidence interval for the mean. 
 }
@@
   The five methods for handling stop words are shown both for the context model component (i.e., no prior term) and the full model.
+  All error bars are the 95\% bootstrapped confidence interval for the mean. 
 }
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsSOSji-crop.pdf}}}
-  \caption{Stop-word techniques for the Bayesian model for StackOverflow}
+  \caption{
+  Stop-word techniques for the Bayesian model for StackOverflow
+  All error bars are the 95\% bootstrapped confidence interval for the mean. 
+}
   \label{figContextStopBSO}
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTSji-crop.pdf}}}
-  \caption{Stop-word techniques for the Bayesian model for Twitter}
+  \caption{
+  Stop-word techniques for the Bayesian model for Twitter
+  All error bars are the 95\% bootstrapped confidence interval for the mean. 
+}
   \label{figContextStopBT}
@@
 
-The number of documents used to build the co-occurrence matrix has been shown to influence model accuracy \parencite{Sahlgren2008}, so this was the second architectural concern that was examined.
+The number of documents used to build the co-occurrence matrix has been shown to influence model accuracy \parencite{Sahlgren2008}.
 The default number of documents used to build the co-occurrence matrix was \num{1000000} for StackOverflow and \num{3000000} for Twitter.
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
-  \caption{Random permutation model accuracy when using log-odds transformation on context for StackOverflow}
+  \caption{
+    Random permutation model accuracy when using log-odds transformation on context for StackOverflow
+    All error bars are the 95\% bootstrapped confidence interval for the mean. 
+  }
   \label{figContextLogoddsSO}
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
-  \caption{Random permutation model accuracy when using log-odds transformation on context for Twitter}
+  \caption{
+  Random permutation model accuracy when using log-odds transformation on context for Twitter
+  All error bars are the 95\% bootstrapped confidence interval for the mean. 
+}
   \label{figContextLogoddsT}
@@
   \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-orderT-crop.pdf}}}
-  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
+  \caption{
+    Model accuracy for various word order methods for the random permutation model for Twitter
+    All error bars are the 95\% bootstrapped confidence interval for the mean. 
+  }
   \label{figContextOrder}
@@
 The general discussion is separated into two parts.
-The first section will describe the overall results, reemphasize key findings, and discuss the theoretical implications of each key finding.
-The final section will describe future directions for the research.
+The first section will describe the overall results,
+and the final section will go through each key finding and discuss theoretical implications and future directions for research.
 
@@
 These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
-However, both models were not 100 percent accurate, and the amount of variance left to predict was is larger than what would be observed if retrieval noise is included in the model.
+However, both models are not 100 percent accurate, and the amount of variance left to predict is larger than what would be observed if retrieval noise is included in the model.
 One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval
@@
 Ideally, both components would be customized to the individual, and it remains an open question how much accuracy is influenced by using an aggregate or custom context component in the models.
-One practical approach that might be considered for future research:
-Use a large corpus to stand up an initial aggregate co-occurrence matrix (e.g., the Wikipedia corpus),
-then overlay the specific co-occurrences observed for each user on top of this base to create a customized matrix for each user,
-and then test model accuracy with and without the custom overlay.
 
@@
 
-One alternative explanation is that the entropy weighting measure was needed because the contextual co-occurrence matrix for this research was not yet stable, even with these large dataset sizes.
+One alternative explanation is that the entropy weighting measure was needed for this research because the contextual co-occurrence matrix was not yet stable, even with these large dataset sizes.
 If the matrix was not yet stable, then by chance certain stop words would contribute more net activation to specific tags and introduce noise into the retrieval process for that tag.
-So further weighting the contribution of these stop words by leveraging the attentional weight term ensured that this noise was not introduced into the system.
+So further weighting the contribution of these stop words by leveraging the attentional weight term ensured that this noise was suppressed in the system.
 If this is the case, it would be intersting to see how large a dataset is required to not need to modify the attentional weight term.
@@
 It is possible that by adding all pairwise combinations of both words and tags in the posts may soften and stabilize the contribution of stop words when computing contextual activation for a tag, given current context.
+
+Nonetheless, it seems likely that the entropy method for computing the attentional weight term $W_{j}$ may be applicable to other declarative memory tasks.
+The entropy method is parameter-free (unlike the frequency-filtering method), completely data driven (much like the co-occurrence matrix), does not require tuning,
+and substantially increases model accuracy for these datasets.
+It is common to handle stop words in some manner when working with Natural Language Processing (NLP) and large text corpuses.
+So, if these large corpuses are used to test declarative memory,
+then the entropy method may be a more cognitively plausible way of dealing with stop words rather than filtering based on a predetermined list or the frequency count in the dataset.
 
@@
 
-\subsubsection{Which Model is Better?}
+\subsubsection{Directly Comparing the Models}
 
 After including the prior and entropy components for both models, overall accuracy was similar for the Bayesian and random permutation models for this task.
-So it is unclear with this research precisely which model is more cognitively plausible.
+So it is unclear with this research precisely which model is more cognitively plausible (or if both are correct and simply working at different levels of abstraction).
 However, there were two areas in particular where model performance showed relatively large differences: accuracy change as a function of corpus size and reliance on entropy weighting.
@@
 Further, if a large-scale domain is picked and it is shown that the Bayesian model does not require entropy weighting for the co-occurrence matrix after the representations are fully stabilized,
-then this would provide a case where the two models are dramatically different: the random permutation model would require explicitly weighting the stop words (entropy weighting) while the Bayesian model would not.
+then this would provide a case where the two models are dramatically different: The random permutation model would require explicitly weighting the stop words (entropy weighting) while the Bayesian model would not.
 

### 2016-02-14 dissProject/preProcess.R
foo

@@ -29,6 +29,8 @@ library(digest)
 library(parallel)
 library(heplots)
 
+wideScreen(350)
+
 PATH = getPathToThisFile()
 FILE = getNameOfThisFile()
 
@@ -1840,6 +1842,13 @@ analyzePrior <- function(modelVsPredTbl) {
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorStd' & datasetName=='SOQgt500r2'])
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorOL2' & datasetName=='TFollowgt10Mr2' & d < 1])
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorOL2' & datasetName=='SOQgt500r2' & d < 1])
+	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorHybrid1' & datasetName=='SOQgt500r2' & d < 1])
+	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorHybrid5' & datasetName=='SOQgt500r2' & d < 1])
+	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorHybrid10' & datasetName=='SOQgt500r2' & d < 1])
+	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorHybrid1' & datasetName=='TFollowgt10Mr2' & d < 1])
+	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorHybrid5' & datasetName=='TFollowgt10Mr2' & d < 1])
+	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorHybrid10' & datasetName=='TFollowgt10Mr2' & d < 1])
+	modelVsPredTbl[, .N, DVName]
 	resTbl = compareOptimalDs(tbl[DVName %in% c(paste0('topHashtagPostPriorStd', c('')), paste0('topHashtagPostPriorOL2', c(''))) & runNum == 2 & predUsedBest == T])
 	resTbl[, mean(meanVal), DVName]
 	#tbl[d == 0.5 & !maxNP, DVName := paste0(DVName, 'DefaultD')]

### 2016-02-14 writeups/psycMethods/psycMethods.tex
foo

@@
       Optimized Learning &	$B_{i} = log \frac{n}{1-d} - d * log L$ \\
+      Hybrid &			$B_{i} = log \big[ \sum_{j=1}^{k} {t_{j}}^{-d} + \frac{(n-k)(t_{n}^{1-d}-t_{k}^{1-d})}{(1-d)(t_{n}-t_{k})} \big]$ \\
       \hline
@@
 
-Both terms are methods for computing the base-level activation ($B_{i}$) of a chunk in declarative memory.
+All three terms are methods for computing the base-level activation ($B_{i}$) of a chunk in declarative memory.
 The standard form has a time dependence and accounts for the fact that the rate of presentations for a given object (e.g., the word ``loop'') can change over time.
 The optimized learning form assumes that this rate is constant.
-
-For both equations, $i$ represents each chunk (e.g., the word ``loop''), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``loop'' appears), and $n$ is the number of times each chunk has been observed.
+The hybrid form is a blend of the standard and optimized form where activation for the most recent $k$ observations are computed using the standard form and the rest use the optimized form.
+
+For all equations, $i$ represents each chunk (e.g., the word ``loop''), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``loop'' appears), and $n$ is the number of times each chunk has been observed.
 Since the optimized learning form assumes a constant presentation rate for each chunk,
@@
 It does not have to keep track of the time each specific chunk instance was observed.
-Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R Common Lisp implementation.
+Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form.
+The hybrid form only has to keep track of the most recent $k$ presentations of each chunk, so it only pays a small computational penalty over the optimized learning form.
+The optimized learning form is set to the default in the ACT-R Common Lisp implementation.
 
@@
 The values were chosen to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d \approx .5$), to a pure recency model ($d=20$).
-The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
-Only values below 1 were examined for the optimized learning form, since values above 1 for this form are not mathematically defined. 
-They are not mathematically defined at a decay rate greater than 1 because for these values the first term to compute $B_{i}$ is the logarithm of a negative number (see Table \ref{tabACTRBLLModel}).
+The decay rate value was varied for the standard form, the optimized learning form, and the hybrid form of $B_{i}$. 
+Only values below 1 were examined for the optimized and hybrid forms, since values above 1 for these forms are not mathematically defined. 
+They are not mathematically defined at a decay rate greater than 1 because for these values the $(1-d)$ denominator flips the ratios to negative values (see Table \ref{tabACTRBLLModel}).
 At a decay rate equal to 1, every base-level activation $B_{i}$ is infinite, which is not psychologically interesting.
 
-Both the standard form and optimized learning form of $B_{i}$ were used to predict each user's specific tag history.
+All three forms of $B_{i}$ were used to predict each user's specific tag history.
 To do this, a user's entire tag history was collected, and then each tag instance from oldest to newest was examined.
@@
 This process was performed across all users in each of the 12 dataset slices for both the Twitter and StackOverflow popular-users datasets.
-It was repeated for each of the decay rate values and the two different forms of $B_{i}$.
+It was repeated for each of the decay rate values and the three different forms of $B_{i}$.
+For the hybrid form, three different values of k were explored (1, 5, 10).
 
@@
 
-A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDsStd} and \ref{figPriorTwitterSliceDsStd} respectively.
+A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for all three forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDsStd} and \ref{figPriorTwitterSliceDsStd} respectively.
 
@@
     \hfill
-    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
+    \subfigure[][Standard form]{\includegraphics[width=5cm]{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
     \hfill
-    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
+    \subfigure[][Optimized learning form]{\includegraphics[width=5cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
+    \hfill
+    \subfigure[][Hybrid form]{\includegraphics[width=5cm]{visNormMean-SOQgt500r2-topHashtagPostPriorHybrid10-crop.pdf}}
     \caption{
@@
     This includes model accuracy as a function of decay rate value for each user in the dataset slice.
-    Model accuracy is shown both when the standard from of base-level activation ($B_{i}$) and the optimized learning form of the equation are used. 
+    Model accuracy is shown for the standard form of base-level activation ($B_{i}$), the optimized learning form, and the hybrid form when k=10.
   }
@@
     \hfill
-    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}
+    \subfigure[][Standard form]{\includegraphics[width=5cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}
     \hfill
-    \subfigure[][Optmized learning form]{\includegraphics[width=8cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
+    \subfigure[][Optmized learning form]{\includegraphics[width=5cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
+    \hfill
+    \subfigure[][Hybrid form]{\includegraphics[width=5cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorHybrid10-crop.pdf}}
     \caption{
@@
 And the relative accuracy gained from using a pure frequency based model to a blend of frequency and recency is less for the optimized learning form than the standard form of $B_{i}$.
+Finally, for the range defined for the hybrid form ($d < 1$), the standard and hybrid forms are shaped similarly across the decay rate values.
 
@@
 Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
-Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{.43}) of $B_{i}$ compared to the standard form (\num{.80}).
+Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{.43}) of $B_{i}$ compared to both the the standard form (\num{.80}) and hybrid form (\num{.80} for all $k$).
 The optimal values for optimized learning are near \num{0.5}, which lines up with default value used in ACT-R.
@@
 This was done by examining model accuracy at a specific decay rate value for each model.
-Accuracy was assessed at each model's previously-determined best-fit values (\num{.43} and \num{.80} for the optimized learning and standard form) and the standard form's default value (\num{0.5}).
+Accuracy was assessed at each model's previously-determined best-fit values (\num{.43} and \num{.80} and \num{.80} for the optimized learning, standard form, and hybrid form) and additionally the standard form's default value (\num{0.5}).
 Since decay rate values were explored at \num{0.1} increments between the 0 and 1 range, the decay rate values used to examine model accuracy were rounded to the nearest decay rate value that was included in the run.
@@
 Accuracy for both sites is included in Figure \ref{figPriorAcc}. 
-Accuracy for the standard form (\numNoZero{.33}) is higher than the optimized learning form (\numNoZero{.28}).
+Accuracy for the standard form is higher than the optimized learning form (\numNoZero{.33} vs. \numNoZero{.28}, \myCI{.034}{.032}{.036}{3603}).
 This makes sense given that the standard form does not assume equal presentation rate for each chunk, and hashtags on Twitter in particular show trends of high rate of use and then shift to low rate of use.
@@
 
+Accuracy for the hybrid form for the largest $k$ ($k=10$) is slightly higher than the smallest $k$ ($k=1$) (\numNoZero{.334} vs. \numNoZero{.329} \myCI{.0026}{.0023}{.0029}{3608}).
+However accuracy cannot be distinguished between the hybrid form with the largest $k$ and the standard form (\numNoZero{.334} vs. \numNoZer{.334} \myCI{.0000}{-.0000}{.0001}{3609}).
+This indicates that for the datasets analyzed in this study that the hybrid form with $k=10$ is a suitable replacement to the standard form (produces the same results) and uses less computational resources than the standard form.
+
 ACT-R sets the default decay rate value to \num{0.5}, and the best-fit value for the standard form is higher (\num{0.8}) for these datasets.
-The figure shows that although model accuracy does slightly improve (\num{0.325} to \num{0.335}) for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
+The figure shows that although model accuracy does slightly improve (\num{0.325} to \num{0.335} \myCi{.0029}{.0022}{.0036}{3609}) for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
 that percentage improvement is not as large (\num{0.28} to \num{0.33}) as the improvement gained from using the standard form compared to the optimized learning form.
@@
 
-Given this, it may be worthwhile to use the standard form instead of the optimized learning form when declarative memory is modeled across a broader range of tasks.
-However, this should not be taken as a strong recommendation,
-since the performance benefits seen in this study could simply be due to the fact that the declarative memory retrievals span across a large period of time, which is not the case for all tasks.
-Nonetheless, the only real reason not to use the standard form is due to computational resource limitations.
-Since this is becoming less of an issue as hardware speed increases, it may be time to reconsider the default form of base-level learning used in ACT-R.
-It may be the case that the optimized learning form of the equation is no longer computationally necessary.
-That said, one may still run up against performance limitations when trying to run large-scale models in real-time settings.
-Nonetheless, if the standard form is used as the default (or simply tried in another experiment),
-the results from this study suggest that the optimal decay rate value should be set slightly higher (\num{0.8}) than the default when optimized learning is used (\num{0.5}).
+However the hybrid form with a reasonable $k$ is just as accurate as the standard form.
+This is due to the fact that the majority of change in activation for a chunk occurs in the short period of time after the presentation of the chunk (a property of a power law distribution).
+Once a chunk's presentations decay to the flat area of the power law distribution, not much change in activation for these presentations occurs, so they can be averaged and their presentation time can be discarded without much loss in fidelity.
+Therefore if computational speed is a concern, one may wish to leverage the hybrid form at instead of the standard form.
+Although qualitatively the standard form produces the cleanest and most natural effect of decay rate on accuracy, as shown in Figure \ref{figPriorSOQSliceDsStd}.
+It also does not have a bifurcation at $k=1$.
+So if one is a purest and computational speed is less of a concern, one may wish to use the standard form over the hybrid form.
 

### 2016-02-14 dissProject/preProcess.R

@@ -1645,9 +1645,15 @@ renameColDVName <- function(tbl) {
 	setkey(tbl, DVName)
 	mapping = c('topHashtagAcrossPriorStd', 'Standard prior model relaxed across posts',
 		    'topHashtagPostPriorStd', 'Standard prior model',
+		    'topHashtagPostPriorHybrid1', 'Hybrid model with k=01',
+		    'topHashtagPostPriorHybrid5', 'Hybrid model with k=05',
+		    'topHashtagPostPriorHybrid10', 'Hybrid model with k=10',
 		    'topHashtagPostPriorStdDefaultD', 'Standard prior model at default d=0.5',
 		    'topHashtagPostPriorStdBestDStd', 'Standard prior model at best-fit d=0.8',
 		    'topHashtagPostPriorOL2BestDOL', 'Optimized learning model at best-fit d=0.4',
+		    'topHashtagPostPriorHybrid1BestDStd', 'Hybrid model at best-fit d=0.8 with k=01',
+		    'topHashtagPostPriorHybrid5BestDStd', 'Hybrid model at best-fit d=0.8 with k=05',
+		    'topHashtagPostPriorHybrid10BestDStd', 'Hybrid model at best-fit d=0.8 with k=10',
 		    'topHashtagPostPriorOL2', 'Optimized learning model',
 		    'topHashtagAcrossPriorStdNoffset', 'Standard prior model relaxed across posts without offset',
 		    makeLogregMapping(),
@@ -1849,7 +1855,8 @@ analyzePrior <- function(modelVsPredTbl) {
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorHybrid5' & datasetName=='TFollowgt10Mr2' & d < 1])
 	visModelVsPredTbl(modelVsPredTbl[DVName=='topHashtagPostPriorHybrid10' & datasetName=='TFollowgt10Mr2' & d < 1])
 	modelVsPredTbl[, .N, DVName]
-	resTbl = compareOptimalDs(tbl[DVName %in% c(paste0('topHashtagPostPriorStd', c('')), paste0('topHashtagPostPriorOL2', c(''))) & runNum == 2 & predUsedBest == T])
[…]

### 2016-02-14 writeups/psycMethods/psycMethods.tex
foo

@@
 
-Confidence intervals were used throughout for statistical inference instead of explicit p-values.
-This was done because it is a more straightforward way to visualize all of the mean differences in accuracy (i.e., bar and line plots with a visual indication of the error) for each of the models, which eases interpretation
-(similar to the argument made by FIXME:textcite).
+In similar spirit to \textcites{masson2003using, tryon2001evaluating}, confidence intervals were used throughout for statistical inference instead of explicit p-values.
+This was done because it is a more straightforward way to visualize all of the mean differences in accuracy (i.e., bar and line plots with a visual indication of the error) for each of the models, which eases interpretation.
 It is also a conservative estimate of reliability, as using confidence intervals on the means essentially ignores the fact that this is a within-subjects design, where accuracy for each model is measured for each subject in each dataset.
@@
 Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
-Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{.43}) of $B_{i}$ compared to both the the standard form (\num{.80}) and hybrid form (\num{.80} for all $k$).
+Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{.43}) of $B_{i}$ compared to both the standard form (\num{.80}) and hybrid form (\num{.80} for all $k$).
 The optimal values for optimized learning are near \num{0.5}, which lines up with default value used in ACT-R.
@@
 Accuracy for the hybrid form for the largest $k$ ($k=10$) is slightly higher than the smallest $k$ ($k=1$) (\numNoZero{.334} vs. \numNoZero{.329} \myCI{.0026}{.0023}{.0029}{3608}).
-However accuracy cannot be distinguished between the hybrid form with the largest $k$ and the standard form (\numNoZero{.334} vs. \numNoZer{.334} \myCI{.0000}{-.0000}{.0001}{3609}).
+However accuracy cannot be distinguished between the hybrid form with the largest $k$ and the standard form (\numNoZero{.334} vs. \numNoZero{.334} \myCI{.0000}{-.0000}{.0001}{3609}).
 This indicates that for the datasets analyzed in this study that the hybrid form with $k=10$ is a suitable replacement to the standard form (produces the same results) and uses less computational resources than the standard form.
@@
 ACT-R sets the default decay rate value to \num{0.5}, and the best-fit value for the standard form is higher (\num{0.8}) for these datasets.
-The figure shows that although model accuracy does slightly improve (\num{0.325} to \num{0.335} \myCi{.0029}{.0022}{.0036}{3609}) for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
+The figure shows that although model accuracy does slightly improve (\num{0.325} to \num{0.335} \myCI{.0029}{.0022}{.0036}{3609}) for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
 that percentage improvement is not as large (\num{0.28} to \num{0.33}) as the improvement gained from using the standard form compared to the optimized learning form.
@@
 Nonetheless, these models are reasonably accurate in this setting, particularly for StackOverflow.
-Given the results in this research, it at least seems worthwhile to use these publicly-available tagging environments as a resonable proxy and testbed to explore and validate modifications to the current theory.
+Given the results in this research, it at least seems worthwhile to use these publicly-available tagging environments as a reasonable proxy to explore and validate modifications to the current theory.
 Any insights gained could be followed up with a more precisely controlled experimental design.
@@
 Even with the large datasets used in this research, there was not enough data on each individual user to build up separate co-occurrence matrices for them.
-That is, the context component of the declarative memory models used in this resaerch was formed as a collection of observed co-occurrences across all individuals.
+That is, the context component of the declarative memory models used in this research was formed as a collection of observed co-occurrences across all individuals.
 Nonetheless, the prior component for the models was customized for each individual (except in the Twitter popular hashtags dataset when it was not possible).
@@
 So further weighting the contribution of these stop words by leveraging the attentional weight term ensured that this noise was suppressed in the system.
-If this is the case, it would be intersting to see how large a dataset is required to not need to modify the attentional weight term.
+If this is the case, it would be interesting to see how large a dataset is required to not need to modify the attentional weight term.
 However even if this is the case, the random permutation model would not be fixed by larger dataset sizes,
@@
 
-Another explanation for the Bayesian model is that additional attential weighting was needed because the contextual matrix was derived only from word by tag co-occurrences and not (word or tag) by (word or tag) co-occurences.
+Another explanation for the Bayesian model is that additional attentional weighting was needed because the contextual matrix was derived only from word by tag co-occurrences and not (word or tag) by (word or tag) co-occurrences.
 This was done for performance reasons as the computational complexity otherwise is simply too high for current hardware.
@@
 All data used for this study were collected from publicly-available datasets and were not subject to an experimental design (i.e., observational data).
-There are advantages and disadvantes to both observational and experimental designs.
+There are advantages and disadvantages to both observational and experimental designs.
 One primary advantage with the observational approach here is size and availability: data for millions of posts and tweets are immediately available for analysis.

### 2016-02-14 writeups/psycMethods/psycMethods.tex
foo

@@
 The initial model results with the new tokenizer were qualitatively similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
-So the same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter to make comparisons across datasets more equitable.
+The same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter to make comparisons across datasets more equitable.
 
@@
 So tags used for the post were converted to their root tag when possible.
+This does make the decision process easier for the StackOverflow dataset.
+However, we chose to include this to help ensure that model predictions would stabilize.
+This provided the best opportunity to reliably observe differences in accuracy when altering the models.
 
@@
 These values are recorded, and tag instances that match what the author actually tagged are marked as a 1, while all others are marked with a 0.
+When an author used multiple tags for a post, multiple 1s are marked for that post.
 This process is repeated across all posts where a user chooses a tag, and across all users in the dataset.
@@
 then it may be worthwhile to experiment with using a compressed random permutation model instead of the full Bayesian model.
+Also newer models of $S_{ji}$ have been introduced that have a more local impact when a new chunk is observed (\parencite{thomson2013constraining}).
+It would also be interesting to see how these local models of $S_{ji}$ perform while varying the number of observed associations.
 
@@
 Both a method for stop-word removal (frequency filtering) and stop-word attenuating (entropy weighting) performed well for both models.
-Also, using a proper method for handling the low-predictor words produced some of the largest improvements in model accuracy compared to the other explorations
+
+It is unclear at this point if this modification is cognitively plausible, even if the modification was cleanly added to the current ACT-R model.
+What can be said from this research is that using a proper method for handling the low-predictor words produced some of the largest improvements in model accuracy compared to the other explorations
 (e.g., compared to word order or properly adding terms).
-Since the ACT-R declarative memory theory already has a component for attentional weight, it may be the case that the neural system performs some form of stop-word weighting when computing activations,
-and this is reflected in the attentional weight term.
-That is, $W_{j}$ can be defined mathematically, has the same general form across tasks, and both the frequency-filtering and entropy-weighting methods for stop-word removal are possible first candidates for the definition.
-
-However, there may be other ways besides adjusting the attentional weight that these models can be modified to achieve the same performance improvement.
+Consequently, it may be worthwhile to explore this area more thoroughly with further research.
+
+Also, there may be other ways besides adjusting the attentional weight that these models can be modified to achieve the same performance improvement.
 The entropy weighting term is likely solving a problem that resides only in contextual activation and not prior activation,
@@
 
-One alternative explanation is that the entropy weighting measure was needed for this research because the contextual co-occurrence matrix was not yet stable, even with these large dataset sizes.
+One alternative explanation is that the entropy weighting measure was needed because the contextual co-occurrence matrix was not yet stable, even with these large dataset sizes.
 If the matrix was not yet stable, then by chance certain stop words would contribute more net activation to specific tags and introduce noise into the retrieval process for that tag.
@@
 If this is the case, it would be interesting to see how large a dataset is required to not need to modify the attentional weight term.
-However even if this is the case, the random permutation model would not be fixed by larger dataset sizes,
+However even if this is the case, the random permutation model would still require entropy weighting for larger dataset sizes,
 since this model simply adds environment vectors to produce the current context vector and has no mechanism (unlike Bayesian normalization) for weighting the contribution of an environment vector.
@@
 
+Also perhaps a different $S_{ji}$ model may not need entropy weighing at all (e.g., the Hebbian-inspired associative learning model described in \textcite{thomson2013constraining}).
+Further research could explore how entropy weighting interacts with other models of $S_{ji}$, and not just the random permutation and full Bayesian model.
+
 Nonetheless, it seems likely that the entropy method for computing the attentional weight term $W_{j}$ may be applicable to other declarative memory tasks.
 The entropy method is parameter-free (unlike the frequency-filtering method), completely data driven (much like the co-occurrence matrix), does not require tuning,
-and substantially increases model accuracy for these datasets.
+and increases model accuracy for these datasets.
 It is common to handle stop words in some manner when working with Natural Language Processing (NLP) and large text corpuses.

### 2016-02-15 writeups/psycMethods/psycMethods.tex
foo

@@
       Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
-      Constant Base Level &	$B_{i} = \frac{p_{i}}{1-p_{i}}$ \\
       Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
@@
 Usually both resonance and decoding will return similar rank orderings of most likely chunks given context.
-\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
 
@@
 
-The fan effect describes a behavior of human memory where memory items are more difficult to retrieve if they have occurred in a larger number of different contexts.
-\textcite{Rutledge2008} used \textcite{Anderson1974}'s classic experimental paradigm to test the fan effect.
-In the task, participants are exposed to a set of person and place pairs (e.g., ``the hippie is in the park'').
-They are later asked to recall if a particular person and place pair was presented to them before.
-The main manipulation is the ``fan'' of the person and place for each pair.
-The fan is defined as the number of different contexts that each item in the pair is presented in.
-For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store,'' then ``hippie'' would have a fan of two.
-Recall accuracy decrease and time to recall increase as the sum total fan of the person and place increases.
-
-Both the ACT-R declarative memory theory and vector-based models produce the fan effect due to the effect that fan has on the co-occurrence matrix.
-As a particular word (e.g., ``hippie'') is seen in more and more contexts, the co-occurrence count of the word across all contexts is spread across multiple contexts.
-This means that when a retrieval for a context given the word occurs, multiple contexts have relatively high activation.
-This makes it more difficult to test if a signal (e.g., if a word appears in a specific context) is higher than the background noise (e.g., all other contexts that the word has appeared in),
-which makes the retrieval more uncertain.
-
-Both vector-based and Bayesian models produce the fan effect since they both are based on co-occurrences.
-However, vector-based models do further processing on that matrix, reduce its dimensionality, and use a correlation instead of a log odds value to compute the activation.
-Nonetheless, vector-based models still produce the fan effect because at their core they are still computing activations based on co-occurrences.
-
 \subsubsection{Base-Level Activation}
@@
 It was repeated for each of the decay rate values and the three different forms of $B_{i}$.
-For the hybrid form, three different values of k were explored (1, 5, 10).
+For the hybrid form, three different values of $k$ were explored (1, 5, 10).
 
@@
 This was done by examining model accuracy at a specific decay rate value for each model.
-Accuracy was assessed at each model's previously-determined best-fit values (\num{.43} and \num{.80} and \num{.80} for the optimized learning, standard form, and hybrid form) and additionally the standard form's default value (\num{0.5}).
+Accuracy was assessed at each model's previously-determined best-fit values (\num{.43}, \num{.80}, and \num{.80} for the optimized learning, standard form, and hybrid form) and additionally the standard form's default value (\num{0.5}).
 Since decay rate values were explored at \num{0.1} increments between the 0 and 1 range, the decay rate values used to examine model accuracy were rounded to the nearest decay rate value that was included in the run.
@@
 However accuracy cannot be distinguished between the hybrid form with the largest $k$ and the standard form (\numNoZero{.334} vs. \numNoZero{.334} \myCI{.0000}{-.0000}{.0001}{3609}).
-This indicates that for the datasets analyzed in this study that the hybrid form with $k=10$ is a suitable replacement to the standard form (produces the same results) and uses less computational resources than the standard form.
+This indicates that for the datasets analyzed in this study the hybrid form with $k=10$ is a suitable replacement to the standard form (produces the same results) and uses less computational resources than the standard form.
 
@@
 Once a chunk's presentations decay to the flat area of the power law distribution, not much change in activation for these presentations occurs, so they can be averaged and their presentation time can be discarded without much loss in fidelity.
-Therefore if computational speed is a concern, one may wish to leverage the hybrid form at instead of the standard form.
+Therefore if computational speed is a concern, one may wish to leverage the hybrid form instead of the standard form.
 Although qualitatively the standard form produces the cleanest and most natural effect of decay rate on accuracy, as shown in Figure \ref{figPriorSOQSliceDsStd}.
 It also does not have a bifurcation at $k=1$.
-So if one is a purest and computational speed is less of a concern, one may wish to use the standard form over the hybrid form.
+So if one is a purest and computational speed is less of a concern, one may still wish to use the standard form over the hybrid form.
 
@@
 
-A simplified version of \textcite{Stanley2013}'s method was used for the results for this research.
+A simplified version of \textcite{Stanley2013}'s method was used for this research.
 Instead of subtracting the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
@@
 
-The transition point where accuracy for the Bayesian model outperformed the random permutation model was much higher for StackOverflow than Twitter.
-This is most likely because StackOverflow has almost \num{100} times more unique tags than the Twitter popular-hashtags dataset.
-The number of words in a StackOverflow post is also much higher than Twitter, which should have lowered the transition point.
-However, the effect of the number of words in a post must be less than the effect of number of unique tags, as the transition point was still much higher for StackOverflow given that it should have lowered due to post size.
-
 The compressed random permutation model should perform better with smaller corpus sizes, less words per document, more unique tags, and fewer number of tags per post.
@@
 then it may be worthwhile to experiment with using a compressed random permutation model instead of the full Bayesian model.
-Also newer models of $S_{ji}$ have been introduced that have a more local impact when a new chunk is observed (\parencite{thomson2013constraining}).
+Also newer models of $S_{ji}$ have been introduced that have a more local impact when a new chunk is observed \parencite{thomson2013constraining}.
 It would also be interesting to see how these local models of $S_{ji}$ perform while varying the number of observed associations.

### 2016-02-15 writeups/psycMethods/psycMethods.tex
foo

@@
 It also does not have a bifurcation at $k=1$.
-So if one is a purest and computational speed is less of a concern, one may still wish to use the standard form over the hybrid form.
+Consequently, if one is a purest and computational speed is less of a concern, one may still wish to use the standard form over the hybrid form.
 
@@
 Instead of subtracting the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
-So for the prior component, the activation value used in the logistic regression is the prior activation with the mean of the top 5 prior activations subtracted from this value.
+For the prior component for example, the activation value used in the logistic regression is the prior activation with the mean of the top 5 prior activations subtracted from this value.
 The same process is used for the context component.

### 2016-02-27 writeups/psycMethods/psycMethods.tex
foo

@@
 
+% Customer citation for a parencite style but with a few leading words in the parentheses
+% Stolen from parencite but with the out parentheses stripped off
+% For example: (e.g., Anderson, 1972), (see Salton, 2014)
+\DeclareCiteCommand{\noparencite}
+  {\usebibmacro{prenote}}
+  {\usebibmacro{citeindex}%
+   \usebibmacro{cite}}
+  {\multicitedelim}
+  {\usebibmacro{postnote}}
+
 % If the toc is too detailed, try limiting the depth of printed sections in the TOC:
@@
 Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
-For example, it is currently unexplored how a vector-based model can be modified to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
+For example, modifying a vector-based model to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
 Large-scale datasets were used throughout the model development and evaluation process.
@@
 
-As a general methodology for this research, we are using a cognitively-constrained exploratory model development process.
-Since we are working with large-scale datasets, we have enough data to explore many different modifications simultaneously without the risk of overfitting.
-However, we are not interested in bringing to bear the entire collection of machine learning algorithms to this prediction task in order to completely optimize a model.
-Rather, we are interested in evaluating how well two cognitive models can explain the results, and then making selective modifications to them to see how performance changes.
-The modifications that we explored were picked such that they influenced model accuracy, could be easily added to (or removed from) the cognitive models without massive modifications, and could be considered cognitively plausible.
+As a general methodology for this research, we used a cognitively-constrained exploratory model development process.
+Since we worked with large-scale datasets, we had enough data to explore many different modifications simultaneously without the risk of overfitting.
+However, we were not interested in bringing to bear the entire collection of machine learning algorithms to this prediction task in order to completely optimize a model.
+Rather, we were interested in evaluating how well two cognitive models could explain the results, and then making selective modifications to them to see how performance changes.
+The modifications that we explored were selected such that they influenced model accuracy, could be easily added to (or removed from) the cognitive models without massive modifications, and could be considered cognitively plausible.
 
@@
 
-The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and we are interested in retrieval models that generalize across tasks.
+The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and were interested in retrieval models that generalize across tasks.
 However, they are also similar on several accounts:
@@
 This rest of this document is separated into 5 sections.
-``Theoretical Background'' describes the vector-based and Bayesian declarative memory models that will be evaluated in this research,
+``Theoretical Background'' describes the vector-based and Bayesian declarative memory models that were evaluated in this research,
 shows how these models are related to the other common declarative memory models,
@@
 Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
-Both ACT-R's Bayesian declarative memory model and LSA compute activations based on a co-occurrence matrix.
-However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
-LSA uses the Singular Value Decomposition (SVD) process to reduce the dimensionality of the word by document co-occurrence matrix, while maintaining as much of the variance in the original full matrix as possible.
-The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.
-
 Both vector-based models and LSA calculate word similarity on a reduced matrix.
 For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors is much less than the number of words.
-For vector-based models, each tag is represented by a compressed vector with a length much less than the number of words.
-
-Information is lost with this compression in both cases: by starting with a compressed representation with vector-based models or by reducing to a compressed representation with LSA.
-However, the way that the compression works in these two cases is not the same.
-LSA uses SVD to find words that have highly correlated tag occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
+LSA uses Singular Value Decomposition (SVD) to find words that have highly correlated document occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
+
+For vector-based models, each tag (document) is represented by a compressed vector with a length much less than the number of words.
 A vector-based approach like random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
@@
 For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
-\citeauthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect.
+\citeauthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect \parencite{Anderson1974}.
 
@@
 
-Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the relative benefits of the two models. 
+Vector-based models and ACT-R's Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the relative benefits of the two models. 
 In particular, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
@@
 However, adding word-order information to ACT-R's default Bayesian declarative memory model is not common.
-Instead, when word order information is used to model reading for example, production rules are defined so that words are read from left to right (e.g., \textcite{Lewis2006}).
+Instead, when word order information is used to model reading for example, production rules are defined so that words are read from left to right (e.g., \noparencite{Lewis2006}).
 
@@
 
-Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
+Vector-based retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
 Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
@@
 (see the strength of association equation in Table \ref{tabACTRModel}).
-The high-frequency words are often removed from co-occurrence matrix for vector-based models to reduce this effect.
+The high-frequency words are often removed from the co-occurrence matrix for vector-based models to reduce this effect.
 
@@
 Also, the general methods used to tokenize and stem the words in all of the datasets will be described.
+All software code created for this research is publicly available on GitHub \parencite{StanleyRepo2014}.
 
@@
 
-A custom tokenization algorithm created by \textcite{Owoputi2013} was used to chunk the words in the text.
+A custom tokenization algorithm created by \textcite{Owoputi2013} was used to parse the words in the text.
 This algorithm was specifically created for tokenizing Twitter content.
@@
 
-In similar spirit to \textcites{masson2003using, tryon2001evaluating}, confidence intervals were used throughout for statistical inference instead of explicit p-values.
+In similar spirit to \textcites{masson2003using, tryon2001evaluating, cumming2005inference}, confidence intervals were used throughout for statistical inference instead of explicit $p$-values.
 This was done because it is a more straightforward way to visualize all of the mean differences in accuracy (i.e., bar and line plots with a visual indication of the error) for each of the models, which eases interpretation.
@@
     \hfill
-    \subfigure[][Optmized learning form]{\includegraphics[width=5cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
+    \subfigure[][Optimized learning form]{\includegraphics[width=5cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
     \hfill
@@
 This makes sense given that the standard form does not assume equal presentation rate for each chunk, and hashtags on Twitter in particular show trends of high rate of use and then shift to low rate of use.
-The standard form of $B_{i}$ is able to more accurately account for these trends.
+The standard form of $B_{i}$ is able to more accurately account for these dynamics.
 
@@
 It also does not have a bifurcation at $k=1$.
-Consequently, if one is a purest and computational speed is less of a concern, one may still wish to use the standard form over the hybrid form.
+Consequently, if one is a purist and computational speed is less of a concern, one may still wish to use the standard form over the hybrid form.
 
@@
 This section will examine how model accuracy is influenced by four architectural concerns:
-method of handling stop words, influence of dataset size, method to combine prior and context, and word order.
+method of handling stop words, influence of dataset size, method of combining prior and context, and word order.
 These particular architectural concerns were examined for the following reasons:
@@
 The same process is used for the context component.
-Decoupling the offset term and computing an offset for each model component in isolation means that an iterative approach is no longer necessary,
+Decoupling the offset term and computing an offset for each model component in isolation means that an iterative approach was no longer necessary,
 since the optimal weights of the terms are no longer needed when computing the offset.
@@
 
-However, a custom user prior for the Twitter popular-hashtags dataset cannot be used, since the entire Twitter dataset cannot be downloaded, so the history of every user in the popular-hashtags dataset cannot be collected.
+However, a custom user prior for the Twitter popular-hashtags dataset could not be used, since the entire Twitter dataset could not be downloaded, so the history of every user in the popular-hashtags dataset could not be collected.
 Instead, the global prior tag history for all users in the Twitter popular-hashtags dataset was used to compute prior activation values for each tag.
@@
 
-The context component for both the random permutation and Bayesian model is based on the same co-occurrence matrix of words and observed tags.
+The context component for both the random permutation and Bayesian model was based on the same co-occurrence matrix of words and observed tags.
 Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
 The matrix for \num{3000000} posts was only used for Twitter, since the number of words in a StackOverflow post is much higher than Twitter on average,
-and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too large for a current high-performance desktop, even with highly optimized representations.
+and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow was too large.
 
@@
 The two removal methods used in \textcite{Sahlgren2008}
-(i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist: \textcite{Salton1988}),
+(i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist: \noparencite{Salton1988}),
 the entropy-weighting method used in \textcite{Stanley2013},
@@
 then it may be worthwhile to experiment with using a compressed random permutation model instead of the full Bayesian model.
-Also newer models of $S_{ji}$ have been introduced that have a more local impact when a new chunk is observed \parencite{thomson2013constraining}.
+Also, newer models of $S_{ji}$ have been introduced that have a more local impact when a new chunk is observed \parencite{thomson2013constraining}.
 It would also be interesting to see how these local models of $S_{ji}$ perform while varying the number of observed associations.
@@
 This was a surprising result.
-One of the main benefits discussed for the random permutation model is that it can easily incorporate word order information into the representation.
+One of the main benefits touted for the random permutation model is that it can easily incorporate word order information into the representation.
 But including word order information does not dramatically improve performance. 
@@
 Given the results in this research, it at least seems worthwhile to use these publicly-available tagging environments as a reasonable proxy to explore and validate modifications to the current theory.
-Any insights gained could be followed up with a more precisely controlled experimental design.
 
@@
 (e.g., compared to word order or properly adding terms).
-Consequently, it may be worthwhile to explore this area more thoroughly with further research.
+Consequently, it may be worthwhile to explore this area more thoroughly in the future.
 
@@
 
-Also perhaps a different $S_{ji}$ model may not need entropy weighing at all (e.g., the Hebbian-inspired associative learning model described in \textcite{thomson2013constraining}).
+Also perhaps a different $S_{ji}$ model may not need entropy weighing at all (e.g., the Hebbian-inspired associative learning model described in \noparencite{thomson2013constraining}).
 Further research could explore how entropy weighting interacts with other models of $S_{ji}$, and not just the random permutation and full Bayesian model.
@@
 One primary advantage with the observational approach here is size and availability: data for millions of posts and tweets are immediately available for analysis.
-However one disadvantage is that there is no way to probe a user at a particular point in time, which could be quite advantageous if possible.
+However one disadvantage is that there is no way to probe a user at a particular point in time, which could be quite advantageous for answering certain questions.
 One hybrid approach worth considering for future research is to use the publicly available datasets to build up a representation of declarative memory for each user.
@@
 After including the prior and entropy components for both models, overall accuracy was similar for the Bayesian and random permutation models for this task.
-So it is unclear with this research precisely which model is more cognitively plausible (or if both are correct and simply working at different levels of abstraction).
+So it is unclear precisely which model is more cognitively plausible (or if both are correct and simply working at different levels of abstraction).
 However, there were two areas in particular where model performance showed relatively large differences: accuracy change as a function of corpus size and reliance on entropy weighting.
@@
 
-This work has shown how large collections of publicly-available behavioral datasets can be used to explore and test different theories of declarative memory and their individual architectural components.
-The end result of this research is [1] a better understanding of how a user's past tagging history influences future tag use,
+This work has shown how large collections of publicly-available behavioral data can be used to explore and test different theories of declarative memory and their individual architectural components.
+This in turn provided [1] a better understanding of how a user's past tagging history influences future tag use,
 [2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on large-scale real-world retrieval tasks,
@@
 
-\subsection{Support Code}
-
-All software code created for this research is publicly available on GitHub \parencite{StanleyRepo2014}.
-
 \clearpage

### 2016-02-27 writeups/psycMethodsResponse/response.tex
foo

\documentclass[answers,12pt]{exam}
\usepackage{xcolor}
\definecolor{SolutionColor}{rgb}{0.1,0.3,1}
\usepackage{lipsum}
\usepackage{enumerate}
\usepackage{alphalph}

\renewcommand{\theenumi}{\AlphAlph{\value{enumi}}}
\renewcommand{\thequestion}{\AlphAlph{\value{question}}}
\renewcommand\questionlabel{\llap{\thequestion)}}

%\pointsinrightmargin
%\boxedpoints
\unframedsolutions
\shadedsolutions
\definecolor{SolutionColor}{rgb}{0.9,0.9,1}
\renewcommand{\solutiontitle}{}

\begin{document}

Dear Editor in Chief, \bigskip

Please read below our answer about the \numquestions{}
questions made by  both refereees. I hope that you agree
with all our comments. \bigskip

Best regards,\bigskip

The author

\begin{questions}

\question This paper has great potential to serve as a meaningful contribution to the 'big data' special issue. Although there was clearly concern about the paper being a novel contribution to the field, it is reasonable to claim that the ties between clear explanation of a theory combined with clear and accessible approach to 'big data' predictive models would constitute a relatively novel contribution.

\begin{solution}
No need to comment
\end{solution}

\question The main concern is for this paper to be highly accessible to an educated-but-untrained audience of psychological researchers -- providing clear and compelling explanation, streamlined language and accessible formulas for ease of reading, and programs and data for hands-on application. It's probably obvious that the more you can do this, the more impactful (visible, citation-worthy) your paper will be.

\begin{solution}
No need to comment directly to this
\end{solution}

\question 2. The expert reviewers agreed that the paper was an impressive effort--and equally important is how they viewed a critical gap between the theory of ACT-R and the data here that are examined with exploratory 'big data' techniques (Rev 1, bullet 1, 3, and especially 4; Rev 2 para 2). In revision, there is opportunity to narrow this gap, such as by explaining further:

\begin{solution}
No need to comment directly to this
\end{solution}

\question (2a) the substantive nature of 'declarative memory' (e.g., maybe you view the content, through analysis, to reflect traces of the 'declarative memory architecture' and changes thereof; Rev 2 3rd para from bottom);

\begin{solution}
Will be addressed in the reviewer's comment
\end{solution}

\question (2b) how exploratory findings inform ACT-R in its instantiations and perhaps even in qualifications/revision to the theory -- there is a balance to strike here: the paper should be a highly accessible example of combining theory with big data (continuing to clarify for the reader -- e.g., Rev 2, para 6, 8, 9), ...

\begin{solution}
Will be addressed in the reviewer's comment
\end{solution}

\question … yet technical details are also important for those with more serious investments in ACT-R; consider including technical details but relegating them to online Supplemental Material (which has no page limits, can be in color, etc.).

\begin{solution}
Added a section and provided a link to the github-hosted repository for this research

\end{solution}

\question (2c) conversely, how ACT-R (which versions) guides the interpretation of exploratory findings (vs. predictive models on big data often being more black-box-like); for instance, clearer explanation of how the formulas instantiate the theory is needed (Rev 1, bullet 6)

\begin{solution}
Will be addressed in reviewer's comment
\end{solution}

\question (2d) how future studies might complement the large-scale field data here with experimental research (Rev 2, para 5)...

\begin{solution}
Will be addressed in reviewer's comment
\end{solution}

\question ...and perhaps research that combines group-level and individual phenomena in a multilevel framework of big-data modeling, analysis and inference (Rev 1, bullet 2)

\begin{solution}
Will be addressed in reviewer's comment
\end{solution}

\question 3. Closer consideration of optimized learning vs. ACT-R learning is warranted, where a reviewer suggests a hybrid approach in the literature that might suggest an important and needed shift in the framing, analysis and implications of the current work (Reviewer 1, bullet 7 -- also be sure to download Rev 1's hand-written PDF, where time is taken to suggest specific edits to the work)

\begin{solution}
Will be addressed in reviewer's comment and suggested edits will be looked at and incorporated
\end{solution}

\question Again, this paper has promise to reflect a novel and important contribution to the journal that brings together an introduction to modern ACT-R, and a strong example of tying theory to big-data models and analyses -- with accessible data sets and annotated code to readers. In doing so, we hope you will take on the challenges enumerated above and by reviewers to develop the paper further -- and in return, we will do our best in providing a more expeditious review process. Note that we reserve the right to bring in an additional expert reviewer as needed, who will have access to the other reviewers' comments.

\begin{solution}
No need to comment
\end{solution}

Reviewer 1:

\question I would first like to commend the authors for an ambitious project. A high degree of scholarship is apparent in their work. Please see my attached review for in-line comments and suggestions. I would like to make some general points that I would like to see.

\begin{solution}
No need to comment
\end{solution}

\question - I found that the paper proceeded too quickly through the initial explanation of ACT-R and vector-based systems. This is especially important because the default version of ACT-R (6) was not described in detail (no Table or Figure was presented)

\begin{solution}
Added table and description for default version of ACT-R 6
\end{solution}

\question and no comparative figure outlining the differences in representation between ACT-R and vector-based systems was presented. I find the lack of discussion of the changes in the ACT-R architecture a major concern. The authors reference their prior work, but I believe it requires a separate section in the document to properly evaluate their current model's performance.

\begin{solution}
Added section that outlines the differences between ACT-R and vector-based systems
\end{solution}

\question - I found throughout the introduction and theory sections that the terminology seemed to waffle between discussion individual user's vs aggregate performance. It wasn't until the methods section that I realized it was due to the difference between how StackOverflow and Twitter information was obtained. I think some extra discussion on the psychological aspects of individual vs group psychology may make your argument stronger.

\begin{solution}
 Added a paragraph in discussion that addresses this and provides one future research idea
\end{solution}

\question - As previously mentioned, I want to highlight that the theoretical background section is covering a lot of ground but the explanations of each are a bit terse. I think the section could be re-worded for clarity's sake.

\begin{solution}
Addressed by adding previous two sections and additional text describing the retrieval process for vector-based systems
\end{solution}

\question - As I read through the paper, I found it unclear to maintain the exact thesis of the work? What is the goal of the research. While the amount of information collection was impressive, the paper came off much like 'throwing XXXX at the wall and seeing what sticks' rather than a targeted hypothesis of what should work and why. As such, the paper is more explanatory than predictive in the traditional sense, to its decrement since the paper is using cognitively-inspired models.

\begin{solution}
Added additional paragraph in intro
\end{solution}

\question - During the interpretation of the various results, I'm concerned that the graphs do not always accord with the discussion. In some cases, I noticed a small drop in accuracy was within the 95%CI bounds between two conditions, making the difference statistically insignificant, however, these results are reported as being lower and a reason attributed (see p35 for an example). This greatly weakens the overall argument. I recommend performing some inferential statistics on the differences in performance to be sure that all reported effects are in fact significant. Even if they only trend a given way, it's important to note the size of the effect in your discussion.

\begin{solution}
Added section in general methodology describing the method of using CIs instead of explicit p-values throughout the paper
Added CI ranges over the difference scores for each of the model comparison accuracies that are written out
Softened text on p35 since the fine-grained distinctions being made (although likely reliable if a more complicated and powerful statistical model was used to test them) were not necessary for the overall conclusions
\end{solution}

\question - I found the use of your entropy measure for attentional weight to be a key thesis of your paper and was intrigued by its use. I feel that this point could/should be expanded upon. What are the behavioral/cognitive consequences, what is its actual cognitive plausibility, why did you focus on attentional weight (Wj) instead of 'fixing' the Sji equation with some associative learning or the Bi with base-level inhibition? There was not a lot of discussion of other existing changes to the ACT-R architecture/theory that could potentially account for your data.

\begin{solution}
Expanded the discussion section for attentional weight and the entropy method	
\end{solution}

\question - I was also a bit concerned with the dichotomy between optimized learning and full-on ACT-R learning. Most models that I've ever written (and everyone else I've collaborated with) use ACT-R's hybrid optimized learning equation that remembers the last 5 or 10 references and then averages the rest. This keeps the model computationally efficient while still having the benefits of local spikes in activation due to local high-frequency events (something you'd see in twitter or StackOverflow during conversations). It seems an oversight to have not tested the optimal hybrid equation. It should be relatively easy for you to run your model varying that parameter and report (briefly) on the results. This could potentially drastically change your interpretation of when to use Bayesian vs vector-based models... which was a section I found specious to begin with.

\begin{solution}
Ran the hybrid form with k=1,5,10 and incorporated the results into the past user behavior section
\end{solution}

\question - This may be a copy-editing issue, but I think the heading levels need to be adjusted. It makes some points (see Word Order in the theoretical background) a bit unclear.

\begin{solution}
Heading level was as intended, but reworded the heading to make it more explicit that the section was a subsection of vector-based memory systems
\end{solution}

\question - From personal preference, I find too many sentences starting with "So" when no conclusion is made. This is made worse since many sentences proceeding "So" then start with "Thus" or "Then". This is stylistic but there are numerous places where phrasing comes off as too informal or 'train-of-thought'.

\begin{solution}
Removed most of them, particularly the ones that were used to start a new paragraph or the ones that the reviewer highlighted in the marked document. Left the rest as I prefer them stylistically over overuse of ‘consequently’, ‘therefore’, ‘thus’, etc.
\end{solution}

\question To summarize:
- Great Scholarship
- Focus on telling a consistent story throughout the paper
- Extend/revise theoretical discussion to include examples, justify changes to ACT-R
- Verify statistics and discussion
- More focus on relevance of attentional weight
- Set optimized learning to hybrid learning

\begin{solution}
All addressed in previous comments
\end{solution}

Reviewer 2:

\question The authors set out to explore large scale "tagging" datasets with the motivation: "Without using large datasets it is difficult to verify and explore modifications to the declarative memory architecture because the datasets are not large enough or rich enough to rigorously test modifications to the theory."

This certainly did wet the present readers appetite, but unfortunately after reading the manuscript, still starving for clearly defined models, testable predictions, or clear conclusions.

\begin{solution}
No need to comment
\end{solution}

\question Overall the topic: "declarative memory" and its role in message tagging should be of significant interest to the journal and its audience. However, the novelty in the present manuscript is limited. Neither the way data is used (basically the problem is framed as a supervised label assignment problem with different sub-populations of tagged messages in Twitter and StackOverflow).

\begin{solution}
Not sure how to address this. I actually find the fact that it looks straightforward a good thing.
\end{solution}

\question The review of the literature and the two model families (ACT-R and vector spaces) is unfortunately not self-contained, rather brief and qualitative. In case the manuscript is re-submitted I suggest that the authors use the opportunity to provide a clear account -from statistical principles to implementation- of the two sets of methods. The original papers are quite frustrating in their mixture of heuristics, arbitrary choice of details, and implementation. I haven't found a paper that gives a clear self-contained account of the relevant theories.

\begin{solution}
Mention the previous additions made (description of ACT-R proper, comparison between the two models, and  additional text describing the retrieval process for vector-based systems)
\end{solution}

\question With respect to the data I think availability is a strong argument. However, there are strong limitations to observational data compared to active experiments (i.e. controlling for message differences by letting multiple users tag the same message)

\begin{solution}
Added a paragraph in the discussion section
\end{solution}

\question The design description lacks crucial details: "Each model was tested by evaluating the accuracy of model-chosen tags on a fresh test subset of the datasets." How was it (sub-)sampled? what was the size?

\begin{solution}
Addressed in the equivalent comment to follow
\end{solution}

\question More importantly I am in serious doubt of the relevance of tagging for "declarative memory". Tagging can best be seen as social behavior/communication. Hence the communication network appears crucial for understanding tag behavior. Specifically, the behavior of the recipient appears crucial to me. In StackOverflow the aim is to get an answer, hence, the main purpose is to get attention from experts, which is more related to expected behavior of the recipient rather than a memory issue. Similarly for Twitter. This issue should at least be discussed.

\begin{solution}
Added a paragraph in the discussion section
\end{solution}

\question It is also not so clear what we have learned from the small differences in performance - can it be explained by differences in model complexity or does it reflect genuine differences in the models of declarative memory?

\begin{solution}
Added a paragraph in the discussion section
\end{solution}

\question It is well known that tags follow heavy tailed (powerlaw) distributions. Hence, normalization and sampling are critical issues. Here we learn "Each model was tested by evaluating the accuracy of model-chosen tags on a fresh test subset of the datasets." How was the test data (sub-)sampled? what were their size and time ordering, how was the tag space controlled /stratified for the test sets?

\begin{solution}
Removed this text at this point, as it was leading the reader to think that I was about to provide more detail about how the sampling was done, and this detail is provided in line for each subsection. For example, much text to follow explicitly defines how each dataset was sampled, and then the methods section for combining prior and context defines how the test data are randomly sampled (including size) apart from the training data, and calls out that data in the test sets are not included to build the co-occurrence matrices
\end{solution}

\question The English is excellent, however the writing could be made considerably more precise; e.g. by clearly defining central concepts, such as "chunk".

\begin{solution}
Added a definition of ‘chunk’ prior to its first use
\end{solution}

\end{questions}

\end{document}

### 2016-02-27 writeups/psycMethods/psycMethods.tex
foo

@@
 
-``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., a chunk) as a vector.
+``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems are an alternative representation of declarative memory.
+They have been successfully used to predict memory retrievals that leverage word order information \parencite{Jones2007}.
+They also have similar structure to memory models that use Latent Semantic Analysis \parencite{Landauer1997} (LSA).
+
+Vector-based models represent a concept (i.e., a chunk) as a vector.
 The representation of the concept is distributed across all elements of the vector.
 This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and viewing the distribution of co-occurrence counts across the rows in the column as the representation of that concept.
-However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.
 
@@
 
-Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
-Both vector-based models and LSA calculate word similarity on a reduced matrix.
+Both vector-based models and LSA calculate word similarity on a reduced co-occurrence matrix.
 For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors is much less than the number of words.
 LSA uses Singular Value Decomposition (SVD) to find words that have highly correlated document occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
-
 For vector-based models, each tag (document) is represented by a compressed vector with a length much less than the number of words.
@@
 
-Accuracy for the hybrid form for the largest $k$ ($k=10$) is slightly higher than the smallest $k$ ($k=1$) (\numNoZero{.334} vs. \numNoZero{.329} \myCI{.0026}{.0023}{.0029}{3608}).
-However accuracy cannot be distinguished between the hybrid form with the largest $k$ and the standard form (\numNoZero{.334} vs. \numNoZero{.334} \myCI{.0000}{-.0000}{.0001}{3609}).
+Accuracy for the hybrid form for the largest $k$ ($k=10$) is slightly higher than the smallest $k$ ($k=1$) (\numNoZero{.334} vs. \numNoZero{.329}, \myCI{.0026}{.0023}{.0029}{3608}).
+However accuracy cannot be distinguished between the hybrid form with the largest $k$ and the standard form (\numNoZero{.334} vs. \numNoZero{.334}, \myCI{.0000}{-.0000}{.0001}{3609}).
 This indicates that for the datasets analyzed in this study the hybrid form with $k=10$ is a suitable replacement to the standard form (produces the same results) and uses less computational resources than the standard form.
@@
 ACT-R sets the default decay rate value to \num{0.5}, and the best-fit value for the standard form is higher (\num{0.8}) for these datasets.
-The figure shows that although model accuracy does slightly improve (\num{0.325} to \num{0.335} \myCI{.0029}{.0022}{.0036}{3609}) for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
+The figure shows that although model accuracy does slightly improve (\num{0.325} to \num{0.335}, \myCI{.0029}{.0022}{.0036}{3609}) for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
 that percentage improvement is not as large (\num{0.28} to \num{0.33}) as the improvement gained from using the standard form compared to the optimized learning form.

### 2016-02-27 dissProject/preProcess.R

@@ -1645,14 +1645,14 @@ renameColDVName <- function(tbl) {
 	setkey(tbl, DVName)
 	mapping = c('topHashtagAcrossPriorStd', 'Standard prior model relaxed across posts',
 		    'topHashtagPostPriorStd', 'Standard prior model',
-		    'topHashtagPostPriorHybrid1', 'Hybrid model with k=01',
-		    'topHashtagPostPriorHybrid5', 'Hybrid model with k=05',
+		    'topHashtagPostPriorHybrid1', 'Hybrid model with k=  1',
+		    'topHashtagPostPriorHybrid5', 'Hybrid model with k=  5',
 		    'topHashtagPostPriorHybrid10', 'Hybrid model with k=10',
 		    'topHashtagPostPriorStdDefaultD', 'Standard prior model at default d=0.5',
 		    'topHashtagPostPriorStdBestDStd', 'Standard prior model at best-fit d=0.8',
 		    'topHashtagPostPriorOL2BestDOL', 'Optimized learning model at best-fit d=0.4',
-		    'topHashtagPostPriorHybrid1BestDStd', 'Hybrid model at best-fit d=0.8 with k=01',
-		    'topHashtagPostPriorHybrid5BestDStd', 'Hybrid model at best-fit d=0.8 with k=05',
+		    'topHashtagPostPriorHybrid1BestDStd', 'Hybrid model at best-fit d=0.8 with k=  1',
+		    'topHashtagPostPriorHybrid5BestDStd', 'Hybrid model at best-fit d=0.8 with k=  5',
 		    'topHashtagPostPriorHybrid10BestDStd', 'Hybrid model at best-fit d=0.8 with k=10',
 		    'topHashtagPostPriorOL2', 'Optimized learning model',
 		    'topHashtagAcrossPriorStdNoffset', 'Standard prior model relaxed across posts without offset',

### 2016-02-27 writeups/psycMethodsResponse/response.tex
foo

@@
 \begin{solution}
-No need to comment
+No need to comment.
 \end{solution}
@@
 \begin{solution}
-No need to comment directly to this
+No need to comment directly to this.
 \end{solution}
@@
 \begin{solution}
-Will be addressed in the reviewer's comment
+Was addressed in the reviewer's comment.
 \end{solution}
@@
 \begin{solution}
-Added a section and provided a link to the github-hosted repository for this research
+Provided a link to the github-hosted repository for this research.
+
+``All software code created for this research is publicly available on GitHub (Stanley, 2014).''
 
@@
 \begin{solution}
-Will be addressed in reviewer's comment
+Was addressed in reviewer's comment.
 \end{solution}
@@
 \begin{solution}
-Will be addressed in reviewer's comment and suggested edits will be looked at and incorporated
+Was addressed in reviewer's comment and suggested edits will be looked at and incorporated.
 \end{solution}
@@
 \begin{solution}
-Added table and description for default version of ACT-R 6
+Added table and description for default version of ACT-R 6.
+
+Expanded the section for the vector-based model and included a description of the retrieval process for that model.
+
+Added a section that directly compares the theory of the vector-based and ACT-R model (``Comparison of Vector-Based Models and ACT-R'').
 \end{solution}
@@
 \begin{solution}
-Added section that outlines the differences between ACT-R and vector-based systems
+Added section that outlines the differences between ACT-R and vector-based systems (``Comparison of Vector-Based Models and ACT-R'').
 \end{solution}
@@
 \begin{solution}
- Added a paragraph in discussion that addresses this and provides one future research idea
+ Added a paragraph in discussion that addresses this and provides one future research idea.
+
+ ``Individual vs.
+Group Representation of Memory.
+Even with the large datasets used in this research, there was not enough data on each individual user to build up separate co-occurrence matrices for them.
+That is, the context component of the declarative memory models used in this research was formed as a collection of observed co-occurrences across all individuals.
+Nonetheless, the prior component for the models was customized for each individual (except in the Twitter popular hashtags dataset when it was
+not possible).
+Ideally, both components would be customized to the individual, and it remains an open question how much accuracy is influenced by using an aggregate or custom context component in the models.''
 \end{solution}
@@
 \begin{solution}
-Addressed by adding previous two sections and additional text describing the retrieval process for vector-based systems
+Added table and description for default version of ACT-R 6.
+
+Expanded the section for the vector-based model and included a description of the retrieval process for that model.
+
+Added a section that directly compares the theory of the vector-based and ACT-R model (``Comparison of Vector-Based Models and ACT-R'').
 \end{solution}
@@
 \begin{solution}
-Added additional paragraph in intro
+Added additional paragraph in introduction.
+
+``As a general methodology for this research, we used a cognitively-constrained
+exploratory model development process.
+Since we worked with large-scale datasets, we had enough data to explore many different modifications simultaneously without the risk of overfitting.
+However, we were not interested in bringing to bear the entire collection of machine learning algorithms to this prediction task in order to completely optimize a model.
+Rather, we were interested in evaluating how well two cognitive models could explain the results, and then making selective modifications to them to see how performance changes.
+The modifications that we explored were selected such that they influenced model accuracy, could be easily added to (or removed from) the cognitive models without massive modifications, and could be considered cognitively plausible.''
 \end{solution}
@@
 \question - During the interpretation of the various results, I'm concerned that the graphs do not always accord with the discussion.
-In some cases, I noticed a small drop in accuracy was within the 95%CI bounds between two conditions, making the difference statistically insignificant, however, these results are reported as being lower and a reason attributed (see p35 for an example).
+In some cases, I noticed a small drop in accuracy was within the 95\%CI bounds between two conditions, making the difference statistically insignificant, however, these results are reported as being lower and a reason attributed (see p35 for an example).
 This greatly weakens the overall argument.
@@
 \begin{solution}
-Added section in general methodology describing the method of using CIs instead of explicit p-values throughout the paper
-Added CI ranges over the difference scores for each of the model comparison accuracies that are written out
-Softened text on p35 since the fine-grained distinctions being made (although likely reliable if a more complicated and powerful statistical model was used to test them) were not necessary for the overall conclusions
+Added section in general methodology describing the method of using CIs instead of explicit p-values throughout the paper.
+
+``In similar spirit to Cumming and Finch (2005), Masson and Loftus (2003), Tryon (2001), confidence intervals were used throughout for statistical inference instead of explicit p-values.
+This was done because it is a more straightforward way to visualize all of the mean differences in accuracy (i.e., bar and line plots with a visual indication of the error) for each of the models, which eases interpretation.
+It is also a conservative estimate of
+reliability, as using confidence intervals on the means essentially ignores the fact that this is a within-subjects design, where accuracy for each model is measured for each subject in each dataset.
+However, we would rather err on the side of making the results easier to interpret at the cost of using conservative estimates for reliability (i.e., use error bars overlaid on a visualization of the means instead of statistical tests on every combination of difference scores).
+In fact, using large datasets for analysis is what empowers us to do this: With this much data, we can use straightforward (although more conservative) inferentials and still see reliable results.
+This allows us to concentrate more on proper data visualization of the main findings.
+Nonetheless, when accuracy differences between two models are called out explicitly in the text, the confidence interval for the mean difference in accuracy is also included.
+Since this interval is drawn across the observed difference scores, it will be smaller than the intervals included in the plots, since these are drawn independently for each model’s overall accuracy.''
+
+Added CI ranges over the difference scores for each of the model comparison accuracies that are written out.
+
+e.g.: ``Accuracy for the hybrid form for the largest k (k = 10) is slightly higher than the smallest k (k = 1) (.334 vs. .329, mean difference = 0.0026, 95\% CI [0.0023, 0.0029]).''
+
+Softened text on p35 since the fine-grained distinctions being made (although likely reliable if a more complicated and powerful statistical model was used to test them) were not necessary for the overall conclusions.
+
+``Model accuracy as a function of co-occurrence matrix size for StackOverflow is depicted in Figure 11.
+Accuracy improves for both the Bayesian and random permutation models as the size of the documents used to generate the co-occurrence matrix increases.
+However for the random permutation model, accuracy begins to plateau for the largest corpus size for all three tested versions.
+The Bayesian model does not plateau nearly as much and accuracy continues to improve even at the largest measured corpus size.''
+
 \end{solution}
@@
 Expanded the discussion section for attentional weight and the entropy method
+
+``Attentional Weight.
+The Bayesian declarative memory model for ACT-R has an attentional weight term Wj that can be used to attenuate or remove stop words.
+The random permutation model was also easily modified to have an attentional weight term by attenuating each word’s set of ones and negative ones values by that word’s attentional weight.
+Both a method for stop-word removal (frequency filtering) and stop-word attenuating (entropy weighting) performed well for both models.
+
+It is unclear at this point if this modification is cognitively plausible, even if the modification was cleanly added to the current ACT-R model.
+What can be said from this research is that using a proper method for handling the low-predictor words produced some of the largest improvements in model accuracy compared to the other explorations (e.g., compared to word order or properly adding terms).
+Consequently, it may be worthwhile to explore this area more thoroughly in the future.
+
+Also, there may be other ways besides adjusting the attentional weight that these models can be modified to achieve the same performance improvement.
+The entropy weighting term is likely solving a problem that resides only in contextual activation and not prior activation, since the activation for the prior component for each word is independent of the presentation of other words, so stop words have no way to interact with tag prior activation in this case.
+
+One alternative explanation is that the entropy weighting measure was needed because the contextual co-occurrence matrix was not yet stable, even with these large dataset sizes.
+If the matrix was not yet stable, then by chance certain stop words would contribute more net activation to specific tags and introduce noise into the retrieval process for that tag.
+So further weighting the contribution of these stop words by leveraging the attentional weight term ensured that this noise was suppressed in the system.
+If this is the case, it would be interesting to see how large a dataset is required to not need to modify the attentional weight term.
+However even if this is the case, the random permutation model would still require entropy weighting for larger dataset sizes, since this model simply adds environment vectors to produce the current context vector and has no mechanism (unlike Bayesian normalization) for weighting the contribution of an environment vector.
+
+Another explanation for the Bayesian model is that additional attentional weighting was needed because the contextual matrix was derived only from word by tag co-occurrences and not (word or tag) by (word or tag) co-occurrences.
+This was done for performance reasons as the computational complexity otherwise is simply too high for current hardware.
+It is possible that by adding all pairwise combinations of both words and tags in the posts may soften and stabilize the contribution of stop words when computing contextual activation for a tag, given current context.
+
+Also perhaps a different Sji model may not need entropy weighing at all (e.g., the Hebbian-inspired associative learning model described in Thomson and Lebiere, 2013).
+Further research could explore how entropy weighting interacts with other models of Sji, and not just the random permutation and full Bayesian model.
+
+Nonetheless, it seems likely that the entropy method for computing the attentional weight term Wj may be applicable to other declarative memory tasks.
+The entropy method is parameter-free (unlike the frequency-filtering method), completely data driven (much like the co-occurrence matrix), does not require tuning, and increases model accuracy for these datasets.
+It is common to handle stop words in some manner when working with Natural Language Processing (NLP) and large text corpuses.
+So, if these large corpuses are used to test declarative memory, then the entropy method may be a more cognitively plausible way of dealing with stop words rather than filtering based on a predetermined list or the frequency count in the dataset.''
+
 \end{solution}
@@
 \begin{solution}
-Ran the hybrid form with k=1,5,10 and incorporated the results into the past user behavior section
+Ran the hybrid form with k=1,5,10 and incorporated the results into the past user behavior section.
+
+``However the hybrid form with a reasonable k is just as accurate as the standard form.
+This is due to the fact that the majority of change in activation for a chunk occurs in the short period of time after the presentation of the chunk (a property of a power law distribution).
+Once a chunk’s presentations decay to the flat area of the power law distribution, not much change in activation for these presentations occurs, so they can be averaged and their presentation time can be discarded without much loss in fidelity.
+Therefore if computational speed is a concern, one may wish to leverage the hybrid form instead of the standard form.
+Although qualitatively the standard form produces the cleanest and most natural effect of decay rate on accuracy, as shown in Figure 1.
+It also does not have a bifurcation at k = 1.
+Consequently, if one is a purist and computational speed is less of a concern, one may still wish to use the standard form over the hybrid form.''
+
 \end{solution}
@@
 \begin{solution}
-Heading level was as intended, but reworded the heading to make it more explicit that the section was a subsection of vector-based memory systems
+Heading level was as intended, but reworded the heading to make it more explicit that the section was a subsection of vector-based memory systems.
 \end{solution}
@@
 Removed most of them, particularly the ones that were used to start a new paragraph or the ones that the reviewer highlighted in the marked document.
-Left the rest as I prefer them stylistically over overuse of ‘consequently’, ‘therefore’, ‘thus’, etc.
+I left a few as I prefer them stylistically over overuse of ‘consequently’, ‘therefore’, ‘thus’, etc.
 \end{solution}

### 2016-02-27 writeups/psycMethodsResponse/response.tex
foo

@@
 
+% http://tex.stackexchange.com/questions/36423/random-unwanted-space-between-paragraphs
+\raggedbottom
+
 \begin{document}
 
-Dear Editor in Chief, \bigskip
+To Lisa	Harlow and Fred	Oswald,	co-editors, \bigskip
 
-Please read below our answer about the \numquestions{}
-questions made by  both refereees. I hope that you agree
-with all our comments. \bigskip
+Please see below our responses to the reviewer comments.
+Along with pasting text from the revised document inline with the questions where applicable,
+I have also attached a version of the .pdf that contains markup for the differences between the original and revised submission. \bigskip
 
-Best regards,\bigskip
+We hope that you and the reviewers find these changes sufficient and that you find the revised document well suited for this special issue. \bigskip
 
-The author
+Respectfully,
+
+-Clayton Stanley
+
+\bigskip
+\bigskip
 
@@
 \begin{solution}
-No need to comment.
+No need to comment directly.
 \end{solution}
@@
 \begin{solution}
-Was addressed in reviewer's comment and suggested edits will be looked at and incorporated.
+Was addressed in reviewer's comment and suggested edits were looked at and incorporated.
 \end{solution}
@@
 \begin{solution}
-No need to comment
+No need to comment directly.
 \end{solution}
@@
 \begin{solution}
- Added a paragraph in discussion that addresses this and provides one future research idea.
+ Added a paragraph in discussion that addresses this.
 
@@
 This is due to the fact that the majority of change in activation for a chunk occurs in the short period of time after the presentation of the chunk (a property of a power law distribution).
-Once a chunk’s presentations decay to the flat area of the power law distribution, not much change in activation for these presentations occurs, so they can be averaged and their presentation time can be discarded without much loss in fidelity.
+Once a chunk's presentations decay to the flat area of the power law distribution, not much change in activation for these presentations occurs, so they can be averaged and their presentation time can be discarded without much loss in fidelity.
 Therefore if computational speed is a concern, one may wish to leverage the hybrid form instead of the standard form.
@@
 \begin{solution}
-All addressed in previous comments
+All addressed in previous comments.
 \end{solution}
@@
 \begin{solution}
-No need to comment
+  Specific issues were addressed in the following comments from the reviewer.
 \end{solution}
@@
 \begin{solution}
-Mention the previous additions made (description of ACT-R proper, comparison between the two models, and  additional text describing the retrieval process for vector-based systems)
+Added table and description for default version of ACT-R 6.
+
+Expanded the section for the vector-based model and included a description of the retrieval process for that model.
+
+Added a section that directly compares the theory of the vector-based and ACT-R model (``Comparison of Vector-Based Models and ACT-R'').
 \end{solution}
@@
 \begin{solution}
-Added a paragraph in the discussion section
+Added a paragraph in the discussion section.
+
+``Observational vs. Experimental Design. All data used for this study were collected from publicly-available datasets and were not subject to an experimental design
+(i.e., observational data).
+There are advantages and disadvantages to both observational and experimental designs.
+One primary advantage with the observational approach here is size and availability: data for millions of posts and tweets are immediately available for analysis.
+However one disadvantage is that there is no way to probe a user at a particular point in time, which could be quite advantageous for answering certain questions.
+One hybrid approach worth considering for future research is to use the publicly available datasets to build up a representation of declarative memory for each user.
+Afterwards, bring a sample of these specific users into a laboratory setting and probe them with specific retrieval requests.
+Multiple users could be asked to tag the same message, and then models could be tested to see if they reliably predict differences in accuracy between the subjects.''
 \end{solution}
@@
 \begin{solution}
-Addressed in the equivalent comment to follow
+  Removed this text at this point, as it was leading the reader to think that I was about to provide more detail about how the sampling was done, and this detail is provided in line for each subsection.
+For example, much text to follow explicitly defines how each dataset was sampled, and then the methods section for combining prior and context defines how the test data are randomly sampled (including size) apart from the training data, and calls out that data in the test sets are not included to build the co-occurrence matrices.
 \end{solution}
@@
 \begin{solution}
-Added a paragraph in the discussion section
+Added a paragraph in the discussion section.
+
+``Tagging as a Declarative Memory Retrieval Process.
+Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter reasonably well.
+The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
+
+These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
+However, both models are not 100 percent accurate, and the amount of variance left to predict is larger than what would be observed if retrieval noise is included in the model.
+One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval (e.g., strategic processes, determining the number of tags to employ, attempts to add humor).
+
+Another possibility is that this is not a declarative memory process at all.
+In particular, it is important to note that tagging on these sites is widely used as a method to communicate specific messages to specific groups of people in the community.
+It is unclear how much this environment overlaps with verbal communication between two people, and the declarative memory requests required to successfully communicate in this setting.
+Nonetheless, these models are reasonably accurate in this setting, particularly for StackOverflow.
+Given the results in this research, it at least seems worthwhile to use these publicly-available tagging environments as a reasonable proxy to explore and validate modifications to the current theory.''
+
 \end{solution}
@@
 \begin{solution}
-Added a paragraph in the discussion section
+Added a paragraph in the discussion section.
+
+``Directly Comparing the Models.
+After including the prior and entropy components for both models, overall accuracy was similar for the Bayesian and random permutation models for this task.
+So it is unclear precisely which model is more cognitively plausible (or if both are correct and simply working at different levels of abstraction).
+However, there were two areas in particular where model performance showed relatively large differences: accuracy change as a function of corpus size and reliance on entropy weighting.
+These areas could be used in future research to more directly challenge the models.
+For example, a domain could be picked where enough data is collected to build a custom co-occurrence matrix for each user.
+Then one could look at model accuracy over time for each user as the co-occurrence matrix is increased to see if the trajectory is more accurately modeled by a random permutation (more accurate with lower N) or Bayesian (more accurate with higher N) model.
+Further, if a large-scale domain is picked and it is shown that the Bayesian model does not require entropy weighting for the co-occurrence matrix after the representations are fully stabilized, then this would provide a case where the two models are dramatically different: The random permutation model would require explicitly weighting the stop words (entropy weighting) while the Bayesian model would not.''
+
 \end{solution}
@@
 Removed this text at this point, as it was leading the reader to think that I was about to provide more detail about how the sampling was done, and this detail is provided in line for each subsection.
-For example, much text to follow explicitly defines how each dataset was sampled, and then the methods section for combining prior and context defines how the test data are randomly sampled (including size) apart from the training data, and calls out that data in the test sets are not included to build the co-occurrence matrices
+For example, much text to follow explicitly defines how each dataset was sampled, and then the methods section for combining prior and context defines how the test data are randomly sampled (including size) apart from the training data, and calls out that data in the test sets are not included to build the co-occurrence matrices.
+
 \end{solution}
@@
 \begin{solution}
-Added a definition of ‘chunk’ prior to its first use
+Added a definition of ‘chunk’ prior to its first use.
+
+``ACT-R (Anderson, 2007) is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
+The declarative memory system is a component of the architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
+At the core of the system is the concept of a “chunk”: the representation of a piece of knowledge (e.g., a word).''
+
 \end{solution}

### 2016-06-06 writeups/psycMethodsResponse2/response.tex
initial commit of revision 2

\documentclass[answers,12pt]{exam}
\usepackage{xcolor}
\definecolor{SolutionColor}{rgb}{0.1,0.3,1}
\usepackage{lipsum}
\usepackage{enumerate}
\usepackage{alphalph}

\renewcommand{\theenumi}{\AlphAlph{\value{enumi}}}
\renewcommand{\thequestion}{\AlphAlph{\value{question}}}
\renewcommand\questionlabel{\llap{\thequestion)}}

%\pointsinrightmargin
%\boxedpoints
\unframedsolutions
\shadedsolutions
\definecolor{SolutionColor}{rgb}{0.9,0.9,1}
\renewcommand{\solutiontitle}{}

% http://tex.stackexchange.com/questions/36423/random-unwanted-space-between-paragraphs
\raggedbottom

\begin{document}

To Lisa	Harlow and Fred	Oswald,	co-editors, \bigskip

Please see below our responses to the reviewer comments.
Along with pasting text from the revised document inline with the questions where applicable,
I have also attached a version of the .pdf that contains markup for the differences between the original and revised submission. \bigskip

We hope that you and the reviewers find these changes sufficient and that you find the revised document well suited for this special issue. \bigskip

Respectfully,

-Clayton Stanley

\bigskip
\bigskip

\begin{questions}

\question Primary concern: Papers on predictive modeling typically avoid confidence intervals and instead seek to optimize out-of-sample RMSE or R2 based on training vs. test samples (e.g., k-fold cross-validation) or sometimes by resampling (bootstrapping); see overviews of predictive modeling such as http://www-bcf.usc.edu/~gareth/ISL/ and http://appliedpredictivemodeling.com/.  The general concern here is that people reading your paper alongside other 'big data' papers may view your confidence interval approach unusual or outdated; the more specific concern here is overfitting the model to the data (making confidence intervals too confident).

\begin{solution}
No need to comment directly.
\end{solution}

\question Lesser concern: Per Rev 3 para 3, consider how prediction might increase when methods dynamically afford an increase in model dimensionality whenever the size of a corpus increases (this could be considered in terms of differences in sizes across corpora and/or across users). It is fine if you simply point this out as future research in your concluding statements, although an example could be incorporated if you so desired. Your choice.

\begin{solution}
No need to comment directly.
\end{solution}

\question Lesser concern: Per Rev 3, para 5 and 7, please incorporate any additional relevant literature (but do not limit yourself to Rev 3 citations; check out what is current and relevant given this quickly changing phenomenon and research area).

\begin{solution}
No need to comment directly.
\end{solution}

\question If you have not already done so as part of your Author Note, please provide the details (2-4 sentences) of prior dissemination of the ideas and data appearing in the manuscript (e.g., if some or all of the data and ideas in the manuscript were presented at a conference or meeting posted on a listserv, shared on a website, etc.).

\begin{solution}
No need to comment directly.
\end{solution}

Reviewer 3:

\question The counterintuitive results are also intriguing (such as the finding that larger training corpora lowered performance for the random permutations models). However, one factor to consider is that RP models depend on the approximate orthogonality of the random projections (the environment vectors for Beagle). As the number of words increases, this will tend to oversaturate the space, and commonly leads to decreased performance. One standard approach here is to scale the dimensionality of the representations as the training set and vocabulary sizes increase. It would be interesting to see if the performance still leveled off if you increased the dimensionality with the corpus size.

\begin{solution}
\end{solution}

\question A number of works have considered the large scale dynamics of hashtag use (e.g. ``\#Bigbirds Never Die: Understanding Social Dynamics of Emergent Hashtags'', 2013), while other work has looked specifically at the impact of tags used by a user's local network connections (``We know what @you \#tag: does the dual role affect hashtag adoption?'', 2012). There is an extensive literature exploring how user features including demographics (e.g. ``User Conditional Hashtag Prediction for Images'', 2015), prior posts, time of day, mood, etc. The relationship between this literature and the work presented in the paper needs to be discussed thoroughly.

\begin{solution}
\end{solution}

\question While the authors discuss that declarative memory might not be the correct model for this task (p 51-52), this seems to be almost an afterthought. Given that there is a flourishing literature which goes beyond this narrow focus, it seems important to acknowledge that upfront in the paper.

\begin{solution}

\end{solution}

\question That said, declarative memory retrieval is still likely to be a critical part of this process and worth exploration on its own. As such, the paper can still work even without incorporating any of those additional features. Many papers in this area report whether not just whether the tag from the model was predicted, but also whether it was in the top n (10, 25, etc) most likely predictions returned by the model (for one example of this, see ``\#TAGSPACE: Semantic Embeddings from Hashtags'', 2014). One possibility would be to treat the declarative memory portion (modeled here) as generating candidates which other processes consider.

\begin{solution}

\end{solution}

\question Questions of cognitive plausibility were also present in the structure of some of the pre-processing techniques used. While the authors acknowledge this, it would have been helpful to see more about how they felt that non-cognitive modeling or standard NLP techniques can help us to better understand and refine the underlying cognitive models. At very least acknowledging this intersection would have helped to eliminate some of the sense that the paper occasionally veered between claims of cognitive plausibility and more conventional computational modeling.

\begin{solution}
\end{solution}

\end{questions}

\end{document}

### 2016-06-06 writeups/psycMethods/psycMethods.tex
foo

@@
 
+\subsection{Compression Level for Random Permutation}
+
+Another parameter in the random permutation model is the number of rows used to represent all of the different words.
+Each word's environment vector is represented by two randomly placed ones and two negative ones across the rows.
+All that is required is that the four positions and signs of each of these environment vectors is unique.
+But this means multiple environment vectors can overlap when looking at a single row.
+This is a lossy form of compression, and allows the number of rows to represent the words to be much less than the total number of words.
+As the number of rows decreases, the amount of cross talk between the environment vectors increases, compression increases, and the amount of information stored decreases.
+As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.
+
+\subsubsection{Method}
+
+Three different row-size values were tested for the random permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
+A low value (\num{100}), a more standard value (\num{2048}), and a high value (\num{10000}) of rows were tested.
+A separate co-occurrence representation was built for each of these row values.
+Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
+Accuracy measurements were the same as were used for previous runs of these datasets: mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.
+
+\subsubsection{Results}
+
+Results for using different levels of compression for the random permutation model for the StackOverflow dataset are depicted in Figure \ref{figContextDimSO}.
+In the figure, the model name that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
+This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random permutation model for all other results.
+The plot is also broken out for different size datasets used to build the co-occurrence matrix.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
+  \caption{
+    Effect of compression on random permutation model for StackOverflow.
+    Error bars represent the 95\% bootstrapped confidence interval for the mean.
+  }
+  \label{figContextDimSO}
+\end{figure}
+
+Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy, which shows that compression can influence model accuracy.
+However, \num{2048} rows is sufficient for these datasets, as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy, even for the highest corpus sizes.
+
+Results are similar for Twitter.
+Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100}, and little to no impact when increasing from \num{2048} to \num{10000}. 
+These results are consistent across all corpus sizes used to build the co-occurrence matrix.
+
+\begin{figure}[!htbp]
+  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeT-crop.pdf}}}
+  \caption{
+    Effect of compression on random permutation model for Twitter.
+    Error bars represent the 95\% bootstrapped confidence interval for the mean.
+  }
+  \label{figContextDimT}
+\end{figure}
+
+\subsubsection{Discussion}
+
+The fact that the random permutation model only needs around \num{2048} rows to represent all of the words in the Twitter and StackOverflow datasets is impressive.
+Twitter has around \num{2383000} unique words for the \num{3000000} tweets in each of the popular-users datasets.
+StackOverflow has \num{5848000} unique words for the \num{1000000} posts used to generate the co-occurrence matrix.
+So it is impressive that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
+This is likely because word use frequency for StackOverflow and Twitter approximately follows Zipf's law.
+\textcite{Stanley2013} showed that frequency of tag use for StackOverflow approximately follows a Zipf's law distribution.
+This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply when moving to words that are used fewer times.
+Since these words have little to no ability to influence predictions, it makes sense that they can be safely removed from the dataset.
+
 \subsection{Combining Prior and Context}

### 2016-06-06 writeups/psycMethods/psycMethods.tex
foo

@@
 
-This section will examine how model accuracy is influenced by four architectural concerns:
-method of handling stop words, influence of dataset size, method of combining prior and context, and word order.
+This section will examine how model accuracy is influenced by five architectural concerns:
+method of handling stop words, influence of dataset size, influence of compression for the random permutation model, method of combining prior and context, and word order.
 These particular architectural concerns were examined for the following reasons:
@@
 so the effect of dataset size on accuracy was examined in more detail.
+\textcite{Sahlgren2008} showed how different levels of compression in the random permutation model influence accuracy, so this was examined for three reasons:
+to replicate the findings, as an additional check to ensure that the random permutation model was implemented correctly, and also to see if the same levels of compression are valid for the much larger datasets used in this research.
 One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random permutation model,
@@
 This is a lossy form of compression, and allows the number of rows to represent the words to be much less than the total number of words.
-As the number of rows decreases, the amount of cross talk between the environment vectors increases, compression increases, and the amount of information stored decreases.
+As the number of rows decreases, the amount of crosstalk between the environment vectors increases, compression increases, and the amount of information stored decreases.
 As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.
@@
 StackOverflow has \num{5848000} unique words for the \num{1000000} posts used to generate the co-occurrence matrix.
-So it is impressive that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
+This means that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
 This is likely because word use frequency for StackOverflow and Twitter approximately follows Zipf's law.
 \textcite{Stanley2013} showed that frequency of tag use for StackOverflow approximately follows a Zipf's law distribution.
-This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply when moving to words that are used fewer times.
-Since these words have little to no ability to influence predictions, it makes sense that they can be safely removed from the dataset.
+This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply.
+Since these low-use words have little to no ability to influence predictions, it makes sense that accuracy does not substantially improve by providing additional dimensions for them.
 
@@
 However, this means that ACT-R's prior base-level learning component (which is a log-odds value) is added to a correlation value, which may not be mathematically appropriate.
-As a third architectural concern, another method to combine the prior and context components for the random permutation model was explored.
+As a fourth architectural concern, another method to combine the prior and context components for the random permutation model was explored.
 

### 2016-06-06 writeups/psycMethodsResponse2/response.tex
foo

@@
 \begin{solution}
-No need to comment directly.
+This work was already done but was not yet included in the submission.
+I have added it to the paper and the text is included below for convenience.
+
 \end{solution}
@@
 \begin{solution}
+Addressed this by including in the paper the work that looked at model accuracy vs. corpus size and number of rows in the random permutation model.
+The text for the section has been previously included in this document.
 \end{solution}

### 2016-06-06 writeups/psycMethodsResponse2/response.tex
foo

@@
 
+FIXME: Add text
+
 \end{solution}
@@
 \begin{solution}
-No need to comment directly.
+The work presented in this paper was done for my dissertation.
+The dissertation paper is available from Rice University's digital archive (https://scholarship.rice.edu/handle/1911/88165).
 \end{solution}
@@
 
+There was some text in the beginning of the discussion section that mentioned this possibility.
+However it was embedded within parentheses and the other alternatives listed were not as important as this suggestion.
+So I broke this specific possibility out of parentheses and mentioned it directly:
+
+``These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
+However, both models are not 100 percent accurate, and the amount of variance left to predict is larger than what would be observed if retrieval noise is included in the model.
+One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval.
+For example, it may be the case that the declarative system is used to provide a set of likely tag candidates to higher-level processes to consider for selection.''
+
 \end{solution}

### 2016-06-06 writeups/psycMethods/psycMethods.tex
foo

@@
 However, both models are not 100 percent accurate, and the amount of variance left to predict is larger than what would be observed if retrieval noise is included in the model.
-One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval
-(e.g., strategic processes, determining the number of tags to employ, attempts to add humor).
+One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval.
+For example, it may be the case that the declarative system is used to provide a set of likely tag candidates to higher-level processes to consider for selection.
 

### 2016-06-07 writeups/psycMethods/psycMethods.tex
foo

@@
 For each of the four Twitter popular-hashtags datasets, model retrievals were done across 15 runs and 500 posts within each run.
-For the StackOverflow randomly-sampled dataset, 5 runs of 100 posts were used,
+For the StackOverflow randomly-sampled dataset, 5 runs of 500 posts were used,
 since the results for StackOverflow were more stable and retrievals took more wall-clock time for this dataset due to the size of the co-occurrence matrix.

### 2016-06-07 dissProject/preProcess.R

@@ -2262,7 +2262,11 @@ bestFitNameToDVName <- function(vect, pre) {
 }
 
 analyzeLogreg <- function(logregTbl, modelVsPredTbl) {
+	modelVsPredTbl[sizeNum == 2][, .N, d]
+	modelVsPredTbl[predUsedBest == T]
+	logregTbl
 	dTbl = modelVsPredTbl[predUsedBest == T][, list(d,  runNum, groupNum, sizeNum, DVName)][grepl('^topHashtagAcross', DVName)]
+	dTbl
 	setkey(dTbl, DVName, runNum, groupNum, sizeNum, d)
 	logregTbl
 	dTbl
@@ -2276,6 +2280,30 @@ analyzeLogreg <- function(logregTbl, modelVsPredTbl) {
 	compareMeanDVLogreg(logregTbl[predUsedBest & (sizeNum == 2) & dsetType == 'stackoverflow' & DVName != '(Intercept)' & bestFitName %in% bestFitNames], coeffStd, figName='coeffSO')
 	bestFitNames = c('actPriorStd_actTweet', 'actPriorStd_actTweetUsercontext', 'actPriorStd_actTweetOrderlessEnthyman', 'actPriorStd_actTweetOrderlessEnthyser')
 	compareMeanDVLogreg(logregTbl[predUsedBest & (sizeNum == 2) & dsetType == 'twitter' & predName != '(Intercept)' & bestFitName %in% bestFitNames], coeffStd, figName='coeffT')
+}
+
+analyzeOverfitting <- function(logregTbl, modelVsPredTbl) {
+	modelVsPredTbl[predUsedBest == T]
+	bdTbl = modelVsPredTbl[!is.na(topHashtag) & predUsedBest == T, .N, .(datasetName, runNum, groupNum, sizeNum, DVName, d)]
+	bdTbl
+	aTbl = modelVsPredTbl[bdTbl, nomatch=0, on=c('datasetName', 'runNum', 'groupNum', 'sizeNum', 'DVName', 'd')]
+	aTbl[, .N, .(dsetType, dsetGroup, modelType)]
+	aTbl[sizeNum == 2][, .N, .(datasetName, dsetSize)]
+	aTbl[dsetType == 'twitter' & sizeNum == 2]
+	aTbl
+
+	aTbl[, .SD
+	     ][!is.na(topHashtag)
[…]

### 2016-06-12 writeups/psycMethods/psycMethods.tex
foo

@@
 
+Other successful models have shown that higher-level processes such as a user's goals for picking a hashtag, demographics, and specifics about how a hashtag has been used in the past can also influence hashtag use.
+\textcite{lin2013bigbirds} showed that there are at least two distinct use cases for hashtags on Twitter: to mark the content (traditional tagging) and to contribute to a specific community that has created a hashtag.
+\textcite{yang2012we} found that demographics such as gender, age, and location can influence a user's selected hashtag.
+\textcite{denton2015user} found that the growth and lifetime of a hashtag can be influenced by factors such as how often the hashtag was retweeted (and by how many unique users), and the number of followers for a user that used a particular hashtag.
+These types of goal-related features, demographic information, and higher-level attributes of previous hashtag use were not directly included in the declarative memory models that were tested for this research.
+They were not included because it would be difficult (although likely not impossible) to fit these features into a modeling framework that is constrained to use only declarative memory processes,
+and this research focused on measuring differences in model accuracy for two declarative memory models.
+Nonetheless many of these higher-level features may indirectly influence a declarative memory model by influencing the specific tags that a user has chosen in the past,
+which in turn boosts base-level activation for these previously-chosen tags, which in turn makes it more likely that they are chosen in the future.
+
 There has been a mix of co-occurrence models created to predict hashtag use on Twitter, and the results are encouraging.

### 2016-06-12 dissProject/preProcess.R

@@ -2284,7 +2284,7 @@ analyzeLogreg <- function(logregTbl, modelVsPredTbl) {
 
 analyzeOverfitting <- function(logregTbl, modelVsPredTbl) {
 	modelVsPredTbl[predUsedBest == T]
-	bdTbl = modelVsPredTbl[!is.na(topHashtag) & predUsedBest == T, .N, .(datasetName, runNum, groupNum, sizeNum, DVName, d)]
+	bdTbl = modelVsPredTbl[!is.na(topHashtag) & predUsedBest == T, .N, .(datasetName, runNum, groupNum, sizeNum, DVName, d, predUsedBest)]
 	bdTbl
 	aTbl = modelVsPredTbl[bdTbl, nomatch=0, on=c('datasetName', 'runNum', 'groupNum', 'sizeNum', 'DVName', 'd')]
 	aTbl[, .N, .(dsetType, dsetGroup, modelType)]
@@ -2303,15 +2303,37 @@ analyzeOverfitting <- function(logregTbl, modelVsPredTbl) {
 	     ]
 
 
+	logregTbl[, bestFitNameAsDVName := bestFitNameToDVName(bestFitName, 'topHashtagAcross')]
+	logregTbl[modelType == 'perm' & DVName == 'actPriorStd', DVName := 'actPriorStdRP']
+	logregTbl[modelType == 'sji' & DVName == 'actPriorStd', DVName := 'actPriorStdBayes']
 	bTbl = logregTbl[, !'Pr(>|z|)', with=F]
+	bdTbl
 	bTbl = bTbl[bdTbl, nomatch=0, on=c(datasetName='datasetName', runNum='runNum', groupNum='groupNum', sizeNum='sizeNum', bestFitNameAsDVName='DVName', d='d')]
 	bTbl
 	bestFitNames = c('actPriorStd_actTitle_actBody', 'actPriorStd_actTitleOrderlessEnthyman_actBodyOrderlessEnthyman',
 			 'actPriorStd_actTweet', 'actPriorStd_actTweetUsercontext', 'actPriorStd_actTweetOrderlessEnthyman', 'actPriorStd_actTweetOrderlessEnthyser')
+
+	bTbl
 	bTbl[predUsedBest & (sizeNum == 2) & predName != '(Intercept)' & bestFitName %in% bestFitNames
-	     ][, ggplot(.SD, aes(DVName, coeffStd))
+	     ][, label := DVName
[…]

### 2016-06-12 writeups/psycMethodsResponse2/response.tex
foo

@@
 \begin{solution}
-No need to comment directly.
+Added a paragraph in the background section to discuss the types of hashtag-prediction models that were mentioned by Rev 3:
+
+fixme: add text
 \end{solution}
@@
 \begin{solution}
-\end{solution}
-
-\question While the authors discuss that declarative memory might not be the correct model for this task (p 51-52), this seems to be almost an afterthought.
-Given that there is a flourishing literature which goes beyond this narrow focus, it seems important to acknowledge that upfront in the paper.
-
-\begin{solution}
-
+Addressed this by reviewing the literature and including a paragraph in the background section.
+The text for this section has been previously included in this document.
 \end{solution}
@@
 
-\question Questions of cognitive plausibility were also present in the structure of some of the pre-processing techniques used.
-While the authors acknowledge this, it would have been helpful to see more about how they felt that non-cognitive modeling or standard NLP techniques can help us to better understand and refine the underlying cognitive models.
-At very least acknowledging this intersection would have helped to eliminate some of the sense that the paper occasionally veered between claims of cognitive plausibility and more conventional computational modeling.
-
-\begin{solution}
-\end{solution}
-
 \end{questions}

### 2016-06-12 writeups/psycMethods/psycMethods.tex
foo

@@
 \textcite{yang2012we} found that demographics such as gender, age, and location can influence a user's selected hashtag.
-\textcite{denton2015user} found that the growth and lifetime of a hashtag can be influenced by factors such as how often the hashtag was retweeted (and by how many unique users), and the number of followers for a user that used a particular hashtag.
+\textcite{denton2015user} found that the growth and lifetime of a hashtag can be influenced by factors such as how often the hashtag was retweeted and the number of followers for a user that used a particular hashtag.
 These types of goal-related features, demographic information, and higher-level attributes of previous hashtag use were not directly included in the declarative memory models that were tested for this research.

### 2016-06-13 writeups/psycMethods/psycMethods.tex
foo

@@
 These types of goal-related features, demographic information, and higher-level attributes of previous hashtag use were not directly included in the declarative memory models that were tested for this research.
-They were not included because it would be difficult (although likely not impossible) to fit these features into a modeling framework that is constrained to use only declarative memory processes,
+They were not included because it would be difficult to fit these features into a modeling framework that is constrained to use only declarative memory processes,
 and this research focused on measuring differences in model accuracy for two declarative memory models.

### 2016-06-13 writeups/psycMethods/psycMethods.tex
foo

@@
 
+\subsubsection{Sampling and Overfitting}
+
+This research centers around evaluating the differences in model accuracy when various components are added, removed, or altered.
+When evaluating the models, overall model accuracy for a specific model and specific dataset is less important than how that accuracy deviates from a comparative model on the same dataset.
+Therefore, model overfitting is less of a concern in this case as long as all candidate models have equal opportunity to be slightly overfit.
+Nonetheless, we did ensure that the models would not be substantially overfit in two ways:
+The posts used to build the co-occurrence matrices were never included as posts to calibrate weights of model terms.
+This is particularly important since the context term can easily latch on to a post with exactly the same co-occurrence structure as a post that was used to build the underlying matrix.
+Second, for each dataset sample, the logistic regression model that calibrated the model terms was fed with a large number of posts (500).
+Each model only had a few terms to calibrate (e.g., prior, context component), and there were on average 1,300 tags per sample that could be used by the logistic regression model for calibration, so the risk of overfitting was low.
+Further, \textcite{Stanley2013} showed that the out-of-sample model and within-sample model accuracy difference for a similar model on a similarly sized dataset for StackOverflow is small (2\% accuracy difference).
+We were therefore more concerned with the risk of model overfitting by using fewer and more similar dataset samples than from optimizing the weights of a few model terms when those weights were calibrated from thousands of data points.
+This is why the available computational resources were channeled to evaluate the models on multiple dataset samples (particularly the 60 different dataset samples used for Twitter) rather than to perform a k-fold cross validation technique for the weights from the logistic regression.
+
 \subsection{Stop Words}

### 2016-06-13 writeups/psycMethods/psycMethods.tex
foo

@@
 We were therefore more concerned with the risk of model overfitting by using fewer and more similar dataset samples than from optimizing the weights of a few model terms when those weights were calibrated from thousands of data points.
-This is why the available computational resources were channeled to evaluate the models on multiple dataset samples (particularly the 60 different dataset samples used for Twitter) rather than to perform a k-fold cross validation technique for the weights from the logistic regression.
+This is why the available computational resources were channeled to evaluate the models on multiple dataset samples (particularly the 60 different dataset samples used for Twitter) rather than to perform a k-fold cross validation or out-of-sample validation technique for the weights from the logistic regression.
 

### 2016-06-16 writeups/psycMethodsResponse2/response.tex
foo

@@
 
-Please see below our responses to the reviewer comments.
+Please see below our responses to the comments.
 Along with pasting text from the revised document inline with the questions where applicable,
@@
 
-We hope that you and the reviewers find these changes sufficient and that you find the revised document well suited for this special issue. \bigskip
+We hope that you find these changes sufficient and that you find the revised document well suited for this special issue. \bigskip
 
@@
 
-\question Primary concern: Papers on predictive modeling typically avoid confidence intervals and instead seek to optimize out-of-sample RMSE or R2 based on training vs.
-test samples (e.g., k-fold cross-validation) or sometimes by resampling (bootstrapping); see overviews of predictive modeling such as http://www-bcf.usc.edu/~gareth/ISL/ and http://appliedpredictivemodeling.com/.
+\question Primary concern: Papers on predictive modeling typically avoid confidence intervals and instead seek to optimize out-of-sample RMSE or R2 based on training versus test samples (e.g., k-fold cross-validation) or sometimes by resampling (bootstrapping); see overviews of predictive modeling such as http://www-bcf.usc.edu/~gareth/ISL/ and http://appliedpredictivemodeling.com/.
 The general concern here is that people reading your paper alongside other 'big data' papers may view your confidence interval approach unusual or outdated; the more specific concern here is overfitting the model to the data (making confidence intervals too confident).
@@
 \begin{solution}
-No need to comment directly.
+Added a paragraph in the Overall Method section to address this concern.
+The text is included below:
+
+FIXME: add text
+
 \end{solution}
@@
 
-fixme: add text
+FIXME: add text
+
 \end{solution}
@@
 \begin{solution}
-Addressed this by including in the paper the work that looked at model accuracy vs. corpus size and number of rows in the random permutation model.
+Addressed this by including in the paper the work that looked at model accuracy versus corpus size and number of rows in the random permutation model.
 The text for the section has been previously included in this document.
@@
 
-``These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
-However, both models are not 100 percent accurate, and the amount of variance left to predict is larger than what would be observed if retrieval noise is included in the model.
-One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval.
-For example, it may be the case that the declarative system is used to provide a set of likely tag candidates to higher-level processes to consider for selection.''
+FIXME: add text
 

### 2016-06-16 writeups/psycMethods/psycMethods.tex
foo

@@
 and this research focused on measuring differences in model accuracy for two declarative memory models.
-Nonetheless many of these higher-level features may indirectly influence a declarative memory model by influencing the specific tags that a user has chosen in the past,
+Nonetheless, many of these higher-level features may indirectly influence a declarative memory model by influencing the specific tags that a user has chosen in the past,
 which in turn boosts base-level activation for these previously-chosen tags, which in turn makes it more likely that they are chosen in the future.

### 2016-06-16 writeups/psycMethods/psycMethods.tex
foo

@@
 We were therefore more concerned with the risk of model overfitting by using fewer and more similar dataset samples than from optimizing the weights of a few model terms when those weights were calibrated from thousands of data points.
-This is why the available computational resources were channeled to evaluate the models on multiple dataset samples (particularly the 60 different dataset samples used for Twitter) rather than to perform a k-fold cross validation or out-of-sample validation technique for the weights from the logistic regression.
+This is why we focused on evaluating the models on multiple dataset samples (particularly the 60 different dataset samples used for Twitter) rather than to perform a k-fold cross validation or out-of-sample validation technique for the weights from the logistic regression.
 

### 2016-06-16 dissProject/preProcess.R
foo

@@ -1510,7 +1510,7 @@ plotBarSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol='dsetG
 	sumTbl
 }
 
-plotLineSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol, title='Model Name') {
+plotLineSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol, title='Model Name', remapSumTbl) {
 	groupCol = as.symbol(groupCol)
 	fillCol = substitute(fillCol)
 	renameCols(sumTbl)
@@ -1522,6 +1522,7 @@ plotLineSumTbl <- function(sumTbl, fillCol, figName, extras=NULL, groupCol, titl
 	lineTypes
 	shapeTypes = c(15, 16, 17, 18, 12)
 	shapeTypes = shapeTypes[1:length(sumTbl[, unique(DVName)])]
+	sumTbl = remapSumTbl(sumTbl)
 	expr = bquote(ggplot(sumTbl, aes(x=factor(.(groupCol)), y=meanVal)) +
 		      geom_line(aes(group=.(fillCol), linetype=.(fillCol), colour=.(fillCol)), size=.6) + 
 		      geom_point(aes(shape=.(fillCol)), size=3) +
@@ -1553,7 +1554,7 @@ compareMeanDV <- function(modelVsPredTbl, DV, extras=NULL, figName='', groupCol=
 				    ), extras))
 }
 
-compareMeanDVBySize <- function(modelVsPredTbl, DV, extras=NULL, figName='', groupCol, CIFun=CIMean) {
+compareMeanDVBySize <- function(modelVsPredTbl, DV, extras=NULL, figName='', groupCol, CIFun=CIMean, remapSumTbl=identity) {
 	DV = substitute(DV)
 	expr = bquote(modelVsPredTbl[, withCI(.(DV), CIFun=CIFun), keyby=list(DVName, .(as.symbol(groupCol)))])
 	sumTbl = eval(expr)
@@ -1562,7 +1563,8 @@ compareMeanDVBySize <- function(modelVsPredTbl, DV, extras=NULL, figName='', gro
 		       extras=c(list(theme(legend.position='top', legend.direction='vertical'),
 				     ylab('Proportion Correct'),
 				     xlab('Corpus Size')),
-				extras))
+				extras),
+		       remapSumTbl=remapSumTbl)
 }
 
[…]

### 2016-06-16 writeups/psycMethods/psycMethods.tex
foo

@@
 A low value (\num{100}), a more standard value (\num{2048}), and a high value (\num{10000}) of rows were tested.
+The \num{2048}-row matrix was the default size used for the random permutation model for all other results.
 A separate co-occurrence representation was built for each of these row values.
@@
 Results for using different levels of compression for the random permutation model for the StackOverflow dataset are depicted in Figure \ref{figContextDimSO}.
-In the figure, the model name that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
-This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random permutation model for all other results.
 The plot is also broken out for different size datasets used to build the co-occurrence matrix.

### 2016-06-16 writeups/psycMethodsResponse2/response.tex
foo

@@
 \begin{solution}
-Added a paragraph in the Overall Method section to address this concern.
+Added a paragraph in the overall method section to address this concern.
 The text is included below:
 
-FIXME: add text
+``This research centers around evaluating the differences in model accuracy when various components are added, removed, or altered.
+When evaluating the models, overall model accuracy for a specific model and specific dataset is less important than how that accuracy deviates from a comparative model on the same dataset.
+Therefore, model overfitting is less of a concern in this case as long as all candidate models have equal opportunity to be slightly overfit.
+Nonetheless, we did ensure that the models would not be substantially overfit in two ways: The posts used to build the co-occurrence matrices were never included as posts to calibrate weights of model terms.
+This is particularly important since the context term can easily latch on to a post with exactly the same co-occurrence structure as a post that was used to build the underlying matrix.
+Second, for each dataset sample, the logistic regression model that calibrated the model terms was fed with a large number of posts (500).
+Each model only had a few terms to calibrate (e.g., prior, context component), and there were on average 1,300 tags per sample that could be used by the logistic regression model for calibration, so the risk of overfitting was low.
+Further, Stanley and Byrne (2013) showed that the out-of-sample model and within-sample model accuracy difference for a similar model on a similarly sized dataset for StackOverflow is small (2\% accuracy difference).
+We were therefore more concerned with the risk of model overfitting by using fewer and more similar dataset samples than from optimizing the weights of a few model terms when those weights were calibrated from thousands of data points.
+This is why we focused on evaluating the models on multiple dataset samples (particularly the 60 different dataset samples used for Twitter) rather than to perform a k-fold cross validation or out-of-sample validation technique for the weights from the logistic regression.''
 
@@
 
-FIXME: Add text
+``Another parameter in the random permutation model is the number of rows used to represent all of the different words.
+Each word’s environment vector is represented by two randomly placed ones and two negative ones across the rows.
+All that is required is that the four positions and signs of each of these environment vectors is unique.
+But this means multiple environment vectors can overlap when looking at a single row.
+This is a lossy form of compression, and allows the number of rows to represent the words to be much less than the total number of words.
+As the number of rows decreases, the amount of crosstalk between the environment vectors increases, compression increases, and the amount of information stored decreases.
+As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.
+
+Method.
+Three different row-size values were tested for the random permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
+A low value (100), a more standard value (2048), and a high value (10,000) of rows were tested.
+The 2048-row matrix was the default size used for the random permutation model for all other results.
+A separate co-occurrence representation was built for each of these row values.
+Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
+Accuracy measurements were the same as were used for previous runs of these datasets: mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.
+
+Results.
+Results for using different levels of compression for the random permutation model for the StackOverflow dataset are depicted in Figure 13.
+The plot is also broken out for different size datasets used to build the co-occurrence matrix.
+
+Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy, which shows that compression can influence model accuracy.
+However, 2048 rows is sufficient for these datasets, as increasing the number of rows to almost an order of magnitude higher (10,000) has little to no positive impact on accuracy, even for the highest corpus sizes.
+
+Results are similar for Twitter.
+Figure 14 shows the same negative impact when reducing from 2048 to 100, and little to no impact when increasing from 2048 to 10,000.
+These results are consistent across all corpus sizes used to build the co-occurrence matrix.
+
+Discussion.
+The fact that the random permutation model only needs around 2048 rows to represent all of the words in the Twitter and StackOverflow datasets is impressive.
+Twitter has around 2,383,000 unique words for the 3,000,000 tweets in each of the popular-users datasets.
+StackOverflow has 5,848,000 unique words for the 1,000,000 posts used to generate the co-occurrence matrix.
+This means that only 2048 rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
+This is likely because word use frequency for StackOverflow and Twitter approximately follows Zipf’s law.
+Stanley and Byrne (2013) showed that frequency of tag use for StackOverflow approximately follows a Zipf’s law distribution.
+This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply.
+Since these low-use words have little to no ability to influence predictions, it makes sense that accuracy does not substantially improve by providing additional dimensions for them.``
 
@@
 
-FIXME: add text
+``Other successful models have shown that higher-level processes such as a user’s goals for picking a hashtag, demographics, and specifics about how a hashtag has been used in the past can also influence hashtag use.
+Lin, Margolin, Keegan, Baronchelli, and Lazer (2013) showed that there are at least two distinct use cases for hashtags on Twitter: to mark the content (traditional tagging) and to contribute to a specific community that has created a hashtag.
+Yang, Sun, Zhang, and Mei (2012) found that demographics such as gender, age, and location can influence a user’s selected hashtag.
+Denton, Weston, Paluri, Bourdev, and Fergus (2015) found that the growth and lifetime of a hashtag can be influenced by factors such as how often the hashtag was retweeted and the number of followers for a user that used a particular hashtag.
+These types of goal-related features, demographic information, and higher-level attributes of previous hashtag use were not directly included in the declarative memory models that were tested for this research.
+They were not included because it would be difficult to fit these features into a modeling framework that is constrained to use only declarative memory processes, and this research focused on measuring differences in model accuracy for two declarative memory models.
+Nonetheless, many of these higher-level features may indirectly influence a declarative memory model by influencing the specific tags that a user has chosen in the past, which in turn boosts base-level activation for these previously-chosen tags, which in turn makes it more likely that they are chosen in the future.''
 
@@
 
-FIXME: add text
+``These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
+However, both models are not 100 percent accurate, and the amount of variance left to predict is larger than what would be observed if retrieval noise is included in the model.
+One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval.
+For example, it may be the case that the declarative system is used to provide a set of likely tag candidates to higher-level processes to consider for selection.''
 

### 2016-06-17 writeups/psycMethodsResponse2/response.tex
foo

@@
 This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply.
-Since these low-use words have little to no ability to influence predictions, it makes sense that accuracy does not substantially improve by providing additional dimensions for them.``
+Since these low-use words have little to no ability to influence predictions, it makes sense that accuracy does not substantially improve by providing additional dimensions for them.''
 

### 2016-09-29 writeups/psycMethodsTranslationalAbstract/translationalAbstract.tex
foo

--- 2016-06-16 writeups/psycMethods/psycMethods.tex
@@
-\documentclass[man,floatsintext,donotrepeattitle]{apa6}
-%\documentclass[man,donotrepeattitle]{apa6}
+\documentclass[answers,12pt]{exam}
+\usepackage{xcolor}
+\definecolor{SolutionColor}{rgb}{0.1,0.3,1}
+\usepackage{lipsum}
+\usepackage{enumerate}
+\usepackage{alphalph}
 
-% http://tex.stackexchange.com/questions/279/how-do-i-ensure-that-figures-appear-in-the-section-theyre-associated-with
-\usepackage{placeins}
+\renewcommand{\theenumi}{\AlphAlph{\value{enumi}}}
+\renewcommand{\thequestion}{\AlphAlph{\value{question}}}
+\renewcommand\questionlabel{\llap{\thequestion)}}
 
-\usepackage[nodayofweek]{datetime}% http://ctan.org/pkg/datetime
-\usepackage{xfrac}
-\usepackage{textcomp}
-\usepackage[american]{babel}
-\usepackage[utf8]{inputenc}
-
-% Single space after period (Mike prefers)
-\frenchspacing
+%\pointsinrightmargin
+%\boxedpoints
+\unframedsolutions
+\shadedsolutions
+\definecolor{SolutionColor}{rgb}{0.9,0.9,1}
+\renewcommand{\solutiontitle}{}
 
@@
 
-% Changing horizontal space between subsubsection title and text, without loading titlesec package.
-% Not loading titlesec because it clashes with apa6 package.
-% Took base \@startsection command from apa6 .cls file, and tweaked horizontal spacing. 
-% http://tex.stackexchange.com/questions/22576/redefining-sectioning-commands
-% http://www.tex.ac.uk/cgi-bin/texfaq2html?label=atsigns
-% http://tex.stackexchange.com/questions/60127/subsubsection-description-and-title-not-separated
-\makeatletter
-\renewcommand{\subsubsection}{%
-  \@startsection
-  {subsubsection}%
-  {3}%
-  {\parindent}%
-  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-  {-.4em}%
-  {\normalfont\normalsize\bfseries\addperi}}
-\renewcommand{\paragraph}{%
-  \@startsection
-  {paragraph}%
-  {4}%
-  {\parindent}%
-  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-  {-.4em}%
-  {\normalfont\normalsize\bfseries\itshape\addperi}}
-\makeatother
-
-% ref: http://tex.stackexchange.com/questions/28516/how-to-change-the-title-of-toc
-\addto\captionsamerican{% Replace ``english'' with the language you use
-  \renewcommand{\contentsname}%
-  {Table of Contents}%
-}
-
-%\usepackage[compact]{titlesec}
-\usepackage{tabu}
-\usepackage{amssymb,amsmath}
-\usepackage{setspace}
-
-% ref: http://tex.stackexchange.com/questions/48509/insert-list-of-figures-in-the-table-of-contents
-\usepackage[nottoc,numbib]{tocbibind}
-
-\usepackage[hyphens]{url}
-
-% ref: http://tex.stackexchange.com/questions/73862/how-can-i-make-a-clickable-table-of-contents
-\usepackage{hyperref}
-\hypersetup{
-  colorlinks,
-  citecolor=black,
-  filecolor=black,
-  linkcolor=black,
-  urlcolor=black
-}
-%\hypersetup{linktocpage}
-
-%\usepackage{breakurl}
-\usepackage{subfigure}
-\usepackage{graphicx}
-
-\usepackage[group-separator={,}]{siunitx}
-\newcommand{\numNoZero}[1]{{\sisetup{add-integer-zero=false}\num{#1}}}
-
-\newcommand{\myCI}[4]{{mean difference = \num{#1}, 95\% CI [\num{#2}, \num{#3}]}}
-
-\RequirePackage[l2tabu, orthodox]{nag}
-\graphicspath{{./figures/dissertation/}} % Specifies the directory where pictures are stored
-
-\usepackage{csquotes}
-\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber,doi=false,uniquename=false]{biblatex}
-\DeclareLanguageMapping{american}{american-apa}
-
-\addbibresource{bibliography.bib}
-
-% Needed to ensure periods are placed at the end of every bibliography entry in the references section
-\AtEveryBibitem{\clearfield{doi}}
-
-% http://tex.stackexchange.com/questions/55780/disable-month-in-biblatex-bibliography
-\AtEveryBibitem{\clearfield{month}}
-\AtEveryCitekey{\clearfield{month}}
-
-% Make sure only urls that are printed in references section are for webpages
-\AtEveryBibitem{
-  \ifentrytype{misc}{}{
-    \clearfield{url}
-  }
-}
-
-% ref: http://tex.stackexchange.com/questions/128592/remove-backslashes-from-url-fields-in-bib-entry
-\DeclareSourcemap{
-  \maps[datatype=bibtex]{
-    \map{
-      \step[fieldsource=url,
-	match=\regexp{\\},
-      replace=\regexp{}]
-    }
-  }
-}
-
-% Do not let biblatex change any of the casing in the bibliography file
-\DeclareRobustCommand{\MakeSentenceCase*}[1]{{#1}}
-
-% Customer citation for a parencite style but with a few leading words in the parentheses
-% Stolen from parencite but with the out parentheses stripped off
-% For example: (e.g., Anderson, 1972), (see Salton, 2014)
-\DeclareCiteCommand{\noparencite}
-  {\usebibmacro{prenote}}
-  {\usebibmacro{citeindex}%
-   \usebibmacro{cite}}
-  {\multicitedelim}
-  {\usebibmacro{postnote}}
-
-% If the toc is too detailed, try limiting the depth of printed sections in the TOC:
-\setcounter{tocdepth}{2}
-
-\title{Comparing vector-based and Bayesian memory models using large-scale datasets:\protect\\User-generated hashtag and tag prediction on Twitter and StackOverflow}
-\shorttitle{Comparing memory models of tag retrieval}
-%\shorttitle{}
-
-\author{Clayton Stanley and Michael D. Byrne}
-\affiliation{Rice University}
-
-\leftheader{Beitzel}
-
-\keywords{ACT-R declarative memory theory, vector-based models, LSA, machine learning}
-
 \begin{document}
 
-\abstract{
+\begin{questions}
+
+\question
+
  The growth of social media and user-created content on online sites provides unique opportunities to study models of human declarative memory.
@@
   and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
-}
 
-%\note{\today}
+\begin{solution}
 
-\maketitle
+The growth of social media and user-created content on online sites provides unique opportunities to study models of long-term memory.
+By framing the task of choosing a hashtag for a tweet and tagging a post on StackOverflow as a long-term memory retrieval problem,
+two long-term memory models were tested on millions of posts and tweets and evaluated on how accurately they predict a user's chosen tags.
+An uncompressed and compressed model (in terms of storage of information in long-term memory) were tested on the large datasets.
+The results show that past user behavior of tag use is a strong predictor of future behavior.
+Furthermore, past behavior was successfully incorporated into the compressed model that previously used only context.
+Also, an attentional weight term in the uncompressed model was linked to a natural language processing method used to attenuate common words (e.g., articles and prepositions).
+Word order was not found to be a strong predictor of tag use, and the compressed model performed comparably to the uncompressed model without including word order.
+This shows that the strength of the compressed model is not in the ability to represent word order, but rather in the way in which information is efficiently compressed.
+The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
+and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
 
-%\newpage
+\end{solution}
 
-\section{Introduction}
-
-With the massive growth of human-created content on social media sites (e.g., Twitter, StackOverflow, Wikipedia) and public access to many of those databases,
-it is now possible to evaluate psychological theories of declarative memory on large-scale real-world tasks where the theory can be tested on hundreds of millions to billions of data points.
-Without using large datasets it is difficult to verify and explore modifications to the declarative memory architecture because the datasets are not large enough or rich enough to rigorously test modifications to the theory.
-
-A few researchers have started to use large-scale databases to evaluate how well declarative memory models scale,
-for both ACT-R Bayesian models \parencites{Fu2007,Pirolli2003,Stanley2013,Douglass2010} and vector-based memory models \parencites{Jones2007,Rutledge2008,Recchia2010,Sahlgren2008}.
-
-This study expands on previous work that has tested retrieval models on large-scale datasets.
-It focuses specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutation vector-based) can be improved by testing these models on two real-world tagging tasks.
-One primary objective is to compare the random permutation vector-based model to ACT-R's Bayesian declarative memory theory when large datasets are used for each model. 
-Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
-For example, modifying a vector-based model to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
-Large-scale datasets were used throughout the model development and evaluation process.
-This provides an ideal testing environment for constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.
-
-As a general methodology for this research, we used a cognitively-constrained exploratory model development process.
-Since we worked with large-scale datasets, we had enough data to explore many different modifications simultaneously without the risk of overfitting.
-However, we were not interested in bringing to bear the entire collection of machine learning algorithms to this prediction task in order to completely optimize a model.
-Rather, we were interested in evaluating how well two cognitive models could explain the results, and then making selective modifications to them to see how performance changes.
-The modifications that we explored were selected such that they influenced model accuracy, could be easily added to (or removed from) the cognitive models without massive modifications, and could be considered cognitively plausible.
-
-\subsection{Research Questions}
-
-The core question motivating this research is:
-What kinds of cognitively-plausible models can best predict the tags that a user on a social media site selects?
-This question can be tested by generating a prediction for the most likely selected tag whenever a user is about to generate a tag, and then comparing the user's chosen tag to the model's prediction.
-The process of suggesting a relevant tag to a user can be thought of as a memory retrieval request for a tag, given prior tag use and current context.
-
-Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag selection \parencite{Efron2010} and StackOverflow tag selection \parencites{Stanley2013}.
-Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
-At least two of the current memory retrieval models are based on a co-occurrence methodology:
-ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.
-
-\subsubsection{StackOverflow and Twitter}
-
-These two memory models were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
-Twitter is a microblogging service where users create 140 character ``tweets'' and broadcast those messages out to the users who follow their posts.
-A user is free to annotate important concepts or words in the tweet by preceding words with ``\#,'' thereby making the word into a hashtag.
-After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
-Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.
-
-The two memory retrieval models were also tested on tag selection for StackOverflow posts.
-StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
-A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
-The tags chosen by a user on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
-Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.
-
-The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and were interested in retrieval models that generalize across tasks.
-However, they are also similar on several accounts:
-Both domains have amassed large amounts of user-created content, users generate tags for posts when creating content on the site (hashtags for Twitter and tags for StackOverflow),
-and the user data from the datasets are publicly available for analysis.
-
-\subsection{Roadmap}
-
-This rest of this document is separated into 5 sections.
-``Theoretical Background'' describes the vector-based and Bayesian declarative memory models that were evaluated in this research,
-shows how these models are related to the other common declarative memory models,
-and identifies examples of similar tagging models that have already been applied to various tag recommendation systems.
-``General Methodology'' presents the overall method for testing the models on the StackOverflow and Twitter datasets.
-``Past User Behavior'' shows the results for the first test of the models, where the model term for past user behavior is examined in isolation, and tags are predicted based purely on a user's past tagging history.
-``Combining Predictors'' adds the context component to the models, and shows how model accuracy is influenced by changing various architectural components of each model.
-The final section provides a general discussion of the overall findings.
-
-\section{Theoretical Background}
-
-\subsection{ACT-R Declarative Memory Theory}
-
-ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
-The declarative memory system is a component of the architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
-At the core of the system is the concept of a ``chunk'': the representation of a piece of knowledge (e.g., a word).
-The equations that make up this system were derived from a rational analysis of declarative memory retrieval \parencite{Anderson1989}.
-That is, given the task of retrieving a chunk of information from declarative memory,
-the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
-what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
-Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of needing retrieval again, given the prior history of exposure to the chunk.
-These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.
-
-\subsubsection{ACT-R DM Model}
-
-A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACTRModel}.
-
-\begin{table}[!ht]
-  \caption{ACT-R declarative memory model}
-  \label{tabACTRModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Activation &	 	$A_{i} = B_{i} + \sum_{j \in c}^{} W_{j} S_{ji}$ \\
-      Attentional Weight &	$W_{j} = \frac{W}{n}$ \\
-      Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
-      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
-      Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base-level activation ($B_{i}$) and strength of association ($S_{ji}$).
-The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
-Each term in Table \ref{tabACTRModel} will be discussed in greater detail to follow.
-
-\subsubsection{Base-Level Activation}
-
-Base-level activation reflects the log prior odds of needing an observed chunk again.
-The default way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
-This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
-Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
-However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
-In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its frequency and recency of use.
-
-\subsubsection{Strength of Association}
-
-Strength of association ($S_{ji}$) reflects the amount of log-odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
-Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
-Context for the Twitter domain are the words in a tweet.
-Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
-the likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
-For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
-So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
-This assumption was described when deriving the $S_{ji}$ equation \parencite{Anderson1989}, and used in recent research working with large-scale datasets \parencites{Stanley2013,Farahat2004,Douglass2010}.
-
-\subsubsection{Connection to Pointwise Mutual Information}
-
-Pointwise Mutual Information (PMI) is another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
-It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
-The index has been used to accurately and efficiently measure strength of association information for large-scale datasets \parencite{Budiu2007,Farahat2004}.
-The PMI index is presented as Equation \eqref{eqPMI}.
-
-\begin{equation}
-  \label{eqPMI}
-  \mathit{PMI}(y,x) = log \frac{p(y|x)}{p(y)}
-\end{equation}
-
-Further, this PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
-Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.
-
-\subsubsection{Attentional Weight}
-
-The attentional weight term allows for some contextual components to be weighted more than others when the total $S_{ji}$ activation is computed.
-This reflects the fact that the declarative memory system has limited attentional resources ($W$) that must be spread across all of the items in context.
-Most ACT-R models simply weight each contextual element equally (e.g., all words in a StackOverflow post have equal weight), so each word's attentional weight is simply normalized by the number of items in context.
-
-\subsection{Vector-Based Memory Systems}
-
-``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems are an alternative representation of declarative memory.
-They have been successfully used to predict memory retrievals that leverage word order information \parencite{Jones2007}.
-They also have similar structure to memory models that use Latent Semantic Analysis \parencite{Landauer1997} (LSA).
-
-Vector-based models represent a concept (i.e., a chunk) as a vector.
-The representation of the concept is distributed across all elements of the vector.
-This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and viewing the distribution of co-occurrence counts across the rows in the column as the representation of that concept.
-
-For vector-based systems, each word is represented by an environment vector $e_{i}$ that is nearly orthogonal to all other words' environment vectors.
-The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix (typically around 2,000 - 4,000), which is how vector-based systems represent information in a much more compact space.
-Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occurred with that $e_{i}$ environment vector.
-These memory vectors can contain environment vectors from position-independent word co-occurrence information ($c_{i}$) and word order information ($o_{i}$), which can be combined into a single representation ($m_{i}$).
-Over time, memory vectors accumulate a distributed representation of the most common environment vectors that co-occurred with them.
-
-\subsubsection{Retrieval Process}
-
-Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
-Decoding takes the memory vector for context and decodes it back into an environment vector.
-The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
-
-Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
-For example, assume that a user has written the word ``zend'' in a post, and she wants to assign a tag for the post.
-``zend'' has co-occurred with the tag \emph{PHP} many times previously, so the unordered memory vector for ``PHP'' ($c_{\mathit{PHP}}$) contains the unordered environment vector for \emph{zend} ($e_{zend}$).
-To determine the most correlated memory vector with context, a memory vector is created from the current context ($e_{zend}$).
-The cosine between that context memory vector and the memory vector for \emph{PHP} will be high, since the \emph{PHP} memory vector ($c_{\mathit{PHP}}$) often contains the context environment vector ($e_{zend}$).
-Since the cosine is high, the tag \emph{PHP} will most likely be returned as the vector that resonates highest with the current context.
-
-The resonance and decoding processes are inversions of each other.
-Usually both resonance and decoding will return similar rank orderings of most likely chunks given context.
-
-\subsubsection{Word Order for Vector-Based Systems}
-
-One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
-With vector-based systems, order information is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
-This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
-Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007}'s BEAGLE model to implement these operations.
-This allows for retrieval requests such as ``playing \#$\Phi$,'' ``just read \#$\Phi$,'' and ``vector-based $\Phi$ systems.''
-
-\subsubsection{Random Permutation Model}
-
-In order to represent word order information, a function must be used that converts an environment vector for a set of words and positions (e.g., ``stack $\Phi$'') into a new uncorrelated environment vector.
-\textcite{Sahlgren2008} used a simple random permutation method for this operation.
-
-A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
-With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few one and negative one values in random locations. 
-In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is shifted to the left one position.
-This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.
-
-\begin{table}[!ht]
-  \caption{Random permutation model}
-  \label{tabRandPermModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Activation &		$A_{i} = r(m_{C},m_{i})$ \\
-      Memory Vector &		$m_{i} = \sum_{i \in all past} c_{i} + \sum_{i \in all past} \sum_{l \in locations} o_{i,l}$ \\
-      Unordered Context &	$c_{i} = e_{i}$ \\
-      Ordered Context &		$o_{i,l} = e_{i^{-l}}$ \\
-      Context Memory Vector &	$m_{C} = \sum_{i \in C} c_{i} + \sum_{i \in C} \sum_{l \in locations} o_{i,l}$ \\
-      Environment Vector & 	$e_{i} = rand$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-\subsubsection{Connection to LSA}
-
-Both vector-based models and LSA calculate word similarity on a reduced co-occurrence matrix.
-For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors is much less than the number of words.
-LSA uses Singular Value Decomposition (SVD) to find words that have highly correlated document occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
-For vector-based models, each tag (document) is represented by a compressed vector with a length much less than the number of words.
-A vector-based approach like random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
-Nonetheless, vector-based models have been shown to perform similar to if not better than LSA \parencites{Sahlgren2008,Jones2007}.
-
-\subsection{Comparison of Vector-Based Models and ACT-R}
-
-Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
-Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
-Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
-For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
-\citeauthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect \parencite{Anderson1974}.
-
-\subsubsection{Base-Level Activation}
-
-Vector-based models and ACT-R's Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the relative benefits of the two models. 
-In particular, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
-
-The frequency and recency of how content has been used in the past can be a strong predictor of the future importance of that information. 
-The recent content in tweets on a news site like Twitter should be more relevant, for example.
-\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
-They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.
-
-ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
-It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
-Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
-ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
-Vector-based models have focused on computing the associative strength of a chunk and specific contextual elements.
-It is unexplored how accuracy changes for these models once the information about a user's prior behavior (e.g., ACT-R's base-level activation component) is added.
-
-\subsubsection{Word Order}
-
-One strength of a vector-based approach over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
-With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
-To represent word order in a Bayesian model, an additional dimension could be added to the co-occurrence matrix that represents word position.
-However, adding word-order information to ACT-R's default Bayesian declarative memory model is not common.
-Instead, when word order information is used to model reading for example, production rules are defined so that words are read from left to right (e.g., \noparencite{Lewis2006}).
-
-\subsubsection{Activation Calculation}
-
-Vector-based retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
-Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
-These two computations are certainly not equivalent, even though they are most likely correlated with one another.
-
-Using a correlation can be problematic due to range effects in observed co-occurrences.
-If there are particular words for a tag that co-occur much more often than all of the other words with that tag, that single large value diminishes the power of all other words to contribute to the correlation calculation.
-This does not happen with Bayesian retrieval models because each pair of word and tag values is normalized by the total number of words that appeared with that tag and total number of tags that appeared with that word
-(see the strength of association equation in Table \ref{tabACTRModel}).
-The high-frequency words are often removed from the co-occurrence matrix for vector-based models to reduce this effect.
-
-\subsection{Hashtag Prediction}
-
-Models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites,
-though most of these models are not based on theories of human declarative memory.
-Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
-A tag-suggestion tool has also been deployed for the \url{tex.stackexchange.org} and \url{superuser.com} StackExchange sites.
-Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
-Recommendation models for StackOverflow and Twitter will be described in more detail to follow.
-
-\subsubsection{StackOverflow}
-
-\textcite{Stanley2013} used a declarative memory model to predict a user's chosen tags for a post on the StackOverflow site.
-They modified the standard ACT-R model in order to more accurately predict tag use on StackOverflow.
-A formal description of the modified ACT-R Bayesian model is included in Table \ref{tabModACTRModel}.
-
-\begin{table}[!ht]
-  \caption{\citeauthor{Stanley2013}'s StackOverflow tag prediction model}
-  \label{tabModACTRModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} - O$ \\
-      Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
-      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
-      Attentional Weight & 	$W_{j} = W \frac{E_{j}} {\sum_{}^{} {E_{j}}} $ \\
-      Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
-      Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-The activation of each tag ($A_{i}$) is a function of four components:
-That tag's base-level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
-This model has been modified from the standard ACT-R model in four ways:
-Two separate contextual components were used (words in title and body of post), an offset term was added, an entropy measure was used to compute the attentional weight ($W_{j}$),
-and the base-level activation of a tag was based on global past tag use and not customized to a user's past tagging history.
-
-The offset term was added to normalize the mean activation across retrievals.
-This allowed the use of a logistic regression statistical technique to calibrate the weights for each of the terms.
-The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation \parencite{Dumais1991} and is parameter-free.
-
-\subsubsection{Twitter}
-
-Several hashtag recommendation models for Twitter have been developed recently.
-One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a tweet as a function of prior hashtag use and context \parencite{Mazzia2009}.
-This is quite similar to an ACT-R declarative retrieval model.
-The main potential difference is that---like \textcite{Stanley2013}---the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
-
-Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used in similar tweets.
-\textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.
-
-Other successful models have shown that higher-level processes such as a user's goals for picking a hashtag, demographics, and specifics about how a hashtag has been used in the past can also influence hashtag use.
-\textcite{lin2013bigbirds} showed that there are at least two distinct use cases for hashtags on Twitter: to mark the content (traditional tagging) and to contribute to a specific community that has created a hashtag.
-\textcite{yang2012we} found that demographics such as gender, age, and location can influence a user's selected hashtag.
-\textcite{denton2015user} found that the growth and lifetime of a hashtag can be influenced by factors such as how often the hashtag was retweeted and the number of followers for a user that used a particular hashtag.
-These types of goal-related features, demographic information, and higher-level attributes of previous hashtag use were not directly included in the declarative memory models that were tested for this research.
-They were not included because it would be difficult to fit these features into a modeling framework that is constrained to use only declarative memory processes,
-and this research focused on measuring differences in model accuracy for two declarative memory models.
-Nonetheless, many of these higher-level features may indirectly influence a declarative memory model by influencing the specific tags that a user has chosen in the past,
-which in turn boosts base-level activation for these previously-chosen tags, which in turn makes it more likely that they are chosen in the future.
-
-There has been a mix of co-occurrence models created to predict hashtag use on Twitter, and the results are encouraging.
-However, there has been little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
-\textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
-The ACT-R declarative memory retrieval theory shows how past experience and current context combine to produce the most likely retrieval.
-Each component is summed together to generate a total activation, and then the likelihood of chunks are ranked by that summed activation.
-This technique is compensatory and allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
-
-\section{General Methodology}
-
-Different subsets of the StackOverflow and Twitter datasets were used to evaluate past user behavior in isolation and to test the context component of each memory model.
-These datasets will be named and described in general terms.
-Also, the general methods used to tokenize and stem the words in all of the datasets will be described.
-All software code created for this research is publicly available on GitHub \parencite{StanleyRepo2014}.
-
-\subsection{Models}
-
-Two models were evaluated for hashtag prediction on both the Twitter and StackOverflow datasets:
-the random permutation vector-based model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} and 
-the modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel}.
-
-\subsection{Datasets}
-
-The most recent quarterly published StackOverflow dataset \parencite{DataDump2014} was used to test the models on user chosen tags for posts.
-This dataset contains around \num{5.7} million posts and \num{34377} unique tags.
-
-Unlike StackOverflow, the entire Twitter dataset is not released to the public.
-Due to this and the scale of Twitter, only subsets of the Twitter dataset were examined.
-The relevant subsets are described below.
-
-\subsubsection{Popular Users Dataset}
-
-A popular-users dataset was collected for both StackOverflow and Twitter.
-This dataset consists of all of the tweets and posts from various samples of the top users for each site.
-This dataset was collected by fetching all past tagging history for each of the top users.
-It was used to explore the growth and decay characteristics of a hashtag for individual users.
-The temporal dynamics of ACT-R's decay rate model for a chunk were evaluated against these data.
-Multiple samples were taken across a broad range of user types for each dataset to ensure that the model results were not dependent on specific attributes in a single sample.
-
-For StackOverflow, top users were identified by their reputation score on the site (based on their upvoted questions and answers) and the total number of questions asked on the site.
-To create the StackOverflow popular-users dataset, a total of 12 slices of users were taken at different levels of reputation and total number of questions asked.
-500 users were sampled that had a reputation higher than \num{500}, \num{1000}, \num{5000}, \num{10000}, \num{50000}, and \num{100000}.
-100 users were sampled that had asked a total number of questions higher than \num{50}, \num{100}, \num{200}, \num{300}, \num{400}, and \num{500}.
-
-For Twitter, top users were those that had the largest number of followers and had authored the largest number of tweets.
-To create the Twitter popular-users dataset, 12 slices were also taken, this time at different levels of number of followers and total number of tweets.
-100 users were sampled that had a total number of followers higher than \num{1000}, \num{5000}, \num{10000}, \num{100000}, \num{1000000}, and \num{10000000},
-and a total number of tweets higher than \num{100}, \num{500}, \num{1000}, \num{5000}, \num{10000}, and \num{50000}.
-
-\subsubsection{Popular Hashtags Dataset}
-
-An additional popular-hashtags dataset was collected for Twitter.
-This dataset was used to test the context component of the models for Twitter.
-The dataset contains approximately three million tweets where at least one of the 400 most popular hashtags are used.
-A quota of three million tweets was selected because previously research testing Bayesian co-occurrence models on large-scale datasets has used on the order of one million documents
-\parencites{Stanley2013,Douglass2010,Budiu2007}.
-Tweets were collected in real time until the tweet quota was reached.
-
-Four popular-hashtags datasets were collected at four different time periods, spanning one month.
-Multiple datasets were collected to ensure that the results were not dependent on a particular sample of hashtags and would generalize across different popular-hashtag datasets and different time periods.
-Each dataset required around four days to collect enough tweets to reach the quota, and one collection was done each week.
-
-\subsubsection{Randomly Sampled Dataset}
-
-Since the entire dataset is available for StackOverflow, randomly-sampled posts were used to test the context models rather than creating a popular-tags dataset (analogous to Twitter).
-If all Twitter data were available, then the models would have been tested on a randomly-sampled dataset for Twitter as well.
-However, since this is not the case for Twitter, the popular-hashtags dataset was used instead of a randomly-sampled dataset for that site,
-while models for StackOverflow were tested on the more challenging dataset where posts were randomly sampled across all posts created on the site.
-
-\subsection{Acquiring Datasets}
-
-\subsubsection{StackOverflow}
-
-The entire StackOverflow dataset is packaged and released quarterly to the public,
-so the most recent packaged dataset was downloaded and the tables were imported into a relational database.
-
-\subsubsection{Twitter}
-
-Twitter provides two types of APIs for acquiring data: one based on streaming and a direct query-based interface.
-The query-based interface was used to collect all tweets for specific popular users.
-At most \num{3000} of the most recent tweets were collected for each user, since Twitter only allows access to this number of tweets for each user.
-
-The streaming interface was used to collect a sample of all of the tweets that included one of the top 400 hashtags at that time.
-Twitter's streaming API only allows at most 400 hashtags to be monitored at a time.
-Otherwise a larger number of hashtags would have been collected.
-Twitter limits the streaming interface to randomly sample about 1\% of all of the real-time data that is generated on Twitter.
-This sampling rate was adequate to collect around three million tweets within a few days.
-
-\subsection{Tokenization}
-
-A custom tokenization algorithm created by \textcite{Owoputi2013} was used to parse the words in the text.
-This algorithm was specifically created for tokenizing Twitter content.
-However, during testing it was also found to be robust and well-suited for StackOverflow content.
-The initial model results with the new tokenizer were qualitatively similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
-The same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter to make comparisons across datasets more equitable.
-
-\subsection{Stemming}
-
-\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing ``ed'' and ``ing'' endings).
-With the StackOverflow dataset however, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
-This can be thought of as semantic stemming, where two words with identical meaning are converted to the same canonical word.
-So tags used for the post were converted to their root tag when possible.
-This does make the decision process easier for the StackOverflow dataset.
-However, we chose to include this to help ensure that model predictions would stabilize.
-This provided the best opportunity to reliably observe differences in accuracy when altering the models.
-
-\subsection{Testing for Reliable Differences}
-
-In similar spirit to \textcites{masson2003using, tryon2001evaluating, cumming2005inference}, confidence intervals were used throughout for statistical inference instead of explicit $p$-values.
-This was done because it is a more straightforward way to visualize all of the mean differences in accuracy (i.e., bar and line plots with a visual indication of the error) for each of the models, which eases interpretation.
-It is also a conservative estimate of reliability, as using confidence intervals on the means essentially ignores the fact that this is a within-subjects design, where accuracy for each model is measured for each subject in each dataset.
-However, we would rather err on the side of making the results easier to interpret at the cost of using conservative estimates for reliability
-(i.e., use error bars overlaid on a visualization of the means instead of statistical tests on every combination of difference scores).
-In fact, using large datasets for analysis is what empowers us to do this:
-With this much data, we can use straightforward (although more conservative) inferentials and still see reliable results.
-This allows us to concentrate more on proper data visualization of the main findings.
-
-Nonetheless, when accuracy differences between two models are called out explicitly in the text, the confidence interval for the mean difference in accuracy is also included.
-Since this interval is drawn across the observed difference scores, it will be smaller than the intervals included in the plots, since these are drawn independently for each model's overall accuracy.
-
-\section{Past User Behavior}
-
-How does a user's prior tag use influence their likelihood of choosing that tag in the future?
-This question was tested by examining how well ACT-R's base-level activation component of declarative memory predicts the tags that authors choose, given each author's past tagging behavior.
-The model predicts that the more recently and frequently used tags have higher prior activations, and are more likely to be needed in the future.
-
-\subsection{Base-level ACT-R model}
-
-The base-level activation component from the full ACT-R declarative memory model is included in Table \ref{tabACTRBLLModel}.
-
-\begin{table}[!ht]
-  \caption{Base-level component of ACT-R declarative memory}
-  \label{tabACTRBLLModel}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{ll}
-      \hline
-      Common Name &  Equation \\
-      \hline
-      Base-Level Activation &	$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
-      Optimized Learning &	$B_{i} = log \frac{n}{1-d} - d * log L$ \\
-      Hybrid &			$B_{i} = log \big[ \sum_{j=1}^{k} {t_{j}}^{-d} + \frac{(n-k)(t_{n}^{1-d}-t_{k}^{1-d})}{(1-d)(t_{n}-t_{k})} \big]$ \\
-      \hline
-    \end{tabu}
-  }
-\end{table}
-
-All three terms are methods for computing the base-level activation ($B_{i}$) of a chunk in declarative memory.
-The standard form has a time dependence and accounts for the fact that the rate of presentations for a given object (e.g., the word ``loop'') can change over time.
-The optimized learning form assumes that this rate is constant.
-The hybrid form is a blend of the standard and optimized form where activation for the most recent $k$ observations are computed using the standard form and the rest use the optimized form.
-
-For all equations, $i$ represents each chunk (e.g., the word ``loop''), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``loop'' appears), and $n$ is the number of times each chunk has been observed.
-Since the optimized learning form assumes a constant presentation rate for each chunk,
-it only has to keep track of the time since the first presentation of a chunk ($L$) and the total number of instances observed for that chunk ($n$). 
-It does not have to keep track of the time each specific chunk instance was observed.
-Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form.
-The hybrid form only has to keep track of the most recent $k$ presentations of each chunk, so it only pays a small computational penalty over the optimized learning form.
-The optimized learning form is set to the default in the ACT-R Common Lisp implementation.
-
-The decay rate parameter ($d$) represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
-Higher values mean that chunk activation decays faster, which means that the most recently presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
-Lower decay rate values flip the importance of frequency and recency.
-
-The base-level ($B_{i}$) component in Table \ref{tabACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{tabModACTRModel}).
-\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log-odds computation of the probability that each chunk is likely to be presented again, given past frequency.
-If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log \left ( n_{i} \right )$,
-which is a nonlinear transformation of the $B_{i}$ used by \citeauthor{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
-Both forms maintain the same rank ordering, so they should produce similar results.
-The $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).
-
-\subsection{Method}
-
-The StackOverflow and Twitter popular-users datasets were used to characterize tag growth and decay for a user on each site.
-Each user's tag behavior from each site was analyzed.
-
-22 different values of the decay rate parameter were explored
-(0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.8, 2, 5, 10, 20).
-The values were chosen to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d \approx .5$), to a pure recency model ($d=20$).
-The decay rate value was varied for the standard form, the optimized learning form, and the hybrid form of $B_{i}$. 
-Only values below 1 were examined for the optimized and hybrid forms, since values above 1 for these forms are not mathematically defined. 
-They are not mathematically defined at a decay rate greater than 1 because for these values the $(1-d)$ denominator flips the ratios to negative values (see Table \ref{tabACTRBLLModel}).
-At a decay rate equal to 1, every base-level activation $B_{i}$ is infinite, which is not psychologically interesting.
-
-All three forms of $B_{i}$ were used to predict each user's specific tag history.
-To do this, a user's entire tag history was collected, and then each tag instance from oldest to newest was examined.
-For each of these tagging instances, the model generated an activation value for each possible tag based on that tag's frequency and recency of past use.
-
-At each tagging instance, the model has activation values for a set of tags.
-Model accuracy was assessed by rank ordering these activations and choosing the model's highest $N$ tags, where $N$ is the number of tags used for that specific post.
-The model's chosen $N$ tags were compared to the user's chosen $N$ tags, and if a model's chosen tag was in the set of user's chosen tags, then the prediction was marked as correct.
-Otherwise, it was marked as incorrect.
-
-This process was performed across all users in each of the 12 dataset slices for both the Twitter and StackOverflow popular-users datasets.
-It was repeated for each of the decay rate values and the three different forms of $B_{i}$.
-For the hybrid form, three different values of $k$ were explored (1, 5, 10).
-
-\subsection{Results}
-
-\subsubsection{Model Performance Across Decay Rate Values}
-
-A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for all three forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDsStd} and \ref{figPriorTwitterSliceDsStd} respectively.
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[][Standard form]{\includegraphics[width=5cm]{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
-    \hfill
-    \subfigure[][Optimized learning form]{\includegraphics[width=5cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
-    \hfill
-    \subfigure[][Hybrid form]{\includegraphics[width=5cm]{visNormMean-SOQgt500r2-topHashtagPostPriorHybrid10-crop.pdf}}
-    \caption{
-    Model performance for a single dataset slice for StackOverflow.
-    This includes model accuracy as a function of decay rate value for each user in the dataset slice.
-    Model accuracy is shown for the standard form of base-level activation ($B_{i}$), the optimized learning form, and the hybrid form when k=10.
-  }
-    \label{figPriorSOQSliceDsStd}
-  }%
-
-\end{figure}
-
-\begin{figure}[!htbp]
-  {%
-    \setlength{\fboxsep}{0pt}%
-    \setlength{\fboxrule}{1pt}%
-    \hfill
-    \subfigure[][Standard form]{\includegraphics[width=5cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}
-    \hfill
-    \subfigure[][Optimized learning form]{\includegraphics[width=5cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
-    \hfill
-    \subfigure[][Hybrid form]{\includegraphics[width=5cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorHybrid10-crop.pdf}}
-    \caption{
-      Model performance for a single dataset slice for Twitter.
-      This includes model accuracy as a function of decay rate value for each user in the dataset slice.
-    }
-    \label{figPriorTwitterSliceDsStd}
-  }%
-\end{figure}
-
-Model accuracies at each decay rate value for each user in the dataset slice are included in the plots.
-Each user's mean accuracy was subtracted in order to compute a normalized mean for that user.
-This was done so that the relative accuracy change for different decay rate values could be more easily visualized.
-The plotting technique is a analogous to what is done when testing the main effect in a repeated measures design (i.e., each subject's mean score is subtracted).
-
-One large effect in all of the plots is that there is an optimal decay rate value between 0 and 20 that produces the highest model accuracy.
-For the standard $B_{i}$ equation, model accuracy for a pure frequency model ($d=0$) is also worse than using a pure recency model ($d=20$), particularly for Twitter.
-It makes sense that a pure recency model for Twitter is relatively more accurate than for StackOverflow, as the hashtag lifetime for Twitter is much less than a tag for a programming language on StackOverflow.
-
-When comparing the standard and optimized learning forms of the equation, it is apparent that a pure recency model does not work well for optimized learning.
-The best-fit decay rate value for the optimized learning form is also not as clear and pronounced as it is for the standard form (less of a peak).
-And the relative accuracy gained from using a pure frequency based model to a blend of frequency and recency is less for the optimized learning form than the standard form of $B_{i}$.
-Finally, for the range defined for the hybrid form ($d < 1$), the standard and hybrid forms are shaped similarly across the decay rate values.
-
-\subsubsection{Aggregate Model Performance}
-
-Aggregate best-fit decay rates were computed by taking the median of all user's best-fit decay rate values across all dataset slices.
-The median was used since there were users in the popular-users dataset that did not have enough observations to generate stable predictions, and best-fit decay rate values for these users could be as high as 20.
-Using the median effectively trims these unstable values from the sample, without having to define a cut point for an outlier removal process.
-
-Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
-Decay rate values that produced the most accurate model performance are lower for the optimized learning form (\num{.43}) of $B_{i}$ compared to both the standard form (\num{.80}) and hybrid form (\num{.80} for all $k$).
-The optimal values for optimized learning are near \num{0.5}, which lines up with default value used in ACT-R.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
-  \caption{
-  Overall best-fit decay rate for StackOverflow and Twitter.
-  Results are shown for the two Twitter and StackOverflow (SO) popular-users datasets.
-  Error bars represent the 95\% bootstrapped confidence interval for the median.
-}
-  \label{figPriorDecay}
-\end{figure}
-
-\FloatBarrier
-
-Finally, model accuracy was analyzed across all dataset slices for StackOverflow and Twitter.
-This was done by examining model accuracy at a specific decay rate value for each model.
-Accuracy was assessed at each model's previously-determined best-fit values (\num{.43}, \num{.80}, and \num{.80} for the optimized learning, standard form, and hybrid form) and additionally the standard form's default value (\num{0.5}).
-Since decay rate values were explored at \num{0.1} increments between the 0 and 1 range, the decay rate values used to examine model accuracy were rounded to the nearest decay rate value that was included in the run.
-Consequently, a best-fit value of \num{.4} was used for the optimized learning model instead of the \num{.43} average found from the dataset slices.
-Aggregate accuracy was computed by taking the mean of every user's model accuracy across all users in all dataset slices.
-
-Accuracy for both sites is included in Figure \ref{figPriorAcc}. 
-Accuracy for the standard form is higher than the optimized learning form (\numNoZero{.33} vs. \numNoZero{.28}, \myCI{.034}{.032}{.036}{3603}).
-This makes sense given that the standard form does not assume equal presentation rate for each chunk, and hashtags on Twitter in particular show trends of high rate of use and then shift to low rate of use.
-The standard form of $B_{i}$ is able to more accurately account for these dynamics.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc--crop.pdf}}}
-  \caption{
-    Overall model accuracy for StackOverflow and Twitter at optimal decay rate values.
-    All error bars are the 95\% bootstrapped confidence interval for the mean. 
-  }
-  \label{figPriorAcc}
-\end{figure}
-
-Accuracy for the hybrid form for the largest $k$ ($k=10$) is slightly higher than the smallest $k$ ($k=1$) (\numNoZero{.334} vs. \numNoZero{.329}, \myCI{.0026}{.0023}{.0029}{3608}).
-However accuracy cannot be distinguished between the hybrid form with the largest $k$ and the standard form (\numNoZero{.334} vs. \numNoZero{.334}, \myCI{.0000}{-.0000}{.0001}{3609}).
-This indicates that for the datasets analyzed in this study the hybrid form with $k=10$ is a suitable replacement to the standard form (produces the same results) and uses less computational resources than the standard form.
-
-ACT-R sets the default decay rate value to \num{0.5}, and the best-fit value for the standard form is higher (\num{0.8}) for these datasets.
-The figure shows that although model accuracy does slightly improve (\num{0.325} to \num{0.335}, \myCI{.0029}{.0022}{.0036}{3609}) for the standard form when increasing the decay rate parameter from \num{0.5} to the optimal value,
-that percentage improvement is not as large (\num{0.28} to \num{0.33}) as the improvement gained from using the standard form compared to the optimized learning form.
-Nonetheless, model accuracy is highest when both the standard form is used and the decay rate value is increased.
-
-\subsection{Discussion}
-
-Given the large tag space used by post authors on StackOverflow and Twitter, it was surprising that model performance is respectable even when only past tag history and no context is used.
-This speaks to the importance of taking into account overall past tagging history, and ensuring that the prior component of activation for each tag is customized to each user's specific past tagging history.
-
-Also apparent for these datasets is that the optimized learning form of $B_{i}$ is not as accurate as the standard form.
-Further, for the optimized learning form there is not much difference between using a pure frequency-based model and a model where frequency and recency are optimally blended.
-This makes sense given that the optimized learning equation in Table \ref{tabACTRBLLModel} collapses to a pure frequency model as the amount of time since the presentation of the first recorded chunk instance increases,
-and also since the timespan of recorded tag use for StackOverflow and Twitter is large (years).
-
-However the hybrid form with a reasonable $k$ is just as accurate as the standard form.
-This is due to the fact that the majority of change in activation for a chunk occurs in the short period of time after the presentation of the chunk (a property of a power law distribution).
-Once a chunk's presentations decay to the flat area of the power law distribution, not much change in activation for these presentations occurs, so they can be averaged and their presentation time can be discarded without much loss in fidelity.
-Therefore if computational speed is a concern, one may wish to leverage the hybrid form instead of the standard form.
-Although qualitatively the standard form produces the cleanest and most natural effect of decay rate on accuracy, as shown in Figure \ref{figPriorSOQSliceDsStd}.
-It also does not have a bifurcation at $k=1$.
-Consequently, if one is a purist and computational speed is less of a concern, one may still wish to use the standard form over the hybrid form.
-
-\section{Combining Predictors}
-
-The declarative memory models analyzed for this research consist of two main components: past behavior and context.
-Past tag use was isolated in the first section to see how well ACT-R's prior component of declarative memory (i.e., base-level learning) fits each user's tag use over time.
-For the next section the past behavior and context declarative memory components were combined as well as looked at in isolation.
-
-This section will examine how model accuracy is influenced by five architectural concerns:
-method of handling stop words, influence of dataset size, influence of compression for the random permutation model, method of combining prior and context, and word order.
-These particular architectural concerns were examined for the following reasons:
-During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.
-Due to research from \textcite{Budiu2007} and \textcite{Recchia2010}, model accuracy for the Bayesian and vector-based models is influenced by the number of documents in the dataset,
-so the effect of dataset size on accuracy was examined in more detail.
-\textcite{Sahlgren2008} showed how different levels of compression in the random permutation model influence accuracy, so this was examined for three reasons:
-to replicate the findings, as an additional check to ensure that the random permutation model was implemented correctly, and also to see if the same levels of compression are valid for the much larger datasets used in this research.
-One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random permutation model,
-so two different ways to add the model terms were tested.
-Finally, previous research has shown that including word order in the vector-based models improves performance \parencites{Jones2007, Sahlgren2008},
-so word order was also examined to see if the effect could be replicated and also to measure the overall importance of word order on model accuracy relative to other architectural manipulations.
-
-\subsection{Overall Method}
-
-\subsubsection{Logistic Regression}
-
-In order to best combine these predictors, a logistic regression statistical technique similar to \textcite{Stanley2013} was used to determine the most optimal weights for each term.
-When a user creates a tag for a post, a retrieval request for the model is made.
-For each request, the model returns a set of tags (all tags that have been observed for that user in the past) and activations associated with each tag.
-An activation value is computed for each model component (e.g., prior tag use, context) for each tag.
-These values are recorded, and tag instances that match what the author actually tagged are marked as a 1, while all others are marked with a 0.
-When an author used multiple tags for a post, multiple 1s are marked for that post.
-This process is repeated across all posts where a user chooses a tag, and across all users in the dataset.
-Once all of the results are aggregated, a logistic regression is run where the optimal weights for each model component are found that maximize the model's ability to correctly label tags as 0 or 1 based on total activation.
-
-Model accuracy was evaluated by rank ordering all recorded tags by model activation using the optimal weights,
-asking the model to tag the top N posts with a 1 (where N is the total number of recorded tagging instances in the dataset sample),
-and then comparing the model's chosen tags with the used tags for each post (i.e., looking at the ``hits'').
-This is equivalent to throttling the threshold in the logistic regression such that the total number of observations labeled as a 1 match the total number of recorded instances of tagging in the dataset.
-
-\subsubsection{Adding Offset Term}
-
-This logistic regression method is a computationally efficient way to find the optimal weights for the predictors.
-However, when the regression can not be constrained to label a specific number of 1s for each observation (as is the case here), to be used properly an offset term must be added as an additional model predictor.
-The offset term keeps the activation for the small number of top tags for each post equal across posts.
-This ensures that the logistic regression only labels a few 1s for each post.
-
-A simplified version of \textcite{Stanley2013}'s method was used for this research.
-Instead of subtracting the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
-For the prior component for example, the activation value used in the logistic regression is the prior activation with the mean of the top 5 prior activations subtracted from this value.
-The same process is used for the context component.
-Decoupling the offset term and computing an offset for each model component in isolation means that an iterative approach was no longer necessary,
-since the optimal weights of the terms are no longer needed when computing the offset.
-
-\subsubsection{Datasets}
-
-Model accuracy was tested on the popular-hashtags dataset for Twitter and a random sample of posts across the entire StackOverflow dataset.
-For each of the four Twitter popular-hashtags datasets, model retrievals were done across 15 runs and 500 posts within each run.
-For the StackOverflow randomly-sampled dataset, 5 runs of 500 posts were used,
-since the results for StackOverflow were more stable and retrievals took more wall-clock time for this dataset due to the size of the co-occurrence matrix.
-All runs contained different posts, and all posts used for runs were not used to generate the co-occurrence matrix for the two retrieval models.
-
-\subsubsection{Computing Priors}
-
-For each randomly-sampled post for StackOverflow, the prior activation is based on that user's prior tagging history, and not on any global prior of specific tag frequency across users.
-This was done so that the prior component used for StackOverflow for both the popular-users and randomly-sampled datasets was the same.
-
-However, a custom user prior for the Twitter popular-hashtags dataset could not be used, since the entire Twitter dataset could not be downloaded, so the history of every user in the popular-hashtags dataset could not be collected.
-Instead, the global prior tag history for all users in the Twitter popular-hashtags dataset was used to compute prior activation values for each tag.
-
-\subsubsection{Co-occurrence Matrices}
-
-The context component for both the random permutation and Bayesian model was based on the same co-occurrence matrix of words and observed tags.
-Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
-The matrix for \num{3000000} posts was only used for Twitter, since the number of words in a StackOverflow post is much higher than Twitter on average,
-and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow was too large.
-
-\subsubsection{Sampling and Overfitting}
-
-This research centers around evaluating the differences in model accuracy when various components are added, removed, or altered.
-When evaluating the models, overall model accuracy for a specific model and specific dataset is less important than how that accuracy deviates from a comparative model on the same dataset.
-Therefore, model overfitting is less of a concern in this case as long as all candidate models have equal opportunity to be slightly overfit.
-Nonetheless, we did ensure that the models would not be substantially overfit in two ways:
-The posts used to build the co-occurrence matrices were never included as posts to calibrate weights of model terms.
-This is particularly important since the context term can easily latch on to a post with exactly the same co-occurrence structure as a post that was used to build the underlying matrix.
-Second, for each dataset sample, the logistic regression model that calibrated the model terms was fed with a large number of posts (500).
-Each model only had a few terms to calibrate (e.g., prior, context component), and there were on average 1,300 tags per sample that could be used by the logistic regression model for calibration, so the risk of overfitting was low.
-Further, \textcite{Stanley2013} showed that the out-of-sample model and within-sample model accuracy difference for a similar model on a similarly sized dataset for StackOverflow is small (2\% accuracy difference).
-We were therefore more concerned with the risk of model overfitting by using fewer and more similar dataset samples than from optimizing the weights of a few model terms when those weights were calibrated from thousands of data points.
-This is why we focused on evaluating the models on multiple dataset samples (particularly the 60 different dataset samples used for Twitter) rather than to perform a k-fold cross validation or out-of-sample validation technique for the weights from the logistic regression.
-
-\subsection{Stop Words}
-
-The method of handling stop words has been shown to influence model accuracy \parencite{Sahlgren2008,Stanley2013,Jones2007}, so this was the first architectural concern that was examined.
-``Stop words'' (e.g., ``the,'' ``a,'' ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
-\textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
-A data-driven frequency-filtering approach was tried, along with using a previously-generated stoplist, and no stop word removal at all.
-However, \citeauthor{Sahlgren2008} did not try to weight the stop words and only tried methods to remove them.
-
-\textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset.
-The technique is formally described in Table \ref{tabModACTRModel} and uses information theory to determine the amount of predictive information contained in each word.
-Words that are often used with a small number of tags have high information content while words used often across all tags carry low information content.
-This method maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
-
-\subsubsection{Method}
-
-The runs from the Twitter popular-hashtags and StackOverflow randomly-sampled datasets were used for the analysis.
-Both the random permutation and Bayesian models were tested.
-Activations for context were computed by using five different methods to handle stop words:
-The two removal methods used in \textcite{Sahlgren2008}
-(i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist: \noparencite{Salton1988}),
-the entropy-weighting method used in \textcite{Stanley2013},
-combining the entropy-weighting and frequency-filtering methods,
-and no removal or weighting of stop words.
-
-\paragraph{Determining Frequency Cutoff}
-
-A cutoff point must be identified and used in order to remove the high-frequency words.
-\textcite{Sahlgren2008} used a cutoff where words that occurred more than \num{15000} times were removed.
-Since the StackOverflow and Twitter datasets are much larger than the datasets used by \citeauthor{Sahlgren2008}, the same cutoff could not be used.
-
-The ideal cutoff was determined by choosing the value where the standard deviation of the counts for each word for each of the tags were all relatively small.
-The plots for the standard deviations across tags that were used to identify the cutoff for StackOverflow and Twitter are shown in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT}.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{memMatSO-crop.pdf}}}
-  \caption{
-    Variance in observation counts of words for each tag for StackOverflow.
-    Plots are shown for the four different methods for handling stop words.
-    Each plot contains the standard deviation of the total counts for each word in context as a function of hashtag.
-    High values mean that there are context words for a hashtag that have a high number of counts relative to the counts for all other context words for that hashtag.
-  }
-  \label{figContextCutoffSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{memMatT-crop.pdf}}}
-  \caption{
-    Variance in observation counts of words for each tag for Twitter.
-  }
-  \label{figContextCutoffT}
-\end{figure}
-
-The ``no processing'' plot shows the size of the spikes when high-frequency words are not removed.
-The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used.
-The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
-The cutoff that produced the frequency plots in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
-
-\subsubsection{Results for the Random Permutation Model}
-
-Plots for the three methods of handling stop words for the random permutation model on the StackOverflow dataset are shown in Figure \ref{figContextStopRPSO}.
-In the plot, if the model name includes ``full'' then all model components are included.
-When the model components are enumerated, then at least one of the components is left out.
-For example, the ``standard prior model relaxed across posts'' means that only the prior model component was used when determining activation. 
-As another example, ``RP only title with freq'' means that only the title context component of the random permutation model was used, and that stop words were handled using the frequency filtering technique.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsSOPerm-crop.pdf}}}
-  \caption{
-    Stop-word techniques for the random permutation model for StackOverflow.
-    The five methods for handling stop words are shown both for the title model component in isolation and the full model.
-    ``Stoplist'', ``freq'', ``entropy'' are the stoplist removal method, removal based on frequency count, and weighting based entropy metric.
-    All error bars are the 95\% bootstrapped confidence interval for the mean. 
-}
-  \label{figContextStopRPSO}
-\end{figure}
-
-The results show that model accuracy improves if stop words are handled in any of the three ways (predefined stoplist filtering, entropy weighting, and frequency filtering).
-Using a predefined stoplist makes the smallest improvement (\numNoZero{.45} to \numNoZero{.49}, \myCI{.032}{.018}{.047}{5})
-while the entropy (\numNoZero{.45} to \numNoZero{.60}, \myCI{.143}{.123}{.163}{5}) and frequency (\numNoZero{.45} to \numNoZero{.63}, \myCI{.178}{.164}{.192}{5}) techniques improve performance the most,
-and frequency shows a slight edge over entropy.
-However, using the frequency-filtering technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.
-
-The three stop word techniques were then tested for the random permutation model on the Twitter dataset, and the results are depicted in Figure \ref{figContextStopRPT}.
-For the random permutation model for Twitter, the entropy weighting technique is the best method for handling stop words (\numNoZero{.25} to \numNoZero{.26}, \myCI{.013}{.011}{.015}{60}).
-Removing stop-words based on a predetermined list actually decreases accuracy (\numNoZero{.25} to \numNoZero{.23}, \myCI{-.017}{-.015}{-.019}{60}),
-and using data-driven frequency filtering has only a small to no effect (\numNoZero{.25} to \numNoZero{.25}, \myCI{.00}{-.003}{.002}{60}).
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTPerm-crop.pdf}}}
-  \caption{
-  Stop-word techniques for the random permutation model for Twitter.
-  The five methods for handling stop words are shown both for the context model component (i.e., no prior term) and the full model.
-  All error bars are the 95\% bootstrapped confidence interval for the mean. 
-}
-  \label{figContextStopRPT}
-\end{figure}
-
-\subsubsection{Results for the Bayesian Model}
-
-Since accuracy measurements when using a predetermined list to remove stop words for the random permutation model were much worse than the frequency and entropy methods, this method was not examined for the Bayesian model.
-The results for the frequency and entropy techniques for the Bayesian model on the StackOverflow dataset are depicted in Figure \ref{figContextStopBSO}.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsSOSji-crop.pdf}}}
-  \caption{
-  Stop-word techniques for the Bayesian model for StackOverflow
-  All error bars are the 95\% bootstrapped confidence interval for the mean. 
-}
-  \label{figContextStopBSO}
-\end{figure}
-
-The results show that entropy weighting produces the most accurate results (\numNoZero{.61} to \numNoZero{.63}, \myCI{.024}{.011}{.037}{5}).
-Also, accuracy actually decreases when using frequency filtering compared to no filtering (\numNoZero{.61} to \numNoZero{.56}, \myCI{-.043}{-.030}{-.057}{5}) for the Bayesian model.
-So entropy weighting is the clear winner in this case.
-
-Similar results are found when the Bayesian model is tested on the Twitter dataset.
-Figure \ref{figContextStopBT} shows that entropy weighting is again the best performer (\numNoZero{.29} to \numNoZero{.32}, \myCI{.028}{.024}{.031}{60}),
-and frequency filtering decreases performance relative to no filtering (\numNoZero{.29} to \numNoZero{.28}, \myCI{-.017}{-.015}{-.019}{60}).
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTSji-crop.pdf}}}
-  \caption{
-  Stop-word techniques for the Bayesian model for Twitter
-  All error bars are the 95\% bootstrapped confidence interval for the mean. 
-}
-  \label{figContextStopBT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-One result that should not be overlooked is how poorly using a predetermined stoplist performed when compared to the two data-driven frequency and entropy approaches.
-This is most likely because the StackOverflow and Twitter datasets are large and contain high-frequency words that are domain specific (e.g., emoticons) and are not included in the predetermined list.
-This shows that data-driven techniques to identify stop words can be much more effective than using a predetermined list.
-
-Another interesting result is how the Bayesian model actually performed worse when stop words were removed based on frequency compared to no stop word removal.
-This was not the case for the random permutation model in this study, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random permutation model,
-and is most likely never the case for the random permutation model.
-Model accuracy most likely decreases for the Bayesian model because this model already has a normalization component built in that the random permutation model does not have.
-When computing activation for the random permutation model, the correlation is calculated directly on a matrix of counts of co-occurrences for words and tags.
-For the Bayesian model this count matrix is converted to a log-odds ($S_{ji}$) matrix,
-where both the number of observations for the word and the tag are normalized when computing the value for each cell (see the $S_{ji}$ computation in Table \ref{tabModACTRModel}).
-Therefore, it is likely the case that a form of frequency attenuation is already built into the Bayesian model,
-and adding a frequency filter on top of this does nothing to improve accuracy and only removes predictive information.
-
-\textcite{Stanley2013} found that the entropy-weighting technique to attenuate stop words for the Bayesian model significantly improved model accuracy.
-The same effect is found for the StackOverflow and Twitter datasets in this study.
-Further, when this entropy-weighting technique is compared directly to the more common frequency-filtering technique, weighting by entropy produces more accurate results.
-This was clearly the case for the Bayesian model.
-Both frequency and entropy techniques performed equally for the random permutation model.
-However, since the entropy-weighting technique is parameter-free and does not require a cutoff to tune, it is also a better choice for the random permutation model for these datasets.
-
-\subsection{Corpus Size}
-
-The number of documents used to build the co-occurrence matrix has been shown to influence model accuracy \parencite{Sahlgren2008}.
-The default number of documents used to build the co-occurrence matrix was \num{1000000} for StackOverflow and \num{3000000} for Twitter.
-In order to see how accuracy changed as function of corpus size, 
-co-occurrence matrices for a range of smaller corpus sizes were built.
-
-\subsubsection{Method}
-
-For StackOverflow, the following number of posts were used to build separate co-occurrence matrices: \num{1000}, \num{10000}, \num{100000}, and the default \num{1000000}.
-For Twitter, these matrix sizes were tested alongside the default \num{3000000}.
-All runs from the four Twitter popular-hashtags datasets and the randomly-sampled StackOverflow dataset were used to test model accuracy.
-
-\subsubsection{Results for StackOverflow}
-
-Model accuracy as a function of co-occurrence matrix size for StackOverflow is depicted in Figure \ref{figContextDocumentSizeSO}.
-Accuracy improves for both the Bayesian and random permutation models as the size of the documents used to generate the co-occurrence matrix increases.
-However for the random permutation model, accuracy begins to plateau for the largest corpus size for all three tested versions. 
-The Bayesian model does not plateau nearly as much and accuracy continues to improve even at the largest measured corpus size.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-freqVsEntropyBySizeSO-crop.pdf}}}
-  \caption{
-    Effect of size of co-occurrence matrix for StackOverflow.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean model accuracy across all runs in dataset.
-  }
-  \label{figContextDocumentSizeSO}
-\end{figure}
-
-\subsubsection{Results for Twitter}
-
-The overall trend of increasing accuracy with corpus size is present for Twitter as well.
-Both entropy weighting and frequency filtering in Figure \ref{figContextDocumentSizeT} plateau for the random permutation model with increasing corpus size.
-Once again, the Bayesian model continues to improve with corpus size.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-freqVsEntropyBySizeT-crop.pdf}}}
-  \caption{
-    Effect of size of co-occurrence matrix for Twitter.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean.
-  }
-  \label{figContextDocumentSizeT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-It is interesting to compare model accuracy between the Bayesian and random permutation model at the smallest and largest corpus sizes for both StackOverflow and Twitter.
-At smaller corpus sizes, the random permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference diminishes as corpus size increases.
-In other words, the Bayesian model needs larger co-occurrence matrices than the random permutation model to work properly.
-As the size of the dataset increases, the compression may start to lose crucial information, and the uncompressed Bayesian model should be used.
-As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random permutation model is actually beneficial, and the compressed random permutation model should be used.
-This noise is beneficial because it helps soften spurious $S_{ji}$ values when the counts for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller corpus sizes.
-
-This result has important implications.
-The corpus size is rarely chosen, and more likely constrained by the number of documents in the dataset being studied.
-If that dataset is relatively small, it may be better to use the compressed random permutation model than the Bayesian model, and vice-versa if the dataset is large.
-
-However, corpus size is not the only factor that determines whether the Bayesian model or random permutation model will perform better.
-This is because the number of observations used to generate the co-occurrence matrix is a function of four components:
-corpus size, number of words in each document, total number of unique tags, and number of tags used per document.
-
-The compressed random permutation model should perform better with smaller corpus sizes, less words per document, more unique tags, and fewer number of tags per post.
-However, since there are so many factors that determine whether the Bayesian or random permutation model will perform better,
-and only two different datasets were examined in this research,
-it is difficult to make any specific recommendations as far as exact transition cut points (e.g., corpus size, words per document) where either the Bayesian or random permutation model should be used.
-Also, it is likely that the specific domain being studied interacts with the factors already mentioned to determine which model is better for that domain.
-Nevertheless, if a researcher is working in a domain with limited data (e.g., small corpus size, large tag space) that produces a sparse co-occurrence matrix,
-then it may be worthwhile to experiment with using a compressed random permutation model instead of the full Bayesian model.
-Also, newer models of $S_{ji}$ have been introduced that have a more local impact when a new chunk is observed \parencite{thomson2013constraining}.
-It would also be interesting to see how these local models of $S_{ji}$ perform while varying the number of observed associations.
-
-\subsection{Compression Level for Random Permutation}
-
-Another parameter in the random permutation model is the number of rows used to represent all of the different words.
-Each word's environment vector is represented by two randomly placed ones and two negative ones across the rows.
-All that is required is that the four positions and signs of each of these environment vectors is unique.
-But this means multiple environment vectors can overlap when looking at a single row.
-This is a lossy form of compression, and allows the number of rows to represent the words to be much less than the total number of words.
-As the number of rows decreases, the amount of crosstalk between the environment vectors increases, compression increases, and the amount of information stored decreases.
-As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.
-
-\subsubsection{Method}
-
-Three different row-size values were tested for the random permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
-A low value (\num{100}), a more standard value (\num{2048}), and a high value (\num{10000}) of rows were tested.
-The \num{2048}-row matrix was the default size used for the random permutation model for all other results.
-A separate co-occurrence representation was built for each of these row values.
-Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
-Accuracy measurements were the same as were used for previous runs of these datasets: mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.
-
-\subsubsection{Results}
-
-Results for using different levels of compression for the random permutation model for the StackOverflow dataset are depicted in Figure \ref{figContextDimSO}.
-The plot is also broken out for different size datasets used to build the co-occurrence matrix.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
-  \caption{
-    Effect of compression on random permutation model for StackOverflow.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean.
-  }
-  \label{figContextDimSO}
-\end{figure}
-
-Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy, which shows that compression can influence model accuracy.
-However, \num{2048} rows is sufficient for these datasets, as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy, even for the highest corpus sizes.
-
-Results are similar for Twitter.
-Figure \ref{figContextDimT} shows the same negative impact when reducing from \num{2048} to \num{100}, and little to no impact when increasing from \num{2048} to \num{10000}. 
-These results are consistent across all corpus sizes used to build the co-occurrence matrix.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeT-crop.pdf}}}
-  \caption{
-    Effect of compression on random permutation model for Twitter.
-    Error bars represent the 95\% bootstrapped confidence interval for the mean.
-  }
-  \label{figContextDimT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-The fact that the random permutation model only needs around \num{2048} rows to represent all of the words in the Twitter and StackOverflow datasets is impressive.
-Twitter has around \num{2383000} unique words for the \num{3000000} tweets in each of the popular-users datasets.
-StackOverflow has \num{5848000} unique words for the \num{1000000} posts used to generate the co-occurrence matrix.
-This means that only \num{2048} rows are needed to represent most of the variance from all of the words in these co-occurrence matrices.
-This is likely because word use frequency for StackOverflow and Twitter approximately follows Zipf's law.
-\textcite{Stanley2013} showed that frequency of tag use for StackOverflow approximately follows a Zipf's law distribution.
-This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply.
-Since these low-use words have little to no ability to influence predictions, it makes sense that accuracy does not substantially improve by providing additional dimensions for them.
-
-\subsection{Combining Prior and Context}
-
-The activation value for the context component of the vector-based model is a correlation, ranging from zero to one.
-This is not a log-odds value.
-For the previous results, prior and context for the random permutation model were combined by simply adding the terms.
-However, this means that ACT-R's prior base-level learning component (which is a log-odds value) is added to a correlation value, which may not be mathematically appropriate.
-As a fourth architectural concern, another method to combine the prior and context components for the random permutation model was explored.
-
-\subsubsection{Method}
-
-For a given retrieval, the random permutation model returns a set of activations that are correlation values for each of the tags.
-Instead of simply adding this correlation value to the prior model term, it may be more appropriate to convert the correlation value to a log-odds value first, and then add this context term to the log-odds prior term.
-So a method was created to convert a distribution of correlations to a distribution of log-odds values.
-To do this, the one-tailed cumulative distribution function value for each correlation was computed.
-That probability was then converted to a log-odds value in the usual way.
-
-This technique has a few advantages: 
-It is simple and computationally cheap.
-Also, it behaves properly at specific points in the distribution.
-For example, if the conversion is performed on the mean correlation value in the set, the result is a log-odds value of zero.
-This makes sense given that log odds can be interpreted as the amount of odds information (positive or negative) that should be adjusted to the prior log odds, given that new piece of information.
-Since the mean correlation value contains no information regarding if it is better or worse than the other correlation values, it is appropriate to assign this correlation value a zero log-odds adjustment value.
-
-The log-odds transformation technique was tested for the random permutation model on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.
-
-\subsubsection{Results}
-
-Model accuracy results after using the log-odds transformation are depicted in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}.
-Using the log-odds transformation technique has a small effect on model accuracy for StackOverflow (\numNoZero{.45} to \numNoZero{.47}, \myCI{.015}{.009}{.021}{5})
-and minimal to no effect on Twitter (\numNoZero{.25} to \numNoZero{.25}, \myCI{-.002}{-.005}{.001}{60}).
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
-  \caption{
-    Random permutation model accuracy when using log-odds transformation on context for StackOverflow
-    All error bars are the 95\% bootstrapped confidence interval for the mean. 
-  }
-  \label{figContextLogoddsSO}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsT-crop.pdf}}}
-  \caption{
-  Random permutation model accuracy when using log-odds transformation on context for Twitter
-  All error bars are the 95\% bootstrapped confidence interval for the mean. 
-}
-  \label{figContextLogoddsT}
-\end{figure}
-
-\subsubsection{Discussion}
-
-Vector-based models such as the random permutation model \parencite{Sahlgren2008} have included only a context component in the past (i.e., no prior component).
-When the prior component is included for the random permutation model, accuracy improves significantly (compare accuracy for full models to context models in Figures \ref{figContextLogoddsSO} and \ref{figContextLogoddsT}).
-However, it is somewhat surprising that it does not matter much how the context and prior terms are added for the random permutation model.
-Model performance is similar when the context component used for the model is based on a correlation compared to a log-odds value.
-
-There is a slight edge in increased accuracy when a log-odds transformation is used for the StackOverflow dataset, although the magnitude of the improvement in accuracy is small (\numNoZero{.45} to \numNoZero{.47}).
-Nonetheless, it does seem appropriate to convert the context to a log-odds adjustment value when context is added as an additional term to the random permutation model.
-Since performance does not decrease when the transformation is made, may slightly increase,
-and leads to a more natural interpretation of each model term as a log-odds value, it is reasonable to include this step for the random permutation model.
-Future research could explore additional methods to properly combine the terms.
-
-\subsection{Word Order}
-
-As a final architectural concern, vector-based models with and without word order were tested to examine the effect of word order on model accuracy.
-Vector-based models like the random permutation model can easily incorporate word order when making predictions.
-This has been touted as one of their primary strengths \parencites{Jones2007, Sahlgren2008}.
-In order to thoroughly test the impact of adding word order into the model, various methods for incorporating word order were tested on the StackOverflow and Twitter datasets.
-
-\subsubsection{Method}
-
-Three different methods for incorporating word order in the random permutation model were tested:
-taking just the sign of the word's position relative to the tag (direction method), using the position (standard method), and using the position but only including words that were used near the tag (window method).
-For the standard method, all words and their respective position are used to build the representation.
-For the direction method, essentially two representations are created: all co-occurrences with words to the left of the tag, and all co-occurrences with words to the right of the tag.
-The window method is similar to the standard method, where each word's relative position is maintained, but it only includes words near the tag to build the representation.
-\textcite{Sahlgren2008} tested similar methods, and found that accuracy was highest when a narrow window was used for the window method (only including words at most two positions to the left and right of the tag).
-The same $+2$ $-2$ range was used to test the window method on the Twitter dataset.
-
-Only the Twitter dataset was tested, since the StackOverflow dataset contains no word order information between words in the title and body of the post and the tags used for the post.
-The author chooses tags after filling out the title and body forms in the post, so the tags are never intermixed with words in the post like they are for Twitter.
-
-The same four Twitter popular-hashtags datasets were used for testing.
-Both an ordered and orderless context component were created for the random permutation models.
-The orderless context component uses the same co-occurrence matrix used by the Bayesian model to create the representation, since the Bayesian model does not contain any order information.
-Models were tested that used only one of the three word order methods for the ordered context component, used only the orderless context component, used a combination of the two components, and combined context with prior.
-
-\subsubsection{Results}
-
-Model accuracy results after incorporating word order into the random permutation model are depicted in Figure \ref{figContextOrder}.
-All bar plots that include an order component where ``window'' or ``direction'' is not included in the title use the standard ordering method (no direction or windowing, all position information is included).
-This is the default ordering method used, so the names for these bar plots were kept consistent to the names used in other figures so that comparisons could be made across figures.
-Base models that do not use entropy weighting were also included so that effect sizes could be compared between entropy weighting and adding word order information.
-
-\begin{figure}[!htbp]
-  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-orderT-crop.pdf}}}
-  \caption{
-    Model accuracy for various word order methods for the random permutation model for Twitter
-    All error bars are the 95\% bootstrapped confidence interval for the mean. 
-  }
-  \label{figContextOrder}
-\end{figure}
-
-Looking first at models where no prior is added, the random permutation model using only orderless context performs similar to the model using only order context with direction.
-All other order-only models perform worse than the order model using direction and worse than the orderless model.
-
-Full models were also plotted against the model where only the prior and orderless context was included (i.e., no order).
-This was to test how much of the variance in model accuracy is shared between the orderless and order models.
-The results show that there is little unique variance in the order-only models, so order information contributes little to overall model performance above and beyond what an orderless model contributes.
-This can be seen by comparing the two-component model with only prior and orderless context to the three-component model with prior, orderless, and any of the three methods for the order context term.
-
-\subsubsection{Discussion}
-
-This was a surprising result.
-One of the main benefits touted for the random permutation model is that it can easily incorporate word order information into the representation.
-But including word order information does not dramatically improve performance. 
-For the random permutation model, it is much more important to properly handle low-predictor stop words and include a prior component than it is to include word order information,
-at least for the Twitter dataset tested in this study.
-
-The original BEAGLE model \parencite{Jones2007} actually showed only minimal improvement when incorporating word order as well.
-In absolute terms, this model only improved by 2 percentage points in accuracy when trained on the TASA and tested on the TOEFL.
-\textcite{Sahlgren2008}'s results for word order for the random permutation model are similar.
-At a row size of around \num{2000}, model accuracy improves by less than a percentage point when order information is incorporated.
-
-Although incorporating word order into the random permutation model does seem to slightly improve performance,
-the strength of the model is more likely to be in how efficiently it can represent the co-occurrence matrix rather than its ability to include word order in the representation.
-That is, the strength of the model is in its compressed unordered context component term and not the ordered component, at least for the Twitter dataset studied.
-
-Also, the reason why the directional version of ordering is the best predictor of the three ordered components is most likely because order information does not have a large effect,
-so the two representations for words to the left and right of a tag are highly correlated.
-This means that this directionally-ordered component is essentially composed of all the co-occurrence observations in the orderless component after being randomly sampled and partitioned into two separate representations.
-The co-occurrences in these representations are highly correlated with each other, and highly correlated with the orderless component,
-which is why model accuracy is similar for the orderless component and direction ordered component, and why accuracy does not improve when adding the direction ordered component to the orderless component.
-
-\section{General Discussion}
-
-The general discussion is separated into two parts.
-The first section will describe the overall results,
-and the final section will go through each key finding and discuss theoretical implications and future directions for research.
-
-\subsection{Overall Results}
-
-Table \ref{tabOverallResults} summarizes model accuracies for all models and datasets tested in the study.
-This includes the prior-only model tested on the popular-users dataset and the full models (context and prior) tested on the popular-hashtags and randomly-sampled datasets.
-After modifying the models by adding an entropy weighting mechanism to both models and a prior term to the random permutation model,
-accuracy for the compressed random permutation model was nearly equal to the Bayesian model.
-The largest difference between the two models was on the Twitter popular-hashtags dataset (7 percent).
-Both models performed nearly equal for the StackOverflow datasets tested,
-and both models were much more accurate when tested on StackOverflow compared to Twitter.
-
-\begin{table}[!ht]
-  \caption{
-    Summary of accuracy for each model and dataset.
-    The Bayesian model includes entropy weighting for the context words,
-    the RP model includes entropy weighting and log odds transformation when combining the context and prior components.
-  }
-  \label{tabOverallResults}
-  {\tabulinesep=1.2mm
-    \begin{tabu}{llll}
-      \hline
-	Site 		& Dataset		& Model						& Accuracy \\
-	\hline
-	StackOverflow	& popular-users		& Standard prior model				& \numNoZero{.28} \\
-			& randomly-sampled	& Bayesian with entropy 			& \numNoZero{.63} \\
-			& 			& RP with entropy and log odds			& \numNoZero{.63} \\
-	Twitter		& popular-users		& Standard prior model				& \numNoZero{.37} \\
-			& popular-hashtags	& Bayesian with entropy				& \numNoZero{.32} \\
-			& 			& RP with entropy and log odds			& \numNoZero{.25} \\
-	\hline
-    \end{tabu}
-  }
-\end{table}
-
-\subsection{Implications and Future Directions}
-
-\subsubsection{Tagging as a Declarative Memory Retrieval Process}
-
-Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter reasonably well.
-The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
-
-These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
-However, both models are not 100 percent accurate, and the amount of variance left to predict is larger than what would be observed if retrieval noise is included in the model.
-One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval.
-For example, it may be the case that the declarative system is used to provide a set of likely tag candidates to higher-level processes to consider for selection.
-
-Another possibility is that this is not a declarative memory process at all.
-In particular, it is important to note that tagging on these sites is widely used as a method to communicate specific messages to specific groups of people in the community.
-It is unclear how much this environment overlaps with verbal communication between two people, and the declarative memory requests required to successfully communicate in this setting.
-Nonetheless, these models are reasonably accurate in this setting, particularly for StackOverflow.
-Given the results in this research, it at least seems worthwhile to use these publicly-available tagging environments as a reasonable proxy to explore and validate modifications to the current theory.
-
-\subsubsection{Representation}
-
-Given that the storage space required to represent the co-occurrences for the random permutation model is dramatically smaller than the uncompressed Bayesian model,
-and that the compression is lossy and essentially random (i.e., no SVD),
-it is quite interesting that the random permutation model performs nearly as well as the full Bayesian model for this task.
-
-These results pose an interesting and important fundamental question about declarative memory:
-How, precisely, is the neural system configured to represent co-occurrences for declarative memory?
-Since at the behavioral level both models perform similarly, it is possible that the neural system is using something similar to either representation to store co-occurrences.
-Vector-based models are simple enough that they can easily be implemented in a neural network.
-Perhaps a form of vector-based compressed storage is being used at the neural level,
-since it can be implemented in neurons and is an efficient way for the machinery to generate predictions that closely match the statistical structure of the environment
-(i.e., that closely match the uncompressed Bayesian representation).
-
-\subsubsection{Compression}
-
-As previously noted, compression for the random permutation model is essentially random.
-That is, no SVD is done to identify the most predictive components,
-so the compressed representation used by the random permutation model does not contain nearly as much of the variance from the original co-occurrence matrix than it could have.
-However, performing a full SVD on a co-occurrence matrix as large as the ones generated for StackOverflow and Twitter is computationally infeasible.
-Also, when deriving the LSA approach, \textcite{Landauer1997} explicitly state that they make no claims that the brain is computing an SVD at the neural level;
-only that the neural machinery is set up such that the representation makes predictions that are similar to what would have been generated if the matrix decomposition for an SVD would have been computed.
-Perhaps the randomly-compressed representation used for the random permutation model is the way in which the brain tries to efficiently achieve the full LSA- and SVD-like approach that \citeauthor{Landauer1997} propose.
-
-\subsubsection{Word Order}
-
-The relatively strong performance of the random permutation model is not due to the model's ability to represent word order.
-Performance is similar for the random permutation and Bayesian models when the co-occurrence matrix for both models is the same, and not based on word order.
-Performance also does not improve by much when word order is added as the last predictor in the random permutation model.
-The random permutation model seems to work so well mainly due to the way that the full co-occurrence representation is compressed.
-
-\subsubsection{Corpus Size}
-
-The random permutation model was more accurate than the Bayesian model at smaller corpus sizes, while the results were flipped for larger document sizes.
-This is interesting from a practical standpoint because the corpus size is rarely chosen in declarative memory research, and is constrained mainly by the amount of data available in the domain being studied.
-If a researcher is working within a domain that has a relatively small corpus size (generally near or less than \num{1000} documents),
-it may be worthwhile to test the random permutation model in this domain, as this compressed model may be more accurate than the uncompressed Bayesian model.
-
-Further, it is interesting to study the effect of model accuracy on corpus size because people require exposure to only a small number of documents before tagging documents appropriately.
-On StackOverflow, each user asked an average of five questions.
-Model accuracy was high for StackOverflow, but the model required training on 1 million documents before achieving that accuracy.
-The user only needed to ``train'' on five questions and the model needed to train on 1 million questions.
-
-This is a large discrepancy, but this does not necessarily invalidate the declarative memory models used for this research.
-First, the random permutation model was still accurate at smaller document sizes, so perhaps this result provides evidence to favor the random permutation model over the Bayesian model.
-However, the smallest corpus size tested for these models was \num{1000} documents, which is much greater than the average number of questions or tweets that a user contributes to the site.
-
-Second, the models may require large amounts of training because the co-occurrence matrices are being built from scratch.
-That is, the models are being trained on the StackOverflow and Twitter domains without any prior general co-occurrence information to start with.
-Presumably, when a person starts using a site, they can leverage all of their past exposure to different word co-occurrences and then simply add on top of that base the domain-specific co-occurrences.
-In that sense, one could argue that the models are undertrained on generic co-occurrences (people are exposed to many more than one million documents in their lifetime) and overtrained on specific co-occurrences.
-It may be interesting to test these memory models on a hybrid co-occurrence matrix,
-where the majority of the observations are taken from a generic domain (e.g., american literature) and a minority are domain-specific (e.g., StackOverflow word, tag pairs).
-
-\subsubsection{Individual vs. Group Representation of Memory}
-
-Even with the large datasets used in this research, there was not enough data on each individual user to build up separate co-occurrence matrices for them.
-That is, the context component of the declarative memory models used in this research was formed as a collection of observed co-occurrences across all individuals.
-Nonetheless, the prior component for the models was customized for each individual (except in the Twitter popular hashtags dataset when it was not possible).
-Ideally, both components would be customized to the individual, and it remains an open question how much accuracy is influenced by using an aggregate or custom context component in the models.
-
-\subsubsection{Attentional Weight}
-
-The Bayesian declarative memory model for ACT-R has an attentional weight term $W_{j}$ that can be used to attenuate or remove stop words.
-The random permutation model was also easily modified to have an attentional weight term by attenuating each word's set of ones and negative ones values by that word's attentional weight.
-Both a method for stop-word removal (frequency filtering) and stop-word attenuating (entropy weighting) performed well for both models.
-
-It is unclear at this point if this modification is cognitively plausible, even if the modification was cleanly added to the current ACT-R model.
-What can be said from this research is that using a proper method for handling the low-predictor words produced some of the largest improvements in model accuracy compared to the other explorations
-(e.g., compared to word order or properly adding terms).
-Consequently, it may be worthwhile to explore this area more thoroughly in the future.
-
-Also, there may be other ways besides adjusting the attentional weight that these models can be modified to achieve the same performance improvement.
-The entropy weighting term is likely solving a problem that resides only in contextual activation and not prior activation,
-since the activation for the prior component for each word is independent of the presentation of other words, so stop words have no way to interact with tag prior activation in this case.
-
-One alternative explanation is that the entropy weighting measure was needed because the contextual co-occurrence matrix was not yet stable, even with these large dataset sizes.
-If the matrix was not yet stable, then by chance certain stop words would contribute more net activation to specific tags and introduce noise into the retrieval process for that tag.
-So further weighting the contribution of these stop words by leveraging the attentional weight term ensured that this noise was suppressed in the system.
-If this is the case, it would be interesting to see how large a dataset is required to not need to modify the attentional weight term.
-However even if this is the case, the random permutation model would still require entropy weighting for larger dataset sizes,
-since this model simply adds environment vectors to produce the current context vector and has no mechanism (unlike Bayesian normalization) for weighting the contribution of an environment vector.
-
-Another explanation for the Bayesian model is that additional attentional weighting was needed because the contextual matrix was derived only from word by tag co-occurrences and not (word or tag) by (word or tag) co-occurrences.
-This was done for performance reasons as the computational complexity otherwise is simply too high for current hardware.
-It is possible that by adding all pairwise combinations of both words and tags in the posts may soften and stabilize the contribution of stop words when computing contextual activation for a tag, given current context.
-
-Also perhaps a different $S_{ji}$ model may not need entropy weighing at all (e.g., the Hebbian-inspired associative learning model described in \noparencite{thomson2013constraining}).
-Further research could explore how entropy weighting interacts with other models of $S_{ji}$, and not just the random permutation and full Bayesian model.
-
-Nonetheless, it seems likely that the entropy method for computing the attentional weight term $W_{j}$ may be applicable to other declarative memory tasks.
-The entropy method is parameter-free (unlike the frequency-filtering method), completely data driven (much like the co-occurrence matrix), does not require tuning,
-and increases model accuracy for these datasets.
-It is common to handle stop words in some manner when working with Natural Language Processing (NLP) and large text corpuses.
-So, if these large corpuses are used to test declarative memory,
-then the entropy method may be a more cognitively plausible way of dealing with stop words rather than filtering based on a predetermined list or the frequency count in the dataset.
-
-\subsubsection{Prior Component}
-
-Adding a prior component to the random permutation model also increased model accuracy for both StackOverflow and Twitter.
-It is unsurprising that this occurred.
-However, it is quite surprising that this is the first time that the prior component has been included and tested in a random permutation model.
-If the random permutation model is explored as a declarative memory theory for other datasets, it should be worthwhile to include the prior component for those explorations.
-Although it did not seem to matter much how the prior component is combined with the context component in this study, 
-there may be better ways than the log-odds transformation to combine the terms.
-Further research could focus on exploring a larger set of methods for combining the context component of the random permutation model with the prior component from the ACT-R theory.
-
-\subsubsection{Activation Calculation}
-
-The Bayesian and vector-based memory models differ in how the co-occurrence matrix is represented (lossy compression for vector-based models).
-They also differ in how the activation of each tag is calculated, given the different representations.
-The vector-based memory models, like LSA, compute a correlation between the context and each tag, while the Bayesian models sum the log likelihoods that each word in the context occurs with each tag.
-These two methods for computing the activation are correlated but not identical.
-Further research could try to tease apart how model accuracy is independently influenced by the representation (compressed or not) and the activation calculation (correlation or log likelihood).
-To do this, four models could be constructed (uncompressed and compressed representation crossed with correlation and log likelihood activation) compared to the two in this study,
-so that the effects of representation and activation calculation could be partialed out.
-This would allow, for example, the method of compression for the random permutation vector-based model to be analyzed independent of activation calculation.
-The results would help determine the specific representation that the declarative memory system uses to represent co-occurrences,
-and also the specific activation calculation that is carried out by the system.
-
-\subsubsection{Observational vs. Experimental Design}
-
-All data used for this study were collected from publicly-available datasets and were not subject to an experimental design (i.e., observational data).
-There are advantages and disadvantages to both observational and experimental designs.
-One primary advantage with the observational approach here is size and availability: data for millions of posts and tweets are immediately available for analysis.
-However one disadvantage is that there is no way to probe a user at a particular point in time, which could be quite advantageous for answering certain questions.
-One hybrid approach worth considering for future research is to use the publicly available datasets to build up a representation of declarative memory for each user.
-Afterwards, bring a sample of these specific users into a laboratory setting and probe them with specific retrieval requests.
-Multiple users could be asked to tag the same message, and then models could be tested to see if they reliably predict differences in accuracy between the subjects.
-
-\subsubsection{Directly Comparing the Models}
-
-After including the prior and entropy components for both models, overall accuracy was similar for the Bayesian and random permutation models for this task.
-So it is unclear precisely which model is more cognitively plausible (or if both are correct and simply working at different levels of abstraction).
-However, there were two areas in particular where model performance showed relatively large differences: accuracy change as a function of corpus size and reliance on entropy weighting.
-These areas could be used in future research to more directly challenge the models.
-For example, a domain could be picked where enough data is collected to build a custom co-occurrence matrix for each user.
-Then one could look at model accuracy over time for each user as the co-occurrence matrix is increased to see if the trajectory is more accurately modeled by a random permutation (more accurate with lower N) or Bayesian (more accurate with higher N) model.
-Further, if a large-scale domain is picked and it is shown that the Bayesian model does not require entropy weighting for the co-occurrence matrix after the representations are fully stabilized,
-then this would provide a case where the two models are dramatically different: The random permutation model would require explicitly weighting the stop words (entropy weighting) while the Bayesian model would not.
-
-\subsection{Summary}
-
-This work has shown how large collections of publicly-available behavioral data can be used to explore and test different theories of declarative memory and their individual architectural components.
-This in turn provided [1] a better understanding of how a user's past tagging history influences future tag use,
-[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on large-scale real-world retrieval tasks,
-and [3] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets.
-
-This work has shown how past user history is important, and how a context-only model like the random permutation model can be modified to include the prior component.
-The work has also shown how the strength of the random permutation model may not be in its ability to represent word order,
-but rather in its ability to efficiently compress the full co-occurrence representation used for the Bayesian model.
-Several different formalisms for ACT-R's attentional weighting term were also explored, and an entropy-weighting technique was found to perform well without requiring any parameter tuning.
-Finally, this work has helped raise the question about how co-occurrence observations are represented at the neural level, since two models with quite different representations performed similarly well.
-The results from the study and the new research questions raised will help to better understand how imposing specific architectural constraints on memory models influence what information is retrieved.
-
-\clearpage
-\printbibliography[heading=bibintoc]
+\end{questions}
 
